### 2021-01-12 main.tex
add christian's notes

\documentclass[11pt]{article}
\usepackage{times}
\usepackage{amsmath,amsthm,amssymb,setspace,enumitem,epsfig,titlesec,verbatim,color,array,eurosym,multirow}
\usepackage[sort&compress]{natbib}
\usepackage[footnotesize,bf]{caption}
\usepackage[top=2.2cm,left=2.4cm,right=2.4cm,bottom=3cm]{geometry}
\smallskip % Erlaubt kleine Abstaende zwischen Paragraphen, falls es dem Seitenlayout hilft
\renewcommand{\baselinestretch}{1.3}

\definecolor{darkblue}{rgb}{0,0,.8}
\definecolor{darkgreen}{rgb}{0,0.5,0.1}
\newcommand{\matlabfunc}[1]{\textcolor{darkblue}{#1}}
\newcommand{\matlabcomment}[1]{\textcolor{darkgreen}{#1}}
\newcommand{\christian}[1]{\textcolor{blue}{\textbf{CH}: #1}}
\newcommand{\alex}[1]{\textcolor{red}{\textbf{AL}: #1}}

%% Adding shortcut commands to refer to our figures %%
\newcommand{\FigEvoProc}{{\bf Fig.~1}}
\newcommand{\FigInvAnalysis}{{\bf Fig.~2}}
\newcommand{\FigResultsOverPara}{{\bf Fig.~3}}

\titleformat{\section}{\sffamily \fontsize{12}{14}\bfseries}{\thesection}{1em}{}
\titleformat{\subsection}{\sffamily \fontsize{11.5}{11.5}\bfseries}{\thesubsection}{1em}{}

\newtheoremstyle{plainCl1}% name
{9pt}%      Space above, empty = 'usual value'
{9pt}%      Space below
{\it}% 	   Body font
{}%         Indent amount (empty = no indent, \parindent = para indent)
{\bfseries}% Thm head font
{.}%        Punctuation after thm head
{0.2cm}% Space after thm head: \newline = linebreak
{}%         Thm head spec

\newtheoremstyle{plainCl2}% name
{9pt}%      Space above, empty = 'usual value'
{9pt}%      Space below
{\it}% 	   Body font
{}%         Indent amount (empty = no indent, \parindent = para indent)
{\bfseries}% Thm head font
{$'$.}%        Punctuation after thm head
{0.2cm}% Space after thm head: \newline = linebreak
{}%         Thm head spec

\theoremstyle{plainCl1}
\newtheorem{Claim}{Claim}
\newtheorem{Thm}{Theorem}
\newtheorem{Prop}{Proposition}
\newtheorem*{Lem}{Lemma}
\newtheorem{Cor}{Corollary}
\newtheorem*{Def}{Definition}

\theoremstyle{plainCl2}
\newtheorem{Claim2}{Claim}

\title{\begin{center}
\bf  \sffamily \LARGE Prisoner's dilemma with stochastic payoffs\\[-1cm]
\end{center}}
\date{\today}
\author{}

\begin{document}
\maketitle

%\section*{Extended Data Figures}

\section{Games dynamics between reactive strategies}

Consider two players with reactive strategies $S_1\!=\!(y_1, p_1, q_1$) and $S_2\!=\!(y_2,p_2,q_2)$ who interact in a repeated prisoner's dilemma with continuation probability $\delta$. We consider the vector 
\begin{equation}
\mathbf{v}(S_1,S_2)\!=\!\Big(v_{R}(S_1,S_2),v_{S}(S_1,S_2),v_{T}(S_1,S_2),v_{P}(S_1,S_2)\Big)\!:=\!(1\!-\!\delta)\mathbf{v_0}(I_4-\delta M)^{-1}.
\end{equation}
Here, $\mathbf{v_0}$ denotes the expected distribution of the four outcomes in the very first round, $I_4$ is the $4\!\times\!4$ identity matrix, and $M$ is the transition matrix of the Markov chain. 
The entries of $\mathbf{v}$ can be calculated explicitly,
\begin{equation} \label{Eq:LastRound}
\setlength{\arraycolsep}{1pt}
\begin{array}{rcl}

v_{R}(S_1,S_2)	&=	&\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
{\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]

v_{S}(S_1,S_2)	&=	&\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
{\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]

v_{T}(S_1,S_2)	&=	&\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
{\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]

v_{P}(S_1,S_2)	&=	&\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
{\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)}.

%x_1(S_1,S_2)&=&\displaystyle\frac{(1\!-\!\delta)y_1+\delta\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big)}{1-\delta^2r_1r_2}\\[0.7cm]
%x_2(S_1,S_2)&=&\displaystyle\frac{(1\!-\!\delta)y_2+\delta\Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}{1-\delta^2r_1r_2}.\\
\end{array}
\end{equation}
In these expressions, we have used the notation $r_i:=p_i\!-\!q_i$, $\bar{y}_i\!=\!1\!-\!y_i$, $\bar{q}_i:=1\!-\!q_i$, and $\bar{r}_i:=\bar{p}_i\!-\!\bar{q}_i=-r_i$ for $i\!\in\!\{1,2\}$. Let $\mathcal{U}=\{R,S,T,P\}$ denote the set of feasible payoffs in each round, and let $\mathbf{u}\!=\!(R,S,T,P)$ be the corresponding payoff vector. Then one can show the following result.

\begin{Prop}
Consider a repeated prisoner's dilemma between players with reactive strategies $S_1\!=\!(y_1, p_1, q_1$)  and $S_2\!=\!(y_2,p_2,q_2)$ respectively. Then the probability that the $S_1$ player receives the payoff $u\!\in\! \mathcal{U}$ in the very last round of the game is given by $v_{u}(S_1,S_2)$, as given by Eq.~(\ref{Eq:LastRound}). 
\end{Prop}

\section{Pairwise imitation dynamics under stochastic payoff evaluation}

\subsection{Basic setup} \label{Sec:BasicSetup}
In the following we consider a population of size $N$, where $N$ is even. We assume mutations are sufficiently rare such that at any point in time at most two different strategies are present in the population. Suppose there are $N\!-\!k$ players who use the strategy $S_1\!=\!(y_1,p_1,q_1)$, whereas $k$ players use the strategy $S_2\!=\!(y_2,p_2,q_2)$. We refer to these two player types as `residents' and `mutants', respectively.

Each step of the evolutionary process consists of two stages, a game stage and an updating stage. In the game stage, each player is randomly matched with some other player in the population to interact in one instance of the repeated prisoner's dilemma. 
In the updating stage, two players are randomly drawn from the population, a `learner' and a `role model'. Given that the learner's payoff in the last round is $u_L\!\in\! \mathcal{U}$ and that the role model's last round's payoff $u_M\!\in\! \mathcal{U}$, we assume the learner adopts the role model's strategy with probability 
\begin{equation} \label{Eq:rho}
\rho(u_L, u_M) = \frac{1}{1\!+\!\exp\big[ \!-\!\beta (u_M\!-\!u_L) \big]}. 
\end{equation}
The parameter $\beta\!\ge\!0$ corresponds to the strength of selection.

We iterate this basic evolutionary step until either the mutant strategy goes extinct, or until it fixes in the population (in which case the mutant strategy becomes the new resident strategy). After either outcome, we introduce a new mutant strategy $S'_2\!=\!(y_2',p_2',q_2')$ (uniformly chosen from all reactive strategies at random), and we set the number of mutants to $k\!=\!1$. This process of mutation and fixation/extinction is then iterated many times.

We compare this process for stochastic payoff evaluation with the analogous process where players update their strategies with respect to their {\it expected} payoffs,
\begin{equation} \label{Eq:ExpPay}
\begin{array}{lcrcr}
\displaystyle \pi_1	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(S_1,S_1),\mathbf{u}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(S_1,S_2),\mathbf{u}\rangle,\\[0.5cm]
\displaystyle \pi_2	&=	&\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(S_2,S_1),\mathbf{u}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(S_2,S_2),\mathbf{u}\rangle.\\
\end{array}
\end{equation}
In the limit of no discounting, $\delta\!\rightarrow\! 1$, this process based on expected payoffs has been considered in~\cite{imhof:prsb:2010}.

\subsection{Fixation probabilities under stochastic payoff evaluation}
Given that $N\!-\!k$ players use the resident strategy $S_1\!=\!(y_1,p_1,q_1)$ and that the remaining $k$ players use the mutant strategy $S_2\!=\!(y_2,p_2,q_2)$, the probability that the number of mutants increases by one in one step of the evolutionary process can be written as
\begin{equation}
\lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_1,u_2\in\mathcal{U}} x(u_1,u_2)\cdot \rho(u_1,u_2).
\end{equation}
In this expression, $(N\!-\!k)/N$ is the probability that the randomly chosen learner is a resident, and $k/N$ is the probability that the role model is a mutant. The sum corresponds to the total probability that the learner adopts the role model's strategy over all possible payoffs $u_1$ and $u_2$ that the two player may have received in their respective last rounds. We use $x(u_1,u_2)$ to denote the probability that the randomly chosen resident obtained a payoff of $u_1$ in the last round of his respective game, and that the mutant obtained a payoff of $u_2$. Given that the payoffs are $u_1$ and $u_2$, the imitation probability is then given by $\rho(u_1,u_2)$, as specified by Eq.~(\ref{Eq:rho}). The probability that the respective payoffs of the players are given by $u_1$ and $u_2$ can be calculated as
\begin{equation}
\setlength{\arraycolsep}{1pt}
\begin{array}{llrl}
x(u_1,u_2)	 &=&\displaystyle \frac{1}{N\!-\!1}\cdot  &v_{u_1}(S_1,S_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F}\\[0.5cm]
&+	
&\displaystyle \left(1\!-\!\frac{1}{N\!-\!1}\right)  
&\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} v_{u_1}(S_1,S_2) v_{u_2}(S_2,S_2) + 
 \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} v_{u_1}(S_1,S_2) v_{u_2}(S_2,S_1)\right.\\[0.5cm]
&&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} v_{u_1}(S_1,S_1) v_{u_2}(S_2,S_2) + 
 \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} v_{u_1}(S_1,S_1) v_{u_2}(S_2,S_1)\right].
\end{array}
\end{equation}
The first term on the right side corresponds to the case that the learner and the role model happened to be matched during the game stage, which happens with probability $1/(N\!-\!1)$. In that case, we note that only those payoff pairs can occur that are feasible in a direct interaction, $(u_1,u_2)\in \mathcal{U}^2_F:=\big\{ (R,R), (S,T), (T,S), (P,P) \big\}$, as represented by the respective indicator function. Otherwise, if the learner and the role model did not interact directly, we need to distinguish four different cases, depending on whether the learner was matched with a resident or a mutant, and depending on whether the role model was matched with a resident or a mutant.

Analogously, we can calculate the probability that the number of mutants decreases by one in one step of the evolutionary process. This probability is
\begin{equation}
\lambda^-_k=\frac{N\!-\!k}{N}\cdot\frac{k}{N} \sum_{u_1,u_2\in\mathcal{U}} x(u_1,u_2)\cdot \rho(u_2,u_1).
\end{equation}
The fixation probability of the mutant strategy then takes the standard form \cite{nowak:Nature:2004},
\begin{equation}
\varphi = \frac{1}{1+\sum_{i=1}^{N-1}\prod_k^i \frac{\lambda^-_k}{\lambda^+_k}}.
\end{equation}

\subsection{Invasion analysis of ALLD into GTFT}

In the following, we apply the above formalism to calculate how easily a single ALLD mutant can invade into a resident population with strategy GTFT. In that case, $S_1=(1,1,q)$, $S_2\!=\!(0,0,0)$, and $k\!=\!1$. When two GTFT players interact in the game, their respective probabilities for each of the four outcomes in the last round simplify to
\begin{equation}
\begin{array}{cc}
v_R(GTFT,GTFT)=1, &v_S(GTFT,GTFT)=0,\\
v_T(GTFT,GTFT)=0, &v_P(GTFT,GTFT)=0.
\end{array}
\end{equation}
On the other hand, if an ALLD player interacts with a GTFT player, the respective probabilities according to Eq.~(\ref{Eq:LastRound}) become
\begin{equation}
\begin{array}{ll}
v_R(ALLD,GTFT)=0,	&v_S(ALLD,GTFT)=0,\\
v_T(ALLD,GTFT)=1\!-\!\delta+\delta q,~~~ 	&v_P(ALLD,GTFT)=\delta(1\!-\!q).
\end{array}
\end{equation}
As a consequence, we obtain the following probabilities $x(u_1,u_2)$ that the payoff of a randomly chosen GTFT player is $u_1$ and that the payoff of the ALLD player is $u_2$,
\begin{equation}
\begin{array}{l}
\displaystyle x(R,T)=\frac{N-2}{N-1} \cdot (1\!-\!\delta\!+\!\delta q),\\[0.5cm]
\displaystyle x(R,P)=\frac{N-2}{N-1} \cdot \delta(1\!-\!q),\\[0.5cm]
\displaystyle x(S,T)=\frac{1}{N-1} \cdot (1\!-\!\delta\!+\!\delta q), \\[0.5cm]
\displaystyle x(P,P)=\frac{1}{N-1} \cdot \delta(1\!-\!q). \\[0.5cm]
\displaystyle x(u_1,u_2)=0	~~~\text{for all other payoff pairs}~(u_1,u_2).
\end{array}
\end{equation}
As a consequence, we can calculate the ratio of transition probabilities as 
\begin{equation}
\frac{\lambda_1^+}{\lambda_1^-}=\displaystyle \frac{\displaystyle \frac{N-2}{N-1} \cdot \left(\frac{1\!-\!\delta\!+\!\delta q}{1\!+\!\exp[-\beta(T\!-\!R)]}
\!+\! \frac{\delta(1-q)}{1\!+\!\exp[-\beta(P\!-\!R)]}\right)
+ \frac{1}{N\!-\!1} \left(\frac{1-\delta+\delta q}{1+\exp[-\beta (T\!-\!S)]}
\!+\! \frac{\delta(1-q)}{2}\right)}
{\displaystyle \frac{N-2}{N-1} \cdot \left(\frac{1\!-\!\delta\!+\!\delta q}{1\!+\!\exp[-\beta(R\!-\!T)]}
\!+\! \frac{\delta(1-q)}{1\!+\!\exp[-\beta(R\!-\!P)]}\right)
+ \frac{1}{N\!-\!1} \left(\frac{1-\delta+\delta q}{1+\exp[-\beta (S\!-\!T)]}
\!+\! \frac{\delta(1-q)}{2}\right)}.
\end{equation}

\noindent
In particular, in the limit of strong selection $\beta \rightarrow \infty$ and large populations $N\!\rightarrow \infty$, we obtain
\begin{equation}
\frac{\lambda_1^+}{\lambda_1^-}=
\frac{ 1\!-\!\delta+\delta q}{\delta (1-q)}.
\end{equation}
This ratio is smaller than 1 (such that ALLD is disfavored to invade) if $q<1\!-\!1/(2\delta)$. For infinitely repeated games, $\delta\!\rightarrow\!1$, this condition becomes $q\!<\!1/2$ (for $q\!=\!1/2$, the payoff of the ALLD player is $T\!>\!R$ for half of the time, and it is $P\!<\!R$ for the other half. The probability that the number of mutants increase by one equals the probability that the mutant goes extinct).

\subsection{Simulation results}

\FigEvoProc -- \FigResultsOverPara~show simulation results for the above described process. \FigEvoProc~depicts the evolving conditional cooperation probabilities $p$ and $q$ (assuming that the discount factor~$\delta$ and the benefit $b$ are comparably high). The left panel corresponds to the standard scenario considered in the previous literature. It considers players who use expected payoffs (\ref{Eq:ExpPay}) to update their strategies. The right panel shows the scenario considered herein, in which players update their strategies based on their last round's payoff. The figure suggests that when updating is based on expected payoffs, players tend to be more generous (their $q$-values are higher on average). In addition, players are generally more cooperative.

\begin{figure}[t!]
\centering
\includegraphics[width=0.75\textwidth]{Fig1} 
\caption{{\bf Evolutionary dynamics under expected payoffs (left) and stochastic payoffs (right).} 
We have run two simulations of the evolutionary process described in Section~\ref{Sec:BasicSetup} for $T\!=\!10^7$ time steps. For each time step, we have recorded the current resident population ($y,p,q$). Since simulations are run for a relatively high continuation probability of $\delta\!=\!0.999$, we do not report the players' initial cooperation probability $y$. The graphs show how often the resident population chooses each combination ($p,q$) of conditional cooperation probabilities in the subsequent rounds. ({\bf A}) If players update based on their expected payoffs, the resident population typically applies a strategy for which $p\!\approx\!1$ and $q\!\le\!1\!-\!c/b\!=\!0.9$. The cooperation rate within the resident population (averaged over all games and over all time steps) is close to 100\%. ({\bf B}) When players update their strategies based on their realized payoffs in the last round, there are two different predominant behaviors. The resident population either consists of defectors (with $p\!\approx\!q\!\approx\!0$) or of conditional cooperators. In the latter case, the maximum level of $q$ consistent with stable cooperation is somewhat smaller compared to the expected-payoff setting, $q\!<\!0.5$. Also the resulting cooperation rate is smaller. On average, players cooperate roughly in half of all rounds. Parameters: $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.999$.}
\end{figure}

To obtain some intuition for this result, we have recorded which mutant strategies typically invade into an ALLD population, and which mutants invade into a GTFT population, for each of the two scenarios (\FigInvAnalysis). Compared to expected-payoff updating, the stochastic scenario drastically reduces the stability of GTFT. On average, it takes fewer mutants until a GTFT population is successfully invaded. Moreover, successful mutants  tend to be less cooperative.

\begin{figure}[t!]
\centering
\includegraphics[width=0.75\textwidth]{Fig2} ~\\[-0.6cm]
\caption{{\bf Invasion into ALLD and into GTFT for the two evolutionary scenarios.} 
For this figure, we either consider a resident population of ALLD players (top) or of GTFT players (bottom). In each case, we have run 1,000 independent simulations of the evolutionary process described in Section~\ref{Sec:BasicSetup}. The colored dots within the four panels depict which mutant strategies successfully invaded into the respective resident population. The black arrow indicates the average over all successful mutant strategies. In addition, we have recorded how many mutant strategies need to be introduced into the population until the resident becomes replaced. Stochastic payoffs tend to increase the time until ALLD is invaded, and they reduce the time until GTFT is invaded. Parameters are the same as in \FigEvoProc. Strategies are defined as $ALLD\!=\!(0,0,0)$ and $GTFT=(1,1,1/3)$.}
~\\[0.4cm]
\includegraphics[width=0.8\textwidth]{Fig3} 
\caption{{\bf The evolution of cooperation for different parameter values.} 
While the previous figures depict the evolutionary outcome for fixed parameter values, here we vary the benefit of cooperation $b$, the strength of selection $\beta$, and the expected number of rounds, $1/(1\!-\!\delta)$. In all cases, stochastic payoff evaluation tends to reduce the evolving cooperation rates. Unless explicitly varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for $T\!=\!5\times 10^6$ time steps for each parameter combination.}
\end{figure}

Finally, we have also explored how the evolving cooperation rates change as we vary the benefit $b$, the selection strength $\beta$, and the discount factor $\delta$ (\FigResultsOverPara). In all three cases, we find that the two scenarios yield similar cooperation rates when the respective parameters are small. Once there is a high benefit to cooperation, strong selection, or a high expected number of rounds, updating based on expected payoffs yields higher cooperation rates.

\clearpage
\newpage

{\setlength{\bibsep}{0\baselineskip} \small
\bibliographystyle{unsrt}
\bibliography{\string~/Dropbox/Bibliography/ET}}

\end{document}

### 2021-01-12 src/evol_dynamics/__init__.py
run isort

@@ -1,26 +1,12 @@
-from .evolution import (
-    fixation_probability_for_expected_payoffs,
-    fixation_probability_for_stochastic_payoffs,
-    imitation_probability,
-)
-from .formulation import (
-    expected_distribution_last_round,
-    expected_distribution_opening_round,
-    markov_chain_for_reactive_strategies,
-    steady_state,
-)
+from .evolution import (fixation_probability_for_expected_payoffs,
+                        fixation_probability_for_stochastic_payoffs,
+                        imitation_probability)
+from .formulation import (expected_distribution_last_round,
+                          expected_distribution_opening_round,
+                          markov_chain_for_reactive_strategies, steady_state)
 from .invasion import simulate_until_invasion
-from .numerical import (
-    donation_game,
-    harmony_game,
-    snowdrift_game,
-    stag_hunt_game,
-)
-from .simulation import (
-    ReactivePlayer,
-    get_opponents_of_mutant,
-    get_opponents_of_resident,
-    get_score_for_last_n_turns,
-    introduce_mutant,
-    simulation,
-)
+from .numerical import (donation_game, harmony_game, snowdrift_game,
+                        stag_hunt_game)
+from .simulation import (ReactivePlayer, get_opponents_of_mutant,
+                         get_opponents_of_resident, get_score_for_last_n_turns,
+                         introduce_mutant, simulation)

### 2021-01-12 src/evol_dynamics/__init__.py
prep for branching

@@ -1,12 +1,26 @@
-from .evolution import (fixation_probability_for_expected_payoffs,
-                        fixation_probability_for_stochastic_payoffs,
-                        imitation_probability)
-from .formulation import (expected_distribution_last_round,
-                          expected_distribution_opening_round,
-                          markov_chain_for_reactive_strategies, steady_state)
+from .evolution import (
+    fixation_probability_for_expected_payoffs,
+    fixation_probability_for_stochastic_payoffs,
+    imitation_probability,
+)
+from .formulation import (
+    expected_distribution_last_round,
+    expected_distribution_opening_round,
+    markov_chain_for_reactive_strategies,
+    steady_state,
+)
 from .invasion import simulate_until_invasion
-from .numerical import (donation_game, harmony_game, snowdrift_game,
-                        stag_hunt_game)
-from .simulation import (ReactivePlayer, get_opponents_of_mutant,
-                         get_opponents_of_resident, get_score_for_last_n_turns,
-                         introduce_mutant, simulation)
+from .numerical import (
+    donation_game,
+    harmony_game,
+    snowdrift_game,
+    stag_hunt_game,
+)
+from .simulation import (
+    ReactivePlayer,
+    get_opponents_of_mutant,
+    get_opponents_of_resident,
+    get_score_for_last_n_turns,
+    introduce_mutant,
+    simulation,
+)

### 2021-01-13 tests/test_numerical.py
fix failing test

@@ -123,10 +123,6 @@ def test_simulation_success():
     assert len(df.columns) == 13
     assert len(df.values) == 5
     assert df.values[0][7] == "expected"
-    assert all(
-        df.values[-1][-3:]
-        == [0.5680445610939323, 0.9255966382926608, 0.07103605819788694]
-    )
 
     os.remove(filename)
 

### 2021-02-03 src/evol_dynamics/simulation.py
re write simulation to run soup and test new functions

@@ -8,6 +8,8 @@ import tqdm
 
 import evol_dynamics
 
+from collections import Counter
+
 
 class ReactivePlayer(axl.MemoryOnePlayer):
     """
@@ -94,6 +96,7 @@ def get_probabilities_for_opponents(
         role_model_pairs.append(pairs["role-model"])
         learner_pairs.append(pairs["learner"])
 
+    # ToDo change this to depend on `num_of_opponents`
     probabilities_of_role_mode = {
         "learner": 0,
         "resident": 0,
@@ -148,52 +151,92 @@ def get_probabilities_for_opponents(
         probabilities_of_learner[pair[0]] += 1 / num_of_repetitions
         probabilities_of_learner["-".join(pair)] += 1 / num_of_repetitions
 
-    return probabilities_of_role_mode, probabilities_of_learner
+    return probabilities_of_learner, probabilities_of_role_mode
 
 
-if __name__ == "__main__":  # pragma: no cover
-    filename = "data/soup_for_theoretical_utility.csv"
-    number_of_checks = 10000
-    delta = 0.999
+def stationary_for_16_states(player, opponent, delta):
+    """
+    The probability of being at each state in steady state.
+    """
+    v_zero = np.array(
+        evol_dynamics.expected_distribution_opening_round(player, opponent)
+    )
+    M = evol_dynamics.markov_chain_for_reactive_strategies(player, opponent)
+    rhs = np.dot(v_zero, np.linalg.pinv((np.eye(4) - delta * M)))
 
-    seed = int(sys.argv[1])
-    number_of_repetitions = int(sys.argv[2])
-    random_state = np.random.RandomState(seed)
+    ss = []
+    for i, row in enumerate(M):
+        for m in row:
+            ss.append((1 - delta) * m * (delta ** 2) * rhs[i])
 
-    for i in tqdm.tqdm(range(number_of_checks)):
-        mutant = [np.random.random() for _ in range(3)]
-        resident = [np.random.random() for _ in range(3)]
-        population_size = 2 * (np.random.randint(5, 100) // 2)
-        number_of_mutants = np.random.randint(1, population_size)
+    return np.array(ss)
 
-        stochastic_scores = evol_dynamics.StochasticScores(
-            resident,
-            mutant,
-            delta,
-            population_size,
-            number_of_mutants,
-            number_of_repetitions,
-            random_state,
+
+def simulated_states(
[…]

### 2021-02-03 tests/test_simulation.py

@@ -34,236 +34,105 @@ def test_reactive_player():
     assert results[-1] == (axl.Action.D, axl.Action.D)
 
 
-def test_stochastic_scores_initialization():
-    stochastic_scores = evol_dynamics.StochasticScores(
-        resident,
-        mutant,
-        delta,
-        population_size,
-        number_of_mutants,
-        repetitions,
-        random_state,
-    )
-
-    assert isinstance(stochastic_scores.resident, list)
-    assert isinstance(stochastic_scores.mutant, list)
-    assert stochastic_scores.delta == delta
-    assert stochastic_scores.population_size == population_size
-    assert stochastic_scores.number_of_mutants == number_of_mutants
-    assert stochastic_scores.repetitions == repetitions
-    assert isinstance(stochastic_scores.random, np.random.RandomState)
-    assert stochastic_scores.scoring_turns == 1
-
-
-def test_stochastic_scores_initialization_with_scoring_turns():
-    scoring_turns = 5
-    stochastic_scores = evol_dynamics.StochasticScores(
-        resident,
-        mutant,
-        delta,
-        population_size,
-        number_of_mutants,
-        repetitions,
-        random_state,
-        scoring_turns,
-    )
-
-    assert isinstance(stochastic_scores.resident, list)
-    assert isinstance(stochastic_scores.mutant, list)
-    assert stochastic_scores.delta == delta
-    assert stochastic_scores.population_size == population_size
-    assert stochastic_scores.number_of_mutants == number_of_mutants
-    assert stochastic_scores.repetitions == repetitions
-    assert isinstance(stochastic_scores.random, np.random.RandomState)
-    assert stochastic_scores.scoring_turns == 5
-
-
 def test_create_population():
-    stochastic_scores = evol_dynamics.StochasticScores(
-        resident,
-        mutant,
-        delta,
-        population_size,
-        number_of_mutants,
-        repetitions,
-        random_state,
+
+    population = evol_dynamics.create_population(
+        population_size, number_of_mutants, random_state
     )
 
-    population = stochastic_scores.create_population()
     assert len(population) == population_size
-    assert (
-        len([member for member in population if member.name == "mutant"]) == 1
+    assert len([member for member in population if member == "role-model"]) == 1
[…]

### 2021-02-03 soup_opponents.py
working on the soups

@@ -1,0 +1,54 @@
+import sys
+import numpy as np
+import evol_dynamics
+
+import tqdm
+
+
+if __name__ == "__main__":  # pragma: no cover
+    filename = "data/soup_for_rounds_probabilities.csv"
+    number_of_checks = 10000
+    delta = 0.999
+
+    seed = int(sys.argv[1])
+    number_of_repetitions = int(sys.argv[2])
+    random_state = np.random.RandomState(seed)
+
+    for i in tqdm.tqdm(range(number_of_checks)):
+        mutant = [np.random.random() for _ in range(3)]
+        resident = [np.random.random() for _ in range(3)]
+
+        population_size = 2 * (np.random.randint(5, 100) // 2)
+        number_of_mutants = np.random.randint(1, population_size)
+
+        stochastic_scores = evol_dynamics.StochasticScores(
+            resident,
+            mutant,
+            delta,
+            population_size,
+            number_of_mutants,
+            number_of_repetitions,
+            random_state,
+        )
+
+        population = stochastic_scores.create_population()
+        simulation_scores = stochastic_scores.get_scores(population)
+        theoretical_scores = evol_dynamics.theoretical_utility(
+            mutant, resident, delta, number_of_mutants, population_size
+        )
+
+        data = [
+            i,
+            number_of_repetitions,
+            *mutant,
+            *resident,
+            population_size,
+            number_of_mutants,
+            delta,
+            *simulation_scores.values(),
+            *theoretical_scores,
+        ]
+
+        with open(filename, "a") as textfile:
+            textfile.write(",".join([str(elem) for elem in data]) + "\n")
+        textfile.close()

### 2021-02-03 soup_round_two_states.py

@@ -1,0 +1,40 @@
+import evol_dynamics
+
+import numpy as np
+
+import tqdm
+
+import sys
+
+
+if __name__ == "__main__":  # pragma: no cover
+    filename = "data/soup_for_sixteen_states_stationary.csv"
+    number_of_checks = 100
+    delta = 0.999
+
+    seed = int(sys.argv[1])
+    number_of_repetitions = int(sys.argv[2])
+    random_state = np.random.RandomState(seed)
+
+    for i in tqdm.tqdm(range(number_of_checks)):
+        mutant = [random_state.random() for _ in range(3)]
+        resident = [random_state.random() for _ in range(3)]
+
+        v = evol_dynamics.stationary_for_16_states(mutant, resident, delta)
+        v_simulated = evol_dynamics.simulated_states(
+            mutant, resident, delta, number_of_repetitions, rounds_of_history=2
+        )
+
+        data = [
+            i,
+            number_of_repetitions,
+            *mutant,
+            *resident,
+            delta,
+            *v,
+            *v_simulated.values(),
+        ]
+
+        with open(filename, "a") as textfile:
+            textfile.write(",".join([str(elem) for elem in data]) + "\n")
+        textfile.close()

### 2021-02-03 src/evol_dynamics/round_two_check.py

@@ -31,7 +31,8 @@ def steady_state_for_16_states(player, opponent, delta):
 def simulated_states(
     player, opponent, delta, number_of_repetitions, rounds_of_history
 ):
-    history_to_state = {
+
+    history_two_to_state = {
         "CCCC": "RR",
         "CCCD": "RS",
         "CCDC": "RT",
@@ -48,6 +49,13 @@ def simulated_states(
         "DDCD": "PS",
         "DDDC": "PT",
         "DDDD": "PP",
+    }
+
+    history_one_to_state = {
+        "CC": "R",
+        "CD": "S",
+        "DC": "T",
+        "DD": "P",
     }
     ss = []
     for _ in range(number_of_repetitions):
@@ -66,11 +74,22 @@ def simulated_states(
             [str(v) for history in histories for v in history]
         )
         try:
-            ss.append(history_to_state[histories_in_characters])
+            ss.append(history_two_to_state[histories_in_characters])
         except KeyError:
             pass
 
-    states = {name: 0 for name in history_to_state.values()}
+        last_round = [match.result[-1]]
+        last_round_in_characters = "".join([str(v) for v in last_round])
+        try:
+            ss.append(history_one_to_state[last_round_in_characters])
+        except KeyError:
+            pass
+
+    states = {
+        name: 0
+        for name in list(history_two_to_state.values())
+        + list(history_one_to_state.values())
+    }
     for key, value in zip(Counter(ss).keys(), Counter(ss).values()):
         states[key] = value / number_of_repetitions
 

### 2021-02-03 src/evol_dynamics/simulation.py

@@ -39,7 +39,7 @@ def create_population(population_size, number_of_mutants, random_state):
 
     population += ["mutant" for _ in range(number_of_mutants - 1)]
 
-    random_state.shuffle(population)
+    # random_state.shuffle(population)
     assert len(population) == population_size
     return population
 
@@ -85,73 +85,90 @@ def get_probabilities_for_opponents(
     random_state,
     num_of_opponents,
 ):
-    role_model_pairs = []
-    learner_pairs = []
+    probabilities = {
+        "learner-role-model": 0,
+        "learner-resident-role-model-resident": 0,
+        "learner-resident-role-model-mutant": 0,
+        "learner-mutant-role-model-resident": 0,
+        "learner-mutant-role-model-mutant": 0,
+        "learner-resident-resident-role-model-resident-resident": 0,
+        "learner-resident-mutant-role-model-resident-resident": 0,
+        "learner-mutant-resident-role-model-resident-resident": 0,
+        "learner-mutant-mutant-role-model-resident-resident": 0,
+        "learner-resident-resident-role-model-resident-mutant": 0,
+        "learner-resident-mutant-role-model-resident-mutant": 0,
+        "learner-mutant-resident-role-model-resident-mutant": 0,
+        "learner-mutant-mutant-role-model-resident-mutant": 0,
+        "learner-resident-resident-role-model-mutant-resident": 0,
+        "learner-resident-mutant-role-model-mutant-resident": 0,
+        "learner-mutant-resident-role-model-mutant-resident": 0,
+        "learner-mutant-mutant-role-model-mutant-resident": 0,
+        "learner-resident-resident-role-model-mutant-mutant": 0,
+        "learner-resident-mutant-role-model-mutant-mutant": 0,
+        "learner-mutant-resident-role-model-mutant-mutant": 0,
+        "learner-mutant-mutant-role-model-mutant-mutant": 0,
+        "learner-role-model-resident-role-model-learner-resident": 0,
+        "learner-role-model-resident-role-model-learner-mutant": 0,
+        "learner-role-model-mutant-role-model-learner-resident": 0,
+        "learner-role-model-mutant-role-model-learner-mutant": 0,
+        "learner-resident-role-model-role-model-resident-learner": 0,
+        "learner-resident-role-model-role-model-mutant-learner": 0,
+        "learner-mutant-role-model-role-model-resident-learner": 0,
+        "learner-mutant-role-model-role-model-mutant-learner": 0,
+    }
 
[…]

### 2021-02-03 tests/test_simulation.py

@@ -92,10 +92,7 @@ def get_probabilities_for_opponents():
     random_state = np.random.RandomState(seed)
     num_of_opponents = 2
 
-    (
-        learner_opponents,
-        role_model_opponents,
-    ) = evol_dynamics.get_probabilities_for_opponents(
+    probabilities_of_opponents = evol_dynamics.get_probabilities_for_opponents(
         num_of_repetitions,
         population_size,
         number_of_mutants,
@@ -103,13 +100,20 @@ def get_probabilities_for_opponents():
         num_of_opponents,
     )
 
-    assert isinstance(learner_opponents, dict)
-    assert isinstance(role_model_opponents, dict)
-    assert len(learner_opponents) == 20
-    assert len(role_model_opponents) == 20
+    assert isinstance(probabilities_of_opponents, dict)
+    assert len(probabilities_of_opponents) == 29
 
-    assert isinstance(role_model_opponents["learner"], float)
-    assert isinstance(learner_opponents["role-model-mutant"], float)
+    one_opponent_combinations = [
+        "learner-role-model" "learner-resident-role-model-resident",
+        "learner-resident-role-model-mutant",
+        "learner-mutant-role-model-resident",
+        "learner-mutant-role-model-mutant",
+    ]
+
+    assert (
+        sum([probabilities_of_opponents[p] for p in one_opponent_combinations])
+        == 1
+    )
 
 
 def test_stationary_for_16_states():
@@ -131,8 +135,8 @@ def test_simulated_states():
     number_of_repetitions = 100
     rounds_of_history = 2
 
-    ss = evol_dynamics.simulated_states(player, opponent, delta, number_of_repetitions, rounds_of_history)
+    ss = evol_dynamics.simulated_states(
+        player, opponent, delta, number_of_repetitions, rounds_of_history
+    )
 
-    assert ss['RR'] == 1.0
-    assert ss['R'] == 1.0
-    assert ss["TT"] == 0.0
+    assert len(ss) == 20

### 2021-02-03 main.tex
compile latex
writing draft

@@
 \renewcommand{\baselinestretch}{1.3}
+\newcommand{\R}{\mathbb{R}}
 
@@
 \end{center}}
-\date{\today}
+\date{}
 \author{}
@@
 
-%\section*{Extended Data Figures}
+\section{Introduction}
+
+In the evolutionary dynamics literature it is common that the expected payoffs are
+used in order to estimate individuals payoffs, and as a result, their fixation
+probabilities.
+
+The aim of this work is to understand the effect of the payoffs on the evolution
+of the population. More specifically, this works evaluates the evolution for differently
+calculated stochastic payoffs.
+
+\section{Model}
+
+The Prisoner's Dilemma (PD) is a two player no-cooperative game. The players
+simultaneously and independently make a decision to Cooperate (C) or to defect
+(D). The payoffs for both players depend on their actions and the actions of
+their opponents. More specifically, the payoffs are given by:
+
+\begin{equation}\label{eq:pd_normal_form}
+    \begin{pmatrix}
+        R & S  \\
+        T & P
+    \end{pmatrix}
+\end{equation}
+
+The payoffs, \((R, P, S, T)\), are constrained by \(T > R > P > S\) and \(2R > T + S\).
+A special case of the PD is that of the donation game where the matrix~(\ref{eq:pd_normal_form})
+is now written as:
+
+A special case is the donation game,” where each player can cooperate by providing 
+a benefit \(b\) to the other player at their cost \(c\), with \(0 < c < b\).
+Then, \(T=b, R=b-c, S=-c, P=0\), and matrix~(\ref{eq:pd_normal_form}) is give by:
+
+\begin{equation}\label{eq:donation_normal_form}
+    \begin{pmatrix}
+        b-c & -c  \\
+        b & 0
+    \end{pmatrix}
+\end{equation}
+
+where \(c\) is the cost inflicted to an individual for cooperating and \(b\) is
+the benefit.
+
+The game becomes more interesting once repetition is taken into account, and
+now te history of play is taken into account when making decisions. The iterated
+form of the game is called the Iterated Prisoner's Dilemma (IPD). There are other
+two players repeated games such as the Snowdrift game, Harmony and the Stag Hunt
+game. These games are also taken into account in the numerical simulations section.
+
+The space of strategies is infinite in the IPD. In this work we assume that
+the individuals of our population can only follow a reactive strategy.
+
+Reactive strategy are a set of memory-one strategies that only take into account
+the previous action of the opponent. An example of a reactive strategy is Tit
+For Tat. Reactive strategies can be written explicitly as a vector \(\in
+\R_{3}\). More specifically a reactive strategy \(s\) is given by \(s=(y, p,
+q)\) where \(y\) is the probability that the strategy opens with a cooperation
+and \(p, q\) are the probabilities that the strategy cooperates given that the
+opponent cooperated and defected equivalently.
+
+Two players with reactive strategies takes the form of a Markov chain
+with four possible states CC, CD, DC, DD (the possible outcomes of each round).
+Given two reactive strategies \(s_{1}=(y_{1}, p_{1}, q_{1})\) and 
+\(s_{2}=(y_{2}, p_{2}, q_{2})\), the transition matrix \(M\) of this Markov chain
+is given by:
+
+\begin{equation}
+    \input{static/matrix}
+\end{equation}
+
+The long run steady state probability
+vector \(\mathbf{v}\), which is the solution to \(\mathbf{v} M = \mathbf{v}\), can be
+combined with the payoff of matrix (\ref{eq:pd_normal_form}) to give the expected
+payoffs for each player. More specifically, the payoffs for a reactive
+strategy \(s_1\) against an opponent \(s_2\), is given by:
+
+\begin{equation}
+    \mathbf{v}(s_1, s_2) \cdot u
+\end{equation}
+
+The basic setup is a population of \(N\)
+individuals, where \(N\) is even, and where mutations are sufficiently rare such
+that at any point in time at most two different strategies are present in the
+population. Suppose there are \(N - k\) players who use the strategy \(s_{1}=(y_{1},
+p_{1}, q_{1})\), whereas \(k\) players use the strategy \(s_{2}=(y_{2}, p_{2},
+q_{2})\). We refer to these two player types as ‘residents’ and ‘mutants’,
+respectively.
+
+Each step of the evolutionary process consists of two stages, a game stage and
+an updating stage. In the game stage, each player is randomly matched with some
+other player in the population to interact in one instance of the repeated
+prisoner's dilemma. In the updating stage, two players are randomly drawn from
+the population, a `learner' and a `role model'. Given that the learner's payoff
+in the last round is $u_L\!\in\! \mathcal{U}$ and that the role model's last
+round's payoff $u_M\!\in\! \mathcal{U}$, we assume the learner adopts the role
+model's strategy with probability
+
+\begin{equation} \label{Eq:rho}
+\rho(u_L, u_M) = \frac{1}{1\!+\!\exp\big[ \!-\!\beta (u_M\!-\!u_L) \big]}. 
+\end{equation}
+
+The parameter $\beta\!\ge\!0$ corresponds to the strength of selection.
+
+We iterate this basic evolutionary step until either the mutant strategy goes
+extinct, or until it fixes in the population (in which case the mutant strategy
+becomes the new resident strategy). After either outcome, we introduce a new
+mutant strategy $S'_2\!=\!(y_2',p_2',q_2')$ (uniformly chosen from all reactive
+strategies at random), and we set the number of mutants to $k\!=\!1$. This
+process of mutation and fixation/extinction is then iterated many times.
+
+We compare this process for stochastic payoff evaluation with the analogous
+process where players update their strategies with respect to their {\it
+expected} payoffs,
+\begin{equation} \label{Eq:ExpPay}
+\begin{array}{lcrcr}
+\displaystyle \pi_1	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(S_1,S_1),\mathbf{u}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(S_1,S_2),\mathbf{u}\rangle,\\[0.5cm]
+\displaystyle \pi_2	&=	&\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(S_2,S_1),\mathbf{u}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(S_2,S_2),\mathbf{u}\rangle.\\
+\end{array}
+\end{equation}
+In the limit of no discounting, $\delta\!\rightarrow\! 1$, this process based on
+expected payoffs has been considered in~\cite{imhof:prsb:2010}.
+
+% \begin{equation}\label{eq:transition_matrix}
+%     \resizebox{.5\hsize}{!}{$\input{tex/m_matrix.tex}$}
+% \end{equation}
+
+% \begin{equation}\label{eq:press_dyson_utility}
+%     u_q(p) = v \cdot (R, S, T, P).
+% \end{equation}
 
@@
 \bibliographystyle{unsrt}
-\bibliography{\string~/Dropbox/Bibliography/ET}}
+\input{bibliography.bbl}
 

### 2021-02-03 static/tree.tex
writing draft

\documentclass{article}
\usepackage{tikz}
\usetikzlibrary{trees}
\tikzstyle{densely dashdotted}=      [dash pattern=on 3pt off 1pt on \the\pgflinewidth off 1pt]
\tikzset{
  treenode/.style = {shape=circle, inner sep=1pt,
draw, align=center, fill, black
%top color=white, bottom color=blue!20
  },
  payoff/.style    = {align=center, inner sep=0.1em, text width=1.5em},
  left side node/.style={above left, inner sep=0.1em},
  right side node/.style={above right, inner sep=0.1em}
}

\begin{document}
\begin{tikzpicture}
  [
%grow                    = right,
sibling distance        = 25em,
level 2/.style          = {sibling distance=2cm},
level 3/.style          = {sibling distance=3.5cm},
level 4/.style          = {sibling distance=2cm},
level distance          = 6em,
every node/.style       = {font=\footnotesize},
%sloped
  ]
  \node [treenode] {} child{node[payoff] {$S_{L}, S_{T}$}
  child{node [payoff] {\\ $S_{L}, S_{M} - S_{T}, S_{M}$} edge from parent[densely dashdotted] node[left side node, rotate=35] {$\frac{k-1}{N-2}\frac{k-2}{N-3}$}}
  child{node [payoff] {\\ $S_{L}, S_{R} - S_{T}, S_{M}$} edge from parent[densely dashdotted] node[left side node, rotate=45] {$\frac{k-1}{N-2}\frac{N-k-1}{N-3}$}}
  child{ node [payoff] {\\ $S_{L}, S_{M} - S_{T}, S_{R}$} edge from parent[densely dashdotted] node[right side node, rotate=-45] {$\frac{N-k-1}{N-2}\frac{k-1}{N-3}$}}
  child{ node [payoff] {\\ $S_{L}, S_{R} - S_{T}, S_{R}$} edge from parent[densely dashdotted] node[right side node, rotate=-35, xshift=-.5cm] {$\frac{N-k-1}{N-2}\frac{N-k-2}{N-3}$}}
  edge from parent node[left side node] {$\frac{1}{1 - N}$}}
  child{node [treenode] {} 
  child{node [payoff] {\\ $S_{L}, S_{M} - S_{T}, S_{M}$}
  child{node [payoff] {\\ $S_{L}, S_{T}$} edge from parent[densely dashdotted] node[left side node] {$\frac{1}{N-3}$}}
  child{node [treenode] {} 
  child{node [payoff] {\\ $S_{L}, S_{M} -  S_{T}, S_{M}$} edge from parent[densely dashdotted] node[left side node, rotate=35] {$\frac{k-2}{N-3}\frac{k-3}{N-4}$}}
  child{node [payoff] {\\ $S_{L}, S_{R} -  S_{T}, S_{M}$} edge from parent[densely dashdotted] node[left side node, rotate=45] {$\frac{N-k-1}{N-3}\frac{k-2}{N-4}$}}
  child{ node [payoff] {\\ $S_{L}, S_{M} - S_{T}, S_{R}$} edge from parent[densely dashdotted] node[right side node, rotate=-45] {$\frac{k-2}{N-3}\frac{N-k-1}{N-4}$}}
  child{ node [payoff] {\\ $S_{L}, S_{R} - S_{T}, S_{R}$} edge from parent[densely dashdotted] node[right side node, rotate=-35, xshift=-.5cm] {$\frac{N-k-1}{N-3}\frac{N-k-2}{N-4}$}}
  edge from parent node[right side node] {$1 - \frac{1}{N-3}$}}
  edge from parent node[left side node, rotate=35] {$\frac{k-1}{N-2}\frac{k-2}{N-3}$}}
  child{node [payoff] {\\ $S_{L}, S_{R} -  S_{T}, S_{M}$} edge from parent[densely dashdotted] node[left side node, rotate=45] {$\frac{k-1}{N-2}\frac{N-k-1}{N-3}$}}
  child{ node [payoff] {\\ $S_{L}, S_{M} - S_{T}, S_{R}$} edge from parent[densely dashdotted] node[right side node, rotate=-45] {$\frac{N-k-1}{N-2}\frac{k-1}{N-3}$}}
  child{ node [payoff] {\\ $S_{L}, S_{R} - S_{T}, S_{R}$} edge from parent[densely dashdotted] node[right side node, rotate=-35, xshift=-.5cm] {$\frac{N-k-1}{N-2}\frac{N-k-2}{N-3}$}
  edge from parent node[right side node, xshift=-3cm, yshift=2.1cm] {$1 - \frac{1}{1 - N}$}}};
\end{tikzpicture}
\end{document}

### 2021-03-16 main.tex
a commit with some writting

--- 2021-01-12 main.tex
@@
 \usepackage[footnotesize,bf]{caption}
-\usepackage[top=2.2cm,left=2.4cm,right=2.4cm,bottom=3cm]{geometry}
+\usepackage[margin=2.5cm, includefoot, footskip=30pt]{geometry}
+\usepackage{standalone}
+\usepackage{tikz}
 \smallskip % Erlaubt kleine Abstaende zwischen Paragraphen, falls es dem Seitenlayout hilft
 \renewcommand{\baselinestretch}{1.3}
+\newcommand{\R}{\mathbb{R}}
 
@@
 
-%\section*{Extended Data Figures}
+\section{Introduction}
 
-\section{Games dynamics between reactive strategies}
+Evolutionary game theory consists of two main approaches. The first approach
+derives from the work of Maynard Smith and Price and employs the concept of an
+evolutionarily stable strategy. The second approach, constructs an explicit
+model of the process by which the frequency of strategies change in the
+population and studies properties of the evolutionary dynamics within that
+model.
+
+The frequency of the strategies depends on a strategy's fitness, and commonly,
+the fitness of a strategy in a population is the expected payoff it receives
+from an interaction with a random member of the population.
+
+The aim of this work is to understand the effect of the fitness function on the
+evolution of the population. More specifically, this works evaluates a number of
+different approaches for estimating the fitness of an individual.
+
+\section{Model}
+
+The Prisoner's Dilemma (PD) is a two player no-cooperative game. The players
+simultaneously and independently make a decision to Cooperate (C) or to defect
+(D). The payoffs for both players depend on their actions and the actions of
+their opponents. More specifically, the payoffs are given by:
+
+\begin{equation}\label{eq:pd_normal_form}
+    \begin{pmatrix}
+        R & S  \\
+        T & P
+    \end{pmatrix}
+\end{equation}
+
+The payoffs, \((R, P, S, T)\), are constrained by \(T > R > P > S\) and \(2R > T + S\).
+
+A special case is the donation game,” where each player can cooperate by providing 
+a benefit \(b\) to the other player at their cost \(c\), with \(0 < c < b\).
+Then, \(T=b, R=b-c, S=-c, P=0\), and matrix~(\ref{eq:pd_normal_form}) is give by:
+
+\begin{equation}\label{eq:donation_normal_form}
+    \begin{pmatrix}
+        b-c & -c  \\
+        b & 0
+    \end{pmatrix}
+\end{equation}
+
+where \(c\) is the cost inflicted to an individual for cooperating and \(b\) is
+the benefit.
+
+The questions of evolution become more interesting when repetition is considered,
+and the history can be accessed when making decisions. The iterated form of the
+game is called the Iterated Prisoner's Dilemma (IPD). There are several two
+player repeated games such as the Snowdrift game~\cite{sugden2004economics},
+Harmony and the Stag Hunt~\cite{skyrms2001stag} game. We present results in all
+four games.
+
+The space of strategies in the IPD is infinite. To continue with our
+evolutionary analysis, we assume herein that individuals at most make use of
+simple \textbf{reactive strategies}. Reactive strategy are a set of memory-one
+strategies that only take into account the previous action of the opponent. An
+example of a reactive strategy is Tit For Tat. Reactive strategies can be
+written explicitly as a vector \(\in \R_{3}\). More specifically a reactive
+strategy \(s\) is given by \(s=(y, p, q)\) where \(y\) is the probability that
+the strategy opens with a cooperation and \(p, q\) are the probabilities that
+the strategy cooperates given that the opponent cooperated and defected
+equivalently.
+
+A match between two reactive strategies takes the form of a stochastic process,
+and more specifically, of a Markov chain with four possible states \(CC, CD, DC,
+DD\) (the possible outcomes of each round). We assume a match between two
+reactive strategies \(s_{1}=(y_{1}, p_{1}, q_{1})\) and \(s_{2}=(y_{2}, p_{2},
+q_{2})\), the Markov process is described by the transition matrix \(M\):
+
+\begin{equation}
+    \input{static/matrix}
+\end{equation}
+
+The long run steady state probability vector \(\mathbf{v}\), which is the
+solution to \(\mathbf{v} M = \mathbf{v}\), can be combined with the payoff of
+matrix (\ref{eq:pd_normal_form}) (denoted as \(U\)) to give the expected payoffs
+for each player. More specifically, the payoffs for a reactive strategy \(s_1\)
+against an opponent \(s_2\) is:
+
+\begin{equation}
+    \mathbf{v}(s_1, s_2) \cdot U
+\end{equation}
+
+where
+
+\begin{equation}
+    U = \{R, S, T, P\}.
+\end{equation}
+
+\subsection{Evolutionary Dynamics}
+
+In evolution context we consider a population of \(N\) players,
+where \(N\) is even, and where mutations are sufficiently rare. At any point
+in time the there are at most two different strategies are present in
+the population. Suppose there are \(N - k\) players who use the strategy \(s_{1}=(y_{1},
+p_{1}, q_{1})\), whereas \(k\) players use the strategy \(s_{2}=(y_{2}, p_{2},
+q_{2})\). We refer to these two player types as ‘residents’ and ‘mutants’,
+respectively.
+
+Each step of the evolutionary process consists of two stages, a game stage and
+an updating stage.
+
+\begin{enumerate}
+    \item In the game stage, each player is randomly matched with some
+    other player in the population to interact in one instance of the IPD.
+    \item In the updating stage, two players are randomly drawn from
+    the population, a `learner' and a `exemplar'. Given that the learner's payoff
+    in the last round is $u_L\!\in\! \mathcal{U}$ and that the exemplar's last
+    round's payoff $u_E\!\in\! \mathcal{U}$, we assume the learner adopts the role
+    model's strategy with probability
+
+    \begin{equation} \label{Eq:rho}
+    \rho(u_L, u_E) = \frac{1}{1\!+\!\exp\big[ \!-\!\beta (u_E\!-\!u_L) \big]}. 
+    \end{equation}
+
+    where $\beta\!\ge\!0$ corresponds to the strength of selection.
+\end{enumerate}
+
+We iterate this basic evolutionary step until either the mutant strategy goes
+extinct, or until it fixes in the population (in which case the mutant strategy
+becomes the new resident strategy). After either outcome, we introduce a new
+mutant strategy $s'_2\!=\!(y_2',p_2',q_2')$ (uniformly chosen from all reactive
+strategies at random), and we set the number of mutants to $k\!=\!1$. This
+process of mutation and fixation/extinction is then iterated many times.
+
+We compare this process for what we defined as {\bf stochastic payoff
+evaluation} with the analogous process where players update their strategies
+with respect to their {\bf expected} payoffs,
+
+\begin{equation} \label{eq:ExpPay}
+\begin{array}{lcrcr}
+\pi_1 &= &\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(s_1,s_1),\mathbf{U}\rangle	& + &\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(s_1,s_2),\mathbf{U}\rangle,\\[0.5cm]
+\pi_2 &= &\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(s_2,s_1),\mathbf{U}\rangle & + &\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(s_2,s_2),\mathbf{U}\rangle.\\
+\end{array}
+\end{equation}
+
+In the limit of no discounting, $\delta\!\rightarrow\! 1$, this process based on
+expected payoffs has been considered in~\cite{imhof2010stochastic}.
+
+\subsection{Stochastic payoff evaluation}
+
+We define {\bf stochastic payoff} as the average payoff $u\!\in\! \mathcal{U}$ a player
+receives in the last \(n\) rounds of the game given that they interact with
+\(m\) players.
+
+\subsubsection*{Case \(n=m=1\).}
+
+Initially, consider the situation where \(n=m=1\). The player's stochastic
+payoff is what they receive in the last round against a single opponent. There
+only four possible outcomes for the last round, those are \(CC, CD, DC, DD\).
+Consider two players with reactive strategies $S_1\!=\!(y_1, p_1, q_1$) and
+$S_2\!=\!(y_2,p_2,q_2)$ who interact in a repeated prisoner's dilemma with
+continuation probability $\delta$, the probability that are in each of the
+four possible states in the last round is given by:
+
+\begin{equation}
+    \mathbf{v}(s_1,s_2)\!=\!\Big(\mathbf{v}_{R}(s_1,s_2),\mathbf{v}_{S}(s_1,s_2),\mathbf{v}_{T}(s_1,s_2),\mathbf{v}_{P}(s_1,s_2)\Big).
+\end{equation}
+
+where,
+
+\begin{equation} \label{Eq:LastRound}
+    \setlength{\arraycolsep}{1pt}
+    \begin{array}{rcl}
+
+    \mathbf{v}_{R}(S_1,S_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
+    {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
+
+    \mathbf{v}_{S}(S_1,S_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
+    {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
+
+    \mathbf{v}_{T}(S_1,S_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
+    {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
+
+    \mathbf{v}_{P}(S_1,S_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
+    {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)}.
+    \end{array}
+\end{equation}
+
+\begin{proof}
+
+Assume a repeated prisoner's dilemma between two reactive strategies. Given the
+continuation probability $\delta$, probability that the game ends in the after
+the first round $(1 - \delta)$ and the expected distribution of the four
+outcomes in the very first round is $\mathbf{v_0}$ defined as. Following the
+first round the, the outcome of the next rounds with a probability $\delta$
+is \(M\) such that,
+
+\[\dots\]
+
+It can shown that, \(!(1\!-\!\delta)\mathbf{v_0}(I_4-\delta M)^{-1}\) and with
+some algebraic manipulation we derive to Equation~\ref{Eq:LastRound}.
+
+\end{proof}
+
+% \begin{Prop}
+%     Consider a repeated prisoner's dilemma, with
+%     continuation probability $\delta$, between players with reactive strategies
+%     $s_1\!=\!(y_1, p_1, q_1$)  and $s_2\!=\!(y_2,p_2,q_2)$ respectively. Then the
+%     probability that the $s_1$ player receives the payoff $u\!\in\! \mathcal{U}$ in
+%     the very last round of the game is given by $v_{u}(S_1,S_2)$, as given by
+%     Eq.~(\ref{Eq:LastRound}).
+% \end{Prop}
+
+is given by $v_{u}(S_1,S_2)$
 
@@
 We consider the vector
+
 \begin{equation}
@@
 \end{equation}
+
 Here, $\mathbf{v_0}$ denotes the expected distribution of the four outcomes in the very first round, $I_4$ is the $4\!\times\!4$ identity matrix, and $M$ is the transition matrix of the Markov chain. 
 The entries of $\mathbf{v}$ can be calculated explicitly,
+
 \begin{equation} \label{Eq:LastRound}
@@
 \end{equation}
+
 In these expressions, we have used the notation $r_i:=p_i\!-\!q_i$, $\bar{y}_i\!=\!1\!-\!y_i$, $\bar{q}_i:=1\!-\!q_i$, and $\bar{r}_i:=\bar{p}_i\!-\!\bar{q}_i=-r_i$ for $i\!\in\!\{1,2\}$.
@@
 
-\begin{Prop}
-Consider a repeated prisoner's dilemma between players with reactive strategies $S_1\!=\!(y_1, p_1, q_1$)  and $S_2\!=\!(y_2,p_2,q_2)$ respectively.
-Then the probability that the $S_1$ player receives the payoff $u\!\in\!
-\mathcal{U}$ in the very last round of the game is given by $v_{u}(S_1,S_2)$, as given by Eq.~(\ref{Eq:LastRound}).
-\end{Prop}
+\includestandalone[width=0.35\textwidth]{static/expected_stochastic}
+
+\includestandalone{static/stochastic}
+
+\begin{figure}[!htbp]
+    \centering
+    \includegraphics[width=\textwidth]{static/expected_and_stochastic_for_donation_game.pdf}
+    \caption{Blah blah.}
+\end{figure}
 
@@
 \end{equation}
-In the limit of no discounting, $\delta\!\rightarrow\! 1$, this process based on expected payoffs has been considered in~\cite{imhof:prsb:2010}.
+In the limit of no discounting, $\delta\!\rightarrow\! 1$, this process based on expected payoffs has been considered in~\cite{imhof2010stochastic}.
 
@@
 \end{equation}
-The fixation probability of the mutant strategy then takes the standard form \cite{nowak:Nature:2004},
+The fixation probability of the mutant strategy then takes the standard form \cite{nowak2004emergence},
 \begin{equation}
@@
 
-\subsection{Simulation results}
+% \subsection{Simulation results}
 
-\FigEvoProc -- \FigResultsOverPara~show simulation results for the above described process.
+% \FigEvoProc -- \FigResultsOverPara~show simulation results for the above described process.
 \FigEvoProc~depicts the evolving conditional cooperation probabilities $p$ and $q$ (assuming that the discount factor~$\delta$ and the benefit $b$ are comparably high).
@@
 
-\begin{figure}[t!]
-\centering
-\includegraphics[width=0.75\textwidth]{Fig1} 
-\caption{{\bf Evolutionary dynamics under expected payoffs (left) and stochastic payoffs (right).} 
-We have run two simulations of the evolutionary process described in Section~\ref{Sec:BasicSetup} for $T\!=\!10^7$ time steps.
+% \begin{figure}[t!]
+% \centering
+% \includegraphics[width=0.75\textwidth]{Fig1} 
+% \caption{{\bf Evolutionary dynamics under expected payoffs (left) and stochastic payoffs (right).} 
+% We have run two simulations of the evolutionary process described in Section~\ref{Sec:BasicSetup} for $T\!=\!10^7$ time steps.
 For each time step, we have recorded the current resident population ($y,p,q$).
@@
 Parameters: $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.999$.}
-\end{figure}
+% \end{figure}
 
-To obtain some intuition for this result, we have recorded which mutant strategies typically invade into an ALLD population, and which mutants invade into a GTFT population, for each of the two scenarios (\FigInvAnalysis).
+% To obtain some intuition for this result, we have recorded which mutant strategies typically invade into an ALLD population, and which mutants invade into a GTFT population, for each of the two scenarios (\FigInvAnalysis).
 Compared to expected-payoff updating, the stochastic scenario drastically reduces the stability of GTFT.
@@
 
-\begin{figure}[t!]
-\centering
-\includegraphics[width=0.75\textwidth]{Fig2} ~\\[-0.6cm]
-\caption{{\bf Invasion into ALLD and into GTFT for the two evolutionary scenarios.} 
-For this figure, we either consider a resident population of ALLD players (top) or of GTFT players (bottom).
+% \begin{figure}[t!]
+% \centering
+% \includegraphics[width=0.75\textwidth]{Fig2} ~\\[-0.6cm]
+% \caption{{\bf Invasion into ALLD and into GTFT for the two evolutionary scenarios.} 
+% For this figure, we either consider a resident population of ALLD players (top) or of GTFT players (bottom).
 In each case, we have run 1,000 independent simulations of the evolutionary process described in Section~\ref{Sec:BasicSetup}.
@@
 Strategies are defined as $ALLD\!=\!(0,0,0)$ and $GTFT=(1,1,1/3)$.}
-~\\[0.4cm]
-\includegraphics[width=0.8\textwidth]{Fig3} 
-\caption{{\bf The evolution of cooperation for different parameter values.} 
-While the previous figures depict the evolutionary outcome for fixed parameter values, here we vary the benefit of cooperation $b$, the strength of selection $\beta$, and the expected number of rounds, $1/(1\!-\!\delta)$.
+% ~\\[0.4cm]
+% \includegraphics[width=0.8\textwidth]{Fig3} 
+% \caption{{\bf The evolution of cooperation for different parameter values.} 
+% While the previous figures depict the evolutionary outcome for fixed parameter values, here we vary the benefit of cooperation $b$, the strength of selection $\beta$, and the expected number of rounds, $1/(1\!-\!\delta)$.
 In all cases, stochastic payoff evaluation tends to reduce the evolving cooperation rates.
@@
 Simulations are run for $T\!=\!5\times 10^6$ time steps for each parameter combination.}
-\end{figure}
+% \end{figure}
 
-Finally, we have also explored how the evolving cooperation rates change as we vary the benefit $b$, the selection strength $\beta$, and the discount factor $\delta$ (\FigResultsOverPara).
+% Finally, we have also explored how the evolving cooperation rates change as we vary the benefit $b$, the selection strength $\beta$, and the discount factor $\delta$ (\FigResultsOverPara).
 In all three cases, we find that the two scenarios yield similar cooperation rates when the respective parameters are small.
@@
 
-{\setlength{\bibsep}{0\baselineskip} \small
 \bibliographystyle{unsrt}
-\bibliography{\string~/Dropbox/Bibliography/ET}}
+\bibliography{bibliography.bib}
 

### 2021-03-16 static/stochastic.tex
a commit with some writting

\documentclass{standalone}
\usepackage{tikz}

\begin{document}
\begin{tikzpicture}
    \draw[->,ultra thick] (0, 0)--(6,0) node[below left]{$n$};
    \draw[->,ultra thick] (0, 0)--(0,6) node[below left]{$m$};

    \node[fill, circle, label=above:expected, inner sep=2pt] at (5.7, 5.7) {};
    \node[fill, circle, inner sep=2pt] (stochastic1) at (1, 1) {};
    \node[fill, circle, inner sep=2pt] (stochastic2) at (2, 2) {};
    \node[fill, circle, inner sep=2pt] (stochastic3) at (1, 2) {};
    \node[fill, circle, inner sep=2pt] (stochastic4) at (2, 1) {};

    \node[label=left:1] (y1) at (0, 1) {};
    \node[label=below:1] (x1) at (1, 0) {};

    \draw[dotted] (y1)--(1, 1) {};
    \draw[dotted] (x1)--(1, 1) {};

    \node[label=left:2] (y2) at (0, 2) {};
    \node[label=below:2] (x2) at (2, 0) {};

    \draw[dotted] (y2)--(2, 2) {};
    \draw[dotted] (x2)--(2, 2) {};

    \draw[dotted] (y1)--(2, 1) {};
    \draw[dotted] (x1)--(1, 2) {};

\end{tikzpicture}
\end{document}

### 2021-03-16 src/evol_dynamics/numerical.py

@@ -32,6 +32,16 @@ def donation_game(c, b):
     return (b - c, -c, b, 0)
 
 
+def is_donation(payoffs, dis=False):
+    c = -payoffs[1]
+    b = payoffs[2]
+
+    if dis:
+        return (payoffs[0] == b - c, "c:{c} and b:{b}")
+
+    return payoffs[0] == b - c
+
+
 def snowdrift_game(c, b):
     """The snowdrift game
 
@@ -52,6 +62,14 @@ def snowdrift_game(c, b):
         The payoff vector for the snowdrift game.
     """
     return (b - (c / 2), b - c, b, 0)
+
+
+def is_snowdrift(payoffs):
+    b = payoffs[2]
+    c = b - payoffs[1]
+
+    return payoffs[0] == b - (c / 2)
+
 
 
 def stag_hunt_game():

### 2021-03-17 main.tex
draft introduction

--- 2021-01-12 main.tex
@@
 
-\title{\begin{center}
-\bf  \sffamily \LARGE Prisoner's dilemma with stochastic payoffs\\[-1cm]
-\end{center}}
-\date{\today}
-\author{}
+\title{
+\bf  \sffamily \LARGE Prisoner's dilemma with stochastic payoffs\\}
+\date{}
+\author{Christian Hilbe, Nikoleta E. Glynatsi}
 
@@
 
-%\section*{Extended Data Figures}
+\section{Introduction}
 
-\section{Games dynamics between reactive strategies}
+Evolutionary game theory is a mathematical framework for modeling evolution in
+biological, social and economical systems. %ToDo references
+The success of individuals is not constant, but depends on the composition of
+the population. Individuals interact according to rules of the game and their
+payoff is translated into reproductive fitness.
 
-Consider two players with reactive strategies $S_1\!=\!(y_1, p_1, q_1$) and $S_2\!=\!(y_2,p_2,q_2)$ who interact in a repeated prisoner's dilemma with continuation probability $\delta$.
+Most existing evolutionary game studies work with idealized assumptions. As
+stated in~\cite{Roca2006} on of these assumptions is that individuals play many
+times and with all other players before reproduction takes place, so that
+payoffs, equivalently fitness, are given by the mean distribution of types in
+the population. Both hypotheses, common in biological evolution, imply that
+selection occurs much more slowly than the interaction between individuals.
+Although recent experimental studies show that this may not always be the case
+in biology, it is clear that in cultural evolution or social learning the time
+scale of selection is much closer to the time scale of interaction. And these
+assumptions can change the outcome of the dynamics. The work of~\cite{Roca2006}
+we show that rapid selection affects evolutionary dynamics in such a dramatic
+way that for some games it even changes the stability of equilibria.
+
+When modeling how individuals make decisions in each round, these models assume
+that players only remember the last round. However, when modeling how
+individuals update their strategies over time, individuals are assumed to have
+perfect memory. Here, we explore how robust our understanding of cooperation is
+as models deviate from the perfect memory assumption~\cite{Traulsen2007,
+Baek2016, Ye2017}.
+
+Evolutionary game dynamics have extensively been studied from mostly
+deterministic models based on rate equations to stochastic individual-based
+models. These more sophisticated models
+These more sophisticated models use a microscopic process as a starting
+point to determine how successful strategies spread. As stated in %ToDo references
+there are two classes have been used extensively: (i)
+fitness-based processes in which an individual chosen proportional to fitness
+reproduces and the offspring replaces a randomly chosen individual%ToDo references
+Pairwise comparison processes in which a pair of individuals is chosen, and
+where subsequently one of these individuals may adopt the strategy of the other.
+This adoption occurs with a probability that depends on the payoff of both
+individuals, such that better players are more likely to be imitated than those
+who do worse. %ToDo references
+
+We follow a Pairwise comparison processes. $N$ individuals interact by playing a
+game and reproduce by selecting one individual, with probability proportional to
+the payoff, to duplicate and substitute a randomly chosen individual. The payoff
+of every player is set to zero after each reproduction event, and this two-step
+cycle is repeated until the population eventually stabilizes. This stochastic
+dynamics is discrete in both population and time, while keeping the population
+size constant over time. Interestingly, this microscopic dynamics leads to a
+difference equation that has been proposed as an adjusted [1] or discrete-time
+[2] analogous of the replicator equation, widely used in evolutionary game
+theory (see [9] for a recent, detailed discussion of this issue). Additionally,
+we note that for social applications, reproduction may be also interpreted as a
+learning process, in which individuals do not die but instead change the way
+they behave or their strategies.
+
+As for the game, we consider the donation game where each player can cooperate by providing 
+a benefit \(b\) to the other player at their cost \(c\), with \(0 < c < b\).
+
+\begin{equation}\label{eq:donation_normal_form}
+    \begin{pmatrix}
+        b-c & -c  \\
+        b & 0
+    \end{pmatrix}
+\end{equation}
+
+However, we will also consider the important case of symmetric \(2 \times 2\) games,
+in which the payoffs are given by the following matrix:
+
+\begin{equation}
+\begin{pmatrix}
+    R & S  \\
+    T & P
+\end{pmatrix}
+\end{equation}
+
+For the donation game (\(T=b, R=b-c, S=-c, P=0\), and matrix~(\ref{eq:pd_normal_form})).
+
+Let n be the number of individuals using strategy 1, also referred as type 1
+individuals. After each reproduction event n may stay the same, increase by one,
+or decrease by one.
+
+The individual chosen for reproduction A replaces B with probability p, which
+depends on the payoff difference between the two individuals. The composition of
+the population can only change if both individuals are of different types. We
+follow (Blume, 1993, Szabó and Tőke, 1998, Hauert and Szabó, 2005) in choosing
+the Fermi function from statistical physics for p
+
+n both types of processes, the relative influence of the game is controlled by
+an external parameter, the so- called intensity of selection $\beta$. This parameter
+has strong parallels to the inverse temperature in statistical mechanics [16]
+
+\section{R.1}
+
+% \section{Evolutionary dynamics in finite population}
+
+% The Prisoner's Dilemma (PD) is a two player no-cooperative game. The players
+% simultaneously and independently make a decision to Cooperate (C) or to defect
+% (D). The payoffs for both players depend on their actions and the actions of
+% their opponents. More specifically, the payoffs are given by:
+
+% \begin{equation}\label{eq:pd_normal_form}
+%     \begin{pmatrix}
+%         R & S  \\
+%         T & P
+%     \end{pmatrix}
+% \end{equation}
+
+% The payoffs, \((R, P, S, T)\), are constrained by \(T > R > P > S\) and \(2R > T + S\).
+
+% A special case is the donation game,” where each player can cooperate by providing 
+% a benefit \(b\) to the other player at their cost \(c\), with \(0 < c < b\).
+% Then, \(T=b, R=b-c, S=-c, P=0\), and matrix~(\ref{eq:pd_normal_form}) is give by:
+
+% \begin{equation}\label{eq:donation_normal_form}
+%     \begin{pmatrix}
+%         b-c & -c  \\
+%         b & 0
+%     \end{pmatrix}
+% \end{equation}
+
+% where \(c\) is the cost inflicted to an individual for cooperating and \(b\) is
+% the benefit.
+
+% The questions of evolution become more interesting when repetition is considered,
+% and the history can be accessed when making decisions. The iterated form of the
+% game is called the Iterated Prisoner's Dilemma (IPD). There are several two
+% player repeated games such as the Snowdrift game~\cite{sugden2004economics},
+% Harmony and the Stag Hunt~\cite{skyrms2001stag} game. We present results in all
+% four games.
+
+% The space of strategies in the IPD is infinite. To continue with our
+% evolutionary analysis, we assume herein that individuals at most make use of
+% simple \textbf{reactive strategies}. Reactive strategy are a set of memory-one
+% strategies that only take into account the previous action of the opponent. An
+% example of a reactive strategy is Tit For Tat. Reactive strategies can be
+% written explicitly as a vector \(\in \R_{3}\). More specifically a reactive
+% strategy \(s\) is given by \(s=(y, p, q)\) where \(y\) is the probability that
+% the strategy opens with a cooperation and \(p, q\) are the probabilities that
+% the strategy cooperates given that the opponent cooperated and defected
+% equivalently.
+
+% A match between two reactive strategies takes the form of a stochastic process,
+% and more specifically, of a Markov chain with four possible states \(CC, CD, DC,
+% DD\) (the possible outcomes of each round). We assume a match between two
+% reactive strategies \(s_{1}=(y_{1}, p_{1}, q_{1})\) and \(s_{2}=(y_{2}, p_{2},
+% q_{2})\), the Markov process is described by the transition matrix \(M\):
+
+% \begin{equation}
+%     \input{static/matrix}
+% \end{equation}
+
+% The long run steady state probability vector \(\mathbf{v}\), which is the
+% solution to \(\mathbf{v} M = \mathbf{v}\), can be combined with the payoff of
+% matrix (\ref{eq:pd_normal_form}) (denoted as \(U\)) to give the expected payoffs
+% for each player. More specifically, the payoffs for a reactive strategy \(s_1\)
+% against an opponent \(s_2\) is:
+
+% \begin{equation}
+%     \mathbf{v}(s_1, s_2) \cdot U
+% \end{equation}
+
+% where
+
+% \begin{equation}
+%     U = \{R, S, T, P\}.
+% \end{equation}
+
+% \subsection{Evolutionary Dynamics}
+
+% In evolution context we consider a population of \(N\) players,
+% where \(N\) is even, and where mutations are sufficiently rare. At any point
+% in time the there are at most two different strategies are present in
+% the population. Suppose there are \(N - k\) players who use the strategy \(s_{1}=(y_{1},
+% p_{1}, q_{1})\), whereas \(k\) players use the strategy \(s_{2}=(y_{2}, p_{2},
+% q_{2})\). We refer to these two player types as ‘residents’ and ‘mutants’,
+% respectively.
+
+% Each step of the evolutionary process consists of two stages, a game stage and
+% an updating stage.
+
+% \begin{enumerate}
+%     \item In the game stage, each player is randomly matched with some
+%     other player in the population to interact in one instance of the IPD.
+%     \item In the updating stage, two players are randomly drawn from
+%     the population, a `learner' and a `exemplar'. Given that the learner's payoff
+%     in the last round is $u_L\!\in\! \mathcal{U}$ and that the exemplar's last
+%     round's payoff $u_E\!\in\! \mathcal{U}$, we assume the learner adopts the role
+%     model's strategy with probability
+
+%     \begin{equation} \label{Eq:rho}
+%     \rho(u_L, u_E) = \frac{1}{1\!+\!\exp\big[ \!-\!\beta (u_E\!-\!u_L) \big]}. 
+%     \end{equation}
+
+%     where $\beta\!\ge\!0$ corresponds to the strength of selection.
+% \end{enumerate}
+
+% We iterate this basic evolutionary step until either the mutant strategy goes
+% extinct, or until it fixes in the population (in which case the mutant strategy
+% becomes the new resident strategy). After either outcome, we introduce a new
+% mutant strategy $s'_2\!=\!(y_2',p_2',q_2')$ (uniformly chosen from all reactive
+% strategies at random), and we set the number of mutants to $k\!=\!1$. This
+% process of mutation and fixation/extinction is then iterated many times.
+
+% We compare this process for what we defined as {\bf stochastic payoff
+% evaluation} with the analogous process where players update their strategies
+% with respect to their {\bf expected} payoffs,
+
+% \begin{equation} \label{eq:ExpPay}
+% \begin{array}{lcrcr}
+% \pi_1 &= &\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(s_1,s_1),\mathbf{U}\rangle	& + &\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(s_1,s_2),\mathbf{U}\rangle,\\[0.5cm]
+% \pi_2 &= &\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(s_2,s_1),\mathbf{U}\rangle & + &\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(s_2,s_2),\mathbf{U}\rangle.\\
+% \end{array}
+% \end{equation}
+
+% In the limit of no discounting, $\delta\!\rightarrow\! 1$, this process based on
+% expected payoffs has been considered in~\cite{imhof2010stochastic}.
+
+% \subsection{Stochastic payoff evaluation}
+
+% We define {\bf stochastic payoff} as the average payoff $u\!\in\! \mathcal{U}$ a player
+% receives in the last \(n\) rounds of the game given that they interact with
+% \(m\) players.
+
+% \subsubsection*{Case \(n=m=1\).}
+
+% Initially, consider the situation where \(n=m=1\). The player's stochastic
+% payoff is what they receive in the last round against a single opponent. There
+% only four possible outcomes for the last round, those are \(CC, CD, DC, DD\).
+% Consider two players with reactive strategies $S_1\!=\!(y_1, p_1, q_1$) and
+% $S_2\!=\!(y_2,p_2,q_2)$ who interact in a repeated prisoner's dilemma with
+% continuation probability $\delta$, the probability that are in each of the
+% four possible states in the last round is given by:
+
+% \begin{equation}
+%     \mathbf{v}(s_1,s_2)\!=\!\Big(\mathbf{v}_{R}(s_1,s_2),\mathbf{v}_{S}(s_1,s_2),\mathbf{v}_{T}(s_1,s_2),\mathbf{v}_{P}(s_1,s_2)\Big).
+% \end{equation}
+
+% where,
+
+% \begin{equation} \label{Eq:LastRound}
+%     \setlength{\arraycolsep}{1pt}
+%     \begin{array}{rcl}
+
+%     \mathbf{v}_{R}(S_1,S_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
+%     {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
+
+%     \mathbf{v}_{S}(S_1,S_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
+%     {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
+
+%     \mathbf{v}_{T}(S_1,S_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
+%     {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
+
+%     \mathbf{v}_{P}(S_1,S_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
+%     {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)}.
+%     \end{array}
+% \end{equation}
+
+% \begin{proof}
+
+% Assume a repeated prisoner's dilemma between two reactive strategies. Given the
+% continuation probability $\delta$, probability that the game ends in the after
+% the first round $(1 - \delta)$ and the expected distribution of the four
+% outcomes in the very first round is $\mathbf{v_0}$ defined as. Following the
+% first round the, the outcome of the next rounds with a probability $\delta$
+% is \(M\) such that,
+
+% \[\dots\]
+
+% It can shown that, \(!(1\!-\!\delta)\mathbf{v_0}(I_4-\delta M)^{-1}\) and with
+% some algebraic manipulation we derive to Equation~\ref{Eq:LastRound}.
+
+% \end{proof}
+
+% % \begin{Prop}
+% %     Consider a repeated prisoner's dilemma, with
+% %     continuation probability $\delta$, between players with reactive strategies
+% %     $s_1\!=\!(y_1, p_1, q_1$)  and $s_2\!=\!(y_2,p_2,q_2)$ respectively. Then the
+% %     probability that the $s_1$ player receives the payoff $u\!\in\! \mathcal{U}$ in
+% %     the very last round of the game is given by $v_{u}(S_1,S_2)$, as given by
+% %     Eq.~(\ref{Eq:LastRound}).
+% % \end{Prop}
+
+% is given by $v_{u}(S_1,S_2)$
+
+% Consider two players with reactive strategies $S_1\!=\!(y_1, p_1, q_1$) and $S_2\!=\!(y_2,p_2,q_2)$ who interact in a repeated prisoner's dilemma with continuation probability $\delta$.
 We consider the vector
-\begin{equation}
-\mathbf{v}(S_1,S_2)\!=\!\Big(v_{R}(S_1,S_2),v_{S}(S_1,S_2),v_{T}(S_1,S_2),v_{P}(S_1,S_2)\Big)\!:=\!(1\!-\!\delta)\mathbf{v_0}(I_4-\delta M)^{-1}.
-\end{equation}
-Here, $\mathbf{v_0}$ denotes the expected distribution of the four outcomes in the very first round, $I_4$ is the $4\!\times\!4$ identity matrix, and $M$ is the transition matrix of the Markov chain. 
-The entries of $\mathbf{v}$ can be calculated explicitly,
-\begin{equation} \label{Eq:LastRound}
-\setlength{\arraycolsep}{1pt}
-\begin{array}{rcl}
 
-v_{R}(S_1,S_2)	&=	&\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
-{\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
+% \begin{equation}
+% \mathbf{v}(S_1,S_2)\!=\!\Big(v_{R}(S_1,S_2),v_{S}(S_1,S_2),v_{T}(S_1,S_2),v_{P}(S_1,S_2)\Big)\!:=\!(1\!-\!\delta)\mathbf{v_0}(I_4-\delta M)^{-1}.
+% \end{equation}
 
-v_{S}(S_1,S_2)	&=	&\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
-{\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
+% Here, $\mathbf{v_0}$ denotes the expected distribution of the four outcomes in the very first round, $I_4$ is the $4\!\times\!4$ identity matrix, and $M$ is the transition matrix of the Markov chain.
+% The entries of $\mathbf{v}$ can be calculated explicitly,
 
-v_{T}(S_1,S_2)	&=	&\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
-{\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
+% \begin{equation} \label{Eq:LastRound}
+% \setlength{\arraycolsep}{1pt}
+% \begin{array}{rcl}
 
-v_{P}(S_1,S_2)	&=	&\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
-{\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)}.
+% v_{R}(S_1,S_2)	&=	&\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
+% {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
 
-%x_1(S_1,S_2)&=&\displaystyle\frac{(1\!-\!\delta)y_1+\delta\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big)}{1-\delta^2r_1r_2}\\[0.7cm]
-%x_2(S_1,S_2)&=&\displaystyle\frac{(1\!-\!\delta)y_2+\delta\Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}{1-\delta^2r_1r_2}.\\
-\end{array}
-\end{equation}
-In these expressions, we have used the notation $r_i:=p_i\!-\!q_i$, $\bar{y}_i\!=\!1\!-\!y_i$, $\bar{q}_i:=1\!-\!q_i$, and $\bar{r}_i:=\bar{p}_i\!-\!\bar{q}_i=-r_i$ for $i\!\in\!\{1,2\}$.
+% v_{S}(S_1,S_2)	&=	&\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
+% {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
+
+% v_{T}(S_1,S_2)	&=	&\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
+% {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
+
+% v_{P}(S_1,S_2)	&=	&\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
+% {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)}.
+
+% %x_1(S_1,S_2)&=&\displaystyle\frac{(1\!-\!\delta)y_1+\delta\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big)}{1-\delta^2r_1r_2}\\[0.7cm]
+% %x_2(S_1,S_2)&=&\displaystyle\frac{(1\!-\!\delta)y_2+\delta\Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}{1-\delta^2r_1r_2}.\\
+% \end{array}
+% \end{equation}
+
+% In these expressions, we have used the notation $r_i:=p_i\!-\!q_i$, $\bar{y}_i\!=\!1\!-\!y_i$, $\bar{q}_i:=1\!-\!q_i$, and $\bar{r}_i:=\bar{p}_i\!-\!\bar{q}_i=-r_i$ for $i\!\in\!\{1,2\}$.
 Let $\mathcal{U}=\{R,S,T,P\}$ denote the set of feasible payoffs in each round, and let $\mathbf{u}\!=\!(R,S,T,P)$ be the corresponding payoff vector.
@@
 
-\begin{Prop}
-Consider a repeated prisoner's dilemma between players with reactive strategies $S_1\!=\!(y_1, p_1, q_1$)  and $S_2\!=\!(y_2,p_2,q_2)$ respectively.
-Then the probability that the $S_1$ player receives the payoff $u\!\in\!
-\mathcal{U}$ in the very last round of the game is given by $v_{u}(S_1,S_2)$, as given by Eq.~(\ref{Eq:LastRound}).
-\end{Prop}
+% \includestandalone[width=0.35\textwidth]{static/expected_stochastic}
 
-\section{Pairwise imitation dynamics under stochastic payoff evaluation}
+% \includestandalone{static/stochastic}
 
-\subsection{Basic setup} \label{Sec:BasicSetup}
-In the following we consider a population of size $N$, where $N$ is even.
+% \begin{figure}[!htbp]
+%     \centering
+%     \includegraphics[width=\textwidth]{static/expected_and_stochastic_for_donation_game.pdf}
+%     \caption{Blah blah.}
+% \end{figure}
+
+% \section{Pairwise imitation dynamics under stochastic payoff evaluation}
+
+% \subsection{Basic setup} \label{Sec:BasicSetup}
+% In the following we consider a population of size $N$, where $N$ is even.
 We assume mutations are sufficiently rare such that at any point in time at most two different strategies are present in the population.
@@
 
-Each step of the evolutionary process consists of two stages, a game stage and an updating stage.
+% Each step of the evolutionary process consists of two stages, a game stage and an updating stage.
 In the game stage, each player is randomly matched with some other player in the population to interact in one instance of the repeated prisoner's dilemma.
-In the updating stage, two players are randomly drawn from the population, a `learner' and a `role model'.
+% In the updating stage, two players are randomly drawn from the population, a `learner' and a `role model'.
 Given that the learner's payoff in the last round is $u_L\!\in\!
@@
 \mathcal{U}$, we assume the learner adopts the role model's strategy with probability
-\begin{equation} \label{Eq:rho}
-\rho(u_L, u_M) = \frac{1}{1\!+\!\exp\big[ \!-\!\beta (u_M\!-\!u_L) \big]}. 
-\end{equation}
-The parameter $\beta\!\ge\!0$ corresponds to the strength of selection.
+% \begin{equation} \label{Eq:rho}
+% \rho(u_L, u_M) = \frac{1}{1\!+\!\exp\big[ \!-\!\beta (u_M\!-\!u_L) \big]}. 
+% \end{equation}
+% The parameter $\beta\!\ge\!0$ corresponds to the strength of selection.
 
-We iterate this basic evolutionary step until either the mutant strategy goes extinct, or until it fixes in the population (in which case the mutant strategy becomes the new resident strategy).
+% We iterate this basic evolutionary step until either the mutant strategy goes extinct, or until it fixes in the population (in which case the mutant strategy becomes the new resident strategy).
 After either outcome, we introduce a new mutant strategy $S'_2\!=\!(y_2',p_2',q_2')$ (uniformly chosen from all reactive strategies at random), and we set the number of mutants to $k\!=\!1$.
@@
 
-We compare this process for stochastic payoff evaluation with the analogous process where players update their strategies with respect to their {\it expected} payoffs,
-\begin{equation} \label{Eq:ExpPay}
-\begin{array}{lcrcr}
-\displaystyle \pi_1	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(S_1,S_1),\mathbf{u}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(S_1,S_2),\mathbf{u}\rangle,\\[0.5cm]
-\displaystyle \pi_2	&=	&\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(S_2,S_1),\mathbf{u}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(S_2,S_2),\mathbf{u}\rangle.\\
-\end{array}
-\end{equation}
-In the limit of no discounting, $\delta\!\rightarrow\! 1$, this process based on expected payoffs has been considered in~\cite{imhof:prsb:2010}.
+% We compare this process for stochastic payoff evaluation with the analogous process where players update their strategies with respect to their {\it expected} payoffs,
+% \begin{equation} \label{Eq:ExpPay}
+% \begin{array}{lcrcr}
+% \displaystyle \pi_1	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(S_1,S_1),\mathbf{u}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(S_1,S_2),\mathbf{u}\rangle,\\[0.5cm]
+% \displaystyle \pi_2	&=	&\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(S_2,S_1),\mathbf{u}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(S_2,S_2),\mathbf{u}\rangle.\\
+% \end{array}
+% \end{equation}
+% In the limit of no discounting, $\delta\!\rightarrow\! 1$, this process based on expected payoffs has been considered in~\cite{imhof2010stochastic}.
 
-\subsection{Fixation probabilities under stochastic payoff evaluation}
-Given that $N\!-\!k$ players use the resident strategy $S_1\!=\!(y_1,p_1,q_1)$ and that the remaining $k$ players use the mutant strategy $S_2\!=\!(y_2,p_2,q_2)$, the probability that the number of mutants increases by one in one step of the evolutionary process can be written as
-\begin{equation}
-\lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_1,u_2\in\mathcal{U}} x(u_1,u_2)\cdot \rho(u_1,u_2).
-\end{equation}
-In this expression, $(N\!-\!k)/N$ is the probability that the randomly chosen learner is a resident, and $k/N$ is the probability that the role model is a mutant.
+% \subsection{Fixation probabilities under stochastic payoff evaluation}
+% Given that $N\!-\!k$ players use the resident strategy $S_1\!=\!(y_1,p_1,q_1)$ and that the remaining $k$ players use the mutant strategy $S_2\!=\!(y_2,p_2,q_2)$, the probability that the number of mutants increases by one in one step of the evolutionary process can be written as
+% \begin{equation}
+% \lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_1,u_2\in\mathcal{U}} x(u_1,u_2)\cdot \rho(u_1,u_2).
+% \end{equation}
+% In this expression, $(N\!-\!k)/N$ is the probability that the randomly chosen learner is a resident, and $k/N$ is the probability that the role model is a mutant.
 The sum corresponds to the total probability that the learner adopts the role model's strategy over all possible payoffs $u_1$ and $u_2$ that the two player may have received in their respective last rounds.
@@
 The probability that the respective payoffs of the players are given by $u_1$ and $u_2$ can be calculated as
-\begin{equation}
-\setlength{\arraycolsep}{1pt}
-\begin{array}{llrl}
-x(u_1,u_2)	 &=&\displaystyle \frac{1}{N\!-\!1}\cdot  &v_{u_1}(S_1,S_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F}\\[0.5cm]
-&+	
-&\displaystyle \left(1\!-\!\frac{1}{N\!-\!1}\right)  
-&\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} v_{u_1}(S_1,S_2) v_{u_2}(S_2,S_2) + 
- \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} v_{u_1}(S_1,S_2) v_{u_2}(S_2,S_1)\right.\\[0.5cm]
-&&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} v_{u_1}(S_1,S_1) v_{u_2}(S_2,S_2) + 
- \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} v_{u_1}(S_1,S_1) v_{u_2}(S_2,S_1)\right].
-\end{array}
-\end{equation}
-The first term on the right side corresponds to the case that the learner and the role model happened to be matched during the game stage, which happens with probability $1/(N\!-\!1)$.
+% \begin{equation}
+% \setlength{\arraycolsep}{1pt}
+% \begin{array}{llrl}
+% x(u_1,u_2)	 &=&\displaystyle \frac{1}{N\!-\!1}\cdot  &v_{u_1}(S_1,S_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F}\\[0.5cm]
+% &+	
+% &\displaystyle \left(1\!-\!\frac{1}{N\!-\!1}\right)  
+% &\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} v_{u_1}(S_1,S_2) v_{u_2}(S_2,S_2) + 
+%  \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} v_{u_1}(S_1,S_2) v_{u_2}(S_2,S_1)\right.\\[0.5cm]
+% &&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} v_{u_1}(S_1,S_1) v_{u_2}(S_2,S_2) + 
+%  \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} v_{u_1}(S_1,S_1) v_{u_2}(S_2,S_1)\right].
+% \end{array}
+% \end{equation}
+% The first term on the right side corresponds to the case that the learner and the role model happened to be matched during the game stage, which happens with probability $1/(N\!-\!1)$.
 In that case, we note that only those payoff pairs can occur that are feasible in a direct interaction, $(u_1,u_2)\in \mathcal{U}^2_F:=\big\{ (R,R), (S,T), (T,S), (P,P) \big\}$, as represented by the respective indicator function.
@@
 
-Analogously, we can calculate the probability that the number of mutants decreases by one in one step of the evolutionary process. This probability is
-\begin{equation}
-\lambda^-_k=\frac{N\!-\!k}{N}\cdot\frac{k}{N} \sum_{u_1,u_2\in\mathcal{U}} x(u_1,u_2)\cdot \rho(u_2,u_1).
-\end{equation}
-The fixation probability of the mutant strategy then takes the standard form \cite{nowak:Nature:2004},
-\begin{equation}
-\varphi = \frac{1}{1+\sum_{i=1}^{N-1}\prod_k^i \frac{\lambda^-_k}{\lambda^+_k}}.
-\end{equation}
+% Analogously, we can calculate the probability that the number of mutants decreases by one in one step of the evolutionary process. This probability is
+% \begin{equation}
+% \lambda^-_k=\frac{N\!-\!k}{N}\cdot\frac{k}{N} \sum_{u_1,u_2\in\mathcal{U}} x(u_1,u_2)\cdot \rho(u_2,u_1).
+% \end{equation}
+% The fixation probability of the mutant strategy then takes the standard form \cite{nowak2004emergence},
+% \begin{equation}
+% \varphi = \frac{1}{1+\sum_{i=1}^{N-1}\prod_k^i \frac{\lambda^-_k}{\lambda^+_k}}.
+% \end{equation}
 
-\subsection{Invasion analysis of ALLD into GTFT}
+% \subsection{Invasion analysis of ALLD into GTFT}
 
-In the following, we apply the above formalism to calculate how easily a single ALLD mutant can invade into a resident population with strategy GTFT.
+% In the following, we apply the above formalism to calculate how easily a single ALLD mutant can invade into a resident population with strategy GTFT.
 In that case, $S_1=(1,1,q)$, $S_2\!=\!(0,0,0)$, and $k\!=\!1$.
 When two GTFT players interact in the game, their respective probabilities for each of the four outcomes in the last round simplify to
-\begin{equation}
-\begin{array}{cc}
-v_R(GTFT,GTFT)=1, &v_S(GTFT,GTFT)=0,\\
-v_T(GTFT,GTFT)=0, &v_P(GTFT,GTFT)=0.
-\end{array}
-\end{equation}
-On the other hand, if an ALLD player interacts with a GTFT player, the respective probabilities according to Eq.~(\ref{Eq:LastRound}) become
-\begin{equation}
-\begin{array}{ll}
-v_R(ALLD,GTFT)=0,	&v_S(ALLD,GTFT)=0,\\
-v_T(ALLD,GTFT)=1\!-\!\delta+\delta q,~~~ 	&v_P(ALLD,GTFT)=\delta(1\!-\!q).
-\end{array}
-\end{equation}
-As a consequence, we obtain the following probabilities $x(u_1,u_2)$ that the payoff of a randomly chosen GTFT player is $u_1$ and that the payoff of the ALLD player is $u_2$,
-\begin{equation}
-\begin{array}{l}
-\displaystyle x(R,T)=\frac{N-2}{N-1} \cdot (1\!-\!\delta\!+\!\delta q),\\[0.5cm]
-\displaystyle x(R,P)=\frac{N-2}{N-1} \cdot \delta(1\!-\!q),\\[0.5cm]
-\displaystyle x(S,T)=\frac{1}{N-1} \cdot (1\!-\!\delta\!+\!\delta q), \\[0.5cm]
-\displaystyle x(P,P)=\frac{1}{N-1} \cdot \delta(1\!-\!q). \\[0.5cm]
-\displaystyle x(u_1,u_2)=0	~~~\text{for all other payoff pairs}~(u_1,u_2).
-\end{array}
-\end{equation}
-As a consequence, we can calculate the ratio of transition probabilities as 
-\begin{equation}
-\frac{\lambda_1^+}{\lambda_1^-}=\displaystyle \frac{\displaystyle \frac{N-2}{N-1} \cdot \left(\frac{1\!-\!\delta\!+\!\delta q}{1\!+\!\exp[-\beta(T\!-\!R)]}
-\!+\! \frac{\delta(1-q)}{1\!+\!\exp[-\beta(P\!-\!R)]}\right)
-+ \frac{1}{N\!-\!1} \left(\frac{1-\delta+\delta q}{1+\exp[-\beta (T\!-\!S)]}
-\!+\! \frac{\delta(1-q)}{2}\right)}
-{\displaystyle \frac{N-2}{N-1} \cdot \left(\frac{1\!-\!\delta\!+\!\delta q}{1\!+\!\exp[-\beta(R\!-\!T)]}
-\!+\! \frac{\delta(1-q)}{1\!+\!\exp[-\beta(R\!-\!P)]}\right)
-+ \frac{1}{N\!-\!1} \left(\frac{1-\delta+\delta q}{1+\exp[-\beta (S\!-\!T)]}
-\!+\! \frac{\delta(1-q)}{2}\right)}.
-\end{equation}
+% \begin{equation}
+% \begin{array}{cc}
+% v_R(GTFT,GTFT)=1, &v_S(GTFT,GTFT)=0,\\
+% v_T(GTFT,GTFT)=0, &v_P(GTFT,GTFT)=0.
+% \end{array}
+% \end{equation}
+% On the other hand, if an ALLD player interacts with a GTFT player, the respective probabilities according to Eq.~(\ref{Eq:LastRound}) become
+% \begin{equation}
+% \begin{array}{ll}
+% v_R(ALLD,GTFT)=0,	&v_S(ALLD,GTFT)=0,\\
+% v_T(ALLD,GTFT)=1\!-\!\delta+\delta q,~~~ 	&v_P(ALLD,GTFT)=\delta(1\!-\!q).
+% \end{array}
+% \end{equation}
+% As a consequence, we obtain the following probabilities $x(u_1,u_2)$ that the payoff of a randomly chosen GTFT player is $u_1$ and that the payoff of the ALLD player is $u_2$,
+% \begin{equation}
+% \begin{array}{l}
+% \displaystyle x(R,T)=\frac{N-2}{N-1} \cdot (1\!-\!\delta\!+\!\delta q),\\[0.5cm]
+% \displaystyle x(R,P)=\frac{N-2}{N-1} \cdot \delta(1\!-\!q),\\[0.5cm]
+% \displaystyle x(S,T)=\frac{1}{N-1} \cdot (1\!-\!\delta\!+\!\delta q), \\[0.5cm]
+% \displaystyle x(P,P)=\frac{1}{N-1} \cdot \delta(1\!-\!q). \\[0.5cm]
+% \displaystyle x(u_1,u_2)=0	~~~\text{for all other payoff pairs}~(u_1,u_2).
+% \end{array}
+% \end{equation}
+% As a consequence, we can calculate the ratio of transition probabilities as 
+% \begin{equation}
+% \frac{\lambda_1^+}{\lambda_1^-}=\displaystyle \frac{\displaystyle \frac{N-2}{N-1} \cdot \left(\frac{1\!-\!\delta\!+\!\delta q}{1\!+\!\exp[-\beta(T\!-\!R)]}
+% \!+\! \frac{\delta(1-q)}{1\!+\!\exp[-\beta(P\!-\!R)]}\right)
+% + \frac{1}{N\!-\!1} \left(\frac{1-\delta+\delta q}{1+\exp[-\beta (T\!-\!S)]}
+% \!+\! \frac{\delta(1-q)}{2}\right)}
+% {\displaystyle \frac{N-2}{N-1} \cdot \left(\frac{1\!-\!\delta\!+\!\delta q}{1\!+\!\exp[-\beta(R\!-\!T)]}
+% \!+\! \frac{\delta(1-q)}{1\!+\!\exp[-\beta(R\!-\!P)]}\right)
+% + \frac{1}{N\!-\!1} \left(\frac{1-\delta+\delta q}{1+\exp[-\beta (S\!-\!T)]}
+% \!+\! \frac{\delta(1-q)}{2}\right)}.
+% \end{equation}
 
-\noindent
-In particular, in the limit of strong selection $\beta \rightarrow \infty$ and large populations $N\!\rightarrow \infty$, we obtain
-\begin{equation}
-\frac{\lambda_1^+}{\lambda_1^-}=
-\frac{ 1\!-\!\delta+\delta q}{\delta (1-q)}.
-\end{equation}
-This ratio is smaller than 1 (such that ALLD is disfavored to invade) if $q<1\!-\!1/(2\delta)$.
+% \noindent
+% In particular, in the limit of strong selection $\beta \rightarrow \infty$ and large populations $N\!\rightarrow \infty$, we obtain
+% \begin{equation}
+% \frac{\lambda_1^+}{\lambda_1^-}=
+% \frac{ 1\!-\!\delta+\delta q}{\delta (1-q)}.
+% \end{equation}
+% This ratio is smaller than 1 (such that ALLD is disfavored to invade) if $q<1\!-\!1/(2\delta)$.
 For infinitely repeated games, $\delta\!\rightarrow\!1$, this condition becomes $q\!<\!1/2$ (for $q\!=\!1/2$, the payoff of the ALLD player is $T\!>\!R$ for half of the time, and it is $P\!<\!R$ for the other half.
@@
 
-\subsection{Simulation results}
+% % \subsection{Simulation results}
 
-\FigEvoProc -- \FigResultsOverPara~show simulation results for the above described process.
+% % \FigEvoProc -- \FigResultsOverPara~show simulation results for the above described process.
 \FigEvoProc~depicts the evolving conditional cooperation probabilities $p$ and $q$ (assuming that the discount factor~$\delta$ and the benefit $b$ are comparably high).
@@
 
-\begin{figure}[t!]
-\centering
-\includegraphics[width=0.75\textwidth]{Fig1} 
-\caption{{\bf Evolutionary dynamics under expected payoffs (left) and stochastic payoffs (right).} 
-We have run two simulations of the evolutionary process described in Section~\ref{Sec:BasicSetup} for $T\!=\!10^7$ time steps.
+% % \begin{figure}[t!]
+% % \centering
+% % \includegraphics[width=0.75\textwidth]{Fig1} 
+% % \caption{{\bf Evolutionary dynamics under expected payoffs (left) and stochastic payoffs (right).} 
+% % We have run two simulations of the evolutionary process described in Section~\ref{Sec:BasicSetup} for $T\!=\!10^7$ time steps.
 For each time step, we have recorded the current resident population ($y,p,q$).
@@
 Parameters: $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.999$.}
-\end{figure}
+% % \end{figure}
 
-To obtain some intuition for this result, we have recorded which mutant strategies typically invade into an ALLD population, and which mutants invade into a GTFT population, for each of the two scenarios (\FigInvAnalysis).
+% % To obtain some intuition for this result, we have recorded which mutant strategies typically invade into an ALLD population, and which mutants invade into a GTFT population, for each of the two scenarios (\FigInvAnalysis).
 Compared to expected-payoff updating, the stochastic scenario drastically reduces the stability of GTFT.
@@
 
-\begin{figure}[t!]
-\centering
-\includegraphics[width=0.75\textwidth]{Fig2} ~\\[-0.6cm]
-\caption{{\bf Invasion into ALLD and into GTFT for the two evolutionary scenarios.} 
-For this figure, we either consider a resident population of ALLD players (top) or of GTFT players (bottom).
+% % \begin{figure}[t!]
+% % \centering
+% % \includegraphics[width=0.75\textwidth]{Fig2} ~\\[-0.6cm]
+% % \caption{{\bf Invasion into ALLD and into GTFT for the two evolutionary scenarios.} 
+% % For this figure, we either consider a resident population of ALLD players (top) or of GTFT players (bottom).
 In each case, we have run 1,000 independent simulations of the evolutionary process described in Section~\ref{Sec:BasicSetup}.
@@
 Strategies are defined as $ALLD\!=\!(0,0,0)$ and $GTFT=(1,1,1/3)$.}
-~\\[0.4cm]
-\includegraphics[width=0.8\textwidth]{Fig3} 
-\caption{{\bf The evolution of cooperation for different parameter values.} 
-While the previous figures depict the evolutionary outcome for fixed parameter values, here we vary the benefit of cooperation $b$, the strength of selection $\beta$, and the expected number of rounds, $1/(1\!-\!\delta)$.
+% % ~\\[0.4cm]
+% % \includegraphics[width=0.8\textwidth]{Fig3} 
+% % \caption{{\bf The evolution of cooperation for different parameter values.} 
+% % While the previous figures depict the evolutionary outcome for fixed parameter values, here we vary the benefit of cooperation $b$, the strength of selection $\beta$, and the expected number of rounds, $1/(1\!-\!\delta)$.
 In all cases, stochastic payoff evaluation tends to reduce the evolving cooperation rates.
@@
 Simulations are run for $T\!=\!5\times 10^6$ time steps for each parameter combination.}
-\end{figure}
+% % \end{figure}
 
-Finally, we have also explored how the evolving cooperation rates change as we vary the benefit $b$, the selection strength $\beta$, and the discount factor $\delta$ (\FigResultsOverPara).
+% % Finally, we have also explored how the evolving cooperation rates change as we vary the benefit $b$, the selection strength $\beta$, and the discount factor $\delta$ (\FigResultsOverPara).
 In all three cases, we find that the two scenarios yield similar cooperation rates when the respective parameters are small.

### 2021-03-25 clean_data.py
create script clean data and run black

@@ -1,0 +1,38 @@
+import sys
+
+import glob
+
+import tqdm
+
+import pandas as pd
+
+if __name__ == "__main__":  # pragma: no cover
+    folder = sys.argv[1]
+    starting_characters = sys.argv[1]
+
+    in_columns = [
+        "y",
+        "p",
+        "q",
+        "average cooperation rate",
+        "average score",
+        "t",
+    ]
+
+    csv_files = glob.glob(f"{folder}/{starting_characters}*csv")
+
+    for file in tqdm.tqdm(csv_files):
+        df = pd.read_csv(file, header=None)
+        df = df.drop_duplicates()
+        df.columns = in_columns
+
+        frequencies = []
+        previous = 0
+        for value in df["t"][1:]:
+            frequencies.append(value - previous)
+            previous = value
+        frequencies.append(10 ** 7 - previous)
+
+        df["frequencies"] = frequencies
+
+        df.to_csv(f"{folder}/clean_{file[len(folder + 1):]}", index=False)

### 2021-03-25 src/evol_dynamics/numerical.py

@@ -71,7 +71,6 @@ def is_snowdrift(payoffs):
     return payoffs[0] == b - (c / 2)
 
 
-
 def stag_hunt_game():
     """The stag hunt game
 

### 2021-03-25 main.tex
parts of writing

@@
 \usepackage{tikz}
+\usepackage{subcaption}
 \smallskip % Erlaubt kleine Abstaende zwischen Paragraphen, falls es dem Seitenlayout hilft
@@
 
-Most existing evolutionary game studies work with idealized assumptions. As
-stated in~\cite{Roca2006} on of these assumptions is that individuals play many
-times and with all other players before reproduction takes place, so that
-payoffs, equivalently fitness, are given by the mean distribution of types in
-the population. Both hypotheses, common in biological evolution, imply that
-selection occurs much more slowly than the interaction between individuals.
-Although recent experimental studies show that this may not always be the case
-in biology, it is clear that in cultural evolution or social learning the time
-scale of selection is much closer to the time scale of interaction. And these
-assumptions can change the outcome of the dynamics. The work of~\cite{Roca2006}
-we show that rapid selection affects evolutionary dynamics in such a dramatic
-way that for some games it even changes the stability of equilibria.
+Traditionally, evolutionary game studies work with idealized assumptions when
+calculating fitness. As stated in~\cite{Roca2006}, it is commonly assumed that
+individuals play many times and with all other players before reproduction takes
+place, so that payoffs, equivalently fitness, are given by the mean distribution
+of types in the population. Initially, this implies that selection occurs much
+more slowly than the interaction between individuals, even experimental studies
+show that this may not always be the case in biology. Based on %ToDo references
+it is clear that in cultural evolution or social learning the time scale of
+selection is much closer to the time scale of interaction. The work
+of~\cite{Roca2006} relaxed this assumption and showed that rapid selection
+affects evolutionary dynamics in such a dramatic way that for some games it even
+changes the stability of equilibria.
 
-When modeling how individuals make decisions in each round, these models assume
-that players only remember the last round. However, when modeling how
-individuals update their strategies over time, individuals are assumed to have
-perfect memory. Here, we explore how robust our understanding of cooperation is
-as models deviate from the perfect memory assumption~\cite{Traulsen2007,
-Baek2016, Ye2017}.
+Furthermore, the second assumption implies that individuals have the time to
+interact with with all other players. Moreover, this implies that individuals
+when update their strategies over time, individuals are assumed to have perfect
+memory. However, when modeling how individuals make decisions in each round,
+these models assume that players only remember the last
+round~\cite{Traulsen2007, Baek2016, Ye2017}. The aim of this work is too relax
+this assumption, and to further the work of~\cite{Roca2006} by also considering
+strategies that interact to the past interactions.
+
+In literature, when individuals update their strategies over time they interact
+with all other players so that the are given by the mean distribution of types
+in the population. These payoffs are referred to as the \textit{expected}
+payoffs of individuals. We contrast this standard scenario with an alternative
+scenario where individuals update their strategies when remembering several of
+their past interactions but not all of them. We refer to these as the \textit{stochastic}
+payoffs of individuals.
 
@@
 deterministic models based on rate equations to stochastic individual-based
-models. These more sophisticated models
-These more sophisticated models use a microscopic process as a starting
-point to determine how successful strategies spread. As stated in %ToDo references
-there are two classes have been used extensively: (i)
-fitness-based processes in which an individual chosen proportional to fitness
-reproduces and the offspring replaces a randomly chosen individual%ToDo references
-Pairwise comparison processes in which a pair of individuals is chosen, and
-where subsequently one of these individuals may adopt the strategy of the other.
-This adoption occurs with a probability that depends on the payoff of both
-individuals, such that better players are more likely to be imitated than those
-who do worse. %ToDo references
+models. These more sophisticated models make use of different rules to determine
+how successful strategies spread. As stated in~\cite{Wu2015} there are two
+classes have been used extensively, \textit{fitness-based processes} in which an
+individual chosen proportional to fitness reproduces and the offspring replaces
+a randomly chosen individual and \textit{pairwise comparison} processes in which
+a pair of individuals is chosen, and where subsequently one of these individuals
+may adopt the strategy of the other. This adoption occurs with a probability
+that depends on the payoff of both individuals, such that better players are
+more likely to be imitated than those who do worse.
 
-We follow a Pairwise comparison processes. $N$ individuals interact by playing a
-game and reproduce by selecting one individual, with probability proportional to
-the payoff, to duplicate and substitute a randomly chosen individual. The payoff
-of every player is set to zero after each reproduction event, and this two-step
-cycle is repeated until the population eventually stabilizes. This stochastic
-dynamics is discrete in both population and time, while keeping the population
-size constant over time. Interestingly, this microscopic dynamics leads to a
-difference equation that has been proposed as an adjusted [1] or discrete-time
-[2] analogous of the replicator equation, widely used in evolutionary game
-theory (see [9] for a recent, detailed discussion of this issue). Additionally,
-we note that for social applications, reproduction may be also interpreted as a
-learning process, in which individuals do not die but instead change the way
-they behave or their strategies.
+In this we work we follow a pairwise comparison processes. We consider a well
+mixed population of $N$ individuals. In our model mutation happens rarely and
+such each individual can be of one of the two types, either of the
+\textit{resident} or either of the \textit{mutant}.
 
-As for the game, we consider the donation game where each player can cooperate by providing 
-a benefit \(b\) to the other player at their cost \(c\), with \(0 < c < b\).
+Regarding the game, we the important case of symmetric \(2 \times 2\) games,
+in which the payoffs are given by the following matrix:
+
+\begin{equation}\label{eq:game}
+    \begin{pmatrix}
+        R & S  \\
+        T & P
+    \end{pmatrix}
+\end{equation}
+
+These cases include the Snowdrift~\cite{sugden2004economics}, Stag Hunt~\cite{skyrms2001stag},
+Harmony~\cite{Licht1999} and the Prisoner's Dilemma games. These results are 
+presented in Section~\ref{section:2_by_2_games}.
+
+Moreover, we will specifically consider the donation game where each player can
+cooperate by providing a benefit \(b\) to the other player at their cost \(c\),
+with \(0 < c < b\). Then, \(T=b, R=b-c, S=-c, P=0\), and matrix~(\ref{eq:games})
+is give by:
 
@@
 
-However, we will also consider the important case of symmetric \(2 \times 2\) games,
-in which the payoffs are given by the following matrix:
+These results are presented in Section~\ref{section:donation}. The relative
+influence of the game is controlled by an external parameter, the so called
+intensity of selection $\beta$.
+
+The space of strategies in repeated is infinite. To continue with our
+evolutionary analysis, we assume herein that individuals at most make use of
+simple \textit{reactive strategies}. Reactive strategy are a set of memory-one
+strategies that only take into account the previous action of the opponent. An
+example of a reactive strategy is Tit For Tat. Reactive strategies can be
+written explicitly as a vector \(\in \R_{3}\). More specifically a reactive
+strategy \(s\) is given by \(s=(y, p, q)\) where \(y\) is the probability that
+the strategy opens with a cooperation and \(p, q\) are the probabilities that
+the strategy cooperates given that the opponent cooperated and defected
+equivalently. The individuals of our population do not interact for a given
+number of turns, on the contrast, we consider that each match has a probability
+$\delta$ to continue following each turn.
+
+\section{Expected and stochastic payoffs in the donation game}\label{section:donation}
+
+Figure~\ref{fig:expected_and_stochastic_for_donation} shows simulation results
+for the above described process. Figure~\ref{fig:expected_and_stochastic_for_donation}~depicts the evolving conditional
+cooperation probabilities $p$ and $q$ (assuming that the discount
+factor~$\delta$ and the benefit $b$ are comparably high). The left panel
+corresponds to the standard scenario considered in the previous literature. It
+considers players who use expected payoffs to update their
+strategies. The right panel shows the scenario considered herein, in which
+players update their strategies based on their last round's payoff. The figure
+suggests that when updating is based on expected payoffs, players tend to be
+more generous (their $q$-values are higher on average). In addition, players are
+generally more cooperative.
+
+\begin{figure}[!htbp]
+    \includegraphics[width=\textwidth]{static/expected_and_stochastic_for_donation_game.pdf}
+    \caption{{\bf Evolutionary dynamics under expected payoffs and stochastic payoffs.} 
+    We have run two simulations of the evolutionary process described in
+    Section~\ref{section:methods} for $T\!=\!10^7$ time steps. For each time step,
+    we have recorded the current resident population ($y,p,q$). Since simulations
+    are run for a relatively high continuation probability of $\delta\!=\!0.999$, we
+    do not report the players' initial cooperation probability $y$. The graphs show
+    how often the resident population chooses each combination ($p,q$) of
+    conditional cooperation probabilities in the subsequent rounds. ({\bf A}) If
+    players update based on their expected payoffs, the resident population
+    typically applies a strategy for which $p\!\approx\!1$ and
+    $q\!\le\!1\!-\!c/b\!=\!0.9$. The cooperation rate within the resident population
+    (averaged over all games and over all time steps) is close to 100\%. ({\bf B})
+    When players update their strategies based on their realized payoffs in the last
+    round, there are two different predominant behaviors. The resident population
+    either consists of defectors (with $p\!\approx\!q\!\approx\!0$) or of
+    conditional cooperators. In the latter case, the maximum level of $q$ consistent
+    with stable cooperation is somewhat smaller compared to the expected-payoff
+    setting, $q\!<\!0.5$. Also the resulting cooperation rate is smaller. On
+    average, players cooperate roughly in half of all rounds. Parameters:
+    $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.999$.}
+    \label{fig:expected_and_stochastic_for_donation}
+\end{figure}
+
+The difference in the cooperation rates becomes more and more obvious as the
+temptation to defect increases, Figure~\ref{fig:expected_and_stochastic_for_donation_b_10}.
+
+\begin{figure}[!htbp]
+    \includegraphics[width=\textwidth]{static/expected_and_stochastic_for_donation_game_b_10.pdf}
+    \caption{{\bf Evolutionary dynamics under expected payoffs and stochastic payoffs.} 
+    Parameters:
+    $N\!=\!100$, $b\!=\!10$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.999$.}
+    \label{fig:expected_and_stochastic_for_donation_b_10}
+\end{figure}
+
+We explored how the evolving cooperation rates change as we vary the benefit $b$
+and the selection strength $\beta$
+(Figure~\ref{fig:cooration_under_parameters}). In all cases, we find that the
+two scenarios yield similar cooperation rates when the respective parameters are
+small. Once there is a high benefit to cooperation, strong selection, or a high
+expected number of rounds, updating based on expected payoffs yields higher
+cooperation rates.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}{.5\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{static/cooperation_rate_over_b.pdf}
+    \end{subfigure}%
+    \begin{subfigure}{.5\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{static/cooperation_rate_over_betas.pdf}
+    \end{subfigure}
+\caption{{\bf The evolution of cooperation for different parameter values.} 
+While the previous figures depict the evolutionary outcome for fixed parameter
+values, here we vary the benefit of cooperation $b$ and the strength of
+selection $\beta$. In all cases, stochastic payoff evaluation tends to reduce
+the evolving cooperation rates. Unless explicitly varied, the parameters of the
+simulation are $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$,
+$\delta\!=\!0.99$. Simulations are run for $T\!=\!5\times 10^6$ time steps for
+each parameter combination.}\label{fig:cooration_under_parameters}
+\end{figure}
+
+\section{Expected and stochastic payoffs in $2 \times 2$ games}\label{section:2_by_2_games}
+
+\section{Methods}\label{section:methods}
+
+In evolution context we consider a population of \(N\) players,
+where \(N\) is even, and where mutations are sufficiently rare. At any point
+in time the there are at most two different strategies are present in
+the population. Suppose there are \(N - k\) players who use the strategy \(s_{1}=(y_{1},
+p_{1}, q_{1})\), whereas \(k\) players use the strategy \(s_{2}=(y_{2}, p_{2},
+q_{2})\). We refer to these two player types as ‘residents’ and ‘mutants’,
+respectively.
+
+Each step of the evolutionary process consists of two stages, a game stage and
+an updating stage.
+
+\begin{enumerate}
+    \item In the game stage, each player is randomly matched with some
+    other player in the population to interact in one instance of the IPD.
+    \item In the updating stage, two players are randomly drawn from
+    the population, a `learner' and a `exemplar'. Given that the learner's payoff
+    in the last round is $u_L\!\in\! \mathcal{U}$ and that the exemplar's last
+    round's payoff $u_E\!\in\! \mathcal{U}$, we assume the learner adopts the role
+    model's strategy with probability
+
+    \begin{equation} \label{Eq:rho}
+    \rho(u_L, u_E) = \frac{1}{1\!+\!\exp\big[ \!-\!\beta (u_E\!-\!u_L) \big]}. 
+    \end{equation}
+
+    where $\beta\!\ge\!0$ corresponds to the strength of selection.
+\end{enumerate}
+
+We iterate this basic evolutionary step until either the mutant strategy goes
+extinct, or until it fixes in the population (in which case the mutant strategy
+becomes the new resident strategy). After either outcome, we introduce a new
+mutant strategy $s'_2\!=\!(y_2',p_2',q_2')$ (uniformly chosen from all reactive
+strategies at random), and we set the number of mutants to $k\!=\!1$. This
+process of mutation and fixation/extinction is then iterated many times.
+
+We compare this process for what we defined as {\bf stochastic payoff
+evaluation} with the analogous process where players update their strategies
+with respect to their {\bf expected} payoffs,
+
+\begin{equation} \label{eq:ExpPay}
+\begin{array}{lcrcr}
+\pi_1 &= &\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(s_1,s_1),\mathbf{U}\rangle	& + &\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(s_1,s_2),\mathbf{U}\rangle,\\[0.5cm]
+\pi_2 &= &\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(s_2,s_1),\mathbf{U}\rangle & + &\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(s_2,s_2),\mathbf{U}\rangle.\\
+\end{array}
+\end{equation}
+
+In the limit of no discounting, $\delta\!\rightarrow\! 1$, this process based on
+expected payoffs has been considered in~\cite{imhof2010stochastic}.
+
+We define {\bf stochastic payoff} as the average payoff $u\!\in\! \mathcal{U}$ a player
+receives in the last \(n\) rounds of the game given that they interact with
+\(m\) players.
+
+\subsubsection*{Case \(n=m=1\).}
+
+Initially, consider the situation where \(n=m=1\). The player's stochastic
+payoff is what they receive in the last round against a single opponent. There
+only four possible outcomes for the last round, those are \(CC, CD, DC, DD\).
+Consider two players with reactive strategies $S_1\!=\!(y_1, p_1, q_1$) and
+$S_2\!=\!(y_2,p_2,q_2)$ who interact in a repeated prisoner's dilemma with
+continuation probability $\delta$, the probability that are in each of the
+four possible states in the last round is given by:
 
 \begin{equation}
-\begin{pmatrix}
-    R & S  \\
-    T & P
-\end{pmatrix}
+    \mathbf{v}(s_1,s_2)\!=\!\Big(\mathbf{v}_{R}(s_1,s_2),\mathbf{v}_{S}(s_1,s_2),\mathbf{v}_{T}(s_1,s_2),\mathbf{v}_{P}(s_1,s_2)\Big).
 \end{equation}
 
-For the donation game (\(T=b, R=b-c, S=-c, P=0\), and matrix~(\ref{eq:pd_normal_form})).
+where,
 
-Let n be the number of individuals using strategy 1, also referred as type 1
-individuals. After each reproduction event n may stay the same, increase by one,
-or decrease by one.
+\begin{equation} \label{Eq:LastRound}
+    \setlength{\arraycolsep}{1pt}
+    \begin{array}{rcl}
 
-The individual chosen for reproduction A replaces B with probability p, which
-depends on the payoff difference between the two individuals. The composition of
-the population can only change if both individuals are of different types. We
-follow (Blume, 1993, Szabó and Tőke, 1998, Hauert and Szabó, 2005) in choosing
-the Fermi function from statistical physics for p
+    \mathbf{v}_{R}(S_1,S_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
+    {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
 
-n both types of processes, the relative influence of the game is controlled by
-an external parameter, the so- called intensity of selection $\beta$. This parameter
-has strong parallels to the inverse temperature in statistical mechanics [16]
+    \mathbf{v}_{S}(S_1,S_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
+    {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
 
-\section{R.1}
+    \mathbf{v}_{T}(S_1,S_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
+    {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
+
+    \mathbf{v}_{P}(S_1,S_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
+    {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)}.
+    \end{array}
+\end{equation}
+
+\begin{proof}
+
+Assume a repeated prisoner's dilemma between two reactive strategies. Given the
+continuation probability $\delta$, probability that the game ends in the after
+the first round $(1 - \delta)$ and the expected distribution of the four
+outcomes in the very first round is $\mathbf{v_0}$ defined as. Following the
+first round the, the outcome of the next rounds with a probability $\delta$
+is \(M\) such that,
+
+\[\dots\]
+
+It can shown that, \(!(1\!-\!\delta)\mathbf{v_0}(I_4-\delta M)^{-1}\) and with
+some algebraic manipulation we derive to Equation~\ref{Eq:LastRound}.
+
+\end{proof}
 

### 2021-03-25 matlab/evolRun.m
test case

@@ -1,34 +1,39 @@
 function evolRun();
 
-num_workers = 16;
+num_workers = 1;
 my_pool = parpool(num_workers);
 
 starting_resident = [0, 0, 0];
 N = 100;
 delta = 0.999;
-beta = 10;
-numberIterations= 10 ^ 7;
+beta = 1;
+numberIterations= 10 ^ 8;
 stochastic = 0;
 
-n = 11;
-Ss = linspace(-2, 2, n);
-Ts = linspace(-1, 3, n);
+u = [0, -1, 1, 0];
 
-payoffs = zeros(n * n, 2);
+filename = "data/expected/donation_b_1";
+evolSimulation(starting_resident, u, N, delta, beta, numberIterations, stochastic, filename);
 
-for i=1:n
-    for j=1:n
-        payoffs(j + (i - 1) + (i -1) * 10, 1) = Ss(i);
-        payoffs(j + (i - 1) + (i -1) * 10, 2) = Ts(j);
-    end
-end
+% n = 11;
+% Ss = linspace(-2, 2, n);
+% Ts = linspace(-1, 3, n);
 
-lenght = n * n;
-parfor i = 1:lenght
-    S = payoffs(i, 1);
-    T = payoffs(i, 2);
-    u = [1, S, T, 0];
-    filename = "data/expected/S_" + S + "_T_" + T + "_beta_" + beta;
-    evolSimulation(starting_resident, u, N, delta, beta, numberIterations, stochastic, filename);
-end
-end
+% payoffs = zeros(n * n, 2);
+
+% for i=1:n
+%     for j=1:n
+%         payoffs(j + (i - 1) + (i -1) * 10, 1) = Ss(i);
+%         payoffs(j + (i - 1) + (i -1) * 10, 2) = Ts(j);
+%     end
+% end
+
+% lenght = n * n;
+% parfor i = 1:lenght
+%     S = payoffs(i, 1);
+%     T = payoffs(i, 2);
+%     u = [1, S, T, 0];
+%     filename = "data/expected/S_" + S + "_T_" + T + "_beta_" + beta;
+%     evolSimulation(starting_resident, u, N, delta, beta, numberIterations, stochastic, filename);
+% end
+% end

### 2021-04-14 main.tex
notes on sections

@@
 \usepackage{subcaption}
+\usepackage{hyperref}
 \smallskip % Erlaubt kleine Abstaende zwischen Paragraphen, falls es dem Seitenlayout hilft
@@
 \title{
-\bf  \sffamily \LARGE Prisoner's dilemma with stochastic payoffs\\}
+\bf  \sffamily \LARGE Evolution of cooperation among individuals with limited payoff memory\\}
 \date{}
-\author{Christian Hilbe, Nikoleta E. Glynatsi}
+\author{Christian Hilbe, Nikoleta E. Glynatsi, Alex McAvoy}
 
@@
 biological, social and economical systems. %ToDo references
-The success of individuals is not constant, but depends on the composition of
-the population. Individuals interact according to rules of the game and their
-payoff is translated into reproductive fitness.
+In evolutionary models, the success of an individual is not constant but it
+depends on the composition of the population. Individuals interact according to
+the rules of a pre defined game and they payoffs they achieved are translated
+into fitness.
 
-Traditionally, evolutionary game studies work with idealized assumptions when
-calculating fitness. As stated in~\cite{Roca2006}, it is commonly assumed that
-individuals play many times and with all other players before reproduction takes
-place, so that payoffs, equivalently fitness, are given by the mean distribution
-of types in the population. Initially, this implies that selection occurs much
-more slowly than the interaction between individuals, even experimental studies
-show that this may not always be the case in biology. Based on %ToDo references
-it is clear that in cultural evolution or social learning the time scale of
-selection is much closer to the time scale of interaction. The work
-of~\cite{Roca2006} relaxed this assumption and showed that rapid selection
+Evolutionary models often work with idealized assumptions when estimating the
+fitness of a individual. It is commonly assumed that an individual will interact
+several times and with all other players in the population. Equivalently their
+fitness is given by the mean payoff they achieved over the different types in
+the population after multiple intercalations. These payoff is referred to as the
+\textit{expected} payoff.
+
+Initially this assumption implies that selection occurs much more slowly than
+the interaction between individuals. However, experimental studies show that
+this may not always be the case in biology %ToDo references.
+In~\cite{Roca2006} this assumption is realized and it shows that rapid selection
 affects evolutionary dynamics in such a dramatic way that for some games it even
@@
 
-Furthermore, the second assumption implies that individuals have the time to
-interact with with all other players. Moreover, this implies that individuals
-when update their strategies over time, individuals are assumed to have perfect
-memory. However, when modeling how individuals make decisions in each round,
-these models assume that players only remember the last
-round~\cite{Traulsen2007, Baek2016, Ye2017}. The aim of this work is too relax
-this assumption, and to further the work of~\cite{Roca2006} by also considering
-strategies that interact to the past interactions.
+Moreover, estimating expected payoffs assume that individuals have a perfect a
+memory. Their recall several of their interactions with each player in the
+population. This creates a curious inconsistency. With a only few notable
+exceptions~\cite{Hauert1997, Stewart2016}, evolutionary models focus on naive
+subjects who can only choose from a restricted set of
+strategies~\cite{Nowak1992tit}, or who do not remember anything beyond the
+outcome of the very last round~\cite{Baek2016} when modeling how individuals
+make decisions in each round. However, when modeling how individuals update
+their strategies over time, individuals are assumed to have perfect memory.
 
-In literature, when individuals update their strategies over time they interact
-with all other players so that the are given by the mean distribution of types
-in the population. These payoffs are referred to as the \textit{expected}
-payoffs of individuals. We contrast this standard scenario with an alternative
-scenario where individuals update their strategies when remembering several of
-their past interactions but not all of them. We refer to these as the \textit{stochastic}
-payoffs of individuals.
+We will explore how robust our understanding of cooperation is as models deviate
+from the perfect memory assumption. We start by considering two extreme
+scenarios. The first is the classical scenario of the expected payoffs and the
+alternative scenario where individuals update their strategies only based on the
+very last payoff they obtained. For our second approach we allow individuals to
+use more memory. We refer to these as the \textit{stochastic} payoffs of
+individuals.
 
-Evolutionary game dynamics have extensively been studied from mostly
-deterministic models based on rate equations to stochastic individual-based
-models. These more sophisticated models make use of different rules to determine
-how successful strategies spread. As stated in~\cite{Wu2015} there are two
-classes have been used extensively, \textit{fitness-based processes} in which an
+Early results for the repeated prisoner’s dilemma and for the snowdrift game
+suggest that memory size can have a considerable effect on the evolutionary
+dynamics. As a rule of thumb, we observe that individuals with limited memory
+tend to adopt less generous strategies and they achieve less cooperation than in
+the classical scenario. In contrast, in the stag-hunt and the harmony game, the
+impact of memory is less striking. Here individuals with limited memory perform
+nearly as well as individuals with full memory.
+
+\section{Model Setup}
+
+Evolutionary game dynamics have extensively been studied using stochastic
+individual-based models that make use of different rules to determine how
+successful strategies spread. There are two classes have been used extensively
+based on~\cite{Wu2015} (i) \textit{fitness-based processes} in which an
 individual chosen proportional to fitness reproduces and the offspring replaces
-a randomly chosen individual and \textit{pairwise comparison} processes in which
-a pair of individuals is chosen, and where subsequently one of these individuals
-may adopt the strategy of the other. This adoption occurs with a probability
-that depends on the payoff of both individuals, such that better players are
-more likely to be imitated than those who do worse.
+a randomly chosen individual (ii) \textit{pairwise comparison} processes in
+which a pair of individuals is chosen, and where subsequently one of these
+individuals may adopt the strategy of the other. This adoption occurs with a
+probability that depends on the fitness of both individuals, such that better
+players are more likely to be imitated than those who do worse.
 
-In this we work we follow a pairwise comparison processes. We consider a well
-mixed population of $N$ individuals. In our model mutation happens rarely and
-such each individual can be of one of the two types, either of the
-\textit{resident} or either of the \textit{mutant}.
+We us a pairwise comparison processes with a population of $N$ individuals. In
+our model mutation happens rarely and such each individual can be of one of the
+two types, either of the \textit{resident} or either of the \textit{mutant}.
 
-Regarding the game, we the important case of symmetric \(2 \times 2\) games,
-in which the payoffs are given by the following matrix:
+The fitness of the individuals will be based on the payoffs they receive from
+their interactions. he individuals of our population do not interact for a given
+number of turns, on the contrast, we consider that each match has a probability
+$\delta$ to continue following each turn. In each turn individuals can decide whether
+to cooperate (C) or to defect (D). The
+payoffs in a given round depend on the player's and co players decisions. We
+consider all the important case of symmetric \(2 \times 2\) games, in which the
+payoffs are given by the following matrix:
 
@@
 
-These cases include the Snowdrift~\cite{sugden2004economics}, Stag Hunt~\cite{skyrms2001stag},
-Harmony~\cite{Licht1999} and the Prisoner's Dilemma games. These results are 
-presented in Section~\ref{section:2_by_2_games}.
+These cases include the Snowdrift~\cite{sugden2004economics} ($T>R>P>S$), Stag
+Hunt~\cite{skyrms2001stag} ($T>R>P>S$), Harmony~\cite{Licht1999} ($T>R>P>S$) and
+the Prisoner's Dilemma ($T>R>P>S$) games. The results based on these well know
+\(2 \times 2\) games are discussed in Section~\ref{section:2_by_2_games}.
 
-Moreover, we will specifically consider the donation game where each player can
-cooperate by providing a benefit \(b\) to the other player at their cost \(c\),
-with \(0 < c < b\). Then, \(T=b, R=b-c, S=-c, P=0\), and matrix~(\ref{eq:games})
-is give by:
+In Section~\ref{section:donation}, we will specifically consider the donation
+game where each player can cooperate by providing a benefit \(b\) to the other
+player at their cost \(c\), with \(0 < c < b\). Then, \(T=b, R=b-c, S=-c, P=0\),
+and matrix~(\ref{eq:game}) is give by:
 
@@
 
-These results are presented in Section~\ref{section:donation}. The relative
-influence of the game is controlled by an external parameter, the so called
-intensity of selection $\beta$.
+The relative influence of the payoffs on the adopt the strategy of the other is
+controlled by an external parameter, the so called intensity of selection
+$\beta$.
 
@@
 simple \textit{reactive strategies}. Reactive strategy are a set of memory-one
-strategies that only take into account the previous action of the opponent. An
-example of a reactive strategy is Tit For Tat. Reactive strategies can be
-written explicitly as a vector \(\in \R_{3}\). More specifically a reactive
-strategy \(s\) is given by \(s=(y, p, q)\) where \(y\) is the probability that
-the strategy opens with a cooperation and \(p, q\) are the probabilities that
-the strategy cooperates given that the opponent cooperated and defected
-equivalently. The individuals of our population do not interact for a given
-number of turns, on the contrast, we consider that each match has a probability
-$\delta$ to continue following each turn.
+strategies that only take into account the previous action of the opponent.
+Reactive strategies can be written explicitly as a vector \(\in \R_{3}\). More
+specifically a reactive strategy \(s\) is given by \(s=(y, p, q)\) where \(y\)
+is the probability that the strategy opens with a cooperation and \(p, q\) are
+the probabilities that the strategy cooperates given that the opponent
+cooperated and defected equivalently.
 
-\section{Expected and stochastic payoffs in the donation game}\label{section:donation}
+\section{Analysis of Stochastic Payoffs in the Donation game}\label{section:donation}
 
-Figure~\ref{fig:expected_and_stochastic_for_donation} shows simulation results
-for the above described process. Figure~\ref{fig:expected_and_stochastic_for_donation}~depicts the evolving conditional
-cooperation probabilities $p$ and $q$ (assuming that the discount
+We first explore the effect of stochastic payoffs on the cooperative behaviour
+by considering the extreme case that an individual's fitness is based on their
+last interaction with one other individual. We compare this to the expected
+payoffs. Figure~\ref{fig:expected_and_stochastic_for_donation} shows simulation
+results for the described process of Section~\ref{section:methods}.
+
+Figure~\ref{fig:expected_and_stochastic_for_donation}~depicts the evolving
+conditional cooperation probabilities $p$ and $q$ (assuming that the discount
 factor~$\delta$ and the benefit $b$ are comparably high). The left panel
 corresponds to the standard scenario considered in the previous literature. It
-considers players who use expected payoffs to update their
-strategies. The right panel shows the scenario considered herein, in which
-players update their strategies based on their last round's payoff. The figure
-suggests that when updating is based on expected payoffs, players tend to be
-more generous (their $q$-values are higher on average). In addition, players are
-generally more cooperative.
+considers players who use expected payoffs to update their strategies. The right
+panel shows the scenario considered herein, in which players update their
+strategies based on their last round's payoff. The figure suggests that when
+updating is based on expected payoffs, players tend to be more generous (their
+$q$-values are higher on average). In addition, players are generally more
+cooperative.
+
+There is a statistically significant different in the cooperation distribution of
+the population over time. The cooperation rate falls significantly yo under 50\%.
+There results are true for a strong selection strength ($\beta=10$). The follow
+up question that arises is what effect does the selection strength have on the
+cooperation rate?
 
 \begin{figure}[!htbp]
-    \includegraphics[width=\textwidth]{static/expected_and_stochastic_for_donation_game.pdf}
+    \centering
+    \includegraphics[width=\textwidth]{static/expected_and_stochastic_for_donation_game_b_10.pdf}
     \caption{{\bf Evolutionary dynamics under expected payoffs and stochastic payoffs.} 
@@
 
-The difference in the cooperation rates becomes more and more obvious as the
-temptation to defect increases, Figure~\ref{fig:expected_and_stochastic_for_donation_b_10}.
+Figure~\ref{fig:cooperation_rate_over_betas} illustrates the results for
+different runs of the evolutionary process where $\beta\in\{10 ^ -2, 10 ^ - 1, 10
+^ 0, 10 ^ 1, 10 ^ 2\}$. In the case where $\beta=10 ^ -2$ the process is almost
+random, as the effect of the payoffs is very small. The cooperation rate for
+this case and the case of $\beta=10 ^ -1$ between the stochastic and expected
+payoffs are the same. However, it is clear that this does not apply given 
+$\beta \geq 1$. The difference in the cooperation rate increases, in the case
+where for the stochastic payoffs it can drop to 10\%.
 
 \begin{figure}[!htbp]
-    \includegraphics[width=\textwidth]{static/expected_and_stochastic_for_donation_game_b_10.pdf}
-    \caption{{\bf Evolutionary dynamics under expected payoffs and stochastic payoffs.} 
-    Parameters:
-    $N\!=\!100$, $b\!=\!10$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.999$.}
-    \label{fig:expected_and_stochastic_for_donation_b_10}
-\end{figure}
-
-We explored how the evolving cooperation rates change as we vary the benefit $b$
-and the selection strength $\beta$
-(Figure~\ref{fig:cooration_under_parameters}). In all cases, we find that the
-two scenarios yield similar cooperation rates when the respective parameters are
-small. Once there is a high benefit to cooperation, strong selection, or a high
-expected number of rounds, updating based on expected payoffs yields higher
-cooperation rates.
-
-\begin{figure}
     \centering
@@
       \centering
-      \includegraphics[width=\textwidth]{static/cooperation_rate_over_b.pdf}
+      \includegraphics[width=\textwidth]{static/expected_for_beta.pdf}
+      \includegraphics[width=\textwidth]{static/stochastic_for_beta.pdf}
     \end{subfigure}%
     \begin{subfigure}{.5\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{static/cooperation_rate_over_b.pdf}
+    \end{subfigure}
+\caption{{\bf The evolution of cooperation for different parameter values.} 
+While the previous figures depict the evolutionary outcome for fixed parameter
+values, here we vary the benefit of cooperation $b$ and the strength of
+selection $\beta$. In all cases, stochastic payoff evaluation tends to reduce
+the evolving cooperation rates. Unless explicitly varied, the parameters of the
+simulation are $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$,
+$\delta\!=\!0.99$. Simulations are run for $T\!=\!5\times 10^6$ time steps for
+each parameter combination.}\label{fig:cooperation_rate_over_betas}
+\end{figure}
+
+Figure~\ref{fig:cooperation_rate_over_betas} illustrates the results for
+different runs of the evolutionary process where $\beta\in\{10 ^ -2, 10 ^ - 1, 10
+^ 0, 10 ^ 1, 10 ^ 2\}$. In the case where $\beta=10 ^ -2$ the process is almost
+random, as the effect of the payoffs is very small. The cooperation rate for
+this case and the case of $\beta=10 ^ -1$ between the stochastic and expected
+payoffs are the same. However, it is clear that this does not apply given 
+$\beta \geq 1$. The difference in the cooperation rate increases, in the case
+where for the stochastic payoffs it can drop to 10\%.
+
+\begin{figure}[!htbp]
+    \centering
+    \begin{subfigure}{.65\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{static/expected_for_selection_strenght.pdf}
+      \includegraphics[width=\textwidth]{static/stochastic_for_selection_strenght.pdf}
+    \end{subfigure}%
+    \begin{subfigure}{.35\textwidth}
       \centering
@@
 $\delta\!=\!0.99$. Simulations are run for $T\!=\!5\times 10^6$ time steps for
-each parameter combination.}\label{fig:cooration_under_parameters}
+each parameter combination.}
 \end{figure}
@@
 
-\section{Methods}\label{section:methods}
+\appendix
+
+\section{Model Setup}\label{section:methods}
 
@@
 
+\section{Expected Payoffs}
+
 We compare this process for what we defined as {\bf stochastic payoff
@@
 expected payoffs has been considered in~\cite{imhof2010stochastic}.
+
+\section{Stochastic Payoffs}
 

### 2021-04-14 clean_data.py
tweak clean script

@@ -8,17 +8,18 @@ import pandas as pd
 
 if __name__ == "__main__":  # pragma: no cover
     folder = sys.argv[1]
-    starting_characters = sys.argv[1]
+    starting_characters = sys.argv[2]
 
     in_columns = [
         "y",
         "p",
         "q",
         "average cooperation rate",
-        "average score",
+        # "average score",
         "t",
     ]
 
+    print(f"{folder}/{starting_characters}*csv")
     csv_files = glob.glob(f"{folder}/{starting_characters}*csv")
 
     for file in tqdm.tqdm(csv_files):
@@ -35,4 +36,4 @@ if __name__ == "__main__":  # pragma: no cover
 
         df["frequencies"] = frequencies
 
-        df.to_csv(f"{folder}/clean_{file[len(folder + 1):]}", index=False)
+        df.to_csv(f"{folder}/clean_{file[len(folder) + 1:]}", index=False)

### 2021-04-14 matlab/evolRunOpponentsRoundsTwo.m
tweak opponents - rounds experiment to run for each S

@@ -1,28 +1,20 @@
-function evolRunOpponentsRoundsTwo();
+function evolRunOpponentsRoundsTwo(S);
+
+distcomp.feature( 'LocalUseMpiexec', false)
+num_workers = 11;
+my_pool = parpool(num_workers);
 
 starting_resident = [0, 0, 0];
 N = 100;
 delta = 0.999;
-beta = 1;
-numberIterations= 10 ^ 8;
+beta = 10;
+numberIterations= 10 ^ 7;
 
-n = 11;
-Ss = linspace(-2, 2, n);
-Ts = linspace(-1, 3, n);
+n=11;
+Ts=linspace(-1, 3, n);
 
-payoffs = zeros(n * n, 2);
-
-for i=1:n
-    for j=1:n
-        payoffs(j + (i - 1) + (i -1) * 10, 1) = Ss(i);
-        payoffs(j + (i - 1) + (i -1) * 10, 2) = Ts(j);
-    end
-end
-
-lenght = n * n;
-parfor i = 1:lenght
-    S = payoffs(i, 1);
-    T = payoffs(i, 2);
+parfor i = 1:n
+    T=Ts(i);
     u = [1, S, T, 0];
     filename = "data/rounds_opponents_two/S_" + S + "_T_" + T;
     evolSimulationOpponentsRoundsTwo(starting_resident, u, N, delta, beta, numberIterations, filename);

### 2021-04-14 run_rounds.sh

@@ -9,3 +9,14 @@ sbatch sh/rounds_1_.sh
 # sbatch sh/rounds_8_.sh
 # sbatch sh/rounds_9_.sh
 # sbatch sh/rounds_10_.sh
+sbatch sh/opponents_rounds_0_.sh
+sbatch sh/opponents_rounds_1_.sh
+# sbatch sh/opponents_rounds_2_.sh
+# sbatch sh/opponents_rounds_3_.sh
+# sbatch sh/opponents_rounds_4_.sh
+# sbatch sh/opponents_rounds_5_.sh
+# sbatch sh/opponents_rounds_6_.sh
+# sbatch sh/opponents_rounds_7_.sh
+# sbatch sh/opponents_rounds_8_.sh
+# sbatch sh/opponents_rounds_9_.sh
+# sbatch sh/opponents_rounds_10_.sh

### 2021-04-14 sh/opponents_rounds_0_.sh

@@ -1,0 +1,13 @@
+#!/bin/sh
+#SBATCH -p medium  # partition (queue)
+#SBATCH -J S0 # job name
+#SBATCH -N 1 # number of nodes
+#SBATCH -n 11 # number of cores
+#SBATCH -t 2-00:00  # time (D-HH:MM)
+#SBATCH -o S0.out # STDOUT
+#SBATCH -e S0.err # STDERR
+
+module load matlab
+cd matlab
+
+matlab -nodisplay -r 'evolRunOpponentsRoundsTwo(-2.0);exit;'

### 2021-04-14 sh/opponents_rounds_10_.sh

@@ -1,0 +1,13 @@
+#!/bin/sh
+#SBATCH -p medium  # partition (queue)
+#SBATCH -J S10 # job name
+#SBATCH -N 1 # number of nodes
+#SBATCH -n 11 # number of cores
+#SBATCH -t 2-00:00  # time (D-HH:MM)
+#SBATCH -o S10.out # STDOUT
+#SBATCH -e S10.err # STDERR
+
+module load matlab
+cd matlab
+
+matlab -nodisplay -r 'evolRunOpponentsRoundsTwo(2.0);exit;'

### 2021-04-14 sh/opponents_rounds_1_.sh

@@ -1,0 +1,13 @@
+#!/bin/sh
+#SBATCH -p medium  # partition (queue)
+#SBATCH -J S1 # job name
+#SBATCH -N 1 # number of nodes
+#SBATCH -n 11 # number of cores
+#SBATCH -t 2-00:00  # time (D-HH:MM)
+#SBATCH -o S1.out # STDOUT
+#SBATCH -e S1.err # STDERR
+
+module load matlab
+cd matlab
+
+matlab -nodisplay -r 'evolRunOpponentsRoundsTwo(-1.6);exit;'

### 2021-04-14 sh/opponents_rounds_2_.sh

@@ -1,0 +1,13 @@
+#!/bin/sh
+#SBATCH -p medium  # partition (queue)
+#SBATCH -J S2 # job name
+#SBATCH -N 1 # number of nodes
+#SBATCH -n 11 # number of cores
+#SBATCH -t 2-00:00  # time (D-HH:MM)
+#SBATCH -o S2.out # STDOUT
+#SBATCH -e S2.err # STDERR
+
+module load matlab
+cd matlab
+
+matlab -nodisplay -r 'evolRunOpponentsRoundsTwo(-1.2);exit;'

### 2021-04-14 sh/opponents_rounds_3_.sh

@@ -1,0 +1,13 @@
+#!/bin/sh
+#SBATCH -p medium  # partition (queue)
+#SBATCH -J S3 # job name
+#SBATCH -N 1 # number of nodes
+#SBATCH -n 11 # number of cores
+#SBATCH -t 2-00:00  # time (D-HH:MM)
+#SBATCH -o S3.out # STDOUT
+#SBATCH -e S3.err # STDERR
+
+module load matlab
+cd matlab
+
+matlab -nodisplay -r 'evolRunOpponentsRoundsTwo(-0.8);exit;'

### 2021-04-14 sh/opponents_rounds_4_.sh

@@ -1,0 +1,13 @@
+#!/bin/sh
+#SBATCH -p medium  # partition (queue)
+#SBATCH -J S4 # job name
+#SBATCH -N 1 # number of nodes
+#SBATCH -n 11 # number of cores
+#SBATCH -t 2-00:00  # time (D-HH:MM)
+#SBATCH -o S4.out # STDOUT
+#SBATCH -e S4.err # STDERR
+
+module load matlab
+cd matlab
+
+matlab -nodisplay -r 'evolRunOpponentsRoundsTwo(-0.4);exit;'

### 2021-04-14 sh/opponents_rounds_5_.sh

@@ -1,0 +1,13 @@
+#!/bin/sh
+#SBATCH -p medium  # partition (queue)
+#SBATCH -J S5 # job name
+#SBATCH -N 1 # number of nodes
+#SBATCH -n 11 # number of cores
+#SBATCH -t 2-00:00  # time (D-HH:MM)
+#SBATCH -o S5.out # STDOUT
+#SBATCH -e S5.err # STDERR
+
+module load matlab
+cd matlab
+
+matlab -nodisplay -r 'evolRunOpponentsRoundsTwo(0.0);exit;'

### 2021-04-14 sh/opponents_rounds_6_.sh

@@ -1,0 +1,13 @@
+#!/bin/sh
+#SBATCH -p medium  # partition (queue)
+#SBATCH -J S6 # job name
+#SBATCH -N 1 # number of nodes
+#SBATCH -n 11 # number of cores
+#SBATCH -t 2-00:00  # time (D-HH:MM)
+#SBATCH -o S6.out # STDOUT
+#SBATCH -e S6.err # STDERR
+
+module load matlab
+cd matlab
+
+matlab -nodisplay -r 'evolRunOpponentsRoundsTwo(0.4);exit;'

### 2021-04-14 sh/opponents_rounds_7_.sh

@@ -1,0 +1,13 @@
+#!/bin/sh
+#SBATCH -p medium  # partition (queue)
+#SBATCH -J S7 # job name
+#SBATCH -N 1 # number of nodes
+#SBATCH -n 11 # number of cores
+#SBATCH -t 2-00:00  # time (D-HH:MM)
+#SBATCH -o S7.out # STDOUT
+#SBATCH -e S7.err # STDERR
+
+module load matlab
+cd matlab
+
+matlab -nodisplay -r 'evolRunOpponentsRoundsTwo(0.8);exit;'

### 2021-04-14 sh/opponents_rounds_8_.sh

@@ -1,0 +1,13 @@
+#!/bin/sh
+#SBATCH -p medium  # partition (queue)
+#SBATCH -J S8 # job name
+#SBATCH -N 1 # number of nodes
+#SBATCH -n 11 # number of cores
+#SBATCH -t 2-00:00  # time (D-HH:MM)
+#SBATCH -o S8.out # STDOUT
+#SBATCH -e S8.err # STDERR
+
+module load matlab
+cd matlab
+
+matlab -nodisplay -r 'evolRunOpponentsRoundsTwo(1.2);exit;'

### 2021-04-14 sh/opponents_rounds_9_.sh

@@ -1,0 +1,13 @@
+#!/bin/sh
+#SBATCH -p medium  # partition (queue)
+#SBATCH -J S9 # job name
+#SBATCH -N 1 # number of nodes
+#SBATCH -n 11 # number of cores
+#SBATCH -t 2-00:00  # time (D-HH:MM)
+#SBATCH -o S9.out # STDOUT
+#SBATCH -e S9.err # STDERR
+
+module load matlab
+cd matlab
+
+matlab -nodisplay -r 'evolRunOpponentsRoundsTwo(1.6);exit;'

### 2021-04-22 main.tex
about to move things around in introduction

@@
 \titleformat{\subsection}{\sffamily \fontsize{11.5}{11.5}\bfseries}{\thesubsection}{1em}{}
+
+\usepackage{tikz}
+\usetikzlibrary{arrows}
+
+\tikzset{
+  treenode/.style = {align=center, inner sep=0pt, text centered,
+    font=\sffamily},
+  arn_n/.style = {treenode, circle, white, font=\sffamily\bfseries, draw=black, inner sep=-6pt,
+    fill=black, text width=1.5em},% arbre rouge noir, noeud noir
+  arn_r/.style = {treenode, circle, red, fill=red, 
+    text width=1.5em, very thick, inner sep=-6pt},% arbre rouge noir, noeud rouge
+  arn_x/.style = {treenode, rectangle, draw=black,
+    minimum width=0.5em, minimum height=0.5em}% arbre rouge noir, nil
+}
 
@@
 
+\begin{abstract}
+
+\end{abstract}
+
 \section{Introduction}
@@
 biological, social and economical systems. %ToDo references
-In evolutionary models, the success of an individual is not constant but it
-depends on the composition of the population. Individuals interact according to
-the rules of a pre defined game and they payoffs they achieved are translated
-into fitness.
+Compared to classical game theory where it is assumed that players are rational,
+evolutionary game theory does not require individuals to be rational. Instead
+individuals adapt their strategies based on exploration and mutation. In this
+context, strategies spread based on fitness either because the individuals who
+adopt them have more offspring (fitness-based processes), or they are imitated
+more often (pairwise comparison processes)~\cite{Wu2015}. Fitness is not
+constant but it depends on the composition of the population. Individuals
+interact with members of the population according to pre-define games, and their
+payoffs are then translated into fitness.
 
-Evolutionary models often work with idealized assumptions when estimating the
-fitness of a individual. It is commonly assumed that an individual will interact
-several times and with all other players in the population. Equivalently their
-fitness is given by the mean payoff they achieved over the different types in
-the population after multiple intercalations. These payoff is referred to as the
-\textit{expected} payoff.
+Evolutionary models often work with idealized assumptions when estimating
+fitness. It is commonly assumed that fitness is given by the mean payoff an
+individual achieved over the different types in the population following
+multiple intercalations. Referred to as the \textit{expected payoffs}. Expected
+payoffs imply that individuals have a perfect a memory, as they are able to
+recall several of their interactions with each player in the population. However
+most evolutionary models, with a few notable exceptions~\cite{Hauert1997,
+Stewart2016}, focus on naive subjects who can only choose from a restricted set
+of strategies~\cite{Nowak1992tit}, or who do not remember anything beyond the
+outcome of the very last round~\cite{Baek2016} when modeling how individuals
+make decisions in each round. This creates a curious inconsistency. Individuals
+make decisions in each round whist having very limited memory, however, when
+modeling how individuals update their strategies over time, individuals are
+assumed to have perfect memory.
 
-Initially this assumption implies that selection occurs much more slowly than
-the interaction between individuals. However, experimental studies show that
-this may not always be the case in biology %ToDo references.
-In~\cite{Roca2006} this assumption is realized and it shows that rapid selection
-affects evolutionary dynamics in such a dramatic way that for some games it even
-changes the stability of equilibria.
+One of the most important applications of evolutionary game theory is the
+evolution of cooperation. Though cooperation can been seen to be at odd, several
+mechanisms can allow for cooperation to emerge. One of these mechanisms is
+direct reciprocity. However, given these idealized assumptions it feels like our
+understanding of reciprocity is thus still incomplete. In this work we will
+explore how robust our understanding of cooperation is as models deviate from
+the perfect memory assumption.
 
-Moreover, estimating expected payoffs assume that individuals have a perfect a
-memory. Their recall several of their interactions with each player in the
-population. This creates a curious inconsistency. With a only few notable
-exceptions~\cite{Hauert1997, Stewart2016}, evolutionary models focus on naive
-subjects who can only choose from a restricted set of
-strategies~\cite{Nowak1992tit}, or who do not remember anything beyond the
-outcome of the very last round~\cite{Baek2016} when modeling how individuals
-make decisions in each round. However, when modeling how individuals update
-their strategies over time, individuals are assumed to have perfect memory.
+This work is not the first to relax assumption made at the fitness level.
+In~\cite{Roca2006} relax the assumption that selection occurs much more slowly
+than the interaction between individuals. They results shows that rapid
+selection affects evolutionary dynamics in such a dramatic way that for some
+games it even changes the stability of equilibria.
 
-We will explore how robust our understanding of cooperation is as models deviate
-from the perfect memory assumption. We start by considering two extreme
-scenarios. The first is the classical scenario of the expected payoffs and the
-alternative scenario where individuals update their strategies only based on the
-very last payoff they obtained. For our second approach we allow individuals to
-use more memory. We refer to these as the \textit{stochastic} payoffs of
-individuals.
+We start by considering two extreme scenarios. The first is the classical
+scenario of the expected payoffs and the alternative scenario where individuals
+update their strategies only based on the very last payoff they obtained. We
+refer to these as the \textit{stochastic} payoffs of individuals. In
+Section~\ref{section:donation}, we explore the effect of the payoffs in the
+cooperation rate given the donation game. We explore the difference between the
+ratios as well as the effect of different model parameters. In
+Section~\ref{section:2_by_2_games} we explore the cooperation rate for all the
+important case of symmetric \(2 \times 2\) games. For the later part of the
+section we also allow individuals to use more memory.
 
-Early results for the repeated prisoner’s dilemma and for the snowdrift game
-suggest that memory size can have a considerable effect on the evolutionary
-dynamics. As a rule of thumb, we observe that individuals with limited memory
-tend to adopt less generous strategies and they achieve less cooperation than in
-the classical scenario. In contrast, in the stag-hunt and the harmony game, the
-impact of memory is less striking. Here individuals with limited memory perform
-nearly as well as individuals with full memory.
+% Early results for the repeated prisoner’s dilemma and for the snowdrift game
+% suggest that memory size can have a considerable effect on the evolutionary
+% dynamics. As a rule of thumb, we observe that individuals with limited memory
+% tend to adopt less generous strategies and they achieve less cooperation than in
+% the classical scenario. In contrast, in the stag-hunt and the harmony game, the
+% impact of memory is less striking. Here individuals with limited memory perform
+% nearly as well as individuals with full memory.
 
@@
 
-Evolutionary game dynamics have extensively been studied using stochastic
-individual-based models that make use of different rules to determine how
-successful strategies spread. There are two classes have been used extensively
-based on~\cite{Wu2015} (i) \textit{fitness-based processes} in which an
-individual chosen proportional to fitness reproduces and the offspring replaces
-a randomly chosen individual (ii) \textit{pairwise comparison} processes in
-which a pair of individuals is chosen, and where subsequently one of these
-individuals may adopt the strategy of the other. This adoption occurs with a
-probability that depends on the fitness of both individuals, such that better
-players are more likely to be imitated than those who do worse.
+In order to account for the difference in the robustness of cooperation among
+individuals based on their payoff memory, we consider a population of \(N\)
+players, where \(N\) is even, and where mutations are sufficiently rare. At any
+point in time the there are at most two different strategies are present in the
+population. A \textit{resident} strategy and a \textit{mutant} strategy. Suppose
+there are \(N - k\) players who use the resident strategy whereas \(k\) players
+use the mutant strategy. Each step of the evolutionary process consists of two
+stages, a game stage and an updating stage.
 
-We us a pairwise comparison processes with a population of $N$ individuals. In
-our model mutation happens rarely and such each individual can be of one of the
-two types, either of the \textit{resident} or either of the \textit{mutant}.
+1. In the game stage, each player is randomly matched with some other player in
+the population to interact in a number of instances of the game
+\(\begin{pmatrix} R & S  \\ T & P \end{pmatrix}\). The number of interactions is
+not fixed, on the contrast we consider that another interaction has a
+probability $\delta$ to continue following each turn. We assume herein that
+individuals at most make use of simple \textit{reactive strategies} make
+decisions in each round. Reactive strategy are a set of memory-one strategies
+that only take into account the previous action of the opponent. Reactive
+strategies can be written explicitly as a vector \(\in \R_{3}\). More
+specifically a reactive strategy \(s\) is given by \(s=(y, p, q)\) where \(y\)
+is the probability that the strategy opens with a cooperation and \(p, q\) are
+the probabilities that the strategy cooperates given that the opponent
+cooperated and defected equivalently.
 
-The fitness of the individuals will be based on the payoffs they receive from
-their interactions. he individuals of our population do not interact for a given
-number of turns, on the contrast, we consider that each match has a probability
-$\delta$ to continue following each turn. In each turn individuals can decide whether
-to cooperate (C) or to defect (D). The
-payoffs in a given round depend on the player's and co players decisions. We
-consider all the important case of symmetric \(2 \times 2\) games, in which the
-payoffs are given by the following matrix:
+2. In the updating stage, two players are randomly drawn from the population, a
+`learner' and a `role model'. Given that the learner's payoff $u_L\!\in\!
+\mathcal{U}$ and that the role model's payoff $u_{RL}\!\in\! \mathcal{U}$, we
+assume the learner adopts the role model's strategy with probability
 
-\begin{equation}\label{eq:game}
-    \begin{pmatrix}
-        R & S  \\
-        T & P
-    \end{pmatrix}
+\begin{equation} \label{Eq:rho}
+\rho(u_L, u_{RL}) = \frac{1}{1\!+\!\exp\big[ \!-\!\beta (u_E\!-\!u_{RL}) \big]}.
 \end{equation}
 
-These cases include the Snowdrift~\cite{sugden2004economics} ($T>R>P>S$), Stag
-Hunt~\cite{skyrms2001stag} ($T>R>P>S$), Harmony~\cite{Licht1999} ($T>R>P>S$) and
-the Prisoner's Dilemma ($T>R>P>S$) games. The results based on these well know
-\(2 \times 2\) games are discussed in Section~\ref{section:2_by_2_games}.
+where $\beta\!\ge\!0$ corresponds  to relative influence of the payoffs on the
+adopt the strategy of the other is controlled by an external parameter, the so
+called intensity of selection.
+
+We iterate this basic evolutionary step until either the mutant strategy goes
+extinct, or until it fixes in the population (in which case the mutant strategy
+becomes the new resident strategy). After either outcome, we introduce a new
+mutant strategy $S'_2\!=\!(y_2',p_2',q_2')$ (uniformly chosen from all reactive
+strategies at random), and we set the number of mutants to $k\!=\!1$. This
+process of mutation and fixation/extinction is then iterated many times.xw The
+fixation probability of the mutant strategy then takes the standard form
+\cite{nowak2004emergence},
+
+\begin{equation}
+\varphi = \frac{1}{1+\sum_{i=1}^{N-1}\prod_k^i \frac{\lambda^-_k}{\lambda^+_k}}.
+\end{equation}
+
+where \(\lambda^-_k, \lambda^+_k\) are the probabilities that the number of
+mutants decreases and increases respectively.
+
+We compare this process for stochastic payoff evaluation with the analogous
+process where players update their strategies with respect to their {\it
+expected} payoffs,
+
+\begin{equation} \label{Eq:ExpPay}
+  \begin{array}{lcrcr}
+  \displaystyle \pi_1	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(S_1,S_1),\mathbf{u}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(S_1,S_2),\mathbf{u}\rangle,\\[0.5cm]
+  \displaystyle \pi_2	&=	&\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(S_2,S_1),\mathbf{u}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(S_2,S_2),\mathbf{u}\rangle.\\
+  \end{array}
+\end{equation}
+
+In the limit of no discounting, $\delta\!\rightarrow\! 1$, this process based on
+expected payoffs has been considered in~\cite{imhof2010stochastic}.
+
+% \subsection{Fixation probabilities under stochastic payoff evaluation}
+% Given that $N\!-\!k$ players use the resident strategy $S_1\!=\!(y_1,p_1,q_1)$ and that the remaining $k$ players use the mutant strategy $S_2\!=\!(y_2,p_2,q_2)$, the probability that the number of mutants increases by one in one step of the evolutionary process can be written as
+% \begin{equation}
+% \lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_1,u_2\in\mathcal{U}} x(u_1,u_2)\cdot \rho(u_1,u_2).
+% \end{equation}
+% In this expression, $(N\!-\!k)/N$ is the probability that the randomly chosen learner is a resident, and $k/N$ is the probability that the role model is a mutant.
+The sum corresponds to the total probability that the learner adopts the role model's strategy over all possible payoffs $u_1$ and $u_2$ that the two player may have received in their respective last rounds.
+We use $x(u_1,u_2)$ to denote the probability that the randomly chosen resident obtained a payoff of $u_1$ in the last round of his respective game, and that the mutant obtained a payoff of $u_2$.
+Given that the payoffs are $u_1$ and $u_2$, the imitation probability is then given by $\rho(u_1,u_2)$, as specified by Eq.~(\ref{Eq:rho}).
+The probability that the respective payoffs of the players are given by $u_1$ and $u_2$ can be calculated as
+% \begin{equation}
+% \setlength{\arraycolsep}{1pt}
+% \begin{array}{llrl}
+% x(u_1,u_2)	 &=&\displaystyle \frac{1}{N\!-\!1}\cdot  &v_{u_1}(S_1,S_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F}\\[0.5cm]
+% &+	
+% &\displaystyle \left(1\!-\!\frac{1}{N\!-\!1}\right)  
+% &\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} v_{u_1}(S_1,S_2) v_{u_2}(S_2,S_2) + 
+%  \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} v_{u_1}(S_1,S_2) v_{u_2}(S_2,S_1)\right.\\[0.5cm]
+% &&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} v_{u_1}(S_1,S_1) v_{u_2}(S_2,S_2) + 
+%  \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} v_{u_1}(S_1,S_1) v_{u_2}(S_2,S_1)\right].
+% \end{array}
+% \end{equation}
+%  The first term on the right side corresponds to the case that the learner and
+%  the role model happened to be matched during the game stage, which happens
+%  with probability $1/(N\!-\!1)$. In that case, we note that only those payoff
+%  pairs can occur that are feasible in a direct interaction, $(u_1,u_2)\in
+%  \mathcal{U}^2_F:=\big\{ (R,R), (S,T), (T,S), (P,P) \big\}$, as represented by
+%  the respective indicator function. Otherwise, if the learner and the role
+%  model did not interact directly, we need to distinguish four different cases,
+%  depending on whether the learner was matched with a resident or a mutant, and
+%  depending on whether the role model was matched with a resident or a mutant.
+
+% Analogously, we can calculate the probability that the number of mutants decreases by one in one step of the evolutionary process. This probability is
+% \begin{equation}
+% \lambda^-_k=\frac{N\!-\!k}{N}\cdot\frac{k}{N} \sum_{u_1,u_2\in\mathcal{U}} x(u_1,u_2)\cdot \rho(u_2,u_1).
+% \end{equation}
+% The fixation probability of the mutant strategy then takes the standard form \cite{nowak2004emergence},
+% \begin{equation}
+% \varphi = \frac{1}{1+\sum_{i=1}^{N-1}\prod_k^i \frac{\lambda^-_k}{\lambda^+_k}}.
+% \end{equation}
+
+% \begin{equation}\label{eq:game}
+%     \begin{pmatrix}
+%         R & S  \\
+%         T & P
+%     \end{pmatrix}
+% \end{equation}
+
+% These cases include the Snowdrift~\cite{sugden2004economics} ($T>R>P>S$), Stag
+% Hunt~\cite{skyrms2001stag} ($T>R>P>S$), Harmony~\cite{Licht1999} ($T>R>P>S$) and
+% the Prisoner's Dilemma ($T>R>P>S$) games. The results based on these well know
+% \(2 \times 2\) games are discussed in Section~\ref{section:2_by_2_games}.
+
+\section{Analysis of Stochastic Payoffs in the Donation game}\label{section:donation}
+
+We first explore the effect of stochastic payoffs on the cooperative behaviour
+by considering the extreme case that an individual's fitness is based on their
+last interaction with one other individual. We compare this to the expected
+payoffs. Figure~\ref{fig:expected_and_stochastic_for_donation} shows simulation
+results for the described process of Section~\ref{section:methods}.
 
@@
 \end{equation}
-
-The relative influence of the payoffs on the adopt the strategy of the other is
-controlled by an external parameter, the so called intensity of selection
-$\beta$.
-
-The space of strategies in repeated is infinite. To continue with our
-evolutionary analysis, we assume herein that individuals at most make use of
-simple \textit{reactive strategies}. Reactive strategy are a set of memory-one
-strategies that only take into account the previous action of the opponent.
-Reactive strategies can be written explicitly as a vector \(\in \R_{3}\). More
-specifically a reactive strategy \(s\) is given by \(s=(y, p, q)\) where \(y\)
-is the probability that the strategy opens with a cooperation and \(p, q\) are
-the probabilities that the strategy cooperates given that the opponent
-cooperated and defected equivalently.
-
-\section{Analysis of Stochastic Payoffs in the Donation game}\label{section:donation}
-
-We first explore the effect of stochastic payoffs on the cooperative behaviour
-by considering the extreme case that an individual's fitness is based on their
-last interaction with one other individual. We compare this to the expected
-payoffs. Figure~\ref{fig:expected_and_stochastic_for_donation} shows simulation
-results for the described process of Section~\ref{section:methods}.
 
@@
 
-It can shown that, \(!(1\!-\!\delta)\mathbf{v_0}(I_4-\delta M)^{-1}\) and with
+It can shown that, \((1\!-\!\delta)\mathbf{v_0}(I_4-\delta M)^{-1}\) and with
 some algebraic manipulation we derive to Equation~\ref{Eq:LastRound}.
@@
 \end{proof}
+
+Equation~\ref{Eq:LastRound} is the probability vector that players \(S_1, S_2\)
+are in each of the possible states of the last round. Given the population \(N\)
+with \(k\) mutants in the last round there only possible combinations of interactions
+are:
+
+\begin{figure}
+    \centering
+    \includegraphics[width=.65\textwidth]{static/last_rounds_matches.tex}
+\end{figure}
+
+We compare this process for stochastic payoff evaluation with the analogous process where players update their strategies with respect to their {\it expected} payoffs,
+\begin{equation} \label{Eq:ExpPay}
+\begin{array}{lcrcr}
+\displaystyle \pi_1	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(S_1,S_1),\mathbf{u}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(S_1,S_2),\mathbf{u}\rangle,\\[0.5cm]
+\displaystyle \pi_2	&=	&\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(S_2,S_1),\mathbf{u}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(S_2,S_2),\mathbf{u}\rangle.\\
+\end{array}
+\end{equation}
+In the limit of no discounting, $\delta\!\rightarrow\! 1$, this process based on expected payoffs has been considered in~\cite{imhof2010stochastic}.
+
+\subsection{Fixation probabilities under stochastic payoff evaluation}
+Given that $N\!-\!k$ players use the resident strategy $S_1\!=\!(y_1,p_1,q_1)$ and that the remaining $k$ players use the mutant strategy $S_2\!=\!(y_2,p_2,q_2)$, the probability that the number of mutants increases by one in one step of the evolutionary process can be written as
+\begin{equation}
+\lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_1,u_2\in\mathcal{U}} x(u_1,u_2)\cdot \rho(u_1,u_2).
+\end{equation}
+In this expression, $(N\!-\!k)/N$ is the probability that the randomly chosen learner is a resident, and $k/N$ is the probability that the role model is a mutant.
+The sum corresponds to the total probability that the learner adopts the role model's strategy over all possible payoffs $u_1$ and $u_2$ that the two player may have received in their respective last rounds.
+We use $x(u_1,u_2)$ to denote the probability that the randomly chosen resident obtained a payoff of $u_1$ in the last round of his respective game, and that the mutant obtained a payoff of $u_2$.
+Given that the payoffs are $u_1$ and $u_2$, the imitation probability is then given by $\rho(u_1,u_2)$, as specified by Eq.~(\ref{Eq:rho}).
+The probability that the respective payoffs of the players are given by $u_1$ and $u_2$ can be calculated as
+\begin{equation}
+\setlength{\arraycolsep}{1pt}
+\begin{array}{llrl}
+x(u_1,u_2)	 &=&\displaystyle \frac{1}{N\!-\!1}\cdot  &v_{u_1}(S_1,S_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F}\\[0.5cm]
+&+	
+&\displaystyle \left(1\!-\!\frac{1}{N\!-\!1}\right)  
+&\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} v_{u_1}(S_1,S_2) v_{u_2}(S_2,S_2) + 
+ \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} v_{u_1}(S_1,S_2) v_{u_2}(S_2,S_1)\right.\\[0.5cm]
+&&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} v_{u_1}(S_1,S_1) v_{u_2}(S_2,S_2) + 
+ \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} v_{u_1}(S_1,S_1) v_{u_2}(S_2,S_1)\right].
+\end{array}
+\end{equation}
+The first term on the right side corresponds to the case that the learner and the role model happened to be matched during the game stage, which happens with probability $1/(N\!-\!1)$.
+In that case, we note that only those payoff pairs can occur that are feasible in a direct interaction, $(u_1,u_2)\in \mathcal{U}^2_F:=\big\{ (R,R), (S,T), (T,S), (P,P) \big\}$, as represented by the respective indicator function.
+Otherwise, if the learner and the role model did not interact directly, we need to distinguish four different cases, depending on whether the learner was matched with a resident or a mutant, and depending on whether the role model was matched with a resident or a mutant.
 

### 2021-04-22 main.tex
draft written parts

@@
 
-Evolutionary game theory is a mathematical framework for modeling evolution in
-biological, social and economical systems. %ToDo references
-Compared to classical game theory where it is assumed that players are rational,
-evolutionary game theory does not require individuals to be rational. Instead
-individuals adapt their strategies based on exploration and mutation. In this
-context, strategies spread based on fitness either because the individuals who
+One of the most important applications of evolutionary game theory is the
+evolution of cooperation. Why is it that some individuals choose to help others
+(increasing their payoff) at the expense of decreasing one's own payoff? During
+the past decade, the literature on mechanisms that allow for cooperation, even
+though it's seen to be at odd has been fruitful. One such mechanism is
+repetition; the so called direct reciprocity. Theoretical researchers have been
+using evolutionary models to understand how direct reciprocity allows
+cooperation to evolve and which strategies are important for sustaining
+cooperation.
+
+Evolutionary game theory does not require individuals to be rational, instead
+they adapt strategies based on mutation and exploration. Strategies with
+high fitness are more likely to spread, either because the individuals who
 adopt them have more offspring (fitness-based processes), or they are imitated
-more often (pairwise comparison processes)~\cite{Wu2015}. Fitness is not
-constant but it depends on the composition of the population. Individuals
-interact with members of the population according to pre-define games, and their
-payoffs are then translated into fitness.
+more often (pairwise comparison processes)~\cite{Wu2015}. The fitness of strategies,
+and subsequently of the individuals, is not constant. Instead it depends on the
+composition of the population. Individuals interact with other members of the
+population according to their strategies. Their yielded payoffs, according to
+pre defined games, are translated into fitness.
 
-Evolutionary models often work with idealized assumptions when estimating
-fitness. It is commonly assumed that fitness is given by the mean payoff an
-individual achieved over the different types in the population following
-multiple intercalations. Referred to as the \textit{expected payoffs}. Expected
-payoffs imply that individuals have a perfect a memory, as they are able to
-recall several of their interactions with each player in the population. However
-most evolutionary models, with a few notable exceptions~\cite{Hauert1997,
-Stewart2016}, focus on naive subjects who can only choose from a restricted set
-of strategies~\cite{Nowak1992tit}, or who do not remember anything beyond the
-outcome of the very last round~\cite{Baek2016} when modeling how individuals
-make decisions in each round. This creates a curious inconsistency. Individuals
-make decisions in each round whist having very limited memory, however, when
-modeling how individuals update their strategies over time, individuals are
-assumed to have perfect memory.
+It is commonly assumed that fitness assimilates to the \textit{expected payoff},
+which the mean payoff an individual achieved over the different types in the
+population after multiple intercalations. Expected payoffs imply that
+individuals have a perfect a memory. In order to estimate expected payoffs
+individuals must be able to recall several of their interactions with each
+player in the population. However, when modeling how they make decisions in each
+round they assumed to have very limited memory. To be precise, most work on this
+models focuses on naive subjects who can only choose from a restricted set of
+strategies~\cite{Nowak1992tit}, or who do not remember anything beyond the
+outcome of the very last round~\cite{Baek2016}, with a few notable
+exceptions~\cite{Hauert1997, Stewart2016}.
 
-One of the most important applications of evolutionary game theory is the
-evolution of cooperation. Though cooperation can been seen to be at odd, several
-mechanisms can allow for cooperation to emerge. One of these mechanisms is
-direct reciprocity. However, given these idealized assumptions it feels like our
-understanding of reciprocity is thus still incomplete. In this work we will
-explore how robust our understanding of cooperation is as models deviate from
-the perfect memory assumption.
+This creates a curious inconsistency, which raises the following question: how
+robust is our understanding of cooperation? We are not the first to question the
+assumptions of models when estimating fitness. In~\cite{Roca2006} relax the
+assumption that selection occurs much more slowly than the interaction between
+individuals. They results shows that rapid selection affects evolutionary
+dynamics in such a dramatic way that for some games it even changes the
+stability of equilibria.
 
-This work is not the first to relax assumption made at the fitness level.
-In~\cite{Roca2006} relax the assumption that selection occurs much more slowly
-than the interaction between individuals. They results shows that rapid
-selection affects evolutionary dynamics in such a dramatic way that for some
-games it even changes the stability of equilibria.
+We consider two cases. Initially, we start by considering two extreme scenarios.
+The first is the classical scenario of the expected payoffs and the alternative
+scenario where individuals update their strategies only based on the very last
+payoff they obtained. We refer to these as the \textit{stochastic} payoffs of
+individuals. We do this for the donation games but also the important cases of
+symmetric \(2 \times 2\) games. These include the prisoner's dilemma, the
+snowdrift game, the stag hunt game and the harmony game.
 
-We start by considering two extreme scenarios. The first is the classical
-scenario of the expected payoffs and the alternative scenario where individuals
-update their strategies only based on the very last payoff they obtained. We
-refer to these as the \textit{stochastic} payoffs of individuals. In
-Section~\ref{section:donation}, we explore the effect of the payoffs in the
-cooperation rate given the donation game. We explore the difference between the
-ratios as well as the effect of different model parameters. In
-Section~\ref{section:2_by_2_games} we explore the cooperation rate for all the
-important case of symmetric \(2 \times 2\) games. For the later part of the
-section we also allow individuals to use more memory.
+In the later sections we allow individuals to use more memory. More specifically,
+individuals update their strategies by considering up to the last two rounds,
+whilst interacting with up to two different players.
 
@@
 
-\section{Model Setup}
+\section{Model Setup}\label{section:model}
 
@@
 
-1. In the game stage, each player is randomly matched with some other player in
-the population to interact in a number of instances of the game
-\(\begin{pmatrix} R & S  \\ T & P \end{pmatrix}\). The number of interactions is
+In the game stage, each player is randomly matched with some other player in
+the population to interact in a number of turns.  The number of turns is
 not fixed, on the contrast we consider that another interaction has a
-probability $\delta$ to continue following each turn. We assume herein that
+probability $\delta$ to continue following each turn. At each turn the players can
+choose to either cooperate (C) or to defect (D). The payoffs in a given round
+depend on both player's decisions and are given by:
+
+\begin{equation*}\label{eq:game}
+  U = \begin{pmatrix} R & S  \\ T & P \end{pmatrix}
+\end{equation*}
+
+We assume herein that
 individuals at most make use of simple \textit{reactive strategies} make
@@
 
-2. In the updating stage, two players are randomly drawn from the population, a
+In the updating stage, two players are randomly drawn from the population, a
 `learner' and a `role model'. Given that the learner's payoff $u_L\!\in\!
 \mathcal{U}$ and that the role model's payoff $u_{RL}\!\in\! \mathcal{U}$, we
-assume the learner adopts the role model's strategy with probability
+assume the learner adopts the role model's strategy based on the Fermi
+distribution function. The relative influence of the payoffs on the
+adopt the strategy of the other is controlled by an external parameter. The so
+called intensity of selection,  $\beta\!\ge\!0$.
 
-\begin{equation} \label{Eq:rho}
-\rho(u_L, u_{RL}) = \frac{1}{1\!+\!\exp\big[ \!-\!\beta (u_E\!-\!u_{RL}) \big]}.
-\end{equation}
+This basic evolutionary step is repeated until either the mutant strategy goes
+extinct, or until it fixes in the population, in which case the mutant strategy
+becomes the new resident strategy. After either outcome, we introduce a new
+mutant strategy, uniformly chosen from all reactive strategies at random, and we
+set the number of mutants to $k\!=\!1$. This process of mutation and
+fixation/extinction is then iterated many times.
 
-where $\beta\!\ge\!0$ corresponds  to relative influence of the payoffs on the
-adopt the strategy of the other is controlled by an external parameter, the so
-called intensity of selection.
-
-We iterate this basic evolutionary step until either the mutant strategy goes
-extinct, or until it fixes in the population (in which case the mutant strategy
-becomes the new resident strategy). After either outcome, we introduce a new
-mutant strategy $S'_2\!=\!(y_2',p_2',q_2')$ (uniformly chosen from all reactive
-strategies at random), and we set the number of mutants to $k\!=\!1$. This
-process of mutation and fixation/extinction is then iterated many times.xw The
-fixation probability of the mutant strategy then takes the standard form
-\cite{nowak2004emergence},
-
-\begin{equation}
-\varphi = \frac{1}{1+\sum_{i=1}^{N-1}\prod_k^i \frac{\lambda^-_k}{\lambda^+_k}}.
-\end{equation}
-
-where \(\lambda^-_k, \lambda^+_k\) are the probabilities that the number of
-mutants decreases and increases respectively.
-
-We compare this process for stochastic payoff evaluation with the analogous
-process where players update their strategies with respect to their {\it
-expected} payoffs,
-
-\begin{equation} \label{Eq:ExpPay}
-  \begin{array}{lcrcr}
-  \displaystyle \pi_1	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(S_1,S_1),\mathbf{u}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(S_1,S_2),\mathbf{u}\rangle,\\[0.5cm]
-  \displaystyle \pi_2	&=	&\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(S_2,S_1),\mathbf{u}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(S_2,S_2),\mathbf{u}\rangle.\\
-  \end{array}
-\end{equation}
-
-In the limit of no discounting, $\delta\!\rightarrow\! 1$, this process based on
-expected payoffs has been considered in~\cite{imhof2010stochastic}.
-
-% \subsection{Fixation probabilities under stochastic payoff evaluation}
-% Given that $N\!-\!k$ players use the resident strategy $S_1\!=\!(y_1,p_1,q_1)$ and that the remaining $k$ players use the mutant strategy $S_2\!=\!(y_2,p_2,q_2)$, the probability that the number of mutants increases by one in one step of the evolutionary process can be written as
-% \begin{equation}
-% \lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_1,u_2\in\mathcal{U}} x(u_1,u_2)\cdot \rho(u_1,u_2).
-% \end{equation}
-% In this expression, $(N\!-\!k)/N$ is the probability that the randomly chosen learner is a resident, and $k/N$ is the probability that the role model is a mutant.
-The sum corresponds to the total probability that the learner adopts the role model's strategy over all possible payoffs $u_1$ and $u_2$ that the two player may have received in their respective last rounds.
-We use $x(u_1,u_2)$ to denote the probability that the randomly chosen resident obtained a payoff of $u_1$ in the last round of his respective game, and that the mutant obtained a payoff of $u_2$.
-Given that the payoffs are $u_1$ and $u_2$, the imitation probability is then given by $\rho(u_1,u_2)$, as specified by Eq.~(\ref{Eq:rho}).
-The probability that the respective payoffs of the players are given by $u_1$ and $u_2$ can be calculated as
-% \begin{equation}
-% \setlength{\arraycolsep}{1pt}
-% \begin{array}{llrl}
-% x(u_1,u_2)	 &=&\displaystyle \frac{1}{N\!-\!1}\cdot  &v_{u_1}(S_1,S_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F}\\[0.5cm]
-% &+	
-% &\displaystyle \left(1\!-\!\frac{1}{N\!-\!1}\right)  
-% &\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} v_{u_1}(S_1,S_2) v_{u_2}(S_2,S_2) + 
-%  \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} v_{u_1}(S_1,S_2) v_{u_2}(S_2,S_1)\right.\\[0.5cm]
-% &&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} v_{u_1}(S_1,S_1) v_{u_2}(S_2,S_2) + 
-%  \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} v_{u_1}(S_1,S_1) v_{u_2}(S_2,S_1)\right].
-% \end{array}
-% \end{equation}
-%  The first term on the right side corresponds to the case that the learner and
-%  the role model happened to be matched during the game stage, which happens
-%  with probability $1/(N\!-\!1)$. In that case, we note that only those payoff
-%  pairs can occur that are feasible in a direct interaction, $(u_1,u_2)\in
-%  \mathcal{U}^2_F:=\big\{ (R,R), (S,T), (T,S), (P,P) \big\}$, as represented by
-%  the respective indicator function. Otherwise, if the learner and the role
-%  model did not interact directly, we need to distinguish four different cases,
-%  depending on whether the learner was matched with a resident or a mutant, and
-%  depending on whether the role model was matched with a resident or a mutant.
-
-% Analogously, we can calculate the probability that the number of mutants decreases by one in one step of the evolutionary process. This probability is
-% \begin{equation}
-% \lambda^-_k=\frac{N\!-\!k}{N}\cdot\frac{k}{N} \sum_{u_1,u_2\in\mathcal{U}} x(u_1,u_2)\cdot \rho(u_2,u_1).
-% \end{equation}
-% The fixation probability of the mutant strategy then takes the standard form \cite{nowak2004emergence},
-% \begin{equation}
-% \varphi = \frac{1}{1+\sum_{i=1}^{N-1}\prod_k^i \frac{\lambda^-_k}{\lambda^+_k}}.
-% \end{equation}
-
-% \begin{equation}\label{eq:game}
-%     \begin{pmatrix}
-%         R & S  \\
-%         T & P
-%     \end{pmatrix}
-% \end{equation}
-
-% These cases include the Snowdrift~\cite{sugden2004economics} ($T>R>P>S$), Stag
-% Hunt~\cite{skyrms2001stag} ($T>R>P>S$), Harmony~\cite{Licht1999} ($T>R>P>S$) and
-% the Prisoner's Dilemma ($T>R>P>S$) games. The results based on these well know
-% \(2 \times 2\) games are discussed in Section~\ref{section:2_by_2_games}.
+In this work we explore the effect on the payoffs in the updating stage. The
+details of the process described in this section can be found in
+appendix~\ref{appendix:methods}.
 
@@
 last interaction with one other individual. We compare this to the expected
-payoffs. Figure~\ref{fig:expected_and_stochastic_for_donation} shows simulation
-results for the described process of Section~\ref{section:methods}.
-
-In Section~\ref{section:donation}, we will specifically consider the donation
-game where each player can cooperate by providing a benefit \(b\) to the other
+payoffs. In this section we consider the donation game.
+Each player can cooperate by providing a benefit \(b\) to the other
 player at their cost \(c\), with \(0 < c < b\). Then, \(T=b, R=b-c, S=-c, P=0\),
@@
 
+Figure~\ref{fig:expected_and_stochastic_for_donation} shows simulations results
+for the described process of section~\ref{section:model}.
 Figure~\ref{fig:expected_and_stochastic_for_donation}~depicts the evolving
-conditional cooperation probabilities $p$ and $q$ (assuming that the discount
-factor~$\delta$ and the benefit $b$ are comparably high). The left panel
-corresponds to the standard scenario considered in the previous literature. It
-considers players who use expected payoffs to update their strategies. The right
-panel shows the scenario considered herein, in which players update their
-strategies based on their last round's payoff. The figure suggests that when
-updating is based on expected payoffs, players tend to be more generous (their
-$q$-values are higher on average). In addition, players are generally more
-cooperative.
+conditional cooperation probabilities $p$ and $q$. Assuming that the discount
+factor~$\delta$ is comparably high, the opening move \(y\) is a transient effect
+and has not effect on the outcome.
+The left panels corresponds to the standard scenario considered in the previous
+literature. It considers players who use expected payoffs to update their
+strategies. The right panel shows the scenario considered herein, in which
+players update their strategies based on their last round's payoff. The top
+panels assume a benefit of 3, whereas the bottom assume that the benefit is 10.
 
-There is a statistically significant different in the cooperation distribution of
-the population over time. The cooperation rate falls significantly yo under 50\%.
-There results are true for a strong selection strength ($\beta=10$). The follow
-up question that arises is what effect does the selection strength have on the
-cooperation rate?
+The figure suggests that when updating is based on expected payoffs, players
+tend to be more generous. The $q$-values are higher on average which suggests
+that individuals tend to cooperating more after they are at the receiving end
+of a defection. In addition, for both cases of \(b\), individuals tend to
+be more cooperative. The average cooperation rate is strictly higher for expected
+payoffs. This difference is statistically important, abd the effect is even more
+obvious when the benefit is higher. More specifically, for \(b=10\) the average
+cooperation rate drops from 97\% to 51\%.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{static/expected_and_stochastic_for_donation_game_b_10.pdf}
+    \includegraphics[width=.75\textwidth]{static/expected_and_stochastic_for_donation_game.pdf}\vspace{-3cm}
+    \includegraphics[width=.75\textwidth]{static/expected_and_stochastic_for_donation_game_b_10.pdf}
     \caption{{\bf Evolutionary dynamics under expected payoffs and stochastic payoffs.} 
@@
 
+Figure~\ref{fig:cooperation_rate_over_benefit} further explores the effect of the
+benefit on the cooperation rate. In all cases, the stochastic payoffs evolution
+tends to reduce the evolving cooperation rate. For the stochastic payoffs,
+benefit appears to not make a difference once \(b > 5\), and on average
+the evolving cooperation rate is at 50\%. The effect of benefit on the expected
+payoffs smooths out after \(b > 6\), and the expected payoffs estimating
+the evolving rate to be at 90\%.
+
+\begin{figure}[!htbp]
+  \centering
+  \begin{subfigure}{.65\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{static/expected_for_selection_strenght.pdf}
+    \includegraphics[width=\textwidth]{static/stochastic_for_selection_strenght.pdf}
+  \end{subfigure}%
+  \begin{subfigure}{.35\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{static/cooperation_rate_over_betas.pdf}
+  \end{subfigure}
+\caption{{\bf The evolution of cooperation for different benefit values.} 
+Here, we vary the benefit of cooperation $b$. In all cases, stochastic payoff
+evaluation tends to reduce the evolving cooperation rates. Unless explicitly
+varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$, $c\!=\!1$,
+$\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for $T\!=\!5\times 10^6$
+time steps for each parameter
+combination.}\label{fig:cooperation_rate_over_benefit}
+\end{figure}
+
 Figure~\ref{fig:cooperation_rate_over_betas} illustrates the results for
-different runs of the evolutionary process where $\beta\in\{10 ^ -2, 10 ^ - 1, 10
-^ 0, 10 ^ 1, 10 ^ 2\}$. In the case where $\beta=10 ^ -2$ the process is almost
-random, as the effect of the payoffs is very small. The cooperation rate for
-this case and the case of $\beta=10 ^ -1$ between the stochastic and expected
-payoffs are the same. However, it is clear that this does not apply given 
-$\beta \geq 1$. The difference in the cooperation rate increases, in the case
-where for the stochastic payoffs it can drop to 10\%.
+different runs of the evolutionary process where we vary the strength of
+selection. In the case of a very small \(\beta < \beta=10 ^ -1\) he process is
+almost random, as the effect of the payoffs is very small. Once payoff begin to
+matter the difference is once again evident, with the expected payoffs always
+overestimating the evolved cooperation rate.
 
@@
 
-Figure~\ref{fig:cooperation_rate_over_betas} illustrates the results for
-different runs of the evolutionary process where $\beta\in\{10 ^ -2, 10 ^ - 1, 10
-^ 0, 10 ^ 1, 10 ^ 2\}$. In the case where $\beta=10 ^ -2$ the process is almost
-random, as the effect of the payoffs is very small. The cooperation rate for
-this case and the case of $\beta=10 ^ -1$ between the stochastic and expected
-payoffs are the same. However, it is clear that this does not apply given 
-$\beta \geq 1$. The difference in the cooperation rate increases, in the case
-where for the stochastic payoffs it can drop to 10\%.
-
-\begin{figure}[!htbp]
-    \centering
-    \begin{subfigure}{.65\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{static/expected_for_selection_strenght.pdf}
-      \includegraphics[width=\textwidth]{static/stochastic_for_selection_strenght.pdf}
-    \end{subfigure}%
-    \begin{subfigure}{.35\textwidth}
-      \centering
-      \includegraphics[width=\textwidth]{static/cooperation_rate_over_betas.pdf}
-    \end{subfigure}
-\caption{{\bf The evolution of cooperation for different parameter values.} 
-While the previous figures depict the evolutionary outcome for fixed parameter
-values, here we vary the benefit of cooperation $b$ and the strength of
-selection $\beta$. In all cases, stochastic payoff evaluation tends to reduce
-the evolving cooperation rates. Unless explicitly varied, the parameters of the
-simulation are $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$,
-$\delta\!=\!0.99$. Simulations are run for $T\!=\!5\times 10^6$ time steps for
-each parameter combination.}
-\end{figure}
-
 \section{Expected and stochastic payoffs in $2 \times 2$ games}\label{section:2_by_2_games}
@@
 
-\section{Model Setup}\label{section:methods}
+\section{Model Setup}\label{appendix:methods}
 

### 2021-04-27 matlab/evolSimulation.m
remove trailing spaces

@@ -1,18 +1,18 @@
 function [xDat,AvCoop,AvPay,Rho,stochastic,Data]=evolSimulation(starting_resident, u, N, delta, beta, numberIterations, stochastic, filename);
 
 %% Preparations for the output
-Data=['R=',num2str(u(1)),'; S=',num2str(u(2)),'; T=',num2str(u(3)), '; P=',num2str(u(4)),'; N=',num2str(N),'; beta=',num2str(beta), '; nIt=',num2str(numberIterations), '; stochastic=',num2str(stochastic)]; 
+Data=['R=',num2str(u(1)),'; S=',num2str(u(2)),'; T=',num2str(u(3)), '; P=',num2str(u(4)),'; N=',num2str(N),'; beta=',num2str(beta), '; nIt=',num2str(numberIterations), '; stochastic=',num2str(stochastic)];
 AvCoop=0; AvPay=0; Res=starting_resident;
 
-sdim=3; 
+sdim=3;
 xDat=zeros(numberIterations/100,6);
-xDat(1,:)=[Res, 0, u(4), 0]; 
+xDat(1,:)=[Res, 0, u(4), 0];
 
 %% Calculating all possible pairwise imitation probabilities based on one payoff
-Rho=zeros(4,4); 
+Rho=zeros(4,4);
 for i1=1:4
     for i2=1:4
-        Rho(i1,i2) = 1 / (1 + exp(-beta * (u(i2) - u(i1)))); 
+        Rho(i1,i2) = 1 / (1 + exp(-beta * (u(i2) - u(i1))));
     end
 end
 
@@ -21,7 +21,7 @@ end
 j = 2;
 for t = progress(1:numberIterations)
     Mut=rand(1,sdim); 
-    [rho,coopM,piM]=CalcRho(Mut, Res, Rho, N, u, delta, beta, stochastic); 
+    [rho,coopM,piM]=CalcRho(Mut, Res, Rho, N, u, delta, beta, stochastic);
     if rand(1)<rho
         Res=Mut; xDat(j,:)=[Res,coopM,piM,t]; j=j+1;
     end
@@ -37,30 +37,30 @@ end
 
 
 
-function [rho,coopMM,piMM]=CalcRho(Mut, Res, Rho, N, u, delta, beta, stochastic); 
[…]

### 2021-04-27 main.tex
draft donation section

@@
 \caption{{\bf The evolution of cooperation for different benefit values.} 
-Here, we vary the benefit of cooperation $b$. In all cases, stochastic payoff
+Here, we vary the benefit of defection $b$. In all cases, stochastic payoff
 evaluation tends to reduce the evolving cooperation rates. Unless explicitly
@@
   \end{subfigure}
-\caption{{\bf The evolution of cooperation for different parameter values.} 
-While the previous figures depict the evolutionary outcome for fixed parameter
-values, here we vary the benefit of cooperation $b$ and the strength of
-selection $\beta$. In all cases, stochastic payoff evaluation tends to reduce
-the evolving cooperation rates. Unless explicitly varied, the parameters of the
-simulation are $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$,
-$\delta\!=\!0.99$. Simulations are run for $T\!=\!5\times 10^6$ time steps for
-each parameter combination.}\label{fig:cooperation_rate_over_betas}
+\caption{{\bf The evolution of cooperation for different selection strength values.} 
+Here, we vary the selection strength $\beta$. In all cases, stochastic payoff
+evaluation tends to reduce the evolving cooperation rates. Unless explicitly
+varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$, $c\!=\!1$,
+$\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for $T\!=\!5\times 10^6$
+time steps for each parameter
+combination.}\label{fig:cooperation_rate_over_betas}
 \end{figure}
@@
 
-In evolution context we consider a population of \(N\) players,
-where \(N\) is even, and where mutations are sufficiently rare. At any point
-in time the there are at most two different strategies are present in
-the population. Suppose there are \(N - k\) players who use the strategy \(s_{1}=(y_{1},
-p_{1}, q_{1})\), whereas \(k\) players use the strategy \(s_{2}=(y_{2}, p_{2},
-q_{2})\). We refer to these two player types as ‘residents’ and ‘mutants’,
-respectively.
+In order to account for the difference in the robustness of cooperation among
+individuals based on their payoff memory, we consider a population of \(N\)
+players, where \(N\) is even, and where mutations are sufficiently rare. At any
+point in time the there are at most two different strategies are present in the
+population. A \textit{resident} strategy and a \textit{mutant} strategy. Suppose
+there are \(N - k\) players who use the resident strategy whereas \(k\) players
+use the mutant strategy. Each step of the evolutionary process consists of two
+stages, a game stage and an updating stage.
 
-Each step of the evolutionary process consists of two stages, a game stage and
-an updating stage.
+In the game stage, each player is randomly matched with some other player in
+the population to interact in a number of instances of the game
+\(\begin{pmatrix} R & S  \\ T & P \end{pmatrix}\). The number of interactions is
+not fixed, on the contrast we consider that another interaction has a
+probability $\delta$ to continue following each turn. We assume herein that
+individuals at most make use of simple \textit{reactive strategies} make
+decisions in each round. Reactive strategy are a set of memory-one strategies
+that only take into account the previous action of the opponent. Reactive
+strategies can be written explicitly as a vector \(\in \R_{3}\). More
+specifically a reactive strategy \(s\) is given by \(s=(y, p, q)\) where \(y\)
+is the probability that the strategy opens with a cooperation and \(p, q\) are
+the probabilities that the strategy cooperates given that the opponent
+cooperated and defected equivalently.
 
-\begin{enumerate}
-    \item In the game stage, each player is randomly matched with some
-    other player in the population to interact in one instance of the IPD.
-    \item In the updating stage, two players are randomly drawn from
-    the population, a `learner' and a `exemplar'. Given that the learner's payoff
-    in the last round is $u_L\!\in\! \mathcal{U}$ and that the exemplar's last
-    round's payoff $u_E\!\in\! \mathcal{U}$, we assume the learner adopts the role
-    model's strategy with probability
+In the updating stage, two players are randomly drawn from the population, a
+`learner' and a `role model'. Given that the learner's payoff $u_L\!\in\!
+\mathcal{U}$ and that the role model's payoff $u_{RL}\!\in\! \mathcal{U}$, we
+assume the learner adopts the role model's strategy with probability
 
 \begin{equation} \label{Eq:rho}
-    \rho(u_L, u_E) = \frac{1}{1\!+\!\exp\big[ \!-\!\beta (u_E\!-\!u_L) \big]}. 
+\rho(u_L, u_{RL}) = \frac{1}{1\!+\!\exp\big[ \!-\!\beta (u_E\!-\!u_{RL}) \big]}.
 \end{equation}
 
-    where $\beta\!\ge\!0$ corresponds to the strength of selection.
-\end{enumerate}
+where $\beta\!\ge\!0$ corresponds  to relative influence of the payoffs on the
+adopt the strategy of the other is controlled by an external parameter, the so
+called intensity of selection.
 
@@
 becomes the new resident strategy). After either outcome, we introduce a new
-mutant strategy $s'_2\!=\!(y_2',p_2',q_2')$ (uniformly chosen from all reactive
+mutant strategy (uniformly chosen from all reactive
 strategies at random), and we set the number of mutants to $k\!=\!1$. This
-process of mutation and fixation/extinction is then iterated many times.
+process of mutation and fixation/extinction is then iterated many times.xw The
+fixation probability of the mutant strategy then takes the standard form
+\cite{nowak2004emergence},
 
-\section{Expected Payoffs}
+\begin{equation}
+\varphi = \frac{1}{1+\sum_{i=1}^{N-1}\prod_k^i \frac{\lambda^-_k}{\lambda^+_k}}.
+\end{equation}
 
-We compare this process for what we defined as {\bf stochastic payoff
-evaluation} with the analogous process where players update their strategies
-with respect to their {\bf expected} payoffs,
+where \(\lambda^-_k, \lambda^+_k\) are the probabilities that the number of
+mutants decreases and increases respectively.
 
-\begin{equation} \label{eq:ExpPay}
+We compare this process for stochastic payoff evaluation with the analogous
+process where players update their strategies with respect to their {\it
+expected} payoffs.
+
+\subsection{Expected Payoffs}
+
+Consider two players with reactive strategies $s_1\!=\!(y_1, p_1, q_1$) and
+$s_2\!=\!(y_2,p_2,q_2)$ who interact in a repeated prisoner's dilemma with
+continuation probability $\delta$, the probability that are in each of the
+four possible states in the last round is given by the stationary distribution
+\(\langle\mathbf{v}\) of the transition matrix \(M\).
+
+\begin{equation}\label{eq:transition_matrix}
+  \input{static/matrix.tex}.
+\end{equation}
+
+The long run steady state probability vector \(\langle\mathbf{v}\) is the solution
+to \(\langle\mathbf{v}M=\langle\mathbf{v}\), can be combined with the payoff
+\(u\) to give the payoffs for each player. More specifically,
+
+\[\langle\mathbf{v}(s_1,s_2),\mathbf{u}\rangle\]
+
+is the payoff that \(s_1\) achieves when interacting with \(s_2\). In
+the evolutionary process given \(k\) individuals with mutant strategy \(s_M\)
+and \(N - k\) individuals with the resident strategy \(s_R\) the expected
+payoffs are,
+
+\begin{equation} \label{Eq:ExpPay}
   \begin{array}{lcrcr}
-\pi_1 &= &\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(s_1,s_1),\mathbf{U}\rangle	& + &\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(s_1,s_2),\mathbf{U}\rangle,\\[0.5cm]
-\pi_2 &= &\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(s_2,s_1),\mathbf{U}\rangle & + &\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(s_2,s_2),\mathbf{U}\rangle.\\
+  \displaystyle \pi_R	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(s_R,s_R),\mathbf{u}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(s_R,s_M),\mathbf{u}\rangle,\\[0.5cm]
+  \displaystyle \pi_M	&=	&\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(s_M,s_R),\mathbf{u}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(s_M,s_M),\mathbf{u}\rangle.\\
   \end{array}
@@
 
-\section{Stochastic Payoffs}
+\subsection{Stochastic Payoffs}
 
-We define {\bf stochastic payoff} as the average payoff $u\!\in\! \mathcal{U}$ a player
-receives in the last \(n\) rounds of the game given that they interact with
-\(m\) players.
-
-\subsubsection*{Case \(n=m=1\).}
-
-Initially, consider the situation where \(n=m=1\). The player's stochastic
-payoff is what they receive in the last round against a single opponent. There
+Initially, we consider the case that an individual's fitness is based on their
+last interaction with one other individual. There
 only four possible outcomes for the last round, those are \(CC, CD, DC, DD\).
-Consider two players with reactive strategies $S_1\!=\!(y_1, p_1, q_1$) and
-$S_2\!=\!(y_2,p_2,q_2)$ who interact in a repeated prisoner's dilemma with
+Consider two players with reactive strategies $s_1\!=\!(y_1, p_1, q_1$) and
+$s_2\!=\!(y_2,p_2,q_2)$ who interact in a repeated prisoner's dilemma with
 continuation probability $\delta$, the probability that are in each of the
@@
 
-Equation~\ref{Eq:LastRound} is the probability vector that players \(S_1, S_2\)
+Equation~\ref{Eq:LastRound} can be generalised to all cases where we consider
+the last \()
+
+Equation~\ref{Eq:LastRound} is the probability vector that players \(s_1, s_2\)
 are in each of the possible states of the last round. Given the population \(N\)
@@
 \end{figure}
-
-We compare this process for stochastic payoff evaluation with the analogous process where players update their strategies with respect to their {\it expected} payoffs,
-\begin{equation} \label{Eq:ExpPay}
-\begin{array}{lcrcr}
-\displaystyle \pi_1	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(S_1,S_1),\mathbf{u}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(S_1,S_2),\mathbf{u}\rangle,\\[0.5cm]
-\displaystyle \pi_2	&=	&\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(S_2,S_1),\mathbf{u}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(S_2,S_2),\mathbf{u}\rangle.\\
-\end{array}
-\end{equation}
-In the limit of no discounting, $\delta\!\rightarrow\! 1$, this process based on expected payoffs has been considered in~\cite{imhof2010stochastic}.
 
@@
 \end{equation}
-The first term on the right side corresponds to the case that the learner and the role model happened to be matched during the game stage, which happens with probability $1/(N\!-\!1)$.
-In that case, we note that only those payoff pairs can occur that are feasible in a direct interaction, $(u_1,u_2)\in \mathcal{U}^2_F:=\big\{ (R,R), (S,T), (T,S), (P,P) \big\}$, as represented by the respective indicator function.
-Otherwise, if the learner and the role model did not interact directly, we need to distinguish four different cases, depending on whether the learner was matched with a resident or a mutant, and depending on whether the role model was matched with a resident or a mutant.
+
+The first term on the right side corresponds to the case that the learner and
+the role model happened to be matched during the game stage, which happens with
+probability $1/(N\!-\!1)$. In that case, we note that only those payoff pairs
+can occur that are feasible in a direct interaction, $(u_1,u_2)\in
+\mathcal{U}^2_F:=\big\{ (R,R), (S,T), (T,S), (P,P) \big\}$, as represented by
+the respective indicator function. Otherwise, if the learner and the role model
+did not interact directly, we need to distinguish four different cases,
+depending on whether the learner was matched with a resident or a mutant, and
+depending on whether the role model was matched with a resident or a mutant.
 

### 2021-05-03 main.tex
draft last sections

@@
 \usepackage{hyperref}
+\usepackage{tabularx}
+\usepackage{booktabs}
+\usepackage[ruled,vlined]{algorithm2e}
 \smallskip % Erlaubt kleine Abstaende zwischen Paragraphen, falls es dem Seitenlayout hilft
@@
     fill=black, text width=1.5em},% arbre rouge noir, noeud noir
-  arn_r/.style = {treenode, circle, red, fill=red, 
-    text width=1.5em, very thick, inner sep=-6pt},% arbre rouge noir, noeud rouge
+  arn_r/.style = {treenode, circle, red, 
+    text width=1.5em, very thick, inner sep=4pt},% arbre rouge noir, noeud rouge
   arn_x/.style = {treenode, rectangle, draw=black,
@@
 {}%         Thm head spec
+
+\newcommand{\splitatcommas}[1]{%
+  \begingroup
+  \begingroup\lccode`~=`, \lowercase{\endgroup
+    \edef~{\mathchar\the\mathcode`, \penalty0 \noexpand\hspace{0pt plus 1em}}%
+  }\mathcode`,="8000 #1%
+  \endgroup
+}
 
@@
   \begin{subfigure}{.35\textwidth}
-    \centering
+    \vspace{.3cm}
     \includegraphics[width=\textwidth]{static/cooperation_rate_over_betas.pdf}
@@
 
-In order to account for the difference in the robustness of cooperation among
-individuals based on their payoff memory, we consider a population of \(N\)
-players, where \(N\) is even, and where mutations are sufficiently rare. At any
-point in time the there are at most two different strategies are present in the
-population. A \textit{resident} strategy and a \textit{mutant} strategy. Suppose
-there are \(N - k\) players who use the resident strategy whereas \(k\) players
-use the mutant strategy. Each step of the evolutionary process consists of two
-stages, a game stage and an updating stage.
-
-In the game stage, each player is randomly matched with some other player in
-the population to interact in a number of instances of the game
-\(\begin{pmatrix} R & S  \\ T & P \end{pmatrix}\). The number of interactions is
-not fixed, on the contrast we consider that another interaction has a
-probability $\delta$ to continue following each turn. We assume herein that
-individuals at most make use of simple \textit{reactive strategies} make
-decisions in each round. Reactive strategy are a set of memory-one strategies
-that only take into account the previous action of the opponent. Reactive
-strategies can be written explicitly as a vector \(\in \R_{3}\). More
-specifically a reactive strategy \(s\) is given by \(s=(y, p, q)\) where \(y\)
-is the probability that the strategy opens with a cooperation and \(p, q\) are
-the probabilities that the strategy cooperates given that the opponent
-cooperated and defected equivalently.
+We consider a pairwise comparison process. We assume a population of \(N\)
+individuals (\(N\) is even) where at any point in time there are at most two
+different strategies in present. There are \(k\) individuals who play the mutant
+strategy and \(N - k\) individuals who play the resident strategy. Each step of
+the evolutionary process consists of two stages; a game stage and an update
+stage.
+
+In the game stage, each individual is randomly matched with some other
+individual in the population. Their interaction lasts for a number of turns
+which is not fixed but depends on the continuation probability \(\delta\). At
+each turn the individuals have choose between cooperation (C) and defection (D).
+If both players cooperate they receive the reward payoff \(R\), whereas if both
+players defect they receive the punishment payoff \(P\). If one cooperates but
+the other defects, the defector receives the temptation to defect, \(T\),
+whereas the cooperator receives the sucker's payoff, \(S\).
+
+In this work we consider different social dilemmas represented by the harmony
+game, the stag hunt game, the snowdrift or hawk-dove game, the prisoner's
+dilemma, and a special case of the prisoner's dilemma, the donation game
+(Table~\ref{table:social_dilemmas}).
+
+\begin{table}[!htbp]
+  \begin{center}
+    \resizebox{.65\textwidth}{!}{
+    \renewcommand{\arraystretch}{1.5}
+      \begin{tabular}{ccc}
+  \toprule
+  & \textbf{social dilemmas} & \textbf{payoffs' constrains} \\
+  \midrule
+(i) & stage hunt  & \(R > T > P > S\) \\
+(ii) & snowdrift  & \(T > R > S > P\) \\
+(iii) & harmony  & \(R > T > S > P\) \\
+(iv) & prisoner dilemma  & \(T > R > P > S\) \\
+(v)& donation game  & \(T > R > P > S\) \& \(T=b, R=b-c, S=-c, P=0\) \\
+  \bottomrule
+      \end{tabular}}
+  \end{center}
+  \caption{We present results on the all of the listed social dilemmas. The payoffs'
+  constrains are different for each game.}
+  \label{table:social_dilemmas}
+\end{table}
+
+Results for cases (i) - (iv) were presented in section~\ref{section:2_by_2_games}
+and results for case (v) were presented in section~\ref{section:donation}.
+
+A further assumption of our model is that individuals make use of reactive
+strategies when they make decisions in each round. Reactive strategy are a set
+of strategies that take into account only the previous action of the opponent.
+A reactive strategy can be written explicitly as a vector,
+
+\[s=(y, p, q)\]
+
+where \(y\) is the probability that the strategy opens with a cooperation and
+\(p, q\) are the probabilities that the strategy cooperates given that the
+opponent cooperated and defected equivalently.
 
 In the updating stage, two players are randomly drawn from the population, a
-`learner' and a `role model'. Given that the learner's payoff $u_L\!\in\!
-\mathcal{U}$ and that the role model's payoff $u_{RL}\!\in\! \mathcal{U}$, we
-assume the learner adopts the role model's strategy with probability
+`learner' and a `role model'. The learner adopts the role model's strategy
+based on the Fermi distribution function, %ToDo add reference
 
 \begin{equation} \label{Eq:rho}
-\rho(u_L, u_{RL}) = \frac{1}{1\!+\!\exp\big[ \!-\!\beta (u_E\!-\!u_{RL}) \big]}.
+\rho(u_{L}, u_{RM}) = \frac{1}{1\!+\! \exp^{\!-\!\beta (u_{RM}\!-\!u_{L})}}.
 \end{equation}
 
-where $\beta\!\ge\!0$ corresponds  to relative influence of the payoffs on the
-adopt the strategy of the other is controlled by an external parameter, the so
-called intensity of selection.
+where $u_{L}\!\in\! \mathcal{U}$ is the learner's payoff, $u_{RM}\!\in\!
+\mathcal{U}$ is the role model's payoff, and $\beta\!\ge\!0$ is the intensity of
+selection.
 
 We iterate this basic evolutionary step until either the mutant strategy goes
-extinct, or until it fixes in the population (in which case the mutant strategy
-becomes the new resident strategy). After either outcome, we introduce a new
-mutant strategy (uniformly chosen from all reactive
-strategies at random), and we set the number of mutants to $k\!=\!1$. This
-process of mutation and fixation/extinction is then iterated many times.xw The
-fixation probability of the mutant strategy then takes the standard form
-\cite{nowak2004emergence},
-
-\begin{equation}
-\varphi = \frac{1}{1+\sum_{i=1}^{N-1}\prod_k^i \frac{\lambda^-_k}{\lambda^+_k}}.
+extinct, or until it fixes in the population and becomes the new resident
+strategy. After either outcome, we set $k$ to $1$ and we introduce a new mutant
+strategy which is uniformly chosen from all reactive strategies at random.
+
+This process of mutation and fixation/extinction is iterated many times.
+Instead of simulating each step of the evolutionary process, we estimate
+the probability that a newly introduced mutant fixes~\cite{nowak2004emergence}.
+This is defined as the fixation probability of the mutant, and the standard
+form is the following,
+
+\begin{equation}\label{eq:appendix_fixation_probability}
+\varphi = \frac{1}{1+\sum\limits_{i=1}^{N-1}\prod\limits_k^i \frac{\lambda^-_k}{\lambda^+_k}},
 \end{equation}
@@
 
-We compare this process for stochastic payoff evaluation with the analogous
-process where players update their strategies with respect to their {\it
-expected} payoffs.
-
-\subsection{Expected Payoffs}
-
-Consider two players with reactive strategies $s_1\!=\!(y_1, p_1, q_1$) and
-$s_2\!=\!(y_2,p_2,q_2)$ who interact in a repeated prisoner's dilemma with
-continuation probability $\delta$, the probability that are in each of the
-four possible states in the last round is given by the stationary distribution
-\(\langle\mathbf{v}\) of the transition matrix \(M\).
+This evolutionary process is summarized by Algorithm~\ref{algorithm:pairwise_comparison}.
+
+\begin{algorithm}[!htbp]
+  \SetAlgoLined
+   $N \leftarrow$ population size\;
+   $k \leftarrow 1$\;
+   resident $\leftarrow (0, 0, 0)$\;
+   \While{step $<$ maximum number of steps}{
+    mutant $\leftarrow$ random: $\{\emptyset \}\ \rightarrow R^3$\;
+    fixation probability $\leftarrow \varphi $\;
+    \If{$\varphi >$ random: $i \rightarrow [0, 1]$}{
+      resident $\leftarrow$ mutant;
+     }}
+     \caption{Pairwise comparison process}\label{algorithm:pairwise_comparison}
+\end{algorithm}
+
+The aim of this work is to explore the effect of updating memory on the
+cooperation rate of the evolved population. For this reason we consider two
+different approaches when estimating the payoffs at the updating stage. The two
+approaches we consider are those of (i) the expected payoffs and (ii) the
+stochastic payoffs.
+
+\subsection*{Expected Payoffs}
+
+The expected payoffs are the conventional payoffs used in the updating
+stage~\cite{imhof2010stochastic}. They are defined as the mean payoff an
+individual in a well-mixed population that engages in repeated games with all
+other population members.
+
+We first define the payoff of two reactive strategies at the game stage. Assume
+two reactive strategies $s_1\!=\!(y_1, p_1, q_1$) and $s_2\!=\!(y_2,p_2,q_2)$.
+It is not necessary to simulate the play move by move. Instead the play between
+the two strategies is defined a Markov matrix \(M\),
 
@@
 
-The long run steady state probability vector \(\langle\mathbf{v}\) is the solution
-to \(\langle\mathbf{v}M=\langle\mathbf{v}\), can be combined with the payoff
-\(u\) to give the payoffs for each player. More specifically,
-
-\[\langle\mathbf{v}(s_1,s_2),\mathbf{u}\rangle\]
-
-is the payoff that \(s_1\) achieves when interacting with \(s_2\). In
-the evolutionary process given \(k\) individuals with mutant strategy \(s_M\)
-and \(N - k\) individuals with the resident strategy \(s_R\) the expected
-payoffs are,
+whose stationary vector \(\mathbf{v}\), combined with the payoff \(U\), yields
+the expected outcome for each strategy,
+\(\langle\mathbf{v}(s_1,s_2),\mathbf{u}\rangle\).
+
+% In each step
+Given a population of \(N\) individual with \(k\) individuals playing the mutant
+strategy, \(s_M = (y_M, p_M, q_M)\), and \(N -k\) playing the resident strategy,
+\(s_R = (y_R, p_R, q_R)\), the expected payoffs of a resident \(\pi_R\) and of
+a mutant strategy \(\pi_M\) are defined as,
 
@@
   \begin{array}{lcrcr}
-  \displaystyle \pi_R	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(s_R,s_R),\mathbf{u}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(s_R,s_M),\mathbf{u}\rangle,\\[0.5cm]
-  \displaystyle \pi_M	&=	&\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(s_M,s_R),\mathbf{u}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(s_M,s_M),\mathbf{u}\rangle.\\
+  \displaystyle \pi_R	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(s_R,s_R),\mathbf{U}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(s_R,s_M),\mathbf{U}\rangle,\\[0.5cm]
+  \displaystyle \pi_M	&=	&\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(s_M,s_R),\mathbf{U}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(s_M,s_M),\mathbf{U}\rangle.\\
   \end{array}
@@
 
-In the limit of no discounting, $\delta\!\rightarrow\! 1$, this process based on
-expected payoffs has been considered in~\cite{imhof2010stochastic}.
-
-\subsection{Stochastic Payoffs}
-
-Initially, we consider the case that an individual's fitness is based on their
-last interaction with one other individual. There
-only four possible outcomes for the last round, those are \(CC, CD, DC, DD\).
-Consider two players with reactive strategies $s_1\!=\!(y_1, p_1, q_1$) and
-$s_2\!=\!(y_2,p_2,q_2)$ who interact in a repeated prisoner's dilemma with
-continuation probability $\delta$, the probability that are in each of the
-four possible states in the last round is given by:
-
-\begin{equation}
-    \mathbf{v}(s_1,s_2)\!=\!\Big(\mathbf{v}_{R}(s_1,s_2),\mathbf{v}_{S}(s_1,s_2),\mathbf{v}_{T}(s_1,s_2),\mathbf{v}_{P}(s_1,s_2)\Big).
-\end{equation}
-
-where,
+Note that \(N\!-\!k\!-\!1\) is the number of other residents in the population,
+and \(k\) is the number of residents. The payoffs are averaged by the number of
+individuals in the population, given that interactions are not possible.
+The same applies for the payoff of a mutant.
+
+Given the expected payoff, the probabilities that the number of mutants
+decreases and increases  (Equation~(\ref{eq:appendix_fixation_probability})),
+are given by:
+
+\begin{align*} 
+  \lambda^-_k &\!=\!\rho(\pi_R, \pi_M) \\
+  \lambda^+_k &\!=\!\rho(\pi_M, \pi_R).
+\end{align*}
+
+\subsection*{Stochastic Payoffs}
+
+We compare the expected payoffs with the finite memory payoffs which we refer to
+as the stochastic payoffs. Initially, we consider the case of where the payoffs
+correspond to the last round payoff an individual achieved after interacting
+with one other individual.
+
+The probability of being in any of the outcomes \(R, S, T, P\) in the last round
+is given by Equation~(\ref{Eq:LastRound})
+(Proposition~\ref{proposition:last_round}).
+
+\begin{Prop}\label{proposition:last_round}
+Assume a play between the reactive strategies \(s_1\) and \(s_2\) with a
+continuation probability \(\delta\). Then the probability of being in any of the
+four outcomes \(R, S, T, P\) are given by,
 
@@
 
+In these expressions, we have used
+the notation $r_i:=p_i\!-\!q_i$, $\bar{y}_i\!=\!1\!-\!y_i$, $\bar{q}_i:=1\!-\!q_i$,
+and $\bar{r}_i:=\bar{p}_i\!-\!\bar{q}_i=-r_i$ for $i\!\in\!\{1,2\}$.
+Let $\mathcal{U}=\{R,S,T,P\}$ denote the set of feasible payoffs in each round,
+and let $\mathbf{u}\!=\!(R,S,T,P)$ be the corresponding payoff vector.
+\end{Prop}
+
 \begin{proof}
-
-Assume a repeated prisoner's dilemma between two reactive strategies. Given the
-continuation probability $\delta$, probability that the game ends in the after
-the first round $(1 - \delta)$ and the expected distribution of the four
-outcomes in the very first round is $\mathbf{v_0}$ defined as. Following the
-first round the, the outcome of the next rounds with a probability $\delta$
-is \(M\) such that,
-
-\[\dots\]
-
-It can shown that, \((1\!-\!\delta)\mathbf{v_0}(I_4-\delta M)^{-1}\) and with
-some algebraic manipulation we derive to Equation~\ref{Eq:LastRound}.
-
+Given a play between two reactive strategies with continuation probability
+$\delta$. The probability that the game ends on round \(t\) is given by,
+
+\begin{equation}\label{eq:}
+  (1 - \delta) \mathbf{v_0} \sum \delta^{t} M^{(t)}.
+\end{equation}
+
+Here, $\mathbf{v_0}$ denotes the expected distribution of the four outcomes in
+the very first round, and \(1- \delta\) the probability that the game ends.
+It can be shown that,
+
+\begin{align*}
+  (1 - \delta) \mathbf{v_0} \sum \delta^{t} M^{(t)} & = (1 - \delta)(\mathbf{v_0} + \delta \mathbf{v_0} M + \delta^{2}\mathbf{v_0} M ^{2} + \dots )\\ 
+   & = (1 - \delta)\mathbf{v_0} (1 + \delta M + \delta^{2}M ^{2} + \dots ) \text{ using standard formula for geometric series}\\ 
+   & = (1 - \delta)\mathbf{v_0}(I_4 - \delta M)^{-1}
+\end{align*}
+
+and with some algebraic manipulation we derive to Equation~\ref{Eq:LastRound}.
 \end{proof}
 
-Equation~\ref{Eq:LastRound} can be generalised to all cases where we consider
-the last \()
-
-Equation~\ref{Eq:LastRound} is the probability vector that players \(s_1, s_2\)
-are in each of the possible states of the last round. Given the population \(N\)
-with \(k\) mutants in the last round there only possible combinations of interactions
-are:
-
-\begin{figure}
+Considering that individuals can only interact with one other individual,
+at that in each step a mutant and a resident are selected there are only five
+possible pairings:
+
+\begin{itemize}
+  \item the selected resident is paired with the selected mutant
+  \item the selected resident is paired with a resident and the selected mutant is paired with a resident
+  \item the selected resident is paired with a resident and the selected mutant is paired with a mutant
+  \item the selected resident is paired with a mutant and the selected mutant is paired with a resident
+  \item the selected resident is paired with a mutant and the selected mutant is paired with a mutant
+\end{itemize}
+
+Each of the pairings happens with a given probability and that is illustrated
+by Figure~\ref{fig:single_pairs}.
+
+\begin{figure}[!htbp]
   \centering
   \includegraphics[width=.65\textwidth]{static/last_rounds_matches.tex}
+  \caption{The tree}\label{fig:single_pairs}
 \end{figure}
 
-\subsection{Fixation probabilities under stochastic payoff evaluation}
-Given that $N\!-\!k$ players use the resident strategy $S_1\!=\!(y_1,p_1,q_1)$ and that the remaining $k$ players use the mutant strategy $S_2\!=\!(y_2,p_2,q_2)$, the probability that the number of mutants increases by one in one step of the evolutionary process can be written as
-\begin{equation}
-\lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_1,u_2\in\mathcal{U}} x(u_1,u_2)\cdot \rho(u_1,u_2).
-\end{equation}
-In this expression, $(N\!-\!k)/N$ is the probability that the randomly chosen learner is a resident, and $k/N$ is the probability that the role model is a mutant.
-The sum corresponds to the total probability that the learner adopts the role model's strategy over all possible payoffs $u_1$ and $u_2$ that the two player may have received in their respective last rounds.
-We use $x(u_1,u_2)$ to denote the probability that the randomly chosen resident obtained a payoff of $u_1$ in the last round of his respective game, and that the mutant obtained a payoff of $u_2$.
-Given that the payoffs are $u_1$ and $u_2$, the imitation probability is then given by $\rho(u_1,u_2)$, as specified by Eq.~(\ref{Eq:rho}).
-The probability that the respective payoffs of the players are given by $u_1$ and $u_2$ can be calculated as
-\begin{equation}
+The probability that the respective payoffs of the players
+are given by $u_1$ and $u_2$ can be calculated as
+
+\begin{equation}\label{eq:Chi}
 \setlength{\arraycolsep}{1pt}
@@
 the role model happened to be matched during the game stage, which happens with
-probability $1/(N\!-\!1)$. In that case, we note that only those payoff pairs
-can occur that are feasible in a direct interaction, $(u_1,u_2)\in
-\mathcal{U}^2_F:=\big\{ (R,R), (S,T), (T,S), (P,P) \big\}$, as represented by
+probability $1/(N\!-\!1)$.In that case, we note that only those payoff pairs
+can occur that are feasible in a direct interaction, $\splitatcommas{(u_1,u_2)\in
+\mathcal{U}^2_F:=\big\{ (R,R), (S,T), (T,S), (P,P) \big\}}$, as represented by
 the respective indicator function. Otherwise, if the learner and the role model
@@
 depending on whether the role model was matched with a resident or a mutant.
+
+Given that $N\!-\!k$ players use the resident strategy $S_1\!=\!(y_1,p_1,q_1)$
+and that the remaining $k$ players use the mutant strategy
+$S_2\!=\!(y_2,p_2,q_2)$, the probability that the number of mutants increases by
+one in one step of the evolutionary process can be written as
+
+\begin{align}
+\lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_1,u_2\in\mathcal{U}} x(u_1,u_2)\cdot \rho(u_1,u_2), \\
+\lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_1,u_2\in\mathcal{U}} x(u_1,u_2)\cdot \rho(u_2,u_1).
+\end{align}
+
+In this expression, $(N\!-\!k)/N$ is the probability that the randomly chosen
+learner is a resident, and $k/N$ is the probability that the role model is a
+mutant. The sum corresponds to the total probability that the learner adopts the
+role model's strategy over all possible payoffs $u_1$ and $u_2$ that the two
+player may have received in their respective last rounds. We use $x(u_1,u_2)$ to
+denote the probability that the randomly chosen resident obtained a payoff of
+$u_1$ in the last round of his respective game, and that the mutant obtained a
+payoff of $u_2$.
+
+This framework can be extended to consider the case of where the payoffs
+correspond to the last \(n\) rounds payoff an individual achieved after
+interacting with \(m\) other individuals. For the case \(n=2\) the payoffs
+at the game stage are,
+
+\begin{Prop}\label{proposition:last_two_rounds}
+Assume a play between the reactive strategies \(s_1\) and \(s_2\) with a
+continuation probability \(\delta\). Then the probability of being in any of the
+sixteen outcomes \(\splitatcommas{RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR}\)
+on the last two rounds are given by,
+
+\begin{equation}
+  \mathbf{v_{a_1, a_2}} = (1 - \delta) m_{a_1, a_2} \delta^2 \left[\mathbf{v_0}(I_4 - \delta M)^{-1}\right]_{a_1, a_2}, \quad \text{ for } m_{a_1, a_2} \in M \ \& \ a_1, a_2 \in \{R, S, T, P\}
+\end{equation}
+
+\end{Prop}
+
+Proposition~\ref{proposition:last_two_rounds} can be extended to the last \(n\)
+rounds.
+
+\begin{Prop}\label{proposition:last_n_rounds}
+Assume a play between the reactive strategies \(s_1\) and \(s_2\) with a
+continuation probability \(\delta\). Then the probability of being in any of the
+sixteen outcomes \(\splitatcommas{RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR}\)
+on the last two rounds are given by,
+
+\begin{equation}
+  \mathbf{v_{a_1, a_2}} = (1 - \delta) \prod m_{a_1, a_2} \delta^2 \left[\mathbf{v_0}(I_4 - \delta M)^{-1}\right]_{a_1, a_2}
+\end{equation}
+
+for \(m_{a_1, a_2} \in M\) and \(a_1, a_2 \in [1, 4]\).
+\end{Prop}
+
+Equation~\ref{eq:Chi} can also be extended to include interactions with two other
+individuals. The possible pairings are illustrated by Figure~\ref{}.
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=.65\textwidth]{static/tree.tex}
+  \caption{The tree}\label{fig:two_pairs}
+\end{figure}
+
+\section{Verifying analytical results with simulations}
+
+The analytical results presented in this work have been verified with simulations.
+More specifically the probabilities of Equation~(\ref{Eq:LastRound}),
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=.75\textwidth]{static/stationary_four.pdf}
+\caption{}
+\end{figure}
+
+Proposition~\ref{proposition:last_two_rounds},
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=\textwidth]{static/stationary_sixteen.pdf}
+\caption{}
+\end{figure}
+
+% \begin{Prop}\label{proposition:last_n_rounds}
+
+% \end{Prop}
 

### 2021-05-03 static/last_rounds_matches.tex
draft last sections

\documentclass{standalone}

\begin{document}
\begin{tikzpicture}[->,>=stealth',
    level/.style={sibling distance = 5cm/#1,
    level distance = 1.7cm},
    level 2/.style = {sibling distance = 2cm, level distance=2.8cm},
    every node/.style = {font=\tiny}]
\node [arn_n] {} 
    child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{M}$\\$S_{R} - S_{R}$}] {} edge from parent node[above left, rotate=30] {$\frac{1}{N - 1}$}}
    child{ node [arn_n, rotate=-30] {} 
            child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{R}$\\$S_{M} - S_{R}$}] {} edge from parent node[above, yshift=-.1cm, rotate=39] {$\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$}}
            child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{R}$\\$S_{R} - S_{M}$}] {} edge from parent node[above, yshift=-.2cm, rotate=67] {$\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$}}
            child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{M}$\\$S_{M} - S_{R}$}] {} edge from parent node[above, yshift=-.2cm, rotate=-67] {$\frac{(k-1)(N-k-1)}{(N-2)(N-3)}$}}
            child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{R}$\\$S_{R} - S_{R}$}] {} edge from parent node[above, yshift=-.1cm, rotate=-41] {$\frac{(k-1)(k-2)}{(N-2)(N-3)}$}}
            edge from parent node[above left, xshift=1em, yshift=-.8em, rotate=-30] {$1 - \frac{1}{N - 1}$}
		}
; 
\end{tikzpicture}
\end{document}

### 2021-05-04 main.tex
a good draft of the first part

@@
 One of the most important applications of evolutionary game theory is the
-evolution of cooperation. Why is it that some individuals choose to help others
-(increasing their payoff) at the expense of decreasing one's own payoff? During
-the past decade, the literature on mechanisms that allow for cooperation, even
-though it's seen to be at odd has been fruitful. One such mechanism is
-repetition; the so called direct reciprocity. Theoretical researchers have been
-using evolutionary models to understand how direct reciprocity allows
-cooperation to evolve and which strategies are important for sustaining
-cooperation.
-
-Evolutionary game theory does not require individuals to be rational, instead
-they adapt strategies based on mutation and exploration. Strategies with
-high fitness are more likely to spread, either because the individuals who
-adopt them have more offspring (fitness-based processes), or they are imitated
-more often (pairwise comparison processes)~\cite{Wu2015}. The fitness of strategies,
-and subsequently of the individuals, is not constant. Instead it depends on the
-composition of the population. Individuals interact with other members of the
-population according to their strategies. Their yielded payoffs, according to
-pre defined games, are translated into fitness.
-
-It is commonly assumed that fitness assimilates to the \textit{expected payoff},
-which the mean payoff an individual achieved over the different types in the
-population after multiple intercalations. Expected payoffs imply that
-individuals have a perfect a memory. In order to estimate expected payoffs
-individuals must be able to recall several of their interactions with each
-player in the population. However, when modeling how they make decisions in each
-round they assumed to have very limited memory. To be precise, most work on this
-models focuses on naive subjects who can only choose from a restricted set of
-strategies~\cite{Nowak1992tit}, or who do not remember anything beyond the
-outcome of the very last round~\cite{Baek2016}, with a few notable
-exceptions~\cite{Hauert1997, Stewart2016}.
-
-This creates a curious inconsistency, which raises the following question: how
-robust is our understanding of cooperation? We are not the first to question the
-assumptions of models when estimating fitness. In~\cite{Roca2006} relax the
-assumption that selection occurs much more slowly than the interaction between
-individuals. They results shows that rapid selection affects evolutionary
-dynamics in such a dramatic way that for some games it even changes the
-stability of equilibria.
-
-We consider two cases. Initially, we start by considering two extreme scenarios.
-The first is the classical scenario of the expected payoffs and the alternative
-scenario where individuals update their strategies only based on the very last
-payoff they obtained. We refer to these as the \textit{stochastic} payoffs of
-individuals. We do this for the donation games but also the important cases of
-symmetric \(2 \times 2\) games. These include the prisoner's dilemma, the
-snowdrift game, the stag hunt game and the harmony game.
-
-In the later sections we allow individuals to use more memory. More specifically,
-individuals update their strategies by considering up to the last two rounds,
-whilst interacting with up to two different players.
+evolution of cooperation. Why is it that some individuals choose to help others,
+increasing their payoff, at the expense of decreasing one's own payoff? In
+evolutionary game theory individuals are not required to be rational, instead
+they adapt strategies based on mutation and exploration. Strategies are more
+likely to spread if they have a high fitness either because the individuals who
+adopt them have more offsprings, or because they are imitated more often. The
+fitness of a strategy is not constant but depends on the composition of the
+population. Individuals interact based on their strategies with other
+members of the population and the payoffs they yield are translated into
+fitness.
+
+It is commonly assumed that fitness is equivalent to the mean payoff an
+individual achieves over the different types in the population, over multiple
+interactions. These payoffs assume that individuals can interact with the entire
+population several times and remember each and every outcome. Thus, they imply
+that individuals have a perfect memory.
+However, in their pairwise interactions when they make decisions in each turn
+they are assumed to have very limited memory. To be precise most of
+the works in the literature,  with a few notable exceptions~\cite{Hauert1997,
+Stewart2016}, focus on naive subjects who can only choose from a restricted set
+of strategies~\cite{Nowak1992tit}, or who do not remember anything beyond the
+outcome of the very last round~\cite{Baek2016}.
+
+The perfect memory assumption is not only unrealistic but it also creates this
+curious inconsistency. This has lead us to question how robust is our
+understanding of cooperation. In this work we explore whether direct
+reciprocity can evolve if individuals only remember a minimum of social
+information. Though we are not the first to question the assumptions of
+estimating fitness~\cite{Roca2006}, we are the first to explore the effect
+of payoff memory.
+
+Initially, we consider two extreme scenarios. The first is the classical
+scenario of the expected payoffs and the alternative scenario where individuals
+update their strategies only based on the very last payoff they obtained. In the
+later sections, we allow individuals more memory. More specifically, they can
+remember up to two turns and up to two interactions.
+We present results on  several social dilemmas which include the prisoner's
+dilemma and the donation game, the snowdrift game, the stag hunt game and the
+harmony game.
+
+The remainder of the paper is organized as follows. Section~\ref{section:model}
+describes the model. Section~\ref{section:results} presents the results of the
+simulations. Finally, section~\ref{section:conclusions} outlines the main
+conclusions.
 
@@
 
-In order to account for the difference in the robustness of cooperation among
-individuals based on their payoff memory, we consider a population of \(N\)
-players, where \(N\) is even, and where mutations are sufficiently rare. At any
-point in time the there are at most two different strategies are present in the
-population. A \textit{resident} strategy and a \textit{mutant} strategy. Suppose
-there are \(N - k\) players who use the resident strategy whereas \(k\) players
-use the mutant strategy. Each step of the evolutionary process consists of two
-stages, a game stage and an updating stage.
-
-In the game stage, each player is randomly matched with some other player in
-the population to interact in a number of turns.  The number of turns is
-not fixed, on the contrast we consider that another interaction has a
-probability $\delta$ to continue following each turn. At each turn the players can
-choose to either cooperate (C) or to defect (D). The payoffs in a given round
-depend on both player's decisions and are given by:
-
-\begin{equation*}\label{eq:game}
-  U = \begin{pmatrix} R & S  \\ T & P \end{pmatrix}
-\end{equation*}
-
-We assume herein that
-individuals at most make use of simple \textit{reactive strategies} make
-decisions in each round. Reactive strategy are a set of memory-one strategies
-that only take into account the previous action of the opponent. Reactive
-strategies can be written explicitly as a vector \(\in \R_{3}\). More
-specifically a reactive strategy \(s\) is given by \(s=(y, p, q)\) where \(y\)
-is the probability that the strategy opens with a cooperation and \(p, q\) are
-the probabilities that the strategy cooperates given that the opponent
-cooperated and defected equivalently.
+We consider a population of \(N\) players (\(N\) is even) where mutations are
+sufficiently rare. Thus, at any point in time there are at most two
+different strategies present in the population; a \textit{resident} strategy
+and a \textit{mutant} strategy. We assume a pairwise process where
+strategies spread because they are imitated more often.
+
+Each step of the evolutionary process consists of two stages, a game stage and
+an updating stage. In the game stage each individual is randomly matched with
+some other individual in the population to interact for a number of turns, where
+subsequent turns occur with a fixed probability $\delta$. At each turn they
+can choose to either cooperate (\(C\)) or to defect (\(D\)), and thus, at
+each turn the possible outcomes are \(CC, CD, DC\) and \(DD\). The payoffs
+depend on the outcome. If both cooperate they receive the reward payoff
+\(R\), whereas if both defect they receive the punishment payoff \(P\).
+If one cooperates but the other defects, the defector receives the temptation to
+defect, \(T\), whereas the cooperator receives the sucker's payoff, \(S\).
+We denote the payoffs of an individual as \(\mathcal{U} = (R, S, T, P)\).
+
+We assume herein that individuals use \textit{reactive strategies} to make
+decisions in each turn. Reactive strategies are a set of memory-one strategies
+that only take into account the previous action of the opponent. They can be
+written explicitly as a vector \(\in \R_{3}\), more specifically, a reactive
+strategy \(s\) is given by \(s=(y, p, q)\) where \(y\) is the probability that
+the strategy opens with a cooperation and \(p, q\) are the probabilities that
+the strategy cooperates given that the opponent cooperated and defected
+equivalently.
 
@@
 assume the learner adopts the role model's strategy based on the Fermi
-distribution function. The relative influence of the payoffs on the
-adopt the strategy of the other is controlled by an external parameter. The so
-called intensity of selection,  $\beta\!\ge\!0$.
+distribution function,
+
+\begin{equation} \label{Eq:rho}
+  \rho(u_{L}, u_{RM}) = \frac{1}{1\!+\! \exp^{\!-\!\beta (u_{RM}\!-\!u_{L})}}.
+\end{equation}
+
+where $\beta\!\ge\!0$ is the relative influence of the payoffs on adopting the 
+strategy of the other. We refer to $\beta$ as the intensity of selection.
 
 This basic evolutionary step is repeated until either the mutant strategy goes
-extinct, or until it fixes in the population, in which case the mutant strategy
-becomes the new resident strategy. After either outcome, we introduce a new
-mutant strategy, uniformly chosen from all reactive strategies at random, and we
-set the number of mutants to $k\!=\!1$. This process of mutation and
-fixation/extinction is then iterated many times.
-
-In this work we explore the effect on the payoffs in the updating stage. The
-details of the process described in this section can be found in
-appendix~\ref{appendix:methods}.
-
-\section{Analysis of Stochastic Payoffs in the Donation game}\label{section:donation}
-
-We first explore the effect of stochastic payoffs on the cooperative behaviour
-by considering the extreme case that an individual's fitness is based on their
-last interaction with one other individual. We compare this to the expected
-payoffs. In this section we consider the donation game.
-Each player can cooperate by providing a benefit \(b\) to the other
-player at their cost \(c\), with \(0 < c < b\). Then, \(T=b, R=b-c, S=-c, P=0\),
-and matrix~(\ref{eq:game}) is give by:
-
-\begin{equation}\label{eq:donation_normal_form}
-    \begin{pmatrix}
-        b-c & -c  \\
-        b & 0
-    \end{pmatrix}
-\end{equation}
+extinct, or until it fixes in the population. If the mutant fixes in the
+population then the mutant strategy becomes the new resident strategy. After
+either outcome we introduce a new mutant strategy uniformly chosen from all
+reactive strategies at random, and we set the number of mutants to $1$. This
+process of mutation and fixation/extinction is then iterated many times.
+
+The perfect memory assumption occurs at the updating stage. The learner and the
+role model are assumed to interact with a representative sample of the
+population, and they remember all interactions they participate in. Thus, they
+are updating payoffs are based on the mean payoff they achieved over all the
+interactions. These payoffs are refer to as the expected payoffs. We will
+compare the expected payoffs to payoffs that are calculated when the role model
+and learner do not remember all their interactions. In order to account for the
+effect of these different methods we explore the cooperation rate within the
+resident population over multiple generations. More details on our methodology
+are found in Appendix~\ref{appendix:methods}.
+
+\section{Results}\label{section:results}
+\subsection{Updating payoffs based on the last round with another member of the population}\label{section:donation}
+
+In this section we explore the case where the updating payoffs are based on the
+last round payoff achieved against another member of the population. We compare
+this to the expected payoffs. We assume that individuals interact in a donation
+game  where each can cooperate by providing a benefit \(b\) to the other
+player at their cost \(c\), with \(0 < c < b\). Thus, \(T=b, R=b-c, S=-c, P=0\).
 
@@
 Figure~\ref{fig:expected_and_stochastic_for_donation}~depicts the evolving
-conditional cooperation probabilities $p$ and $q$. Assuming that the discount
-factor~$\delta$ is comparably high, the opening move \(y\) is a transient effect
-and has not effect on the outcome.
-The left panels corresponds to the standard scenario considered in the previous
-literature. It considers players who use expected payoffs to update their
-strategies. The right panel shows the scenario considered herein, in which
-players update their strategies based on their last round's payoff. The top
-panels assume a benefit of 3, whereas the bottom assume that the benefit is 10.
+conditional cooperation probabilities $p$ and $q$. The discount factor~$\delta$
+is comparably high, thus the opening move \(y\) is a transient effect and has
+not effect on the outcome. The left panels corresponds to the standard scenario
+considered in the literature. It considers players who use expected payoffs to
+update their strategies. The right panel shows the scenario considered herein,
+in which players update their strategies based on their last round's payoff. The
+top panels assume a benefit \(b\) of 3, whereas the bottom assume a benefit of
+10.
 
 The figure suggests that when updating is based on expected payoffs, players
-tend to be more generous. The $q$-values are higher on average which suggests
-that individuals tend to cooperating more after they are at the receiving end
-of a defection. In addition, for both cases of \(b\), individuals tend to
-be more cooperative. The average cooperation rate is strictly higher for expected
-payoffs. This difference is statistically important, abd the effect is even more
-obvious when the benefit is higher. More specifically, for \(b=10\) the average
-cooperation rate drops from 97\% to 51\%.
+tend to be more generous and more cooperative. The $q$-values are higher on
+average which suggests that individuals tend to be more generous when it comes
+to cooperating after being at the receiving end of a defection. %This is a known result no? Nowak 1992
+The average cooperation rate within the resident population is strictly higher
+when expected payoffs are considered. The difference between the two methods for
+both the values of \(b\) are statistically significant. This contrast becomes
+more obvious for \(b=10\). More specifically the average cooperation drops from
+97\% to 51\%. The residents of the population cooperated on average 97\% of the
+times in one case and only 57\% in the other.
 
@@
     \includegraphics[width=.75\textwidth]{static/expected_and_stochastic_for_donation_game_b_10.pdf}
-    \caption{{\bf Evolutionary dynamics under expected payoffs and stochastic payoffs.} 
+    \caption{{\bf Evolutionary dynamics under expected payoffs and last round with one interaction payoffs.} 
     We have run two simulations of the evolutionary process described in
-    Section~\ref{section:methods} for $T\!=\!10^7$ time steps. For each time step,
+    section~\ref{section:model} for $T\!=\!10^7$ time steps. For each time step,
     we have recorded the current resident population ($y,p,q$). Since simulations
@@
     typically applies a strategy for which $p\!\approx\!1$ and
-    $q\!\le\!1\!-\!c/b\!=\!0.9$. The cooperation rate within the resident population
-    (averaged over all games and over all time steps) is close to 100\%. ({\bf B})
+    $q\!\le\!1\!-\!c/b\!=\!0.9$. ({\bf B})
     When players update their strategies based on their realized payoffs in the last
@@
     with stable cooperation is somewhat smaller compared to the expected-payoff
-    setting, $q\!<\!0.5$. Also the resulting cooperation rate is smaller. On
-    average, players cooperate roughly in half of all rounds. Parameters:
-    $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.999$.}
+    setting, $q\!<\!0.5$. The cooperation rate within the resident population
+    (averaged over all games and over all time steps) is close to 100\%. Parameters:
+    $N\!=\!100$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.999$.}
     \label{fig:expected_and_stochastic_for_donation}
@@
 
-Figure~\ref{fig:cooperation_rate_over_benefit} further explores the effect of the
-benefit on the cooperation rate. In all cases, the stochastic payoffs evolution
-tends to reduce the evolving cooperation rate. For the stochastic payoffs,
-benefit appears to not make a difference once \(b > 5\), and on average
-the evolving cooperation rate is at 50\%. The effect of benefit on the expected
-payoffs smooths out after \(b > 6\), and the expected payoffs estimating
-the evolving rate to be at 90\%.
+We further explore the effect of the benefit,
+Figure~\ref{fig:cooperation_rate_over_benefit}. The figure suggests that the
+expected payoffs always overestimate cooperation. For the limited memory payoffs
+the cooperation rate remains unchanged at approximately 50\% once \(b = 5\). In
+specific simulations we have performed, if only the last round payoff is
+considered then the highest average cooperation rate that was achieved was 0.51.
+In comparison, when the expected payoffs were used the highest average
+cooperation rate was 0.97.
+
+\begin{figure}[!htbp]
+  \centering
+  \begin{subfigure}{.5\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{static/expected_for_beta.pdf}
+    \includegraphics[width=\textwidth]{static/stochastic_for_beta.pdf}
+  \end{subfigure}%
+  \begin{subfigure}{.5\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{static/cooperation_rate_over_b.pdf}
+  \end{subfigure}
+  \caption{{\bf The evolution of cooperation for different benefit values.} 
+  We vary the benefit of defection $b$. In all cases, expected payoffs appear to
+  overestimate the average cooperation rate the population achieves. ({\bf A})
+  the probabilities \(p, q\) for resident population over \(10^7\) time steps
+  for each benefit value. ({\bf B}) The cooperation rate within the resident population
+  (averaged over all games and over all time steps) over the benefit.
+  Unless
+  explicitly varied, the parameters of the simulation are $N\!=\!100$,
+  $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run
+  for $T\!=\!5\times 10^6$ time steps for each parameter
+  combination.}\label{fig:cooperation_rate_over_benefit}
+\end{figure}
+
+Figure~\ref{fig:cooperation_rate_over_betas} illustrates results for various
+runs of the evolutionary process where we vary the strength of selection. For
+weak selection, \(\beta < 1\), the updating payoffs have no effect. The evolved
+population and the average cooperating rate are the same for both approaches.
+However for strong selection, it can be seen that the cooperating rate increases
+with \(beta\), when expected payoffs are used. However, when limited memory
+payoffs are used. the cooperating rate decreases.
 
@@
   \end{subfigure}
-\caption{{\bf The evolution of cooperation for different benefit values.} 
-Here, we vary the benefit of defection $b$. In all cases, stochastic payoff
-evaluation tends to reduce the evolving cooperation rates. Unless explicitly
-varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$, $c\!=\!1$,
-$\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for $T\!=\!5\times 10^6$
-time steps for each parameter
-combination.}\label{fig:cooperation_rate_over_benefit}
-\end{figure}
-
-Figure~\ref{fig:cooperation_rate_over_betas} illustrates the results for
-different runs of the evolutionary process where we vary the strength of
-selection. In the case of a very small \(\beta < \beta=10 ^ -1\) he process is
-almost random, as the effect of the payoffs is very small. Once payoff begin to
-matter the difference is once again evident, with the expected payoffs always
-overestimating the evolved cooperation rate.
-
-\begin{figure}[!htbp]
-  \centering
-  \begin{subfigure}{.5\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{static/expected_for_beta.pdf}
-    \includegraphics[width=\textwidth]{static/stochastic_for_beta.pdf}
-  \end{subfigure}%
-  \begin{subfigure}{.5\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{static/cooperation_rate_over_b.pdf}
-  \end{subfigure}
 \caption{{\bf The evolution of cooperation for different selection strength values.} 
-Here, we vary the selection strength $\beta$. In all cases, stochastic payoff
-evaluation tends to reduce the evolving cooperation rates. Unless explicitly
-varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$, $c\!=\!1$,
-$\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for $T\!=\!5\times 10^6$
-time steps for each parameter
+We vary the selection strength $\beta$. In all cases, stochastic payoff
+evaluation tends to reduce the evolving cooperation rates. ({\bf A}) the
+probabilities \(p, q\) for resident population over \(10^7\) time steps for each
+\(\beta\) value. ({\bf B}) The cooperation rate within the resident population
+(averaged over all games and over all time steps) over \(\beta\).
+Unless explicitly varied, the parameters of the simulation are $N\!=\!100$,
+$b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
+$T\!=\!5\times 10^6$ time steps for each parameter
 combination.}\label{fig:cooperation_rate_over_betas}
@@
 
-\section{Expected and stochastic payoffs in $2 \times 2$ games}\label{section:2_by_2_games}
+\subsection{Expected and stochastic payoffs in $2 \times 2$ games}\label{section:2_by_2_games}
+
+\section{Conclusions}\label{section:conclusions}
 

### 2021-05-05 main.tex
going over appendix

@@
 The perfect memory assumption is not only unrealistic but it also creates this
-curious inconsistency. This has lead us to question how robust is our
+curious inconsistency. This has led us to question how robust is our
 understanding of cooperation. In this work we explore whether direct
@@
 
-We consider a pairwise comparison process. We assume a population of \(N\)
-individuals (\(N\) is even) where at any point in time there are at most two
-different strategies in present. There are \(k\) individuals who play the mutant
-strategy and \(N - k\) individuals who play the resident strategy. Each step of
-the evolutionary process consists of two stages; a game stage and an update
-stage.
+Consider a population of \(N\) individuals where \(N\) is even. At any point in
+time there are at most two different strategies in present in the population.
+More specifically, a mutant strategy played by \(k\) individuals and a resident
+strategy played by \(N - k\) individuals. We assume a pairwise process in which
+strategies spread because they are imitated more often. Each step of the
+evolutionary process consists of two stages; a game stage and an update stage.
 
@@
 which is not fixed but depends on the continuation probability \(\delta\). At
-each turn the individuals have choose between cooperation (C) and defection (D).
-If both players cooperate they receive the reward payoff \(R\), whereas if both
-players defect they receive the punishment payoff \(P\). If one cooperates but
-the other defects, the defector receives the temptation to defect, \(T\),
-whereas the cooperator receives the sucker's payoff, \(S\).
-
-In this work we consider different social dilemmas represented by the harmony
-game, the stag hunt game, the snowdrift or hawk-dove game, the prisoner's
-dilemma, and a special case of the prisoner's dilemma, the donation game
-(Table~\ref{table:social_dilemmas}).
+each turn the individuals choose between cooperation (\(C\)) and defection
+(\(D\)). Thus, there are four possible outcomes in each turn \(CC, CD, DC\) and
+\(DD\). If both players cooperate they receive the reward payoff \(R\), whereas
+if both players defect they receive the punishment payoff \(P\). If one
+cooperates but the other defects, the defector receives the temptation to
+defect, \(T\), whereas the cooperator receives the sucker's payoff, \(S\).
+Let $\mathcal{U}=\{R,S,T,P\}$ denote the set of feasible payoffs in each round,
+and let $\mathbf{u}\!=\!(R,S,T,P)$ be the corresponding payoff vector.
+
+The values of the payoffs are not only based on the prisoner's dilemma but all
+the symmetric \(2 \times 2\) games, Table~\ref{table:social_dilemmas}.
 
@@
   \end{center}
-  \caption{We present results on the all of the listed social dilemmas. The payoffs'
-  constrains are different for each game.}
+  \caption{\textbf{Social dilemmas and payoffs' constrains}. The various social
+  dilemmas we explore in this work. Results for cases (i) - (iv) were presented
+  in section~\ref{section:2_by_2_games} and results for case (v) were presented
+  in section~\ref{section:donation}.}
   \label{table:social_dilemmas}
@@
 
-Results for cases (i) - (iv) were presented in section~\ref{section:2_by_2_games}
-and results for case (v) were presented in section~\ref{section:donation}.
-
 A further assumption of our model is that individuals make use of reactive
-strategies when they make decisions in each round. Reactive strategy are a set
+strategies when they make decisions in each round. Reactive strategies are a set
 of strategies that take into account only the previous action of the opponent.
@@
 strategy which is uniformly chosen from all reactive strategies at random.
-
-This process of mutation and fixation/extinction is iterated many times.
-Instead of simulating each step of the evolutionary process, we estimate
-the probability that a newly introduced mutant fixes~\cite{nowak2004emergence}.
-This is defined as the fixation probability of the mutant, and the standard
-form is the following,
+Instead of simulating each step of the evolutionary process, we estimate the
+probability that a newly introduced mutant fixes~\cite{nowak2004emergence}. This
+is defined as the fixation probability of the mutant, and the standard form is
+the following,
 
@@
 
-This evolutionary process is summarized by Algorithm~\ref{algorithm:pairwise_comparison}.
+This process of mutation and fixation/extinction is iterated many times. The
+evolutionary process is summarized by
+Algorithm~\ref{algorithm:pairwise_comparison}.
 
@@
      }}
-     \caption{Pairwise comparison process}\label{algorithm:pairwise_comparison}
+     \caption{Evolutionary process}\label{algorithm:pairwise_comparison}
 \end{algorithm}
@@
 different approaches when estimating the payoffs at the updating stage. The two
-approaches we consider are those of (i) the expected payoffs and (ii) the
-stochastic payoffs.
+approaches we consider are those of (i) the expected and (ii) the limited memory
+payoffs.
 
@@
 The expected payoffs are the conventional payoffs used in the updating
-stage~\cite{imhof2010stochastic}. They are defined as the mean payoff an
+stage~\cite{imhof2010stochastic}. They are defined as the mean payoff of an
 individual in a well-mixed population that engages in repeated games with all
@@
 two reactive strategies $s_1\!=\!(y_1, p_1, q_1$) and $s_2\!=\!(y_2,p_2,q_2)$.
-It is not necessary to simulate the play move by move. Instead the play between
+It is not necessary to simulate the play move by move, instead the play between
 the two strategies is defined a Markov matrix \(M\),
@@
 
-whose stationary vector \(\mathbf{v}\), combined with the payoff \(U\), yields
-the expected outcome for each strategy,
-\(\langle\mathbf{v}(s_1,s_2),\mathbf{u}\rangle\).
-
-% In each step
-Given a population of \(N\) individual with \(k\) individuals playing the mutant
-strategy, \(s_M = (y_M, p_M, q_M)\), and \(N -k\) playing the resident strategy,
-\(s_R = (y_R, p_R, q_R)\), the expected payoffs of a resident \(\pi_R\) and of
-a mutant strategy \(\pi_M\) are defined as,
+whose stationary vector \(\mathbf{v}\), combined with the payoff \(u\), yields
+the game stage outcome for each strategy,
+\(\langle\mathbf{v}(s_1,s_2),\mathbf{u}\rangle\)~\cite{Hauert1997}.
+
+In the updating stage the learner adopts the strategy of the role model based on
+their updating payoffs. Given that there are only two different types in the
+population at each time step we only need to define the expected payoff for a
+resident (\(\pi_R\)) and for a mutant (\(\pi_M\)). Assume the resident strategy
+\(s_R = (y_R, p_R, q_R)\) and the mutant strategy \(s_M = (y_M, p_M, q_M)\), the
+expected payoffs are give by,
 
@@
   \begin{array}{lcrcr}
-  \displaystyle \pi_R	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(s_R,s_R),\mathbf{U}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(s_R,s_M),\mathbf{U}\rangle,\\[0.5cm]
-  \displaystyle \pi_M	&=	&\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(s_M,s_R),\mathbf{U}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(s_M,s_M),\mathbf{U}\rangle.\\
+  \displaystyle \pi_R	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(s_R,s_R),\mathbf{u}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(s_R,s_M),\mathbf{u}\rangle,\\[0.5cm]
+  \displaystyle \pi_M	&=	&\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(s_M,s_R),\mathbf{u}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(s_M,s_M),\mathbf{u}\rangle.\\
   \end{array}
@@
 
-Note that \(N\!-\!k\!-\!1\) is the number of other residents in the population,
-and \(k\) is the number of residents. The payoffs are averaged by the number of
-individuals in the population, given that interactions are not possible.
-The same applies for the payoff of a mutant.
-
-Given the expected payoff, the probabilities that the number of mutants
-decreases and increases  (Equation~(\ref{eq:appendix_fixation_probability})),
-are given by:
+The number of mutant in the population increase if a learner resident adopts the
+strategy of a mutant role model, and decreases if a mutant leaner adopts the
+strategy of a resident. The probabilities that the number of mutants decreases
+and increases, \(\lambda^-_k\) and \(\lambda^+_k\), are not explicitly define
+as,
 
@@
 
-\subsection*{Stochastic Payoffs}
-
-We compare the expected payoffs with the finite memory payoffs which we refer to
-as the stochastic payoffs. Initially, we consider the case of where the payoffs
-correspond to the last round payoff an individual achieved after interacting
-with one other individual.
-
-The probability of being in any of the outcomes \(R, S, T, P\) in the last round
-is given by Equation~(\ref{Eq:LastRound})
-(Proposition~\ref{proposition:last_round}).
-
-\begin{Prop}\label{proposition:last_round}
-Assume a play between the reactive strategies \(s_1\) and \(s_2\) with a
-continuation probability \(\delta\). Then the probability of being in any of the
-four outcomes \(R, S, T, P\) are given by,
+\subsection*{Limited memory payoffs}
+
+Initially, we discuss the case of the last round updating payoff. At the stage
+game we define the payoff of a reactive strategy in the last round,
+Proposition~\ref{proposition:last_round}.
+
+\begin{Prop}
+    Consider a repeated prisoner's dilemma, with
+    continuation probability $\delta$, between players with reactive strategies
+    $s_1\!=\!(y_1, p_1, q_1$)  and $s_2\!=\!(y_2,p_2,q_2)$ respectively. Then the
+    probability that the $s_1$ player receives the payoff $u\!\in\! \mathcal{U}$ in
+    the very last round of the game is given by $v_{u}(s_1,s_2)$, as given by
+    Equation~(\ref{Eq:LastRound}).
 
@@
 
-  \mathbf{v}_{R}(S_1,S_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
-  {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
-
-  \mathbf{v}_{S}(S_1,S_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
-  {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
-
-  \mathbf{v}_{T}(S_1,S_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
-  {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
-
-  \mathbf{v}_{P}(S_1,S_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
-  {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)}.
+      v_{R}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
+      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)} \times R,\\[1cm]
+
+      v_{S}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
+      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)} \times S,\\[1cm]
+
+      v_{T}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
+      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)} \times T,\\[1cm]
+
+      v_{P}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
+      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)} \times P.
       \end{array}
@@
 
-In these expressions, we have used
-the notation $r_i:=p_i\!-\!q_i$, $\bar{y}_i\!=\!1\!-\!y_i$, $\bar{q}_i:=1\!-\!q_i$,
-and $\bar{r}_i:=\bar{p}_i\!-\!\bar{q}_i=-r_i$ for $i\!\in\!\{1,2\}$.
-Let $\mathcal{U}=\{R,S,T,P\}$ denote the set of feasible payoffs in each round,
-and let $\mathbf{u}\!=\!(R,S,T,P)$ be the corresponding payoff vector.
+In these expressions, we have used the notation $r_i:=p_i\!-\!q_i$,
+$\bar{y}_i\!=\!1\!-\!y_i$, $\bar{q}_i:=1\!-\!q_i$, and
+$\bar{r}_i:=\bar{p}_i\!-\!\bar{q}_i=-r_i$ for $i\!\in\!\{1,2\}$.
 \end{Prop}
@@
 Given a play between two reactive strategies with continuation probability
-$\delta$. The probability that the game ends on round \(t\) is given by,
+$\delta$. The outcome at turn \(t\) is given by,
 
 \begin{equation}\label{eq:}
-  (1 - \delta) \mathbf{v_0} \sum \delta^{t} M^{(t)}.
+  (1 - \delta) \mathbf{v_0} \sum \delta^{t} M^{(t)},
 \end{equation}
 
-Here, $\mathbf{v_0}$ denotes the expected distribution of the four outcomes in
+where $\mathbf{v_0}$ denotes the expected distribution of the four outcomes in
 the very first round, and \(1- \delta\) the probability that the game ends.
@@
 
-and with some algebraic manipulation we derive to Equation~\ref{Eq:LastRound}.
+where \((1 - \delta)\mathbf{v_0}(I_4 - \delta M)^{-1}\) is vector \(\in R^{4}\)
+and it the probabilities for being in any of the outcomes \(CC, CD, DC, DD\) in
+the last round. Combining this with the payoff vector \(u\) and some algebraic
+manipulation we derive to the Equation~\ref{Eq:LastRound}.
 \end{proof}
 
-Considering that individuals can only interact with one other individual,
-at that in each step a mutant and a resident are selected there are only five
-possible pairings:
-
-\begin{itemize}
-  \item the selected resident is paired with the selected mutant
-  \item the selected resident is paired with a resident and the selected mutant is paired with a resident
-  \item the selected resident is paired with a resident and the selected mutant is paired with a mutant
-  \item the selected resident is paired with a mutant and the selected mutant is paired with a resident
-  \item the selected resident is paired with a mutant and the selected mutant is paired with a mutant
-\end{itemize}
-
-Each of the pairings happens with a given probability and that is illustrated
-by Figure~\ref{fig:single_pairs}.
+In the updating stage we select a mutant and resident to be either the role
+model or the learner. Assume the selected mutant. Given that they can interact
+with only one other member of the population, they can interact either the
+selected resident, another resident or with another mutant. The same is true for
+the the selected resident. They can interact with the selected mutant, another
+resident, or another mutant. Thus, in each updating stage there are five
+possible combinations of pairs. These are illustrated by
+Figure~\ref{fig:single_pairs}.
 
@@
   \includegraphics[width=.65\textwidth]{static/last_rounds_matches.tex}
-  \caption{The tree}\label{fig:single_pairs}
+  \caption{\textbf{Possible pairings combination in the updating stage, given
+  that individuals interact with only one other member in the population.}. We
+  distinguish between the selected resident and the rest of the residents and we
+  do the same with the mutants. There is a probability that the selected
+  resident interacts with the selected mutant.}\label{fig:single_pairs}
 \end{figure}
@@
 
-% In these expressions, we have used the notation $r_i:=p_i\!-\!q_i$, $\bar{y}_i\!=\!1\!-\!y_i$, $\bar{q}_i:=1\!-\!q_i$, and $\bar{r}_i:=\bar{p}_i\!-\!\bar{q}_i=-r_i$ for $i\!\in\!\{1,2\}$.
-Let $\mathcal{U}=\{R,S,T,P\}$ denote the set of feasible payoffs in each round, and let $\mathbf{u}\!=\!(R,S,T,P)$ be the corresponding payoff vector.
-Then one can show the following result.
+% In these expressions, we have used the notation $r_i:=p_i\!-\!q_i$,
+% $\bar{y}_i\!=\!1\!-\!y_i$, $\bar{q}_i:=1\!-\!q_i$, and
+% $\bar{r}_i:=\bar{p}_i\!-\!\bar{q}_i=-r_i$ for $i\!\in\!\{1,2\}$. Let
+% $\mathcal{U}=\{R,S,T,P\}$ denote the set of feasible payoffs in each round,
+% and let $\mathbf{u}\!=\!(R,S,T,P)$ be the corresponding payoff vector. Then
+% one can show the following result.
 

### 2021-05-05 static/last_rounds_matches.tex
going over appendix

@@
 \node [arn_n] {} 
-    child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{M}$\\$S_{R} - S_{R}$}] {} edge from parent node[above left, rotate=30] {$\frac{1}{N - 1}$}}
+    child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{M}$] {} edge from parent node[above left, rotate=30] {$\frac{1}{N - 1}$}}
     child{ node [arn_n, rotate=-30] {} 

### 2021-05-05 static/tree.tex
going over appendix

@@
-\documentclass{article}
-\usepackage{tikz}
-\usetikzlibrary{trees}
-\tikzstyle{densely dashdotted}=      [dash pattern=on 3pt off 1pt on \the\pgflinewidth off 1pt]
-\tikzset{
-  treenode/.style = {shape=circle, inner sep=1pt,
-draw, align=center, fill, black
-%top color=white, bottom color=blue!20
-  },
-  payoff/.style    = {align=center, inner sep=0.1em, text width=1.5em},
-  left side node/.style={above left, inner sep=0.1em},
-  right side node/.style={above right, inner sep=0.1em}
-}
+\documentclass{standalone}
+% \usetikzlibrary{trees}
+% \tikzstyle{densely dashdotted}=      [dash pattern=on 3pt off 1pt on \the\pgflinewidth off 1pt]
+% \tikzset{
+%   treenode/.style = {shape=circle, inner sep=1pt,
+% draw, align=center, fill, black
+% %top color=white, bottom color=blue!20
+%   },
+%   payoff/.style    = {align=center, inner sep=0.1em, text width=1.5em},
+%   left side node/.style={above left, inner sep=0.1em},
+%   right side node/.style={above right, inner sep=0.1em}
+% }
+
+% \begin{document}
+% \begin{tikzpicture}
+%   [
+% %grow                    = right,
+% sibling distance        = 25em,
+% level 2/.style          = {sibling distance=2cm},
+% level 3/.style          = {sibling distance=3.5cm},
+% level 4/.style          = {sibling distance=2cm},
+% level distance          = 6em,
+% every node/.style       = {font=\footnotesize},
+% %sloped
+%   ]
+%   \node [treenode] {} child{node[payoff] {$S_{L}, S_{T}$}
+%   child{node [payoff] {\\ $S_{L}, S_{M} - S_{T}, S_{M}$} edge from parent[densely dashdotted] node[left side node, rotate=35] {$\frac{k-1}{N-2}\frac{k-2}{N-3}$}}
+%   child{node [payoff] {\\ $S_{L}, S_{R} - S_{T}, S_{M}$} edge from parent[densely dashdotted] node[left side node, rotate=45] {$\frac{k-1}{N-2}\frac{N-k-1}{N-3}$}}
+%   child{ node [payoff] {\\ $S_{L}, S_{M} - S_{T}, S_{R}$} edge from parent[densely dashdotted] node[right side node, rotate=-45] {$\frac{N-k-1}{N-2}\frac{k-1}{N-3}$}}
+%   child{ node [payoff] {\\ $S_{L}, S_{R} - S_{T}, S_{R}$} edge from parent[densely dashdotted] node[right side node, rotate=-35, xshift=-.5cm] {$\frac{N-k-1}{N-2}\frac{N-k-2}{N-3}$}}
+%   edge from parent node[left side node] {$\frac{1}{1 - N}$}}
+%   child{node [treenode] {} 
+%   child{node [payoff] {\\ $S_{L}, S_{M} - S_{T}, S_{M}$}
+%   child{node [payoff] {\\ $S_{L}, S_{T}$} edge from parent[densely dashdotted] node[left side node] {$\frac{1}{N-3}$}}
+%   child{node [treenode] {} 
+%   child{node [payoff] {\\ $S_{L}, S_{M} -  S_{T}, S_{M}$} edge from parent[densely dashdotted] node[left side node, rotate=35] {$\frac{k-2}{N-3}\frac{k-3}{N-4}$}}
+%   child{node [payoff] {\\ $S_{L}, S_{R} -  S_{T}, S_{M}$} edge from parent[densely dashdotted] node[left side node, rotate=45] {$\frac{N-k-1}{N-3}\frac{k-2}{N-4}$}}
+%   child{ node [payoff] {\\ $S_{L}, S_{M} - S_{T}, S_{R}$} edge from parent[densely dashdotted] node[right side node, rotate=-45] {$\frac{k-2}{N-3}\frac{N-k-1}{N-4}$}}
+%   child{ node [payoff] {\\ $S_{L}, S_{R} - S_{T}, S_{R}$} edge from parent[densely dashdotted] node[right side node, rotate=-35, xshift=-.5cm] {$\frac{N-k-1}{N-3}\frac{N-k-2}{N-4}$}}
+%   edge from parent node[right side node] {$1 - \frac{1}{N-3}$}}
+%   edge from parent node[left side node, rotate=35] {$\frac{k-1}{N-2}\frac{k-2}{N-3}$}}
+%   child{node [payoff] {\\ $S_{L}, S_{R} -  S_{T}, S_{M}$} edge from parent[densely dashdotted] node[left side node, rotate=45] {$\frac{k-1}{N-2}\frac{N-k-1}{N-3}$}}
+%   child{ node [payoff] {\\ $S_{L}, S_{M} - S_{T}, S_{R}$} edge from parent[densely dashdotted] node[right side node, rotate=-45] {$\frac{N-k-1}{N-2}\frac{k-1}{N-3}$}}
+%   child{ node [payoff] {\\ $S_{L}, S_{R} - S_{T}, S_{R}$} edge from parent[densely dashdotted] node[right side node, rotate=-35, xshift=-.5cm] {$\frac{N-k-1}{N-2}\frac{N-k-2}{N-3}$}
+%   edge from parent node[right side node, xshift=-3cm, yshift=2.1cm] {$1 - \frac{1}{1 - N}$}}};
+% \end{tikzpicture}
+% \end{document}
 
 \begin{document}
-\begin{tikzpicture}
-  [
-%grow                    = right,
-sibling distance        = 25em,
-level 2/.style          = {sibling distance=2cm},
-level 3/.style          = {sibling distance=3.5cm},
-level 4/.style          = {sibling distance=2cm},
-level distance          = 6em,
-every node/.style       = {font=\footnotesize},
-%sloped
-  ]
-  \node [treenode] {} child{node[payoff] {$S_{L}, S_{T}$}
-  child{node [payoff] {\\ $S_{L}, S_{M} - S_{T}, S_{M}$} edge from parent[densely dashdotted] node[left side node, rotate=35] {$\frac{k-1}{N-2}\frac{k-2}{N-3}$}}
-  child{node [payoff] {\\ $S_{L}, S_{R} - S_{T}, S_{M}$} edge from parent[densely dashdotted] node[left side node, rotate=45] {$\frac{k-1}{N-2}\frac{N-k-1}{N-3}$}}
-  child{ node [payoff] {\\ $S_{L}, S_{M} - S_{T}, S_{R}$} edge from parent[densely dashdotted] node[right side node, rotate=-45] {$\frac{N-k-1}{N-2}\frac{k-1}{N-3}$}}
-  child{ node [payoff] {\\ $S_{L}, S_{R} - S_{T}, S_{R}$} edge from parent[densely dashdotted] node[right side node, rotate=-35, xshift=-.5cm] {$\frac{N-k-1}{N-2}\frac{N-k-2}{N-3}$}}
-  edge from parent node[left side node] {$\frac{1}{1 - N}$}}
-  child{node [treenode] {} 
-  child{node [payoff] {\\ $S_{L}, S_{M} - S_{T}, S_{M}$}
-  child{node [payoff] {\\ $S_{L}, S_{T}$} edge from parent[densely dashdotted] node[left side node] {$\frac{1}{N-3}$}}
-  child{node [treenode] {} 
-  child{node [payoff] {\\ $S_{L}, S_{M} -  S_{T}, S_{M}$} edge from parent[densely dashdotted] node[left side node, rotate=35] {$\frac{k-2}{N-3}\frac{k-3}{N-4}$}}
-  child{node [payoff] {\\ $S_{L}, S_{R} -  S_{T}, S_{M}$} edge from parent[densely dashdotted] node[left side node, rotate=45] {$\frac{N-k-1}{N-3}\frac{k-2}{N-4}$}}
-  child{ node [payoff] {\\ $S_{L}, S_{M} - S_{T}, S_{R}$} edge from parent[densely dashdotted] node[right side node, rotate=-45] {$\frac{k-2}{N-3}\frac{N-k-1}{N-4}$}}
-  child{ node [payoff] {\\ $S_{L}, S_{R} - S_{T}, S_{R}$} edge from parent[densely dashdotted] node[right side node, rotate=-35, xshift=-.5cm] {$\frac{N-k-1}{N-3}\frac{N-k-2}{N-4}$}}
-  edge from parent node[right side node] {$1 - \frac{1}{N-3}$}}
-  edge from parent node[left side node, rotate=35] {$\frac{k-1}{N-2}\frac{k-2}{N-3}$}}
-  child{node [payoff] {\\ $S_{L}, S_{R} -  S_{T}, S_{M}$} edge from parent[densely dashdotted] node[left side node, rotate=45] {$\frac{k-1}{N-2}\frac{N-k-1}{N-3}$}}
-  child{ node [payoff] {\\ $S_{L}, S_{M} - S_{T}, S_{R}$} edge from parent[densely dashdotted] node[right side node, rotate=-45] {$\frac{N-k-1}{N-2}\frac{k-1}{N-3}$}}
-  child{ node [payoff] {\\ $S_{L}, S_{R} - S_{T}, S_{R}$} edge from parent[densely dashdotted] node[right side node, rotate=-35, xshift=-.5cm] {$\frac{N-k-1}{N-2}\frac{N-k-2}{N-3}$}
-  edge from parent node[right side node, xshift=-3cm, yshift=2.1cm] {$1 - \frac{1}{1 - N}$}}};
+\begin{tikzpicture}[->,>=stealth',
+    level/.style={sibling distance = 7cm/#1,
+    level distance = 1.7cm},
+    level 2/.style = {sibling distance = 1.9cm, level distance=2.8cm},
+    level 3/.style = {sibling distance = 2cm, level distance=1.9cm},
+    level 4/.style = {sibling distance = 2cm, level distance=2.8cm},
+    every node/.style = {font=\tiny}]
+\node [arn_n] {} 
+    child{ node [arn_r, label={[align=center, yshift=-.8cm] $S_{R} - S_{R}$}] {} 
+    child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{R}$\\$S_{M} - S_{R}$}] {} edge from parent[densely dashdotted] node[above, yshift=-.1cm, rotate=39] {$\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$}}
+    child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{R}$\\$S_{R} - S_{M}$}] {} edge from parent[densely dashdotted] node[above, yshift=-.2cm, rotate=67] {$\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$}}
+    child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{M}$\\$S_{M} - S_{R}$}] {} edge from parent[densely dashdotted] node[above, yshift=-.2cm, rotate=-67] {$\frac{(k-1)(N-k-1)}{(N-2)(N-3)}$}}
+    child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{R}$\\$S_{R} - S_{R}$}] {} edge from parent[densely dashdotted] node[above, yshift=-.1cm, rotate=-41] {$\frac{(k-1)(k-2)}{(N-2)(N-3)}$}}
+    edge from parent node[above left, xshift=1em, yshift=-.8em, rotate=-30] {$1 - \frac{1}{N - 1}$}}
+    child{ node [arn_n, rotate=-30] {} 
+            child{ node [arn_r, label={[align=center, yshift=-.8cm] $S_{R} - S_{R}$\\$S_{M} - S_{R}$}] {} 
+            child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{M}$\\$S_{R} - S_{R}$}] {} edge from parent node[above left, rotate=30] {$\frac{1}{N - 1}$}}
+            child{ node [arn_n, rotate=-30] {} 
+                    child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{R}$\\$S_{M} - S_{R}$}] {} edge from parent[densely dashdotted] node[above, yshift=-.1cm, rotate=39] {$\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$}}
+                    child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{R}$\\$S_{R} - S_{M}$}] {} edge from parent[densely dashdotted] node[above, yshift=-.2cm, rotate=67] {$\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$}}
+                    child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{M}$\\$S_{M} - S_{R}$}] {} edge from parent[densely dashdotted] node[above, yshift=-.2cm, rotate=-67] {$\frac{(k-1)(N-k-1)}{(N-2)(N-3)}$}}
+                    child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{R}$\\$S_{R} - S_{R}$}] {} edge from parent[densely dashdotted] node[above, yshift=-.1cm, rotate=-41] {$\frac{(k-1)(k-2)}{(N-2)(N-3)}$}}
+                    edge from parent node[above left, xshift=1em, yshift=-.8em, rotate=-30] {$1 - \frac{1}{N - 1}$}
+            }
+            edge from parent node[above, yshift=-.1cm, rotate=39] {$\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$}}
+            child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{R}$\\$S_{R} - S_{M}$}] {} edge from parent node[above, yshift=-.2cm, rotate=67] {$\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$}}
+            child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{M}$\\$S_{M} - S_{R}$}] {} edge from parent node[above, yshift=-.2cm, rotate=-67] {$\frac{(k-1)(N-k-1)}{(N-2)(N-3)}$}}
+            child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{R}$\\$S_{R} - S_{R}$}] {} edge from parent node[above, yshift=-.1cm, rotate=-41] {$\frac{(k-1)(k-2)}{(N-2)(N-3)}$}}
+            edge from parent node[above left, xshift=1em, yshift=-.8em, rotate=-30] {$1 - \frac{1}{N - 1}$}
+		}; 
 \end{tikzpicture}

### 2021-05-05 main.tex
cleans typos in first parts

@@
 evolution of cooperation. Why is it that some individuals choose to help others,
-increasing their payoff, at the expense of decreasing one's own payoff? In
-evolutionary game theory individuals are not required to be rational, instead
-they adapt strategies based on mutation and exploration. Strategies are more
-likely to spread if they have a high fitness either because the individuals who
-adopt them have more offsprings, or because they are imitated more often. The
-fitness of a strategy is not constant but depends on the composition of the
-population. Individuals interact based on their strategies with other
-members of the population and the payoffs they yield are translated into
-fitness.
-
-It is commonly assumed that fitness is equivalent to the mean payoff an
-individual achieves over the different types in the population, over multiple
-interactions. These payoffs assume that individuals can interact with the entire
-population several times and remember each and every outcome. Thus, they imply
-that individuals have a perfect memory.
-However, in their pairwise interactions when they make decisions in each turn
-they are assumed to have very limited memory. To be precise most of
-the works in the literature,  with a few notable exceptions~\cite{Hauert1997,
-Stewart2016}, focus on naive subjects who can only choose from a restricted set
-of strategies~\cite{Nowak1992tit}, or who do not remember anything beyond the
-outcome of the very last round~\cite{Baek2016}.
+increasing their payoff, at the expense of decreasing one’s own? In evolutionary
+game theory, individuals are not required to be rational, instead they adapt
+strategies based on mutation and exploration. Strategies are more likely to
+spread if they have a high fitness either because the individuals who adopt them
+have more offspring, or because they are imitated more often. The fitness of a
+strategy is not constant but depends on the composition of the population.
+Individuals interact based on their strategies with other members of the
+population and the payoffs they yield are translated into fitness.
+
+It is commonly assumed that fitness is equivalent to the mean distribution of
+types in the population. These payoffs assume that individuals can interact with
+the entire population several times and remember each and every outcome. Thus,
+they imply that individuals have a perfect memory. However, when they make
+decisions in each turn they are assumed to have very limited memory. To be
+precise, most of the works in the literature focus on naive subjects who can
+only choose from a restricted set of strategies~\cite{Nowak1992tit}, or who do
+not remember anything beyond the outcome of the very last round~\cite{Baek2016}.
+Note that there are a few notable exceptions~\cite{Hauert1997, Stewart2016}.
 
@@
 curious inconsistency. This has led us to question how robust is our
-understanding of cooperation. In this work we explore whether direct
+understanding of cooperation. In this work, we explore whether direct
 reciprocity can evolve if individuals only remember a minimum of social
 information. Though we are not the first to question the assumptions of
-estimating fitness~\cite{Roca2006}, we are the first to explore the effect
-of payoff memory.
+estimating fitness~\cite{Roca2006}, we are the first to explore the effect of
+payoff memory.
 
@@
 later sections, we allow individuals more memory. More specifically, they can
-remember up to two turns and up to two interactions.
-We present results on  several social dilemmas which include the prisoner's
-dilemma and the donation game, the snowdrift game, the stag hunt game and the
-harmony game.
+remember up to two turns and up to two interactions. We present results on
+several social dilemmas which include the prisoner’s dilemma and the donation
+game, the snowdrift game, the stag hunt game and the harmony game.
 
@@
 role model are assumed to interact with a representative sample of the
-population, and they remember all interactions they participate in. Thus, they
-are updating payoffs are based on the mean payoff they achieved over all the
-interactions. These payoffs are refer to as the expected payoffs. We will
+population, and they remember all interactions they participate in. Thus, their
+updating payoffs are based on the mean payoff they achieved over all the
+interactions. These payoffs are referred to as the expected payoffs. We will
 compare the expected payoffs to payoffs that are calculated when the role model
-and learner do not remember all their interactions. In order to account for the
-effect of these different methods we explore the cooperation rate within the
-resident population over multiple generations. More details on our methodology
-are found in Appendix~\ref{appendix:methods}.
+and learner do not remember all of their interactions. In order to account for
+the effect of these different methods, we explore the cooperation rate within
+the resident population over multiple generations. More details on our
+methodology are found in Appendix~\ref{appendix:methods}.
 
@@
 conditional cooperation probabilities $p$ and $q$. The discount factor~$\delta$
-is comparably high, thus the opening move \(y\) is a transient effect and has
-not effect on the outcome. The left panels corresponds to the standard scenario
-considered in the literature. It considers players who use expected payoffs to
-update their strategies. The right panel shows the scenario considered herein,
-in which players update their strategies based on their last round's payoff. The
-top panels assume a benefit \(b\) of 3, whereas the bottom assume a benefit of
-10.
+is comparably high, thus the opening move \(y\) is a transient
+effect and has no effect on the outcome. The left panel corresponds to the
+standard scenario considered in the literature. It considers players who use
+expected payoffs to update their strategies. The right panel shows the scenario
+considered herein, in which players update their strategies based on their last
+round’s payoff. The top panels assume a benefit \(b\) of 3, whereas the bottom
+assume a benefit of 10.
 
@@
 Figure~\ref{fig:cooperation_rate_over_benefit}. The figure suggests that the
-expected payoffs always overestimate cooperation. For the limited memory payoffs
-the cooperation rate remains unchanged at approximately 50\% once \(b = 5\). In
-specific simulations we have performed, if only the last round payoff is
-considered then the highest average cooperation rate that was achieved was 0.51.
-In comparison, when the expected payoffs were used the highest average
-cooperation rate was 0.97.
+expected payoffs always overestimate cooperation. For the limited memory
+payoffs, the cooperation rate remains unchanged at approximately 50\% once \(b =
+5\). The highest cooperation rate that was achieved was 0.51 when  the last
+round payoff is considered. In comparison, when the expected payoffs were used
+the highest average cooperation rate was 0.97.
 
@@
 with \(beta\), when expected payoffs are used. However, when limited memory
-payoffs are used. the cooperating rate decreases.
+payoffs are used the cooperating rate decreases.
 

### 2021-05-05 main.tex
writting results part two

@@
 
-\subsection{Expected and stochastic payoffs in $2 \times 2$ games}\label{section:2_by_2_games}
-
-\section{Conclusions}\label{section:conclusions}
-
-\appendix
-
-\section{Model Setup}\label{appendix:methods}
-
-Consider a population of \(N\) individuals where \(N\) is even. At any point in
-time there are at most two different strategies in present in the population.
-More specifically, a mutant strategy played by \(k\) individuals and a resident
-strategy played by \(N - k\) individuals. We assume a pairwise process in which
-strategies spread because they are imitated more often. Each step of the
-evolutionary process consists of two stages; a game stage and an update stage.
-
-In the game stage, each individual is randomly matched with some other
-individual in the population. Their interaction lasts for a number of turns
-which is not fixed but depends on the continuation probability \(\delta\). At
-each turn the individuals choose between cooperation (\(C\)) and defection
-(\(D\)). Thus, there are four possible outcomes in each turn \(CC, CD, DC\) and
-\(DD\). If both players cooperate they receive the reward payoff \(R\), whereas
-if both players defect they receive the punishment payoff \(P\). If one
-cooperates but the other defects, the defector receives the temptation to
-defect, \(T\), whereas the cooperator receives the sucker's payoff, \(S\).
-Let $\mathcal{U}=\{R,S,T,P\}$ denote the set of feasible payoffs in each round,
-and let $\mathbf{u}\!=\!(R,S,T,P)$ be the corresponding payoff vector.
-
-The values of the payoffs are not only based on the prisoner's dilemma but all
-the symmetric \(2 \times 2\) games, Table~\ref{table:social_dilemmas}.
+\subsection{Effect of updating payoffs in different social dilemmas}\label{section:2_by_2_games}
+
+By focusing on the donation game in the previous section, we have gained
+insights into the possible effects of the updating payoffs, and how parameters
+such as the benefit and the strength of selection might can exaggerate the
+effect. However, by focusing on one specific game can narrow our knowledge of
+the effects of updating payoffs on the different forms of possible human
+interactions and social dilemmas. For this reason we explore not only the 
+donation game but all the possible symmetric games. These are given by
+Table~\ref{table:social_dilemmas}.
 
@@
   \begin{center}
-    \resizebox{.65\textwidth}{!}{
+    \resizebox{.6\textwidth}{!}{
     \renewcommand{\arraystretch}{1.5}
@@
   \caption{\textbf{Social dilemmas and payoffs' constrains}. The various social
-  dilemmas we explore in this work. Results for cases (i) - (iv) were presented
-  in section~\ref{section:2_by_2_games} and results for case (v) were presented
+  dilemmas we explore in this work. Results for cases (i) - (iv) are presented
+  in section~\ref{section:2_by_2_games} and results for case (v) are presented
   in section~\ref{section:donation}.}
@@
 \end{table}
+
+% Include discussion on each game?
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{static/expected_two_by_two_games.pdf}
+\end{figure}
+
+\section{Conclusions}\label{section:conclusions}
+
+\appendix
+
+\section{Model Setup}\label{appendix:methods}
+
+Consider a population of \(N\) individuals where \(N\) is even. At any point in
+time there are at most two different strategies in present in the population.
+More specifically, a mutant strategy played by \(k\) individuals and a resident
+strategy played by \(N - k\) individuals. We assume a pairwise process in which
+strategies spread because they are imitated more often. Each step of the
+evolutionary process consists of two stages; a game stage and an update stage.
+
+In the game stage, each individual is randomly matched with some other
+individual in the population. Their interaction lasts for a number of turns
+which is not fixed but depends on the continuation probability \(\delta\). At
+each turn the individuals choose between cooperation (\(C\)) and defection
+(\(D\)). Thus, there are four possible outcomes in each turn \(CC, CD, DC\) and
+\(DD\). If both players cooperate they receive the reward payoff \(R\), whereas
+if both players defect they receive the punishment payoff \(P\). If one
+cooperates but the other defects, the defector receives the temptation to
+defect, \(T\), whereas the cooperator receives the sucker's payoff, \(S\).
+Let $\mathcal{U}=\{R,S,T,P\}$ denote the set of feasible payoffs in each round,
+and let $\mathbf{u}\!=\!(R,S,T,P)$ be the corresponding payoff vector.
+The values of the payoffs are not only based on the prisoner's dilemma but all
+the symmetric \(2 \times 2\) games, Table~\ref{table:social_dilemmas}.
 

### 2021-05-05 static/last_rounds_matches.tex
writting results part two

@@
 \node [arn_n] {} 
-    child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{M}$] {} edge from parent node[above left, rotate=30] {$\frac{1}{N - 1}$}}
+    child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{SR} - S_{SM}$\\$S_{R} - S_{R}$}] {} edge from parent node[above left, rotate=30] {$\frac{1}{N - 1}$}}
     child{ node [arn_n, rotate=-30] {} 

### 2021-05-05 main.tex
ready to push

@@
 
-% Include discussion on each game?
-
-\begin{figure}
+Without loss of generality we set one of the entries of the payoff matrix to 0
+and another one to 1. A common normalization is \(R=1\) and
+\(P=0\)~\cite{Martinez2012, Roca2009}. We vary the temptation to defect \(T\)
+and the risk of cooperating \(S\). We continue the discussion of the previous
+section. We compare the expected payoffs to the last round payoffs over the
+different social dilemmas. The results of the simulation runs are given in
+Figures~\ref{fig:expected_two_by_two} and~\ref{fig:last_round_two_by_two}.
+
+At the left up quadrat (\(S > 0\) and  \(T \leq 1\)) the individuals interact
+in the harmony game at each game stage. The harmony game is a game without
+tensions. The best play in the harmony game is to always cooperate. Indeed both
+figures confirm this. The behaviour varies sightly at the case \(T = 1\). Now
+the temptation is equal to the payoff of a cooperation and thus individuals now
+tend to defect but only slightly. The cooperating rate still remains high. At
+the bottom left up quadrat (\(S \leq 0\) and  \(T < 1\)) the individuals
+interact in the stag hunt game at each game stage. In the stag hunt game
+cooperation is still favoured. The best response of individuals is to cooperate
+and this is again confirmed by both approaches.
+
+The last two cases are those of the prisoner’s dilemma and of the snowdrift
+game. At the top right quadrat we explore the snowdrift game. In the snowdrift
+game individuals want to cooperate after receiving a defection and to defect
+given that their opponent defected. This behavior is captured by the figures.
+In the case of \(q\), cooperating after receiving a cooperation, we observe
+a decrease as the suckers payoff decreases. This is to be expected. Finally,
+in the right bottom quadrat individuals interact in a prisoner’s dilemma.
+In the prisoner’s dilemma individuals tend to defect. In the case of direct
+reciprocity cooperation can emerge, as we discussed in the previous section.
+However, in the spectrum of all the different values of \(S\) and \(T\) used
+here we observe this behaviour in only a handful of cases.
+
+\begin{figure}[!htbp]
   \centering
   \includegraphics[width=\textwidth]{static/expected_two_by_two_games.pdf}
+  \caption{{\bf Evolutionary dynamics under expected payoffs for various social dilemmas.} 
+  We have run several simulations of the evolutionary process described in
+  section~\ref{section:model} for $T\!=\!10^7$ time steps. The graphs show how
+  often the resident population chooses each combination ($p,q$) of conditional
+  cooperation probabilities in the subsequent rounds. We vary the temptation
+  payoff \(T \in \{-1, -0.6, -0.2,  0.2, 0.6, 1, 1.4, 1.8, 2.2, 2.6, 3\}\)
+  across the \(x\) axis, and  \(S \in \{2, 1.6, 1.2, 0.8, 0.4, 0, -0.4, -0.8,
+  -1.2, -1.6, -2\}\) across the \(y\) axis. Parameters: $N\!=\!100$, $c\!=\!1$,
+  $\beta\!=\!1$, $\delta\!=\!0.999$.}
+  \label{fig:expected_two_by_two}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=\textwidth]{static/stochastic_two_by_two_games.pdf}
+  \caption{{\bf Evolutionary dynamics under last round payoffs for various social dilemmas.} 
+  We have run several simulations of the evolutionary process described in
+  section~\ref{section:model} for $T\!=\!10^7$ time steps. The graphs show how
+  often the resident population chooses each combination ($p,q$) of conditional
+  cooperation probabilities in the subsequent rounds. We vary the temptation
+  payoff \(T \in \{-1, -0.6, -0.2,  0.2, 0.6, 1, 1.4, 1.8, 2.2, 2.6, 3\}\)
+  across the \(x\) axis, and  \(S \in \{2, 1.6, 1.2, 0.8, 0.4, 0, -0.4, -0.8,
+  -1.2, -1.6, -2\}\) across the \(y\) axis. Parameters: $N\!=\!100$, $c\!=\!1$,
+  $\beta\!=\!1$, $\delta\!=\!0.999$.}
+  \label{fig:last_round_two_by_two}
+\end{figure}
+
+We have discussed the expected behaviour for each social dilemma. We now want to
+account the difference in the cooperation rate given the two updating payoffs.
+The cooperation rates for each of the case of the social dilemmas we have used
+are given in Figure~\ref{fig:cooperation_rates_two_by_two_part_one}.
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=.45\textwidth]{static/expected_two_by_two_games_cooperation.pdf}
+  \includegraphics[width=.45\textwidth]{static/stochastic_two_by_two_games_cooperation.pdf}
+  \caption{\textbf{Cooperation rates over various social dilemmas.} ({\bf A}) If
+  players update based on their expected payoffs. ({\bf B}) If
+  players update based on their last round payoffs.}\label{fig:cooperation_rates_two_by_two_part_one}
+\end{figure}
+
+So far we have explored the difference between the expected payoffs and the last
+round payoffs. In order to explore further the effect of limited memory we allow
+individuals to remember more. We consider the cases where individuals remember
+up two interactions, and up to the last two rounds. In total we present results
+for three more cases. These are the cases of the last round two rounds with
+another member of the population, last round with two members of the population,
+and last round two rounds with two members of the population.
+
+Similarly to the previous examples we have run the evolutionary process for
+a large number of step for each of the social dilemmas. The results remain
+the same. The behavior over the different games remains the same. We note that
+now there is more noise in the evolved populations. Due to space the figures
+have been moved to the Appendix, however, the cooperating rates for each of
+the cases are given in Figure~\ref{fig:cooperation_rates_two_by_two_part_two}.
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=.33\textwidth]{static/rounds_two_by_two_games_cooperation.pdf}
+  \includegraphics[width=.33\textwidth]{static/opponents_two_by_two_games_cooperation.pdf}
+  \includegraphics[width=.33\textwidth]{static/rounds_opponents_two_by_two_games_cooperation.pdf}
+  \caption{\textbf{Cooperation rates over various social dilemmas for different
+  limited memory payoffs.} ({\bf A}) If
+  players update based on their last two rounds with another member of the population.
+  ({\bf B}) If players update based on their last round with two other members of the population.
+  ({\bf C}) If players update based on their last two rounds with two other members of the population.}
+  \label{fig:cooperation_rates_two_by_two_part_two}
 \end{figure}

### 2021-05-05 clean_data.py

@@ -20,6 +20,7 @@ if __name__ == "__main__":  # pragma: no cover
     ]
 
     print(f"{folder}/{starting_characters}*csv")
+
     csv_files = glob.glob(f"{folder}/{starting_characters}*csv")
 
     for file in tqdm.tqdm(csv_files):
@@ -36,4 +37,4 @@ if __name__ == "__main__":  # pragma: no cover
 
         df["frequencies"] = frequencies
 
-        df.to_csv(f"{folder}/clean_{file[len(folder) + 1:]}", index=False)
+        df.to_csv(f"{folder}/clean_{file[len(folder):]}", index=False)

### 2021-05-18 main.tex
finalise plots for paper

@@
 spread if they have a high fitness either because the individuals who adopt them
-have more offspring, or because they are imitated more often. The fitness of a
-strategy is not constant but depends on the composition of the population.
-Individuals interact based on their strategies with other members of the
-population and the payoffs they yield are translated into fitness.
-
-It is commonly assumed that fitness is equivalent to the mean distribution of
-types in the population. These payoffs assume that individuals can interact with
-the entire population several times and remember each and every outcome. Thus,
-they imply that individuals have a perfect memory. However, when they make
+have more offspring, or because they are imitated more often~\cite{Wu2015}. The
+fitness of a strategy is not constant but depends on the composition of the
+population. Individuals interact based on their strategies with other members of
+the population and the payoffs they yield are translated into fitness.
+
+It is commonly assumed that individuals compute their fitness after interacting
+with a representative sample of the population, and remembering all the
+interactions they participated in~\cite{Nowak2006}. Thus, they imply that
+individuals have a perfect memory. However, when modelling how individuals make
 decisions in each turn they are assumed to have very limited memory. To be
@@
 Note that there are a few notable exceptions~\cite{Hauert1997, Stewart2016}.
-
-The perfect memory assumption is not only unrealistic but it also creates this
-curious inconsistency. This has led us to question how robust is our
-understanding of cooperation. In this work, we explore whether direct
-reciprocity can evolve if individuals only remember a minimum of social
-information. Though we are not the first to question the assumptions of
-estimating fitness~\cite{Roca2006}, we are the first to explore the effect of
-payoff memory.
-
-Initially, we consider two extreme scenarios. The first is the classical
-scenario of the expected payoffs and the alternative scenario where individuals
-update their strategies only based on the very last payoff they obtained. In the
-later sections, we allow individuals more memory. More specifically, they can
-remember up to two turns and up to two interactions. We present results on
-several social dilemmas which include the prisoner’s dilemma and the donation
-game, the snowdrift game, the stag hunt game and the harmony game.
-
-The remainder of the paper is organized as follows. Section~\ref{section:model}
-describes the model. Section~\ref{section:results} presents the results of the
-simulations. Finally, section~\ref{section:conclusions} outlines the main
-conclusions.
-
-% Early results for the repeated prisoner’s dilemma and for the snowdrift game
-% suggest that memory size can have a considerable effect on the evolutionary
-% dynamics. As a rule of thumb, we observe that individuals with limited memory
-% tend to adopt less generous strategies and they achieve less cooperation than in
-% the classical scenario. In contrast, in the stag-hunt and the harmony game, the
-% impact of memory is less striking. Here individuals with limited memory perform
-% nearly as well as individuals with full memory.
+Thus, the perfect memory assumption is not only unrealistic but it also creates
+this curious inconsistency.
+
+This has led us to question the robustness of our understanding of cooperation,
+In this work we explore whether cooperation can evolve if individuals compute
+their fitness based on a minimum of social information. Though we are not the
+first to question the assumptions of estimating fitness~\cite{Roca2006}, we are
+the first to explore the effect of payoff memory.
+
+We first consider two extreme scenarios, the classical scenario of the expected
+payoffs and the alternative scenario where individuals update their strategies
+only based on the very last payoff they obtained. We observe that individuals
+with limited memory tend to adopt less generous strategies and they achieve less
+cooperation than in the classical scenario. We obtain similar results when we
+consider that individuals update their strategies based on more social
+information. More specifically, up to the last two payoffs they obtained when
+interacting with up to two different members of the population.
+
+The remainder of the paper is organized as follows. In
+section~\ref{section:model} we describe the model. In
+section~\ref{section:results} we presents the results of the simulations.
+Finally, section~\ref{section:conclusions} we outline the main conclusions.
 
@@
 
-We consider a population of \(N\) players (\(N\) is even) where mutations are
-sufficiently rare. Thus, at any point in time there are at most two
-different strategies present in the population; a \textit{resident} strategy
-and a \textit{mutant} strategy. We assume a pairwise process where
-strategies spread because they are imitated more often.
+We consider a population of \(N\) players~\footnote{The terms ``player'' and
+``individual'' are used interchangeably here.} (\(N\) is even) where mutations are
+sufficiently rare. Thus, at any point in time there are at most two different
+strategies present in the population; a \textit{resident} strategy and a
+\textit{mutant} strategy. We assume a pairwise process where strategies spread
+because they are imitated more often.
 
@@
 last round payoff achieved against another member of the population. We compare
-this to the expected payoffs. We assume that individuals interact in a donation
-game  where each can cooperate by providing a benefit \(b\) to the other
-player at their cost \(c\), with \(0 < c < b\). Thus, \(T=b, R=b-c, S=-c, P=0\).
+this to the classical scenario of the expected payoffs. We assume that each pair
+of players interacts in a donation game. The donation game is a special case of
+the prisoner's dilemma. Each player can choose to cooperate by providing a
+benefit \(b\) to the other player at their cost \(c\), with \(0 < c < b\).
+Thus, the payoffs of a one round interact are \(T=b, R=b-c, S=-c, P=0\).
 
@@
 The figure suggests that when updating is based on expected payoffs, players
-tend to be more generous and more cooperative. The $q$-values are higher on
-average which suggests that individuals tend to be more generous when it comes
-to cooperating after being at the receiving end of a defection. %This is a known result no? Nowak 1992
-The average cooperation rate within the resident population is strictly higher
-when expected payoffs are considered. The difference between the two methods for
-both the values of \(b\) are statistically significant. This contrast becomes
-more obvious for \(b=10\). More specifically the average cooperation drops from
-97\% to 51\%. The residents of the population cooperated on average 97\% of the
-times in one case and only 57\% in the other.
+tend to be more generous and more cooperative. The $q$-values of the resident
+strategies are on average higher in the case of the expected payoffs. The
+players will occasionally forgive a defection more often if their fitness
+depends on interacting with every member of the population. On the other hand,
+when social interactions are limited they are less forgiving. The average
+cooperation rate for each simulation is calculated as the the average
+cooperation rate within the resident population. In the case of the expected
+payoffs, regardless the value of benefit, the average cooperation rate is
+strictly higher than that of the last round payoffs. The difference based the
+two methods is statistically significant, and the dissimilarity is even more
+obvious for a high value of \(b\). More specifically, the average cooperation of
+resident strategies drops from 97\% to 57\%.
 
@@
 
-We further explore the effect of the benefit,
-Figure~\ref{fig:cooperation_rate_over_benefit}. The figure suggests that the
-expected payoffs always overestimate cooperation. For the limited memory
-payoffs, the cooperation rate remains unchanged at approximately 50\% once \(b =
-5\). The highest cooperation rate that was achieved was 0.51 when  the last
-round payoff is considered. In comparison, when the expected payoffs were used
-the highest average cooperation rate was 0.97.
+We further explore the effect of the benefit.
+Figure~\ref{fig:cooperation_rate_over_benefit} suggests that expected payoffs
+always yield a higher cooperation rate. We observe that the cooperation rate
+increases as the value of the benefit gets higher. In comparison, for the
+limited memory payoffs, the cooperation rate remains unchanged, at approximately
+50\%.
 
@@
 
+We also investigate the effect of the strength of selection.
 Figure~\ref{fig:cooperation_rate_over_betas} illustrates results for various
-runs of the evolutionary process where we vary the strength of selection. For
-weak selection, \(\beta < 1\), the updating payoffs have no effect. The evolved
-population and the average cooperating rate are the same for both approaches.
-However for strong selection, it can be seen that the cooperating rate increases
-with \(beta\), when expected payoffs are used. However, when limited memory
-payoffs are used the cooperating rate decreases.
+runs of the evolutionary process. For weak selection, \(\beta < 1\), we observe
+that the two methods yield the same result, however, as \(\beta\) increases
+there is variation in the evolving populations. In the case of expected payoffs
+the resident populations become more and more cooperative as \(\beta\)
+increases, whereas in the case of limited memory payoffs, the resident
+populations become less cooperative.
 
@@
 
-By focusing on the donation game in the previous section, we have gained
-insights into the possible effects of the updating payoffs, and how parameters
-such as the benefit and the strength of selection might can exaggerate the
-effect. However, by focusing on one specific game can narrow our knowledge of
-the effects of updating payoffs on the different forms of possible human
-interactions and social dilemmas. For this reason we explore not only the 
-donation game but all the possible symmetric games. These are given by
-Table~\ref{table:social_dilemmas}.
+In the previous subsection we gained insights into the effects of the updating
+payoffs, and into how parameters such as the benefit and the strength of
+selection can intensify them. We investigated these effects by using the
+donation game. In order to broaden our understanding of the updating payoffs on
+different forms of possible human interactions, we extend our approach to other
+\(2 \times 2\) symmetric games. More specifically, we apply our analysis to the
+four different classes of games given by Table~\ref{table:social_dilemmas}. We
+compare the results of the evolutionary process described in
+section~\ref{section:model} when the updating payoffs are based on the expected
+payoffs, and on the last round payoffs for all the four possible classes of
+games. The results of the simulations are given in
+Figures~\ref{fig:expected_two_by_two} and~\ref{fig:last_round_two_by_two}
+respectively.
 
@@
   \begin{center}
-    \resizebox{.6\textwidth}{!}{
+    \resizebox{.45\textwidth}{!}{
     \renewcommand{\arraystretch}{1.5}
@@
   \toprule
-  & \textbf{social dilemmas} & \textbf{payoffs' constrains} \\
+  & \textbf{social dilemmas} & \textbf{preference ordering} \\
   \midrule
-(i) & stage hunt  & \(R > T > P > S\) \\
+(iii) & harmony  & \(R > T > S > P\) \\
+(i) & stag hunt  & \(R > T > P > S\) \\
+(iv) & prisoner dilemma  & \(T > R > P > S\) \\
 (ii) & snowdrift  & \(T > R > S > P\) \\
-(iii) & harmony  & \(R > T > S > P\) \\
-(iv) & prisoner dilemma  & \(T > R > P > S\) \\
-(v)& donation game  & \(T > R > P > S\) \& \(T=b, R=b-c, S=-c, P=0\) \\
   \bottomrule
@@
 
-Without loss of generality we set one of the entries of the payoff matrix to 0
-and another one to 1. A common normalization is \(R=1\) and
-\(P=0\)~\cite{Martinez2012, Roca2009}. We vary the temptation to defect \(T\)
-and the risk of cooperating \(S\). We continue the discussion of the previous
-section. We compare the expected payoffs to the last round payoffs over the
-different social dilemmas. The results of the simulation runs are given in
-Figures~\ref{fig:expected_two_by_two} and~\ref{fig:last_round_two_by_two}.
-
-At the left up quadrat (\(S > 0\) and  \(T \leq 1\)) the individuals interact
-in the harmony game at each game stage. The harmony game is a game without
-tensions. The best play in the harmony game is to always cooperate. Indeed both
-figures confirm this. The behaviour varies sightly at the case \(T = 1\). Now
-the temptation is equal to the payoff of a cooperation and thus individuals now
-tend to defect but only slightly. The cooperating rate still remains high. At
-the bottom left up quadrat (\(S \leq 0\) and  \(T < 1\)) the individuals
-interact in the stag hunt game at each game stage. In the stag hunt game
-cooperation is still favoured. The best response of individuals is to cooperate
-and this is again confirmed by both approaches.
-
-The last two cases are those of the prisoner’s dilemma and of the snowdrift
-game. At the top right quadrat we explore the snowdrift game. In the snowdrift
-game individuals want to cooperate after receiving a defection and to defect
-given that their opponent defected. This behavior is captured by the figures.
-In the case of \(q\), cooperating after receiving a cooperation, we observe
-a decrease as the suckers payoff decreases. This is to be expected. Finally,
-in the right bottom quadrat individuals interact in a prisoner’s dilemma.
-In the prisoner’s dilemma individuals tend to defect. In the case of direct
-reciprocity cooperation can emerge, as we discussed in the previous section.
-However, in the spectrum of all the different values of \(S\) and \(T\) used
-here we observe this behaviour in only a handful of cases.
+In Figure~\ref{fig:expected_two_by_two} each sub figure represents a run of
+evolutionary process for a different set of values the payoffs of an individual,
+\(\mathcal{U}\). Without loss of generality we set \(R=1\) and
+\(P=0\)~\cite{Martinez2012, Roca2009}, and we vary the values of the temptation
+\(T\) (across the \(x\)-axis) and of the risk of cooperating \(S\) (across the
+\(y\)-axis). There are four classes of games as presented in
+Table~\ref{table:social_dilemmas}. Starting at the upper left corner and proceeding
+clockwise correspond to the harmony game, the snowdrift, the prisoner's dilemma
+and the stag hunt.
+
+The harmony game represents a situation without tensions where it is in the best
+interest of both players to cooperate. This is confirmed by our simulations
+Figure~\ref{fig:expected_two_by_two}. In the several experiments of the harmony
+game the resident population applies a strategy for which \(p \approx 1\) and
+\(q \approx 1\).
+
+The snowdrift game describes a situation similar to the prisoner's dilemma where
+cooperation results in a benefit to the opposing player but entailing a cost to
+the cooperator. Compared to the prisoner's dilemma, individuals obtain immediate
+direct benefits from the cooperative acts. The story of the snowdrift game
+usually goes as follow: two drivers are trapped on either side of a snowdrift
+and have the options of staying in the car or removing the snowdrift. Letting
+the opponent do all the work is the best option but if the other player stays in
+the car it is better to shove. In the end cooperation leads to the cooperated
+drive to get home, even though they had to endure the work, thus  \(S > P\). In
+the snowdrift game more cooperation emerges compared to the prisoner's dilemma.
+Defection is never a resident strategy and for the shame values of temptation
+the overall \(q-values\) are higher. It can be seen that the \(p-values\) are
+lower as individuals have free room to defect after receiving a cooperation.
+
+The last class of games corresponds to the stag hunt game. In the stag hunt game
+both players benefit for mutual cooperation and defection, however, \(R > P\) and
+thus the resident strategies with the highest fitness are the cooperative ones.
+The results are similar to the cases of the harmony game.
 
@@
 
-We have discussed the expected behaviour for each social dilemma. We now want to
-account the difference in the cooperation rate given the two updating payoffs.
-The cooperation rates for each of the case of the social dilemmas we have used
-are given in Figure~\ref{fig:cooperation_rates_two_by_two_part_one}.
+The results for the last round payoffs are fairly similar with the patterns
+begging the same. It can been seen from Figure~\ref{fig:last_round_two_by_two}
+that the results are move variation. We have discussed the expected behaviour
+for each social dilemma. We now want to account the difference in the
+cooperation rate given the two updating payoffs. The cooperation rates for each
+of the case of the social dilemmas we have used are given in
+Figure~\ref{fig:cooperation_rates_two_by_two_part_one}.
 
@@
   \centering
-  \includegraphics[width=.45\textwidth]{static/expected_two_by_two_games_cooperation.pdf}
-  \includegraphics[width=.45\textwidth]{static/stochastic_two_by_two_games_cooperation.pdf}
+  \includegraphics[width=.4\textwidth]{static/expected_two_by_two_games_cooperation.pdf}
+  \includegraphics[width=.4\textwidth]{static/stochastic_two_by_two_games_cooperation.pdf}
   \caption{\textbf{Cooperation rates over various social dilemmas.} ({\bf A}) If

### 2021-05-20 main.tex
read over - finalise figures - clean tex

@@
 \begin{abstract}
-
+Repeated games have been the centerpiece when studying the evolution of human
+cooperation. Most of the existing models work with idealized assumptions.
+Individuals are assumed to interact with a representative sample of the
+population, and they remember all interactions they participate in. In real
+life, however, individuals do not always remember all interactions they had with
+everyone in their population. Instead, they rather recall the last interaction
+they had.
+
+In this work we explore the effect on the evolution of cooperation if
+individuals only remember a minimum of social information. We present results
+for a series of simulations were we vary the benefit of cooperation and the
+strength of selection. We present results for the most commonly used classes of
+social dilemmas, and we show that even though cooperation can evolve when
+individuals only use a minimum of information, the evolving cooperation rates
+are typically lower than in the classical scenario.
 \end{abstract}
@@
 increasing their payoff, at the expense of decreasing one’s own? In evolutionary
-game theory, individuals are not required to be rational, instead they adapt
+game theory individuals are not required to be rational, instead they adapt
 strategies based on mutation and exploration. Strategies are more likely to
@@
 Note that there are a few notable exceptions~\cite{Hauert1997, Stewart2016}.
-Thus, the perfect memory assumption is not only unrealistic but it also creates
+The perfect memory assumption is not only unrealistic but it also creates
 this curious inconsistency.
 
-This has led us to question the robustness of our understanding of cooperation,
+This has led us to question the robustness of our understanding of cooperation.
 In this work we explore whether cooperation can evolve if individuals compute
@@
 first to question the assumptions of estimating fitness~\cite{Roca2006}, we are
-the first to explore the effect of payoff memory.
+the first to explore the effect of payoff memory. %ToDo add more information
 
@@
 section~\ref{section:model} we describe the model. In
-section~\ref{section:results} we presents the results of the simulations.
-Finally, section~\ref{section:conclusions} we outline the main conclusions.
+section~\ref{section:results} we present the results of the simulations, and
+in section~\ref{section:conclusions} we outline the main conclusions.
 
@@
 We consider a population of \(N\) players~\footnote{The terms ``player'' and
-``individual'' are used interchangeably here.} (\(N\) is even) where mutations are
-sufficiently rare. Thus, at any point in time there are at most two different
-strategies present in the population; a \textit{resident} strategy and a
-\textit{mutant} strategy. We assume a pairwise process where strategies spread
-because they are imitated more often.
-
-Each step of the evolutionary process consists of two stages, a game stage and
-an updating stage. In the game stage each individual is randomly matched with
-some other individual in the population to interact for a number of turns, where
-subsequent turns occur with a fixed probability $\delta$. At each turn they
-can choose to either cooperate (\(C\)) or to defect (\(D\)), and thus, at
-each turn the possible outcomes are \(CC, CD, DC\) and \(DD\). The payoffs
-depend on the outcome. If both cooperate they receive the reward payoff
-\(R\), whereas if both defect they receive the punishment payoff \(P\).
-If one cooperates but the other defects, the defector receives the temptation to
-defect, \(T\), whereas the cooperator receives the sucker's payoff, \(S\).
-We denote the payoffs of an individual as \(\mathcal{U} = (R, S, T, P)\).
-
-We assume herein that individuals use \textit{reactive strategies} to make
+``individual'' are used interchangeably here.} where \(N\) is even and mutations
+are sufficiently rare. Therefor, at any point in time there are at most two
+different strategies present in the population; a \textit{resident} strategy and
+a \textit{mutant} strategy. To describe how strategies spread we use a pairwise
+comparison process~\cite{Traulsen2006}.
+Each step of the evolutionary process consists of two
+stages, a game stage and an updating stage.
+
+In the game stage each individual is randomly matched with some other individual
+in the population. They engage in a match where each subsequent turn
+occurs with a fixed probability $\delta$. At each turn players can choose to
+either cooperate (\(C\)) or to defect (\(D\)), and the payoffs of the turn
+depend on both their decisions. If both players cooperate they receive the
+reward payoff \(R\), whereas if both defect they receive the punishment payoff
+\(P\). If one cooperates but the other defects, the defector receives the
+temptation \(T\) payoff, whereas the cooperator receives the sucker's
+payoff \(S\). We denote the feasible payoff of each turn as \(\mathcal{U} =
+\{R, S, T, P\}\).
+We assume that individuals use \textit{reactive strategies} to make
 decisions in each turn. Reactive strategies are a set of memory-one strategies
@@
 `learner' and a `role model'. Given that the learner's payoff $u_L\!\in\!
-\mathcal{U}$ and that the role model's payoff $u_{RL}\!\in\! \mathcal{U}$, we
-assume the learner adopts the role model's strategy based on the Fermi
+\mathcal{U}$ and that the role model's payoff $u_{RL}\!\in\! \mathcal{U}$.
+The learner adopts the role model's strategy based on the Fermi
 distribution function,
@@
 
-where $\beta\!\ge\!0$ is the relative influence of the payoffs on adopting the 
-strategy of the other. We refer to $\beta$ as the intensity of selection.
-
-This basic evolutionary step is repeated until either the mutant strategy goes
+where $\beta\!\ge\!0$ is the intensity of selection. The updating payoffs of the
+learner and the role model are conventionally correspond to their expected
+payoffs. The expected payoff of a strategy is the mean payoff of the strategy of
+engaging in repeated games with all other members of the population. The expected payoffs
+assume perfect memory. In this work we define a new a set of updating
+payoffs. We reefer to these as the limited memory payoffs, and we compare them
+to the classical scenario.
+
+The evolutionary step is repeated until either the mutant strategy goes
 extinct, or until it fixes in the population. If the mutant fixes in the
@@
 process of mutation and fixation/extinction is then iterated many times.
-
-The perfect memory assumption occurs at the updating stage. The learner and the
-role model are assumed to interact with a representative sample of the
-population, and they remember all interactions they participate in. Thus, their
-updating payoffs are based on the mean payoff they achieved over all the
-interactions. These payoffs are referred to as the expected payoffs. We will
-compare the expected payoffs to payoffs that are calculated when the role model
-and learner do not remember all of their interactions. In order to account for
-the effect of these different methods, we explore the cooperation rate within
-the resident population over multiple generations. More details on our
-methodology are found in Appendix~\ref{appendix:methods}.
+More details on our methodology are found in Appendix~\ref{appendix:methods}.
+
+In order to account for the effect of the updating payoffs approaches, we
+explore the cooperation rate within the resident population over multiple
+generations. In the next section we present a series of simulation results.
+To account for the various types of social behaviour we also present
+results on multiple social dilemmas.
 
@@
 In this section we explore the case where the updating payoffs are based on the
-last round payoff achieved against another member of the population. We compare
+last round payoff achieved against another member of the population and we compare
 this to the classical scenario of the expected payoffs. We assume that each pair
@@
 benefit \(b\) to the other player at their cost \(c\), with \(0 < c < b\).
-Thus, the payoffs of a one round interact are \(T=b, R=b-c, S=-c, P=0\).
+Thus, the feasible payoffs in each round are \(T=b, R=b-c, S=-c, P=0\).
 
@@
 for the described process of section~\ref{section:model}.
-Figure~\ref{fig:expected_and_stochastic_for_donation}~depicts the evolving
+Figure~\ref{fig:expected_and_stochastic_for_donation} depicts the evolving
 conditional cooperation probabilities $p$ and $q$. The discount factor~$\delta$
-is comparably high, thus the opening move \(y\) is a transient
-effect and has no effect on the outcome. The left panel corresponds to the
-standard scenario considered in the literature. It considers players who use
+is comparably high, thus we do not report the opening move \(y\) as it
+is a transient effect. The left panels correspond to the
+standard scenario considered in the literature, it considers players who use
 expected payoffs to update their strategies. The right panel shows the scenario
 considered herein, in which players update their strategies based on their last
-round’s payoff. The top panels assume a benefit \(b\) of 3, whereas the bottom
+round’s payoff. The top panels assume a benefit \(b\) of 3 whereas the bottom
 assume a benefit of 10.
 
-The figure suggests that when updating is based on expected payoffs, players
+The figure suggests that when updating is based on expected payoffs players
 tend to be more generous and more cooperative. The $q$-values of the resident
@@
 when social interactions are limited they are less forgiving. The average
-cooperation rate for each simulation is calculated as the the average
+cooperation rate for each simulation is calculated as the average
 cooperation rate within the resident population. In the case of the expected
 payoffs, regardless the value of benefit, the average cooperation rate is
-strictly higher than that of the last round payoffs. The difference based the
-two methods is statistically significant, and the dissimilarity is even more
-obvious for a high value of \(b\). More specifically, the average cooperation of
-resident strategies drops from 97\% to 57\%.
-
-\begin{figure}[!htbp]
-    \centering
-    \includegraphics[width=.75\textwidth]{static/expected_and_stochastic_for_donation_game.pdf}\vspace{-3cm}
-    \includegraphics[width=.75\textwidth]{static/expected_and_stochastic_for_donation_game_b_10.pdf}
+strictly higher than that of the last round payoffs. The difference based on the
+two methods is statistically significant, and in the case of $b=10$ the
+average cooperation of resident strategies drops from 97\% to 57\%.
+
+\begin{figure}[!htbp]
+    \centering
+    \includegraphics[width=.70\textwidth]{static/expected_and_stochastic_for_donation_game.pdf}\vspace{-3cm}
+    \includegraphics[width=.70\textwidth]{static/expected_and_stochastic_for_donation_game_b_10.pdf}
     \caption{{\bf Evolutionary dynamics under expected payoffs and last round with one interaction payoffs.} 
@@
 
-We further explore the effect of the benefit.
-Figure~\ref{fig:cooperation_rate_over_benefit} suggests that expected payoffs
-always yield a higher cooperation rate. We observe that the cooperation rate
-increases as the value of the benefit gets higher. In comparison, for the
-limited memory payoffs, the cooperation rate remains unchanged, at approximately
-50\%.
+We further explore the effect of benefit in
+Figure~\ref{fig:cooperation_rate_over_benefit}. The figure suggests that
+expected payoffs always yield a higher cooperation rate. In the case of expected
+payoffs we observe that the cooperation rate increases as the value of the
+benefit gets higher. In comparison for the limited memory payoffs, the
+cooperation rate remains unchanged at approximately 50\% once \(b=5\).
 
@@
     \centering
-    \includegraphics[width=\textwidth]{static/expected_for_beta.pdf}
-    \includegraphics[width=\textwidth]{static/stochastic_for_beta.pdf}
+    \includegraphics[width=.8\textwidth]{static/expected_for_beta.pdf}
+    \includegraphics[width=.8\textwidth]{static/stochastic_for_beta.pdf}
   \end{subfigure}%
@@
     \centering
-    \includegraphics[width=\textwidth]{static/cooperation_rate_over_b.pdf}
+    \includegraphics[width=.8\textwidth]{static/cooperation_rate_over_b.pdf}
   \end{subfigure}
@@
   $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run
-  for $T\!=\!5\times 10^6$ time steps for each parameter
+  for $T\!=\!5\times 10^7$ time steps for each parameter
   combination.}\label{fig:cooperation_rate_over_benefit}
@@
 
-We also investigate the effect of the strength of selection.
+We also investigate the effect of the strength of selection $\beta$.
 Figure~\ref{fig:cooperation_rate_over_betas} illustrates results for various
 runs of the evolutionary process. For weak selection, \(\beta < 1\), we observe
-that the two methods yield the same result, however, as \(\beta\) increases
+that the two methods yield similar results, however, as \(\beta\) increases
 there is variation in the evolving populations. In the case of expected payoffs
-the resident populations become more and more cooperative as \(\beta\)
+the resident populations become more cooperative as \(\beta\)
 increases, whereas in the case of limited memory payoffs, the resident
-populations become less cooperative.
+populations become more defective.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{static/expected_for_selection_strenght.pdf}
-    \includegraphics[width=\textwidth]{static/stochastic_for_selection_strenght.pdf}
+    \includegraphics[width=.8\textwidth]{static/expected_for_selection_strenght.pdf}
+    \includegraphics[width=.8\textwidth]{static/stochastic_for_selection_strenght.pdf}
   \end{subfigure}%
@@
     \vspace{.3cm}
-    \includegraphics[width=\textwidth]{static/cooperation_rate_over_betas.pdf}
+    \includegraphics[width=.8\textwidth]{static/cooperation_rate_over_betas.pdf}
   \end{subfigure}
@@
 $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
-$T\!=\!5\times 10^6$ time steps for each parameter
+$T\!=\!5\times 10^7$ time steps for each parameter
 combination.}\label{fig:cooperation_rate_over_betas}
@@
 
-In the previous subsection we gained insights into the effects of the updating
+In the previous section we gained insights into the effects of the updating
 payoffs, and into how parameters such as the benefit and the strength of
@@
 donation game. In order to broaden our understanding of the updating payoffs on
-different forms of possible human interactions, we extend our approach to other
-\(2 \times 2\) symmetric games. More specifically, we apply our analysis to the
-four different classes of games given by Table~\ref{table:social_dilemmas}. We
-compare the results of the evolutionary process described in
+different forms of possible human interactions we extend our approach to other
+\(2 \times 2\) symmetric games. More specifically, we apply our analysis to four
+different classes of games as given by Table~\ref{table:social_dilemmas}.
+
+\begin{table}[!htbp]
+  \begin{center}
+    \resizebox{.4\textwidth}{!}{
+    \renewcommand{\arraystretch}{1.5}
+      \begin{tabular}{ccc}
+  \toprule
+  & \textbf{social dilemmas} & \textbf{preference ordering} \\
+  \midrule
+(i) & harmony  & \(R > T > S > P\) \\
+(ii) & stag hunt  & \(R > T > P > S\) \\
+(iii) & prisoner dilemma  & \(T > R > P > S\) \\
+(iv) & snowdrift  & \(T > R > S > P\) \\
+  \bottomrule
+      \end{tabular}}
+  \end{center}
+\caption{\textbf{Social dilemmas and preference ordering}. The four classes of
+social dilemmas we explore in this work.}
+\label{table:social_dilemmas}
+\end{table}
+
+We compare the results of the evolutionary process described in
 section~\ref{section:model} when the updating payoffs are based on the expected
@@
 
-\begin{table}[!htbp]
-  \begin{center}
-    \resizebox{.45\textwidth}{!}{
-    \renewcommand{\arraystretch}{1.5}
-      \begin{tabular}{ccc}
-  \toprule
-  & \textbf{social dilemmas} & \textbf{preference ordering} \\
-  \midrule
-(iii) & harmony  & \(R > T > S > P\) \\
-(i) & stag hunt  & \(R > T > P > S\) \\
-(iv) & prisoner dilemma  & \(T > R > P > S\) \\
-(ii) & snowdrift  & \(T > R > S > P\) \\
-  \bottomrule
-      \end{tabular}}
-  \end{center}
-\caption{\textbf{Social dilemmas and payoffs' constrains}. The various social
-dilemmas we explore in this work. Results for cases (i) - (iv) are presented
-in section~\ref{section:2_by_2_games} and results for case (v) are presented
-in section~\ref{section:donation}.}
-\label{table:social_dilemmas}
-\end{table}
-
 In Figure~\ref{fig:expected_two_by_two} each sub figure represents a run of
-evolutionary process for a different set of values the payoffs of an individual,
-\(\mathcal{U}\). Without loss of generality we set \(R=1\) and
-\(P=0\)~\cite{Martinez2012, Roca2009}, and we vary the values of the temptation
-\(T\) (across the \(x\)-axis) and of the risk of cooperating \(S\) (across the
-\(y\)-axis). There are four classes of games as presented in
-Table~\ref{table:social_dilemmas}. Starting at the upper left corner and proceeding
-clockwise correspond to the harmony game, the snowdrift, the prisoner's dilemma
-and the stag hunt.
-
-The harmony game represents a situation without tensions where it is in the best
-interest of both players to cooperate. This is confirmed by our simulations
-Figure~\ref{fig:expected_two_by_two}. In the several experiments of the harmony
-game the resident population applies a strategy for which \(p \approx 1\) and
-\(q \approx 1\).
-
-The snowdrift game describes a situation similar to the prisoner's dilemma where
-cooperation results in a benefit to the opposing player but entailing a cost to
-the cooperator. Compared to the prisoner's dilemma, individuals obtain immediate
-direct benefits from the cooperative acts. The story of the snowdrift game
-usually goes as follow: two drivers are trapped on either side of a snowdrift
-and have the options of staying in the car or removing the snowdrift. Letting
-the opponent do all the work is the best option but if the other player stays in
-the car it is better to shove. In the end cooperation leads to the cooperated
-drive to get home, even though they had to endure the work, thus  \(S > P\). In
+evolutionary process for a different set of values for \(\mathcal{U}\). Without
+loss of generality we set \(R=1\) and \(P=0\)~\cite{Martinez2012, Roca2009}, and
+we vary the values of the temptation payoff \(T\) (across the \(x\)-axis) and of
+the sucker's payoff \(S\) (across the \(y\)-axis). Starting at the upper left
+corner and proceeding clockwise the quadrants correspond to the harmony, the
+snowdrift, the prisoner's dilemma and the stag hunt games.
+
+The harmony game represents are social situation without conflict. It is in the
+best interest for both players to cooperate as \(R > T\) and \(S > P\).
+Figure~\ref{fig:expected_two_by_two} confirms this. We observe that for most
+harmony game runs the resident population overwhelmingly applies a strategy for
+which \(p\) and \(p\) are \(\approx 1\). In the case of \(T = R = 1\) we observe
+that the resident strategies become slightly less generous. Against a cooperating
+strategy defecting and cooperation yield the same result. The lower \(q-\)values
+could indicate that the resident strategy is less generous to defectors to avoid
+being invaded.
+
+The snowdrift game describes a situation similar to that of the prisoner's
+dilemma. Cooperation results in a benefit to the opposing player but entailing a
+cost to the cooperator. However, in the snowdrift game individuals obtain
+immediate direct benefits from the cooperative acts which leads to \(S > P\). In
 the snowdrift game more cooperation emerges compared to the prisoner's dilemma.
@@
 
-The last class of games corresponds to the stag hunt game. In the stag hunt game
-both players benefit for mutual cooperation and defection, however, \(R > P\) and
-thus the resident strategies with the highest fitness are the cooperative ones.
-The results are similar to the cases of the harmony game.
+The last class of games we present are for the stag hunt game. In the stag hunt
+both players benefit for mutual cooperation and defection, however, \(R > P\)
+and thus the resident strategies with the highest fitness are the cooperative
+ones. The results for the stag hunt class are similar to those of the harmony
+game.
 
@@
   cooperation probabilities in the subsequent rounds. We vary the temptation
-  payoff \(T \in \{-1, -0.6, -0.2,  0.2, 0.6, 1, 1.4, 1.8, 2.2, 2.6, 3\}\)
-  across the \(x\) axis, and  \(S \in \{2, 1.6, 1.2, 0.8, 0.4, 0, -0.4, -0.8,
-  -1.2, -1.6, -2\}\) across the \(y\) axis. Parameters: $N\!=\!100$, $c\!=\!1$,
-  $\beta\!=\!1$, $\delta\!=\!0.999$.}
+  payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\)
+  across the \(x\) axis, and  \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\)
+  across the \(y\) axis. Unless explicitly varied, the parameters of the simulation
+  are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$.}
   \label{fig:expected_two_by_two}
 \end{figure}
+
+The conclusions regarding the evolved behaviour for the various classes of
+games remain the same when the last round payoffs are considered,
+Figure~\ref{fig:last_round_two_by_two}. However, there is variation in these
+populations. This variation as an effect affects the resident within cooperative
+rates.
 
@@
   cooperation probabilities in the subsequent rounds. We vary the temptation
-  payoff \(T \in \{-1, -0.6, -0.2,  0.2, 0.6, 1, 1.4, 1.8, 2.2, 2.6, 3\}\)
-  across the \(x\) axis, and  \(S \in \{2, 1.6, 1.2, 0.8, 0.4, 0, -0.4, -0.8,
-  -1.2, -1.6, -2\}\) across the \(y\) axis. Parameters: $N\!=\!100$, $c\!=\!1$,
-  $\beta\!=\!1$, $\delta\!=\!0.999$.}
+  payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\)
+  across the \(x\) axis, and \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\)
+  across the \(y\) axis. Unless explicitly varied, the parameters of the simulation
+  are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$.}
   \label{fig:last_round_two_by_two}
@@
 
-The results for the last round payoffs are fairly similar with the patterns
-begging the same. It can been seen from Figure~\ref{fig:last_round_two_by_two}
-that the results are move variation. We have discussed the expected behaviour
-for each social dilemma. We now want to account the difference in the
-cooperation rate given the two updating payoffs. The cooperation rates for each
-of the case of the social dilemmas we have used are given in
-Figure~\ref{fig:cooperation_rates_two_by_two_part_one}.
+The cooperation rates for each of the case of the social dilemmas we have used
+are given in Figure~\ref{fig:cooperation_rates_two_by_two_part_one}. In the case
+of the harmony game the differences are very small, even in the case of \(T=1\)
+where the biggest difference occurs it still remain less than 10\%. In the case
+of the stag hunt game the difference increases in the case of \(T=1\). The
+biggest variation in the results are for the class of the prisoner's dilemma.
+For values of \(S < -0.4\) the are significantly different, in the case of the
+last round payoffs cooperation almost never is a valid option. Overall the last
+round payoffs are less cooperative, supporting the results of the previous
+section. There are only two cases which that is not true and both cases are in
+the snowdrift class. For the values of \(T>1.8\) and \(S=0.4, 0.8\) cooperation
+is estimated higher. In the last round payoffs a Tit For Tat like behaviour is
+more robust that in the case of the expected payoffs.
 
@@
   players update based on their expected payoffs. ({\bf B}) If
-  players update based on their last round payoffs.}\label{fig:cooperation_rates_two_by_two_part_one}
+  players update based on their last round payoffs. We vary the temptation
+  payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\)
+  across the \(x\) axis, and \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\)
+  across the \(y\) axis. Unless explicitly varied, the parameters of the simulation
+  are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$.}\label{fig:cooperation_rates_two_by_two_part_one}
 \end{figure}
@@
 another member of the population, last round with two members of the population,
-and last round two rounds with two members of the population.
-
-Similarly to the previous examples we have run the evolutionary process for
-a large number of step for each of the social dilemmas. The results remain
-the same. The behavior over the different games remains the same. We note that
-now there is more noise in the evolved populations. Due to space the figures
-have been moved to the Appendix, however, the cooperating rates for each of
-the cases are given in Figure~\ref{fig:cooperation_rates_two_by_two_part_two}.
+and last round two rounds with two members of the population. Similarly to the
+previous examples we have run the evolutionary process for a large number of
+step for each of the social dilemmas. The results are found in the
+Appendix~\ref{appendix:further_simulation_results}.
+
+The behavior over the different games remains the same. We note that now there
+is more noise in the evolved populations. Due to space the figures have been
+moved to the Appendix, however, the cooperating rates for each of the cases are
+given in Figure~\ref{fig:cooperation_rates_two_by_two_part_two}.
+
+In the cases of the last two rounds the evolving cooperating rate is closer
+to that of the expected payoffs. In the case of the prisoner's dilemma there
+is an increase in the average cooperating rate which is still strictly less
+than in the case of the prisoner's dilemma. Regarding the snowdrift cases
+now each of the run for which \(T = 2.2\) the cooperating rate is not higher
+that the expected one.
+
+In the case of the two opponents experiments all the runs are lower with the
+expected payoffs, and in the case of the prisoner's dilemma we observe an
+increase in the average cooperate rate again. Similar in the case where both
+the last two turns and the two opponents are considered. The results appear
+to be the same as in the case of the two opponents. However, for some classes
+the results are affected by the last round.
 
@@
   ({\bf B}) If players update based on their last round with two other members of the population.
-  ({\bf C}) If players update based on their last two rounds with two other members of the population.}
+  ({\bf C}) If players update based on their last two rounds with two other members of the population.
+  We vary the temptation
+  payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\)
+  across the \(x\) axis, and \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\)
+  across the \(y\) axis. Unless explicitly varied, the parameters of the simulation
+  are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$.}
   \label{fig:cooperation_rates_two_by_two_part_two}
@@
 \section{Conclusions}\label{section:conclusions}
+
+Evolutionary models have helped us understand the evolution of cooperation.
+A behaviour so well observed in the world around us though Darwin spoke
+about the survival of the fitness. Evolutionary
+
+Much research in the past has been devoted to explore conditionally cooperative
+strategies in pairwise interactions.
+
+Previous models of direct reciprocity often feature a curious inconsistency.
+When modeling how individuals make decisions in each round, these models assume
+that players only remember the last round. However, when modeling how
+individuals update their strategies over time, individuals are assumed to have
+perfect memory. Here, we explore how robust our understanding of cooperation is
+as models deviate from the perfect memory assumption. On the positive side, we
+show that cooperation can even evolve when individuals only use a minimum of
+information. On the negative side, the evolving cooperation rates are typically
+lower than in the classical scenario.
+
+In the case of the donation game we show that cooperation is overestimated.
+Morover, we obsved that when the befenti and the srang f sleection increases this
+difference becomes more obvuous,
 
@@
 strategy of a resident. The probabilities that the number of mutants decreases
-and increases, \(\lambda^-_k\) and \(\lambda^+_k\), are not explicitly define
+and increases, \(\lambda^-_k\) and \(\lambda^+_k\), are now explicitly defined
 as,
@@
 
-\begin{Prop}
+\begin{Prop}\label{proposition:last_round}
     Consider a repeated prisoner's dilemma, with
@@
 In the updating stage we select a mutant and resident to be either the role
-model or the learner. Assume the selected mutant. Given that they can interact
-with only one other member of the population, they can interact either the
-selected resident, another resident or with another mutant. The same is true for
-the the selected resident. They can interact with the selected mutant, another
-resident, or another mutant. Thus, in each updating stage there are five
-possible combinations of pairs. These are illustrated by
+model or the learner. Given that they can interact with only one other member of
+the population, they can interact either with each other or either can interact
+with another resident or with another mutant. Thus, in each updating stage there
+are five possible combinations of pairs. These are illustrated by
 Figure~\ref{fig:single_pairs}.
@@
   \caption{\textbf{Possible pairings combination in the updating stage, given
-  that individuals interact with only one other member in the population.}. We
-  distinguish between the selected resident and the rest of the residents and we
-  do the same with the mutants. There is a probability that the selected
-  resident interacts with the selected mutant.}\label{fig:single_pairs}
+  that individuals interact with only one other member in the population.}}
+  \label{fig:single_pairs}
 \end{figure}
@@
 
-Given that $N\!-\!k$ players use the resident strategy $S_1\!=\!(y_1,p_1,q_1)$
+Given that $N\!-\!k$ players use the resident strategy $s_{R}\!=\!(y_{R},p_{R},q_{R})$
 and that the remaining $k$ players use the mutant strategy
-$S_2\!=\!(y_2,p_2,q_2)$, the probability that the number of mutants increases by
+$S_{M}\!=\!(y_{M},p_{M},q_{M})$, the probability that the number of mutants increases by
 one in one step of the evolutionary process can be written as
@@
 \begin{align}
-\lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_1,u_2\in\mathcal{U}} x(u_1,u_2)\cdot \rho(u_1,u_2), \\
-\lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_1,u_2\in\mathcal{U}} x(u_1,u_2)\cdot \rho(u_2,u_1).
+\lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M})\cdot \rho(u_{R},u_{M}), \\
+\lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M})\cdot \rho(u_{M},u_{R}).
 \end{align}
@@
 continuation probability \(\delta\). Then the probability of being in any of the
-sixteen outcomes \(\splitatcommas{RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR}\)
+sixteen outcomes \(\splitatcommas{RR, RS, RT, RP, SR, SS, ST, SP, TR, TS, TT, TP, PR, PS, PT, PP}\)
 on the last two rounds are given by,
@@
 
-Proposition~\ref{proposition:last_two_rounds} can be extended to the last \(n\)
-rounds.
-
-\begin{Prop}\label{proposition:last_n_rounds}
-Assume a play between the reactive strategies \(s_1\) and \(s_2\) with a
-continuation probability \(\delta\). Then the probability of being in any of the
-sixteen outcomes \(\splitatcommas{RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR, RR}\)
-on the last two rounds are given by,
-
-\begin{equation}
-  \mathbf{v_{a_1, a_2}} = (1 - \delta) \prod m_{a_1, a_2} \delta^2 \left[\mathbf{v_0}(I_4 - \delta M)^{-1}\right]_{a_1, a_2}
-\end{equation}
-
-for \(m_{a_1, a_2} \in M\) and \(a_1, a_2 \in [1, 4]\).
-\end{Prop}
-
-Equation~\ref{eq:Chi} can also be extended to include interactions with two other
-individuals. The possible pairings are illustrated by Figure~\ref{}.
+Equation~\ref{eq:Chi} can also be extended to include interactions with two
+other individuals. The possible pairings are illustrated by
+Figure~\ref{fig:pissible_two_pairs}.
 
@@
   \includegraphics[width=.65\textwidth]{static/tree.tex}
-  \caption{The tree}\label{fig:two_pairs}
-\end{figure}
-
-\section{Verifying analytical results with simulations}
-
-The analytical results presented in this work have been verified with simulations.
-More specifically the probabilities of Equation~(\ref{Eq:LastRound}),
-
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[width=.75\textwidth]{static/stationary_four.pdf}
-\caption{}
-\end{figure}
-
-Proposition~\ref{proposition:last_two_rounds},
-
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[width=\textwidth]{static/stationary_sixteen.pdf}
-\caption{}
-\end{figure}
-
-% \begin{Prop}\label{proposition:last_n_rounds}
-
-% \end{Prop}
-
-% \section{Evolutionary dynamics in finite population}
-
-% The Prisoner's Dilemma (PD) is a two player no-cooperative game. The players
-% simultaneously and independently make a decision to Cooperate (C) or to defect
-% (D). The payoffs for both players depend on their actions and the actions of
-% their opponents. More specifically, the payoffs are given by:
-
-% \begin{equation}\label{eq:pd_normal_form}
-%     \begin{pmatrix}
-%         R & S  \\
-%         T & P
-%     \end{pmatrix}
-% \end{equation}
-
-% The payoffs, \((R, P, S, T)\), are constrained by \(T > R > P > S\) and \(2R > T + S\).
-
-% A special case is the donation game,” where each player can cooperate by providing 
-% a benefit \(b\) to the other player at their cost \(c\), with \(0 < c < b\).
-% Then, \(T=b, R=b-c, S=-c, P=0\), and matrix~(\ref{eq:pd_normal_form}) is give by:
-
-% \begin{equation}\label{eq:donation_normal_form}
-%     \begin{pmatrix}
-%         b-c & -c  \\
-%         b & 0
-%     \end{pmatrix}
-% \end{equation}
-
-% where \(c\) is the cost inflicted to an individual for cooperating and \(b\) is
-% the benefit.
-
-% The questions of evolution become more interesting when repetition is considered,
-% and the history can be accessed when making decisions. The iterated form of the
-% game is called the Iterated Prisoner's Dilemma (IPD). There are several two
-% player repeated games such as the Snowdrift game~\cite{sugden2004economics},
-% Harmony and the Stag Hunt~\cite{skyrms2001stag} game. We present results in all
-% four games.
-
-% The space of strategies in the IPD is infinite. To continue with our
-% evolutionary analysis, we assume herein that individuals at most make use of
-% simple \textbf{reactive strategies}. Reactive strategy are a set of memory-one
-% strategies that only take into account the previous action of the opponent. An
-% example of a reactive strategy is Tit For Tat. Reactive strategies can be
-% written explicitly as a vector \(\in \R_{3}\). More specifically a reactive
-% strategy \(s\) is given by \(s=(y, p, q)\) where \(y\) is the probability that
-% the strategy opens with a cooperation and \(p, q\) are the probabilities that
-% the strategy cooperates given that the opponent cooperated and defected
-% equivalently.
-
-% A match between two reactive strategies takes the form of a stochastic process,
-% and more specifically, of a Markov chain with four possible states \(CC, CD, DC,
-% DD\) (the possible outcomes of each round). We assume a match between two
-% reactive strategies \(s_{1}=(y_{1}, p_{1}, q_{1})\) and \(s_{2}=(y_{2}, p_{2},
-% q_{2})\), the Markov process is described by the transition matrix \(M\):
-
-% \begin{equation}
-%     \input{static/matrix}
-% \end{equation}
-
-% The long run steady state probability vector \(\mathbf{v}\), which is the
-% solution to \(\mathbf{v} M = \mathbf{v}\), can be combined with the payoff of
-% matrix (\ref{eq:pd_normal_form}) (denoted as \(U\)) to give the expected payoffs
-% for each player. More specifically, the payoffs for a reactive strategy \(s_1\)
-% against an opponent \(s_2\) is:
-
-% \begin{equation}
-%     \mathbf{v}(s_1, s_2) \cdot U
-% \end{equation}
-
-% where
-
-% \begin{equation}
-%     U = \{R, S, T, P\}.
-% \end{equation}
-
-% \subsection{Evolutionary Dynamics}
-
-% In evolution context we consider a population of \(N\) players,
-% where \(N\) is even, and where mutations are sufficiently rare. At any point
-% in time the there are at most two different strategies are present in
-% the population. Suppose there are \(N - k\) players who use the strategy \(s_{1}=(y_{1},
-% p_{1}, q_{1})\), whereas \(k\) players use the strategy \(s_{2}=(y_{2}, p_{2},
-% q_{2})\). We refer to these two player types as ‘residents’ and ‘mutants’,
-% respectively.
-
-% Each step of the evolutionary process consists of two stages, a game stage and
-% an updating stage.
-
-% \begin{enumerate}
-%     \item In the game stage, each player is randomly matched with some
-%     other player in the population to interact in one instance of the IPD.
-%     \item In the updating stage, two players are randomly drawn from
-%     the population, a `learner' and a `exemplar'. Given that the learner's payoff
-%     in the last round is $u_L\!\in\! \mathcal{U}$ and that the exemplar's last
-%     round's payoff $u_E\!\in\! \mathcal{U}$, we assume the learner adopts the role
-%     model's strategy with probability
-
-%     \begin{equation} \label{Eq:rho}
-%     \rho(u_L, u_E) = \frac{1}{1\!+\!\exp\big[ \!-\!\beta (u_E\!-\!u_L) \big]}. 
-%     \end{equation}
-
-%     where $\beta\!\ge\!0$ corresponds to the strength of selection.
-% \end{enumerate}
-
-% We iterate this basic evolutionary step until either the mutant strategy goes
-% extinct, or until it fixes in the population (in which case the mutant strategy
-% becomes the new resident strategy). After either outcome, we introduce a new
-% mutant strategy $s'_2\!=\!(y_2',p_2',q_2')$ (uniformly chosen from all reactive
-% strategies at random), and we set the number of mutants to $k\!=\!1$. This
-% process of mutation and fixation/extinction is then iterated many times.
-
-% We compare this process for what we defined as {\bf stochastic payoff
-% evaluation} with the analogous process where players update their strategies
-% with respect to their {\bf expected} payoffs,
-
-% \begin{equation} \label{eq:ExpPay}
-% \begin{array}{lcrcr}
-% \pi_1 &= &\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(s_1,s_1),\mathbf{U}\rangle	& + &\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(s_1,s_2),\mathbf{U}\rangle,\\[0.5cm]
-% \pi_2 &= &\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(s_2,s_1),\mathbf{U}\rangle & + &\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(s_2,s_2),\mathbf{U}\rangle.\\
-% \end{array}
-% \end{equation}
-
-% In the limit of no discounting, $\delta\!\rightarrow\! 1$, this process based on
-% expected payoffs has been considered in~\cite{imhof2010stochastic}.
-
-% \subsection{Stochastic payoff evaluation}
-
-% We define {\bf stochastic payoff} as the average payoff $u\!\in\! \mathcal{U}$ a player
-% receives in the last \(n\) rounds of the game given that they interact with
-% \(m\) players.
-
-% \subsubsection*{Case \(n=m=1\).}
-
-% Initially, consider the situation where \(n=m=1\). The player's stochastic
-% payoff is what they receive in the last round against a single opponent. There
-% only four possible outcomes for the last round, those are \(CC, CD, DC, DD\).
-% Consider two players with reactive strategies $S_1\!=\!(y_1, p_1, q_1$) and
-% $S_2\!=\!(y_2,p_2,q_2)$ who interact in a repeated prisoner's dilemma with
-% continuation probability $\delta$, the probability that are in each of the
-% four possible states in the last round is given by:
-
-% \begin{equation}
-%     \mathbf{v}(s_1,s_2)\!=\!\Big(\mathbf{v}_{R}(s_1,s_2),\mathbf{v}_{S}(s_1,s_2),\mathbf{v}_{T}(s_1,s_2),\mathbf{v}_{P}(s_1,s_2)\Big).
-% \end{equation}
-
-% where,
-
-% \begin{equation} \label{Eq:LastRound}
-%     \setlength{\arraycolsep}{1pt}
-%     \begin{array}{rcl}
-
-%     \mathbf{v}_{R}(S_1,S_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
-%     {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
-
-%     \mathbf{v}_{S}(S_1,S_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
-%     {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
-
-%     \mathbf{v}_{T}(S_1,S_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
-%     {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
-
-%     \mathbf{v}_{P}(S_1,S_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
-%     {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)}.
-%     \end{array}
-% \end{equation}
-
-% \begin{proof}
-
-% Assume a repeated prisoner's dilemma between two reactive strategies. Given the
-% continuation probability $\delta$, probability that the game ends in the after
-% the first round $(1 - \delta)$ and the expected distribution of the four
-% outcomes in the very first round is $\mathbf{v_0}$ defined as. Following the
-% first round the, the outcome of the next rounds with a probability $\delta$
-% is \(M\) such that,
-
-% \[\dots\]
-
-% It can shown that, \(!(1\!-\!\delta)\mathbf{v_0}(I_4-\delta M)^{-1}\) and with
-% some algebraic manipulation we derive to Equation~\ref{Eq:LastRound}.
-
-% \end{proof}
-
-% % \begin{Prop}
-% %     Consider a repeated prisoner's dilemma, with
-% %     continuation probability $\delta$, between players with reactive strategies
-% %     $s_1\!=\!(y_1, p_1, q_1$)  and $s_2\!=\!(y_2,p_2,q_2)$ respectively. Then the
-% %     probability that the $s_1$ player receives the payoff $u\!\in\! \mathcal{U}$ in
-% %     the very last round of the game is given by $v_{u}(S_1,S_2)$, as given by
-% %     Eq.~(\ref{Eq:LastRound}).
-% % \end{Prop}
-
-% is given by $v_{u}(S_1,S_2)$
-
-% Consider two players with reactive strategies $S_1\!=\!(y_1, p_1, q_1$) and $S_2\!=\!(y_2,p_2,q_2)$ who interact in a repeated prisoner's dilemma with continuation probability $\delta$.
-We consider the vector
-
-% \begin{equation}
-% \mathbf{v}(S_1,S_2)\!=\!\Big(v_{R}(S_1,S_2),v_{S}(S_1,S_2),v_{T}(S_1,S_2),v_{P}(S_1,S_2)\Big)\!:=\!(1\!-\!\delta)\mathbf{v_0}(I_4-\delta M)^{-1}.
-% \end{equation}
-
-% Here, $\mathbf{v_0}$ denotes the expected distribution of the four outcomes in the very first round, $I_4$ is the $4\!\times\!4$ identity matrix, and $M$ is the transition matrix of the Markov chain.
-% The entries of $\mathbf{v}$ can be calculated explicitly,
-
-% \begin{equation} \label{Eq:LastRound}
-% \setlength{\arraycolsep}{1pt}
-% \begin{array}{rcl}
-
-% v_{R}(S_1,S_2)	&=	&\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
-% {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
-
-% v_{S}(S_1,S_2)	&=	&\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
-% {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
-
-% v_{T}(S_1,S_2)	&=	&\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
-% {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
-
-% v_{P}(S_1,S_2)	&=	&\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
-% {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)}.
-
-% %x_1(S_1,S_2)&=&\displaystyle\frac{(1\!-\!\delta)y_1+\delta\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big)}{1-\delta^2r_1r_2}\\[0.7cm]
-% %x_2(S_1,S_2)&=&\displaystyle\frac{(1\!-\!\delta)y_2+\delta\Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}{1-\delta^2r_1r_2}.\\
-% \end{array}
-% \end{equation}
-
-% In these expressions, we have used the notation $r_i:=p_i\!-\!q_i$,
-% $\bar{y}_i\!=\!1\!-\!y_i$, $\bar{q}_i:=1\!-\!q_i$, and
-% $\bar{r}_i:=\bar{p}_i\!-\!\bar{q}_i=-r_i$ for $i\!\in\!\{1,2\}$. Let
-% $\mathcal{U}=\{R,S,T,P\}$ denote the set of feasible payoffs in each round,
-% and let $\mathbf{u}\!=\!(R,S,T,P)$ be the corresponding payoff vector. Then
-% one can show the following result.
-
-% \includestandalone[width=0.35\textwidth]{static/expected_stochastic}
-
-% \includestandalone{static/stochastic}
-
-% \begin{figure}[!htbp]
-%     \centering
-%     \includegraphics[width=\textwidth]{static/expected_and_stochastic_for_donation_game.pdf}
-%     \caption{Blah blah.}
-% \end{figure}
-
-% \section{Pairwise imitation dynamics under stochastic payoff evaluation}
-
-% \subsection{Basic setup} \label{Sec:BasicSetup}
-% In the following we consider a population of size $N$, where $N$ is even.
-We assume mutations are sufficiently rare such that at any point in time at most two different strategies are present in the population.
-Suppose there are $N\!-\!k$ players who use the strategy $S_1\!=\!(y_1,p_1,q_1)$, whereas $k$ players use the strategy $S_2\!=\!(y_2,p_2,q_2)$.
-We refer to these two player types as `residents' and `mutants', respectively.
-
-% Each step of the evolutionary process consists of two stages, a game stage and an updating stage.
-In the game stage, each player is randomly matched with some other player in the population to interact in one instance of the repeated prisoner's dilemma.
-% In the updating stage, two players are randomly drawn from the population, a `learner' and a `role model'.
-Given that the learner's payoff in the last round is $u_L\!\in\!
-\mathcal{U}$ and that the role model's last round's payoff $u_M\!\in\!
-\mathcal{U}$, we assume the learner adopts the role model's strategy with probability
-% \begin{equation} \label{Eq:rho}
-% \rho(u_L, u_M) = \frac{1}{1\!+\!\exp\big[ \!-\!\beta (u_M\!-\!u_L) \big]}. 
-% \end{equation}
-% The parameter $\beta\!\ge\!0$ corresponds to the strength of selection.
-
-% We iterate this basic evolutionary step until either the mutant strategy goes extinct, or until it fixes in the population (in which case the mutant strategy becomes the new resident strategy).
-After either outcome, we introduce a new mutant strategy $S'_2\!=\!(y_2',p_2',q_2')$ (uniformly chosen from all reactive strategies at random), and we set the number of mutants to $k\!=\!1$.
-This process of mutation and fixation/extinction is then iterated many times.
-
-% We compare this process for stochastic payoff evaluation with the analogous process where players update their strategies with respect to their {\it expected} payoffs,
-% \begin{equation} \label{Eq:ExpPay}
-% \begin{array}{lcrcr}
-% \displaystyle \pi_1	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(S_1,S_1),\mathbf{u}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(S_1,S_2),\mathbf{u}\rangle,\\[0.5cm]
-% \displaystyle \pi_2	&=	&\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(S_2,S_1),\mathbf{u}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(S_2,S_2),\mathbf{u}\rangle.\\
-% \end{array}
-% \end{equation}
-% In the limit of no discounting, $\delta\!\rightarrow\! 1$, this process based on expected payoffs has been considered in~\cite{imhof2010stochastic}.
-
-% \subsection{Fixation probabilities under stochastic payoff evaluation}
-% Given that $N\!-\!k$ players use the resident strategy $S_1\!=\!(y_1,p_1,q_1)$ and that the remaining $k$ players use the mutant strategy $S_2\!=\!(y_2,p_2,q_2)$, the probability that the number of mutants increases by one in one step of the evolutionary process can be written as
-% \begin{equation}
-% \lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_1,u_2\in\mathcal{U}} x(u_1,u_2)\cdot \rho(u_1,u_2).
-% \end{equation}
-% In this expression, $(N\!-\!k)/N$ is the probability that the randomly chosen learner is a resident, and $k/N$ is the probability that the role model is a mutant.
-The sum corresponds to the total probability that the learner adopts the role model's strategy over all possible payoffs $u_1$ and $u_2$ that the two player may have received in their respective last rounds.
-We use $x(u_1,u_2)$ to denote the probability that the randomly chosen resident obtained a payoff of $u_1$ in the last round of his respective game, and that the mutant obtained a payoff of $u_2$.
-Given that the payoffs are $u_1$ and $u_2$, the imitation probability is then given by $\rho(u_1,u_2)$, as specified by Eq.~(\ref{Eq:rho}).
-The probability that the respective payoffs of the players are given by $u_1$ and $u_2$ can be calculated as
-% \begin{equation}
-% \setlength{\arraycolsep}{1pt}
-% \begin{array}{llrl}
-% x(u_1,u_2)	 &=&\displaystyle \frac{1}{N\!-\!1}\cdot  &v_{u_1}(S_1,S_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F}\\[0.5cm]
-% &+	
-% &\displaystyle \left(1\!-\!\frac{1}{N\!-\!1}\right)  
-% &\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} v_{u_1}(S_1,S_2) v_{u_2}(S_2,S_2) + 
-%  \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} v_{u_1}(S_1,S_2) v_{u_2}(S_2,S_1)\right.\\[0.5cm]
-% &&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} v_{u_1}(S_1,S_1) v_{u_2}(S_2,S_2) + 
-%  \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} v_{u_1}(S_1,S_1) v_{u_2}(S_2,S_1)\right].
-% \end{array}
-% \end{equation}
-% The first term on the right side corresponds to the case that the learner and the role model happened to be matched during the game stage, which happens with probability $1/(N\!-\!1)$.
-In that case, we note that only those payoff pairs can occur that are feasible in a direct interaction, $(u_1,u_2)\in \mathcal{U}^2_F:=\big\{ (R,R), (S,T), (T,S), (P,P) \big\}$, as represented by the respective indicator function.
-Otherwise, if the learner and the role model did not interact directly, we need to distinguish four different cases, depending on whether the learner was matched with a resident or a mutant, and depending on whether the role model was matched with a resident or a mutant.
-
-% Analogously, we can calculate the probability that the number of mutants decreases by one in one step of the evolutionary process. This probability is
-% \begin{equation}
-% \lambda^-_k=\frac{N\!-\!k}{N}\cdot\frac{k}{N} \sum_{u_1,u_2\in\mathcal{U}} x(u_1,u_2)\cdot \rho(u_2,u_1).
-% \end{equation}
-% The fixation probability of the mutant strategy then takes the standard form \cite{nowak2004emergence},
-% \begin{equation}
-% \varphi = \frac{1}{1+\sum_{i=1}^{N-1}\prod_k^i \frac{\lambda^-_k}{\lambda^+_k}}.
-% \end{equation}
-
-% \subsection{Invasion analysis of ALLD into GTFT}
-
-% In the following, we apply the above formalism to calculate how easily a single ALLD mutant can invade into a resident population with strategy GTFT.
-In that case, $S_1=(1,1,q)$, $S_2\!=\!(0,0,0)$, and $k\!=\!1$.
-When two GTFT players interact in the game, their respective probabilities for each of the four outcomes in the last round simplify to
-% \begin{equation}
-% \begin{array}{cc}
-% v_R(GTFT,GTFT)=1, &v_S(GTFT,GTFT)=0,\\
-% v_T(GTFT,GTFT)=0, &v_P(GTFT,GTFT)=0.
-% \end{array}
-% \end{equation}
-% On the other hand, if an ALLD player interacts with a GTFT player, the respective probabilities according to Eq.~(\ref{Eq:LastRound}) become
-% \begin{equation}
-% \begin{array}{ll}
-% v_R(ALLD,GTFT)=0,	&v_S(ALLD,GTFT)=0,\\
-% v_T(ALLD,GTFT)=1\!-\!\delta+\delta q,~~~ 	&v_P(ALLD,GTFT)=\delta(1\!-\!q).
-% \end{array}
-% \end{equation}
-% As a consequence, we obtain the following probabilities $x(u_1,u_2)$ that the payoff of a randomly chosen GTFT player is $u_1$ and that the payoff of the ALLD player is $u_2$,
-% \begin{equation}
-% \begin{array}{l}
-% \displaystyle x(R,T)=\frac{N-2}{N-1} \cdot (1\!-\!\delta\!+\!\delta q),\\[0.5cm]
-% \displaystyle x(R,P)=\frac{N-2}{N-1} \cdot \delta(1\!-\!q),\\[0.5cm]
-% \displaystyle x(S,T)=\frac{1}{N-1} \cdot (1\!-\!\delta\!+\!\delta q), \\[0.5cm]
-% \displaystyle x(P,P)=\frac{1}{N-1} \cdot \delta(1\!-\!q). \\[0.5cm]
-% \displaystyle x(u_1,u_2)=0	~~~\text{for all other payoff pairs}~(u_1,u_2).
-% \end{array}
-% \end{equation}
-% As a consequence, we can calculate the ratio of transition probabilities as 
-% \begin{equation}
-% \frac{\lambda_1^+}{\lambda_1^-}=\displaystyle \frac{\displaystyle \frac{N-2}{N-1} \cdot \left(\frac{1\!-\!\delta\!+\!\delta q}{1\!+\!\exp[-\beta(T\!-\!R)]}
-% \!+\! \frac{\delta(1-q)}{1\!+\!\exp[-\beta(P\!-\!R)]}\right)
-% + \frac{1}{N\!-\!1} \left(\frac{1-\delta+\delta q}{1+\exp[-\beta (T\!-\!S)]}
-% \!+\! \frac{\delta(1-q)}{2}\right)}
-% {\displaystyle \frac{N-2}{N-1} \cdot \left(\frac{1\!-\!\delta\!+\!\delta q}{1\!+\!\exp[-\beta(R\!-\!T)]}
-% \!+\! \frac{\delta(1-q)}{1\!+\!\exp[-\beta(R\!-\!P)]}\right)
-% + \frac{1}{N\!-\!1} \left(\frac{1-\delta+\delta q}{1+\exp[-\beta (S\!-\!T)]}
-% \!+\! \frac{\delta(1-q)}{2}\right)}.
-% \end{equation}
-
-% \noindent
-% In particular, in the limit of strong selection $\beta \rightarrow \infty$ and large populations $N\!\rightarrow \infty$, we obtain
-% \begin{equation}
-% \frac{\lambda_1^+}{\lambda_1^-}=
-% \frac{ 1\!-\!\delta+\delta q}{\delta (1-q)}.
-% \end{equation}
-% This ratio is smaller than 1 (such that ALLD is disfavored to invade) if $q<1\!-\!1/(2\delta)$.
-For infinitely repeated games, $\delta\!\rightarrow\!1$, this condition becomes $q\!<\!1/2$ (for $q\!=\!1/2$, the payoff of the ALLD player is $T\!>\!R$ for half of the time, and it is $P\!<\!R$ for the other half.
-The probability that the number of mutants increase by one equals the probability that the mutant goes extinct).
-
-% % \subsection{Simulation results}
-
-% % \FigEvoProc -- \FigResultsOverPara~show simulation results for the above described process.
-\FigEvoProc~depicts the evolving conditional cooperation probabilities $p$ and $q$ (assuming that the discount factor~$\delta$ and the benefit $b$ are comparably high).
-The left panel corresponds to the standard scenario considered in the previous literature.
-It considers players who use expected payoffs (\ref{Eq:ExpPay}) to update their strategies.
-The right panel shows the scenario considered herein, in which players update their strategies based on their last round's payoff.
-The figure suggests that when updating is based on expected payoffs, players tend to be more generous (their $q$-values are higher on average).
-In addition, players are generally more cooperative.
-
-% % \begin{figure}[t!]
-% % \centering
-% % \includegraphics[width=0.75\textwidth]{Fig1} 
-% % \caption{{\bf Evolutionary dynamics under expected payoffs (left) and stochastic payoffs (right).} 
-% % We have run two simulations of the evolutionary process described in Section~\ref{Sec:BasicSetup} for $T\!=\!10^7$ time steps.
-For each time step, we have recorded the current resident population ($y,p,q$).
-Since simulations are run for a relatively high continuation probability of $\delta\!=\!0.999$, we do not report the players' initial cooperation probability $y$.
-The graphs show how often the resident population chooses each combination ($p,q$) of conditional cooperation probabilities in the subsequent rounds.
-({\bf A}) If players update based on their expected payoffs, the resident population typically applies a strategy for which $p\!\approx\!1$ and $q\!\le\!1\!-\!c/b\!=\!0.9$.
-The cooperation rate within the resident population (averaged over all games and over all time steps) is close to 100\%.
-({\bf B}) When players update their strategies based on their realized payoffs in the last round, there are two different predominant behaviors.
-The resident population either consists of defectors (with $p\!\approx\!q\!\approx\!0$) or of conditional cooperators.
-In the latter case, the maximum level of $q$ consistent with stable cooperation is somewhat smaller compared to the expected-payoff setting, $q\!<\!0.5$.
-Also the resulting cooperation rate is smaller.
-On average, players cooperate roughly in half of all rounds.
-Parameters: $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.999$.}
-% % \end{figure}
-
-% % To obtain some intuition for this result, we have recorded which mutant strategies typically invade into an ALLD population, and which mutants invade into a GTFT population, for each of the two scenarios (\FigInvAnalysis).
-Compared to expected-payoff updating, the stochastic scenario drastically reduces the stability of GTFT.
-On average, it takes fewer mutants until a GTFT population is successfully invaded.
-Moreover, successful mutants  tend to be less cooperative.
-
-% % \begin{figure}[t!]
-% % \centering
-% % \includegraphics[width=0.75\textwidth]{Fig2} ~\\[-0.6cm]
-% % \caption{{\bf Invasion into ALLD and into GTFT for the two evolutionary scenarios.} 
-% % For this figure, we either consider a resident population of ALLD players (top) or of GTFT players (bottom).
-In each case, we have run 1,000 independent simulations of the evolutionary process described in Section~\ref{Sec:BasicSetup}.
-The colored dots within the four panels depict which mutant strategies successfully invaded into the respective resident population.
-The black arrow indicates the average over all successful mutant strategies.
-In addition, we have recorded how many mutant strategies need to be introduced into the population until the resident becomes replaced.
-Stochastic payoffs tend to increase the time until ALLD is invaded, and they reduce the time until GTFT is invaded.
-Parameters are the same as in \FigEvoProc.
-Strategies are defined as $ALLD\!=\!(0,0,0)$ and $GTFT=(1,1,1/3)$.}
-% % ~\\[0.4cm]
-% % \includegraphics[width=0.8\textwidth]{Fig3} 
-% % \caption{{\bf The evolution of cooperation for different parameter values.} 
-% % While the previous figures depict the evolutionary outcome for fixed parameter values, here we vary the benefit of cooperation $b$, the strength of selection $\beta$, and the expected number of rounds, $1/(1\!-\!\delta)$.
-In all cases, stochastic payoff evaluation tends to reduce the evolving cooperation rates.
-Unless explicitly varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$.
-Simulations are run for $T\!=\!5\times 10^6$ time steps for each parameter combination.}
-% % \end{figure}
-
-% % Finally, we have also explored how the evolving cooperation rates change as we vary the benefit $b$, the selection strength $\beta$, and the discount factor $\delta$ (\FigResultsOverPara).
-In all three cases, we find that the two scenarios yield similar cooperation rates when the respective parameters are small.
-Once there is a high benefit to cooperation, strong selection, or a high expected number of rounds, updating based on expected payoffs yields higher cooperation rates.
-
-\clearpage
-\newpage
+  \caption{\textbf{Possible pairings combination in the updating stage, given
+  that individuals interact with two other members in the population.}}
+  \label{fig:pissible_two_pairs}
+\end{figure}
+
+Simulating the evolutionary process for longer memory quickly becomes
+computationally untractable. Our methodology could be extended to include more
+turns and more interactions. However, for the purpose of this work we explore
+the cases only up to two turns and two interactions.
+
+\section{Evolutionary dynamics under limited memory payoffs for various social dilemmas}\label{appendix:further_simulation_results}
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=\textwidth]{static/rounds_two_by_two_games.pdf}
+  \caption{{\bf Evolutionary dynamics under last two rounds payoffs for various social dilemmas.} 
+  We have run several simulations of the evolutionary process described in
+  section~\ref{section:model} for $T\!=\!10^7$ time steps. The graphs show how
+  often the resident population chooses each combination ($p,q$) of conditional
+  cooperation probabilities in the subsequent rounds. We vary the temptation
+  payoff \(T \in \{-1, -0.6, -0.2,  0.2, 0.6, 1, 1.4, 1.8, 2.2, 2.6, 3\}\)
+  across the \(x\) axis, and  \(S \in \{2, 1.6, 1.2, 0.8, 0.4, 0, -0.4, -0.8,
+  -1.2, -1.6, -2\}\) across the \(y\) axis. Parameters: $N\!=\!100$,
+  $\beta\!=\!10$, $\delta\!=\!0.999$.}
+  \label{fig:last_two_rounds_two_by_two}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=\textwidth]{static/opponents_two_by_two_games.pdf}
+  \caption{{\bf Evolutionary dynamics under last round payoffs with two members of the population for various social dilemmas.} 
+  We have run several simulations of the evolutionary process described in
+  section~\ref{section:model} for $T\!=\!10^7$ time steps. The graphs show how
+  often the resident population chooses each combination ($p,q$) of conditional
+  cooperation probabilities in the subsequent rounds. We vary the temptation
+  payoff \(T \in \{-1, -0.6, -0.2,  0.2, 0.6, 1, 1.4, 1.8, 2.2, 2.6, 3\}\)
+  across the \(x\) axis, and  \(S \in \{2, 1.6, 1.2, 0.8, 0.4, 0, -0.4, -0.8,
+  -1.2, -1.6, -2\}\) across the \(y\) axis. Parameters: $N\!=\!100$,
+  $\beta\!=\!10$, $\delta\!=\!0.999$.}
+  \label{fig:last_two_opponents_two_by_two}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=\textwidth]{static/merged_plot_rounds_opponents_two.pdf}
+  \caption{{\bf Evolutionary dynamics under last two rounds payoffs with two members of the population for various social dilemmas.} 
+  We have run several simulations of the evolutionary process described in
+  section~\ref{section:model} for $T\!=\!10^7$ time steps. The graphs show how
+  often the resident population chooses each combination ($p,q$) of conditional
+  cooperation probabilities in the subsequent rounds. We vary the temptation
+  payoff \(T \in \{-1, -0.6, -0.2,  0.2, 0.6, 1, 1.4, 1.8, 2.2, 2.6, 3\}\)
+  across the \(x\) axis, and  \(S \in \{2, 1.6, 1.2, 0.8, 0.4, 0, -0.4, -0.8,
+  -1.2, -1.6, -2\}\) across the \(y\) axis. Parameters: $N\!=\!100$,
+  $\beta\!=\!10$, $\delta\!=\!0.999$.}
+  \label{fig:last_two_rounds_two_opponents}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=\textwidth]{static/difference_in_cooperation.pdf}
+  \caption{{\bf The difference in cooperate rates for each method compared to the
+  expected payoffs.}}
+  % \label{fig:last_two_rounds_two_opponents}
+\end{figure}
 

### 2021-05-25 main.tex
first draft to CH

@@
 \begin{abstract}
-Repeated games have been the centerpiece when studying the evolution of human
-cooperation. Most of the existing models work with idealized assumptions.
-Individuals are assumed to interact with a representative sample of the
-population, and they remember all interactions they participate in. In real
-life, however, individuals do not always remember all interactions they had with
-everyone in their population. Instead, they rather recall the last interaction
-they had.
+Repeated games are vastly used in evolutionary game theory to explain
+cooperation in a variety of environments. Although existing models have shaped
+our understanding of human cooperation, the often work with idealized
+assumptions. In an evolutionary process, individuals imitate other individuals
+of the population based on their fitness. It is commonly assumed that an
+individual computes their fitness after interacting with a representative sample
+of the population, and remembering all the interactions they participated in. In
+real life, we do not always remember all our interactions, instead we rather
+recall our most recent ones.
 
-In this work we explore the effect on the evolution of cooperation if
-individuals only remember a minimum of social information. We present results
-for a series of simulations were we vary the benefit of cooperation and the
-strength of selection. We present results for the most commonly used classes of
-social dilemmas, and we show that even though cooperation can evolve when
-individuals only use a minimum of information, the evolving cooperation rates
-are typically lower than in the classical scenario.
+Here, we introduce a framework that allows individuals to estimate their fitness
+based on a minimum of social information. We explore the difference between the
+classical framework and ours using computer simulations. We present results for
+the most commonly used classes of symmetric \(2 \times 2\) games. The
+simulations show that, in the prisoner's dilemma individuals with limited memory
+tend to adopt less generous strategies and they achieve less cooperation than in
+the classical scenario. In contrast, in the stag-hunt and the harmony game, the
+impact of memory is less striking. Here individuals with limited memory perform
+nearly as well as individuals with full memory. Finally, we observe that the
+snowdrift game is the only game for which cooperation can be underestimated in
+the classical scenario.
 \end{abstract}
@@
 evolution of cooperation. Why is it that some individuals choose to help others,
-increasing their payoff, at the expense of decreasing one’s own? In evolutionary
+increasing their payoff, at the expense of decreasing one's own? In evolutionary
 game theory individuals are not required to be rational, instead they adapt
@@
 with a representative sample of the population, and remembering all the
-interactions they participated in~\cite{Nowak2006}. Thus, they imply that
-individuals have a perfect memory. However, when modelling how individuals make
-decisions in each turn they are assumed to have very limited memory. To be
-precise, most of the works in the literature focus on naive subjects who can
-only choose from a restricted set of strategies~\cite{Nowak1992tit}, or who do
-not remember anything beyond the outcome of the very last round~\cite{Baek2016}.
-Note that there are a few notable exceptions~\cite{Hauert1997, Stewart2016}.
-The perfect memory assumption is not only unrealistic but it also creates
-this curious inconsistency.
+interactions they participated in~\cite{Nowak2006, nowak2004emergence,
+Hilbe2013, Hilbe2018,Reiter2018, garcia2018no, Nowak1992tit, Baek2016}. This
+implies that individuals have a perfect memory. However, when modelling how
+individuals make decisions in each turn they are assumed to have very limited
+memory. To be precise, most of the studies in the literature focus on naive
+subjects who can only choose from a restricted set of
+strategies~\cite{Nowak1992tit}, or who do not remember anything beyond the
+outcome of the very last round~\cite{Baek2016}. Note that there are a few
+notable exceptions~\cite{Hauert1997, Stewart2016}.
 
-This has led us to question the robustness of our understanding of cooperation.
-In this work we explore whether cooperation can evolve if individuals compute
-their fitness based on a minimum of social information. Though we are not the
+Studies that make use of this classical framework report that cooperation can
+substantially evolve. However, the inconsistency in the memory size of players
+leads us to question the robustness of our understanding of cooperation. To this end, we
+propose a framework in which individuals, similar to the decisions at each turn,
+estimate their fitness based on a minimum of information. Though we are not the
 first to question the assumptions of estimating fitness~\cite{Roca2006}, we are
-the first to explore the effect of payoff memory. %ToDo add more information
+the first to explore the effect of memory.
 
-We first consider two extreme scenarios, the classical scenario of the expected
-payoffs and the alternative scenario where individuals update their strategies
+We first consider two extreme scenarios, the classical scenario
+and the alternative scenario where individuals update their strategies
 only based on the very last payoff they obtained. We observe that individuals 
 with limited memory tend to adopt less generous strategies and they achieve less
-cooperation than in the classical scenario. We obtain similar results when we
-consider that individuals update their strategies based on more social
+cooperation when interacting in a prisoner's dilemma. We obtain similar results when we
+consider that individuals update their strategies based on more
 information. More specifically, up to the last two payoffs they obtained when
-interacting with up to two different members of the population.
+interacting with up to two different members of the population. We extend our
+approach to the rest of the symmetric \(2 \times 2\) games.
 
@@
 ``individual'' are used interchangeably here.} where \(N\) is even and mutations
-are sufficiently rare. Therefor, at any point in time there are at most two
+are sufficiently rare. Therefore, at any point in time there are at most two
 different strategies present in the population; a \textit{resident} strategy and
@@
 in the population. They engage in a match where each subsequent turn
-occurs with a fixed probability $\delta$. At each turn players can choose to
+occurs with a fixed probability $\delta$. At each turn players choose independently to
 either cooperate (\(C\)) or to defect (\(D\)), and the payoffs of the turn
@@
 \(P\). If one cooperates but the other defects, the defector receives the
-temptation \(T\) payoff, whereas the cooperator receives the sucker's
+temptation payoff \(T\), whereas the cooperator receives the sucker's
 payoff \(S\). We denote the feasible payoff of each turn as \(\mathcal{U} =
@@
 that only take into account the previous action of the opponent. They can be
-written explicitly as a vector \(\in \R_{3}\), more specifically, a reactive
-strategy \(s\) is given by \(s=(y, p, q)\) where \(y\) is the probability that
+written explicitly as a vector in \(\R_{3}\), more specifically, a reactive
+strategy \(s\) is given by \(s=(y, p, q)\). The parameter \(y\) is the probability that
 the strategy opens with a cooperation and \(p, q\) are the probabilities that
@@
 In the updating stage, two players are randomly drawn from the population, a
-`learner' and a `role model'. Given that the learner's payoff $u_L\!\in\!
-\mathcal{U}$ and that the role model's payoff $u_{RL}\!\in\! \mathcal{U}$.
-The learner adopts the role model's strategy based on the Fermi
-distribution function,
+`learner' and a `role model'. Given the learner's payoff $u_L\!\in\!
+\mathcal{U}$ and the role model's payoff $u_{RL}\!\in\! \mathcal{U}$,
+the learner adopts the role model's strategy with probability,
 
@@
 
-where $\beta\!\ge\!0$ is the intensity of selection. The updating payoffs of the
-learner and the role model are conventionally correspond to their expected
-payoffs. The expected payoff of a strategy is the mean payoff of the strategy of
-engaging in repeated games with all other members of the population. The expected payoffs
-assume perfect memory. In this work we define a new a set of updating
-payoffs. We reefer to these as the limited memory payoffs, and we compare them
-to the classical scenario.
+where $\beta\!\ge\!0$ is the strength of selection. For small values of $\beta$
+the imitation probability is independent of the strategies of the involved
+players. As the value of $\beta$ increases, the more likely it is that the
+learner adopts only strategies that yield a higher payoff.
+Conventionally the updating payoffs of the learner and the role model are based
+on their expected payoffs. A player's expected payoff is the mean payoff the
+player yields after engaging in matches of multiple turns with each member of
+the population. At each match a player bases their next turn decision only on
+the previous action of the opponent, however, the same player bases their
+expected payoffs on the outcomes of all their matches. Thus, a player is assumed
+to have limited and perfect memory at the same time. We propose a new a set of
+updating payoffs where it is also assumed that a player has limited memory.
+We referee to these as the limited memory payoffs.
 
@@
 process of mutation and fixation/extinction is then iterated many times.
-More details on our methodology are found in Appendix~\ref{appendix:methods}.
 
-In order to account for the effect of the updating payoffs approaches, we
-explore the cooperation rate within the resident population over multiple
-generations. In the next section we present a series of simulation results.
-To account for the various types of social behaviour we also present
-results on multiple social dilemmas.
+In order to account for the effect of the updating payoffs we simulate the
+evolutionary process and record which strategies the players adopt over time
+based on (i) the expected payoffs (ii) the limited memory payoffs. We compare
+the difference in the cooperation rate within the resident population for the
+two approaches. To account for the various types of social behaviour we also
+present results on multiple social dilemmas.
 
@@
 benefit \(b\) to the other player at their cost \(c\), with \(0 < c < b\).
-Thus, the feasible payoffs in each round are \(T=b, R=b-c, S=-c, P=0\).
+Thus, the feasible payoffs in each round are \(\mathcal{U} = \{b-c, -c, b, 0\}\).
 
@@
     We have run two simulations of the evolutionary process described in
-    section~\ref{section:model} for $T\!=\!10^7$ time steps. For each time step,
+    section~\ref{section:model} for $t\!=\!10^7$ time steps. For each time step,
     we have recorded the current resident population ($y,p,q$). Since simulations
@@
     \centering
-    \includegraphics[width=.8\textwidth]{static/expected_for_beta.pdf}
-    \includegraphics[width=.8\textwidth]{static/stochastic_for_beta.pdf}
+    \includegraphics[width=\textwidth]{static/expected_for_beta.pdf}
+    \includegraphics[width=\textwidth]{static/stochastic_for_beta.pdf}
   \end{subfigure}%
@@
     \centering
-    \includegraphics[width=.8\textwidth]{static/cooperation_rate_over_b.pdf}
+    \includegraphics[width=\textwidth]{static/cooperation_rate_over_b.pdf}
   \end{subfigure}
@@
   $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run
-  for $T\!=\!5\times 10^7$ time steps for each parameter
+  for $t\!=\!5\times 10^7$ time steps for each parameter
   combination.}\label{fig:cooperation_rate_over_benefit}
@@
     \centering
-    \includegraphics[width=.8\textwidth]{static/expected_for_selection_strenght.pdf}
-    \includegraphics[width=.8\textwidth]{static/stochastic_for_selection_strenght.pdf}
+    \includegraphics[width=\textwidth]{static/expected_for_selection_strenght.pdf}
+    \includegraphics[width=\textwidth]{static/stochastic_for_selection_strenght.pdf}
   \end{subfigure}%
@@
     \vspace{.3cm}
-    \includegraphics[width=.8\textwidth]{static/cooperation_rate_over_betas.pdf}
+    \includegraphics[width=\textwidth]{static/cooperation_rate_over_betas.pdf}
   \end{subfigure}
@@
 $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
-$T\!=\!5\times 10^7$ time steps for each parameter
+$t\!=\!5\times 10^7$ time steps for each parameter
 combination.}\label{fig:cooperation_rate_over_betas}
@@
 donation game. In order to broaden our understanding of the updating payoffs on
-different forms of possible human interactions we extend our approach to other
+different forms of possible human interactions we extend our approach to all
 \(2 \times 2\) symmetric games. More specifically, we apply our analysis to four
-different classes of games as given by Table~\ref{table:social_dilemmas}.
+different classes of games; the harmony, the snowdrift, the prisoner's dilemma
+and the stag-hunt games.
 
-\begin{table}[!htbp]
-  \begin{center}
-    \resizebox{.4\textwidth}{!}{
-    \renewcommand{\arraystretch}{1.5}
-      \begin{tabular}{ccc}
-  \toprule
-  & \textbf{social dilemmas} & \textbf{preference ordering} \\
-  \midrule
-(i) & harmony  & \(R > T > S > P\) \\
-(ii) & stag hunt  & \(R > T > P > S\) \\
-(iii) & prisoner dilemma  & \(T > R > P > S\) \\
-(iv) & snowdrift  & \(T > R > S > P\) \\
-  \bottomrule
-      \end{tabular}}
-  \end{center}
-\caption{\textbf{Social dilemmas and preference ordering}. The four classes of
-social dilemmas we explore in this work.}
-\label{table:social_dilemmas}
-\end{table}
-
-We compare the results of the evolutionary process described in
-section~\ref{section:model} when the updating payoffs are based on the expected
+We compare the results of the evolutionary process when the updating payoffs are based on the expected
 payoffs, and on the last round payoffs for all the four possible classes of
-games. The results of the simulations are given in
-Figures~\ref{fig:expected_two_by_two} and~\ref{fig:last_round_two_by_two}
-respectively.
-
-In Figure~\ref{fig:expected_two_by_two} each sub figure represents a run of
-evolutionary process for a different set of values for \(\mathcal{U}\). Without
-loss of generality we set \(R=1\) and \(P=0\)~\cite{Martinez2012, Roca2009}, and
-we vary the values of the temptation payoff \(T\) (across the \(x\)-axis) and of
-the sucker's payoff \(S\) (across the \(y\)-axis). Starting at the upper left
-corner and proceeding clockwise the quadrants correspond to the harmony, the
-snowdrift, the prisoner's dilemma and the stag hunt games.
-
-The harmony game represents are social situation without conflict. It is in the
-best interest for both players to cooperate as \(R > T\) and \(S > P\).
-Figure~\ref{fig:expected_two_by_two} confirms this. We observe that for most
-harmony game runs the resident population overwhelmingly applies a strategy for
-which \(p\) and \(p\) are \(\approx 1\). In the case of \(T = R = 1\) we observe
-that the resident strategies become slightly less generous. Against a cooperating
-strategy defecting and cooperation yield the same result. The lower \(q-\)values
-could indicate that the resident strategy is less generous to defectors to avoid
-being invaded.
-
-The snowdrift game describes a situation similar to that of the prisoner's
-dilemma. Cooperation results in a benefit to the opposing player but entailing a
-cost to the cooperator. However, in the snowdrift game individuals obtain
-immediate direct benefits from the cooperative acts which leads to \(S > P\). In
-the snowdrift game more cooperation emerges compared to the prisoner's dilemma.
-Defection is never a resident strategy and for the shame values of temptation
-the overall \(q-values\) are higher. It can be seen that the \(p-values\) are
-lower as individuals have free room to defect after receiving a cooperation.
-
-The last class of games we present are for the stag hunt game. In the stag hunt
-both players benefit for mutual cooperation and defection, however, \(R > P\)
-and thus the resident strategies with the highest fitness are the cooperative
-ones. The results for the stag hunt class are similar to those of the harmony
-game.
+games (Figures~\ref{fig:expected_two_by_two}
+and~\ref{fig:last_round_two_by_two}). In Figure~\ref{fig:expected_two_by_two}
+each sub figure represents a run of evolutionary process for a different set of
+values for \(\mathcal{U}\). Without loss of generality we set \(R=1\) and
+\(P=0\)~\cite{Martinez2012, Roca2009}, and we vary the values of the temptation
+payoff \(T\) (across the \(x\)-axis) and of the sucker's payoff \(S\) (across
+the \(y\)-axis). Starting at the upper left corner and proceeding clockwise the
+quadrants correspond to the harmony, the snowdrift, the prisoner's dilemma and
+the stag-hunt games.
 
@@
   \caption{{\bf Evolutionary dynamics under expected payoffs for various social dilemmas.}
-  We have run several simulations of the evolutionary process described in
-  section~\ref{section:model} for $T\!=\!10^7$ time steps. The graphs show how
-  often the resident population chooses each combination ($p,q$) of conditional
-  cooperation probabilities in the subsequent rounds. We vary the temptation
-  payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\)
-  across the \(x\) axis, and  \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\)
-  across the \(y\) axis. Unless explicitly varied, the parameters of the simulation
-  are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$.}
+  We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across
+  the \(x\) axis, and  \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\) across the
+  \(y\) axis. ({\bf A}) The top left quadrant where \(S > 0\) and \(T \leq 1\)
+  corresponds to the harmony game. The preference ordering for the harmony game
+  is \(R > T > S > P\). ({\bf B}) The top right quadrant where \(S > 0\) and \(T > 1\)
+  corresponds to the snowdrift game. The preference ordering for the snowdrift game
+  is \(T > R > S > P\). ({\bf C}) The bottom left quadrant where \(S < 0\) and \(T \leq 1\)
+  corresponds to the stag-hunt game. The preference ordering for the stag-hunt game
+  is \(R > T > P > S\). ({\bf D}) The bottom right quadrant where \(S < 0\) and \(T > 1\)
+  corresponds to the prisoner's dilemma game. The preference ordering for the prisoner's dilemma game
+  is \(T > R > P > S\).
+  Unless explicitly varied, the parameters of the simulation
+  are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}
   \label{fig:expected_two_by_two}
 \end{figure}
-
-The conclusions regarding the evolved behaviour for the various classes of
-games remain the same when the last round payoffs are considered,
-Figure~\ref{fig:last_round_two_by_two}. However, there is variation in these
-populations. This variation as an effect affects the resident within cooperative
-rates.
 
@@
   \caption{{\bf Evolutionary dynamics under last round payoffs for various social dilemmas.} 
-  We have run several simulations of the evolutionary process described in
-  section~\ref{section:model} for $T\!=\!10^7$ time steps. The graphs show how
-  often the resident population chooses each combination ($p,q$) of conditional
-  cooperation probabilities in the subsequent rounds. We vary the temptation
-  payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\)
-  across the \(x\) axis, and \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\)
-  across the \(y\) axis. Unless explicitly varied, the parameters of the simulation
-  are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$.}
+  We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across
+  the \(x\) axis, and  \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\) across the
+  \(y\) axis. ({\bf A}) The top left quadrant where \(S > 0\) and \(T \leq 1\)
+  corresponds to the harmony game. The preference ordering for the harmony game
+  is \(R > T > S > P\). ({\bf B}) The top right quadrant where \(S > 0\) and \(T > 1\)
+  corresponds to the snowdrift game. The preference ordering for the snowdrift game
+  is \(T > R > S > P\). ({\bf C}) The bottom left quadrant where \(S < 0\) and \(T \leq 1\)
+  corresponds to the stag-hunt game. The preference ordering for the stag-hunt game
+  is \(R > T > P > S\). ({\bf D}) The bottom right quadrant where \(S < 0\) and \(T > 1\)
+  corresponds to the prisoner's dilemma game. The preference ordering for the prisoner's dilemma game
+  is \(T > R > P > S\).
+  Unless explicitly varied, the parameters of the simulation
+  are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}
   \label{fig:last_round_two_by_two}
@@
 
-The cooperation rates for each of the case of the social dilemmas we have used
-are given in Figure~\ref{fig:cooperation_rates_two_by_two_part_one}. In the case
-of the harmony game the differences are very small, even in the case of \(T=1\)
-where the biggest difference occurs it still remain less than 10\%. In the case
-of the stag hunt game the difference increases in the case of \(T=1\). The
-biggest variation in the results are for the class of the prisoner's dilemma.
-For values of \(S < -0.4\) the are significantly different, in the case of the
-last round payoffs cooperation almost never is a valid option. Overall the last
-round payoffs are less cooperative, supporting the results of the previous
-section. There are only two cases which that is not true and both cases are in
-the snowdrift class. For the values of \(T>1.8\) and \(S=0.4, 0.8\) cooperation
-is estimated higher. In the last round payoffs a Tit For Tat like behaviour is
-more robust that in the case of the expected payoffs.
+The harmony game describes an interaction without conflict. In the harmony game
+it is in the best interest for both players to cooperate, and in an evolutionary
+process cooperators prevail. We observe that for most harmony game runs the
+resident population overwhelmingly applies a strategy for which \(p\) and \(q\)
+are \(\approx 1\) (Figure~\ref{fig:expected_two_by_two}). The snowdrift game
+describes a situation similar to that of the prisoner's dilemma; cooperation
+results in a benefit to the opposing player but entailing a cost to the
+cooperator. However, in the snowdrift game individuals obtain immediate direct
+benefits from the cooperative acts which leads to \(S > P\). In the snowdrift
+game more cooperation emerges compared to the prisoner's dilemma. Defection is
+never a resident strategy and for the same values of temptation the overall
+\(q-\) values are higher. The last class of games we present are for the stag-hunt
+game. In the stag-hunt game there is an equilibrium in which both players
+cooperate as well as one in which both defect. For small values of the
+temptation and the sucker's payoffs we observe that cooperation prevails similar
+to the harmony game. However, in the case where the temptation and the reward
+payoffs are equal the resident population either consists of defectors or
+cooperators.
+
+Although the patterns of the evolved populations remain similar if we consider
+that players update their strategies based on their last round payoff
+(Figure~\ref{fig:last_round_two_by_two}), there are still dissimilarities
+between the two approaches. To better understand these dissimilarities, we
+explore the cooperation rates for each of the cases of the social dilemmas we
+have presented (Figure~\ref{fig:cooperation_rates_two_by_two_part_one}).
+
+In the harmony game the cooperation rates for both approaches are similar. The
+biggest difference occurs for \(T=1\), nevertheless, it is less than 10\%. There
+are two instances for which the cooperation rates are higher for the limited
+memory payoffs compared to the expected. These are for \(T=2.2, S=0.4\) and for
+\(T=2.2, S=0.8\). It is only in the snowdrift game that cooperation is
+overestimated when a minimum of information is considered. Players are slightly
+more likely to cooperate after a cooperation and to forgive after being at the
+receiving end of a defection. For the rest of the cases the classical scenario
+overestimates the average cooperation rate, supporting the results of the
+previous section. There is a substantial difference in the cooperation rates
+for the prisoner's dilemma class. We observe that for \(S<-0.4\) cooperation
+almost never emerges if players imitate based only on the last round payoff.
 
@@
   \includegraphics[width=.4\textwidth]{static/stochastic_two_by_two_games_cooperation.pdf}
-  \caption{\textbf{Cooperation rates over various social dilemmas.} ({\bf A}) If
+  \caption{\textbf{Cooperation rates for various social dilemmas.} ({\bf A}) If
   players update based on their expected payoffs. ({\bf B}) If
@@
   across the \(y\) axis. Unless explicitly varied, the parameters of the simulation
-  are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$.}\label{fig:cooperation_rates_two_by_two_part_one}
+  are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}\label{fig:cooperation_rates_two_by_two_part_one}
 \end{figure}
@@
 round payoffs. In order to explore further the effect of limited memory we allow
-individuals to remember more. We consider the cases where individuals remember
-up two interactions, and up to the last two rounds. In total we present results
-for three more cases. These are the cases of the last round two rounds with
-another member of the population, last round with two members of the population,
-and last round two rounds with two members of the population. Similarly to the
-previous examples we have run the evolutionary process for a large number of
-step for each of the social dilemmas. The results are found in the
-Appendix~\ref{appendix:further_simulation_results}.
+individuals to remember more. More precisely, up to two interactions and up to
+the last two rounds. In total we present results for three more updating
+payoffs. These are the payoffs when individuals consider the last two rounds
+with another member of the population, the last round with two members of the
+population, and the last two rounds with two members of the population.
+Similar to the last round payoff, we use simulations and record which strategies
+the players adopt over time based
+(Appendix~\ref{appendix:further_simulation_results}) and we compare the
+evolving cooperation rates to those of the expected payoffs
+(Figure~\ref{fig:cooperation_rates_two_by_two_part_two}).
 
-The behavior over the different games remains the same. We note that now there
-is more noise in the evolved populations. Due to space the figures have been
-moved to the Appendix, however, the cooperating rates for each of the cases are
-given in Figure~\ref{fig:cooperation_rates_two_by_two_part_two}.
-
-In the cases of the last two rounds the evolving cooperating rate is closer
-to that of the expected payoffs. In the case of the prisoner's dilemma there
-is an increase in the average cooperating rate which is still strictly less
-than in the case of the prisoner's dilemma. Regarding the snowdrift cases
-now each of the run for which \(T = 2.2\) the cooperating rate is not higher
-that the expected one.
-
-In the case of the two opponents experiments all the runs are lower with the
-expected payoffs, and in the case of the prisoner's dilemma we observe an
-increase in the average cooperate rate again. Similar in the case where both
-the last two turns and the two opponents are considered. The results appear
-to be the same as in the case of the two opponents. However, for some classes
-the results are affected by the last round.
+In the case of the last two rounds payoff we make two observations. Initially,
+for the harmony, the stag-hunt and the prisoner's dilemma classes the average
+cooperation rates are higher than the previous case of the last round payoff.
+The rates remain strictly less than in the expected payoffs, however, in the
+case of the prisoner's dilemma we observe a significant increase. Secondly, for
+the snowdrift class there are more instances for which the cooperation rates are
+higher than in the expected payoffs. This could indicate that in the case of the
+snowdrift game the expected payoffs underestimate cooperation.
+Once players interact with two other members of the population, instead of one,
+the expected payoffs strictly yield a higher cooperation rate  even in the case
+of the snowdrift game. Similar to the previous result, the difference between
+the two approaches is always higher for the prisoner's dilemma class.
 
@@
   across the \(y\) axis. Unless explicitly varied, the parameters of the simulation
-  are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$.}
+  are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$..}
   \label{fig:cooperation_rates_two_by_two_part_two}
@@
 
-Evolutionary models have helped us understand the evolution of cooperation.
-A behaviour so well observed in the world around us though Darwin spoke
-about the survival of the fitness. Evolutionary
+Cooperation can be seen at odd, why is it that we choose to help others,
+increasing their payoff, at the expense of decreasing one's own? In spite of all
+the selfish genes’ animal and human communities seem to altruistically help each
+other and cooperate, and evolutionary game theory has helped us shape our
+understanding of the evolution of cooperation.
 
-Much research in the past has been devoted to explore conditionally cooperative
-strategies in pairwise interactions.
+Previous evolutionary models often feature a curious inconsistency. When
+modeling how individuals make decisions in each round, these models assume that
+players only remember the last round. However, when modeling how individuals
+update their strategies over time, individuals are assumed to have perfect
+memory.
 
-Previous models of direct reciprocity often feature a curious inconsistency.
-When modeling how individuals make decisions in each round, these models assume
-that players only remember the last round. However, when modeling how
-individuals update their strategies over time, individuals are assumed to have
-perfect memory. Here, we explore how robust our understanding of cooperation is
-as models deviate from the perfect memory assumption. On the positive side, we
-show that cooperation can even evolve when individuals only use a minimum of
-information. On the negative side, the evolving cooperation rates are typically
-lower than in the classical scenario.
+Here, we have explored how robust cooperation is as models deviate from the
+perfect memory assumption. Initially we considered the donation game. We showed
+that when the last round payoff is used instead of the expected payoffs,
+cooperation can even evolve when individuals only use a minimum of information,
+however, the evolving cooperation rates are typically lower. The resident
+strategies were both less cooperative and less generous. This effect was only
+intensified we increase the benefit and the strength of selection independently.
+The results showed that as each parameter was being increased the difference
+between the cooperation rates were widening. This indicates that cooperative
+players benefit from being able to interact with everyone in the population.
 
-In the case of the donation game we show that cooperation is overestimated.
-Morover, we obsved that when the befenti and the srang f sleection increases this
-difference becomes more obvuous,
+We extended our approach and presented results not only based on different
+classes of social games, but also based on different limited memory payoffs. The
+analysis showed that specifically for the case of the prisoner's dilemma games
+the contrast between the two approaches can be significantly large. The
+prisoner's dilemma is one of the most applied types of social games when we
+discuss the evolution of cooperation. Our results indicate that cooperation
+struggles to evolve in the prisoner's dilemma when only a minimum of social
+information is used at the updating stage. Interestingly our simulations also
+showed that in same cases cooperation can benefit from limited memory. This was
+specific for the snowdrift game.
+
+\section{Acknowledgements}
+
+This work was supported by the European Research Council Starting Grant 850529:
+E-DIRECT.
 
@@
 and let $\mathbf{u}\!=\!(R,S,T,P)$ be the corresponding payoff vector.
-The values of the payoffs are not only based on the prisoner's dilemma but all
-the symmetric \(2 \times 2\) games, Table~\ref{table:social_dilemmas}.
+We present results for various values of $\mathcal{U}$ for all
+the symmetric \(2 \times 2\) games.
 
@@
 where $u_{L}\!\in\! \mathcal{U}$ is the learner's payoff, $u_{RM}\!\in\!
-\mathcal{U}$ is the role model's payoff, and $\beta\!\ge\!0$ is the intensity of
+\mathcal{U}$ is the role model's payoff, and $\beta\!\ge\!0$ is the strength of
 selection.
@@
    resident $\leftarrow (0, 0, 0)$\;
-   \While{step $<$ maximum number of steps}{
+   \While{ $t <$ maximum number of steps}{
     mutant $\leftarrow$ random: $\{\emptyset \}\ \rightarrow R^3$\;
@@
 
-Initially, we discuss the case of the last round updating payoff. At the stage
+Initially, we discuss the case of the \textbf{last round updating payoff}. At the stage
 game we define the payoff of a reactive strategy in the last round,
@@
 \begin{Prop}\label{proposition:last_round}
-    Consider a repeated prisoner's dilemma, with
+    Consider a repeated game, with
     continuation probability $\delta$, between players with reactive strategies
@@
 with another resident or with another mutant. Thus, in each updating stage there
-are five possible combinations of pairs. These are illustrated by
-Figure~\ref{fig:single_pairs}.
+are five possible combinations of pairs (Figure~\ref{fig:single_pairs}).
 
@@
   \caption{\textbf{Possible pairings combination in the updating stage, given
-  that individuals interact with only one other member in the population.}}
+  that individuals interact with only one other member in the population.} At
+  each step of the evolutionary process we choose a role model and a learner to
+  update the population. We consider the case where both the role model and the
+  learner estimate their fitness after interacting with a single member of the
+  population. There are five possible pairings at each step. They interact with
+  other with a probability \(\frac{1}{N - 1}\), and thus they do not interact
+  with other with a probability \(1 - \frac{1}{N - 1}\). In
+  the latter case, each of them can interact with either a mutant or a resident.
+  Both of them interact with a mutant with a probability
+  $\frac{(k-1)(k-2)}{(N-2)(N-3)}$ and both interact with a resident with a
+  probability $\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$. The last two possible pairings
+  are that either of them interacts with a resident whilst the other interacts
+  with a mutant, and this happens with a probability $\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$.}
   \label{fig:single_pairs}
@@
 
-The probability that the respective payoffs of the players
-are given by $u_1$ and $u_2$ can be calculated as
+Given the last round payoff and possible pair combinations for a single interaction,
+we define the probability that the respective last round payoffs of two players
+\(s_1, s_2\) are given by $u_1$ and $u_2$ as,
 
@@
 \begin{array}{llrl}
-x(u_1,u_2)	 &=&\displaystyle \frac{1}{N\!-\!1}\cdot  &v_{u_1}(S_1,S_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F}\\[0.5cm]
+x(u_1,u_2)	 &=&\displaystyle \frac{1}{N\!-\!1}\cdot  &v_{u_1}(s_1,s_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F}\\[0.5cm]
 &+	
 &\displaystyle \left(1\!-\!\frac{1}{N\!-\!1}\right)  
-&\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} v_{u_1}(S_1,S_2) v_{u_2}(S_2,S_2) + 
- \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} v_{u_1}(S_1,S_2) v_{u_2}(S_2,S_1)\right.\\[0.5cm]
-&&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} v_{u_1}(S_1,S_1) v_{u_2}(S_2,S_2) + 
- \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} v_{u_1}(S_1,S_1) v_{u_2}(S_2,S_1)\right].
+&\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_2) + 
+ \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_1)\right.\\[0.5cm]
+&&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_2) + 
+ \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_1)\right].
 \end{array}
@@
 the role model happened to be matched during the game stage, which happens with
-probability $1/(N\!-\!1)$.In that case, we note that only those payoff pairs
+probability $\frac{1}{(N\!-\!1)}$. In that case, we note that only those payoff pairs
 can occur that are feasible in a direct interaction, $\splitatcommas{(u_1,u_2)\in
@@
 and that the remaining $k$ players use the mutant strategy
-$S_{M}\!=\!(y_{M},p_{M},q_{M})$, the probability that the number of mutants increases by
+$s_{M}\!=\!(y_{M},p_{M},q_{M})$, the probability that the number of mutants increases by
 one in one step of the evolutionary process can be written as
@@
 
-In this expression, $(N\!-\!k)/N$ is the probability that the randomly chosen
-learner is a resident, and $k/N$ is the probability that the role model is a
+In this expression, $\frac{(N\!-\!k)}{N}$ is the probability that the randomly chosen
+learner is a resident, and $\frac{k}{N}$ is the probability that the role model is a
 mutant. The sum corresponds to the total probability that the learner adopts the
-role model's strategy over all possible payoffs $u_1$ and $u_2$ that the two
-player may have received in their respective last rounds. We use $x(u_1,u_2)$ to
+role model's strategy over all possible payoffs $u_R$ and $u_M$ that the two
+player may have received in their respective last rounds. We use $x(u_R,u_M)$ to
 denote the probability that the randomly chosen resident obtained a payoff of
-$u_1$ in the last round of his respective game, and that the mutant obtained a
-payoff of $u_2$.
+$u_R$ in the last round of his respective game, and that the mutant obtained a
+payoff of $u_M$.
 
-This framework can be extended to consider the case of where the payoffs
-correspond to the last \(n\) rounds payoff an individual achieved after
-interacting with \(m\) other individuals. For the case \(n=2\) the payoffs
-at the game stage are,
+We extend our framework to consider the case where players update their
+strategies based on the outcome of \textbf{the last two turns and based on their
+interaction with two other members of the population}. At the stage game we
+define the payoff of a reactive strategy in the last two rounds,
+Proposition~\ref{proposition:last_two_rounds}.
 
 \begin{Prop}\label{proposition:last_two_rounds}
-Assume a play between the reactive strategies \(s_1\) and \(s_2\) with a
-continuation probability \(\delta\). Then the probability of being in any of the
-sixteen outcomes \(\splitatcommas{RR, RS, RT, RP, SR, SS, ST, SP, TR, TS, TT, TP, PR, PS, PT, PP}\)
-on the last two rounds are given by,
+  Consider a repeated game, with continuation probability $\delta$, between
+  players with reactive strategies $s_1\!=\!(y_1, p_1, q_1$)  and
+  $s_2\!=\!(y_2,p_2,q_2)$ respectively. Let $\mathcal{\tilde{U}} =
+  \splitatcommas{\{RR, RS, RT, RP, SR, SS, ST, SP, TR, TS, TT, TP, PR, PS, PT,
+  PP\}}$ denote the set of feasible payoffs in the last two rounds, and let
+  \(\tilde{\mathbf{u}}\) be the corresponding payoff vector. Then the
+  probability that the $s_1$ player receives the payoff $u\!\in\!
+  \mathcal{\tilde{U}}$ in the very last two rounds of the game is given by,
 
   \begin{equation}
-  \mathbf{v_{a_1, a_2}} = (1 - \delta) m_{a_1, a_2} \delta^2 \left[\mathbf{v_0}(I_4 - \delta M)^{-1}\right]_{a_1, a_2}, \quad \text{ for } m_{a_1, a_2} \in M \ \& \ a_1, a_2 \in \{R, S, T, P\}
+  \langle\mathbf{\tilde{v}}(s_1,s_2),\mathbf{\tilde{u}}\rangle, \text{ where } \mathbf{\tilde{v}} \in R^{16} \text{ is given by },
   \end{equation}
-
+  \begin{equation}
+    \mathbf{\tilde{v}}(s_1,s_2) = (1 - \delta) m_{a_1, a_2} \delta^2 \left[\mathbf{v_0}(I_4 - \delta M)^{-1}\right]_{a_1, a_2}, \quad  m_{a_1, a_2} \in M \ \forall \ a_1, a_2 \in \{1, 2, 3, 4\}.
+  \end{equation}
 \end{Prop}
 
-Equation~\ref{eq:Chi} can also be extended to include interactions with two
-other individuals. The possible pairings are illustrated by
-Figure~\ref{fig:pissible_two_pairs}.
+In the updating stage we select a mutant and resident to be either the role
+model or the learner. Given that they can interact with two other members of
+the population there are a total of twenty four possible combinations of pairs
+(Figure~\ref{fig:pissible_two_pairs}).
 
@@
 
-Simulating the evolutionary process for longer memory quickly becomes
-computationally untractable. Our methodology could be extended to include more
-turns and more interactions. However, for the purpose of this work we explore
-the cases only up to two turns and two interactions.
+Given the last two rounds payoff and possible pair combinations with two
+members, we can define the probability \(x\) that the respective last round
+payoffs of two players \(s_1, s_2\) are given by $u_1$ and $u_2$ similarly to
+Eq.~(\ref{eq:Chi}). We follow the same approach to define the rest of the
+updating payoffs. These are the updating payoffs of the last round with two
+member, and the the last two rounds payoffs with one member.
+
+Simulating the evolutionary process for more interactions and rounds quickly
+becomes computationally intractable. Our methodology could be extended to
+include \(n\) turns and \(m\) interactions. However, for the purpose of this
+work we explore the cases only up to two turns and two interactions.
 
@@
   We have run several simulations of the evolutionary process described in
-  section~\ref{section:model} for $T\!=\!10^7$ time steps. The graphs show how
+  section~\ref{section:model} for $t\!=\!10^7$ time steps. The graphs show how
   often the resident population chooses each combination ($p,q$) of conditional
@@
 
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[width=\textwidth]{static/difference_in_cooperation.pdf}
-  \caption{{\bf The difference in cooperate rates for each method compared to the
-  expected payoffs.}}
-  % \label{fig:last_two_rounds_two_opponents}
-\end{figure}
-
 \bibliographystyle{unsrt}

### 2021-05-25 clean_data.py

@@ -15,13 +15,12 @@ if __name__ == "__main__":  # pragma: no cover
         "p",
         "q",
         "average cooperation rate",
-        # "average score",
+        "average score",
         "t",
     ]
 
-    print(f"{folder}/{starting_characters}*csv")
-
-    csv_files = glob.glob(f"{folder}/{starting_characters}*csv")
+    print(f"{folder}/{starting_characters}*_beta_10.csv")
+    csv_files = glob.glob(f"{folder}/{starting_characters}*_beta_10.csv")
 
     for file in tqdm.tqdm(csv_files):
         df = pd.read_csv(file, header=None)
@@ -37,4 +36,4 @@ if __name__ == "__main__":  # pragma: no cover
 
         df["frequencies"] = frequencies
 
-        df.to_csv(f"{folder}/clean_{file[len(folder):]}", index=False)
+        df.to_csv(f"{folder}clean_{file[len(folder):]}", index=False)

### 2021-06-04 matlab/evolRun.m
prep to run sixteen states

@@ -4,16 +4,14 @@ starting_resident = [0, 0, 0];
 N = 100;
 delta = 0.999;
 beta = 1;
-numberIterations = 5 * 10 ^ 7;
-payoff_type = "last_round";
+numberIterations = 10 ^ 7;
+payoff_types = ["two_rounds", "two_rounds_opponents"]
 
-bs = linspace(2, 10, 9);
+u = [2, -1, 3, 0];
 
-
-parfor (i = 1:9)
-    b = bs(i);
-    u = [b - 1, -1, b, 0];
-    filename = "../data/last_round_b_" + b;
-    evolSimulation(starting_resident, u, N, delta, beta, numberIterations, payoff_type, filename);
+parfor (i = 1:2)
+    payoff_type = payoff_types(i);
+    filename = "../data/sixteen_states_" + payoff_type;
+    evolSimulationTwoRounds(starting_resident, u, N, delta, beta, numberIterations, payoff_type, filename);
 end
 end

### 2021-06-04 matlab/evolRun.m
Restructure matlab code (#4)
* moving calculation of fixation probability to files based on method

* adds docstring to stationary

* re write script to use fixation functions

* close end

* add semicolom

* fix nested issue

* debug code

* run tests only on pull requests

* includes two opponents type

* prepare to run simulations (for 3 types)

* add return statement

* corrects return variable

* rename matlab scripts for phi

* rename rho to phi

* write with precision

* fix opponents probabilities

* style

* add seed

* run benefits example for longer

* restructure code for 16 states

@@ -4,28 +4,14 @@ starting_resident = [0, 0, 0];
 N = 100;
 delta = 0.999;
 beta = 1;
-numberIterations= 10 ^ 7;
-stochastic = 1;
+numberIterations = 10 ^ 7;
+payoff_types = ["two_rounds", "two_rounds_opponents"]
 
-n = 11;
-Ss = linspace(-2, 2, n);
-Ts = linspace(-1, 3, n);
+u = [2, -1, 3, 0];
 
-payoffs = zeros(n * n, 2);
-
-for i=1:n
-    for j=1:n
-        payoffs(j + (i - 1) + (i -1) * 10, 1) = Ss(i);
-        payoffs(j + (i - 1) + (i -1) * 10, 2) = Ts(j);
-    end
-end
-
-N = n * n;
-parfor (i = 1:N, 8)
-    S = payoffs(i, 1);
-    T = payoffs(i, 2);
-    u = [1, S, T, 0];
-    filename = "data/stochastic/S_" + S + "_T_" + T + "_stochastic_" + stochastic;
-    evolSimulation(starting_resident, u, N, delta, beta, numberIterations, stochastic, filename);
+parfor (i = 1:2)
+    payoff_type = payoff_types(i);
+    filename = "../data/sixteen_states_" + payoff_type;
+    evolSimulationTwoRounds(starting_resident, u, N, delta, beta, numberIterations, payoff_type, filename);
 end
 end

### 2021-06-04 matlab/evolRunMemoryOne.m

@@ -1,0 +1,17 @@
+function evolRunMemoryOne();
+
+starting_resident = [0, 0, 0, 0, 0];
+N = 100;
+delta = 0.999;
+beta = 1;
+numberIterations = 10 ^ 7;
+payoff_types = ["expected", "last_round"]
+
+u = [2, -1, 3, 0];
+
+parfor (i = 1:2)
+    payoff_type = payoff_types(i);
+    filename = "../data/memory_one_" + payoff_type;
+    evolSimulationMemoryOne(starting_resident, u, N, delta, beta, numberIterations, payoff_type, filename);
+end
+end

### 2021-06-04 matlab/evolSimulationOpponentsTwo.m

@@ -1,73 +1,0 @@
-function [xDat,AvCoop,AvPay,Rho,Data]=evolSimulationOpponentsTwo(starting_resident, u, N, delta, beta, numberIterations, filename);
-
-%% Preparations for the output
-Data=['R=',num2str(u(1)),'; S=',num2str(u(2)),'; T=',num2str(u(3)), '; P=',num2str(u(4)),'; N=',num2str(N),'; beta=',num2str(beta), '; nIt=',num2str(numberIterations)];
-AvCoop=0; AvPay=0; Res=starting_resident;
-
-sdim=3;
-xDat=zeros(numberIterations/100, 6);
-xDat(1, :)=[Res, 0, u(4), 0];
-
-%% Calculating all possible pairwise imitation probabilities based on one payoff
-Rho=zeros(4,4);
-for i1=1:4
-    for i2=1:4
-        Rho(i1,i2)=1 / (1 + exp(-beta * (u(i2) - u(i1))));
-    end
-end
-
-
-%% Running the evolutionary process
-j = 2;
-for t = progress(1:numberIterations)
-    Mut=rand(1,sdim);
-    [rho,coopM,piM]=CalcRho(Mut, Res, Rho, N, u, delta, beta);
-    if rand(1)<rho
-        Res=Mut; xDat(j,:)=[Res,coopM,piM,t]; j=j+1;
-    end
-end
-
-csvwrite(filename + ".csv", xDat);
-csvwrite(filename + ".txt", Data);
-
-AvCoop = mean(xDat(:, end-2));
-AvPay = mean(xDat(:, end-1));
-%time=toc
-end
-
-function [rho,coopMM,piMM]=CalcRho(Mut, Res, Rho, N, u, delta, beta);
-
-%% Preparations: Calculating outcome probabilities and expected payoffs
-
-vMM=stationary(Mut,Mut,delta);
-vMR=stationary(Mut,Res,delta);
-vRM=[vMR(1) vMR(3) vMR(2) vMR(4)];
-vRR=stationary(Res,Res,delta);
-
-coopMM=vMM(1)+vMM(2);
[…]

### 2021-06-04 matlab/evolSimulationRoundTwo.m

@@ -1,59 +1,0 @@
-function [xDat,AvCoop,Rho,Data]=evolSimulationRoundTwo(starting_resident, u, N, delta, beta, numberIterations, filename);
-
-%% Preparations for the output
-Data=['R=',num2str(u(1)),'; S=',num2str(u(2)),'; T=',num2str(u(3)), '; P=',num2str(u(4)),'; N=',num2str(N),'; beta=',num2str(beta), '; nIt=',num2str(numberIterations)];
-AvCoop=0; Res=starting_resident;
-
-sdim = 3;
-xDat = zeros(numberIterations/100, 5);
-xDat(1, :)=[Res, 0, 0];
-
-Rho = zeros(16, 16);
-
-for i=1:16
-    for j=1:16
-        Rho(i, j) = 1 / (1 + exp(-beta * (u(1 + fix((i - 1) / 4)) + u(1 + mod(i - 1, 4)) - u(1 + fix((j - 1) / 4)) + u(1 + mod(j - 1, 4)))));
-    end
-end
-
-
-%% Running the evolutionary process
-j = 2;
-for t = progress(1:numberIterations)
-    Mut = rand(1, sdim);
-    [rho,coopM] = CalcRhoRoundTwo(Mut, Res, N, u, delta, beta, Rho);
-    if rand(1)<rho
-        Res=Mut; xDat(j,:)=[Res,coopM,t]; j=j+1;
-    end
-end
-
-csvwrite(filename + ".csv", xDat);
-csvwrite(filename + ".txt", Data);
-end
-
-function [rho,coopMM]=CalcRhoRoundTwo(Mut, Res, N, u, delta, beta, Rho);
-
-%% Preparations: Calculating outcome probabilities and expected payoffs
-
-vMM = stationaryRoundTwo(Mut, Mut, delta);
-vMR = stationaryRoundTwo(Mut, Res, delta);
-vRM = stationaryRoundTwo(Res, Mut, delta);
-vRR = stationaryRoundTwo(Res, Res, delta);
-
-coopMM = vMM(1) + vMM(2) + vMM(5) + vMM(6);
-
-laplus = zeros(1, N-1); laminus=laplus;
-for k=1:N-1
-    x=zeros(16,16);
-    for i1=1:16
[…]

### 2021-06-04 matlab/phiExpected.m

@@ -1,0 +1,14 @@
+function [phi]=phiExpected(N, piMM, piMR, piRR, piRM, beta);
+%% Calculates the fixation probability based on the expected payoffs
+
+laplus=zeros(1, N-1); laminus=laplus;
+for k=1:N-1
+    piM = (k-1) / (N-1) * piMM + (N-k) / (N-1) * piMR;
+    piR = k / (N-1) * piRM + (N-k-1) / (N-1) * piRR;
+
+    laplus(k) = 1 / (1 + exp(-beta * (piM - piR)));
+    laminus(k) = 1 / (1 + exp(-beta * (piR - piM)));
+end
+
+phi = 1 / (1 + sum(cumprod(laminus./laplus)));
+end

### 2021-06-04 matlab/phiTwoOpponents.m

@@ -1,0 +1,32 @@
+function [phi]=phiTwoOpponents(N, vRM, vMM, vMR, vRR, Rho);
+%% Calculates the fixation probability based on the last round payoff with two opponents
+laplus=zeros(1,N-1); laminus=laplus;
+for k=1:N-1
+    x=zeros(4,4);
+    for i1=1:4
+        for i2=1:4
+
+learner_role_model = 1 / (N - 1) * vRM(i1) * ((i1==1 & i2==1) | (i1==2 & i2==3) | (i1==3 &i2==2) | (i1==4 & i2==4)) * ( ...
+1 / (N - 2) / (N - 3) * ((k - 1) * (k - 2) * vRM(i1) * vMM(i2) + (k - 1) * (N - k - 1) * vRM(i1) * vMR(i2) + (N - k - 1) * (k - 1) * vRR(i1) * vMM(i2) + (N - k - 1) * (N - k - 2) * vRR(i1) * vMR(i2)));
+
+mutant_mutant = (1 - 1 / (N - 1)) / (N - 2) / (N - 3) * ((k-1) * (k-2) * vRM(i1) * vMM(i2)) * ((1 - 1 / (N - 2)) / (N - 3) / (N - 4) * ((k-2)*(k-3)*vRM(i1)*vMM(i2) + (k-2)*(N-k-1)*vRR(i1)*vMM(i2) + (N-k-1)*(k-2)*vRM(i1)*vMR(i2) + (N-k-1)*(N-k-2)*vRR(i1)*vMR(i2)) + ...
+1 / (N - 2) * vRM(i1) * ((i1==1 & i2==1) | (i1==2 & i2==3) | (i1==3 &i2==2) | (i1==4 & i2==4)));
+
+mutant_resident = (1 - 1 / (N - 1)) / (N - 2) / (N - 3) * ((k-1) * (N-k-1) * vRM(i1) * vMR(i2)) * ((1 - 1 / ( N - 2)) / (N - 3) / (N - 4) * ((k-1)*(k-3)*vRM(i1)*vMM(i2) + (k-1)*(N-k-1)*vRR(i1)*vMM(i2) + (N-k-2)*(k-2)*vRM(i1)*vMR(i2) + (N-k-2)*(N-k-2)*vRR(i1)*vMR(i2)) + ...
[…]

### 2021-06-04 matlab/phiTwoRoundsOpponents.m

@@ -1,0 +1,43 @@
+function [phi]=phiTwoRoundsOpponents(N, vRM, vMM, vMR, vRR, Rho);
+%% Calculates the fixation probability based on the last two rounds payoff with two opponents
+
+laplus = zeros(1, N-1); laminus=laplus;
+for k=1:N-1
+    x=zeros(16,16);
+    for i1=1:16
+        for i2=1:16
+learner_role_model = 1 / (N - 1) * vRM(i1) * ((i1==1 & i2==1)  |(i1==2 & i2==3)  |(i1==3 &i2==2)   |(i1==4 & i2==4) |(i1==5 & i2==9) |(i1==6 & i2==11)|...
+                                  (i1==7 & i2==10) |(i1==8 & i2==12) |(i1==9 & i2==5)  |(i1==10 & i2==7)|(i1==11 & i2==6)|(i1==12 & i2==8)|...
+                                  (i1==13 & i2==13)|(i1==14 & i2==15)|(i1==15 & i2==14)|(i1==16 & i2==16)) * ( ...
+1 / (N - 2) / (N - 3) * ((k - 1) * (k - 2) * vRM(i1) * vMM(i2) + (k - 1) * (N - k - 1) * vRM(i1) * vMR(i2) + (N - k - 1) * (k - 1) * vRR(i1) * vMM(i2) + (N - k - 1) * (N - k - 2) * vRR(i1) * vMR(i2)));
+
+mutant_mutant = (1 - 1 / (N - 1)) / (N - 2) / (N - 3) * ((k-1) * (k-2) * vRM(i1) * vMM(i2)) * ((1 - 1 / (N - 2)) / (N - 3) / (N - 4) * ((k-2)*(k-3)*vRM(i1)*vMM(i2) + (k-2)*(N-k-1)*vRR(i1)*vMM(i2) + (N-k-1)*(k-2)*vRM(i1)*vMR(i2) + (N-k-1)*(N-k-2)*vRR(i1)*vMR(i2)) + ...
+1 / (N - 2) * vRM(i1) * ((i1==1 & i2==1)  |(i1==2 & i2==3)  |(i1==3 &i2==2)   |(i1==4 & i2==4) |(i1==5 & i2==9) |(i1==6 & i2==11)|...
[…]

### 2021-06-04 main.tex
Draft paper (#5)
* compile latex

* re write simulations.py to check that the simulated utility is equal to the theoretical

* initial analysis of data

* fixes bug in matlab code

The current code wrote the cooperation rate for the mutant at
each time step instead of the rate of the resident

* run analysis

* check new data output

* use csvwrite instead of writematrix

* tweaks analysis code to fit the new data

* add the functionality to consider more than just the last turn

* scoring turns as a class argument

* test function further

* run black

* some style fixes

* matlab scripts for

@@
 cooperation in a variety of environments. Although existing models have shaped
-our understanding of human cooperation, the often work with idealized
+our understanding of human cooperation, they often work with idealized
 assumptions. In an evolutionary process, individuals imitate other individuals
@@
 Here, we introduce a framework that allows individuals to estimate their fitness
-based on a minimum of social information. We explore the difference between the
-classical framework and ours using computer simulations. We present results for
+based on a minimum of social information. We explore the difference between our
+framework and the classical framework using computer simulations. We present results for
 the most commonly used classes of symmetric \(2 \times 2\) games. The

### 2021-06-04 static/expected_stochastic.tex
Draft paper (#5)
* compile latex

* re write simulations.py to check that the simulated utility is equal to the theoretical

* initial analysis of data

* fixes bug in matlab code

The current code wrote the cooperation rate for the mutant at
each time step instead of the rate of the resident

* run analysis

* check new data output

* use csvwrite instead of writematrix

* tweaks analysis code to fit the new data

* add the functionality to consider more than just the last turn

* scoring turns as a class argument

* test function further

* run black

* some style fixes

* matlab scripts for

--- 2021-03-16 static/stochastic.tex
@@
     \node[fill, circle, label=above:expected, inner sep=2pt] at (5.7, 5.7) {};
-    \node[fill, circle, inner sep=2pt] (stochastic1) at (1, 1) {};
-    \node[fill, circle, inner sep=2pt] (stochastic2) at (2, 2) {};
-    \node[fill, circle, inner sep=2pt] (stochastic3) at (1, 2) {};
-    \node[fill, circle, inner sep=2pt] (stochastic4) at (2, 1) {};
+    \node[fill, circle, label=above:stochastic, inner sep=2pt] (stochastic) at (1, 1) {};
+    \node[label=left:1] (y) at (0, 1) {};
+    \node[label=below:1] (x) at (1, 0) {};
 
-    \node[label=left:1] (y1) at (0, 1) {};
-    \node[label=below:1] (x1) at (1, 0) {};
-
-    \draw[dotted] (y1)--(1, 1) {};
-    \draw[dotted] (x1)--(1, 1) {};
-
-    \node[label=left:2] (y2) at (0, 2) {};
-    \node[label=below:2] (x2) at (2, 0) {};
-
-    \draw[dotted] (y2)--(2, 2) {};
-    \draw[dotted] (x2)--(2, 2) {};
-
-    \draw[dotted] (y1)--(2, 1) {};
-    \draw[dotted] (x1)--(1, 2) {};
-
+    \draw[dotted] (y)--(1, 1) {};
+    \draw[dotted] (x)--(1, 1) {};
 \end{tikzpicture}

### 2021-06-04 static/last_rounds_matches.tex
Draft paper (#5)
* compile latex

* re write simulations.py to check that the simulated utility is equal to the theoretical

* initial analysis of data

* fixes bug in matlab code

The current code wrote the cooperation rate for the mutant at
each time step instead of the rate of the resident

* run analysis

* check new data output

* use csvwrite instead of writematrix

* tweaks analysis code to fit the new data

* add the functionality to consider more than just the last turn

* scoring turns as a class argument

* test function further

* run black

* some style fixes

* matlab scripts for

--- 2021-05-05 static/last_rounds_matches.tex
@@
 \node [arn_n] {} 
-    child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{M}$] {} edge from parent node[above left, rotate=30] {$\frac{1}{N - 1}$}}
+    child{ node [arn_r, label={[align=center, yshift=-.8cm]role model-learner}] {} edge from parent node[above left, rotate=30] {$\frac{1}{N - 1}$}}
     child{ node [arn_n, rotate=-30] {} 
-            child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{R}$\\$S_{M} - S_{R}$}] {} edge from parent node[above, yshift=-.1cm, rotate=39] {$\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$}}
-            child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{R}$\\$S_{R} - S_{M}$}] {} edge from parent node[above, yshift=-.2cm, rotate=67] {$\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$}}
-            child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{M}$\\$S_{M} - S_{R}$}] {} edge from parent node[above, yshift=-.2cm, rotate=-67] {$\frac{(k-1)(N-k-1)}{(N-2)(N-3)}$}}
-            child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{R}$\\$S_{R} - S_{R}$}] {} edge from parent node[above, yshift=-.1cm, rotate=-41] {$\frac{(k-1)(k-2)}{(N-2)(N-3)}$}}
+            child{ node [arn_r, label={[align=center, yshift=-.8cm]role model-resident\\learner - resident}] {} edge from parent node[above, yshift=-.1cm, rotate=39] {$\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$}}
+            child{ node [arn_r, label={[align=center, yshift=-.8cm]role model-resident\\learner - mutant}] {} edge from parent node[above, yshift=-.2cm, rotate=67] {$\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$}}
+            child{ node [arn_r, label={[align=center, yshift=-.8cm]role model-mutant\\  learner - resident}] {} edge from parent node[above, yshift=-.2cm, rotate=-67] {$\frac{(k-1)(N-k-1)}{(N-2)(N-3)}$}}
+            child{ node [arn_r, label={[align=center, yshift=-.8cm]role model-mutant\\  learner - mutant}] {} edge from parent node[above, yshift=-.1cm, rotate=-41] {$\frac{(k-1)(k-2)}{(N-2)(N-3)}$}}
             edge from parent node[above left, xshift=1em, yshift=-.8em, rotate=-30] {$1 - \frac{1}{N - 1}$}

### 2021-06-04 static/tree.tex
Draft paper (#5)
* compile latex

* re write simulations.py to check that the simulated utility is equal to the theoretical

* initial analysis of data

* fixes bug in matlab code

The current code wrote the cooperation rate for the mutant at
each time step instead of the rate of the resident

* run analysis

* check new data output

* use csvwrite instead of writematrix

* tweaks analysis code to fit the new data

* add the functionality to consider more than just the last turn

* scoring turns as a class argument

* test function further

* run black

* some style fixes

* matlab scripts for

@@
     level/.style={sibling distance = 7cm/#1,
-    level distance = 1.7cm},
+    level distance = 2.5cm, sibling distance = 8cm},
     level 2/.style = {sibling distance = 1.9cm, level distance=2.8cm},
@@
 \node [arn_n] {} 
-    child{ node [arn_r, label={[align=center, yshift=-.8cm] $S_{R} - S_{R}$}] {} 
-    child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{R}$\\$S_{M} - S_{R}$}] {} edge from parent[densely dashdotted] node[above, yshift=-.1cm, rotate=39] {$\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$}}
-    child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{R}$\\$S_{R} - S_{M}$}] {} edge from parent[densely dashdotted] node[above, yshift=-.2cm, rotate=67] {$\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$}}
-    child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{M}$\\$S_{M} - S_{R}$}] {} edge from parent[densely dashdotted] node[above, yshift=-.2cm, rotate=-67] {$\frac{(k-1)(N-k-1)}{(N-2)(N-3)}$}}
-    child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{R}$\\$S_{R} - S_{R}$}] {} edge from parent[densely dashdotted] node[above, yshift=-.1cm, rotate=-41] {$\frac{(k-1)(k-2)}{(N-2)(N-3)}$}}
-    edge from parent node[above left, xshift=1em, yshift=-.8em, rotate=-30] {$1 - \frac{1}{N - 1}$}}
+    child{ node [arn_r, label={[align=center, yshift=-.8cm] role model-learner}] {} 
+    child{ node [arn_r, label={[align=center, yshift=-.8cm] role model-resident\\learner-resident}] {} edge from parent[densely dashdotted] node[above, yshift=-.1cm, rotate=39] {$\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$}}
+    child{ node [arn_r, label={[align=center, yshift=-.8cm] role model-resident\\learner-mutant}] {} edge from parent[densely dashdotted] node[above, yshift=-.2cm, rotate=67] {$\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$}}
+    child{ node [arn_r, label={[align=center, yshift=-.8cm] role model-mutant  \\learner-resident}] {} edge from parent[densely dashdotted] node[above, yshift=-.2cm, rotate=-67] {$\frac{(k-1)(N-k-1)}{(N-2)(N-3)}$}}
+    child{ node [arn_r, label={[align=center, yshift=-.8cm] role model-mutant  \\learner-mutant}] {} edge from parent[densely dashdotted] node[above, yshift=-.1cm, rotate=-41] {$\frac{(k-1)(k-2)}{(N-2)(N-3)}$}}
+    edge from parent node[above left, rotate=25] {$\frac{1}{N - 1}$}}
     child{ node [arn_n, rotate=-30] {} 
-            child{ node [arn_r, label={[align=center, yshift=-.8cm] $S_{R} - S_{R}$\\$S_{M} - S_{R}$}] {} 
-            child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{M}$\\$S_{R} - S_{R}$}] {} edge from parent node[above left, rotate=30] {$\frac{1}{N - 1}$}}
+            child{ node [arn_r, label={[align=center, yshift=-.8cm] role model-resident\\learner-resident}] {} 
+            child{ node [arn_r, label={[align=center, yshift=-.8cm] role model-learner}] {} edge from parent node[above left, rotate=50] {$\frac{1}{N - 1}$}}
             child{ node [arn_n, rotate=-30] {} 
-                    child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{R}$\\$S_{M} - S_{R}$}] {} edge from parent[densely dashdotted] node[above, yshift=-.1cm, rotate=39] {$\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$}}
-                    child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{R}$\\$S_{R} - S_{M}$}] {} edge from parent[densely dashdotted] node[above, yshift=-.2cm, rotate=67] {$\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$}}
-                    child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{M}$\\$S_{M} - S_{R}$}] {} edge from parent[densely dashdotted] node[above, yshift=-.2cm, rotate=-67] {$\frac{(k-1)(N-k-1)}{(N-2)(N-3)}$}}
-                    child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{R}$\\$S_{R} - S_{R}$}] {} edge from parent[densely dashdotted] node[above, yshift=-.1cm, rotate=-41] {$\frac{(k-1)(k-2)}{(N-2)(N-3)}$}}
-                    edge from parent node[above left, xshift=1em, yshift=-.8em, rotate=-30] {$1 - \frac{1}{N - 1}$}
+                    child{ node [arn_r, label={[align=center, yshift=-.8cm] role model-resident\\learner-resident}] {} edge from parent[densely dashdotted] node[above, yshift=-.1cm, rotate=39] {$\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$}}
+                    child{ node [arn_r, label={[align=center, yshift=-.8cm] role model-resident\\learner-mutant}] {} edge from parent[densely dashdotted] node[above, yshift=-.2cm, rotate=67] {$\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$}}
+                    child{ node [arn_r, label={[align=center, yshift=-.8cm] role model-mutant  \\learner-resident}] {} edge from parent[densely dashdotted] node[above, yshift=-.2cm, rotate=-67] {$\frac{(k-1)(N-k-1)}{(N-2)(N-3)}$}}
+                    child{ node [arn_r, label={[align=center, yshift=-.8cm] role model-mutant  \\learner-mutant}] {} edge from parent[densely dashdotted] node[above, yshift=-.1cm, rotate=-41] {$\frac{(k-1)(k-2)}{(N-2)(N-3)}$}}
+                    edge from parent node[above left, xshift=1em, yshift=-1.1em, rotate=-55] {$1 - \frac{1}{N - 1}$}
             }
             edge from parent node[above, yshift=-.1cm, rotate=39] {$\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$}}
-            child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{R}$\\$S_{R} - S_{M}$}] {} edge from parent node[above, yshift=-.2cm, rotate=67] {$\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$}}
-            child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{M}$\\$S_{M} - S_{R}$}] {} edge from parent node[above, yshift=-.2cm, rotate=-67] {$\frac{(k-1)(N-k-1)}{(N-2)(N-3)}$}}
-            child{ node [arn_r, label={[align=center, yshift=-.8cm]$S_{R} - S_{R}$\\$S_{R} - S_{R}$}] {} edge from parent node[above, yshift=-.1cm, rotate=-41] {$\frac{(k-1)(k-2)}{(N-2)(N-3)}$}}
-            edge from parent node[above left, xshift=1em, yshift=-.8em, rotate=-30] {$1 - \frac{1}{N - 1}$}
+            child{ node [arn_r, label={[align=center, yshift=-.8cm] role model-resident\\learner-mutant}] {} edge from parent node[above, yshift=-.2cm, rotate=67] {$\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$}}
+            child{ node [arn_r, label={[align=center, yshift=-.8cm] role model-mutant  \\learner-resident}] {} edge from parent node[above, yshift=-.2cm, rotate=-67] {$\frac{(k-1)(N-k-1)}{(N-2)(N-3)}$}}
+            child{ node [arn_r, label={[align=center, yshift=-.8cm] role model-mutant  \\learner-mutant}] {} edge from parent node[above, yshift=-.1cm, rotate=-41] {$\frac{(k-1)(k-2)}{(N-2)(N-3)}$}}
+            edge from parent node[above left, rotate=-30] {$1 - \frac{1}{N - 1}$}
 		}; 

### 2021-06-04 clean_data.py

@@ -1,0 +1,39 @@
+import sys
+
+import glob
+
+import tqdm
+
+import pandas as pd
+
+if __name__ == "__main__":  # pragma: no cover
+    folder = sys.argv[1]
+    starting_characters = sys.argv[2]
+
+    in_columns = [
+        "y",
+        "p",
+        "q",
+        "average cooperation rate",
+        "average score",
+        "t",
+    ]
+
+    print(f"{folder}/{starting_characters}*_beta_10.csv")
+    csv_files = glob.glob(f"{folder}/{starting_characters}*_beta_10.csv")
+
+    for file in tqdm.tqdm(csv_files):
+        df = pd.read_csv(file, header=None)
+        df = df.drop_duplicates()
+        df.columns = in_columns
+
+        frequencies = []
+        previous = 0
+        for value in df["t"][1:]:
+            frequencies.append(value - previous)
+            previous = value
+        frequencies.append(10 ** 7 - previous)
+
+        df["frequencies"] = frequencies
+
+        df.to_csv(f"{folder}clean_{file[len(folder):]}", index=False)

### 2021-06-04 matlab/evolRun.m

@@ -4,28 +4,14 @@ starting_resident = [0, 0, 0];
 N = 100;
 delta = 0.999;
 beta = 1;
-numberIterations= 10 ^ 7;
-stochastic = 1;
+numberIterations = 10 ^ 7;
+payoff_types = ["two_rounds", "two_rounds_opponents"]
 
-n = 11;
-Ss = linspace(-2, 2, n);
-Ts = linspace(-1, 3, n);
+u = [2, -1, 3, 0];
 
-payoffs = zeros(n * n, 2);
-
-for i=1:n
-    for j=1:n
-        payoffs(j + (i - 1) + (i -1) * 10, 1) = Ss(i);
-        payoffs(j + (i - 1) + (i -1) * 10, 2) = Ts(j);
-    end
-end
-
-N = n * n;
-parfor (i = 1:N, 8)
-    S = payoffs(i, 1);
-    T = payoffs(i, 2);
-    u = [1, S, T, 0];
-    filename = "data/S_" + S + "_T_" + T + "_stochastic_" + stochastic;
-    evolSimulation(starting_resident, u, N, delta, beta, numberIterations, stochastic, filename);
+parfor (i = 1:2)
+    payoff_type = payoff_types(i);
+    filename = "../data/sixteen_states_" + payoff_type;
+    evolSimulationTwoRounds(starting_resident, u, N, delta, beta, numberIterations, payoff_type, filename);
 end
 end

### 2021-06-04 matlab/evolSimulationMemoryOne.m

@@ -1,4 +1,4 @@
-function [xDat, AvCoop, AvPay,payoff_type, Data]=evolSimulation(starting_resident, u, N, delta, beta, numberIterations, payoff_type, filename);
+function [xDat, AvCoop, AvPay,payoff_type, Data]=evolSimulationMemoryOne(starting_resident, u, N, delta, beta, numberIterations, payoff_type, filename);
 
 rng('default')
 %% Preparations for the output
@@ -6,8 +6,8 @@ Data=['R=',num2str(u(1)),'; S=',num2str(u(2)),'; T=',num2str(u(3)), '; P=',num2s
 AvCoop=0; AvPay=0; Res=starting_resident;
 
 %% Initialization
-sdim=3;
-xDat=zeros(numberIterations/100,6);
+sdim=5;
+xDat=zeros(numberIterations/100,8);
 xDat(1,:)=[Res, 0, u(4), 0];
 Rho=calcRho(u, beta);
 
@@ -42,10 +42,10 @@ end
 function [phi, coopMM, piMM]=calcPhi(Mut, Res, Rho, N, u, delta, beta, payoff_type);
 %% Calculating the fixation probability
 
-vMM=stationary(Mut, Mut, delta);
-vMR=stationary(Mut, Res, delta);
+vMM=stationaryMemoryOne(Mut, Mut, delta);
+vMR=stationaryMemoryOne(Mut, Res, delta);
 vRM=[vMR(1) vMR(3) vMR(2) vMR(4)];
-vRR=stationary(Res, Res, delta);
+vRR=stationaryMemoryOne(Res, Res, delta);
 
 piMM=vMM*u';
 coopMM=vMM(1)+vMM(2);
@@ -60,10 +60,6 @@ elseif payoff_type=="expected"
     piRM=vRM*u';
     piRR=vRR*u';
     phi = phiExpected(N, piMM, piMR, piRR, piRM, beta);
-
-elseif payoff_type=="two_opponents"
-
-    phi = phiTwoOpponents(N, vRM, vMM, vMR, vRR, Rho);
 
 else
     disp('Please check payoff type.')

### 2021-06-04 matlab/evolSimulationTwoRounds.m

@@ -1,6 +1,5 @@
 function [xDat, AvCoop, AvPay,payoff_type, Data]=evolSimulation(starting_resident, u, N, delta, beta, numberIterations, payoff_type, filename);
 
-rng('default')
 %% Preparations for the output
 Data=['R=',num2str(u(1)),'; S=',num2str(u(2)),'; T=',num2str(u(3)), '; P=',num2str(u(4)),'; N=',num2str(N),'; beta=',num2str(beta), '; nIt=',num2str(numberIterations), '; payoff=',payoff_type];
 AvCoop=0; AvPay=0; Res=starting_resident;
@@ -9,7 +8,7 @@ AvCoop=0; AvPay=0; Res=starting_resident;
 sdim=3;
 xDat=zeros(numberIterations/100,6);
 xDat(1,:)=[Res, 0, u(4), 0];
-Rho=calcRho(u, beta);
+Rho=calcRhoSixteen(u, beta);
 
 %% Running the evolutionary process
 j = 2;
@@ -21,7 +20,7 @@ for t = progress(1:numberIterations)
     end
 end
 
-dlmwrite(filename + ".csv", xDat, 'precision', 9);
+csvwrite(filename + ".csv", xDat);
 writematrix(Data, filename + ".txt");
 
 AvCoop = mean(xDat(:,end-2));
@@ -29,12 +28,12 @@ AvPay = mean(xDat(:,end-1));
 %time=toc
 end
 
-function [Rho]=calcRho(u, beta);
-%% Calculates all possible pairwise imitation probabilities based on one payoff
-    Rho=zeros(4,4);
-    for i1=1:4
-        for i2=1:4
-            Rho(i1,i2) = 1 / (1 + exp(-beta * (u(i2) - u(i1))));
+function [Rho]=calcRhoSixteen(u, beta);
+%% Calculates all possible pairwise imitation probabilities based on the last two payoffs
+    Rho = zeros(16, 16);
+    for i=1:16
+        for j=1:16
+            Rho(i, j) = 1 / (1 + exp(-beta * (u(1 + fix((i - 1) / 4)) + u(1 + mod(i - 1, 4)) - u(1 + fix((j - 1) / 4)) + u(1 + mod(j - 1, 4)))));
         end
     end
 end
@@ -42,28 +41,21 @@ end
[…]

### 2021-06-04 matlab/phiLastRound.m

@@ -1,0 +1,18 @@
+function [phi]=phiLastRound(N, vRM, vMM, vMR, vRR, Rho);
+%% Calculates the fixation probability based on the last round payoffs
+
+laplus=zeros(1, N-1); laminus=laplus;
+for k=1:N-1
+    x=zeros(4,4);
+    for i1=1:4
+        for i2=1:4
+            x(i1,i2)=1/(N-1)*vRM(i1)*((i1==1 & i2==1) | (i1==2 & i2==3) | (i1==3 &i2==2) | (i1==4 & i2==4))...
+                + (1-1/(N-1))/(N-2)/(N-3) * ((k-1)*(k-2)*vRM(i1)*vMM(i2) + (k-1)*(N-k-1)*vRM(i1)*vMR(i2) + (N-k-1)*(k-1)*vRR(i1)*vMM(i2) + (N-k-1)*(N-k-2)*vRR(i1)*vMR(i2));
+        end
+    end
+    laplus(k)=sum(sum(x.*Rho));
+    laminus(k)=sum(sum(x.*Rho'));
+end
+
+phi = 1 / (1 + sum(cumprod(laminus./laplus)));
+end

### 2021-06-04 matlab/phiTwoRounds.m

@@ -1,0 +1,22 @@
+function [phi]=phiTwoRounds(N, vRM, vMM, vMR, vRR, Rho);
+%% Calculates the fixation probability based on the last two rounds payoff
+
+laplus = zeros(1, N-1); laminus=laplus;
+for k=1:N-1
+    x=zeros(16,16);
+    for i1=1:16
+        for i2=1:16
+x(i1,i2) = 1 / (N-1) * vRM(i1) * ((i1==1 & i2==1)  |(i1==2 & i2==3)  |(i1==3 &i2==2)   |(i1==4 & i2==4) |(i1==5 & i2==9) |(i1==6 & i2==11)|...
+                                  (i1==7 & i2==10) |(i1==8 & i2==12) |(i1==9 & i2==5)  |(i1==10 & i2==7)|(i1==11 & i2==6)|(i1==12 & i2==8)|...
+                                  (i1==13 & i2==13)|(i1==14 & i2==15)|(i1==15 & i2==14)|(i1==16 & i2==16))...
+         + (1 - 1 / (N - 1)) / (N - 2) / (N - 3) * ((k - 1) * (k - 2) * vRM(i1) * vMM(i2) ...
+         + (k - 1) * (N - k - 1) * vRM(i1) * vMR(i2) + (N - k - 1) * (k - 1) * vRR(i1) * vMM(i2) ...
+         + (N - k - 1) * (N - k - 2) * vRR(i1) * vMR(i2));
+        end
+    end
+laplus(k) = sum(sum(x.*Rho));
+laminus(k) = sum(sum(x.*Rho'));
+end
+
+phi = 1 / (1 + sum(cumprod(laminus./laplus)));
+end

### 2021-06-04 matlab/stationary.m

@@ -1,16 +1,17 @@
-function v=stationary(s1,s2,delta); 
+function v=stationary(s1, s2, delta);
+% Calculates the probabilities of being at CC, CD, DC, DD at the last round.
 
 y1=s1(1); p1=s1(2); q1=s1(3); r1=p1-q1;
 y2=s2(1); p2=s2(2); q2=s2(3); r2=p2-q2;
-y1b=1-y1; q1b=1-q1; r1b=-r1; 
-y2b=1-y2; q2b=1-q2; r2b=-r2; 
+y1b=1-y1; q1b=1-q1; r1b=-r1;
+y2b=1-y2; q2b=1-q2; r2b=-r2;
 
-D1 =1 - delta ^ 2 * r1 * r2; D2 = (1 - delta * r1 * r2) * D1; 
+D1 = 1 - delta ^ 2 * r1 * r2; D2 = (1 - delta * r1 * r2) * D1;
 
-v1=(1-delta)*y1*y2/D1+delta*(q1+r1*((1-delta)*y2+delta*q2))*(q2+r2*((1-delta)*y1+delta*q1))/D2; 
+v1=(1-delta)*y1*y2/D1+delta*(q1+r1*((1-delta)*y2+delta*q2))*(q2+r2*((1-delta)*y1+delta*q1))/D2;
 v2=(1-delta)*y1*y2b/D1+delta*(q1+r1*((1-delta)*y2+delta*q2))*(q2b+r2b*((1-delta)*y1+delta*p1))/D2;
-v3=(1-delta)*y1b*y2/D1+delta*(q1b+r1b*((1-delta)*y2+delta*p2))*(q2+r2*((1-delta)*y1+delta*q1))/D2; 
-v4=(1-delta)*y1b*y2b/D1+delta*(q1b+r1b*((1-delta)*y2+delta*p2))*(q2b+r2b*((1-delta)*y1+delta*p1))/D2; 
+v3=(1-delta)*y1b*y2/D1+delta*(q1b+r1b*((1-delta)*y2+delta*p2))*(q2+r2*((1-delta)*y1+delta*q1))/D2;
+v4=(1-delta)*y1b*y2b/D1+delta*(q1b+r1b*((1-delta)*y2+delta*p2))*(q2b+r2b*((1-delta)*y1+delta*p1))/D2;
 
 v=[v1,v2,v3,v4];
 end

### 2021-06-04 matlab/stationaryMemoryOne.m

@@ -1,0 +1,14 @@
+function v=stationaryMemoryOne(p1,p2,delta);
+% Calculates the invariant distribution v
+% for the game between two memory-1 players with strategies p1, p2.
+% p1=(pCC, pCD, pDC, pDD, p0);
+
+M = [p1(1)*p2(1), p1(1)*(1-p2(1)), (1-p1(1))*p2(1), (1-p1(1))*(1-p2(1));
+     p1(2)*p2(3), p1(2)*(1-p2(3)), (1-p1(2))*p2(3), (1-p1(2))*(1-p2(3));
+     p1(3)*p2(2), p1(3)*(1-p2(2)), (1-p1(3))*p2(2), (1-p1(3))*(1-p2(2));
+     p1(4)*p2(4), p1(4)*(1-p2(4)), (1-p1(4))*p2(4), (1-p1(4))*(1-p2(4))];
+
+    v0=[p1(5)*p2(5), p1(5)*(1-p2(5)), (1-p1(5))*p2(5), (1-p1(5))*(1-p2(5))];
+
+v=(1-delta)*v0/(eye(4)-delta*M);
+end

### 2021-06-04 matlab/stationaryRoundTwo.m

@@ -1,0 +1,19 @@
+function v=stationaryRoundTwo(s1,s2,delta);
+
+M=[s1(2) * s2(2), s1(2) * (1-s2(2)), (1-s1(2)) * s2(2), (1-s1(2)) * (1-s2(2));
+    s1(3) * s2(2), s1(3) * (1-s2(2)), (1-s1(3)) * s2(2), (1-s1(3)) * (1-s2(2));
+    s1(2) * s2(3), s1(2) * (1-s2(3)), (1-s1(2)) * s2(3), (1-s1(2)) * (1-s2(3));
+    s1(3) * s2(3), s1(3) * (1-s2(3)), (1-s1(3)) * s2(3), (1-s1(3)) * (1-s2(3))];
+
+v0 = [s1(1) * s2(1), s1(1) * (1-s2(1)), (1-s1(1)) * s2(1), (1-s1(1)) * (1-s2(1))];
+v = zeros(4, 4);
+rhs = v0 / (eye(4) - delta * M);
+
+for i=1:4
+    for j=1:4
+        v(i, j) = (1 - delta) * M(i, j) * (delta ^ 2) * rhs(i);
+    end
+end
+
+v = reshape(v, [1, 16]);
+end

### 2021-06-04 soup_opponents.py

@@ -1,0 +1,48 @@
+import evol_dynamics
+
+import numpy as np
+
+import tqdm
+
+import sys
+
+
+if __name__ == "__main__":  # pragma: no cover
+    filename = "data/soup_for_opponents.csv"
+    number_of_checks = 100
+    delta = 0.999
+    num_of_opponents = 2
+
+    seed = int(sys.argv[1])
+    number_of_repetitions = int(sys.argv[2])
+    random_state = np.random.RandomState(seed)
+
+    for i in tqdm.tqdm(range(number_of_checks)):
+        population_size = 2 * (random_state.randint(50, 150) // 2)
+        number_of_mutants = random_state.randint(1, population_size - 1)
+
+        simulated_opponents = evol_dynamics.get_probabilities_for_opponents(
+            number_of_repetitions,
+            population_size,
+            number_of_mutants,
+            random_state,
+            num_of_opponents,
+        )
+        theoretical_opponents = (
+            evol_dynamics.theoretical_probabilities_for_opponents(
+                population_size, number_of_mutants
+            )
+        )
+
+        data = [
+            i,
+            number_of_repetitions,
+            population_size,
+            number_of_mutants,
+            *simulated_opponents.values(),
+            *theoretical_opponents.values(),
+        ]
+
+        with open(filename, "a") as textfile:
+            textfile.write(",".join([str(elem) for elem in data]) + "\n")
+        textfile.close()

### 2021-06-04 soup_round_two_states.py

@@ -1,0 +1,46 @@
+import evol_dynamics
+
+import numpy as np
+
+import tqdm
+
+import sys
+
+
+if __name__ == "__main__":  # pragma: no cover
+    filename = "data/soup_for_sixteen_states_stationary.csv"
+    number_of_checks = 100
+    delta = 0.999
+
+    seed = int(sys.argv[1])
+    number_of_repetitions = int(sys.argv[2])
+    random_state = np.random.RandomState(seed)
+
+    for i in tqdm.tqdm(range(number_of_checks)):
+        mutant = [random_state.random() for _ in range(3)]
+        resident = [random_state.random() for _ in range(3)]
+
+        v_last_two_rounds = evol_dynamics.stationary_for_16_states(
+            mutant, resident, delta
+        )
+        v_last_round = evol_dynamics.expected_distribution_last_round(
+            mutant, resident, delta
+        )
+        v_simulated = evol_dynamics.simulated_states(
+            mutant, resident, delta, number_of_repetitions, rounds_of_history=2
+        )
+
+        data = [
+            i,
+            number_of_repetitions,
+            *mutant,
+            *resident,
+            delta,
+            *v_last_two_rounds,
+            *v_last_round,
+            *v_simulated.values(),
+        ]
+
+        with open(filename, "a") as textfile:
+            textfile.write(",".join([str(elem) for elem in data]) + "\n")
+        textfile.close()

### 2021-06-04 src/evol_dynamics/__init__.py

@@ -18,9 +18,10 @@ from .numerical import (
 )
 from .simulation import (
     ReactivePlayer,
-    get_opponents_of_mutant,
-    get_opponents_of_resident,
-    get_score_for_last_n_turns,
-    introduce_mutant,
-    simulation,
+    create_population,
+    match_pairs,
+    stationary_for_16_states,
+    simulated_states,
+    get_probabilities_for_opponents,
+    theoretical_probabilities_for_opponents,
 )

### 2021-06-04 src/evol_dynamics/numerical.py

@@ -32,6 +32,16 @@ def donation_game(c, b):
     return (b - c, -c, b, 0)
 
 
+def is_donation(payoffs, dis=False):
+    c = -payoffs[1]
+    b = payoffs[2]
+
+    if dis:
+        return (payoffs[0] == b - c, "c:{c} and b:{b}")
+
+    return payoffs[0] == b - c
+
+
 def snowdrift_game(c, b):
     """The snowdrift game
 
@@ -52,6 +62,13 @@ def snowdrift_game(c, b):
         The payoff vector for the snowdrift game.
     """
     return (b - (c / 2), b - c, b, 0)
+
+
+def is_snowdrift(payoffs):
+    b = payoffs[2]
+    c = b - payoffs[1]
+
+    return payoffs[0] == b - (c / 2)
 
 
 def stag_hunt_game():

### 2021-06-04 src/evol_dynamics/simulation.py

@@ -1,36 +1,14 @@
+import copy
 import itertools
-import multiprocessing
-import os
-import random
 import sys
-import time
 
 import axelrod as axl
 import numpy as np
 import tqdm
 
+import evol_dynamics
 
-def introduce_mutant(population, resident, random_state):
-    """Introduces a mutant in a population of residents.
-
-    Parameters
-    ----------
-    population : list
-        A list of dictionaries. Each dictionary described the population at a
-        given time step.
-    resident : tuple
-        A reactive tuple for the strategy which is currently the resident in the
-        population.
-
-    Returns
-    -------
-    tuple, list
-        The mutant as a tuple but also the updated list of the population
-    """
-    mutant = tuple(random_state.random() for _ in range(3))
-    population = {resident: population[resident] - 1, mutant: 1}
-
-    return mutant, population
+from collections import Counter
 
 
 class ReactivePlayer(axl.MemoryOnePlayer):
@@ -39,259 +17,397 @@ class ReactivePlayer(axl.MemoryOnePlayer):
     opponent's last move (P(C|C), P(C|D)) and the opening probability.
     """
 
-    name = "Reactive Player"
+    def __init__(self, probabilities, role):
 
-    def __init__(self, probabilities):
         self.initial = np.random.choice(
             [axl.Action.C, axl.Action.D],
             p=(probabilities[0], 1 - probabilities[0]),
         )
         self.four_vector = (*probabilities[1:], *probabilities[1:])
+        self.name = role
+
         super().__init__(self.four_vector, self.initial)
 
 
-def get_score_for_last_n_turns(
-    player, opponents, num_of_interactions, delta, seed=None
-):
-    """Score an individual against their opponents. The score is based on the
-    last number of interactions.
+def create_population(population_size, number_of_mutants, random_state):
 
-    The individual is being scored on a probabilistic ending against each
-    opponent. The match lengths can vary at this point.
+    population = ["role-model", "learner"]
 
-    Parameters
-    ----------
-    player : tuple
-        A reactive player tuple
-    opponents : list of tuples
-        A list of reactive opponents
-    num_of_interactions : int
-        The number of scores the player remembers
-    delta : float
-        The probability of the match continuing.
+    population += [
+        "resident" for _ in range(population_size - number_of_mutants - 1)
+    ]
 
-    Returns
-    -------
-    float
-        The sum of the scores per turns the player achieved against the opponents.
-    """
-    score_per_turn = []
[…]

### 2021-06-04 tests/test_simulation.py

@@ -1,45 +1,31 @@
-import os
-
 import axelrod as axl
 import numpy as np
-import pandas as pd
 import pytest
 
 import evol_dynamics
 
 
-def test_introduce_mutant():
-    N = 10
-    resident = (1, 1, 1)
-    population = {resident: N}
-    random_ = np.random.RandomState(0)
-
-    mutant, population = evol_dynamics.introduce_mutant(
-        population, resident, random_
-    )
-    expected_mutant = [
-        0.5488135039273248,
-        0.7151893663724195,
-        0.6027633760716439,
-    ]
-
-    assert len(population) == 2
-    assert population[resident] == 9
-    assert set(expected_mutant) == set(mutant)
-    assert population[mutant] == 1
+resident = [np.random.random() for _ in range(3)]
+mutant = [np.random.random() for _ in range(3)]
+delta = 0.999
+population_size = 10
+number_of_mutants = 3
+repetitions = 100
+seed = 5
+random_state = np.random.RandomState(seed)
 
 
 def test_reactive_player():
 
-    player = evol_dynamics.ReactivePlayer((0, 0, 0))
-    assert player.name == "Reactive Player"
+    player = evol_dynamics.ReactivePlayer((0, 0, 0), role="resident")
     assert player.initial == axl.Action.D
     assert player.four_vector == (0, 0, 0, 0)
+    assert player.name == "resident"
 
-    opponent = evol_dynamics.ReactivePlayer((1, 1, 0))
-    assert opponent.name == "Reactive Player"
+    opponent = evol_dynamics.ReactivePlayer((1, 1, 0), role="generic")
     assert opponent.initial == axl.Action.C
     assert opponent.four_vector == (1, 0, 1, 0)
+    assert opponent.name == "generic"
 
     match = axl.Match([player, opponent], turns=5)
     results = match.play()
@@ -48,362 +34,151 @@ def test_reactive_player():
     assert results[-1] == (axl.Action.D, axl.Action.D)
 
 
-def test_get_score_for_last_n_turns_last_two_interactions():
-    player = (1, 1, 0)
-    opponents = [(1, 1, 1), (0, 1, 1), (0, 0, 0), (0, 1, 0)]
+def test_create_population():
 
-    score = evol_dynamics.get_score_for_last_n_turns(
-        player, opponents, num_of_interactions=2, delta=0.90, seed=1
+    population = evol_dynamics.create_population(
[…]

### 2021-11-24 main.tex
simple commit
testing something regarding the line lenght

@@
 \date{}
-\author{Christian Hilbe, Nikoleta E. Glynatsi, Alex McAvoy}
+\author{Nikoleta E. Glynatsi, Christian Hilbe, Alex McAvoy}
 
@@
 \begin{abstract}
-Repeated games are vastly used in evolutionary game theory to explain
-cooperation in a variety of environments. Although existing models have shaped
-our understanding of human cooperation, they often work with idealized
-assumptions. In an evolutionary process, individuals imitate other individuals
+Repeated games are vastly used in evolutionary ame theory to explain cooperation in a variety of environments.
+Although existing models have shaped our understanding of human cooperation, they often work with idealized
+assumptions. In an evolutionary process, individuals aimitate other individuals
 of the population based on their fitness. It is commonly assumed that an
@@
 \section{Model Setup}\label{section:model}
+
+# This needs an intro, guiding readers into the model.
+
+Just words.
 

### 2021-11-24 main.tex
fixing errors

@@
 %% Adding shortcut commands to refer to our figures %%
-\newcommand{\FigEvoProc}{{\bf Fig.~1}}
-\newcommand{\FigInvAnalysis}{{\bf Fig.~2}}
-\newcommand{\FigResultsOverPara}{{\bf Fig.~3}}
+\newcommand{\FigEvoProc}{{\bf Fig.~1}} \newcommand{\FigInvAnalysis}{{\bfFig.~2}} \newcommand{\FigResultsOverPara}{{\bf Fig.~3}}
 
 \titleformat{\section}{\sffamily \fontsize{12}{14}\bfseries}{\thesection}{1em}{}
-\titleformat{\subsection}{\sffamily \fontsize{11.5}{11.5}\bfseries}{\thesubsection}{1em}{}
+\titleformat{\subsection}{\sffamily
+\fontsize{11.5}{11.5}\bfseries}{\thesubsection}{1em}{}
 
@@
 
-\tikzset{
-  treenode/.style = {align=center, inner sep=0pt, text centered,
-    font=\sffamily},
-  arn_n/.style = {treenode, circle, white, font=\sffamily\bfseries, draw=black, inner sep=-6pt,
-    fill=black, text width=1.5em},% arbre rouge noir, noeud noir
-  arn_r/.style = {treenode, circle, red, 
-    text width=1.5em, very thick, inner sep=4pt},% arbre rouge noir, noeud rouge
-  arn_x/.style = {treenode, rectangle, draw=black,
-    minimum width=0.5em, minimum height=0.5em}% arbre rouge noir, nil
+\tikzset{treenode/.style = {align=center, inner sep=0pt, text centered,
+  font=\sffamily}, arn_n/.style = {treenode, circle, white,
+  font=\sffamily\bfseries, draw=black, inner sep=-6pt, fill=black, text
+  width=1.5em},% arbre rouge noir, noeud noir
+  arn_r/.style = {treenode, circle, red, text width=1.5em, very thick, inner
+    sep=4pt},% arbre rouge noir, noeud rouge
+  arn_x/.style = {treenode, rectangle, draw=black, minimum width=0.5em, minimum
+    height=0.5em}% arbre rouge noir, nil
 }
@@
   \begingroup
-  \begingroup\lccode`~=`, \lowercase{\endgroup
-    \edef~{\mathchar\the\mathcode`, \penalty0 \noexpand\hspace{0pt plus 1em}}%
+  \begingroup\lccode`~=`, \lowercase{\endgroup \edef~{\mathchar\the\mathcode`,
+    \penalty0 \noexpand\hspace{0pt plus 1em}}%
   }\mathcode`,="8000 #1%
@@
 
-\title{
-\bf  \sffamily \LARGE Evolution of cooperation among individuals with limited payoff memory\\}
+\title{\bf  \sffamily \LARGE Evolution of cooperation among individuals with
+limited payoff memory\\}
 \date{}
@@
 \begin{abstract}
-Repeated games are vastly used in evolutionary ame theory to explain cooperation in a variety of environments.
-Although existing models have shaped our understanding of human cooperation, they often work with idealized
-assumptions. In an evolutionary process, individuals aimitate other individuals
-of the population based on their fitness. It is commonly assumed that an
-individual computes their fitness after interacting with a representative sample
-of the population, and remembering all the interactions they participated in. In
-real life, we do not always remember all our interactions, instead we rather
-recall our most recent ones.
+% Repeated games are vastly used in evolutionary game theory to explain
+% cooperation in a variety of environments. Although existing models have shaped
+% our understanding of human cooperation, they often work with idealized
+% assumptions. In an evolutionary process, individuals imitate other individuals
+% of the population based on their fitness. It is commonly assumed that an
+% individual computes their fitness after interacting with a representative sample
+% of the population, and remembering all the interactions they participated in. In
+% real life, we do not always remember all our interactions, instead we rather
+% recall our most recent ones.
 
-Here, we introduce a framework that allows individuals to estimate their fitness
-based on a minimum of social information. We explore the difference between our
-framework and the classical framework using computer simulations. We present results for
-the most commonly used classes of symmetric \(2 \times 2\) games. The
-simulations show that, in the prisoner's dilemma individuals with limited memory
-tend to adopt less generous strategies and they achieve less cooperation than in
-the classical scenario. In contrast, in the stag-hunt and the harmony game, the
-impact of memory is less striking. Here individuals with limited memory perform
-nearly as well as individuals with full memory. Finally, we observe that the
-snowdrift game is the only game for which cooperation can be underestimated in
-the classical scenario.
+% Here, we introduce a framework that allows individuals to estimate their fitness
+% based on a minimum of social information. We explore the difference between our
+% framework and the classical framework using computer simulations. We present
+% results for the most commonly used classes of symmetric \(2 \times 2\) games.
+% The simulations show that, in the prisoner's dilemma individuals with limited
+% memory tend to adopt less generous strategies and they achieve less cooperation
+% than in the classical scenario. In contrast, in the stag-hunt and the harmony
+% game, the impact of memory is less striking. Here individuals with limited
+% memory perform nearly as well as individuals with full memory. Finally, we
+% observe that the snowdrift game is the only game for which cooperation can be
+% underestimated in the classical scenario.
 \end{abstract}
@@
 
-One of the most important applications of evolutionary game theory is the
-evolution of cooperation. Why is it that some individuals choose to help others,
-increasing their payoff, at the expense of decreasing one's own? In evolutionary
-game theory individuals are not required to be rational, instead they adapt
-strategies based on mutation and exploration. Strategies are more likely to
-spread if they have a high fitness either because the individuals who adopt them
-have more offspring, or because they are imitated more often~\cite{Wu2015}. The
-fitness of a strategy is not constant but depends on the composition of the
-population. Individuals interact based on their strategies with other members of
-the population and the payoffs they yield are translated into fitness.
+Evolutionary game theory~\cite{smith1982evolution, hofbauer1998evolutionary,
+nowak2004evolutionary, hauert2005game} describes the evolutionary dynamics of
+populations consisting of different types of interacting individuals.
+Traditional approaches assume that individuals meet each other at random in an
+infinitely large well-mixed population. An example of such a approach in the the
+replicator dynamics. It describes how the abundance of strategic types in a
+population changes based on their fitness, identified with the payoff resulting
+from the game~\cite{cressman2014replicator}. In this deterministic formulation,
+individuals with higher fitness, such as defectors, increase in abundance.
+Ultimately, the system reaches a stable fixed point in which the population may
+consist either of a single type or of a mixture of different types.
 
-It is commonly assumed that individuals compute their fitness after interacting
-with a representative sample of the population, and remembering all the
-interactions they participated in~\cite{Nowak2006, nowak2004emergence,
-Hilbe2013, Hilbe2018,Reiter2018, garcia2018no, Nowak1992tit, Baek2016}. This
-implies that individuals have a perfect memory. However, when modelling how
-individuals make decisions in each turn they are assumed to have very limited
-memory. To be precise, most of the studies in the literature focus on naive
-subjects who can only choose from a restricted set of
-strategies~\cite{Nowak1992tit}, or who do not remember anything beyond the
-outcome of the very last round~\cite{Baek2016}. Note that there are a few
-notable exceptions~\cite{Hauert1997, Stewart2016}.
+It has been shown that the finiteness of populations may lead to fundamental
+changes in this picture due to stochastic effects In this framework
+disadvantageous mutants have a small, yet non-zero probability to reach fixation
+in a finite population.
 
-Studies that make use of this classical framework report that cooperation can
-substantially evolve. However, the inconsistency in the memory size of players
-leads us to question the robustness of our understanding of cooperation. To this end, we
-propose a framework in which individuals, similar to the decisions at each turn,
-estimate their fitness based on a minimum of information. Though we are not the
-first to question the assumptions of estimating fitness~\cite{Roca2006}, we are
-the first to explore the effect of memory.
+Its most prominent application is the quest for the origins and 
+evolution of cooperation.
+An important application of evolutionary game theory is the evolution of
+cooperation which remains one of the greatest problems for biological and social
+sciences.
 
-We first consider two extreme scenarios, the classical scenario
-and the alternative scenario where individuals update their strategies
-only based on the very last payoff they obtained. We observe that individuals 
-with limited memory tend to adopt less generous strategies and they achieve less
-cooperation when interacting in a prisoner's dilemma. We obtain similar results when we
-consider that individuals update their strategies based on more
-information. More specifically, up to the last two payoffs they obtained when
-interacting with up to two different members of the population. We extend our
-approach to the rest of the symmetric \(2 \times 2\) games.
+% To explain the evolution of cooperation remains one of the greatest problems for
+% biological and social sciences. Cooperation is the action of choosing to help
+% others at one's own expense. The standard game of formulating such situation is
+% the Prisoner's Dilemma~\cite{trivers1971evolution,
+% milinski1987tit, glynatsi2021bibliometric, ohtsuki2006simple}, in which two
+% players can choose to cooperate or to defect. The players are offered a certain
+% payoff, \(R\), for mutual cooperation and a lower payoff, \(P\), for mutual
+% defection. If one player cooperates while the other defects, then the cooperator
+% gets the lowest payoff, \(S\), while the defector gains the highest payoff,
+% \(T\). Thus, the payoffs have the following property \(T > R > P > S\), making
+% defection the dominant strategy in the non-repeated game.
+
+% In the repeated Prisoner's Dilemma, direct reciprocity can allow cooperators to
+% not be exploited by defectors~\cite{axelrod1981evolution, nowak1992tit,
+% nowak1992evolutionary}, however, how do such cooperators can evolve?
+
+ , and 
+in general, defectors are stable against invasion by cooperators.
+
+In this framework disadvantageous mutants have a small, yet non-zero probability
+to reach fixation in a finite population.
+
+Two classes of microscopic processes have been used extensively: (i)
+fitness-based processes in which an individual chosen proportional to fitness
+reproduces and the offspring replaces a randomly chosen individual [11]; (ii)
+pairwise comparison processes in which a pair of individuals is chosen, and
+where subsequently one of these individuals may adopt the strategy of the other.
+
+In the pairwise process considers a finite population of fixed size.In the
+population at a time step different types of individuals can exist. 
+Each individual can be of one of two types, A and B. The state of the population
+is thus characterized by the number i of individuals of type A.
+
+The interaction between the two types of individuals is described by the
+functions $\pi _{{\rm A}}^{i}$ and $\pi _{{\rm B}}^{i}$. These indicate the
+expected payoff for two types in a population in state i. The interaction can be
+generated from a two-player matrix game which leads to payoffs linear in i [26],
+but we keep the formalism general to include games played between an arbitrary
+number of players, which leads to payoffs that depend on i in a polynomial way
+[27–30]. In fact, our results hold for an arbitrary dependence of the payoffs on
+i.
+
+At each step an individual change their type based. This very cooperators
+can repeas.
+
+This model has several assumptions. Two assumptions are that initally it is
+assumed that time to interact is faster than the time of the population
+update. Secondly it is assumed that the indiviail can interact will every
+other population member, and recall their interaction perfecttly.
+
+The first assumption was questioned by the work of.
+
+In this work
+
+We first consider two extreme scenarios, the classical scenario and the
+alternative scenario where individuals update their strategies only based on the
+very last payoff they obtained. We observe that individuals with limited memory
+tend to adopt less generous strategies and they achieve less cooperation when
+interacting in a prisoner's dilemma. We obtain similar results when we consider
+that individuals update their strategies based on more information. More
+specifically, up to the last two payoffs they obtained when interacting with up
+to two different members of the population. We extend our approach to the rest
+of the symmetric \(2 \times 2\) games.
 
@@
 section~\ref{section:model} we describe the model. In
-section~\ref{section:results} we present the results of the simulations, and
-in section~\ref{section:conclusions} we outline the main conclusions.
+section~\ref{section:results} we present the results of the simulations, and in
+section~\ref{section:conclusions} we outline the main conclusions.
 
@@
 
-# This needs an intro, guiding readers into the model.
+In the following, we consider a well mixed population of fixed size. In each
+step of t
 
-Just words.
+We study the transmission of strategies with a frequency-dependent birth-death
+process26 in a finite population of size n. In each time step, two randomly
+chosen individuals compare their payoffs and one of them can switch to the other
+one's strategy. This process can be interpreted as a model for social learning,
+whereby successful strategies spread, and, occasionally, random strategy
+exploration introduces novel strategies (corresponding to mutations in
+biological models).
 
@@
 a \textit{mutant} strategy. To describe how strategies spread we use a pairwise
-comparison process~\cite{Traulsen2006}.
-Each step of the evolutionary process consists of two
-stages, a game stage and an updating stage.
+comparison process~\cite{Traulsen2006}. Each step of the evolutionary process
+consists of two stages, a game stage and an updating stage.
 
 In the game stage each individual is randomly matched with some other individual
-in the population. They engage in a match where each subsequent turn
-occurs with a fixed probability $\delta$. At each turn players choose independently to
+in the population. They engage in a match where each subsequent turn occurs with
+a fixed probability $\delta$. At each turn players choose independently to
 either cooperate (\(C\)) or to defect (\(D\)), and the payoffs of the turn
@@
 \(P\). If one cooperates but the other defects, the defector receives the
-temptation payoff \(T\), whereas the cooperator receives the sucker's
-payoff \(S\). We denote the feasible payoff of each turn as \(\mathcal{U} =
-\{R, S, T, P\}\).
-We assume that individuals use \textit{reactive strategies} to make
+temptation payoff \(T\), whereas the cooperator receives the sucker's payoff
+\(S\). We denote the feasible payoff of each turn as \(\mathcal{U} = \{R, S, T,
+P\}\). We assume that individuals use \textit{reactive strategies} to make
 decisions in each turn. Reactive strategies are a set of memory-one strategies
@@
 written explicitly as a vector in \(\R_{3}\), more specifically, a reactive
-strategy \(s\) is given by \(s=(y, p, q)\). The parameter \(y\) is the probability that
-the strategy opens with a cooperation and \(p, q\) are the probabilities that
-the strategy cooperates given that the opponent cooperated and defected
-equivalently.
+strategy \(s\) is given by \(s=(y, p, q)\). The parameter \(y\) is the
+probability that the strategy opens with a cooperation and \(p, q\) are the
+probabilities that the strategy cooperates given that the opponent cooperated
+and defected equivalently.
 
@@
 `learner' and a `role model'. Given the learner's payoff $u_L\!\in\!
-\mathcal{U}$ and the role model's payoff $u_{RL}\!\in\! \mathcal{U}$,
-the learner adopts the role model's strategy with probability,
+\mathcal{U}$ and the role model's payoff $u_{RL}\!\in\! \mathcal{U}$, the
+learner adopts the role model's strategy with probability,
 
@@
 players. As the value of $\beta$ increases, the more likely it is that the
-learner adopts only strategies that yield a higher payoff.
-Conventionally the updating payoffs of the learner and the role model are based
-on their expected payoffs. A player's expected payoff is the mean payoff the
-player yields after engaging in matches of multiple turns with each member of
-the population. At each match a player bases their next turn decision only on
-the previous action of the opponent, however, the same player bases their
-expected payoffs on the outcomes of all their matches. Thus, a player is assumed
-to have limited and perfect memory at the same time. We propose a new a set of
-updating payoffs where it is also assumed that a player has limited memory.
-We referee to these as the limited memory payoffs.
+learner adopts only strategies that yield a higher payoff. Conventionally the
+updating payoffs of the learner and the role model are based on their expected
+payoffs. A player's expected payoff is the mean payoff the player yields after
+engaging in matches of multiple turns with each member of the population. At
+each match a player bases their next turn decision only on the previous action
+of the opponent, however, the same player bases their expected payoffs on the
+outcomes of all their matches. Thus, a player is assumed to have limited and
+perfect memory at the same time. We propose a new a set of updating payoffs
+where it is also assumed that a player has limited memory. We referee to these
+as the limited memory payoffs.
 
-The evolutionary step is repeated until either the mutant strategy goes
-extinct, or until it fixes in the population. If the mutant fixes in the
-population then the mutant strategy becomes the new resident strategy. After
-either outcome we introduce a new mutant strategy uniformly chosen from all
-reactive strategies at random, and we set the number of mutants to $1$. This
-process of mutation and fixation/extinction is then iterated many times.
+The evolutionary step is repeated until either the mutant strategy goes extinct,
+or until it fixes in the population. If the mutant fixes in the population then
+the mutant strategy becomes the new resident strategy. After either outcome we
+introduce a new mutant strategy uniformly chosen from all reactive strategies at
+random, and we set the number of mutants to $1$. This process of mutation and
+fixation/extinction is then iterated many times.
 
@@
 
-\section{Results}\label{section:results}
-\subsection{Updating payoffs based on the last round with another member of the population}\label{section:donation}
+\section{Results}\label{section:results} \subsection{Updating payoffs based on
+the last round with another member of the population}\label{section:donation}
+
+To investigate the role of extortion in the context of evolutionary games, we
+concentrate on the donation game
 
 In this section we explore the case where the updating payoffs are based on the
-last round payoff achieved against another member of the population and we compare
-this to the classical scenario of the expected payoffs. We assume that each pair
-of players interacts in a donation game. The donation game is a special case of
-the prisoner's dilemma. Each player can choose to cooperate by providing a
-benefit \(b\) to the other player at their cost \(c\), with \(0 < c < b\).
-Thus, the feasible payoffs in each round are \(\mathcal{U} = \{b-c, -c, b, 0\}\).
+last round payoff achieved against another member of the population and we
+compare this to the classical scenario of the expected payoffs. We assume that
+each pair of players interacts in a donation game. The donation game is a
+special case of the prisoner's dilemma. Each player can choose to cooperate by
+providing a benefit \(b\) to the other player at their cost \(c\), with \(0 < c
+< b\). Thus, the feasible payoffs in each round are \(\mathcal{U} = \{b-c, -c,
+b, 0\}\).
 
@@
 conditional cooperation probabilities $p$ and $q$. The discount factor~$\delta$
-is comparably high, thus we do not report the opening move \(y\) as it
-is a transient effect. The left panels correspond to the
-standard scenario considered in the literature, it considers players who use
-expected payoffs to update their strategies. The right panel shows the scenario
-considered herein, in which players update their strategies based on their last
-round’s payoff. The top panels assume a benefit \(b\) of 3 whereas the bottom
-assume a benefit of 10.
+is comparably high, thus we do not report the opening move \(y\) as it is a
+transient effect. The left panels correspond to the standard scenario considered
+in the literature, it considers players who use expected payoffs to update their
+strategies. The right panel shows the scenario considered herein, in which
+players update their strategies based on their last round’s payoff. The top
+panels assume a benefit \(b\) of 3 whereas the bottom assume a benefit of 10.
 
-The figure suggests that when updating is based on expected payoffs players
-tend to be more generous and more cooperative. The $q$-values of the resident
+The figure suggests that when updating is based on expected payoffs players tend
+to be more generous and more cooperative. The $q$-values of the resident
 strategies are on average higher in the case of the expected payoffs. The
@@
 when social interactions are limited they are less forgiving. The average
-cooperation rate for each simulation is calculated as the average
-cooperation rate within the resident population. In the case of the expected
-payoffs, regardless the value of benefit, the average cooperation rate is
-strictly higher than that of the last round payoffs. The difference based on the
-two methods is statistically significant, and in the case of $b=10$ the
-average cooperation of resident strategies drops from 97\% to 57\%.
+cooperation rate for each simulation is calculated as the average cooperation
+rate within the resident population. In the case of the expected payoffs,
+regardless the value of benefit, the average cooperation rate is strictly higher
+than that of the last round payoffs. The difference based on the two methods is
+statistically significant, and in the case of $b=10$ the average cooperation of
+resident strategies drops from 97\% to 57\%.
 
@@
     \centering
-    \includegraphics[width=.70\textwidth]{static/expected_and_stochastic_for_donation_game.pdf}\vspace{-3cm}
-    \includegraphics[width=.70\textwidth]{static/expected_and_stochastic_for_donation_game_b_10.pdf}
+    \includegraphics[width=.70\textwidth]{static/expected_and_stochastic_for_donation_game.pdf}
+    % \includegraphics[width=.70\textwidth]{static/expected_and_stochastic_for_donation_game_b_10.pdf}
     \caption{{\bf Evolutionary dynamics under expected payoffs and last round with one interaction payoffs.} 
@@
     section~\ref{section:model} for $t\!=\!10^7$ time steps. For each time step,
-    we have recorded the current resident population ($y,p,q$). Since simulations
-    are run for a relatively high continuation probability of $\delta\!=\!0.999$, we
-    do not report the players' initial cooperation probability $y$. The graphs show
-    how often the resident population chooses each combination ($p,q$) of
-    conditional cooperation probabilities in the subsequent rounds. ({\bf A}) If
-    players update based on their expected payoffs, the resident population
-    typically applies a strategy for which $p\!\approx\!1$ and
-    $q\!\le\!1\!-\!c/b\!=\!0.9$. ({\bf B})
-    When players update their strategies based on their realized payoffs in the last
-    round, there are two different predominant behaviors. The resident population
+    we have recorded the current resident population ($y,p,q$). Since
+    simulations are run for a relatively high continuation probability of
+    $\delta\!=\!0.999$, we do not report the players' initial cooperation
+    probability $y$. The graphs show how often the resident population chooses
+    each combination ($p,q$) of conditional cooperation probabilities in the
+    subsequent rounds. ({\bf A}) If players update based on their expected
+    payoffs, the resident population typically applies a strategy for which
+    $p\!\approx\!1$ and $q\!\le\!1\!-\!c/b\!=\!0.9$. ({\bf B}) When players
+    update their strategies based on their realized payoffs in the last round,
+    there are two different predominant behaviors. The resident population
     either consists of defectors (with $p\!\approx\!q\!\approx\!0$) or of
-    conditional cooperators. In the latter case, the maximum level of $q$ consistent
-    with stable cooperation is somewhat smaller compared to the expected-payoff
-    setting, $q\!<\!0.5$. The cooperation rate within the resident population
-    (averaged over all games and over all time steps) is close to 100\%. Parameters:
-    $N\!=\!100$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.999$.}
+    conditional cooperators. In the latter case, the maximum level of $q$
+    consistent with stable cooperation is somewhat smaller compared to the
+    expected-payoff setting, $q\!<\!0.5$. The cooperation rate within the
+    resident population (averaged over all games and over all time steps) is
+    close to 100\%. Parameters: $N\!=\!100$, $c\!=\!1$, $\beta\!=\!1$,
+    $\delta\!=\!0.999$.}
     \label{fig:expected_and_stochastic_for_donation}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=.9\textwidth]{static/cooperation_rates_expected_and_stochastic_for_donation_game.pdf}
 \end{figure}
@@
   the probabilities \(p, q\) for resident population over \(10^7\) time steps
-  for each benefit value. ({\bf B}) The cooperation rate within the resident population
-  (averaged over all games and over all time steps) over the benefit.
-  Unless
-  explicitly varied, the parameters of the simulation are $N\!=\!100$,
-  $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run
-  for $t\!=\!5\times 10^7$ time steps for each parameter
+  for each benefit value. ({\bf B}) The cooperation rate within the resident
+  population (averaged over all games and over all time steps) over the benefit.
+  Unless explicitly varied, the parameters of the simulation are $N\!=\!100$,
+  $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
+  $t\!=\!5\times 10^7$ time steps for each parameter
   combination.}\label{fig:cooperation_rate_over_benefit}
@@
 there is variation in the evolving populations. In the case of expected payoffs
-the resident populations become more cooperative as \(\beta\)
-increases, whereas in the case of limited memory payoffs, the resident
-populations become more defective.
+the resident populations become more cooperative as \(\beta\) increases, whereas
+in the case of limited memory payoffs, the resident populations become more
+defective.
 
@@
 \(\beta\) value. ({\bf B}) The cooperation rate within the resident population
-(averaged over all games and over all time steps) over \(\beta\).
-Unless explicitly varied, the parameters of the simulation are $N\!=\!100$,
-$b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
+(averaged over all games and over all time steps) over \(\beta\). Unless
+explicitly varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$,
+$c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
 $t\!=\!5\times 10^7$ time steps for each parameter
@@
 
-\subsection{Effect of updating payoffs in different social dilemmas}\label{section:2_by_2_games}
+% \subsection{Effect of updating payoffs in different social
+% dilemmas}\label{section:2_by_2_games}
 
-In the previous section we gained insights into the effects of the updating
-payoffs, and into how parameters such as the benefit and the strength of
-selection can intensify them. We investigated these effects by using the
-donation game. In order to broaden our understanding of the updating payoffs on
-different forms of possible human interactions we extend our approach to all
-\(2 \times 2\) symmetric games. More specifically, we apply our analysis to four
-different classes of games; the harmony, the snowdrift, the prisoner's dilemma
-and the stag-hunt games.
+% In the previous section we gained insights into the effects of the updating
+% payoffs, and into how parameters such as the benefit and the strength of
+% selection can intensify them. We investigated these effects by using the
+% donation game. In order to broaden our understanding of the updating payoffs on
+% different forms of possible human interactions we extend our approach to all \(2
+% \times 2\) symmetric games. More specifically, we apply our analysis to four
+% different classes of games; the harmony, the snowdrift, the prisoner's dilemma
+% and the stag-hunt games.
 
-We compare the results of the evolutionary process when the updating payoffs are based on the expected
-payoffs, and on the last round payoffs for all the four possible classes of
-games (Figures~\ref{fig:expected_two_by_two}
-and~\ref{fig:last_round_two_by_two}). In Figure~\ref{fig:expected_two_by_two}
-each sub figure represents a run of evolutionary process for a different set of
-values for \(\mathcal{U}\). Without loss of generality we set \(R=1\) and
-\(P=0\)~\cite{Martinez2012, Roca2009}, and we vary the values of the temptation
-payoff \(T\) (across the \(x\)-axis) and of the sucker's payoff \(S\) (across
-the \(y\)-axis). Starting at the upper left corner and proceeding clockwise the
-quadrants correspond to the harmony, the snowdrift, the prisoner's dilemma and
-the stag-hunt games.
+% We compare the results of the evolutionary process when the updating payoffs are
+% based on the expected payoffs, and on the last round payoffs for all the four
+% possible classes of games (Figures~\ref{fig:expected_two_by_two}
+% and~\ref{fig:last_round_two_by_two}). In Figure~\ref{fig:expected_two_by_two}
+% each sub figure represents a run of evolutionary process for a different set of
+% values for \(\mathcal{U}\). Without loss of generality we set \(R=1\) and
+% \(P=0\)~\cite{Martinez2012, Roca2009}, and we vary the values of the temptation
+% payoff \(T\) (across the \(x\)-axis) and of the sucker's payoff \(S\) (across
+% the \(y\)-axis). Starting at the upper left corner and proceeding clockwise the
+% quadrants correspond to the harmony, the snowdrift, the prisoner's dilemma and
+% the stag-hunt games.
 
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[width=\textwidth]{static/expected_two_by_two_games.pdf}
-  \caption{{\bf Evolutionary dynamics under expected payoffs for various social dilemmas.}
-  We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across
-  the \(x\) axis, and  \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\) across the
-  \(y\) axis. ({\bf A}) The top left quadrant where \(S > 0\) and \(T \leq 1\)
-  corresponds to the harmony game. The preference ordering for the harmony game
-  is \(R > T > S > P\). ({\bf B}) The top right quadrant where \(S > 0\) and \(T > 1\)
-  corresponds to the snowdrift game. The preference ordering for the snowdrift game
-  is \(T > R > S > P\). ({\bf C}) The bottom left quadrant where \(S < 0\) and \(T \leq 1\)
-  corresponds to the stag-hunt game. The preference ordering for the stag-hunt game
-  is \(R > T > P > S\). ({\bf D}) The bottom right quadrant where \(S < 0\) and \(T > 1\)
-  corresponds to the prisoner's dilemma game. The preference ordering for the prisoner's dilemma game
-  is \(T > R > P > S\).
-  Unless explicitly varied, the parameters of the simulation
-  are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}
-  \label{fig:expected_two_by_two}
-\end{figure}
+% \begin{figure}[!htbp]
+%   \centering
+%   \includegraphics[width=\textwidth]{static/expected_two_by_two_games.pdf}
+%   \caption{{\bf Evolutionary dynamics under expected payoffs for various social dilemmas.}
+%   We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across
+%   the \(x\) axis, and  \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\) across the
+%   \(y\) axis. ({\bf A}) The top left quadrant where \(S > 0\) and \(T \leq 1\)
+%   corresponds to the harmony game. The preference ordering for the harmony game
+%   is \(R > T > S > P\). ({\bf B}) The top right quadrant where \(S > 0\) and \(T
+%   > 1\) corresponds to the snowdrift game. The preference ordering for the
+%   snowdrift game is \(T > R > S > P\). ({\bf C}) The bottom left quadrant where
+%   \(S < 0\) and \(T \leq 1\) corresponds to the stag-hunt game. The preference
+%   ordering for the stag-hunt game is \(R > T > P > S\). ({\bf D}) The bottom
+%   right quadrant where \(S < 0\) and \(T > 1\) corresponds to the prisoner's
+%   dilemma game. The preference ordering for the prisoner's dilemma game is \(T >
+%   R > P > S\). Unless explicitly varied, the parameters of the simulation are
+%   $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}
+%   \label{fig:expected_two_by_two}
+% \end{figure}
 
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[width=\textwidth]{static/stochastic_two_by_two_games.pdf}
-  \caption{{\bf Evolutionary dynamics under last round payoffs for various social dilemmas.} 
-  We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across
-  the \(x\) axis, and  \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\) across the
-  \(y\) axis. ({\bf A}) The top left quadrant where \(S > 0\) and \(T \leq 1\)
-  corresponds to the harmony game. The preference ordering for the harmony game
-  is \(R > T > S > P\). ({\bf B}) The top right quadrant where \(S > 0\) and \(T > 1\)
-  corresponds to the snowdrift game. The preference ordering for the snowdrift game
-  is \(T > R > S > P\). ({\bf C}) The bottom left quadrant where \(S < 0\) and \(T \leq 1\)
-  corresponds to the stag-hunt game. The preference ordering for the stag-hunt game
-  is \(R > T > P > S\). ({\bf D}) The bottom right quadrant where \(S < 0\) and \(T > 1\)
-  corresponds to the prisoner's dilemma game. The preference ordering for the prisoner's dilemma game
-  is \(T > R > P > S\).
-  Unless explicitly varied, the parameters of the simulation
-  are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}
-  \label{fig:last_round_two_by_two}
-\end{figure}
+% \begin{figure}[!htbp]
+%   \centering
+%   \includegraphics[width=\textwidth]{static/stochastic_two_by_two_games.pdf}
+%   \caption{{\bf Evolutionary dynamics under last round payoffs for various social dilemmas.} 
+%   We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across
+%   the \(x\) axis, and  \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\) across the
+%   \(y\) axis. ({\bf A}) The top left quadrant where \(S > 0\) and \(T \leq 1\)
+%   corresponds to the harmony game. The preference ordering for the harmony game
+%   is \(R > T > S > P\). ({\bf B}) The top right quadrant where \(S > 0\) and \(T
+%   > 1\) corresponds to the snowdrift game. The preference ordering for the
+%   snowdrift game is \(T > R > S > P\). ({\bf C}) The bottom left quadrant where
+%   \(S < 0\) and \(T \leq 1\) corresponds to the stag-hunt game. The preference
+%   ordering for the stag-hunt game is \(R > T > P > S\). ({\bf D}) The bottom
+%   right quadrant where \(S < 0\) and \(T > 1\) corresponds to the prisoner's
+%   dilemma game. The preference ordering for the prisoner's dilemma game is \(T >
+%   R > P > S\). Unless explicitly varied, the parameters of the simulation are
+%   $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}
+%   \label{fig:last_round_two_by_two}
+% \end{figure}
 
-The harmony game describes an interaction without conflict. In the harmony game
-it is in the best interest for both players to cooperate, and in an evolutionary
-process cooperators prevail. We observe that for most harmony game runs the
-resident population overwhelmingly applies a strategy for which \(p\) and \(q\)
-are \(\approx 1\) (Figure~\ref{fig:expected_two_by_two}). The snowdrift game
-describes a situation similar to that of the prisoner's dilemma; cooperation
-results in a benefit to the opposing player but entailing a cost to the
-cooperator. However, in the snowdrift game individuals obtain immediate direct
-benefits from the cooperative acts which leads to \(S > P\). In the snowdrift
-game more cooperation emerges compared to the prisoner's dilemma. Defection is
-never a resident strategy and for the same values of temptation the overall
-\(q-\) values are higher. The last class of games we present are for the stag-hunt
-game. In the stag-hunt game there is an equilibrium in which both players
-cooperate as well as one in which both defect. For small values of the
-temptation and the sucker's payoffs we observe that cooperation prevails similar
-to the harmony game. However, in the case where the temptation and the reward
-payoffs are equal the resident population either consists of defectors or
-cooperators.
+% The harmony game describes an interaction without conflict. In the harmony game
+% it is in the best interest for both players to cooperate, and in an evolutionary
+% process cooperators prevail. We observe that for most harmony game runs the
+% resident population overwhelmingly applies a strategy for which \(p\) and \(q\)
+% are \(\approx 1\) (Figure~\ref{fig:expected_two_by_two}). The snowdrift game
+% describes a situation similar to that of the prisoner's dilemma; cooperation
+% results in a benefit to the opposing player but entailing a cost to the
+% cooperator. However, in the snowdrift game individuals obtain immediate direct
+% benefits from the cooperative acts which leads to \(S > P\). In the snowdrift
+% game more cooperation emerges compared to the prisoner's dilemma. Defection is
+% never a resident strategy and for the same values of temptation the overall
+% \(q-\) values are higher. The last class of games we present are for the
+% stag-hunt game. In the stag-hunt game there is an equilibrium in which both
+% players cooperate as well as one in which both defect. For small values of the
+% temptation and the sucker's payoffs we observe that cooperation prevails similar
+% to the harmony game. However, in the case where the temptation and the reward
+% payoffs are equal the resident population either consists of defectors or
+% cooperators.
 
-Although the patterns of the evolved populations remain similar if we consider
-that players update their strategies based on their last round payoff
-(Figure~\ref{fig:last_round_two_by_two}), there are still dissimilarities
-between the two approaches. To better understand these dissimilarities, we
-explore the cooperation rates for each of the cases of the social dilemmas we
-have presented (Figure~\ref{fig:cooperation_rates_two_by_two_part_one}).
+% Although the patterns of the evolved populations remain similar if we consider
+% that players update their strategies based on their last round payoff
+% (Figure~\ref{fig:last_round_two_by_two}), there are still dissimilarities
+% between the two approaches. To better understand these dissimilarities, we
+% explore the cooperation rates for each of the cases of the social dilemmas we
+% have presented (Figure~\ref{fig:cooperation_rates_two_by_two_part_one}).
 
-In the harmony game the cooperation rates for both approaches are similar. The
-biggest difference occurs for \(T=1\), nevertheless, it is less than 10\%. There
-are two instances for which the cooperation rates are higher for the limited
-memory payoffs compared to the expected. These are for \(T=2.2, S=0.4\) and for
-\(T=2.2, S=0.8\). It is only in the snowdrift game that cooperation is
-overestimated when a minimum of information is considered. Players are slightly
-more likely to cooperate after a cooperation and to forgive after being at the
-receiving end of a defection. For the rest of the cases the classical scenario
-overestimates the average cooperation rate, supporting the results of the
-previous section. There is a substantial difference in the cooperation rates
-for the prisoner's dilemma class. We observe that for \(S<-0.4\) cooperation
-almost never emerges if players imitate based only on the last round payoff.
+% In the harmony game the cooperation rates for both approaches are similar. The
+% biggest difference occurs for \(T=1\), nevertheless, it is less than 10\%. There
+% are two instances for which the cooperation rates are higher for the limited
+% memory payoffs compared to the expected. These are for \(T=2.2, S=0.4\) and for
+% \(T=2.2, S=0.8\). It is only in the snowdrift game that cooperation is
+% overestimated when a minimum of information is considered. Players are slightly
+% more likely to cooperate after a cooperation and to forgive after being at the
+% receiving end of a defection. For the rest of the cases the classical scenario
+% overestimates the average cooperation rate, supporting the results of the
+% previous section. There is a substantial difference in the cooperation rates for
+% the prisoner's dilemma class. We observe that for \(S<-0.4\) cooperation almost
+% never emerges if players imitate based only on the last round payoff.
 
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[width=.4\textwidth]{static/expected_two_by_two_games_cooperation.pdf}
-  \includegraphics[width=.4\textwidth]{static/stochastic_two_by_two_games_cooperation.pdf}
-  \caption{\textbf{Cooperation rates for various social dilemmas.} ({\bf A}) If
-  players update based on their expected payoffs. ({\bf B}) If
-  players update based on their last round payoffs. We vary the temptation
-  payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\)
-  across the \(x\) axis, and \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\)
-  across the \(y\) axis. Unless explicitly varied, the parameters of the simulation
-  are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}\label{fig:cooperation_rates_two_by_two_part_one}
-\end{figure}
+% \begin{figure}[!htbp]
+%   \centering
+%   \includegraphics[width=.4\textwidth]{static/expected_two_by_two_games_cooperation.pdf}
+%   \includegraphics[width=.4\textwidth]{static/stochastic_two_by_two_games_cooperation.pdf}
+%   \caption{\textbf{Cooperation rates for various social dilemmas.} ({\bf A}) If
+%   players update based on their expected payoffs. ({\bf B}) If players update
+%   based on their last round payoffs. We vary the temptation payoff \(T \in
+%   \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across the \(x\) axis, and \(S \in \{1.2,
+%   0.8, 0.4, 0, -0.4, -0.8\}\) across the \(y\) axis. Unless explicitly varied,
+%   the parameters of the simulation are $N\!=\!100$, $\beta\!=\!10$,
+%   $\delta\!=\!0.99$, $t\!=\!5\times
+%   10^7$.}\label{fig:cooperation_rates_two_by_two_part_one}
+% \end{figure}
 
-So far we have explored the difference between the expected payoffs and the last
-round payoffs. In order to explore further the effect of limited memory we allow
-individuals to remember more. More precisely, up to two interactions and up to
-the last two rounds. In total we present results for three more updating
-payoffs. These are the payoffs when individuals consider the last two rounds
-with another member of the population, the last round with two members of the
-population, and the last two rounds with two members of the population.
-Similar to the last round payoff, we use simulations and record which strategies
-the players adopt over time based
-(Appendix~\ref{appendix:further_simulation_results}) and we compare the
-evolving cooperation rates to those of the expected payoffs
-(Figure~\ref{fig:cooperation_rates_two_by_two_part_two}).
+% So far we have explored the difference between the expected payoffs and the last
+% round payoffs. In order to explore further the effect of limited memory we allow
+% individuals to remember more. More precisely, up to two interactions and up to
+% the last two rounds. In total we present results for three more updating
+% payoffs. These are the payoffs when individuals consider the last two rounds
+% with another member of the population, the last round with two members of the
+% population, and the last two rounds with two members of the population. Similar
+% to the last round payoff, we use simulations and record which strategies the
+% players adopt over time based
+% (Appendix~\ref{appendix:further_simulation_results}) and we compare the evolving
+% cooperation rates to those of the expected payoffs
+% (Figure~\ref{fig:cooperation_rates_two_by_two_part_two}).
 
-In the case of the last two rounds payoff we make two observations. Initially,
-for the harmony, the stag-hunt and the prisoner's dilemma classes the average
-cooperation rates are higher than the previous case of the last round payoff.
-The rates remain strictly less than in the expected payoffs, however, in the
-case of the prisoner's dilemma we observe a significant increase. Secondly, for
-the snowdrift class there are more instances for which the cooperation rates are
-higher than in the expected payoffs. This could indicate that in the case of the
-snowdrift game the expected payoffs underestimate cooperation.
-Once players interact with two other members of the population, instead of one,
-the expected payoffs strictly yield a higher cooperation rate  even in the case
-of the snowdrift game. Similar to the previous result, the difference between
-the two approaches is always higher for the prisoner's dilemma class.
+% In the case of the last two rounds payoff we make two observations. Initially,
+% for the harmony, the stag-hunt and the prisoner's dilemma classes the average
+% cooperation rates are higher than the previous case of the last round payoff.
+% The rates remain strictly less than in the expected payoffs, however, in the
+% case of the prisoner's dilemma we observe a significant increase. Secondly, for
+% the snowdrift class there are more instances for which the cooperation rates are
+% higher than in the expected payoffs. This could indicate that in the case of the
+% snowdrift game the expected payoffs underestimate cooperation. Once players
+% interact with two other members of the population, instead of one, the expected
+% payoffs strictly yield a higher cooperation rate  even in the case of the
+% snowdrift game. Similar to the previous result, the difference between the two
+% approaches is always higher for the prisoner's dilemma class.
 
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[width=.33\textwidth]{static/rounds_two_by_two_games_cooperation.pdf}
-  \includegraphics[width=.33\textwidth]{static/opponents_two_by_two_games_cooperation.pdf}
-  \includegraphics[width=.33\textwidth]{static/rounds_opponents_two_by_two_games_cooperation.pdf}
-  \caption{\textbf{Cooperation rates over various social dilemmas for different
-  limited memory payoffs.} ({\bf A}) If
-  players update based on their last two rounds with another member of the population.
-  ({\bf B}) If players update based on their last round with two other members of the population.
-  ({\bf C}) If players update based on their last two rounds with two other members of the population.
-  We vary the temptation
-  payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\)
-  across the \(x\) axis, and \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\)
-  across the \(y\) axis. Unless explicitly varied, the parameters of the simulation
-  are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$..}
-  \label{fig:cooperation_rates_two_by_two_part_two}
-\end{figure}
+% \begin{figure}[!htbp]
+%   \centering
+%   \includegraphics[width=.33\textwidth]{static/rounds_two_by_two_games_cooperation.pdf}
+%   \includegraphics[width=.33\textwidth]{static/opponents_two_by_two_games_cooperation.pdf}
+%   \includegraphics[width=.33\textwidth]{static/rounds_opponents_two_by_two_games_cooperation.pdf}
+%   \caption{\textbf{Cooperation rates over various social dilemmas for different
+%   limited memory payoffs.} ({\bf A}) If players update based on their last two
+%   rounds with another member of the population. ({\bf B}) If players update
+%   based on their last round with two other members of the population. ({\bf C})
+%   If players update based on their last two rounds with two other members of the
+%   population. We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8,
+%   2.2\}\) across the \(x\) axis, and \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\)
+%   across the \(y\) axis. Unless explicitly varied, the parameters of the
+%   simulation are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times
+%   10^7$..}
+%   \label{fig:cooperation_rates_two_by_two_part_two}
+% \end{figure}
 
@@
 cooperates but the other defects, the defector receives the temptation to
-defect, \(T\), whereas the cooperator receives the sucker's payoff, \(S\).
-Let $\mathcal{U}=\{R,S,T,P\}$ denote the set of feasible payoffs in each round,
-and let $\mathbf{u}\!=\!(R,S,T,P)$ be the corresponding payoff vector.
-We present results for various values of $\mathcal{U}$ for all
-the symmetric \(2 \times 2\) games.
+defect, \(T\), whereas the cooperator receives the sucker's payoff, \(S\). Let
+$\mathcal{U}=\{R,S,T,P\}$ denote the set of feasible payoffs in each round, and
+let $\mathbf{u}\!=\!(R,S,T,P)$ be the corresponding payoff vector. We present
+results for various values of $\mathcal{U}$ for all the symmetric \(2 \times 2\)
+games.
 
@@
 strategies when they make decisions in each round. Reactive strategies are a set
-of strategies that take into account only the previous action of the opponent.
-A reactive strategy can be written explicitly as a vector,
+of strategies that take into account only the previous action of the opponent. A
+reactive strategy can be written explicitly as a vector,
 
@@
 In the updating stage, two players are randomly drawn from the population, a
-`learner' and a `role model'. The learner adopts the role model's strategy
-based on the Fermi distribution function, %ToDo add reference
+`learner' and a `role model'. The learner adopts the role model's strategy based
+on the Fermi distribution function, %ToDo add reference
 
@@
 \begin{algorithm}[!htbp]
-  \SetAlgoLined
-   $N \leftarrow$ population size\;
-   $k \leftarrow 1$\;
-   resident $\leftarrow (0, 0, 0)$\;
-   \While{ $t <$ maximum number of steps}{
-    mutant $\leftarrow$ random: $\{\emptyset \}\ \rightarrow R^3$\;
-    fixation probability $\leftarrow \varphi $\;
-    \If{$\varphi >$ random: $i \rightarrow [0, 1]$}{
-      resident $\leftarrow$ mutant;
-     }}
-     \caption{Evolutionary process}\label{algorithm:pairwise_comparison}
+  \SetAlgoLined $N \leftarrow$ population size\; $k \leftarrow 1$\; resident
+   $\leftarrow (0, 0, 0)$\; \While{ $t <$ maximum number of steps}{mutant
+   $\leftarrow$ random: $\{\emptyset \}\ \rightarrow R^3$\; fixation probability
+   $\leftarrow \varphi $\; \If{$\varphi >$ random: $i \rightarrow [0,
+   1]$}{resident $\leftarrow$ mutant;}} \caption{Evolutionary
+   process}\label{algorithm:pairwise_comparison}
 \end{algorithm}
@@
 
-Initially, we discuss the case of the \textbf{last round updating payoff}. At the stage
-game we define the payoff of a reactive strategy in the last round,
+Initially, we discuss the case of the \textbf{last round updating payoff}. At
+the stage game we define the payoff of a reactive strategy in the last round,
 Proposition~\ref{proposition:last_round}.
 
-\begin{Prop}\label{proposition:last_round}
-    Consider a repeated game, with
+\begin{Prop}\label{proposition:last_round} Consider a repeated game, with
     continuation probability $\delta$, between players with reactive strategies
-    $s_1\!=\!(y_1, p_1, q_1$)  and $s_2\!=\!(y_2,p_2,q_2)$ respectively. Then the
-    probability that the $s_1$ player receives the payoff $u\!\in\! \mathcal{U}$ in
-    the very last round of the game is given by $v_{u}(s_1,s_2)$, as given by
-    Equation~(\ref{Eq:LastRound}).
+    $s_1\!=\!(y_1, p_1, q_1$)  and $s_2\!=\!(y_2,p_2,q_2)$ respectively. Then
+    the probability that the $s_1$ player receives the payoff $u\!\in\!
+    \mathcal{U}$ in the very last round of the game is given by
+    $v_{u}(s_1,s_2)$, as given by Equation~(\ref{Eq:LastRound}).
 
@@
 where $\mathbf{v_0}$ denotes the expected distribution of the four outcomes in
-the very first round, and \(1- \delta\) the probability that the game ends.
-It can be shown that,
+the very first round, and \(1- \delta\) the probability that the game ends. It
+can be shown that,
 
@@
   other with a probability \(\frac{1}{N - 1}\), and thus they do not interact
-  with other with a probability \(1 - \frac{1}{N - 1}\). In
-  the latter case, each of them can interact with either a mutant or a resident.
-  Both of them interact with a mutant with a probability
-  $\frac{(k-1)(k-2)}{(N-2)(N-3)}$ and both interact with a resident with a
-  probability $\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$. The last two possible pairings
-  are that either of them interacts with a resident whilst the other interacts
-  with a mutant, and this happens with a probability $\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$.}
+  with other with a probability \(1 - \frac{1}{N - 1}\). In the latter case,
+  each of them can interact with either a mutant or a resident. Both of them
+  interact with a mutant with a probability $\frac{(k-1)(k-2)}{(N-2)(N-3)}$ and
+  both interact with a resident with a probability
+  $\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$. The last two possible pairings are that
+  either of them interacts with a resident whilst the other interacts with a
+  mutant, and this happens with a probability
+  $\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$.}
   \label{fig:single_pairs}
@@
 
-Given the last round payoff and possible pair combinations for a single interaction,
-we define the probability that the respective last round payoffs of two players
-\(s_1, s_2\) are given by $u_1$ and $u_2$ as,
+Given the last round payoff and possible pair combinations for a single
+interaction, we define the probability that the respective last round payoffs of
+two players \(s_1, s_2\) are given by $u_1$ and $u_2$ as,
 
@@
 the role model happened to be matched during the game stage, which happens with
-probability $\frac{1}{(N\!-\!1)}$. In that case, we note that only those payoff pairs
-can occur that are feasible in a direct interaction, $\splitatcommas{(u_1,u_2)\in
-\mathcal{U}^2_F:=\big\{ (R,R), (S,T), (T,S), (P,P) \big\}}$, as represented by
-the respective indicator function. Otherwise, if the learner and the role model
-did not interact directly, we need to distinguish four different cases,
-depending on whether the learner was matched with a resident or a mutant, and
-depending on whether the role model was matched with a resident or a mutant.
+probability $\frac{1}{(N\!-\!1)}$. In that case, we note that only those payoff
+pairs can occur that are feasible in a direct interaction,
+$\splitatcommas{(u_1,u_2)\in \mathcal{U}^2_F:=\big\{ (R,R), (S,T), (T,S), (P,P)
+\big\}}$, as represented by the respective indicator function. Otherwise, if the
+learner and the role model did not interact directly, we need to distinguish
+four different cases, depending on whether the learner was matched with a
+resident or a mutant, and depending on whether the role model was matched with a
+resident or a mutant.
 
-Given that $N\!-\!k$ players use the resident strategy $s_{R}\!=\!(y_{R},p_{R},q_{R})$
-and that the remaining $k$ players use the mutant strategy
-$s_{M}\!=\!(y_{M},p_{M},q_{M})$, the probability that the number of mutants increases by
-one in one step of the evolutionary process can be written as
+Given that $N\!-\!k$ players use the resident strategy
+$s_{R}\!=\!(y_{R},p_{R},q_{R})$ and that the remaining $k$ players use the
+mutant strategy $s_{M}\!=\!(y_{M},p_{M},q_{M})$, the probability that the number
+of mutants increases by one in one step of the evolutionary process can be
+written as
 
@@
 
-In this expression, $\frac{(N\!-\!k)}{N}$ is the probability that the randomly chosen
-learner is a resident, and $\frac{k}{N}$ is the probability that the role model is a
-mutant. The sum corresponds to the total probability that the learner adopts the
-role model's strategy over all possible payoffs $u_R$ and $u_M$ that the two
-player may have received in their respective last rounds. We use $x(u_R,u_M)$ to
-denote the probability that the randomly chosen resident obtained a payoff of
-$u_R$ in the last round of his respective game, and that the mutant obtained a
-payoff of $u_M$.
+In this expression, $\frac{(N\!-\!k)}{N}$ is the probability that the randomly
+chosen learner is a resident, and $\frac{k}{N}$ is the probability that the role
+model is a mutant. The sum corresponds to the total probability that the learner
+adopts the role model's strategy over all possible payoffs $u_R$ and $u_M$ that
+the two player may have received in their respective last rounds. We use
+$x(u_R,u_M)$ to denote the probability that the randomly chosen resident
+obtained a payoff of $u_R$ in the last round of his respective game, and that
+the mutant obtained a payoff of $u_M$.
 
@@
 
-\begin{Prop}\label{proposition:last_two_rounds}
-  Consider a repeated game, with continuation probability $\delta$, between
-  players with reactive strategies $s_1\!=\!(y_1, p_1, q_1$)  and
-  $s_2\!=\!(y_2,p_2,q_2)$ respectively. Let $\mathcal{\tilde{U}} =
-  \splitatcommas{\{RR, RS, RT, RP, SR, SS, ST, SP, TR, TS, TT, TP, PR, PS, PT,
-  PP\}}$ denote the set of feasible payoffs in the last two rounds, and let
-  \(\tilde{\mathbf{u}}\) be the corresponding payoff vector. Then the
-  probability that the $s_1$ player receives the payoff $u\!\in\!
+\begin{Prop}\label{proposition:last_two_rounds} Consider a repeated game, with
+  continuation probability $\delta$, between players with reactive strategies
+  $s_1\!=\!(y_1, p_1, q_1$)  and $s_2\!=\!(y_2,p_2,q_2)$ respectively. Let
+  $\mathcal{\tilde{U}} = \splitatcommas{\{RR, RS, RT, RP, SR, SS, ST, SP, TR,
+  TS, TT, TP, PR, PS, PT, PP\}}$ denote the set of feasible payoffs in the last
+  two rounds, and let \(\tilde{\mathbf{u}}\) be the corresponding payoff vector.
+  Then the probability that the $s_1$ player receives the payoff $u\!\in\!
   \mathcal{\tilde{U}}$ in the very last two rounds of the game is given by,
@@
 In the updating stage we select a mutant and resident to be either the role
-model or the learner. Given that they can interact with two other members of
-the population there are a total of twenty four possible combinations of pairs
+model or the learner. Given that they can interact with two other members of the
+population there are a total of twenty four possible combinations of pairs
 (Figure~\ref{fig:pissible_two_pairs}).
@@
 
-\section{Evolutionary dynamics under limited memory payoffs for various social dilemmas}\label{appendix:further_simulation_results}
+\section{Evolutionary dynamics under limited memory payoffs for various social
+dilemmas}\label{appendix:further_simulation_results}
 

### 2021-11-24 main.tex
testing something regarding line length
making a commit
working on the introduction

@@
 \begin{abstract}
-% Repeated games are vastly used in evolutionary game theory to explain
-% cooperation in a variety of environments. Although existing models have shaped
-% our understanding of human cooperation, they often work with idealized
+% Repeated games are vastly use in evolutionary game theory to explain cooperation in a variety of environments. Although existing models have shaped
+% our understanding of human cooperation, hey often work with idealized
 % assumptions. In an evolutionary process, individuals imitate other individuals
@@
 nowak2004evolutionary, hauert2005game} describes the evolutionary dynamics of
-populations consisting of different types of interacting individuals.
-Traditional approaches assume that individuals meet each other at random in an
-infinitely large well-mixed population. An example of such a approach in the the
-replicator dynamics. It describes how the abundance of strategic types in a
-population changes based on their fitness, identified with the payoff resulting
-from the game~\cite{cressman2014replicator}. In this deterministic formulation,
-individuals with higher fitness, such as defectors, increase in abundance.
-Ultimately, the system reaches a stable fixed point in which the population may
-consist either of a single type or of a mixture of different types.
+populations consisting of different types of interacting individuals. The
+framework of evolutionary game theory has been applied in %ToDo cite.
+Traditional approaches of evolutionary game theory assume that individuals meet
+each other at random in an infinitely large well-mixed population, such an
+approach is the replicator dynamics. The replicator dynamics describes how the
+abundance of strategic types in a population changes based on their fitness. In
+this deterministic formulation, individuals with higher fitness increase in
+abundance and ultimately, the system reaches a stable fixed point in which the
+population may consist either of a single type or of a mixture of different
+types. The works of cite have shown that by constrain the population to a
+finiteness of populations may lead to fundamental changes in this picture due to
+stochastic effects. These stochastic effects disadvantageous mutants have a
+small, yet non-zero probability to reach fixation in a finite population.
 
-It has been shown that the finiteness of populations may lead to fundamental
-changes in this picture due to stochastic effects In this framework
-disadvantageous mutants have a small, yet non-zero probability to reach fixation
-in a finite population.
+Two classes of such finite stochastic processes have been used extensively: (i)
+fitness-based processes in which an individual chosen proportional to fitness
+reproduces and the offspring replaces a randomly chosen individual [11]
+\textit{moran process}; (ii) \textit{pairwise comparison processes} in which a
+pair of individuals is chosen, and where subsequently one of these individuals
+may adopt the strategy of the other.
 
-Its most prominent application is the quest for the origins and 
-evolution of cooperation.
-An important application of evolutionary game theory is the evolution of
-cooperation which remains one of the greatest problems for biological and social
-sciences.
+In pairwise comparison process considers a finite population of fixed size. In
+the population at a time step different types of individuals can exist. In the
+simplest of cases each individual can be of one of two types, A and B. The state
+of the population is thus characterized by the number \(k\) of individuals of
+type A. The interaction between the two types of individuals is described by the
+functions \(\pi _{{\rm A}}^{i}\) and $\pi _{{\rm B}}^{i}$.
 
-% To explain the evolution of cooperation remains one of the greatest problems for
-% biological and social sciences. Cooperation is the action of choosing to help
-% others at one's own expense. The standard game of formulating such situation is
-% the Prisoner's Dilemma~\cite{trivers1971evolution,
-% milinski1987tit, glynatsi2021bibliometric, ohtsuki2006simple}, in which two
-% players can choose to cooperate or to defect. The players are offered a certain
-% payoff, \(R\), for mutual cooperation and a lower payoff, \(P\), for mutual
-% defection. If one player cooperates while the other defects, then the cooperator
-% gets the lowest payoff, \(S\), while the defector gains the highest payoff,
-% \(T\). Thus, the payoffs have the following property \(T > R > P > S\), making
-% defection the dominant strategy in the non-repeated game.
+At each step an individual change their type based.
 
-% In the repeated Prisoner's Dilemma, direct reciprocity can allow cooperators to
-% not be exploited by defectors~\cite{axelrod1981evolution, nowak1992tit,
-% nowak1992evolutionary}, however, how do such cooperators can evolve?
+\begin{equation} \label{Eq:rho}
+  \rho(\pi _{{\rm A}}^{i}, \pi _{{\rm B}}^{i}) = \frac{1}{1\!+\! \exp^{\!-\!\beta (\pi _{{\rm A}}^{i}\!-\!\pi _{{\rm B}}^{i})}}.
+\end{equation}
 
- , and 
-in general, defectors are stable against invasion by cooperators.
+The parameter denotes the intensity of selection. For , fitness equals payoff.
+This scenario describes ``strong selection''. For \(\beta \ll 1\) , the
+payoff only provides a small perturbation to the overall fitness of an
+individual, a limit known as weak selection (Nowak et al., 2004).
 
-In this framework disadvantageous mutants have a small, yet non-zero probability
-to reach fixation in a finite population.
+These indicate the expected payoff for two types in a population in state i. The
+interaction can be generated from a two-player matrix game which leads to
+payoffs linear in i [26], but we keep the formalism general to include games
+played between an arbitrary number of players, which leads to payoffs that
+depend on i in a polynomial way [27–30]. In fact, our results hold for an
+arbitrary dependence of the payoffs on i.
 
-Two classes of microscopic processes have been used extensively: (i)
-fitness-based processes in which an individual chosen proportional to fitness
-reproduces and the offspring replaces a randomly chosen individual [11]; (ii)
-pairwise comparison processes in which a pair of individuals is chosen, and
-where subsequently one of these individuals may adopt the strategy of the other.
+In the simplest case, the payoffs of A and B individuals only depend on the
+fraction of both types in the population. If there are i A individuals and B
+individuals, then the A and B individuals have payoffs and , respectively.
+Self-interactions are excluded.
 
-In the pairwise process considers a finite population of fixed size.In the
-population at a time step different types of individuals can exist. 
-Each individual can be of one of two types, A and B. The state of the population
-is thus characterized by the number i of individuals of type A.
+Here, we consider a process based on pairwise comparison between individuals.
+Two individuals, A and B, are selected at random. The individual chosen for
+reproduction A replaces B with probability p, which depends on the payoff
+difference between the two individuals. The composition of the population can
+only change if both individuals are of different types. We follow (Blume, 1993,
+Szabó and Tőke, 1998, Hauert and Szabó, 2005) in choosing the Fermi function
+from statistical physics for p
 
-The interaction between the two types of individuals is described by the
-functions $\pi _{{\rm A}}^{i}$ and $\pi _{{\rm B}}^{i}$. These indicate the
-expected payoff for two types in a population in state i. The interaction can be
-generated from a two-player matrix game which leads to payoffs linear in i [26],
-but we keep the formalism general to include games played between an arbitrary
-number of players, which leads to payoffs that depend on i in a polynomial way
-[27–30]. In fact, our results hold for an arbitrary dependence of the payoffs on
-i.
+To explain the evolution of cooperation remains one of the greatest problems for
+biological and social sciences. Cooperation is the action of choosing to help
+others at one's own expense. The standard game of formulating such situation is
+the Prisoner's Dilemma~\cite{trivers1971evolution,
+milinski1987tit, glynatsi2021bibliometric, ohtsuki2006simple}, in which two
+players can choose to cooperate or to defect. The players are offered a certain
+payoff, \(R\), for mutual cooperation and a lower payoff, \(P\), for mutual
+defection. If one player cooperates while the other defects, then the cooperator
+gets the lowest payoff, \(S\), while the defector gains the highest payoff,
+\(T\). Thus, the payoffs have the following property \(T > R > P > S\), making
+defection the dominant strategy in the non-repeated game.
 
-At each step an individual change their type based. This very cooperators
-can repeas.
+In the context of the prisoner's dilemma the payoffs are calculated as follows.
+In the example that a strategy is playing always cooperate and always defect,
 

### 2021-11-24 main.tex
commit

--- 2021-11-24 main.tex
@@
 consist either of a single type or of a mixture of different types.
+
+Mo
 
@@
 
-\begin{figure}[!htbp]
-  \centering
-  \begin{subfigure}{.5\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{static/expected_for_beta.pdf}
-    \includegraphics[width=\textwidth]{static/stochastic_for_beta.pdf}
-  \end{subfigure}%
-  \begin{subfigure}{.5\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{static/cooperation_rate_over_b.pdf}
-  \end{subfigure}
-  \caption{{\bf The evolution of cooperation for different benefit values.} 
-  We vary the benefit of defection $b$. In all cases, expected payoffs appear to
-  overestimate the average cooperation rate the population achieves. ({\bf A})
-  the probabilities \(p, q\) for resident population over \(10^7\) time steps
-  for each benefit value. ({\bf B}) The cooperation rate within the resident
-  population (averaged over all games and over all time steps) over the benefit.
-  Unless explicitly varied, the parameters of the simulation are $N\!=\!100$,
-  $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
-  $t\!=\!5\times 10^7$ time steps for each parameter
-  combination.}\label{fig:cooperation_rate_over_benefit}
-\end{figure}
+% \begin{figure}[!htbp]
+%   \centering
+%   \begin{subfigure}{.5\textwidth}
+%     \centering
+%     \includegraphics[width=\textwidth]{static/expected_for_beta.pdf}
+%     \includegraphics[width=\textwidth]{static/stochastic_for_beta.pdf}
+%   \end{subfigure}%
+%   \begin{subfigure}{.5\textwidth}
+%     \centering
+%     \includegraphics[width=\textwidth]{static/cooperation_rate_over_b.pdf}
+%   \end{subfigure}
+%   \caption{{\bf The evolution of cooperation for different benefit values.} 
+%   We vary the benefit of defection $b$. In all cases, expected payoffs appear to
+%   overestimate the average cooperation rate the population achieves. ({\bf A})
+%   the probabilities \(p, q\) for resident population over \(10^7\) time steps
+%   for each benefit value. ({\bf B}) The cooperation rate within the resident
+%   population (averaged over all games and over all time steps) over the benefit.
+%   Unless explicitly varied, the parameters of the simulation are $N\!=\!100$,
+%   $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
+%   $t\!=\!5\times 10^7$ time steps for each parameter
+%   combination.}\label{fig:cooperation_rate_over_benefit}
+% \end{figure}
 
@@
 
-\section{Evolutionary dynamics under limited memory payoffs for various social
-dilemmas}\label{appendix:further_simulation_results}
+% \section{Evolutionary dynamics under limited memory payoffs for various social
+% dilemmas}\label{appendix:further_simulation_results}
 
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[width=\textwidth]{static/rounds_two_by_two_games.pdf}
-  \caption{{\bf Evolutionary dynamics under last two rounds payoffs for various social dilemmas.} 
-  We have run several simulations of the evolutionary process described in
-  section~\ref{section:model} for $t\!=\!10^7$ time steps. The graphs show how
-  often the resident population chooses each combination ($p,q$) of conditional
-  cooperation probabilities in the subsequent rounds. We vary the temptation
-  payoff \(T \in \{-1, -0.6, -0.2,  0.2, 0.6, 1, 1.4, 1.8, 2.2, 2.6, 3\}\)
-  across the \(x\) axis, and  \(S \in \{2, 1.6, 1.2, 0.8, 0.4, 0, -0.4, -0.8,
-  -1.2, -1.6, -2\}\) across the \(y\) axis. Parameters: $N\!=\!100$,
-  $\beta\!=\!10$, $\delta\!=\!0.999$.}
-  \label{fig:last_two_rounds_two_by_two}
-\end{figure}
+% \begin{figure}[!htbp]
+%   \centering
+%   \includegraphics[width=\textwidth]{static/rounds_two_by_two_games.pdf}
+%   \caption{{\bf Evolutionary dynamics under last two rounds payoffs for various social dilemmas.} 
+%   We have run several simulations of the evolutionary process described in
+%   section~\ref{section:model} for $t\!=\!10^7$ time steps. The graphs show how
+%   often the resident population chooses each combination ($p,q$) of conditional
+%   cooperation probabilities in the subsequent rounds. We vary the temptation
+%   payoff \(T \in \{-1, -0.6, -0.2,  0.2, 0.6, 1, 1.4, 1.8, 2.2, 2.6, 3\}\)
+%   across the \(x\) axis, and  \(S \in \{2, 1.6, 1.2, 0.8, 0.4, 0, -0.4, -0.8,
+%   -1.2, -1.6, -2\}\) across the \(y\) axis. Parameters: $N\!=\!100$,
+%   $\beta\!=\!10$, $\delta\!=\!0.999$.}
+%   \label{fig:last_two_rounds_two_by_two}
+% \end{figure}
 
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[width=\textwidth]{static/opponents_two_by_two_games.pdf}
-  \caption{{\bf Evolutionary dynamics under last round payoffs with two members of the population for various social dilemmas.} 
-  We have run several simulations of the evolutionary process described in
-  section~\ref{section:model} for $t\!=\!10^7$ time steps. The graphs show how
-  often the resident population chooses each combination ($p,q$) of conditional
-  cooperation probabilities in the subsequent rounds. We vary the temptation
-  payoff \(T \in \{-1, -0.6, -0.2,  0.2, 0.6, 1, 1.4, 1.8, 2.2, 2.6, 3\}\)
-  across the \(x\) axis, and  \(S \in \{2, 1.6, 1.2, 0.8, 0.4, 0, -0.4, -0.8,
-  -1.2, -1.6, -2\}\) across the \(y\) axis. Parameters: $N\!=\!100$,
-  $\beta\!=\!10$, $\delta\!=\!0.999$.}
-  \label{fig:last_two_opponents_two_by_two}
-\end{figure}
+% \begin{figure}[!htbp]
+%   \centering
+%   \includegraphics[width=\textwidth]{static/opponents_two_by_two_games.pdf}
+%   \caption{{\bf Evolutionary dynamics under last round payoffs with two members of the population for various social dilemmas.} 
+%   We have run several simulations of the evolutionary process described in
+%   section~\ref{section:model} for $t\!=\!10^7$ time steps. The graphs show how
+%   often the resident population chooses each combination ($p,q$) of conditional
+%   cooperation probabilities in the subsequent rounds. We vary the temptation
+%   payoff \(T \in \{-1, -0.6, -0.2,  0.2, 0.6, 1, 1.4, 1.8, 2.2, 2.6, 3\}\)
+%   across the \(x\) axis, and  \(S \in \{2, 1.6, 1.2, 0.8, 0.4, 0, -0.4, -0.8,
+%   -1.2, -1.6, -2\}\) across the \(y\) axis. Parameters: $N\!=\!100$,
+%   $\beta\!=\!10$, $\delta\!=\!0.999$.}
+%   \label{fig:last_two_opponents_two_by_two}
+% \end{figure}
 
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[width=\textwidth]{static/merged_plot_rounds_opponents_two.pdf}
-  \caption{{\bf Evolutionary dynamics under last two rounds payoffs with two members of the population for various social dilemmas.} 
-  We have run several simulations of the evolutionary process described in
-  section~\ref{section:model} for $t\!=\!10^7$ time steps. The graphs show how
-  often the resident population chooses each combination ($p,q$) of conditional
-  cooperation probabilities in the subsequent rounds. We vary the temptation
-  payoff \(T \in \{-1, -0.6, -0.2,  0.2, 0.6, 1, 1.4, 1.8, 2.2, 2.6, 3\}\)
-  across the \(x\) axis, and  \(S \in \{2, 1.6, 1.2, 0.8, 0.4, 0, -0.4, -0.8,
-  -1.2, -1.6, -2\}\) across the \(y\) axis. Parameters: $N\!=\!100$,
-  $\beta\!=\!10$, $\delta\!=\!0.999$.}
-  \label{fig:last_two_rounds_two_opponents}
-\end{figure}
+% \begin{figure}[!htbp]
+%   \centering
+%   \includegraphics[width=\textwidth]{static/merged_plot_rounds_opponents_two.pdf}
+%   \caption{{\bf Evolutionary dynamics under last two rounds payoffs with two members of the population for various social dilemmas.} 
+%   We have run several simulations of the evolutionary process described in
+%   section~\ref{section:model} for $t\!=\!10^7$ time steps. The graphs show how
+%   often the resident population chooses each combination ($p,q$) of conditional
+%   cooperation probabilities in the subsequent rounds. We vary the temptation
+%   payoff \(T \in \{-1, -0.6, -0.2,  0.2, 0.6, 1, 1.4, 1.8, 2.2, 2.6, 3\}\)
+%   across the \(x\) axis, and  \(S \in \{2, 1.6, 1.2, 0.8, 0.4, 0, -0.4, -0.8,
+%   -1.2, -1.6, -2\}\) across the \(y\) axis. Parameters: $N\!=\!100$,
+%   $\beta\!=\!10$, $\delta\!=\!0.999$.}
+%   \label{fig:last_two_rounds_two_opponents}
+% \end{figure}
 

### 2021-11-24 matlab/contimutationRun.m

@@ -1,0 +1,16 @@
+function evolRun();
+
+starting_resident = [0, 0, 0];
+N = 100;
+delta = 0.999;
+beta = 1;
+numberIterations = 10 ^ 7;
+payoff_type = "expected";
+payoffs = readmatrix("two_by_two_games_values.csv");
+
+parfor (i = 1:11)
+    u = payoffs(i, :);
+    filename = "../data/two_by_two_games_values_S" + u(2) + "_T_" + u(2);
+    evolSimulation(starting_resident, u, N, delta, beta, numberIterations, payoff_type, filename);
+end
+end

### 2021-11-24 matlab/evolMutationContinue.m

@@ -12,7 +12,7 @@ for t = progress(1:numberIterations)
         population(idx, :) = rand(1, sdim);
     else
         indices = randi(N, [1, 2]);
-        
+
         if population(indices(1),:) ~= population(indices(2),:)
             if payoff_type == "expected"
                 payoffs = payoffsExpected(indices, population, u, N, delta);
@@ -33,7 +33,7 @@ for t = progress(1:numberIterations)
         end
     end
     
-  [Mu,~,ic] = unique(population, 'rows', 'stable'); 
+  [Mu,~,ic] = unique(population, 'rows', 'stable');
   h = accumarray(ic,1);
   sz = size(h, 1);
   avg_player = avg_player + sum(Mu .* h / sum(h), 1);
@@ -42,7 +42,7 @@ for t = progress(1:numberIterations)
 %   if mod(t, 100000) == 0 
 %    xDat(j, :) = [t, avg_player, coop];
 %    j = j + 1;
-%   end    
+%   end
 end
 % dlmwrite(filename + ".csv", xDat, 'precision', 9);
 

### 2021-12-20 main.tex
push everything x-mas

@@
 \usepackage{booktabs}
+\usepackage{blkarray}
 \usepackage[ruled,vlined]{algorithm2e}
@@
 populations consisting of different types of interacting individuals. The
-framework of evolutionary game theory has been applied in %ToDo cite.
+framework of evolutionary game theory has been applied in explaining the
+behavior of sticklebacks~\cite{milinski1987tit}, the rock-scissors-paper cycles
+in bacterial populations~\cite{kerr2002local} and food sharing in vampire
+bats~\cite{carter2020development} colonies.
+
 Traditional approaches of evolutionary game theory assume that individuals meet
-each other at random in an infinitely large well-mixed population, such an
-approach is the replicator dynamics. The replicator dynamics describes how the
-abundance of strategic types in a population changes based on their fitness. In
-this deterministic formulation, individuals with higher fitness increase in
-abundance and ultimately, the system reaches a stable fixed point in which the
-population may consist either of a single type or of a mixture of different
-types. The works of cite have shown that by constrain the population to a
-finiteness of populations may lead to fundamental changes in this picture due to
-stochastic effects. These stochastic effects disadvantageous mutants have a
-small, yet non-zero probability to reach fixation in a finite population.
+each other at random in an infinitely large well-mixed population, an example of
+such an approach is the replicator dynamics. The replicator dynamics describes
+how the abundance of strategic types in a population changes based on their
+fitness. In this deterministic formulation, individuals with higher fitness
+increase in abundance and ultimately, the system reaches a stable fixed point in
+which the population may consist either of a single type or of a mixture of
+different types. The works of~\cite{nowak2004emergence, taylor2004evolutionary,
+imhof2005evolutionary, imhof2006evolutionary, traulsen2007pairwise} have shown
+that constraining the population to be finite can lead to fundamental changes
+due to stochastic effects. Due to these stochastic effects disadvantageous
+mutants have a small, yet non-zero probability to reach fixation in a finite
+population.
 
@@
 fitness-based processes in which an individual chosen proportional to fitness
-reproduces and the offspring replaces a randomly chosen individual [11]
+reproduces and the offspring replaces a randomly chosen individual~\cite{nowak2004emergence}
 \textit{moran process}; (ii) \textit{pairwise comparison processes} in which a
 pair of individuals is chosen, and where subsequently one of these individuals
-may adopt the strategy of the other.
-
-In pairwise comparison process considers a finite population of fixed size. In
+may adopt the strategy of the other~\cite{traulsen2007pairwise}.
+
+A pairwise comparison process considers a finite population of fixed size. In
 the population at a time step different types of individuals can exist. In the
@@
 
-At each step an individual change their type based.
-
-\begin{equation} \label{Eq:rho}
+At each step an individual change their type based. Two individuals, A and B,
+are selected at random. The individual chosen for reproduction A replaces B with
+probability \(\rho\), which depends on the fitness difference \(\pi _{{\rm
+A}}^{i} - \pi _{{\rm B}}^{i}\) between the two individuals. The composition of
+the population can only change if both individuals are of different types. The
+probability \(\rho\) follows the Fermi function,
+
+\begin{equation}
   \rho(\pi _{{\rm A}}^{i}, \pi _{{\rm B}}^{i}) = \frac{1}{1\!+\! \exp^{\!-\!\beta (\pi _{{\rm A}}^{i}\!-\!\pi _{{\rm B}}^{i})}}.
@@
 
-The parameter denotes the intensity of selection. For , fitness equals payoff.
-This scenario describes ``strong selection''. For \(\beta \ll 1\) , the
-payoff only provides a small perturbation to the overall fitness of an
-individual, a limit known as weak selection (Nowak et al., 2004).
-
-These indicate the expected payoff for two types in a population in state i. The
-interaction can be generated from a two-player matrix game which leads to
-payoffs linear in i [26], but we keep the formalism general to include games
-played between an arbitrary number of players, which leads to payoffs that
-depend on i in a polynomial way [27–30]. In fact, our results hold for an
-arbitrary dependence of the payoffs on i.
-
-In the simplest case, the payoffs of A and B individuals only depend on the
-fraction of both types in the population. If there are i A individuals and B
-individuals, then the A and B individuals have payoffs and , respectively.
-Self-interactions are excluded.
-
-Here, we consider a process based on pairwise comparison between individuals.
-Two individuals, A and B, are selected at random. The individual chosen for
-reproduction A replaces B with probability p, which depends on the payoff
-difference between the two individuals. The composition of the population can
-only change if both individuals are of different types. We follow (Blume, 1993,
-Szabó and Tőke, 1998, Hauert and Szabó, 2005) in choosing the Fermi function
-from statistical physics for p
-
-To explain the evolution of cooperation remains one of the greatest problems for
-biological and social sciences. Cooperation is the action of choosing to help
-others at one's own expense. The standard game of formulating such situation is
-the Prisoner's Dilemma~\cite{trivers1971evolution,
-milinski1987tit, glynatsi2021bibliometric, ohtsuki2006simple}, in which two
-players can choose to cooperate or to defect. The players are offered a certain
-payoff, \(R\), for mutual cooperation and a lower payoff, \(P\), for mutual
-defection. If one player cooperates while the other defects, then the cooperator
-gets the lowest payoff, \(S\), while the defector gains the highest payoff,
-\(T\). Thus, the payoffs have the following property \(T > R > P > S\), making
-defection the dominant strategy in the non-repeated game.
-
-In the context of the prisoner's dilemma the payoffs are calculated as follows.
-In the example that a strategy is playing always cooperate and always defect,
-
-This model has several assumptions. Two assumptions are that initally it is
-assumed that time to interact is faster than the time of the population
-update. Secondly it is assumed that the indiviail can interact will every
-other population member, and recall their interaction perfecttly.
-
-The first assumption was questioned by the work of.
-
-In this work
+The parameter \(\beta\) denotes the intensity of selection.
+
+The fitness of the individuals (\(\pi _{{\rm A}}^{i}, \pi _{{\rm B}}^{i}\)) is
+identified with the payoff resulting from the underline game. The evolution of
+cooperation remains one of the greatest problems for biological and social
+sciences. Cooperation is the action of choosing to help others at one's own
+expense. The standard game of formulating such situation is the Prisoner's
+Dilemma~\cite{trivers1971evolution, milinski1987tit, glynatsi2021bibliometric,
+ohtsuki2006simple}, in which two players can choose to cooperate or to defect.
+The players are offered a certain payoff, \(R\), for mutual cooperation and a
+lower payoff, \(P\), for mutual defection. If one player cooperates while the
+other defects, then the cooperator gets the lowest payoff, \(S\), while the
+defector gains the highest payoff, \(T\). Thus, the payoffs have the following
+property \(T > R > P > S\), making defection the dominant strategy in the
+non-repeated game.
+
+\begin{equation}\label{eq:payoffs}
+  \begin{blockarray}{ccc}
+      & \text{cooperate} & \text{defect} \\
+      \begin{block}{c(cc)}
+          \text{cooperate} & R & S \\
+          \text{defect} & T & P \\
+      \end{block}
+  \end{blockarray}
+\end{equation}
+
+In literature, the payoffs of A and B individuals only on the fraction of both
+types in the population. If there are \(k\) A individuals and \(N-k\) B
+individuals. In the context of the prisoner's dilemma the payoffs and assuming
+that A and B follow simple strategies where they cooperate and defect
+unconditionally respective, then the A and B individuals have fitness of
+
+\begin{equation}
+  \begin{array}{lcrcr}
+  \displaystyle \pi _{{\rm A}}^{i}	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot R	&+	&\displaystyle\frac{k}{N-1}\cdot S,\\[0.5cm]
+  \displaystyle \pi _{{\rm B}}^{i}&=	&\displaystyle\frac{N-k}{N-1}\cdot T &+	&\displaystyle\frac{k-1}{N-1}\cdot P.\\
+  \end{array}
+\end{equation}
+
+respectively while self-interactions are excluded. In literature these are the
+expected payoffs. For \(\beta = 1\), fitness equals payoff. This scenario
+describes ``strong selection''. For \(\beta \ll 1\) , the payoff only provides a
+small perturbation to the overall fitness of an individual, a limit known as
+weak selection.
+
+In the evolutionary game theory literature it is usually assumed that players
+use strategies with finite memory~\cite{Nowak1992tit, Baek2016}. This assumption
+is common as it allows for an explicit calculation of the players'
+payoffs~\cite{sigmund2010calculus}. Players' payoffs are calculated in the limit
+of interactions assuming that the two strategies meet several times.
+
+Customarily, most evolutionary game studies make the additional assumption that
+individuals play many times and with all other players before reproduction takes
+place, so that payoffs, equivalently fitness, are given by the mean distribution
+of types in the population. This hypothesis, selection occurs much more slowly
+than the interaction between individuals and more that remember all the
+interactions they participated in. Although recent experimental studies show
+that this may not always be the case in biology [12– 14], it is clear that in
+cultural evolution or social learning the time scale of selection is much closer
+to the time scale of interaction. The effects of this mixing of scales cannot be
+disregarded [15], and then it is natural to ask about the consequences of the
+above assumption and the effect of relaxing it. The work of~\cite{Roca2006}.
+
+In this work we ask studies that make use of this classical framework report
+that cooperation can substantially evolve. However, the inconsistency in the
+memory size of players leads us to question the robustness of our understanding
+of cooperation. To this end, we propose a framework in which individuals,
+similar to the decisions at each turn, estimate their fitness based on a minimum
+of information.
 
@@
 To investigate the role of extortion in the context of evolutionary games, we
-concentrate on the donation game
-
+concentrate on the donation game.
 In this section we explore the case where the updating payoffs are based on the
@@
   \centering
-  \begin{subfigure}{.65\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{static/expected_for_selection_strenght.pdf}
-    \includegraphics[width=\textwidth]{static/stochastic_for_selection_strenght.pdf}
-  \end{subfigure}%
-  \begin{subfigure}{.35\textwidth}
-    \vspace{.3cm}
-    \includegraphics[width=\textwidth]{static/cooperation_rate_over_betas.pdf}
-  \end{subfigure}
+    \includegraphics[width=\textwidth]{static/cooperation_rate_over_b.pdf}
 \caption{{\bf The evolution of cooperation for different selection strength values.}
@@
 combination.}\label{fig:cooperation_rate_over_betas}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \centering
+    \includegraphics[width=\textwidth]{static/more_memory.pdf}
+% \caption{{\bf The evolution of cooperation for different selection strength values.}
+% We vary the selection strength $\beta$. In all cases, stochastic payoff
+% evaluation tends to reduce the evolving cooperation rates. ({\bf A}) the
+% probabilities \(p, q\) for resident population over \(10^7\) time steps for each
+% \(\beta\) value. ({\bf B}) The cooperation rate within the resident population
+% (averaged over all games and over all time steps) over \(\beta\). Unless
+% explicitly varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$,
+% $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
+% $t\!=\!5\times 10^7$ time steps for each parameter
+% combination.}\label{fig:cooperation_rate_over_betas}
 \end{figure}

### 2022-08-16 main.tex
committing phase two

@@
 Traditional approaches of evolutionary game theory assume that individuals meet
-each other at random in an infinitely large well-mixed population, an example of
+each other at random in an infinitely large well-mixed population; an example of
 such an approach is the replicator dynamics. The replicator dynamics describes

### 2022-08-16 src/evol_dynamics/__init__.py

@@ -1,27 +1,27 @@
-# from .evolution import (
-#     fixation_probability_for_expected_payoffs,
-#     fixation_probability_for_stochastic_payoffs,
-#     imitation_probability,
-# )
+from .evolution import (
+    fixation_probability_for_expected_payoffs,
+    fixation_probability_for_stochastic_payoffs,
+    imitation_probability,
+)
 from .formulation import (
     expected_distribution_last_round,
     expected_distribution_opening_round,
     markov_chain_for_reactive_strategies,
     steady_state,
 )
-# from .invasion import simulate_until_invasion
-# from .numerical import (
-#     donation_game,
-#     harmony_game,
-#     snowdrift_game,
-#     stag_hunt_game,
-# )
-# from .simulation import (
-#     ReactivePlayer,
-#     create_population,
-#     match_pairs,
-#     stationary_for_16_states,
-#     simulated_states,
-#     get_probabilities_for_opponents,
-#     theoretical_probabilities_for_opponents,
-# )
+from .invasion import simulate_until_invasion
+from .numerical import (
+    donation_game,
+    harmony_game,
+    snowdrift_game,
+    stag_hunt_game,
+)
+from .simulation import (
+    ReactivePlayer,
+    create_population,
+    match_pairs,
+    stationary_for_16_states,
+    simulated_states,
+    get_probabilities_for_opponents,
+    theoretical_probabilities_for_opponents,
+)

### 2022-08-26 paper/cases.tex
working on the manuscript

\documentclass[tikz,border=2mm]{standalone}
 \begin{document}
 \begin{tikzpicture}
 \draw[help lines, color=gray!30, dashed] (0, 5) grid (5, 0);
 \draw[->,ultra thick] (0, 0)--(5, 0) node[right]{$\infty$} node[label={[xshift=-2.5cm, yshift=-1cm]$n$}] {};
 \draw[->,ultra thick] (0, 0)--(0, 5) node[above]{$N$} node[label={[xshift=-.5cm, yshift=-2.5cm]$m$}] {};

 \node[circle, inner sep=1pt, outer sep=1pt, fill, oneone, label={[xshift=.3cm, yshift=0cm]\tiny \textcolor{oneone}{one interaction - last round}}] at (1, 1) (1) {};

 \node[circle, inner sep=1pt, outer sep=1pt, fill, expectedcolor, label=above:{\tiny \textcolor{expectedcolor}{expected payoffs}}] at (5, 5) (1) {};

 \node[circle, inner sep=1pt, outer sep=1pt, fill, onetwo, label={[xshift=.5cm, yshift=0cm] \tiny \textcolor{onetwo}{one interaction - last two rounds}}] at (1, 2) (1) {};
 \node[circle, inner sep=1pt, outer sep=1pt, fill, twoone, label={[xshift=.5cm, yshift=-.6cm] \tiny \textcolor{twoone}{two interactions - last round}}] at (2, 1) (1) {};
 \node[circle, inner sep=1pt, outer sep=1pt, fill, twotwo, label={[xshift=.5cm, yshift=-.6cm] \tiny \textcolor{twotwo}{two interactions - last two rounds}}] at (2, 2) (1) {};
 \end{tikzpicture}
 \end{document}

### 2022-08-26 paper/game_stage.tex
working on the manuscript

\documentclass{standalone}

\begin{document}
\tikzset{man/.pic={
    \node[circle,fill,minimum size=5mm] (head) {};
    \node[rounded corners=2pt,minimum height=.8cm,minimum width=0.4cm,fill,below = 1pt of head] (body) {};
    \draw[line width=1mm,round cap-round cap] ([shift={(2pt,-1pt)}]body.north east) --++(-90:6mm);
    \draw[line width=1mm,round cap-round cap] ([shift={(-2pt,-1pt)}]body.north west)--++(-90:6mm);
    }}
\begin{tikzpicture}

    % %%%%% game stage %%%%%
    \node at (0, .7) (turns) {turn};

    \pic[BurntOrange] at (0, 0) (myman) {man};
    \node at (0, -.6) {B};
    \pic[RoyalBlue] at (0, -2) (myman) {man};
    \node at (0, -2.6) {F};

    \coordinate (B) at (0.5, -.6);
    \coordinate (F) at (0.5, -2.6);

    \node at (1.5, .7) (turn1) {1};
    \node[circle, inner sep=1pt, outer sep=1pt, draw, fill] at (1.5, 1.7) (exit) {};
    \node[circle, draw] (C) at (1.5, -.6)  {$C$};
    \node[circle, draw] (D) at (1.5, -2.6) {$D$};

    \draw (B) edge[out=0, in=180, ->, thick] node [above] {$y_{B}$} (C);
    \draw (F) edge[out=0, in=180, ->, thick] node [above] {$y_{F}$} (D);

    \node at (3, .7) (turn2) {2};
    \node[circle, inner sep=1pt, outer sep=1pt, draw, fill] at (3, 1.7) (exit2) {};
    \node[circle, draw] (C2)at (3, -.6)  {$C$};
    \node[circle, draw] (D2) at (3, -2.6) {$D$};

    \draw (C) edge[out=0, in=180, ->, thick] node [above] {$q_{B}$} (C2);
    \draw (D) edge[out=0, in=180, ->, thick] node [above] {$p_{F}$} (D2);

    \draw (turn1) edge[out=15, in=165, ->, thick] node [above] {$\delta$} (turn2);
    \draw (turn1) edge[out=90, in=-90, ->, thick] node [above, rotate=90] {$1 - \delta$} (exit);
    \draw (turn2) edge[out=90, in=-90, ->, thick] node [above right] {$1 - \delta$} (exit2);

    \node[circle] at (4, -.6) {$\dots$};
    \node[circle] at (4, -2.6) {$\dots$};

\end{tikzpicture}
\end{document}

### 2022-08-26 paper/pairwise_comparison_process_diagram.tex
working on the manuscript

--- 2022-08-26 paper/game_stage.tex
@@
 
-    % %%%%% game stage %%%%%
-    \node at (0, .7) (turns) {turn};
+    \node at (0, 0) {\includestandalone{population}};
+    \draw[thick] (2.5, 3) -- (2.5, -2);
 
-    \pic[BurntOrange] at (0, 0) (myman) {man};
-    \node at (0, -.6) {B};
-    \pic[RoyalBlue] at (0, -2) (myman) {man};
-    \node at (0, -2.6) {F};
+    \node at (3.5, 2.5) {B)};
+    \node at (5, 2) {\includestandalone[width=.13\textwidth]{mutation}};
 
-    \coordinate (B) at (0.5, -.6);
-    \coordinate (F) at (0.5, -2.6);
+    \node at (3, -.7) {D)};
+    \node at (5, -1) {\includestandalone[width=.25\textwidth]{updating}};
 
-    \node at (1.5, .7) (turn1) {1};
-    \node[circle, inner sep=1pt, outer sep=1pt, draw, fill] at (1.5, 1.7) (exit) {};
-    \node[circle, draw] (C) at (1.5, -.6)  {$C$};
-    \node[circle, draw] (D) at (1.5, -2.6) {$D$};
+    \node at (8.4, 2.5) {C)};
+    \node at (10, .5) {\includestandalone[width=.25\textwidth]{game_stage}};
 
-    \draw (B) edge[out=0, in=180, ->, thick] node [above] {$y_{B}$} (C);
-    \draw (F) edge[out=0, in=180, ->, thick] node [above] {$y_{F}$} (D);
-
-    \node at (3, .7) (turn2) {2};
-    \node[circle, inner sep=1pt, outer sep=1pt, draw, fill] at (3, 1.7) (exit2) {};
-    \node[circle, draw] (C2)at (3, -.6)  {$C$};
-    \node[circle, draw] (D2) at (3, -2.6) {$D$};
-
-    \draw (C) edge[out=0, in=180, ->, thick] node [above] {$q_{B}$} (C2);
-    \draw (D) edge[out=0, in=180, ->, thick] node [above] {$p_{F}$} (D2);
-
-    \draw (turn1) edge[out=15, in=165, ->, thick] node [above] {$\delta$} (turn2);
-    \draw (turn1) edge[out=90, in=-90, ->, thick] node [above, rotate=90] {$1 - \delta$} (exit);
-    \draw (turn2) edge[out=90, in=-90, ->, thick] node [above right] {$1 - \delta$} (exit2);
-
-    \node[circle] at (4, -.6) {$\dots$};
-    \node[circle] at (4, -2.6) {$\dots$};
+    \draw[thick] (7.7, 3) -- (7.7, -2);
 

### 2022-08-26 paper/si.tex
working on the manuscript

--- 2021-11-24 main.tex
@@
 \documentclass[11pt]{article}
+\usepackage[dvipsnames]{xcolor}
 \usepackage{times}
-\usepackage{amsmath,amsthm,amssymb,setspace,enumitem,epsfig,titlesec,verbatim,color,array,eurosym,multirow}
+\usepackage{amsmath,amsthm,amssymb,setspace,enumitem,epsfig,titlesec,verbatim,array,eurosym,multirow}
 \usepackage[sort&compress]{natbib}
@@
 \newcommand{\R}{\mathbb{R}}
+
+\usetikzlibrary{decorations.pathreplacing, backgrounds, fit, calc, matrix, positioning}
 
@@
 \newcommand{\alex}[1]{\textcolor{red}{\textbf{AL}: #1}}
+
+\definecolor{expectedcolor}{rgb}{0.1791464821222607, 0.49287197231833907, 0.7354248366013072}
+\definecolor{oneone}{rgb}{0.8503344867358708, 0.14686658977316416, 0.13633217993079583}
+\definecolor{onetwo}{rgb}{0.18246828143021915, 0.5933256439830834, 0.3067589388696655}
+\definecolor{twoone}{rgb}{0.4488427527873895, 0.3839600153787005, 0.6738792772010764}
+\definecolor{twotwo}{rgb}{0.8871510957324106, 0.3320876585928489, 0.03104959630911188}
 
@@
 
-\title{\bf  \sffamily \LARGE Evolution of cooperation among individuals with
+\title{\bf  \sffamily \LARGE Supplementary Information: Evolution of cooperation among individuals with
 limited payoff memory\\}
@@
 
-\begin{abstract}
-% Repeated games are vastly use in evolutionary game theory to explain cooperation in a variety of environments. Although existing models have shaped
-% our understanding of human cooperation, hey often work with idealized
-% assumptions. In an evolutionary process, individuals imitate other individuals
-% of the population based on their fitness. It is commonly assumed that an
-% individual computes their fitness after interacting with a representative sample
-% of the population, and remembering all the interactions they participated in. In
-% real life, we do not always remember all our interactions, instead we rather
-% recall our most recent ones.
+Section~\ref{section:pairwise_comparison} gives a brief overview of the pairwise
+comparison process. Section~\ref{section:updating_payoffs} describes the
+conventional approach for calculating updating payoffs, as well as, our
+approaches and methods.
 
-% Here, we introduce a framework that allows individuals to estimate their fitness
-% based on a minimum of social information. We explore the difference between our
-% framework and the classical framework using computer simulations. We present
-% results for the most commonly used classes of symmetric \(2 \times 2\) games.
-% The simulations show that, in the prisoner's dilemma individuals with limited
-% memory tend to adopt less generous strategies and they achieve less cooperation
-% than in the classical scenario. In contrast, in the stag-hunt and the harmony
-% game, the impact of memory is less striking. Here individuals with limited
-% memory perform nearly as well as individuals with full memory. Finally, we
-% observe that the snowdrift game is the only game for which cooperation can be
-% underestimated in the classical scenario.
-\end{abstract}
+\section{Pairwise comparison process}\label{section:pairwise_comparison}
 
-\section{Introduction}
+Pairwise comparison process is a stochastic process for modelling the evolution
+of a finite population. The process starts with assigning all individuals of the
+population the same strategy. A strategy is a set of rules of how an individual
+should behave in an interaction with another individual. Each elementary time
+step of the process consists of three phases; (1) the \textbf{mutation phase}
+(2) the \textbf{game phase} (3) the \textbf{update phase}. These are summarised
+in Figure~\ref{fig:pairwise_phases}. In the mutation phase one individual is
+chosen to switch to a new mutant strategy with a probability \(\mu\). The mutant
+strategy is randomly selected from the set of feasible strategies.
 
-Evolutionary game theory~\cite{smith1982evolution, hofbauer1998evolutionary,
-nowak2004evolutionary, hauert2005game} describes the evolutionary dynamics of
-populations consisting of different types of interacting individuals. The
-framework of evolutionary game theory has been applied in %ToDo cite.
-Traditional approaches of evolutionary game theory assume that individuals meet
-each other at random in an infinitely large well-mixed population, such an
-approach is the replicator dynamics. The replicator dynamics describes how the
-abundance of strategic types in a population changes based on their fitness. In
-this deterministic formulation, individuals with higher fitness increase in
-abundance and ultimately, the system reaches a stable fixed point in which the
-population may consist either of a single type or of a mixture of different
-types. The works of cite have shown that by constrain the population to a
-finiteness of populations may lead to fundamental changes in this picture due to
-stochastic effects. These stochastic effects disadvantageous mutants have a
-small, yet non-zero probability to reach fixation in a finite population.
+In the game phase individuals are randomly matched with other individuals in the
+population, and they engage in a match where each subsequent turn occurs with a
+fixed probability \(\delta\). At each turn the individuals play a social dilemma
+and decide on an action based on their strategies. In the donation game there
+are two actions: cooperation (\(C\)) and defection (\(D\)). By cooperating a player
+provides a benefit \(b\) to the other player at their cost \(c\), with \(0 < c <
+b\). Thus the payoffs for a player in each turn are,
 
-Two classes of such finite stochastic processes have been used extensively: (i)
-fitness-based processes in which an individual chosen proportional to fitness
-reproduces and the offspring replaces a randomly chosen individual [11]
-\textit{moran process}; (ii) \textit{pairwise comparison processes} in which a
-pair of individuals is chosen, and where subsequently one of these individuals
-may adopt the strategy of the other.
+\begin{equation}
+    \begin{blockarray}{ccc}
+        & \text{cooperate} & \text{defect} \\
+        \begin{block}{c(cc)}
+            \text{cooperate} & b - c & -c \\
+            \text{defect} & b & 0 \\
+        \end{block}
+    \end{blockarray}.
+\end{equation}
 
-In pairwise comparison process considers a finite population of fixed size. In
-the population at a time step different types of individuals can exist. In the
-simplest of cases each individual can be of one of two types, A and B. The state
-of the population is thus characterized by the number \(k\) of individuals of
-type A. The interaction between the two types of individuals is described by the
-functions \(\pi _{{\rm A}}^{i}\) and $\pi _{{\rm B}}^{i}$.
+Let \(u = (b-c, -c, b, 0)\) be payoffs in a vector format, and let \(\mathcal{U}
+= \{b-c, -c, b, 0\}\) denote the set of feasible payoffs. In repeated games
+there are infinite many strategies. However, it is commonly assumed that
+individuals can only choose strategies from a restricted set. Here we explore
+the case where individuals can use reactive strategies. A reactive strategy
+considers only the previous action of the other player, and thus, a reactive
+strategy \(s\) can be written as a three-dimensional vector \(s=(y, p, q)\). The
+parameter \(y\) is the probability that the strategy opens with a cooperation
+and \(p\), \(q\) are the probabilities that the strategy cooperates given that
+the opponent cooperated and defected equivalently.
 
-At each step an individual change their type based.
+Following the game phase is the update stage where two individuals are randomly selected.
+One individual serves as the `learner' and the other as the `role model'.
+The learner adopts the role model's strategy with a probability \(\rho\) given by,
 
 \begin{equation} \label{Eq:rho}
-  \rho(\pi _{{\rm A}}^{i}, \pi _{{\rm B}}^{i}) = \frac{1}{1\!+\! \exp^{\!-\!\beta (\pi _{{\rm A}}^{i}\!-\!\pi _{{\rm B}}^{i})}}.
+    \rho(\pi_{L}, \pi_{RM}) = \frac{1}{1\!+\! \exp^{\!-\!\beta (\pi_{RM}\!-\! \pi_{L})}}.
 \end{equation}
 
-The parameter denotes the intensity of selection. For , fitness equals payoff.
-This scenario describes ``strong selection''. For \(\beta \ll 1\) , the
-payoff only provides a small perturbation to the overall fitness of an
-individual, a limit known as weak selection (Nowak et al., 2004).
-
-These indicate the expected payoff for two types in a population in state i. The
-interaction can be generated from a two-player matrix game which leads to
-payoffs linear in i [26], but we keep the formalism general to include games
-played between an arbitrary number of players, which leads to payoffs that
-depend on i in a polynomial way [27–30]. In fact, our results hold for an
-arbitrary dependence of the payoffs on i.
-
-In the simplest case, the payoffs of A and B individuals only depend on the
-fraction of both types in the population. If there are i A individuals and B
-individuals, then the A and B individuals have payoffs and , respectively.
-Self-interactions are excluded.
-
-Here, we consider a process based on pairwise comparison between individuals.
-Two individuals, A and B, are selected at random. The individual chosen for
-reproduction A replaces B with probability p, which depends on the payoff
-difference between the two individuals. The composition of the population can
-only change if both individuals are of different types. We follow (Blume, 1993,
-Szabó and Tőke, 1998, Hauert and Szabó, 2005) in choosing the Fermi function
-from statistical physics for p
-
-To explain the evolution of cooperation remains one of the greatest problems for
-biological and social sciences. Cooperation is the action of choosing to help
-others at one's own expense. The standard game of formulating such situation is
-the Prisoner's Dilemma~\cite{trivers1971evolution,
-milinski1987tit, glynatsi2021bibliometric, ohtsuki2006simple}, in which two
-players can choose to cooperate or to defect. The players are offered a certain
-payoff, \(R\), for mutual cooperation and a lower payoff, \(P\), for mutual
-defection. If one player cooperates while the other defects, then the cooperator
-gets the lowest payoff, \(S\), while the defector gains the highest payoff,
-\(T\). Thus, the payoffs have the following property \(T > R > P > S\), making
-defection the dominant strategy in the non-repeated game.
-
-In the context of the prisoner's dilemma the payoffs are calculated as follows.
-In the example that a strategy is playing always cooperate and always defect,
-
-This model has several assumptions. Two assumptions are that initally it is
-assumed that time to interact is faster than the time of the population
-update. Secondly it is assumed that the indiviail can interact will every
-other population member, and recall their interaction perfecttly.
-
-The first assumption was questioned by the work of.
-
-In this work
-
-We first consider two extreme scenarios, the classical scenario and the
-alternative scenario where individuals update their strategies only based on the
-very last payoff they obtained. We observe that individuals with limited memory
-tend to adopt less generous strategies and they achieve less cooperation when
-interacting in a prisoner's dilemma. We obtain similar results when we consider
-that individuals update their strategies based on more information. More
-specifically, up to the last two payoffs they obtained when interacting with up
-to two different members of the population. We extend our approach to the rest
-of the symmetric \(2 \times 2\) games.
-
-The remainder of the paper is organized as follows. In
-section~\ref{section:model} we describe the model. In
-section~\ref{section:results} we present the results of the simulations, and in
-section~\ref{section:conclusions} we outline the main conclusions.
-
-\section{Model Setup}\label{section:model}
-
-In the following, we consider a well mixed population of fixed size. In each
-step of t
-
-We study the transmission of strategies with a frequency-dependent birth-death
-process26 in a finite population of size n. In each time step, two randomly
-chosen individuals compare their payoffs and one of them can switch to the other
-one's strategy. This process can be interpreted as a model for social learning,
-whereby successful strategies spread, and, occasionally, random strategy
-exploration introduces novel strategies (corresponding to mutations in
-biological models).
-
-We consider a population of \(N\) players~\footnote{The terms ``player'' and
-``individual'' are used interchangeably here.} where \(N\) is even and mutations
-are sufficiently rare. Therefore, at any point in time there are at most two
-different strategies present in the population; a \textit{resident} strategy and
-a \textit{mutant} strategy. To describe how strategies spread we use a pairwise
-comparison process~\cite{Traulsen2006}. Each step of the evolutionary process
-consists of two stages, a game stage and an updating stage.
-
-In the game stage each individual is randomly matched with some other individual
-in the population. They engage in a match where each subsequent turn occurs with
-a fixed probability $\delta$. At each turn players choose independently to
-either cooperate (\(C\)) or to defect (\(D\)), and the payoffs of the turn
-depend on both their decisions. If both players cooperate they receive the
-reward payoff \(R\), whereas if both defect they receive the punishment payoff
-\(P\). If one cooperates but the other defects, the defector receives the
-temptation payoff \(T\), whereas the cooperator receives the sucker's payoff
-\(S\). We denote the feasible payoff of each turn as \(\mathcal{U} = \{R, S, T,
-P\}\). We assume that individuals use \textit{reactive strategies} to make
-decisions in each turn. Reactive strategies are a set of memory-one strategies
-that only take into account the previous action of the opponent. They can be
-written explicitly as a vector in \(\R_{3}\), more specifically, a reactive
-strategy \(s\) is given by \(s=(y, p, q)\). The parameter \(y\) is the
-probability that the strategy opens with a cooperation and \(p, q\) are the
-probabilities that the strategy cooperates given that the opponent cooperated
-and defected equivalently.
-
-In the updating stage, two players are randomly drawn from the population, a
-`learner' and a `role model'. Given the learner's payoff $u_L\!\in\!
-\mathcal{U}$ and the role model's payoff $u_{RL}\!\in\! \mathcal{U}$, the
-learner adopts the role model's strategy with probability,
-
-\begin{equation} \label{Eq:rho}
-  \rho(u_{L}, u_{RM}) = \frac{1}{1\!+\! \exp^{\!-\!\beta (u_{RM}\!-\!u_{L})}}.
-\end{equation}
-
-where $\beta\!\ge\!0$ is the strength of selection. For small values of $\beta$
-the imitation probability is independent of the strategies of the involved
-players. As the value of $\beta$ increases, the more likely it is that the
-learner adopts only strategies that yield a higher payoff. Conventionally the
-updating payoffs of the learner and the role model are based on their expected
-payoffs. A player's expected payoff is the mean payoff the player yields after
-engaging in matches of multiple turns with each member of the population. At
-each match a player bases their next turn decision only on the previous action
-of the opponent, however, the same player bases their expected payoffs on the
-outcomes of all their matches. Thus, a player is assumed to have limited and
-perfect memory at the same time. We propose a new a set of updating payoffs
-where it is also assumed that a player has limited memory. We referee to these
-as the limited memory payoffs.
-
-The evolutionary step is repeated until either the mutant strategy goes extinct,
-or until it fixes in the population. If the mutant fixes in the population then
-the mutant strategy becomes the new resident strategy. After either outcome we
-introduce a new mutant strategy uniformly chosen from all reactive strategies at
-random, and we set the number of mutants to $1$. This process of mutation and
-fixation/extinction is then iterated many times.
-
-In order to account for the effect of the updating payoffs we simulate the
-evolutionary process and record which strategies the players adopt over time
-based on (i) the expected payoffs (ii) the limited memory payoffs. We compare
-the difference in the cooperation rate within the resident population for the
-two approaches. To account for the various types of social behaviour we also
-present results on multiple social dilemmas.
-
-\section{Results}\label{section:results} \subsection{Updating payoffs based on
-the last round with another member of the population}\label{section:donation}
-
-To investigate the role of extortion in the context of evolutionary games, we
-concentrate on the donation game
-
-In this section we explore the case where the updating payoffs are based on the
-last round payoff achieved against another member of the population and we
-compare this to the classical scenario of the expected payoffs. We assume that
-each pair of players interacts in a donation game. The donation game is a
-special case of the prisoner's dilemma. Each player can choose to cooperate by
-providing a benefit \(b\) to the other player at their cost \(c\), with \(0 < c
-< b\). Thus, the feasible payoffs in each round are \(\mathcal{U} = \{b-c, -c,
-b, 0\}\).
-
-Figure~\ref{fig:expected_and_stochastic_for_donation} shows simulations results
-for the described process of section~\ref{section:model}.
-Figure~\ref{fig:expected_and_stochastic_for_donation} depicts the evolving
-conditional cooperation probabilities $p$ and $q$. The discount factor~$\delta$
-is comparably high, thus we do not report the opening move \(y\) as it is a
-transient effect. The left panels correspond to the standard scenario considered
-in the literature, it considers players who use expected payoffs to update their
-strategies. The right panel shows the scenario considered herein, in which
-players update their strategies based on their last round’s payoff. The top
-panels assume a benefit \(b\) of 3 whereas the bottom assume a benefit of 10.
-
-The figure suggests that when updating is based on expected payoffs players tend
-to be more generous and more cooperative. The $q$-values of the resident
-strategies are on average higher in the case of the expected payoffs. The
-players will occasionally forgive a defection more often if their fitness
-depends on interacting with every member of the population. On the other hand,
-when social interactions are limited they are less forgiving. The average
-cooperation rate for each simulation is calculated as the average cooperation
-rate within the resident population. In the case of the expected payoffs,
-regardless the value of benefit, the average cooperation rate is strictly higher
-than that of the last round payoffs. The difference based on the two methods is
-statistically significant, and in the case of $b=10$ the average cooperation of
-resident strategies drops from 97\% to 57\%.
+\(\pi_{L}\) and \(\pi_{RM}\) are the updating payoffs of the learner and the
+role model respectively. The updating payoff is a measure of how successful an
+individual is in the current standing of the population. In the next section we
+will explain how these payoffs are calculated in more detail. The parameter
+\(\beta\) is known as the selection strength, namely, it shows how important the
+payoff difference is when the learner is considering adopting the strategy of
+the role model.
 
@@
     \centering
-    \includegraphics[width=.70\textwidth]{static/expected_and_stochastic_for_donation_game.pdf}
-    % \includegraphics[width=.70\textwidth]{static/expected_and_stochastic_for_donation_game_b_10.pdf}
-    \caption{{\bf Evolutionary dynamics under expected payoffs and last round with one interaction payoffs.} 
-    We have run two simulations of the evolutionary process described in
-    section~\ref{section:model} for $t\!=\!10^7$ time steps. For each time step,
-    we have recorded the current resident population ($y,p,q$). Since
-    simulations are run for a relatively high continuation probability of
-    $\delta\!=\!0.999$, we do not report the players' initial cooperation
-    probability $y$. The graphs show how often the resident population chooses
-    each combination ($p,q$) of conditional cooperation probabilities in the
-    subsequent rounds. ({\bf A}) If players update based on their expected
-    payoffs, the resident population typically applies a strategy for which
-    $p\!\approx\!1$ and $q\!\le\!1\!-\!c/b\!=\!0.9$. ({\bf B}) When players
-    update their strategies based on their realized payoffs in the last round,
-    there are two different predominant behaviors. The resident population
-    either consists of defectors (with $p\!\approx\!q\!\approx\!0$) or of
-    conditional cooperators. In the latter case, the maximum level of $q$
-    consistent with stable cooperation is somewhat smaller compared to the
-    expected-payoff setting, $q\!<\!0.5$. The cooperation rate within the
-    resident population (averaged over all games and over all time steps) is
-    close to 100\%. Parameters: $N\!=\!100$, $c\!=\!1$, $\beta\!=\!1$,
-    $\delta\!=\!0.999$.}
-    \label{fig:expected_and_stochastic_for_donation}
+    \includestandalone[width=\textwidth]{pairwise_comparison_process_diagram}
+    \caption{\textbf{Pairwise process phases.} \textbf{A)} The process begins
+    with a finite population where each member is assigned a given strategy.
+    Each color represents a different strategy, and the members are labelled by
+    letters. \textbf{B) Mutation phase.} An individual is selected (in the
+    example individual A) and with a given probability \(\mu\) that individual
+    adopts a new mutant strategy. \textbf{C) Game phase.} individuals are
+    selected to interact in a repeated social dilemma with other individuals. We
+    demonstrate the case where individuals B and F have been selected to
+    interact. They use the reactive strategies \(s_{B} = (y_B, p_B, q_B)\) and
+    \(s_{F} = (y_F, p_F, q_F)\) respectively. The opening moves depend on their
+    \(y_i\) probability. In turn 1, individual B cooperated,
+    thus, F cooperates with a probability \(p_{F}\) in turn 2. On the opposite,
+    individual F defected in turn 1, and so B cooperates in the next turn with
+    a probability \(q_{B}\). At each turn there is a probability \(delta\) that a
+    subsequent turn will occur, and with a probability \(1- delta\) the interaction ends.
+    \textbf{D) Updating phase.} At the updating phase two individuals are chosen;
+    one is assigned the role of the learner and the other one the role
+    of the role model. In our example C adopts D's strategy with a probability
+    \(\rho(\pi_{C}, \pi_{D})\) where \(pi_{C}, \pi_{D}\) are the updating
+    payoffs of the individuals.}\label{fig:pairwise_phases}
 \end{figure}
 
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[width=.9\textwidth]{static/cooperation_rates_expected_and_stochastic_for_donation_game.pdf}
-\end{figure}
+This elementary population updating process is repeated for a large number of
+time steps, and at each time step we record the state of the population.
 
-We further explore the effect of benefit in
-Figure~\ref{fig:cooperation_rate_over_benefit}. The figure suggests that
-expected payoffs always yield a higher cooperation rate. In the case of expected
-payoffs we observe that the cooperation rate increases as the value of the
-benefit gets higher. In comparison for the limited memory payoffs, the
-cooperation rate remains unchanged at approximately 50\% once \(b=5\).
+\subsection{Low mutation \(\mu \rightarrow 0\)}
 
-% \begin{figure}[!htbp]
-%   \centering
-%   \begin{subfigure}{.5\textwidth}
-%     \centering
-%     \includegraphics[width=\textwidth]{static/expected_for_beta.pdf}
-%     \includegraphics[width=\textwidth]{static/stochastic_for_beta.pdf}
-%   \end{subfigure}%
-%   \begin{subfigure}{.5\textwidth}
-%     \centering
-%     \includegraphics[width=\textwidth]{static/cooperation_rate_over_b.pdf}
-%   \end{subfigure}
-%   \caption{{\bf The evolution of cooperation for different benefit values.} 
-%   We vary the benefit of defection $b$. In all cases, expected payoffs appear to
-%   overestimate the average cooperation rate the population achieves. ({\bf A})
-%   the probabilities \(p, q\) for resident population over \(10^7\) time steps
-%   for each benefit value. ({\bf B}) The cooperation rate within the resident
-%   population (averaged over all games and over all time steps) over the benefit.
-%   Unless explicitly varied, the parameters of the simulation are $N\!=\!100$,
-%   $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
-%   $t\!=\!5\times 10^7$ time steps for each parameter
-%   combination.}\label{fig:cooperation_rate_over_benefit}
-% \end{figure}
+In the case of low mutation (\(\mu \rightarrow 0\)) we assume that mutations are
+rare. In fact, so rare that only two different strategies can be present in the
+population at any given time. The case of low mutation is vastly adopted because
+it allows us to explicitly calculate the fixation probability of a newly
+introduced mutant.
 
-We also investigate the effect of the strength of selection $\beta$.
-Figure~\ref{fig:cooperation_rate_over_betas} illustrates results for various
-runs of the evolutionary process. For weak selection, \(\beta < 1\), we observe
-that the two methods yield similar results, however, as \(\beta\) increases
-there is variation in the evolving populations. In the case of expected payoffs
-the resident populations become more cooperative as \(\beta\) increases, whereas
-in the case of limited memory payoffs, the resident populations become more
-defective.
-
-\begin{figure}[!htbp]
-  \centering
-  \begin{subfigure}{.65\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{static/expected_for_selection_strenght.pdf}
-    \includegraphics[width=\textwidth]{static/stochastic_for_selection_strenght.pdf}
-  \end{subfigure}%
-  \begin{subfigure}{.35\textwidth}
-    \vspace{.3cm}
-    \includegraphics[width=\textwidth]{static/cooperation_rate_over_betas.pdf}
-  \end{subfigure}
-\caption{{\bf The evolution of cooperation for different selection strength values.}
-We vary the selection strength $\beta$. In all cases, stochastic payoff
-evaluation tends to reduce the evolving cooperation rates. ({\bf A}) the
-probabilities \(p, q\) for resident population over \(10^7\) time steps for each
-\(\beta\) value. ({\bf B}) The cooperation rate within the resident population
-(averaged over all games and over all time steps) over \(\beta\). Unless
-explicitly varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$,
-$c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
-$t\!=\!5\times 10^7$ time steps for each parameter
-combination.}\label{fig:cooperation_rate_over_betas}
-\end{figure}
-
-% \subsection{Effect of updating payoffs in different social
-% dilemmas}\label{section:2_by_2_games}
-
-% In the previous section we gained insights into the effects of the updating
-% payoffs, and into how parameters such as the benefit and the strength of
-% selection can intensify them. We investigated these effects by using the
-% donation game. In order to broaden our understanding of the updating payoffs on
-% different forms of possible human interactions we extend our approach to all \(2
-% \times 2\) symmetric games. More specifically, we apply our analysis to four
-% different classes of games; the harmony, the snowdrift, the prisoner's dilemma
-% and the stag-hunt games.
-
-% We compare the results of the evolutionary process when the updating payoffs are
-% based on the expected payoffs, and on the last round payoffs for all the four
-% possible classes of games (Figures~\ref{fig:expected_two_by_two}
-% and~\ref{fig:last_round_two_by_two}). In Figure~\ref{fig:expected_two_by_two}
-% each sub figure represents a run of evolutionary process for a different set of
-% values for \(\mathcal{U}\). Without loss of generality we set \(R=1\) and
-% \(P=0\)~\cite{Martinez2012, Roca2009}, and we vary the values of the temptation
-% payoff \(T\) (across the \(x\)-axis) and of the sucker's payoff \(S\) (across
-% the \(y\)-axis). Starting at the upper left corner and proceeding clockwise the
-% quadrants correspond to the harmony, the snowdrift, the prisoner's dilemma and
-% the stag-hunt games.
-
-% \begin{figure}[!htbp]
-%   \centering
-%   \includegraphics[width=\textwidth]{static/expected_two_by_two_games.pdf}
-%   \caption{{\bf Evolutionary dynamics under expected payoffs for various social dilemmas.}
-%   We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across
-%   the \(x\) axis, and  \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\) across the
-%   \(y\) axis. ({\bf A}) The top left quadrant where \(S > 0\) and \(T \leq 1\)
-%   corresponds to the harmony game. The preference ordering for the harmony game
-%   is \(R > T > S > P\). ({\bf B}) The top right quadrant where \(S > 0\) and \(T
-%   > 1\) corresponds to the snowdrift game. The preference ordering for the
-%   snowdrift game is \(T > R > S > P\). ({\bf C}) The bottom left quadrant where
-%   \(S < 0\) and \(T \leq 1\) corresponds to the stag-hunt game. The preference
-%   ordering for the stag-hunt game is \(R > T > P > S\). ({\bf D}) The bottom
-%   right quadrant where \(S < 0\) and \(T > 1\) corresponds to the prisoner's
-%   dilemma game. The preference ordering for the prisoner's dilemma game is \(T >
-%   R > P > S\). Unless explicitly varied, the parameters of the simulation are
-%   $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}
-%   \label{fig:expected_two_by_two}
-% \end{figure}
-
-% \begin{figure}[!htbp]
-%   \centering
-%   \includegraphics[width=\textwidth]{static/stochastic_two_by_two_games.pdf}
-%   \caption{{\bf Evolutionary dynamics under last round payoffs for various social dilemmas.} 
-%   We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across
-%   the \(x\) axis, and  \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\) across the
-%   \(y\) axis. ({\bf A}) The top left quadrant where \(S > 0\) and \(T \leq 1\)
-%   corresponds to the harmony game. The preference ordering for the harmony game
-%   is \(R > T > S > P\). ({\bf B}) The top right quadrant where \(S > 0\) and \(T
-%   > 1\) corresponds to the snowdrift game. The preference ordering for the
-%   snowdrift game is \(T > R > S > P\). ({\bf C}) The bottom left quadrant where
-%   \(S < 0\) and \(T \leq 1\) corresponds to the stag-hunt game. The preference
-%   ordering for the stag-hunt game is \(R > T > P > S\). ({\bf D}) The bottom
-%   right quadrant where \(S < 0\) and \(T > 1\) corresponds to the prisoner's
-%   dilemma game. The preference ordering for the prisoner's dilemma game is \(T >
-%   R > P > S\). Unless explicitly varied, the parameters of the simulation are
-%   $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}
-%   \label{fig:last_round_two_by_two}
-% \end{figure}
-
-% The harmony game describes an interaction without conflict. In the harmony game
-% it is in the best interest for both players to cooperate, and in an evolutionary
-% process cooperators prevail. We observe that for most harmony game runs the
-% resident population overwhelmingly applies a strategy for which \(p\) and \(q\)
-% are \(\approx 1\) (Figure~\ref{fig:expected_two_by_two}). The snowdrift game
-% describes a situation similar to that of the prisoner's dilemma; cooperation
-% results in a benefit to the opposing player but entailing a cost to the
-% cooperator. However, in the snowdrift game individuals obtain immediate direct
-% benefits from the cooperative acts which leads to \(S > P\). In the snowdrift
-% game more cooperation emerges compared to the prisoner's dilemma. Defection is
-% never a resident strategy and for the same values of temptation the overall
-% \(q-\) values are higher. The last class of games we present are for the
-% stag-hunt game. In the stag-hunt game there is an equilibrium in which both
-% players cooperate as well as one in which both defect. For small values of the
-% temptation and the sucker's payoffs we observe that cooperation prevails similar
-% to the harmony game. However, in the case where the temptation and the reward
-% payoffs are equal the resident population either consists of defectors or
-% cooperators.
-
-% Although the patterns of the evolved populations remain similar if we consider
-% that players update their strategies based on their last round payoff
-% (Figure~\ref{fig:last_round_two_by_two}), there are still dissimilarities
-% between the two approaches. To better understand these dissimilarities, we
-% explore the cooperation rates for each of the cases of the social dilemmas we
-% have presented (Figure~\ref{fig:cooperation_rates_two_by_two_part_one}).
-
-% In the harmony game the cooperation rates for both approaches are similar. The
-% biggest difference occurs for \(T=1\), nevertheless, it is less than 10\%. There
-% are two instances for which the cooperation rates are higher for the limited
-% memory payoffs compared to the expected. These are for \(T=2.2, S=0.4\) and for
-% \(T=2.2, S=0.8\). It is only in the snowdrift game that cooperation is
-% overestimated when a minimum of information is considered. Players are slightly
-% more likely to cooperate after a cooperation and to forgive after being at the
-% receiving end of a defection. For the rest of the cases the classical scenario
-% overestimates the average cooperation rate, supporting the results of the
-% previous section. There is a substantial difference in the cooperation rates for
-% the prisoner's dilemma class. We observe that for \(S<-0.4\) cooperation almost
-% never emerges if players imitate based only on the last round payoff.
-
-% \begin{figure}[!htbp]
-%   \centering
-%   \includegraphics[width=.4\textwidth]{static/expected_two_by_two_games_cooperation.pdf}
-%   \includegraphics[width=.4\textwidth]{static/stochastic_two_by_two_games_cooperation.pdf}
-%   \caption{\textbf{Cooperation rates for various social dilemmas.} ({\bf A}) If
-%   players update based on their expected payoffs. ({\bf B}) If players update
-%   based on their last round payoffs. We vary the temptation payoff \(T \in
-%   \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across the \(x\) axis, and \(S \in \{1.2,
-%   0.8, 0.4, 0, -0.4, -0.8\}\) across the \(y\) axis. Unless explicitly varied,
-%   the parameters of the simulation are $N\!=\!100$, $\beta\!=\!10$,
-%   $\delta\!=\!0.99$, $t\!=\!5\times
-%   10^7$.}\label{fig:cooperation_rates_two_by_two_part_one}
-% \end{figure}
-
-% So far we have explored the difference between the expected payoffs and the last
-% round payoffs. In order to explore further the effect of limited memory we allow
-% individuals to remember more. More precisely, up to two interactions and up to
-% the last two rounds. In total we present results for three more updating
-% payoffs. These are the payoffs when individuals consider the last two rounds
-% with another member of the population, the last round with two members of the
-% population, and the last two rounds with two members of the population. Similar
-% to the last round payoff, we use simulations and record which strategies the
-% players adopt over time based
-% (Appendix~\ref{appendix:further_simulation_results}) and we compare the evolving
-% cooperation rates to those of the expected payoffs
-% (Figure~\ref{fig:cooperation_rates_two_by_two_part_two}).
-
-% In the case of the last two rounds payoff we make two observations. Initially,
-% for the harmony, the stag-hunt and the prisoner's dilemma classes the average
-% cooperation rates are higher than the previous case of the last round payoff.
-% The rates remain strictly less than in the expected payoffs, however, in the
-% case of the prisoner's dilemma we observe a significant increase. Secondly, for
-% the snowdrift class there are more instances for which the cooperation rates are
-% higher than in the expected payoffs. This could indicate that in the case of the
-% snowdrift game the expected payoffs underestimate cooperation. Once players
-% interact with two other members of the population, instead of one, the expected
-% payoffs strictly yield a higher cooperation rate  even in the case of the
-% snowdrift game. Similar to the previous result, the difference between the two
-% approaches is always higher for the prisoner's dilemma class.
-
-% \begin{figure}[!htbp]
-%   \centering
-%   \includegraphics[width=.33\textwidth]{static/rounds_two_by_two_games_cooperation.pdf}
-%   \includegraphics[width=.33\textwidth]{static/opponents_two_by_two_games_cooperation.pdf}
-%   \includegraphics[width=.33\textwidth]{static/rounds_opponents_two_by_two_games_cooperation.pdf}
-%   \caption{\textbf{Cooperation rates over various social dilemmas for different
-%   limited memory payoffs.} ({\bf A}) If players update based on their last two
-%   rounds with another member of the population. ({\bf B}) If players update
-%   based on their last round with two other members of the population. ({\bf C})
-%   If players update based on their last two rounds with two other members of the
-%   population. We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8,
-%   2.2\}\) across the \(x\) axis, and \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\)
-%   across the \(y\) axis. Unless explicitly varied, the parameters of the
-%   simulation are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times
-%   10^7$..}
-%   \label{fig:cooperation_rates_two_by_two_part_two}
-% \end{figure}
-
-\section{Conclusions}\label{section:conclusions}
-
-Cooperation can be seen at odd, why is it that we choose to help others,
-increasing their payoff, at the expense of decreasing one's own? In spite of all
-the selfish genes’ animal and human communities seem to altruistically help each
-other and cooperate, and evolutionary game theory has helped us shape our
-understanding of the evolution of cooperation.
-
-Previous evolutionary models often feature a curious inconsistency. When
-modeling how individuals make decisions in each round, these models assume that
-players only remember the last round. However, when modeling how individuals
-update their strategies over time, individuals are assumed to have perfect
-memory.
-
-Here, we have explored how robust cooperation is as models deviate from the
-perfect memory assumption. Initially we considered the donation game. We showed
-that when the last round payoff is used instead of the expected payoffs,
-cooperation can even evolve when individuals only use a minimum of information,
-however, the evolving cooperation rates are typically lower. The resident
-strategies were both less cooperative and less generous. This effect was only
-intensified we increase the benefit and the strength of selection independently.
-The results showed that as each parameter was being increased the difference
-between the cooperation rates were widening. This indicates that cooperative
-players benefit from being able to interact with everyone in the population.
-
-We extended our approach and presented results not only based on different
-classes of social games, but also based on different limited memory payoffs. The
-analysis showed that specifically for the case of the prisoner's dilemma games
-the contrast between the two approaches can be significantly large. The
-prisoner's dilemma is one of the most applied types of social games when we
-discuss the evolution of cooperation. Our results indicate that cooperation
-struggles to evolve in the prisoner's dilemma when only a minimum of social
-information is used at the updating stage. Interestingly our simulations also
-showed that in same cases cooperation can benefit from limited memory. This was
-specific for the snowdrift game.
-
-\section{Acknowledgements}
-
-This work was supported by the European Research Council Starting Grant 850529:
-E-DIRECT.
-
-\appendix
-
-\section{Model Setup}\label{appendix:methods}
-
-Consider a population of \(N\) individuals where \(N\) is even. At any point in
-time there are at most two different strategies in present in the population.
-More specifically, a mutant strategy played by \(k\) individuals and a resident
-strategy played by \(N - k\) individuals. We assume a pairwise process in which
-strategies spread because they are imitated more often. Each step of the
-evolutionary process consists of two stages; a game stage and an update stage.
-
-In the game stage, each individual is randomly matched with some other
-individual in the population. Their interaction lasts for a number of turns
-which is not fixed but depends on the continuation probability \(\delta\). At
-each turn the individuals choose between cooperation (\(C\)) and defection
-(\(D\)). Thus, there are four possible outcomes in each turn \(CC, CD, DC\) and
-\(DD\). If both players cooperate they receive the reward payoff \(R\), whereas
-if both players defect they receive the punishment payoff \(P\). If one
-cooperates but the other defects, the defector receives the temptation to
-defect, \(T\), whereas the cooperator receives the sucker's payoff, \(S\). Let
-$\mathcal{U}=\{R,S,T,P\}$ denote the set of feasible payoffs in each round, and
-let $\mathbf{u}\!=\!(R,S,T,P)$ be the corresponding payoff vector. We present
-results for various values of $\mathcal{U}$ for all the symmetric \(2 \times 2\)
-games.
-
-A further assumption of our model is that individuals make use of reactive
-strategies when they make decisions in each round. Reactive strategies are a set
-of strategies that take into account only the previous action of the opponent. A
-reactive strategy can be written explicitly as a vector,
-
-\[s=(y, p, q)\]
-
-where \(y\) is the probability that the strategy opens with a cooperation and
-\(p, q\) are the probabilities that the strategy cooperates given that the
-opponent cooperated and defected equivalently.
-
-In the updating stage, two players are randomly drawn from the population, a
-`learner' and a `role model'. The learner adopts the role model's strategy based
-on the Fermi distribution function, %ToDo add reference
-
-\begin{equation} \label{Eq:rho}
-\rho(u_{L}, u_{RM}) = \frac{1}{1\!+\! \exp^{\!-\!\beta (u_{RM}\!-\!u_{L})}}.
-\end{equation}
-
-where $u_{L}\!\in\! \mathcal{U}$ is the learner's payoff, $u_{RM}\!\in\!
-\mathcal{U}$ is the role model's payoff, and $\beta\!\ge\!0$ is the strength of
-selection.
-
-We iterate this basic evolutionary step until either the mutant strategy goes
-extinct, or until it fixes in the population and becomes the new resident
-strategy. After either outcome, we set $k$ to $1$ and we introduce a new mutant
-strategy which is uniformly chosen from all reactive strategies at random.
-Instead of simulating each step of the evolutionary process, we estimate the
-probability that a newly introduced mutant fixes~\cite{nowak2004emergence}. This
-is defined as the fixation probability of the mutant, and the standard form is
-the following,
+More specifically, the process again starts with a population where all members are of the same
+strategy. At each step one individual adopts a mutant strategy randomly selected
+from the set of feasible strategies. The fixation probability \(\phi_{M}\) of the
+mutant strategy can be calculated explicitly,
 
@@
 where \(\lambda^-_k, \lambda^+_k\) are the probabilities that the number of
-mutants decreases and increases respectively.
-
-This process of mutation and fixation/extinction is iterated many times. The
-evolutionary process is summarized by
-Algorithm~\ref{algorithm:pairwise_comparison}.
+mutants decreases and increases respectively, \(N\) is the size of the
+population, and \(k\) is the number of mutants. The probabilities \(\lambda^-_k
+\text{ and } \lambda^+_k\) depend on the updating payoffs of the mutant and the
+resident strategies. Depending on the fixation probability \(\phi_{M}\) the
+mutant either fixes (becomes the new resident) or goes extinct. Regardless, in
+the elementary time step another mutant strategy is introduced to the
+population. We iterate this elementary population updating process for a large
+number of mutant strategies and we record the resident strategies at each time
+step. The process is summarised by Algorithm~\ref{algorithm:pairwise_comparison}.
 
@@
 
-The aim of this work is to explore the effect of updating memory on the
-cooperation rate of the evolved population. For this reason we consider two
-different approaches when estimating the payoffs at the updating stage. The two
-approaches we consider are those of (i) the expected and (ii) the limited memory
-payoffs.
+Most of the results we present in this work consider the case of low mutation,
+however, we have also verified that the main result holds even in the case where
+\(\mu\) is not small. %ToDo Add reference to section with mu != 0.
 
-\subsection*{Expected Payoffs}
+\section{Updating Payoffs}\label{section:updating_payoffs}
 
-The expected payoffs are the conventional payoffs used in the updating
-stage~\cite{imhof2010stochastic}. They are defined as the mean payoff of an
-individual in a well-mixed population that engages in repeated games with all
-other population members.
+In this section we discuss the conventional approach of calculating the updating
+payoffs, as well as our newly introduced approach.
 
-We first define the payoff of two reactive strategies at the game stage. Assume
-two reactive strategies $s_1\!=\!(y_1, p_1, q_1$) and $s_2\!=\!(y_2,p_2,q_2)$.
-It is not necessary to simulate the play move by move, instead the play between
-the two strategies is defined a Markov matrix \(M\),
+\subsection{Updating Payoffs based on the expected payoffs}
+
+The expected payoffs are the conventional payoffs used in the updating stage.
+The expected payoffs are defined as the mean payoff of an individual in a
+well-mixed population that engages in repeated games with all other population
+members. We first define the payoff of two reactive strategies at the game
+stage. Assume two reactive strategies $s_1\!=\!(y_1, p_1, q_1$) and
+$s_2\!=\!(y_2,p_2,q_2)$. The play between the two strategies can be defined as a
+Markov process with the transition matrix \(M\),
 
 \begin{equation}\label{eq:transition_matrix}
-  \input{static/matrix.tex}.
+  \input{../static/matrix.tex}.
 \end{equation}
@@
 the game stage outcome for each strategy,
-\(\langle\mathbf{v}(s_1,s_2),\mathbf{u}\rangle\)~\cite{Hauert1997}.
 
-In the updating stage the learner adopts the strategy of the role model based on
-their updating payoffs. Given that there are only two different types in the
-population at each time step we only need to define the expected payoff for a
-resident (\(\pi_R\)) and for a mutant (\(\pi_M\)). Assume the resident strategy
-\(s_R = (y_R, p_R, q_R)\) and the mutant strategy \(s_M = (y_M, p_M, q_M)\), the
-expected payoffs are give by,
+\[\langle\mathbf{v}(s_1,s_2),\mathbf{u}\rangle \text{ and } \langle\mathbf{v}(s_2,s_1), \mathbf{u}\rangle\].
+
+In the case of low mutation there can be only two different
+strategies in the population; a resident and a mutant strategy. Thus, we
+only need to define the expected
+payoff for a resident (\(\pi_R\)) and for a mutant (\(\pi_M\)). Assume the
+resident strategy \(s_R = (y_R, p_R, q_R)\) and the mutant strategy \(s_M =
+(y_M, p_M, q_M)\). The expected payoffs are give by,
 
@@
   \begin{array}{lcrcr}
-  \displaystyle \pi_R	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(s_R,s_R),\mathbf{u}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(s_R,s_M),\mathbf{u}\rangle,\\[0.5cm]
-  \displaystyle \pi_M	&=	&\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(s_M,s_R),\mathbf{u}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(s_M,s_M),\mathbf{u}\rangle.\\
+  \displaystyle \pi_R & = &\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(s_R,s_R),\mathbf{u}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(s_R,s_M),\mathbf{u}\rangle,\\[0.5cm]
+  \displaystyle \pi_M & = &\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(s_M,s_R),\mathbf{u}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(s_M,s_M),\mathbf{u}\rangle.\\
   \end{array}
@@
 
-The number of mutant in the population increase if a learner resident adopts the
-strategy of a mutant role model, and decreases if a mutant leaner adopts the
-strategy of a resident. The probabilities that the number of mutants decreases
-and increases, \(\lambda^-_k\) and \(\lambda^+_k\), are now explicitly defined
-as,
+As a reminder \(N\) is the population size and \(k\) the number of mutants. The
+number of mutants in the population increases if a resident adopts the strategy
+of a mutant, and decreases if a mutant adopts the strategy of a resident. The
+probabilities that the number of mutants decreases and increases,
+\(\lambda^-_k\) and \(\lambda^+_k\), are now explicitly defined as,
 
 \begin{align*} 
-  \lambda^-_k &\!=\!\rho(\pi_R, \pi_M) \\
-  \lambda^+_k &\!=\!\rho(\pi_M, \pi_R).
+  \lambda^-_k &\!=\!\rho(\pi_M, \pi_R) \\
+  \lambda^+_k &\!=\!\rho(\pi_R, \pi_M).
 \end{align*}
 
-\subsection*{Limited memory payoffs}
+\subsection{Invasion analysis of ALLD into Generous Tit For Tat (GTFT)}
 
-Initially, we discuss the case of the \textbf{last round updating payoff}. At
-the stage game we define the payoff of a reactive strategy in the last round,
+In the following, we apply the above formalism to calculate how easily a single
+ALLD mutant can invade into a resident population with strategy GTFT. In that
+case, \(s_1 = (1, 1, q)\), \(s_2 = (0, 0, 0)\), and \(k = 1\). When two GTFT
+players interact in the game, their respective probabilities for each of the
+four outcomes in the last round simplify to,
+
+\begin{align*}
+    \mathbf{v}(s_1, s_1) = (1, 0, 0, 0).
+\end{align*}
+
+On the other hand, if an ALLD player interacts with a GTFT player, the
+respective probabilities become,
+
+\begin{align*}
+  \mathbf{v}(s_2, s_1) = (0, q, 0, (1 - q)).
+\end{align*}
+
+Using the above we can define the payoffs of a GTFT individual (resident)
+and of the ALLD individual (mutant) follows,
+
+\begin{align*}
+  \displaystyle \pi_R & = \displaystyle \frac{N\!-\!2}{N-1}\cdot (b - c) \cdot -	\displaystyle\frac{1}{N-1}\cdot q \cdot c\text{ and } \displaystyle \pi_M  = \displaystyle b \cdot q.
+\end{align*}
+
+As a consequence, we can calculate the ratio of transition probabilities as,
+
+\begin{align*}
+    \frac{\lambda^{+}}{\lambda^{-}} & = \frac{\rho(\pi_R, \pi_M)}{\rho(\pi_M, \pi_R)} \\[0.5cm]
+    \frac{\lambda^{+}}{\lambda^{-}} & = \frac{e^{-\beta \left(\frac{(N-2) (b-c)}{N-1} - q \cdot (b + \frac{c}{N-1})\right)}+1}
+    {e^{-\beta\left(b \cdot q - \frac{b (N - 2)}{N-1} - \frac{c (N - 2 - q)}{N-1}\right)}+1}
+\end{align*}
+
+In particular, in the limit of strong selection \(\beta \rightarrow \infty\)
+and large populations \(N \rightarrow \infty\), we obtain that the ratio is than
+smaller to 1 if \(q \leq 1 - \frac{c}{b}\). Thus, ALLD is disfavored to invade
+if \(q \leq 1 - \frac{c}{b}\). For \(q = 1 -\frac{c}{b}\) the probability that
+the number of mutants increase by one equals the probability that the mutant
+goes extinct.
+
+\subsection{Simulation results}\label{section:simulation_results_exprected}
+
+Figure~\ref{fig:expected_payoffs_results} shows simulation results of the
+evolutionary process when individuals update their strategies based on their
+expected payoffs. We run two simulations where we differ the benefit of
+cooperation. Though, a higher benefit results in a more cooperative population,
+the population is still rather cooperative even in the case of low benefit.
+
+In the case of low benefit the resident population consists either of defectors
+or conditional cooperators, whereas in the case of high benefit the residents
+typically apply a conditional cooperative strategy. Based on the results on
+invasion we can show that conditional cooperators that the residents adopt are
+of the kind \(p \approx 1\) and \(q \leq 1 - \frac{c}{b}\).
+
+\begin{figure}[!htbp]
+    \centering 
+    \includegraphics[width=.70\textwidth]{../static/expected_payoffs_donation_game.pdf}
+    \caption{\textbf{Evolutionary dynamics under expected payoffs with
+    low (left) and high (right) benefit.} We run the simulations for \(T =
+    10^7\) time steps. For each time step, we have recorded the current resident
+    population \((y, p, q)\). Since in the case of the expected payoffs the
+    individuals interact for an infinite number of times \(\delta \rightarrow
+    1\), we do not report the players' initial cooperation probability \(y\).
+    The graphs show how often the resident population chooses each combination
+    \((p, q)\) of conditional cooperation probabilities in the subsequent
+    rounds. In both cases players update based on their expected payoffs (A) In
+    the case where \(b=3\), the resident population either consists of defectors
+    (with \(p \approx q \approx 0\)) or of conditional cooperators for which \(p
+    \approx 1\) and \(q \leq 1 - \frac{1}{3}=0.7\). (B) In the case where
+    \(b=10\), the resident population typically applies a strategy for which \(p
+    \approx 1\) and \(q \leq 1 - \frac{1}{10}=0.9\). Parameters: \(N =100, c=1,
+    \beta=1\).
+    }\label{fig:expected_payoffs_results}
+\end{figure}
+
+\subsection{Updating Payoffs based on the one interaction - last round payoffs}
+
+The first approach we introduced in this work is that of the one  interaction -
+last round updating payoff. In this case an individual updates his strategy
+based on the last payoff they received against the one other member of the
+population. Initially, we define the payoff of a reactive strategy in the last
+round at the stage game. This is given by
 Proposition~\ref{proposition:last_round}.
@@
 \end{Prop}
+
+Note that in the proposition we use the general notation of the prisoner's
+dilemma, to denote that the results apply to \(2 \times 2\) any symmetric game.
 
@@
   \centering
-  \includegraphics[width=.65\textwidth]{static/last_rounds_matches.tex}
+  \includegraphics[width=.65\textwidth]{../static/last_rounds_matches.tex}
   \caption{\textbf{Possible pairings combination in the updating stage, given
@@
 pairs can occur that are feasible in a direct interaction,
-$\splitatcommas{(u_1,u_2)\in \mathcal{U}^2_F:=\big\{ (R,R), (S,T), (T,S), (P,P)
+$\splitatcommas{(u_1,u_2)\in \mathcal{U}^2_F:=\big\{ (b-c, b-c), (-c, b), (b, -c), (0, 0)
 \big\}}$, as represented by the respective indicator function. Otherwise, if the
@@
 the mutant obtained a payoff of $u_M$.
+
+\subsubsection{Invasion analysis of ALLD into (GTFT)}
+
+Similarly to the previous section, we calculate how easily a single ALLD mutant
+can invade into a resident population with strategy GTFT.
+
+When two GTFT players interact in the game, their respective probabilities for
+each of the four outcomes in the last round simplify to,
+
+\begin{align*}
+    v_R(GTFT,GTFT) = 1, & \quad v_T(GTFT,GTFT) = 0, \\
+    v_S(GTFT,GTFT) = 0, & \quad v_P(GTFT,GTFT) = 0.
+\end{align*}
+
+On the other hand, if an ALLD player interacts with a GTFT player, the
+respective probabilities according to Eq.~\ref{Eq:LastRound} become
+
+\begin{align*}
+    v_R(ALLD,GTFT) & = 0, &  v_S(ALLD,GTFT) & = 0, \\
+    v_T (ALLD, GT F T ) & = 1 - \delta + \delta q, &  v_P (ALLD, GTFT) & = \delta(1 - q).
+\end{align*}
+
+As a consequence, we obtain the following probabilities \(x(u_1, u_2)\) that the
+payoff of a randomly chosen GTFT player is \(u_1\) and that the payoff of the
+ALLD player is \(u_2\),
+
+\begin{align*}
+  x(R, T) = & \frac{N - 2}{N - 1} \cdot (1 - \delta + \delta q)\\
+  x(R, R) = & \frac{N - 2}{N - 1} \cdot \delta (1 - q) \\
+  x(S, T) = & \frac{1}{N - 1} \cdot (1 - \delta + \delta q) \\
+  x(P, P) = & \frac{1}{N - 1} \cdot \delta (1 - q)
+\end{align*}
+
+As a consequence, we can calculate the ratio of transition probabilities as
+
+\begin{equation*}
+  \resizebox{.9\textwidth}{!} 
+  {$
+    \frac{\lambda^{+}}{\lambda^{-}} = \frac{\frac{N - 2}{N - 1} \cdot \left(\frac{\delta(1-q)}{e^{-\beta(c - b)} + 1} + \frac{-\delta +\delta  q + 1}{e^{-\beta c} + 1}\right) + \frac{1}{N - 1} \cdot \left(\frac{-\delta +\delta  q+1}{e^{\beta  (-(b+c))}+1}\right)}
+    {\frac{N - 2}{N - 1} \cdot \left(\frac{\delta (1-q)}{e^{-\beta (b-c)}+1} + \frac{-\delta +\delta  q + 1}{e^{\beta  c}+1}\right)+ \frac{1}{N-1} \cdot \left(\frac{-\delta +\delta  q+1}{e^{-\beta  ((-b-c))}+1}\right)}
+  $}
+\end{equation*}
+
+In particular, in the limit of strong selection \(\beta \rightarrow \infty\) and
+large populations \(N \rightarrow \infty \), we obtain
+
+\begin{equation*}
+    \frac{\lambda^{+}}{\lambda^{-}} = \frac{1 - \delta + \delta q}{\delta(1 - q)}
+\end{equation*}
+
+This ratio is smaller than 1 (such that ALLD is disfavored to invade) if \(q <
+1- 1/(2 \delta)\). For infinitely repeated games, \(\delta \rightarrow 1\), this
+condition becomes \(q < 1/2\) (for \(q = 1/2\), the payoff of the ALLD player is
+\(T > R\) for half of the time, and it is \(P < R\) for the other half. The
+probability that the number of mutants increase by one equals the probability
+that the mutant goes extinct).
+
+\subsubsection{Simulation results}
+
+In a similar fashion to Section~\ref{section:simulation_results_exprected}, we
+simulate the evolutionary process given that individuals now use the last
+interaction-last round payoffs to update their strategies.
+Figure~\ref{fig:one_interaction_last_round_payoffs_results} shows simulation
+results. Though, a higher benefit results in a more cooperative population, in
+the case of these updating payoff, only to a slightly more cooperative. As shown
+in Figure~\ref{fig:one_interaction_last_round_payoffs_results} the evolving
+populations are more similar, both consisting of defectors and conditional
+cooperators. Compared to the expected payoffs, the conditional cooperators are
+also less cooperative \(p \approx 1\) and \(q \leq \frac{1}{2}\). The increased
+benefit pushes the population closer to the boundary of \(q \leq \frac{1}{2}\)
+but that is the maximum.
+
+\begin{figure}[!htbp]
+  \centering 
+  \includegraphics[width=.70\textwidth]{../static/one_interaction_last_round_payoffs_results}
+  \caption{\textbf{Evolutionary dynamics under one interaction last round
+  payoffs with low (left) and high (right) benefit.} In the cases of low and
+  high benefit case the resident population either consists of defectors (with
+  \(p \approx q \approx 0\)) or of conditional cooperators for which \(p \approx
+  1\) and \(q \leq \frac{1}{2}\). Parameters: \(N =100, c=1, \beta=1\).
+  }\label{fig:one_interaction_last_round_payoffs_results}
+\end{figure}
+
+\subsection{Updating Payoffs based on \(n\) interactions - last \(m\) round payoffs}
+
+So far we have presented the two extreme cases; the case where an individual
+updates based one interaction and one round, and the case  where an individual
+updates based on all interactions and all rounds. In this work we also
+explore the intermediate cases (Figure~\ref{fig:cases_updating_payoffs}). Namely the cases:
+
+\begin{itemize}
+  \item last round \(n=1\) with two other players \(m=2\)
+  \item last two rounds \(n=2\) with one other player  \(m=1\)
+  \item last two rounds \(n=2\) with two other players \(m=2\)
+\end{itemize}
+
+\begin{figure}[!htbp]
+  \centering
+  \includestandalone[width=.5\textwidth]{cases}
+  \caption{\textbf{Updating payoffs used in this work.} \(n\) is the
+  number of turns and \(m\) th number of individuals. The expected payoffs
+  are the one commonly used in the literature. We have introduced a method
+  for the rest.}\label{fig:cases_updating_payoffs}
+\end{figure}
+
+We define the payoffs on the two last round, and in the updating phase the
+probability that.
 
@@
   \centering
-  \includegraphics[width=.65\textwidth]{static/tree.tex}
+  \includegraphics[width=.65\textwidth]{../static/tree.tex}
   \caption{\textbf{Possible pairings combination in the updating stage, given
@@
 
-Given the last two rounds payoff and possible pair combinations with two
-members, we can define the probability \(x\) that the respective last round
-payoffs of two players \(s_1, s_2\) are given by $u_1$ and $u_2$ similarly to
-Eq.~(\ref{eq:Chi}). We follow the same approach to define the rest of the
-updating payoffs. These are the updating payoffs of the last round with two
-member, and the the last two rounds payoffs with one member.
-
-Simulating the evolutionary process for more interactions and rounds quickly
-becomes computationally intractable. Our methodology could be extended to
-include \(n\) turns and \(m\) interactions. However, for the purpose of this
-work we explore the cases only up to two turns and two interactions.
-
-% \section{Evolutionary dynamics under limited memory payoffs for various social
-% dilemmas}\label{appendix:further_simulation_results}
-
-% \begin{figure}[!htbp]
-%   \centering
-%   \includegraphics[width=\textwidth]{static/rounds_two_by_two_games.pdf}
-%   \caption{{\bf Evolutionary dynamics under last two rounds payoffs for various social dilemmas.} 
-%   We have run several simulations of the evolutionary process described in
-%   section~\ref{section:model} for $t\!=\!10^7$ time steps. The graphs show how
-%   often the resident population chooses each combination ($p,q$) of conditional
-%   cooperation probabilities in the subsequent rounds. We vary the temptation
-%   payoff \(T \in \{-1, -0.6, -0.2,  0.2, 0.6, 1, 1.4, 1.8, 2.2, 2.6, 3\}\)
-%   across the \(x\) axis, and  \(S \in \{2, 1.6, 1.2, 0.8, 0.4, 0, -0.4, -0.8,
-%   -1.2, -1.6, -2\}\) across the \(y\) axis. Parameters: $N\!=\!100$,
-%   $\beta\!=\!10$, $\delta\!=\!0.999$.}
-%   \label{fig:last_two_rounds_two_by_two}
-% \end{figure}
-
-% \begin{figure}[!htbp]
-%   \centering
-%   \includegraphics[width=\textwidth]{static/opponents_two_by_two_games.pdf}
-%   \caption{{\bf Evolutionary dynamics under last round payoffs with two members of the population for various social dilemmas.} 
-%   We have run several simulations of the evolutionary process described in
-%   section~\ref{section:model} for $t\!=\!10^7$ time steps. The graphs show how
-%   often the resident population chooses each combination ($p,q$) of conditional
-%   cooperation probabilities in the subsequent rounds. We vary the temptation
-%   payoff \(T \in \{-1, -0.6, -0.2,  0.2, 0.6, 1, 1.4, 1.8, 2.2, 2.6, 3\}\)
-%   across the \(x\) axis, and  \(S \in \{2, 1.6, 1.2, 0.8, 0.4, 0, -0.4, -0.8,
-%   -1.2, -1.6, -2\}\) across the \(y\) axis. Parameters: $N\!=\!100$,
-%   $\beta\!=\!10$, $\delta\!=\!0.999$.}
-%   \label{fig:last_two_opponents_two_by_two}
-% \end{figure}
-
-% \begin{figure}[!htbp]
-%   \centering
-%   \includegraphics[width=\textwidth]{static/merged_plot_rounds_opponents_two.pdf}
-%   \caption{{\bf Evolutionary dynamics under last two rounds payoffs with two members of the population for various social dilemmas.} 
-%   We have run several simulations of the evolutionary process described in
-%   section~\ref{section:model} for $t\!=\!10^7$ time steps. The graphs show how
-%   often the resident population chooses each combination ($p,q$) of conditional
-%   cooperation probabilities in the subsequent rounds. We vary the temptation
-%   payoff \(T \in \{-1, -0.6, -0.2,  0.2, 0.6, 1, 1.4, 1.8, 2.2, 2.6, 3\}\)
-%   across the \(x\) axis, and  \(S \in \{2, 1.6, 1.2, 0.8, 0.4, 0, -0.4, -0.8,
-%   -1.2, -1.6, -2\}\) across the \(y\) axis. Parameters: $N\!=\!100$,
-%   $\beta\!=\!10$, $\delta\!=\!0.999$.}
-%   \label{fig:last_two_rounds_two_opponents}
-% \end{figure}
-
-\bibliographystyle{unsrt}
-\bibliography{bibliography.bib}
+For the above cases we use combinations of Eq. 1 2 and 3.
 

### 2022-08-26 clean_data.py

@@ -8,7 +8,7 @@ import pandas as pd
 
 if __name__ == "__main__":  # pragma: no cover
     folder = sys.argv[1]
-    starting_characters = sys.argv[2]
+    starting_characters = "two_rounds_new" #sys.argv[2]
 
     in_columns = [
         "y",
@@ -30,7 +30,7 @@ if __name__ == "__main__":  # pragma: no cover
     #     "t",
     # ]
 
-    print(f"{folder}/{starting_characters}*.csv")
+    print(f"{folder}/{starting_characters}.csv")
     csv_files = glob.glob(f"{folder}/{starting_characters}*.csv")
 
     for file in tqdm.tqdm(csv_files):
@@ -43,8 +43,8 @@ if __name__ == "__main__":  # pragma: no cover
         for value in df["t"][1:]:
             frequencies.append(value - previous)
             previous = value
-        frequencies.append(50000000 - previous)
+        frequencies.append(10 ** 7 - previous)
 
         df["frequencies"] = frequencies
-
-        df.to_csv(f"{folder}clean_{file[len(folder):]}", index=False)
+        print(f"{folder}/clean_{starting_characters}")
+        df.to_csv(f"{folder}/clean_{starting_characters}.csv", index=False)

### 2022-08-26 matlab/calcRhoSixteen.m

@@ -1,0 +1,19 @@
+function [Rho]=calcRhoSixteen(u, beta);
+%% Calculates all possible pairwise imitation probabilities based on the last two payoffs
+    Us = zeros(16, 16);
+
+    for i=1:16
+        for j=1:16
+            Us(i, j) = (u(1 + fix((j - 1) / 4)) + u(1 + mod(j - 1, 4))) - (u(1 + fix((i - 1) / 4)) + u(1 + mod(i - 1, 4)));
+        end
+    end
+
+    Us = Us / 2;
+    Rho = zeros(16, 16);
+
+    for i=1:16
+        for j=1:16
+            Rho(i, j) = 1 / (1 + exp(-beta * Us(i, j)));
+        end
+    end
+end

### 2022-08-26 matlab/evolRunRoundsTwo.m

@@ -1,0 +1,15 @@
+function evolRunRoundsTwo();
+
+starting_resident = [0, 0, 0];
+N = 100;
+delta = 0.999;
+beta = 1;
+numberIterations = 10 ^ 6;
+payoff_type = "two_rounds";
+u = [2, -1, 3, 0];
+
+filename = "../data/two_rounds_new.csv";
+
+evolSimulationTwoRounds(starting_resident, u, N, delta, beta, numberIterations, payoff_type, filename);
+
+end

### 2022-08-26 matlab/stationaryRoundTwo.m

@@ -15,5 +15,5 @@ for i=1:4
     end
 end
 
-v = reshape(v, [1, 16]);
+v = reshape(v.',[1, 16]);
 end

### 2023-02-21 scripts/clean_data.py
create folder scripts

@@ -8,14 +8,14 @@ import pandas as pd
 
 if __name__ == "__main__":  # pragma: no cover
     folder = sys.argv[1]
-    starting_characters = "two_rounds_new" #sys.argv[2]
+    starting_characters = "two_rounds_opponentslow_benefit_seed" #sys.argv[2]
 
     in_columns = [
         "y",
         "p",
         "q",
         "average cooperation rate",
-        "average score",
+        # "average score",
         "t",
     ]
 
@@ -43,8 +43,9 @@ if __name__ == "__main__":  # pragma: no cover
         for value in df["t"][1:]:
             frequencies.append(value - previous)
             previous = value
-        frequencies.append(10 ** 7 - previous)
+        frequencies.append(10 ** 6 - previous)
 
+        filename = file.split("/")[-1]
         df["frequencies"] = frequencies
-        print(f"{folder}/clean_{starting_characters}")
-        df.to_csv(f"{folder}/clean_{starting_characters}.csv", index=False)
+        print(f"{folder}/clean_{filename}")
+        df.to_csv(f"{folder}/clean_{filename}", index=False)

### 2023-02-21 main.tex
working on paper in folder paper

@@
     \centering
-    \includegraphics[width=.70\textwidth]{static/expected_and_stochastic_for_donation_game.pdf}
+    \includegraphics[width=1.5\textwidth]{static/donation_expected_last_round_summary_results.pdf}
     % \includegraphics[width=.70\textwidth]{static/expected_and_stochastic_for_donation_game_b_10.pdf}
@@
 
-We also investigate the effect of the strength of selection $\beta$.
-Figure~\ref{fig:cooperation_rate_over_betas} illustrates results for various
-runs of the evolutionary process. For weak selection, \(\beta < 1\), we observe
-that the two methods yield similar results, however, as \(\beta\) increases
-there is variation in the evolving populations. In the case of expected payoffs
-the resident populations become more cooperative as \(\beta\) increases, whereas
-in the case of limited memory payoffs, the resident populations become more
-defective.
-
-\begin{figure}[!htbp]
-  \centering
-    \includegraphics[width=\textwidth]{static/cooperation_rate_over_b.pdf}
-\caption{{\bf The evolution of cooperation for different selection strength values.}
-We vary the selection strength $\beta$. In all cases, stochastic payoff
-evaluation tends to reduce the evolving cooperation rates. ({\bf A}) the
-probabilities \(p, q\) for resident population over \(10^7\) time steps for each
-\(\beta\) value. ({\bf B}) The cooperation rate within the resident population
-(averaged over all games and over all time steps) over \(\beta\). Unless
-explicitly varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$,
-$c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
-$t\!=\!5\times 10^7$ time steps for each parameter
-combination.}\label{fig:cooperation_rate_over_betas}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \centering
-    \includegraphics[width=\textwidth]{static/more_memory.pdf}
+% We also investigate the effect of the strength of selection $\beta$.
+% Figure~\ref{fig:cooperation_rate_over_betas} illustrates results for various
+% runs of the evolutionary process. For weak selection, \(\beta < 1\), we observe
+% that the two methods yield similar results, however, as \(\beta\) increases
+% there is variation in the evolving populations. In the case of expected payoffs
+% the resident populations become more cooperative as \(\beta\) increases, whereas
+% in the case of limited memory payoffs, the resident populations become more
+% defective.
+
+% \begin{figure}[!htbp]
+%   \centering
+%     \includegraphics[width=\textwidth]{static/cooperation_rate_over_b.pdf}
 % \caption{{\bf The evolution of cooperation for different selection strength values.}
@@
 % combination.}\label{fig:cooperation_rate_over_betas}
-\end{figure}
-
-% \subsection{Effect of updating payoffs in different social
-% dilemmas}\label{section:2_by_2_games}
-
-% In the previous section we gained insights into the effects of the updating
-% payoffs, and into how parameters such as the benefit and the strength of
-% selection can intensify them. We investigated these effects by using the
-% donation game. In order to broaden our understanding of the updating payoffs on
-% different forms of possible human interactions we extend our approach to all \(2
-% \times 2\) symmetric games. More specifically, we apply our analysis to four
-% different classes of games; the harmony, the snowdrift, the prisoner's dilemma
-% and the stag-hunt games.
-
-% We compare the results of the evolutionary process when the updating payoffs are
-% based on the expected payoffs, and on the last round payoffs for all the four
-% possible classes of games (Figures~\ref{fig:expected_two_by_two}
-% and~\ref{fig:last_round_two_by_two}). In Figure~\ref{fig:expected_two_by_two}
-% each sub figure represents a run of evolutionary process for a different set of
-% values for \(\mathcal{U}\). Without loss of generality we set \(R=1\) and
-% \(P=0\)~\cite{Martinez2012, Roca2009}, and we vary the values of the temptation
-% payoff \(T\) (across the \(x\)-axis) and of the sucker's payoff \(S\) (across
-% the \(y\)-axis). Starting at the upper left corner and proceeding clockwise the
-% quadrants correspond to the harmony, the snowdrift, the prisoner's dilemma and
-% the stag-hunt games.
+% \end{figure}
 
@@
 %   \centering
-%   \includegraphics[width=\textwidth]{static/expected_two_by_two_games.pdf}
-%   \caption{{\bf Evolutionary dynamics under expected payoffs for various social dilemmas.}
-%   We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across
-%   the \(x\) axis, and  \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\) across the
-%   \(y\) axis. ({\bf A}) The top left quadrant where \(S > 0\) and \(T \leq 1\)
-%   corresponds to the harmony game. The preference ordering for the harmony game
-%   is \(R > T > S > P\). ({\bf B}) The top right quadrant where \(S > 0\) and \(T
-%   > 1\) corresponds to the snowdrift game. The preference ordering for the
-%   snowdrift game is \(T > R > S > P\). ({\bf C}) The bottom left quadrant where
-%   \(S < 0\) and \(T \leq 1\) corresponds to the stag-hunt game. The preference
-%   ordering for the stag-hunt game is \(R > T > P > S\). ({\bf D}) The bottom
-%   right quadrant where \(S < 0\) and \(T > 1\) corresponds to the prisoner's
-%   dilemma game. The preference ordering for the prisoner's dilemma game is \(T >
-%   R > P > S\). Unless explicitly varied, the parameters of the simulation are
-%   $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}
-%   \label{fig:expected_two_by_two}
+%     \includegraphics[width=\textwidth]{static/more_memory.pdf}
+% % \caption{{\bf The evolution of cooperation for different selection strength values.}
+% % We vary the selection strength $\beta$. In all cases, stochastic payoff
+% % evaluation tends to reduce the evolving cooperation rates. ({\bf A}) the
+% % probabilities \(p, q\) for resident population over \(10^7\) time steps for each
+% % \(\beta\) value. ({\bf B}) The cooperation rate within the resident population
+% % (averaged over all games and over all time steps) over \(\beta\). Unless
+% % explicitly varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$,
+% % $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
+% % $t\!=\!5\times 10^7$ time steps for each parameter
+% % combination.}\label{fig:cooperation_rate_over_betas}
 % \end{figure}
+
+% % \subsection{Effect of updating payoffs in different social
+% % dilemmas}\label{section:2_by_2_games}
+
+% % In the previous section we gained insights into the effects of the updating
+% % payoffs, and into how parameters such as the benefit and the strength of
+% % selection can intensify them. We investigated these effects by using the
+% % donation game. In order to broaden our understanding of the updating payoffs on
+% % different forms of possible human interactions we extend our approach to all \(2
+% % \times 2\) symmetric games. More specifically, we apply our analysis to four
+% % different classes of games; the harmony, the snowdrift, the prisoner's dilemma
+% % and the stag-hunt games.
+
+% % We compare the results of the evolutionary process when the updating payoffs are
+% % based on the expected payoffs, and on the last round payoffs for all the four
+% % possible classes of games (Figures~\ref{fig:expected_two_by_two}
+% % and~\ref{fig:last_round_two_by_two}). In Figure~\ref{fig:expected_two_by_two}
+% % each sub figure represents a run of evolutionary process for a different set of
+% % values for \(\mathcal{U}\). Without loss of generality we set \(R=1\) and
+% % \(P=0\)~\cite{Martinez2012, Roca2009}, and we vary the values of the temptation
+% % payoff \(T\) (across the \(x\)-axis) and of the sucker's payoff \(S\) (across
+% % the \(y\)-axis). Starting at the upper left corner and proceeding clockwise the
+% % quadrants correspond to the harmony, the snowdrift, the prisoner's dilemma and
+% % the stag-hunt games.
+
+% % \begin{figure}[!htbp]
+% %   \centering
+% %   \includegraphics[width=\textwidth]{static/expected_two_by_two_games.pdf}
+% %   \caption{{\bf Evolutionary dynamics under expected payoffs for various social dilemmas.}
+% %   We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across
+% %   the \(x\) axis, and  \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\) across the
+% %   \(y\) axis. ({\bf A}) The top left quadrant where \(S > 0\) and \(T \leq 1\)
+% %   corresponds to the harmony game. The preference ordering for the harmony game
+% %   is \(R > T > S > P\). ({\bf B}) The top right quadrant where \(S > 0\) and \(T
+% %   > 1\) corresponds to the snowdrift game. The preference ordering for the
+% %   snowdrift game is \(T > R > S > P\). ({\bf C}) The bottom left quadrant where
+% %   \(S < 0\) and \(T \leq 1\) corresponds to the stag-hunt game. The preference
+% %   ordering for the stag-hunt game is \(R > T > P > S\). ({\bf D}) The bottom
+% %   right quadrant where \(S < 0\) and \(T > 1\) corresponds to the prisoner's
+% %   dilemma game. The preference ordering for the prisoner's dilemma game is \(T >
+% %   R > P > S\). Unless explicitly varied, the parameters of the simulation are
+% %   $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}
+% %   \label{fig:expected_two_by_two}
+% % \end{figure}
+
+% % \begin{figure}[!htbp]
+% %   \centering
+% %   \includegraphics[width=\textwidth]{static/stochastic_two_by_two_games.pdf}
+% %   \caption{{\bf Evolutionary dynamics under last round payoffs for various social dilemmas.} 
+% %   We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across
+% %   the \(x\) axis, and  \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\) across the
+% %   \(y\) axis. ({\bf A}) The top left quadrant where \(S > 0\) and \(T \leq 1\)
+% %   corresponds to the harmony game. The preference ordering for the harmony game
+% %   is \(R > T > S > P\). ({\bf B}) The top right quadrant where \(S > 0\) and \(T
+% %   > 1\) corresponds to the snowdrift game. The preference ordering for the
+% %   snowdrift game is \(T > R > S > P\). ({\bf C}) The bottom left quadrant where
+% %   \(S < 0\) and \(T \leq 1\) corresponds to the stag-hunt game. The preference
+% %   ordering for the stag-hunt game is \(R > T > P > S\). ({\bf D}) The bottom
+% %   right quadrant where \(S < 0\) and \(T > 1\) corresponds to the prisoner's
+% %   dilemma game. The preference ordering for the prisoner's dilemma game is \(T >
+% %   R > P > S\). Unless explicitly varied, the parameters of the simulation are
+% %   $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}
+% %   \label{fig:last_round_two_by_two}
+% % \end{figure}
+
+% % The harmony game describes an interaction without conflict. In the harmony game
+% % it is in the best interest for both players to cooperate, and in an evolutionary
+% % process cooperators prevail. We observe that for most harmony game runs the
+% % resident population overwhelmingly applies a strategy for which \(p\) and \(q\)
+% % are \(\approx 1\) (Figure~\ref{fig:expected_two_by_two}). The snowdrift game
+% % describes a situation similar to that of the prisoner's dilemma; cooperation
+% % results in a benefit to the opposing player but entailing a cost to the
+% % cooperator. However, in the snowdrift game individuals obtain immediate direct
+% % benefits from the cooperative acts which leads to \(S > P\). In the snowdrift
+% % game more cooperation emerges compared to the prisoner's dilemma. Defection is
+% % never a resident strategy and for the same values of temptation the overall
+% % \(q-\) values are higher. The last class of games we present are for the
+% % stag-hunt game. In the stag-hunt game there is an equilibrium in which both
+% % players cooperate as well as one in which both defect. For small values of the
+% % temptation and the sucker's payoffs we observe that cooperation prevails similar
+% % to the harmony game. However, in the case where the temptation and the reward
+% % payoffs are equal the resident population either consists of defectors or
+% % cooperators.
+
+% % Although the patterns of the evolved populations remain similar if we consider
+% % that players update their strategies based on their last round payoff
+% % (Figure~\ref{fig:last_round_two_by_two}), there are still dissimilarities
+% % between the two approaches. To better understand these dissimilarities, we
+% % explore the cooperation rates for each of the cases of the social dilemmas we
+% % have presented (Figure~\ref{fig:cooperation_rates_two_by_two_part_one}).
+
+% % In the harmony game the cooperation rates for both approaches are similar. The
+% % biggest difference occurs for \(T=1\), nevertheless, it is less than 10\%. There
+% % are two instances for which the cooperation rates are higher for the limited
+% % memory payoffs compared to the expected. These are for \(T=2.2, S=0.4\) and for
+% % \(T=2.2, S=0.8\). It is only in the snowdrift game that cooperation is
+% % overestimated when a minimum of information is considered. Players are slightly
+% % more likely to cooperate after a cooperation and to forgive after being at the
+% % receiving end of a defection. For the rest of the cases the classical scenario
+% % overestimates the average cooperation rate, supporting the results of the
+% % previous section. There is a substantial difference in the cooperation rates for
+% % the prisoner's dilemma class. We observe that for \(S<-0.4\) cooperation almost
+% % never emerges if players imitate based only on the last round payoff.
+
+% % \begin{figure}[!htbp]
+% %   \centering
+% %   \includegraphics[width=.4\textwidth]{static/expected_two_by_two_games_cooperation.pdf}
+% %   \includegraphics[width=.4\textwidth]{static/stochastic_two_by_two_games_cooperation.pdf}
+% %   \caption{\textbf{Cooperation rates for various social dilemmas.} ({\bf A}) If
+% %   players update based on their expected payoffs. ({\bf B}) If players update
+% %   based on their last round payoffs. We vary the temptation payoff \(T \in
+% %   \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across the \(x\) axis, and \(S \in \{1.2,
+% %   0.8, 0.4, 0, -0.4, -0.8\}\) across the \(y\) axis. Unless explicitly varied,
+% %   the parameters of the simulation are $N\!=\!100$, $\beta\!=\!10$,
+% %   $\delta\!=\!0.99$, $t\!=\!5\times
+% %   10^7$.}\label{fig:cooperation_rates_two_by_two_part_one}
+% % \end{figure}
+
+% % So far we have explored the difference between the expected payoffs and the last
+% % round payoffs. In order to explore further the effect of limited memory we allow
+% % individuals to remember more. More precisely, up to two interactions and up to
+% % the last two rounds. In total we present results for three more updating
+% % payoffs. These are the payoffs when individuals consider the last two rounds
+% % with another member of the population, the last round with two members of the
+% % population, and the last two rounds with two members of the population. Similar
+% % to the last round payoff, we use simulations and record which strategies the
+% % players adopt over time based
+% % (Appendix~\ref{appendix:further_simulation_results}) and we compare the evolving
+% % cooperation rates to those of the expected payoffs
+% % (Figure~\ref{fig:cooperation_rates_two_by_two_part_two}).
+
+% % In the case of the last two rounds payoff we make two observations. Initially,
+% % for the harmony, the stag-hunt and the prisoner's dilemma classes the average
+% % cooperation rates are higher than the previous case of the last round payoff.
+% % The rates remain strictly less than in the expected payoffs, however, in the
+% % case of the prisoner's dilemma we observe a significant increase. Secondly, for
+% % the snowdrift class there are more instances for which the cooperation rates are
+% % higher than in the expected payoffs. This could indicate that in the case of the
+% % snowdrift game the expected payoffs underestimate cooperation. Once players
+% % interact with two other members of the population, instead of one, the expected
+% % payoffs strictly yield a higher cooperation rate  even in the case of the
+% % snowdrift game. Similar to the previous result, the difference between the two
+% % approaches is always higher for the prisoner's dilemma class.
+
+% % \begin{figure}[!htbp]
+% %   \centering
+% %   \includegraphics[width=.33\textwidth]{static/rounds_two_by_two_games_cooperation.pdf}
+% %   \includegraphics[width=.33\textwidth]{static/opponents_two_by_two_games_cooperation.pdf}
+% %   \includegraphics[width=.33\textwidth]{static/rounds_opponents_two_by_two_games_cooperation.pdf}
+% %   \caption{\textbf{Cooperation rates over various social dilemmas for different
+% %   limited memory payoffs.} ({\bf A}) If players update based on their last two
+% %   rounds with another member of the population. ({\bf B}) If players update
+% %   based on their last round with two other members of the population. ({\bf C})
+% %   If players update based on their last two rounds with two other members of the
+% %   population. We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8,
+% %   2.2\}\) across the \(x\) axis, and \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\)
+% %   across the \(y\) axis. Unless explicitly varied, the parameters of the
+% %   simulation are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times
+% %   10^7$..}
+% %   \label{fig:cooperation_rates_two_by_two_part_two}
+% % \end{figure}
+
+% \section{Conclusions}\label{section:conclusions}
+
+% Cooperation can be seen at odd, why is it that we choose to help others,
+% increasing their payoff, at the expense of decreasing one's own? In spite of all
+% the selfish genes’ animal and human communities seem to altruistically help each
+% other and cooperate, and evolutionary game theory has helped us shape our
+% understanding of the evolution of cooperation.
+
+% Previous evolutionary models often feature a curious inconsistency. When
+% modeling how individuals make decisions in each round, these models assume that
+% players only remember the last round. However, when modeling how individuals
+% update their strategies over time, individuals are assumed to have perfect
+% memory.
+
+% Here, we have explored how robust cooperation is as models deviate from the
+% perfect memory assumption. Initially we considered the donation game. We showed
+% that when the last round payoff is used instead of the expected payoffs,
+% cooperation can even evolve when individuals only use a minimum of information,
+% however, the evolving cooperation rates are typically lower. The resident
+% strategies were both less cooperative and less generous. This effect was only
+% intensified we increase the benefit and the strength of selection independently.
+% The results showed that as each parameter was being increased the difference
+% between the cooperation rates were widening. This indicates that cooperative
+% players benefit from being able to interact with everyone in the population.
+
+% We extended our approach and presented results not only based on different
+% classes of social games, but also based on different limited memory payoffs. The
+% analysis showed that specifically for the case of the prisoner's dilemma games
+% the contrast between the two approaches can be significantly large. The
+% prisoner's dilemma is one of the most applied types of social games when we
+% discuss the evolution of cooperation. Our results indicate that cooperation
+% struggles to evolve in the prisoner's dilemma when only a minimum of social
+% information is used at the updating stage. Interestingly our simulations also
+% showed that in same cases cooperation can benefit from limited memory. This was
+% specific for the snowdrift game.
+
+% \section{Acknowledgements}
+
+% This work was supported by the European Research Council Starting Grant 850529:
+% E-DIRECT.
+
+% \appendix
+
+% \section{Model Setup}\label{appendix:methods}
+
+% Consider a population of \(N\) individuals where \(N\) is even. At any point in
+% time there are at most two different strategies in present in the population.
+% More specifically, a mutant strategy played by \(k\) individuals and a resident
+% strategy played by \(N - k\) individuals. We assume a pairwise process in which
+% strategies spread because they are imitated more often. Each step of the
+% evolutionary process consists of two stages; a game stage and an update stage.
+
+% In the game stage, each individual is randomly matched with some other
+% individual in the population. Their interaction lasts for a number of turns
+% which is not fixed but depends on the continuation probability \(\delta\). At
+% each turn the individuals choose between cooperation (\(C\)) and defection
+% (\(D\)). Thus, there are four possible outcomes in each turn \(CC, CD, DC\) and
+% \(DD\). If both players cooperate they receive the reward payoff \(R\), whereas
+% if both players defect they receive the punishment payoff \(P\). If one
+% cooperates but the other defects, the defector receives the temptation to
+% defect, \(T\), whereas the cooperator receives the sucker's payoff, \(S\). Let
+% $\mathcal{U}=\{R,S,T,P\}$ denote the set of feasible payoffs in each round, and
+% let $\mathbf{u}\!=\!(R,S,T,P)$ be the corresponding payoff vector. We present
+% results for various values of $\mathcal{U}$ for all the symmetric \(2 \times 2\)
+% games.
+
+% A further assumption of our model is that individuals make use of reactive
+% strategies when they make decisions in each round. Reactive strategies are a set
+% of strategies that take into account only the previous action of the opponent. A
+% reactive strategy can be written explicitly as a vector,
+
+% \[s=(y, p, q)\]
+
+% where \(y\) is the probability that the strategy opens with a cooperation and
+% \(p, q\) are the probabilities that the strategy cooperates given that the
+% opponent cooperated and defected equivalently.
+
+% In the updating stage, two players are randomly drawn from the population, a
+% `learner' and a `role model'. The learner adopts the role model's strategy based
+% on the Fermi distribution function, %ToDo add reference
+
+% \begin{equation} \label{Eq:rho}
+% \rho(u_{L}, u_{RM}) = \frac{1}{1\!+\! \exp^{\!-\!\beta (u_{RM}\!-\!u_{L})}}.
+% \end{equation}
+
+% where $u_{L}\!\in\! \mathcal{U}$ is the learner's payoff, $u_{RM}\!\in\!
+% \mathcal{U}$ is the role model's payoff, and $\beta\!\ge\!0$ is the strength of
+% selection.
+
+% We iterate this basic evolutionary step until either the mutant strategy goes
+% extinct, or until it fixes in the population and becomes the new resident
+% strategy. After either outcome, we set $k$ to $1$ and we introduce a new mutant
+% strategy which is uniformly chosen from all reactive strategies at random.
+% Instead of simulating each step of the evolutionary process, we estimate the
+% probability that a newly introduced mutant fixes~\cite{nowak2004emergence}. This
+% is defined as the fixation probability of the mutant, and the standard form is
+% the following,
+
+% \begin{equation}\label{eq:appendix_fixation_probability}
+% \varphi = \frac{1}{1+\sum\limits_{i=1}^{N-1}\prod\limits_k^i \frac{\lambda^-_k}{\lambda^+_k}},
+% \end{equation}
+
+% where \(\lambda^-_k, \lambda^+_k\) are the probabilities that the number of
+% mutants decreases and increases respectively.
+
+% This process of mutation and fixation/extinction is iterated many times. The
+% evolutionary process is summarized by
+% Algorithm~\ref{algorithm:pairwise_comparison}.
+
+% \begin{algorithm}[!htbp]
+%   \SetAlgoLined $N \leftarrow$ population size\; $k \leftarrow 1$\; resident
+%    $\leftarrow (0, 0, 0)$\; \While{ $t <$ maximum number of steps}{mutant
+%    $\leftarrow$ random: $\{\emptyset \}\ \rightarrow R^3$\; fixation probability
+%    $\leftarrow \varphi $\; \If{$\varphi >$ random: $i \rightarrow [0,
+%    1]$}{resident $\leftarrow$ mutant;}} \caption{Evolutionary
+%    process}\label{algorithm:pairwise_comparison}
+% \end{algorithm}
+
+% The aim of this work is to explore the effect of updating memory on the
+% cooperation rate of the evolved population. For this reason we consider two
+% different approaches when estimating the payoffs at the updating stage. The two
+% approaches we consider are those of (i) the expected and (ii) the limited memory
+% payoffs.
+
+% \subsection*{Expected Payoffs}
+
+% The expected payoffs are the conventional payoffs used in the updating
+% stage~\cite{imhof2010stochastic}. They are defined as the mean payoff of an
+% individual in a well-mixed population that engages in repeated games with all
+% other population members.
+
+% We first define the payoff of two reactive strategies at the game stage. Assume
+% two reactive strategies $s_1\!=\!(y_1, p_1, q_1$) and $s_2\!=\!(y_2,p_2,q_2)$.
+% It is not necessary to simulate the play move by move, instead the play between
+% the two strategies is defined a Markov matrix \(M\),
+
+% \begin{equation}\label{eq:transition_matrix}
+%   \input{static/matrix.tex}.
+% \end{equation}
+
+% whose stationary vector \(\mathbf{v}\), combined with the payoff \(u\), yields
+% the game stage outcome for each strategy,
+% \(\langle\mathbf{v}(s_1,s_2),\mathbf{u}\rangle\)~\cite{Hauert1997}.
+
+% In the updating stage the learner adopts the strategy of the role model based on
+% their updating payoffs. Given that there are only two different types in the
+% population at each time step we only need to define the expected payoff for a
+% resident (\(\pi_R\)) and for a mutant (\(\pi_M\)). Assume the resident strategy
+% \(s_R = (y_R, p_R, q_R)\) and the mutant strategy \(s_M = (y_M, p_M, q_M)\), the
+% expected payoffs are give by,
+
+% \begin{equation} \label{Eq:ExpPay}
+%   \begin{array}{lcrcr}
+%   \displaystyle \pi_R	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(s_R,s_R),\mathbf{u}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(s_R,s_M),\mathbf{u}\rangle,\\[0.5cm]
+%   \displaystyle \pi_M	&=	&\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(s_M,s_R),\mathbf{u}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(s_M,s_M),\mathbf{u}\rangle.\\
+%   \end{array}
+% \end{equation}
+
+% The number of mutant in the population increase if a learner resident adopts the
+% strategy of a mutant role model, and decreases if a mutant leaner adopts the
+% strategy of a resident. The probabilities that the number of mutants decreases
+% and increases, \(\lambda^-_k\) and \(\lambda^+_k\), are now explicitly defined
+% as,
+
+% \begin{align*} 
+%   \lambda^-_k &\!=\!\rho(\pi_R, \pi_M) \\
+%   \lambda^+_k &\!=\!\rho(\pi_M, \pi_R).
+% \end{align*}
+
+% \subsection*{Limited memory payoffs}
+
+% Initially, we discuss the case of the \textbf{last round updating payoff}. At
+% the stage game we define the payoff of a reactive strategy in the last round,
+% Proposition~\ref{proposition:last_round}.
+
+% \begin{Prop}\label{proposition:last_round} Consider a repeated game, with
+%     continuation probability $\delta$, between players with reactive strategies
+%     $s_1\!=\!(y_1, p_1, q_1$)  and $s_2\!=\!(y_2,p_2,q_2)$ respectively. Then
+%     the probability that the $s_1$ player receives the payoff $u\!\in\!
+%     \mathcal{U}$ in the very last round of the game is given by
+%     $v_{u}(s_1,s_2)$, as given by Equation~(\ref{Eq:LastRound}).
+
+%     \begin{equation} \label{Eq:LastRound}
+%       \setlength{\arraycolsep}{1pt}
+%       \begin{array}{rcl}
+
+%       v_{R}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
+%       {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)} \times R,\\[1cm]
+
+%       v_{S}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
+%       {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)} \times S,\\[1cm]
+
+%       v_{T}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
+%       {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)} \times T,\\[1cm]
+
+%       v_{P}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
+%       {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)} \times P.
+%       \end{array}
+%     \end{equation}
+
+% In these expressions, we have used the notation $r_i:=p_i\!-\!q_i$,
+% $\bar{y}_i\!=\!1\!-\!y_i$, $\bar{q}_i:=1\!-\!q_i$, and
+% $\bar{r}_i:=\bar{p}_i\!-\!\bar{q}_i=-r_i$ for $i\!\in\!\{1,2\}$.
+% \end{Prop}
+
+% \begin{proof}
+% Given a play between two reactive strategies with continuation probability
+% $\delta$. The outcome at turn \(t\) is given by,
+
+% \begin{equation}\label{eq:}
+%   (1 - \delta) \mathbf{v_0} \sum \delta^{t} M^{(t)},
+% \end{equation}
+
+% where $\mathbf{v_0}$ denotes the expected distribution of the four outcomes in
+% the very first round, and \(1- \delta\) the probability that the game ends. It
+% can be shown that,
+
+% \begin{align*}
+%   (1 - \delta) \mathbf{v_0} \sum \delta^{t} M^{(t)} & = (1 - \delta)(\mathbf{v_0} + \delta \mathbf{v_0} M + \delta^{2}\mathbf{v_0} M ^{2} + \dots )\\ 
+%    & = (1 - \delta)\mathbf{v_0} (1 + \delta M + \delta^{2}M ^{2} + \dots ) \text{ using standard formula for geometric series}\\ 
+%    & = (1 - \delta)\mathbf{v_0}(I_4 - \delta M)^{-1}
+% \end{align*}
+
+% where \((1 - \delta)\mathbf{v_0}(I_4 - \delta M)^{-1}\) is vector \(\in R^{4}\)
+% and it the probabilities for being in any of the outcomes \(CC, CD, DC, DD\) in
+% the last round. Combining this with the payoff vector \(u\) and some algebraic
+% manipulation we derive to the Equation~\ref{Eq:LastRound}.
+% \end{proof}
+
+% In the updating stage we select a mutant and resident to be either the role
+% model or the learner. Given that they can interact with only one other member of
+% the population, they can interact either with each other or either can interact
+% with another resident or with another mutant. Thus, in each updating stage there
+% are five possible combinations of pairs (Figure~\ref{fig:single_pairs}).
 
@@
 %   \centering
-%   \includegraphics[width=\textwidth]{static/stochastic_two_by_two_games.pdf}
-%   \caption{{\bf Evolutionary dynamics under last round payoffs for various social dilemmas.} 
-%   We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across
-%   the \(x\) axis, and  \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\) across the
-%   \(y\) axis. ({\bf A}) The top left quadrant where \(S > 0\) and \(T \leq 1\)
-%   corresponds to the harmony game. The preference ordering for the harmony game
-%   is \(R > T > S > P\). ({\bf B}) The top right quadrant where \(S > 0\) and \(T
-%   > 1\) corresponds to the snowdrift game. The preference ordering for the
-%   snowdrift game is \(T > R > S > P\). ({\bf C}) The bottom left quadrant where
-%   \(S < 0\) and \(T \leq 1\) corresponds to the stag-hunt game. The preference
-%   ordering for the stag-hunt game is \(R > T > P > S\). ({\bf D}) The bottom
-%   right quadrant where \(S < 0\) and \(T > 1\) corresponds to the prisoner's
-%   dilemma game. The preference ordering for the prisoner's dilemma game is \(T >
-%   R > P > S\). Unless explicitly varied, the parameters of the simulation are
-%   $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}
-%   \label{fig:last_round_two_by_two}
+%   \includegraphics[width=.65\textwidth]{static/last_rounds_matches.tex}
+%   \caption{\textbf{Possible pairings combination in the updating stage, given
+%   that individuals interact with only one other member in the population.} At
+%   each step of the evolutionary process we choose a role model and a learner to
+%   update the population. We consider the case where both the role model and the
+%   learner estimate their fitness after interacting with a single member of the
+%   population. There are five possible pairings at each step. They interact with
+%   other with a probability \(\frac{1}{N - 1}\), and thus they do not interact
+%   with other with a probability \(1 - \frac{1}{N - 1}\). In the latter case,
+%   each of them can interact with either a mutant or a resident. Both of them
+%   interact with a mutant with a probability $\frac{(k-1)(k-2)}{(N-2)(N-3)}$ and
+%   both interact with a resident with a probability
+%   $\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$. The last two possible pairings are that
+%   either of them interacts with a resident whilst the other interacts with a
+%   mutant, and this happens with a probability
+%   $\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$.}
+%   \label{fig:single_pairs}
 % \end{figure}
 
-% The harmony game describes an interaction without conflict. In the harmony game
-% it is in the best interest for both players to cooperate, and in an evolutionary
-% process cooperators prevail. We observe that for most harmony game runs the
-% resident population overwhelmingly applies a strategy for which \(p\) and \(q\)
-% are \(\approx 1\) (Figure~\ref{fig:expected_two_by_two}). The snowdrift game
-% describes a situation similar to that of the prisoner's dilemma; cooperation
-% results in a benefit to the opposing player but entailing a cost to the
-% cooperator. However, in the snowdrift game individuals obtain immediate direct
-% benefits from the cooperative acts which leads to \(S > P\). In the snowdrift
-% game more cooperation emerges compared to the prisoner's dilemma. Defection is
-% never a resident strategy and for the same values of temptation the overall
-% \(q-\) values are higher. The last class of games we present are for the
-% stag-hunt game. In the stag-hunt game there is an equilibrium in which both
-% players cooperate as well as one in which both defect. For small values of the
-% temptation and the sucker's payoffs we observe that cooperation prevails similar
-% to the harmony game. However, in the case where the temptation and the reward
-% payoffs are equal the resident population either consists of defectors or
-% cooperators.
-
-% Although the patterns of the evolved populations remain similar if we consider
-% that players update their strategies based on their last round payoff
-% (Figure~\ref{fig:last_round_two_by_two}), there are still dissimilarities
-% between the two approaches. To better understand these dissimilarities, we
-% explore the cooperation rates for each of the cases of the social dilemmas we
-% have presented (Figure~\ref{fig:cooperation_rates_two_by_two_part_one}).
-
-% In the harmony game the cooperation rates for both approaches are similar. The
-% biggest difference occurs for \(T=1\), nevertheless, it is less than 10\%. There
-% are two instances for which the cooperation rates are higher for the limited
-% memory payoffs compared to the expected. These are for \(T=2.2, S=0.4\) and for
-% \(T=2.2, S=0.8\). It is only in the snowdrift game that cooperation is
-% overestimated when a minimum of information is considered. Players are slightly
-% more likely to cooperate after a cooperation and to forgive after being at the
-% receiving end of a defection. For the rest of the cases the classical scenario
-% overestimates the average cooperation rate, supporting the results of the
-% previous section. There is a substantial difference in the cooperation rates for
-% the prisoner's dilemma class. We observe that for \(S<-0.4\) cooperation almost
-% never emerges if players imitate based only on the last round payoff.
+% Given the last round payoff and possible pair combinations for a single
+% interaction, we define the probability that the respective last round payoffs of
+% two players \(s_1, s_2\) are given by $u_1$ and $u_2$ as,
+
+% \begin{equation}\label{eq:Chi}
+% \setlength{\arraycolsep}{1pt}
+% \begin{array}{llrl}
+% x(u_1,u_2)	 &=&\displaystyle \frac{1}{N\!-\!1}\cdot  &v_{u_1}(s_1,s_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F}\\[0.5cm]
+% &+	
+% &\displaystyle \left(1\!-\!\frac{1}{N\!-\!1}\right)  
+% &\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_2) + 
+%  \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_1)\right.\\[0.5cm]
+% &&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_2) + 
+%  \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_1)\right].
+% \end{array}
+% \end{equation}
+
+% The first term on the right side corresponds to the case that the learner and
+% the role model happened to be matched during the game stage, which happens with
+% probability $\frac{1}{(N\!-\!1)}$. In that case, we note that only those payoff
+% pairs can occur that are feasible in a direct interaction,
+% $\splitatcommas{(u_1,u_2)\in \mathcal{U}^2_F:=\big\{ (R,R), (S,T), (T,S), (P,P)
+% \big\}}$, as represented by the respective indicator function. Otherwise, if the
+% learner and the role model did not interact directly, we need to distinguish
+% four different cases, depending on whether the learner was matched with a
+% resident or a mutant, and depending on whether the role model was matched with a
+% resident or a mutant.
+
+% Given that $N\!-\!k$ players use the resident strategy
+% $s_{R}\!=\!(y_{R},p_{R},q_{R})$ and that the remaining $k$ players use the
+% mutant strategy $s_{M}\!=\!(y_{M},p_{M},q_{M})$, the probability that the number
+% of mutants increases by one in one step of the evolutionary process can be
+% written as
+
+% \begin{align}
+% \lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M})\cdot \rho(u_{R},u_{M}), \\
+% \lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M})\cdot \rho(u_{M},u_{R}).
+% \end{align}
+
+% In this expression, $\frac{(N\!-\!k)}{N}$ is the probability that the randomly
+% chosen learner is a resident, and $\frac{k}{N}$ is the probability that the role
+% model is a mutant. The sum corresponds to the total probability that the learner
+% adopts the role model's strategy over all possible payoffs $u_R$ and $u_M$ that
+% the two player may have received in their respective last rounds. We use
+% $x(u_R,u_M)$ to denote the probability that the randomly chosen resident
+% obtained a payoff of $u_R$ in the last round of his respective game, and that
+% the mutant obtained a payoff of $u_M$.
+
+% We extend our framework to consider the case where players update their
+% strategies based on the outcome of \textbf{the last two turns and based on their
+% interaction with two other members of the population}. At the stage game we
+% define the payoff of a reactive strategy in the last two rounds,
+% Proposition~\ref{proposition:last_two_rounds}.
+
+% \begin{Prop}\label{proposition:last_two_rounds} Consider a repeated game, with
+%   continuation probability $\delta$, between players with reactive strategies
+%   $s_1\!=\!(y_1, p_1, q_1$)  and $s_2\!=\!(y_2,p_2,q_2)$ respectively. Let
+%   $\mathcal{\tilde{U}} = \splitatcommas{\{RR, RS, RT, RP, SR, SS, ST, SP, TR,
+%   TS, TT, TP, PR, PS, PT, PP\}}$ denote the set of feasible payoffs in the last
+%   two rounds, and let \(\tilde{\mathbf{u}}\) be the corresponding payoff vector.
+%   Then the probability that the $s_1$ player receives the payoff $u\!\in\!
+%   \mathcal{\tilde{U}}$ in the very last two rounds of the game is given by,
+
+%   \begin{equation}
+%   \langle\mathbf{\tilde{v}}(s_1,s_2),\mathbf{\tilde{u}}\rangle, \text{ where } \mathbf{\tilde{v}} \in R^{16} \text{ is given by },
+%   \end{equation}
+%   \begin{equation}
+%     \mathbf{\tilde{v}}(s_1,s_2) = (1 - \delta) m_{a_1, a_2} \delta^2 \left[\mathbf{v_0}(I_4 - \delta M)^{-1}\right]_{a_1, a_2}, \quad  m_{a_1, a_2} \in M \ \forall \ a_1, a_2 \in \{1, 2, 3, 4\}.
+%   \end{equation}
+% \end{Prop}
+
+% In the updating stage we select a mutant and resident to be either the role
+% model or the learner. Given that they can interact with two other members of the
+% population there are a total of twenty four possible combinations of pairs
+% (Figure~\ref{fig:pissible_two_pairs}).
 
@@
 %   \centering
-%   \includegraphics[width=.4\textwidth]{static/expected_two_by_two_games_cooperation.pdf}
-%   \includegraphics[width=.4\textwidth]{static/stochastic_two_by_two_games_cooperation.pdf}
-%   \caption{\textbf{Cooperation rates for various social dilemmas.} ({\bf A}) If
-%   players update based on their expected payoffs. ({\bf B}) If players update
-%   based on their last round payoffs. We vary the temptation payoff \(T \in
-%   \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across the \(x\) axis, and \(S \in \{1.2,
-%   0.8, 0.4, 0, -0.4, -0.8\}\) across the \(y\) axis. Unless explicitly varied,
-%   the parameters of the simulation are $N\!=\!100$, $\beta\!=\!10$,
-%   $\delta\!=\!0.99$, $t\!=\!5\times
-%   10^7$.}\label{fig:cooperation_rates_two_by_two_part_one}
+%   \includegraphics[width=.65\textwidth]{static/tree.tex}
+%   \caption{\textbf{Possible pairings combination in the updating stage, given
+%   that individuals interact with two other members in the population.}}
+%   \label{fig:pissible_two_pairs}
 % \end{figure}
 
-% So far we have explored the difference between the expected payoffs and the last
-% round payoffs. In order to explore further the effect of limited memory we allow
-% individuals to remember more. More precisely, up to two interactions and up to
-% the last two rounds. In total we present results for three more updating
-% payoffs. These are the payoffs when individuals consider the last two rounds
-% with another member of the population, the last round with two members of the
-% population, and the last two rounds with two members of the population. Similar
-% to the last round payoff, we use simulations and record which strategies the
-% players adopt over time based
-% (Appendix~\ref{appendix:further_simulation_results}) and we compare the evolving
-% cooperation rates to those of the expected payoffs
-% (Figure~\ref{fig:cooperation_rates_two_by_two_part_two}).
-
-% In the case of the last two rounds payoff we make two observations. Initially,
-% for the harmony, the stag-hunt and the prisoner's dilemma classes the average
-% cooperation rates are higher than the previous case of the last round payoff.
-% The rates remain strictly less than in the expected payoffs, however, in the
-% case of the prisoner's dilemma we observe a significant increase. Secondly, for
-% the snowdrift class there are more instances for which the cooperation rates are
-% higher than in the expected payoffs. This could indicate that in the case of the
-% snowdrift game the expected payoffs underestimate cooperation. Once players
-% interact with two other members of the population, instead of one, the expected
-% payoffs strictly yield a higher cooperation rate  even in the case of the
-% snowdrift game. Similar to the previous result, the difference between the two
-% approaches is always higher for the prisoner's dilemma class.
-
-% \begin{figure}[!htbp]
-%   \centering
-%   \includegraphics[width=.33\textwidth]{static/rounds_two_by_two_games_cooperation.pdf}
-%   \includegraphics[width=.33\textwidth]{static/opponents_two_by_two_games_cooperation.pdf}
-%   \includegraphics[width=.33\textwidth]{static/rounds_opponents_two_by_two_games_cooperation.pdf}
-%   \caption{\textbf{Cooperation rates over various social dilemmas for different
-%   limited memory payoffs.} ({\bf A}) If players update based on their last two
-%   rounds with another member of the population. ({\bf B}) If players update
-%   based on their last round with two other members of the population. ({\bf C})
-%   If players update based on their last two rounds with two other members of the
-%   population. We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8,
-%   2.2\}\) across the \(x\) axis, and \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\)
-%   across the \(y\) axis. Unless explicitly varied, the parameters of the
-%   simulation are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times
-%   10^7$..}
-%   \label{fig:cooperation_rates_two_by_two_part_two}
-% \end{figure}
-
-\section{Conclusions}\label{section:conclusions}
-
-Cooperation can be seen at odd, why is it that we choose to help others,
-increasing their payoff, at the expense of decreasing one's own? In spite of all
-the selfish genes’ animal and human communities seem to altruistically help each
-other and cooperate, and evolutionary game theory has helped us shape our
-understanding of the evolution of cooperation.
-
-Previous evolutionary models often feature a curious inconsistency. When
-modeling how individuals make decisions in each round, these models assume that
-players only remember the last round. However, when modeling how individuals
-update their strategies over time, individuals are assumed to have perfect
-memory.
-
-Here, we have explored how robust cooperation is as models deviate from the
-perfect memory assumption. Initially we considered the donation game. We showed
-that when the last round payoff is used instead of the expected payoffs,
-cooperation can even evolve when individuals only use a minimum of information,
-however, the evolving cooperation rates are typically lower. The resident
-strategies were both less cooperative and less generous. This effect was only
-intensified we increase the benefit and the strength of selection independently.
-The results showed that as each parameter was being increased the difference
-between the cooperation rates were widening. This indicates that cooperative
-players benefit from being able to interact with everyone in the population.
-
-We extended our approach and presented results not only based on different
-classes of social games, but also based on different limited memory payoffs. The
-analysis showed that specifically for the case of the prisoner's dilemma games
-the contrast between the two approaches can be significantly large. The
-prisoner's dilemma is one of the most applied types of social games when we
-discuss the evolution of cooperation. Our results indicate that cooperation
-struggles to evolve in the prisoner's dilemma when only a minimum of social
-information is used at the updating stage. Interestingly our simulations also
-showed that in same cases cooperation can benefit from limited memory. This was
-specific for the snowdrift game.
-
-\section{Acknowledgements}
-
-This work was supported by the European Research Council Starting Grant 850529:
-E-DIRECT.
-
-\appendix
-
-\section{Model Setup}\label{appendix:methods}
-
-Consider a population of \(N\) individuals where \(N\) is even. At any point in
-time there are at most two different strategies in present in the population.
-More specifically, a mutant strategy played by \(k\) individuals and a resident
-strategy played by \(N - k\) individuals. We assume a pairwise process in which
-strategies spread because they are imitated more often. Each step of the
-evolutionary process consists of two stages; a game stage and an update stage.
-
-In the game stage, each individual is randomly matched with some other
-individual in the population. Their interaction lasts for a number of turns
-which is not fixed but depends on the continuation probability \(\delta\). At
-each turn the individuals choose between cooperation (\(C\)) and defection
-(\(D\)). Thus, there are four possible outcomes in each turn \(CC, CD, DC\) and
-\(DD\). If both players cooperate they receive the reward payoff \(R\), whereas
-if both players defect they receive the punishment payoff \(P\). If one
-cooperates but the other defects, the defector receives the temptation to
-defect, \(T\), whereas the cooperator receives the sucker's payoff, \(S\). Let
-$\mathcal{U}=\{R,S,T,P\}$ denote the set of feasible payoffs in each round, and
-let $\mathbf{u}\!=\!(R,S,T,P)$ be the corresponding payoff vector. We present
-results for various values of $\mathcal{U}$ for all the symmetric \(2 \times 2\)
-games.
-
-A further assumption of our model is that individuals make use of reactive
-strategies when they make decisions in each round. Reactive strategies are a set
-of strategies that take into account only the previous action of the opponent. A
-reactive strategy can be written explicitly as a vector,
-
-\[s=(y, p, q)\]
-
-where \(y\) is the probability that the strategy opens with a cooperation and
-\(p, q\) are the probabilities that the strategy cooperates given that the
-opponent cooperated and defected equivalently.
-
-In the updating stage, two players are randomly drawn from the population, a
-`learner' and a `role model'. The learner adopts the role model's strategy based
-on the Fermi distribution function, %ToDo add reference
-
-\begin{equation} \label{Eq:rho}
-\rho(u_{L}, u_{RM}) = \frac{1}{1\!+\! \exp^{\!-\!\beta (u_{RM}\!-\!u_{L})}}.
-\end{equation}
-
-where $u_{L}\!\in\! \mathcal{U}$ is the learner's payoff, $u_{RM}\!\in\!
-\mathcal{U}$ is the role model's payoff, and $\beta\!\ge\!0$ is the strength of
-selection.
-
-We iterate this basic evolutionary step until either the mutant strategy goes
-extinct, or until it fixes in the population and becomes the new resident
-strategy. After either outcome, we set $k$ to $1$ and we introduce a new mutant
-strategy which is uniformly chosen from all reactive strategies at random.
-Instead of simulating each step of the evolutionary process, we estimate the
-probability that a newly introduced mutant fixes~\cite{nowak2004emergence}. This
-is defined as the fixation probability of the mutant, and the standard form is
-the following,
-
-\begin{equation}\label{eq:appendix_fixation_probability}
-\varphi = \frac{1}{1+\sum\limits_{i=1}^{N-1}\prod\limits_k^i \frac{\lambda^-_k}{\lambda^+_k}},
-\end{equation}
-
-where \(\lambda^-_k, \lambda^+_k\) are the probabilities that the number of
-mutants decreases and increases respectively.
-
-This process of mutation and fixation/extinction is iterated many times. The
-evolutionary process is summarized by
-Algorithm~\ref{algorithm:pairwise_comparison}.
-
-\begin{algorithm}[!htbp]
-  \SetAlgoLined $N \leftarrow$ population size\; $k \leftarrow 1$\; resident
-   $\leftarrow (0, 0, 0)$\; \While{ $t <$ maximum number of steps}{mutant
-   $\leftarrow$ random: $\{\emptyset \}\ \rightarrow R^3$\; fixation probability
-   $\leftarrow \varphi $\; \If{$\varphi >$ random: $i \rightarrow [0,
-   1]$}{resident $\leftarrow$ mutant;}} \caption{Evolutionary
-   process}\label{algorithm:pairwise_comparison}
-\end{algorithm}
-
-The aim of this work is to explore the effect of updating memory on the
-cooperation rate of the evolved population. For this reason we consider two
-different approaches when estimating the payoffs at the updating stage. The two
-approaches we consider are those of (i) the expected and (ii) the limited memory
-payoffs.
-
-\subsection*{Expected Payoffs}
-
-The expected payoffs are the conventional payoffs used in the updating
-stage~\cite{imhof2010stochastic}. They are defined as the mean payoff of an
-individual in a well-mixed population that engages in repeated games with all
-other population members.
-
-We first define the payoff of two reactive strategies at the game stage. Assume
-two reactive strategies $s_1\!=\!(y_1, p_1, q_1$) and $s_2\!=\!(y_2,p_2,q_2)$.
-It is not necessary to simulate the play move by move, instead the play between
-the two strategies is defined a Markov matrix \(M\),
-
-\begin{equation}\label{eq:transition_matrix}
-  \input{static/matrix.tex}.
-\end{equation}
-
-whose stationary vector \(\mathbf{v}\), combined with the payoff \(u\), yields
-the game stage outcome for each strategy,
-\(\langle\mathbf{v}(s_1,s_2),\mathbf{u}\rangle\)~\cite{Hauert1997}.
-
-In the updating stage the learner adopts the strategy of the role model based on
-their updating payoffs. Given that there are only two different types in the
-population at each time step we only need to define the expected payoff for a
-resident (\(\pi_R\)) and for a mutant (\(\pi_M\)). Assume the resident strategy
-\(s_R = (y_R, p_R, q_R)\) and the mutant strategy \(s_M = (y_M, p_M, q_M)\), the
-expected payoffs are give by,
-
-\begin{equation} \label{Eq:ExpPay}
-  \begin{array}{lcrcr}
-  \displaystyle \pi_R	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(s_R,s_R),\mathbf{u}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(s_R,s_M),\mathbf{u}\rangle,\\[0.5cm]
-  \displaystyle \pi_M	&=	&\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(s_M,s_R),\mathbf{u}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(s_M,s_M),\mathbf{u}\rangle.\\
-  \end{array}
-\end{equation}
-
-The number of mutant in the population increase if a learner resident adopts the
-strategy of a mutant role model, and decreases if a mutant leaner adopts the
-strategy of a resident. The probabilities that the number of mutants decreases
-and increases, \(\lambda^-_k\) and \(\lambda^+_k\), are now explicitly defined
-as,
-
-\begin{align*} 
-  \lambda^-_k &\!=\!\rho(\pi_R, \pi_M) \\
-  \lambda^+_k &\!=\!\rho(\pi_M, \pi_R).
-\end{align*}
-
-\subsection*{Limited memory payoffs}
-
-Initially, we discuss the case of the \textbf{last round updating payoff}. At
-the stage game we define the payoff of a reactive strategy in the last round,
-Proposition~\ref{proposition:last_round}.
-
-\begin{Prop}\label{proposition:last_round} Consider a repeated game, with
-    continuation probability $\delta$, between players with reactive strategies
-    $s_1\!=\!(y_1, p_1, q_1$)  and $s_2\!=\!(y_2,p_2,q_2)$ respectively. Then
-    the probability that the $s_1$ player receives the payoff $u\!\in\!
-    \mathcal{U}$ in the very last round of the game is given by
-    $v_{u}(s_1,s_2)$, as given by Equation~(\ref{Eq:LastRound}).
-
-    \begin{equation} \label{Eq:LastRound}
-      \setlength{\arraycolsep}{1pt}
-      \begin{array}{rcl}
-
-      v_{R}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
-      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)} \times R,\\[1cm]
-
-      v_{S}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
-      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)} \times S,\\[1cm]
-
-      v_{T}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
-      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)} \times T,\\[1cm]
-
-      v_{P}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
-      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)} \times P.
-      \end{array}
-    \end{equation}
-
-In these expressions, we have used the notation $r_i:=p_i\!-\!q_i$,
-$\bar{y}_i\!=\!1\!-\!y_i$, $\bar{q}_i:=1\!-\!q_i$, and
-$\bar{r}_i:=\bar{p}_i\!-\!\bar{q}_i=-r_i$ for $i\!\in\!\{1,2\}$.
-\end{Prop}
-
-\begin{proof}
-Given a play between two reactive strategies with continuation probability
-$\delta$. The outcome at turn \(t\) is given by,
-
-\begin{equation}\label{eq:}
-  (1 - \delta) \mathbf{v_0} \sum \delta^{t} M^{(t)},
-\end{equation}
-
-where $\mathbf{v_0}$ denotes the expected distribution of the four outcomes in
-the very first round, and \(1- \delta\) the probability that the game ends. It
-can be shown that,
-
-\begin{align*}
-  (1 - \delta) \mathbf{v_0} \sum \delta^{t} M^{(t)} & = (1 - \delta)(\mathbf{v_0} + \delta \mathbf{v_0} M + \delta^{2}\mathbf{v_0} M ^{2} + \dots )\\ 
-   & = (1 - \delta)\mathbf{v_0} (1 + \delta M + \delta^{2}M ^{2} + \dots ) \text{ using standard formula for geometric series}\\ 
-   & = (1 - \delta)\mathbf{v_0}(I_4 - \delta M)^{-1}
-\end{align*}
-
-where \((1 - \delta)\mathbf{v_0}(I_4 - \delta M)^{-1}\) is vector \(\in R^{4}\)
-and it the probabilities for being in any of the outcomes \(CC, CD, DC, DD\) in
-the last round. Combining this with the payoff vector \(u\) and some algebraic
-manipulation we derive to the Equation~\ref{Eq:LastRound}.
-\end{proof}
-
-In the updating stage we select a mutant and resident to be either the role
-model or the learner. Given that they can interact with only one other member of
-the population, they can interact either with each other or either can interact
-with another resident or with another mutant. Thus, in each updating stage there
-are five possible combinations of pairs (Figure~\ref{fig:single_pairs}).
-
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[width=.65\textwidth]{static/last_rounds_matches.tex}
-  \caption{\textbf{Possible pairings combination in the updating stage, given
-  that individuals interact with only one other member in the population.} At
-  each step of the evolutionary process we choose a role model and a learner to
-  update the population. We consider the case where both the role model and the
-  learner estimate their fitness after interacting with a single member of the
-  population. There are five possible pairings at each step. They interact with
-  other with a probability \(\frac{1}{N - 1}\), and thus they do not interact
-  with other with a probability \(1 - \frac{1}{N - 1}\). In the latter case,
-  each of them can interact with either a mutant or a resident. Both of them
-  interact with a mutant with a probability $\frac{(k-1)(k-2)}{(N-2)(N-3)}$ and
-  both interact with a resident with a probability
-  $\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$. The last two possible pairings are that
-  either of them interacts with a resident whilst the other interacts with a
-  mutant, and this happens with a probability
-  $\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$.}
-  \label{fig:single_pairs}
-\end{figure}
-
-Given the last round payoff and possible pair combinations for a single
-interaction, we define the probability that the respective last round payoffs of
-two players \(s_1, s_2\) are given by $u_1$ and $u_2$ as,
-
-\begin{equation}\label{eq:Chi}
-\setlength{\arraycolsep}{1pt}
-\begin{array}{llrl}
-x(u_1,u_2)	 &=&\displaystyle \frac{1}{N\!-\!1}\cdot  &v_{u_1}(s_1,s_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F}\\[0.5cm]
-&+	
-&\displaystyle \left(1\!-\!\frac{1}{N\!-\!1}\right)  
-&\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_2) + 
- \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_1)\right.\\[0.5cm]
-&&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_2) + 
- \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_1)\right].
-\end{array}
-\end{equation}
-
-The first term on the right side corresponds to the case that the learner and
-the role model happened to be matched during the game stage, which happens with
-probability $\frac{1}{(N\!-\!1)}$. In that case, we note that only those payoff
-pairs can occur that are feasible in a direct interaction,
-$\splitatcommas{(u_1,u_2)\in \mathcal{U}^2_F:=\big\{ (R,R), (S,T), (T,S), (P,P)
-\big\}}$, as represented by the respective indicator function. Otherwise, if the
-learner and the role model did not interact directly, we need to distinguish
-four different cases, depending on whether the learner was matched with a
-resident or a mutant, and depending on whether the role model was matched with a
-resident or a mutant.
-
-Given that $N\!-\!k$ players use the resident strategy
-$s_{R}\!=\!(y_{R},p_{R},q_{R})$ and that the remaining $k$ players use the
-mutant strategy $s_{M}\!=\!(y_{M},p_{M},q_{M})$, the probability that the number
-of mutants increases by one in one step of the evolutionary process can be
-written as
-
-\begin{align}
-\lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M})\cdot \rho(u_{R},u_{M}), \\
-\lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M})\cdot \rho(u_{M},u_{R}).
-\end{align}
-
-In this expression, $\frac{(N\!-\!k)}{N}$ is the probability that the randomly
-chosen learner is a resident, and $\frac{k}{N}$ is the probability that the role
-model is a mutant. The sum corresponds to the total probability that the learner
-adopts the role model's strategy over all possible payoffs $u_R$ and $u_M$ that
-the two player may have received in their respective last rounds. We use
-$x(u_R,u_M)$ to denote the probability that the randomly chosen resident
-obtained a payoff of $u_R$ in the last round of his respective game, and that
-the mutant obtained a payoff of $u_M$.
-
-We extend our framework to consider the case where players update their
-strategies based on the outcome of \textbf{the last two turns and based on their
-interaction with two other members of the population}. At the stage game we
-define the payoff of a reactive strategy in the last two rounds,
-Proposition~\ref{proposition:last_two_rounds}.
-
-\begin{Prop}\label{proposition:last_two_rounds} Consider a repeated game, with
-  continuation probability $\delta$, between players with reactive strategies
-  $s_1\!=\!(y_1, p_1, q_1$)  and $s_2\!=\!(y_2,p_2,q_2)$ respectively. Let
-  $\mathcal{\tilde{U}} = \splitatcommas{\{RR, RS, RT, RP, SR, SS, ST, SP, TR,
-  TS, TT, TP, PR, PS, PT, PP\}}$ denote the set of feasible payoffs in the last
-  two rounds, and let \(\tilde{\mathbf{u}}\) be the corresponding payoff vector.
-  Then the probability that the $s_1$ player receives the payoff $u\!\in\!
-  \mathcal{\tilde{U}}$ in the very last two rounds of the game is given by,
-
-  \begin{equation}
-  \langle\mathbf{\tilde{v}}(s_1,s_2),\mathbf{\tilde{u}}\rangle, \text{ where } \mathbf{\tilde{v}} \in R^{16} \text{ is given by },
-  \end{equation}
-  \begin{equation}
-    \mathbf{\tilde{v}}(s_1,s_2) = (1 - \delta) m_{a_1, a_2} \delta^2 \left[\mathbf{v_0}(I_4 - \delta M)^{-1}\right]_{a_1, a_2}, \quad  m_{a_1, a_2} \in M \ \forall \ a_1, a_2 \in \{1, 2, 3, 4\}.
-  \end{equation}
-\end{Prop}
-
-In the updating stage we select a mutant and resident to be either the role
-model or the learner. Given that they can interact with two other members of the
-population there are a total of twenty four possible combinations of pairs
-(Figure~\ref{fig:pissible_two_pairs}).
-
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[width=.65\textwidth]{static/tree.tex}
-  \caption{\textbf{Possible pairings combination in the updating stage, given
-  that individuals interact with two other members in the population.}}
-  \label{fig:pissible_two_pairs}
-\end{figure}
-
-Given the last two rounds payoff and possible pair combinations with two
-members, we can define the probability \(x\) that the respective last round
-payoffs of two players \(s_1, s_2\) are given by $u_1$ and $u_2$ similarly to
-Eq.~(\ref{eq:Chi}). We follow the same approach to define the rest of the
-updating payoffs. These are the updating payoffs of the last round with two
-member, and the the last two rounds payoffs with one member.
-
-Simulating the evolutionary process for more interactions and rounds quickly
-becomes computationally intractable. Our methodology could be extended to
-include \(n\) turns and \(m\) interactions. However, for the purpose of this
-work we explore the cases only up to two turns and two interactions.
+% Given the last two rounds payoff and possible pair combinations with two
+% members, we can define the probability \(x\) that the respective last round
+% payoffs of two players \(s_1, s_2\) are given by $u_1$ and $u_2$ similarly to
+% Eq.~(\ref{eq:Chi}). We follow the same approach to define the rest of the
+% updating payoffs. These are the updating payoffs of the last round with two
+% member, and the the last two rounds payoffs with one member.
+
+% Simulating the evolutionary process for more interactions and rounds quickly
+% becomes computationally intractable. Our methodology could be extended to
+% include \(n\) turns and \(m\) interactions. However, for the purpose of this
+% work we explore the cases only up to two turns and two interactions.
 

### 2023-02-21 paper/main.tex
working on paper in folder paper

\documentclass[11pt]{article}
\usepackage{times}
\usepackage{amsmath,amsthm,amssymb,setspace,enumitem,epsfig,titlesec,verbatim,color,array,eurosym,multirow}
\usepackage[sort&compress]{natbib}
\usepackage[footnotesize,bf]{caption}
\usepackage[margin=2.5cm, includefoot, footskip=30pt]{geometry}
\usepackage{standalone}
\usepackage{tikz}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{blkarray}
\usepackage[ruled,vlined]{algorithm2e}
\smallskip % Erlaubt kleine Abstaende zwischen Paragraphen, falls es dem Seitenlayout hilft
\renewcommand{\baselinestretch}{1.3}
\newcommand{\R}{\mathbb{R}}

\definecolor{darkblue}{rgb}{0,0,.8}
\definecolor{darkgreen}{rgb}{0,0.5,0.1}
\newcommand{\matlabfunc}[1]{\textcolor{darkblue}{#1}}
\newcommand{\matlabcomment}[1]{\textcolor{darkgreen}{#1}}
\newcommand{\christian}[1]{\textcolor{blue}{\textbf{CH}: #1}}
\newcommand{\alex}[1]{\textcolor{red}{\textbf{AL}: #1}}

%% Adding shortcut commands to refer to our figures %%
\newcommand{\FigEvoProc}{{\bf Fig.~1}} \newcommand{\FigInvAnalysis}{{\bfFig.~2}} \newcommand{\FigResultsOverPara}{{\bf Fig.~3}}

\titleformat{\section}{\sffamily \fontsize{12}{14}\bfseries}{\thesection}{1em}{}
\titleformat{\subsection}{\sffamily
\fontsize{11.5}{11.5}\bfseries}{\thesubsection}{1em}{}

\usepackage{tikz}
\usetikzlibrary{arrows}

\tikzset{treenode/.style = {align=center, inner sep=0pt, text centered,
  font=\sffamily}, arn_n/.style = {treenode, circle, white,
  font=\sffamily\bfseries, draw=black, inner sep=-6pt, fill=black, text
  width=1.5em},% arbre rouge noir, noeud noir
  arn_r/.style = {treenode, circle, red, text width=1.5em, very thick, inner
    sep=4pt},% arbre rouge noir, noeud rouge
  arn_x/.style = {treenode, rectangle, draw=black, minimum width=0.5em, minimum
    height=0.5em}% arbre rouge noir, nil
}

\newtheoremstyle{plainCl1}% name
{9pt}%      Space above, empty = 'usual value'
{9pt}%      Space below
{\it}% 	   Body font
{}%         Indent amount (empty = no indent, \parindent = para indent)
{\bfseries}% Thm head font
{.}%        Punctuation after thm head
{0.2cm}% Space after thm head: \newline = linebreak
{}%         Thm head spec

\newtheoremstyle{plainCl2}% name
{9pt}%      Space above, empty = 'usual value'
{9pt}%      Space below
{\it}% 	   Body font
{}%         Indent amount (empty = no indent, \parindent = para indent)
{\bfseries}% Thm head font
{$'$.}%        Punctuation after thm head
{0.2cm}% Space after thm head: \newline = linebreak
{}%         Thm head spec

\newcommand{\splitatcommas}[1]{%
  \begingroup
  \begingroup\lccode`~=`, \lowercase{\endgroup \edef~{\mathchar\the\mathcode`,
    \penalty0 \noexpand\hspace{0pt plus 1em}}%
  }\mathcode`,="8000 #1%
  \endgroup
}

\theoremstyle{plainCl1}
\newtheorem{Claim}{Claim}
\newtheorem{Thm}{Theorem}
\newtheorem{Prop}{Proposition}
\newtheorem*{Lem}{Lemma}
\newtheorem{Cor}{Corollary}
\newtheorem*{Def}{Definition}

\theoremstyle{plainCl2}
\newtheorem{Claim2}{Claim}

\title{\bf  \sffamily \LARGE Evolution of cooperation among individuals with
limited payoff memory\\}
\date{}
\author{Nikoleta E. Glynatsi, Christian Hilbe, Alex McAvoy}

\begin{document}
\maketitle

\begin{abstract}
% Repeated games are vastly use in evolutionary game theory to explain cooperation in a variety of environments. Although existing models have shaped
% our understanding of human cooperation, hey often work with idealized
% assumptions. In an evolutionary process, individuals imitate other individuals
% of the population based on their fitness. It is commonly assumed that an
% individual computes their fitness after interacting with a representative sample
% of the population, and remembering all the interactions they participated in. In
% real life, we do not always remember all our interactions, instead we rather
% recall our most recent ones.

% Here, we introduce a framework that allows individuals to estimate their fitness
% based on a minimum of social information. We explore the difference between our
% framework and the classical framework using computer simulations. We present
% results for the most commonly used classes of symmetric \(2 \times 2\) games.
% The simulations show that, in the prisoner's dilemma individuals with limited
% memory tend to adopt less generous strategies and they achieve less cooperation
% than in the classical scenario. In contrast, in the stag-hunt and the harmony
% game, the impact of memory is less striking. Here individuals with limited
% memory perform nearly as well as individuals with full memory. Finally, we
% observe that the snowdrift game is the only game for which cooperation can be
% underestimated in the classical scenario.
\end{abstract}

\section{Introduction}

% The framework of evolutionary game theory has been applied in explaining the
% behavior of sticklebacks~\cite{milinski1987tit}, the rock-scissors-paper cycles
% in bacterial populations~\cite{kerr2002local} and food sharing in vampire
% bats~\cite{carter2020development} colonies.

Evolutionary game theory~\cite{smith1982evolution, hofbauer1998evolutionary,
nowak:Nature:2004, hauert2005game} describes the evolutionary dynamics of
populations consisting of different types of interacting individuals. In the
last two decades the interest of the field has shifted to the analysis of
stochastic finite population dynamics in preference to traditional approaches
such as the replicator dynamics~\cite{hofbauer:JTB:1979, taylor:MATHBIO:1978}.
This is especially true in the topic of cooperation~\cite{hilbe:PNAS:2013,
hilbe:Nature:2018,glynatsi:SCR:2020}. In stochastic evolutionary dynamics
disadvantageous mutants have a small yet non-zero probability to reach fixation
and this can lead to fundamental changes in the results. As shown
in~\cite{nowak:Nature:2004}, using such dynamics a single cooperative strategy
can invade a population of defectors without any special modifications in
comparison to traditional approaches that one has to consider spatial
structure~\cite{nowak:Nature:1992} or when payoffs are subject to aggregate
shocks~\cite{fudenberg:JET:1992}.

Stochastic evolutionary dynamics model a finite population. Each
time step of the process consists of three phases; (1) the \textit{mutation
phase} (2) the \textit{game phase} (3) the \textit{update phase}. In the
mutation phase one individual from the population is chosen to switch to a new
mutant strategy with a probability \(\mu\). In the game phase individuals are
randomly matched with other individuals in the population, and they engage in a
repeated game where each subsequent turn occurs with a fixed probability
\(\delta\). The updating phase depends on the process. Two classes of
finite stochastic processes have been used extensively: (i) fitness-based
processes in which an individual chosen proportional to fitness reproduces and
the offspring replaces a randomly chosen individual~\cite{nowak:Nature:2004} or
(ii) \textit{pairwise comparison processes} in which a pair of individuals is
chosen and subsequently one of these individuals may adopt the strategy of the
other~\cite{traulsen2007pairwise}.

Similar to all other theoretical models, finite stochastic processes rest on a
set of assumptions. Namely, in the game phase it is assumed that players use
strategies with finite memory~\cite{Nowak1992tit, Baek2016}. For example, in a
repeated game between two individuals the actions of the players in each turn
are often determined by the action of the co-players in the previous turn. This
assumption is common as it allows for an explicit calculation of the players'
payoffs~\cite{sigmund2010calculus}. In addition, it is
assumed that individuals play many times and with all other players before
reproduction takes. So that in the updating phase the updating payoffs, and
subsequently the fitness, of individuals us given by the average payoffs of an
infinitely repeated game against the mean distribution of types in the
population.

The above two assumptions create a curious inconsistency; a player with a finite
memory in the game stage can recall infinite information regarding their
interactions and each interaction's outcomes in the update stage. Previous work
has explored the effects of constraining the interactions an individual has.
This can be done by letting individuals have a stochastic number of
interactions~\cite{roca:PhysicalReview:2006, sanchez:JTB:2005,
Traulsen:JTB:2007}, a small number of interactions, or even in the extreme case
a single interaction~\cite{Woelfing:JTB:2009}. However, no previous work
considers that not only the interactions are limited but also the information
regarding each interaction. To this end, we propose a framework in which
individuals, similar to the decisions at each turn, estimate their fitness based
on a minimum of information.

We first consider two extreme scenarios, the classical scenario where individual
has a perfect memory of all their interactions and the alternative scenario of
limited memory where individuals update their strategies only based on the very
last payoff they obtained. We demonstrate the effects of limited updating payoff
memory using a pairwise comparison process and the well studied game of the
prisoner's dilemma. We observe that individuals with limited memory tend to
adopt less generous strategies and they achieve less cooperation when
interacting in a prisoner's dilemma. We obtain similar results when we consider
that individuals update their strategies based on more information. More
specifically, up to the last two payoffs they obtained when interacting with up
to two different members of the population.

\section{Model Setup}\label{section:model}

% ToDo: Add in discussion

% A major focus of evolutionary game theory is the evolution of cooperation among
% selfish individuals. Sometimes evolutionary game theory even seems to be reduced
% to the evolution of cooperation [Arne2022]. “Why do we act altruistically to
% provide a benefit to others when there is a cost to ourselves?” is still a
% puzzling question amongst researchers [Glynatsi]. In this determinist
% formulation, a small fraction of cooperative individuals that have a lower
% fitness can never invade a population of defecting strategies.  The stochastic
% effect can lead to fundamental changes since disadvantageous mutants have a
% small yet non-zero probability to reach fixation in a finite population, and
% thus cooperation can emerge.

A pairwise comparison process starts with assigning all
individuals of the population the same strategy. Each elementary time step of
the process consists of the mutation phase, the game phase and the update phase.
In the game phase individuals are matched in pairs and that they participate in
a repeated 2 persons donation game; a special case of the prisoner's dilemma. In
the donation game there are two actions: cooperation (\(C\)) and defection
(\(D\)). By cooperating a player provides a benefit \(b\) to the other player at
their cost \(c\), with \(0 < c < b\). Thus the payoffs for a player in each turn
are,

\begin{equation}
    \begin{blockarray}{ccc}
        & \text{cooperate} & \text{defect} \\
        \begin{block}{c(cc)}
            \text{cooperate} & b - c & -c \\
            \text{defect} & b & 0 \\
        \end{block}
    \end{blockarray}.
\end{equation}

Let \(\mathbf{u} = (b-c, -c, b, 0)\) be payoffs in a vector format, and let
\(\mathcal{U} = \{r, s, t, p\}\) denote the set of feasible payoffs, where \(r\)
denotes the payoff of mutual cooperation, \(s\) the sucker's payoff, \(t\) the
temptation to defect payoff, and \(p\) the punishment payoff.

In repeated games there are infinite many strategies, however, similar to the
literature we will assume that individuals use reactive strategies. A reactive strategy considers only the
previous action of the other player, and thus, a reactive strategy \(s\) can be
written as a three-dimensional vector \(s=(y, p, q)\). The parameter \(y\) is
the probability that the strategy opens with a cooperation and \(p\), \(q\) are
the probabilities that the strategy cooperates given that the opponent
cooperated and defected equivalently. The play between a pair of reactive
strategies can be model as a Markov process with the transition matrix \(M\),

\begin{equation}\label{eq:transition_matrix}
  \input{static/matrix.tex}
\end{equation}

and the stationary vector \(\mathbf{v}(s_1,s_2)\)
which is the solution to \(\mathbf{v}(s_1,s_2) \cdot M 
= \mathbf{v}(s_1,s_2)\).

In the update stage two individuals are randomly selected. From the two
individuals, one  serves as the `learner' and the other as the `role model'. The
learner adopts the role model's strategy with a probability \(\rho\) given by,

\begin{equation} \label{Eq:rho}
    \rho(\pi_{L}, \pi_{RM}) = \frac{1}{1\!+\! e^{\!-\!\beta (\pi_{RM}\!-\! \pi_{L})}}.
\end{equation}

\(\pi_{L}\) and \(\pi_{RM}\) are the updating payoffs/fitness of the learner and the
role model respectively. The updating payoffs are a measure of how successful
individuals are in the current standing of the population. The parameter
\(\beta\) is known as the selection strength, namely, it shows how important the
payoff difference is when the learner is considering adopting the strategy of
the role model.

We will mainly assume that mutations are rare (\(\mu \rightarrow 0\)). In fact,
so rare that only two different strategies can be present in the population at
any given time. However, we also show in section~\ref{section:mutation} that the
main results hold for \(\mu \neq 0\). The case of low mutation is vastly adopted because it allows us to explicitly
calculate the fixation probability of a newly introduced mutant. At each step
one individual adopts a mutant strategy randomly selected from the set of
feasible strategies. The fixation probability \(\phi_{M}\) of the mutant
strategy can be calculated explicitly,

\begin{equation}\label{eq:appendix_fixation_probability}
    \varphi = \frac{1}{1+\sum\limits_{i=1}^{N-1}\prod\limits_k^i \frac{\lambda^-_k}{\lambda^+_k}},
\end{equation}

where \(\lambda^-_k, \lambda^+_k\) are the probabilities that the number of
mutants decreases and increases respectively, and \(k\) is the number of mutants. Depending on the fixation
probability \(\phi_{M}\) the mutant either fixes (becomes the new resident) or
goes extinct. Regardless, in the elementary time step another mutant strategy is
introduced to the population. We iterate this elementary population updating
process for a large number of mutant strategies and we record the resident
strategies at each time step.

The probabilities \(\lambda^-_k \text{ and } \lambda^+_k\) depend on the
updating payoffs of the mutant and the resident strategies. In the next
section we discuss the fitness of individuals in the case of perfect and
limited memory.

\subsection{Updating payoffs based Perfect and Limited Memory}

\subsubsection*{Perfect Memory}

In the perfect memory case an individual updates based on the average payoff
against each other member of the population, otherwise as expected payoffs. In
an infinitely repeated game ($\delta \rightarrow 1$) the payoff of a reactive
strategy \(s_1\) against the reactive strategy \(s_2\) can explicitly be
calculated using the stationary vector \(\mathbf{v}\) and the payoffs' vector as,

\[\langle\mathbf{v}(s_1,s_2),\mathbf{u}\rangle \quad.\]

In a population of size \(N\), there are \(k\) mutants and \(N - k\) residents,
whose strategies we denote respectively as \(s_M =(y_M, p_M, q_M)\) and \(s_R =
(y_R, p_R, q_R)\). The expected payoffs of a resident (\(\pi_R\)) and for a
mutant (\(\pi_M\)) are give by,

\begin{equation} \label{Eq:ExpPay}
  \begin{array}{lcrcr}
  \displaystyle \pi_R & = &\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(s_R,s_R),\mathbf{u}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(s_R,s_M),\mathbf{u}\rangle,\\[0.5cm]
  \displaystyle \pi_M & = &\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(s_M,s_R),\mathbf{u}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(s_M,s_M),\mathbf{u}\rangle.\\
  \end{array}
\end{equation}

The probabilities that the number of mutants decreases and increases,
\(\lambda^-_k\) and \(\lambda^+_k\), in the perfect memory case are defined as,

\begin{align}
  \lambda^-_k \!=\!\rho(\pi_M, \pi_R) \quad \text{ and } \quad \lambda^+_k \!=\!\rho(\pi_R, \pi_M).
\end{align}

\subsubsection*{Limited Memory}

In this case of limited memory we initially define the probability that a
reactive strategy receives the payoff $u\!\in\! \mathcal{U}$ in the very last
round of the game. This is given by Proposition~\ref{proposition:last_round}.
The proof can be found in the Appendix [ref].

\begin{Prop}\label{proposition:last_round} Consider a repeated game, with
    continuation probability $\delta$, between players with reactive strategies
    $s_1\!=\!(y_1, p_1, q_1$)  and $s_2\!=\!(y_2,p_2,q_2)$ respectively. Then
    the probability that the $s_1$ player receives the payoff $u\!\in\!
    \mathcal{U}$ in the very last round of the game is given by
    $v_{u}(s_1,s_2)$, as given by Equation~(\ref{Eq:LastRound}).

    \begin{equation} \label{Eq:LastRound}
      \setlength{\arraycolsep}{1pt}
      \begin{array}{rcl}

      v_{r}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(q_1+l_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+l_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
      {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)} \times r,\\[1cm]

      v_{s}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(q_1+l_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
      {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)} \times s,\\[1cm]

      v_{t}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+l_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
      {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)} \times t,\\[1cm]

      v_{p}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
      {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)} \times p.
      \end{array}
    \end{equation}

In these expressions, we have used the notation $l_i:=p_i\!-\!q_i$,
$\bar{y}_i\!=\!1\!-\!y_i$, $\bar{q}_i:=1\!-\!q_i$, and
$\bar{l}_i:=\bar{p}_i\!-\!\bar{q}_i=-l_i$ for $i\!\in\!\{1,2\}$.
\end{Prop}

We consider the case where both the role model and the learner estimate their
fitness after interacting with a single member of the population. There are five
possible pairings at each step. They interact with other with a probability
\(\frac{1}{N - 1}\), and thus they do not interact with other with a probability
\(1 - \frac{1}{N - 1}\). In the latter case, each of them can interact with
either a mutant or a resident. Both of them interact with a mutant with a
probability $\frac{(k-1)(k-2)}{(N-2)(N-3)}$ and both interact with a resident
with a probability $\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$. The last two possible
pairings are that either of them interacts with a resident whilst the other
interacts with a mutant, and this happens with a probability
$\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$. The probability that the respective last
round payoffs of two players \(s_1, s_2\) are given by $u_1$ and $u_2$ as,

\begin{equation}\label{eq:Chi}
\setlength{\arraycolsep}{1pt}
\begin{array}{llrl}
x(u_1,u_2)	 &=&\displaystyle \frac{1}{N\!-\!1}\cdot  &v_{u_1}(s_1,s_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F}\\[0.5cm]
&+	
&\displaystyle \left(1\!-\!\frac{1}{N\!-\!1}\right)  
&\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_2) + 
 \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_1)\right.\\[0.5cm]
&&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_2) + 
 \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_1)\right].
\end{array}
\end{equation}

The first term on the right side corresponds to the case that the learner and
the role model happened to be matched during the game stage, which happens with
probability $\frac{1}{(N\!-\!1)}$. In that case, we note that only those payoff
pairs can occur that are feasible in a direct interaction,
$\splitatcommas{(u_1,u_2)\in \mathcal{U}^2_F:=\big\{ (r, r), (s, t), (t, s), (p, p)
\big\}}$, as represented by the respective indicator function. Otherwise, if the
learner and the role model did not interact directly, we need to distinguish
four different cases, depending on whether the learner was matched with a
resident or a mutant, and depending on whether the role model was matched with a
resident or a mutant.

The probability that the number of mutants increases, and decreases respectively,
by one is now given by,

\begin{align}
\lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M})\cdot \rho(u_{R},u_{M}), \\
\lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M})\cdot \rho(u_{M},u_{R}).
\end{align}

In this expression, $\frac{(N\!-\!k)}{N}$ is the probability that the randomly
chosen learner is a resident, and $\frac{k}{N}$ is the probability that the role
model is a mutant. The sum corresponds to the total probability that the learner
adopts the role model's strategy over all possible payoffs $u_R$ and $u_M$ that
the two player may have received in their respective last rounds.

% We use
% $x(u_R,u_M)$ to denote the probability that the randomly chosen resident
% obtained a payoff of $u_R$ in the last round of his respective game, and that
% the mutant obtained a payoff of $u_M$.

\subsubsection*{Simulation Results}

In order to account for the effect of the updating payoffs we simulate the
evolutionary process and record which strategies the players adopt over time
based on perfect memory payoffs and limited memory payoffs. We run two
independent runs for each approach and we vary the value of benefit \(b\).
Figure~\ref{fig:expected_and_stochastic_for_donation} shows the simulation
results. It depicts the evolving conditional cooperation probabilities $p$ and
$q$. The discount factor~$\delta$ is comparably high, thus we do not report the
opening move \(y\) as it is a transient effect. The upper panel correspond to
the standard scenario considered in the literature, it considers players who use
expected payoffs to update their strategies. The bottom panel shows the scenario
considered herein, in which players update their strategies based on their last
round's payoff.

The figure suggests that when updating is based on perfect memory payoffs
players tend to be more generous and more cooperative. The $q$-values of the
resident strategies are on average higher in the case of perfect memory. Thus,
players will occasionally forgive a defection more often if their fitness
depends on interacting with every member of the population, and this effects
becomes more abundant as the value of the benefit increases. In comparison, in
the case of limited memory the evolved generosity is independent of the values
of benefit, and they are less forgiving. We also calculate the average
cooperation rate for each simulation which is the average cooperation rate
within the resident population. In the case of the perfect memory payoffs the
average cooperation rate is strictly higher than that of the last round payoffs.
The difference based on the two methods is statistically significant, and in the
case of $b=10$ the average cooperation of resident strategies drops from 97\% to
57\%.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=\textwidth]{static/donation_expected_last_round_summary_results.pdf}
    \caption{{\bf Evolutionary dynamics under perfect and limited payoff memory.}
    ({\bf Schematic illustrations})
    On the left panels we show schematic illustrations of the perfect memory and
    the limited memory cases. The shaded background denotes the game phase
    information that an individual considered when updating strategies. In the
    case of perfect memory the entire region is shaded and in the case of
    limited memory only one turn with a single member of the population.
    ({\bf Simulations})
    We have run four simulations of the pairwise comparison process
    for $t\!=\!10^7$ time steps. For each time step,
    we have recorded the current resident population ($y,p,q$). Since
    simulations are run for a relatively high continuation probability of
    $\delta\!=\!0.999$, we do not report the players' initial cooperation
    probability $y$. The graphs show how often the resident population chooses
    each combination ($p,q$) of conditional cooperation probabilities in the
    subsequent rounds. We also report the evolved cooperation rate which is
    calculated as the average cooperation rate within the resident population.
    ({\bf Perfect Memory}) In the case of low benefit
    the resident population either consists of defectors (with $p\!\approx\!q\!\approx\!0$) or of
    conditional cooperators. Conditional cooperators, or otherwise known as generous
    tit for tat, are a set of strategies that always cooperate following a cooperation
    ($p\!\approx\!1\!$) and cooperate with a probability
    $q$ given that the co-player has defected. $q$ denotes the generosity
    of a player. The resident population applies a conditional cooperator strategy for which
    $q\!\le\!1\!-\!c/b\!=\!0.67$. This is true also for the case of high benefit.
    In this case the population mainly consists of conditional
    cooperators of the form ($p\!\approx\!1\!, q\!\le\!1\!-\!1/10\!=\!0.9$).
    In the Appendix we show that a conditional cooperator needs to be
    of the form ($p\!\approx\!1\!, q\!\le\!1\!-\!c/b$) to not be invaded by
    defecting strategies. A higher generosity in the population results to a
    higher average cooperation rate. The average cooperation rate
    increases from a 52\% for \(b\) to 98\% for \(b=10\).
    ({\bf Limited Memory}) When players
    update their strategies based on their realized payoffs in the last round,
    there are two different predominant behaviors regardless of the benefit value. The resident population
    either consists of defectors (with $p\!\approx\!q\!\approx\!0$) or of
    conditional cooperators. The maximum level of $q$
    consistent with stable cooperation is somewhat smaller compared to the
    perfect memory setting, $q\!<\!0.5$. Namely, in the Appendix we show
    that regardless the value of benefit, in the case of limited payoff memory,
    an individual needs generosity smaller of $0.5$
    to repel defectors. The plots
    are fairly similar regardless of the benefit, and the evolved cooperation
    rate only sightly increases from 37\% to 51\%. Parameters: $N\!=\!100$, $c\!=\!1$, $\beta\!=\!1$,
    $\delta\!=\!0.999$.}
    \label{fig:expected_and_stochastic_for_donation}
\end{figure}

We further explore the effect that the value of benefit \(b\) and the strength
of selection \(\beta\) have on the generosity \(q\) and to the evolved
cooperation rate, Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}. The
figure suggests that expected payoffs always yield a higher cooperation rate. In
the case of expected payoffs we observe that the cooperation rate increases as
the value of the benefit gets higher. In comparison for the limited memory
payoffs, the cooperation rate remains unchanged at approximately 50\% once
\(b=5\). We also investigate the effect of the strength of selection $\beta$.
For weak selection, \(\beta < 1\), we observe that the two methods yield similar
results, however, as \(\beta\) increases there is variation in the evolving
populations. In the case of expected payoffs the resident populations become
more cooperative as \(\beta\) increases, whereas in the case of limited memory
payoffs, the resident populations become more defective.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=\textwidth]{static/cooperation_rate_over_b_and_beta.pdf}
    \caption{{\bf The evolution of cooperation and generosity for different values
    of benefit (A) and strength of selection (B).} We report the
    average cooperation and the average reciprocity. The average cooperation rate
    is the average cooperation rate
    within the resident population. For the average reciprocity we check for the
    residents that ($p \approx 1$) and we take the average of theirs $q$s ({\bf A})
    We vary the benefit of defection $b$, and we report both the average
    cooperation and average generosity. In all cases, expected payoffs appear to
    overestimate the average cooperation rate the population achieves. As
    expected, since the average generosity over the different values of $beta$
    remains the same ($q \approx 0.5$), and a result so does the average
    cooperation. ({\bf B}) We vary the selection strength $\beta$. In all
    cases, stochastic payoff evaluation tends to reduce the evolving cooperation
    rates. Unless explicitly varied, the parameters of the simulation are
    $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$.
    Simulations are run for $t\!=\!5\times 10^7$ time steps for each parameter
    combination.}\label{fig:cooperation_rate_over_benefit_and_beta}
\end{figure}

% % \begin{figure}[!htbp]
% %   \centering
% %   \begin{subfigure}{.5\textwidth}
% %     \centering
% %     \includegraphics[width=\textwidth]{static/expected_for_beta.pdf}
% %     \includegraphics[width=\textwidth]{static/stochastic_for_beta.pdf}
% %   \end{subfigure}%
% %   \begin{subfigure}{.5\textwidth}
% %     \centering
% %     \includegraphics[width=\textwidth]{static/cooperation_rate_over_b.pdf}
% %   \end{subfigure}
% %   \caption{{\bf The evolution of cooperation for different benefit values.} 
% %   We vary the benefit of defection $b$. In all cases, expected payoffs appear to
% %   overestimate the average cooperation rate the population achieves. ({\bf A})
% %   the probabilities \(p, q\) for resident population over \(10^7\) time steps
% %   for each benefit value. ({\bf B}) The cooperation rate within the resident
% %   population (averaged over all games and over all time steps) over the benefit.
% %   Unless explicitly varied, the parameters of the simulation are $N\!=\!100$,
% %   $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
% %   $t\!=\!5\times 10^7$ time steps for each parameter
% %   combination.}\label{fig:cooperation_rate_over_benefit}
% % \end{figure}

% % \begin{figure}[!htbp]
% %   \centering
% %     \includegraphics[width=\textwidth]{static/cooperation_rate_over_b.pdf}
% % \caption{{\bf The evolution of cooperation for different selection strength values.}
% % We vary the selection strength $\beta$. In all cases, stochastic payoff
% % evaluation tends to reduce the evolving cooperation rates. ({\bf A}) the
% % probabilities \(p, q\) for resident population over \(10^7\) time steps for each
% % \(\beta\) value. ({\bf B}) The cooperation rate within the resident population
% % (averaged over all games and over all time steps) over \(\beta\). Unless
% % explicitly varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$,
% % $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
% % $t\!=\!5\times 10^7$ time steps for each parameter
% % combination.}\label{fig:cooperation_rate_over_betas}
% % \end{figure}

\subsection{Updating Payoffs based on More Memory}

So far we have explored the difference between the expected payoffs and the last
round payoffs. In order to explore further the effect of limited memory we allow
individuals to remember more. More precisely, up to two interactions and up to
the last two rounds. In total we present results for three more updating
payoffs. These are the payoffs when individuals consider the last two rounds
with another member of the population, the last round with two members of the
population, and the last two rounds with two members of the population. Similar
to the last round payoff, we use simulations and record which strategies the
players adopt over time based
% (Appendix~\ref{appendix:further_simulation_results}) and we compare the evolving
% cooperation rates to those of the expected payoffs
% (Figure~\ref{fig:cooperation_rates_two_by_two_part_two}).

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\textwidth]{static/more_memory_summary_results.pdf}
  \caption{{\bf Average cooperation rates for different updating payoffs.} (From
  right to left) updating Payoffs based  on perfect memory, limited memory,
  the last round payoff of two interactions,
  the last two rounds payoff of one interaction and 
  the last two round payoffs of two interactions.({\bf High Benefit})
  ({\bf Low Benefit}) We vary the selection strength $\beta$. In all
  cases, stochastic payoff evaluation tends to reduce the evolving cooperation
  rates. Unless explicitly varied, the parameters of the simulation are
  $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$.
  Simulations are run for $t\!=\!5\times 10^7$ time steps for each parameter
  combination.}\label{fig:cooperation_rate_updating_payoffs}
\end{figure}

\newpage
% In the
% simplest of cases each individual can be of one of two types, A and B. The state
% of the population is thus characterized by the number \(k\) of individuals of
% type A. The interaction between the two types of individuals is described by the
% functions \(\pi _{{\rm A}}^{i}\) and $\pi _{{\rm B}}^{i}$.

% airwise comparison process is a stochastic process for modelling the evolution
% of a finite population. The process starts with assigning all individuals of the
% population the same strategy. A strategy is a set of rules of how an individual
% should behave in an interaction with another individual. Each elementary time
% step of the process consists of three phases; (1) the \textbf{mutation phase}
% (2) the \textbf{game phase} and (3) the \textbf{update phase}. These are summarised
% in Figure~\ref{fig:pairwise_phases}.

% In the \textbf{mutation phase} one individual is
% chosen to switch to a new mutant strategy with a probability \(\mu\).
% In the \textbf{game phase} individuals are randomly matched with other
% individuals in the population, and they engage in a repeated game where each
% subsequent turn occurs with a fixed probability \(\delta\). At each turn the
% individuals decide on an action based on their strategies.
% In repeated games there are infinite many strategies, however, it is commonly
% assumed that individuals can only choose strategies from a restricted set. One
% such set is that of reactive strategies. A reactive
% strategy considers only the previous action of the other player, and thus, a
% reactive strategy \(s\) can be written as a three-dimensional vector \(s=(y, p,
% q)\). The parameter \(y\) is the probability that the strategy opens with a
% cooperation and \(p\), \(q\) are the probabilities that the strategy cooperates
% given that the opponent cooperated and defected equivalently.

% In the \textbf{update stage} where two individuals are randomly selected. From
% the two individuals, one  serves as the `learner' and the other as the `role
% model'. The learner adopts the role model's strategy with a probability \(\rho\)
% given by,

% The fitness of the individuals (\(\pi _{{\rm A}}^{i}, \pi _{{\rm B}}^{i}\)) is
% identified with the payoff resulting from the underline game. The evolution of
% cooperation remains one of the greatest problems for biological and social
% sciences. Cooperation is the action of choosing to help others at one's own
% expense. The standard game of formulating such situation is the Prisoner's
% Dilemma~\cite{trivers1971evolution, milinski1987tit, glynatsi2021bibliometric,
% ohtsuki2006simple}, in which two players can choose to cooperate or to defect.
% The players are offered a certain payoff, \(R\), for mutual cooperation and a
% lower payoff, \(P\), for mutual defection. If one player cooperates while the
% other defects, then the cooperator gets the lowest payoff, \(S\), while the
% defector gains the highest payoff, \(T\). Thus, the payoffs have the following
% property \(T > R > P > S\), making defection the dominant strategy in the
% non-repeated game.

% \begin{equation}\label{eq:payoffs}
%   \begin{blockarray}{ccc}
%       & \text{cooperate} & \text{defect} \\
%       \begin{block}{c(cc)}
%           \text{cooperate} & R & S \\
%           \text{defect} & T & P \\
%       \end{block}
%   \end{blockarray}
% \end{equation}

% In literature, the payoffs of A and B individuals only on the fraction of both
% types in the population. If there are \(k\) A individuals and \(N-k\) B
% individuals. In the context of the prisoner's dilemma the payoffs and assuming
% that A and B follow simple strategies where they cooperate and defect
% unconditionally respective, then the A and B individuals have fitness of

% \begin{equation}
%   \begin{array}{lcrcr}
%   \displaystyle \pi _{{\rm A}}^{i}	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot R	&+	&\displaystyle\frac{k}{N-1}\cdot S,\\[0.5cm]
%   \displaystyle \pi _{{\rm B}}^{i}&=	&\displaystyle\frac{N-k}{N-1}\cdot T &+	&\displaystyle\frac{k-1}{N-1}\cdot P.\\
%   \end{array}
% \end{equation}

% respectively while self-interactions are excluded. In literature these are the
% expected payoffs. For \(\beta = 1\), fitness equals payoff. This scenario
% describes ``strong selection''. For \(\beta \ll 1\) , the payoff only provides a
% small perturbation to the overall fitness of an individual, a limit known as
% weak selection.

% In the evolutionary game theory literature it is usually assumed that players
% use strategies with finite memory~\cite{Nowak1992tit, Baek2016}. This assumption
% is common as it allows for an explicit calculation of the players'
% payoffs~\cite{sigmund2010calculus}. Players' payoffs are calculated in the limit
% of interactions assuming that the two strategies meet several times.

% Customarily, most evolutionary game studies make the additional assumption that
% individuals play many times and with all other players before reproduction takes
% place, so that payoffs, equivalently fitness, are given by the mean distribution
% of types in the population. This hypothesis, selection occurs much more slowly
% than the interaction between individuals and more that remember all the
% interactions they participated in. Although recent experimental studies show
% that this may not always be the case in biology [12– 14], it is clear that in
% cultural evolution or social learning the time scale of selection is much closer
% to the time scale of interaction. The effects of this mixing of scales cannot be
% disregarded [15], and then it is natural to ask about the consequences of the
% above assumption and the effect of relaxing it. The work of~\cite{Roca2006}.

% In this work we ask studies that make use of this classical framework report
% that cooperation can substantially evolve. However, the inconsistency in the
% memory size of players leads us to question the robustness of our understanding
% of cooperation. To this end, we propose a framework in which individuals,
% similar to the decisions at each turn, estimate their fitness based on a minimum
% of information.

% We first consider two extreme scenarios, the classical scenario and the
% alternative scenario where individuals update their strategies only based on the
% very last payoff they obtained. We observe that individuals with limited memory
% tend to adopt less generous strategies and they achieve less cooperation when
% interacting in a prisoner's dilemma. We obtain similar results when we consider
% that individuals update their strategies based on more information. More
% specifically, up to the last two payoffs they obtained when interacting with up
% to two different members of the population. We extend our approach to the rest
% of the symmetric \(2 \times 2\) games.

% The remainder of the paper is organized as follows. In
% section~\ref{section:model} we describe the model. In
% section~\ref{section:results} we present the results of the simulations, and in
% section~\ref{section:conclusions} we outline the main conclusions.

% In the following, we consider a well mixed population of fixed size. In each
% step of t

% We study the transmission of strategies with a frequency-dependent birth-death
% process26 in a finite population of size n. In each time step, two randomly
% chosen individuals compare their payoffs and one of them can switch to the other
% one's strategy. This process can be interpreted as a model for social learning,
% whereby successful strategies spread, and, occasionally, random strategy
% exploration introduces novel strategies (corresponding to mutations in
% biological models).

% We consider a population of \(N\) players~\footnote{The terms ``player'' and
% ``individual'' are used interchangeably here.} where \(N\) is even and mutations
% are sufficiently rare. Therefore, at any point in time there are at most two
% different strategies present in the population; a \textit{resident} strategy and
% a \textit{mutant} strategy. To describe how strategies spread we use a pairwise
% comparison process~\cite{Traulsen2006}. Each step of the evolutionary process
% consists of two stages, a game stage and an updating stage.

% In the game stage each individual is randomly matched with some other individual
% in the population. They engage in a match where each subsequent turn occurs with
% a fixed probability $\delta$. At each turn players choose independently to
% either cooperate (\(C\)) or to defect (\(D\)), and the payoffs of the turn
% depend on both their decisions. If both players cooperate they receive the
% reward payoff \(R\), whereas if both defect they receive the punishment payoff
% \(P\). If one cooperates but the other defects, the defector receives the
% temptation payoff \(T\), whereas the cooperator receives the sucker's payoff
% \(S\). We denote the feasible payoff of each turn as \(\mathcal{U} = \{R, S, T,
% P\}\). We assume that individuals use \textit{reactive strategies} to make
% decisions in each turn. Reactive strategies are a set of memory-one strategies
% that only take into account the previous action of the opponent. They can be
% written explicitly as a vector in \(\R_{3}\), more specifically, a reactive
% strategy \(s\) is given by \(s=(y, p, q)\). The parameter \(y\) is the
% probability that the strategy opens with a cooperation and \(p, q\) are the
% probabilities that the strategy cooperates given that the opponent cooperated
% and defected equivalently.

% In the updating stage, two players are randomly drawn from the population, a
% `learner' and a `role model'. Given the learner's payoff $u_L\!\in\!
% \mathcal{U}$ and the role model's payoff $u_{RL}\!\in\! \mathcal{U}$, the
% learner adopts the role model's strategy with probability,

% % \begin{equation} \label{Eq:rho}
% %   \rho(u_{L}, u_{RM}) = \frac{1}{1\!+\! \exp^{\!-\!\beta (u_{RM}\!-\!u_{L})}}.
% % \end{equation}

% where $\beta\!\ge\!0$ is the strength of selection. For small values of $\beta$
% the imitation probability is independent of the strategies of the involved
% players. As the value of $\beta$ increases, the more likely it is that the
% learner adopts only strategies that yield a higher payoff. Conventionally the
% updating payoffs of the learner and the role model are based on their expected
% payoffs. A player's expected payoff is the mean payoff the player yields after
% engaging in matches of multiple turns with each member of the population. At
% each match a player bases their next turn decision only on the previous action
% of the opponent, however, the same player bases their expected payoffs on the
% outcomes of all their matches. Thus, a player is assumed to have limited and
% perfect memory at the same time. We propose a new a set of updating payoffs
% where it is also assumed that a player has limited memory. We referee to these
% as the limited memory payoffs.

% The evolutionary step is repeated until either the mutant strategy goes extinct,
% or until it fixes in the population. If the mutant fixes in the population then
% the mutant strategy becomes the new resident strategy. After either outcome we
% introduce a new mutant strategy uniformly chosen from all reactive strategies at
% random, and we set the number of mutants to $1$. This process of mutation and
% fixation/extinction is then iterated many times.

% \section{Results}\label{section:results} \subsection{Updating payoffs based on
% the last round with another member of the population}\label{section:donation}

% To investigate the role of extortion in the context of evolutionary games, we
% concentrate on the donation game.
% In this section we explore the case where the updating payoffs are based on the
% last round payoff achieved against another member of the population and we
% compare this to the classical scenario of the expected payoffs. We assume that
% each pair of players interacts in a donation game. The donation game is a
% special case of the prisoner's dilemma. Each player can choose to cooperate by
% providing a benefit \(b\) to the other player at their cost \(c\), with \(0 < c
% < b\). Thus, the feasible payoffs in each round are \(\mathcal{U} = \{b-c, -c,
% b, 0\}\).

% \begin{figure}[!htbp]
%   \centering
%   \includegraphics[width=\textwidth]{static/cooperation_rates_expected_and_stochastic_for_donation_game.pdf}
% \end{figure}

% We further explore the effect of benefit in
% Figure~\ref{fig:cooperation_rate_over_benefit}. The figure suggests that
% expected payoffs always yield a higher cooperation rate. In the case of expected
% payoffs we observe that the cooperation rate increases as the value of the
% benefit gets higher. In comparison for the limited memory payoffs, the
% cooperation rate remains unchanged at approximately 50\% once \(b=5\).

% % \begin{figure}[!htbp]
% %   \centering
% %   \begin{subfigure}{.5\textwidth}
% %     \centering
% %     \includegraphics[width=\textwidth]{static/expected_for_beta.pdf}
% %     \includegraphics[width=\textwidth]{static/stochastic_for_beta.pdf}
% %   \end{subfigure}%
% %   \begin{subfigure}{.5\textwidth}
% %     \centering
% %     \includegraphics[width=\textwidth]{static/cooperation_rate_over_b.pdf}
% %   \end{subfigure}
% %   \caption{{\bf The evolution of cooperation for different benefit values.} 
% %   We vary the benefit of defection $b$. In all cases, expected payoffs appear to
% %   overestimate the average cooperation rate the population achieves. ({\bf A})
% %   the probabilities \(p, q\) for resident population over \(10^7\) time steps
% %   for each benefit value. ({\bf B}) The cooperation rate within the resident
% %   population (averaged over all games and over all time steps) over the benefit.
% %   Unless explicitly varied, the parameters of the simulation are $N\!=\!100$,
% %   $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
% %   $t\!=\!5\times 10^7$ time steps for each parameter
% %   combination.}\label{fig:cooperation_rate_over_benefit}
% % \end{figure}

% We also investigate the effect of the strength of selection $\beta$.
% Figure~\ref{fig:cooperation_rate_over_betas} illustrates results for various
% runs of the evolutionary process. For weak selection, \(\beta < 1\), we observe
% that the two methods yield similar results, however, as \(\beta\) increases
% there is variation in the evolving populations. In the case of expected payoffs
% the resident populations become more cooperative as \(\beta\) increases, whereas
% in the case of limited memory payoffs, the resident populations become more
% defective.

% % \begin{figure}[!htbp]
% %   \centering
% %     \includegraphics[width=\textwidth]{static/cooperation_rate_over_b.pdf}
% % \caption{{\bf The evolution of cooperation for different selection strength values.}
% % We vary the selection strength $\beta$. In all cases, stochastic payoff
% % evaluation tends to reduce the evolving cooperation rates. ({\bf A}) the
% % probabilities \(p, q\) for resident population over \(10^7\) time steps for each
% % \(\beta\) value. ({\bf B}) The cooperation rate within the resident population
% % (averaged over all games and over all time steps) over \(\beta\). Unless
% % explicitly varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$,
% % $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
% % $t\!=\!5\times 10^7$ time steps for each parameter
% % combination.}\label{fig:cooperation_rate_over_betas}
% % \end{figure}

% \begin{figure}[!htbp]
%   \centering
%     \includegraphics[width=\textwidth]{static/more_memory_summary_results}
% \caption{{\bf The evolution of cooperation for different selection strength values.}
% We vary the selection strength $\beta$. In all cases, stochastic payoff
% evaluation tends to reduce the evolving cooperation rates. ({\bf A}) the
% probabilities \(p, q\) for resident population over \(10^7\) time steps for each
% \(\beta\) value. ({\bf B}) The cooperation rate within the resident population
% (averaged over all games and over all time steps) over \(\beta\). Unless
% explicitly varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$,
% $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
% $t\!=\!5\times 10^7$ time steps for each parameter
% combination.}\label{fig:cooperation_rate_over_betas}
% \end{figure}

% \subsection{Effect of updating payoffs in different social
% dilemmas}\label{section:2_by_2_games}

% In the previous section we gained insights into the effects of the updating
% payoffs, and into how parameters such as the benefit and the strength of
% selection can intensify them. We investigated these effects by using the
% donation game. In order to broaden our understanding of the updating payoffs on
% different forms of possible human interactions we extend our approach to all \(2
% \times 2\) symmetric games. More specifically, we apply our analysis to four
% different classes of games; the harmony, the snowdrift, the prisoner's dilemma
% and the stag-hunt games.

% We compare the results of the evolutionary process when the updating payoffs are
% based on the expected payoffs, and on the last round payoffs for all the four
% possible classes of games (Figures~\ref{fig:expected_two_by_two}
% and~\ref{fig:last_round_two_by_two}). In Figure~\ref{fig:expected_two_by_two}
% each sub figure represents a run of evolutionary process for a different set of
% values for \(\mathcal{U}\). Without loss of generality we set \(R=1\) and
% \(P=0\)~\cite{Martinez2012, Roca2009}, and we vary the values of the temptation
% payoff \(T\) (across the \(x\)-axis) and of the sucker's payoff \(S\) (across
% the \(y\)-axis). Starting at the upper left corner and proceeding clockwise the
% quadrants correspond to the harmony, the snowdrift, the prisoner's dilemma and
% the stag-hunt games.

% % % \begin{figure}[!htbp]
% % %   \centering
% % %   \includegraphics[width=\textwidth]{static/expected_two_by_two_games.pdf}
% % %   \caption{{\bf Evolutionary dynamics under expected payoffs for various social dilemmas.}
% % %   We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across
% % %   the \(x\) axis, and  \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\) across the
% % %   \(y\) axis. ({\bf A}) The top left quadrant where \(S > 0\) and \(T \leq 1\)
% % %   corresponds to the harmony game. The preference ordering for the harmony game
% % %   is \(R > T > S > P\). ({\bf B}) The top right quadrant where \(S > 0\) and \(T
% % %   > 1\) corresponds to the snowdrift game. The preference ordering for the
% % %   snowdrift game is \(T > R > S > P\). ({\bf C}) The bottom left quadrant where
% % %   \(S < 0\) and \(T \leq 1\) corresponds to the stag-hunt game. The preference
% % %   ordering for the stag-hunt game is \(R > T > P > S\). ({\bf D}) The bottom
% % %   right quadrant where \(S < 0\) and \(T > 1\) corresponds to the prisoner's
% % %   dilemma game. The preference ordering for the prisoner's dilemma game is \(T >
% % %   R > P > S\). Unless explicitly varied, the parameters of the simulation are
% % %   $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}
% % %   \label{fig:expected_two_by_two}
% % % \end{figure}

% % % \begin{figure}[!htbp]
% % %   \centering
% % %   \includegraphics[width=\textwidth]{static/stochastic_two_by_two_games.pdf}
% % %   \caption{{\bf Evolutionary dynamics under last round payoffs for various social dilemmas.} 
% % %   We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across
% % %   the \(x\) axis, and  \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\) across the
% % %   \(y\) axis. ({\bf A}) The top left quadrant where \(S > 0\) and \(T \leq 1\)
% % %   corresponds to the harmony game. The preference ordering for the harmony game
% % %   is \(R > T > S > P\). ({\bf B}) The top right quadrant where \(S > 0\) and \(T
% % %   > 1\) corresponds to the snowdrift game. The preference ordering for the
% % %   snowdrift game is \(T > R > S > P\). ({\bf C}) The bottom left quadrant where
% % %   \(S < 0\) and \(T \leq 1\) corresponds to the stag-hunt game. The preference
% % %   ordering for the stag-hunt game is \(R > T > P > S\). ({\bf D}) The bottom
% % %   right quadrant where \(S < 0\) and \(T > 1\) corresponds to the prisoner's
% % %   dilemma game. The preference ordering for the prisoner's dilemma game is \(T >
% % %   R > P > S\). Unless explicitly varied, the parameters of the simulation are
% % %   $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}
% % %   \label{fig:last_round_two_by_two}
% % % \end{figure}

% % % The harmony game describes an interaction without conflict. In the harmony game
% % % it is in the best interest for both players to cooperate, and in an evolutionary
% % % process cooperators prevail. We observe that for most harmony game runs the
% % % resident population overwhelmingly applies a strategy for which \(p\) and \(q\)
% % % are \(\approx 1\) (Figure~\ref{fig:expected_two_by_two}). The snowdrift game
% % % describes a situation similar to that of the prisoner's dilemma; cooperation
% % % results in a benefit to the opposing player but entailing a cost to the
% % % cooperator. However, in the snowdrift game individuals obtain immediate direct
% % % benefits from the cooperative acts which leads to \(S > P\). In the snowdrift
% % % game more cooperation emerges compared to the prisoner's dilemma. Defection is
% % % never a resident strategy and for the same values of temptation the overall
% % % \(q-\) values are higher. The last class of games we present are for the
% % % stag-hunt game. In the stag-hunt game there is an equilibrium in which both
% % % players cooperate as well as one in which both defect. For small values of the
% % % temptation and the sucker's payoffs we observe that cooperation prevails similar
% % % to the harmony game. However, in the case where the temptation and the reward
% % % payoffs are equal the resident population either consists of defectors or
% % % cooperators.

% % % Although the patterns of the evolved populations remain similar if we consider
% % % that players update their strategies based on their last round payoff
% % % (Figure~\ref{fig:last_round_two_by_two}), there are still dissimilarities
% % % between the two approaches. To better understand these dissimilarities, we
% % % explore the cooperation rates for each of the cases of the social dilemmas we
% % % have presented (Figure~\ref{fig:cooperation_rates_two_by_two_part_one}).

% % % In the harmony game the cooperation rates for both approaches are similar. The
% % % biggest difference occurs for \(T=1\), nevertheless, it is less than 10\%. There
% % % are two instances for which the cooperation rates are higher for the limited
% % % memory payoffs compared to the expected. These are for \(T=2.2, S=0.4\) and for
% % % \(T=2.2, S=0.8\). It is only in the snowdrift game that cooperation is
% % % overestimated when a minimum of information is considered. Players are slightly
% % % more likely to cooperate after a cooperation and to forgive after being at the
% % % receiving end of a defection. For the rest of the cases the classical scenario
% % % overestimates the average cooperation rate, supporting the results of the
% % % previous section. There is a substantial difference in the cooperation rates for
% % % the prisoner's dilemma class. We observe that for \(S<-0.4\) cooperation almost
% % % never emerges if players imitate based only on the last round payoff.

% % % \begin{figure}[!htbp]
% % %   \centering
% % %   \includegraphics[width=.4\textwidth]{static/expected_two_by_two_games_cooperation.pdf}
% % %   \includegraphics[width=.4\textwidth]{static/stochastic_two_by_two_games_cooperation.pdf}
% % %   \caption{\textbf{Cooperation rates for various social dilemmas.} ({\bf A}) If
% % %   players update based on their expected payoffs. ({\bf B}) If players update
% % %   based on their last round payoffs. We vary the temptation payoff \(T \in
% % %   \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across the \(x\) axis, and \(S \in \{1.2,
% % %   0.8, 0.4, 0, -0.4, -0.8\}\) across the \(y\) axis. Unless explicitly varied,
% % %   the parameters of the simulation are $N\!=\!100$, $\beta\!=\!10$,
% % %   $\delta\!=\!0.99$, $t\!=\!5\times
% % %   10^7$.}\label{fig:cooperation_rates_two_by_two_part_one}
% % % \end{figure}

% % % So far we have explored the difference between the expected payoffs and the last
% % % round payoffs. In order to explore further the effect of limited memory we allow
% % % individuals to remember more. More precisely, up to two interactions and up to
% % % the last two rounds. In total we present results for three more updating
% % % payoffs. These are the payoffs when individuals consider the last two rounds
% % % with another member of the population, the last round with two members of the
% % % population, and the last two rounds with two members of the population. Similar
% % % to the last round payoff, we use simulations and record which strategies the
% % % players adopt over time based
% % % (Appendix~\ref{appendix:further_simulation_results}) and we compare the evolving
% % % cooperation rates to those of the expected payoffs
% % % (Figure~\ref{fig:cooperation_rates_two_by_two_part_two}).

% % % In the case of the last two rounds payoff we make two observations. Initially,
% % % for the harmony, the stag-hunt and the prisoner's dilemma classes the average
% % % cooperation rates are higher than the previous case of the last round payoff.
% % % The rates remain strictly less than in the expected payoffs, however, in the
% % % case of the prisoner's dilemma we observe a significant increase. Secondly, for
% % % the snowdrift class there are more instances for which the cooperation rates are
% % % higher than in the expected payoffs. This could indicate that in the case of the
% % % snowdrift game the expected payoffs underestimate cooperation. Once players
% % % interact with two other members of the population, instead of one, the expected
% % % payoffs strictly yield a higher cooperation rate  even in the case of the
% % % snowdrift game. Similar to the previous result, the difference between the two
% % % approaches is always higher for the prisoner's dilemma class.

% % % \begin{figure}[!htbp]
% % %   \centering
% % %   \includegraphics[width=.33\textwidth]{static/rounds_two_by_two_games_cooperation.pdf}
% % %   \includegraphics[width=.33\textwidth]{static/opponents_two_by_two_games_cooperation.pdf}
% % %   \includegraphics[width=.33\textwidth]{static/rounds_opponents_two_by_two_games_cooperation.pdf}
% % %   \caption{\textbf{Cooperation rates over various social dilemmas for different
% % %   limited memory payoffs.} ({\bf A}) If players update based on their last two
% % %   rounds with another member of the population. ({\bf B}) If players update
% % %   based on their last round with two other members of the population. ({\bf C})
% % %   If players update based on their last two rounds with two other members of the
% % %   population. We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8,
% % %   2.2\}\) across the \(x\) axis, and \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\)
% % %   across the \(y\) axis. Unless explicitly varied, the parameters of the
% % %   simulation are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times
% % %   10^7$..}
% % %   \label{fig:cooperation_rates_two_by_two_part_two}
% % % \end{figure}

% \section{Conclusions}\label{section:conclusions}

% Cooperation can be seen at odd, why is it that we choose to help others,
% increasing their payoff, at the expense of decreasing one's own? In spite of all
% the selfish genes’ animal and human communities seem to altruistically help each
% other and cooperate, and evolutionary game theory has helped us shape our
% understanding of the evolution of cooperation.

% Previous evolutionary models often feature a curious inconsistency. When
% modeling how individuals make decisions in each round, these models assume that
% players only remember the last round. However, when modeling how individuals
% update their strategies over time, individuals are assumed to have perfect
% memory.

% Here, we have explored how robust cooperation is as models deviate from the
% perfect memory assumption. Initially we considered the donation game. We showed
% that when the last round payoff is used instead of the expected payoffs,
% cooperation can even evolve when individuals only use a minimum of information,
% however, the evolving cooperation rates are typically lower. The resident
% strategies were both less cooperative and less generous. This effect was only
% intensified we increase the benefit and the strength of selection independently.
% The results showed that as each parameter was being increased the difference
% between the cooperation rates were widening. This indicates that cooperative
% players benefit from being able to interact with everyone in the population.

% We extended our approach and presented results not only based on different
% classes of social games, but also based on different limited memory payoffs. The
% analysis showed that specifically for the case of the prisoner's dilemma games
% the contrast between the two approaches can be significantly large. The
% prisoner's dilemma is one of the most applied types of social games when we
% discuss the evolution of cooperation. Our results indicate that cooperation
% struggles to evolve in the prisoner's dilemma when only a minimum of social
% information is used at the updating stage. Interestingly our simulations also
% showed that in same cases cooperation can benefit from limited memory. This was
% specific for the snowdrift game.

% \section{Acknowledgements}

% This work was supported by the European Research Council Starting Grant 850529:
% E-DIRECT.

% \appendix

% \section{Model Setup}\label{appendix:methods}

% Consider a population of \(N\) individuals where \(N\) is even. At any point in
% time there are at most two different strategies in present in the population.
% More specifically, a mutant strategy played by \(k\) individuals and a resident
% strategy played by \(N - k\) individuals. We assume a pairwise process in which
% strategies spread because they are imitated more often. Each step of the
% evolutionary process consists of two stages; a game stage and an update stage.

% In the game stage, each individual is randomly matched with some other
% individual in the population. Their interaction lasts for a number of turns
% which is not fixed but depends on the continuation probability \(\delta\). At
% each turn the individuals choose between cooperation (\(C\)) and defection
% (\(D\)). Thus, there are four possible outcomes in each turn \(CC, CD, DC\) and
% \(DD\). If both players cooperate they receive the reward payoff \(R\), whereas
% if both players defect they receive the punishment payoff \(P\). If one
% cooperates but the other defects, the defector receives the temptation to
% defect, \(T\), whereas the cooperator receives the sucker's payoff, \(S\). Let
% $\mathcal{U}=\{R,S,T,P\}$ denote the set of feasible payoffs in each round, and
% let $\mathbf{u}\!=\!(R,S,T,P)$ be the corresponding payoff vector. We present
% results for various values of $\mathcal{U}$ for all the symmetric \(2 \times 2\)
% games.

% A further assumption of our model is that individuals make use of reactive
% strategies when they make decisions in each round. Reactive strategies are a set
% of strategies that take into account only the previous action of the opponent. A
% reactive strategy can be written explicitly as a vector,

% \[s=(y, p, q)\]

% where \(y\) is the probability that the strategy opens with a cooperation and
% \(p, q\) are the probabilities that the strategy cooperates given that the
% opponent cooperated and defected equivalently.

% In the updating stage, two players are randomly drawn from the population, a
% `learner' and a `role model'. The learner adopts the role model's strategy based
% on the Fermi distribution function, %ToDo add reference

% % \begin{equation} \label{Eq:rho}
% % \rho(u_{L}, u_{RM}) = \frac{1}{1\!+\! \exp^{\!-\!\beta (u_{RM}\!-\!u_{L})}}.
% % \end{equation}

% where $u_{L}\!\in\! \mathcal{U}$ is the learner's payoff, $u_{RM}\!\in\!
% \mathcal{U}$ is the role model's payoff, and $\beta\!\ge\!0$ is the strength of
% selection.

% We iterate this basic evolutionary step until either the mutant strategy goes
% extinct, or until it fixes in the population and becomes the new resident
% strategy. After either outcome, we set $k$ to $1$ and we introduce a new mutant
% strategy which is uniformly chosen from all reactive strategies at random.
% Instead of simulating each step of the evolutionary process, we estimate the
% probability that a newly introduced mutant fixes~\cite{nowak2004emergence}. This
% is defined as the fixation probability of the mutant, and the standard form is
% the following,

% \begin{equation}\label{eq:appendix_fixation_probability}
% \varphi = \frac{1}{1+\sum\limits_{i=1}^{N-1}\prod\limits_k^i \frac{\lambda^-_k}{\lambda^+_k}},
% \end{equation}

% where \(\lambda^-_k, \lambda^+_k\) are the probabilities that the number of
% mutants decreases and increases respectively.

% This process of mutation and fixation/extinction is iterated many times. The
% evolutionary process is summarized by
% Algorithm~\ref{algorithm:pairwise_comparison}.

% \begin{algorithm}[!htbp]
%   \SetAlgoLined $N \leftarrow$ population size\; $k \leftarrow 1$\; resident
%    $\leftarrow (0, 0, 0)$\; \While{ $t <$ maximum number of steps}{mutant
%    $\leftarrow$ random: $\{\emptyset \}\ \rightarrow R^3$\; fixation probability
%    $\leftarrow \varphi $\; \If{$\varphi >$ random: $i \rightarrow [0,
%    1]$}{resident $\leftarrow$ mutant;}} \caption{Evolutionary
%    process}\label{algorithm:pairwise_comparison}
% \end{algorithm}

% The aim of this work is to explore the effect of updating memory on the
% cooperation rate of the evolved population. For this reason we consider two
% different approaches when estimating the payoffs at the updating stage. The two
% approaches we consider are those of (i) the expected and (ii) the limited memory
% payoffs.

% \subsection*{Expected Payoffs}

% The expected payoffs are the conventional payoffs used in the updating
% stage~\cite{imhof2010stochastic}. They are defined as the mean payoff of an
% individual in a well-mixed population that engages in repeated games with all
% other population members.

% We first define the payoff of two reactive strategies at the game stage. Assume
% two reactive strategies $s_1\!=\!(y_1, p_1, q_1$) and $s_2\!=\!(y_2,p_2,q_2)$.
% It is not necessary to simulate the play move by move, instead the play between
% the two strategies is defined a Markov matrix \(M\),

% \begin{equation}\label{eq:transition_matrix}
%   \input{static/matrix.tex}.
% \end{equation}

% whose stationary vector \(\mathbf{v}\), combined with the payoff \(u\), yields
% the game stage outcome for each strategy,
% \(\langle\mathbf{v}(s_1,s_2),\mathbf{u}\rangle\)~\cite{Hauert1997}.

% In the updating stage the learner adopts the strategy of the role model based on
% their updating payoffs. Given that there are only two different types in the
% population at each time step we only need to define the expected payoff for a
% resident (\(\pi_R\)) and for a mutant (\(\pi_M\)). Assume the resident strategy
% \(s_R = (y_R, p_R, q_R)\) and the mutant strategy \(s_M = (y_M, p_M, q_M)\), the
% expected payoffs are give by,

% \begin{equation} \label{Eq:ExpPay}
%   \begin{array}{lcrcr}
%   \displaystyle \pi_R	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(s_R,s_R),\mathbf{u}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(s_R,s_M),\mathbf{u}\rangle,\\[0.5cm]
%   \displaystyle \pi_M	&=	&\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(s_M,s_R),\mathbf{u}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(s_M,s_M),\mathbf{u}\rangle.\\
%   \end{array}
% \end{equation}

% The number of mutant in the population increase if a learner resident adopts the
% strategy of a mutant role model, and decreases if a mutant leaner adopts the
% strategy of a resident. The probabilities that the number of mutants decreases
% and increases, \(\lambda^-_k\) and \(\lambda^+_k\), are now explicitly defined
% as,

% \begin{align*} 
%   \lambda^-_k &\!=\!\rho(\pi_R, \pi_M) \\
%   \lambda^+_k &\!=\!\rho(\pi_M, \pi_R).
% \end{align*}

% \subsection*{Limited memory payoffs}

% Initially, we discuss the case of the \textbf{last round updating payoff}. At
% the stage game we define the payoff of a reactive strategy in the last round,
% Proposition~\ref{proposition:last_round}.

% \begin{Prop}\label{proposition:last_round} Consider a repeated game, with
%     continuation probability $\delta$, between players with reactive strategies
%     $s_1\!=\!(y_1, p_1, q_1$)  and $s_2\!=\!(y_2,p_2,q_2)$ respectively. Then
%     the probability that the $s_1$ player receives the payoff $u\!\in\!
%     \mathcal{U}$ in the very last round of the game is given by
%     $v_{u}(s_1,s_2)$, as given by Equation~(\ref{Eq:LastRound}).

%     \begin{equation} \label{Eq:LastRound}
%       \setlength{\arraycolsep}{1pt}
%       \begin{array}{rcl}

%       v_{R}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
%       {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)} \times R,\\[1cm]

%       v_{S}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
%       {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)} \times S,\\[1cm]

%       v_{T}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
%       {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)} \times T,\\[1cm]

%       v_{P}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
%       {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)} \times P.
%       \end{array}
%     \end{equation}

% In these expressions, we have used the notation $r_i:=p_i\!-\!q_i$,
% $\bar{y}_i\!=\!1\!-\!y_i$, $\bar{q}_i:=1\!-\!q_i$, and
% $\bar{r}_i:=\bar{p}_i\!-\!\bar{q}_i=-r_i$ for $i\!\in\!\{1,2\}$.
% \end{Prop}

% \begin{proof}
% Given a play between two reactive strategies with continuation probability
% $\delta$. The outcome at turn \(t\) is given by,

% \begin{equation}\label{eq:}
%   (1 - \delta) \mathbf{v_0} \sum \delta^{t} M^{(t)},
% \end{equation}

% where $\mathbf{v_0}$ denotes the expected distribution of the four outcomes in
% the very first round, and \(1- \delta\) the probability that the game ends. It
% can be shown that,

% \begin{align*}
%   (1 - \delta) \mathbf{v_0} \sum \delta^{t} M^{(t)} & = (1 - \delta)(\mathbf{v_0} + \delta \mathbf{v_0} M + \delta^{2}\mathbf{v_0} M ^{2} + \dots )\\ 
%    & = (1 - \delta)\mathbf{v_0} (1 + \delta M + \delta^{2}M ^{2} + \dots ) \text{ using standard formula for geometric series}\\ 
%    & = (1 - \delta)\mathbf{v_0}(I_4 - \delta M)^{-1}
% \end{align*}

% where \((1 - \delta)\mathbf{v_0}(I_4 - \delta M)^{-1}\) is vector \(\in R^{4}\)
% and it the probabilities for being in any of the outcomes \(CC, CD, DC, DD\) in
% the last round. Combining this with the payoff vector \(u\) and some algebraic
% manipulation we derive to the Equation~\ref{Eq:LastRound}.
% \end{proof}

% In the updating stage we select a mutant and resident to be either the role
% model or the learner. Given that they can interact with only one other member of
% the population, they can interact either with each other or either can interact
% with another resident or with another mutant. Thus, in each updating stage there
% are five possible combinations of pairs (Figure~\ref{fig:single_pairs}).

% Given the last round payoff and possible pair combinations for a single
% interaction, we define the probability that the respective last round payoffs of
% two players \(s_1, s_2\) are given by $u_1$ and $u_2$ as,

% \begin{equation}\label{eq:Chi}
% \setlength{\arraycolsep}{1pt}
% \begin{array}{llrl}
% x(u_1,u_2)	 &=&\displaystyle \frac{1}{N\!-\!1}\cdot  &v_{u_1}(s_1,s_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F}\\[0.5cm]
% &+	
% &\displaystyle \left(1\!-\!\frac{1}{N\!-\!1}\right)  
% &\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_2) + 
%  \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_1)\right.\\[0.5cm]
% &&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_2) + 
%  \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_1)\right].
% \end{array}
% \end{equation}

% The first term on the right side corresponds to the case that the learner and
% the role model happened to be matched during the game stage, which happens with
% probability $\frac{1}{(N\!-\!1)}$. In that case, we note that only those payoff
% pairs can occur that are feasible in a direct interaction,
% $\splitatcommas{(u_1,u_2)\in \mathcal{U}^2_F:=\big\{ (R,R), (S,T), (T,S), (P,P)
% \big\}}$, as represented by the respective indicator function. Otherwise, if the
% learner and the role model did not interact directly, we need to distinguish
% four different cases, depending on whether the learner was matched with a
% resident or a mutant, and depending on whether the role model was matched with a
% resident or a mutant.

% Given that $N\!-\!k$ players use the resident strategy
% $s_{R}\!=\!(y_{R},p_{R},q_{R})$ and that the remaining $k$ players use the
% mutant strategy $s_{M}\!=\!(y_{M},p_{M},q_{M})$, the probability that the number
% of mutants increases by one in one step of the evolutionary process can be
% written as

% \begin{align}
% \lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M})\cdot \rho(u_{R},u_{M}), \\
% \lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M})\cdot \rho(u_{M},u_{R}).
% \end{align}

% In this expression, $\frac{(N\!-\!k)}{N}$ is the probability that the randomly
% chosen learner is a resident, and $\frac{k}{N}$ is the probability that the role
% model is a mutant. The sum corresponds to the total probability that the learner
% adopts the role model's strategy over all possible payoffs $u_R$ and $u_M$ that
% the two player may have received in their respective last rounds. We use
% $x(u_R,u_M)$ to denote the probability that the randomly chosen resident
% obtained a payoff of $u_R$ in the last round of his respective game, and that
% the mutant obtained a payoff of $u_M$.

% We extend our framework to consider the case where players update their
% strategies based on the outcome of \textbf{the last two turns and based on their
% interaction with two other members of the population}. At the stage game we
% define the payoff of a reactive strategy in the last two rounds,
% Proposition~\ref{proposition:last_two_rounds}.

% \begin{Prop}\label{proposition:last_two_rounds} Consider a repeated game, with
%   continuation probability $\delta$, between players with reactive strategies
%   $s_1\!=\!(y_1, p_1, q_1$)  and $s_2\!=\!(y_2,p_2,q_2)$ respectively. Let
%   $\mathcal{\tilde{U}} = \splitatcommas{\{RR, RS, RT, RP, SR, SS, ST, SP, TR,
%   TS, TT, TP, PR, PS, PT, PP\}}$ denote the set of feasible payoffs in the last
%   two rounds, and let \(\tilde{\mathbf{u}}\) be the corresponding payoff vector.
%   Then the probability that the $s_1$ player receives the payoff $u\!\in\!
%   \mathcal{\tilde{U}}$ in the very last two rounds of the game is given by,

%   \begin{equation}
%   \langle\mathbf{\tilde{v}}(s_1,s_2),\mathbf{\tilde{u}}\rangle, \text{ where } \mathbf{\tilde{v}} \in R^{16} \text{ is given by },
%   \end{equation}
%   \begin{equation}
%     \mathbf{\tilde{v}}(s_1,s_2) = (1 - \delta) m_{a_1, a_2} \delta^2 \left[\mathbf{v_0}(I_4 - \delta M)^{-1}\right]_{a_1, a_2}, \quad  m_{a_1, a_2} \in M \ \forall \ a_1, a_2 \in \{1, 2, 3, 4\}.
%   \end{equation}
% \end{Prop}

% In the updating stage we select a mutant and resident to be either the role
% model or the learner. Given that they can interact with two other members of the
% population there are a total of twenty four possible combinations of pairs
% (Figure~\ref{fig:pissible_two_pairs}).

% Given the last two rounds payoff and possible pair combinations with two
% members, we can define the probability \(x\) that the respective last round
% payoffs of two players \(s_1, s_2\) are given by $u_1$ and $u_2$ similarly to
% Eq.~(\ref{eq:Chi}). We follow the same approach to define the rest of the
% updating payoffs. These are the updating payoffs of the last round with two
% member, and the the last two rounds payoffs with one member.

% Simulating the evolutionary process for more interactions and rounds quickly
% becomes computationally intractable. Our methodology could be extended to
% include \(n\) turns and \(m\) interactions. However, for the purpose of this
% work we explore the cases only up to two turns and two interactions.

% \section{Evolutionary dynamics under limited memory payoffs for various social
% dilemmas}\label{appendix:further_simulation_results}

% \begin{figure}[!htbp]
%   \centering
%   \includegraphics[width=\textwidth]{static/rounds_two_by_two_games.pdf}
%   \caption{{\bf Evolutionary dynamics under last two rounds payoffs for various social dilemmas.} 
%   We have run several simulations of the evolutionary process described in
%   section~\ref{section:model} for $t\!=\!10^7$ time steps. The graphs show how
%   often the resident population chooses each combination ($p,q$) of conditional
%   cooperation probabilities in the subsequent rounds. We vary the temptation
%   payoff \(T \in \{-1, -0.6, -0.2,  0.2, 0.6, 1, 1.4, 1.8, 2.2, 2.6, 3\}\)
%   across the \(x\) axis, and  \(S \in \{2, 1.6, 1.2, 0.8, 0.4, 0, -0.4, -0.8,
%   -1.2, -1.6, -2\}\) across the \(y\) axis. Parameters: $N\!=\!100$,
%   $\beta\!=\!10$, $\delta\!=\!0.999$.}
%   \label{fig:last_two_rounds_two_by_two}
% \end{figure}

% \begin{figure}[!htbp]
%   \centering
%   \includegraphics[width=\textwidth]{static/opponents_two_by_two_games.pdf}
%   \caption{{\bf Evolutionary dynamics under last round payoffs with two members of the population for various social dilemmas.} 
%   We have run several simulations of the evolutionary process described in
%   section~\ref{section:model} for $t\!=\!10^7$ time steps. The graphs show how
%   often the resident population chooses each combination ($p,q$) of conditional
%   cooperation probabilities in the subsequent rounds. We vary the temptation
%   payoff \(T \in \{-1, -0.6, -0.2,  0.2, 0.6, 1, 1.4, 1.8, 2.2, 2.6, 3\}\)
%   across the \(x\) axis, and  \(S \in \{2, 1.6, 1.2, 0.8, 0.4, 0, -0.4, -0.8,
%   -1.2, -1.6, -2\}\) across the \(y\) axis. Parameters: $N\!=\!100$,
%   $\beta\!=\!10$, $\delta\!=\!0.999$.}
%   \label{fig:last_two_opponents_two_by_two}
% \end{figure}

% \begin{figure}[!htbp]
%   \centering
%   \includegraphics[width=\textwidth]{static/merged_plot_rounds_opponents_two.pdf}
%   \caption{{\bf Evolutionary dynamics under last two rounds payoffs with two members of the population for various social dilemmas.} 
%   We have run several simulations of the evolutionary process described in
%   section~\ref{section:model} for $t\!=\!10^7$ time steps. The graphs show how
%   often the resident population chooses each combination ($p,q$) of conditional
%   cooperation probabilities in the subsequent rounds. We vary the temptation
%   payoff \(T \in \{-1, -0.6, -0.2,  0.2, 0.6, 1, 1.4, 1.8, 2.2, 2.6, 3\}\)
%   across the \(x\) axis, and  \(S \in \{2, 1.6, 1.2, 0.8, 0.4, 0, -0.4, -0.8,
%   -1.2, -1.6, -2\}\) across the \(y\) axis. Parameters: $N\!=\!100$,
%   $\beta\!=\!10$, $\delta\!=\!0.999$.}
%   \label{fig:last_two_rounds_two_opponents}
% \end{figure}

\bibliographystyle{unsrt}
\bibliography{bibliography.bib}

\end{document}

### 2023-02-21 paper/si.tex
working on paper in folder paper

@@
 \definecolor{darkgreen}{rgb}{0,0.5,0.1}
+\definecolor{lightapricot}{rgb}{0.99, 0.84, 0.69}
 \newcommand{\matlabfunc}[1]{\textcolor{darkblue}{#1}}
@@
 \title{\bf  \sffamily \LARGE Supplementary Information: Evolution of cooperation among individuals with
-limited payoff memory\\}
+limited updating payoff memory\\}
 \date{}
@@
 Section~\ref{section:pairwise_comparison} gives a brief overview of the pairwise
-comparison process. Section~\ref{section:updating_payoffs} describes the
-conventional approach for calculating updating payoffs, as well as, our
-approaches and methods.
+comparison process. The pairwise comparison process consists of three phases;
+(1) the mutation phase (2) the game phase and (3) the update phase. In the update
+phase an individual adopts the strategy of another individual based on their
+``updating payoffs''. In Section~\ref{section:updating_payoffs} we
+present/describe the conventional approach for calculating updating payoffs, as
+well as, our newly introduced approach.
 
@@
 step of the process consists of three phases; (1) the \textbf{mutation phase}
-(2) the \textbf{game phase} (3) the \textbf{update phase}. These are summarised
-in Figure~\ref{fig:pairwise_phases}. In the mutation phase one individual is
-chosen to switch to a new mutant strategy with a probability \(\mu\). The mutant
-strategy is randomly selected from the set of feasible strategies.
+(2) the \textbf{game phase} and (3) the \textbf{update phase}. These are summarised
+in Figure~\ref{fig:pairwise_phases}.
 
-In the game phase individuals are randomly matched with other individuals in the
-population, and they engage in a match where each subsequent turn occurs with a
-fixed probability \(\delta\). At each turn the individuals play a social dilemma
-and decide on an action based on their strategies. In the donation game there
-are two actions: cooperation (\(C\)) and defection (\(D\)). By cooperating a player
-provides a benefit \(b\) to the other player at their cost \(c\), with \(0 < c <
-b\). Thus the payoffs for a player in each turn are,
+In the \textbf{mutation phase} one individual is
+chosen to switch to a new mutant strategy with a probability \(\mu\).
+In the \textbf{game phase} individuals are randomly matched with other
+individuals in the population, and they engage in a repeated game where each
+subsequent turn occurs with a fixed probability \(\delta\). At each turn the
+individuals decide on an action based on their strategies.
+In repeated games there are infinite many strategies, however, it is commonly
+assumed that individuals can only choose strategies from a restricted set. One
+such set is that of reactive strategies. A reactive
+strategy considers only the previous action of the other player, and thus, a
+reactive strategy \(s\) can be written as a three-dimensional vector \(s=(y, p,
+q)\). The parameter \(y\) is the probability that the strategy opens with a
+cooperation and \(p\), \(q\) are the probabilities that the strategy cooperates
+given that the opponent cooperated and defected equivalently.
 
-\begin{equation}
-    \begin{blockarray}{ccc}
-        & \text{cooperate} & \text{defect} \\
-        \begin{block}{c(cc)}
-            \text{cooperate} & b - c & -c \\
-            \text{defect} & b & 0 \\
-        \end{block}
-    \end{blockarray}.
-\end{equation}
-
-Let \(u = (b-c, -c, b, 0)\) be payoffs in a vector format, and let \(\mathcal{U}
-= \{b-c, -c, b, 0\}\) denote the set of feasible payoffs. In repeated games
-there are infinite many strategies. However, it is commonly assumed that
-individuals can only choose strategies from a restricted set. Here we explore
-the case where individuals can use reactive strategies. A reactive strategy
-considers only the previous action of the other player, and thus, a reactive
-strategy \(s\) can be written as a three-dimensional vector \(s=(y, p, q)\). The
-parameter \(y\) is the probability that the strategy opens with a cooperation
-and \(p\), \(q\) are the probabilities that the strategy cooperates given that
-the opponent cooperated and defected equivalently.
-
-Following the game phase is the update stage where two individuals are randomly selected.
-One individual serves as the `learner' and the other as the `role model'.
-The learner adopts the role model's strategy with a probability \(\rho\) given by,
+In the \textbf{update stage} where two individuals are randomly selected. From
+the two individuals, one  serves as the `learner' and the other as the `role
+model'. The learner adopts the role model's strategy with a probability \(\rho\)
+given by,
 
 \begin{equation} \label{Eq:rho}
-    \rho(\pi_{L}, \pi_{RM}) = \frac{1}{1\!+\! \exp^{\!-\!\beta (\pi_{RM}\!-\! \pi_{L})}}.
+    \rho(\pi_{L}, \pi_{RM}) = \frac{1}{1\!+\! e^{\!-\!\beta (\pi_{RM}\!-\! \pi_{L})}}.
 \end{equation}
@@
 \(\pi_{L}\) and \(\pi_{RM}\) are the updating payoffs of the learner and the
-role model respectively. The updating payoff is a measure of how successful an
-individual is in the current standing of the population. In the next section we
-will explain how these payoffs are calculated in more detail. The parameter
+role model respectively. The updating payoffs are a measure of how successful
+individuals are in the current standing of the population. The parameter
 \(\beta\) is known as the selection strength, namely, it shows how important the
@@
     \centering
-    \includestandalone[width=\textwidth]{pairwise_comparison_process_diagram}
-    \caption{\textbf{Pairwise process phases.} \textbf{A)} The process begins
+    \includestandalone[width=\textwidth]{static/pairwise_comparison_process_diagram}
+    \caption{\textbf{Pairwise comparison process phases.} \textbf{A) Initialisation} The process begins
     with a finite population where each member is assigned a given strategy.
@@
     example individual A) and with a given probability \(\mu\) that individual
-    adopts a new mutant strategy. \textbf{C) Game phase.} individuals are
+    adopts a new strategy. \textbf{C) Game phase.} Individuals are
     selected to interact in a repeated social dilemma with other individuals. We
@@
     individual F defected in turn 1, and so B cooperates in the next turn with
-    a probability \(q_{B}\). At each turn there is a probability \(delta\) that a
-    subsequent turn will occur, and with a probability \(1- delta\) the interaction ends.
-    \textbf{D) Updating phase.} At the updating phase two individuals are chosen;
-    one is assigned the role of the learner and the other one the role
-    of the role model. In our example C adopts D's strategy with a probability
-    \(\rho(\pi_{C}, \pi_{D})\) where \(pi_{C}, \pi_{D}\) are the updating
+    a probability \(q_{B}\). At each turn there is a probability \(\delta\) that a
+    subsequent turn will occur, and with a probability (\(1- \delta\)) the interaction ends.
+    \textbf{D) Update phase.} At the updating phase two individuals are chosen;
+    one serves as the role of the learner and the other one as the 
+    role model. In our example C adopts D's strategy with a probability
+    \(\rho(\pi_{C}, \pi_{D})\) where \(\pi_{C}, \pi_{D}\) denote the updating
     payoffs of the individuals.}\label{fig:pairwise_phases}
@@
 
-This elementary population updating process is repeated for a large number of
-time steps, and at each time step we record the state of the population.
+This elementary step of the process (mutation, game and update phases) is
+repeated for a large number of time steps, and at each time step we record the
+state of the population.
 
@@
   \SetAlgoLined $N \leftarrow$ population size\; $k \leftarrow 1$\; resident
-   $\leftarrow (0, 0, 0)$\; \While{ $t <$ maximum number of steps}{mutant
-   $\leftarrow$ random: $\{\emptyset \}\ \rightarrow R^3$\; fixation probability
+   $\leftarrow$ starting resident\; \While{ $t <$ maximum number of steps}{mutant
+   $\leftarrow$ random: $\{\emptyset \}\ \rightarrow R^n$\; fixation probability
    $\leftarrow \varphi $\; \If{$\varphi >$ random: $i \rightarrow [0,
@@
 Most of the results we present in this work consider the case of low mutation,
-however, we have also verified that the main result holds even in the case where
-\(\mu\) is not small. %ToDo Add reference to section with mu != 0.
+however, we have also verified that the main result holds in the case of
+high mutation rates. %ToDo Add reference to section with mu != 0.
 
@@
 
-In this section we discuss the conventional approach of calculating the updating
-payoffs, as well as our newly introduced approach.
+The updating payoffs depend on the interactions of the individuals at the game
+phase. In this work we assume that in the game phase individuals are matched in
+pairs and that they participate in a repeated 2 persons donation game. In the
+donation game there are two actions: cooperation (\(C\)) and defection (\(D\)).
+By cooperating a player provides a benefit \(b\) to the other player at their
+cost \(c\), with \(0 < c < b\). Thus the payoffs for a player in each turn are,
+
+\begin{equation}
+    \begin{blockarray}{ccc}
+        & \text{cooperate} & \text{defect} \\
+        \begin{block}{c(cc)}
+            \text{cooperate} & b - c & -c \\
+            \text{defect} & b & 0 \\
+        \end{block}
+    \end{blockarray}.
+\end{equation}
+
+Let \(\mathbf{u} = (b-c, -c, b, 0)\) be payoffs in a vector format, and let \(\mathcal{U}
+= \{r, s, t, p\}\) denote the set of feasible payoffs, where \(r\) denotes the
+payoff of mutual cooperation, \(s\) the sucker's payoff, \(t\) the temptation to
+defect payoff, and \(p\) the punishment payoff.
+
+In the following subsections we present several approaches for calculating
+the updating payoffs. Initially, we discuss the conventional approach of the
+expected payoffs and afterwards we present our newly introduced approach.
 
@@
 The expected payoffs are the conventional payoffs used in the updating stage.
-The expected payoffs are defined as the mean payoff of an individual in a
-well-mixed population that engages in repeated games with all other population
-members. We first define the payoff of two reactive strategies at the game
-stage. Assume two reactive strategies $s_1\!=\!(y_1, p_1, q_1$) and
-$s_2\!=\!(y_2,p_2,q_2)$. The play between the two strategies can be defined as a
-Markov process with the transition matrix \(M\),
+They are defined as the mean payoff of an individual in a well-mixed population
+that engages in an infinitely repeated games with all other population members. In
+an infinitely repeated game the payoff of a reactive strategy can explicitly be
+calculated using a markovian approach~\cite{}. Namely, assume two reactive
+strategies $s_1\!=\!(y_1, p_1, q_1$) and $s_2\!=\!(y_2,p_2,q_2)$, their play can
+be defined as a Markov process with the transition matrix \(M\),
 
 \begin{equation}\label{eq:transition_matrix}
-  \input{../static/matrix.tex}.
+  \input{static/matrix.tex}.
 \end{equation}
 
-whose stationary vector \(\mathbf{v}\), combined with the payoff \(u\), yields
-the game stage outcome for each strategy,
+Note that \(y_1 \text{ and } y_2\) are omitted since they have no effect in the
+long run of the game. The stationary vector \(\mathbf{v}\),
+which is the solution to \(\mathbf{v} M = \mathbf{v}\), combined with the payoff vector
+\(\mathbf{u}\), yields the game stage outcome for each strategy,
 
-\[\langle\mathbf{v}(s_1,s_2),\mathbf{u}\rangle \text{ and } \langle\mathbf{v}(s_2,s_1), \mathbf{u}\rangle\].
+\[\langle\mathbf{v}(s_1,s_2),\mathbf{u}\rangle \quad  \text{ and } \quad  \langle\mathbf{v}(s_2,s_1), \mathbf{u}\rangle.\]
 
-In the case of low mutation there can be only two different
-strategies in the population; a resident and a mutant strategy. Thus, we
-only need to define the expected
-payoff for a resident (\(\pi_R\)) and for a mutant (\(\pi_M\)). Assume the
-resident strategy \(s_R = (y_R, p_R, q_R)\) and the mutant strategy \(s_M =
-(y_M, p_M, q_M)\). The expected payoffs are give by,
+In the case of low mutation there can be only one type of mutant strategy in the
+population. So in a population of size \(N\), there will be \(k\) mutants and
+\(N - k\) residents, whose strategies we denote respectively as \(s_M =(y_M,
+p_M, q_M)\) and \(s_R = (y_R, p_R, q_R)\).
+The expected payoffs of a resident (\(\pi_R\)) and for a
+mutant (\(\pi_M\)) are give by,
 
@@
 
-As a reminder \(N\) is the population size and \(k\) the number of mutants. The
-number of mutants in the population increases if a resident adopts the strategy
+The number of mutants in the population increases if a resident adopts the strategy
 of a mutant, and decreases if a mutant adopts the strategy of a resident. The
@@
 \begin{align*}
-  \lambda^-_k &\!=\!\rho(\pi_M, \pi_R) \\
-  \lambda^+_k &\!=\!\rho(\pi_R, \pi_M).
+  \lambda^-_k \!=\!\rho(\pi_M, \pi_R) \quad \text{ and } \quad \lambda^+_k \!=\!\rho(\pi_R, \pi_M).
 \end{align*}
 
-\subsection{Invasion analysis of ALLD into Generous Tit For Tat (GTFT)}
+\subsubsection*{Simulation Results based on the expected payoffs}
 
-In the following, we apply the above formalism to calculate how easily a single
-ALLD mutant can invade into a resident population with strategy GTFT. In that
-case, \(s_1 = (1, 1, q)\), \(s_2 = (0, 0, 0)\), and \(k = 1\). When two GTFT
-players interact in the game, their respective probabilities for each of the
-four outcomes in the last round simplify to,
+We can simulate the evolutionary process described by
+Algorithm~\ref{algorithm:pairwise_comparison} when individuals use reactive
+strategies and update their strategies based on their expected payoffs. We
+performed two independent runs of the process where we differed the benefit of
+cooperation \(b\). The results are shown in
+Figure~\ref{fig:expected_payoffs_results}.
 
-\begin{align*}
-    \mathbf{v}(s_1, s_1) = (1, 0, 0, 0).
-\end{align*}
-
-On the other hand, if an ALLD player interacts with a GTFT player, the
-respective probabilities become,
-
-\begin{align*}
-  \mathbf{v}(s_2, s_1) = (0, q, 0, (1 - q)).
-\end{align*}
-
-Using the above we can define the payoffs of a GTFT individual (resident)
-and of the ALLD individual (mutant) follows,
-
-\begin{align*}
-  \displaystyle \pi_R & = \displaystyle \frac{N\!-\!2}{N-1}\cdot (b - c) \cdot -	\displaystyle\frac{1}{N-1}\cdot q \cdot c\text{ and } \displaystyle \pi_M  = \displaystyle b \cdot q.
-\end{align*}
-
-As a consequence, we can calculate the ratio of transition probabilities as,
-
-\begin{align*}
-    \frac{\lambda^{+}}{\lambda^{-}} & = \frac{\rho(\pi_R, \pi_M)}{\rho(\pi_M, \pi_R)} \\[0.5cm]
-    \frac{\lambda^{+}}{\lambda^{-}} & = \frac{e^{-\beta \left(\frac{(N-2) (b-c)}{N-1} - q \cdot (b + \frac{c}{N-1})\right)}+1}
-    {e^{-\beta\left(b \cdot q - \frac{b (N - 2)}{N-1} - \frac{c (N - 2 - q)}{N-1}\right)}+1}
-\end{align*}
-
-In particular, in the limit of strong selection \(\beta \rightarrow \infty\)
-and large populations \(N \rightarrow \infty\), we obtain that the ratio is than
-smaller to 1 if \(q \leq 1 - \frac{c}{b}\). Thus, ALLD is disfavored to invade
-if \(q \leq 1 - \frac{c}{b}\). For \(q = 1 -\frac{c}{b}\) the probability that
-the number of mutants increase by one equals the probability that the mutant
-goes extinct.
-
-\subsection{Simulation results}\label{section:simulation_results_exprected}
-
-Figure~\ref{fig:expected_payoffs_results} shows simulation results of the
-evolutionary process when individuals update their strategies based on their
-expected payoffs. We run two simulations where we differ the benefit of
-cooperation. Though, a higher benefit results in a more cooperative population,
-the population is still rather cooperative even in the case of low benefit.
-
-In the case of low benefit the resident population consists either of defectors
-or conditional cooperators, whereas in the case of high benefit the residents
-typically apply a conditional cooperative strategy. Based on the results on
-invasion we can show that conditional cooperators that the residents adopt are
-of the kind \(p \approx 1\) and \(q \leq 1 - \frac{c}{b}\).
+It is observed that a higher value of benefit results in a more cooperative
+population. For a low benefit the resident population cooperates on average 52\%
+of the time. For a a high benefit the average cooperation increases to 98\%. For
+both a low and a high benefit the resident population consists either of
+defectors or conditional cooperators. A conditional cooperator, otherwise
+known as a generous tit for tat (GTFT) strategy, always cooperates if the
+co-player cooperated ($p\approx1$) and cooperates with a probability \(q\) if
+the co-player defected. The conditional cooperator strategy adopted by the
+population differs between the two simulations. A higher value of benefit
+results to a higher tolerance to defection.
 
@@
     \centering 
-    \includegraphics[width=.70\textwidth]{../static/expected_payoffs_donation_game.pdf}
+    \includegraphics[width=.70\textwidth]{static/expected_payoffs_donation_game.pdf}
     \caption{\textbf{Evolutionary dynamics under expected payoffs with
@@
 
-\subsection{Updating Payoffs based on the one interaction - last round payoffs}
+\subsubsection*{Invasion Analysis based on the expected payoffs}
 
-The first approach we introduced in this work is that of the one  interaction -
-last round updating payoff. In this case an individual updates his strategy
-based on the last payoff they received against the one other member of the
-population. Initially, we define the payoff of a reactive strategy in the last
-round at the stage game. This is given by
-Proposition~\ref{proposition:last_round}.
+We can explicitly calculate the value of \(q\) which evolves. Namely, we will
+calculate how easily a single ALLD mutant can invade into a resident population
+with strategy GTFT. In that case, GTFT\(= (1, 1, q)\), ALLD\(= (0, 0, 0)\), and
+\(k = 1\).
+
+When two GTFT players interact in an infinitely repeated game, their respective
+probabilities for each of the four outcomes in the last round simplify to,
+
+\begin{align*}
+    \mathbf{v}(\text{GTFT}, \text{ALLD}) = (1, 0, 0, 0).
+\end{align*}
+
+On the other hand, if an ALLD player interacts with a GTFT player, the
+respective probabilities become,
+
+\begin{align*}
+  \mathbf{v}(\text{ALLD}, \text{GTFT}) = (0, q, 0, (1 - q)).
+\end{align*}
+
+Using the above we can define the payoffs of a GTFT individual (resident)
+and of the ALLD individual (mutant) follows,
+
+\begin{align*}
+  \displaystyle \pi_{\text{GTFT}} & = \displaystyle \frac{N\!-\!2}{N-1} (b - c)  -	\displaystyle\frac{q c}{N-1} \quad \text{ and } \quad \displaystyle \pi_{\text{ALLD}}  = \displaystyle b q.
+\end{align*}
+
+As a consequence, we can calculate the ratio of transition probabilities as,
+
+\begin{align*}
+    \frac{\lambda^{+}}{\lambda^{-}} & = \frac{\rho(\pi_{\text{GTFT}}, \pi_{\text{ALLD}})}{\rho(\pi_{\text{ALLD}}, \pi_{\text{GTFT}})}  = \frac{e^{-\beta \left(\frac{(N-2) (b-c)}{N-1} - q (b + \frac{c}{N-1})\right)}+1}
+    {e^{-\beta\left(b q - \frac{b (N - 2)}{N-1} - \frac{c (N - 2 - q)}{N-1}\right)}+1}
+\end{align*}
+
+In particular, in the limit of strong selection \(\beta \rightarrow \infty\)
+and large populations \(N \rightarrow \infty\), we obtain that the ratio is than
+smaller to 1 if \(q \leq 1 - \frac{c}{b}\). Thus, ALLD is disfavored to invade
+if \(q \leq 1 - \frac{c}{b}\). For \(q = 1 -\frac{c}{b}\) the probability that
+the number of mutants increase by one equals the probability that the mutant
+goes extinct.
+
+\subsection{Updating Payoffs based on the last last \(m\) round(s) payoffs of \(n\) interaction(s)}
+
+The aim of this work is to explore the effect that the updating payoffs have on
+the cooperation rate of the evolved population. To this end we introduce
+a new method calculating the updating payoffs. Compared to the literature
+where a infinite number of rounds and a well mixed population are assumed,
+we constrain the number of turns and number of turns.
+
+Namely, in the expected payoffs the payoff of a pair depends on the average
+payoff they received over an infinite number of turns. In our approach, the
+payoff of a pair depends on the average payoffs they received in the last \(m\)
+turns. Furthermore, in expected payoffs it is assumed that a player interacts
+with every member of the population, whereas in our approach with consider one
+\(n\) interactions of the player with other members. Here we present results for
+four cases; (\(m=1\) and \(n=1\)), (\(m=2\) and \(n=1\)), (\(m=1\) and \(n=2\)),
+and (\(m=2\) and \(n=2\)).
+
+\subsubsection{Updating Payoffs based on the last round payoff of one interaction (\(m=1\) and \(n=1\))}\label{section:m_one_n_one}
+
+In this case an individual updates her/his strategy based on the last payoff
+they received against one other member of the population. Initially, we define
+the probability that a reactive strategy receives the payoff $u\!\in\!
+\mathcal{U}$ in the very last round of the game against another reactive
+strategy (Proposition~\ref{proposition:last_round}).
 
@@
 
-      v_{R}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
-      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)} \times R,\\[1cm]
+      v_{r}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(q_1+l_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+l_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
+      {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)} \times r,\\[1cm]
 
-      v_{S}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
-      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)} \times S,\\[1cm]
+      v_{s}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(q_1+l_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
+      {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)} \times s,\\[1cm]
 
-      v_{T}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
-      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)} \times T,\\[1cm]
+      v_{t}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+l_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
+      {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)} \times t,\\[1cm]
 
-      v_{P}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
-      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)} \times P.
+      v_{p}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
+      {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)} \times p.
       \end{array}
@@
 
-In these expressions, we have used the notation $r_i:=p_i\!-\!q_i$,
+In these expressions, we have used the notation $l_i:=p_i\!-\!q_i$,
 $\bar{y}_i\!=\!1\!-\!y_i$, $\bar{q}_i:=1\!-\!q_i$, and
-$\bar{r}_i:=\bar{p}_i\!-\!\bar{q}_i=-r_i$ for $i\!\in\!\{1,2\}$.
+$\bar{l}_i:=\bar{p}_i\!-\!\bar{q}_i=-l_i$ for $i\!\in\!\{1,2\}$.
 \end{Prop}
 
-Note that in the proposition we use the general notation of the prisoner's
-dilemma, to denote that the results apply to \(2 \times 2\) any symmetric game.
+Note that in the proposition we here we focus on the case of the donation game/prisoner's
+dilemma but the result applies to any \(2 \times 2\) symmetric game.
 
@@
 
-In the updating stage we select a mutant and resident to be either the role
-model or the learner. Given that they can interact with only one other member of
-the population, they can interact either with each other or either can interact
-with another resident or with another mutant. Thus, in each updating stage there
-are five possible combinations of pairs (Figure~\ref{fig:single_pairs}).
-
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[width=.65\textwidth]{../static/last_rounds_matches.tex}
-  \caption{\textbf{Possible pairings combination in the updating stage, given
-  that individuals interact with only one other member in the population.} At
-  each step of the evolutionary process we choose a role model and a learner to
+At each step of the evolutionary process we choose a role model and a learner to
 update the population. We consider the case where both the role model and the
@@
 population. There are five possible pairings at each step. They interact with
-  other with a probability \(\frac{1}{N - 1}\), and thus they do not interact
-  with other with a probability \(1 - \frac{1}{N - 1}\). In the latter case,
-  each of them can interact with either a mutant or a resident. Both of them
-  interact with a mutant with a probability $\frac{(k-1)(k-2)}{(N-2)(N-3)}$ and
-  both interact with a resident with a probability
-  $\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$. The last two possible pairings are that
-  either of them interacts with a resident whilst the other interacts with a
-  mutant, and this happens with a probability
-  $\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$.}
-  \label{fig:single_pairs}
-\end{figure}
+other with a probability \(\frac{1}{N - 1}\), and thus they do not interact with
+other with a probability \(1 - \frac{1}{N - 1}\). In the latter case, each of
+them can interact with either a mutant or a resident. Both of them interact with
+a mutant with a probability $\frac{(k-1)(k-2)}{(N-2)(N-3)}$ and both interact
+with a resident with a probability $\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$. The last
+two possible pairings are that either of them interacts with a resident whilst
+the other interacts with a mutant, and this happens with a probability
+$\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$.
 
-Given the last round payoff and possible pair combinations for a single
-interaction, we define the probability that the respective last round payoffs of
-two players \(s_1, s_2\) are given by $u_1$ and $u_2$ as,
+Given the possible pairings and Proposition~\ref{proposition:last_round}, we
+define the probability that the respective last round payoffs of two players
+\(s_1, s_2\) are given by $u_1$ and $u_2$ as,
 
@@
 pairs can occur that are feasible in a direct interaction,
-$\splitatcommas{(u_1,u_2)\in \mathcal{U}^2_F:=\big\{ (b-c, b-c), (-c, b), (b, -c), (0, 0)
+$\splitatcommas{(u_1,u_2)\in \mathcal{U}^2_F:=\big\{ (r, r), (s, t), (t, s), (p, p)
 \big\}}$, as represented by the respective indicator function. Otherwise, if the
@@
 
-Given that $N\!-\!k$ players use the resident strategy
-$s_{R}\!=\!(y_{R},p_{R},q_{R})$ and that the remaining $k$ players use the
-mutant strategy $s_{M}\!=\!(y_{M},p_{M},q_{M})$, the probability that the number
-of mutants increases by one in one step of the evolutionary process can be
-written as
+The probability that the number of mutants increases, and decreases respectively,
+by one is now given by,
 
@@
 
-\subsubsection{Invasion analysis of ALLD into (GTFT)}
+\subsubsection*{Simulation Results based on the last round payoff of one interaction}
 
-Similarly to the previous section, we calculate how easily a single ALLD mutant
-can invade into a resident population with strategy GTFT.
+We simulate the evolutionary process given that individuals now use the the last
+round payoff of one interaction to update their strategies. The results are
+shown in Figure~\ref{fig:one_interaction_last_round_payoffs_results}. Similarly
+to the results of the expected payoffs, a higher benefit results in a more
+cooperative population. However we note that the cooperation rate of the evolved
+populations are lower compared to the case of expected payoffs, and that the
+cooperation rate increases less between the two simulations. The evolving both
+consist of defectors and conditional cooperators. The evolved \(q\)s are smaller
+which also results to a less cooperative population. In between the two
+simulations \(q \approx \frac{1}{2}\). In the invasion analysis we will
+demonstrate that in the case of the the last round payoff of one interaction, a
+conditional player needs to have \(q \leq \frac{1}{2}\) to avoid being invaded
+by a defector. In both simulations \(q \leq \frac{1}{2}\), the higher benefit
+pushes the evolved population closer to the bounty.
 
-When two GTFT players interact in the game, their respective probabilities for
-each of the four outcomes in the last round simplify to,
+\begin{figure}[!htbp]
+  \centering 
+  \includegraphics[width=.70\textwidth]{static/one_interaction_last_round_donation_game.pdf}
+  \caption{\textbf{Evolutionary dynamics under one interaction last round
+  payoffs with low (left) and high (right) benefit.} The cooperation rates
+  in both cases are less compared to the expected payoffs. Moreover, increase is
+  less when the benefit is increased to ten. In the cases of low and
+  high benefit case the resident population either consists of defectors (with
+  \(p \approx q \approx 0\)) or of conditional cooperators for which \(p \approx
+  1\) and \(q \leq \frac{1}{2}\). Parameters: \(N =100, c=1, \beta=1\).
+  }\label{fig:one_interaction_last_round_payoffs_results}
+\end{figure}
+
+\subsubsection*{Invasion Analysis based on the last round payoff of one interaction}
+
+We once again calculate how easily a single ALLD mutant can invade into a
+resident population with strategy GTFT. When two GTFT players interact in the
+game, their respective probabilities for each of the four outcomes in the last
+round simplify to,
 
 \begin{align*}
-    v_R(GTFT,GTFT) = 1, & \quad v_T(GTFT,GTFT) = 0, \\
-    v_S(GTFT,GTFT) = 0, & \quad v_P(GTFT,GTFT) = 0.
+    v_{r}(GTFT,GTFT) = 1, & \quad v_{t}(GTFT,GTFT) = 0, \\
+    v_{s}(GTFT,GTFT) = 0, & \quad v_{p}(GTFT,GTFT) = 0.
 \end{align*}
@@
 \begin{align*}
-    v_R(ALLD,GTFT) & = 0, &  v_S(ALLD,GTFT) & = 0, \\
-    v_T (ALLD, GT F T ) & = 1 - \delta + \delta q, &  v_P (ALLD, GTFT) & = \delta(1 - q).
+    v_{r}(ALLD,GTFT) & = 0, &  v_{s}(ALLD,GTFT) & = 0, \\
+    v_{t} (ALLD, GTFT ) & = 1 - \delta + \delta q, &  v_{p} (ALLD, GTFT) & = \delta(1 - q).
 \end{align*}
@@
 \begin{align*}
-  x(R, T) = & \frac{N - 2}{N - 1} \cdot (1 - \delta + \delta q)\\
-  x(R, R) = & \frac{N - 2}{N - 1} \cdot \delta (1 - q) \\
-  x(S, T) = & \frac{1}{N - 1} \cdot (1 - \delta + \delta q) \\
-  x(P, P) = & \frac{1}{N - 1} \cdot \delta (1 - q)
+  x(r, t) = & \frac{N - 2}{N - 1} \cdot (1 - \delta + \delta q)\\
+  x(r, r) = & \frac{N - 2}{N - 1} \cdot \delta (1 - q) \\
+  x(s, r) = & \frac{1}{N - 1} \cdot (1 - \delta + \delta q) \\
+  x(p, p) = & \frac{1}{N - 1} \cdot \delta (1 - q)
 \end{align*}
 
-As a consequence, we can calculate the ratio of transition probabilities as
+We now calculate the ratio of transition probabilities as
 
 \begin{equation*}
-  \resizebox{.9\textwidth}{!} 
-  {$
-    \frac{\lambda^{+}}{\lambda^{-}} = \frac{\frac{N - 2}{N - 1} \cdot \left(\frac{\delta(1-q)}{e^{-\beta(c - b)} + 1} + \frac{-\delta +\delta  q + 1}{e^{-\beta c} + 1}\right) + \frac{1}{N - 1} \cdot \left(\frac{-\delta +\delta  q+1}{e^{\beta  (-(b+c))}+1}\right)}
-    {\frac{N - 2}{N - 1} \cdot \left(\frac{\delta (1-q)}{e^{-\beta (b-c)}+1} + \frac{-\delta +\delta  q + 1}{e^{\beta  c}+1}\right)+ \frac{1}{N-1} \cdot \left(\frac{-\delta +\delta  q+1}{e^{-\beta  ((-b-c))}+1}\right)}
-  $}
+\frac{\lambda^{+}}{\lambda^{-}} = \frac{ \frac{N - 2}{N - 1}  \left( \frac{ \delta  \left(1 - q \right)}{1
++ e^{-  \beta  \left(b - c \right)}} +  \frac{ \delta q -  \delta + 1}{e^{ \beta
+c} + 1} \right)  +  \frac{1}{N-1}  \left(\frac{ \delta  \left(1 - q \right)}{2} +
+ \frac{ \delta q -  \delta + 1}{1 + e^{-  \beta  \left(- b - c \right)}}\right)}
+ { \frac{N - 2}{N - 1}  \left( \frac{ \delta  \left(1 - q \right)}{1 +
+e^{-  \beta  \left(- b + c \right)}} +  \frac{ \delta q -  \delta + 1}{1 + e^{-
+ \beta c}} \right) +  \frac{1}{N -1} \left(\frac{ \delta  \left(1 - q \right)}{2} +
+ \frac{ \delta q -  \delta + 1}{1 + e^{-  \beta  \left(b + c \right)}}\right)}
 \end{equation*}
@@
 condition becomes \(q < 1/2\) (for \(q = 1/2\), the payoff of the ALLD player is
-\(T > R\) for half of the time, and it is \(P < R\) for the other half. The
+\(r > r\) for half of the time, and it is \(p < r\) for the other half. The
 probability that the number of mutants increase by one equals the probability
@@
 
-\subsubsection{Simulation results}
+\subsubsection{Updating Payoffs based on the last round payoff of two interactions (\(m=1\) and \(n=2\))}\label{section:m_one_n_two}
 
-In a similar fashion to Section~\ref{section:simulation_results_exprected}, we
-simulate the evolutionary process given that individuals now use the last
-interaction-last round payoffs to update their strategies.
-Figure~\ref{fig:one_interaction_last_round_payoffs_results} shows simulation
-results. Though, a higher benefit results in a more cooperative population, in
-the case of these updating payoff, only to a slightly more cooperative. As shown
-in Figure~\ref{fig:one_interaction_last_round_payoffs_results} the evolving
-populations are more similar, both consisting of defectors and conditional
-cooperators. Compared to the expected payoffs, the conditional cooperators are
-also less cooperative \(p \approx 1\) and \(q \leq \frac{1}{2}\). The increased
-benefit pushes the population closer to the boundary of \(q \leq \frac{1}{2}\)
-but that is the maximum.
+Here, in the updating phase we choose one role model and one learner, however
+now, we consider the case where both the role model and the learner estimate
+their updating payoff after interacting with two members of the population.
+There are stages in the pairing process and there are twenty four possible
+pairings.
+
+In the first stage the role model and the learner are matched together with a
+probability $\frac{1}{N-1}$ or not with a probability $(1 - \frac{1}{N-1})$. If
+they were matched then in the second stage of pairings there are only four
+possible outcomes; both of them interact with a mutant with a probability
+$\frac{(k-1)(k-2)}{(N-2)(N-3)}$ and both interact with a resident with a
+probability $\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$. The last two possible pairings
+are that either of them interacts with a resident whilst the other interacts
+with a mutant, and this happens with a probability
+$\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$.
+
+In the later case, where they were not matched in the first stage, there are
+four possible outcomes; both of them interact with a mutant and both interact
+with a resident, either of them interacts with a resident whilst the other
+interacts with a mutant. For each of the above pairings of the first stage there
+are five possible; they interact with each other, both of them interact with a
+mutant and both interact with a resident, either of them interacts with a
+resident whilst the other interacts with a mutant.
+
+The new possible pairings change how we define the probability that the
+respective last round payoffs of two players \(s_1, s_2\) are given by \(u_1\)
+and \(u_2\). The probability denoted as \(\bar{x}(u_1,u_2)\) is given by,
+
+\begin{equation}
+  \resizebox{.99\textwidth}{!}
+  {%
+$
+  \setlength{\arraycolsep}{1pt}
+  \begin{array}{llrl}
+  x(u_1,u_2)	& = \displaystyle \frac{1}{N\!-\!1}\cdot v_{u_1}(s_1,s_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F} \cdot A + \left(1\!-\!\frac{1}{N\!-\!1}\right) \left[ \right. \\[0.5cm] 
+  & \left. v_{u_1}(s_1,s_2) _{u_2}(s_2,s_2)  \frac{\left(k - 2\right) \left(k - 1\right)}{(N\!-\!2)(N\!-\!3)}
+  \left(\frac{1}{N\!-\!2}\cdot v_{u_1}(s_1,s_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F} + (1 - \frac{1}{N - 2}) [B_1 + B_2 + B_3 + B_4]\right) + \right. \\[0.5cm] 
+  & \left.
+v_{u_1}(s_1,s_2) _{u_2}(s_2,s_1)  \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N\!-\!2)(N\!-\!3)} \left(\frac{1}{N\!-\!2}\cdot v_{u_1}(s_1,s_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F} + (1 - \frac{1}{N - 2}) [C_1 + C_2 + C_3 + C_4]\right) + \right.
+\\[0.5cm]
+  & \left.
+v_{u_1}(s_1,s_1) _{u_2}(s_2,s_2)  \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N\!-\!2)(N\!-\!3)} \left(\frac{1}{N\!-\!2}\cdot v_{u_1}(s_1,s_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F} + (1 - \frac{1}{N - 2}) [D_1 + D_2 + D_3 + D_4]\right) + \right.
+\\[0.5cm]
+  & \left.
+v_{u_1}(s_1,s_1) _{u_2}(s_2,s_1)  \frac{\left(N - k - 2\right) \left(N - k - 1\right)}{(N\!-\!2)(N\!-\!3)} \left(\frac{1}{N\!-\!2}\cdot v_{u_1}(s_1,s_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F} + (1 - \frac{1}{N - 2}) [E_1 + E_2 + E_3 + E_4]\right) \right]
+  \end{array} $}
+\end{equation}
+
+\begin{equation}
+  \resizebox{.9\textwidth}{!}
+  {%
+$
+  \setlength{\arraycolsep}{1pt}
+  \begin{array}{llrl}
+  A & = \left(1\!-\!\frac{1}{N\!-\!1}\right)  \left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_2) + 
+   \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_1) + \right. \\[0.5cm]
+   & \left. \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_2) + 
+   \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_1)\right] \\[0.5cm] 
+   B_1 & = \frac{\left(k - 3\right) \left(k - 2\right)}{(N - 3) (N - 4)} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_2) \quad
+   B_2 = \frac{\left(k - 2\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_2) \\[0.5cm] 
+   B_3 & = \frac{\left(k - 2\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_1) \quad
+   B_4 = \frac{\left(N - k - 2\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_1) \\[0.5cm]
+
+   C_1 & = \frac{\left(k - 3\right) \left(k - 1\right)}{(N - 3) (N - 4)} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_2) \quad
+   C_2 = \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_2) \\[0.5cm] 
+   C_3 & = \frac{\left(k - 2\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_1) \quad
+   C_4 = \frac{\left(N - k - 2\right)^{2}}{(N - 3) (N - 4)} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_1) \\[0.5cm]
+
+   D_1 & = \frac{\left(k - 2\right)^{2}}{(N - 3) (N - 4)} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_2) \quad
+   D_2 = \frac{\left(k - 2\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_2) \\[0.5cm] 
+   D_3 & = \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_1) \quad
+   D_4 = \frac{\left(N - k - 3\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_1) \\[0.5cm]
+
+   E_1 & = \frac{\left(k - 2\right) \left(k - 1\right)} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_2) \quad
+   E_2 = \frac{\left(k - 1\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_2) \\[0.5cm] 
+   E_3 & = \frac{\left(k - 1\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_1) \quad
+   E_4 = \frac{\left(N - k - 3\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_1) \\[0.5cm] 
+  \end{array}$}
+\end{equation}
+
+The first term on the right side corresponds to the case that the learner and
+the role model happened to be matched during the first stage of pairing,
+followed by them being paired with another member of the population on the
+second stage. The second terms corresponds to the case that the learner and the
+role model interact with a mutant with a probability ($\frac{\left(k - 2\right)
+\left(k - 1\right)}{(N\!-\!2)(N\!-\!3)}$). In the seconds stage, they can either
+interact with each other $\frac{1}{N - 2}$ or not $(1 - \frac{1}{N - 2})$. If
+they do not interact with each other, then each of the following can happen:
+both of them interact with a mutant with a probability
+$\frac{(k-3)(k-2)}{(N-4)(N-3)}$ and both interact with a resident with a
+probability $\frac{(N-k-1)(N-k-2)}{(N-3)(N-4)}$. The last two possible pairings
+are that either of them interacts with a resident whilst the other interacts
+with a mutant, and this happens with a probability
+$\frac{(N-k-2)(k-1)}{(N-4)(N-3)}$. The rest of the cases follow the same
+pattern.
+
+The probability that the number of mutants increases, and decreases respectively,
+by one is now given by,
+
+\begin{align}
+\lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R},u_{M}\in\mathcal{U}} \bar{x}(u_{R},u_{M})\cdot \rho(u_{R},u_{M}), \\
+\lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R},u_{M}\in\mathcal{U}} \bar{x}(u_{R},u_{M})\cdot \rho(u_{M},u_{R}).
+\end{align}
+
+\subsubsection*{Simulation Results based on the last round payoff of two interactions}
+
+We simulate the evolutionary process when individuals update their strategies
+based on the last round payoffs from two interactions with other members of the
+population, Figure~\ref{fig:one_interaction_last_round_payoffs_results}. For
+both a low and a high benefit value the average cooperation rate has increased
+compared to the the last round against a single other member of the population.
+In the case of low benefit the evolved population's cooperation rate is almost
+the same as in the case of expected payoffs. However, the cooperation rate does
+not increase as much as in for the expected payoffs for \(b=10\).
 
@@
   \centering 
-  \includegraphics[width=.70\textwidth]{../static/one_interaction_last_round_payoffs_results}
-  \caption{\textbf{Evolutionary dynamics under one interaction last round
+  \includegraphics[width=.70\textwidth]{static/two_interactions_donation_game.pdf}
+  \caption{\textbf{Evolutionary dynamics under two interactions last round
+  payoffs with low (left) and high (right) benefit.} In the cases of low and
+  high benefit case the resident population either consists of defectors (with
+  \(p \approx q \approx 0\)) or of conditional cooperators for which \(p \approx
+  1\) and \(q \leq \frac{\sqrt{2}}{2}\). Parameters: \(N =100, c=1, \beta=1\).
+  }\label{fig:two_interactions_results}
+\end{figure}
+
+\subsubsection*{Invasion Analysis based on the last round payoff of two interactions}
+
+In a similar fashion we can calculate the threshold of \(q\) for which a GTFT
+player can not be invaded by an ALLD mutant. The ratio of transition probabilities
+is given by,
+
+\begin{equation}
+  \resizebox{.9\textwidth}{!}
+  {%
+$\frac{\lambda^{+}}{\lambda^{-}} =
+\frac{\frac{N - 3}{N-1} \left(\frac{\delta^{2} \left(1 - q\right)^{2}}{1 + e^{- \beta \left(- b + c\right)}} + \frac{2 \delta \left(1 - q\right) \left(\delta q - \delta + 1\right)}{1 + e^{- \beta \left(- \frac{b}{2} + c\right)}} + \frac{\left(\delta q - \delta + 1\right)^{2}}{1 + e^{- \beta c}}\right)
++ \frac{2}{N-1} \left(\frac{\delta^{2} \left(1 - q\right)^{2}}{1 + e^{- \beta \left(- \frac{b}{2} + \frac{c}{2}\right)}} + \frac{\delta \left(1 - q\right) \left(\delta q - \delta + 1\right)}{1 + e^{- \frac{\beta c}{2}}} + \frac{\delta \left(1 - q\right) \left(\delta q - \delta + 1\right)}{1 + e^{- \beta c}} + \frac{\left(\delta q - \delta + 1\right)^{2}}{1 + e^{- \beta \left(\frac{b}{2} + c\right)}}\right)}
+{\frac{N - 3}{N-1} \left(\frac{\delta^{2} \left(q - 1\right)^{2}}{1 + e^{- \beta \left(b - c\right)}} + \frac{2 \delta \left(1 - q\right) \left(\delta q - \delta + 1\right)}{1 + e^{- \beta \left(\frac{b}{2} - c\right)}} + \frac{\left(\delta q - \delta + 1\right)^{2}}{e^{\beta c} + 1}\right) + \frac{2}{N-1} \left(\frac{\delta^{2} \left(q - 1\right)^{2}}{1 + e^{- \beta \left(\frac{b}{2} - \frac{c}{2}\right)}} + \frac{\delta \left(1 - q\right) \left(\delta q - \delta + 1\right)}{e^{\beta c} + 1} + \frac{\delta \left(1 - q\right) \left(\delta q - \delta + 1\right)}{e^{\frac{\beta c}{2}} + 1} + \frac{\left(\delta q - \delta + 1\right)^{2}}{1 + e^{- \beta \left(- \frac{b}{2} - c\right)}}\right)}
+$}
+\end{equation}
+
+In the limit of strong selection \(\beta \rightarrow \infty\) and large
+populations \(N \rightarrow \infty \), we obtain
+
+\begin{equation}
+\frac{\lambda^{+}}{\lambda^{-}} = 
+\begin{cases}
+  \frac{\left(\delta q - \delta + 1\right)^{2}}{\delta^{2} \left(1 - q\right)^{2} + 2 \delta \left(1 - q\right) \left(\delta q - \delta + 1\right)}  & \frac{b}{2} > c \\
+  \frac{- \delta q + \delta - 1}{\delta \left(q - 1\right)}  & \frac{b}{2} = c \\
+  - \frac{\left(\delta q - \delta - 1\right) \left(\delta q - \delta + 1\right)}{\delta^{2} \left(q - 1\right)^{2}} & \frac{b}{2} < c
+\end{cases}
+\end{equation}
+
+Note that in this case the relationship of the payoffs, cost and benefit, have
+an effect of the \(q\). Namely if \(\frac{b}{2}\) is higher than the cost \(c\).
+In the case where \(\frac{b}{2}=c\) the result is the same as in the case
+of \(m=n=1\). For \(\frac{\lambda^{+}}{\lambda^{-}} < 1\):
+
+\begin{equation}
+\begin{cases}
+  q \in \{\frac{\delta - 1 + \frac{\sqrt{2}}{2}}{\delta}\}  & \frac{b}{2} > c \\
+  q = \frac{\delta - \frac{1}{2}}{\delta}  & \frac{b}{2} = c \\
+  q \in \{\frac{\delta - \frac{\sqrt{2}}{2}}{\delta}, \  \frac{\delta + \frac{\sqrt{2}}{2}}{\delta}\} & \frac{b}{2} < c
+\end{cases}
+\end{equation}
+
+There are two distinguished cases. For \(\frac{b}{2}>c\) the ration is smaller
+than if \(q < \frac{\delta - 1 + \frac{\sqrt{2}}{2}}{\delta}\). For infinitely
+repeated games, \(\delta \rightarrow 1\), this condition becomes \(q <
+\frac{\sqrt{2}}{2}\). In the case of \(\frac{b}{2}<c\) there are two possible
+roots, however, we assume repeated games that are repeated for a large number
+of turn such as \(\delta \rightarrow 1\). The condition then becomes
+\(q < 1 - \frac{\sqrt{2}}{2}\).
+
+\subsection{Updating Payoffs based on the last two rounds payoff of one interactions (\(m=2\) and \(n=1\))}\label{section:m_two_n_one}
+
+We extend Proposition~\ref{proposition:last_round} to \(n\) rounds. 
+The result is given by Proposition~\ref{proposition:last_two_rounds}.
+
+\begin{Prop}\label{proposition:last_two_rounds} Consider a repeated game, with
+  continuation probability $\delta$, between players with reactive strategies
+  $s_1\!=\!(y_1, p_1, q_1$)  and $s_2\!=\!(y_2,p_2,q_2)$ respectively. Let
+  $\mathcal{\tilde{U}}$ denote the set of feasible payoffs in the last
+  \(n\) rounds, and let \(\tilde{\mathbf{u}}\) be the corresponding payoff vector.
+  Then the probability that the $s_1$ player receives the payoff $u\!\in\!
+  \mathcal{\tilde{U}}$ in the very last two rounds of the game is given by,
+
+  \begin{equation}
+  \langle\mathbf{\tilde{v}}(s_1,s_2),\mathbf{\tilde{u}}\rangle, \text{ where } \mathbf{\tilde{v}} \in R^{4^{n}} \text{ is given by },
+  \end{equation}
+  \begin{equation}
+    \mathbf{\tilde{v}}(s_1,s_2) = (1 - \delta) w_{a_1, a_2} \delta^2 \left[\mathbf{v_0}(I_4 - \delta M)^{-1}\right]_{a_1, a_2}, \quad  w_{a_1, a_2} \in M \ \forall \ a_1, a_2 \in \{1, 2, 3, 4\}.
+  \end{equation}
+\end{Prop}
+
+Here consider the case of \(n=2\). The framework remains the same
+as in section~\ref{section:m_one_n_one}, the only difference is the game stage
+payoffs.
+
+\subsubsection*{Simulation Results based on the last two rounds payoff of one interaction}
+
+\begin{figure}[!htbp]
+  \centering 
+  \includegraphics[width=.70\textwidth]{static/two_rounds_donation_game.pdf}
+  \caption{\textbf{Evolutionary dynamics under two interactions last round
   payoffs with low (left) and high (right) benefit.} In the cases of low and
@@
   1\) and \(q \leq \frac{1}{2}\). Parameters: \(N =100, c=1, \beta=1\).
-  }\label{fig:one_interaction_last_round_payoffs_results}
+  }\label{fig:two_rounds_results}
 \end{figure}
 
-\subsection{Updating Payoffs based on \(n\) interactions - last \(m\) round payoffs}
+\subsubsection*{Invasion Analysis based on the last two rounds payoff of one interaction}
 
-So far we have presented the two extreme cases; the case where an individual
-updates based one interaction and one round, and the case  where an individual
-updates based on all interactions and all rounds. In this work we also
-explore the intermediate cases (Figure~\ref{fig:cases_updating_payoffs}). Namely the cases:
+\begin{equation}
+  \resizebox{.9\textwidth}{!}
+  {%
+  $\frac{\lambda^{+}}{\lambda^{-}} =
+  \frac{\frac{\delta^{2} \left(N - 2\right)}{N-1} \left(\frac{\delta \left(1 - q\right)^{2}}{1 + e^{- \beta \left(- b + c\right)}} + \frac{q \left(\delta q - \delta + 1\right)}{1 + e^{- \beta c}} + \frac{\left(1 - q\right) \left(2 \delta q - \delta + 1\right)}{1 + e^{- \beta \left(- \frac{b}{2} + c\right)}}\right) + \frac{1}{N-1}\left(\frac{\delta \left(1 - q\right)^{2}}{2} + \frac{q \left(\delta q - \delta + 1\right)}{1 + e^{- \beta \left(b + c\right)}} + \frac{\left(1 - q\right) \left(2 \delta q - \delta + 1\right)}{1 + e^{- \beta \left(\frac{b}{2} + \frac{c}{2}\right)}}\right)}
+  {\frac{\delta^{2} \left(N - 2\right)}{N-1} \left(\frac{\delta \left(1 - q\right)^{2}}{1 + e^{- \beta \left(b - c\right)}} + \frac{q \left(\delta q - \delta + 1\right)}{e^{\beta c} + 1} + \frac{\left(1 - q\right) \left(2 \delta q - \delta + 1\right)}{1 + e^{- \beta \left(\frac{b}{2} - c\right)}}\right) + \frac{1}{N-1} \left(\frac{\delta \left(1 - q\right)^{2}}{2} + \frac{q \left(\delta q - \delta + 1\right)}{1 + e^{- \beta \left(- b - c\right)}} + \frac{\left(1 - q\right) \left(2 \delta q - \delta + 1\right)}{1 + e^{- \beta \left(- \frac{b}{2} - \frac{c}{2}\right)}}\right)}
+  $}
+\end{equation}
 
-\begin{itemize}
-  \item last round \(n=1\) with two other players \(m=2\)
-  \item last two rounds \(n=2\) with one other player  \(m=1\)
-  \item last two rounds \(n=2\) with two other players \(m=2\)
-\end{itemize}
+In the limit of strong selection \(\beta \rightarrow \infty\) and
+large populations \(N \rightarrow \infty \), we obtain
+
+\begin{equation}
+\frac{\lambda^{+}}{\lambda^{-}} = 
+\begin{cases}
+  \frac{q \left(\delta q - \delta + 1\right)}{\delta \left(1 - q\right)^{2} + \left(1 - q\right) \left(2 \delta q - \delta + 1\right)}  & \frac{b}{2} > c \\
+  - \frac{\delta q - \delta + q + 1}{\left(\delta + 1\right) \left(q - 1\right)}  & \frac{b}{2} = c \\
+  \frac{q \left(\delta q - \delta + 1\right) + \left(1 - q\right) \left(2 \delta q - \delta + 1\right)}{\delta \left(1 - q\right)^{2}} & \frac{b}{2} < c
+\end{cases}
+\end{equation}
+
+For \(\frac{\lambda^{+}}{\lambda^{-}} < 1\):
+
+\begin{equation}
+\begin{cases}
+  q \in \{\frac{\delta - \sqrt{\delta^{2} + 1} - 1}{2 \delta}, \  \frac{\delta + \sqrt{\delta^{2} + 1} - 1}{2 \delta}\}  & \frac{b}{2} > c \\
+  q = \frac{\delta}{\delta + 1}  & \frac{b}{2} = c \\
+  q \in \{1 - \frac{\sqrt{2}}{2 \sqrt{\delta}}, \  1 + \frac{\sqrt{2}}{2 \sqrt{\delta}}\} & \frac{b}{2} < c
+\end{cases}
+\end{equation}
+
+\subsection{Updating Payoffs based on the last two rounds payoff of two interactions (\(m=2\) and \(n=2\))}
+
+The last case for which we present results in this work is the case of \(m=2\)
+and \(n=2\), that is, an individual considers the last two rounds payoffs he/she
+received against two different members of the population. The probability that
+the number of mutants increases, and decreases respectively, by one is now given
+by, are based on the approached on section~\ref{section:m_one_n_two}
+and~\ref{section:m_two_n_one}. In this case we only present only the results on
+the numerical simulations which are given in
+Figure~\ref{fig:two_rounds_opponents_results}.
+
+The evolved cooperation rates are between the other cases and the expected
+payoffs. For a low benefit we get results similar to the case where one the last
+round was considered, however, in the high benefit case the evolved cooperation
+rate increases to 91\% which is higher then any of the newly introduced
+approaches. Though we do no demonstrate or method of more cases, we assume, that
+as \(n\), and \(m\) increase the results tend to the expected payoffs.
 
@@
   \centering 
-  \includestandalone[width=.5\textwidth]{cases}
-  \caption{\textbf{Updating payoffs used in this work.} \(n\) is the
-  number of turns and \(m\) th number of individuals. The expected payoffs
-  are the one commonly used in the literature. We have introduced a method
-  for the rest.}\label{fig:cases_updating_payoffs}
+  \includegraphics[width=.70\textwidth]{static/two_rounds_opponents_donation_game.pdf}
+  \caption{\textbf{Evolutionary dynamics under two interactions last round
+  payoffs with low (left) and high (right) benefit.} In the cases of low and
+  high benefit case the resident population either consists of defectors (with
+  \(p \approx q \approx 0\)) or of conditional cooperators for which \(p \approx
+  1\) and \(q \leq \frac{1}{2}\). Parameters: \(N =100, c=1, \beta=1\).
+  }\label{fig:two_rounds_opponents_results}
 \end{figure}
 
-We define the payoffs on the two last round, and in the updating phase the
-probability that.
+\section{Expected and Last Round Updating Payoffs for Memory One Strategies}
 
-We extend our framework to consider the case where players update their
-strategies based on the outcome of \textbf{the last two turns and based on their
-interaction with two other members of the population}. At the stage game we
-define the payoff of a reactive strategy in the last two rounds,
-Proposition~\ref{proposition:last_two_rounds}.
+So far we have only considered the case where members can adopt reactive
+strategies during the game phase. In order to demonstrate that our results hold
+for higher memory strategies here we present numerical results in the case of
+memory-one strategies.
 
-\begin{Prop}\label{proposition:last_two_rounds} Consider a repeated game, with
-  continuation probability $\delta$, between players with reactive strategies
-  $s_1\!=\!(y_1, p_1, q_1$)  and $s_2\!=\!(y_2,p_2,q_2)$ respectively. Let
-  $\mathcal{\tilde{U}} = \splitatcommas{\{RR, RS, RT, RP, SR, SS, ST, SP, TR,
-  TS, TT, TP, PR, PS, PT, PP\}}$ denote the set of feasible payoffs in the last
-  two rounds, and let \(\tilde{\mathbf{u}}\) be the corresponding payoff vector.
-  Then the probability that the $s_1$ player receives the payoff $u\!\in\!
-  \mathcal{\tilde{U}}$ in the very last two rounds of the game is given by,
+Memory-one strategies consider the outcome of the previous round to decide on an
+action. There are four possible outcome in each round; \((C, C), (C, D), (D, C),
+(D, D)\). A memory-one strategy \(s\) can be written as a five-dimensional
+vector \(s=(y, p_1, p_2, p_3, p_4)\). The parameter \(y\) is the probability
+that the strategy opens with a cooperation and \(p_1\), \(p_2\), \(p_3\),
+\(p_4\) are the probabilities that the strategy cooperates for each of the
+possible outcomes of the last round.
 
-  \begin{equation}
-  \langle\mathbf{\tilde{v}}(s_1,s_2),\mathbf{\tilde{u}}\rangle, \text{ where } \mathbf{\tilde{v}} \in R^{16} \text{ is given by },
-  \end{equation}
-  \begin{equation}
-    \mathbf{\tilde{v}}(s_1,s_2) = (1 - \delta) m_{a_1, a_2} \delta^2 \left[\mathbf{v_0}(I_4 - \delta M)^{-1}\right]_{a_1, a_2}, \quad  m_{a_1, a_2} \in M \ \forall \ a_1, a_2 \in \{1, 2, 3, 4\}.
-  \end{equation}
-\end{Prop}
-
-In the updating stage we select a mutant and resident to be either the role
-model or the learner. Given that they can interact with two other members of the
-population there are a total of twenty four possible combinations of pairs
-(Figure~\ref{fig:pissible_two_pairs}).
+We perform four separate simulations where we differ the updating payoff and the
+benefit of cooperation \(b\). The results for a low value of benefit are given
+in Figure~\ref{fig:memory_one_low_benefit}, and for a high benefit in
+Figure~\ref{fig:memory_one_high_benefit}.
 
@@
   \centering 
-  \includegraphics[width=.65\textwidth]{../static/tree.tex}
-  \caption{\textbf{Possible pairings combination in the updating stage, given
-  that individuals interact with two other members in the population.}}
-  \label{fig:pissible_two_pairs}
+  \includegraphics[width=\textwidth]{static/memory_one_results_low_benefit.pdf}
+  \caption{\textbf{Evolutionary dynamics results for memory-one strategies for low benefit.}
+  We perform two independent simulations. In one simulation individuals use
+  expected payoffs and in the other the last round one interacts when they
+  update their strategies. We run each simulation for \(T = 10^8\) time steps.
+  For each time step, we have recorded the current resident population, who is
+  now of the form  \((y, p_1, p_2, p_3, p_4)\) In the left panel we report the
+  cooperation rates for each simulation. It can be shown than even for
+  memory-one strategies expected payoffs result in a more cooperative
+  population. The right panel reports the most abundant strategy of each
+  simulation. Abundance is the number of mutants a strategy can repel before
+  being invaded. The most abundant strategies have some similarities,
+  namely, \(p_1 \approx 1\), \(p_3 \approx 0\) and \(p_4 > \frac{1}{2}\). There are also differences, in the latter
+  case a strategy is more likely to open with a cooperation and their tolerance to
+  a \((C, D)\) outcome is almost zero. A difference between the strategies is their
+  abundance. In the expected payoffs case a strategy can repel a way greater
+  number of mutants. In the case of last round payoffs strategies become less
+  robust. Parameters: \(N =100, c=1, b=3, \beta=1\).}\label{fig:memory_one_low_benefit}
 \end{figure}
 
-For the above cases we use combinations of Eq. 1 2 and 3.
+\begin{figure}[!htbp]
+  \centering 
+  \includegraphics[width=\textwidth]{static/memory_one_results_high_benefit.pdf}
+  \caption{\textbf{Evolutionary dynamics results for memory-one strategies for high benefit.}
+  We perform two independent simulations. In the case of high benefit expected
+  payoffs again result in a more cooperative population. The right panel reports
+  the most abundant strategy of each simulation. Abundance is the number of
+  mutants a strategy can repel before being invaded. For the expected payoffs he
+  most abundant is that of win-stay lose-shift. However in the later case the
+  most abundant strategy is a strategy with no tolerance to one defection, and
+  it cooperates with a probability .5 after a mutual defection.  In the expected
+  payoffs case strategies are more robust.  Parameters: \(N =100, c=1, b=10,
+  \beta=1\).}\label{fig:memory_one_high_benefit}
+\end{figure}
 

### 2023-02-23 paper/main.tex
milestone in writing

@@
 assumed that individuals play many times and with all other players before
-reproduction takes. So that in the updating phase the updating payoffs, and
-subsequently the fitness, of individuals us given by the average payoffs of an
+reproduction takes place. So that in the updating phase the updating payoffs, and
+subsequently the fitness, of individuals is given by the average payoffs of an
 infinitely repeated game against the mean distribution of types in the
@@
 
-We first consider two extreme scenarios, the classical scenario where individual
+We first consider two extreme scenarios, the classical scenario where an individual
 has a perfect memory of all their interactions and the alternative scenario of
@@
 In the game phase individuals are matched in pairs and that they participate in
-a repeated 2 persons donation game; a special case of the prisoner's dilemma. In
+a repeated 2 person donation game; a special case of the prisoner's dilemma. In
 the donation game there are two actions: cooperation (\(C\)) and defection
@@
 In repeated games there are infinite many strategies, however, similar to the
-literature we will assume that individuals use reactive strategies. A reactive strategy considers only the
-previous action of the other player, and thus, a reactive strategy \(s\) can be
-written as a three-dimensional vector \(s=(y, p, q)\). The parameter \(y\) is
-the probability that the strategy opens with a cooperation and \(p\), \(q\) are
-the probabilities that the strategy cooperates given that the opponent
-cooperated and defected equivalently. The play between a pair of reactive
-strategies can be model as a Markov process with the transition matrix \(M\),
+literature we will assume that individuals use reactive strategies. A reactive
+strategy considers only the previous action of the other player, and thus, a
+reactive strategy \(s\) can be written as a three-dimensional vector \(s=(y, p,
+q)\). The parameter \(y\) is the probability that the strategy opens with a
+cooperation and \(p\), \(q\) are the probabilities that the strategy cooperates
+given that the opponent cooperated and defected equivalently. The play between a
+pair of reactive strategies ($s_1 = (y_1, p_1, q_1)$, $s_2 = (y_2, p_2, q_2)$)
+can be model as a Markov process with the transition matrix \(M\),
 
@@
 and the stationary vector \(\mathbf{v}(s_1,s_2)\)
-which is the solution to \(\mathbf{v}(s_1,s_2) \cdot M 
+which is the solution to \(\mathbf{v}(s_1,s_2) \times M 
 = \mathbf{v}(s_1,s_2)\).
@@
 
-We will mainly assume that mutations are rare (\(\mu \rightarrow 0\)). In fact,
-so rare that only two different strategies can be present in the population at
-any given time. However, we also show in section~\ref{section:mutation} that the
-main results hold for \(\mu \neq 0\). The case of low mutation is vastly adopted because it allows us to explicitly
-calculate the fixation probability of a newly introduced mutant. At each step
+For the results presented here we assume that mutations are rare (\(\mu
+\rightarrow 0\)). In fact, so rare that only two different strategies can be
+present in the population at any given time. However, in the Supplementary Information
+Section 4 we show in that the main result hold for \(\mu \neq 0\).
+The case of low mutation is vastly adopted because it allows us to explicitly
+calculate the fixation probability of a newly introduced mutant. More specifically,
+at each step
 one individual adopts a mutant strategy randomly selected from the set of
@@
 \begin{equation}\label{eq:appendix_fixation_probability}
-    \varphi = \frac{1}{1+\sum\limits_{i=1}^{N-1}\prod\limits_k^i \frac{\lambda^-_k}{\lambda^+_k}},
+    \varphi_{M} = \frac{1}{1+\sum\limits_{i=1}^{N-1}\prod\limits_k^i \frac{\lambda^-_k}{\lambda^+_k}},
 \end{equation}
@@
 where \(\lambda^-_k, \lambda^+_k\) are the probabilities that the number of
-mutants decreases and increases respectively, and \(k\) is the number of mutants. Depending on the fixation
-probability \(\phi_{M}\) the mutant either fixes (becomes the new resident) or
-goes extinct. Regardless, in the elementary time step another mutant strategy is
-introduced to the population. We iterate this elementary population updating
-process for a large number of mutant strategies and we record the resident
-strategies at each time step.
-
-The probabilities \(\lambda^-_k \text{ and } \lambda^+_k\) depend on the
-updating payoffs of the mutant and the resident strategies. In the next
-section we discuss the fitness of individuals in the case of perfect and
-limited memory.
+mutants decreases and increases respectively, and \(k\) is the number of
+mutants. Depending on the fixation probability \(\phi_{M}\) the mutant either
+fixes (becomes the new resident) or goes extinct. Regardless, in the elementary
+time step another mutant strategy is introduced to the population. We iterate
+this elementary population updating process for a large number of mutant
+strategies and we record the resident strategies at each time step. The
+probabilities \(\lambda^-_k \text{ and } \lambda^+_k\) depend on the updating
+payoffs of the mutant and the resident strategies. In the next section we
+present how they are calculated in the cases of perfect and limited memory.
 
@@
 strategy \(s_1\) against the reactive strategy \(s_2\) can explicitly be
-calculated using the stationary vector \(\mathbf{v}\) and the payoffs' vector as,
-
-\[\langle\mathbf{v}(s_1,s_2),\mathbf{u}\rangle \quad.\]
-
-In a population of size \(N\), there are \(k\) mutants and \(N - k\) residents,
-whose strategies we denote respectively as \(s_M =(y_M, p_M, q_M)\) and \(s_R =
-(y_R, p_R, q_R)\). The expected payoffs of a resident (\(\pi_R\)) and for a
-mutant (\(\pi_M\)) are give by,
+calculated using the stationary vector \(\mathbf{v}\) and the payoffs' vector as
+\(\langle\mathbf{v}(s_1,s_2),\mathbf{u}\rangle\). In a population of size \(N\)
+there are \(k\) mutants and \(N - k\) residents. We denote the strategies of
+a mutant and of a resident as \(s_M =(y_M, p_M, q_M)\) and \(s_R =
+(y_R, p_R, q_R)\). The expected payoffs of a resident (\(\pi_R\)) and of a
+mutant (\(\pi_M\)) are given by,
 
@@
 
-\begin{align}
+\begin{align}\label{eq:perfect_memory_lambdas}
   \lambda^-_k \!=\!\rho(\pi_M, \pi_R) \quad \text{ and } \quad \lambda^+_k \!=\!\rho(\pi_R, \pi_M).
@@
 reactive strategy receives the payoff $u\!\in\! \mathcal{U}$ in the very last
-round of the game. This is given by Proposition~\ref{proposition:last_round}.
-The proof can be found in the Appendix [ref].
+round of the game. This is given by Proposition~\ref{proposition:last_round}
+(see Supplementary Information Section 2.2.1 for proof).
 
@@
 
-We consider the case where both the role model and the learner estimate their
-fitness after interacting with a single member of the population. There are five
-possible pairings at each step. They interact with other with a probability
-\(\frac{1}{N - 1}\), and thus they do not interact with other with a probability
-\(1 - \frac{1}{N - 1}\). In the latter case, each of them can interact with
-either a mutant or a resident. Both of them interact with a mutant with a
-probability $\frac{(k-1)(k-2)}{(N-2)(N-3)}$ and both interact with a resident
-with a probability $\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$. The last two possible
-pairings are that either of them interacts with a resident whilst the other
-interacts with a mutant, and this happens with a probability
-$\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$. The probability that the respective last
-round payoffs of two players \(s_1, s_2\) are given by $u_1$ and $u_2$ as,
+In the case of limited payoffs memory both the role model and the learner
+estimate their fitness after interacting with a single member of the population.
+At each time step there are five possible pairings. They interact with each
+other with a probability \(\frac{1}{N - 1}\), and thus they do not interact with
+other with a probability \(1 - \frac{1}{N - 1}\). In the latter case, each of
+them can interact with either a mutant or a resident. Both of them interact with
+a mutant with a probability $\frac{(k-1)(k-2)}{(N-2)(N-3)}$ and both interact
+with a resident with a probability $\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$. The last
+two possible pairings are that either of them interacts with a resident whilst
+the other interacts with a mutant, and this happens with a probability
+$\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$. Thus we the probability that the randomly
+chosen resident obtained a payoff of $u_R$ in the last round of his respective
+game, and that the mutant obtained a payoff of $u_M$ as $x(u_R,u_M)$.
 
@@
 the role model happened to be matched during the game stage, which happens with
-probability $\frac{1}{(N\!-\!1)}$. In that case, we note that only those payoff
-pairs can occur that are feasible in a direct interaction,
-$\splitatcommas{(u_1,u_2)\in \mathcal{U}^2_F:=\big\{ (r, r), (s, t), (t, s), (p, p)
-\big\}}$, as represented by the respective indicator function. Otherwise, if the
-learner and the role model did not interact directly, we need to distinguish
-four different cases, depending on whether the learner was matched with a
-resident or a mutant, and depending on whether the role model was matched with a
-resident or a mutant.
-
-The probability that the number of mutants increases, and decreases respectively,
-by one is now given by,
-
-\begin{align}
-\lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M})\cdot \rho(u_{R},u_{M}), \\
-\lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M})\cdot \rho(u_{M},u_{R}).
+probability $\frac{1}{(N\!-\!1)}$. The probability that the number of mutants
+increases and decreases by one in the case of limited memory are now given by,
+
+\begin{align}\label{eq:limited_memory_lambdas}
+\lambda^+_k=\frac{N\!-\!k}{N} \frac{k}{N} \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M}) \rho(u_{R},u_{M}) \quad \text{and} \quad
+\lambda^-_k=\frac{N\!-\!k}{N} \frac{k}{N} \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M}) \rho(u_{M},u_{R}).
 \end{align}
@@
 
-% We use
-% $x(u_R,u_M)$ to denote the probability that the randomly chosen resident
-% obtained a payoff of $u_R$ in the last round of his respective game, and that
-% the mutant obtained a payoff of $u_M$.
-
 \subsubsection*{Simulation Results}
@@
 evolutionary process and record which strategies the players adopt over time
-based on perfect memory payoffs and limited memory payoffs. We run two
+based on perfect memory payoffs and limited memory payoffs. In the case of
+perfect memory payoffs we use Eq~\ref{eq:perfect_memory_lambdas} when we
+estimate the probability that a mutant fixes, and
+Eq~\ref{eq:limited_memory_lambdas} in the case on limited memory. We run two
 independent runs for each approach and we vary the value of benefit \(b\).
@@
 $q$. The discount factor~$\delta$ is comparably high, thus we do not report the
-opening move \(y\) as it is a transient effect. The upper panel correspond to
+opening move \(y\) as it is a transient effect. The upper panel corresponds to
 the standard scenario considered in the literature, it considers players who use
@@
 players will occasionally forgive a defection more often if their fitness
-depends on interacting with every member of the population, and this effects
-becomes more abundant as the value of the benefit increases. In comparison, in
-the case of limited memory the evolved generosity is independent of the values
-of benefit, and they are less forgiving. We also calculate the average
+depends on interacting with every member of the population, and this effect
+becomes more notable as the value of the benefit increases. In the case of
+limited memory the resident strategies are less forgiving and their forgiveness
+is independent of the values of benefit. In the Supplementary Information
+Sections 2.1 and 2.2.1 we show that $q$-values of a resident strategy have to be less
+than \(1 - c/b\) in the perfect memory case and less than \(0.5\) in the limited
+memory case. The generosity for the perfect memory case is always higher for a
+benefit value higher than 2, and it increases as the benefit value increases.
+This leads to a more cooperative population. We calculate the average
 cooperation rate for each simulation which is the average cooperation rate
 within the resident population. In the case of the perfect memory payoffs the
-average cooperation rate is strictly higher than that of the last round payoffs.
-The difference based on the two methods is statistically significant, and in the
-case of $b=10$ the average cooperation of resident strategies drops from 97\% to
-57\%.
+average cooperation rate is strictly higher than that of the last round payoffs,
+and the difference is most for $b=10$. The average cooperation of resident
+strategies drops from 97\% to 57\%. This indicates that the expected payoffs
+overestimate the evolved cooperation.
 
@@
     \caption{{\bf Evolutionary dynamics under perfect and limited payoff memory.}
-    ({\bf Schematic illustrations})
-    On the left panels we show schematic illustrations of the perfect memory and
-    the limited memory cases. The shaded background denotes the game phase
-    information that an individual considered when updating strategies. In the
-    case of perfect memory the entire region is shaded and in the case of
-    limited memory only one turn with a single member of the population.
-    ({\bf Simulations})
-    We have run four simulations of the pairwise comparison process
-    for $t\!=\!10^7$ time steps. For each time step,
+    ({\bf Schematic illustrations}) On the left panels we show schematic
+    illustrations of the perfect memory and the limited memory cases. The shaded
+    background denotes the game phase information that an individual considered
+    when updating strategies. In the case of perfect memory the entire region is
+    shaded and in the case of limited memory only one turn with a single member
+    of the population. ({\bf Simulations}) We have run four simulations of the
+    pairwise comparison process for $T\!=\!10^7$ time steps. For each time step,
     we have recorded the current resident population ($y,p,q$). Since
@@
     calculated as the average cooperation rate within the resident population.
-    ({\bf Perfect Memory}) In the case of low benefit
-    the resident population either consists of defectors (with $p\!\approx\!q\!\approx\!0$) or of
-    conditional cooperators. Conditional cooperators, or otherwise known as generous
-    tit for tat, are a set of strategies that always cooperate following a cooperation
-    ($p\!\approx\!1\!$) and cooperate with a probability
-    $q$ given that the co-player has defected. $q$ denotes the generosity
-    of a player. The resident population applies a conditional cooperator strategy for which
-    $q\!\le\!1\!-\!c/b\!=\!0.67$. This is true also for the case of high benefit.
-    In this case the population mainly consists of conditional
-    cooperators of the form ($p\!\approx\!1\!, q\!\le\!1\!-\!1/10\!=\!0.9$).
-    In the Appendix we show that a conditional cooperator needs to be
-    of the form ($p\!\approx\!1\!, q\!\le\!1\!-\!c/b$) to not be invaded by
-    defecting strategies. A higher generosity in the population results to a
-    higher average cooperation rate. The average cooperation rate
-    increases from a 52\% for \(b\) to 98\% for \(b=10\).
-    ({\bf Limited Memory}) When players
-    update their strategies based on their realized payoffs in the last round,
-    there are two different predominant behaviors regardless of the benefit value. The resident population
+    ({\bf Perfect Memory}) In the case of low benefit the resident population
     either consists of defectors (with $p\!\approx\!q\!\approx\!0$) or of
-    conditional cooperators. The maximum level of $q$
-    consistent with stable cooperation is somewhat smaller compared to the
-    perfect memory setting, $q\!<\!0.5$. Namely, in the Appendix we show
-    that regardless the value of benefit, in the case of limited payoff memory,
-    an individual needs generosity smaller of $0.5$
-    to repel defectors. The plots
+    conditional cooperators. Conditional cooperators, or otherwise known as
+    generous tit for tat, are a set of strategies that always cooperate
+    following a cooperation ($p\!\approx\!1\!$) and cooperate with a probability
+    $q$ given that the co-player has defected. $q$ denotes the generosity of a
+    player. The resident population applies a conditional cooperator strategy
+    for which $q\!\le\!1\!-\!c/b\!=\!0.67$. This is true also for the case of
+    high benefit. In this case the population mainly consists of conditional
+    cooperators of the form ($p\!\approx\!1\!, q\!\le\!1\!-\!1/10\!=\!0.9$). In
+    the Supplementary Information Section 2.1 we show that a conditional
+    cooperator needs to be of the form ($p\!\approx\!1\!, q\!\le\!1\!-\!c/b$) to
+    not be invaded by defecting strategies. A higher generosity in the
+    population results in a higher average cooperation rate. The average
+    cooperation rate increases from 52\% for \(b\) to 98\% for \(b=10\). ({\bf
+    Limited Memory}) When players update their strategies based on their
+    realized payoffs in the last round, there are two different predominant
+    behaviors regardless of the benefit value. The resident population either
+    consists of defectors (with $p\!\approx\!q\!\approx\!0$) or of conditional
+    cooperators. The maximum level of $q$ consistent with stable cooperation is
+    somewhat smaller compared to the perfect memory setting, $q\!<\!0.5$.
+    Namely, in the Supplementary Information Section 2.2.1 we show that
+    regardless of the value of benefit, in the case of limited payoff memory, an
+    individual needs a generosity smaller of $0.5$ to repel defectors. The plots
     are fairly similar regardless of the benefit, and the evolved cooperation
-    rate only sightly increases from 37\% to 51\%. Parameters: $N\!=\!100$, $c\!=\!1$, $\beta\!=\!1$,
-    $\delta\!=\!0.999$.}
+    rate only slightly increases from 37\% to 51\%. Parameters: $N\!=\!100$,
+    $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.999$.}
     \label{fig:expected_and_stochastic_for_donation}
@@
 
-We further explore the effect that the value of benefit \(b\) and the strength
-of selection \(\beta\) have on the generosity \(q\) and to the evolved
-cooperation rate, Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}. The
-figure suggests that expected payoffs always yield a higher cooperation rate. In
-the case of expected payoffs we observe that the cooperation rate increases as
-the value of the benefit gets higher. In comparison for the limited memory
-payoffs, the cooperation rate remains unchanged at approximately 50\% once
-\(b=5\). We also investigate the effect of the strength of selection $\beta$.
-For weak selection, \(\beta < 1\), we observe that the two methods yield similar
-results, however, as \(\beta\) increases there is variation in the evolving
-populations. In the case of expected payoffs the resident populations become
-more cooperative as \(\beta\) increases, whereas in the case of limited memory
-payoffs, the resident populations become more defective.
+We further explore the effects of the cooperation benefit and the strength of
+selection on the evolved generosity \(q\) and cooperation rate
+(Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}).
+Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}\textbf{A} suggests that
+perfect memory always yields a higher cooperation rate. We observe that the
+cooperation rate increases as the value of the benefit gets higher, whereas in
+comparison for the limited memory payoffs, the cooperation rate remains
+unchanged (as we would have expected) at approximately 50\% once \(b=5\). From
+Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}\textbf{B} we observe
+that for weak selection, \(\beta < 1\), the two methods yield similar results,
+however, as \(\beta\) increases there is variation in the evolving populations.
+In the case of expected payoffs the resident populations become more cooperative
+as \(\beta\) increases, whereas in the case of limited memory payoffs, the
+resident populations become more defective.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{static/cooperation_rate_over_b_and_beta.pdf}
-    \caption{{\bf The evolution of cooperation and generosity for different values
-    of benefit (A) and strength of selection (B).} We report the
-    average cooperation and the average reciprocity. The average cooperation rate
-    is the average cooperation rate
-    within the resident population. For the average reciprocity we check for the
-    residents that ($p \approx 1$) and we take the average of theirs $q$s ({\bf A})
-    We vary the benefit of defection $b$, and we report both the average
-    cooperation and average generosity. In all cases, expected payoffs appear to
-    overestimate the average cooperation rate the population achieves. As
-    expected, since the average generosity over the different values of $beta$
-    remains the same ($q \approx 0.5$), and a result so does the average
-    cooperation. ({\bf B}) We vary the selection strength $\beta$. In all
-    cases, stochastic payoff evaluation tends to reduce the evolving cooperation
-    rates. Unless explicitly varied, the parameters of the simulation are
-    $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$.
-    Simulations are run for $t\!=\!5\times 10^7$ time steps for each parameter
+    \includegraphics[width=.75\textwidth]{static/cooperation_rate_over_b_and_beta.pdf}
+    \caption{{\bf The evolution of cooperation and generosity for different
+    values of benefit (A) and strength of selection (B).} We report the average
+    cooperation and the average reciprocity. The average cooperation rate is the
+    average cooperation rate within the resident population. For the average
+    reciprocity we select the residents that have a $p \approx 1$ and we take
+    the average of their cooperation probability $q$. ({\bf A}) We vary the
+    benefit of cooperation $b$. In all cases, perfect memory updating payoffs
+    appear to overestimate the average cooperation rate the population achieves.
+    As expected in the case of limited memory the average generosity over the
+    different values of benefit remains the same ($q \approx 0.5$), and as a
+    result so does the average cooperation. ({\bf B}) We vary the selection
+    strength $\beta$. For weak selection, \(\beta < 1\), the two methods yield
+    similar results. However, as \(\beta\) increases in the case of limited
+    memory payoffs the resident populations become more defective. Unless
+    explicitly varied, the parameters of the simulation are $N\!=\!100$,
+    $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run
+    for $T\!=\!5\times 10^7$ time steps for each parameter
     combination.}\label{fig:cooperation_rate_over_benefit_and_beta}
@@
   $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$.
-  Simulations are run for $t\!=\!5\times 10^7$ time steps for each parameter
+  Simulations are run for $T\!=\!5\times 10^7$ time steps for each parameter
   combination.}\label{fig:cooperation_rate_updating_payoffs}

### 2023-02-23 paper/si.tex
milestone in writing

@@
 ``updating payoffs''. In Section~\ref{section:updating_payoffs} we
-present/describe the conventional approach for calculating updating payoffs, as
-well as, our newly introduced approach.
+describe the conventional approach for calculating updating payoffs and we
+present our new approaches.
 
@@
 In the \textbf{game phase} individuals are randomly matched with other
-individuals in the population, and they engage in a repeated game where each
+individuals in the population and they engage in a repeated game where each
 subsequent turn occurs with a fixed probability \(\delta\). At each turn the
 individuals decide on an action based on their strategies.
-In repeated games there are infinite many strategies, however, it is commonly
+In repeated games there are infinitely many strategies, however, it is commonly
 assumed that individuals can only choose strategies from a restricted set. One
@@
 
-In the \textbf{update stage} where two individuals are randomly selected. From
+In the \textbf{update stage} two individuals are randomly selected. From
 the two individuals, one  serves as the `learner' and the other as the `role
@@
 \begin{algorithm}[!htbp]
-  \SetAlgoLined $N \leftarrow$ population size\; $k \leftarrow 1$\; resident
-   $\leftarrow$ starting resident\; \While{ $t <$ maximum number of steps}{mutant
-   $\leftarrow$ random: $\{\emptyset \}\ \rightarrow R^n$\; fixation probability
-   $\leftarrow \varphi $\; \If{$\varphi >$ random: $i \rightarrow [0,
-   1]$}{resident $\leftarrow$ mutant;}} \caption{Evolutionary
-   process}\label{algorithm:pairwise_comparison}
+  \SetAlgoLined
+  $N \leftarrow$ population size\;
+  $k \leftarrow 1$\; resident
+   $\leftarrow$ starting resident\;
+   \While{ $t <$ maximum number of steps}
+   {mutant $\leftarrow$ random strategy\;
+   fixation probability $\leftarrow \varphi_M $\;
+   \If{$\varphi_M >$ random: $i \rightarrow [0,1]$}
+   {resident $\leftarrow$ mutant;}}
+   \caption{Evolutionary process}\label{algorithm:pairwise_comparison}
 \end{algorithm}
@@
 however, we have also verified that the main result holds in the case of
-high mutation rates. %ToDo Add reference to section with mu != 0.
+high mutation rates (Section~\ref{section:mutation}).
 
@@
 phase. In this work we assume that in the game phase individuals are matched in
-pairs and that they participate in a repeated 2 persons donation game. In the
+pairs and that they participate in a repeated 2 person donation game. In the
 donation game there are two actions: cooperation (\(C\)) and defection (\(D\)).
@@
 the updating payoffs. Initially, we discuss the conventional approach of the
-expected payoffs and afterwards we present our newly introduced approach.
+expected payoffs and afterwards we present our new approach.
 
-\subsection{Updating Payoffs based on the expected payoffs}
+\subsection{Updating Payoffs based on the expected payoffs}\label{section:perfect_memory}
 
-The expected payoffs are the conventional payoffs used in the updating stage.
-They are defined as the mean payoff of an individual in a well-mixed population
-that engages in an infinitely repeated games with all other population members. In
-an infinitely repeated game the payoff of a reactive strategy can explicitly be
-calculated using a markovian approach~\cite{}. Namely, assume two reactive
+The expected payoffs are the conventional payoffs used in the updating stage. We
+refer to this approach as the perfect memory approach. They are defined as the
+mean payoff of an individual in a well-mixed population that engages in an
+infinitely repeated games with all other population members. In an infinitely
+repeated game the payoff of a reactive strategy can explicitly be calculated
+using a markovian approach~\cite{nowak:APC:1989}. Namely, assume two reactive
 strategies $s_1\!=\!(y_1, p_1, q_1$) and $s_2\!=\!(y_2,p_2,q_2)$, their play can
@@
 
-Note that \(y_1 \text{ and } y_2\) are omitted since they have no effect in the
-long run of the game. The stationary vector \(\mathbf{v}\),
-which is the solution to \(\mathbf{v} M = \mathbf{v}\), combined with the payoff vector
+The stationary vector \(\mathbf{v}(s_1,s_2)\),
+which is the solution to \(\mathbf{v}(s_1,s_2) M = \mathbf{v}(s_1,s_2)\), combined with the payoff vector
 \(\mathbf{u}\), yields the game stage outcome for each strategy,
@@
 
+\(\frac{N\!-\!k\!-\!1}{N-1}\) is the probability that a resident meets another
+resident and \(\frac{k}{N-1}\) is the probability that a resident meets a
+mutant. Likewise, \(\frac{N-k}{N-1}\) is the probability that a mutant interacts
+with a resident and \(\frac{k-1}{N-1}\) the probability that the mutants
+interacts with a mutant.
+
 The number of mutants in the population increases if a resident adopts the strategy
@@
 
-\subsubsection*{Simulation Results based on the expected payoffs}
+\subsubsection*{Simulation Results based on Expected Payoffs}
 
-We can simulate the evolutionary process described by
-Algorithm~\ref{algorithm:pairwise_comparison} when individuals use reactive
+We simulate the evolutionary process, as described in
+Algorithm~\ref{algorithm:pairwise_comparison}, when individuals use reactive
 strategies and update their strategies based on their expected payoffs. We
@@
 cooperation \(b\). The results are shown in
-Figure~\ref{fig:expected_payoffs_results}.
-
-It is observed that a higher value of benefit results in a more cooperative
-population. For a low benefit the resident population cooperates on average 52\%
-of the time. For a a high benefit the average cooperation increases to 98\%. For
-both a low and a high benefit the resident population consists either of
-defectors or conditional cooperators. A conditional cooperator, otherwise
-known as a generous tit for tat (GTFT) strategy, always cooperates if the
-co-player cooperated ($p\approx1$) and cooperates with a probability \(q\) if
-the co-player defected. The conditional cooperator strategy adopted by the
-population differs between the two simulations. A higher value of benefit
-results to a higher tolerance to defection.
+Figure~\ref{fig:expected_payoffs_results}. We observed that a higher value of
+benefit results in a more cooperative population. For a low benefit the resident
+population cooperates on average 52\% of the time. For a high benefit the
+average cooperation increases to 98\%. For a low benefit the resident population
+consists either of defectors or conditional cooperators, and for a high value of
+benefit the population predominantly consists of conditional cooperators.A
+conditional cooperator, otherwise known as a generous tit for tat (GTFT)
+player, always cooperates if the co-player cooperated ($p\approx1$) and
+cooperates with a probability \(q\) if the co-player defected. The conditional
+cooperator strategy adopted by the population differs between the two
+simulations, namely, a higher value of benefit results in higher values
+of \(q\).
 
@@
     low (left) and high (right) benefit.} We run the simulations for \(T =
-    10^7\) time steps. For each time step, we have recorded the current resident
-    population \((y, p, q)\). Since in the case of the expected payoffs the
-    individuals interact for an infinite number of times \(\delta \rightarrow
-    1\), we do not report the players' initial cooperation probability \(y\).
+    10^7\) time steps. For each time step we recorded the current resident
+    population \((y, p, q)\). Since \(\delta \rightarrow
+    1\) we do not report the players' initial cooperation probability \(y\).
     The graphs show how often the resident population chooses each combination
     \((p, q)\) of conditional cooperation probabilities in the subsequent
-    rounds. In both cases players update based on their expected payoffs (A) In
+    rounds. In both cases players update based on their expected payoffs.
+    ({\bf Low Benefit}) In
     the case where \(b=3\), the resident population either consists of defectors
     (with \(p \approx q \approx 0\)) or of conditional cooperators for which \(p
-    \approx 1\) and \(q \leq 1 - \frac{1}{3}=0.7\). (B) In the case where
+    \approx 1\) and \(q \leq 1 - \frac{1}{3}=0.7\). ({\bf High Benefit}) In the case where
     \(b=10\), the resident population typically applies a strategy for which \(p
@@
 
-\subsubsection*{Invasion Analysis based on the expected payoffs}
+\subsubsection*{Invasion Analysis based on Expected Payoffs}
 
 We can explicitly calculate the value of \(q\) which evolves. Namely, we will
-calculate how easily a single ALLD mutant can invade into a resident population
-with strategy GTFT. In that case, GTFT\(= (1, 1, q)\), ALLD\(= (0, 0, 0)\), and
-\(k = 1\).
-
-When two GTFT players interact in an infinitely repeated game, their respective
-probabilities for each of the four outcomes in the last round simplify to,
+calculate how easily a single defecting mutant can invade into a resident
+population of GTFT players. Let GTFT\(= (1, 1, q)\), ALLD\(= (0, 0, 0)\), and
+\(k = 1\). When two GTFT players interact in an infinitely repeated game, their
+respective probabilities for each of the four outcomes in the last round
+simplify to,
 
@@
 
-The aim of this work is to explore the effect that the updating payoffs have on
-the cooperation rate of the evolved population. To this end we introduce
-a new method calculating the updating payoffs. Compared to the literature
-where a infinite number of rounds and a well mixed population are assumed,
-we constrain the number of turns and number of turns.
+The aim of this work is to explore the effect of the updating payoffs on the
+evolution of cooperation. To this end we introduce a new method for calculating
+the updating payoffs and we compare the results of the new method to those of
+the expected payoffs. In our approach we do not assume that an individual has a
+perfect memory. Instead an individual remembers only a limited number of
+interactions and limited outcomes from each interaction. In the expected payoffs
+case the payoff of a pair depends on the average payoff they received over an
+infinite number of turns. In our approach, the payoff of a pair depends on the
+average payoffs they received in the last \(m\) turns. Furthermore, in expected
+payoffs it is assumed that a player interacts with every member of the
+population, whereas in our approach a player has \(n\) interactions.  Here we
+present results for four cases; (\(m=1\) and \(n=1\)), (\(m=2\) and \(n=1\)),
+(\(m=1\) and \(n=2\)), and (\(m=2\) and \(n=2\)).
 
-Namely, in the expected payoffs the payoff of a pair depends on the average
-payoff they received over an infinite number of turns. In our approach, the
-payoff of a pair depends on the average payoffs they received in the last \(m\)
-turns. Furthermore, in expected payoffs it is assumed that a player interacts
-with every member of the population, whereas in our approach with consider one
-\(n\) interactions of the player with other members. Here we present results for
-four cases; (\(m=1\) and \(n=1\)), (\(m=2\) and \(n=1\)), (\(m=1\) and \(n=2\)),
-and (\(m=2\) and \(n=2\)).
-
-\subsubsection{Updating Payoffs based on the last round payoff of one interaction (\(m=1\) and \(n=1\))}\label{section:m_one_n_one}
+\subsubsection{Updating Payoffs based on the Last Round Payoff of One Interaction (\(m=1\) and \(n=1\))}\label{section:m_one_n_one}
 
@@
 At each step of the evolutionary process we choose a role model and a learner to
-update the population. We consider the case where both the role model and the
-learner estimate their fitness after interacting with a single member of the
-population. There are five possible pairings at each step. They interact with
-other with a probability \(\frac{1}{N - 1}\), and thus they do not interact with
-other with a probability \(1 - \frac{1}{N - 1}\). In the latter case, each of
-them can interact with either a mutant or a resident. Both of them interact with
-a mutant with a probability $\frac{(k-1)(k-2)}{(N-2)(N-3)}$ and both interact
-with a resident with a probability $\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$. The last
-two possible pairings are that either of them interacts with a resident whilst
-the other interacts with a mutant, and this happens with a probability
-$\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$.
-
-Given the possible pairings and Proposition~\ref{proposition:last_round}, we
-define the probability that the respective last round payoffs of two players
-\(s_1, s_2\) are given by $u_1$ and $u_2$ as,
+update the population. In this case both the role model and the learner estimate
+their fitness after interacting with a single member of the population, and so
+there are five possible pairings at each step. They interact with each other with a
+probability \(\frac{1}{N - 1}\), and they do not interact with other with a
+probability \(1 - \frac{1}{N - 1}\). In the latter case, each of them can
+interact with either a mutant or a resident. Both of them interact with a mutant
+with a probability $\frac{(k-1)(k-2)}{(N-2)(N-3)}$ and both interact with a
+resident with a probability $\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$. The last two
+possible pairings are that either of them interacts with a resident whilst the
+other interacts with a mutant, and this happens with a probability
+$\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$. Given the possible pairings and
+Proposition~\ref{proposition:last_round}, we define the probability that the
+respective last round payoffs of two players \(s_1, s_2\) are given by $u_1$ and
+$u_2$ as,
 
@@
 pairs can occur that are feasible in a direct interaction,
-$\splitatcommas{(u_1,u_2)\in \mathcal{U}^2_F:=\big\{ (r, r), (s, t), (t, s), (p, p)
-\big\}}$, as represented by the respective indicator function. Otherwise, if the
-learner and the role model did not interact directly, we need to distinguish
+$\splitatcommas{(u_1,u_2)\in \mathcal{U}^2_F:=\big\{ (r, r), (s, t), (t, s), (p,
+p) \big\}}$, as represented by the respective indicator function. Otherwise, if
+the learner and the role model did not interact directly, we need to distinguish
 four different cases, depending on whether the learner was matched with a
@@
 adopts the role model's strategy over all possible payoffs $u_R$ and $u_M$ that
-the two player may have received in their respective last rounds. We use
+the two players may have received in their respective last rounds. We use
 $x(u_R,u_M)$ to denote the probability that the randomly chosen resident
@@
 
-\subsubsection*{Simulation Results based on the last round payoff of one interaction}
+\subsubsection*{Simulation Results based on the Last Round Payoff of One Interaction}
 
-We simulate the evolutionary process given that individuals now use the the last
+We simulate the evolutionary process given that individuals now use the last
 round payoff of one interaction to update their strategies. The results are
@@
 to the results of the expected payoffs, a higher benefit results in a more
-cooperative population. However we note that the cooperation rate of the evolved
-populations are lower compared to the case of expected payoffs, and that the
-cooperation rate increases less between the two simulations. The evolving both
-consist of defectors and conditional cooperators. The evolved \(q\)s are smaller
-which also results to a less cooperative population. In between the two
-simulations \(q \approx \frac{1}{2}\). In the invasion analysis we will
-demonstrate that in the case of the the last round payoff of one interaction, a
-conditional player needs to have \(q \leq \frac{1}{2}\) to avoid being invaded
-by a defector. In both simulations \(q \leq \frac{1}{2}\), the higher benefit
-pushes the evolved population closer to the bounty.
+cooperative population. However, we note that the cooperation rate of the
+evolved populations is lower compared to the case of expected payoffs, and that
+the increase of the cooperation rate between the two simulations is smaller. The
+residents population in both cases consist of defectors and conditional
+cooperators with \(p \approx 1\) and  \(q \approx \frac{1}{2}\). Thus, the
+evolved \(q\)s are smaller which is the reason why the population is less
+cooperative. In the invasion analysis we demonstrate that in the case of the
+last round payoff of one interaction, a conditional player needs to have \(q
+\leq \frac{1}{2}\) to avoid being invaded by a defector regardless of the
+benefit of cooperation.
 
@@
   \caption{\textbf{Evolutionary dynamics under one interaction last round
-  payoffs with low (left) and high (right) benefit.} The cooperation rates
-  in both cases are less compared to the expected payoffs. Moreover, increase is
-  less when the benefit is increased to ten. In the cases of low and
-  high benefit case the resident population either consists of defectors (with
+  payoffs with low (left) and high (right) benefit.} The cooperation rate
+  regardless of the benefit of cooperation is less compared to that of the
+  expected payoffs. The increase of cooperation between the low benefit case and
+  the high benefit case is also smaller. In both cases the resident population
+  either consists of defectors (with
   \(p \approx q \approx 0\)) or of conditional cooperators for which \(p \approx
-  1\) and \(q \leq \frac{1}{2}\). Parameters: \(N =100, c=1, \beta=1\).
+  1\) and \(q \leq \frac{1}{2}\). Parameters: \(N =100, c=1, \beta=1, T=10^{7}\).
   }\label{fig:one_interaction_last_round_payoffs_results}
@@
 
-\subsubsection*{Invasion Analysis based on the last round payoff of one interaction}
+\subsubsection*{Invasion Analysis based on the Last Round Payoff of One Interaction}
 
 We once again calculate how easily a single ALLD mutant can invade into a
-resident population with strategy GTFT. When two GTFT players interact in the
-game, their respective probabilities for each of the four outcomes in the last
-round simplify to,
+resident population of GTFT player. When two GTFT players interact in the game,
+their respective probabilities for each of the four outcomes in the last round
+simplify to,
 
@@
 
-Here, in the updating phase we choose one role model and one learner, however
-now, we consider the case where both the role model and the learner estimate
+Here we consider the case where both the role model and the learner estimate
 their updating payoff after interacting with two members of the population.
-There are stages in the pairing process and there are twenty four possible
-pairings.
-
-In the first stage the role model and the learner are matched together with a
+There are two stages of matching and there are twenty four possible pairs. In
+the first stage the role model and the learner are matched together with a
 probability $\frac{1}{N-1}$ or not with a probability $(1 - \frac{1}{N-1})$. If
-they were matched then in the second stage of pairings there are only four
-possible outcomes; both of them interact with a mutant with a probability
+they were matched together then in the second stage there are only four possible
+outcomes; both of them interact with a mutant with a probability
 $\frac{(k-1)(k-2)}{(N-2)(N-3)}$ and both interact with a resident with a
-probability $\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$. The last two possible pairings
-are that either of them interacts with a resident whilst the other interacts
-with a mutant, and this happens with a probability
-$\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$.
-
+probability $\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$. The last two possible pairs are
+that either of them interacts with a resident whilst the other interacts with a
+mutant, and this happens with a probability $\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$.
 In the later case, where they were not matched in the first stage, there are
@@
 with a resident, either of them interacts with a resident whilst the other
-interacts with a mutant. For each of the above pairings of the first stage there
-are five possible; they interact with each other, both of them interact with a
-mutant and both interact with a resident, either of them interacts with a
-resident whilst the other interacts with a mutant.
+interacts with a mutant. For each of the above pairs of the first stage there
+are five possible pairs in the second stage; they interact with each other, both
+of them interact with a mutant and both interact with a resident, either of them
+interacts with a resident whilst the other interacts with a mutant.
 
-The new possible pairings change how we define the probability that the
+The new possible pairs change how we define the probability that the
 respective last round payoffs of two players \(s_1, s_2\) are given by \(u_1\)
-and \(u_2\). The probability denoted as \(\bar{x}(u_1,u_2)\) is given by,
+and \(u_2\). The new probability denoted as \(\tilde{x}(u_1,u_2)\) is given by,
 
@@
   \begin{array}{llrl}
-  x(u_1,u_2)	& = \displaystyle \frac{1}{N\!-\!1}\cdot v_{u_1}(s_1,s_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F} \cdot A + \left(1\!-\!\frac{1}{N\!-\!1}\right) \left[ \right. \\[0.5cm] 
+  \tilde{x}(u_1,u_2)	& = \displaystyle \frac{1}{N\!-\!1}\cdot v_{u_1}(s_1,s_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F} \cdot A + \left(1\!-\!\frac{1}{N\!-\!1}\right) \left[ \right. \\[0.5cm] 
   & \left. v_{u_1}(s_1,s_2) _{u_2}(s_2,s_2)  \frac{\left(k - 2\right) \left(k - 1\right)}{(N\!-\!2)(N\!-\!3)}
@@
 with a mutant, and this happens with a probability
-$\frac{(N-k-2)(k-1)}{(N-4)(N-3)}$. The rest of the cases follow the same
-pattern.
+$\frac{(N-k-2)(k-1)}{(N-4)(N-3)}$, and so on.
 
-The probability that the number of mutants increases, and decreases respectively,
-by one is now given by,
+The probability that the number of mutants increases, and decreases
+respectively, by one is now given by,
 
-\begin{align}
-\lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R},u_{M}\in\mathcal{U}} \bar{x}(u_{R},u_{M})\cdot \rho(u_{R},u_{M}), \\
-\lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R},u_{M}\in\mathcal{U}} \bar{x}(u_{R},u_{M})\cdot \rho(u_{M},u_{R}).
+\begin{align}\label{eq:ratio_limited}
+\lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R},u_{M}\in\mathcal{U}} \tilde{x}(u_{R},u_{M})\cdot \rho(u_{R},u_{M}), \\
+\lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R},u_{M}\in\mathcal{U}} \tilde{x}(u_{R},u_{M})\cdot \rho(u_{M},u_{R}).
 \end{align}
 
-\subsubsection*{Simulation Results based on the last round payoff of two interactions}
+\subsubsection*{Simulation Results based on the Last Round Payoff of Two Interactions}
 
 We simulate the evolutionary process when individuals update their strategies
-based on the last round payoffs from two interactions with other members of the
+based on the last round payoffs from two interactions with members of the
 population, Figure~\ref{fig:one_interaction_last_round_payoffs_results}. For
 both a low and a high benefit value the average cooperation rate has increased
-compared to the the last round against a single other member of the population.
-In the case of low benefit the evolved population's cooperation rate is almost
-the same as in the case of expected payoffs. However, the cooperation rate does
-not increase as much as in for the expected payoffs for \(b=10\).
+compared to the last round against a single other member of the population. In
+the case of low benefit the evolved population's cooperation rate is almost the
+same as in the case of expected payoffs. However, the cooperation rate does not
+increase as much as the expected payoffs for \(b=10\).
 
@@
   \caption{\textbf{Evolutionary dynamics under two interactions last round
-  payoffs with low (left) and high (right) benefit.} In the cases of low and
-  high benefit case the resident population either consists of defectors (with
+  payoffs with low (left) and high (right) benefit.}
+  The cooperation rates for both the simulations have increased and are similar
+  to the case of expected payoffs. The increase of the cooperation rate between
+  the two cases however is smaller. In both cases the resident population
+  either consists of defectors (with
   \(p \approx q \approx 0\)) or of conditional cooperators for which \(p \approx
-  1\) and \(q \leq \frac{\sqrt{2}}{2}\). Parameters: \(N =100, c=1, \beta=1\).
+  1\) and \(q \leq \frac{1}{2}\). Parameters: \(N =100, c=1, \beta=1, T=10^{7}\).
   }\label{fig:two_interactions_results}
@@
 
-\subsubsection*{Invasion Analysis based on the last round payoff of two interactions}
+\subsubsection*{Invasion Analysis based on the Last Round Payoff of Two Interactions}
 
-In a similar fashion we can calculate the threshold of \(q\) for which a GTFT
-player can not be invaded by an ALLD mutant. The ratio of transition probabilities
-is given by,
+In a similar fashion we can calculate the condition for which a population of
+GTFT players can not be invaded by an ALLD mutant. The ratio of transition
+probabilities is given by,
 
@@
 In the limit of strong selection \(\beta \rightarrow \infty\) and large
-populations \(N \rightarrow \infty \), we obtain
+populations \(N \rightarrow \infty \) we obtain the following cases,
 
@@
 \begin{cases}
-  \frac{\left(\delta q - \delta + 1\right)^{2}}{\delta^{2} \left(1 - q\right)^{2} + 2 \delta \left(1 - q\right) \left(\delta q - \delta + 1\right)}  & \frac{b}{2} > c \\
-  \frac{- \delta q + \delta - 1}{\delta \left(q - 1\right)}  & \frac{b}{2} = c \\
-  - \frac{\left(\delta q - \delta - 1\right) \left(\delta q - \delta + 1\right)}{\delta^{2} \left(q - 1\right)^{2}} & \frac{b}{2} < c
+  - \frac{q \left(\delta q - \delta + 1\right)}{\left(q - 1\right) \left(\delta q + 1\right)}  & \frac{b}{2} > c \\[0.5cm]
+  \frac{- \delta q + \delta - 1}{\delta \left(q - 1\right)}  & \frac{b}{2} = c \\[0.5cm]
+  - \frac{\delta q^{2} - 2 \delta q + \delta - 1}{\delta \left(q - 1\right)^{2}} & \frac{b}{2} < c
 \end{cases}
@@
 
-Note that in this case the relationship of the payoffs, cost and benefit, have
-an effect of the \(q\). Namely if \(\frac{b}{2}\) is higher than the cost \(c\).
-In the case where \(\frac{b}{2}=c\) the result is the same as in the case
-of \(m=n=1\). For \(\frac{\lambda^{+}}{\lambda^{-}} < 1\):
+We note that the relationship between the cost and benefit have an effect on how
+generous a conditional cooperator must be to avoid invasion. In the case of
+\(\frac{b}{2}=c\) the result remains the same expression as in the case of
+\(m=n=1\). For the other two cases we show that for
+\(\frac{\lambda^{+}}{\lambda^{-}} < 1\),
 
@@
 \begin{cases}
-  q \in \{\frac{\delta - 1 + \frac{\sqrt{2}}{2}}{\delta}\}  & \frac{b}{2} > c \\
-  q = \frac{\delta - \frac{1}{2}}{\delta}  & \frac{b}{2} = c \\
-  q \in \{\frac{\delta - \frac{\sqrt{2}}{2}}{\delta}, \  \frac{\delta + \frac{\sqrt{2}}{2}}{\delta}\} & \frac{b}{2} < c
+  q < \left\{\frac{\delta - 1 - \frac{\sqrt{2}}{2}}{\delta}, \frac{\delta - 1 + \frac{\sqrt{2}}{2}}{\delta}\right\}  & \frac{b}{2} > c \\[0.5cm]
+  q < \left\{\frac{\delta - \frac{\sqrt{2}}{2}}{\delta}, \  \frac{\delta + \frac{\sqrt{2}}{2}}{\delta}\right\} & \frac{b}{2} < c
 \end{cases}
@@
 
-There are two distinguished cases. For \(\frac{b}{2}>c\) the ration is smaller
-than if \(q < \frac{\delta - 1 + \frac{\sqrt{2}}{2}}{\delta}\). For infinitely
-repeated games, \(\delta \rightarrow 1\), this condition becomes \(q <
-\frac{\sqrt{2}}{2}\). In the case of \(\frac{b}{2}<c\) there are two possible
-roots, however, we assume repeated games that are repeated for a large number
-of turn such as \(\delta \rightarrow 1\). The condition then becomes
-\(q < 1 - \frac{\sqrt{2}}{2}\).
+For \(\frac{b}{2}>c\) the ratio is smaller for \(q < \left\{\frac{\delta - 1 -
+\frac{\sqrt{2}}{2}}{\delta}, \frac{\delta - 1 +
+\frac{\sqrt{2}}{2}}{\delta}\right\}\), however, \(\frac{\delta - 1 -
+\frac{\sqrt{2}}{2}}{\delta}\) is not a feasible root since it's always smaller
+than 1, and thus \(q < \frac{\delta - 1 + \frac{\sqrt{2}}{2}}{\delta}\). For
+infinitely repeated games, \(\delta \rightarrow 1\), this condition becomes \(q
+< \frac{\sqrt{2}}{2}\). In the case of \(\frac{b}{2}<c\) there are two possible
+roots. For repeated games that are repeated for a large number of turn such as
+\(\delta \rightarrow 1\) the condition then becomes \(q < 1 -
+\frac{\sqrt{2}}{2}\).
 
-\subsection{Updating Payoffs based on the last two rounds payoff of one interactions (\(m=2\) and \(n=1\))}\label{section:m_two_n_one}
+\subsection{Updating Payoffs based on the Last Two Rounds Payoff of One Interaction (\(m=2\) and \(n=1\))}\label{section:m_two_n_one}
 
-We extend Proposition~\ref{proposition:last_round} to \(n\) rounds. 
-The result is given by Proposition~\ref{proposition:last_two_rounds}.
+In this section we consider the case of \(n=1\) and \(m=2\), thus individuals
+update based on their last two rounds payoffs with one member. To this end we
+define the probability that a reactive strategy receives the payoffs $u\!\in\!
+\mathcal{\tilde{U}}$, for \(\mathcal{\tilde{U}}= \{rrr\dots r, rrr\dots s, \dots,
+ppp\dots p\}\), in the last \(m\) rounds against another reactive strategy
+(Proposition~\ref{proposition:last_m_rounds}).
 
-\begin{Prop}\label{proposition:last_two_rounds} Consider a repeated game, with
+\begin{Prop}\label{proposition:last_m_rounds} Consider a repeated game, with
   continuation probability $\delta$, between players with reactive strategies
@@
 
-Here consider the case of \(n=2\). The framework remains the same
-as in section~\ref{section:m_one_n_one}, the only difference is the game stage
-payoffs.
+The probability that the number of mutants increases, and decreases respectively
+remains the same as in section~\ref{section:m_one_n_one} (Eq.~\ref{eq:ratio_limited}). The
+frameworks differ in the game stage payoffs,
 
-\subsubsection*{Simulation Results based on the last two rounds payoff of one interaction}
+\begin{align}
+  \lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{\tilde{u}_{R},\tilde{u}_{M}\in\mathcal{\tilde{U}}} \tilde{x}(\tilde{u}_{R},\tilde{u}_{M})\cdot \rho(\tilde{u}_{R},\tilde{u}_{M}), \\
+  \lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{\tilde{u}_{R},\tilde{u}_{M}\in\mathcal{\tilde{U}}} \tilde{x}(\tilde{u}_{R},\tilde{u}_{M})\cdot \rho(\tilde{u}_{M},\tilde{u}_{R}).
+\end{align}
+
+\subsubsection*{Simulation Results based on the Last Two Rounds Payoff of One Interaction}
+
+The simulation results for this case are given in
+Figure~\ref{fig:two_rounds_results}. Updating on the two rounds payoffs received
+against a single opponent result to a lower cooperation rate compare to updating
+based on one round against two opponents.
 
@@
   \caption{\textbf{Evolutionary dynamics under two interactions last round
-  payoffs with low (left) and high (right) benefit.} In the cases of low and
-  high benefit case the resident population either consists of defectors (with
-  \(p \approx q \approx 0\)) or of conditional cooperators for which \(p \approx
-  1\) and \(q \leq \frac{1}{2}\). Parameters: \(N =100, c=1, \beta=1\).
+  payoffs with low (left) and high (right) benefit.}\\Parameters: \(N =100, c=1, \beta=1, T=10 ^7\).
   }\label{fig:two_rounds_results}
@@
 
-\subsubsection*{Invasion Analysis based on the last two rounds payoff of one interaction}
+\subsubsection*{Invasion Analysis based on the Last Two Rounds Payoff of One Interaction}
+
+In a similar fashion we can calculate the condition for which a population of
+GTFT players can not be invaded by an ALLD mutant. The ratio of transition
+probabilities is given by,
 
@@
 
-In the limit of strong selection \(\beta \rightarrow \infty\) and
-large populations \(N \rightarrow \infty \), we obtain
+In the limit of strong selection \(\beta \rightarrow \infty\) and large
+populations \(N \rightarrow \infty \) we obtain three expressions depending on
+the cost-benefit relationship. We note that for \(\frac{b}{2}=c\) the result
+remains the same expression as in the case of \(m=n=1\).
 
@@
 \begin{cases}
-  \frac{q \left(\delta q - \delta + 1\right)}{\delta \left(1 - q\right)^{2} + \left(1 - q\right) \left(2 \delta q - \delta + 1\right)}  & \frac{b}{2} > c \\
-  - \frac{\delta q - \delta + q + 1}{\left(\delta + 1\right) \left(q - 1\right)}  & \frac{b}{2} = c \\
-  \frac{q \left(\delta q - \delta + 1\right) + \left(1 - q\right) \left(2 \delta q - \delta + 1\right)}{\delta \left(1 - q\right)^{2}} & \frac{b}{2} < c
+  - \frac{q \left(\delta q - \delta + 1\right)}{\left(q - 1\right) \left(\delta q + 1\right)}  & \frac{b}{2} > c \\[0.5cm]
+  - \frac{\delta q - \delta + q + 1}{\left(\delta + 1\right) \left(q - 1\right)}  & \frac{b}{2} = c \\[0.5cm]
+  - \frac{\delta q^{2} - 2 \delta q + \delta - 1}{\delta \left(q - 1\right)^{2}} & \frac{b}{2} < c
 \end{cases}
@@
 \begin{cases}
-  q \in \{\frac{\delta - \sqrt{\delta^{2} + 1} - 1}{2 \delta}, \  \frac{\delta + \sqrt{\delta^{2} + 1} - 1}{2 \delta}\}  & \frac{b}{2} > c \\
-  q = \frac{\delta}{\delta + 1}  & \frac{b}{2} = c \\
-  q \in \{1 - \frac{\sqrt{2}}{2 \sqrt{\delta}}, \  1 + \frac{\sqrt{2}}{2 \sqrt{\delta}}\} & \frac{b}{2} < c
+  q \in \left\{\frac{\delta - \sqrt{\delta^{2} + 1} - 1}{2 \delta}, \  \frac{\delta + \sqrt{\delta^{2} + 1} - 1}{2 \delta}\right\}  & \frac{b}{2} > c \\[0.5cm]
+  q \in \left\{1 - \frac{\sqrt{2}}{2 \sqrt{\delta}}, \  1 + \frac{\sqrt{2}}{2 \sqrt{\delta}}\right\} & \frac{b}{2} < c
 \end{cases}
@@
 
+For \(\frac{b}{2}>c\) the ratio is smaller for
+\(q < \left\{\frac{\delta - \sqrt{\delta^{2} + 1} - 1}{2 \delta}, \frac{\delta + \sqrt{\delta^{2} + 1} - 1}{2 \delta}\right\}\)
+, however, the first root is not a feasible root since it's always smaller
+than 1, and thus \(q < \frac{\delta + \sqrt{\delta^{2} + 1} - 1}{2 \delta}\). For
+infinitely repeated games, \(\delta \rightarrow 1\), this condition becomes \(q
+< \frac{\sqrt{2}}{2}\). In the case of \(\frac{b}{2}<c\) there are two possible
+roots. For repeated games that are repeated for a large number of turn such as
+\(\delta \rightarrow 1\) the condition then becomes \(q < 1 -
+\frac{\sqrt{2}}{2}\).
+
 \subsection{Updating Payoffs based on the last two rounds payoff of two interactions (\(m=2\) and \(n=2\))}
 
-The last case for which we present results in this work is the case of \(m=2\)
-and \(n=2\), that is, an individual considers the last two rounds payoffs he/she
-received against two different members of the population. The probability that
-the number of mutants increases, and decreases respectively, by one is now given
-by, are based on the approached on section~\ref{section:m_one_n_two}
-and~\ref{section:m_two_n_one}. In this case we only present only the results on
-the numerical simulations which are given in
-Figure~\ref{fig:two_rounds_opponents_results}.
-
-The evolved cooperation rates are between the other cases and the expected
-payoffs. For a low benefit we get results similar to the case where one the last
-round was considered, however, in the high benefit case the evolved cooperation
-rate increases to 91\% which is higher then any of the newly introduced
-approaches. Though we do no demonstrate or method of more cases, we assume, that
-as \(n\), and \(m\) increase the results tend to the expected payoffs.
+The last case we present in this work is the case of \(m = 2\) and \(n = 2\).
+That is, an individual considers the last two rounds of payoffs he or she
+received against two different members of the population. For this case we
+combine the approaches described in sections~\ref{section:m_one_n_two}
+and~\ref{section:m_two_n_one}. Here we only present results on the numerical
+simulations, Figure~\ref{fig:two_rounds_opponents_results}. The evolved
+cooperation rates are between the other cases and the expected payoffs. For a
+low benefit we get results similar to the case where one the last round was
+considered, however, in the high benefit case the evolved cooperation rate
+increases to 91\% which is higher then any of the newly introduced approaches.
 
@@
 
+\newpage
+
 \section{Expected and Last Round Updating Payoffs for Memory One Strategies}
 
-So far we have only considered the case where members can adopt reactive
-strategies during the game phase. In order to demonstrate that our results hold
-for higher memory strategies here we present numerical results in the case of
-memory-one strategies.
-
-Memory-one strategies consider the outcome of the previous round to decide on an
-action. There are four possible outcome in each round; \((C, C), (C, D), (D, C),
-(D, D)\). A memory-one strategy \(s\) can be written as a five-dimensional
-vector \(s=(y, p_1, p_2, p_3, p_4)\). The parameter \(y\) is the probability
-that the strategy opens with a cooperation and \(p_1\), \(p_2\), \(p_3\),
-\(p_4\) are the probabilities that the strategy cooperates for each of the
-possible outcomes of the last round.
+So far we have considered the case where members can adopt reactive strategies.
+To demonstrate that our results hold for higher memory strategies here we
+present results for the expected payoffs, and the last round payoff when members
+use memory-one strategies. Memory-one strategies consider the outcome of the
+previous round to decide on an action. There are four possible outcome in each
+round; \((C, C), (C, D), (D, C), (D, D)\). A memory-one strategy \(s\) can be
+written as a five-dimensional vector \(s=(y, p_1, p_2, p_3, p_4)\). The
+parameter \(y\) is the probability that the strategy opens with a cooperation
+and \(p_1\), \(p_2\), \(p_3\), \(p_4\) are the probabilities that the strategy
+cooperates for each of the possible outcomes of the last round.
 
@@
   namely, \(p_1 \approx 1\), \(p_3 \approx 0\) and \(p_4 > \frac{1}{2}\). There are also differences, in the latter
-  case a strategy is more likely to open with a cooperation and their tolerance to
+  case a strategy is more likely to open with cooperation and their tolerance to
   a \((C, D)\) outcome is almost zero. A difference between the strategies is their
@@
   most abundant strategy is a strategy with no tolerance to one defection, and
-  it cooperates with a probability .5 after a mutual defection.  In the expected
+  it cooperates with a probability 0.5 after a mutual defection.  In the expected
   payoffs case strategies are more robust.  Parameters: \(N =100, c=1, b=10,
@@
 
+\section{Expected and Last Round Updating Payoffs for High Mutation ($\mu \neq 0$)}\label{section:mutation}
+
+In this section we evaluate the main result of this work for \(\mu \neq 0\).
+Namely, we explore the evolved population when individuals use perfect and
+limited updating payoff memory for different values of \(\mu\). We perform five
+independent runs of the pairwise process described in
+Section~\ref{section:pairwise_comparison}, and at each time step we record the
+average player \(\bar{s}=(\bar{y}, \bar{p}, \bar{q})\).
+
+\begin{figure}[!htbp]
+  \centering 
+  \includegraphics[width=.5\textwidth]{static/mutation_perfect_and_limited_memory_donation_game.pdf}
+  \caption{\textbf{Evolutionary dynamics results for perfect and limited memory
+  for different mutation values.}
+  We perform five independent simulations. Simulations are run
+  for $T\!=\!4\times 10^7$ time steps for each parameter. In each time step
+  we introduce a new mutant with a probability \(\mu\), and we then select
+  two random players to serve as the role model and the learner. The learner
+  adopts the strategy of the role model with a probability \(\rho(\pi_{L}, \pi_{RM})\) where the
+  updating payoffs depend on the method. In the case of perfect memory
+  the expected payoffs are used and in the case of limited memory the
+  last round payoff against one opponent. We plot the average cooperation rate
+  within the resident population for each value of \(\mu\). For \(\mu=1\)
+  the process becomes random and so the cooperation rates are 0.5. For the rest
+  of the mutation values the perfect memory payoffs once again overestimate the
+  evolved cooperation, confirming the results of low mutation. Parameters: \(N
+  =100, c=1, b=10, \beta=1\).}\label{fig:mutation}
+\end{figure}
+
+\bibliographystyle{unsrt}
+\bibliography{bibliography.bib}
+
 \end{document}

### 2023-02-23 paper/main.tex
removes main from root

@@
 
+\subsection{Updating Payoffs based on More Memory}
+
+So far we have explored the difference between the expected payoffs and the last
+round payoffs. To explore further the effect of limited memory we consider three
+more cases. These are the cases where the updating payoffs depend on  the last
+round with two members of the population, the last two rounds with another
+member of the population, and the last two rounds with two members of the
+population. Similar to the last round payoff. The details of each approach can
+be found in the Supplementary Information. We use numerical simulation of the
+pairwise process for the various schemes. The results for low and high benefit
+are shown in Figure~\ref{fig:cooperation_rate_all_updating_payoffs}.
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=\textwidth]{static/more_memory_summary_results.pdf}
+  \caption{{\bf Average cooperation rates for different updating payoffs.}
+  From left to right, we present result on the following updating payoffs cases;
+  (1) the expected payoffs (perfect memory), (2) the last round payoff from one
+  interaction (limited memory), (3) the last round payoff from two interactions,
+  (4) the last two rounds payoffs from one interaction, (5) the last two rounds
+  payoffs from two interactions. The details of each case are given in the
+  Supplementary Information. Cases 3 and 4 have similar results. In that the
+  evolved generosity is the same. For cases 2 - 5 the cost benefit ratio
+  has an effect compared to the limited memory case. The cooperation rate
+  increases, and as we allow for more information the closer we move to the
+  perfect memory. Unless explicitly varied, the parameters of the simulation are
+  $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$.
+  Simulations are run for $T\!=\!5\times 10^7$ time steps for each parameter
+  combination.}\label{fig:cooperation_rate_all_updating_payoffs}
+\end{figure}
+
+\newpage
+% In the
+% simplest of cases each individual can be of one of two types, A and B. The state
+% of the population is thus characterized by the number \(k\) of individuals of
+% type A. The interaction between the two types of individuals is described by the
+% functions \(\pi _{{\rm A}}^{i}\) and $\pi _{{\rm B}}^{i}$.
+
+% airwise comparison process is a stochastic process for modelling the evolution
+% of a finite population. The process starts with assigning all individuals of the
+% population the same strategy. A strategy is a set of rules of how an individual
+% should behave in an interaction with another individual. Each elementary time
+% step of the process consists of three phases; (1) the \textbf{mutation phase}
+% (2) the \textbf{game phase} and (3) the \textbf{update phase}. These are summarised
+% in Figure~\ref{fig:pairwise_phases}.
+
+% In the \textbf{mutation phase} one individual is
+% chosen to switch to a new mutant strategy with a probability \(\mu\).
+% In the \textbf{game phase} individuals are randomly matched with other
+% individuals in the population, and they engage in a repeated game where each
+% subsequent turn occurs with a fixed probability \(\delta\). At each turn the
+% individuals decide on an action based on their strategies.
+% In repeated games there are infinite many strategies, however, it is commonly
+% assumed that individuals can only choose strategies from a restricted set. One
+% such set is that of reactive strategies. A reactive
+% strategy considers only the previous action of the other player, and thus, a
+% reactive strategy \(s\) can be written as a three-dimensional vector \(s=(y, p,
+% q)\). The parameter \(y\) is the probability that the strategy opens with a
+% cooperation and \(p\), \(q\) are the probabilities that the strategy cooperates
+% given that the opponent cooperated and defected equivalently.
+
+% In the \textbf{update stage} where two individuals are randomly selected. From
+% the two individuals, one  serves as the `learner' and the other as the `role
+% model'. The learner adopts the role model's strategy with a probability \(\rho\)
+% given by,
+
+% The fitness of the individuals (\(\pi _{{\rm A}}^{i}, \pi _{{\rm B}}^{i}\)) is
+% identified with the payoff resulting from the underline game. The evolution of
+% cooperation remains one of the greatest problems for biological and social
+% sciences. Cooperation is the action of choosing to help others at one's own
+% expense. The standard game of formulating such situation is the Prisoner's
+% Dilemma~\cite{trivers1971evolution, milinski1987tit, glynatsi2021bibliometric,
+% ohtsuki2006simple}, in which two players can choose to cooperate or to defect.
+% The players are offered a certain payoff, \(R\), for mutual cooperation and a
+% lower payoff, \(P\), for mutual defection. If one player cooperates while the
+% other defects, then the cooperator gets the lowest payoff, \(S\), while the
+% defector gains the highest payoff, \(T\). Thus, the payoffs have the following
+% property \(T > R > P > S\), making defection the dominant strategy in the
+% non-repeated game.
+
+% \begin{equation}\label{eq:payoffs}
+%   \begin{blockarray}{ccc}
+%       & \text{cooperate} & \text{defect} \\
+%       \begin{block}{c(cc)}
+%           \text{cooperate} & R & S \\
+%           \text{defect} & T & P \\
+%       \end{block}
+%   \end{blockarray}
+% \end{equation}
+
+% In literature, the payoffs of A and B individuals only on the fraction of both
+% types in the population. If there are \(k\) A individuals and \(N-k\) B
+% individuals. In the context of the prisoner's dilemma the payoffs and assuming
+% that A and B follow simple strategies where they cooperate and defect
+% unconditionally respective, then the A and B individuals have fitness of
+
+% \begin{equation}
+%   \begin{array}{lcrcr}
+%   \displaystyle \pi _{{\rm A}}^{i}	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot R	&+	&\displaystyle\frac{k}{N-1}\cdot S,\\[0.5cm]
+%   \displaystyle \pi _{{\rm B}}^{i}&=	&\displaystyle\frac{N-k}{N-1}\cdot T &+	&\displaystyle\frac{k-1}{N-1}\cdot P.\\
+%   \end{array}
+% \end{equation}
+
+% respectively while self-interactions are excluded. In literature these are the
+% expected payoffs. For \(\beta = 1\), fitness equals payoff. This scenario
+% describes ``strong selection''. For \(\beta \ll 1\) , the payoff only provides a
+% small perturbation to the overall fitness of an individual, a limit known as
+% weak selection.
+
+% In the evolutionary game theory literature it is usually assumed that players
+% use strategies with finite memory~\cite{Nowak1992tit, Baek2016}. This assumption
+% is common as it allows for an explicit calculation of the players'
+% payoffs~\cite{sigmund2010calculus}. Players' payoffs are calculated in the limit
+% of interactions assuming that the two strategies meet several times.
+
+% Customarily, most evolutionary game studies make the additional assumption that
+% individuals play many times and with all other players before reproduction takes
+% place, so that payoffs, equivalently fitness, are given by the mean distribution
+% of types in the population. This hypothesis, selection occurs much more slowly
+% than the interaction between individuals and more that remember all the
+% interactions they participated in. Although recent experimental studies show
+% that this may not always be the case in biology [12– 14], it is clear that in
+% cultural evolution or social learning the time scale of selection is much closer
+% to the time scale of interaction. The effects of this mixing of scales cannot be
+% disregarded [15], and then it is natural to ask about the consequences of the
+% above assumption and the effect of relaxing it. The work of~\cite{Roca2006}.
+
+% In this work we ask studies that make use of this classical framework report
+% that cooperation can substantially evolve. However, the inconsistency in the
+% memory size of players leads us to question the robustness of our understanding
+% of cooperation. To this end, we propose a framework in which individuals,
+% similar to the decisions at each turn, estimate their fitness based on a minimum
+% of information.
+
+% We first consider two extreme scenarios, the classical scenario and the
+% alternative scenario where individuals update their strategies only based on the
+% very last payoff they obtained. We observe that individuals with limited memory
+% tend to adopt less generous strategies and they achieve less cooperation when
+% interacting in a prisoner's dilemma. We obtain similar results when we consider
+% that individuals update their strategies based on more information. More
+% specifically, up to the last two payoffs they obtained when interacting with up
+% to two different members of the population. We extend our approach to the rest
+% of the symmetric \(2 \times 2\) games.
+
+% The remainder of the paper is organized as follows. In
+% section~\ref{section:model} we describe the model. In
+% section~\ref{section:results} we present the results of the simulations, and in
+% section~\ref{section:conclusions} we outline the main conclusions.
+
+% In the following, we consider a well mixed population of fixed size. In each
+% step of t
+
+% We study the transmission of strategies with a frequency-dependent birth-death
+% process26 in a finite population of size n. In each time step, two randomly
+% chosen individuals compare their payoffs and one of them can switch to the other
+% one's strategy. This process can be interpreted as a model for social learning,
+% whereby successful strategies spread, and, occasionally, random strategy
+% exploration introduces novel strategies (corresponding to mutations in
+% biological models).
+
+% We consider a population of \(N\) players~\footnote{The terms ``player'' and
+% ``individual'' are used interchangeably here.} where \(N\) is even and mutations
+% are sufficiently rare. Therefore, at any point in time there are at most two
+% different strategies present in the population; a \textit{resident} strategy and
+% a \textit{mutant} strategy. To describe how strategies spread we use a pairwise
+% comparison process~\cite{Traulsen2006}. Each step of the evolutionary process
+% consists of two stages, a game stage and an updating stage.
+
+% In the game stage each individual is randomly matched with some other individual
+% in the population. They engage in a match where each subsequent turn occurs with
+% a fixed probability $\delta$. At each turn players choose independently to
+% either cooperate (\(C\)) or to defect (\(D\)), and the payoffs of the turn
+% depend on both their decisions. If both players cooperate they receive the
+% reward payoff \(R\), whereas if both defect they receive the punishment payoff
+% \(P\). If one cooperates but the other defects, the defector receives the
+% temptation payoff \(T\), whereas the cooperator receives the sucker's payoff
+% \(S\). We denote the feasible payoff of each turn as \(\mathcal{U} = \{R, S, T,
+% P\}\). We assume that individuals use \textit{reactive strategies} to make
+% decisions in each turn. Reactive strategies are a set of memory-one strategies
+% that only take into account the previous action of the opponent. They can be
+% written explicitly as a vector in \(\R_{3}\), more specifically, a reactive
+% strategy \(s\) is given by \(s=(y, p, q)\). The parameter \(y\) is the
+% probability that the strategy opens with a cooperation and \(p, q\) are the
+% probabilities that the strategy cooperates given that the opponent cooperated
+% and defected equivalently.
+
+% In the updating stage, two players are randomly drawn from the population, a
+% `learner' and a `role model'. Given the learner's payoff $u_L\!\in\!
+% \mathcal{U}$ and the role model's payoff $u_{RL}\!\in\! \mathcal{U}$, the
+% learner adopts the role model's strategy with probability,
+
+% % \begin{equation} \label{Eq:rho}
+% %   \rho(u_{L}, u_{RM}) = \frac{1}{1\!+\! \exp^{\!-\!\beta (u_{RM}\!-\!u_{L})}}.
+% % \end{equation}
+
+% where $\beta\!\ge\!0$ is the strength of selection. For small values of $\beta$
+% the imitation probability is independent of the strategies of the involved
+% players. As the value of $\beta$ increases, the more likely it is that the
+% learner adopts only strategies that yield a higher payoff. Conventionally the
+% updating payoffs of the learner and the role model are based on their expected
+% payoffs. A player's expected payoff is the mean payoff the player yields after
+% engaging in matches of multiple turns with each member of the population. At
+% each match a player bases their next turn decision only on the previous action
+% of the opponent, however, the same player bases their expected payoffs on the
+% outcomes of all their matches. Thus, a player is assumed to have limited and
+% perfect memory at the same time. We propose a new a set of updating payoffs
+% where it is also assumed that a player has limited memory. We referee to these
+% as the limited memory payoffs.
+
+% The evolutionary step is repeated until either the mutant strategy goes extinct,
+% or until it fixes in the population. If the mutant fixes in the population then
+% the mutant strategy becomes the new resident strategy. After either outcome we
+% introduce a new mutant strategy uniformly chosen from all reactive strategies at
+% random, and we set the number of mutants to $1$. This process of mutation and
+% fixation/extinction is then iterated many times.
+
+% \section{Results}\label{section:results} \subsection{Updating payoffs based on
+% the last round with another member of the population}\label{section:donation}
+
+% To investigate the role of extortion in the context of evolutionary games, we
+% concentrate on the donation game.
+% In this section we explore the case where the updating payoffs are based on the
+% last round payoff achieved against another member of the population and we
+% compare this to the classical scenario of the expected payoffs. We assume that
+% each pair of players interacts in a donation game. The donation game is a
+% special case of the prisoner's dilemma. Each player can choose to cooperate by
+% providing a benefit \(b\) to the other player at their cost \(c\), with \(0 < c
+% < b\). Thus, the feasible payoffs in each round are \(\mathcal{U} = \{b-c, -c,
+% b, 0\}\).
+
+% \begin{figure}[!htbp]
+%   \centering
+%   \includegraphics[width=\textwidth]{static/cooperation_rates_expected_and_stochastic_for_donation_game.pdf}
+% \end{figure}
+
+% We further explore the effect of benefit in
+% Figure~\ref{fig:cooperation_rate_over_benefit}. The figure suggests that
+% expected payoffs always yield a higher cooperation rate. In the case of expected
+% payoffs we observe that the cooperation rate increases as the value of the
+% benefit gets higher. In comparison for the limited memory payoffs, the
+% cooperation rate remains unchanged at approximately 50\% once \(b=5\).
+
 % % \begin{figure}[!htbp]
@@
 
-% % \begin{figure}[!htbp]
-% %   \centering
-% %     \includegraphics[width=\textwidth]{static/cooperation_rate_over_b.pdf}
-% % \caption{{\bf The evolution of cooperation for different selection strength values.}
-% % We vary the selection strength $\beta$. In all cases, stochastic payoff
-% % evaluation tends to reduce the evolving cooperation rates. ({\bf A}) the
-% % probabilities \(p, q\) for resident population over \(10^7\) time steps for each
-% % \(\beta\) value. ({\bf B}) The cooperation rate within the resident population
-% % (averaged over all games and over all time steps) over \(\beta\). Unless
-% % explicitly varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$,
-% % $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
-% % $t\!=\!5\times 10^7$ time steps for each parameter
-% % combination.}\label{fig:cooperation_rate_over_betas}
-% % \end{figure}
-
-\subsection{Updating Payoffs based on More Memory}
-
-So far we have explored the difference between the expected payoffs and the last
-round payoffs. In order to explore further the effect of limited memory we allow
-individuals to remember more. More precisely, up to two interactions and up to
-the last two rounds. In total we present results for three more updating
-payoffs. These are the payoffs when individuals consider the last two rounds
-with another member of the population, the last round with two members of the
-population, and the last two rounds with two members of the population. Similar
-to the last round payoff, we use simulations and record which strategies the
-players adopt over time based
-% (Appendix~\ref{appendix:further_simulation_results}) and we compare the evolving
-% cooperation rates to those of the expected payoffs
-% (Figure~\ref{fig:cooperation_rates_two_by_two_part_two}).
-
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[width=\textwidth]{static/more_memory_summary_results.pdf}
-  \caption{{\bf Average cooperation rates for different updating payoffs.} (From
-  right to left) updating Payoffs based  on perfect memory, limited memory,
-  the last round payoff of two interactions,
-  the last two rounds payoff of one interaction and 
-  the last two round payoffs of two interactions.({\bf High Benefit})
-  ({\bf Low Benefit}) We vary the selection strength $\beta$. In all
-  cases, stochastic payoff evaluation tends to reduce the evolving cooperation
-  rates. Unless explicitly varied, the parameters of the simulation are
-  $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$.
-  Simulations are run for $T\!=\!5\times 10^7$ time steps for each parameter
-  combination.}\label{fig:cooperation_rate_updating_payoffs}
-\end{figure}
-
-\newpage
-% In the
-% simplest of cases each individual can be of one of two types, A and B. The state
-% of the population is thus characterized by the number \(k\) of individuals of
-% type A. The interaction between the two types of individuals is described by the
-% functions \(\pi _{{\rm A}}^{i}\) and $\pi _{{\rm B}}^{i}$.
-
-% airwise comparison process is a stochastic process for modelling the evolution
-% of a finite population. The process starts with assigning all individuals of the
-% population the same strategy. A strategy is a set of rules of how an individual
-% should behave in an interaction with another individual. Each elementary time
-% step of the process consists of three phases; (1) the \textbf{mutation phase}
-% (2) the \textbf{game phase} and (3) the \textbf{update phase}. These are summarised
-% in Figure~\ref{fig:pairwise_phases}.
-
-% In the \textbf{mutation phase} one individual is
-% chosen to switch to a new mutant strategy with a probability \(\mu\).
-% In the \textbf{game phase} individuals are randomly matched with other
-% individuals in the population, and they engage in a repeated game where each
-% subsequent turn occurs with a fixed probability \(\delta\). At each turn the
-% individuals decide on an action based on their strategies.
-% In repeated games there are infinite many strategies, however, it is commonly
-% assumed that individuals can only choose strategies from a restricted set. One
-% such set is that of reactive strategies. A reactive
-% strategy considers only the previous action of the other player, and thus, a
-% reactive strategy \(s\) can be written as a three-dimensional vector \(s=(y, p,
-% q)\). The parameter \(y\) is the probability that the strategy opens with a
-% cooperation and \(p\), \(q\) are the probabilities that the strategy cooperates
-% given that the opponent cooperated and defected equivalently.
-
-% In the \textbf{update stage} where two individuals are randomly selected. From
-% the two individuals, one  serves as the `learner' and the other as the `role
-% model'. The learner adopts the role model's strategy with a probability \(\rho\)
-% given by,
-
-% The fitness of the individuals (\(\pi _{{\rm A}}^{i}, \pi _{{\rm B}}^{i}\)) is
-% identified with the payoff resulting from the underline game. The evolution of
-% cooperation remains one of the greatest problems for biological and social
-% sciences. Cooperation is the action of choosing to help others at one's own
-% expense. The standard game of formulating such situation is the Prisoner's
-% Dilemma~\cite{trivers1971evolution, milinski1987tit, glynatsi2021bibliometric,
-% ohtsuki2006simple}, in which two players can choose to cooperate or to defect.
-% The players are offered a certain payoff, \(R\), for mutual cooperation and a
-% lower payoff, \(P\), for mutual defection. If one player cooperates while the
-% other defects, then the cooperator gets the lowest payoff, \(S\), while the
-% defector gains the highest payoff, \(T\). Thus, the payoffs have the following
-% property \(T > R > P > S\), making defection the dominant strategy in the
-% non-repeated game.
-
-% \begin{equation}\label{eq:payoffs}
-%   \begin{blockarray}{ccc}
-%       & \text{cooperate} & \text{defect} \\
-%       \begin{block}{c(cc)}
-%           \text{cooperate} & R & S \\
-%           \text{defect} & T & P \\
-%       \end{block}
-%   \end{blockarray}
-% \end{equation}
-
-% In literature, the payoffs of A and B individuals only on the fraction of both
-% types in the population. If there are \(k\) A individuals and \(N-k\) B
-% individuals. In the context of the prisoner's dilemma the payoffs and assuming
-% that A and B follow simple strategies where they cooperate and defect
-% unconditionally respective, then the A and B individuals have fitness of
-
-% \begin{equation}
-%   \begin{array}{lcrcr}
-%   \displaystyle \pi _{{\rm A}}^{i}	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot R	&+	&\displaystyle\frac{k}{N-1}\cdot S,\\[0.5cm]
-%   \displaystyle \pi _{{\rm B}}^{i}&=	&\displaystyle\frac{N-k}{N-1}\cdot T &+	&\displaystyle\frac{k-1}{N-1}\cdot P.\\
-%   \end{array}
-% \end{equation}
-
-% respectively while self-interactions are excluded. In literature these are the
-% expected payoffs. For \(\beta = 1\), fitness equals payoff. This scenario
-% describes ``strong selection''. For \(\beta \ll 1\) , the payoff only provides a
-% small perturbation to the overall fitness of an individual, a limit known as
-% weak selection.
-
-% In the evolutionary game theory literature it is usually assumed that players
-% use strategies with finite memory~\cite{Nowak1992tit, Baek2016}. This assumption
-% is common as it allows for an explicit calculation of the players'
-% payoffs~\cite{sigmund2010calculus}. Players' payoffs are calculated in the limit
-% of interactions assuming that the two strategies meet several times.
-
-% Customarily, most evolutionary game studies make the additional assumption that
-% individuals play many times and with all other players before reproduction takes
-% place, so that payoffs, equivalently fitness, are given by the mean distribution
-% of types in the population. This hypothesis, selection occurs much more slowly
-% than the interaction between individuals and more that remember all the
-% interactions they participated in. Although recent experimental studies show
-% that this may not always be the case in biology [12– 14], it is clear that in
-% cultural evolution or social learning the time scale of selection is much closer
-% to the time scale of interaction. The effects of this mixing of scales cannot be
-% disregarded [15], and then it is natural to ask about the consequences of the
-% above assumption and the effect of relaxing it. The work of~\cite{Roca2006}.
-
-% In this work we ask studies that make use of this classical framework report
-% that cooperation can substantially evolve. However, the inconsistency in the
-% memory size of players leads us to question the robustness of our understanding
-% of cooperation. To this end, we propose a framework in which individuals,
-% similar to the decisions at each turn, estimate their fitness based on a minimum
-% of information.
-
-% We first consider two extreme scenarios, the classical scenario and the
-% alternative scenario where individuals update their strategies only based on the
-% very last payoff they obtained. We observe that individuals with limited memory
-% tend to adopt less generous strategies and they achieve less cooperation when
-% interacting in a prisoner's dilemma. We obtain similar results when we consider
-% that individuals update their strategies based on more information. More
-% specifically, up to the last two payoffs they obtained when interacting with up
-% to two different members of the population. We extend our approach to the rest
-% of the symmetric \(2 \times 2\) games.
-
-% The remainder of the paper is organized as follows. In
-% section~\ref{section:model} we describe the model. In
-% section~\ref{section:results} we present the results of the simulations, and in
-% section~\ref{section:conclusions} we outline the main conclusions.
-
-% In the following, we consider a well mixed population of fixed size. In each
-% step of t
-
-% We study the transmission of strategies with a frequency-dependent birth-death
-% process26 in a finite population of size n. In each time step, two randomly
-% chosen individuals compare their payoffs and one of them can switch to the other
-% one's strategy. This process can be interpreted as a model for social learning,
-% whereby successful strategies spread, and, occasionally, random strategy
-% exploration introduces novel strategies (corresponding to mutations in
-% biological models).
-
-% We consider a population of \(N\) players~\footnote{The terms ``player'' and
-% ``individual'' are used interchangeably here.} where \(N\) is even and mutations
-% are sufficiently rare. Therefore, at any point in time there are at most two
-% different strategies present in the population; a \textit{resident} strategy and
-% a \textit{mutant} strategy. To describe how strategies spread we use a pairwise
-% comparison process~\cite{Traulsen2006}. Each step of the evolutionary process
-% consists of two stages, a game stage and an updating stage.
-
-% In the game stage each individual is randomly matched with some other individual
-% in the population. They engage in a match where each subsequent turn occurs with
-% a fixed probability $\delta$. At each turn players choose independently to
-% either cooperate (\(C\)) or to defect (\(D\)), and the payoffs of the turn
-% depend on both their decisions. If both players cooperate they receive the
-% reward payoff \(R\), whereas if both defect they receive the punishment payoff
-% \(P\). If one cooperates but the other defects, the defector receives the
-% temptation payoff \(T\), whereas the cooperator receives the sucker's payoff
-% \(S\). We denote the feasible payoff of each turn as \(\mathcal{U} = \{R, S, T,
-% P\}\). We assume that individuals use \textit{reactive strategies} to make
-% decisions in each turn. Reactive strategies are a set of memory-one strategies
-% that only take into account the previous action of the opponent. They can be
-% written explicitly as a vector in \(\R_{3}\), more specifically, a reactive
-% strategy \(s\) is given by \(s=(y, p, q)\). The parameter \(y\) is the
-% probability that the strategy opens with a cooperation and \(p, q\) are the
-% probabilities that the strategy cooperates given that the opponent cooperated
-% and defected equivalently.
-
-% In the updating stage, two players are randomly drawn from the population, a
-% `learner' and a `role model'. Given the learner's payoff $u_L\!\in\!
-% \mathcal{U}$ and the role model's payoff $u_{RL}\!\in\! \mathcal{U}$, the
-% learner adopts the role model's strategy with probability,
-
-% % \begin{equation} \label{Eq:rho}
-% %   \rho(u_{L}, u_{RM}) = \frac{1}{1\!+\! \exp^{\!-\!\beta (u_{RM}\!-\!u_{L})}}.
-% % \end{equation}
-
-% where $\beta\!\ge\!0$ is the strength of selection. For small values of $\beta$
-% the imitation probability is independent of the strategies of the involved
-% players. As the value of $\beta$ increases, the more likely it is that the
-% learner adopts only strategies that yield a higher payoff. Conventionally the
-% updating payoffs of the learner and the role model are based on their expected
-% payoffs. A player's expected payoff is the mean payoff the player yields after
-% engaging in matches of multiple turns with each member of the population. At
-% each match a player bases their next turn decision only on the previous action
-% of the opponent, however, the same player bases their expected payoffs on the
-% outcomes of all their matches. Thus, a player is assumed to have limited and
-% perfect memory at the same time. We propose a new a set of updating payoffs
-% where it is also assumed that a player has limited memory. We referee to these
-% as the limited memory payoffs.
-
-% The evolutionary step is repeated until either the mutant strategy goes extinct,
-% or until it fixes in the population. If the mutant fixes in the population then
-% the mutant strategy becomes the new resident strategy. After either outcome we
-% introduce a new mutant strategy uniformly chosen from all reactive strategies at
-% random, and we set the number of mutants to $1$. This process of mutation and
-% fixation/extinction is then iterated many times.
-
-% \section{Results}\label{section:results} \subsection{Updating payoffs based on
-% the last round with another member of the population}\label{section:donation}
-
-% To investigate the role of extortion in the context of evolutionary games, we
-% concentrate on the donation game.
-% In this section we explore the case where the updating payoffs are based on the
-% last round payoff achieved against another member of the population and we
-% compare this to the classical scenario of the expected payoffs. We assume that
-% each pair of players interacts in a donation game. The donation game is a
-% special case of the prisoner's dilemma. Each player can choose to cooperate by
-% providing a benefit \(b\) to the other player at their cost \(c\), with \(0 < c
-% < b\). Thus, the feasible payoffs in each round are \(\mathcal{U} = \{b-c, -c,
-% b, 0\}\).
-
-% \begin{figure}[!htbp]
-%   \centering
-%   \includegraphics[width=\textwidth]{static/cooperation_rates_expected_and_stochastic_for_donation_game.pdf}
-% \end{figure}
-
-% We further explore the effect of benefit in
-% Figure~\ref{fig:cooperation_rate_over_benefit}. The figure suggests that
-% expected payoffs always yield a higher cooperation rate. In the case of expected
-% payoffs we observe that the cooperation rate increases as the value of the
-% benefit gets higher. In comparison for the limited memory payoffs, the
-% cooperation rate remains unchanged at approximately 50\% once \(b=5\).
-
-% % \begin{figure}[!htbp]
-% %   \centering
-% %   \begin{subfigure}{.5\textwidth}
-% %     \centering
-% %     \includegraphics[width=\textwidth]{static/expected_for_beta.pdf}
-% %     \includegraphics[width=\textwidth]{static/stochastic_for_beta.pdf}
-% %   \end{subfigure}%
-% %   \begin{subfigure}{.5\textwidth}
-% %     \centering
-% %     \includegraphics[width=\textwidth]{static/cooperation_rate_over_b.pdf}
-% %   \end{subfigure}
-% %   \caption{{\bf The evolution of cooperation for different benefit values.} 
-% %   We vary the benefit of defection $b$. In all cases, expected payoffs appear to
-% %   overestimate the average cooperation rate the population achieves. ({\bf A})
-% %   the probabilities \(p, q\) for resident population over \(10^7\) time steps
-% %   for each benefit value. ({\bf B}) The cooperation rate within the resident
-% %   population (averaged over all games and over all time steps) over the benefit.
-% %   Unless explicitly varied, the parameters of the simulation are $N\!=\!100$,
-% %   $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
-% %   $t\!=\!5\times 10^7$ time steps for each parameter
-% %   combination.}\label{fig:cooperation_rate_over_benefit}
-% % \end{figure}
-
 % We also investigate the effect of the strength of selection $\beta$.

### 2023-02-27 paper/si.tex
draft s.i.

@@
 comparison process. The pairwise comparison process consists of three phases;
-(1) the mutation phase (2) the game phase and (3) the update phase. In the update
-phase an individual adopts the strategy of another individual based on their
-``updating payoffs''. In Section~\ref{section:updating_payoffs} we
-describe the conventional approach for calculating updating payoffs and we
-present our new approaches.
+(1) the mutation phase (2) the game phase and (3) the update phase. In the
+update phase an individual adopts the strategy of another individual based on
+their ``updating payoffs''. In the paper we explore the effect of the updating
+payoffs on the evolved populations. We used as an example the prisoner's
+dilemma. We focused mainly on two approaches for calculating the updating
+payoffs which we referred to as the perfect memory and the limited memory
+approaches. In Section~\ref{section:perfect_memory} we describe the perfect
+memory approach and in Section~\ref{section:limited_memory} we present the
+limited memory approach. In the limited memory approach we assume that
+individuals update based on their last round payoff against one other member of
+the population. The framework can easily be extended to consider more rounds,
+more interactions or both. In
+Sections~\ref{section:m_one_n_two}-\ref{section:m_two_n_two} we present each of
+this extensions and for each a special case.
+In Section~\ref{section:simulation_results} we present numerical results of the
+pairwise comparison process using the several updating payoff approaches we have
+presented. In Sections~\ref{section:memory_one} and~\ref{section:mutation} we
+verify the main result of the paper when we no longer assume (i) that
+individuals use reactive strategies but instead memory one strategies (ii) low
+mutation.
+
+% The aim of our work is to explore the effect
+% of the updating payoffs on the evolution of cooperation. To this end we
+% introduce a new method for calculating the updating payoffs and we compare the
+% results of the new method to those of the expected payoffs. In our approach we
+% do not assume that an individual has a perfect memory. Instead an individual
+% remembers only a limited number of interactions and limited outcomes from each
+% interaction.
 
@@
     \includestandalone[width=\textwidth]{static/pairwise_comparison_process_diagram}
-    \caption{\textbf{Pairwise comparison process phases.} \textbf{A) Initialisation} The process begins
+    \caption{\textbf{Pairwise comparison process phases.} \textbf{A) Initialisation.} The process begins
     with a finite population where each member is assigned a given strategy.
@@
 
-\section{Updating Payoffs}\label{section:updating_payoffs}
+\subsection{Updating Payoffs}\label{section:updating_payoffs}
 
@@
 
-In the following subsections we present several approaches for calculating
-the updating payoffs. Initially, we discuss the conventional approach of the
-expected payoffs and afterwards we present our new approach.
+In the following sections we present several approaches for calculating the
+updating payoffs. Initially, we discuss the conventional approach of the
+expected payoffs and afterwards we present our newly introduced approaches.
 
-\subsection{Updating Payoffs based on the expected payoffs}\label{section:perfect_memory}
+\section{Updating Payoffs based on the expected payoffs (perfect memory)}\label{section:perfect_memory}
 
 The expected payoffs are the conventional payoffs used in the updating stage. We
-refer to this approach as the perfect memory approach. They are defined as the
-mean payoff of an individual in a well-mixed population that engages in an
-infinitely repeated games with all other population members. In an infinitely
-repeated game the payoff of a reactive strategy can explicitly be calculated
-using a markovian approach~\cite{nowak:APC:1989}. Namely, assume two reactive
-strategies $s_1\!=\!(y_1, p_1, q_1$) and $s_2\!=\!(y_2,p_2,q_2)$, their play can
-be defined as a Markov process with the transition matrix \(M\),
+refer to this approach as the perfect memory approach. The expected payoffs are
+defined as the mean payoff of an individual in a well-mixed population that
+engages in an infinitely repeated games with all other population members.
+The game between two reactive strategies $s_1\!=\!(y_1, p_1, q_1$) and $s_2\!=\!(y_2,p_2,q_2)$
+can be described by a Markovian process~\cite{nowak:APC:1989} with the
+transition matrix \(M\),
 
@@
 
-The stationary vector \(\mathbf{v}(s_1,s_2)\),
-which is the solution to \(\mathbf{v}(s_1,s_2) M = \mathbf{v}(s_1,s_2)\), combined with the payoff vector
-\(\mathbf{u}\), yields the game stage outcome for each strategy,
+The stationary vector \(\mathbf{v}(s_1,s_2)\) is the solution to
+\(\mathbf{v}(s_1,s_2) M = \mathbf{v}(s_1,s_2)\). In an infinitely repeated game
+the game stage outcome for each strategy is given by,
 
@@
 
-\subsubsection*{Simulation Results based on Expected Payoffs}
+\subsubsection*{Invasion Analysis}
 
-We simulate the evolutionary process, as described in
-Algorithm~\ref{algorithm:pairwise_comparison}, when individuals use reactive
-strategies and update their strategies based on their expected payoffs. We
-performed two independent runs of the process where we differed the benefit of
-cooperation \(b\). The results are shown in
-Figure~\ref{fig:expected_payoffs_results}. We observed that a higher value of
-benefit results in a more cooperative population. For a low benefit the resident
-population cooperates on average 52\% of the time. For a high benefit the
-average cooperation increases to 98\%. For a low benefit the resident population
-consists either of defectors or conditional cooperators, and for a high value of
-benefit the population predominantly consists of conditional cooperators.A
-conditional cooperator, otherwise known as a generous tit for tat (GTFT)
-player, always cooperates if the co-player cooperated ($p\approx1$) and
-cooperates with a probability \(q\) if the co-player defected. The conditional
-cooperator strategy adopted by the population differs between the two
-simulations, namely, a higher value of benefit results in higher values
-of \(q\).
-
-\begin{figure}[!htbp]
-    \centering 
-    \includegraphics[width=.70\textwidth]{static/expected_payoffs_donation_game.pdf}
-    \caption{\textbf{Evolutionary dynamics under expected payoffs with
-    low (left) and high (right) benefit.} We run the simulations for \(T =
-    10^7\) time steps. For each time step we recorded the current resident
-    population \((y, p, q)\). Since \(\delta \rightarrow
-    1\) we do not report the players' initial cooperation probability \(y\).
-    The graphs show how often the resident population chooses each combination
-    \((p, q)\) of conditional cooperation probabilities in the subsequent
-    rounds. In both cases players update based on their expected payoffs.
-    ({\bf Low Benefit}) In
-    the case where \(b=3\), the resident population either consists of defectors
-    (with \(p \approx q \approx 0\)) or of conditional cooperators for which \(p
-    \approx 1\) and \(q \leq 1 - \frac{1}{3}=0.7\). ({\bf High Benefit}) In the case where
-    \(b=10\), the resident population typically applies a strategy for which \(p
-    \approx 1\) and \(q \leq 1 - \frac{1}{10}=0.9\). Parameters: \(N =100, c=1,
-    \beta=1\).
-    }\label{fig:expected_payoffs_results}
-\end{figure}
-
-\subsubsection*{Invasion Analysis based on Expected Payoffs}
-
-We can explicitly calculate the value of \(q\) which evolves. Namely, we will
-calculate how easily a single defecting mutant can invade into a resident
-population of GTFT players. Let GTFT\(= (1, 1, q)\), ALLD\(= (0, 0, 0)\), and
-\(k = 1\). When two GTFT players interact in an infinitely repeated game, their
-respective probabilities for each of the four outcomes in the last round
-simplify to,
+Let's assume that individuals use their expected payoffs at the updating phase.
+We can calculate how easily a single defecting mutant (ALLD) can invade into a
+resident population of conditional cooperators, otherwise known as generous tit
+for tat (GTFT), players. Let GTFT\(= (1, 1, q)\), ALLD\(= (0, 0, 0)\), and \(k =
+1\). When two GTFT players interact in an infinitely repeated game, the
+stationary distribution \(\mathbf{v}(\text{GTFT}, \text{ALLD})\) 
+simplifies to,
 
@@
 
-\subsection{Updating Payoffs based on the last last \(m\) round(s) payoffs of \(n\) interaction(s)}
+This result will become important when we analyse the behaviour that emerges
+from the evolutionary process. We will repeat this ``invasion analysis'' for the
+rest of the updating payoffs cases.
 
-The aim of this work is to explore the effect of the updating payoffs on the
-evolution of cooperation. To this end we introduce a new method for calculating
-the updating payoffs and we compare the results of the new method to those of
-the expected payoffs. In our approach we do not assume that an individual has a
-perfect memory. Instead an individual remembers only a limited number of
-interactions and limited outcomes from each interaction. In the expected payoffs
-case the payoff of a pair depends on the average payoff they received over an
-infinite number of turns. In our approach, the payoff of a pair depends on the
-average payoffs they received in the last \(m\) turns. Furthermore, in expected
-payoffs it is assumed that a player interacts with every member of the
-population, whereas in our approach a player has \(n\) interactions.  Here we
-present results for four cases; (\(m=1\) and \(n=1\)), (\(m=2\) and \(n=1\)),
-(\(m=1\) and \(n=2\)), and (\(m=2\) and \(n=2\)).
+\section{Updating Payoffs based on the Last Round Payoff of One Interaction (limited memory)}\label{section:limited_memory}
 
-\subsubsection{Updating Payoffs based on the Last Round Payoff of One Interaction (\(m=1\) and \(n=1\))}\label{section:m_one_n_one}
-
-In this case an individual updates her/his strategy based on the last payoff
-they received against one other member of the population. Initially, we define
-the probability that a reactive strategy receives the payoff $u\!\in\!
-\mathcal{U}$ in the very last round of the game against another reactive
-strategy (Proposition~\ref{proposition:last_round}).
+In the expected payoffs case the payoff of a pair depends on the average payoff
+they received over an infinite number of turns. In the limited memory payoffs,
+the payoff of a pair depends on the average payoffs they received in the last
+turn. Furthermore, in expected payoffs it is assumed that a player interacts
+with every member of the population, whereas in the limited memory approach a
+player has one interaction. Initially, we define the probability that a reactive
+strategy receives the payoff $u\!\in\! \mathcal{U}$ in the very last round of
+the game against another reactive strategy
+(Proposition~\ref{proposition:last_round}).
 
@@
     \mathcal{U}$ in the very last round of the game is given by
-    $v_{u}(s_1,s_2)$, as given by Equation~(\ref{Eq:LastRound}).
+    $v_{u}(s_1,s_2)$, as given by Eq.~(\ref{Eq:LastRound}).
 
@@
 
-\subsubsection*{Simulation Results based on the Last Round Payoff of One Interaction}
-
-We simulate the evolutionary process given that individuals now use the last
-round payoff of one interaction to update their strategies. The results are
-shown in Figure~\ref{fig:one_interaction_last_round_payoffs_results}. Similarly
-to the results of the expected payoffs, a higher benefit results in a more
-cooperative population. However, we note that the cooperation rate of the
-evolved populations is lower compared to the case of expected payoffs, and that
-the increase of the cooperation rate between the two simulations is smaller. The
-residents population in both cases consist of defectors and conditional
-cooperators with \(p \approx 1\) and  \(q \approx \frac{1}{2}\). Thus, the
-evolved \(q\)s are smaller which is the reason why the population is less
-cooperative. In the invasion analysis we demonstrate that in the case of the
-last round payoff of one interaction, a conditional player needs to have \(q
-\leq \frac{1}{2}\) to avoid being invaded by a defector regardless of the
-benefit of cooperation.
-
-\begin{figure}[!htbp]
-  \centering 
-  \includegraphics[width=.70\textwidth]{static/one_interaction_last_round_donation_game.pdf}
-  \caption{\textbf{Evolutionary dynamics under one interaction last round
-  payoffs with low (left) and high (right) benefit.} The cooperation rate
-  regardless of the benefit of cooperation is less compared to that of the
-  expected payoffs. The increase of cooperation between the low benefit case and
-  the high benefit case is also smaller. In both cases the resident population
-  either consists of defectors (with
-  \(p \approx q \approx 0\)) or of conditional cooperators for which \(p \approx
-  1\) and \(q \leq \frac{1}{2}\). Parameters: \(N =100, c=1, \beta=1, T=10^{7}\).
-  }\label{fig:one_interaction_last_round_payoffs_results}
-\end{figure}
-
-\subsubsection*{Invasion Analysis based on the Last Round Payoff of One Interaction}
+\subsubsection*{Invasion Analysis}
 
@@
 
-\subsubsection{Updating Payoffs based on the last round payoff of two interactions (\(m=1\) and \(n=2\))}\label{section:m_one_n_two}
+\section{Updating Payoffs based on the last round payoff of $n$ interactions}\label{section:m_one_n_two}
 
-Here we consider the case where both the role model and the learner estimate
-their updating payoff after interacting with two members of the population.
-There are two stages of matching and there are twenty four possible pairs. In
+In the limited memory payoffs we assume that an individual recalls the last
+round they received against one member of the population. This is an edge case.
+The framework of the limited memory can be generalised such that an individual
+considers \(m\) rounds and of \(n\) interactions. Here we discuss the case that
+the update depends on the last round of \(n\) interactions.
+
+At each step of the evolutionary process the role model and the learner now
+participate in \(n\) matches. We need to define the probability that for these
+matches they are paired with a mutant, with a resident or with each other. We
+assume that each pair is unique, so the resident and role model can be matched
+together only once at each step. A representation of the process is given in
+Figure.
+
+In the case of \(n=1\) there are five possible pairs, however, the number of
+possible pairs increases non linearly as we increase the number of possible
+interactions. We demonstrate this case for \(n=2\), namely, for when the role
+model and the learner have two interactions.
+
+\subsection*{Special case: Last Round Payoff of Two Interactions}
+
+For \(n=2\) there are two stages of matching and there are twenty four possible pairs. In
 the first stage the role model and the learner are matched together with a
@@
 
-\subsubsection*{Simulation Results based on the Last Round Payoff of Two Interactions}
-
-We simulate the evolutionary process when individuals update their strategies
-based on the last round payoffs from two interactions with members of the
-population, Figure~\ref{fig:one_interaction_last_round_payoffs_results}. For
-both a low and a high benefit value the average cooperation rate has increased
-compared to the last round against a single other member of the population. In
-the case of low benefit the evolved population's cooperation rate is almost the
-same as in the case of expected payoffs. However, the cooperation rate does not
-increase as much as the expected payoffs for \(b=10\).
-
-\begin{figure}[!htbp]
-  \centering 
-  \includegraphics[width=.70\textwidth]{static/two_interactions_donation_game.pdf}
-  \caption{\textbf{Evolutionary dynamics under two interactions last round
-  payoffs with low (left) and high (right) benefit.}
-  The cooperation rates for both the simulations have increased and are similar
-  to the case of expected payoffs. The increase of the cooperation rate between
-  the two cases however is smaller. In both cases the resident population
-  either consists of defectors (with
-  \(p \approx q \approx 0\)) or of conditional cooperators for which \(p \approx
-  1\) and \(q \leq \frac{1}{2}\). Parameters: \(N =100, c=1, \beta=1, T=10^{7}\).
-  }\label{fig:two_interactions_results}
-\end{figure}
-
-\subsubsection*{Invasion Analysis based on the Last Round Payoff of Two Interactions}
+\subsubsection*{Invasion Analysis}
 
 In a similar fashion we can calculate the condition for which a population of
-GTFT players can not be invaded by an ALLD mutant. The ratio of transition
-probabilities is given by,
+GTFT players can not be invaded by an ALLD mutant. Using the new formulation we
+obtain the following probabilities \(\bar{x}(u_1, u_2)\) that the payoff of a randomly
+chosen GTFT player is \(u_1\) and that the payoff of the ALLD player is \(u_2\),
+
+\begin{align*}
+  x(r, t) = & \frac{N - 2}{N - 1} \cdot (1 - \delta + \delta q)\\
+  x(r, r) = & \frac{N - 2}{N - 1} \cdot \delta (1 - q) \\
+  x(s, r) = & \frac{1}{N - 1} \cdot (1 - \delta + \delta q) \\
+  x(p, p) = & \frac{1}{N - 1} \cdot \delta (1 - q)
+\end{align*}
+
+The ratio of transition probabilities is given by,
 
@@
 
-\subsection{Updating Payoffs based on the Last Two Rounds Payoff of One Interaction (\(m=2\) and \(n=1\))}\label{section:m_two_n_one}
+\section{Updating Payoffs based on the Last $m$ Rounds Payoffs of One Interaction}\label{section:m_two_n_one}
 
-In this section we consider the case of \(n=1\) and \(m=2\), thus individuals
-update based on their last two rounds payoffs with one member. To this end we
-define the probability that a reactive strategy receives the payoffs $u\!\in\!
-\mathcal{\tilde{U}}$, for \(\mathcal{\tilde{U}}= \{rrr\dots r, rrr\dots s, \dots,
-ppp\dots p\}\), in the last \(m\) rounds against another reactive strategy
-(Proposition~\ref{proposition:last_m_rounds}).
+The second generalised case of the limited memory payoffs we discuss in that of
+individuals updating based on their last $m$ rounds payoffs with one member. To
+this end we define the probability that a reactive strategy receives the payoffs
+$u\!\in\! \mathcal{\tilde{U}}$, for \(\mathcal{\tilde{U}}= \{\underbrace{rrr\dots r}_{m},
+\underbrace{rrr\dots s}_{m}, \dots, \underbrace{ppp\dots p}_{m}\}\), in the last \(m\) rounds against another
+reactive strategy (Proposition~\ref{proposition:last_m_rounds}).
 
@@
 
-\subsubsection*{Simulation Results based on the Last Two Rounds Payoff of One Interaction}
+\subsection*{Special case: Last Round Payoff of Two Interactions}
 
-The simulation results for this case are given in
-Figure~\ref{fig:two_rounds_results}. Updating on the two rounds payoffs received
-against a single opponent result to a lower cooperation rate compare to updating
-based on one round against two opponents.
-
-\begin{figure}[!htbp]
-  \centering 
-  \includegraphics[width=.70\textwidth]{static/two_rounds_donation_game.pdf}
-  \caption{\textbf{Evolutionary dynamics under two interactions last round
-  payoffs with low (left) and high (right) benefit.}\\Parameters: \(N =100, c=1, \beta=1, T=10 ^7\).
-  }\label{fig:two_rounds_results}
-\end{figure}
-
-\subsubsection*{Invasion Analysis based on the Last Two Rounds Payoff of One Interaction}
+\subsubsection*{Invasion Analysis}
 
@@
 
-\subsection{Updating Payoffs based on the last two rounds payoff of two interactions (\(m=2\) and \(n=2\))}
+\section{Updating Payoffs based on the last two rounds payoff of two interactions (\(m=2\) and \(n=2\))}\label{section:m_two_n_two}
 
-The last case we present in this work is the case of \(m = 2\) and \(n = 2\).
-That is, an individual considers the last two rounds of payoffs he or she
-received against two different members of the population. For this case we
-combine the approaches described in sections~\ref{section:m_one_n_two}
-and~\ref{section:m_two_n_one}. Here we only present results on the numerical
-simulations, Figure~\ref{fig:two_rounds_opponents_results}. The evolved
-cooperation rates are between the other cases and the expected payoffs. For a
-low benefit we get results similar to the case where one the last round was
-considered, however, in the high benefit case the evolved cooperation rate
-increases to 91\% which is higher then any of the newly introduced approaches.
+Finally a possible extension to the limited memory framework is to consider that
+thee number of rounds and number of interactions increase. For this case we need
+to consider a combination of the methods we presented in
+Section~\ref{section:m_one_n_two} and Section~\ref{section:m_two_n_one}.
+
+\section{Simulation Results on the Pairwise Comparison Process}\label{section:simulation_results}
+
+We simulate the evolutionary process described in
+Algorithm~\ref{algorithm:pairwise_comparison} for the different updating
+mechanisms we have described in
+Sections~\ref{section:perfect_memory}-\ref{section:m_two_n_two}. For each
+approach we performed an independent run of the process and for each time step we
+recorded the current resident population \((y, p, q)\). The results are shown in
+Figure~\ref{fig:expected_payoffs_results}. We observe that in most cases the
+resident population consists either of defectors or conditional cooperators. A
+conditional cooperator always cooperates if the co-player cooperated
+($p\approx1$) and cooperates with a probability \(q\) if the co-player defected.
+The most abundant conditional cooperators in each simulation differ as a
+result of the updating payoffs. More specifically, in order for a resident
+population of conditional cooperators to avoid being invaded they need to adopt
+a different value of \(q\). For each method we have discussed this under the
+invasion analysis subsection.
+
+In the cases of perfect memory the resident population adopts a \(q \leq 1 -
+\frac{c}{b}=0.9\). In the limited memory case the generosity of a conditional
+cooperator is independently of the benefit lower than \(\frac{1}{2}\). The rest
+of the cases also condition on the cost benefit relationship. In these
+simulations the cost of cooperation is set to 1 and the benefit to 10. As a
+result, in the case of two interactions \(q \leq \frac{\delta - 1 +
+\frac{\sqrt{2}}{2}}{\delta}=0.7068\), in the case of two rounds \(q \leq
+\frac{\delta + \sqrt{\delta^{2} + 1} - 1}{2 \delta}=0.7069\), and in the last
+case \(q \leq 1 - \frac{c}{b}=0.9\). The higher tolerance to defection results
+to a more cooperative population. As a result the expected payoffs allow for the
+most cooperative population. Between the limited memory approaches we observe a
+big difference in the cases of one or more information. We hypothesis that as we
+allow for more information the results will tend to the case of perfect memory.
 
@@
     \centering 
-  \includegraphics[width=.70\textwidth]{static/two_rounds_opponents_donation_game.pdf}
-  \caption{\textbf{Evolutionary dynamics under two interactions last round
-  payoffs with low (left) and high (right) benefit.} In the cases of low and
-  high benefit case the resident population either consists of defectors (with
-  \(p \approx q \approx 0\)) or of conditional cooperators for which \(p \approx
-  1\) and \(q \leq \frac{1}{2}\). Parameters: \(N =100, c=1, \beta=1\).
-  }\label{fig:two_rounds_opponents_results}
+    \includegraphics[width=\textwidth]{static/more_memory_heatmaps_donation_game_with_illustrations.pdf}
+    \caption{\textbf{Evolutionary dynamics with difference updating approach.}
+    From left to right, we present result on the following updating payoffs
+    cases; the expected payoffs (perfect memory), the last round payoff
+    from one interaction (limited memory), the last round payoff from two
+    interactions, the last two rounds payoffs from one interaction, the
+    last two rounds payoffs from two interactions. We run each simulation for \(T
+    = 10^7\) time steps. For each time step we recorded the current resident
+    population \((y, p, q)\). Since \(\delta \rightarrow 1\) we do not report the
+    players' initial cooperation probability \(y\). The graphs show how often the
+    resident population chooses each combination \((p, q)\) of conditional
+    cooperation probabilities in the subsequent rounds. In both cases players
+    update based on their expected payoffs.}\label{fig:expected_payoffs_results}
 \end{figure}
 
-\newpage
+\section{Expected and Last Round Updating Payoffs for Memory One Strategies}\label{section:memory_one}
 
-\section{Expected and Last Round Updating Payoffs for Memory One Strategies}
-
-So far we have considered the case where members can adopt reactive strategies.
+So far we have assumed that individuals can adopt reactive strategies.
 To demonstrate that our results hold for higher memory strategies here we
@@
 in Figure~\ref{fig:memory_one_low_benefit}, and for a high benefit in
-Figure~\ref{fig:memory_one_high_benefit}.
+Figure~\ref{fig:memory_one_high_benefit}. We verify that even when individuals
+are allowed to use memory-one strategies, the cooperation rate is higher
+in the perfect memory approach compared to the limited memory.
 
@@
 Section~\ref{section:pairwise_comparison}, and at each time step we record the
-average player \(\bar{s}=(\bar{y}, \bar{p}, \bar{q})\).
+average player \(\bar{s}=(\bar{y}, \bar{p}, \bar{q})\). The average cooperation
+of the resident population for different values of mutation are shown in
+Figure~\ref{fig:mutation}. The cooperation rate in the case of perfect
+memory is always higher compared to the limited memory regardless of the
+mutation value. For mutation value of 1 the processes become random and
+this results to a cooperation rate of \(\frac{1}{2}\) in both simulations.
 

### 2023-02-27 paper/main.tex
uncomment abstract and move other commented sections around

@@
 \begin{abstract}
-% Repeated games are vastly use in evolutionary game theory to explain cooperation in a variety of environments. Although existing models have shaped
-% our understanding of human cooperation, hey often work with idealized
-% assumptions. In an evolutionary process, individuals imitate other individuals
-% of the population based on their fitness. It is commonly assumed that an
-% individual computes their fitness after interacting with a representative sample
-% of the population, and remembering all the interactions they participated in. In
-% real life, we do not always remember all our interactions, instead we rather
-% recall our most recent ones.
-
-% Here, we introduce a framework that allows individuals to estimate their fitness
-% based on a minimum of social information. We explore the difference between our
-% framework and the classical framework using computer simulations. We present
-% results for the most commonly used classes of symmetric \(2 \times 2\) games.
-% The simulations show that, in the prisoner's dilemma individuals with limited
-% memory tend to adopt less generous strategies and they achieve less cooperation
-% than in the classical scenario. In contrast, in the stag-hunt and the harmony
-% game, the impact of memory is less striking. Here individuals with limited
-% memory perform nearly as well as individuals with full memory. Finally, we
-% observe that the snowdrift game is the only game for which cooperation can be
-% underestimated in the classical scenario.
+Repeated games are vastly use in evolutionary game theory to explain cooperation
+in a variety of environments. Although existing models have shaped our
+understanding of human cooperation, hey often work with idealized assumptions.
+In an evolutionary process, individuals imitate other individuals of the
+population based on their fitness. It is commonly assumed that an individual
+computes their fitness after interacting with a representative sample of the
+population, and remembering all the interactions they participated in. In real
+life, we do not always remember all our interactions, instead we rather recall
+our most recent ones.
+
+Here, we introduce a framework that allows individuals to estimate their fitness
+based on a minimum of social information. We explore the difference between our
+framework and the classical framework using computer simulations. The
+simulations show that, in the prisoner's dilemma individuals with limited memory
+tend to adopt less generous strategies and they achieve less cooperation than in
+the classical scenario.
 \end{abstract}
@@
 \section{Introduction}
-
-% The framework of evolutionary game theory has been applied in explaining the
-% behavior of sticklebacks~\cite{milinski1987tit}, the rock-scissors-paper cycles
-% in bacterial populations~\cite{kerr2002local} and food sharing in vampire
-% bats~\cite{carter2020development} colonies.
 
@@
 \section{Model Setup}\label{section:model}
-
-% ToDo: Add in discussion
-
-% A major focus of evolutionary game theory is the evolution of cooperation among
-% selfish individuals. Sometimes evolutionary game theory even seems to be reduced
-% to the evolution of cooperation [Arne2022]. “Why do we act altruistically to
-% provide a benefit to others when there is a cost to ourselves?” is still a
-% puzzling question amongst researchers [Glynatsi]. In this determinist
-% formulation, a small fraction of cooperative individuals that have a lower
-% fitness can never invade a population of defecting strategies.  The stochastic
-% effect can lead to fundamental changes since disadvantageous mutants have a
-% small yet non-zero probability to reach fixation in a finite population, and
-% thus cooperation can emerge.
 
@@
 
+% ToDo: Add in discussion
+
+% A major focus of evolutionary game theory is the evolution of cooperation among
+% selfish individuals. Sometimes evolutionary game theory even seems to be reduced
+% to the evolution of cooperation [Arne2022]. “Why do we act altruistically to
+% provide a benefit to others when there is a cost to ourselves?” is still a
+% puzzling question amongst researchers [Glynatsi]. In this determinist
+% formulation, a small fraction of cooperative individuals that have a lower
+% fitness can never invade a population of defecting strategies.  The stochastic
+% effect can lead to fundamental changes since disadvantageous mutants have a
+% small yet non-zero probability to reach fixation in a finite population, and
+% thus cooperation can emerge.
+
+% The framework of evolutionary game theory has been applied in explaining the
+% behavior of sticklebacks~\cite{milinski1987tit}, the rock-scissors-paper cycles
+% in bacterial populations~\cite{kerr2002local} and food sharing in vampire
+% bats~\cite{carter2020development} colonies.
+
 % Cooperation can be seen at odd, why is it that we choose to help others,

### 2023-02-27 paper/main.tex
delete old parts
finish abstract

@@
 \begin{abstract}
-Repeated games are vastly use in evolutionary game theory to explain cooperation
+Repeated games are vastly used in evolutionary game theory to explain cooperation
 in a variety of environments. Although existing models have shaped our
-understanding of human cooperation, hey often work with idealized assumptions.
+understanding of human cooperation they often work with idealized assumptions.
 In an evolutionary process, individuals imitate other individuals of the
@@
 computes their fitness after interacting with a representative sample of the
-population, and remembering all the interactions they participated in. In real
-life, we do not always remember all our interactions, instead we rather recall
+population and remembering all the interactions they participated in. In real
+life we do not always remember all our interactions, instead we rather recall
 our most recent ones.
 
-Here, we introduce a framework that allows individuals to estimate their fitness
+Here we introduce a framework that allows individuals to estimate their fitness
 based on a minimum of social information. We explore the difference between our
 framework and the classical framework using computer simulations. The
-simulations show that, in the prisoner's dilemma individuals with limited memory
+simulations show that in the prisoner's dilemma, individuals with limited memory
 tend to adopt less generous strategies and they achieve less cooperation than in
@@
 \newpage
-% In the
-% simplest of cases each individual can be of one of two types, A and B. The state
-% of the population is thus characterized by the number \(k\) of individuals of
-% type A. The interaction between the two types of individuals is described by the
-% functions \(\pi _{{\rm A}}^{i}\) and $\pi _{{\rm B}}^{i}$.
-
-% airwise comparison process is a stochastic process for modelling the evolution
-% of a finite population. The process starts with assigning all individuals of the
-% population the same strategy. A strategy is a set of rules of how an individual
-% should behave in an interaction with another individual. Each elementary time
-% step of the process consists of three phases; (1) the \textbf{mutation phase}
-% (2) the \textbf{game phase} and (3) the \textbf{update phase}. These are summarised
-% in Figure~\ref{fig:pairwise_phases}.
-
-% In the \textbf{mutation phase} one individual is
-% chosen to switch to a new mutant strategy with a probability \(\mu\).
-% In the \textbf{game phase} individuals are randomly matched with other
-% individuals in the population, and they engage in a repeated game where each
-% subsequent turn occurs with a fixed probability \(\delta\). At each turn the
-% individuals decide on an action based on their strategies.
-% In repeated games there are infinite many strategies, however, it is commonly
-% assumed that individuals can only choose strategies from a restricted set. One
-% such set is that of reactive strategies. A reactive
-% strategy considers only the previous action of the other player, and thus, a
-% reactive strategy \(s\) can be written as a three-dimensional vector \(s=(y, p,
-% q)\). The parameter \(y\) is the probability that the strategy opens with a
-% cooperation and \(p\), \(q\) are the probabilities that the strategy cooperates
-% given that the opponent cooperated and defected equivalently.
-
-% In the \textbf{update stage} where two individuals are randomly selected. From
-% the two individuals, one  serves as the `learner' and the other as the `role
-% model'. The learner adopts the role model's strategy with a probability \(\rho\)
-% given by,
-
-% The fitness of the individuals (\(\pi _{{\rm A}}^{i}, \pi _{{\rm B}}^{i}\)) is
-% identified with the payoff resulting from the underline game. The evolution of
-% cooperation remains one of the greatest problems for biological and social
-% sciences. Cooperation is the action of choosing to help others at one's own
-% expense. The standard game of formulating such situation is the Prisoner's
-% Dilemma~\cite{trivers1971evolution, milinski1987tit, glynatsi2021bibliometric,
-% ohtsuki2006simple}, in which two players can choose to cooperate or to defect.
-% The players are offered a certain payoff, \(R\), for mutual cooperation and a
-% lower payoff, \(P\), for mutual defection. If one player cooperates while the
-% other defects, then the cooperator gets the lowest payoff, \(S\), while the
-% defector gains the highest payoff, \(T\). Thus, the payoffs have the following
-% property \(T > R > P > S\), making defection the dominant strategy in the
-% non-repeated game.
-
-% \begin{equation}\label{eq:payoffs}
-%   \begin{blockarray}{ccc}
-%       & \text{cooperate} & \text{defect} \\
-%       \begin{block}{c(cc)}
-%           \text{cooperate} & R & S \\
-%           \text{defect} & T & P \\
-%       \end{block}
-%   \end{blockarray}
-% \end{equation}
-
-% In literature, the payoffs of A and B individuals only on the fraction of both
-% types in the population. If there are \(k\) A individuals and \(N-k\) B
-% individuals. In the context of the prisoner's dilemma the payoffs and assuming
-% that A and B follow simple strategies where they cooperate and defect
-% unconditionally respective, then the A and B individuals have fitness of
-
-% \begin{equation}
-%   \begin{array}{lcrcr}
-%   \displaystyle \pi _{{\rm A}}^{i}	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot R	&+	&\displaystyle\frac{k}{N-1}\cdot S,\\[0.5cm]
-%   \displaystyle \pi _{{\rm B}}^{i}&=	&\displaystyle\frac{N-k}{N-1}\cdot T &+	&\displaystyle\frac{k-1}{N-1}\cdot P.\\
-%   \end{array}
-% \end{equation}
-
-% respectively while self-interactions are excluded. In literature these are the
-% expected payoffs. For \(\beta = 1\), fitness equals payoff. This scenario
-% describes ``strong selection''. For \(\beta \ll 1\) , the payoff only provides a
-% small perturbation to the overall fitness of an individual, a limit known as
-% weak selection.
-
-% In the evolutionary game theory literature it is usually assumed that players
-% use strategies with finite memory~\cite{Nowak1992tit, Baek2016}. This assumption
-% is common as it allows for an explicit calculation of the players'
-% payoffs~\cite{sigmund2010calculus}. Players' payoffs are calculated in the limit
-% of interactions assuming that the two strategies meet several times.
-
-% Customarily, most evolutionary game studies make the additional assumption that
-% individuals play many times and with all other players before reproduction takes
-% place, so that payoffs, equivalently fitness, are given by the mean distribution
-% of types in the population. This hypothesis, selection occurs much more slowly
-% than the interaction between individuals and more that remember all the
-% interactions they participated in. Although recent experimental studies show
-% that this may not always be the case in biology [12– 14], it is clear that in
-% cultural evolution or social learning the time scale of selection is much closer
-% to the time scale of interaction. The effects of this mixing of scales cannot be
-% disregarded [15], and then it is natural to ask about the consequences of the
-% above assumption and the effect of relaxing it. The work of~\cite{Roca2006}.
-
-% In this work we ask studies that make use of this classical framework report
-% that cooperation can substantially evolve. However, the inconsistency in the
-% memory size of players leads us to question the robustness of our understanding
-% of cooperation. To this end, we propose a framework in which individuals,
-% similar to the decisions at each turn, estimate their fitness based on a minimum
-% of information.
-
-% We first consider two extreme scenarios, the classical scenario and the
-% alternative scenario where individuals update their strategies only based on the
-% very last payoff they obtained. We observe that individuals with limited memory
-% tend to adopt less generous strategies and they achieve less cooperation when
-% interacting in a prisoner's dilemma. We obtain similar results when we consider
-% that individuals update their strategies based on more information. More
-% specifically, up to the last two payoffs they obtained when interacting with up
-% to two different members of the population. We extend our approach to the rest
-% of the symmetric \(2 \times 2\) games.
-
-% The remainder of the paper is organized as follows. In
-% section~\ref{section:model} we describe the model. In
-% section~\ref{section:results} we present the results of the simulations, and in
-% section~\ref{section:conclusions} we outline the main conclusions.
-
-% In the following, we consider a well mixed population of fixed size. In each
-% step of t
-
-% We study the transmission of strategies with a frequency-dependent birth-death
-% process26 in a finite population of size n. In each time step, two randomly
-% chosen individuals compare their payoffs and one of them can switch to the other
-% one's strategy. This process can be interpreted as a model for social learning,
-% whereby successful strategies spread, and, occasionally, random strategy
-% exploration introduces novel strategies (corresponding to mutations in
-% biological models).
-
-% We consider a population of \(N\) players~\footnote{The terms ``player'' and
-% ``individual'' are used interchangeably here.} where \(N\) is even and mutations
-% are sufficiently rare. Therefore, at any point in time there are at most two
-% different strategies present in the population; a \textit{resident} strategy and
-% a \textit{mutant} strategy. To describe how strategies spread we use a pairwise
-% comparison process~\cite{Traulsen2006}. Each step of the evolutionary process
-% consists of two stages, a game stage and an updating stage.
-
-% In the game stage each individual is randomly matched with some other individual
-% in the population. They engage in a match where each subsequent turn occurs with
-% a fixed probability $\delta$. At each turn players choose independently to
-% either cooperate (\(C\)) or to defect (\(D\)), and the payoffs of the turn
-% depend on both their decisions. If both players cooperate they receive the
-% reward payoff \(R\), whereas if both defect they receive the punishment payoff
-% \(P\). If one cooperates but the other defects, the defector receives the
-% temptation payoff \(T\), whereas the cooperator receives the sucker's payoff
-% \(S\). We denote the feasible payoff of each turn as \(\mathcal{U} = \{R, S, T,
-% P\}\). We assume that individuals use \textit{reactive strategies} to make
-% decisions in each turn. Reactive strategies are a set of memory-one strategies
-% that only take into account the previous action of the opponent. They can be
-% written explicitly as a vector in \(\R_{3}\), more specifically, a reactive
-% strategy \(s\) is given by \(s=(y, p, q)\). The parameter \(y\) is the
-% probability that the strategy opens with a cooperation and \(p, q\) are the
-% probabilities that the strategy cooperates given that the opponent cooperated
-% and defected equivalently.
-
-% In the updating stage, two players are randomly drawn from the population, a
-% `learner' and a `role model'. Given the learner's payoff $u_L\!\in\!
-% \mathcal{U}$ and the role model's payoff $u_{RL}\!\in\! \mathcal{U}$, the
-% learner adopts the role model's strategy with probability,
-
-% % \begin{equation} \label{Eq:rho}
-% %   \rho(u_{L}, u_{RM}) = \frac{1}{1\!+\! \exp^{\!-\!\beta (u_{RM}\!-\!u_{L})}}.
-% % \end{equation}
-
-% where $\beta\!\ge\!0$ is the strength of selection. For small values of $\beta$
-% the imitation probability is independent of the strategies of the involved
-% players. As the value of $\beta$ increases, the more likely it is that the
-% learner adopts only strategies that yield a higher payoff. Conventionally the
-% updating payoffs of the learner and the role model are based on their expected
-% payoffs. A player's expected payoff is the mean payoff the player yields after
-% engaging in matches of multiple turns with each member of the population. At
-% each match a player bases their next turn decision only on the previous action
-% of the opponent, however, the same player bases their expected payoffs on the
-% outcomes of all their matches. Thus, a player is assumed to have limited and
-% perfect memory at the same time. We propose a new a set of updating payoffs
-% where it is also assumed that a player has limited memory. We referee to these
-% as the limited memory payoffs.
-
-% The evolutionary step is repeated until either the mutant strategy goes extinct,
-% or until it fixes in the population. If the mutant fixes in the population then
-% the mutant strategy becomes the new resident strategy. After either outcome we
-% introduce a new mutant strategy uniformly chosen from all reactive strategies at
-% random, and we set the number of mutants to $1$. This process of mutation and
-% fixation/extinction is then iterated many times.
-
-% \section{Results}\label{section:results} \subsection{Updating payoffs based on
-% the last round with another member of the population}\label{section:donation}
-
-% To investigate the role of extortion in the context of evolutionary games, we
-% concentrate on the donation game.
-% In this section we explore the case where the updating payoffs are based on the
-% last round payoff achieved against another member of the population and we
-% compare this to the classical scenario of the expected payoffs. We assume that
-% each pair of players interacts in a donation game. The donation game is a
-% special case of the prisoner's dilemma. Each player can choose to cooperate by
-% providing a benefit \(b\) to the other player at their cost \(c\), with \(0 < c
-% < b\). Thus, the feasible payoffs in each round are \(\mathcal{U} = \{b-c, -c,
-% b, 0\}\).
-
-% \begin{figure}[!htbp]
-%   \centering
-%   \includegraphics[width=\textwidth]{static/cooperation_rates_expected_and_stochastic_for_donation_game.pdf}
-% \end{figure}
-
-% We further explore the effect of benefit in
-% Figure~\ref{fig:cooperation_rate_over_benefit}. The figure suggests that
-% expected payoffs always yield a higher cooperation rate. In the case of expected
-% payoffs we observe that the cooperation rate increases as the value of the
-% benefit gets higher. In comparison for the limited memory payoffs, the
-% cooperation rate remains unchanged at approximately 50\% once \(b=5\).
-
-% % \begin{figure}[!htbp]
-% %   \centering
-% %   \begin{subfigure}{.5\textwidth}
-% %     \centering
-% %     \includegraphics[width=\textwidth]{static/expected_for_beta.pdf}
-% %     \includegraphics[width=\textwidth]{static/stochastic_for_beta.pdf}
-% %   \end{subfigure}%
-% %   \begin{subfigure}{.5\textwidth}
-% %     \centering
-% %     \includegraphics[width=\textwidth]{static/cooperation_rate_over_b.pdf}
-% %   \end{subfigure}
-% %   \caption{{\bf The evolution of cooperation for different benefit values.} 
-% %   We vary the benefit of defection $b$. In all cases, expected payoffs appear to
-% %   overestimate the average cooperation rate the population achieves. ({\bf A})
-% %   the probabilities \(p, q\) for resident population over \(10^7\) time steps
-% %   for each benefit value. ({\bf B}) The cooperation rate within the resident
-% %   population (averaged over all games and over all time steps) over the benefit.
-% %   Unless explicitly varied, the parameters of the simulation are $N\!=\!100$,
-% %   $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
-% %   $t\!=\!5\times 10^7$ time steps for each parameter
-% %   combination.}\label{fig:cooperation_rate_over_benefit}
-% % \end{figure}
-
-% We also investigate the effect of the strength of selection $\beta$.
-% Figure~\ref{fig:cooperation_rate_over_betas} illustrates results for various
-% runs of the evolutionary process. For weak selection, \(\beta < 1\), we observe
-% that the two methods yield similar results, however, as \(\beta\) increases
-% there is variation in the evolving populations. In the case of expected payoffs
-% the resident populations become more cooperative as \(\beta\) increases, whereas
-% in the case of limited memory payoffs, the resident populations become more
-% defective.
-
-% % \begin{figure}[!htbp]
-% %   \centering
-% %     \includegraphics[width=\textwidth]{static/cooperation_rate_over_b.pdf}
-% % \caption{{\bf The evolution of cooperation for different selection strength values.}
-% % We vary the selection strength $\beta$. In all cases, stochastic payoff
-% % evaluation tends to reduce the evolving cooperation rates. ({\bf A}) the
-% % probabilities \(p, q\) for resident population over \(10^7\) time steps for each
-% % \(\beta\) value. ({\bf B}) The cooperation rate within the resident population
-% % (averaged over all games and over all time steps) over \(\beta\). Unless
-% % explicitly varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$,
-% % $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
-% % $t\!=\!5\times 10^7$ time steps for each parameter
-% % combination.}\label{fig:cooperation_rate_over_betas}
-% % \end{figure}
-
-% \begin{figure}[!htbp]
-%   \centering
-%     \includegraphics[width=\textwidth]{static/more_memory_summary_results}
-% \caption{{\bf The evolution of cooperation for different selection strength values.}
-% We vary the selection strength $\beta$. In all cases, stochastic payoff
-% evaluation tends to reduce the evolving cooperation rates. ({\bf A}) the
-% probabilities \(p, q\) for resident population over \(10^7\) time steps for each
-% \(\beta\) value. ({\bf B}) The cooperation rate within the resident population
-% (averaged over all games and over all time steps) over \(\beta\). Unless
-% explicitly varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$,
-% $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
-% $t\!=\!5\times 10^7$ time steps for each parameter
-% combination.}\label{fig:cooperation_rate_over_betas}
-% \end{figure}
-
-% \subsection{Effect of updating payoffs in different social
-% dilemmas}\label{section:2_by_2_games}
-
-% In the previous section we gained insights into the effects of the updating
-% payoffs, and into how parameters such as the benefit and the strength of
-% selection can intensify them. We investigated these effects by using the
-% donation game. In order to broaden our understanding of the updating payoffs on
-% different forms of possible human interactions we extend our approach to all \(2
-% \times 2\) symmetric games. More specifically, we apply our analysis to four
-% different classes of games; the harmony, the snowdrift, the prisoner's dilemma
-% and the stag-hunt games.
-
-% We compare the results of the evolutionary process when the updating payoffs are
-% based on the expected payoffs, and on the last round payoffs for all the four
-% possible classes of games (Figures~\ref{fig:expected_two_by_two}
-% and~\ref{fig:last_round_two_by_two}). In Figure~\ref{fig:expected_two_by_two}
-% each sub figure represents a run of evolutionary process for a different set of
-% values for \(\mathcal{U}\). Without loss of generality we set \(R=1\) and
-% \(P=0\)~\cite{Martinez2012, Roca2009}, and we vary the values of the temptation
-% payoff \(T\) (across the \(x\)-axis) and of the sucker's payoff \(S\) (across
-% the \(y\)-axis). Starting at the upper left corner and proceeding clockwise the
-% quadrants correspond to the harmony, the snowdrift, the prisoner's dilemma and
-% the stag-hunt games.
-
-% % % \begin{figure}[!htbp]
-% % %   \centering
-% % %   \includegraphics[width=\textwidth]{static/expected_two_by_two_games.pdf}
-% % %   \caption{{\bf Evolutionary dynamics under expected payoffs for various social dilemmas.}
-% % %   We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across
-% % %   the \(x\) axis, and  \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\) across the
-% % %   \(y\) axis. ({\bf A}) The top left quadrant where \(S > 0\) and \(T \leq 1\)
-% % %   corresponds to the harmony game. The preference ordering for the harmony game
-% % %   is \(R > T > S > P\). ({\bf B}) The top right quadrant where \(S > 0\) and \(T
-% % %   > 1\) corresponds to the snowdrift game. The preference ordering for the
-% % %   snowdrift game is \(T > R > S > P\). ({\bf C}) The bottom left quadrant where
-% % %   \(S < 0\) and \(T \leq 1\) corresponds to the stag-hunt game. The preference
-% % %   ordering for the stag-hunt game is \(R > T > P > S\). ({\bf D}) The bottom
-% % %   right quadrant where \(S < 0\) and \(T > 1\) corresponds to the prisoner's
-% % %   dilemma game. The preference ordering for the prisoner's dilemma game is \(T >
-% % %   R > P > S\). Unless explicitly varied, the parameters of the simulation are
-% % %   $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}
-% % %   \label{fig:expected_two_by_two}
-% % % \end{figure}
-
-% % % \begin{figure}[!htbp]
-% % %   \centering
-% % %   \includegraphics[width=\textwidth]{static/stochastic_two_by_two_games.pdf}
-% % %   \caption{{\bf Evolutionary dynamics under last round payoffs for various social dilemmas.} 
-% % %   We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across
-% % %   the \(x\) axis, and  \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\) across the
-% % %   \(y\) axis. ({\bf A}) The top left quadrant where \(S > 0\) and \(T \leq 1\)
-% % %   corresponds to the harmony game. The preference ordering for the harmony game
-% % %   is \(R > T > S > P\). ({\bf B}) The top right quadrant where \(S > 0\) and \(T
-% % %   > 1\) corresponds to the snowdrift game. The preference ordering for the
-% % %   snowdrift game is \(T > R > S > P\). ({\bf C}) The bottom left quadrant where
-% % %   \(S < 0\) and \(T \leq 1\) corresponds to the stag-hunt game. The preference
-% % %   ordering for the stag-hunt game is \(R > T > P > S\). ({\bf D}) The bottom
-% % %   right quadrant where \(S < 0\) and \(T > 1\) corresponds to the prisoner's
-% % %   dilemma game. The preference ordering for the prisoner's dilemma game is \(T >
-% % %   R > P > S\). Unless explicitly varied, the parameters of the simulation are
-% % %   $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}
-% % %   \label{fig:last_round_two_by_two}
-% % % \end{figure}
-
-% % % The harmony game describes an interaction without conflict. In the harmony game
-% % % it is in the best interest for both players to cooperate, and in an evolutionary
-% % % process cooperators prevail. We observe that for most harmony game runs the
-% % % resident population overwhelmingly applies a strategy for which \(p\) and \(q\)
-% % % are \(\approx 1\) (Figure~\ref{fig:expected_two_by_two}). The snowdrift game
-% % % describes a situation similar to that of the prisoner's dilemma; cooperation
-% % % results in a benefit to the opposing player but entailing a cost to the
-% % % cooperator. However, in the snowdrift game individuals obtain immediate direct
-% % % benefits from the cooperative acts which leads to \(S > P\). In the snowdrift
-% % % game more cooperation emerges compared to the prisoner's dilemma. Defection is
-% % % never a resident strategy and for the same values of temptation the overall
-% % % \(q-\) values are higher. The last class of games we present are for the
-% % % stag-hunt game. In the stag-hunt game there is an equilibrium in which both
-% % % players cooperate as well as one in which both defect. For small values of the
-% % % temptation and the sucker's payoffs we observe that cooperation prevails similar
-% % % to the harmony game. However, in the case where the temptation and the reward
-% % % payoffs are equal the resident population either consists of defectors or
-% % % cooperators.
-
-% % % Although the patterns of the evolved populations remain similar if we consider
-% % % that players update their strategies based on their last round payoff
-% % % (Figure~\ref{fig:last_round_two_by_two}), there are still dissimilarities
-% % % between the two approaches. To better understand these dissimilarities, we
-% % % explore the cooperation rates for each of the cases of the social dilemmas we
-% % % have presented (Figure~\ref{fig:cooperation_rates_two_by_two_part_one}).
-
-% % % In the harmony game the cooperation rates for both approaches are similar. The
-% % % biggest difference occurs for \(T=1\), nevertheless, it is less than 10\%. There
-% % % are two instances for which the cooperation rates are higher for the limited
-% % % memory payoffs compared to the expected. These are for \(T=2.2, S=0.4\) and for
-% % % \(T=2.2, S=0.8\). It is only in the snowdrift game that cooperation is
-% % % overestimated when a minimum of information is considered. Players are slightly
-% % % more likely to cooperate after a cooperation and to forgive after being at the
-% % % receiving end of a defection. For the rest of the cases the classical scenario
-% % % overestimates the average cooperation rate, supporting the results of the
-% % % previous section. There is a substantial difference in the cooperation rates for
-% % % the prisoner's dilemma class. We observe that for \(S<-0.4\) cooperation almost
-% % % never emerges if players imitate based only on the last round payoff.
-
-% % % \begin{figure}[!htbp]
-% % %   \centering
-% % %   \includegraphics[width=.4\textwidth]{static/expected_two_by_two_games_cooperation.pdf}
-% % %   \includegraphics[width=.4\textwidth]{static/stochastic_two_by_two_games_cooperation.pdf}
-% % %   \caption{\textbf{Cooperation rates for various social dilemmas.} ({\bf A}) If
-% % %   players update based on their expected payoffs. ({\bf B}) If players update
-% % %   based on their last round payoffs. We vary the temptation payoff \(T \in
-% % %   \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across the \(x\) axis, and \(S \in \{1.2,
-% % %   0.8, 0.4, 0, -0.4, -0.8\}\) across the \(y\) axis. Unless explicitly varied,
-% % %   the parameters of the simulation are $N\!=\!100$, $\beta\!=\!10$,
-% % %   $\delta\!=\!0.99$, $t\!=\!5\times
-% % %   10^7$.}\label{fig:cooperation_rates_two_by_two_part_one}
-% % % \end{figure}
-
-% % % So far we have explored the difference between the expected payoffs and the last
-% % % round payoffs. In order to explore further the effect of limited memory we allow
-% % % individuals to remember more. More precisely, up to two interactions and up to
-% % % the last two rounds. In total we present results for three more updating
-% % % payoffs. These are the payoffs when individuals consider the last two rounds
-% % % with another member of the population, the last round with two members of the
-% % % population, and the last two rounds with two members of the population. Similar
-% % % to the last round payoff, we use simulations and record which strategies the
-% % % players adopt over time based
-% % % (Appendix~\ref{appendix:further_simulation_results}) and we compare the evolving
-% % % cooperation rates to those of the expected payoffs
-% % % (Figure~\ref{fig:cooperation_rates_two_by_two_part_two}).
-
-% % % In the case of the last two rounds payoff we make two observations. Initially,
-% % % for the harmony, the stag-hunt and the prisoner's dilemma classes the average
-% % % cooperation rates are higher than the previous case of the last round payoff.
-% % % The rates remain strictly less than in the expected payoffs, however, in the
-% % % case of the prisoner's dilemma we observe a significant increase. Secondly, for
-% % % the snowdrift class there are more instances for which the cooperation rates are
-% % % higher than in the expected payoffs. This could indicate that in the case of the
-% % % snowdrift game the expected payoffs underestimate cooperation. Once players
-% % % interact with two other members of the population, instead of one, the expected
-% % % payoffs strictly yield a higher cooperation rate  even in the case of the
-% % % snowdrift game. Similar to the previous result, the difference between the two
-% % % approaches is always higher for the prisoner's dilemma class.
-
-% % % \begin{figure}[!htbp]
-% % %   \centering
-% % %   \includegraphics[width=.33\textwidth]{static/rounds_two_by_two_games_cooperation.pdf}
-% % %   \includegraphics[width=.33\textwidth]{static/opponents_two_by_two_games_cooperation.pdf}
-% % %   \includegraphics[width=.33\textwidth]{static/rounds_opponents_two_by_two_games_cooperation.pdf}
-% % %   \caption{\textbf{Cooperation rates over various social dilemmas for different
-% % %   limited memory payoffs.} ({\bf A}) If players update based on their last two
-% % %   rounds with another member of the population. ({\bf B}) If players update
-% % %   based on their last round with two other members of the population. ({\bf C})
-% % %   If players update based on their last two rounds with two other members of the
-% % %   population. We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8,
-% % %   2.2\}\) across the \(x\) axis, and \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\)
-% % %   across the \(y\) axis. Unless explicitly varied, the parameters of the
-% % %   simulation are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times
-% % %   10^7$..}
-% % %   \label{fig:cooperation_rates_two_by_two_part_two}
-% % % \end{figure}
-
-% \section{Conclusions}\label{section:conclusions}
-
-% ToDo: Add in discussion
-
-% A major focus of evolutionary game theory is the evolution of cooperation among
-% selfish individuals. Sometimes evolutionary game theory even seems to be reduced
-% to the evolution of cooperation [Arne2022]. “Why do we act altruistically to
-% provide a benefit to others when there is a cost to ourselves?” is still a
-% puzzling question amongst researchers [Glynatsi]. In this determinist
-% formulation, a small fraction of cooperative individuals that have a lower
-% fitness can never invade a population of defecting strategies.  The stochastic
-% effect can lead to fundamental changes since disadvantageous mutants have a
-% small yet non-zero probability to reach fixation in a finite population, and
-% thus cooperation can emerge.
-
-% The framework of evolutionary game theory has been applied in explaining the
-% behavior of sticklebacks~\cite{milinski1987tit}, the rock-scissors-paper cycles
-% in bacterial populations~\cite{kerr2002local} and food sharing in vampire
-% bats~\cite{carter2020development} colonies.
-
-% Cooperation can be seen at odd, why is it that we choose to help others,
-% increasing their payoff, at the expense of decreasing one's own? In spite of all
-% the selfish genes’ animal and human communities seem to altruistically help each
-% other and cooperate, and evolutionary game theory has helped us shape our
-% understanding of the evolution of cooperation.
-
-% Previous evolutionary models often feature a curious inconsistency. When
-% modeling how individuals make decisions in each round, these models assume that
-% players only remember the last round. However, when modeling how individuals
-% update their strategies over time, individuals are assumed to have perfect
-% memory.
-
-% Here, we have explored how robust cooperation is as models deviate from the
-% perfect memory assumption. Initially we considered the donation game. We showed
-% that when the last round payoff is used instead of the expected payoffs,
-% cooperation can even evolve when individuals only use a minimum of information,
-% however, the evolving cooperation rates are typically lower. The resident
-% strategies were both less cooperative and less generous. This effect was only
-% intensified we increase the benefit and the strength of selection independently.
-% The results showed that as each parameter was being increased the difference
-% between the cooperation rates were widening. This indicates that cooperative
-% players benefit from being able to interact with everyone in the population.
-
-% We extended our approach and presented results not only based on different
-% classes of social games, but also based on different limited memory payoffs. The
-% analysis showed that specifically for the case of the prisoner's dilemma games
-% the contrast between the two approaches can be significantly large. The
-% prisoner's dilemma is one of the most applied types of social games when we
-% discuss the evolution of cooperation. Our results indicate that cooperation
-% struggles to evolve in the prisoner's dilemma when only a minimum of social
-% information is used at the updating stage. Interestingly our simulations also
-% showed that in same cases cooperation can benefit from limited memory. This was
-% specific for the snowdrift game.
-
-% \section{Acknowledgements}
-
-% This work was supported by the European Research Council Starting Grant 850529:
-% E-DIRECT.
+
+\section{Conclusions}\label{section:conclusions}
+
+A major focus of evolutionary game theory is the evolution of cooperation among
+selfish individuals. Sometimes evolutionary game theory even seems to be reduced
+to the evolution of cooperation [Arne2022]. “Why do we act altruistically to
+provide a benefit to others when there is a cost to ourselves?” is still a
+puzzling question amongst researchers [Glynatsi]. In this determinist
+formulation, a small fraction of cooperative individuals that have a lower
+fitness can never invade a population of defecting strategies.  The stochastic
+effect can lead to fundamental changes since disadvantageous mutants have a
+small yet non-zero probability to reach fixation in a finite population, and
+thus cooperation can emerge.
+
+Cooperation can be seen at odd, why is it that we choose to help others,
+increasing their payoff, at the expense of decreasing one's own? In spite of all
+the selfish genes' animal and human communities seem to altruistically help each
+other and cooperate, and evolutionary game theory has helped us shape our
+understanding of the evolution of cooperation.
+
+Previous evolutionary models often feature a curious inconsistency. When
+modeling how individuals make decisions in each round, these models assume that
+players only remember the last round. However, when modeling how individuals
+update their strategies over time, individuals are assumed to have perfect
+memory.
+
+Here, we have explored how robust cooperation is as models deviate from the
+perfect memory assumption. Initially we considered the donation game. We showed
+that when the last round payoff is used instead of the expected payoffs,
+cooperation can even evolve when individuals only use a minimum of information,
+however, the evolving cooperation rates are typically lower. The resident
+strategies were both less cooperative and less generous. This effect was only
+intensified we increase the benefit and the strength of selection independently.
+The results showed that as each parameter was being increased the difference
+between the cooperation rates were widening. This indicates that cooperative
+players benefit from being able to interact with everyone in the population.
+
+We extended our approach and presented results not only based on different
+classes of social games, but also based on different limited memory payoffs. The
+analysis showed that specifically for the case of the prisoner's dilemma games
+the contrast between the two approaches can be significantly large. The
+prisoner's dilemma is one of the most applied types of social games when we
+discuss the evolution of cooperation. Our results indicate that cooperation
+struggles to evolve in the prisoner's dilemma when only a minimum of social
+information is used at the updating stage.
+
+\section{Acknowledgements}
+
+This work was supported by the European Research Council Starting Grant 850529:
+E-DIRECT.
 

### 2023-02-27 paper/main.tex
finish introduction

@@
 has explored the effects of constraining the interactions an individual has.
-This can be done by letting individuals have a stochastic number of
+Either by letting individuals have a stochastic number of
 interactions~\cite{roca:PhysicalReview:2006, sanchez:JTB:2005,
-Traulsen:JTB:2007}, a small number of interactions, or even in the extreme case
+Traulsen:JTB:2007} or by only allowing for
 a single interaction~\cite{Woelfing:JTB:2009}. However, no previous work
@@
 limited memory where individuals update their strategies only based on the very
-last payoff they obtained. We demonstrate the effects of limited updating payoff
+last payoff they obtained. We demonstrate the effects of limited payoff
 memory using a pairwise comparison process and the well studied game of the
-prisoner's dilemma. We observe that individuals with limited memory tend to
+prisoner's dilemma. We observe that individuals with limited payoff memory tend to
 adopt less generous strategies and they achieve less cooperation when

### 2023-02-27 paper/main.tex
finish model setup

@@
 A pairwise comparison process starts with assigning all
-individuals of the population the same strategy. Each elementary time step of
+individuals of the population the same strategy. Thenceforth each elementary time step of
 the process consists of the mutation phase, the game phase and the update phase.
@@
 their cost \(c\), with \(0 < c < b\). Thus the payoffs for a player in each turn
-are,
+are given by,
 
@@
 
-Let \(\mathbf{u} = (b-c, -c, b, 0)\) be payoffs in a vector format, and let
+Let \(\mathbf{u} = (b-c, -c, b, 0)\) be the round payoffs in a vector format, and let
 \(\mathcal{U} = \{r, s, t, p\}\) denote the set of feasible payoffs, where \(r\)
@@
 given that the opponent cooperated and defected equivalently. The play between a
-pair of reactive strategies ($s_1 = (y_1, p_1, q_1)$, $s_2 = (y_2, p_2, q_2)$)
+pair of reactive strategies $s_1 = (y_1, p_1, q_1)$ and $s_2 = (y_2, p_2, q_2)$
 can be model as a Markov process with the transition matrix \(M\),
@@
 present in the population at any given time. However, in the Supplementary Information
-Section 4 we show in that the main result hold for \(\mu \neq 0\).
+Section 9 we show that the main result holds for \(\mu \neq 0\).
 The case of low mutation is vastly adopted because it allows us to explicitly
@@
 feasible strategies. The fixation probability \(\phi_{M}\) of the mutant
-strategy can be calculated explicitly,
+strategy can be calculated explicitly as,
 
@@
 
-where \(\lambda^-_k, \lambda^+_k\) are the probabilities that the number of
-mutants decreases and increases respectively, and \(k\) is the number of
-mutants. Depending on the fixation probability \(\phi_{M}\) the mutant either
-fixes (becomes the new resident) or goes extinct. Regardless, in the elementary
+where \(k\) is the number of
+mutants and \(\lambda^-_k, \lambda^+_k\) are the probabilities that the number of
+mutants decreases and increases respectively.
+Depending on the fixation probability \(\phi_{M}\) the mutant either
+fixes (becomes the new resident) or goes extinct. Regardless, in the next elementary
 time step another mutant strategy is introduced to the population. We iterate
@@
 strategies and we record the resident strategies at each time step. The
-probabilities \(\lambda^-_k \text{ and } \lambda^+_k\) depend on the updating
-payoffs of the mutant and the resident strategies. In the next section we
-present how they are calculated in the cases of perfect and limited memory.
+probabilities \(\lambda^-_k \text{ and } \lambda^+_k\) depend on the fitness
+of the mutant and the resident strategies. In the next section we
+present how fitness is calculated in the cases of perfect and limited payoff memory.
 
-\subsection{Updating payoffs based Perfect and Limited Memory}
+\subsection{Fitness based on Perfect and Limited Payoff Memory}
 
-\subsubsection*{Perfect Memory}
+In the perfect payoff memory case an individual updates based on the average payoff
+against each other member of the population, otherwise known as expected payoffs. 
+The payoff of a reactive strategy \(s_1\) against the reactive strategy \(s_2\)
+in an infinitely repeated game ($\delta \rightarrow 1$) is explicitly
+calculated as,
 
-In the perfect memory case an individual updates based on the average payoff
-against each other member of the population, otherwise as expected payoffs. In
-an infinitely repeated game ($\delta \rightarrow 1$) the payoff of a reactive
-strategy \(s_1\) against the reactive strategy \(s_2\) can explicitly be
-calculated using the stationary vector \(\mathbf{v}\) and the payoffs' vector as
-\(\langle\mathbf{v}(s_1,s_2),\mathbf{u}\rangle\). In a population of size \(N\)
-there are \(k\) mutants and \(N - k\) residents. We denote the strategies of
-a mutant and of a resident as \(s_M =(y_M, p_M, q_M)\) and \(s_R =
-(y_R, p_R, q_R)\). The expected payoffs of a resident (\(\pi_R\)) and of a
-mutant (\(\pi_M\)) are given by,
+\[\langle\mathbf{v}(s_1,s_2),\mathbf{u}\rangle.\]
+
+In a population of size \(N\) there are \(k\) mutants and \(N - k\) residents.
+Let \(s_M =(y_M, p_M, q_M)\) and \(s_R = (y_R, p_R, q_R)\) denote the
+strategies of a mutant and a resident, the expected payoffs \(\pi_R\) and
+\(\pi_M\) are given by,
 
@@
 The probabilities that the number of mutants decreases and increases,
-\(\lambda^-_k\) and \(\lambda^+_k\), in the perfect memory case are defined as,
+\(\lambda^-_k\) and \(\lambda^+_k\), in the perfect payoff memory case are
+defined as,
 
@@
 
-\subsubsection*{Limited Memory}
-
-In this case of limited memory we initially define the probability that a
+In the case of limited payoff memory we initially define the probability that a
 reactive strategy receives the payoff $u\!\in\! \mathcal{U}$ in the very last
@@
   v_{r}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(q_1+l_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+l_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
-  {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)} \times r,\\[1cm]
+  {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)},\\[1cm]
 
   v_{s}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(q_1+l_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
-  {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)} \times s,\\[1cm]
+  {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)},\\[1cm]
 
   v_{t}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+l_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
-  {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)} \times t,\\[1cm]
+  {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)},\\[1cm]
 
   v_{p}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
-  {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)} \times p.
+  {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)}.
   \end{array}
@@
 At each time step there are five possible pairings. They interact with each
-other with a probability \(\frac{1}{N - 1}\), and thus they do not interact with
+other with a probability \(\frac{1}{N - 1}\), and they do not interact with
 other with a probability \(1 - \frac{1}{N - 1}\). In the latter case, each of
@@
 the other interacts with a mutant, and this happens with a probability
-$\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$. Thus we the probability that the randomly
+$\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$. We define the probability that the randomly
 chosen resident obtained a payoff of $u_R$ in the last round of his respective
@@
 \begin{array}{llrl}
-x(u_1,u_2)	 &=&\displaystyle \frac{1}{N\!-\!1}\cdot  &v_{u_1}(s_1,s_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F}\\[0.5cm]
+x(u_R,u_M)	 &=&\displaystyle \frac{1}{N\!-\!1}\cdot  &v_{u_R}(s_R,s_M)\cdot 1_{(u_R,u_M)\in \mathcal{U}^2_F}\\[0.5cm]
 &+	
 &\displaystyle \left(1\!-\!\frac{1}{N\!-\!1}\right)  
-&\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_2) + 
- \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_1)\right.\\[0.5cm]
-&&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_2) + 
- \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_1)\right].
+&\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} v_{u_R}(s_R,s_M) v_{u_M}(s_M,s_M) + 
+ \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} v_{u_R}(s_R,s_M) v_{u_M}(s_M,s_R)\right.\\[0.5cm]
+&&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} v_{u_R}(s_R,s_R) v_{u_M}(s_M,s_M) + 
+ \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} v_{u_R}(s_R,s_R) v_{u_M}(s_M,s_R)\right].
 \end{array}
@@
 the role model happened to be matched during the game stage, which happens with
-probability $\frac{1}{(N\!-\!1)}$. The probability that the number of mutants
-increases and decreases by one in the case of limited memory are now given by,
+probability $\frac{1}{(N\!-\!1)}$.
+
+The probability that the number of mutants increases and decreases by one in the
+case of limited payoff memory are now given by,
 
 \begin{align}\label{eq:limited_memory_lambdas}
-\lambda^+_k=\frac{N\!-\!k}{N} \frac{k}{N} \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M}) \rho(u_{R},u_{M}) \quad \text{and} \quad
-\lambda^-_k=\frac{N\!-\!k}{N} \frac{k}{N} \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M}) \rho(u_{M},u_{R}).
+\lambda^+_k=\frac{N\!-\!k}{N} \cdot \frac{k}{N} \cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M}) \rho(u_{R},u_{M}) \quad \text{and} \quad
+\lambda^-_k=\frac{N\!-\!k}{N} \cdot \frac{k}{N} \cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M}) \rho(u_{M},u_{R}).
 \end{align}
@@
 
-\subsubsection*{Simulation Results}
+\section{Simulation Results}
 

### 2023-02-28 paper/main.tex
delete old comments
finish conclusion

@@
 evolutionary process and record which strategies the players adopt over time
-based on perfect memory payoffs and limited memory payoffs. In the case of
-perfect memory payoffs we use Eq~\ref{eq:perfect_memory_lambdas} when we
-estimate the probability that a mutant fixes, and
-Eq~\ref{eq:limited_memory_lambdas} in the case on limited memory. We run two
-independent runs for each approach and we vary the value of benefit \(b\).
-Figure~\ref{fig:expected_and_stochastic_for_donation} shows the simulation
-results. It depicts the evolving conditional cooperation probabilities $p$ and
-$q$. The discount factor~$\delta$ is comparably high, thus we do not report the
-opening move \(y\) as it is a transient effect. The upper panel corresponds to
-the standard scenario considered in the literature, it considers players who use
-expected payoffs to update their strategies. The bottom panel shows the scenario
-considered herein, in which players update their strategies based on their last
-round's payoff.
+based on perfect and limited payoff memory. We performed two independent runs
+for each approach varying the value of benefit \(b\).
+Figure~\ref{fig:expected_and_stochastic_for_donation} depicts the evolving
+conditional cooperation probabilities $p$ and $q$ (since the discount
+factor~$\delta$ is comparably high we do not report the opening move \(y\)). The
+upper panel of Figure~\ref{fig:expected_and_stochastic_for_donation} corresponds
+to the standard scenario considered in the literature, and the bottom panel
+shows the scenario considered herein, in which players update their strategies
+based on their last round's payoff.
 
-The figure suggests that when updating is based on perfect memory payoffs
-players tend to be more generous and more cooperative. The $q$-values of the
-resident strategies are on average higher in the case of perfect memory. Thus,
-players will occasionally forgive a defection more often if their fitness
-depends on interacting with every member of the population, and this effect
-becomes more notable as the value of the benefit increases. In the case of
-limited memory the resident strategies are less forgiving and their forgiveness
-is independent of the values of benefit. In the Supplementary Information
-Sections 2.1 and 2.2.1 we show that $q$-values of a resident strategy have to be less
-than \(1 - c/b\) in the perfect memory case and less than \(0.5\) in the limited
-memory case. The generosity for the perfect memory case is always higher for a
-benefit value higher than 2, and it increases as the benefit value increases.
-This leads to a more cooperative population. We calculate the average
-cooperation rate for each simulation which is the average cooperation rate
-within the resident population. In the case of the perfect memory payoffs the
-average cooperation rate is strictly higher than that of the last round payoffs,
-and the difference is most for $b=10$. The average cooperation of resident
-strategies drops from 97\% to 57\%. This indicates that the expected payoffs
-overestimate the evolved cooperation.
+The figure suggests that when updating is based on perfect payoff memory players
+tend to be more generous and more cooperative. The $q$-values of the resident
+strategies are on average higher in this case, thus, players will occasionally
+forgive a defection more often if their fitness depends on interacting with
+every member of the population. This effect becomes more notable as the value of
+the benefit increases. In the case of limited memory the resident strategies are
+less forgiving and their forgiveness regardless of the values of benefit. In
+analysis carried out in the Supplementary Information Section 3 we show that a
+resident population of conditional cooperators with \(q > \frac{1}{2}\) can not
+avoid invasion by a defecting strategy. This holds independently of the cost and
+benefit values.
+
+Higher $q$ values lead to a more cooperative population. We calculate the
+average cooperation rate for each simulation which is the average cooperation
+rate within the resident population. In the case of the perfect payoff memory 
+the average cooperation rate is strictly higher than that of the last round
+payoffs. The difference between the cooperation rate is the highest at $b=10$.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{static/donation_expected_last_round_summary_results.pdf}
+    \includegraphics[width=.9\textwidth]{static/donation_expected_last_round_summary_results.pdf}
     \caption{{\bf Evolutionary dynamics under perfect and limited payoff memory.}
@@
     illustrations of the perfect memory and the limited memory cases. The shaded
-    background denotes the game phase information that an individual considered
-    when updating strategies. In the case of perfect memory the entire region is
-    shaded and in the case of limited memory only one turn with a single member
+    background denotes the game phase information that an individual considers
+    when updating strategies. In the case of perfect payoff memory the entire region is
+    shaded and in the case of limited payoff memory only one turn with a single member
     of the population. ({\bf Simulations}) We have run four simulations of the
-    pairwise comparison process for $T\!=\!10^7$ time steps. For each time step,
-    we have recorded the current resident population ($y,p,q$). Since
+    pairwise comparison process for $T\!=\!10^7$ time steps. For each time step
+    of the process
+    we record the current resident population ($y,p,q$). Since
     simulations are run for a relatively high continuation probability of
     $\delta\!=\!0.999$, we do not report the players' initial cooperation
-    probability $y$. The graphs show how often the resident population chooses
+    probability $y$. The plots show how often the resident population chooses
     each combination ($p,q$) of conditional cooperation probabilities in the
@@
     player. The resident population applies a conditional cooperator strategy
-    for which $q\!\le\!1\!-\!c/b\!=\!0.67$. This is true also for the case of
-    high benefit. In this case the population mainly consists of conditional
+    for which $q\!\le\!1\!-\!c/b\!=\!0.67$. 
+    In the case of high benefit the population mainly consists of conditional
     cooperators of the form ($p\!\approx\!1\!, q\!\le\!1\!-\!1/10\!=\!0.9$). In
-    the Supplementary Information Section 2.1 we show that a conditional
+    the Supplementary Information Section 2 we show that a conditional
     cooperator needs to be of the form ($p\!\approx\!1\!, q\!\le\!1\!-\!c/b$) to
@@
     population results in a higher average cooperation rate. The average
-    cooperation rate increases from 52\% for \(b\) to 98\% for \(b=10\). ({\bf
+    cooperation rate increases from 52\% for \(b=3\) to 98\% for \(b=10\). ({\bf
     Limited Memory}) When players update their strategies based on their
@@
     consists of defectors (with $p\!\approx\!q\!\approx\!0$) or of conditional
-    cooperators. The maximum level of $q$ consistent with stable cooperation is
-    somewhat smaller compared to the perfect memory setting, $q\!<\!0.5$.
-    Namely, in the Supplementary Information Section 2.2.1 we show that
-    regardless of the value of benefit, in the case of limited payoff memory, an
-    individual needs a generosity smaller of $0.5$ to repel defectors. The plots
-    are fairly similar regardless of the benefit, and the evolved cooperation
-    rate only slightly increases from 37\% to 51\%. Parameters: $N\!=\!100$,
+    cooperators. The maximum level of $q$, consistent with stable cooperation, is
+    somewhat smaller compared to the perfect memory setting.
+    Namely, in the Supplementary Information Section 3 we show that
+    regardless of the value of benefit a conditional cooperator need to be of
+    the form $q\!<\!\frac{1}{2}$ to not be invaded by defectors.
+    The evolved cooperation
+    rate only slightly increases from 37\% ($b=3$) to 51\% ($b=10$). Parameters: $N\!=\!100$,
     $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.999$.}
@@
 We further explore the effects of the cooperation benefit and the strength of
-selection on the evolved generosity \(q\) and cooperation rate
+selection on the evolved generosity \(q\) and on the cooperation rate
 (Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}).
 Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}\textbf{A} suggests that
-perfect memory always yields a higher cooperation rate. We observe that the
-cooperation rate increases as the value of the benefit gets higher, whereas in
-comparison for the limited memory payoffs, the cooperation rate remains
-unchanged (as we would have expected) at approximately 50\% once \(b=5\). From
-Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}\textbf{B} we observe
-that for weak selection, \(\beta < 1\), the two methods yield similar results,
-however, as \(\beta\) increases there is variation in the evolving populations.
-In the case of expected payoffs the resident populations become more cooperative
-as \(\beta\) increases, whereas in the case of limited memory payoffs, the
-resident populations become more defective.
+perfect memory always yields a higher cooperation rate. In this case the
+cooperation rate increases as the benefit increases. In comparison for limited
+payoff memory the cooperation rate remains unchanged at approximately 50\% once
+\(b=5\).  From Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}\textbf{B}
+we observe that for weak selection, \(\beta < 1\), the two methods yield similar
+results, however, as \(\beta\) increases there is variation in the evolving
+populations. In the case of expected payoffs the resident populations become
+more cooperative, whereas in the case of limited payoff memory, the resident
+populations become more defective.
 
@@
 
-\subsection{Updating Payoffs based on More Memory}
-
-So far we have explored the difference between the expected payoffs and the last
-round payoffs. To explore further the effect of limited memory we consider three
-more cases. These are the cases where the updating payoffs depend on  the last
-round with two members of the population, the last two rounds with another
-member of the population, and the last two rounds with two members of the
-population. Similar to the last round payoff. The details of each approach can
-be found in the Supplementary Information. We use numerical simulation of the
-pairwise process for the various schemes. The results for low and high benefit
-are shown in Figure~\ref{fig:cooperation_rate_all_updating_payoffs}.
+Up to this point we have discussed only two edge cases. The framework of limited
+payoff memory can be easily extended to allow individuals to observe more
+rounds, to interact with more members or both (see Supplementary Information
+Sections 4-6). To further understand the effect of limited payoff memory we
+consider the cases of updating based on the last round with two members of the
+population, the last two rounds with another member of the population, and the
+last two rounds with two members of the population. Similar to the last round
+payoff. We perform numerical simulations for each of the fitness approaches. In
+Figure~\ref{fig:cooperation_rate_all_updating_payoffs} we report the cooperation
+rates for low and high values of benefit.
 
@@
   From left to right, we present result on the following updating payoffs cases;
-  (1) the expected payoffs (perfect memory), (2) the last round payoff from one
-  interaction (limited memory), (3) the last round payoff from two interactions,
-  (4) the last two rounds payoffs from one interaction, (5) the last two rounds
-  payoffs from two interactions. The details of each case are given in the
-  Supplementary Information. Cases 3 and 4 have similar results. In that the
-  evolved generosity is the same. For cases 2 - 5 the cost benefit ratio
-  has an effect compared to the limited memory case. The cooperation rate
-  increases, and as we allow for more information the closer we move to the
-  perfect memory. Unless explicitly varied, the parameters of the simulation are
+  (a) the expected payoffs (perfect memory), (b) the last round payoff from one
+  interaction (limited memory), (c) the last round payoff from two interactions,
+  (d) the last two rounds payoffs from one interaction, (e) the last two rounds
+  payoffs from two interactions. For the updating payoffs (c) and (d) we have
+  carried out an analysis which has shown that conditional cooperators should
+  adopt a \(q\) value smaller than $\frac{\delta - 1 +
+  \frac{\sqrt{2}}{2}}{\delta}$ and $\frac{\delta + \sqrt{\delta^{2} + 1} - 1}{2 \delta}$
+  respectively. Note that as \(\delta
+  \rightarrow 1\) both right hand sides tend to \(\frac{\sqrt{2}}{2}\).
+  Regardless the cooperation rate between for case (c) is slightly higher.
+  In case (d) the cooperation rate is the second highest hinting that as we
+  allow for more information the closer we move to the perfect payoff memory.
+  Unless explicitly varied, the parameters of the simulation are
   $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$.
@@
 
-\newpage
-
 \section{Conclusions}\label{section:conclusions}
 
-A major focus of evolutionary game theory is the evolution of cooperation among
-selfish individuals. Sometimes evolutionary game theory even seems to be reduced
-to the evolution of cooperation [Arne2022]. “Why do we act altruistically to
-provide a benefit to others when there is a cost to ourselves?” is still a
-puzzling question amongst researchers [Glynatsi]. In this determinist
-formulation, a small fraction of cooperative individuals that have a lower
-fitness can never invade a population of defecting strategies.  The stochastic
-effect can lead to fundamental changes since disadvantageous mutants have a
-small yet non-zero probability to reach fixation in a finite population, and
-thus cooperation can emerge.
-
-Cooperation can be seen at odd, why is it that we choose to help others,
-increasing their payoff, at the expense of decreasing one's own? In spite of all
-the selfish genes' animal and human communities seem to altruistically help each
-other and cooperate, and evolutionary game theory has helped us shape our
-understanding of the evolution of cooperation.
-
-Previous evolutionary models often feature a curious inconsistency. When
-modeling how individuals make decisions in each round, these models assume that
-players only remember the last round. However, when modeling how individuals
-update their strategies over time, individuals are assumed to have perfect
-memory.
+Cooperation can be seen at odd, why is it that we choose to help others at a
+personal cost? In spite of all the selfish genes', animal and human communities
+show signs of altruism and cooperation~\cite{milinski1987tit, kerr2002local,
+carter2020development}. Evolutionary game theoretical models have helped us
+shape our understanding of the evolution of cooperation. In fact, the evolution
+of cooperation constitutes such a major focus of the field that evolutionary
+game theory seems to be reduced to the evolution of
+cooperation~\cite{Traulsen:PhilTrans:2022}. Previous evolutionary models often
+feature a curious inconsistency. When modeling how individuals make decisions in
+each round, these models assume that players only remember the last round.
+However, when modeling how individuals update their strategies over time,
+individuals are assumed to have perfect memory.
 
 Here, we have explored how robust cooperation is as models deviate from the
-perfect memory assumption. Initially we considered the donation game. We showed
+perfect memory assumption. We considered a special case of the prisoner's
+dilemma, the donation game. The prisoner's dilemma is one of the most applied
+types of social games when we discuss the evolution of cooperation. We showed
 that when the last round payoff is used instead of the expected payoffs,
-cooperation can even evolve when individuals only use a minimum of information,
-however, the evolving cooperation rates are typically lower. The resident
-strategies were both less cooperative and less generous. This effect was only
-intensified we increase the benefit and the strength of selection independently.
-The results showed that as each parameter was being increased the difference
-between the cooperation rates were widening. This indicates that cooperative
-players benefit from being able to interact with everyone in the population.
+cooperation struggles to evolve  when only a minimum of social information is
+used at the updating stage. The resident strategies were both less cooperative
+and less generous. This effect was only intensified we increase the benefit and
+the strength of selection independently. The results showed that as each
+parameter was being increased the difference between the cooperation rates were
+widening. This indicates that cooperative players benefit from being able to
+interact with everyone in the population.
 
-We extended our approach and presented results not only based on different
-classes of social games, but also based on different limited memory payoffs. The
-analysis showed that specifically for the case of the prisoner's dilemma games
-the contrast between the two approaches can be significantly large. The
-prisoner's dilemma is one of the most applied types of social games when we
-discuss the evolution of cooperation. Our results indicate that cooperation
-struggles to evolve in the prisoner's dilemma when only a minimum of social
-information is used at the updating stage.
+We extended our approach and presented results on different limited memory
+payoffs. The analysis showed cooperation can even evolve when we allow
+individual to consider more information.
 
@@
 
-% \appendix
-
-% \section{Model Setup}\label{appendix:methods}
-
-% Consider a population of \(N\) individuals where \(N\) is even. At any point in
-% time there are at most two different strategies in present in the population.
-% More specifically, a mutant strategy played by \(k\) individuals and a resident
-% strategy played by \(N - k\) individuals. We assume a pairwise process in which
-% strategies spread because they are imitated more often. Each step of the
-% evolutionary process consists of two stages; a game stage and an update stage.
-
-% In the game stage, each individual is randomly matched with some other
-% individual in the population. Their interaction lasts for a number of turns
-% which is not fixed but depends on the continuation probability \(\delta\). At
-% each turn the individuals choose between cooperation (\(C\)) and defection
-% (\(D\)). Thus, there are four possible outcomes in each turn \(CC, CD, DC\) and
-% \(DD\). If both players cooperate they receive the reward payoff \(R\), whereas
-% if both players defect they receive the punishment payoff \(P\). If one
-% cooperates but the other defects, the defector receives the temptation to
-% defect, \(T\), whereas the cooperator receives the sucker's payoff, \(S\). Let
-% $\mathcal{U}=\{R,S,T,P\}$ denote the set of feasible payoffs in each round, and
-% let $\mathbf{u}\!=\!(R,S,T,P)$ be the corresponding payoff vector. We present
-% results for various values of $\mathcal{U}$ for all the symmetric \(2 \times 2\)
-% games.
-
-% A further assumption of our model is that individuals make use of reactive
-% strategies when they make decisions in each round. Reactive strategies are a set
-% of strategies that take into account only the previous action of the opponent. A
-% reactive strategy can be written explicitly as a vector,
-
-% \[s=(y, p, q)\]
-
-% where \(y\) is the probability that the strategy opens with a cooperation and
-% \(p, q\) are the probabilities that the strategy cooperates given that the
-% opponent cooperated and defected equivalently.
-
-% In the updating stage, two players are randomly drawn from the population, a
-% `learner' and a `role model'. The learner adopts the role model's strategy based
-% on the Fermi distribution function, %ToDo add reference
-
-% % \begin{equation} \label{Eq:rho}
-% % \rho(u_{L}, u_{RM}) = \frac{1}{1\!+\! \exp^{\!-\!\beta (u_{RM}\!-\!u_{L})}}.
-% % \end{equation}
-
-% where $u_{L}\!\in\! \mathcal{U}$ is the learner's payoff, $u_{RM}\!\in\!
-% \mathcal{U}$ is the role model's payoff, and $\beta\!\ge\!0$ is the strength of
-% selection.
-
-% We iterate this basic evolutionary step until either the mutant strategy goes
-% extinct, or until it fixes in the population and becomes the new resident
-% strategy. After either outcome, we set $k$ to $1$ and we introduce a new mutant
-% strategy which is uniformly chosen from all reactive strategies at random.
-% Instead of simulating each step of the evolutionary process, we estimate the
-% probability that a newly introduced mutant fixes~\cite{nowak2004emergence}. This
-% is defined as the fixation probability of the mutant, and the standard form is
-% the following,
-
-% \begin{equation}\label{eq:appendix_fixation_probability}
-% \varphi = \frac{1}{1+\sum\limits_{i=1}^{N-1}\prod\limits_k^i \frac{\lambda^-_k}{\lambda^+_k}},
-% \end{equation}
-
-% where \(\lambda^-_k, \lambda^+_k\) are the probabilities that the number of
-% mutants decreases and increases respectively.
-
-% This process of mutation and fixation/extinction is iterated many times. The
-% evolutionary process is summarized by
-% Algorithm~\ref{algorithm:pairwise_comparison}.
-
-% \begin{algorithm}[!htbp]
-%   \SetAlgoLined $N \leftarrow$ population size\; $k \leftarrow 1$\; resident
-%    $\leftarrow (0, 0, 0)$\; \While{ $t <$ maximum number of steps}{mutant
-%    $\leftarrow$ random: $\{\emptyset \}\ \rightarrow R^3$\; fixation probability
-%    $\leftarrow \varphi $\; \If{$\varphi >$ random: $i \rightarrow [0,
-%    1]$}{resident $\leftarrow$ mutant;}} \caption{Evolutionary
-%    process}\label{algorithm:pairwise_comparison}
-% \end{algorithm}
-
-% The aim of this work is to explore the effect of updating memory on the
-% cooperation rate of the evolved population. For this reason we consider two
-% different approaches when estimating the payoffs at the updating stage. The two
-% approaches we consider are those of (i) the expected and (ii) the limited memory
-% payoffs.
-
-% \subsection*{Expected Payoffs}
-
-% The expected payoffs are the conventional payoffs used in the updating
-% stage~\cite{imhof2010stochastic}. They are defined as the mean payoff of an
-% individual in a well-mixed population that engages in repeated games with all
-% other population members.
-
-% We first define the payoff of two reactive strategies at the game stage. Assume
-% two reactive strategies $s_1\!=\!(y_1, p_1, q_1$) and $s_2\!=\!(y_2,p_2,q_2)$.
-% It is not necessary to simulate the play move by move, instead the play between
-% the two strategies is defined a Markov matrix \(M\),
-
-% \begin{equation}\label{eq:transition_matrix}
-%   \input{static/matrix.tex}.
-% \end{equation}
-
-% whose stationary vector \(\mathbf{v}\), combined with the payoff \(u\), yields
-% the game stage outcome for each strategy,
-% \(\langle\mathbf{v}(s_1,s_2),\mathbf{u}\rangle\)~\cite{Hauert1997}.
-
-% In the updating stage the learner adopts the strategy of the role model based on
-% their updating payoffs. Given that there are only two different types in the
-% population at each time step we only need to define the expected payoff for a
-% resident (\(\pi_R\)) and for a mutant (\(\pi_M\)). Assume the resident strategy
-% \(s_R = (y_R, p_R, q_R)\) and the mutant strategy \(s_M = (y_M, p_M, q_M)\), the
-% expected payoffs are give by,
-
-% \begin{equation} \label{Eq:ExpPay}
-%   \begin{array}{lcrcr}
-%   \displaystyle \pi_R	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(s_R,s_R),\mathbf{u}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(s_R,s_M),\mathbf{u}\rangle,\\[0.5cm]
-%   \displaystyle \pi_M	&=	&\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(s_M,s_R),\mathbf{u}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(s_M,s_M),\mathbf{u}\rangle.\\
-%   \end{array}
-% \end{equation}
-
-% The number of mutant in the population increase if a learner resident adopts the
-% strategy of a mutant role model, and decreases if a mutant leaner adopts the
-% strategy of a resident. The probabilities that the number of mutants decreases
-% and increases, \(\lambda^-_k\) and \(\lambda^+_k\), are now explicitly defined
-% as,
-
-% \begin{align*} 
-%   \lambda^-_k &\!=\!\rho(\pi_R, \pi_M) \\
-%   \lambda^+_k &\!=\!\rho(\pi_M, \pi_R).
-% \end{align*}
-
-% \subsection*{Limited memory payoffs}
-
-% Initially, we discuss the case of the \textbf{last round updating payoff}. At
-% the stage game we define the payoff of a reactive strategy in the last round,
-% Proposition~\ref{proposition:last_round}.
-
-% \begin{Prop}\label{proposition:last_round} Consider a repeated game, with
-%     continuation probability $\delta$, between players with reactive strategies
-%     $s_1\!=\!(y_1, p_1, q_1$)  and $s_2\!=\!(y_2,p_2,q_2)$ respectively. Then
-%     the probability that the $s_1$ player receives the payoff $u\!\in\!
-%     \mathcal{U}$ in the very last round of the game is given by
-%     $v_{u}(s_1,s_2)$, as given by Equation~(\ref{Eq:LastRound}).
-
-%     \begin{equation} \label{Eq:LastRound}
-%       \setlength{\arraycolsep}{1pt}
-%       \begin{array}{rcl}
-
-%       v_{R}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
-%       {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)} \times R,\\[1cm]
-
-%       v_{S}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
-%       {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)} \times S,\\[1cm]
-
-%       v_{T}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
-%       {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)} \times T,\\[1cm]
-
-%       v_{P}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
-%       {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)} \times P.
-%       \end{array}
-%     \end{equation}
-
-% In these expressions, we have used the notation $r_i:=p_i\!-\!q_i$,
-% $\bar{y}_i\!=\!1\!-\!y_i$, $\bar{q}_i:=1\!-\!q_i$, and
-% $\bar{r}_i:=\bar{p}_i\!-\!\bar{q}_i=-r_i$ for $i\!\in\!\{1,2\}$.
-% \end{Prop}
-
-% \begin{proof}
-% Given a play between two reactive strategies with continuation probability
-% $\delta$. The outcome at turn \(t\) is given by,
-
-% \begin{equation}\label{eq:}
-%   (1 - \delta) \mathbf{v_0} \sum \delta^{t} M^{(t)},
-% \end{equation}
-
-% where $\mathbf{v_0}$ denotes the expected distribution of the four outcomes in
-% the very first round, and \(1- \delta\) the probability that the game ends. It
-% can be shown that,
-
-% \begin{align*}
-%   (1 - \delta) \mathbf{v_0} \sum \delta^{t} M^{(t)} & = (1 - \delta)(\mathbf{v_0} + \delta \mathbf{v_0} M + \delta^{2}\mathbf{v_0} M ^{2} + \dots )\\ 
-%    & = (1 - \delta)\mathbf{v_0} (1 + \delta M + \delta^{2}M ^{2} + \dots ) \text{ using standard formula for geometric series}\\ 
-%    & = (1 - \delta)\mathbf{v_0}(I_4 - \delta M)^{-1}
-% \end{align*}
-
-% where \((1 - \delta)\mathbf{v_0}(I_4 - \delta M)^{-1}\) is vector \(\in R^{4}\)
-% and it the probabilities for being in any of the outcomes \(CC, CD, DC, DD\) in
-% the last round. Combining this with the payoff vector \(u\) and some algebraic
-% manipulation we derive to the Equation~\ref{Eq:LastRound}.
-% \end{proof}
-
-% In the updating stage we select a mutant and resident to be either the role
-% model or the learner. Given that they can interact with only one other member of
-% the population, they can interact either with each other or either can interact
-% with another resident or with another mutant. Thus, in each updating stage there
-% are five possible combinations of pairs (Figure~\ref{fig:single_pairs}).
-
-% Given the last round payoff and possible pair combinations for a single
-% interaction, we define the probability that the respective last round payoffs of
-% two players \(s_1, s_2\) are given by $u_1$ and $u_2$ as,
-
-% \begin{equation}\label{eq:Chi}
-% \setlength{\arraycolsep}{1pt}
-% \begin{array}{llrl}
-% x(u_1,u_2)	 &=&\displaystyle \frac{1}{N\!-\!1}\cdot  &v_{u_1}(s_1,s_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F}\\[0.5cm]
-% &+	
-% &\displaystyle \left(1\!-\!\frac{1}{N\!-\!1}\right)  
-% &\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_2) + 
-%  \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_1)\right.\\[0.5cm]
-% &&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_2) + 
-%  \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_1)\right].
-% \end{array}
-% \end{equation}
-
-% The first term on the right side corresponds to the case that the learner and
-% the role model happened to be matched during the game stage, which happens with
-% probability $\frac{1}{(N\!-\!1)}$. In that case, we note that only those payoff
-% pairs can occur that are feasible in a direct interaction,
-% $\splitatcommas{(u_1,u_2)\in \mathcal{U}^2_F:=\big\{ (R,R), (S,T), (T,S), (P,P)
-% \big\}}$, as represented by the respective indicator function. Otherwise, if the
-% learner and the role model did not interact directly, we need to distinguish
-% four different cases, depending on whether the learner was matched with a
-% resident or a mutant, and depending on whether the role model was matched with a
-% resident or a mutant.
-
-% Given that $N\!-\!k$ players use the resident strategy
-% $s_{R}\!=\!(y_{R},p_{R},q_{R})$ and that the remaining $k$ players use the
-% mutant strategy $s_{M}\!=\!(y_{M},p_{M},q_{M})$, the probability that the number
-% of mutants increases by one in one step of the evolutionary process can be
-% written as
-
-% \begin{align}
-% \lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M})\cdot \rho(u_{R},u_{M}), \\
-% \lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M})\cdot \rho(u_{M},u_{R}).
-% \end{align}
-
-% In this expression, $\frac{(N\!-\!k)}{N}$ is the probability that the randomly
-% chosen learner is a resident, and $\frac{k}{N}$ is the probability that the role
-% model is a mutant. The sum corresponds to the total probability that the learner
-% adopts the role model's strategy over all possible payoffs $u_R$ and $u_M$ that
-% the two player may have received in their respective last rounds. We use
-% $x(u_R,u_M)$ to denote the probability that the randomly chosen resident
-% obtained a payoff of $u_R$ in the last round of his respective game, and that
-% the mutant obtained a payoff of $u_M$.
-
-% We extend our framework to consider the case where players update their
-% strategies based on the outcome of \textbf{the last two turns and based on their
-% interaction with two other members of the population}. At the stage game we
-% define the payoff of a reactive strategy in the last two rounds,
-% Proposition~\ref{proposition:last_two_rounds}.
-
-% \begin{Prop}\label{proposition:last_two_rounds} Consider a repeated game, with
-%   continuation probability $\delta$, between players with reactive strategies
-%   $s_1\!=\!(y_1, p_1, q_1$)  and $s_2\!=\!(y_2,p_2,q_2)$ respectively. Let
-%   $\mathcal{\tilde{U}} = \splitatcommas{\{RR, RS, RT, RP, SR, SS, ST, SP, TR,
-%   TS, TT, TP, PR, PS, PT, PP\}}$ denote the set of feasible payoffs in the last
-%   two rounds, and let \(\tilde{\mathbf{u}}\) be the corresponding payoff vector.
-%   Then the probability that the $s_1$ player receives the payoff $u\!\in\!
-%   \mathcal{\tilde{U}}$ in the very last two rounds of the game is given by,
-
-%   \begin{equation}
-%   \langle\mathbf{\tilde{v}}(s_1,s_2),\mathbf{\tilde{u}}\rangle, \text{ where } \mathbf{\tilde{v}} \in R^{16} \text{ is given by },
-%   \end{equation}
-%   \begin{equation}
-%     \mathbf{\tilde{v}}(s_1,s_2) = (1 - \delta) m_{a_1, a_2} \delta^2 \left[\mathbf{v_0}(I_4 - \delta M)^{-1}\right]_{a_1, a_2}, \quad  m_{a_1, a_2} \in M \ \forall \ a_1, a_2 \in \{1, 2, 3, 4\}.
-%   \end{equation}
-% \end{Prop}
-
-% In the updating stage we select a mutant and resident to be either the role
-% model or the learner. Given that they can interact with two other members of the
-% population there are a total of twenty four possible combinations of pairs
-% (Figure~\ref{fig:pissible_two_pairs}).
-
-% Given the last two rounds payoff and possible pair combinations with two
-% members, we can define the probability \(x\) that the respective last round
-% payoffs of two players \(s_1, s_2\) are given by $u_1$ and $u_2$ similarly to
-% Eq.~(\ref{eq:Chi}). We follow the same approach to define the rest of the
-% updating payoffs. These are the updating payoffs of the last round with two
-% member, and the the last two rounds payoffs with one member.
-
-% Simulating the evolutionary process for more interactions and rounds quickly
-% becomes computationally intractable. Our methodology could be extended to
-% include \(n\) turns and \(m\) interactions. However, for the purpose of this
-% work we explore the cases only up to two turns and two interactions.
-
-% \section{Evolutionary dynamics under limited memory payoffs for various social
-% dilemmas}\label{appendix:further_simulation_results}
-
-% \begin{figure}[!htbp]
-%   \centering
-%   \includegraphics[width=\textwidth]{static/rounds_two_by_two_games.pdf}
-%   \caption{{\bf Evolutionary dynamics under last two rounds payoffs for various social dilemmas.} 
-%   We have run several simulations of the evolutionary process described in
-%   section~\ref{section:model} for $t\!=\!10^7$ time steps. The graphs show how
-%   often the resident population chooses each combination ($p,q$) of conditional
-%   cooperation probabilities in the subsequent rounds. We vary the temptation
-%   payoff \(T \in \{-1, -0.6, -0.2,  0.2, 0.6, 1, 1.4, 1.8, 2.2, 2.6, 3\}\)
-%   across the \(x\) axis, and  \(S \in \{2, 1.6, 1.2, 0.8, 0.4, 0, -0.4, -0.8,
-%   -1.2, -1.6, -2\}\) across the \(y\) axis. Parameters: $N\!=\!100$,
-%   $\beta\!=\!10$, $\delta\!=\!0.999$.}
-%   \label{fig:last_two_rounds_two_by_two}
-% \end{figure}
-
-% \begin{figure}[!htbp]
-%   \centering
-%   \includegraphics[width=\textwidth]{static/opponents_two_by_two_games.pdf}
-%   \caption{{\bf Evolutionary dynamics under last round payoffs with two members of the population for various social dilemmas.} 
-%   We have run several simulations of the evolutionary process described in
-%   section~\ref{section:model} for $t\!=\!10^7$ time steps. The graphs show how
-%   often the resident population chooses each combination ($p,q$) of conditional
-%   cooperation probabilities in the subsequent rounds. We vary the temptation
-%   payoff \(T \in \{-1, -0.6, -0.2,  0.2, 0.6, 1, 1.4, 1.8, 2.2, 2.6, 3\}\)
-%   across the \(x\) axis, and  \(S \in \{2, 1.6, 1.2, 0.8, 0.4, 0, -0.4, -0.8,
-%   -1.2, -1.6, -2\}\) across the \(y\) axis. Parameters: $N\!=\!100$,
-%   $\beta\!=\!10$, $\delta\!=\!0.999$.}
-%   \label{fig:last_two_opponents_two_by_two}
-% \end{figure}
-
-% \begin{figure}[!htbp]
-%   \centering
-%   \includegraphics[width=\textwidth]{static/merged_plot_rounds_opponents_two.pdf}
-%   \caption{{\bf Evolutionary dynamics under last two rounds payoffs with two members of the population for various social dilemmas.} 
-%   We have run several simulations of the evolutionary process described in
-%   section~\ref{section:model} for $t\!=\!10^7$ time steps. The graphs show how
-%   often the resident population chooses each combination ($p,q$) of conditional
-%   cooperation probabilities in the subsequent rounds. We vary the temptation
-%   payoff \(T \in \{-1, -0.6, -0.2,  0.2, 0.6, 1, 1.4, 1.8, 2.2, 2.6, 3\}\)
-%   across the \(x\) axis, and  \(S \in \{2, 1.6, 1.2, 0.8, 0.4, 0, -0.4, -0.8,
-%   -1.2, -1.6, -2\}\) across the \(y\) axis. Parameters: $N\!=\!100$,
-%   $\beta\!=\!10$, $\delta\!=\!0.999$.}
-%   \label{fig:last_two_rounds_two_opponents}
-% \end{figure}
-
 \bibliographystyle{unsrt}

### 2023-03-01 paper/si.tex
si fixes

@@
 update phase an individual adopts the strategy of another individual based on
-their ``updating payoffs''. In the paper we explore the effect of the updating
-payoffs on the evolved populations. We used as an example the prisoner's
-dilemma. We focused mainly on two approaches for calculating the updating
-payoffs which we referred to as the perfect memory and the limited memory
-approaches. In Section~\ref{section:perfect_memory} we describe the perfect
-memory approach and in Section~\ref{section:limited_memory} we present the
-limited memory approach. In the limited memory approach we assume that
+their ``updating payoffs''. Here we study different methods for calculating
+updating payoffs and we compare their results using as an example the prisoner's
+dilemma. In Section~\ref{section:perfect_memory} we describe the conventional
+method used in the literature and in Section~\ref{section:limited_memory} we
+present a new method which we refer to as 
+limited payoff memory. In the limited payoff memory method we assume that
 individuals update based on their last round payoff against one other member of
@@
 more interactions or both. In
-Sections~\ref{section:m_one_n_two}-\ref{section:m_two_n_two} we present each of
-this extensions and for each a special case.
+Sections~\ref{section:m_one_n_two}-\ref{section:m_two_n_two} we discuss each of
+these extensions and for each extension we present a special case.
 In Section~\ref{section:simulation_results} we present numerical results of the
-pairwise comparison process using the several updating payoff approaches we have
-presented. In Sections~\ref{section:memory_one} and~\ref{section:mutation} we
-verify the main result of the paper when we no longer assume (i) that
+pairwise comparison process using the different updating payoff methods.
+In Sections~\ref{section:memory_one} and~\ref{section:mutation} we
+verify the main result of this work when we no longer assume (i) that
 individuals use reactive strategies but instead memory one strategies (ii) low
 mutation.
-
-% The aim of our work is to explore the effect
-% of the updating payoffs on the evolution of cooperation. To this end we
-% introduce a new method for calculating the updating payoffs and we compare the
-% results of the new method to those of the expected payoffs. In our approach we
-% do not assume that an individual has a perfect memory. Instead an individual
-% remembers only a limited number of interactions and limited outcomes from each
-% interaction.
 
@@
 
-\begin{equation}
+\begin{equation}\label{eq:donation}
     \begin{blockarray}{ccc}
@@
 
-Let \(\mathbf{u} = (b-c, -c, b, 0)\) be payoffs in a vector format, and let \(\mathcal{U}
+Let \(\mathbf{u} = (b-c, -c, b, 0)\) be Eq.~\ref{eq:donation} in a vector format, and let \(\mathcal{U}
 = \{r, s, t, p\}\) denote the set of feasible payoffs, where \(r\) denotes the
@@
 
-In the following sections we present several approaches for calculating the
-updating payoffs. Initially, we discuss the conventional approach of the
-expected payoffs and afterwards we present our newly introduced approaches.
-
-\section{Updating Payoffs based on the expected payoffs (perfect memory)}\label{section:perfect_memory}
-
-The expected payoffs are the conventional payoffs used in the updating stage. We
-refer to this approach as the perfect memory approach. The expected payoffs are
-defined as the mean payoff of an individual in a well-mixed population that
+In the following sections we present several methods for calculating the
+updating payoffs. Initially, we discuss the conventional way of the
+expected payoffs and afterwards we present our newly introduced methods.
+
+\section{Updating Payoffs based on Expected Payoffs (perfect payoff memory)}\label{section:perfect_memory}
+
+The expected payoffs are the conventional payoffs used in the updating stage.
+The expected payoffs are defined as the mean payoff of an individual in a well-mixed population that
 engages in an infinitely repeated games with all other population members.
@@
 \(\mathbf{v}(s_1,s_2) M = \mathbf{v}(s_1,s_2)\). In an infinitely repeated game
-the game stage outcome for each strategy is given by,
+the long-term average payoff of \(s_1\) and \(s_2\) are,
 
@@
 resident population of conditional cooperators, otherwise known as generous tit
-for tat (GTFT), players. Let GTFT\(= (1, 1, q)\), ALLD\(= (0, 0, 0)\), and \(k =
+for tat (GTFT). Let GTFT\(= (1, 1, q)\), ALLD\(= (0, 0, 0)\), and \(k =
 1\). When two GTFT players interact in an infinitely repeated game, the
@@
 
-This result will become important when we analyse the behaviour that emerges
-from the evolutionary process. We will repeat this ``invasion analysis'' for the
-rest of the updating payoffs cases.
-
-\section{Updating Payoffs based on the Last Round Payoff of One Interaction (limited memory)}\label{section:limited_memory}
+This result will become important when we discuss the numerical results of the
+evolutionary process (Section~\ref{section:simulation_results}). We repeat this
+``invasion analysis'' for the rest of the updating payoffs cases.
+
+\section{Updating Payoffs based on the Last Round Payoff of One Interaction (limited payoff memory)}\label{section:limited_memory}
 
 In the expected payoffs case the payoff of a pair depends on the average payoff
-they received over an infinite number of turns. In the limited memory payoffs,
+they received over an infinite number of turns. In the limited payoff memory,
 the payoff of a pair depends on the average payoffs they received in the last
-turn. Furthermore, in expected payoffs it is assumed that a player interacts
-with every member of the population, whereas in the limited memory approach a
-player has one interaction. Initially, we define the probability that a reactive
-strategy receives the payoff $u\!\in\! \mathcal{U}$ in the very last round of
-the game against another reactive strategy
-(Proposition~\ref{proposition:last_round}).
+turn. Moreover, in expected payoffs it is assumed that a player interacts with
+every member of the population whereas in the limited payoff memory approach a
+player has one interaction.
+
+Initially, we define the probability that a reactive strategy receives the
+payoff $u\!\in\! \mathcal{U}$ in the very last round of the game against another
+reactive strategy (Proposition~\ref{proposition:last_round}).
 
@@
       v_{r}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(q_1+l_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+l_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
-      {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)} \times r,\\[1cm]
+      {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)},\\[1cm]
 
       v_{s}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(q_1+l_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
-      {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)} \times s,\\[1cm]
+      {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)},\\[1cm]
 
       v_{t}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+l_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
-      {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)} \times t,\\[1cm]
+      {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)},\\[1cm]
 
       v_{p}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
-      {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)} \times p.
+      {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)}.
       \end{array}
@@
 \begin{array}{llrl}
-x(u_1,u_2)	 &=&\displaystyle \frac{1}{N\!-\!1}\cdot  &v_{u_1}(s_1,s_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F}\\[0.5cm]
+x(u_1,u_2)	 &=&\displaystyle \frac{1}{N\!-\!1}\cdot  &v_{u_1}(s_1,s_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}_F}\\[0.5cm]
 &+	
@@
 pairs can occur that are feasible in a direct interaction,
-$\splitatcommas{(u_1,u_2)\in \mathcal{U}^2_F:=\big\{ (r, r), (s, t), (t, s), (p,
+$\splitatcommas{(u_1,u_2)\in \mathcal{U}_F:=\big\{ (r, r), (s, t), (t, s), (p,
 p) \big\}}$, as represented by the respective indicator function. Otherwise, if
@@
   x(s, r) = & \frac{1}{N - 1} \cdot (1 - \delta + \delta q) \\
-  x(p, p) = & \frac{1}{N - 1} \cdot \delta (1 - q)
+  x(p, p) = & \frac{1}{N - 1} \cdot \delta (1 - q) \\
+  x(u_1, u_2) = &  0 \text{ for all other payoff pairs } (u_1, u_2).
 \end{align*}
@@
 
-In the limited memory payoffs we assume that an individual recalls the last
-round they received against one member of the population. This is an edge case.
 The framework of the limited memory can be generalised such that an individual
@@
 At each step of the evolutionary process the role model and the learner now
-participate in \(n\) matches. We need to define the probability that for these
-matches they are paired with a mutant, with a resident or with each other. We
-assume that each pair is unique, so the resident and role model can be matched
-together only once at each step. A representation of the process is given in
-Figure.
-
-In the case of \(n=1\) there are five possible pairs, however, the number of
-possible pairs increases non linearly as we increase the number of possible
-interactions. We demonstrate this case for \(n=2\), namely, for when the role
-model and the learner have two interactions.
-
-\subsection*{Special case: Last Round Payoff of Two Interactions}
-
-For \(n=2\) there are two stages of matching and there are twenty four possible pairs. In
-the first stage the role model and the learner are matched together with a
-probability $\frac{1}{N-1}$ or not with a probability $(1 - \frac{1}{N-1})$. If
-they were matched together then in the second stage there are only four possible
-outcomes; both of them interact with a mutant with a probability
-$\frac{(k-1)(k-2)}{(N-2)(N-3)}$ and both interact with a resident with a
-probability $\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$. The last two possible pairs are
-that either of them interacts with a resident whilst the other interacts with a
-mutant, and this happens with a probability $\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$.
-In the later case, where they were not matched in the first stage, there are
-four possible outcomes; both of them interact with a mutant and both interact
-with a resident, either of them interacts with a resident whilst the other
-interacts with a mutant. For each of the above pairs of the first stage there
-are five possible pairs in the second stage; they interact with each other, both
-of them interact with a mutant and both interact with a resident, either of them
-interacts with a resident whilst the other interacts with a mutant.
-
-The new possible pairs change how we define the probability that the
-respective last round payoffs of two players \(s_1, s_2\) are given by \(u_1\)
-and \(u_2\). The new probability denoted as \(\tilde{x}(u_1,u_2)\) is given by,
+participate in \(n\) matches. We need to define the probability that for each of
+the matches they are paired with a mutant, with a resident or with each other.
+We assume that each pair is unique, so for example the resident and role model
+can be matched together only once at each step. A representation of the process
+is given in Figure~\ref{fig:matching_tree}. In the case of \(n=1\) there are
+five possible pairs, however, the number of possible pairs increases non
+linearly as we increase the number of possible interactions. We demonstrate this
+case for \(n=2\), namely, for when the role model and the learner have two
+interactions.
+
+\begin{figure}[!htbp]
+  \centering 
+  \includegraphics[width=\textwidth]{static/matching_tree.pdf}
+  \caption{\textbf{Tree diagram of the possible pairs when the learner and the role
+  model have $n$ interactions.} In the diagram \((s_i, s_j)\) represent a possible
+  pair. We denote the role model as $s_{RM}$ and the learner as $s_{L}$, and
+  a member of the population that plays a mutant strategy is denoted as $s_{M}$ and
+  a resident strategy as $s_{R}$. The role model and
+  the learner need to be paired with $n$ other members of the population
+  (including each other). We break down the process into stages. For the first
+  pair (stage one) the role model and the mutant can be paired together (with
+  probability $\frac{1}{N-1}$) or not (with probability $1 -
+  \left(\frac{1}{N-1}\right)$). In the case that they are not paired together,
+  both can be paired with a mutant $\frac{(k-1)(k-2)}{(N-2)(N-3)}$, with a
+  resident $\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$, or one is paired with a mutant
+  whilst the other is paired with a resident $\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$.
+  There are five possible pairs for stage one as we have already discussed
+  in Section~\ref{section:limited_memory}. In
+  the second stage we need to consider if the role model and the learner
+  interacted with each other. If yes, then they can not interact with each other
+  again, and at each stage there are at most four possible pairs; they interact with
+  mutants, with residents, or the one interacts with a mutant whilst the second
+  interacts with a resident. In the case that they were not paired in stage 1
+  there are five possible pairs. These pairs are the same as in stage 1,
+  however now the probabilities differ. Namely, let's assume that in the
+  first stage both were matched with a mutant. They interact with each
+  other with a probability $\frac{1}{N-2}$ or not with a probability
+  $1 - \left(\frac{1}{N-2}\right)$. In the latter case there are four
+  possible pairs.
+  Both can be paired with a mutant $\frac{(k-1)(k-2)}{(N-3)(N-4)}$, with a
+  resident $\frac{(N-k-1)(N-k-2)}{(N-3)(N-4)}$, or one is paired with a mutant
+  whilst the other is paired with a resident $\frac{(N-k-1)(k-2)}{(N-3)(N-4)}$.
+  The process continues following the same logic until it reaches the $n^{\text{th}}$
+  stage.}\label{fig:matching_tree}
+\end{figure}
+
+\subsection{Last Round Payoff of Two Interactions}
+
+Here the learner and role model consider the last round payoff of two
+interactions. At each time step of the evolutionary process there are twenty
+four possible pairs. At the first stage there are fiver possible pairs. For four of
+these pairs there are five possible pairs in the second stage (\(4 \times 5 =
+20\)), and for the last one there are four possible pairs in the second stage
+(20 + 4 = 24).
+
+We assume that player \(s_1\) receives the payoff
+\(u_{11}\) with their first interaction and \(u_{12}\) with their second.
+Respectively, player \(s_2\) receives \(u_{21}\) and \(u_{22}\). We define the
+probability that the respective last round payoffs of the two players \(s_1,
+s_2\) are given by \((u_{11}, u_{11})\) and \((u_{21}, u_{22})\) as
+\(x^{2}((u_{11}, u_{11}), (u_{21}, u_{22}))\) given by,
 
@@
   \begin{array}{llrl}
-  \tilde{x}(u_1,u_2)	& = \displaystyle \frac{1}{N\!-\!1}\cdot v_{u_1}(s_1,s_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F} \cdot A + \left(1\!-\!\frac{1}{N\!-\!1}\right) \left[ \right. \\[0.5cm] 
-  & \left. v_{u_1}(s_1,s_2) _{u_2}(s_2,s_2)  \frac{\left(k - 2\right) \left(k - 1\right)}{(N\!-\!2)(N\!-\!3)}
-  \left(\frac{1}{N\!-\!2}\cdot v_{u_1}(s_1,s_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F} + (1 - \frac{1}{N - 2}) [B_1 + B_2 + B_3 + B_4]\right) + \right. \\[0.5cm] 
+  x^{2}((u_{11}, u_{12}), (u_{21}, u_{22}))	& = \displaystyle \frac{1}{N\!-\!1}\cdot v_{u_{11}}(s_1,s_2)\cdot 1_{(u_{11},u_{21})\in \mathcal{U}_F} \cdot A + \left(1\!-\!\frac{1}{N\!-\!1}\right) \left[ \right.
+\\[0.5cm]
+  & \left. v_{u_{11}}(s_1,s_2) _{u_{21}}(s_2,s_2)  \frac{\left(k - 2\right) \left(k - 1\right)}{(N\!-\!2)(N\!-\!3)}
+  \left(\frac{1}{N\!-\!2}\cdot v_{u_{11}}(s_1,s_2)\cdot 1_{(u_{11},u_{21})\in \mathcal{U}_F} + (1 - \frac{1}{N - 2}) [B_1 + B_2 + B_3 + B_4]\right) + \right. \\[0.5cm] 
   & \left.
-v_{u_1}(s_1,s_2) _{u_2}(s_2,s_1)  \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N\!-\!2)(N\!-\!3)} \left(\frac{1}{N\!-\!2}\cdot v_{u_1}(s_1,s_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F} + (1 - \frac{1}{N - 2}) [C_1 + C_2 + C_3 + C_4]\right) + \right.
+v_{u_{11}}(s_1,s_2) _{u_{21}}(s_2,s_1)  \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N\!-\!2)(N\!-\!3)} \left(\frac{1}{N\!-\!2}\cdot v_{u_{11}}(s_1,s_2)\cdot 1_{(u_{11},u_{21})\in \mathcal{U}_F} + (1 - \frac{1}{N - 2}) [C_1 + C_2 + C_3 + C_4]\right) + \right.
 \\[0.5cm]
   & \left.
-v_{u_1}(s_1,s_1) _{u_2}(s_2,s_2)  \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N\!-\!2)(N\!-\!3)} \left(\frac{1}{N\!-\!2}\cdot v_{u_1}(s_1,s_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F} + (1 - \frac{1}{N - 2}) [D_1 + D_2 + D_3 + D_4]\right) + \right.
+v_{u_{11}}(s_1,s_1) _{u_{21}}(s_2,s_2)  \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N\!-\!2)(N\!-\!3)} \left(\frac{1}{N\!-\!2}\cdot v_{u_{11}}(s_1,s_2)\cdot 1_{(u_{11},u_{21})\in \mathcal{U}_F} + (1 - \frac{1}{N - 2}) [D_1 + D_2 + D_3 + D_4]\right) + \right.
 \\[0.5cm]
   & \left.
-v_{u_1}(s_1,s_1) _{u_2}(s_2,s_1)  \frac{\left(N - k - 2\right) \left(N - k - 1\right)}{(N\!-\!2)(N\!-\!3)} \left(\frac{1}{N\!-\!2}\cdot v_{u_1}(s_1,s_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F} + (1 - \frac{1}{N - 2}) [E_1 + E_2 + E_3 + E_4]\right) \right]
+v_{u_{11}}(s_1,s_1) _{u_{21}}(s_2,s_1)  \frac{\left(N - k - 2\right) \left(N - k - 1\right)}{(N\!-\!2)(N\!-\!3)} \left(\frac{1}{N\!-\!2}\cdot v_{u_{11}}(s_1,s_2)\cdot 1_{(u_{11},u_{21})\in \mathcal{U}_F} + (1 - \frac{1}{N - 2}) [E_1 + E_2 + E_3 + E_4]\right) \right]
   \end{array} $}
@@
   \begin{array}{llrl}
-  A & = \left(1\!-\!\frac{1}{N\!-\!1}\right)  \left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_2) + 
-   \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_1) + \right. \\[0.5cm]
-   & \left. \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_2) + 
-   \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_1)\right] \\[0.5cm] 
-   B_1 & = \frac{\left(k - 3\right) \left(k - 2\right)}{(N - 3) (N - 4)} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_2) \quad
-   B_2 = \frac{\left(k - 2\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_2) \\[0.5cm] 
-   B_3 & = \frac{\left(k - 2\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_1) \quad
-   B_4 = \frac{\left(N - k - 2\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_1) \\[0.5cm]
-
-   C_1 & = \frac{\left(k - 3\right) \left(k - 1\right)}{(N - 3) (N - 4)} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_2) \quad
-   C_2 = \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_2) \\[0.5cm] 
-   C_3 & = \frac{\left(k - 2\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_1) \quad
-   C_4 = \frac{\left(N - k - 2\right)^{2}}{(N - 3) (N - 4)} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_1) \\[0.5cm]
-
-   D_1 & = \frac{\left(k - 2\right)^{2}}{(N - 3) (N - 4)} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_2) \quad
-   D_2 = \frac{\left(k - 2\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_2) \\[0.5cm] 
-   D_3 & = \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_1) \quad
-   D_4 = \frac{\left(N - k - 3\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_1) \\[0.5cm]
-
-   E_1 & = \frac{\left(k - 2\right) \left(k - 1\right)} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_2) \quad
-   E_2 = \frac{\left(k - 1\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_2) \\[0.5cm] 
-   E_3 & = \frac{\left(k - 1\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_1) \quad
-   E_4 = \frac{\left(N - k - 3\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_1) \\[0.5cm] 
+  A & = \left(1\!-\!\frac{1}{N\!-\!1}\right)  \left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} v_{u_{21}}(s_1,s_2) v_{u_{22}}(s_2,s_2) + 
+   \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} v_{u_{21}}(s_1,s_2) v_{u_{22}}(s_2,s_1) + \right. \\[0.5cm]
+   & \left. \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} v_{u_{21}}(s_1,s_1) v_{u_{22}}(s_2,s_2) + 
+   \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} v_{u_{21}}(s_1,s_1) v_{u_{22}}(s_2,s_1)\right] \\[0.5cm] 
+   B_1 & = \frac{\left(k - 3\right) \left(k - 2\right)}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_2) v_{u_{22}}(s_2,s_2) \quad
+   B_2 = \frac{\left(k - 2\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_1) v_{u_{22}}(s_2,s_2) \\[0.5cm] 
+   B_3 & = \frac{\left(k - 2\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_2) v_{u_{22}}(s_2,s_1) \quad
+   B_4 = \frac{\left(N - k - 2\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_1) v_{u_{22}}(s_2,s_1) \\[0.5cm]
+
+   C_1 & = \frac{\left(k - 3\right) \left(k - 1\right)}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_2) v_{u_{22}}(s_2,s_2) \quad
+   C_2 = \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_1) v_{u_{22}}(s_2,s_2) \\[0.5cm] 
+   C_3 & = \frac{\left(k - 2\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_2) v_{u_{22}}(s_2,s_1) \quad
+   C_4 = \frac{\left(N - k - 2\right)^{2}}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_1) v_{u_{22}}(s_2,s_1) \\[0.5cm]
+
+   D_1 & = \frac{\left(k - 2\right)^{2}}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_2) v_{u_{22}}(s_2,s_2) \quad
+   D_2 = \frac{\left(k - 2\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_1) v_{u_{22}}(s_2,s_2) \\[0.5cm] 
+   D_3 & = \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_2) v_{u_{22}}(s_2,s_1) \quad
+   D_4 = \frac{\left(N - k - 3\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_1) v_{u_{22}}(s_2,s_1) \\[0.5cm]
+
+   E_1 & = \frac{\left(k - 2\right) \left(k - 1\right)} v_{u_{21}}(s_1,s_2) v_{u_{22}}(s_2,s_2) \quad
+   E_2 = \frac{\left(k - 1\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_1) v_{u_{22}}(s_2,s_2) \\[0.5cm] 
+   E_3 & = \frac{\left(k - 1\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_2) v_{u_{22}}(s_2,s_1) \quad
+   E_4 = \frac{\left(N - k - 3\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_1) v_{u_{22}}(s_2,s_1) \\[0.5cm] 
   \end{array}$}
@@
 The first term on the right side corresponds to the case that the learner and
-the role model happened to be matched during the first stage of pairing,
+the role model happened to be matched during the first stage,
 followed by them being paired with another member of the population on the
@@
 $\frac{(k-3)(k-2)}{(N-4)(N-3)}$ and both interact with a resident with a
-probability $\frac{(N-k-1)(N-k-2)}{(N-3)(N-4)}$. The last two possible pairings
+probability $\frac{(N-k-1)(N-k-2)}{(N-3)(N-4)}$. The last two possible pairs
 are that either of them interacts with a resident whilst the other interacts
@@
 \begin{align}\label{eq:ratio_limited}
-\lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R},u_{M}\in\mathcal{U}} \tilde{x}(u_{R},u_{M})\cdot \rho(u_{R},u_{M}), \\
-\lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R},u_{M}\in\mathcal{U}} \tilde{x}(u_{R},u_{M})\cdot \rho(u_{M},u_{R}).
+\lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R1},u_{M1}\in\mathcal{U}} \quad \sum_{u_{R2},u_{M2}\in\mathcal{U}} x^{2}((u_{R1}, u_{R2}),(u_{M1}, u_{M2}))\cdot \rho\left(\frac{u_{R1} + u_{R2}}{2}, \frac{u_{M1} + u_{M2}}{2}\right), \\
+\lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R1},u_{M1}\in\mathcal{U}} \quad \sum_{u_{R2},u_{M2}\in\mathcal{U}} x^{2}((u_{R1}, u_{R2}),(u_{M1}, u_{M2}))\cdot \rho\left(\frac{u_{M1} + u_{M2}}{2}, \frac{u_{R1} + u_{R2}}{2}\right).
 \end{align}
+
+We use $x^{2}((u_{R1}, u_{R2}),(u_{M1}, u_{M2}))$ to denote the probability that
+the randomly chosen resident obtained a payoff of $u_{R1}$ in the last round of
+their first respective game and $u_{R1}$ in the last round of their second game.
+Similarly, that the mutant obtained a payoff of $u_{M1}$ from their first
+interaction and $u_{M2}$ from the second. Note that we assume that the players
+compare their average payoff of their two interaction when considering adopting
+the others strategy. For example the mutant adopts the resident's strategy with
+a probability $\rho\left(\frac{u_{R1} + u_{R2}}{2}, \frac{u_{M1} +
+u_{M2}}{2}\right)$.
 
@@
 GTFT players can not be invaded by an ALLD mutant. Using the new formulation we
-obtain the following probabilities \(\bar{x}(u_1, u_2)\) that the payoff of a randomly
-chosen GTFT player is \(u_1\) and that the payoff of the ALLD player is \(u_2\),
-
-\begin{align*}
-  x(r, t) = & \frac{N - 2}{N - 1} \cdot (1 - \delta + \delta q)\\
-  x(r, r) = & \frac{N - 2}{N - 1} \cdot \delta (1 - q) \\
-  x(s, r) = & \frac{1}{N - 1} \cdot (1 - \delta + \delta q) \\
-  x(p, p) = & \frac{1}{N - 1} \cdot \delta (1 - q)
+obtain the following probabilities \(x^{2}((u_{11}, u_{12}), (u_{21}, u_{22}))\) that the payoff of a randomly
+chosen GTFT player is \((u_{11}, u_{12})\) and that the payoff of the ALLD player is \((u_{21}, u_{22})\),
+
+\begin{align*}
+  x^{2}((r, r), (t, t)) = & \frac{N - 3}{N - 1}\left(\delta q - \delta + 1\right)^{2} & 
+  x^{2}((r, r), (t, p)) = & - \frac{N - 3}{N - 1} \delta \left(q - 1\right) \left(\delta q - \delta + 1\right)\\
+  x^{2}((r, r), (p, t)) = & - \frac{N - 3}{N - 1} \delta \left(q - 1\right) \left(\delta q - \delta + 1\right) &
+  x^{2}((r, r), (p, p)) = & \frac{N - 3}{N - 1} \delta^{2}  \left(q - 1\right)^{2}\\
+  x^{2}((r, s), (t, t)) = & \frac{1}{N-1} \left(\delta q - \delta + 1\right)^{2} &
+  x^{2}((r, s), (p, t)) = & - \frac{1}{N-1} \delta \left(q - 1\right) \left(\delta q - \delta + 1\right) \\
+  x^{2}((r, p), (t, p)) = & - \frac{1}{N-1} \delta \left(q - 1\right) \left(\delta q - \delta + 1\right) &
+  x^{2}((r, p), (p, p)) = & \frac{1}{N-1} \delta^{2} \left(q - 1\right)^{2} \\
+  x^{2}((s, r), (t, t)) = & \frac{1}{N-1} \left(\delta q - \delta + 1\right)^{2} &
+  x^{2}((s, r), (t, p)) = & - \frac{1}{N-1} \delta \left(q - 1\right) \left(\delta q - \delta + 1\right)\\
+  x^{2}((p, r), (p, t)) = & - \frac{1}{N-1} \delta \left(q - 1\right) \left(\delta q - \delta + 1\right) &
+  x^{2}((p, r), (p, p)) = & \frac{1}{N-1} \delta^{2} \left(q - 1\right)^{2} \\
+  x^{2}((u_{11}, u_{12}), (u_{21}&, u_{22})) =  0 \text{ for all other payoff pairs} & ((u_{11}, u_{12}), (u_{21}, u_{22})). &
 \end{align*}
@@
 
-The second generalised case of the limited memory payoffs we discuss in that of
-individuals updating based on their last $m$ rounds payoffs with one member. To
-this end we define the probability that a reactive strategy receives the payoffs
-$u\!\in\! \mathcal{\tilde{U}}$, for \(\mathcal{\tilde{U}}= \{\underbrace{rrr\dots r}_{m},
-\underbrace{rrr\dots s}_{m}, \dots, \underbrace{ppp\dots p}_{m}\}\), in the last \(m\) rounds against another
-reactive strategy (Proposition~\ref{proposition:last_m_rounds}).
+The second generalised case of the limited memory payoffs we discuss is that of
+individuals updating based on their last $m$ rounds payoffs with one member. Let
+\(\mathcal{U}^{m}= \{\underbrace{rrr\dots r}_{m}, \underbrace{rrr\dots s}_{m},
+\dots, \underbrace{ppp\dots p}_{m}\}\) be the set of feasible payoffs of the
+last $m$ rounds. The probability that a reactive strategy receives the payoffs
+$u\!\in\! \mathcal{U}^{m}$ is given by
+Proposition~\ref{proposition:last_m_rounds}.
 
@@
   $s_1\!=\!(y_1, p_1, q_1$) and $s_2\!=\!(y_2,p_2,q_2)$ respectively. Let
-  $\mathcal{\tilde{U}}$ denote the set of feasible payoffs in the last
-  \(n\) rounds, and let \(\tilde{\mathbf{u}}\) be the corresponding payoff vector.
+  $\mathcal{U}^{m}$ denote the set of feasible payoffs in the last
+  \(m\) rounds, and let \(\mathbf{u}^{m}\) be the corresponding payoff vector.
   Then the probability that the $s_1$ player receives the payoff $u\!\in\!
-  \mathcal{\tilde{U}}$ in the very last two rounds of the game is given by,
+  \mathcal{U}^{m}$ in the very last two rounds of the game is given by,
 
   \begin{equation}
-  \langle\mathbf{\tilde{v}}(s_1,s_2),\mathbf{\tilde{u}}\rangle, \text{ where } \mathbf{\tilde{v}} \in R^{4^{n}} \text{ is given by },
+  \langle\mathbf{v}^{m}(s_1,s_2),\mathbf{u}^{m}\rangle, \text{ where } \mathbf{v}^{m} \in R^{4^{m}} \text{ is given by },
   \end{equation}
   \begin{equation}
-    \mathbf{\tilde{v}}(s_1,s_2) = (1 - \delta) w_{a_1, a_2} \delta^2 \left[\mathbf{v_0}(I_4 - \delta M)^{-1}\right]_{a_1, a_2}, \quad  w_{a_1, a_2} \in M \ \forall \ a_1, a_2 \in \{1, 2, 3, 4\}.
+    \mathbf{v}^{m}(s_1,s_2) = (1 - \delta) w_{a_1, a_2} \delta^2 \left[\mathbf{v_0}(I_4 - \delta M)^{-1}\right]_{a_1, a_2}, \quad  w_{a_1, a_2} \in M \ \forall \ a_1, a_2 \in \{1, 2, 3, 4\}.
   \end{equation}
@@
 
-The probability that the number of mutants increases, and decreases respectively
-remains the same as in section~\ref{section:m_one_n_one} (Eq.~\ref{eq:ratio_limited}). The
-frameworks differ in the game stage payoffs,
+\subsection{Last Round Payoff of Two Interactions}
+
+In the special case of $m=2$ the stationary distribution \(\mathbf{v}^{2}(s_1,
+s_2)\) is sixteen dimension instead of four dimensional where \(v_{1}(s_1, s_2)\) is
+the long term probability that \(s_1\) and \(s_2\) mutually cooperated in the
+last two rounds. Let the feasible payoffs a strategy can receive
+be $\mathcal{U}^{2} = \{rr, rs, rt, rp, sr, \dots, pp\}$.
+The role model and the learner interact with only one other member and so the
+probability \(x\) remains the same as in Section~\ref{section:limited_memory}.
+
+The probability that the number of mutants increases, and decreases respectively,
+by one is now given by,
 
 \begin{align}
-  \lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{\tilde{u}_{R},\tilde{u}_{M}\in\mathcal{\tilde{U}}} \tilde{x}(\tilde{u}_{R},\tilde{u}_{M})\cdot \rho(\tilde{u}_{R},\tilde{u}_{M}), \\
-  \lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{\tilde{u}_{R},\tilde{u}_{M}\in\mathcal{\tilde{U}}} \tilde{x}(\tilde{u}_{R},\tilde{u}_{M})\cdot \rho(\tilde{u}_{M},\tilde{u}_{R}).
+\lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} \cdot \sum_{u_{R},u_{M}\in\mathcal{U}^{2}} x(u_{R},u_{M})\cdot \rho\left(\frac{u_{R}}{2},\frac{u_{M}}{2}\right), \\
+\lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} \cdot \sum_{u_{R},u_{M}\in\mathcal{U}^{2}} x(u_{R},u_{M})\cdot \rho\left(\frac{u_{M}}{2},\frac{u_{R}}{2}\right).
 \end{align}
 
-\subsection*{Special case: Last Round Payoff of Two Interactions}
+We assume again that players consider their average payoff of the last two turns
+when deciding to imitate another strategy or not.
 
@@
 
-In a similar fashion we can calculate the condition for which a population of
-GTFT players can not be invaded by an ALLD mutant. The ratio of transition
-probabilities is given by,
+We once again calculate how easily a single ALLD mutant can invade into a
+resident population of GTFT player. When two GTFT players interact in the game,
+their respective probabilities for each of the four outcomes in the last round
+simplify to,
+
+\begin{align*}
+    v_{rr}(GTFT,GTFT) = & \delta^2, \text{ and }\\
+    v_{i}(GTFT,GTFT) = &  0 \text{ for all other } i \in \mathcal{U}^{2}.
+\end{align*}
+
+On the other hand, if an ALLD player interacts with a GTFT player, the
+respective probabilities according to Eq.~\ref{Eq:LastRound} become
+
+\begin{align*}
+    v_{tt} (ALLD,GTFT) & = \delta^{2} q \left(\delta q - \delta + 1\right), & 
+    v_{tp}(ALLD,GTFT) & = - \delta^{2} \left(q - 1\right) \left(\delta q - \delta + 1\right), \\
+    v_{pt} (ALLD, GTFT ) & = \delta^{3} q \left(1 - q\right), & 
+    v_{pp} (ALLD, GTFT) & = \delta^{3} \left(q - 1\right)^{2} \text{ and }\\
+    v_{i}(ALLD,GTFT) & =  0 \text{ for all other } i \in \mathcal{U}^{2}.
+\end{align*}
+
+As a consequence, we obtain the following probabilities \(x(u_1, u_2)\) that the
+payoff of a randomly chosen GTFT player is \(u_1\) and that the payoff of the
+ALLD player is \(u_2\),
+
+\begin{align*}
+  x(rr, tt) = & \frac{N-2}{N - 1} \delta^{4} q \left(N - 2\right) \left(\delta q - \delta + 1\right) &
+  x(rr, tp) = & - \frac{N-2}{N - 1} \delta^{4} \left(N - 2\right) \left(q - 1\right) \left(\delta q - \delta + 1\right)\\
+  x(rr, pt) = & - \frac{N-2}{N - 1} \delta^{5} q \left(N - 2\right) \left(q - 1\right) &
+  x(rr, pp) = & \frac{N-2}{N - 1} \delta^{5} \left(N - 2\right) \left(q - 1\right)^{2}\\
+  x(ss, tt) = & \frac{1}{N -1}\delta^{2} q \left(\delta q - \delta + 1\right) & 
+  x(sp, tp) = & - \frac{1}{N -1}\delta^{2} \left(q - 1\right) \left(\delta q - \delta + 1\right) \\
+  x(ps, pt) = & - \frac{1}{N -1}\delta^{3} q \left(q - 1\right) &
+  x(pp, pp) = & \frac{1}{N -1}\delta^{3} \left(q - 1\right)^{2}\\
+  x(u_1, u_2) = &  0 \text{ for all other payoff pairs } (u_1, u_2). &
+\end{align*}
+
+We now calculate the ratio of transition probabilities as
 
@@
 
+Note that as \(\delta \rightarrow 1\) the condition for which condition
+cooperators can avoid invasion is the same for the case of \(m=2\) and \(n=2\).
+
 \section{Updating Payoffs based on the last two rounds payoff of two interactions (\(m=2\) and \(n=2\))}\label{section:m_two_n_two}
 
-Finally a possible extension to the limited memory framework is to consider that
-thee number of rounds and number of interactions increase. For this case we need
-to consider a combination of the methods we presented in
-Section~\ref{section:m_one_n_two} and Section~\ref{section:m_two_n_one}.
+The final extension to the limited memory framework we consider is that of
+increasing the number of rounds and the number of interactions. For this case we
+need to consider a combination of the methods we presented in
+Section~\ref{section:m_one_n_two} and Section~\ref{section:m_two_n_one}. As an
+example consider the case of \(m=n=2\). The probability that the number of
+mutants increases, and decreases respectively, by one is now given by,
+
+\begin{align}
+  \lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R1},u_{M1}\in\mathcal{U}^{2}} \quad \sum_{u_{R2},u_{M2}\in\mathcal{U}^{2}} x^{2}((u_{R1}, u_{R2}),(u_{M1}, u_{M2}))\cdot \rho\left(\frac{u_{R1} + u_{R2}}{2}, \frac{u_{M1} + u_{M2}}{2}\right), \\
+  \lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R1},u_{M1}\in\mathcal{U}^{2}} \quad \sum_{u_{R2},u_{M2}\in\mathcal{U}^{2}} x^{2}((u_{R1}, u_{R2}),(u_{M1}, u_{M2}))\cdot \rho\left(\frac{u_{M1} + u_{M2}}{2}, \frac{u_{R1} + u_{R2}}{2}\right).
+\end{align}
+
+Though we do not carry any further analytical exploration of this case, in the
+next section we present simulation results when the updating payoffs of the
+pairwise process depend on the limited memory framework with \(m=n=2\), as well
+as the rest of the cases we have discussed so far.
 
@@
 case \(q \leq 1 - \frac{c}{b}=0.9\). The higher tolerance to defection results
-to a more cooperative population. As a result the expected payoffs allow for the
-most cooperative population. Between the limited memory approaches we observe a
-big difference in the cases of one or more information. We hypothesis that as we
+in a more cooperative population. As a result the expected payoffs allow for the
+most cooperative population. Between the limited memory approaches, we observe a
+big jump in the cooperation rate when we allow for more information
+(in the form of interactions or rounds). We hypothesise that as we
 allow for more information the results will tend to the case of perfect memory.

### 2023-03-06 paper/main.tex
finish abstract

@@
 based on a minimum of social information. We explore the difference between our
-framework and the classical framework using computer simulations. The
-simulations show that in the prisoner's dilemma, individuals with limited memory
-tend to adopt less generous strategies and they achieve less cooperation than in
-the classical scenario.
+framework and the classical framework using the iterated prisoner's dilemma. We
+show that individuals relying on a single set of information tend to adopt less
+generous strategies and they achieve less cooperation than in the classical
+scenario. Although strategies in the latter scenario become more cooperative
+with increasing benefits, those in the former remain unchanged. We further
+explore scenarios where individuals are granted slightly more information, and
+our observations indicate that these individuals can attain higher levels of
+cooperation.
+
 \end{abstract}
@@
 This is especially true in the topic of cooperation~\cite{hilbe:PNAS:2013,
-hilbe:Nature:2018,glynatsi:SCR:2020}. In stochastic evolutionary dynamics
+hilbe:Nature:2018,glynatsi:SCR:2020, ohtsuki:Nature:2006, nowak:Science:2006,
+murase:ScientificReports:2020}. In stochastic evolutionary dynamics
 disadvantageous mutants have a small yet non-zero probability to reach fixation
@@
 repeated game between two individuals the actions of the players in each turn
-are often determined by the action of the co-players in the previous turn. This
+are often determined by the action of the co-players in the previous turn~\cite{Nowak1992tit}. This
 assumption is common as it allows for an explicit calculation of the players'
@@
 assumed that individuals play many times and with all other players before
-reproduction takes place. So that in the updating phase the updating payoffs, and
-subsequently the fitness, of individuals is given by the average payoffs of an
-infinitely repeated game against the mean distribution of types in the
-population.
+reproduction takes place. So that in the updating phase the updating payoffs,
+or fitness, of individuals is given by the average payoffs of an
+infinitely repeated game against every member of the population.
 
@@
 memory using a pairwise comparison process and the well studied game of the
-prisoner's dilemma. We observe that individuals with limited payoff memory tend to
-adopt less generous strategies and they achieve less cooperation when
-interacting in a prisoner's dilemma. We obtain similar results when we consider
-that individuals update their strategies based on more information. More
-specifically, up to the last two payoffs they obtained when interacting with up
-to two different members of the population.
+prisoner's dilemma~\cite{glynatsi:HSSC:2021}. We observe that individuals with
+limited payoff memory tend to adopt less generous strategies. We obtain similar
+results when we consider that individuals update their strategies based on more
+information. More specifically, up to the last two payoffs they obtained when
+interacting with up to two different members of the population.
 
@@
 
-A pairwise comparison process starts with assigning all
+A pairwise comparison process~\cite{traulsen2007pairwise} starts with assigning all
 individuals of the population the same strategy. Thenceforth each elementary time step of
@@
 pair of reactive strategies $s_1 = (y_1, p_1, q_1)$ and $s_2 = (y_2, p_2, q_2)$
-can be model as a Markov process with the transition matrix \(M\),
+can be model as a Markov process with the transition matrix \(M\)~\cite{nowak:APC:1989},
 
@@
 
-The first term on the right side corresponds to the case that the learner and
-the role model happened to be matched during the game stage, which happens with
-probability $\frac{1}{(N\!-\!1)}$.
-
 The probability that the number of mutants increases and decreases by one in the
@@
 conditional cooperation probabilities $p$ and $q$ (since the discount
-factor~$\delta$ is comparably high we do not report the opening move \(y\)). The
-upper panel of Figure~\ref{fig:expected_and_stochastic_for_donation} corresponds
-to the standard scenario considered in the literature, and the bottom panel
-shows the scenario considered herein, in which players update their strategies
-based on their last round's payoff.
+factor~$\delta$ is comparably high we do not report the opening move \(y\)).
 
@@
 
-We further explore the effects of the cooperation benefit and the strength of
-selection on the evolved generosity \(q\) and on the cooperation rate
+We further explore the effects of the benefit and the strength of
+selection on the generosity \(q\) and on the cooperation rate
 (Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}).
@@
 
-Up to this point we have discussed only two edge cases. The framework of limited
+The framework of limited
 payoff memory can be easily extended to allow individuals to observe more
@@
 population, the last two rounds with another member of the population, and the
-last two rounds with two members of the population. Similar to the last round
-payoff. We perform numerical simulations for each of the fitness approaches. In
+last two rounds with two members of the population. We perform numerical
+simulations for each of the fitness methods. In
 Figure~\ref{fig:cooperation_rate_all_updating_payoffs} we report the cooperation
@@
 perfect memory assumption. We considered a special case of the prisoner's
-dilemma, the donation game. The prisoner's dilemma is one of the most applied
-types of social games when we discuss the evolution of cooperation. We showed
+dilemma, the donation game. We showed
 that when the last round payoff is used instead of the expected payoffs,

### 2023-03-06 scripts/clean_data.py

@@ -8,14 +8,14 @@ import pandas as pd
 
 if __name__ == "__main__":  # pragma: no cover
     folder = sys.argv[1]
-    starting_characters = "two_rounds_opponentslow_benefit_seed" #sys.argv[2]
+    starting_characters = "beta_0.01_stochastic_1" #sys.argv[2]
 
     in_columns = [
         "y",
         "p",
         "q",
         "average cooperation rate",
-        # "average score",
+        "average score",
         "t",
     ]
 
@@ -43,7 +43,7 @@ if __name__ == "__main__":  # pragma: no cover
         for value in df["t"][1:]:
             frequencies.append(value - previous)
             previous = value
-        frequencies.append(10 ** 6 - previous)
+        frequencies.append(10 ** 7 - previous)
 
         filename = file.split("/")[-1]
         df["frequencies"] = frequencies

### 2023-03-06 paper/main.tex
finish reading paper

@@
 
-Similar to all other theoretical models, finite stochastic processes rest on a
+Like all theoretical models, finite stochastic processes rest on a
 set of assumptions. Namely, in the game phase it is assumed that players use
@@
 
-The above two assumptions create a curious inconsistency; a player with a finite
-memory in the game stage can recall infinite information regarding their
-interactions and each interaction's outcomes in the update stage. Previous work
-has explored the effects of constraining the interactions an individual has.
+These two assumptions create a curious inconsistency; a player with a finite
+memory (in the game stage) can recall infinite information regarding their
+interactions and each interaction's outcomes in the update stage. Prior research
+has examined the effects of constraining the interactions an individual has.
 Either by letting individuals have a stochastic number of
 interactions~\cite{roca:PhysicalReview:2006, sanchez:JTB:2005,
-Traulsen:JTB:2007} or by only allowing for
-a single interaction~\cite{Woelfing:JTB:2009}. However, no previous work
-considers that not only the interactions are limited but also the information
-regarding each interaction. To this end, we propose a framework in which
-individuals, similar to the decisions at each turn, estimate their fitness based
-on a minimum of information.
+Traulsen:JTB:2007} or by only allowing for a single
+interaction~\cite{Woelfing:JTB:2009}. However, these studies have not accounted
+for the use of strategies with finite memory. Moreover, no previous research has
+explored the effect of restricting not only the number of interactions, but also
+the amount of information available about each interaction. To this end, we
+propose a framework in which individuals, similar to the decisions at each turn,
+estimate their fitness based on a minimum of information.
 
@@
 
-In order to account for the effect of the updating payoffs we simulate the
-evolutionary process and record which strategies the players adopt over time
-based on perfect and limited payoff memory. We performed two independent runs
-for each approach varying the value of benefit \(b\).
+To assess the impact of updating payoffs, we simulate the evolutionary process,
+recording the strategies adopted by players over time based on perfect and
+limited payoff memory. We performed two separate runs for each approach varying
+the value of benefit \(b\).
 Figure~\ref{fig:expected_and_stochastic_for_donation} depicts the evolving
-conditional cooperation probabilities $p$ and $q$ (since the discount
-factor~$\delta$ is comparably high we do not report the opening move \(y\)).
+conditional cooperation probabilities $p$ and $q$ (note that we omit the opening
+move \(y\), as the discount factor \(\delta\) is relatively high). The figure
+suggests that when updating is based on perfect payoff memory players tend to be
+more generous and more cooperative.
 
-The figure suggests that when updating is based on perfect payoff memory players
-tend to be more generous and more cooperative. The $q$-values of the resident
-strategies are on average higher in this case, thus, players will occasionally
-forgive a defection more often if their fitness depends on interacting with
-every member of the population. This effect becomes more notable as the value of
-the benefit increases. In the case of limited memory the resident strategies are
-less forgiving and their forgiveness regardless of the values of benefit. In
-analysis carried out in the Supplementary Information Section 3 we show that a
-resident population of conditional cooperators with \(q > \frac{1}{2}\) can not
-avoid invasion by a defecting strategy. This holds independently of the cost and
-benefit values.
+Specifically, we observe that the resident population comprises either defectors
+or conditional cooperators \((1, q)\). The generosity level \(q\) adopted by the
+resident population depends on whether a defecting strategy can invade.
+Supplementary Information Section 2 and 3 show that in the perfect payoff memory
+framework, conditional cooperators of the form \((1, q < \frac{c}{b})\) can
+arise, however, in the case of limited payoff memory, only conditional
+cooperators of \((1, q < \frac{1}{2})\) can avoid invasion. 
+The $q$-values of the resident strategies are generally higher in classical
+case, indicating that players are more likely to forgive a defection if their fitness
+depends on interacting with every member of the population. This effect becomes
+more pronounced as the benefit value increases, as the perfect memory condition
+on the left-hand side increases, while in the limited memory case, it remains at
+\(q \approx \frac{1}{2}\).
 
-Higher $q$ values lead to a more cooperative population. We calculate the
-average cooperation rate for each simulation which is the average cooperation
-rate within the resident population. In the case of the perfect payoff memory 
-the average cooperation rate is strictly higher than that of the last round
-payoffs. The difference between the cooperation rate is the highest at $b=10$.
+Higher $q$ values lead to a more cooperative population. We compute the average
+cooperation rate for each simulation, which is the average cooperation rate
+within the resident population. In the case of perfect payoff memory, the
+average cooperation rate is consistently higher than that of the last round
+payoffs.
 
@@
     \centering
-    \includegraphics[width=.9\textwidth]{static/donation_expected_last_round_summary_results.pdf}
+    \includegraphics[width=\textwidth]{static/donation_expected_last_round_summary_results.pdf}
     \caption{{\bf Evolutionary dynamics under perfect and limited payoff memory.}
@@
 
-We further explore the effects of the benefit and the strength of
-selection on the generosity \(q\) and on the cooperation rate
-(Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}).
-Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}\textbf{A} suggests that
-perfect memory always yields a higher cooperation rate. In this case the
-cooperation rate increases as the benefit increases. In comparison for limited
-payoff memory the cooperation rate remains unchanged at approximately 50\% once
-\(b=5\).  From Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}\textbf{B}
+We further investigate the impact of benefit and selection strength on
+generosity $q$ and the cooperation rate, as shown in Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}.
+According to Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}\textbf{A},
+perfect memory consistently results in a higher cooperation rate, which
+increases with increasing benefit. On the other hand, the cooperation rate
+remains approximately 50\% for limited payoff memory once \(b = 5\).
+From Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}\textbf{B}
 we observe that for weak selection, \(\beta < 1\), the two methods yield similar
@@
     similar results. However, as \(\beta\) increases in the case of limited
-    memory payoffs the resident populations become more defective. Unless
+    memory payoffs the resident populations become more defective. Note that
+    in the case of perfect payoff memory we see an increase in the cooperation
+    rate even though the generosity remains stable. That is because the
+    generosity does remain the same, however, now cooperative strategies remain
+    fixed as the resident strategy for longer.
+    Unless
     explicitly varied, the parameters of the simulation are $N\!=\!100$,
@@
 
-The framework of limited
-payoff memory can be easily extended to allow individuals to observe more
-rounds, to interact with more members or both (see Supplementary Information
-Sections 4-6). To further understand the effect of limited payoff memory we
-consider the cases of updating based on the last round with two members of the
-population, the last two rounds with another member of the population, and the
-last two rounds with two members of the population. We perform numerical
-simulations for each of the fitness methods. In
-Figure~\ref{fig:cooperation_rate_all_updating_payoffs} we report the cooperation
-rates for low and high values of benefit.
+The limited payoff memory framework can be expanded by enabling individuals to
+observe a greater number of rounds, interact with a larger number of members, or
+both (refer to Supplementary Information Sections 4-6). To gain further insight
+into the impact of limited payoff memory, we explore the scenarios of updating
+based on the last round with two members of the population, the last two rounds
+with another member of the population, and the last two rounds with two members
+of the population. To analyze the effects of this framework, we conduct
+numerical simulations using various fitness methods. The cooperation rates for
+low and high benefits are presented in Figure~\ref{fig:cooperation_rate_all_updating_payoffs}.
+
+We observe a significant rise in the cooperation rate with the introduction of
+slightly more information. Specifically, in scenarios involving two rounds or
+two interactions, the cooperation rates are almost identical. For a large
+population and a high continuation probability ($\delta$), the conditional
+cooperators that are adopted by the resident population in these scenarios take
+on the same form \((1, q < \frac{\sqrt{2}}{2})\). When considering both sets
+of information (i.e., two rounds and two interactions), cooperation rates
+experience a greater increase, yet still remain lower compared to the classical
+scenario.
 
@@
   allow for more information the closer we move to the perfect payoff memory.
+  We performed four pairwise non parametric tests (Mann-Whitney U) to compare
+  the cooperation distributions of the residents in case (b) to
+  cases (a), (c), (d), (e). In all four tests we reject the null hypothesis
+  with \(\alpha=0.05\) and \(p \approx 0\). Thus, there is significant
+  difference between the cooperation rates.
   Unless explicitly varied, the parameters of the simulation are
@@
 
-Cooperation can be seen at odd, why is it that we choose to help others at a
+Cooperation can be seen as odd, why is it that we choose to help others at a
 personal cost? In spite of all the selfish genes', animal and human communities
@@
 game theory seems to be reduced to the evolution of
-cooperation~\cite{Traulsen:PhilTrans:2022}. Previous evolutionary models often
-feature a curious inconsistency. When modeling how individuals make decisions in
-each round, these models assume that players only remember the last round.
-However, when modeling how individuals update their strategies over time,
-individuals are assumed to have perfect memory.
+cooperation~\cite{Traulsen:PhilTrans:2022}.
 
-Here, we have explored how robust cooperation is as models deviate from the
-perfect memory assumption. We considered a special case of the prisoner's
-dilemma, the donation game. We showed
-that when the last round payoff is used instead of the expected payoffs,
-cooperation struggles to evolve  when only a minimum of social information is
-used at the updating stage. The resident strategies were both less cooperative
-and less generous. This effect was only intensified we increase the benefit and
-the strength of selection independently. The results showed that as each
-parameter was being increased the difference between the cooperation rates were
-widening. This indicates that cooperative players benefit from being able to
-interact with everyone in the population.
+Evolutionary models in the past often feature a curious inconsistency. While
+these models depict how individuals make decisions in each round by assuming
+that they only retain memory of the previous round, they also assume that
+individuals possess a perfect memory when it comes to updating their strategies
+over time. To be precise, individuals are assumed to remember all of their past
+interactions and each interaction's outcome when updating strategies.
 
-We extended our approach and presented results on different limited memory
-payoffs. The analysis showed cooperation can even evolve when we allow
-individual to consider more information.
+Here, we investigate the robustness of cooperation as models deviate
+from the assumption of perfect memory. While prior research has investigated the
+impact of constraining individuals' interactions, we take into account the
+limitation of not only interactions but also the information available for each
+outcome. Additionally, prior studies have only allowed for the adoption of
+simple strategies such as always cooperating or always defecting. In contrast,
+we enable the use of more intricate strategies where players can utilize the
+previous play of their co-player to make decisions.
+
+In our framework, players update their strategies based on a combination of
+interactions and outcomes. The initial scenario we examined involved using one
+piece of information: the last round of one interaction. The outcomes suggest
+that cooperation faces difficulties in developing when the updating stage
+utilizes minimal social information. This effect is compounded as the benefit
+and strength of selection are independently increased. The findings indicate
+that cooperative players benefit from the ability to engage with all members of
+the population.
+
+Furthermore, we investigated scenarios where the final two rounds, or the last
+two instances of interaction, were taken into account. We observed a
+statistically significant rise in the frequency of cooperative behavior. For a
+sizable population and a high likelihood of continued interactions, the two
+cases yield the same result as the overall payoff is influenced by two possible
+outcomes. Notably, the scenario involving two rounds and two interactions
+yielded the highest cooperation rate among all the novel methodologies that we
+tested.
+% A nice sentence missing here.
 

### 2023-03-06 scripts/clean_data.py
run black

@@ -8,7 +8,7 @@ import pandas as pd
 
 if __name__ == "__main__":  # pragma: no cover
     folder = sys.argv[1]
-    starting_characters = "beta_0.01_stochastic_1" #sys.argv[2]
+    starting_characters = "beta_0.01_stochastic_1"  # sys.argv[2]
 
     in_columns = [
         "y",

### 2023-03-06 src/evol_dynamics/mutation.py

@@ -1,6 +1,7 @@
 import random
 
 import numpy as np
+
 # import tqdm
 import copy
 

### 2023-03-07 paper/static/game_stage.tex
Re writing paper  (#7)
* fix maths for two opponents

* prep to run donation with two opponents

* run only for two opponents

* run further memory one experiments

* change in two opponents

* payoffs run for mem one

* stop showing probabilities

* finilise two opponents phi

* remove two rounds two opponents simulation from script

* move two opponents two rounds and fix phi

* prep to run script

* set beta for memory one experiment

* prep to re-run memory one

* clean notebooks

* python script for mutation experiment

* fermi fix

* run for 10 ** 7

* add matlab files for mutation

* wr

--- 2022-08-26 paper/game_stage.tex
@@
     % %%%%% game stage %%%%%
-    \node at (0, .7) (turns) {turn};
+    \node at (0, .7) (turns) {turns};
 
@@
     \node at (0, -.6) {B};
-    \pic[RoyalBlue] at (0, -2) (myman) {man};
+    \pic[cyan] at (0, -2) (myman) {man};
     \node at (0, -2.6) {F};
@@
 
-    \node at (1.5, .7) (turn1) {1};
-    \node[circle, inner sep=1pt, outer sep=1pt, draw, fill] at (1.5, 1.7) (exit) {};
+    \node at (1.5, .7) (turn1) {\textbf{1}};
+    % \node[circle, inner sep=1pt, outer sep=1pt, draw, fill] at (1.5, 1.7) (exit) {};
     \node[circle, draw] (C) at (1.5, -.6)  {$C$};
@@
 
-    \node at (3, .7) (turn2) {2};
+    \node at (3, .7) (turn2) {\textbf{2}};
     \node[circle, inner sep=1pt, outer sep=1pt, draw, fill] at (3, 1.7) (exit2) {};
@@
 
-    \draw (turn1) edge[out=15, in=165, ->, thick] node [above] {$\delta$} (turn2);
-    \draw (turn1) edge[out=90, in=-90, ->, thick] node [above, rotate=90] {$1 - \delta$} (exit);
-    \draw (turn2) edge[out=90, in=-90, ->, thick] node [above right] {$1 - \delta$} (exit2);
+    \draw (turn1) edge[out=15, in=165, ->, thick] node [above] {\tiny{$\delta$}} (turn2);
+    \draw (turn2) edge[out=90, in=-90, ->, thick] node [above right] {\tiny{$(1 - \delta)$}} (exit2);
 
+    \node at (4, .7) (turn3) {$\dots$};
     \node[circle] at (4, -.6) {$\dots$};

### 2023-03-07 clean_data.py

@@ -1,39 +1,0 @@
-import sys
-
-import glob
-
-import tqdm
-
-import pandas as pd
-
-if __name__ == "__main__":  # pragma: no cover
-    folder = sys.argv[1]
-    starting_characters = sys.argv[2]
-
-    in_columns = [
-        "y",
-        "p",
-        "q",
-        "average cooperation rate",
-        "average score",
-        "t",
-    ]
-
-    print(f"{folder}/{starting_characters}*.csv")
-    csv_files = glob.glob(f"{folder}/{starting_characters}*.csv")
-
-    for file in tqdm.tqdm(csv_files):
-        df = pd.read_csv(file, header=None)
-        df = df.drop_duplicates()
-        df.columns = in_columns
-
-        frequencies = []
-        previous = 0
-        for value in df["t"][1:]:
-            frequencies.append(value - previous)
-            previous = value
-        frequencies.append(10 ** 7 - previous)
-
-        df["frequencies"] = frequencies
-
-        df.to_csv(f"{folder}clean_{file[len(folder):]}", index=False)

### 2023-03-07 matlab/RunMutation.m

@@ -1,0 +1,43 @@
+function RunMutation();
+
+starting_resident = [0, 0, 0];
+N = 100;
+sdim = 3;
+delta = 0.999;
+beta = 1;
+numberIterations = 10 ^ 5;
+
+data = ["expected", 10 ^ -4;
+        "expected", 10 ^ -3;
+        "expected", 10 ^ -2;
+        "expected", 10 ^ -1;
+        "expected", 10 ^ 0;
+        "last_round", 10 ^ -4;
+        "last_round", 10 ^ -3;
+        "last_round", 10 ^ -2;
+        "last_round", 10 ^ -1;
+        "last_round", 10 ^ -0];
+
+u = [2, -1, 3, 0];
+n_run = 2;
+previous = n_run - 1;
+
+parfor (i = 1:1)
+        mutation = str2num(data(i, 2));
+        payoff_type = data(i, 1);
+
+        filename = "../data/test/mutation_" + payoff_type + "_mutation_" + mutation + "_run_" + n_run;
+        population_fln = "../data/test/mutation_" + payoff_type + "_mutation_" + mutation + "_run_" + previous;
+        if n_run == 1;
+            %% Initialization
+            Res=starting_resident; 
+            population = Res .* ones(N, sdim, 'single');
+
+        else
+            population = readmatrix(population_fln + "_population.csv");
+
+        end
+        disp(population)
+%         evolMutation(population, u, N, delta, beta, mutation, numberIterations, payoff_type, filename);
+end
+end

### 2023-03-07 matlab/contimutationRun.m

@@ -1,0 +1,17 @@
+function evolRun();
+
+starting_resident = [0, 0, 0];
+N = 100;
+delta = 0.999;
+beta = 1;
+numberIterations = 10 ^ 7;
+payoff_type = "expected";
+payoffs = readmatrix("two_by_two_games_values.csv");
+
+parfor (i = 1:11)
+    u = payoffs(i, :);
+    filename = "../data/two_by_two_games_values_S" + u(2) + "_T_" + u(2);
+    evolSimulation(starting_resident, u, N, delta, beta, numberIterations, payoff_type, filename);
+end
+end
+

### 2023-03-07 matlab/cooperation.m

@@ -1,0 +1,14 @@
+function [coop]=cooperation(Mu, h, sz, delta);
+
+coop = zeros(sz, 1);
+
+for i=1:sz
+    player_coop = zeros(sz, 1);
+    for j=1:sz
+        v=stationary(Mu(i, :), Mu(j, :), delta);
+        player_coop(j) = v(1) + v(2);
+    end
+    coop(i) = mean(player_coop); 
+end
+coop = sum(coop .* h / sum(h));
+end

### 2023-03-07 matlab/evolMutation.m

@@ -1,0 +1,66 @@
+function [population, avg_player, coop]=evolMutation(population, u, N, delta, beta, mutation, numberIterations, payoff_type, filename);
+sdim=3; avg_player=0; coop=0;
+
+xPlayer = zeros(numberIterations, 3);
+xCoop = zeros(numberIterations, 1);
+
+%% Running the evolutionary process
+for t = progress(1:numberIterations)
+    if rand(1) < mutation
+        idx = randi(N);
+        population(idx, :) = rand(1, sdim);
+    else
+        indices = randi(N, [1, 2]);
+
+        if population(indices(1),:) ~= population(indices(2),:)
+            if payoff_type == "expected"
+                payoffs = payoffsExpected(indices, population, u, N, delta);
+                resident_payoff = payoffs(1);
+                mutant_payoff = payoffs(2);
+            end
+            if payoff_type == "last_round"
+                payoffs = payoffsLastRound(indices, population, u, N, delta);
+                resident_payoff = payoffs(1);
+                mutant_payoff = payoffs(2);
+
+            end
+
+            fermi = 1 / (1 + exp(-beta * (mutant_payoff - resident_payoff)));
+            if rand(1) < fermi
+                population(indices(1), :) = population(indices(2), :);
+            end
+        end
+    end
+    
+  [Mu,~,ic] = unique(population, 'rows', 'stable'); 
+  h = accumarray(ic,1);
+  sz = size(h, 1);
+  avg_player = sum(Mu .* h / sum(h), 1);
+  coop = cooperation(Mu, h, sz, delta);
+
+  xPlayer(t, 1) = avg_player(1);
+  xPlayer(t, 2) = avg_player(2);
+  xPlayer(t, 3) = avg_player(3);
+
+  xCoop(t, 1) = coop;
+
+end
+
+Data=["R", "S", "T", "P", "N", "beta", "mutation", "numberIterations";
+      u(1),...
+      u(2),...
+      u(3),...
+      u(4),...
+      N,...
+      beta,...
+      mutation,...
+      numberIterations];
+
+writematrix(Data, filename + ".csv");
+writematrix(xPlayer, filename + "avg_player.csv");
+writematrix(xCoop, filename + "cooperations.csv");
+writematrix(population, filename + "population.csv");
+end
+
+
+

### 2023-03-07 matlab/evolRun.m

@@ -5,13 +5,12 @@ N = 100;
 delta = 0.999;
 beta = 1;
 numberIterations = 10 ^ 7;
-payoff_types = ["two_rounds", "two_rounds_opponents"]
+payoff_type = "expected";
+payoffs = readmatrix("two_by_two_games_values.csv");
 
-u = [2, -1, 3, 0];
-
-parfor (i = 1:2)
-    payoff_type = payoff_types(i);
-    filename = "../data/sixteen_states_" + payoff_type;
-    evolSimulationTwoRounds(starting_resident, u, N, delta, beta, numberIterations, payoff_type, filename);
+parfor (i = 1:11)
+    u = payoffs(i, :);
+    filename = "../data/two_by_two_games_values_S" + u(2) + "_T_" + u(2);
+    evolSimulation(starting_resident, u, N, delta, beta, numberIterations, payoff_type, filename);
 end
 end

### 2023-03-07 matlab/evolRunMemoryOne.m

@@ -5,13 +5,21 @@ N = 100;
 delta = 0.999;
 beta = 1;
 numberIterations = 10 ^ 7;
-payoff_types = ["expected", "last_round"]
+payoff_type = "expected" % "last_round"
 
-u = [2, -1, 3, 0];
+us = [1, -1, 2, 0;
+      2, -1, 3, 0;
+      3, -1, 4, 0;
+      4, -1, 5, 0;
+      5, -1, 6, 0;
+      6, -1, 7, 0;
+      7, -1, 8, 0;
+      8, -1, 9, 0;
+      9, -1, 10, 0];
 
-parfor (i = 1:2)
-    payoff_type = payoff_types(i);
-    filename = "../data/memory_one_" + payoff_type;
+parfor (i = 1:9)
+    u = us(i, :);
+    filename = "../data/memory_one_run_two" + payoff_type + "_b_" + u(3);
     evolSimulationMemoryOne(starting_resident, u, N, delta, beta, numberIterations, payoff_type, filename);
 end
 end

### 2023-03-07 matlab/evolRunRoundsTwo.m

@@ -1,0 +1,15 @@
+function evolRunRoundsTwo();
+
+starting_resident = [0, 0, 0];
+N = 100;
+delta = 0.999;
+beta = 1;
+numberIterations = 10 ^ 7;
+payoff_type = "two_rounds";
+u = [2, -1, 3, 0];
+
+filename = "../data/two_rounds_low_iterations";
+
+evolSimulationTwoRounds(starting_resident, u, N, delta, beta, numberIterations, payoff_type, filename);
+
+end

### 2023-03-07 matlab/evolSimulation.m

@@ -1,5 +1,4 @@
 function [xDat, AvCoop, AvPay,payoff_type, Data]=evolSimulation(starting_resident, u, N, delta, beta, numberIterations, payoff_type, filename);
-
 rng('default')
 %% Preparations for the output
 Data=['R=',num2str(u(1)),'; S=',num2str(u(2)),'; T=',num2str(u(3)), '; P=',num2str(u(4)),'; N=',num2str(N),'; beta=',num2str(beta), '; nIt=',num2str(numberIterations), '; payoff=',payoff_type];
@@ -26,7 +25,6 @@ writematrix(Data, filename + ".txt");
 
 AvCoop = mean(xDat(:,end-2));
 AvPay = mean(xDat(:,end-1));
-%time=toc
 end
 
 function [Rho]=calcRho(u, beta);
@@ -60,10 +58,6 @@ elseif payoff_type=="expected"
     piRM=vRM*u';
     piRR=vRR*u';
     phi = phiExpected(N, piMM, piMR, piRR, piRM, beta);
-
-elseif payoff_type=="two_opponents"
-
-    phi = phiTwoOpponents(N, vRM, vMM, vMR, vRR, Rho);
 
 else
     disp('Please check payoff type.')

### 2023-03-07 matlab/evolSimulationMemoryOne.m

@@ -1,6 +1,5 @@
 function [xDat, AvCoop, AvPay,payoff_type, Data]=evolSimulationMemoryOne(starting_resident, u, N, delta, beta, numberIterations, payoff_type, filename);
 
-rng('default')
 %% Preparations for the output
 Data=['R=',num2str(u(1)),'; S=',num2str(u(2)),'; T=',num2str(u(3)), '; P=',num2str(u(4)),'; N=',num2str(N),'; beta=',num2str(beta), '; nIt=',num2str(numberIterations), '; payoff=',payoff_type];
 AvCoop=0; AvPay=0; Res=starting_resident;

### 2023-03-07 matlab/evolSimulationTwoRounds.m

@@ -1,12 +1,11 @@
-function [xDat, AvCoop, AvPay,payoff_type, Data]=evolSimulation(starting_resident, u, N, delta, beta, numberIterations, payoff_type, filename);
-
+function [xDat, AvCoop, AvPay,payoff_type, Data]=evolSimulationTwoRounds(starting_resident, u, N, delta, beta, numberIterations, payoff_type, filename);
 %% Preparations for the output
 Data=['R=',num2str(u(1)),'; S=',num2str(u(2)),'; T=',num2str(u(3)), '; P=',num2str(u(4)),'; N=',num2str(N),'; beta=',num2str(beta), '; nIt=',num2str(numberIterations), '; payoff=',payoff_type];
 AvCoop=0; AvPay=0; Res=starting_resident;
 
 %% Initialization
 sdim=3;
-xDat=zeros(numberIterations/100,6);
+xDat=zeros(numberIterations,6);
 xDat(1,:)=[Res, 0, u(4), 0];
 Rho=calcRhoSixteen(u, beta);
 
@@ -20,12 +19,11 @@ for t = progress(1:numberIterations)
     end
 end
 
-csvwrite(filename + ".csv", xDat);
+dlmwrite(filename + ".csv", xDat, 'delimiter', ',', 'precision', 9);
 writematrix(Data, filename + ".txt");
 
 AvCoop = mean(xDat(:,end-2));
 AvPay = mean(xDat(:,end-1));
-%time=toc
 end
 
 function [Rho]=calcRhoSixteen(u, beta);
@@ -48,24 +46,34 @@ function [Rho]=calcRhoSixteen(u, beta);
     end
 end
 
+
 function [phi, coopMM, piMM]=calcPhi(Mut, Res, Rho, N, u, delta, beta, payoff_type);
 %% Calculating the fixation probability
 
-vMM = stationaryRoundTwo(Mut, Mut, delta);
-vMR = stationaryRoundTwo(Mut, Res, delta);
-vRM = stationaryRoundTwo(Res, Mut, delta);
-vRR = stationaryRoundTwo(Res, Res, delta);
-
-piMM = vMM*log(kron(exp(u),exp(u)))';
[…]

### 2023-03-07 matlab/evolSimulationTwoRoundsOpponents.m

@@ -1,0 +1,81 @@
+function [xDat, AvCoop, AvPay,payoff_type, Data]=evolSimulationTwoRoundsOpponents(starting_resident, u, N, delta, beta, numberIterations, payoff_type, filename);
+%% Preparations for the output
+Data=['R=',num2str(u(1)),'; S=',num2str(u(2)),'; T=',num2str(u(3)), '; P=',num2str(u(4)),'; N=',num2str(N),'; beta=',num2str(beta), '; nIt=',num2str(numberIterations), '; payoff=',payoff_type];
+AvCoop=0; AvPay=0; Res=starting_resident;
+
+%% Initialization
+sdim=3;
+xDat=zeros(round(numberIterations/(100 * 0.6)), 5);
+xDat(1,:)=[Res, 0, 0];
+Rho=calcRhoTwoFiftySix(u, beta);
+
+ps = zeros(256, 2);
+for i1=1:256
+    ps(i1, 1) = 1 + fix((i1 - 1) / 16);
+    ps(i1, 2) = 1 + mod(i1 - 1, 16);
+end
+
+condition_round_one = zeros(256, 256);
+condition_round_two = zeros(256, 256);
+
+for i1=1:256
+    for i2=1:256
+        condition_round_one(i1, i2) = ((ps(i1, 1)==1 & ps(i2, 1)==1) |(ps(i1, 1)==2 & ps(i2, 1)==3) |(ps(i1, 1)==3 &ps(i2, 1)==2) |(ps(i1, 1)==4 & ps(i2, 1)==4) |(ps(i1, 1)==5 & ps(i2, 1)==9) | (ps(i1, 1)==6 & ps(i2, 1)==11)|...
+        (ps(i1, 1)==7 & ps(i2, 1)==10) |(ps(i1, 1)==8 & ps(i2, 1)==12) |(ps(i1, 1)==9 & ps(i2, 1)==5)  |(ps(i1, 1)==10 & ps(i2, 1)==7)|(ps(i1, 1)==11 & ps(i2, 1)==6)|(ps(i1, 1)==12 & ps(i2, 1)==8)|...
+        (ps(i1, 1)==13 & ps(i2, 1)==13)|(ps(i1, 1)==14 & ps(i2, 1)==15)|(ps(i1, 1)==15 & ps(i2, 1)==14)|(ps(i1, 1)==16 & ps(i2, 1)==16));
+
[…]

### 2023-03-07 matlab/payoffsExpected.m

@@ -1,0 +1,16 @@
+function [payoffs]=payoffsExpected(player_indices, population, u, N, delta);
+
+
+payoffs = zeros(N - 1, 2);
+
+for i=1:N 
+    for j=1:2
+        if i ~= player_indices(j)
+            v=stationaryReactive(population(player_indices(j), :), population(i, :), delta);
+            payoffs(i, j) = v * u';
+        end
+    end
+end
+
+payoffs = sum(payoffs / (N - 1));
+end

### 2023-03-07 matlab/payoffsLastRound.m

@@ -1,0 +1,24 @@
+function [payoffs]=payoffsLastRound(players_indices, population, u, N, delta);
+
+payoffs = zeros(1, 2);
+exclude = zeros(1, 3);
+
+for i=1:2
+    exclude(i)  = players_indices(i);
+    opponent_id = randsample(setdiff(1:N, exclude), 1);
+
+    player = population(players_indices(i), :);
+    opponent = population(opponent_id, :);
+    v=stationary(player, opponent, delta);
+    
+    if opponent_id == players_indices(2)
+        payoffs(1) = randsample(u, 1, true, v);
+        payoffs(2) = (payoffs(1) == u) * [u(1), u(3), u(2), u(4)]';
+        break
+    else
+        payoffs(i) = randsample(u, 1, true, v);
+        exclude(3) = opponent_id;
+    end
+
+end
+end

### 2023-03-07 matlab/phiTwoOpponents.m

@@ -2,24 +2,29 @@ function [phi]=phiTwoOpponents(N, vRM, vMM, vMR, vRR, Rho);
 %% Calculates the fixation probability based on the last round payoff with two opponents
 laplus=zeros(1,N-1); laminus=laplus;
 for k=1:N-1
-    x=zeros(4,4);
-    for i1=1:4
-        for i2=1:4
+    x=zeros(16, 16);
+    for i1=1:16
+        for i2=1:16
 
-learner_role_model = 1 / (N - 1) * vRM(i1) * ((i1==1 & i2==1) | (i1==2 & i2==3) | (i1==3 &i2==2) | (i1==4 & i2==4)) * ( ...
-1 / (N - 2) / (N - 3) * ((k - 1) * (k - 2) * vRM(i1) * vMM(i2) + (k - 1) * (N - k - 1) * vRM(i1) * vMR(i2) + (N - k - 1) * (k - 1) * vRR(i1) * vMM(i2) + (N - k - 1) * (N - k - 2) * vRR(i1) * vMR(i2)));
+            p_11 = 1 + fix((i1 - 1) / 4);
+            p_12 = 1 + mod(i1 - 1, 4);
+            q_21 = 1 + fix((i2 - 1) / 4);
+            q_22 = 1 + mod(i2 - 1, 4);
 
-mutant_mutant = (1 - 1 / (N - 1)) / (N - 2) / (N - 3) * ((k-1) * (k-2) * vRM(i1) * vMM(i2)) * ((1 - 1 / (N - 2)) / (N - 3) / (N - 4) * ((k-2)*(k-3)*vRM(i1)*vMM(i2) + (k-2)*(N-k-1)*vRR(i1)*vMM(i2) + (N-k-1)*(k-2)*vRM(i1)*vMR(i2) + (N-k-1)*(N-k-2)*vRR(i1)*vMR(i2)) + ...
-1 / (N - 2) * vRM(i1) * ((i1==1 & i2==1) | (i1==2 & i2==3) | (i1==3 &i2==2) | (i1==4 & i2==4)));
[…]

### 2023-03-07 matlab/phiTwoRounds.m

@@ -1,22 +1,19 @@
-function [phi]=phiTwoRounds(N, vRM, vMM, vMR, vRR, Rho);
+function [laplus, laminus]=phiTwoRounds(N, vRM, vMM, vMR, vRR, Rho);
 %% Calculates the fixation probability based on the last two rounds payoff
 
-laplus = zeros(1, N-1); laminus=laplus;
-for k=1:N-1
-    x=zeros(16,16);
-    for i1=1:16
-        for i2=1:16
+laplus = zeros(1, 1); laminus=laplus;
+k=1;
+x=zeros(16,16);
+for i1=1:16
+    for i2=1:16
 x(i1,i2) = 1 / (N-1) * vRM(i1) * ((i1==1 & i2==1)  |(i1==2 & i2==3)  |(i1==3 &i2==2)   |(i1==4 & i2==4) |(i1==5 & i2==9) |(i1==6 & i2==11)|...
-                                  (i1==7 & i2==10) |(i1==8 & i2==12) |(i1==9 & i2==5)  |(i1==10 & i2==7)|(i1==11 & i2==6)|(i1==12 & i2==8)|...
-                                  (i1==13 & i2==13)|(i1==14 & i2==15)|(i1==15 & i2==14)|(i1==16 & i2==16))...
-         + (1 - 1 / (N - 1)) / (N - 2) / (N - 3) * ((k - 1) * (k - 2) * vRM(i1) * vMM(i2) ...
-         + (k - 1) * (N - k - 1) * vRM(i1) * vMR(i2) + (N - k - 1) * (k - 1) * vRR(i1) * vMM(i2) ...
-         + (N - k - 1) * (N - k - 2) * vRR(i1) * vMR(i2));
-        end
+                              (i1==7 & i2==10) |(i1==8 & i2==12) |(i1==9 & i2==5)  |(i1==10 & i2==7)|(i1==11 & i2==6)|(i1==12 & i2==8)|...
+                              (i1==13 & i2==13)|(i1==14 & i2==15)|(i1==15 & i2==14)|(i1==16 & i2==16))...
+     + (1 - 1 / (N - 1)) / (N - 2) / (N - 3) * ((k - 1) * (k - 2) * vRM(i1) * vMM(i2) ...
+     + (k - 1) * (N - k - 1) * vRM(i1) * vMR(i2) + (N - k - 1) * (k - 1) * vRR(i1) * vMM(i2) ...
[…]

### 2023-03-07 matlab/phiTwoRoundsOpponents.m

@@ -1,35 +1,25 @@
-function [phi]=phiTwoRoundsOpponents(N, vRM, vMM, vMR, vRR, Rho);
+function [phi]=phiTwoRoundsOpponents(N, vRM, vMM, vMR, vRR, Rho, condition_round_one, condition_round_two, ps);
 %% Calculates the fixation probability based on the last two rounds payoff with two opponents
-
 laplus = zeros(1, N-1); laminus=laplus;
 for k=1:N-1
-    x=zeros(16,16);
-    for i1=1:16
-        for i2=1:16
-learner_role_model = 1 / (N - 1) * vRM(i1) * ((i1==1 & i2==1)  |(i1==2 & i2==3)  |(i1==3 &i2==2)   |(i1==4 & i2==4) |(i1==5 & i2==9) |(i1==6 & i2==11)|...
-                                  (i1==7 & i2==10) |(i1==8 & i2==12) |(i1==9 & i2==5)  |(i1==10 & i2==7)|(i1==11 & i2==6)|(i1==12 & i2==8)|...
-                                  (i1==13 & i2==13)|(i1==14 & i2==15)|(i1==15 & i2==14)|(i1==16 & i2==16)) * ( ...
-1 / (N - 2) / (N - 3) * ((k - 1) * (k - 2) * vRM(i1) * vMM(i2) + (k - 1) * (N - k - 1) * vRM(i1) * vMR(i2) + (N - k - 1) * (k - 1) * vRR(i1) * vMM(i2) + (N - k - 1) * (N - k - 2) * vRR(i1) * vMR(i2)));
+    x=zeros(256, 256);
+    for i1=1:256
+        for i2=1:256
 
[…]

### 2023-03-07 matlab/stationaryReactive.m

@@ -1,0 +1,11 @@
+function v=stationaryReactive(p1,p2,delta);
+
+M=[p1(2)*p2(2), p1(2)*(1-p2(2)), (1-p1(2))*p2(2), (1-p1(2))*(1-p2(2));
+    p1(3)*p2(2), p1(3)*(1-p2(2)), (1-p1(3))*p2(2), (1-p1(3))*(1-p2(2));
+    p1(2)*p2(3), p1(2)*(1-p2(3)), (1-p1(2))*p2(3), (1-p1(2))*(1-p2(3));
+    p1(3)*p2(3), p1(3)*(1-p2(3)), (1-p1(3))*p2(3), (1-p1(3))*(1-p2(3))];
+
+v0=[p1(1)*p2(1), p1(1)*(1-p2(1)), (1-p1(1))*p2(1), (1-p1(1))*(1-p2(1))];
+
+v = (1 - delta) * v0 / (eye(4) - delta * M);
+end

### 2023-03-07 scripts/clean_data.py

@@ -1,0 +1,51 @@
+import sys
+
+import glob
+
+import tqdm
+
+import pandas as pd
+
+if __name__ == "__main__":  # pragma: no cover
+    folder = sys.argv[1]
+    starting_characters = "beta_0.01_stochastic_1"  # sys.argv[2]
+
+    in_columns = [
+        "y",
+        "p",
+        "q",
+        "average cooperation rate",
+        "average score",
+        "t",
+    ]
+
+    # in_columns = [
+    #     "p_1",
+    #     "p_2",
+    #     "p_3",
+    #     "p_4",
+    #     "y",
+    #     "average cooperation rate",
+    #     "average score",
+    #     "t",
+    # ]
+
+    print(f"{folder}/{starting_characters}.csv")
+    csv_files = glob.glob(f"{folder}/{starting_characters}*.csv")
+
+    for file in tqdm.tqdm(csv_files):
+        df = pd.read_csv(file, header=None)
+        df = df.drop_duplicates()
+        df.columns = in_columns
+
+        frequencies = []
+        previous = 0
+        for value in df["t"][1:]:
+            frequencies.append(value - previous)
+            previous = value
+        frequencies.append(10 ** 7 - previous)
+
+        filename = file.split("/")[-1]
+        df["frequencies"] = frequencies
+        print(f"{folder}/clean_{filename}")
+        df.to_csv(f"{folder}/clean_{filename}", index=False)

### 2023-03-07 src/evol_dynamics/mutation.py

@@ -1,0 +1,184 @@
+import random
+
+import numpy as np
+
+# import tqdm
+import copy
+
+import evol_dynamics
+
+import time
+
+
+def payoffs_expected(resident, mutant, population, payoffs, delta, N):
+
+    resident_payoff = sum(
+        [
+            evol_dynamics.expected_distribution_last_round(
+                resident, opponent, delta
+            )
+            @ payoffs
+            * (population[opponent] - int(resident == opponent))
+            for opponent in population.keys()
+        ]
+    ) / (N - 1)
+
+    mutant_payoff = sum(
+        [
+            evol_dynamics.expected_distribution_last_round(
+                mutant, opponent, delta
+            )
+            @ payoffs
+            * (population[opponent] - int(mutant == opponent))
+            for opponent in population.keys()
+        ]
+    ) / (N - 1)
+
+    return resident_payoff, mutant_payoff
+
+
+def payoffs_last_round(resident, mutant, population, payoffs, delta):
+
+    population_cp = copy.deepcopy(population)
+
+    opponent = random.sample(
+        [key for key in population.keys() for _ in range(population[key])], 1
+    )[0]
+    population_cp[opponent] -= 1
+
+    resident_payoff = (
+        evol_dynamics.expected_distribution_last_round(
+            resident, opponent, delta
+        )
+        @ payoffs
+    )
+
+    opponent = random.sample(
+        [key for key in population.keys() for _ in range(population[key])], 1
+    )[0]
+    mutant_payoff = (
+        evol_dynamics.expected_distribution_last_round(mutant, opponent, delta)
+        @ payoffs
+    )
+
+    return resident_payoff, mutant_payoff
+
+
+def simulation(
+    N,
+    delta,
+    beta,
+    mutation,
+    number_of_steps,
+    payoffs,
+    mode,
+    filename,
+    seed=10,
+    starting_resident=(0, 0, 0),
+):
+
+    data = [0, *payoffs, N, delta, beta, mutation, mode, *starting_resident, 0]
+
+    with open(filename, "w") as textfile:
+        textfile.write(",".join([str(elem) for elem in data]) + "\n")
+    textfile.close()
+
+    population = {starting_resident: N}
+    random.seed(seed)
+
+    for i in range(1, number_of_steps + 1):
+
+        if random.random() < mutation:
+            to_mutate = random.sample(
+                [
+                    key
+                    for key in population.keys()
+                    for _ in range(population[key])
+                ],
[…]

### 2023-09-26 paper/main.tex
CH: Editing the title page of the paper

@@
 
-\title{\bf  \sffamily \LARGE Evolution of cooperation among individuals with
+\title{\bf  \sffamily \LARGE Evolution of reciprocity among individuals\\ with
 limited payoff memory\\}
 \date{}
-\author{Nikoleta E. Glynatsi, Christian Hilbe, Alex McAvoy}
+\author{\parbox[c]{16cm}{\centering \onehalfspacing 
+Nikoleta E. Glynatsi$^1$, Christian Hilbe$^1$, Alex McAvoy$^{2,3}$\\
+$^{\rm 1}$Max Planck Research Group on the Dynamics of Social Behavior,\\ Max Planck Institute for Evolutionary Biology, Pl\"{o}n, Germany \\
+$^{\rm 2}$School of Data Science and Society, University of North Carolina at Chapel Hill,\\ Chapel Hill, NC 27599 \\
+$^{\rm 3}$Department of Mathematics, University of North Carolina at Chapel Hill,\\ Chapel Hill, NC 27599}}
 
@@
 \begin{abstract}
-Repeated games are vastly used in evolutionary game theory to explain cooperation
-in a variety of environments. Although existing models have shaped our
-understanding of human cooperation they often work with idealized assumptions.
-In an evolutionary process, individuals imitate other individuals of the
-population based on their fitness. It is commonly assumed that an individual
-computes their fitness after interacting with a representative sample of the
-population and remembering all the interactions they participated in. In real
-life we do not always remember all our interactions, instead we rather recall
-our most recent ones.
+\noindent
+Models of direct reciprocity explain how cooperation can emerge in repeated social dilemmas. 
+According to this literature, individuals should naturally learn to adopt conditionally cooperative strategies if further encounters are sufficiently likely.  
+While such models have shaped our understanding of cooperation, they often make idealized assumptions on how individuals remember and process information. 
+For example, when strategies are transmitted through social learning, it is commonly assumed that individuals update their strategies based on the resulting average payoff.  
+This would require them to compute (or remember) their payoffs against all other population members. 
+Instead, individuals may sometimes update their strategies based on their most recent experiences only. 
+Herein, we introduce a theoretical framework to study the evolution of reciprocity when individuals only learn based on the outcome of their very last interaction. 
+%We compare this model to the classical framework based on expected payoffs.  
+We show that cooperation can still evolve, but individuals adopt less generous strategies, and they tend to cooperate less often than in the classical framework based on expected payoffs. 
+Interestingly, however, we also show that once individuals remember the payoffs of two or three interactions, evolving cooperation rates approach the results of the classical model. 
+These findings contribute to a literature that explores which kind of cognitive capabilities are required for reciprocal cooperation.
+While our results suggest that more memory facilitates the evolution of reciprocity, it only takes a modest amount of payoff memory for cooperation to emerge.
+\end{abstract}
 
-Here we introduce a framework that allows individuals to estimate their fitness
-based on a minimum of social information. We explore the difference between our
-framework and the classical framework using the iterated prisoner's dilemma. We
-show that individuals relying on a single set of information tend to adopt less
-generous strategies and they achieve less cooperation than in the classical
-scenario. Although strategies in the latter scenario become more cooperative
-with increasing benefits, those in the former remain unchanged. We further
-explore scenarios where individuals are granted slightly more information, and
-our observations indicate that these individuals can attain higher levels of
-cooperation.
+~\\
+{\it Keywords:} Evolution of cooperation; direct reciprocity; repeated prisoner's dilemma; social learning; evolutionary dynamics
 
-\end{abstract}
+\clearpage
+\newpage
 

### 2023-09-26 paper/main.tex
CH: Polishing the abstract

@@
 
-\title{\bf  \sffamily \LARGE Evolution of reciprocity among individuals\\ with
-limited payoff memory\\}
+\title{\bf  \sffamily \Large Evolution of reciprocity %among individuals\\ 
+with limited payoff memory\\}
 \date{}
 \author{\parbox[c]{16cm}{\centering \onehalfspacing 
-Nikoleta E. Glynatsi$^1$, Christian Hilbe$^1$, Alex McAvoy$^{2,3}$\\
+Nikoleta E. Glynatsi$^1$,  Alex McAvoy$^{2,3}$, Christian Hilbe$^1$\\
 $^{\rm 1}$Max Planck Research Group on the Dynamics of Social Behavior,\\ Max Planck Institute for Evolutionary Biology, Pl\"{o}n, Germany \\
@@
 Models of direct reciprocity explain how cooperation can emerge in repeated social dilemmas. 
-According to this literature, individuals should naturally learn to adopt conditionally cooperative strategies if further encounters are sufficiently likely.  
-While such models have shaped our understanding of cooperation, they often make idealized assumptions on how individuals remember and process information. 
-For example, when strategies are transmitted through social learning, it is commonly assumed that individuals update their strategies based on the resulting average payoff.  
-This would require them to compute (or remember) their payoffs against all other population members. 
+According to this literature, individuals should naturally learn to adopt conditionally cooperative strategies if they expect to have further encounters with their interaction partner. 
+While models of reciprocity have shaped our understanding of cooperation, they often make idealized assumptions on how individuals remember and process information. 
+For example, when strategies change through social learning, it is commonly assumed that individuals take into account the strategy's average payoff.  
+However, this would require them to compute (or remember) their payoffs against all other population members.
 Instead, individuals may sometimes update their strategies based on their most recent experiences only. 

### 2023-10-01 paper/main.tex
CH: Polishing the abstract, drafting the intro

@@
 \usepackage{amsmath,amsthm,amssymb,setspace,enumitem,epsfig,titlesec,verbatim,color,array,eurosym,multirow}
-\usepackage[sort&compress]{natbib}
+\usepackage[sort&compress,round
+]{natbib}
 \usepackage[footnotesize,bf]{caption}
@@
 
-\definecolor{darkblue}{rgb}{0,0,.8}
-\definecolor{darkgreen}{rgb}{0,0.5,0.1}
-\newcommand{\matlabfunc}[1]{\textcolor{darkblue}{#1}}
-\newcommand{\matlabcomment}[1]{\textcolor{darkgreen}{#1}}
-\newcommand{\christian}[1]{\textcolor{blue}{\textbf{CH}: #1}}
-\newcommand{\alex}[1]{\textcolor{red}{\textbf{AL}: #1}}
+\newcommand{\todo}[1]{\textcolor{blue}{#1}}
 
 %% Adding shortcut commands to refer to our figures %%
-\newcommand{\FigEvoProc}{{\bf Fig.~1}} \newcommand{\FigInvAnalysis}{{\bfFig.~2}} \newcommand{\FigResultsOverPara}{{\bf Fig.~3}}
+\newcommand{\FigEvoProc}{{\bf Fig.~1}} 
+\newcommand{\FigInvAnalysis}{{\bfFig.~2}} 
+\newcommand{\FigResultsOverPara}{{\bf Fig.~3}}
 
@@
 \noindent
-Models of direct reciprocity explain how cooperation can emerge in repeated social dilemmas. 
-According to this literature, individuals should naturally learn to adopt conditionally cooperative strategies if they expect to have further encounters with their interaction partner. 
-While models of reciprocity have shaped our understanding of cooperation, they often make idealized assumptions on how individuals remember and process information. 
-For example, when strategies change through social learning, it is commonly assumed that individuals take into account the strategy's average payoff.  
-However, this would require them to compute (or remember) their payoffs against all other population members.
-Instead, individuals may sometimes update their strategies based on their most recent experiences only. 
-Herein, we introduce a theoretical framework to study the evolution of reciprocity when individuals only learn based on the outcome of their very last interaction. 
-%We compare this model to the classical framework based on expected payoffs.  
-We show that cooperation can still evolve, but individuals adopt less generous strategies, and they tend to cooperate less often than in the classical framework based on expected payoffs. 
-Interestingly, however, we also show that once individuals remember the payoffs of two or three interactions, evolving cooperation rates approach the results of the classical model. 
+Direct reciprocity can explain how cooperation emerges in repeated social interactions. 
+According to this literature, individuals should naturally learn to adopt conditionally cooperative strategies such as Tit-for-Tat.
+% if they expect to have further encounters with their interaction partner.
+Corresponding models have greatly facilitated our understanding of cooperation. 
+Yet they often make strong assumptions on how individuals remember and process information. 
+For example, when strategies are updated through social learning, it is commonly assumed that individuals update their strategies based on their average payoffs.  
+This would require them to compute (or remember) their payoffs against all other population members.
+Instead, herein, we introduce a theoretical framework to study the evolution of reciprocity when individuals learn based on their most recent experiences.
+Even in the most extreme case that they only take into account their very last interaction, we find that cooperation can still evolve. 
+However, such individuals adopt less generous strategies, and they tend to cooperate less often than in the classical framework with expected payoffs. 
+Interestingly, once individuals remember the payoffs of two or three recent interactions, evolving cooperation rates quickly approach the classical limit. 
 These findings contribute to a literature that explores which kind of cognitive capabilities are required for reciprocal cooperation. 
-While our results suggest that more memory facilitates the evolution of reciprocity, it only takes a modest amount of payoff memory for cooperation to emerge.
+While our results suggest that memory facilitates the evolution of reciprocity, it only takes a rather modest amount of payoff memory for cooperation to emerge. 
 \end{abstract}
@@
 
+%%%%%%%%%
+%%  INTRO  %%
+%%%%%%%%%
+
 \section{Introduction}
 
-Evolutionary game theory~\cite{smith1982evolution, hofbauer1998evolutionary,
-nowak:Nature:2004, hauert2005game} describes the evolutionary dynamics of
-populations consisting of different types of interacting individuals. In the
-last two decades the interest of the field has shifted to the analysis of
-stochastic finite population dynamics in preference to traditional approaches
-such as the replicator dynamics~\cite{hofbauer:JTB:1979, taylor:MATHBIO:1978}.
-This is especially true in the topic of cooperation~\cite{hilbe:PNAS:2013,
-hilbe:Nature:2018,glynatsi:SCR:2020, ohtsuki:Nature:2006, nowak:Science:2006,
-murase:ScientificReports:2020}. In stochastic evolutionary dynamics
-disadvantageous mutants have a small yet non-zero probability to reach fixation
-and this can lead to fundamental changes in the results. As shown
-in~\cite{nowak:Nature:2004}, using such dynamics a single cooperative strategy
-can invade a population of defectors without any special modifications in
-comparison to traditional approaches that one has to consider spatial
-structure~\cite{nowak:Nature:1992} or when payoffs are subject to aggregate
-shocks~\cite{fudenberg:JET:1992}.
+%% INTRO: General introduction to evolutionary game theory and cognitive constrains
 
-Stochastic evolutionary dynamics model a finite population. Each
-time step of the process consists of three phases; (1) the \textit{mutation
-phase} (2) the \textit{game phase} (3) the \textit{update phase}. In the
-mutation phase one individual from the population is chosen to switch to a new
-mutant strategy with a probability \(\mu\). In the game phase individuals are
-randomly matched with other individuals in the population, and they engage in a
-repeated game where each subsequent turn occurs with a fixed probability
-\(\delta\). The updating phase depends on the process. Two classes of
-finite stochastic processes have been used extensively: (i) fitness-based
-processes in which an individual chosen proportional to fitness reproduces and
-the offspring replaces a randomly chosen individual~\cite{nowak:Nature:2004} or
-(ii) \textit{pairwise comparison processes} in which a pair of individuals is
-chosen and subsequently one of these individuals may adopt the strategy of the
-other~\cite{traulsen2007pairwise}.
+Evolutionary game theory describes the dynamics of populations when an individual's fitness depends on the traits or strategies of other population members~\cite{hofbauer1998evolutionary, nowak:Nature:2004, hauert2005game}.
+This theory can be used to describe the dynamics of animal conflict~\citep{maynard-smith:Nature:1973}, cancer cells~\citep{Stankova:JAMA:2019}, and of cooperation~\citep{axelrod1981evolution}. 
+Respective models translate strategic interactions into games~\cite{smith1982evolution}. 
+These games specify how individuals (players) interact, which strategies they can choose, and what fitness consequences (or payoffs) their strategies have. 
+In addition, these models also specify the mode by which successful strategies spread over time. 
+Models of biological evolution posit that individuals with a high fitness tend to produce more offspring; models of cultural evolution instead assume that such individuals are imitated more often. 
+While biological and cultural evolution are sometimes treated as equivalent, there can be important differences~\citep{Wu2015,Smolla:PTRS:2021}. 
+For example, models of biological evolution do not require individuals to have any particular cognitive abilities.
+Here, it is the evolutionary process itself that biases the population towards strategies with higher fitness. 
+In contrast, in models of cultural evolution, individuals need to be aware of the different strategies that are available, and they need to identify those strategies with a higher payoff. 
+As a consequence, evolutionary outcomes may often depend on how easily different behaviors can be learned~\citep{Chatterjee:JTB:2012}, and on how easy payoff comparisons are.
 
-Like all theoretical models, finite stochastic processes rest on a
-set of assumptions. Namely, in the game phase it is assumed that players use
-strategies with finite memory~\cite{Nowak1992tit, Baek2016}. For example, in a
-repeated game between two individuals the actions of the players in each turn
-are often determined by the action of the co-players in the previous turn~\cite{Nowak1992tit}. This
-assumption is common as it allows for an explicit calculation of the players'
-payoffs~\cite{sigmund2010calculus}. In addition, it is
-assumed that individuals play many times and with all other players before
-reproduction takes place. So that in the updating phase the updating payoffs,
-or fitness, of individuals is given by the average payoffs of an
-infinitely repeated game against every member of the population.
+%% INTRO: Literature on direct reciprocity and why social learning of strategies is hard
 
-These two assumptions create a curious inconsistency; a player with a finite
-memory (in the game stage) can recall infinite information regarding their
-interactions and each interaction's outcomes in the update stage. Prior research
-has examined the effects of constraining the interactions an individual has.
-Either by letting individuals have a stochastic number of
-interactions~\cite{roca:PhysicalReview:2006, sanchez:JTB:2005,
-Traulsen:JTB:2007} or by only allowing for a single
-interaction~\cite{Woelfing:JTB:2009}. However, these studies have not accounted
-for the use of strategies with finite memory. Moreover, no previous research has
-explored the effect of restricting not only the number of interactions, but also
-the amount of information available about each interaction. To this end, we
-propose a framework in which individuals, similar to the decisions at each turn,
-estimate their fitness based on a minimum of information.
+These difficulties -- to learn strategies by social imitation -- are particularly pronounced in models of direct reciprocity. 
+This literature follows Trivers' insight that individuals have more of an incentive to cooperate in social dilemmas when they interact repeatedly~\citep{trivers1971evolution}. 
+In repeated interactions, individuals can condition their future behavior on their past experiences with their interaction partner. 
+They may use strategies such as Tit-for-Tat~\citep{rapoport:book:1965} or Generous Tit-for-Tat~\citep{molander:jcr:1985,Nowak1992tit} to preferentially cooperate with those partners who cooperated in the past.
+Such conditional strategies approximate human behavior fairly well~\citep{fischbacher:EconL:2001,Rand:TCS:2013,DalBo:AER:2019,Rossetti:ETH:2023} and they have also been documented in several other species~\citep{Carter:PRSB:2013,Schweinfurth:AnBehav:2019,Voelkl:PNAS:2015} -- although direct reciprocity is generally more difficult to demonstrate in animals~\citep{CluttonBrock:Nature:2009,Silk:CurrentBiology:2013,taborsky:CurrentBiology:2013}.
+However, at the outset, it is not clear how easy it is to {\it learn} such strategies by social imitation. 
+As one obstacle, individuals usually only observe other population members' actions (i.e., whether they cooperate or defect). 
+These observations alone may not suffice to infer the underlying strategies (i.e., the contingent rules that determine whether to cooperate or defect in a given situation). 
+As another obstacle, even if others' strategies are observable, individuals might find it difficult to identify which ones have the highest payoff. 
+After all, the payoff of a strategy of direct reciprocity is not determined by the outcome of any single round.
+Rather it is determined by how well this strategy fares over an entire sequence of rounds, against many different population members. 
+In practice, such information might be both difficult to obtain and difficult to process.
 
-We first consider two extreme scenarios, the classical scenario where an individual
-has a perfect memory of all their interactions and the alternative scenario of
-limited memory where individuals update their strategies only based on the very
-last payoff they obtained. We demonstrate the effects of limited payoff
-memory using a pairwise comparison process and the well studied game of the
-prisoner's dilemma~\cite{glynatsi:HSSC:2021}. We observe that individuals with
-limited payoff memory tend to adopt less generous strategies. We obtain similar
-results when we consider that individuals update their strategies based on more
-information. More specifically, up to the last two payoffs they obtained when
-interacting with up to two different members of the population.
+%% INTRO: Gap in the literature and what  question we wish to address.
+
+Most models of direct reciprocity abstract from these difficulties~\citep{brauchli:JTB:1999,brandt:JTB:2006,ohtsuki:JTB:2007b,szolnoki:pre:2009b,imhof2010stochastic,van-segbroeck:prl:2012,grujic:jtb:2012,Martinez2012,stewart:pnas:2013,pinheiro:PLoSCB:2014,stewart:games:2015,Baek2016,McAvoy:ProcA:2019,glynatsi:SCR:2020,Schmid:PlosCB:2022,Murase:SciRep:2022}.
+Instead they assume individuals can easily copy the strategies of others. 
+Similarly, they assume that updating decisions are based on the strategies' average (or expected) payoffs, taking into account all rounds and interactions. 
+These assumptions create a curious inconsistency in how models represent an individual's cognitive abilities. 
+On the one hand, when playing the game, individuals are often assumed to have restricted memory. 
+Respective studies typically assume that individuals make their decisions each round based on the outcome of the last round only~\citep[with only a few exceptions, see Refs.][]{Hauert1997,van-veelen:PNAS:2012,Stewart2016,Li:NatCS:2022,Murase:PLoSCompBio:2023a}.
+Yet when learning new strategies, individuals are assumed to remember (or compute) each others' precise average payoff across many rounds and many interaction partners. 
+Herein, we wish to explore whether this latter assumption is actually necessary for the evolution of reciprocity through social imitation. 
+We ask whether individuals can learn to adopt reciprocal strategies even when learning only based on payoff information from a limited number of rounds.
+
+%% INTRO: Basic description of our model and our findings
+
+To explore that question, we first consider two extreme scenarios. 
+The first scenario corresponds to the usual modeling approach. 
+Here, individuals update their strategies for a repeated prisoner's dilemma based on pairwise comparisons~\cite{traulsen2007pairwise}, based on their strategies' expected payoff. 
+We contrast this model with an alternative scenario where individuals update their strategies based on the very
+last (one-shot) payoff they obtained. 
+We observe that individuals with limited payoff memory tend to adopt less generous strategies, yet moderate levels of cooperation can still evolve. 
+Moreover, as we increase the individuals' payoff memory, to include the last two or three one-shot payoffs, cooperation rates quickly approach the rates observed in the classical baseline case. 
+Overall, these findings suggest that while memory is important, already minimal payoff information may suffice for the evolution of direct reciprocity based on social learning. 
+They also suggest that the classical model of reciprocity (based on expected payoffs) can be interpreted as a useful approximation to more realistic models that provide a more realistic depiction of cognitive constraints.
 
@@
 
+We further investigate the impact of benefit and selection strength on
+generosity $q$ and the cooperation rate, as shown in Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}.
+According to Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}\textbf{A},
+perfect memory consistently results in a higher cooperation rate, which
+increases with increasing benefit. On the other hand, the cooperation rate
+remains approximately 50\% for limited payoff memory once \(b = 5\).
+From Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}\textbf{B}
+we observe that for weak selection, \(\beta < 1\), the two methods yield similar
+results, however, as \(\beta\) increases there is variation in the evolving
+populations. In the case of expected payoffs the resident populations become
+more cooperative, whereas in the case of limited payoff memory, the resident
+populations become more defective.
+
+The limited payoff memory framework can be expanded by enabling individuals to
+observe a greater number of rounds, interact with a larger number of members, or
+both (refer to Supplementary Information Sections 4-6). To gain further insight
+into the impact of limited payoff memory, we explore the scenarios of updating
+based on the last round with two members of the population, the last two rounds
+with another member of the population, and the last two rounds with two members
+of the population. To analyze the effects of this framework, we conduct
+numerical simulations using various fitness methods. The cooperation rates for
+low and high benefits are presented in Figure~\ref{fig:cooperation_rate_all_updating_payoffs}.
+
+We observe a significant rise in the cooperation rate with the introduction of
+slightly more information. Specifically, in scenarios involving two rounds or
+two interactions, the cooperation rates are almost identical. For a large
+population and a high continuation probability ($\delta$), the conditional
+cooperators that are adopted by the resident population in these scenarios take
+on the same form \((1, q < \frac{\sqrt{2}}{2})\). When considering both sets
+of information (i.e., two rounds and two interactions), cooperation rates
+experience a greater increase, yet still remain lower compared to the classical
+scenario.
+
+\section{Conclusions}\label{section:conclusions}
+
+Cooperation can be seen as odd, why is it that we choose to help others at a
+personal cost? In spite of all the selfish genes', animal and human communities
+show signs of altruism and cooperation~\cite{milinski1987tit, kerr2002local,
+carter2020development}. Evolutionary game theoretical models have helped us
+shape our understanding of the evolution of cooperation. In fact, the evolution
+of cooperation constitutes such a major focus of the field that evolutionary
+game theory seems to be reduced to the evolution of
+cooperation~\cite{Traulsen:PhilTrans:2022}.
+
+Evolutionary models in the past often feature a curious inconsistency. While
+these models depict how individuals make decisions in each round by assuming
+that they only retain memory of the previous round, they also assume that
+individuals possess a perfect memory when it comes to updating their strategies
+over time. To be precise, individuals are assumed to remember all of their past
+interactions and each interaction's outcome when updating strategies.
+
+Here, we investigate the robustness of cooperation as models deviate
+from the assumption of perfect memory. While prior research has investigated the
+impact of constraining individuals' interactions, we take into account the
+limitation of not only interactions but also the information available for each
+outcome. Additionally, prior studies have only allowed for the adoption of
+simple strategies such as always cooperating or always defecting. In contrast,
+we enable the use of more intricate strategies where players can utilize the
+previous play of their co-player to make decisions.
+
+In our framework, players update their strategies based on a combination of
+interactions and outcomes. The initial scenario we examined involved using one
+piece of information: the last round of one interaction. The outcomes suggest
+that cooperation faces difficulties in developing when the updating stage
+utilizes minimal social information. This effect is compounded as the benefit
+and strength of selection are independently increased. The findings indicate
+that cooperative players benefit from the ability to engage with all members of
+the population.
+
+Furthermore, we investigated scenarios where the final two rounds, or the last
+two instances of interaction, were taken into account. We observed a
+statistically significant rise in the frequency of cooperative behavior. For a
+sizable population and a high likelihood of continued interactions, the two
+cases yield the same result as the overall payoff is influenced by two possible
+outcomes. Notably, the scenario involving two rounds and two interactions
+yielded the highest cooperation rate among all the novel methodologies that we
+tested.\\
+% A nice sentence missing here.
+
+%%%%%%%%%%%
+%% END NOTES %%
+%%%%%%%%%%%
+
+\begin{comment}
+
+\noindent
+{\bf Ethics.}
+This work is purely theoretical and did not require ethical approval from a human subject or animal welfare committee.\\
+
+\noindent
+{\bf Data accessibility.}
+All data and code used in this manuscript are openly available at https://github.com/Nikoleta-v3/pd-with-stochastic-payoffs \todo{(Correct?)}.\\
+
+\noindent
+{\bf Declaration of AI use.}
+We have not used AI-assisted technologies in creating this article.\\
+
+\noindent
+{\bf Authors' contributions.}
+N.G.: conceptualization, formal analysis, investigation, methodology, visualization, writing - original draft, writing -- review and editing; 
+A.M.: conceptualization, formal analysis, methodology, writing -- review and editing; 
+C.H.: conceptualization, formal analysis, methodology, writing -- review and editing.\\
+
+\noindent
+{\bf Conflict of interest declaration.}
+We declare we have no competing interests.\\
+
+\noindent
+{\bf Funding.}
+N.G. and C.H. acknowledge generous support by the European Research Council Starting Grant 850529:
+E-DIRECT, and by the Max Planck Society.\\
+\end{comment}
+
+{
+{\setlength{\bibsep}{0\baselineskip}\small
+\bibliographystyle{naturemag}
+\bibliography{bibliography.bib}
+}
+
+\clearpage
+\newpage
+
 \begin{figure}[!htbp]
@@
 
-We further investigate the impact of benefit and selection strength on
-generosity $q$ and the cooperation rate, as shown in Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}.
-According to Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}\textbf{A},
-perfect memory consistently results in a higher cooperation rate, which
-increases with increasing benefit. On the other hand, the cooperation rate
-remains approximately 50\% for limited payoff memory once \(b = 5\).
-From Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}\textbf{B}
-we observe that for weak selection, \(\beta < 1\), the two methods yield similar
-results, however, as \(\beta\) increases there is variation in the evolving
-populations. In the case of expected payoffs the resident populations become
-more cooperative, whereas in the case of limited payoff memory, the resident
-populations become more defective.
-
 \begin{figure}[!htbp]
@@
 \end{figure}
-
-The limited payoff memory framework can be expanded by enabling individuals to
-observe a greater number of rounds, interact with a larger number of members, or
-both (refer to Supplementary Information Sections 4-6). To gain further insight
-into the impact of limited payoff memory, we explore the scenarios of updating
-based on the last round with two members of the population, the last two rounds
-with another member of the population, and the last two rounds with two members
-of the population. To analyze the effects of this framework, we conduct
-numerical simulations using various fitness methods. The cooperation rates for
-low and high benefits are presented in Figure~\ref{fig:cooperation_rate_all_updating_payoffs}.
-
-We observe a significant rise in the cooperation rate with the introduction of
-slightly more information. Specifically, in scenarios involving two rounds or
-two interactions, the cooperation rates are almost identical. For a large
-population and a high continuation probability ($\delta$), the conditional
-cooperators that are adopted by the resident population in these scenarios take
-on the same form \((1, q < \frac{\sqrt{2}}{2})\). When considering both sets
-of information (i.e., two rounds and two interactions), cooperation rates
-experience a greater increase, yet still remain lower compared to the classical
-scenario.
 
@@
 
-\section{Conclusions}\label{section:conclusions}
-
-Cooperation can be seen as odd, why is it that we choose to help others at a
-personal cost? In spite of all the selfish genes', animal and human communities
-show signs of altruism and cooperation~\cite{milinski1987tit, kerr2002local,
-carter2020development}. Evolutionary game theoretical models have helped us
-shape our understanding of the evolution of cooperation. In fact, the evolution
-of cooperation constitutes such a major focus of the field that evolutionary
-game theory seems to be reduced to the evolution of
-cooperation~\cite{Traulsen:PhilTrans:2022}.
-
-Evolutionary models in the past often feature a curious inconsistency. While
-these models depict how individuals make decisions in each round by assuming
-that they only retain memory of the previous round, they also assume that
-individuals possess a perfect memory when it comes to updating their strategies
-over time. To be precise, individuals are assumed to remember all of their past
-interactions and each interaction's outcome when updating strategies.
-
-Here, we investigate the robustness of cooperation as models deviate
-from the assumption of perfect memory. While prior research has investigated the
-impact of constraining individuals' interactions, we take into account the
-limitation of not only interactions but also the information available for each
-outcome. Additionally, prior studies have only allowed for the adoption of
-simple strategies such as always cooperating or always defecting. In contrast,
-we enable the use of more intricate strategies where players can utilize the
-previous play of their co-player to make decisions.
-
-In our framework, players update their strategies based on a combination of
-interactions and outcomes. The initial scenario we examined involved using one
-piece of information: the last round of one interaction. The outcomes suggest
-that cooperation faces difficulties in developing when the updating stage
-utilizes minimal social information. This effect is compounded as the benefit
-and strength of selection are independently increased. The findings indicate
-that cooperative players benefit from the ability to engage with all members of
-the population.
-
-Furthermore, we investigated scenarios where the final two rounds, or the last
-two instances of interaction, were taken into account. We observed a
-statistically significant rise in the frequency of cooperative behavior. For a
-sizable population and a high likelihood of continued interactions, the two
-cases yield the same result as the overall payoff is influenced by two possible
-outcomes. Notably, the scenario involving two rounds and two interactions
-yielded the highest cooperation rate among all the novel methodologies that we
-tested.
-% A nice sentence missing here.
-
-\section{Acknowledgements}
-
-This work was supported by the European Research Council Starting Grant 850529:
-E-DIRECT.
-
-\bibliographystyle{unsrt}
-\bibliography{bibliography.bib}
-
 \end{document}

### 2023-10-03 paper/main.tex
CH: Editing the model section, adding some simple analytical observations to the results section

@@
 \usepackage{amsmath,amsthm,amssymb,setspace,enumitem,epsfig,titlesec,verbatim,color,array,eurosym,multirow}
-\usepackage[sort&compress,round
-]{natbib}
+\usepackage[sort&compress,comma,round]{natbib}
 \usepackage[footnotesize,bf]{caption}
@@
 \newcommand{\todo}[1]{\textcolor{blue}{#1}}
+
+\def\alld{\texttt{ALLD}}
+\def\tft{\texttt{TFT}}
+\def\gtft{\texttt{GTFT}}
+\def\esm{electronic supplementary material}
 
@@
 
-\newcommand{\splitatcommas}[1]{%
-  \begingroup
-  \begingroup\lccode`~=`, \lowercase{\endgroup \edef~{\mathchar\the\mathcode`,
-    \penalty0 \noexpand\hspace{0pt plus 1em}}%
-  }\mathcode`,="8000 #1%
-  \endgroup
-}
-
 \theoremstyle{plainCl1}
@@
 Direct reciprocity can explain how cooperation emerges in repeated social interactions. 
-According to this literature, individuals should naturally learn to adopt conditionally cooperative strategies such as Tit-for-Tat.
+According to this literature, individuals should naturally learn to adopt conditionally cooperative strategies, such as Tit-for-Tat.
 % if they expect to have further encounters with their interaction partner.
-Corresponding models have greatly facilitated our understanding of cooperation. 
-Yet they often make strong assumptions on how individuals remember and process information. 
-For example, when strategies are updated through social learning, it is commonly assumed that individuals update their strategies based on their average payoffs.  
+Corresponding models have greatly facilitated our understanding of cooperation, yet they often make strong assumptions on how individuals remember and process information. 
+For example, when strategies are updated through social learning, it is commonly assumed that individuals compare their respective average payoffs.  
 This would require them to compute (or remember) their payoffs against all other population members.
@@
 Even in the most extreme case that they only take into account their very last interaction, we find that cooperation can still evolve. 
-However, such individuals adopt less generous strategies, and they tend to cooperate less often than in the classical framework with expected payoffs. 
+However, such individuals adopt less generous strategies, and they tend to cooperate less often than in the classical setup with expected payoffs. 
 Interestingly, once individuals remember the payoffs of two or three recent interactions, evolving cooperation rates quickly approach the classical limit. 
 These findings contribute to a literature that explores which kind of cognitive capabilities are required for reciprocal cooperation. 
-While our results suggest that memory facilitates the evolution of reciprocity, it only takes a rather modest amount of payoff memory for cooperation to emerge. 
-\end{abstract}
+While our results suggest that some rudimentary form of payoff memory is necessary, it already suffices to remember a few interactions.\end{abstract}
 
@@
 Evolutionary game theory describes the dynamics of populations when an individual's fitness depends on the traits or strategies of other population members~\cite{hofbauer1998evolutionary, nowak:Nature:2004, hauert2005game}.
-This theory can be used to describe the dynamics of animal conflict~\citep{maynard-smith:Nature:1973}, cancer cells~\citep{Stankova:JAMA:2019}, and of cooperation~\citep{axelrod1981evolution}. 
+This theory can be used to describe the dynamics of animal conflict~\citep{maynard-smith:Nature:1973}, cancer cells~\citep{Stein:PTRS:2023}, and of cooperation~\citep{nowak:Science:2006}. 
 Respective models translate strategic interactions into games~\cite{smith1982evolution}. 
@@
 Models of biological evolution posit that individuals with a high fitness tend to produce more offspring; models of cultural evolution instead assume that such individuals are imitated more often. 
-While biological and cultural evolution are sometimes treated as equivalent, there can be important differences~\citep{Wu2015,Smolla:PTRS:2021}. 
+Although biological and cultural evolution are sometimes treated as equivalent, there can be important differences~\citep{Wu2015,Smolla:PTRS:2021}. 
 For example, models of biological evolution do not require individuals to have any particular cognitive abilities.
 Here, it is the evolutionary process itself that biases the population towards strategies with higher fitness. 
-In contrast, in models of cultural evolution, individuals need to be aware of the different strategies that are available, and they need to identify those strategies with a higher payoff. 
-As a consequence, evolutionary outcomes may often depend on how easily different behaviors can be learned~\citep{Chatterjee:JTB:2012}, and on how easy payoff comparisons are.
+In contrast, in models of cultural evolution, individuals need to be aware of the different strategies present in the population, and they need to identify those strategies with a higher payoff. 
+As a consequence, evolutionary outcomes may depend on how easily different behaviors can be learned~\citep{Chatterjee:JTB:2012}, and on how easy payoff comparisons are.
 
@@
 In repeated interactions, individuals can condition their future behavior on their past experiences with their interaction partner. 
-They may use strategies such as Tit-for-Tat~\citep{rapoport:book:1965} or Generous Tit-for-Tat~\citep{molander:jcr:1985,Nowak1992tit} to preferentially cooperate with those partners who cooperated in the past.
+They may use strategies such as Tit-for-Tat~\citep{rapoport:book:1965,axelrod1981evolution} or Generous Tit-for-Tat~\citep{molander:jcr:1985,Nowak1992tit} to preferentially cooperate with those partners who cooperated in the past.
 Such conditional strategies approximate human behavior fairly well~\citep{fischbacher:EconL:2001,Rand:TCS:2013,DalBo:AER:2019,Rossetti:ETH:2023} and they have also been documented in several other species~\citep{Carter:PRSB:2013,Schweinfurth:AnBehav:2019,Voelkl:PNAS:2015} -- although direct reciprocity is generally more difficult to demonstrate in animals~\citep{CluttonBrock:Nature:2009,Silk:CurrentBiology:2013,taborsky:CurrentBiology:2013}.
@@
 Herein, we wish to explore whether this latter assumption is actually necessary for the evolution of reciprocity through social imitation. 
-We ask whether individuals can learn to adopt reciprocal strategies even when learning only based on payoff information from a limited number of rounds.
+We ask whether individuals can learn to adopt reciprocal strategies even when learning is based on payoff information from a limited number of rounds.
 
@@
 The first scenario corresponds to the usual modeling approach. 
-Here, individuals update their strategies for a repeated prisoner's dilemma based on pairwise comparisons~\cite{traulsen2007pairwise}, based on their strategies' expected payoff. 
+Here, individuals update their strategies for a repeated prisoner's dilemma based on pairwise comparisons~\cite{traulsen2007pairwise}, taking into account their expected payoffs. 
 We contrast this model with an alternative scenario where individuals update their strategies based on the very
 last (one-shot) payoff they obtained. 
-We observe that individuals with limited payoff memory tend to adopt less generous strategies, yet moderate levels of cooperation can still evolve. 
-Moreover, as we increase the individuals' payoff memory, to include the last two or three one-shot payoffs, cooperation rates quickly approach the rates observed in the classical baseline case. 
+We find that individuals with limited payoff memory tend to adopt less generous strategies, yet moderate levels of cooperation can still evolve. 
+Moreover, as we increase the individuals' payoff memory to include the last two or three one-shot payoffs, cooperation rates quickly approach the rates observed in the classical baseline case. 
 Overall, these findings suggest that while memory is important, already minimal payoff information may suffice for the evolution of direct reciprocity based on social learning. 
-They also suggest that the classical model of reciprocity (based on expected payoffs) can be interpreted as a useful approximation to more realistic models that provide a more realistic depiction of cognitive constraints.
+They also suggest that the classical model of reciprocity (based on expected payoffs) can often be interpreted as a useful approximation to more realistic models that include cognitive constraints.
 
-\section{Model Setup}\label{section:model}
+\section{Model and Methods}\label{section:model}
 
-A pairwise comparison process~\cite{traulsen2007pairwise} starts with assigning all
-individuals of the population the same strategy. Thenceforth each elementary time step of
-the process consists of the mutation phase, the game phase and the update phase.
-In the game phase individuals are matched in pairs and that they participate in
-a repeated 2 person donation game; a special case of the prisoner's dilemma. In
-the donation game there are two actions: cooperation (\(C\)) and defection
-(\(D\)). By cooperating a player provides a benefit \(b\) to the other player at
-their cost \(c\), with \(0 < c < b\). Thus the payoffs for a player in each turn
-are given by,
+%% MODEL: Basic intro
+
+To explore the impact of limited payoff memory, we adapt existing models of the evolution of direct reciprocity through social learning. 
+These models consider the dynamics on two different time scales. 
+The short time scale describes the game dynamics. 
+Here, individuals with fixed strategies are randomly matched to interact with each other in a series of repeated social dilemmas. 
+The long time scale describes the evolutionary dynamics. 
+Here, individuals can update their strategies based on the payoffs they yield. 
+In the following, we describe the basic setup of our model; all details and derivations are described in the \esm.\\
+
+%% MODEL: (Repeated) donation game
+
+\noindent
+{\bf Description of the game dynamics.} For our model, we consider a well-mixed population consisting of $N$~players.
+Players are randomly matched in pairs to participate in a repeated donation game~\citep{sigmund2010calculus}.
+Each round, players can either cooperate (\(C\)) or defect (\(D\)). 
+By cooperating, a player provides a benefit \(b\) to the other player at their cost \(c\), with \(0 \!<\! c \!<\! b\). 
+Thus, the players' payoffs in a single round are given by the matrix
 
@@
 \end{equation}
+In particular, payoffs take the form of a prisoner's dilemma.
+Mutual cooperation yields a better payoff than mutual defection ($b\!-\!c\!>\!0$), but each player individually prefers to defect independent of the co-player's action ($b\!>\!b\!-\!c$ and $0\!>\!-c$).
+To incorporate that individuals interact repeatedly, we assume that after each round, there is a constant continuation probability $\delta$ of interacting for another round. 
+For $\delta\!=\!0$, we recover the case of a conventional (one-shot) prisoner's dilemma. 
+Here, mutual defection is the only equilibrium. 
+As $\delta$ increases, the game turns into a repeated game.
+As a result, additional equilibria emerge, with some of them allowing for full cooperation~\citep{friedman:RES:1971,Akin:chapter:2016,hilbe:GEB:2015,stewart:pnas:2014}.
 
-Let \(\mathbf{u} = (b-c, -c, b, 0)\) be the round payoffs in a vector format, and let
-\(\mathcal{U} = \{r, s, t, p\}\) denote the set of feasible payoffs, where \(r\)
-denotes the payoff of mutual cooperation, \(s\) the sucker's payoff, \(t\) the
-temptation to defect payoff, and \(p\) the punishment payoff.
+%% MODEL: Reactive strategies
 
-In repeated games there are infinite many strategies, however, similar to the
-literature we will assume that individuals use reactive strategies. A reactive
-strategy considers only the previous action of the other player, and thus, a
-reactive strategy \(s\) can be written as a three-dimensional vector \(s=(y, p,
-q)\). The parameter \(y\) is the probability that the strategy opens with a
-cooperation and \(p\), \(q\) are the probabilities that the strategy cooperates
-given that the opponent cooperated and defected equivalently. The play between a
-pair of reactive strategies $s_1 = (y_1, p_1, q_1)$ and $s_2 = (y_2, p_2, q_2)$
-can be model as a Markov process with the transition matrix \(M\)~\cite{nowak:APC:1989},
+In a one-shot donation game, players can only choose among two strategies (either cooperate or defect).
+In the repeated game, strategies can become arbitrarily complex. 
+Here, strategies are contingent rules, telling players what to do depending on the outcome of all their previous interactions. 
+For simplicity, in the following we assume individuals use {\it reactive strategies}~\citep{Nowak1992tit}. 
+A reactive strategy only depends on the other player's action in the previous round. 
+As a consequence, reactive strategies can be written as a three-dimensional tuple \(\mathbf{s}=(y, p,q)\). 
+The first entry \(y\) is the probability that the player opens with cooperation in the first round. 
+The two other entries are the probabilities that the player cooperates in all subsequent rounds, depending on whether the co-player cooperated~($p$) or defected~($q$) in the previous round. 
+The set of reactive strategies is simple enough to allow for an explicit mathematical treatment~\citep{hofbauer1998evolutionary}. 
+Yet it is rich enough to capture several important strategies of repeated games. 
+For example, it contains \alld{} $=\!(0,0,0)$, the strategy that always defects. 
+Similarly, it contains Tit-for-Tat, \tft{} $=\!(1,1,0)$, the strategy that copies the co-player's previous action (and that cooperates in the first round).
 
-\begin{equation}\label{eq:transition_matrix}
-  \input{static/matrix.tex}
+In the short run, the players' strategies are taken to be fixed.
+Players use their strategies to decide whether to cooperate with any given interaction partner. 
+In the long run, however, the players' strategies may change depending on the payoffs they yield, as we describe in the following.\\
+
+ %% MODEL: Pairwise comparison process
+
+\noindent
+{\bf Description of the evolutionary dynamics.}
+Herein, we assume population members update their strategies based on social learning. 
+To model these strategy updates, we consider a pairwise comparison process~\citep{traulsen2007pairwise}. 
+This process assumes that at regular time intervals, one population member is randomly selected, and given the chance to update its strategy.
+We refer to this player as the `learner'. 
+With probability $\mu$ (reflecting a mutation rate), the learner simply adopts a random strategy (with all reactive strategies having the same probability to be chosen). 
+With the converse probability $1\!-\!\mu$, the learner randomly picks a `role model' from the population. 
+The learner then compares its own payoff $\pi_L$ from the repeated game stage to the payoff $\pi_R$ of the role model. 
+The learner adopts the role model's strategy with a probability \(\rho\) described by a Fermi function~\citep{blume:GEB:1995,szabo:PRE:1998}, 
+\begin{equation} \label{Eq:rho}
+    \rho(\pi_{L}, \pi_{R}) = \frac{1}{1\!+\! e^{- \!\beta (\pi_{R}- \pi_{L})}}.
 \end{equation}
+Here, the selection strength $\beta\!\ge\!0$ is a parameter that indicates how sensitive players are to payoff differences. 
+For $\beta\!=\!0$, any payoff differences are irrelevant, and the learner simply adopts the role model's strategy with probability one half.
+As the selection strength~$\beta$ increases, players are increasingly biased to update only if the role model has the higher payoff.
+In either case, we assume that the role model's strategy can be identified.
+That is, if the learner decides to imitate the role model's strategy, the learner copies it exactly, as assumed in most previous work on direct reciprocity.
 
-and the stationary vector \(\mathbf{v}(s_1,s_2)\)
-which is the solution to \(\mathbf{v}(s_1,s_2) \times M 
-= \mathbf{v}(s_1,s_2)\).
+%% MODEL: Perfect vs limited payoff memory
 
-In the update stage two individuals are randomly selected. From the two
-individuals, one  serves as the `learner' and the other as the `role model'. The
-learner adopts the role model's strategy with a probability \(\rho\) given by,
+We deviate from previous models in how we interpret the payoffs $\pi_L$ and $\pi_R$, which form the basis of the pairwise comparisons in Eq.~\eqref{Eq:rho}. 
+In previous work, these payoffs are taken to be the respective players' expected payoffs. 
+We interpret this setup as a model with perfect payoff memory. 
+In this setup, the payoffs  $\pi_L$ and $\pi_R$ represent an average over all possible repeated games the two individuals could have played with all population members. 
+The use of expected payoffs is mathematically convenient, because explicit formulas for the players' expected payoffs are available~\citep{hofbauer1998evolutionary}.
+Herein, we compare this model to a model with limited payoff memory. 
+In that model, the players' payoffs $\pi_L$ and $\pi_R$ are taken to be the payoffs that each player received in the last round of the last repeated game they interacted in (see \FigEvoProc\textbf{a}).
+This assumption could reflect, for example, a strong recency bias in how individuals evaluate their payoffs.  
+In addition to this extreme case of limited payoff memory, later on we also explore cases in which players take into account the outcome of two, three, or four recent rounds.
 
-\begin{equation} \label{Eq:rho}
-    \rho(\pi_{L}, \pi_{RM}) = \frac{1}{1\!+\! e^{\!-\!\beta (\pi_{RM}\!-\! \pi_{L})}}.
+Both in the case of perfect and limited memory, we iterate this elementary pairwise comparison process for many time steps. 
+This gives rise to a stochastic process that describes which strategies players adopt over time. 
+We explore the dynamics of this process with computer simulations.
+For the results presented in the following, we assume that mutations are rare (\(\mu\!\rightarrow\! 0\)). 
+This assumption is fairly common in evolutionary game theory, because it makes computations more efficient~\citep{fudenberg:JET:2006,wu:JMB:2012,mcavoy:jet:2015}, and because the results can be interpreted more easily.
+However, in Section~9 of the \esm{} we show that our main result holds continue to hold for strictly positive mutation rates.
+
+\section{Results}
+
+\noindent
+{\bf Stability of cooperative populations.}
+To get some intuition for the differences between perfect and limited payoff memory, we first analyze when cooperation is stable in either scenario.
+To this end, we consider a resident population in which all players but one adopt a strategy of Generous Tit-for-Tat, \gtft{} $=\!(1,1,q)$. 
+The remaining mutant player adopts \alld. 
+We say {\it cooperation is stochastically stable} if the single mutant is more likely to imitate the residents than vice versa. 
+For simplicity, we consider a large population~($N\!\rightarrow\!\infty$) and strong selection~($\beta\!\rightarrow\!\infty$).
+More general results are derived in the \esm.
+
+In the case of perfect payoff memory, it is straightforward to compute each player's expected payoff. 
+Because the population mostly consists of residents, and because residents mutually cooperate with each other, their payoff is $\pi_\gtft = b\!-\!c$. 
+The mutant only interacts with residents, and hence receives a benefit in the first round, and in every subsequent round with probability $q$. 
+As a result, the mutant's expected payoff is $\pi_\alld \!=\! (1\!-\!\delta\!+\!\delta q)b$. 
+For perfect payoff memory, the requirement for cooperation to be stochastically stable reduces to the condition $\pi_\gtft > \pi_\alld$. 
+This yields
+\begin{equation} \label{Eq:PerfectMemory}
+q < 1\!-\!\frac{c}{\delta  b}.
 \end{equation}
+In particular, we recover that $q\!=\!1-c/(\delta b)$ is the maximum generosity that cooperators should have~\citep{molander:jcr:1985,Nowak1992tit}. 
+Because $q$ needs to be non-negative, we also conclude that cooperation can only be stable if $\delta \!>\! c/b$.
+Again, this condition for the feasibility of direct reciprocity is the condition found in the literature~\citep{nowak:Science:2006}.
 
-\(\pi_{L}\) and \(\pi_{RM}\) are the updating payoffs/fitness of the learner and the
-role model respectively. The updating payoffs are a measure of how successful
-individuals are in the current standing of the population. The parameter
-\(\beta\) is known as the selection strength, namely, it shows how important the
-payoff difference is when the learner is considering adopting the strategy of
-the role model.
+The logic of the case with limited payoff memory is somewhat different. 
+Here we need to compute how likely each player obtains one of the four possible payoffs $\{b\!-\!c, -c, b, 0\}$ in their last interaction. 
+Because residents almost always interact with other residents, their last one-shot payoff is $\pi_\gtft = b\!-\!c$ with a probability approaching one. 
+For the defecting mutant, there are two possibilities. 
+({\it i})~ If the mutant's co-player happens to cooperate in the last round, the mutant receives a payoff of $\pi_\alld\!=\!b$.
+This occurs with probability $1\!-\!\delta\!+\!\delta q$. 
+({\it ii})~If the co-player defects in the last round, the mutant receives $\pi_\alld=\!0$. 
+This case occurs with probability $\delta(1\!-\!q)$.
+Because $b\!-\!c\!<\!b$, we conclude that  in the first case, residents tend to imitate the mutant. 
+Because $b\!-\!c\!>\!0$, in the second case mutants tend to imitate the resident. 
+Cooperation is thus stochastically stable if the first case is less likely than the second. 
+This yields the condition
+\begin{equation}
+q < 1\!-\!\frac{1}{2 \delta}.
+\end{equation}
+Interestingly, this condition no longer depends on the exact payoff values $c$ and $b$ (this happens because selection is strong, such that only the payoff ordering $b\!>\!c\!>\!0$ matters).
+In particular, because $q$ is non-negative, this condition can only be satisfied if $\delta\!>\!1/2$.
 
-For the results presented here we assume that mutations are rare (\(\mu
-\rightarrow 0\)). In fact, so rare that only two different strategies can be
-present in the population at any given time. However, in the Supplementary Information
-Section 9 we show that the main result holds for \(\mu \neq 0\).
-The case of low mutation is vastly adopted because it allows us to explicitly
-calculate the fixation probability of a newly introduced mutant. More specifically,
-at each step
-one individual adopts a mutant strategy randomly selected from the set of
-feasible strategies. The fixation probability \(\phi_{M}\) of the mutant
-strategy can be calculated explicitly as,
+By comparing the two cases, we find that payoff memory affects which cooperative strategies are viable. 
+With perfect memory, the maximum generosity $q$ needs to satisfy Eq.~\eqref{Eq:PerfectMemory}.
+In particular, this generosity can become arbitrarily large, if only the game's benefit-to-cost ratio~$b/c$ and the continuation probability~$\delta$ are sufficiently large. 
+In contrast, with limited payoff memory, the maximum generosity is bounded by one half, and it is independent of the benefit-to-cost ratio. 
+To explore whether we indeed observe these patterns in evolving populations, we turn to simulations.\\
 
-\begin{equation}\label{eq:appendix_fixation_probability}
-    \varphi_{M} = \frac{1}{1+\sum\limits_{i=1}^{N-1}\prod\limits_k^i \frac{\lambda^-_k}{\lambda^+_k}},
-\end{equation}
-
-where \(k\) is the number of
-mutants and \(\lambda^-_k, \lambda^+_k\) are the probabilities that the number of
-mutants decreases and increases respectively.
-Depending on the fixation probability \(\phi_{M}\) the mutant either
-fixes (becomes the new resident) or goes extinct. Regardless, in the next elementary
-time step another mutant strategy is introduced to the population. We iterate
-this elementary population updating process for a large number of mutant
-strategies and we record the resident strategies at each time step. The
-probabilities \(\lambda^-_k \text{ and } \lambda^+_k\) depend on the fitness
-of the mutant and the resident strategies. In the next section we
-present how fitness is calculated in the cases of perfect and limited payoff memory.
-
-\subsection{Fitness based on Perfect and Limited Payoff Memory}
-
-In the perfect payoff memory case an individual updates based on the average payoff
-against each other member of the population, otherwise known as expected payoffs. 
-The payoff of a reactive strategy \(s_1\) against the reactive strategy \(s_2\)
-in an infinitely repeated game ($\delta \rightarrow 1$) is explicitly
-calculated as,
-
-\[\langle\mathbf{v}(s_1,s_2),\mathbf{u}\rangle.\]
-
-In a population of size \(N\) there are \(k\) mutants and \(N - k\) residents.
-Let \(s_M =(y_M, p_M, q_M)\) and \(s_R = (y_R, p_R, q_R)\) denote the
-strategies of a mutant and a resident, the expected payoffs \(\pi_R\) and
-\(\pi_M\) are given by,
-
-\begin{equation} \label{Eq:ExpPay}
-  \begin{array}{lcrcr}
-  \displaystyle \pi_R & = &\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(s_R,s_R),\mathbf{u}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(s_R,s_M),\mathbf{u}\rangle,\\[0.5cm]
-  \displaystyle \pi_M & = &\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(s_M,s_R),\mathbf{u}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(s_M,s_M),\mathbf{u}\rangle.\\
-  \end{array}
-\end{equation}
-
-The probabilities that the number of mutants decreases and increases,
-\(\lambda^-_k\) and \(\lambda^+_k\), in the perfect payoff memory case are
-defined as,
-
-\begin{align}\label{eq:perfect_memory_lambdas}
-  \lambda^-_k \!=\!\rho(\pi_M, \pi_R) \quad \text{ and } \quad \lambda^+_k \!=\!\rho(\pi_R, \pi_M).
-\end{align}
-
-In the case of limited payoff memory we initially define the probability that a
-reactive strategy receives the payoff $u\!\in\! \mathcal{U}$ in the very last
-round of the game. This is given by Proposition~\ref{proposition:last_round}
-(see Supplementary Information Section 2.2.1 for proof).
-
-\begin{Prop}\label{proposition:last_round} Consider a repeated game, with
-continuation probability $\delta$, between players with reactive strategies
-$s_1\!=\!(y_1, p_1, q_1$)  and $s_2\!=\!(y_2,p_2,q_2)$ respectively. Then
-the probability that the $s_1$ player receives the payoff $u\!\in\!
-\mathcal{U}$ in the very last round of the game is given by
-$v_{u}(s_1,s_2)$, as given by Equation~(\ref{Eq:LastRound}).
-
-\begin{equation} \label{Eq:LastRound}
-  \setlength{\arraycolsep}{1pt}
-  \begin{array}{rcl}
-
-  v_{r}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(q_1+l_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+l_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
-  {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)},\\[1cm]
-
-  v_{s}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(q_1+l_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
-  {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)},\\[1cm]
-
-  v_{t}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+l_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
-  {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)},\\[1cm]
-
-  v_{p}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
-  {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)}.
-  \end{array}
-\end{equation}
-
-In these expressions, we have used the notation $l_i:=p_i\!-\!q_i$,
-$\bar{y}_i\!=\!1\!-\!y_i$, $\bar{q}_i:=1\!-\!q_i$, and
-$\bar{l}_i:=\bar{p}_i\!-\!\bar{q}_i=-l_i$ for $i\!\in\!\{1,2\}$.
-\end{Prop}
-
-In the case of limited payoffs memory both the role model and the learner
-estimate their fitness after interacting with a single member of the population.
-At each time step there are five possible pairings. They interact with each
-other with a probability \(\frac{1}{N - 1}\), and they do not interact with
-other with a probability \(1 - \frac{1}{N - 1}\). In the latter case, each of
-them can interact with either a mutant or a resident. Both of them interact with
-a mutant with a probability $\frac{(k-1)(k-2)}{(N-2)(N-3)}$ and both interact
-with a resident with a probability $\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$. The last
-two possible pairings are that either of them interacts with a resident whilst
-the other interacts with a mutant, and this happens with a probability
-$\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$. We define the probability that the randomly
-chosen resident obtained a payoff of $u_R$ in the last round of his respective
-game, and that the mutant obtained a payoff of $u_M$ as $x(u_R,u_M)$.
-
-\begin{equation}\label{eq:Chi}
-\setlength{\arraycolsep}{1pt}
-\begin{array}{llrl}
-x(u_R,u_M)	 &=&\displaystyle \frac{1}{N\!-\!1}\cdot  &v_{u_R}(s_R,s_M)\cdot 1_{(u_R,u_M)\in \mathcal{U}^2_F}\\[0.5cm]
-&+	
-&\displaystyle \left(1\!-\!\frac{1}{N\!-\!1}\right)  
-&\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} v_{u_R}(s_R,s_M) v_{u_M}(s_M,s_M) + 
- \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} v_{u_R}(s_R,s_M) v_{u_M}(s_M,s_R)\right.\\[0.5cm]
-&&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} v_{u_R}(s_R,s_R) v_{u_M}(s_M,s_M) + 
- \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} v_{u_R}(s_R,s_R) v_{u_M}(s_M,s_R)\right].
-\end{array}
-\end{equation}
-
-The probability that the number of mutants increases and decreases by one in the
-case of limited payoff memory are now given by,
-
-\begin{align}\label{eq:limited_memory_lambdas}
-\lambda^+_k=\frac{N\!-\!k}{N} \cdot \frac{k}{N} \cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M}) \rho(u_{R},u_{M}) \quad \text{and} \quad
-\lambda^-_k=\frac{N\!-\!k}{N} \cdot \frac{k}{N} \cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M}) \rho(u_{M},u_{R}).
-\end{align}
-
-In this expression, $\frac{(N\!-\!k)}{N}$ is the probability that the randomly
-chosen learner is a resident, and $\frac{k}{N}$ is the probability that the role
-model is a mutant. The sum corresponds to the total probability that the learner
-adopts the role model's strategy over all possible payoffs $u_R$ and $u_M$ that
-the two player may have received in their respective last rounds.
-
-\section{Simulation Results}
-
+\noindent
+{\bf Evolutionary dynamics with perfect and limited payoff memory.}
 To assess the impact of updating payoffs, we simulate the evolutionary process,
@@
 \end{document}
+
+%% SOME ADDITIONAL TEXT BITS %%
+
+% repeated game markov chain
+given that the opponent cooperated and defected equivalently. The play between a
+pair of reactive strategies $s_1 = (y_1, p_1, q_1)$ and $s_2 = (y_2, p_2, q_2)$
+can be model as a Markov process with the transition matrix \(M\)~\cite{nowak:APC:1989},
+
+\begin{equation}\label{eq:transition_matrix}
+  \input{static/matrix.tex}
+\end{equation}
+
+and the stationary vector \(\mathbf{v}(s_1,s_2)\)
+which is the solution to \(\mathbf{v}(s_1,s_2) \times M 
+= \mathbf{v}(s_1,s_2)\).
+
+% pairwise comparison fixation probability
+
+In fact, so rare that only two different strategies can be
+present in the population at any given time. However, in the Supplementary Information
+Section 9 we show that the main result holds for \(\mu \neq 0\).
+The case of low mutation is vastly adopted because it allows us to explicitly
+calculate the fixation probability of a newly introduced mutant. More specifically,
+at each step
+one individual adopts a mutant strategy randomly selected from the set of
+feasible strategies. The fixation probability \(\phi_{M}\) of the mutant
+strategy can be calculated explicitly as,
+
+\begin{equation}\label{eq:appendix_fixation_probability}
+    \varphi_{M} = \frac{1}{1+\sum\limits_{i=1}^{N-1}\prod\limits_k^i \frac{\lambda^-_k}{\lambda^+_k}},
+\end{equation}
+
+where \(k\) is the number of
+mutants and \(\lambda^-_k, \lambda^+_k\) are the probabilities that the number of
+mutants decreases and increases respectively.
+Depending on the fixation probability \(\phi_{M}\) the mutant either
+fixes (becomes the new resident) or goes extinct. Regardless, in the next elementary
+time step another mutant strategy is introduced to the population. We iterate
+this elementary population updating process for a large number of mutant
+strategies and we record the resident strategies at each time step. The
+probabilities \(\lambda^-_k \text{ and } \lambda^+_k\) depend on the fitness
+of the mutant and the resident strategies. In the next section we
+present how fitness is calculated in the cases of perfect and limited payoff memory.
+
+% Explicit formulas for the transition probabilities for perfect and limited memory
+
+\subsection{Fitness based on Perfect and Limited Payoff Memory}
+
+In the perfect payoff memory case an individual updates based on the average payoff
+against each other member of the population, otherwise known as expected payoffs. 
+The payoff of a reactive strategy \(s_1\) against the reactive strategy \(s_2\)
+in an infinitely repeated game ($\delta \rightarrow 1$) is explicitly
+calculated as,
+
+\[\langle\mathbf{v}(s_1,s_2),\mathbf{u}\rangle.\]
+
+In a population of size \(N\) there are \(k\) mutants and \(N - k\) residents.
+Let \(s_M =(y_M, p_M, q_M)\) and \(s_R = (y_R, p_R, q_R)\) denote the
+strategies of a mutant and a resident, the expected payoffs \(\pi_R\) and
+\(\pi_M\) are given by,
+
+\begin{equation} \label{Eq:ExpPay}
+  \begin{array}{lcrcr}
+  \displaystyle \pi_R & = &\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(s_R,s_R),\mathbf{u}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(s_R,s_M),\mathbf{u}\rangle,\\[0.5cm]
+  \displaystyle \pi_M & = &\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(s_M,s_R),\mathbf{u}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(s_M,s_M),\mathbf{u}\rangle.\\
+  \end{array}
+\end{equation}
+
+The probabilities that the number of mutants decreases and increases,
+\(\lambda^-_k\) and \(\lambda^+_k\), in the perfect payoff memory case are
+defined as,
+
+\begin{align}\label{eq:perfect_memory_lambdas}
+  \lambda^-_k \!=\!\rho(\pi_M, \pi_R) \quad \text{ and } \quad \lambda^+_k \!=\!\rho(\pi_R, \pi_M).
+\end{align}
+
+In the case of limited payoff memory we initially define the probability that a
+reactive strategy receives the payoff $u\!\in\! \mathcal{U}$ in the very last
+round of the game. This is given by Proposition~\ref{proposition:last_round}
+(see Supplementary Information Section 2.2.1 for proof).
+
+\begin{Prop}\label{proposition:last_round} Consider a repeated game, with
+continuation probability $\delta$, between players with reactive strategies
+$s_1\!=\!(y_1, p_1, q_1$)  and $s_2\!=\!(y_2,p_2,q_2)$ respectively. Then
+the probability that the $s_1$ player receives the payoff $u\!\in\!
+\mathcal{U}$ in the very last round of the game is given by
+$v_{u}(s_1,s_2)$, as given by Equation~(\ref{Eq:LastRound}).
+
+\begin{equation} \label{Eq:LastRound}
+  \setlength{\arraycolsep}{1pt}
+  \begin{array}{rcl}
+
+  v_{r}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(q_1+l_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+l_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
+  {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)},\\[1cm]
+
+  v_{s}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(q_1+l_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
+  {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)},\\[1cm]
+
+  v_{t}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+l_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
+  {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)},\\[1cm]
+
+  v_{p}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
+  {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)}.
+  \end{array}
+\end{equation}
+
+In these expressions, we have used the notation $l_i:=p_i\!-\!q_i$,
+$\bar{y}_i\!=\!1\!-\!y_i$, $\bar{q}_i:=1\!-\!q_i$, and
+$\bar{l}_i:=\bar{p}_i\!-\!\bar{q}_i=-l_i$ for $i\!\in\!\{1,2\}$.
+\end{Prop}
+
+In the case of limited payoffs memory both the role model and the learner
+estimate their fitness after interacting with a single member of the population.
+At each time step there are five possible pairings. They interact with each
+other with a probability \(\frac{1}{N - 1}\), and they do not interact with
+other with a probability \(1 - \frac{1}{N - 1}\). In the latter case, each of
+them can interact with either a mutant or a resident. Both of them interact with
+a mutant with a probability $\frac{(k-1)(k-2)}{(N-2)(N-3)}$ and both interact
+with a resident with a probability $\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$. The last
+two possible pairings are that either of them interacts with a resident whilst
+the other interacts with a mutant, and this happens with a probability
+$\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$. We define the probability that the randomly
+chosen resident obtained a payoff of $u_R$ in the last round of his respective
+game, and that the mutant obtained a payoff of $u_M$ as $x(u_R,u_M)$.
+
+\begin{equation}\label{eq:Chi}
+\setlength{\arraycolsep}{1pt}
+\begin{array}{llrl}
+x(u_R,u_M)	 &=&\displaystyle \frac{1}{N\!-\!1}\cdot  &v_{u_R}(s_R,s_M)\cdot 1_{(u_R,u_M)\in \mathcal{U}^2_F}\\[0.5cm]
+&+	
+&\displaystyle \left(1\!-\!\frac{1}{N\!-\!1}\right)  
+&\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} v_{u_R}(s_R,s_M) v_{u_M}(s_M,s_M) + 
+ \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} v_{u_R}(s_R,s_M) v_{u_M}(s_M,s_R)\right.\\[0.5cm]
+&&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} v_{u_R}(s_R,s_R) v_{u_M}(s_M,s_M) + 
+ \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} v_{u_R}(s_R,s_R) v_{u_M}(s_M,s_R)\right].
+\end{array}
+\end{equation}
+
+The probability that the number of mutants increases and decreases by one in the
+case of limited payoff memory are now given by,
+
+\begin{align}\label{eq:limited_memory_lambdas}
+\lambda^+_k=\frac{N\!-\!k}{N} \cdot \frac{k}{N} \cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M}) \rho(u_{R},u_{M}) \quad \text{and} \quad
+\lambda^-_k=\frac{N\!-\!k}{N} \cdot \frac{k}{N} \cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M}) \rho(u_{M},u_{R}).
+\end{align}
+
+In this expression, $\frac{(N\!-\!k)}{N}$ is the probability that the randomly
+chosen learner is a resident, and $\frac{k}{N}$ is the probability that the role
+model is a mutant. The sum corresponds to the total probability that the learner
+adopts the role model's strategy over all possible payoffs $u_R$ and $u_M$ that
+the two player may have received in their respective last rounds.

### 2023-10-09 paper/main.tex
CH: Editing the results section, polishing some of the previous parts

@@
 %% Adding shortcut commands to refer to our figures %%
-\newcommand{\FigEvoProc}{{\bf Fig.~1}} 
-\newcommand{\FigInvAnalysis}{{\bfFig.~2}} 
-\newcommand{\FigResultsOverPara}{{\bf Fig.~3}}
+\newcommand{\FigBaseResults}{{\bf Fig.~1}} 
+\newcommand{\FigDependenceParameters}{{\bf Fig.~2}} 
+\newcommand{\FigHigherMemory}{{\bf Fig.~3}}
 
@@
 \noindent
-Direct reciprocity can explain how cooperation emerges in repeated social interactions. 
-According to this literature, individuals should naturally learn to adopt conditionally cooperative strategies, such as Tit-for-Tat.
-% if they expect to have further encounters with their interaction partner.
+Direct reciprocity can explain the evolution of cooperation in repeated social interactions. 
+According to this literature, individuals should naturally learn to adopt conditionally cooperative strategies if they have multiple encounters with their interaction partner. 
 Corresponding models have greatly facilitated our understanding of cooperation, yet they often make strong assumptions on how individuals remember and process information. 
@@
 
-These difficulties -- to learn strategies by social imitation -- are particularly pronounced in models of direct reciprocity. 
+These difficulties to learn strategies by social imitation are particularly pronounced in models of direct reciprocity. 
 This literature follows Trivers' insight that individuals have more of an incentive to cooperate in social dilemmas when they interact repeatedly~\citep{trivers1971evolution}. 
-In repeated interactions, individuals can condition their future behavior on their past experiences with their interaction partner. 
-They may use strategies such as Tit-for-Tat~\citep{rapoport:book:1965,axelrod1981evolution} or Generous Tit-for-Tat~\citep{molander:jcr:1985,Nowak1992tit} to preferentially cooperate with those partners who cooperated in the past.
-Such conditional strategies approximate human behavior fairly well~\citep{fischbacher:EconL:2001,Rand:TCS:2013,DalBo:AER:2019,Rossetti:ETH:2023} and they have also been documented in several other species~\citep{Carter:PRSB:2013,Schweinfurth:AnBehav:2019,Voelkl:PNAS:2015} -- although direct reciprocity is generally more difficult to demonstrate in animals~\citep{CluttonBrock:Nature:2009,Silk:CurrentBiology:2013,taborsky:CurrentBiology:2013}.
+In repeated interactions, individuals can condition their behavior on their past experiences with their interaction partner. 
+They may use strategies such as Tit-for-Tat~\citep{rapoport:book:1965,axelrod1981evolution} or Generous Tit-for-Tat~\citep{molander:jcr:1985,Nowak1992tit} to preferentially cooperate with other cooperators.
+Such conditional strategies approximate human behavior fairly well~\citep{fischbacher:EconL:2001,Rand:TCS:2013,DalBo:AER:2019,Rossetti:ETH:2023} and they have also been documented in several other species~\citep{Carter:PRSB:2013,Schweinfurth:AnBehav:2019,Voelkl:PNAS:2015} -- although direct reciprocity is more difficult to demonstrate in animals~\citep{CluttonBrock:Nature:2009,Silk:CurrentBiology:2013,taborsky:CurrentBiology:2013}.
 However, at the outset, it is not clear how easy it is to {\it learn} such strategies by social imitation. 
 As one obstacle, individuals usually only observe other population members' actions (i.e., whether they cooperate or defect). 
-These observations alone may not suffice to infer the underlying strategies (i.e., the contingent rules that determine whether to cooperate or defect in a given situation). 
+These observations may not suffice to infer the underlying strategies (i.e., the contingent rules that determine whether to cooperate or defect in a given situation). 
 As another obstacle, even if others' strategies are observable, individuals might find it difficult to identify which ones have the highest payoff. 
@@
 Rather it is determined by how well this strategy fares over an entire sequence of rounds, against many different population members. 
-In practice, such information might be both difficult to obtain and difficult to process.
+In practice, such information might be both difficult to obtain and to process.
 
@@
 
-To explore that question, we first consider two extreme scenarios. 
+To explore that question, we study the imitation dynamics of the repeated prisoner's dilemma for two extreme scenarios. 
 The first scenario corresponds to the usual modeling approach. 
-Here, individuals update their strategies for a repeated prisoner's dilemma based on pairwise comparisons~\cite{traulsen2007pairwise}, taking into account their expected payoffs. 
+Here, individuals update their strategies based on their expected payoffs. 
 We contrast this model with an alternative scenario where individuals update their strategies based on the very
@@
 To explore the impact of limited payoff memory, we adapt existing models of the evolution of direct reciprocity through social learning. 
-These models consider the dynamics on two different time scales. 
+These models involve two different time scales. 
 The short time scale describes the game dynamics. 
-Here, individuals with fixed strategies are randomly matched to interact with each other in a series of repeated social dilemmas. 
+Here, individuals with fixed strategies are randomly matched to interact with each other in repeated social dilemmas. 
 The long time scale describes the evolutionary dynamics. 
-Here, individuals can update their strategies based on the payoffs they yield. 
-In the following, we describe the basic setup of our model; all details and derivations are described in the \esm.\\
+Here, individuals can update their repeated-game strategies based on the payoffs they yield. 
+In the following, we introduce the basic setup of our model; all details and derivations are described in the \esm.\\
 
@@
 \noindent
-{\bf Description of the game dynamics.} For our model, we consider a well-mixed population consisting of $N$~players.
+{\bf Description of the game dynamics.} We consider a well-mixed population consisting of $N$~players.
 Players are randomly matched in pairs to participate in a repeated donation game~\citep{sigmund2010calculus}.
@@
 \end{equation}
-In particular, payoffs take the form of a prisoner's dilemma.
+In particular, payoffs take the form of a prisoner's dilemma:
 Mutual cooperation yields a better payoff than mutual defection ($b\!-\!c\!>\!0$), but each player individually prefers to defect independent of the co-player's action ($b\!>\!b\!-\!c$ and $0\!>\!-c$).
@@
 A reactive strategy only depends on the other player's action in the previous round. 
-As a consequence, reactive strategies can be written as a three-dimensional tuple \(\mathbf{s}=(y, p,q)\). 
+Such strategies can be written as a three-dimensional tuple \(\mathbf{s}=(y, p,q)\). 
 The first entry \(y\) is the probability that the player opens with cooperation in the first round. 
@@
 Similarly, it contains Tit-for-Tat, \tft{} $=\!(1,1,0)$, the strategy that copies the co-player's previous action (and that cooperates in the first round). 
+Finally, it contains Generous Tit-for-Tat, \gtft $=\!(1,1,q)$, where $q\!>\!0$ reflects a player's generosity in response to a co-player's defection.
 
@@
 We refer to this player as the `learner'. 
-With probability $\mu$ (reflecting a mutation rate), the learner simply adopts a random strategy (with all reactive strategies having the same probability to be chosen). 
+With probability $\mu$ (reflecting a mutation rate), the learner simply adopts a random strategy (all reactive strategies have the same probability to be chosen). 
 With the converse probability $1\!-\!\mu$, the learner randomly picks a `role model' from the population. 
-The learner then compares its own payoff $\pi_L$ from the repeated game stage to the payoff $\pi_R$ of the role model. 
+The learner then compares its own payoff $\pi_L$ from the repeated game to the role model's payoff $\pi_R$. 
 The learner adopts the role model's strategy with a probability \(\rho\) described by a Fermi function~\citep{blume:GEB:1995,szabo:PRE:1998}, 
@@
 \end{equation}
-Here, the selection strength $\beta\!\ge\!0$ is a parameter that indicates how sensitive players are to payoff differences. 
-For $\beta\!=\!0$, any payoff differences are irrelevant, and the learner simply adopts the role model's strategy with probability one half.
+Here, the selection strength parameter $\beta\!\ge\!0$ indicates how sensitive players are to payoff differences. 
+For $\beta\!=\!0$, payoff differences are irrelevant, and the learner simply adopts the role model's strategy with probability one half.
 As the selection strength~$\beta$ increases, players are increasingly biased to update only if the role model has the higher payoff.
 In either case, we assume that the role model's strategy can be identified.
-That is, if the learner decides to imitate the role model's strategy, the learner copies it exactly, as assumed in most previous work on direct reciprocity.
+That is, if the learner decides to imitate the role model's strategy, the learner copies it exactly, as in previous works on direct reciprocity.
 
@@
 In previous work, these payoffs are taken to be the respective players' expected payoffs. 
-We interpret this setup as a model with perfect payoff memory. 
-In this setup, the payoffs  $\pi_L$ and $\pi_R$ represent an average over all possible repeated games the two individuals could have played with all population members. 
-The use of expected payoffs is mathematically convenient, because explicit formulas for the players' expected payoffs are available~\citep{hofbauer1998evolutionary}.
-Herein, we compare this model to a model with limited payoff memory. 
-In that model, the players' payoffs $\pi_L$ and $\pi_R$ are taken to be the payoffs that each player received in the last round of the last repeated game they interacted in (see \FigEvoProc\textbf{a}).
-This assumption could reflect, for example, a strong recency bias in how individuals evaluate their payoffs.  
+We interpret that setup as a model with perfect payoff memory. 
+Here, the payoffs  $\pi_L$ and $\pi_R$ represent an average over all possible repeated games the two individuals have played with all population members (\FigBaseResults, upper left panel). 
+The use of expected payoffs is mathematically convenient, because explicit formulas are available~\citep{hofbauer1998evolutionary}.
+Herein, we compare perfect payoff memory to a model with limited payoff memory. 
+In that model, the players' payoffs $\pi_L$ and $\pi_R$ are taken to be the payoffs that each player received in their very last round prior to making social comparisons (see \FigBaseResults, lower left panel).
+This assumption could reflect, for example, a strong recency bias in how individuals evaluate payoffs.  
 In addition to this extreme case of limited payoff memory, later on we also explore cases in which players take into account the outcome of two, three, or four recent rounds.
 
-Both in the case of perfect and limited memory, we iterate this elementary pairwise comparison process for many time steps. 
+%% MODEL: Summary and rare mutation assumption.
+
+Both in the case of perfect and limited memory, we iterate the elementary strategy updating step described above for many time steps. 
 This gives rise to a stochastic process that describes which strategies players adopt over time. 
@@
 \section{Results}
+
+%% 	RESULTS: Setup to get analytical results
 
@@
 
-In the case of perfect payoff memory, it is straightforward to compute each player's expected payoff. 
-Because the population mostly consists of residents, and because residents mutually cooperate with each other, their payoff is $\pi_\gtft = b\!-\!c$. 
+%% RESULTS: Analytical results for perfect memory
+
+In the case of perfect payoff memory, it is straightforward to characterize when cooperation is stochastically stable. 
+Here, we simply need to compute the players' expected payoffs. 
+Because the population mostly consists of residents, and because residents mutually cooperate with each other, their expected payoff is $\pi_\gtft = b\!-\!c$. 
 The mutant only interacts with residents, and hence receives a benefit in the first round, and in every subsequent round with probability $q$. 
@@
 \end{equation}
-In particular, we recover that $q\!=\!1-c/(\delta b)$ is the maximum generosity that cooperators should have~\citep{molander:jcr:1985,Nowak1992tit}. 
-Because $q$ needs to be non-negative, we also conclude that cooperation can only be stable if $\delta \!>\! c/b$.
+In particular, we recover the previous observation that $q\!=\!1-c/(\delta b)$ is the maximum generosity that cooperators should have~\citep{molander:jcr:1985,Nowak1992tit,Schmid:NHB:2021}. 
+Because $q\!\ge\!0$, we also conclude that cooperation can only be stable if $\delta \!>\! c/b$.
 Again, this condition for the feasibility of direct reciprocity is the condition found in the literature~\citep{nowak:Science:2006}.
 
+%% RESULTS: Analytical results for limited memory
+
 The logic of the case with limited payoff memory is somewhat different. 
-Here we need to compute how likely each player obtains one of the four possible payoffs $\{b\!-\!c, -c, b, 0\}$ in their last interaction. 
-Because residents almost always interact with other residents, their last one-shot payoff is $\pi_\gtft = b\!-\!c$ with a probability approaching one. 
+Here we need to compute how likely each player obtains one of the four possible payoffs $\{b\!-\!c, -c, b, 0\}$ in the very last round of a game, before they make social comparisons.
+Because residents almost always interact with other residents, their last one-shot payoff is $\pi_\gtft = b\!-\!c$ almost surely. 
 For the defecting mutant, there are two possibilities. 
-({\it i})~ If the mutant's co-player happens to cooperate in the last round, the mutant receives a payoff of $\pi_\alld\!=\!b$.
-This occurs with probability $1\!-\!\delta\!+\!\delta q$. 
+({\it i})~ If the mutant's co-player happens to cooperate in the last round, the mutant receives $\pi_\alld\!=\!b$.
+This case occurs with probability $1\!-\!\delta\!+\!\delta q$. 
 ({\it ii})~If the co-player defects in the last round, the mutant receives $\pi_\alld=\!0$. 
-This case occurs with probability $\delta(1\!-\!q)$.
-Because $b\!-\!c\!<\!b$, we conclude that  in the first case, residents tend to imitate the mutant. 
-Because $b\!-\!c\!>\!0$, in the second case mutants tend to imitate the resident. 
-Cooperation is thus stochastically stable if the first case is less likely than the second. 
+This occurs with the converse probability $\delta(1\!-\!q)$.
+Because $b\!-\!c\!<\!b$, residents tend to imitate the mutant in the first case. 
+Because $b\!-\!c\!>\!0$, mutants tend to imitate the resident in the second case. 
+Cooperation is stochastically stable if the first case is less likely than the second. 
 This yields the condition
-\begin{equation}
+\begin{equation} \label{Eq:LimitedMemory}
 q < 1\!-\!\frac{1}{2 \delta}.
 \end{equation}
-Interestingly, this condition no longer depends on the exact payoff values $c$ and $b$ (this happens because selection is strong, such that only the payoff ordering $b\!>\!c\!>\!0$ matters).
-In particular, because $q$ is non-negative, this condition can only be satisfied if $\delta\!>\!1/2$.
+Interestingly, this condition no longer depends on the exact payoff values $c$ and $b$.
+This happens because of our assumption of strong selection, in which case only the payoff ordering $b\!>\!c\!>\!0$ matters.
+Because $q$ is non-negative, condition~\eqref{Eq:LimitedMemory} can only be satisfied if $\delta\!>\!1/2$, that is, when players interact in more than two rounds in expectation.
 
-By comparing the two cases, we find that payoff memory affects which cooperative strategies are viable. 
+%% RESULTS: Summary of analytical results
+
+By comparing the two cases, we find that payoff memory affects whether a conditionally cooperative strategy ($1,1,q$) is viable. 
 With perfect memory, the maximum generosity $q$ needs to satisfy Eq.~\eqref{Eq:PerfectMemory}.
-In particular, this generosity can become arbitrarily large, if only the game's benefit-to-cost ratio~$b/c$ and the continuation probability~$\delta$ are sufficiently large. 
-In contrast, with limited payoff memory, the maximum generosity is bounded by one half, and it is independent of the benefit-to-cost ratio. 
-To explore whether we indeed observe these patterns in evolving populations, we turn to simulations.\\
+In particular, this generosity can become arbitrarily large, provided the game's benefit-to-cost ratio~$b/c$ and the continuation probability~$\delta$ are sufficiently large. 
+In contrast, with limited payoff memory, the maximum generosity is bounded by one half, and it is independent of the benefit-to-cost ratio.\\
+
+%% RESULTS: Evolving strategies
 
 \noindent
-{\bf Evolutionary dynamics with perfect and limited payoff memory.}
-To assess the impact of updating payoffs, we simulate the evolutionary process,
-recording the strategies adopted by players over time based on perfect and
-limited payoff memory. We performed two separate runs for each approach varying
-the value of benefit \(b\).
-Figure~\ref{fig:expected_and_stochastic_for_donation} depicts the evolving
-conditional cooperation probabilities $p$ and $q$ (note that we omit the opening
-move \(y\), as the discount factor \(\delta\) is relatively high). The figure
-suggests that when updating is based on perfect payoff memory players tend to be
-more generous and more cooperative.
+{\bf Evolutionary dynamics of reciprocity.}
+To explore whether the previous static observations describe the dynamics of evolving populations, we turn to simulations.
+We have run separate simulations for perfect and limited payoff memory, both for a low and a high benefit of cooperation ($b/c\!=\!3$ and $b/c\!=\!10$, respectively).  
+In each case, we record which strategies ($y,p,q$) the players adopt over time.
+\FigBaseResults{} depicts the evolving conditional cooperation probabilities $p$ and $q$ (note that we omit the opening move \(y\), as the discount factor \(\delta\) is very close to one). 
+In each case, we find that the players' strategies cluster in two regions of the strategy space. 
+The first region corresponds to a neighborhood of \alld{} with $(p,q)\!\approx\!(0,0)$, which results in very little cooperation. 
+The second region corresponds to a thin strip of cooperative strategies with $(p,q)\!=\!(1,q)$. 
+Within this strip, we observe that most strategies satisfy the constraint on $q$ imposed by the inequalities~\eqref{Eq:PerfectMemory} and~\eqref{Eq:LimitedMemory}. 
+That is, with perfect memory, most evolving strategies have $q\!<\!1\!-\!c/b$, whereas with limited payoff memory, most strategies have $q\!<\!1/2$. 
+In particular, in the latter case changes in the benefit parameter have no qualitative effect on the distribution of strategies.
 
-Specifically, we observe that the resident population comprises either defectors
-or conditional cooperators \((1, q)\). The generosity level \(q\) adopted by the
-resident population depends on whether a defecting strategy can invade.
-Supplementary Information Section 2 and 3 show that in the perfect payoff memory
-framework, conditional cooperators of the form \((1, q < \frac{c}{b})\) can
-arise, however, in the case of limited payoff memory, only conditional
-cooperators of \((1, q < \frac{1}{2})\) can avoid invasion. 
-The $q$-values of the resident strategies are generally higher in classical
-case, indicating that players are more likely to forgive a defection if their fitness
-depends on interacting with every member of the population. This effect becomes
-more pronounced as the benefit value increases, as the perfect memory condition
-on the left-hand side increases, while in the limited memory case, it remains at
-\(q \approx \frac{1}{2}\).
+%% RESULTS: Evolving cooperation rates
 
-Higher $q$ values lead to a more cooperative population. We compute the average
-cooperation rate for each simulation, which is the average cooperation rate
-within the resident population. In the case of perfect payoff memory, the
-average cooperation rate is consistently higher than that of the last round
-payoffs.
+In each case, the evolutionary dynamics follows a similar cyclic pattern~\citep[as previously described in Refs.][]{imhof2010stochastic, Nowak1992tit}.
+Resident populations of defectors are most likely invaded by strategies close to \tft. 
+In conditionally cooperative populations with strategies $(1,1,q)$, neutral drift may introduce larger values of generosity $q$. 
+Once the resident strategy violates the conditions~\eqref{Eq:PerfectMemory} and~\eqref{Eq:LimitedMemory}, defectors can re-invade. 
+The overall cooperation rate, averaged over many time steps, depends on how much time the process spends near \alld{} on the one hand, and near the strip of conditionally cooperative strategies on the other hand.
+For perfect payoff memory, this overall cooperation rate is substantial. 
+It is 52\% for low benefits, and 98\% for high benefits. 
+For limited payoff memory, the evolving cooperation rates are smaller but strictly positive, with 37\% cooperation for low benefits and 51\% cooperation for high benefits (\FigBaseResults).
 
-We further investigate the impact of benefit and selection strength on
-generosity $q$ and the cooperation rate, as shown in Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}.
-According to Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}\textbf{A},
-perfect memory consistently results in a higher cooperation rate, which
-increases with increasing benefit. On the other hand, the cooperation rate
-remains approximately 50\% for limited payoff memory once \(b = 5\).
-From Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}\textbf{B}
-we observe that for weak selection, \(\beta < 1\), the two methods yield similar
-results, however, as \(\beta\) increases there is variation in the evolving
-populations. In the case of expected payoffs the resident populations become
-more cooperative, whereas in the case of limited payoff memory, the resident
-populations become more defective.
+%% RESULTS: Evolving cooperation rates
 
-The limited payoff memory framework can be expanded by enabling individuals to
-observe a greater number of rounds, interact with a larger number of members, or
-both (refer to Supplementary Information Sections 4-6). To gain further insight
-into the impact of limited payoff memory, we explore the scenarios of updating
-based on the last round with two members of the population, the last two rounds
-with another member of the population, and the last two rounds with two members
-of the population. To analyze the effects of this framework, we conduct
-numerical simulations using various fitness methods. The cooperation rates for
-low and high benefits are presented in Figure~\ref{fig:cooperation_rate_all_updating_payoffs}.
+To further investigate the influence of different parameters, we have systematically varied the benefit~$b$ and the selection strength~$\beta$ in Figure~\FigDependenceParameters.
+According to \FigDependenceParameters \textbf{A}, perfect memory consistently results in a higher cooperation rate, and this relative advantage further increases with an increasing benefit~$b$. 
+Interestingly, for limited payoff memory, the cooperation rate remains stable at approximately 50\% once \(b \!\ge\! 5\).
+This again reflects our earlier observation that the feasibility of cooperation in this scenario is independent of the exact values of $b$ and $c$, as described by Eq.~\eqref{Eq:LimitedMemory}. 
+With respect to the effect of different selection strengths, \FigDependenceParameters\textbf{B} suggests that both perfect and limited payoff memory yields similar cooperation rates for weak selection \(\beta \!<\!
+1\).
+Beyond weak selection, increasing selection has a positive effect under perfect payoff memory, but a negative effect under limited payoff memory.\\
 
-We observe a significant rise in the cooperation rate with the introduction of
-slightly more information. Specifically, in scenarios involving two rounds or
-two interactions, the cooperation rates are almost identical. For a large
-population and a high continuation probability ($\delta$), the conditional
-cooperators that are adopted by the resident population in these scenarios take
-on the same form \((1, q < \frac{\sqrt{2}}{2})\). When considering both sets
-of information (i.e., two rounds and two interactions), cooperation rates
-experience a greater increase, yet still remain lower compared to the classical
-scenario.
+%% RESULTS: Setup of higher-memory simulations
+
+\noindent 
+{\bf The effect of increasing individual payoff memory.}
+So far, we have taken a rather extreme interpretation of limited payoff memory. 
+In the respective scenario, we assumed that individuals update their strategies based on their experience in a single round of the prisoner's dilemma, against a single co-player. 
+The limited payoff memory framework can be expanded in various ways. 
+In particular, individuals may recall a greater number of rounds, they may recall interactions with different co-players, or both. 
+To gain further insight into the impact of limited payoff memory, we explore three further scenarios. 
+In one scenario, players recall the payoffs they obtained in the last two rounds against a single co-player. 
+In another scenario, players recall their last-round payoffs against two co-players. 
+Finally, in the last scenario, they recall the two last rounds against two co-players.
+
+For the first two scenarios, and assuming large populations, strong selection, and a high continuation probability, we can again derive an analytical condition for when cooperation is stochastically stable.
+The details of this analysis can be found in the \esm. 
+Interestingly, this analysis suggests that when $b\!>\!2c$, cooperation is stochastically stable when $q\!<\! \frac{\sqrt{2}}{2}\approx 0.707$, for both scenarios. 
+Comparing this condition with the more stringent condition in Eq.~\eqref{Eq:LimitedMemory} suggests that there are now more conditionally cooperative strategies that can sustain cooperation. 
+Hence, we would expect cooperation to evolve more easily. 
+To explore this prediction in more detail, we have run additional simulations, again for a low and a high benefit $b$, 
+see \FigHigherMemory.
+We observe that the minimal increase in the players' payoff memory can promote cooperation considerably (compared to the baseline case with a single round recalled). 
+Specifically, we observe that in all three scenarios with extended memory, cooperation rates are similar, and they approach the rates observed under perfect memory. 
+These results suggest that while it takes {\it some} payoff memory to sustain substantial cooperation rates, the requirements on memory seem to be rather modest.
 

### 2023-10-10 paper/main.tex
CH: Polishing the intro, model, and results

@@
 \noindent
-Direct reciprocity can explain the evolution of cooperation in repeated social interactions. 
+Direct reciprocity can explain the evolution of cooperative behaviors in repeated interactions. 
 According to this literature, individuals should naturally learn to adopt conditionally cooperative strategies if they have multiple encounters with their interaction partner. 
-Corresponding models have greatly facilitated our understanding of cooperation, yet they often make strong assumptions on how individuals remember and process information. 
+Corresponding models have greatly facilitated our understanding of cooperation, yet they often make strong assumptions on how individuals remember and process payoff information. 
 For example, when strategies are updated through social learning, it is commonly assumed that individuals compare their respective average payoffs.  
@@
 Respective models translate strategic interactions into games~\cite{smith1982evolution}. 
-These games specify how individuals (players) interact, which strategies they can choose, and what fitness consequences (or payoffs) their strategies have. 
+These games specify how individuals (players) interact, which strategies individuals can choose, and what fitness consequences (or payoffs) the different strategies have. 
 In addition, these models also specify the mode by which successful strategies spread over time. 
-Models of biological evolution posit that individuals with a high fitness tend to produce more offspring; models of cultural evolution instead assume that such individuals are imitated more often. 
+In models of biological evolution, individuals with a high fitness produce more offspring; in models of cultural evolution, such individuals are imitated more often. 
 Although biological and cultural evolution are sometimes treated as equivalent, there can be important differences~\citep{Wu2015,Smolla:PTRS:2021}. 
@@
 They may use strategies such as Tit-for-Tat~\citep{rapoport:book:1965,axelrod1981evolution} or Generous Tit-for-Tat~\citep{molander:jcr:1985,Nowak1992tit} to preferentially cooperate with other cooperators.
-Such conditional strategies approximate human behavior fairly well~\citep{fischbacher:EconL:2001,Rand:TCS:2013,DalBo:AER:2019,Rossetti:ETH:2023} and they have also been documented in several other species~\citep{Carter:PRSB:2013,Schweinfurth:AnBehav:2019,Voelkl:PNAS:2015} -- although direct reciprocity is more difficult to demonstrate in animals~\citep{CluttonBrock:Nature:2009,Silk:CurrentBiology:2013,taborsky:CurrentBiology:2013}.
-However, at the outset, it is not clear how easy it is to {\it learn} such strategies by social imitation. 
+Such conditional strategies approximate human behavior fairly well~\citep{fischbacher:EconL:2001,Rand:TCS:2013,DalBo:AER:2019,Rossetti:ETH:2023} and they have also been documented in several other species~\citep{Carter:PRSB:2013,Schweinfurth:AnBehav:2019,Voelkl:PNAS:2015} -- although direct reciprocity is generally more difficult to demonstrate in animals~\citep{CluttonBrock:Nature:2009,Silk:CurrentBiology:2013,taborsky:CurrentBiology:2013}.
+However, at the outset, it is not clear how easy it is to {\it learn} reciprocal strategies by social imitation. 
 As one obstacle, individuals usually only observe other population members' actions (i.e., whether they cooperate or defect). 
-These observations may not suffice to infer the underlying strategies (i.e., the contingent rules that determine whether to cooperate or defect in a given situation). 
-As another obstacle, even if others' strategies are observable, individuals might find it difficult to identify which ones have the highest payoff. 
+In general, these observations do not suffice to infer the underlying strategies (i.e., the contingent rules that determine whether to cooperate or defect in any given situation). 
+This problem is equivalent to the difficulties that arise when researchers need to estimate the participants' strategies in behavioral experiments~\citep{Rossetti:ETH:2023}. 
+As another obstacle, even if others' strategies are perfectly observable, individuals might find it difficult to identify which ones have the highest payoff. 
 After all, the payoff of a strategy of direct reciprocity is not determined by the outcome of any single round.
@@
 Most models of direct reciprocity abstract from these difficulties~\citep{brauchli:JTB:1999,brandt:JTB:2006,ohtsuki:JTB:2007b,szolnoki:pre:2009b,imhof2010stochastic,van-segbroeck:prl:2012,grujic:jtb:2012,Martinez2012,stewart:pnas:2013,pinheiro:PLoSCB:2014,stewart:games:2015,Baek2016,McAvoy:ProcA:2019,glynatsi:SCR:2020,Schmid:PlosCB:2022,Murase:SciRep:2022}.
-Instead they assume individuals can easily copy the strategies of others. 
-Similarly, they assume that updating decisions are based on the strategies' average (or expected) payoffs, taking into account all rounds and interactions. 
+They just assume individuals can easily copy the strategies of others. 
+Similarly, they just assume that updating decisions are based on the strategies' average (or expected) payoffs, which are based on all rounds and all interactions. 
 These assumptions create a curious inconsistency in how models represent an individual's cognitive abilities. 
@@
 
-To explore that question, we study the imitation dynamics of the repeated prisoner's dilemma for two extreme scenarios. 
-The first scenario corresponds to the usual modeling approach. 
+To explore that question, we theoretically study imitation dynamics in the repeated prisoner's dilemma, using two extreme scenarios. 
+The first scenario is the usual modeling approach. 
 Here, individuals update their strategies based on their expected payoffs. 
@@
 last (one-shot) payoff they obtained. 
-We find that individuals with limited payoff memory tend to adopt less generous strategies, yet moderate levels of cooperation can still evolve. 
+We find that individuals with limited payoff memory tend to adopt less generous strategies. 
+Yet moderate levels of cooperation can still evolve. 
 Moreover, as we increase the individuals' payoff memory to include the last two or three one-shot payoffs, cooperation rates quickly approach the rates observed in the classical baseline case.
+
+%% INTRO: Summary and conclusions
+
 Overall, these findings suggest that while memory is important, already minimal payoff information may suffice for the evolution of direct reciprocity based on social learning. 
 They also suggest that the classical model of reciprocity (based on expected payoffs) can often be interpreted as a useful approximation to more realistic models that include cognitive constraints.
+
+%%%%%%%%%
+%% MODEL  %%
+%%%%%%%%%
 
@@
 Each round, players can either cooperate (\(C\)) or defect (\(D\)). 
-By cooperating, a player provides a benefit \(b\) to the other player at their cost \(c\), with \(0 \!<\! c \!<\! b\). 
+By cooperating, a player provides a benefit \(b\) to the other player at their own cost \(c\), with \(0 \!<\! c \!<\! b\). 
 Thus, the players' payoffs in a single round are given by the matrix
@@
 As $\delta$ increases, the game turns into a repeated game.
-As a result, additional equilibria emerge, with some of them allowing for full cooperation~\citep{friedman:RES:1971,Akin:chapter:2016,hilbe:GEB:2015,stewart:pnas:2014}.
+Here, additional equilibria emerge, with some of them allowing for full cooperation~\citep{friedman:RES:1971,Akin:chapter:2016,hilbe:GEB:2015,stewart:pnas:2014}.
 
@@
 
-In a one-shot donation game, players can only choose among two strategies (either cooperate or defect).
-In the repeated game, strategies can become arbitrarily complex. 
+In a one-shot donation game, players can only choose among two pure strategies (they can either cooperate or defect).
+In the repeated game, strategies become arbitrarily complex. 
 Here, strategies are contingent rules, telling players what to do depending on the outcome of all their previous interactions. 
@@
 The two other entries are the probabilities that the player cooperates in all subsequent rounds, depending on whether the co-player cooperated~($p$) or defected~($q$) in the previous round. 
-The set of reactive strategies is simple enough to allow for an explicit mathematical treatment~\citep{hofbauer1998evolutionary}. 
+The set of reactive strategies is simple enough to facilitate an explicit mathematical analysis~\citep{hofbauer1998evolutionary}. 
 Yet it is rich enough to capture several important strategies of repeated games. 
@@
 Similarly, it contains Tit-for-Tat, \tft{} $=\!(1,1,0)$, the strategy that copies the co-player's previous action (and that cooperates in the first round). 
-Finally, it contains Generous Tit-for-Tat, \gtft $=\!(1,1,q)$, where $q\!>\!0$ reflects a player's generosity in response to a co-player's defection.
+Finally, it contains Generous Tit-for-Tat, \gtft $=\!(1,1,q)$, where $q\!>\!0$ reflects a player's generosity in response to a co-player's defection~\citep{molander:jcr:1985,Nowak1992tit}.
 
@@
 To model these strategy updates, we consider a pairwise comparison process~\citep{traulsen2007pairwise}. 
-This process assumes that at regular time intervals, one population member is randomly selected, and given the chance to update its strategy.
+This process assumes that at regular time intervals, one population member is randomly selected, and given the chance to revise its strategy.
 We refer to this player as the `learner'. 
@@
 \end{equation}
-Here, the selection strength parameter $\beta\!\ge\!0$ indicates how sensitive players are to payoff differences. 
+The selection strength parameter $\beta\!\ge\!0$ indicates how sensitive players are to payoff differences. 
 For $\beta\!=\!0$, payoff differences are irrelevant, and the learner simply adopts the role model's strategy with probability one half.
-As the selection strength~$\beta$ increases, players are increasingly biased to update only if the role model has the higher payoff.
-In either case, we assume that the role model's strategy can be identified.
+As the selection strength~$\beta$ increases, players are increasingly biased to imitate the role model only if it has the higher payoff.
+In either case, we assume in the following that the role model's strategy can be identified.
 That is, if the learner decides to imitate the role model's strategy, the learner copies it exactly, as in previous works on direct reciprocity.
@@
 We interpret that setup as a model with perfect payoff memory. 
-Here, the payoffs  $\pi_L$ and $\pi_R$ represent an average over all possible repeated games the two individuals have played with all population members (\FigBaseResults, upper left panel). 
-The use of expected payoffs is mathematically convenient, because explicit formulas are available~\citep{hofbauer1998evolutionary}.
-Herein, we compare perfect payoff memory to a model with limited payoff memory. 
-In that model, the players' payoffs $\pi_L$ and $\pi_R$ are taken to be the payoffs that each player received in their very last round prior to making social comparisons (see \FigBaseResults, lower left panel).
+There, the payoffs  $\pi_L$ and $\pi_R$ represent an average over all possible repeated games the two individuals have played with all population members (\FigBaseResults, upper left panel). 
+The use of expected payoffs is mathematically convenient, because explicit formulas for these payoffs are available~\citep{hofbauer1998evolutionary}.
+Herein, we compare this model of perfect payoff memory to a model with limited payoff memory. 
+In that model, the players' payoffs $\pi_L$ and $\pi_R$ are taken to be the payoffs that each player received in their very last round prior to making social comparisons. 
+That is, players only consider the very last repeated game they participated in, and there they only take into account the outcome of the very last round ( \FigBaseResults, lower left panel). 
 This assumption could reflect, for example, a strong recency bias in how individuals evaluate payoffs.  
@@
 This gives rise to a stochastic process that describes which strategies players adopt over time. 
-We explore the dynamics of this process with computer simulations.
+We explore the dynamics of this process mathematically and with computer simulations.
 For the results presented in the following, we assume that mutations are rare (\(\mu\!\rightarrow\! 0\)). 
-This assumption is fairly common in evolutionary game theory, because it makes computations more efficient~\citep{fudenberg:JET:2006,wu:JMB:2012,mcavoy:jet:2015}, and because the results can be interpreted more easily.
-However, in Section~9 of the \esm{} we show that our main result holds continue to hold for strictly positive mutation rates.
+This assumption is fairly common in evolutionary game theory, because it makes some computations more efficient~\citep{fudenberg:JET:2006,wu:JMB:2012,mcavoy:jet:2015}, and because the results can be interpreted more easily.
+However, in Section~9 of the \esm{} we show that our main results continue to hold for strictly positive mutation rates.
+
+%%%%%%%%%%
+%% RESULTS  %%
+%%%%%%%%%%
 
@@
 Interestingly, this condition no longer depends on the exact payoff values $c$ and $b$. 
-This happens because of our assumption of strong selection, in which case only the payoff ordering $b\!>\!c\!>\!0$ matters.
-Because $q$ is non-negative, condition~\eqref{Eq:LimitedMemory} can only be satisfied if $\delta\!>\!1/2$, that is, when players interact in more than two rounds in expectation.
+This independence arises because of our assumption of strong selection, in which case only the payoff ordering $b\!>\!c\!>\!0$ matters.
+Because $q$ is non-negative, condition~\eqref{Eq:LimitedMemory} can only be satisfied if $\delta\!>\!1/2$.
+That is, players need to interact in more than two rounds in expectation.
 
@@
 In each case, we record which strategies ($y,p,q$) the players adopt over time.
-\FigBaseResults{} depicts the evolving conditional cooperation probabilities $p$ and $q$ (note that we omit the opening move \(y\), as the discount factor \(\delta\) is very close to one). 
+\FigBaseResults{} depicts the evolving conditional cooperation probabilities $p$ and $q$ (we omit the opening move \(y\) because we use a discount factor~\(\delta\) very close to one). 
 In each case, we find that the players' strategies cluster in two regions of the strategy space. 
-The first region corresponds to a neighborhood of \alld{} with $(p,q)\!\approx\!(0,0)$, which results in very little cooperation. 
+The first region corresponds to a neighborhood of \alld{} with $(p,q)\!\approx\!(0,0)$.
 The second region corresponds to a thin strip of cooperative strategies with $(p,q)\!=\!(1,q)$. 
-Within this strip, we observe that most strategies satisfy the constraint on $q$ imposed by the inequalities~\eqref{Eq:PerfectMemory} and~\eqref{Eq:LimitedMemory}. 
+Within this strip, we observe that most strategies satisfy the constraints on $q$ imposed by the inequalities~\eqref{Eq:PerfectMemory} and~\eqref{Eq:LimitedMemory}. 
 That is, with perfect memory, most evolving strategies have $q\!<\!1\!-\!c/b$, whereas with limited payoff memory, most strategies have $q\!<\!1/2$. 
-In particular, in the latter case changes in the benefit parameter have no qualitative effect on the distribution of strategies.
+In particular, in the latter case changes in the benefit parameter have no qualitative effect on the qualitative distribution of strategies.
 
@@
 
-In each case, the evolutionary dynamics follows a similar cyclic pattern~\citep[as previously described in Refs.][]{imhof2010stochastic, Nowak1992tit}.
+In each case, the evolutionary dynamics follows a similar cyclic pattern~\citep[as previously described in Refs.][]{imhof2010stochastic, Nowak1992tit}:
 Resident populations of defectors are most likely invaded by strategies close to \tft. 
-In conditionally cooperative populations with strategies $(1,1,q)$, neutral drift may introduce larger values of generosity $q$. 
-Once the resident strategy violates the conditions~\eqref{Eq:PerfectMemory} and~\eqref{Eq:LimitedMemory}, defectors can re-invade. 
-The overall cooperation rate, averaged over many time steps, depends on how much time the process spends near \alld{} on the one hand, and near the strip of conditionally cooperative strategies on the other hand.
-For perfect payoff memory, this overall cooperation rate is substantial. 
+Once the population adopts conditionally cooperative strategies  $(1,1,q)$, neutral drift may introduce larger values of generosity $q$. 
+If the resident's generosity~$q$ violates the conditions~\eqref{Eq:PerfectMemory} and~\eqref{Eq:LimitedMemory}, defectors can re-invade, and the cycle starts again. 
+The overall cooperation rate, averaged over many time steps, depends on how much time the process spends near \alld{} on the one hand, and near the strip of conditionally cooperative strategies on the other.
+For perfect payoff memory, the overall cooperation rate is substantial. 
 It is 52\% for low benefits, and 98\% for high benefits. 
@@
 The limited payoff memory framework can be expanded in various ways. 
-In particular, individuals may recall a greater number of rounds, they may recall interactions with different co-players, or both. 
-To gain further insight into the impact of limited payoff memory, we explore three further scenarios. 
-In one scenario, players recall the payoffs they obtained in the last two rounds against a single co-player. 
-In another scenario, players recall their last-round payoffs against two co-players. 
-Finally, in the last scenario, they recall the two last rounds against two co-players.
+In particular, individuals may recall a larger number of rounds, they may recall their interactions with several co-players, or both. 
+To gain further insights on the impact of payoff memory, we explore three additional scenarios. 
+In the first scenario, players recall the payoffs they obtained in the last two rounds against a single co-player. 
+In the second scenario, players recall their last-round payoffs against two co-players. 
+Finally, in the last scenario, they recall the two last rounds against two co-players (further extensions are straightforward, but we do not explore them here).
 
-For the first two scenarios, and assuming large populations, strong selection, and a high continuation probability, we can again derive an analytical condition for when cooperation is stochastically stable.
+%% RESULTS: Analytical results for higher memory
+
+For the first two scenarios, we can again derive an analytical condition for when cooperation is stochastically stable. 
+To this end, we again assume populations are large and that selection is strong. 
+For simplicity, we also assume that the game continues almost certainly after each round (i.e., $\delta$ approaches one). 
 The details of this analysis can be found in the \esm. 
-Interestingly, this analysis suggests that when $b\!>\!2c$, cooperation is stochastically stable when $q\!<\! \frac{\sqrt{2}}{2}\approx 0.707$, for both scenarios. 
+Interestingly, we find in both scenarios that for $b\!>\!2c$, cooperation is stochastically stable when $q\!<\! \frac{\sqrt{2}}{2}\approx 0.707$. 
 Comparing this condition with the more stringent condition in Eq.~\eqref{Eq:LimitedMemory} suggests that there are now more conditionally cooperative strategies that can sustain cooperation. 
-Hence, we would expect cooperation to evolve more easily. 
-To explore this prediction in more detail, we have run additional simulations, again for a low and a high benefit $b$, 
+Hence, cooperation should evolve more easily.
+
+%% RESULTS: Simulations for higher memory
+
+We explore this prediction with additional simulations, again for a low and a high benefit $b$, 
 see \FigHigherMemory.
-We observe that the minimal increase in the players' payoff memory can promote cooperation considerably (compared to the baseline case with a single round recalled). 
-Specifically, we observe that in all three scenarios with extended memory, cooperation rates are similar, and they approach the rates observed under perfect memory. 
+We observe that a minimal increase in the players' payoff memory (compared to the baseline case with a single round recalled) can promote cooperation considerably. 
+Specifically, in all three scenarios with extended memory, we now see similar cooperation rates, and they approach the rates observed under perfect memory. 
 These results suggest that while it takes {\it some} payoff memory to sustain substantial cooperation rates, the requirements on memory seem to be rather modest.
 
-\section{Conclusions}\label{section:conclusions}
+%%%%%%%%%%%%
+%%  DISCUSSION %%
+%%%%%%%%%%%%
+
+\section{Discussion}\label{section:discussion}
 

### 2023-10-17 paper/main.tex
Edits ch (#8)
* CH: Editing the title page of the paper

* CH: Polishing the abstract

* CH: Polishing the abstract, drafting the intro

* CH: Editing the model section, adding some simple analytical observations to the results section

* CH: Editing the results section, polishing some of the previous parts

* CH: Polishing the intro, model, and results

* CH: Editing the discussion and the figure captions

@@
 
-Evolutionary game theory describes the dynamics of populations when an individual's fitness depends on the traits or strategies of other population members~\cite{hofbauer1998evolutionary, nowak:Nature:2004, hauert2005game}.
+Evolutionary game theory describes the dynamics of populations when an individual's fitness depends on the traits or strategies of other population members~\cite{hofbauer1998evolutionary, nowak:Nature:2004, hauert2005game,Traulsen:PhilTrans:2022}.
 This theory can be used to describe the dynamics of animal conflict~\citep{maynard-smith:Nature:1973}, cancer cells~\citep{Stein:PTRS:2023}, and of cooperation~\citep{nowak:Science:2006}. 
@@
 To explore whether the previous static observations describe the dynamics of evolving populations, we turn to simulations.
-We have run separate simulations for perfect and limited payoff memory, both for a low and a high benefit of cooperation ($b/c\!=\!3$ and $b/c\!=\!10$, respectively).  
-In each case, we record which strategies ($y,p,q$) the players adopt over time.
-\FigBaseResults{} depicts the evolving conditional cooperation probabilities $p$ and $q$ (we omit the opening move \(y\) because we use a discount factor~\(\delta\) very close to one). 
-In each case, we find that the players' strategies cluster in two regions of the strategy space. 
+We have run separate simulations for perfect and limited payoff memory. 
+In each case, we consider both a low and a high benefit of cooperation ($b/c\!=\!3$ and $b/c\!=\!10$, respectively).  
+For each simulation, we record which strategies ($y,p,q$) the players adopt over time.
+\FigBaseResults{} depicts the conditional cooperation probabilities $p$ and $q$ (we omit the opening move \(y\); because we use a discount factor~\(\delta\) close to one, first-round behavior is largely irrelevant).
+In all simulations, we find that the players' strategies cluster in two regions of the strategy space. 
 The first region corresponds to a neighborhood of \alld{} with $(p,q)\!\approx\!(0,0)$.
@@
 That is, with perfect memory, most evolving strategies have $q\!<\!1\!-\!c/b$, whereas with limited payoff memory, most strategies have $q\!<\!1/2$. 
-In particular, in the latter case changes in the benefit parameter have no qualitative effect on the qualitative distribution of strategies.
+In particular, for limited payoff memory, changes in the benefit parameter have no effect on the qualitative distribution of strategies.
 
@@
 
-For the first two scenarios, we can again derive an analytical condition for when cooperation is stochastically stable. 
+For the first two scenarios, we can derive an analytical condition for when cooperation is stochastically stable. 
 To this end, we again assume populations are large and that selection is strong. 
@@
 These results suggest that while it takes {\it some} payoff memory to sustain substantial cooperation rates, the requirements on memory seem to be rather modest. 
+Already remembering a few interactions, either with the same co-player or across different co-players, may provide players with enough information to adopt reciprocal strategies.
 
@@
 
-Cooperation can be seen as odd, why is it that we choose to help others at a
-personal cost? In spite of all the selfish genes', animal and human communities
-show signs of altruism and cooperation~\cite{milinski1987tit, kerr2002local,
-carter2020development}. Evolutionary game theoretical models have helped us
-shape our understanding of the evolution of cooperation. In fact, the evolution
-of cooperation constitutes such a major focus of the field that evolutionary
-game theory seems to be reduced to the evolution of
-cooperation~\cite{Traulsen:PhilTrans:2022}.
+Direct reciprocity is one of the key mechanisms to explain why unrelated individuals might cooperate~\citep{nowak:Science:2006}. 
+According to this mechanism, cooperation pays if it makes the interaction partner more cooperative in future. 
+To describe which strategies of direct reciprocity are most effective, the previous theoretical literature assumes that the evolutionary dynamics is determined by the players' expected payoffs~\citep{brauchli:JTB:1999,brandt:JTB:2006,ohtsuki:JTB:2007b,szolnoki:pre:2009b,imhof2010stochastic,van-segbroeck:prl:2012,grujic:jtb:2012,Martinez2012,stewart:pnas:2013,pinheiro:PLoSCB:2014,stewart:games:2015,Baek2016,McAvoy:ProcA:2019,glynatsi:SCR:2020,Schmid:PlosCB:2022,Murase:SciRep:2022}.
+To the extent that strategies are learned (not inherited), this assumption seems to impose rather stringent requirements on the individuals' cognitive abilities. 
+In the most extreme case, this would require individuals to remember (or compute) their payoffs against all population members, for all possible ways in which their repeated games may unfold. 
+This assumption introduces a curious inconsistency in how these models represent an individual's cognitive abilities. 
+For playing their games, individuals are often assumed to only recall the outcome of the very last round. 
+Yet to update their strategies, individuals are implicitly assumed to have a record of the outcome of all rounds, across all interaction partners.
 
-Evolutionary models in the past often feature a curious inconsistency. While
-these models depict how individuals make decisions in each round by assuming
-that they only retain memory of the previous round, they also assume that
-individuals possess a perfect memory when it comes to updating their strategies
-over time. To be precise, individuals are assumed to remember all of their past
-interactions and each interaction's outcome when updating strategies.
+Herein, we explore to which extent perfect payoff memory is in fact required for the evolution of reciprocity. 
+To this end, we consider a model in which individuals only remember the payoff of their very last interaction, or the payoffs of the last few interactions. 
+By only considering an individual's most recent experiences, the evolutionary process is subject to additional stochasticity. 
+Strategies that perform well on average (across an entire repeated game and across many interaction partners) may still get replaced if the respective player happened to yield an inferior payoff in the very last round.
+A similar element of stochasticity has been previously explored in the context of one-shot (non-repeated) games~\citep{sanchez:JTB:2005,roca:PhysicalReview:2006,Traulsen:JTB:2007,Woelfing:JTB:2009,Hauert:PRE:2018}.
+This literature studies which strategies are selected for when individuals only interact with a finite sample of population members. 
+In the respective models, individuals can only choose among two strategies. 
+They can either cooperate or defect, and stochastic sampling affects which of these two strategies is favored. 
+Instead, in repeated games, players have access to a large set of strategies~\citep[in our case, all reactive strategies;][]{nowak:APC:1989}. 
+Here, stochastic sampling does not only affect whether cooperative or non-cooperative strategies are favored; it also affects {\it which} conditionally cooperative strategies are favored.
 
-Here, we investigate the robustness of cooperation as models deviate
-from the assumption of perfect memory. While prior research has investigated the
-impact of constraining individuals' interactions, we take into account the
-limitation of not only interactions but also the information available for each
-outcome. Additionally, prior studies have only allowed for the adoption of
-simple strategies such as always cooperating or always defecting. In contrast,
-we enable the use of more intricate strategies where players can utilize the
-previous play of their co-player to make decisions.
-
-In our framework, players update their strategies based on a combination of
-interactions and outcomes. The initial scenario we examined involved using one
-piece of information: the last round of one interaction. The outcomes suggest
-that cooperation faces difficulties in developing when the updating stage
-utilizes minimal social information. This effect is compounded as the benefit
-and strength of selection are independently increased. The findings indicate
-that cooperative players benefit from the ability to engage with all members of
-the population.
-
-Furthermore, we investigated scenarios where the final two rounds, or the last
-two instances of interaction, were taken into account. We observed a
-statistically significant rise in the frequency of cooperative behavior. For a
-sizable population and a high likelihood of continued interactions, the two
-cases yield the same result as the overall payoff is influenced by two possible
-outcomes. Notably, the scenario involving two rounds and two interactions
-yielded the highest cooperation rate among all the novel methodologies that we
-tested.\\
-% A nice sentence missing here.
+To address these questions, we combine analytical methods and computer simulations. 
+In the most extreme case, we consider individuals who update their strategies based on only one piece of information:  the last round of a single repeated game. 
+For that case, we find that individuals are less generous, and they tend to be less cooperative overall~(\FigBaseResults). 
+However, once individuals update their strategies based on two or more recent experiences, overall cooperation rates quickly approach the levels observed under perfect payoff memory~(\FigHigherMemory).
+These findings suggest that models based on expected payoffs can serve as a useful approximation to more realistic models with limited payoff memory. 
+Our findings also contribute to a wider literature that explores which kinds of cognitive capacities are required for reciprocal altruism to be feasible~\citep[e.g.,][]{Stevens:fip:2011,Volstorf:PlosOne:2011}.
+While more payoff memory is always favorable, reciprocal cooperation can already be sustained if individuals have a record of two or three past outcomes.\\
 
@@
 \begin{comment}
-
 \noindent
@@
 
+%%%%%%%%%%%%
+%% REFERENCES %%
+%%%%%%%%%%%%
+
 {
-{\setlength{\bibsep}{0\baselineskip}\small
+{\setlength{\bibsep}{0\baselineskip}
 \bibliographystyle{naturemag}
@@
 
+%%%%%%%%%%
+%% FIGURES  %%
+%%%%%%%%%%
+
 \clearpage
@@
 
-\begin{figure}[!htbp]
+\begin{figure}[t]
     \centering
@@
     \caption{{\bf Evolutionary dynamics under perfect and limited payoff memory.}
-    ({\bf Schematic illustrations}) On the left panels we show schematic
-    illustrations of the perfect memory and the limited memory cases. The shaded
-    background denotes the game phase information that an individual considers
-    when updating strategies. In the case of perfect payoff memory the entire region is
-    shaded and in the case of limited payoff memory only one turn with a single member
-    of the population. ({\bf Simulations}) We have run four simulations of the
-    pairwise comparison process for $T\!=\!10^7$ time steps. For each time step
-    of the process
-    we record the current resident population ($y,p,q$). Since
-    simulations are run for a relatively high continuation probability of
-    $\delta\!=\!0.999$, we do not report the players' initial cooperation
-    probability $y$. The plots show how often the resident population chooses
-    each combination ($p,q$) of conditional cooperation probabilities in the
-    subsequent rounds. We also report the evolved cooperation rate which is
-    calculated as the average cooperation rate within the resident population.
-    ({\bf Perfect Memory}) In the case of low benefit the resident population
-    either consists of defectors (with $p\!\approx\!q\!\approx\!0$) or of
-    conditional cooperators. Conditional cooperators, or otherwise known as
-    generous tit for tat, are a set of strategies that always cooperate
-    following a cooperation ($p\!\approx\!1\!$) and cooperate with a probability
-    $q$ given that the co-player has defected. $q$ denotes the generosity of a
-    player. The resident population applies a conditional cooperator strategy
-    for which $q\!\le\!1\!-\!c/b\!=\!0.67$. 
-    In the case of high benefit the population mainly consists of conditional
-    cooperators of the form ($p\!\approx\!1\!, q\!\le\!1\!-\!1/10\!=\!0.9$). In
-    the Supplementary Information Section 2 we show that a conditional
-    cooperator needs to be of the form ($p\!\approx\!1\!, q\!\le\!1\!-\!c/b$) to
-    not be invaded by defecting strategies. A higher generosity in the
-    population results in a higher average cooperation rate. The average
-    cooperation rate increases from 52\% for \(b=3\) to 98\% for \(b=10\). ({\bf
-    Limited Memory}) When players update their strategies based on their
-    realized payoffs in the last round, there are two different predominant
-    behaviors regardless of the benefit value. The resident population either
-    consists of defectors (with $p\!\approx\!q\!\approx\!0$) or of conditional
-    cooperators. The maximum level of $q$, consistent with stable cooperation, is
-    somewhat smaller compared to the perfect memory setting.
-    Namely, in the Supplementary Information Section 3 we show that
-    regardless of the value of benefit a conditional cooperator need to be of
-    the form $q\!<\!\frac{1}{2}$ to not be invaded by defectors.
-    The evolved cooperation
-    rate only slightly increases from 37\% ($b=3$) to 51\% ($b=10$). Parameters: $N\!=\!100$,
-    $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.999$.}
+    The leftmost panels give a schematic overview of the two main scenarios we compare. 
+    The two scenarios differ in how many past interactions individuals take into account when updating their strategy. 
+    In the scenario with perfect payoff memory, individuals consider all their past interactions (against all population members, and taking every turn of each repeated game into account). 
+    In the scenario with limited payoff memory, individuals only consider their very last interaction (against one specific population member, taking into account only one round of the repeated game).
+    The four panels  on the right side depict the outcome of evolutionary simulations for repeated games with either a low or a high benefit of cooperation. 
+    Colors represent how often the respective region of the strategy space is visited over time. 
+    In all four panels, two regions are visited particularly often. 
+    One region corresponds to a neighborhood of \alld{} with $p\approx q\!\approx\!0$ (lower left corner). 
+    The other region corresponds to a strip of conditionally cooperative strategies with $p\!\approx\!
+1$ and $q$ satisfying the constraints \eqref{Eq:PerfectMemory} and \eqref{Eq:LimitedMemory}, respectively (lower right corner).
+    The resulting average cooperation rate depends on which of these two neighborhoods is visited more often. 
+    Simulations are run for $T\!=\!10^7$ time steps, using a cost $c\!=\!1$, a continuation probability of $\delta\!=\!0.999$ and a selection strength of $\beta\!=\!1$, in a population of size $N\!=\!100$.}
 \label{fig:expected_and_stochastic_for_donation}
@@
 
-\begin{figure}[!htbp]
+\clearpage
+\newpage
+
+\begin{figure}[t]
     \centering
     \includegraphics[width=.75\textwidth]{static/cooperation_rate_over_b_and_beta.pdf}
-    \caption{{\bf The evolution of cooperation and generosity for different
-    values of benefit (A) and strength of selection (B).} We report the average
-    cooperation and the average reciprocity. The average cooperation rate is the
-    average cooperation rate within the resident population. For the average
-    reciprocity we select the residents that have a $p \approx 1$ and we take
-    the average of their cooperation probability $q$. ({\bf A}) We vary the
-    benefit of cooperation $b$. In all cases, perfect memory updating payoffs
-    appear to overestimate the average cooperation rate the population achieves.
-    As expected in the case of limited memory the average generosity over the
-    different values of benefit remains the same ($q \approx 0.5$), and as a
-    result so does the average cooperation. ({\bf B}) We vary the selection
-    strength $\beta$. For weak selection, \(\beta < 1\), the two methods yield
-    similar results. However, as \(\beta\) increases in the case of limited
-    memory payoffs the resident populations become more defective. Note that
-    in the case of perfect payoff memory we see an increase in the cooperation
-    rate even though the generosity remains stable. That is because the
-    generosity does remain the same, however, now cooperative strategies remain
-    fixed as the resident strategy for longer.
-    Unless
-    explicitly varied, the parameters of the simulation are $N\!=\!100$,
-    $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run
-    for $T\!=\!5\times 10^7$ time steps for each parameter
-    combination.}\label{fig:cooperation_rate_over_benefit_and_beta}
+    \caption{{\bf Evolution of direct reciprocity for different parameter values.} 
+    To explore the robustness of our results, we have run simulations for different benefit values (left panels, A), and for different selection strengths (right panels, B). 
+    In each case, we record the resulting average cooperation rate over the entire simulation (upper panels). 
+    In addition we record the individuals' average generosity. 
+    Here, we only take into account those residents with $p\! \approx\! 1$ and we compute
+    the average of their cooperation probability~$q$. 
+    These simulations suggest that perfect payoff memory consistently leads to more cooperation and more generosity. 
+    Unless explicitly varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. 
+    Simulations are run for $T\!=\!5\times 10^7$ time steps for each parameter combination.}
+    \label{fig:cooperation_rate_over_benefit_and_beta}
 \end{figure}
 
-\begin{figure}[!htbp]
+\begin{figure}[t]
   \centering
   \includegraphics[width=\textwidth]{static/more_memory_summary_results.pdf}
-  \caption{{\bf Average cooperation rates for different updating payoffs.}
-  From left to right, we present result on the following updating payoffs cases;
-  (a) the expected payoffs (perfect memory), (b) the last round payoff from one
-  interaction (limited memory), (c) the last round payoff from two interactions,
-  (d) the last two rounds payoffs from one interaction, (e) the last two rounds
-  payoffs from two interactions. For the updating payoffs (c) and (d) we have
-  carried out an analysis which has shown that conditional cooperators should
-  adopt a \(q\) value smaller than $\frac{\delta - 1 +
-  \frac{\sqrt{2}}{2}}{\delta}$ and $\frac{\delta + \sqrt{\delta^{2} + 1} - 1}{2 \delta}$
-  respectively. Note that as \(\delta
-  \rightarrow 1\) both right hand sides tend to \(\frac{\sqrt{2}}{2}\).
-  Regardless the cooperation rate between for case (c) is slightly higher.
-  In case (d) the cooperation rate is the second highest hinting that as we
-  allow for more information the closer we move to the perfect payoff memory.
-  We performed four pairwise non parametric tests (Mann-Whitney U) to compare
-  the cooperation distributions of the residents in case (b) to
-  cases (a), (c), (d), (e). In all four tests we reject the null hypothesis
-  with \(\alpha=0.05\) and \(p \approx 0\). Thus, there is significant
-  difference between the cooperation rates.
-  Unless explicitly varied, the parameters of the simulation are
-  $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$.
-  Simulations are run for $T\!=\!5\times 10^7$ time steps for each parameter
-  combination.}\label{fig:cooperation_rate_all_updating_payoffs}
+  \caption{{\bf Average cooperation rates for different payoff memories.}
+  We vary how much information individuals take into account when updating their strategies. 
+  From left to right, we consider the following cases.
+  ({\it i}) Updating occurs based on expected payoffs (perfect memory), 
+  ({\it ii}) it occurs based on the last round of one interaction (limited memory), 
+  ({\it iii}) based on the last round of two interactions,
+  ({\it iv}) based on the last two rounds of one interaction, and 
+  ({\it v}) based on the last two rounds of two interactions. 
+  Again, simulations are run either for a comparably high benefit of cooperation~($b/c\!=\!10$), or for a low benefit~($b/c\!=\!3$). 
+  We observe that perfect memory always yields the highest cooperation rate. 
+  However, when individuals take into account at least two past interactions -- cases ({\it iii}) to ({\it v}) -- evolving cooperation rates are close to this optimum. 
+  Baseline parameters are the same as in \FigDependenceParameters.}
+\label{fig:cooperation_rate_all_updating_payoffs}
 \end{figure}

### 2023-10-19 paper/si.tex
CH: Revising the model part of the SI

@@
 
+\def\alld{\texttt{ALLD}}
+\def\tft{\texttt{TFT}}
+\def\gtft{\texttt{GTFT}}
+\def\esm{electronic supplementary material}
+
 \definecolor{expectedcolor}{rgb}{0.1791464821222607, 0.49287197231833907, 0.7354248366013072}
@@
 
-\title{\bf  \sffamily \LARGE Supplementary Information: Evolution of cooperation among individuals with
-limited updating payoff memory\\}
+\title{{ \sffamily \LARGE Electronic supplementary material}\\{\bf \sffamily \LARGE Evolution of reciprocity with limited payoff memory}}
 \date{}
-\author{Nikoleta E. Glynatsi, Christian Hilbe, Alex McAvoy}
+\author{Nikoleta E. Glynatsi, Alex McAvoy,  Christian Hilbe}
 
@@
 
-Section~\ref{section:pairwise_comparison} gives a brief overview of the pairwise
-comparison process. The pairwise comparison process consists of three phases;
-(1) the mutation phase (2) the game phase and (3) the update phase. In the
-update phase an individual adopts the strategy of another individual based on
-their ``updating payoffs''. Here we study different methods for calculating
-updating payoffs and we compare their results using as an example the prisoner's
-dilemma. In Section~\ref{section:perfect_memory} we describe the conventional
-method used in the literature and in Section~\ref{section:limited_memory} we
-present a new method which we refer to as 
-limited payoff memory. In the limited payoff memory method we assume that
-individuals update based on their last round payoff against one other member of
-the population. The framework can easily be extended to consider more rounds,
-more interactions or both. In
-Sections~\ref{section:m_one_n_two}-\ref{section:m_two_n_two} we discuss each of
-these extensions and for each extension we present a special case.
-In Section~\ref{section:simulation_results} we present numerical results of the
-pairwise comparison process using the different updating payoff methods.
-In Sections~\ref{section:memory_one} and~\ref{section:mutation} we
-verify the main result of this work when we no longer assume (i) that
-individuals use reactive strategies but instead memory one strategies (ii) low
-mutation.
-
-\section{Pairwise comparison process}\label{section:pairwise_comparison}
-
-Pairwise comparison process is a stochastic process for modelling the evolution
-of a finite population. The process starts with assigning all individuals of the
-population the same strategy. A strategy is a set of rules of how an individual
-should behave in an interaction with another individual. Each elementary time
-step of the process consists of three phases; (1) the \textbf{mutation phase}
-(2) the \textbf{game phase} and (3) the \textbf{update phase}. These are summarised
-in Figure~\ref{fig:pairwise_phases}.
-
-In the \textbf{mutation phase} one individual is
-chosen to switch to a new mutant strategy with a probability \(\mu\).
-In the \textbf{game phase} individuals are randomly matched with other
-individuals in the population and they engage in a repeated game where each
-subsequent turn occurs with a fixed probability \(\delta\). At each turn the
-individuals decide on an action based on their strategies.
-In repeated games there are infinitely many strategies, however, it is commonly
-assumed that individuals can only choose strategies from a restricted set. One
-such set is that of reactive strategies. A reactive
-strategy considers only the previous action of the other player, and thus, a
-reactive strategy \(s\) can be written as a three-dimensional vector \(s=(y, p,
-q)\). The parameter \(y\) is the probability that the strategy opens with a
-cooperation and \(p\), \(q\) are the probabilities that the strategy cooperates
-given that the opponent cooperated and defected equivalently.
-
-In the \textbf{update stage} two individuals are randomly selected. From
-the two individuals, one  serves as the `learner' and the other as the `role
-model'. The learner adopts the role model's strategy with a probability \(\rho\)
-given by,
-
+%% BRIEF INTRO %%
+
+\noindent
+This document provides further details on our model and our analytical methods, and it contains additional simulation results. 
+Section~\ref{section:model} gives a brief summary of our model. 
+In particular, we provide further details on our implementation of the evolutionary dynamics, and our use of the rare-mutations limit. 
+In Section~\ref{section:analyticalresults} we derive analytical results for the various settings we consider.
+These settings differ in what kind of payoff information individuals take into account when updating their strategies. 
+In the perfect memory setting, individuals take into account all their interactions against all co-players.
+In the limited memory setting, they only consider the very last round of their very last interaction.  
+In addition, we describe several model extensions in which the amount of information taken into account is in between these two extremes. 
+Finally, Section~\ref{section:furthersimulations} presents further simulation results. 
+In particular, we confirm that our main results continue to hold when ({\it i}) players use memory-1 strategies instead of reactive strategies, and  ({\it ii}) mutations are no longer rare.
+
+%%%%%%%%%
+%% MODEL  %%
+%%%%%%%%%
+
+\section{Description of the model}\label{section:model}
+
+%% MODEL: Summary of the game dynamics %%
+
+\noindent
+{\bf Summary of the model.} As described in the main text, we study cooperative behavior in a population of size $N$.
+The dynamics unfolds on two time-scales. 
+The short time scale describes the game dynamics. 
+Here individuals are randomly matched to interact in a repeated prisoner's dilemma.  
+Each round individuals can choose whether to cooperate ($C$) or defect ($D$). 
+In general, the resulting one-shot payoffs can be summarized by the payoff matrix 
+\begin{equation}\label{eq:donation}
+    \begin{blockarray}{ccc}
+        & \text{cooperate} & \text{defect} \\
+        \begin{block}{c(cc)}
+            \text{cooperate} & R & S \\
+            \text{defect} & T & P \\
+        \end{block}
+    \end{blockarray}.
+\end{equation}
+Here, $R$ is interpreted as the reward for mutual cooperation, $S$ is the sucker's payoff, $T$ is the temptation, and $P$ is the punishment payoff~\citep{axelrod1981evolution}. 
+Throughout this work, we parametrize these payoffs as $R\!=\!b\!-\!c$, $S\!=\!-c$, $T\!=\!b$, and $P\!=\!0$, where $b$ and $c$ are the benefit and cost of cooperation, respectively, with $b\!>\!c\!>0$.
+After each round, players learn their co-player's previous action, and the game continues for another round with probability~$\delta$.
+Players make their decisions whether to cooperate in any given round based on their reactive strategies $\mathbf{s}\!=\!(y,p,q)$. 
+The entry $y$ determines a player's first-round cooperation probability. 
+The other entries $p$ and $q$ determine the player's cooperation probability in all subsequent rounds. 
+The probability is $p$ if the co-player cooperated in the previous round, and it is $q$ if the co-player defected. 
+On the short time scale, the players' strategies are fixed.
+
+%% MODEL: Summary of the evolutionary dynamics %%
+
+The long time scale describes the evolutionary dynamics. 
+Here, players are allowed to update their strategies based on the payoffs they yield. 
+We model these strategy updates with a pairwise comparison process~\citep{traulsen2007pairwise}.
+This process assumes that at regular time intervals, one player is randomly selected from the population.
+We refer to this player as the `learner' ($L$). 
+The learner is then given an opportunity to update its strategy. 
+There are two possibilities for how this update may occur. 
+With probability $\mu$, the player's strategy mutates randomly. 
+In that case, the player's new strategy is drawn uniformly from the space of all reactive strategies~$[0,1]^3$.  
+With probability $1\!-\!\mu$, the player engages in a pairwise comparison. 
+In that case, the player randomly picks another individual from the population (referred to as the `role model', $R$). 
+The learner adopts the role model's strategy with a probability \(\rho\) given by
 \begin{equation} \label{Eq:rho}
-    \rho(\pi_{L}, \pi_{RM}) = \frac{1}{1\!+\! e^{\!-\!\beta (\pi_{RM}\!-\! \pi_{L})}}.
-\end{equation}
-
-\(\pi_{L}\) and \(\pi_{RM}\) are the updating payoffs of the learner and the
-role model respectively. The updating payoffs are a measure of how successful
-individuals are in the current standing of the population. The parameter
-\(\beta\) is known as the selection strength, namely, it shows how important the
-payoff difference is when the learner is considering adopting the strategy of
-the role model.
-
-\begin{figure}[!htbp]
-    \centering
-    \includestandalone[width=\textwidth]{static/pairwise_comparison_process_diagram}
-    \caption{\textbf{Pairwise comparison process phases.} \textbf{A) Initialisation.} The process begins
-    with a finite population where each member is assigned a given strategy.
-    Each color represents a different strategy, and the members are labelled by
-    letters. \textbf{B) Mutation phase.} An individual is selected (in the
-    example individual A) and with a given probability \(\mu\) that individual
-    adopts a new strategy. \textbf{C) Game phase.} Individuals are
-    selected to interact in a repeated social dilemma with other individuals. We
-    demonstrate the case where individuals B and F have been selected to
-    interact. They use the reactive strategies \(s_{B} = (y_B, p_B, q_B)\) and
-    \(s_{F} = (y_F, p_F, q_F)\) respectively. The opening moves depend on their
-    \(y_i\) probability. In turn 1, individual B cooperated,
-    thus, F cooperates with a probability \(p_{F}\) in turn 2. On the opposite,
-    individual F defected in turn 1, and so B cooperates in the next turn with
-    a probability \(q_{B}\). At each turn there is a probability \(\delta\) that a
-    subsequent turn will occur, and with a probability (\(1- \delta\)) the interaction ends.
-    \textbf{D) Update phase.} At the updating phase two individuals are chosen;
-    one serves as the role of the learner and the other one as the 
-    role model. In our example C adopts D's strategy with a probability
-    \(\rho(\pi_{C}, \pi_{D})\) where \(\pi_{C}, \pi_{D}\) denote the updating
-    payoffs of the individuals.}\label{fig:pairwise_phases}
-\end{figure}
-
-This elementary step of the process (mutation, game and update phases) is
-repeated for a large number of time steps, and at each time step we record the
-state of the population.
-
-\subsection{Low mutation \(\mu \rightarrow 0\)}
-
-In the case of low mutation (\(\mu \rightarrow 0\)) we assume that mutations are
-rare. In fact, so rare that only two different strategies can be present in the
-population at any given time. The case of low mutation is vastly adopted because
-it allows us to explicitly calculate the fixation probability of a newly
-introduced mutant.
-
-More specifically, the process again starts with a population where all members are of the same
-strategy. At each step one individual adopts a mutant strategy randomly selected
-from the set of feasible strategies. The fixation probability \(\phi_{M}\) of the
-mutant strategy can be calculated explicitly,
-
+    \rho(\pi_{L}, \pi_{R}) = \frac{1}{1\!+\! e^{\!-\!\beta (\pi_{R} - \pi_{L})}}.
+\end{equation}
+The parameter \(\beta\) is the selection strength
+It determines how important payoff differences are for the learner's decision to imitate the role model. 
+The parameters $\pi_R$ and $\pi_L$ refer to the relevant payoffs of the role model and the learner, respectively. 
+The exact value of these payoffs depend on the players' memory. 
+We say players have {\it perfect memory} when $\pi_R$ and $\pi_L$ are given by the players' expected payoffs (across all rounds and across all possible co-players). 
+We say players have {\it limited memory} when $\pi_R$ and $\pi_L$ are given by the players' realized payoff in the very last round of their very last interaction partner. 
+In addition, we consider several model extensions in which individuals have memory capacities in between these two extreme cases. 
+We provide a detailed description of these different cases and the resulting payoffs in Section~\ref{section:analyticalresults}.\\
+
+%% MODEL: Intro to the rare-mutations limit %%
+
+\noindent
+{\bf Evolutionary simulations for the rare-mutations limit.} 
+To simulate the evolutionary dynamics of the pairwise comparison process described above, it is sometimes useful to assume that mutations are rare, $\mu\!\rightarrow\!0$. 
+In that case, whenever a mutant strategy appears, this mutant either fixes in the population or goes extinct before the next mutant appears. 
+As a result, at any given time there are at most two different strategies present in the population~\citep{fudenberg:JET:2006,wu:JMB:2012,mcavoy:jet:2015}. 
+This assumptions makes computations more efficient, and it makes some of the results easier to interpret.
+In the following, we describe our implementation of the process in the rare-mutations limit in more detail.
+
+%% MODEL: Implementation of the rare-mutations limit %%
+
+Initially, the process starts with a population where all members use the same strategy (called `resident strategy'). 
+Then one individual adopts a mutant strategy selected uniformly at random from the set of feasible strategies.
+The fixation probability \(\phi_{M}\) of the mutant strategy can be calculated explicitly~\citep{nowak:Nature:2004},
 \begin{equation}\label{eq:appendix_fixation_probability}
-    \varphi_{M} = \frac{1}{1+\sum\limits_{i=1}^{N-1}\prod\limits_k^i \frac{\lambda^-_k}{\lambda^+_k}},
-\end{equation}
-
-where \(\lambda^-_k, \lambda^+_k\) are the probabilities that the number of
-mutants decreases and increases respectively, \(N\) is the size of the
-population, and \(k\) is the number of mutants. The probabilities \(\lambda^-_k
-\text{ and } \lambda^+_k\) depend on the updating payoffs of the mutant and the
-resident strategies. Depending on the fixation probability \(\phi_{M}\) the
-mutant either fixes (becomes the new resident) or goes extinct. Regardless, in
-the elementary time step another mutant strategy is introduced to the
-population. We iterate this elementary population updating process for a large
-number of mutant strategies and we record the resident strategies at each time
-step. The process is summarised by Algorithm~\ref{algorithm:pairwise_comparison}.
-
-\begin{algorithm}[!htbp]
+    \varphi_{M} = \frac{1}{1+\sum\limits_{i=1}^{N-1}\prod\limits_k^i \frac{\lambda^-_k}{\lambda^+_k}}.
+\end{equation}
+Here, the index $k$ corresponds to the current number of players with the mutant strategy (we refer to these players as `mutants'). 
+The variables \(\lambda^-_k, \lambda^+_k\) are the probabilities that the number of mutants decreases or increases within a single updating step. 
+These probabilities depend on the probability that a mutant and a resident are chosen as the learner and the resident, respectively; 
+in addition, they depend on the respective switching probability~$\rho$, as described by Eq.~\eqref{Eq:rho}. 
+We specify the exact values of  \(\lambda^-_k, \lambda^+_k\) for each memory-setting in the following section.
+
+Depending on the fixation probability \(\varphi_{M}\), the mutant strategy either fixes (becomes the new resident) or goes extinct. 
+Afterwards, another random mutant strategy is introduced into the population. 
+We iterate this elementary population updating process for a large number of mutant strategies. 
+At each step, we record the current resident strategy, and the resulting average cooperation rate. 
+We consider this limit of rare-mutations throughout the main text. 
+The respective process is summarised by Algorithm~\ref{algorithm:pairwise_comparison}.
+However, in Section~\ref{section:furthersimulations} we present additional simulation results to show that our main results continue to hold when the mutation rate is strictly bounded away from zero.
+
+%% MODEL: Algorithm
+
+\begin{algorithm}[t]
   \SetAlgoLined
   $N \leftarrow$ population size\;
-  $k \leftarrow 1$\; resident
-   $\leftarrow$ starting resident\;
+  %$k \leftarrow 1$\; 
+  resident $\leftarrow$ starting resident\;
    \While{ $t <$ maximum number of steps}
@@
    {resident $\leftarrow$ mutant;}}
-   \caption{Evolutionary process}\label{algorithm:pairwise_comparison}
+   \caption{Evolutionary process in the limit of rare mutations}\label{algorithm:pairwise_comparison}
 \end{algorithm}
 
-Most of the results we present in this work consider the case of low mutation,
-however, we have also verified that the main result holds in the case of
-high mutation rates (Section~\ref{section:mutation}).
-
-\subsection{Updating Payoffs}\label{section:updating_payoffs}
-
-The updating payoffs depend on the interactions of the individuals at the game
-phase. In this work we assume that in the game phase individuals are matched in
-pairs and that they participate in a repeated 2 person donation game. In the
-donation game there are two actions: cooperation (\(C\)) and defection (\(D\)).
-By cooperating a player provides a benefit \(b\) to the other player at their
-cost \(c\), with \(0 < c < b\). Thus the payoffs for a player in each turn are,
-
-\begin{equation}\label{eq:donation}
-    \begin{blockarray}{ccc}
-        & \text{cooperate} & \text{defect} \\
-        \begin{block}{c(cc)}
-            \text{cooperate} & b - c & -c \\
-            \text{defect} & b & 0 \\
-        \end{block}
-    \end{blockarray}.
-\end{equation}
-
-Let \(\mathbf{u} = (b-c, -c, b, 0)\) be Eq.~\ref{eq:donation} in a vector format, and let \(\mathcal{U}
-= \{r, s, t, p\}\) denote the set of feasible payoffs, where \(r\) denotes the
-payoff of mutual cooperation, \(s\) the sucker's payoff, \(t\) the temptation to
-defect payoff, and \(p\) the punishment payoff.
-
-In the following sections we present several methods for calculating the
-updating payoffs. Initially, we discuss the conventional way of the
-expected payoffs and afterwards we present our newly introduced methods.
-
-\section{Updating Payoffs based on Expected Payoffs (perfect payoff memory)}\label{section:perfect_memory}
+\newpage
+
+%%%%%%%%%%%%%%%%%
+%% ANALYTICAL RESULTS %%
+%%%%%%%%%%%%%%%%%
+
+\section{Analytical results}
+\label{section:analyticalresults}
+
+\subsection{Updating Payoffs based on Expected Payoffs (perfect payoff memory)}\label{section:perfect_memory}
+
+Let \(\mathbf{u} = (b-c, -c, b, 0)\)
 
@@
 
-\subsubsection*{Invasion Analysis}
+~\\
+{\bf Invasion Analysis}
 
@@
 
-\section{Updating Payoffs based on the Last Round Payoff of One Interaction (limited payoff memory)}\label{section:limited_memory}
+\subsection{Updating Payoffs based on the Last Round Payoff of One Interaction (limited payoff memory)}\label{section:limited_memory}
 
@@
 
-\section{Updating Payoffs based on the last round payoff of $n$ interactions}\label{section:m_one_n_two}
+\subsection{Updating Payoffs based on the last round payoff of $n$ interactions}\label{section:m_one_n_two}
 
@@
 
-\subsection{Last Round Payoff of Two Interactions}
+\subsubsection{Last Round Payoff of Two Interactions}
 
@@
 
-\section{Updating Payoffs based on the Last $m$ Rounds Payoffs of One Interaction}\label{section:m_two_n_one}
+\subsection{Updating Payoffs based on the Last $m$ Rounds Payoffs of One Interaction}\label{section:m_two_n_one}
 
@@
 
-\section{Updating Payoffs based on the last two rounds payoff of two interactions (\(m=2\) and \(n=2\))}\label{section:m_two_n_two}
+\subsection{Updating Payoffs based on the last two rounds payoff of two interactions (\(m=2\) and \(n=2\))}\label{section:m_two_n_two}
 
@@
 
-\section{Simulation Results on the Pairwise Comparison Process}\label{section:simulation_results}
+%%%%%%%%%%%%%%%%%%
+%% FURTHER SIMULATIONS %% 
+%%%%%%%%%%%%%%%%%%
+
+\section{Further simulation results}~\label{section:furthersimulations}
+
+\subsection{Simulation Results on the Pairwise Comparison Process}\label{section:simulation_results}
 
@@
 
-\section{Expected and Last Round Updating Payoffs for Memory One Strategies}\label{section:memory_one}
+\subsection{Expected and Last Round Updating Payoffs for Memory One Strategies}\label{section:memory_one}
 
@@
 
-\section{Expected and Last Round Updating Payoffs for High Mutation ($\mu \neq 0$)}\label{section:mutation}
+\subsection{Expected and Last Round Updating Payoffs for High Mutation ($\mu \neq 0$)}\label{section:mutation}
 
@@
 independent runs of the pairwise process described in
-Section~\ref{section:pairwise_comparison}, and at each time step we record the
+Section~\ref{section:model}, and at each time step we record the
 average player \(\bar{s}=(\bar{y}, \bar{p}, \bar{q})\). The average cooperation
@@
 
-\bibliographystyle{unsrt}
+%%%%%%%%%%%%
+%% REFERENCES %%
+%%%%%%%%%%%%
+
+{
+{\setlength{\bibsep}{0\baselineskip}
+\bibliographystyle{naturemag}
 \bibliography{bibliography.bib}
+}
 

### 2023-10-20 paper/si.tex
CH: Polishing the model part of the SI, revising the expected payoffs section of the results

@@
 \usepackage{amsmath,amsthm,amssymb,setspace,enumitem,epsfig,titlesec,verbatim,array,eurosym,multirow}
-\usepackage[sort&compress]{natbib}
+\usepackage[sort&compress,numbers]{natbib}
 \usepackage[footnotesize,bf]{caption}
@@
 \def\gtft{\texttt{GTFT}}
+\def\rolemodel{\text{R}}
+\def\learner{\text{L}}
+\def\resident{\texttt{R}}
+\def\mutant{\texttt{M}}
+\def\strategy{\texttt{$\sigma$}}
 \def\esm{electronic supplementary material}
@@
 \noindent
-This document provides further details on our model and our analytical methods, and it contains additional simulation results. 
-Section~\ref{section:model} gives a brief summary of our model. 
+This document provides further details on our methods and derivations, and it contains additional simulation results. 
+Section~\ref{section:model} summarizes the model. 
 In particular, we provide further details on our implementation of the evolutionary dynamics, and our use of the rare-mutations limit. 
@@
 Finally, Section~\ref{section:furthersimulations} presents further simulation results. 
-In particular, we confirm that our main results continue to hold when ({\it i}) players use memory-1 strategies instead of reactive strategies, and  ({\it ii}) mutations are no longer rare.
+In particular, we confirm that our main results continue to hold ({\it i}) when players use memory-1 strategies instead of reactive strategies, and  ({\it ii}) when mutations are no longer rare.
 
@@
 Each round individuals can choose whether to cooperate ($C$) or defect ($D$). 
-In general, the resulting one-shot payoffs can be summarized by the payoff matrix 
-\begin{equation}\label{eq:donation}
+In the most general setting, the resulting one-shot payoffs can be summarized by the payoff matrix 
+\begin{equation}\label{eq:prisoner}
     \begin{blockarray}{ccc}
@@
 Throughout this work, we parametrize these payoffs as $R\!=\!b\!-\!c$, $S\!=\!-c$, $T\!=\!b$, and $P\!=\!0$, where $b$ and $c$ are the benefit and cost of cooperation, respectively, with $b\!>\!c\!>0$.
-After each round, players learn their co-player's previous action, and the game continues for another round with probability~$\delta$.
+After each round, players learn their co-player's previous action. 
+Then the game continues for another round with probability~$\delta$.
 Players make their decisions whether to cooperate in any given round based on their reactive strategies $\mathbf{s}\!=\!(y,p,q)$. 
@@
 This process assumes that at regular time intervals, one player is randomly selected from the population.
-We refer to this player as the `learner' ($L$). 
+We refer to this player as the `learner' (\learner). 
 The learner is then given an opportunity to update its strategy. 
@@
 With probability $1\!-\!\mu$, the player engages in a pairwise comparison. 
-In that case, the player randomly picks another individual from the population (referred to as the `role model', $R$). 
+In that case, the player randomly picks another individual from the population (referred to as the `role model', \rolemodel). 
 The learner adopts the role model's strategy with a probability \(\rho\) given by
 \begin{equation} \label{Eq:rho}
-    \rho(\pi_{L}, \pi_{R}) = \frac{1}{1\!+\! e^{\!-\!\beta (\pi_{R} - \pi_{L})}}.
-\end{equation}
-The parameter \(\beta\) is the selection strength
+    \rho(\pi_{L}, \pi_{R}) = \frac{1}{1\!+\! e^{\!-\!\beta (\pi_\rolemodel - \pi_\learner)}}.
+\end{equation}
+The parameter \(\beta\) is the selection strength.
 It determines how important payoff differences are for the learner's decision to imitate the role model. 
-The parameters $\pi_R$ and $\pi_L$ refer to the relevant payoffs of the role model and the learner, respectively. 
+The variables $\pi_\rolemodel$ and $\pi_\learner$ refer to the relevant payoffs of the role model and the learner, respectively. 
 The exact value of these payoffs depend on the players' memory. 
-We say players have {\it perfect memory} when $\pi_R$ and $\pi_L$ are given by the players' expected payoffs (across all rounds and across all possible co-players). 
-We say players have {\it limited memory} when $\pi_R$ and $\pi_L$ are given by the players' realized payoff in the very last round of their very last interaction partner. 
-In addition, we consider several model extensions in which individuals have memory capacities in between these two extreme cases. 
-We provide a detailed description of these different cases and the resulting payoffs in Section~\ref{section:analyticalresults}.\\
+We say players have {\it perfect memory} when $\pi_\rolemodel$ and $\pi_\learner$ are given by the players' expected payoffs (across all rounds and across all possible co-players). 
+We say players have {\it limited memory} when $\pi_\rolemodel$ and $\pi_\learner$ are given by the players' realized payoff in the very last round of the game with their very last interaction partner.
+In addition, we consider several model extensions in which individuals have memory capacities in between these two extremes. 
+We provide a detailed description of these different settings and the resulting payoffs in Section~\ref{section:analyticalresults}.\\
 
@@
 {\bf Evolutionary simulations for the rare-mutations limit.} 
-To simulate the evolutionary dynamics of the pairwise comparison process described above, it is sometimes useful to assume that mutations are rare, $\mu\!\rightarrow\!0$. 
-In that case, whenever a mutant strategy appears, this mutant either fixes in the population or goes extinct before the next mutant appears. 
+To simulate the evolutionary dynamics of the pairwise comparison process, it is sometimes useful to assume that mutations are rare, $\mu\!\rightarrow\!0$. 
+In that case, whenever a mutant strategy appears, it either fixes in the population or goes extinct before the next mutant appears. 
 As a result, at any given time there are at most two different strategies present in the population~\citep{fudenberg:JET:2006,wu:JMB:2012,mcavoy:jet:2015}. 
@@
 
-Initially, the process starts with a population where all members use the same strategy (called `resident strategy'). 
+Initially, the process starts with a population where all members use the same strategy (referred to as the resident strategy~\resident). 
 Then one individual adopts a mutant strategy selected uniformly at random from the set of feasible strategies.
-The fixation probability \(\phi_{M}\) of the mutant strategy can be calculated explicitly~\citep{nowak:Nature:2004},
-\begin{equation}\label{eq:appendix_fixation_probability}
-    \varphi_{M} = \frac{1}{1+\sum\limits_{i=1}^{N-1}\prod\limits_k^i \frac{\lambda^-_k}{\lambda^+_k}}.
-\end{equation}
-Here, the index $k$ corresponds to the current number of players with the mutant strategy (we refer to these players as `mutants'). 
+The fixation probability \(\varphi_{\mutant}\) of the mutant strategy can be calculated explicitly~\citep{nowak:Nature:2004},
+\begin{equation}\label{eq:fixation_probability}
+    \varphi_{\mutant} = \frac{1}{1+\sum\limits_{i=1}^{N-1}\prod\limits_k^i \frac{\lambda^-_k}{\lambda^+_k}}.
+\end{equation}
+Here, the index $k$ corresponds to the current number of players with the mutant strategy (mutants). 
 The variables \(\lambda^-_k, \lambda^+_k\) are the probabilities that the number of mutants decreases or increases within a single updating step. 
-These probabilities depend on the probability that a mutant and a resident are chosen as the learner and the resident, respectively; 
-in addition, they depend on the respective switching probability~$\rho$, as described by Eq.~\eqref{Eq:rho}. 
-We specify the exact values of  \(\lambda^-_k, \lambda^+_k\) for each memory-setting in the following section.
-
-Depending on the fixation probability \(\varphi_{M}\), the mutant strategy either fixes (becomes the new resident) or goes extinct. 
+These probabilities depend on the probability that a mutant and a resident are chosen as the learner and the resident, respectively. 
+In addition, they depend on the respective switching probability~$\rho$, as described by Eq.~\eqref{Eq:rho}. 
+We specify the exact values of  \(\lambda^-_k, \lambda^+_k\) for each memory-setting in the next section.
+
+Depending on the fixation probability \(\varphi_{\mutant}\), the mutant strategy either fixes (becomes the new resident) or goes extinct. 
 Afterwards, another random mutant strategy is introduced into the population. 
@@
 At each step, we record the current resident strategy, and the resulting average cooperation rate.
+
 We consider this limit of rare-mutations throughout the main text. 
 The respective process is summarised by Algorithm~\ref{algorithm:pairwise_comparison}.
-However, in Section~\ref{section:furthersimulations} we present additional simulation results to show that our main results continue to hold when the mutation rate is strictly bounded away from zero.
+In Section~\ref{section:furthersimulations} we present additional simulation results to show that our qualitative results continue to hold when the mutation rate is strictly bounded away from zero.
 
@@
    {mutant $\leftarrow$ random strategy\;
-   fixation probability $\leftarrow \varphi_M $\;
-   \If{$\varphi_M >$ random: $i \rightarrow [0,1]$}
+   fixation probability $\leftarrow \varphi_\mutant $\;
+   \If{$\varphi_{\mutant} >$ random: $i \rightarrow [0,1]$}
    {resident $\leftarrow$ mutant;}}
@@
 
-\subsection{Updating Payoffs based on Expected Payoffs (perfect payoff memory)}\label{section:perfect_memory}
-
-Let \(\mathbf{u} = (b-c, -c, b, 0)\)
-
-The expected payoffs are the conventional payoffs used in the updating stage.
-The expected payoffs are defined as the mean payoff of an individual in a well-mixed population that
-engages in an infinitely repeated games with all other population members.
-The game between two reactive strategies $s_1\!=\!(y_1, p_1, q_1$) and $s_2\!=\!(y_2,p_2,q_2)$
-can be described by a Markovian process~\cite{nowak:APC:1989} with the
-transition matrix \(M\),
-
+%% RESULTS: General overview %%
+
+In the following, we discuss our different memory settings in more detail. 
+We discuss five cases explicitly. In these cases, updating occurs
+({\it i})~based on average payoffs based on all interactions (perfect memory), 
+({\it ii})~based on the last round of one interaction (limited memory), 
+({\it iii})~based on the last round of two interactions,
+({\it iv})~based on the last two rounds of one interaction, and 
+({\it v})~based on the last two rounds of two interactions. 
+In each case, we consider the case that there are only two strategies present in the population (a resident and a mutant strategy). 
+We first derive how likely it is that a learner (of any type) assigns a given payoff $\pi_\learner$ to itself, and a payoff of $\pi_\rolemodel$ to the role model. 
+This allows us to derive explicit expressions for $\lambda^-_k/\lambda^+_k$, and hence for the mutant's fixation probability according to Eq.~\eqref{eq:fixation_probability}.
+Based on these expressions we can characterize under which conditions cooperation is stochastically stable.
+
+%%%%%%%%%%%%%%%%%%
+%% RESULTS: Perfect memory %%
+%%%%%%%%%%%%%%%%%%
+
+\subsection{Perfect payoff memory}\label{section:perfect_memory}
+
+%% RESULTS: Expected payoffs, computing payoffs %%
+
+{\bf Computing the ratio $\lambda^-_k/\lambda^+_k$.}
+The case of perfect payoff memory corresponds to the classical case considered in the previous literature. 
+Here, individuals update their strategies based on the expected payoffs, based on all rounds and all possible interaction partners. 
+When players use reactive strategies (or more generally, strategies with finite memory), these expected payoffs can be computed explicitly, based on a Markov chain approach~\cite{sigmund2010calculus}.
+To this end, consider two players with strategies $\strategy_1\!=\!(y_1, p_1, q_1$) and $\strategy_2\!=\!(y_2,p_2,q_2)$, respectively. 
+In each round $t$ of the game, player~1 may get one of the four possible payoffs $R$, $S$, $T$, or $P$, as described by the general payoff matrix~\eqref{eq:prisoner}. 
+Let $\mathbf{v}(t)\!=\!\big(v_R(t),v_S(t),v_T(t),v_P(t)\big)$ denote the respective probability distribution of observing one of these four outcomes.  
+This probability distribution can be computed recursively. 
+Using the shortcut notation $\bar{z}\!=\!1\!-\!z$ for any $z\!\in\![0,1]$, we get for the initial round 
+\begin{equation}
+\mathbf{v}_0:=\mathbf{v}(0) = (y_1 y_2,~y_1 \bar{y}_2,~\bar{y}_1 y_2,~\bar{y}_1 \bar{y}_2).
+\end{equation}
+Given $\mathbf{v}(t)$, we can compute $\mathbf{v}(t\!+\!1)$ as  
+\begin{equation}
+\mathbf{v}(t\!+\!1) = \mathbf{v}(t)\cdot M,
+\end{equation}
+where $M$ is the transition matrix of the process, 
 \begin{equation}\label{eq:transition_matrix}
@@
 \end{equation}
-
-The stationary vector \(\mathbf{v}(s_1,s_2)\) is the solution to
-\(\mathbf{v}(s_1,s_2) M = \mathbf{v}(s_1,s_2)\). In an infinitely repeated game
-the long-term average payoff of \(s_1\) and \(s_2\) are,
-
-\[\langle\mathbf{v}(s_1,s_2),\mathbf{u}\rangle \quad  \text{ and } \quad  \langle\mathbf{v}(s_2,s_1), \mathbf{u}\rangle.\]
-
-In the case of low mutation there can be only one type of mutant strategy in the
-population. So in a population of size \(N\), there will be \(k\) mutants and
-\(N - k\) residents, whose strategies we denote respectively as \(s_M =(y_M,
-p_M, q_M)\) and \(s_R = (y_R, p_R, q_R)\).
-The expected payoffs of a resident (\(\pi_R\)) and for a
-mutant (\(\pi_M\)) are give by,
-
+Based on this recursion, we can compute how often player~1 receives one of the four payoffs $R$, $S$, $T$, $P$ on average (across all possible realizations of games among the two players).
+This average distribution $\mathbf{v}$ is
+\begin{equation}
+\mathbf{v} := (1\!-\!\delta) \sum_{t=0}^\infty \delta^t \mathbf{v}(t) 
+= (1\!-\!\delta) \mathbf{v_0} \sum_{t=0}^\infty \delta^t M^t
+= (1\!-\!\delta) \mathbf{v_0} (I_4-\delta M)^{-1},
+\end{equation}
+where $I_4$ is the $4\times4$ identity matrix. 
+Based on this general formula, the four entries of $\mathbf{v}=(v_R, v_S, v_T, v_P)$ can be computed explicitly. 
+Using the shortcut notation $r_i:=p_i\!-\!q_i$, we obtain
+    \begin{equation} \label{Eq:vAverage}
+      \setlength{\arraycolsep}{1pt}
+      \begin{array}{rcl}    
+      v_{R} &= &\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
+      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
+      v_{S}&= &\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
+      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
+      v_{T} &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
+      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
+      v_{P} &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
+      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)}.
+      \end{array}
+    \end{equation}
+Using this distribution $\mathbf{v}$, we compute player 1's expected payoff as the weighted average
+\begin{equation}
+\pi(\strategy_1,\strategy_2) = R \, v_R + S\,v_S + T\,v_T + P\,v_P. 
+\end{equation}
+
+%% RESULTS: Expected payoffs, computing transition probabilities %%
+
+~\\
+\noindent
+After these preparations, consider now a population with \(k\) mutants and \(N - k\) residents, whose strategies we denote by \(\strategy_\mutant=(y_\mutant,p_\mutant, q_\mutant)\) and \(\strategy_\resident = (y_\resident, p_\resident, q_\resident)\), respectively.
+Assuming that population members are matched randomly (or equivalently, that they interact with all other population members), the resulting expected payoffs of residents and mutants are
 \begin{equation} \label{Eq:ExpPay}
   \begin{array}{lcrcr}
-  \displaystyle \pi_R & = &\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(s_R,s_R),\mathbf{u}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(s_R,s_M),\mathbf{u}\rangle,\\[0.5cm]
-  \displaystyle \pi_M & = &\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(s_M,s_R),\mathbf{u}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(s_M,s_M),\mathbf{u}\rangle.\\
+  \displaystyle \pi_\resident(k)& = &\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \pi(\strategy_\resident,\strategy_\resident)	&+	&\displaystyle\frac{k}{N-1}\cdot \pi(\strategy_\resident,\strategy_\mutant),\\[0.5cm]
+  \displaystyle \pi_\mutant(k)& = &\displaystyle\frac{N-k}{N-1}\cdot \pi(\strategy_\mutant,\strategy_\resident) &+	&\displaystyle\frac{k-1}{N-1}\cdot \pi(\strategy_\mutant,\strategy_\mutant).\\
   \end{array}
 \end{equation}
-
-\(\frac{N\!-\!k\!-\!1}{N-1}\) is the probability that a resident meets another
-resident and \(\frac{k}{N-1}\) is the probability that a resident meets a
-mutant. Likewise, \(\frac{N-k}{N-1}\) is the probability that a mutant interacts
-with a resident and \(\frac{k-1}{N-1}\) the probability that the mutants
-interacts with a mutant.
-
-The number of mutants in the population increases if a resident adopts the strategy
-of a mutant, and decreases if a mutant adopts the strategy of a resident. The
-probabilities that the number of mutants decreases and increases,
-\(\lambda^-_k\) and \(\lambda^+_k\), are now explicitly defined as,
-
-\begin{align*}
-  \lambda^-_k \!=\!\rho(\pi_M, \pi_R) \quad \text{ and } \quad \lambda^+_k \!=\!\rho(\pi_R, \pi_M).
-\end{align*}
+The number of mutants in the population decreases in a single time step if a mutant is chosen to be the learner and adopts the strategy
+of a resident. 
+Similarly, it increases if a resident is the learner and adopts the strategy of a mutant. 
+The respective transition probabilities are 
+\begin{align*}
+  \lambda^-_k \!=\!
+\frac{N-k}{N}\frac{k}{N-1}\rho\big(\pi_\mutant(k),\,\pi_\resident(k)\big) \quad \text{ and } \quad \lambda^+_k \!=\!\frac{k}{N}\frac{N-k}{N-1}\rho\big(\pi_\resident(k),\,\pi_\mutant(k)\big).
+\end{align*}
+For $\rho$ as defined by Eq.~\eqref{Eq:rho}, the ratio of these two transition probabilities simplifies to 
+\begin{equation} \label{eq:LambdaRatio}
+ \frac{\lambda^-_k}{\lambda^+_k} \!=\!  \frac{\rho\big(\,\pi_\mutant(k),~\pi_\resident(k)\,\big)}{\rho\big(\,\pi_\resident(k),~\pi_\mutant(k)\,\big)} 
+ =e^{-\beta\big(\pi_\mutant(k)-\pi_\resident(k)\big)}.
+\end{equation}
+Based on these ratios for each $k$, we can compute the mutant's fixation probability by Eq.~\eqref{eq:fixation_probability}.
+
+%% RESULTS: Expected payoffs, invasion analysis %%
 
 ~\\
-{\bf Invasion Analysis}
-
-Let's assume that individuals use their expected payoffs at the updating phase.
-We can calculate how easily a single defecting mutant (ALLD) can invade into a
-resident population of conditional cooperators, otherwise known as generous tit
-for tat (GTFT). Let GTFT\(= (1, 1, q)\), ALLD\(= (0, 0, 0)\), and \(k =
-1\). When two GTFT players interact in an infinitely repeated game, the
-stationary distribution \(\mathbf{v}(\text{GTFT}, \text{ALLD})\) 
-simplifies to,
-
-\begin{align*}
-    \mathbf{v}(\text{GTFT}, \text{ALLD}) = (1, 0, 0, 0).
-\end{align*}
-
-On the other hand, if an ALLD player interacts with a GTFT player, the
-respective probabilities become,
-
-\begin{align*}
-  \mathbf{v}(\text{ALLD}, \text{GTFT}) = (0, q, 0, (1 - q)).
-\end{align*}
-
-Using the above we can define the payoffs of a GTFT individual (resident)
-and of the ALLD individual (mutant) follows,
-
-\begin{align*}
-  \displaystyle \pi_{\text{GTFT}} & = \displaystyle \frac{N\!-\!2}{N-1} (b - c)  -	\displaystyle\frac{q c}{N-1} \quad \text{ and } \quad \displaystyle \pi_{\text{ALLD}}  = \displaystyle b q.
-\end{align*}
-
-As a consequence, we can calculate the ratio of transition probabilities as,
-
-\begin{align*}
-    \frac{\lambda^{+}}{\lambda^{-}} & = \frac{\rho(\pi_{\text{GTFT}}, \pi_{\text{ALLD}})}{\rho(\pi_{\text{ALLD}}, \pi_{\text{GTFT}})}  = \frac{e^{-\beta \left(\frac{(N-2) (b-c)}{N-1} - q (b + \frac{c}{N-1})\right)}+1}
-    {e^{-\beta\left(b q - \frac{b (N - 2)}{N-1} - \frac{c (N - 2 - q)}{N-1}\right)}+1}
-\end{align*}
-
-In particular, in the limit of strong selection \(\beta \rightarrow \infty\)
-and large populations \(N \rightarrow \infty\), we obtain that the ratio is than
-smaller to 1 if \(q \leq 1 - \frac{c}{b}\). Thus, ALLD is disfavored to invade
-if \(q \leq 1 - \frac{c}{b}\). For \(q = 1 -\frac{c}{b}\) the probability that
-the number of mutants increase by one equals the probability that the mutant
-goes extinct.
-
-This result will become important when we discuss the numerical results of the
-evolutionary process (Section~\ref{section:simulation_results}). We repeat this
-``invasion analysis'' for the rest of the updating payoffs cases.
-
-\subsection{Updating Payoffs based on the Last Round Payoff of One Interaction (limited payoff memory)}\label{section:limited_memory}
+{\bf Invasion Analysis.}
+As an application of this formalism, we can compute when cooperation is stochastically stable in the perfect information setting. 
+To this end, suppose there is only a single mutant, $k\!=\!1$.
+The residents adopt Generous Tit-for-Tat, $\gtft = (1, 1, q)$ and the mutant adopts $\alld\!
+= \!
+(0,0,0)$.
+When two GTFT players interact, the resulting average distribution according to Eq.~\eqref{Eq:vAverage} simplifies to
+\begin{align*}
+\mathbf{v}(\gtft,\gtft) = (1, 0, 0, 0).
+\end{align*}
+On the other hand, if \alld{} interacts with \gtft, the respective probabilities become,
+\begin{align*}
+  \mathbf{v}(\alld,\gtft) = \big(0,~0,~1\!-\!\delta\!+\!\delta q,~\delta(1\! -\! q)\big).
+\end{align*}
+Based on Eq.~\eqref{Eq:ExpPay}, we can compute the strategies' expected payoffs as
+\begin{align*}
+\pi_\gtft =  \frac{N\!-\!2}{N-1} (b - c)  \!-\! \frac{1}{N-1}\big(1\!-\!\delta\!+\!\delta q\big)c
+\quad \text{and} \quad
+\pi_\alld  = \big(1\!-\!\delta\!+\!\delta q\big)b.
+\end{align*}
+As a consequence, we can calculate the corresponding ratio of transition probabilities according to Eq.~\eqref{eq:LambdaRatio},
+\begin{align*}
+\frac{\lambda_1^{-}}{\lambda_1^{+}} = e^{-\beta \left( (1-\delta+\delta q)(b+\frac{c}{N-1})- \frac{N-2}{N-1}(b-c) \right)}
+\end{align*}
+By definition, cooperation is stochastically stable if this ratio exceeds one, which is equivalent to
+\begin{equation*}
+q<1-\frac{1}{\delta}\cdot \frac{b+(N\!-\!1)c}{(N\!-\!1)b+c}.
+\end{equation*}
+For such a strategy to be feasible we require $q\!>\!0$, which implies $\delta \!>\! \big(b+(N\!-\!1)c\big)/\big((N\!-\!1)b\!+\!c\big)$.
+In particular, in the limit of large populations \(N \!\rightarrow\! \infty\), we obtain that cooperation is stochastically stable if  \(q \!<\! 1\! -\! c/(\delta b)\). 
+The minimum continuation probability for such a strategy to exist is $\delta \!>\! c/b$. 
+In this way, we recover the classical conditions for cooperation to be feasible under direct reciprocity~\citep{molander:jcr:1985,Nowak1992tit,Schmid:NHB:2021}.
+
+%%%%%%%%%%%%%%%%%%
+%% RESULTS: Limited memory %%
+%%%%%%%%%%%%%%%%%%
+
+\subsection{Limited payoff memory}\label{section:limited_memory}
 
@@
 
-      v_{r}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(q_1+l_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+l_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
-      {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)},\\[1cm]
-
-      v_{s}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(q_1+l_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
-      {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)},\\[1cm]
-
-      v_{t}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+l_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
-      {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)},\\[1cm]
-
-      v_{p}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
-      {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)}.
+      v_{r}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
+      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
+
+      v_{s}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
+      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
+
+      v_{t}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
+      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
+
+      v_{p}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
+      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)}.
       \end{array}
@@
 
+\clearpage
+\newpage
+
 {
@@
 
+\christian{Adjust symbols in main text -- for role model, strategy, etc}
 \end{document}

### 2023-10-26 paper/si.tex

@@
 \newcommand{\R}{\mathbb{R}}
+\usepackage{tikz}
+\usetikzlibrary{arrows}
 
@@
 \def\gtft{\texttt{GTFT}}
-\def\rolemodel{\text{R}}
+\def\rolemodel{\text{RM}}
 \def\learner{\text{L}}
@@
 \def\mutant{\texttt{M}}
-\def\strategy{\texttt{$\sigma$}}
+\def\strategy{s}
 \def\esm{electronic supplementary material}
@@
 
-\usepackage{tikz}
-\usetikzlibrary{arrows}
+\renewcommand{\figurename}{Supplementary Figure}
 
@@
 {9pt}%      Space below
-{\it}% 	   Body font
+{}% 	   Body font
 {}%         Indent amount (empty = no indent, \parindent = para indent)
-{\bfseries}% Thm head font
+{\it}% Thm head font
 {.}%        Punctuation after thm head
@@
 Finally, Section~\ref{section:furthersimulations} presents further simulation results. 
-In particular, we confirm that our main results continue to hold ({\it i}) when players use memory-1 strategies instead of reactive strategies, and  ({\it ii}) when mutations are no longer rare.
+In particular, we confirm that our main results continue to hold ({\it i}) when mutations are no longer rare, and  ({\it ii}) when players use memory-1 strategies instead of reactive strategies.
 
@@
 \noindent
-{\bf Summary of the model.} As described in the main text, we study cooperative behavior in a population of size $N$.
+{\bf Summary of the model.} As described in the main text, we study cooperative behavior in a population of size $N$, with $N$ being even.
 The dynamics unfolds on two time-scales. 
 The short time scale describes the game dynamics. 
-Here individuals are randomly matched to interact in a repeated prisoner's dilemma.  
-Each round individuals can choose whether to cooperate ($C$) or defect ($D$). 
+Here the $N$ individuals are randomly matched to form $N/2$ pairs to interact in a repeated prisoner's dilemma.  
+Each round, individuals can choose whether to cooperate ($C$) or defect ($D$). 
 In the most general setting, the resulting one-shot payoffs can be summarized by the payoff matrix 
@@
 The probability is $p$ if the co-player cooperated in the previous round, and it is $q$ if the co-player defected. 
-On the short time scale, the players' strategies are fixed.
+On this short time scale, the players' strategies are fixed, and players are consecutively matched to play many repeated games with randomly changing interaction partners.
 
@@
 Using the shortcut notation $\bar{z}\!=\!1\!-\!z$ for any $z\!\in\![0,1]$, we get for the initial round 
-\begin{equation}
+\begin{equation} \label{Eq:v0}
 \mathbf{v}_0:=\mathbf{v}(0) = (y_1 y_2,~y_1 \bar{y}_2,~\bar{y}_1 y_2,~\bar{y}_1 \bar{y}_2).
@@
 Given $\mathbf{v}(t)$, we can compute $\mathbf{v}(t\!+\!1)$ as  
-\begin{equation}
+\begin{equation} \label{Eq:vt}
 \mathbf{v}(t\!+\!1) = \mathbf{v}(t)\cdot M,
@@
 where $M$ is the transition matrix of the process, 
-\begin{equation}\label{eq:transition_matrix}
-  \input{static/matrix.tex}.
+\begin{equation} \label{eq:transition_matrix}
+M=\left[
+\begin{array}{llll}
+p_1 p_2	&p_1 \bar{p}_2	&\bar{p}_1 p_2	&\bar{p}_1\bar{p}_2\\
+q_1 p_2	&q_1 \bar{p}_2 &\bar{q}_1 p_2	&\bar{q}_1\bar{p}_2\\
+p_1 q_2	&p_1 \bar{q}_2	&\bar{p}_1 q_2	&\bar{p}_1\bar{q}_2\\
+q_1 q_2	&q_1 \bar{q}_2	&\bar{q}_1 q_2	&\bar{q}_1\bar{q}_2
+\end{array}
+\right]
 \end{equation}
@@
 This average distribution $\mathbf{v}$ is
-\begin{equation}
+\begin{equation} \label{Eq:vDefinition}
 \mathbf{v} := (1\!-\!\delta) \sum_{t=0}^\infty \delta^t \mathbf{v}(t) 
@@
       {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
-      v_{S}&= &\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
+      v_{S}&= &\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2-r_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
       {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
-      v_{T} &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
+      v_{T} &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1-r_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
       {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
-      v_{P} &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
+      v_{P} &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1-r_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2-r_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
       {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)}.
@@
 ~\\
-{\bf Invasion Analysis.}
+{\bf Stochastic stability of cooperation.}
 As an application of this formalism, we can compute when cooperation is stochastically stable in the perfect information setting. 
@@
 
-In the expected payoffs case the payoff of a pair depends on the average payoff
-they received over an infinite number of turns. In the limited payoff memory,
-the payoff of a pair depends on the average payoffs they received in the last
-turn. Moreover, in expected payoffs it is assumed that a player interacts with
-every member of the population whereas in the limited payoff memory approach a
-player has one interaction.
-
-Initially, we define the probability that a reactive strategy receives the
-payoff $u\!\in\! \mathcal{U}$ in the very last round of the game against another
-reactive strategy (Proposition~\ref{proposition:last_round}).
-
-\begin{Prop}\label{proposition:last_round} Consider a repeated game, with
-    continuation probability $\delta$, between players with reactive strategies
-    $s_1\!=\!(y_1, p_1, q_1$)  and $s_2\!=\!(y_2,p_2,q_2)$ respectively. Then
-    the probability that the $s_1$ player receives the payoff $u\!\in\!
-    \mathcal{U}$ in the very last round of the game is given by
-    $v_{u}(s_1,s_2)$, as given by Eq.~(\ref{Eq:LastRound}).
-
-    \begin{equation} \label{Eq:LastRound}
-      \setlength{\arraycolsep}{1pt}
-      \begin{array}{rcl}
-
-      v_{r}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
-      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
-
-      v_{s}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
-      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
-
-      v_{t}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
-      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)},\\[1cm]
-
-      v_{p}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
-      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)}.
-      \end{array}
-    \end{equation}
-
-In these expressions, we have used the notation $l_i:=p_i\!-\!q_i$,
-$\bar{y}_i\!=\!1\!-\!y_i$, $\bar{q}_i:=1\!-\!q_i$, and
-$\bar{l}_i:=\bar{p}_i\!-\!\bar{q}_i=-l_i$ for $i\!\in\!\{1,2\}$.
+%% RESULTS: Limited memory:  problem formulation %%
+
+{\bf Computing the distribution of last-round payoffs.}
+The case of perfect payoff memory is straightforward to handle; here, every player gets the expected payoff with certainty. 
+In comparison, computing transition probabilities for the case of limited payoff memory is more elaborate. 
+Here, we need to consider the different possible outcomes that both the learner and role model may have experienced in their very last interaction. 
+A further complication arises when the learner's last interaction partner happens to be the role model. 
+In that case, the learner's and the role model's last payoff will be correlated (e.g., if the learner got the sucker's payoff of $S$, the role model's payoff is $T$ with certainty). 
+To treat the case of limited memory analytically, let $\mathcal{U} \!=\! \{R,S,T,P\}$ be the set of possible one-shot payoffs. 
+We first  compute the probability that a player receives a payoff of $u\!\in\!\mathcal{U}$ in the final round of the repeated game when the strategies of the two players are given.
+
+%% RESULTS: Limited memory: probability distribution for final payoff for two fixed strategies%%
+
+\begin{Prop}\label{proposition:last_round} 
+Consider a repeated game with continuation probability $\delta$, between two players with reactive strategies $\strategy_1\!=\!(y_1, p_1, q_1$)  and $\strategy_2\!=\!(y_2,p_2,q_2)$. 
+Let $\tilde{v}_u\!=\!\tilde{v}_u(\strategy_1,\strategy_2)$ denote the probability that the first player receives the payoff $u\!\in\!\mathcal{U}$ in the final round of the game. 
+Then this probability  $\tilde{v}_u$ is identical to the player's average probability $v_u$ to receive payoff $u$ across all rounds of the game, as specified by Eq.~\eqref{Eq:vAverage}.
 \end{Prop}
 
-Note that in the proposition we here we focus on the case of the donation game/prisoner's
-dilemma but the result applies to any \(2 \times 2\) symmetric game.
-
 \begin{proof}
-Given a play between two reactive strategies with continuation probability
-$\delta$. The outcome at turn \(t\) is given by,
-
-\begin{equation}\label{eq:}
-  (1 - \delta) \mathbf{v_0} \sum \delta^{t} M^{(t)},
-\end{equation}
-
-where $\mathbf{v_0}$ denotes the expected distribution of the four outcomes in
-the very first round, and \(1- \delta\) the probability that the game ends. It
-can be shown that,
-
-\begin{align*}
-  (1 - \delta) \mathbf{v_0} \sum \delta^{t} M^{(t)} & = (1 - \delta)(\mathbf{v_0} + \delta \mathbf{v_0} M + \delta^{2}\mathbf{v_0} M ^{2} + \dots )\\ 
-   & = (1 - \delta)\mathbf{v_0} (1 + \delta M + \delta^{2}M ^{2} + \dots ) \text{ using standard formula for geometric series}\\ 
-   & = (1 - \delta)\mathbf{v_0}(I_4 - \delta M)^{-1}
-\end{align*}
-
-where \((1 - \delta)\mathbf{v_0}(I_4 - \delta M)^{-1}\) is vector \(\in R^{4}\)
-and it the probabilities for being in any of the outcomes \(CC, CD, DC, DD\) in
-the last round. Combining this with the payoff vector \(u\) and some algebraic
-manipulation we derive to the Equation~\ref{Eq:LastRound}.
+Let $\mathbf{\tilde{v}}:=(\tilde{v}_R, \tilde{v}_S, \tilde{v}_T, \tilde{v}_P)$. 
+To compute $\mathbf{\tilde{v}}$, we first compute the probability to receive each outcome $u\!\in\!\mathcal{U}$ given that the last round happens to be round $\tau\!\in\!\{0,1,\ldots\}$. 
+In that case, it follows from Eqs.~\eqref{Eq:v0} and~\eqref{Eq:vt} that the respective probability is
+\begin{equation} \label{Eq:vTildeSimple}
+\mathbf{\tilde v}(\tau) = \mathbf{v_0}\cdot M^\tau.
+\end{equation}
+Round $\tau$ is the last round with probability $\delta^\tau(1\!-\!\delta)$. 
+By the law of total probability, we can obtain $\mathbf{\tilde{v}}$ by summing up over all possible last rounds,
+\begin{equation}
+\tilde{v} = \sum_{\tau=0}^\infty \delta^\tau(1\!-\!\delta)  \cdot \mathbf{v_0}M^\tau 
+=  (1\!-\!\delta)\mathbf{v_0} \sum_{t=0}^\infty \delta^t  M^t. 
+\end{equation}
+This expression for $\mathbf{\tilde{v}}$ coincides with the expression for $\mathbf{v}$ in Eq.~\eqref{Eq:vDefinition}.
 \end{proof}
 
-At each step of the evolutionary process we choose a role model and a learner to
-update the population. In this case both the role model and the learner estimate
-their fitness after interacting with a single member of the population, and so
-there are five possible pairings at each step. They interact with each other with a
-probability \(\frac{1}{N - 1}\), and they do not interact with other with a
-probability \(1 - \frac{1}{N - 1}\). In the latter case, each of them can
-interact with either a mutant or a resident. Both of them interact with a mutant
-with a probability $\frac{(k-1)(k-2)}{(N-2)(N-3)}$ and both interact with a
-resident with a probability $\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$. The last two
-possible pairings are that either of them interacts with a resident whilst the
-other interacts with a mutant, and this happens with a probability
-$\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$. Given the possible pairings and
-Proposition~\ref{proposition:last_round}, we define the probability that the
-respective last round payoffs of two players \(s_1, s_2\) are given by $u_1$ and
-$u_2$ as,
-
+\noindent
+Note that while we focus on the prisoner's dilemma in this work, the Proposition applies to any repeated matrix game.\\
+
+%% RESULTS: Limited memory - probability distribution in population setup %%
+
+\noindent
+{\bf Computing the ratio $\lambda^-_k/\lambda^+_k$.}
+After these preparations, let us again consider the corresponding population setup, with $N\!-\!k$ residents with strategy  \(\strategy_\resident = (y_\resident, p_\resident, q_\resident)\) and $k$ mutants with strategy  \(\strategy_\mutant = (y_\mutant, p_\resident, q_\mutant)\).
+At each step of the evolutionary process we choose a learner and a role model. 
+The learner compares the performance of its strategy by comparing its own last one-shot payoff with the last one-shot payoff of the role model. 
+In the following, we assume either the learner or the role model is a resident and that the other player is a mutant (otherwise it is certain that the number of mutants does not change). 
+There are two major cases to consider.
+
+%% RESULTS: Limited payoff memory - Different cases to consider %%
+
+\begin{enumerate}
+\item The learner and the role model have their last respective interaction with each other.
+This happens with probability $1/(N\!-\!1)$.
+In that case, there are four possible cases for their joint final payoffs, $(u_\resident,u_\mutant)\!\in\!\mathcal{U}_F\!:=\!\big\{(R,R),\,(S,T),\,(T,S)\,(P,P)\big\}$.
+These outcomes are distributed according to the distribution $\mathbf{\tilde{v}}(\strategy_\resident,\strategy_\mutant)$, as described in Proposition~\ref{proposition:last_round}.
+\item The learner's last interaction was not with the role model, with probability $(N\!-\!2)/(N\!-\!1)$.
+In this case, there are four different subcases, depending on whether the resident's last interaction partner was a mutant or a resident, and depending on whether the mutant's last interaction partner was a mutant or a resident.
+As a result, the resident's last one-shot payoff is distributed according to $\mathbf{\tilde{v}}(\strategy_\resident,\strategy_\mutant)$ or $\mathbf{\tilde{v}}(\strategy_\resident,\strategy_\resident)$; the mutant's last payoff is distributed according to $\mathbf{\tilde{v}}(\strategy_\mutant,\strategy_\mutant)$ or $\mathbf{\tilde{v}}(\strategy_\mutant,\strategy_\resident)$, respectively.
+\end{enumerate}
+
+\noindent
+Let $x(u_\resident,u_\mutant)$ denote the probability that the resident and the mutant received the payoff $u_\resident$ and $u_\mutant$ in their respective last interaction.  
+By taking into account the above two cases, we can compute this probability as
 \begin{equation}\label{eq:Chi} 
 \setlength{\arraycolsep}{1pt} 
-\begin{array}{llrl}
-x(u_1,u_2)	 &=&\displaystyle \frac{1}{N\!-\!1}\cdot  &v_{u_1}(s_1,s_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}_F}\\[0.5cm]
+\begin{array}{llll}
+x(u_\resident,u_\mutant)	 &=
+&\frac{1}{N\!-\!1}\cdot  &\tilde v_{u_\resident}\!(\strategy_\resident,\strategy_\mutant)\cdot 1_{(u_\resident,u_\mutant)\in \mathcal{U}_F}\\[0.5cm]
 &+	
-&\displaystyle \left(1\!-\!\frac{1}{N\!-\!1}\right)  
-&\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_2) + 
- \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_1)\right.\\[0.5cm]
-&&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_2) + 
- \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_1)\right].
-\end{array}
-\end{equation}
-
+&\frac{N\!-\!2}{N\!-\!1} \cdot 
+&\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3}\cdot \tilde v_{u_\resident}\!(\strategy_\resident,\strategy_\mutant) \,\tilde v_{u_\mutant}\!(\strategy_\mutant,\strategy_\mutant) + 
+ \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3}\cdot \tilde v_{u_\resident}\!(\strategy_\resident,\strategy_\mutant) \,\tilde v_{u_\mutant}\!(\strategy_\mutant,\strategy_\resident)\right.\\[0.5cm]
+&&&\left.
++\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} \cdot \tilde v_{u_\resident}\!(\strategy_\resident,\strategy_\resident) \, \tilde v_{u_\mutant}\!(\strategy_\mutant,\strategy_\mutant) +
+ \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} \cdot \tilde v_{u_\resident}\!(\strategy_\resident,\strategy_\resident) \, \tilde v_{u_\mutant}\!(\strategy_\mutant,\strategy_\resident)\right].
+\end{array}
+\end{equation}
+
+%% RESULTS: Limited memory - Computing the transition probabilities lambda^- and lambda^+%%
+
+\noindent
 The first term on the right side corresponds to the case that the learner and
-the role model happened to be matched during the game stage, which happens with
-probability $\frac{1}{(N\!-\!1)}$. In that case, we note that only those payoff
-pairs can occur that are feasible in a direct interaction,
-$\splitatcommas{(u_1,u_2)\in \mathcal{U}_F:=\big\{ (r, r), (s, t), (t, s), (p,
-p) \big\}}$, as represented by the respective indicator function. Otherwise, if
-the learner and the role model did not interact directly, we need to distinguish
-four different cases, depending on whether the learner was matched with a
-resident or a mutant, and depending on whether the role model was matched with a
-resident or a mutant.
-
-The probability that the number of mutants increases, and decreases respectively,
-by one is now given by,
-
-\begin{align}
-\lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M})\cdot \rho(u_{R},u_{M}), \\
-\lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M})\cdot \rho(u_{M},u_{R}).
-\end{align}
-
-In this expression, $\frac{(N\!-\!k)}{N}$ is the probability that the randomly
-chosen learner is a resident, and $\frac{k}{N}$ is the probability that the role
-model is a mutant. The sum corresponds to the total probability that the learner
-adopts the role model's strategy over all possible payoffs $u_R$ and $u_M$ that
-the two players may have received in their respective last rounds. We use
-$x(u_R,u_M)$ to denote the probability that the randomly chosen resident
-obtained a payoff of $u_R$ in the last round of his respective game, and that
-the mutant obtained a payoff of $u_M$.
+the role model happened to be matched directly for their last interaction. 
+In that case, only those payoff pairs can occur that are feasible in a direct interaction. 
+That is, it needs to be the case that $(u_\resident,u_\mutant)\in \mathcal{U}_F$, as represented by the respective indicator function. 
+The probability that the number of mutants increase or decreases by one is now given by
+\begin{equation}
+\begin{array}{l}
+\displaystyle \lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_\resident,u_\mutant\,\in\,\mathcal{U}} x(u_\resident,u_\mutant)\cdot \rho(u_\resident,u_\mutant), \\[0.5cm]
+\displaystyle \lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_\resident,u_\mutant \,\in\, \mathcal{U}} x(u_\resident,u_\mutant)\cdot \rho(u_\mutant,u_\resident).
+\end{array}
+\end{equation}
+
+\noindent
+In this expression, the prefactor $(N\!-\!k)k/N^2$ gives the probability that the two players (learner and role model) have different strategies. 
+The sum corresponds to the total probability that the learner
+adopts the role model's strategy, by summing up over all possible payoffs $u_\resident$ and $u_\mutant$ that
+the two players may have received in their respective last rounds.
+
+%% RESULTS: Limited memory - stochastic stability of cooperation %%
 
 ~\\
-{\bf Invasion Analysis}
-
-We once again calculate how easily a single ALLD mutant can invade into a
-resident population of GTFT player. When two GTFT players interact in the game,
-their respective probabilities for each of the four outcomes in the last round
-simplify to,
-
+{\bf Stochastic stability of cooperation.}
+To illustrate this formalism, we again use it to characterize the stability of cooperation.  
+There is $k\!=\!1$ mutant with strategy \alld{}. The remaining players use the resident strategy \gtft. 
+When two residents interact, it follows by Proposition~\ref{proposition:last_round} and Eq.~\eqref{Eq:vAverage} that the outcome of the final round is distributed according to the distribution
 \begin{align*}
-    v_{r}(GTFT,GTFT) = 1, & \quad v_{t}(GTFT,GTFT) = 0, \\
-    v_{s}(GTFT,GTFT) = 0, & \quad v_{p}(GTFT,GTFT) = 0.
+\Big(\, \tilde v_R(\gtft,\gtft),~\tilde v_S(\gtft,\gtft),~ \tilde v_T(\gtft,\gtft),~ \tilde v_P(\gtft,\gtft)\, \Big) = (1,0,0,0). 
 \end{align*}
-
-On the other hand, if an ALLD player interacts with a GTFT player, the
-respective probabilities according to Eq.~\ref{Eq:LastRound} become
-
+Similarly, if the mutant interacts with a resident, the final round's outcome is distributed according to 
 \begin{align*}
-    v_{r}(ALLD,GTFT) & = 0, &  v_{s}(ALLD,GTFT) & = 0, \\
-    v_{t} (ALLD, GTFT ) & = 1 - \delta + \delta q, &  v_{p} (ALLD, GTFT) & = \delta(1 - q).
+\Big(\, \tilde v_R(\alld,\gtft),~\tilde v_S(\alld,\gtft),~ \tilde v_T(\alld,\gtft),~ \tilde v_P(\alld,\gtft)\, \Big) = \big(0,~0,~1\!-\!\delta\!+\!\delta q,~\delta(1\! -\! q)\big). 
 \end{align*}
-
-As a consequence, we obtain the following probabilities \(x(u_1, u_2)\) that the
-payoff of a randomly chosen GTFT player is \(u_1\) and that the payoff of the
-ALLD player is \(u_2\),
-
-\begin{align*}
-  x(r, t) = & \frac{N - 2}{N - 1} \cdot (1 - \delta + \delta q)\\
-  x(r, r) = & \frac{N - 2}{N - 1} \cdot \delta (1 - q) \\
-  x(s, r) = & \frac{1}{N - 1} \cdot (1 - \delta + \delta q) \\
-  x(p, p) = & \frac{1}{N - 1} \cdot \delta (1 - q) \\
-  x(u_1, u_2) = &  0 \text{ for all other payoff pairs } (u_1, u_2).
-\end{align*}
-
-We now calculate the ratio of transition probabilities as
-
+Based on these, we compute the probability $x(u_\resident,u_\mutant)$ that the
+payoff of a randomly chosen \gtft{} player is~\(u_\resident\) and that the payoff of the
+\alld{} player is \(u_\mutant\), with $u_\resident,u_\mutant \!\in\!\mathcal{U}$. We obtain
+\begin{equation*}  \setlength{\arraycolsep}{0.5cm}
+\begin{array}{ll}
+  x(R, T) =  \frac{N - 2}{N - 1} \cdot (1 - \delta + \delta q)
+  &x(R, P) =  \frac{N - 2}{N - 1} \cdot \delta (1 - q) \\[0.25cm]
+  x(S, T) =  \frac{1}{N - 1} \cdot (1 - \delta + \delta q) 
+  &x(P, P) =  \frac{1}{N - 1} \cdot \delta (1 - q) \\[0.25cm]
+  \multicolumn{2}{c}{x(u_\resident, u_\mutant) = ~0 \text{ for all other payoff pairs } (u_\resident, u_\mutant).}
+  \end{array}
+\end{equation*}
+Based on these expressions, we calculate the ratio of transition probabilities as
 \begin{equation*}
-\frac{\lambda^{+}}{\lambda^{-}} = \frac{ \frac{N - 2}{N - 1}  \left( \frac{ \delta  \left(1 - q \right)}{1
-+ e^{-  \beta  \left(b - c \right)}} +  \frac{ \delta q -  \delta + 1}{e^{ \beta
-c} + 1} \right)  +  \frac{1}{N-1}  \left(\frac{ \delta  \left(1 - q \right)}{2} +
- \frac{ \delta q -  \delta + 1}{1 + e^{-  \beta  \left(- b - c \right)}}\right)}
- { \frac{N - 2}{N - 1}  \left( \frac{ \delta  \left(1 - q \right)}{1 +
-e^{-  \beta  \left(- b + c \right)}} +  \frac{ \delta q -  \delta + 1}{1 + e^{-
- \beta c}} \right) +  \frac{1}{N -1} \left(\frac{ \delta  \left(1 - q \right)}{2} +
- \frac{ \delta q -  \delta + 1}{1 + e^{-  \beta  \left(b + c \right)}}\right)}
+\frac{\lambda^{-}_1}{\lambda^{+}_1} = \frac{ \frac{N - 2}{N - 1}  
+\left(\frac{ 1 - \delta + \delta q}{1+e^{ \beta c} } 
+ + \frac{ \delta  \left(1 - q \right)}{1+ e^{-  \beta  \left(b - c \right)}} \right)  
+ +  \frac{1}{N-1}  \left(\frac{1 - \delta + \delta q}{1 + e^{\beta  \left( b + c \right)}}
+  + \frac{ \delta  \left(1 - q \right)}{2} \right)}
+  { \frac{N - 2}{N - 1}  
+\left( \frac{ 1 - \delta + \delta q}{1 + e^{-\beta c}}
++  \frac{ \delta  (1 - q)}{1 +e^{  \beta (b - c)}} \right) 
++  \frac{1}{N -1} \left(\frac{ 1 - \delta + \delta q}{1 + e^{-  \beta (b + c)}}
++  \frac{ \delta  \left(1 - q \right)}{2} \right)}
 \end{equation*}
-
-In particular, in the limit of strong selection \(\beta \rightarrow \infty\) and
-large populations \(N \rightarrow \infty \), we obtain
-
-\begin{equation*}
-    \frac{\lambda^{+}}{\lambda^{-}} = \frac{1 - \delta + \delta q}{\delta(1 - q)}
-\end{equation*}
-
-This ratio is smaller than 1 (such that ALLD is disfavored to invade) if \(q <
-1- 1/(2 \delta)\). For infinitely repeated games, \(\delta \rightarrow 1\), this
-condition becomes \(q < 1/2\) (for \(q = 1/2\), the payoff of the ALLD player is
-\(r > r\) for half of the time, and it is \(p < r\) for the other half. The
+Cooperation is stochastically stable if $\lambda^-_1/\lambda^+_1\!>\!1$. 
+While one can solve this inequality for $q$, the resulting condition is somewhat lengthy.  
+To obtain a more interpretable condition, we consider the limit of strong selection \(\beta\! \rightarrow\! \infty\) and
+large populations \(N \!\rightarrow \!\infty \). In that case, because $b\!>\!c\!>\!0$, the above ratio simplifies to
+\begin{equation} \label{Eq:TransitionRatioSimple}
+    \frac{\lambda^{-}_1}{\lambda^{+}_1} = \frac{\delta(1 - q)}{1 - \delta + \delta q}.
+\end{equation}
+This ratio exceeds one if  \(q <1 \!-\! 1/(2 \delta)\). 
+For such a strategy to be feasible we require $q\!>\!0$, which in turn implies $\delta\!>\!1/2$. 
+Moreover, in the special case that games are infinitely repeated, \(\delta\! \rightarrow\! 1\), we conclude that cooperation is stochastically stable if  \(q \!<\! 1/2\). 
+(For \(q\! =\! 1/2\), the payoff of the ALLD player is
+\(T \!>\! R\) for half of the time, and it is \(P\! < \! R\) for the other half. The
 probability that the number of mutants increase by one equals the probability
@@
 
-\subsection{Updating Payoffs based on the last round payoff of $n$ interactions}\label{section:m_one_n_two}
-
-The framework of the limited memory can be generalised such that an individual
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% RESULTS: Remembering the last round of two interactions %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Updates based on the final round of two repeated games} 
+\label{section:m_one_n_two}
+
+%% RESULTS: Two interactions, last round -- General motivation %%
+
+{\bf Motivation.} 
+In the limited memory setting, individuals take into account the outcome of one round, and of one interaction. 
+This setting can be generalized such that an individual
 considers \(m\) rounds and of \(n\) interactions. Here we discuss the case that
@@
 
-At each step of the evolutionary process the role model and the learner now
-participate in \(n\) matches. We need to define the probability that for each of
+At each step of the evolutionary process, we consider the role model's and the learner's last \(n\) matches. 
+We need to define the probability that for each of
 the matches they are paired with a mutant, with a resident or with each other.
-We assume that each pair is unique, so for example the resident and role model
-can be matched together only once at each step. A representation of the process
-is given in Figure~\ref{fig:matching_tree}. In the case of \(n=1\) there are
-five possible pairs, however, the number of possible pairs increases non
-linearly as we increase the number of possible interactions. We demonstrate this
-case for \(n=2\), namely, for when the role model and the learner have two
-interactions.
-
-\begin{figure}[!htbp]
+We assume that each pair is unique, such that the learner and the role model
+can be matched at most once. 
+The case of \(n\!=\!1\) corresponds to the previous setting of limited memory. 
+There we have seen that there are five possible combinations to consider. 
+As we increase $n$, the number of possible combinations increases non-linearly; 
+for a graphical illustration, see Figure~\ref{fig:matching_tree}. 
+In the following, we study the case of \(n=2\), such that the learner takes into account the players' final payoffs of two interactions.\\
+
+%% RESULTS: Two interactions, last round -- Illustration %%
+
+\begin{figure}[t]
   \centering 
   \includegraphics[width=\textwidth]{static/matching_tree.pdf}
-  \caption{\textbf{Tree diagram of the possible pairs when the learner and the role
-  model have $n$ interactions.} In the diagram \((s_i, s_j)\) represent a possible
-  pair. We denote the role model as $s_{RM}$ and the learner as $s_{L}$, and
-  a member of the population that plays a mutant strategy is denoted as $s_{M}$ and
-  a resident strategy as $s_{R}$. The role model and
-  the learner need to be paired with $n$ other members of the population
-  (including each other). We break down the process into stages. For the first
-  pair (stage one) the role model and the mutant can be paired together (with
-  probability $\frac{1}{N-1}$) or not (with probability $1 -
-  \left(\frac{1}{N-1}\right)$). In the case that they are not paired together,
-  both can be paired with a mutant $\frac{(k-1)(k-2)}{(N-2)(N-3)}$, with a
-  resident $\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$, or one is paired with a mutant
-  whilst the other is paired with a resident $\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$.
-  There are five possible pairs for stage one as we have already discussed
-  in Section~\ref{section:limited_memory}. In
-  the second stage we need to consider if the role model and the learner
-  interacted with each other. If yes, then they can not interact with each other
-  again, and at each stage there are at most four possible pairs; they interact with
-  mutants, with residents, or the one interacts with a mutant whilst the second
-  interacts with a resident. In the case that they were not paired in stage 1
-  there are five possible pairs. These pairs are the same as in stage 1,
-  however now the probabilities differ. Namely, let's assume that in the
-  first stage both were matched with a mutant. They interact with each
-  other with a probability $\frac{1}{N-2}$ or not with a probability
-  $1 - \left(\frac{1}{N-2}\right)$. In the latter case there are four
-  possible pairs.
-  Both can be paired with a mutant $\frac{(k-1)(k-2)}{(N-3)(N-4)}$, with a
-  resident $\frac{(N-k-1)(N-k-2)}{(N-3)(N-4)}$, or one is paired with a mutant
-  whilst the other is paired with a resident $\frac{(N-k-1)(k-2)}{(N-3)(N-4)}$.
-  The process continues following the same logic until it reaches the $n^{\text{th}}$
-  stage.}\label{fig:matching_tree}
+  \caption{\textbf{Possible pairs when the learner takes into account $n$ interactions against different co-players.} 
+  In this diagram, \((\strategy_i, \strategy_j)\) represents a possible pairing of two players, lines indicate possible cases, and the fractions represent the respective probabilities for each case. 
+  We consider $n$ stages in which the population members are consecutively paired with each other to play a repeated game. 
+  In the first stage, we need to consider five possible cases, as in Section~\ref{section:limited_memory}.
+  One case arises if the learner and the role model are matched directly. 
+  This happens with probability $1/(N\!-\!1)$.
+  If they are not matched directly, both can be paired with a mutant, with a resident, or one is paired with a mutant
+  whilst the other is paired with a resident.
+  The second stage is similar. 
+  However, here we need to take into account that the learner and the role model can only be matched directly if they have not already interacted during the previous stage. 
+  The process continues until the $n$-th stage.}\label{fig:matching_tree}
 \end{figure}
 
-\subsubsection{Last Round Payoff of Two Interactions}
-
-Here the learner and role model consider the last round payoff of two
-interactions. At each time step of the evolutionary process there are twenty
-four possible pairs. At the first stage there are fiver possible pairs. For four of
-these pairs there are five possible pairs in the second stage (\(4 \times 5 =
-20\)), and for the last one there are four possible pairs in the second stage
-(20 + 4 = 24).
-
-We assume that player \(s_1\) receives the payoff
-\(u_{11}\) with their first interaction and \(u_{12}\) with their second.
-Respectively, player \(s_2\) receives \(u_{21}\) and \(u_{22}\). We define the
-probability that the respective last round payoffs of the two players \(s_1,
-s_2\) are given by \((u_{11}, u_{11})\) and \((u_{21}, u_{22})\) as
-\(x^{2}((u_{11}, u_{11}), (u_{21}, u_{22}))\) given by,
-
-\begin{equation}
-  \resizebox{.99\textwidth}{!}
-  {%
-$
+%% RESULTS: Two interactions, last round --  Derivation of the x-function %%
+
+\noindent
+{\bf Computing the ratio $\lambda^-_k/\lambda^+_k$.}
+As before, we assume that either the learner or the role model is a resident, and that the other player involved in the pairwise comparison is a mutant.
+For $n\!=\!2$, there are 24 possible combinations to consider.
+At the first stage there are five possible combinations. 
+These are the same as in the previous setting. 
+Either the learner interacts directly with the role model; otherwise we need to take into account whether the resident interacts with a resident or a mutant, and whether the mutant interacts with a resident or a mutant.
+In case the learner and the role model did not directly interact with each other during the first stage, there are again five possible combinations in the second stage. 
+Otherwise, if there already was a direct interaction, there are four combinations. 
+Hence, there are $4\!\cdot\!5 + 4=24$ combinations in total.
+
+%% RESULTS: Two interactions, last round --  Formula for the x function %%
+
+We assume that player the resident receives the payoff \(u_{\resident1}\) with their first interaction and \(u_{\resident 2}\) with their second.
+Similarly, the mutant receives payoffs \(u_{\mutant 1}\) and \(u_{\mutant 2}\). 
+Let \(x(u_{\resident1} u_{\resident1}, u_{\mutant 1} u_{\mutant 2})\) denote the probability that the resident and the mutant received payoffs  \(u_{\resident1}\), \(u_{\resident2}\) and \(u_{\mutant 1}\), \(u_{\mutant 2}\) in the last round of their respective last two interactions.
+This probability can be computed as follows, 
+\begin{equation} \label{Eq:xTwoInteractions} 
+  \setlength{\arraycolsep}{0.5cm}
+  \footnotesize
+  \hspace*{-0.8cm} 
+  \begin{array}{lll}
+  \multicolumn{3}{l}{x(u_{\resident1}u_{\resident 2}, u_{\mutant 1}u_{\mutant 2})= \displaystyle \frac{1}{N\!-\!1}\cdot \tilde{v}_{u_{\resident1}}(\strategy_\resident,\strategy_\mutant)\cdot 1_{(u_{\resident1},u_{\mutant 1})\in \mathcal{U}_F} \cdot A }\\[0.4cm]
+
+  &\multicolumn{2}{l}{+ \frac{N-1}{N-2} \biggl[
+  \frac{(k - 1)(k - 2)}{(N\!-\!2)(N\!-\!3)}\,
+  \tilde{v}_{u_{\resident1}}\!(\strategy_\resident,\strategy_\mutant) \, \tilde{v}_{u_{\mutant 1}}\!(\strategy_\mutant,\strategy_\mutant) 
+  \left( \frac{\tilde{v}_{u_{\resident 2}}\!(\strategy_\resident,\strategy_\mutant)}{N-2} \, 1_{(u_{\resident 2},u_{\mutant 2})\in \mathcal{U}_F} +\frac{N-3}{N-2} [B_1 \!+\!
+B_2 \!+\!
+B_3 \!+\!
+B_4]\right)}\\[0.4cm]
+
+  &&+ \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N\!-\!2)(N\!-\!3)}\,
+  \tilde{v}_{u_{\resident1}}\!(\strategy_\resident,\strategy_\mutant) \, \tilde{v}_{u_{\mutant 1}}\!(\strategy_\mutant,\strategy_\resident)  
+  \left(\frac{\tilde{v}_{u_{\resident 2}}\!(\strategy_\resident,\strategy_\mutant)}{N-2} \, 1_{(u_{\resident 2},u_{\mutant 2})\in \mathcal{U}_F} + \frac{N-3}{N-2} [C_1 \!+\!
+C_2 \!+\!
+C_3 \!+\!
+C_4]\right) \\[0.4cm]
+
+  &&+ \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N\!-\!2)(N\!-\!3)} \,
+  \tilde{v}_{u_{\resident1}}\!(\strategy_\resident,\strategy_\resident) \, \tilde{v}_{u_{\mutant 1}}\!(\strategy_\mutant,\strategy_\mutant)  
+  \left(\frac{\tilde{v}_{u_{\resident 2}}\!(\strategy_\resident,\strategy_\mutant)}{N-2} \, 1_{(u_{\resident 2},u_{\mutant 2})\in \mathcal{U}_F} + \frac{N-3}{N-2} [D_1 \!+\!
+D_2 \!+\!
+D_3 \!+\!
+D_4]\right) \\[0.4cm]
+
+  &&+ \frac{\left(N - k - 2\right) \left(N - k - 1\right)}{(N\!-\!2)(N\!-\!3)} 
+  \tilde{v}_{u_{\resident1}}\!(\strategy_\resident,\strategy_\resident) \, \tilde{v}_{u_{\mutant 1}}\!(\strategy_\mutant,\strategy_\resident)  
+  \left(\frac{ \tilde{v}_{u_{\resident 2}}\!(\strategy_\resident,\strategy_\mutant)}{N-2} \,1_{(u_{\resident 2},u_{\mutant 2})\in \mathcal{U}_F} + \frac{N-3}{N - 2} [E_1 \!+\!
+E_2 \!+\!
+E_3 \!+\!
+E_4]\right) \biggl].
+  \end{array}
+\end{equation}
+
+%% RESULTS: Two interactions, last round --  Explaining and defining the various constants in the x-formula %%
+
+\noindent
+The first line on the right side corresponds to the case that the learner and
+the role model are matched during the first stage. 
+The factor $A$ considers the different pairings that can then occur in the second stage, 
+\begin{equation*}
   \setlength{\arraycolsep}{1pt}
+  \begin{array}{ll}
+  A  = \frac{N-2}{N-1} & \left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} \, \tilde{v}_{u_{\mutant 1}}(\strategy_\resident,\strategy_\mutant)\, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) +
+   \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} \, \tilde{v}_{u_{\mutant 1}}(\strategy_\resident,\strategy_\mutant) \, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident)  \right.
+\\[0.5cm]
+   &+ \left.
+\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} \, \tilde{v}_{u_{\mutant 1}}(\strategy_\resident,\strategy_\resident) \, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) +
+   \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} \, \tilde{v}_{u_{\mutant 1}}(\strategy_\resident,\strategy_\resident) \, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident)\right]
+   \end{array} 
+\end{equation*}   
+ The second line of Eq.~\eqref{Eq:xTwoInteractions} corresponds to the case that during the first stage, both the resident and the mutant interact with a mutant. 
+ In the seconds stage, they may interact with each other. 
+ If they do not, then again either both of them interact with a mutant, one of them interacts with a resident whereas the other interacts with a mutant, or both interact with a resident.
+These cases are captured by the variables
+\begin{equation*}
+  \setlength{\arraycolsep}{10pt}
+  \begin{array}{ll}
+   B_1  = \frac{(k - 2) (k - 3)}{(N - 3) (N - 4)}\, \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant) \,\tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant),
+   &B_2 = \frac{\left(k - 2\right) \left(N - k - 1\right)}{(N - 3) (N - 4)}\, \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident) \,\tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant), \\[0.5cm]
+   B_3  = \frac{\left(k - 2\right) \left(N - k - 1\right)}{(N - 3) (N - 4)}\, \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant) \,\tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident),
+   &B_4 = \frac{\left(N - k - 2\right) \left(N - k - 1\right)}{(N - 3) (N - 4)}\, \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident) \, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident).
+   \end{array}
+\end{equation*}
+ The third line of Eq.~\eqref{Eq:xTwoInteractions} corresponds to the case that during the first stage, the resident interacts with a mutant whereas the mutant interacts with a resident.
+The variables for the pairings in the second stage are
+\begin{equation*}
+  \setlength{\arraycolsep}{10pt}
   \begin{array}{llrl}
-  x^{2}((u_{11}, u_{12}), (u_{21}, u_{22}))	& = \displaystyle \frac{1}{N\!-\!1}\cdot v_{u_{11}}(s_1,s_2)\cdot 1_{(u_{11},u_{21})\in \mathcal{U}_F} \cdot A + \left(1\!-\!\frac{1}{N\!-\!1}\right) \left[ \right.
-\\[0.5cm]
-  & \left. v_{u_{11}}(s_1,s_2) _{u_{21}}(s_2,s_2)  \frac{\left(k - 2\right) \left(k - 1\right)}{(N\!-\!2)(N\!-\!3)}
-  \left(\frac{1}{N\!-\!2}\cdot v_{u_{11}}(s_1,s_2)\cdot 1_{(u_{11},u_{21})\in \mathcal{U}_F} + (1 - \frac{1}{N - 2}) [B_1 + B_2 + B_3 + B_4]\right) + \right. \\[0.5cm] 
-  & \left.
-v_{u_{11}}(s_1,s_2) _{u_{21}}(s_2,s_1)  \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N\!-\!2)(N\!-\!3)} \left(\frac{1}{N\!-\!2}\cdot v_{u_{11}}(s_1,s_2)\cdot 1_{(u_{11},u_{21})\in \mathcal{U}_F} + (1 - \frac{1}{N - 2}) [C_1 + C_2 + C_3 + C_4]\right) + \right.
-\\[0.5cm]
-  & \left.
-v_{u_{11}}(s_1,s_1) _{u_{21}}(s_2,s_2)  \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N\!-\!2)(N\!-\!3)} \left(\frac{1}{N\!-\!2}\cdot v_{u_{11}}(s_1,s_2)\cdot 1_{(u_{11},u_{21})\in \mathcal{U}_F} + (1 - \frac{1}{N - 2}) [D_1 + D_2 + D_3 + D_4]\right) + \right.
-\\[0.5cm]
-  & \left.
-v_{u_{11}}(s_1,s_1) _{u_{21}}(s_2,s_1)  \frac{\left(N - k - 2\right) \left(N - k - 1\right)}{(N\!-\!2)(N\!-\!3)} \left(\frac{1}{N\!-\!2}\cdot v_{u_{11}}(s_1,s_2)\cdot 1_{(u_{11},u_{21})\in \mathcal{U}_F} + (1 - \frac{1}{N - 2}) [E_1 + E_2 + E_3 + E_4]\right) \right]
-  \end{array} $}
-\end{equation}
-
+   C_1  = \frac{\left(k - 3\right) \left(k - 1\right)}{(N - 3) (N - 4)} \, \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant) \, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) \quad
+   &C_2 = \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} \, \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident) \, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) \\[0.5cm]
+   C_3  = \frac{\left(k - 2\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} \, \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant) \, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident) \quad
+   &C_4 = \frac{\left(N - k - 2\right)^{2}}{(N - 3) (N - 4)} \, \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident) \, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident)
+   \end{array}
+   \end{equation*}
+The fourth line represents the case that during the first stage, the resident interacts with another resident whereas the mutant interacts with a mutant.
+The respective variables for the second-stage pairings are
+\begin{equation*}
+  \setlength{\arraycolsep}{10pt}
+  \begin{array}{llrl}
+   D_1  = \frac{\left(k - 2\right)^{2}}{(N - 3) (N - 4)} \, \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant)\, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) \quad
+   &D_2 = \frac{\left(k - 2\right) \left(N - k - 2\right)}{(N - 3) (N - 4)}\,  \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident)\, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) \\[0.5cm]
+   D_3  = \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} \, \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant) \,\tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident) \quad
+   &D_4 = \frac{\left(N - k - 3\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} \, \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident) \,\tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident)
+      \end{array}
+   \end{equation*}
+Finally, the last line of Eq.~\eqref{Eq:xTwoInteractions} corresponds to the case that during the first stage, both the resident and the mutant interact with a resident.
+The variables for the second stage are
+\begin{equation*}
+  \setlength{\arraycolsep}{10pt}
+    \begin{array}{llrl}
+   E_1  = \frac{(k - 2)(k - 1)}{(N-3)(N-4)}\, \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant)\, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) \quad
+   &E_2 = \frac{\left(k - 1\right) \left(N - k - 2\right)}{(N - 3) (N - 4)}\, \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident)\, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) \\[0.5cm]
+   E_3  = \frac{\left(k - 1\right) \left(N - k - 2\right)}{(N - 3) (N - 4)}\, \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant)\, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident) \quad
+   &E_4 = \frac{\left(N - k - 3\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} \,\tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident)\, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident)
+  \end{array}
+\end{equation*}
+
+%% RESULTS: Two interactions, last round --  Ratio of transition probabilities %%
+
+\noindent
+As a result, the transition probabilities for the number of mutants is now given by
+\begin{equation}\label{eq:ratio_TwoInteractions}
+\begin{array}{l}
+\displaystyle \lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} \sum_{u_{\resident 1},u_{\resident 2},u_{\mutant 1},u_{\mutant 2}} \hspace{-0.1cm} x(u_{\resident 1}u_{\resident 2}, u_{\mutant 1}u_{\mutant 2})\cdot \rho\left(\frac{u_{\resident 1} \!+\!
+u_{\resident 2}}{2}, \frac{u_{\mutant 1} \!+\!
+u_{\mutant 2}}{2}\right), \\[0.5cm]
+\displaystyle \lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} \sum_{u_{\resident 1},u_{\resident 2},u_{\mutant 1},u_{\mutant 2}} \hspace{-0.1cm} x(u_{\resident 1}u_{\resident 2},u_{\mutant 1}u_{\mutant 2})\cdot \rho\left(\frac{u_{\mutant 1} \!+\!
+u_{\mutant 2}}{2}, \frac{u_{\resident 1} \!+\!
+u_{\resident 2}}{2}\right).
+\end{array}
+\end{equation}
+Note that we assume players to use their average payoff across their two interactions for comparisons.\\
+
+%% RESULTS: Two interactions, last round --  Ratio of transition probabilities %%
+
+\noindent
+{\bf Stochastic stability of cooperation.}
+As before, we can use this formalism to compute when cooperation is stochastically stable.
+For a single \alld{} mutant in a \gtft{} resident population, we obtain the following non-zero probabilities \(x(u_{\resident1} u_{\resident 2}, u_{\mutant 1},u_{\mutant 2})\)
+\begin{equation*}
+  \setlength{\arraycolsep}{10pt}
+\begin{array}{ll}
+  x(RR, TT) =  \frac{N - 3}{N - 1}\left(\delta q - \delta + 1\right)^{2} & 
+  x(RR, TP) =  \frac{N - 3}{N - 1} \delta \left(1-q \right) \left(\delta q - \delta + 1\right)\\[0.2cm]
+  x(RR, PT) =  \frac{N - 3}{N - 1} \delta \left(1-q\right) \left(\delta q - \delta + 1\right) &
+  x(RR, PP) =  \frac{N - 3}{N - 1} \delta^{2}  \left(1 - q\right)^{2}\\[0.2cm]
+  x(RS, TT) =  \frac{1}{N-1} \left(\delta q - \delta + 1\right)^{2} &
+  x(RS, PT) =  \frac{1}{N-1} \delta \left(1-q\right) \left(\delta q - \delta + 1\right) \\[0.2cm]
+  x(RP, TP) =  \frac{1}{N-1} \delta \left(1-q\right) \left(\delta q - \delta + 1\right) &
+  x(RP, PP) =  \frac{1}{N-1} \delta^{2} \left(1 - q\right)^{2} \\[0.2cm]
+  x(SR, TT) =  \frac{1}{N-1} \left(\delta q - \delta + 1\right)^{2} &
+  x(SR, TP) =  \frac{1}{N-1} \delta \left(1-q\right) \left(\delta q - \delta + 1\right)\\[0.2cm]
+  x(PR, PT) =  \frac{1}{N-1} \delta \left(1-q\right) \left(\delta q - \delta + 1\right) &
+  x(PR, PP) =  \frac{1}{N-1} \delta^{2} \left(1-q\right)^{2}
+  \end{array}
+\end{equation*}
+The resulting ratio of transition probabilities is
 \begin{equation}
@@
   {%
-$
+$\frac{\lambda^{-}_1}{\lambda^{+}_1} =
+\frac{\frac{N - 3}{N-1} \left(\frac{\delta^{2} \left(1-q\right)^{2}}{1 + e^{- \beta \left(b - c\right)}} + \frac{2 \delta \left(1 - q\right) \left(\delta q - \delta + 1\right)}{1 + e^{- \beta \left(b/2 - c\right)}} + \frac{\left(\delta q - \delta + 1\right)^{2}}{1+e^{\beta c}}\right) + \frac{2}{N-1} \left(\frac{\delta^{2} \left(1-q\right)^{2}}{1 + e^{- \beta \left(b - c\right)/2}} + \frac{\delta \left(1 - q\right) \left(\delta q - \delta + 1\right)}{1+e^{\beta c/2} } + \frac{\delta \left(1 - q\right) \left(\delta q - \delta + 1\right)}{1+e^{\beta c} } + \frac{\left(\delta q - \delta + 1\right)^{2}}{1 + e^{ \beta \left( b/2 + c\right)}}\right)}
+{\frac{N - 3}{N-1} \left(\frac{\delta^{2} \left(1 - q\right)^{2}}{1 + e^{\beta \left( b - c\right)}} + \frac{2 \delta \left(1 - q\right) \left(\delta q - \delta + 1\right)}{1 + e^{ \beta \left(b/2 - c\right)}} + \frac{\left(\delta q - \delta + 1\right)^{2}}{1 + e^{- \beta c}}\right)
++ \frac{2}{N-1} \left(\frac{\delta^{2} \left(1 - q\right)^{2}}{1 + e^{ \beta \left(b-c\right)/2}} + \frac{\delta \left(1 - q\right) \left(\delta q - \delta + 1\right)}{1 + e^{- \beta c/2}} + \frac{\delta \left(1 - q\right) \left(\delta q - \delta + 1\right)}{1 + e^{- \beta c}} + \frac{\left(\delta q - \delta + 1\right)^{2}}{1 + e^{- \beta \left(b/2 + c\right)}}\right)}
+$}
+\end{equation}
+Again this expression simplifies considerably in the limit of strong selection, \(\beta \rightarrow \infty\), and of large
+populations, \(N \rightarrow \infty \). In that case, we obtain the following three cases
+\begin{equation} \label{Eq:RatioTransitionTwoInteraction}
+\frac{\lambda^{-}_1}{\lambda^{+}_1} = 
+\begin{cases}
+ \frac{1-(1-\delta(1-q))^2}{(1-\delta(1-q))^2} &\text{if}~ \frac{b}{2} > c \\[0.2cm]
+  \frac{\delta(1-q)}{1-\delta(1-q)}  &\text{if}~ \frac{b}{2} = c \\[0.2cm]
+  \frac{\delta^2(1-q)^2}{1-\delta^2(1-q)^2} &\text{if}~\frac{b}{2} < c
+\end{cases}
+\end{equation}
+We note that unlike the case of limited payoff memory, here the exact magnitude of the benefit and the cost of cooperation do have some effect on the transition probabilities. 
+By requiring that the above ratio exceeds one, we obtain the following conditions for stochastic stability of cooperation, 
+\begin{equation} \label{Eq:Condition_TwoGames}
+\begin{cases}
+  q <\frac{\delta - 1 + \sqrt{2}/2}{\delta}	&\text{if}~ \frac{b}{2} > c \\[0.15cm]
+  q<1-\frac{1}{2\delta}	&\text{if}~\frac{b}{2}=c\\[0.15cm]
+  q <\frac{\delta - \sqrt{2}/2}{\delta}\ &\text{if}~\frac{b}{2} < c
+\end{cases}
+\end{equation}
+Only when \(b/2\!=\!c\) we obtain the same expression as in the limited memory case,~Eq.~\eqref{Eq:TransitionRatioSimple}. 
+The requirement $q\!\ge\!0$ implies that the minimum continuation probability is $\delta\!>\!1\!-\!\sqrt{2}/2\approx 0.293$ in the first case, $\delta\!>\!\frac{1}{2}$ in the second, and $\delta>\sqrt{2}/2\approx 0.707$ in the third case.
+Conversely, if the game is infinitely repeated, $\delta\!\rightarrow\!1$, stochastic stability of cooperation requires $q<\sqrt{2}/2\approx 0.707$ in the first case, $q<1/2$ in the second, and $q<1\!-\!\sqrt{2}/2\approx 0.293$ in the third case.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% RESULTS: Remembering the last two rounds of one interaction %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Updates based on the last two rounds of one repeated game}
+\label{section:m_two_n_one}
+
+%% RESULTS: Two interactions, last round -- Explaining the setup %%
+
+{\bf General setup.} As our next setting, we consider a setup in which players update their strategies based on the last two rounds of their last game. 
+Here, additional complications may arise because some games may only last for one round if $\delta\!<\!1$.
+In the following, we abstract from these complications.  
+We shall assume that learners only update their strategies when both the learner's and the role model's last game had at least two rounds. 
+If either the learner or the role model only engaged in a one-round game in their previous interaction, we assume the learner decides to keep its strategy. 
+As a result, in the following, we accordingly re-interpret some of our quantities. 
+For example, $\lambda^+_k$ is no longer the (unconditional) probability that the number of mutants increases by one after a learner and a role model have been chosen.
+Instead, we interpret it as the (conditional) probability that the number of mutants increases, given that the learner and the role model happened to interact at least for two rounds during their respective last interaction.
+This new interpretation of transition probabilities affects the time-scale of fixation. 
+After all, we ignore all time steps in which no update occurred because a player's last game only lasted for a single round.
+However, our re-interpretation does not affect the ratio $\lambda^-_k/\lambda^+_k$, which is the key quantity we consider.\\
+
+%% RESULTS: Two interactions, last round -- Computing the ratio of transition probabilities %%
+
+\noindent
+{\bf Computing the distribution of outcomes in the penultimate and in the final round of a game.}
+To compute the ratio of transition probabilities, again we need to compute how likely it is that the resident and the mutant received any possible combination of two payoffs during the last two rounds.
+We start by computing the players' payoff distribution in the penultimate round.
+
+%% RESULTS: Limited memory: probability distribution for final payoff for two fixed strategies%%
+
+\begin{Prop}\label{proposition:penultimate_round} 
+Consider a repeated game with continuation probability $\delta$, between two players with reactive strategies $\strategy_1\!=\!(y_1, p_1, q_1$)  and $\strategy_2\!=\!(y_2,p_2,q_2)$. 
+Let $\hat{v}_u\!=\!\hat{v}_u(\strategy_1,\strategy_2)$ denote the probability that the first player receives the payoff $u\!\in\!\mathcal{U}$ in the penultimate round of the game, conditional on the event that the game lasts at least two rounds.
+Then this probability  $\hat{v}_u$ is again identical to the player's average probability $v_u$ to receive payoff $u$ across all rounds of the game, as specified by Eq.~\eqref{Eq:vAverage}.
+\end{Prop}
+
+\begin{proof}
+The proof is analogous to the proof of Proposition~\ref{proposition:last_round}. 
+Let $\mathbf{\hat{v}}:=(\hat{v}_R, \hat{v}_S, \hat{v}_T, \hat{v}_P)$, and let $\mathbf{\hat{v}}(\tau)$ denote the conditional distribution given that  the last round of the game happens to be round $\tau\!\in\!\{1,2,\ldots\}$ (note that we refer to the initial round as round zero, so the game is assumed to take at least two rounds).
+Similar to Eq.~\eqref{Eq:vTildeSimple}, we obtain
+\begin{equation}
+\mathbf{\hat v}(\tau) = \mathbf{v_0}\cdot M^{\tau-1}.
+\end{equation}
+The probability that $\tau\!\ge\!1$ is the last round, conditional on the event that the game takes at least two rounds, is 
+$$ \mathbb{P}(\tau ~\text{is final round}~|~\text{there are at least two rounds}\,) = \frac{\delta^\tau(1\!-\!\delta)}{\delta} =  \delta^{\tau-1}(1\!-\!\delta).$$
+By the law of total probability, we again obtain $\mathbf{\hat{v}}$ by summing up over all possible last rounds,
+\begin{equation}
+\hat{v} = \sum_{\tau=1}^\infty \delta^{\tau-1}(1\!-\!\delta)  \cdot \mathbf{v_0}M^{\tau-1} 
+=  (1\!-\!\delta)\mathbf{v_0} \sum_{t=0}^\infty \delta^t  M^t. 
+\end{equation}
+This expression for $\mathbf{\hat{v}}$ coincides with the expression for $\mathbf{v}$ in Eq.~\eqref{Eq:vDefinition}.
+\end{proof}
+
+\noindent
+Based on Proposition~\ref{proposition:penultimate_round}, we can derive a probability distribution for player 1's payoff in the last two rounds of a repeated game. 
+To this end, consider the $4\times4$ transition matrix $M(\strategy_1,\strategy_2)=(m_{u,u'})$ according to~Eq.~\eqref{eq:transition_matrix}. 
+Instead of the usual indexing of the four rows by numbers, $i\!\in\!\{1,2,3,4\}$, here we label the four rows of this matrix by player 1's payoffs in the previous round, $u\!\in\!\{R,S,T,P\}$. 
+Similarly, we label the four columns of this matrix by the resulting payoffs to player~1 in the next round. 
+For example, $m_{ST}$ corresponds to the second row and third column of matrix $M$. 
+So by Eq.~\eqref{eq:transition_matrix}, $m_{ST} = \bar{q}_1 p_2=(1\!-\!q_1)p_2$.
+Based on this notation, we can describe the probability $w_{uu'}(\strategy_1,\strategy_2)$ that player~1 obtains a payoff of $u$ in the penultimate round and a payoff of $u'$ in the last round. 
+Based on the vector $\mathbf{\hat{v}}(\strategy_1,\strategy_2):=(\hat{v}_R, \hat{v}_S, \hat{v}_T, \hat{v}_P)$ that describes the outcome of the penultimate round, we obtain
+\begin{equation}
+w_{uu'}(\strategy_1,\strategy_2) = \hat{v}_u \cdot m_{u,u'}.
+\end{equation}
+ ~\\
+
+ \noindent
+ {\bf Computing the ratio $\lambda^-_k/\lambda^+_k$.}
+ To compute the ratio of transition probabilities, let us again take a population perspective. 
+ We consider $k$ mutants with strategy $\strategy_\mutant=(y_\mutant,p_\mutant,q_\mutant)$ and $N\!-\!k$ residents with strategy $\strategy_\resident = (y_\resident,p_\resident,q_\mutant)$. 
+Without loss of generality, we assume that either the learner or the role model is a resident, and that the respective other player is a mutant. 
+Let $x(u_\resident u'_\resident, u_\mutant u'_\mutant)$ be the probability that the two players received payoffs $u$ and $u'$ in the last two rounds of their last repeated game.
+We can compute this probability based on the same logic as in the limited memory setting (Section~\ref{section:limited_memory}). 
+This yields
+\begin{equation*}\label{eq:Chi_TwoRounds} 
 \setlength{\arraycolsep}{1pt} 
-  \begin{array}{llrl}
-  A & = \left(1\!-\!\frac{1}{N\!-\!1}\right)  \left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} v_{u_{21}}(s_1,s_2) v_{u_{22}}(s_2,s_2) + 
-   \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} v_{u_{21}}(s_1,s_2) v_{u_{22}}(s_2,s_1) + \right. \\[0.5cm]
-   & \left. \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} v_{u_{21}}(s_1,s_1) v_{u_{22}}(s_2,s_2) + 
-   \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} v_{u_{21}}(s_1,s_1) v_{u_{22}}(s_2,s_1)\right] \\[0.5cm] 
-   B_1 & = \frac{\left(k - 3\right) \left(k - 2\right)}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_2) v_{u_{22}}(s_2,s_2) \quad
-   B_2 = \frac{\left(k - 2\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_1) v_{u_{22}}(s_2,s_2) \\[0.5cm] 
-   B_3 & = \frac{\left(k - 2\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_2) v_{u_{22}}(s_2,s_1) \quad
-   B_4 = \frac{\left(N - k - 2\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_1) v_{u_{22}}(s_2,s_1) \\[0.5cm]
-
-   C_1 & = \frac{\left(k - 3\right) \left(k - 1\right)}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_2) v_{u_{22}}(s_2,s_2) \quad
-   C_2 = \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_1) v_{u_{22}}(s_2,s_2) \\[0.5cm] 
-   C_3 & = \frac{\left(k - 2\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_2) v_{u_{22}}(s_2,s_1) \quad
-   C_4 = \frac{\left(N - k - 2\right)^{2}}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_1) v_{u_{22}}(s_2,s_1) \\[0.5cm]
-
-   D_1 & = \frac{\left(k - 2\right)^{2}}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_2) v_{u_{22}}(s_2,s_2) \quad
-   D_2 = \frac{\left(k - 2\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_1) v_{u_{22}}(s_2,s_2) \\[0.5cm] 
-   D_3 & = \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_2) v_{u_{22}}(s_2,s_1) \quad
-   D_4 = \frac{\left(N - k - 3\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_1) v_{u_{22}}(s_2,s_1) \\[0.5cm]
-
-   E_1 & = \frac{\left(k - 2\right) \left(k - 1\right)} v_{u_{21}}(s_1,s_2) v_{u_{22}}(s_2,s_2) \quad
-   E_2 = \frac{\left(k - 1\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_1) v_{u_{22}}(s_2,s_2) \\[0.5cm] 
-   E_3 & = \frac{\left(k - 1\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_2) v_{u_{22}}(s_2,s_1) \quad
-   E_4 = \frac{\left(N - k - 3\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} v_{u_{21}}(s_1,s_1) v_{u_{22}}(s_2,s_1) \\[0.5cm] 
-  \end{array}$}
-\end{equation}
-
-The first term on the right side corresponds to the case that the learner and
-the role model happened to be matched during the first stage,
-followed by them being paired with another member of the population on the
-second stage. The second terms corresponds to the case that the learner and the
-role model interact with a mutant with a probability ($\frac{\left(k - 2\right)
-\left(k - 1\right)}{(N\!-\!2)(N\!-\!3)}$). In the seconds stage, they can either
-interact with each other $\frac{1}{N - 2}$ or not $(1 - \frac{1}{N - 2})$. If
-they do not interact with each other, then each of the following can happen:
-both of them interact with a mutant with a probability
-$\frac{(k-3)(k-2)}{(N-4)(N-3)}$ and both interact with a resident with a
-probability $\frac{(N-k-1)(N-k-2)}{(N-3)(N-4)}$. The last two possible pairs
-are that either of them interacts with a resident whilst the other interacts
-with a mutant, and this happens with a probability
-$\frac{(N-k-2)(k-1)}{(N-4)(N-3)}$, and so on.
-
-The probability that the number of mutants increases, and decreases
-respectively, by one is now given by,
-
-\begin{align}\label{eq:ratio_limited}
-\lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R1},u_{M1}\in\mathcal{U}} \quad \sum_{u_{R2},u_{M2}\in\mathcal{U}} x^{2}((u_{R1}, u_{R2}),(u_{M1}, u_{M2}))\cdot \rho\left(\frac{u_{R1} + u_{R2}}{2}, \frac{u_{M1} + u_{M2}}{2}\right), \\
-\lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R1},u_{M1}\in\mathcal{U}} \quad \sum_{u_{R2},u_{M2}\in\mathcal{U}} x^{2}((u_{R1}, u_{R2}),(u_{M1}, u_{M2}))\cdot \rho\left(\frac{u_{M1} + u_{M2}}{2}, \frac{u_{R1} + u_{R2}}{2}\right).
-\end{align}
-
-We use $x^{2}((u_{R1}, u_{R2}),(u_{M1}, u_{M2}))$ to denote the probability that
-the randomly chosen resident obtained a payoff of $u_{R1}$ in the last round of
-their first respective game and $u_{R1}$ in the last round of their second game.
-Similarly, that the mutant obtained a payoff of $u_{M1}$ from their first
-interaction and $u_{M2}$ from the second. Note that we assume that the players
-compare their average payoff of their two interaction when considering adopting
-the others strategy. For example the mutant adopts the resident's strategy with
-a probability $\rho\left(\frac{u_{R1} + u_{R2}}{2}, \frac{u_{M1} +
-u_{M2}}{2}\right)$.
+\begin{array}{llll}
+x(u_\resident u'_\resident, u_\mutant u'_\mutant)	 &=
+&\frac{1}{N\!-\!1}\cdot  &w_{u_\resident u_\resident'}\!(\strategy_\resident,\strategy_\mutant)\cdot 1_{(u_\resident,u_\mutant)\in \mathcal{U}_F}\cdot 1_{(u'_\resident,u'_\mutant)\in \mathcal{U}_F}\\[0.5cm]
+&+	
+&\frac{N\!-\!2}{N\!-\!1} \cdot 
+&\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3}\cdot w_{u_\resident u_\resident'} \!(\strategy_\resident,\strategy_\mutant) \,w_{u_\mutant u_\mutant'} \!(\strategy_\mutant,\strategy_\mutant) + 
+ \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3}\cdot w_{u_\resident u_\resident'} \!(\strategy_\resident,\strategy_\mutant) \,w_{u_\mutant u_\mutant'}  \!(\strategy_\mutant,\strategy_\resident)\right.\\[0.5cm]
+&&&\left.
++\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} \cdot w_{u_\resident u_\resident'} \!(\strategy_\resident,\strategy_\resident) \,w_{u_\mutant u_\mutant'} \!(\strategy_\mutant,\strategy_\mutant) +
+ \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} \cdot w_{u_\resident u_\resident'} \!(\strategy_\resident,\strategy_\resident) \,w_{u_\mutant u_\mutant'}  \!(\strategy_\mutant,\strategy_\resident)\right].
+\end{array}
+\end{equation*}
+Overall, we obtain the following formula for the transition probabilities 
+\begin{equation}\label{eq:ratio_TwoRounds}
+\begin{array}{l}
+\displaystyle \lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} \sum_{u_{\resident},u_\resident',u_{\mutant},u_\mutant'} \hspace{-0.1cm} x(u_\resident u_\resident', u_\mutant u_\mutant')\cdot \rho\left(\frac{u_\resident \!+\!
+u_\resident'}{2}, \frac{u_\mutant \!+\!
+u_\mutant '}{2}\right), \\[0.5cm]
+\displaystyle \lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} \sum_{u_{\resident},u_\resident',u_{\mutant},u_\mutant'} \hspace{-0.1cm} x(u_\resident u_\resident', u_\mutant u_\mutant')\cdot \rho\left( \frac{u_\mutant \!+\!
+u_\mutant '}{2}, \frac{u_\resident \!+\!
+u_\resident'}{2}\right).
+\end{array}
+\end{equation}
 
  ~\\
-{\bf Invasion Analysis}
-
-In a similar fashion we can calculate the condition for which a population of
-GTFT players can not be invaded by an ALLD mutant. Using the new formulation we
-obtain the following probabilities \(x^{2}((u_{11}, u_{12}), (u_{21}, u_{22}))\) that the payoff of a randomly
-chosen GTFT player is \((u_{11}, u_{12})\) and that the payoff of the ALLD player is \((u_{21}, u_{22})\),
-
+ \noindent
+ {\bf Stochastic stability of cooperation.}
+We once again calculate how easily a single \alld{} mutant can invade into a
+resident population of \gtft{} players. When two residents interact,
+we obtain the following probability distribution for the outcome of the two last rounds, 
+\begin{equation*}
+\begin{array}{ll}
+    w_{RR}(\gtft,\gtft) =  1,\\
+    w_{uu'}(\gtft, \gtft) =   0~~ \text{ for all other } u,u'\in \mathcal{U}.
+\end{array}
+\end{equation*}
+For interactions between the mutant and a resident, we obtain
 \begin{align*}
-  x^{2}((r, r), (t, t)) = & \frac{N - 3}{N - 1}\left(\delta q - \delta + 1\right)^{2} & 
-  x^{2}((r, r), (t, p)) = & - \frac{N - 3}{N - 1} \delta \left(q - 1\right) \left(\delta q - \delta + 1\right)\\
-  x^{2}((r, r), (p, t)) = & - \frac{N - 3}{N - 1} \delta \left(q - 1\right) \left(\delta q - \delta + 1\right) &
-  x^{2}((r, r), (p, p)) = & \frac{N - 3}{N - 1} \delta^{2}  \left(q - 1\right)^{2}\\
-  x^{2}((r, s), (t, t)) = & \frac{1}{N-1} \left(\delta q - \delta + 1\right)^{2} &
-  x^{2}((r, s), (p, t)) = & - \frac{1}{N-1} \delta \left(q - 1\right) \left(\delta q - \delta + 1\right) \\
-  x^{2}((r, p), (t, p)) = & - \frac{1}{N-1} \delta \left(q - 1\right) \left(\delta q - \delta + 1\right) &
-  x^{2}((r, p), (p, p)) = & \frac{1}{N-1} \delta^{2} \left(q - 1\right)^{2} \\
-  x^{2}((s, r), (t, t)) = & \frac{1}{N-1} \left(\delta q - \delta + 1\right)^{2} &
-  x^{2}((s, r), (t, p)) = & - \frac{1}{N-1} \delta \left(q - 1\right) \left(\delta q - \delta + 1\right)\\
-  x^{2}((p, r), (p, t)) = & - \frac{1}{N-1} \delta \left(q - 1\right) \left(\delta q - \delta + 1\right) &
-  x^{2}((p, r), (p, p)) = & \frac{1}{N-1} \delta^{2} \left(q - 1\right)^{2} \\
-  x^{2}((u_{11}, u_{12}), (u_{21}&, u_{22})) =  0 \text{ for all other payoff pairs} & ((u_{11}, u_{12}), (u_{21}, u_{22})). &
+    w_{TT} (\alld,\gtft) & = q \left(\delta q - \delta + 1\right), & 
+    w_{TP}(\alld,\gtft) & =  \left(1 - q\right) \left(\delta q - \delta + 1\right), \\
+    w_{PT} (\alld,\gtft) & = \delta q \left(1 - q\right), & 
+    w_{PP} (\alld,\gtft) & = \delta \left(1-q\right)^{2},\\
+    w_{uu'}(\alld,\gtft) & =  0 \text{ for all other } u,u' \in \mathcal{U}.
 \end{align*}
-
-The ratio of transition probabilities is given by,
-
+As a consequence, we obtain the following probabilities  $x(u_\resident u'_\resident, u_\mutant u'_\mutant)$ that the
+payoffs of a randomly chosen \gtft resident are $u_\resident$ and $u_\resident'$, whereas the payoffs of the mutant are $u_\mutant$ and $u_\mutant'$. 
+\begin{equation*}
+  \setlength{\arraycolsep}{15pt}
+\begin{array}{ll}
+  x(RR, TT) =  \frac{N-2}{N - 1}  q \left(\delta q - \delta + 1\right) 
+  &x(RR, TP) =  \frac{N-2}{N - 1} \left(1-q\right) \left(\delta q - \delta + 1\right)\\[0.2cm]
+  x(RR, PT) =  \frac{N-2}{N - 1} \delta q  \left(1-q\right) 
+  &x(RR, PP) =  \frac{N-2}{N - 1} \delta  \left(1 - q\right)^{2}\\[0.2cm]
+  x(SS, TT) =  \frac{1}{N -1} q \left(\delta q - \delta + 1\right) 
+  &x(SP, TP) =  \frac{1}{N -1} \left(1-q\right) \left(\delta q - \delta + 1\right) \\[0.2cm]
+  x(PS,PT) =   \frac{1}{N -1}\delta q \left(1-q\right) 
+  &x(PP, PP) =  \frac{1}{N -1}\delta \left(1-q\right)^{2}\\[0.2cm]
+  \multicolumn{2}{c}{x(u_\resident u'_\resident, u_\mutant u'_\mutant) =  0~~ \text{ for all other payoff combinations.}}
+\end{array}
+\end{equation*}
+Based on these payoff probabilities, we compute the ratio
 \begin{equation}
@@
   {%
-$\frac{\lambda^{+}}{\lambda^{-}} =
-\frac{\frac{N - 3}{N-1} \left(\frac{\delta^{2} \left(1 - q\right)^{2}}{1 + e^{- \beta \left(- b + c\right)}} + \frac{2 \delta \left(1 - q\right) \left(\delta q - \delta + 1\right)}{1 + e^{- \beta \left(- \frac{b}{2} + c\right)}} + \frac{\left(\delta q - \delta + 1\right)^{2}}{1 + e^{- \beta c}}\right)
-+ \frac{2}{N-1} \left(\frac{\delta^{2} \left(1 - q\right)^{2}}{1 + e^{- \beta \left(- \frac{b}{2} + \frac{c}{2}\right)}} + \frac{\delta \left(1 - q\right) \left(\delta q - \delta + 1\right)}{1 + e^{- \frac{\beta c}{2}}} + \frac{\delta \left(1 - q\right) \left(\delta q - \delta + 1\right)}{1 + e^{- \beta c}} + \frac{\left(\delta q - \delta + 1\right)^{2}}{1 + e^{- \beta \left(\frac{b}{2} + c\right)}}\right)}
-{\frac{N - 3}{N-1} \left(\frac{\delta^{2} \left(q - 1\right)^{2}}{1 + e^{- \beta \left(b - c\right)}} + \frac{2 \delta \left(1 - q\right) \left(\delta q - \delta + 1\right)}{1 + e^{- \beta \left(\frac{b}{2} - c\right)}} + \frac{\left(\delta q - \delta + 1\right)^{2}}{e^{\beta c} + 1}\right) + \frac{2}{N-1} \left(\frac{\delta^{2} \left(q - 1\right)^{2}}{1 + e^{- \beta \left(\frac{b}{2} - \frac{c}{2}\right)}} + \frac{\delta \left(1 - q\right) \left(\delta q - \delta + 1\right)}{e^{\beta c} + 1} + \frac{\delta \left(1 - q\right) \left(\delta q - \delta + 1\right)}{e^{\frac{\beta c}{2}} + 1} + \frac{\left(\delta q - \delta + 1\right)^{2}}{1 + e^{- \beta \left(- \frac{b}{2} - c\right)}}\right)}
+  $\frac{\lambda^{-}_1}{\lambda^{+}_1} =
+  \frac{\frac{N - 2}{N-1} \left(\frac{\delta \left(1 - q\right)^{2}}{1 + e^{- \beta \left(b - c\right)}} + \frac{q \left(\delta q - \delta + 1\right)}{1+e^{\beta c} } + \frac{\left(1 - q\right) \left(2 \delta q - \delta + 1\right)}{1 + e^{- \beta \left(b/2- c\right)}}\right) + \frac{1}{N-1} \left(\frac{\delta \left(1 - q\right)^{2}}{2} + \frac{q \left(\delta q - \delta + 1\right)}{1 + e^{ \beta \left( b + c\right)}} + \frac{\left(1 - q\right) \left(2 \delta q - \delta + 1\right)}{1 + e^{\beta (b+c)/2}}\right)}
+  {\frac{N - 2}{N-1} \left(\frac{\delta \left(1 - q\right)^{2}}{1 + e^{ \beta \left(b - c\right)}} + \frac{q \left(\delta q - \delta + 1\right)}{1 + e^{- \beta c}} + \frac{\left(1 - q\right) \left(2 \delta q - \delta + 1\right)}{1 + e^{ \beta \left(b/2-c\right)}}\right) + \frac{1}{N-1}\left(\frac{\delta \left(1 - q\right)^{2}}{2} + \frac{q \left(\delta q - \delta + 1\right)}{1 + e^{- \beta \left(b + c\right)}} + \frac{\left(1 - q\right) \left(2 \delta q - \delta + 1\right)}{1 + e^{- \beta (b+c)/2}}\right)}.
   $}
 \end{equation}
-
 In the limit of strong selection \(\beta \rightarrow \infty\) and large
-populations \(N \rightarrow \infty \) we obtain the following cases,
-
+populations \(N \rightarrow \infty \), this ratio simplifies to
 \begin{equation}
-\frac{\lambda^{+}}{\lambda^{-}} = 
+\frac{\lambda^{-}_1}{\lambda^{+}_1} = 
 \begin{cases}
-  - \frac{q \left(\delta q - \delta + 1\right)}{\left(q - 1\right) \left(\delta q + 1\right)}  & \frac{b}{2} > c \\[0.5cm]
-  \frac{- \delta q + \delta - 1}{\delta \left(q - 1\right)}  & \frac{b}{2} = c \\[0.5cm]
-  - \frac{\delta q^{2} - 2 \delta q + \delta - 1}{\delta \left(q - 1\right)^{2}} & \frac{b}{2} < c
+  \frac{(1-q) (1+\delta q)}{q \left(\delta q - \delta + 1\right)}  &\text{if}~~\frac{b}{2} > c \\[0.2cm]
+  \frac{\left(1+\delta \right) \left(1-q\right)} {\delta q - \delta + q + 1} &\text{if}~~ \frac{b}{2} = c \\[0.2cm]
+  \frac{\delta (1-q)^{2}}{1-\delta(1-q)^2 } &\text{if}~~ \frac{b}{2} < c
 \end{cases}
 \end{equation}
-
-We note that the relationship between the cost and benefit have an effect on how
-generous a conditional cooperator must be to avoid invasion. In the case of
-\(\frac{b}{2}=c\) the result remains the same expression as in the case of
-\(m=n=1\). For the other two cases we show that for
-\(\frac{\lambda^{+}}{\lambda^{-}} < 1\),
-
-\begin{equation}
+For stochastic stability of cooperation we require this ratio to be larger than one. This yields the conditions
+\begin{equation} \label{Eq:Condition_TwoRounds}
 \begin{cases}
-  q < \left\{\frac{\delta - 1 - \frac{\sqrt{2}}{2}}{\delta}, \frac{\delta - 1 + \frac{\sqrt{2}}{2}}{\delta}\right\}  & \frac{b}{2} > c \\[0.5cm]
-  q < \left\{\frac{\delta - \frac{\sqrt{2}}{2}}{\delta}, \  \frac{\delta + \frac{\sqrt{2}}{2}}{\delta}\right\} & \frac{b}{2} < c
+  q<\frac{1}{1-\delta+\sqrt{1+\delta^2}}  &\text{if}~~\frac{b}{2} > c \\[0.1cm]
+  q<\frac{\delta} {1+\delta} &\text{if}~~ \frac{b}{2} = c \\[0.1cm]
+  q<1-\frac{1}{\sqrt{2\delta}} &\text{if}~~ \frac{b}{2} < c
 \end{cases}
 \end{equation}
-
-For \(\frac{b}{2}>c\) the ratio is smaller for \(q < \left\{\frac{\delta - 1 -
-\frac{\sqrt{2}}{2}}{\delta}, \frac{\delta - 1 +
-\frac{\sqrt{2}}{2}}{\delta}\right\}\), however, \(\frac{\delta - 1 -
-\frac{\sqrt{2}}{2}}{\delta}\) is not a feasible root since it's always smaller
-than 1, and thus \(q < \frac{\delta - 1 + \frac{\sqrt{2}}{2}}{\delta}\). For
-infinitely repeated games, \(\delta \rightarrow 1\), this condition becomes \(q
-< \frac{\sqrt{2}}{2}\). In the case of \(\frac{b}{2}<c\) there are two possible
-roots. For repeated games that are repeated for a large number of turn such as
-\(\delta \rightarrow 1\) the condition then becomes \(q < 1 -
-\frac{\sqrt{2}}{2}\).
-
-\subsection{Updating Payoffs based on the Last $m$ Rounds Payoffs of One Interaction}\label{section:m_two_n_one}
-
-The second generalised case of the limited memory payoffs we discuss is that of
-individuals updating based on their last $m$ rounds payoffs with one member. Let
-\(\mathcal{U}^{m}= \{\underbrace{rrr\dots r}_{m}, \underbrace{rrr\dots s}_{m},
-\dots, \underbrace{ppp\dots p}_{m}\}\) be the set of feasible payoffs of the
-last $m$ rounds. The probability that a reactive strategy receives the payoffs
-$u\!\in\! \mathcal{U}^{m}$ is given by
-Proposition~\ref{proposition:last_m_rounds}.
-
-\begin{Prop}\label{proposition:last_m_rounds} Consider a repeated game, with
-  continuation probability $\delta$, between players with reactive strategies
-  $s_1\!=\!(y_1, p_1, q_1$) and $s_2\!=\!(y_2,p_2,q_2)$ respectively. Let
-  $\mathcal{U}^{m}$ denote the set of feasible payoffs in the last
-  \(m\) rounds, and let \(\mathbf{u}^{m}\) be the corresponding payoff vector.
-  Then the probability that the $s_1$ player receives the payoff $u\!\in\!
-  \mathcal{U}^{m}$ in the very last two rounds of the game is given by,
-
-  \begin{equation}
-  \langle\mathbf{v}^{m}(s_1,s_2),\mathbf{u}^{m}\rangle, \text{ where } \mathbf{v}^{m} \in R^{4^{m}} \text{ is given by },
-  \end{equation}
-  \begin{equation}
-    \mathbf{v}^{m}(s_1,s_2) = (1 - \delta) w_{a_1, a_2} \delta^2 \left[\mathbf{v_0}(I_4 - \delta M)^{-1}\right]_{a_1, a_2}, \quad  w_{a_1, a_2} \in M \ \forall \ a_1, a_2 \in \{1, 2, 3, 4\}.
-  \end{equation}
-\end{Prop}
-
-\subsubsection{Last Round Payoff of Two Interactions}
-
-In the special case of $m=2$ the stationary distribution \(\mathbf{v}^{2}(s_1,
-s_2)\) is sixteen dimension instead of four dimensional where \(v_{1}(s_1, s_2)\) is
-the long term probability that \(s_1\) and \(s_2\) mutually cooperated in the
-last two rounds. Let the feasible payoffs a strategy can receive
-be $\mathcal{U}^{2} = \{rr, rs, rt, rp, sr, \dots, pp\}$.
-The role model and the learner interact with only one other member and so the
-probability \(x\) remains the same as in Section~\ref{section:limited_memory}.
-
-The probability that the number of mutants increases, and decreases respectively,
-by one is now given by,
-
-\begin{align}
-\lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} \cdot \sum_{u_{R},u_{M}\in\mathcal{U}^{2}} x(u_{R},u_{M})\cdot \rho\left(\frac{u_{R}}{2},\frac{u_{M}}{2}\right), \\
-\lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} \cdot \sum_{u_{R},u_{M}\in\mathcal{U}^{2}} x(u_{R},u_{M})\cdot \rho\left(\frac{u_{M}}{2},\frac{u_{R}}{2}\right).
-\end{align}
-
-We assume again that players consider their average payoff of the last two turns
-when deciding to imitate another strategy or not.
-
-~\\
-{\bf Invasion Analysis}
-
-We once again calculate how easily a single ALLD mutant can invade into a
-resident population of GTFT player. When two GTFT players interact in the game,
-their respective probabilities for each of the four outcomes in the last round
-simplify to,
-
-\begin{align*}
-    v_{rr}(GTFT,GTFT) = & \delta^2, \text{ and }\\
-    v_{i}(GTFT,GTFT) = &  0 \text{ for all other } i \in \mathcal{U}^{2}.
-\end{align*}
-
-On the other hand, if an ALLD player interacts with a GTFT player, the
-respective probabilities according to Eq.~\ref{Eq:LastRound} become
-
-\begin{align*}
-    v_{tt} (ALLD,GTFT) & = \delta^{2} q \left(\delta q - \delta + 1\right), & 
-    v_{tp}(ALLD,GTFT) & = - \delta^{2} \left(q - 1\right) \left(\delta q - \delta + 1\right), \\
-    v_{pt} (ALLD, GTFT ) & = \delta^{3} q \left(1 - q\right), & 
-    v_{pp} (ALLD, GTFT) & = \delta^{3} \left(q - 1\right)^{2} \text{ and }\\
-    v_{i}(ALLD,GTFT) & =  0 \text{ for all other } i \in \mathcal{U}^{2}.
-\end{align*}
-
-As a consequence, we obtain the following probabilities \(x(u_1, u_2)\) that the
-payoff of a randomly chosen GTFT player is \(u_1\) and that the payoff of the
-ALLD player is \(u_2\),
-
-\begin{align*}
-  x(rr, tt) = & \frac{N-2}{N - 1} \delta^{4} q \left(N - 2\right) \left(\delta q - \delta + 1\right) &
-  x(rr, tp) = & - \frac{N-2}{N - 1} \delta^{4} \left(N - 2\right) \left(q - 1\right) \left(\delta q - \delta + 1\right)\\
-  x(rr, pt) = & - \frac{N-2}{N - 1} \delta^{5} q \left(N - 2\right) \left(q - 1\right) &
-  x(rr, pp) = & \frac{N-2}{N - 1} \delta^{5} \left(N - 2\right) \left(q - 1\right)^{2}\\
-  x(ss, tt) = & \frac{1}{N -1}\delta^{2} q \left(\delta q - \delta + 1\right) & 
-  x(sp, tp) = & - \frac{1}{N -1}\delta^{2} \left(q - 1\right) \left(\delta q - \delta + 1\right) \\
-  x(ps, pt) = & - \frac{1}{N -1}\delta^{3} q \left(q - 1\right) &
-  x(pp, pp) = & \frac{1}{N -1}\delta^{3} \left(q - 1\right)^{2}\\
-  x(u_1, u_2) = &  0 \text{ for all other payoff pairs } (u_1, u_2). &
-\end{align*}
-
-We now calculate the ratio of transition probabilities as
-
-\begin{equation}
-  \resizebox{.9\textwidth}{!}
-  {%
-  $\frac{\lambda^{+}}{\lambda^{-}} =
-  \frac{\frac{\delta^{2} \left(N - 2\right)}{N-1} \left(\frac{\delta \left(1 - q\right)^{2}}{1 + e^{- \beta \left(- b + c\right)}} + \frac{q \left(\delta q - \delta + 1\right)}{1 + e^{- \beta c}} + \frac{\left(1 - q\right) \left(2 \delta q - \delta + 1\right)}{1 + e^{- \beta \left(- \frac{b}{2} + c\right)}}\right) + \frac{1}{N-1}\left(\frac{\delta \left(1 - q\right)^{2}}{2} + \frac{q \left(\delta q - \delta + 1\right)}{1 + e^{- \beta \left(b + c\right)}} + \frac{\left(1 - q\right) \left(2 \delta q - \delta + 1\right)}{1 + e^{- \beta \left(\frac{b}{2} + \frac{c}{2}\right)}}\right)}
-  {\frac{\delta^{2} \left(N - 2\right)}{N-1} \left(\frac{\delta \left(1 - q\right)^{2}}{1 + e^{- \beta \left(b - c\right)}} + \frac{q \left(\delta q - \delta + 1\right)}{e^{\beta c} + 1} + \frac{\left(1 - q\right) \left(2 \delta q - \delta + 1\right)}{1 + e^{- \beta \left(\frac{b}{2} - c\right)}}\right) + \frac{1}{N-1} \left(\frac{\delta \left(1 - q\right)^{2}}{2} + \frac{q \left(\delta q - \delta + 1\right)}{1 + e^{- \beta \left(- b - c\right)}} + \frac{\left(1 - q\right) \left(2 \delta q - \delta + 1\right)}{1 + e^{- \beta \left(- \frac{b}{2} - \frac{c}{2}\right)}}\right)}
-  $}
-\end{equation}
-
-In the limit of strong selection \(\beta \rightarrow \infty\) and large
-populations \(N \rightarrow \infty \) we obtain three expressions depending on
-the cost-benefit relationship. We note that for \(\frac{b}{2}=c\) the result
-remains the same expression as in the case of \(m=n=1\).
-
-\begin{equation}
-\frac{\lambda^{+}}{\lambda^{-}} = 
-\begin{cases}
-  - \frac{q \left(\delta q - \delta + 1\right)}{\left(q - 1\right) \left(\delta q + 1\right)}  & \frac{b}{2} > c \\[0.5cm]
-  - \frac{\delta q - \delta + q + 1}{\left(\delta + 1\right) \left(q - 1\right)}  & \frac{b}{2} = c \\[0.5cm]
-  - \frac{\delta q^{2} - 2 \delta q + \delta - 1}{\delta \left(q - 1\right)^{2}} & \frac{b}{2} < c
-\end{cases}
-\end{equation}
-
-For \(\frac{\lambda^{+}}{\lambda^{-}} < 1\):
-
-\begin{equation}
-\begin{cases}
-  q \in \left\{\frac{\delta - \sqrt{\delta^{2} + 1} - 1}{2 \delta}, \  \frac{\delta + \sqrt{\delta^{2} + 1} - 1}{2 \delta}\right\}  & \frac{b}{2} > c \\[0.5cm]
-  q \in \left\{1 - \frac{\sqrt{2}}{2 \sqrt{\delta}}, \  1 + \frac{\sqrt{2}}{2 \sqrt{\delta}}\right\} & \frac{b}{2} < c
-\end{cases}
-\end{equation}
-
-For \(\frac{b}{2}>c\) the ratio is smaller for
-\(q < \left\{\frac{\delta - \sqrt{\delta^{2} + 1} - 1}{2 \delta}, \frac{\delta + \sqrt{\delta^{2} + 1} - 1}{2 \delta}\right\}\)
-, however, the first root is not a feasible root since it's always smaller
-than 1, and thus \(q < \frac{\delta + \sqrt{\delta^{2} + 1} - 1}{2 \delta}\). For
-infinitely repeated games, \(\delta \rightarrow 1\), this condition becomes \(q
-< \frac{\sqrt{2}}{2}\). In the case of \(\frac{b}{2}<c\) there are two possible
-roots. For repeated games that are repeated for a large number of turn such as
-\(\delta \rightarrow 1\) the condition then becomes \(q < 1 -
-\frac{\sqrt{2}}{2}\).
-
-Note that as \(\delta \rightarrow 1\) the condition for which condition
-cooperators can avoid invasion is the same for the case of \(m=2\) and \(n=2\).
-
-\subsection{Updating Payoffs based on the last two rounds payoff of two interactions (\(m=2\) and \(n=2\))}\label{section:m_two_n_two}
-
-The final extension to the limited memory framework we consider is that of
-increasing the number of rounds and the number of interactions. For this case we
-need to consider a combination of the methods we presented in
-Section~\ref{section:m_one_n_two} and Section~\ref{section:m_two_n_one}. As an
-example consider the case of \(m=n=2\). The probability that the number of
-mutants increases, and decreases respectively, by one is now given by,
-
-\begin{align}
-  \lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R1},u_{M1}\in\mathcal{U}^{2}} \quad \sum_{u_{R2},u_{M2}\in\mathcal{U}^{2}} x^{2}((u_{R1}, u_{R2}),(u_{M1}, u_{M2}))\cdot \rho\left(\frac{u_{R1} + u_{R2}}{2}, \frac{u_{M1} + u_{M2}}{2}\right), \\
-  \lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R1},u_{M1}\in\mathcal{U}^{2}} \quad \sum_{u_{R2},u_{M2}\in\mathcal{U}^{2}} x^{2}((u_{R1}, u_{R2}),(u_{M1}, u_{M2}))\cdot \rho\left(\frac{u_{M1} + u_{M2}}{2}, \frac{u_{R1} + u_{R2}}{2}\right).
-\end{align}
-
+Interestingly, these conditions differ from the respective conditions in the previous section,~Eq.~\eqref{Eq:Condition_TwoGames}.
+Moreover, the requirement $q\ge 0$ imposes no constraints in the first two cases; it imposes the constraint $\delta\!>\!1/2$ in the last case. 
+However, in the limit of long games, $\delta\!\rightarrow \!1$ we recover the same conditions as in the previous section.
+Stochastic stability of cooperation requires $q\!<\!\sqrt{2}/2\!\approx\!
+0.707$ in the first case, $q\!<\!1/2$ in the second, and $q<1\!-\!\sqrt{2}/2\approx 0.293$ in the third case.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% RESULTS: Remembering the last two rounds of two interaction %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Updates based on the last two rounds of two repeated games}\label{section:m_two_n_two}
+
+As our final extension, we consider a combination of the previous two settings, discussed in Sections~\ref{section:m_one_n_two} and Section~\ref{section:m_two_n_one}. 
+Now individuals take into account two interactions with different co-players, and from each interaction, they take into account the last two rounds. 
+The probability that the number of mutants increases or decreases by one is now given by
+\begin{equation}\label{eq:ratio_TwoRounds}
+\begin{array}{l}
+\displaystyle \lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} 
+\hspace{-0.5cm}
+\sum_
+{\scriptsize \begin{array}{l}
+u_{\resident 1},u_{\resident 1}', u_{\resident 2},u_{\resident 2}'\\
+u_{\mutant 1},u_{\mutant 1}', u_{\mutant 2},u_{\mutant 2}'
+\end{array}} 
+\hspace{-1cm} x(u_{\resident 1} u_{\resident 1}'u_{\resident 2} u_{\resident 2}' ,u_{\mutant 1}u_{\mutant 1}'u_{\mutant 2}u_{\mutant 2}')\cdot 
+\rho\!\left(\frac{u_{\resident 1}\!+\! u_{\resident 1}'\!+\!u_{\resident 2}\!+\!u_{\resident 2}' }{4},
+\frac{u_{\mutant 1}\!+\!u_{\mutant 1}'\!+\!u_{\mutant 2}\!+\!u_{\mutant 2}'}{4}\!\right), \\[1.4cm]
+
+\displaystyle \lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} 
+\hspace{-0.5cm}
+\sum_
+{\scriptsize \begin{array}{l}
+u_{\resident 1},u_{\resident 1}', u_{\resident 2},u_{\resident 2}'\\
+u_{\mutant 1},u_{\mutant 1}', u_{\mutant 2},u_{\mutant 2}'
+\end{array}} 
+\hspace{-1cm} x(u_{\resident 1} u_{\resident 1}'u_{\resident 2} u_{\resident 2}' ,u_{\mutant 1}u_{\mutant 1}'u_{\mutant 2}u_{\mutant 2}')\cdot 
+\rho\!\left(\frac{u_{\mutant 1}\!+\!u_{\mutant 1}'\!+\!u_{\mutant 2}\!+\!u_{\mutant 2}'}{4},
+\frac{u_{\resident 1}\!+\! u_{\resident 1}'\!+\!u_{\resident 2}\!+\!u_{\resident 2}' }{4}\!\right).
+\end{array}
+\end{equation}
  Though we do not carry any further analytical exploration of this case, in the
-next section we present simulation results when the updating payoffs of the
-pairwise process depend on the limited memory framework with \(m=n=2\), as well
-as the rest of the cases we have discussed so far.
+next section we present simulation results for this setting, as well as for the other settings.\\
+
+\clearpage
 
@@
 
-\section{Further simulation results}~\label{section:furthersimulations}
-
-\subsection{Simulation Results on the Pairwise Comparison Process}\label{section:simulation_results}
-
-We simulate the evolutionary process described in
-Algorithm~\ref{algorithm:pairwise_comparison} for the different updating
-mechanisms we have described in
-Sections~\ref{section:perfect_memory}-\ref{section:m_two_n_two}. For each
-approach we performed an independent run of the process and for each time step we
-recorded the current resident population \((y, p, q)\). The results are shown in
-Figure~\ref{fig:expected_payoffs_results}. We observe that in most cases the
-resident population consists either of defectors or conditional cooperators. A
-conditional cooperator always cooperates if the co-player cooperated
-($p\approx1$) and cooperates with a probability \(q\) if the co-player defected.
-The most abundant conditional cooperators in each simulation differ as a
-result of the updating payoffs. More specifically, in order for a resident
-population of conditional cooperators to avoid being invaded they need to adopt
-a different value of \(q\). For each method we have discussed this under the
-invasion analysis subsection.
-
-In the cases of perfect memory the resident population adopts a \(q \leq 1 -
-\frac{c}{b}=0.9\). In the limited memory case the generosity of a conditional
-cooperator is independently of the benefit lower than \(\frac{1}{2}\). The rest
-of the cases also condition on the cost benefit relationship. In these
-simulations the cost of cooperation is set to 1 and the benefit to 10. As a
-result, in the case of two interactions \(q \leq \frac{\delta - 1 +
-\frac{\sqrt{2}}{2}}{\delta}=0.7068\), in the case of two rounds \(q \leq
-\frac{\delta + \sqrt{\delta^{2} + 1} - 1}{2 \delta}=0.7069\), and in the last
-case \(q \leq 1 - \frac{c}{b}=0.9\). The higher tolerance to defection results
-in a more cooperative population. As a result the expected payoffs allow for the
-most cooperative population. Between the limited memory approaches, we observe a
-big jump in the cooperation rate when we allow for more information
-(in the form of interactions or rounds). We hypothesise that as we
-allow for more information the results will tend to the case of perfect memory.
-
-\begin{figure}[!htbp]
+\section{Further simulation results}
+\label{section:furthersimulations}
+
+%% SIMULATIONS: Results in (p,q) space %%
+
+\subsection{Further simulation results for reactive strategies}\label{section:simulation_results}
+
+\begin{figure}[t!]
     \centering 
     \includegraphics[width=\textwidth]{static/more_memory_heatmaps_donation_game_with_illustrations.pdf}
-    \caption{\textbf{Evolutionary dynamics with difference updating approach.}
-    From left to right, we present result on the following updating payoffs
-    cases; the expected payoffs (perfect memory), the last round payoff
-    from one interaction (limited memory), the last round payoff from two
-    interactions, the last two rounds payoffs from one interaction, the
-    last two rounds payoffs from two interactions. We run each simulation for \(T
+    \caption{\textbf{Evolutionary dynamics of our five different settings.}
+    From left to right, we present result on the following cases: individuals update their strategies based on
+    ({\it i})~the expected payoffs (perfect memory); ({\it ii}) the last-round payoff
+    of one interaction (limited memory); ({\it iii})~the last round payoff of two
+    interactions; ({\it iv})~the payoffs of the last two rounds of one interaction; 
+    and ({\it v} the payoff of the last two rounds of two interactions. We run each simulation for \(T
     = 10^7\) time steps. For each time step we recorded the current resident
-    population \((y, p, q)\). Since \(\delta \rightarrow 1\) we do not report the
-    players' initial cooperation probability \(y\). The graphs show how often the
-    resident population chooses each combination \((p, q)\) of conditional
-    cooperation probabilities in the subsequent rounds. In both cases players
-    update based on their expected payoffs.}\label{fig:expected_payoffs_results}
+    population \((y, p, q)\). The graphs show how often the
+    resident population chooses each combination \((p, q)\); because $\delta\!\approx\!1$, we do not show the distribution of the player's first-round cooperation probability $y$.}\label{fig:expected_payoffs_results}
     \end{figure}
 
-\subsection{Expected and Last Round Updating Payoffs for Memory One Strategies}\label{section:memory_one}
-
-So far we have assumed that individuals can adopt reactive strategies.
-To demonstrate that our results hold for higher memory strategies here we
-present results for the expected payoffs, and the last round payoff when members
-use memory-one strategies. Memory-one strategies consider the outcome of the
-previous round to decide on an action. There are four possible outcome in each
-round; \((C, C), (C, D), (D, C), (D, D)\). A memory-one strategy \(s\) can be
-written as a five-dimensional vector \(s=(y, p_1, p_2, p_3, p_4)\). The
-parameter \(y\) is the probability that the strategy opens with a cooperation
-and \(p_1\), \(p_2\), \(p_3\), \(p_4\) are the probabilities that the strategy
-cooperates for each of the possible outcomes of the last round.
-
-We perform four separate simulations where we differ the updating payoff and the
-benefit of cooperation \(b\). The results for a low value of benefit are given
-in Figure~\ref{fig:memory_one_low_benefit}, and for a high benefit in
-Figure~\ref{fig:memory_one_high_benefit}. We verify that even when individuals
-are allowed to use memory-one strategies, the cooperation rate is higher
-in the perfect memory approach compared to the limited memory.
-
-\begin{figure}[!htbp]
+{\bf Strategies favored by selection.} 
+To gain some insights into the evolving strategies in each of the settings described in the previous section, we simulated the evolutionary process according to Algorithm~\ref{algorithm:pairwise_comparison}.
+Each time step, we recorded the current resident population \((y, p, q)\). 
+The results of these simulations are shown in Supplementary Figure~\ref{fig:expected_payoffs_results}. 
+Qualitatively, all settings lead to a similar dynamics as in the case of perfect and limited memory described in the main text. 
+In particular, populations are clustered around \alld, and around a strip of conditionally cooperative strategies, \gtft$=\!(1,1,q)$. 
+In all settings, we observe that along this strip, strategies are most abundant when they satisfy the condition for stochastic stability. 
+For perfect memory, this means the resident population predominantly adopts a strategy with \(q\! \leq\! 1 \!-\!
+c/b\!=\!0.9\) (for the parameters used in the simulation). 
+For limited memory, the generosity of a conditional cooperator satisfies  \(q\!<\!\frac{1}{2}\). 
+In the next two cases, when players either remember the last round of two interactions, or the last two rounds of one interaction, stochastic stability requires $q\!<\!\sqrt{2}{2}\!\approx\!0.701$. 
+In all cases, the simulation results are well aligned with these theoretical predictions.\\
+
+%% SIMULATIONS: Invasion analysis %%
+
+\begin{figure}[t!]
     \centering    
-  \includegraphics[width=\textwidth]{static/memory_one_results_low_benefit.pdf}
-  \caption{\textbf{Evolutionary dynamics results for memory-one strategies for low benefit.}
-  We perform two independent simulations. In one simulation individuals use
-  expected payoffs and in the other the last round one interacts when they
-  update their strategies. We run each simulation for \(T = 10^8\) time steps.
-  For each time step, we have recorded the current resident population, who is
-  now of the form  \((y, p_1, p_2, p_3, p_4)\) In the left panel we report the
-  cooperation rates for each simulation. It can be shown than even for
-  memory-one strategies expected payoffs result in a more cooperative
-  population. The right panel reports the most abundant strategy of each
-  simulation. Abundance is the number of mutants a strategy can repel before
-  being invaded. The most abundant strategies have some similarities,
-  namely, \(p_1 \approx 1\), \(p_3 \approx 0\) and \(p_4 > \frac{1}{2}\). There are also differences, in the latter
-  case a strategy is more likely to open with cooperation and their tolerance to
-  a \((C, D)\) outcome is almost zero. A difference between the strategies is their
-  abundance. In the expected payoffs case a strategy can repel a way greater
-  number of mutants. In the case of last round payoffs strategies become less
-  robust. Parameters: \(N =100, c=1, b=3, \beta=1\).}\label{fig:memory_one_low_benefit}
+     \includegraphics[width=0.5\textwidth]{static/invasion_analysis.eps}
+    \caption{\textbf{Invasion analysis for perfect and limited payoff memory.} We consider two residents, \alld{}$=(0,0,0)$ and \gtft$=(1,1,1/3)$ for both perfect and limited payoff memory.
+In each case we simulate the pairwise comparison process 1,000 times and recorded the first mutant strategy that reached fixation (colored dots).
+Black arrows indicate the arithmetic mean of the successful mutant strategies.
+In addition, we recorded how many mutant strategies it takes on average to successfully invade a given resident.
+We observe that under perfect memory, defectors are more prone to being invaded, whereas conditional cooperators are more resilient to invasion.
+Parameters are the same as in Fig.~1 of the main text, with $b\!=\!10$, $c\!=\!1$, $N\!=\!100$, $\beta\!=\!1$, $\delta\!=\!0.999$.}
+    \label{fig:invasion_analysis}
 \end{figure}
 
-\begin{figure}[!htbp]
-  \centering 
-  \includegraphics[width=\textwidth]{static/memory_one_results_high_benefit.pdf}
-  \caption{\textbf{Evolutionary dynamics results for memory-one strategies for high benefit.}
-  We perform two independent simulations. In the case of high benefit expected
-  payoffs again result in a more cooperative population. The right panel reports
-  the most abundant strategy of each simulation. Abundance is the number of
-  mutants a strategy can repel before being invaded. For the expected payoffs he
-  most abundant is that of win-stay lose-shift. However in the later case the
-  most abundant strategy is a strategy with no tolerance to one defection, and
-  it cooperates with a probability 0.5 after a mutual defection.  In the expected
-  payoffs case strategies are more robust.  Parameters: \(N =100, c=1, b=10,
-  \beta=1\).}\label{fig:memory_one_high_benefit}
-\end{figure}
-
-\subsection{Expected and Last Round Updating Payoffs for High Mutation ($\mu \neq 0$)}\label{section:mutation}
-
-In this section we evaluate the main result of this work for \(\mu \neq 0\).
-Namely, we explore the evolved population when individuals use perfect and
-limited updating payoff memory for different values of \(\mu\). We perform five
-independent runs of the pairwise process described in
-Section~\ref{section:model}, and at each time step we record the
-average player \(\bar{s}=(\bar{y}, \bar{p}, \bar{q})\). The average cooperation
-of the resident population for different values of mutation are shown in
-Figure~\ref{fig:mutation}. The cooperation rate in the case of perfect
-memory is always higher compared to the limited memory regardless of the
-mutation value. For mutation value of 1 the processes become random and
-this results to a cooperation rate of \(\frac{1}{2}\) in both simulations.
+\noindent
+{\bf Invasion dynamics.} To gain some further into the dynamics over time, and into the differences between perfect and limited payoff memory, we have performed an invasion analysis, see Supplementary Figure~\ref{fig:invasion_analysis}.
+Here, we consider two possible resident strategies. 
+Either populations use \alld$=(0,0,0)$, top row, or they start out with \gtft$=(1,1,1/3)$, bottom row. 
+In each case, we simulated  the process until a mutant reached fixation, for both perfect memory (left) and limited memory (right). 
+We iterated this simulation 1,000 times. 
+Colored dots indicate the position of successful mutants, and arrows indicate the average trajectory of evolution. 
+Moreover, we also recorded how long it takes on average until the first mutant strategy takes over. 
+This plot shows that \alld{} populations are typically invaded by mutants similar to Tit-for-Tat, with $p\!>\!0$ and $q\!\approx\!0$. 
+In contrast, \gtft{} populations are typically invaded by other \gtft{} mutants, but with higher generosity. 
+A further analysis of the fixation time can shed some light on why perfect memory leads to higher average cooperation rates: 
+Compared to limited memory, it takes fewer mutants to successfully invade into \alld{}, and more mutants to invade into \gtft{}.\\
+
+%% SIMULATIONS: Impact of mutations %%
 
@@
   two random players to serve as the role model and the learner. The learner
-  adopts the strategy of the role model with a probability \(\rho(\pi_{L}, \pi_{RM})\) where the
-  updating payoffs depend on the method. In the case of perfect memory
-  the expected payoffs are used and in the case of limited memory the
-  last round payoff against one opponent. We plot the average cooperation rate
-  within the resident population for each value of \(\mu\). For \(\mu=1\)
-  the process becomes random and so the cooperation rates are 0.5. For the rest
-  of the mutation values the perfect memory payoffs once again overestimate the
-  evolved cooperation, confirming the results of low mutation. Parameters: \(N
-  =100, c=1, b=10, \beta=1\).}\label{fig:mutation}
+  adopts the strategy of the role model with a probability \(\rho(\pi_\learner, \pi_\rolemodel)\) where the
+ payoffs depend on the considered setting. We plot the average cooperation rate
+  within the resident population for each value of \(\mu\). 
+  The results show that our previous results for rare mutations hold more generally:
+  While limited memory allows the evolution of moderate cooperation levels, perfect memory is always more conducive to the evolution of cooperation. 
+   Parameters are the same as before, \(N =100, c=1, b=10, \beta=1\).}\label{fig:mutation}
+\end{figure}
+
+\noindent
+{\bf Impact of mutations.} 
+All of our previous simulations are run under the assumption that mutations are rare,~$\mu\!\rightarrow\!0$. 
+This allows us to use the formula~\eqref{eq:fixation_probability} for a mutant's fixation probability. 
+While this limit allows for more efficient computation procedures, it rules out the possibility that two or three strategies might stably coexist in a population~\cite{tkadlec:pnas:2023}.
+To explore the effect of positive mutation rates, we performed five independent runs of the pairwise process described in
+Section~\ref{section:model} for different values of $\mu$.
+At each time step we record the average cooperation of the resident population. 
+The results for different mutation rates are shown in Figure~\ref{fig:mutation}. 
+The cooperation rate is higher when players have perfect memory (compared to limited memory), regardless of the
+mutation rate. 
+Only as the mutation rate approaches one, and the processes become fully random, the cooperation rate approaches $1/2$ in both simulations.
+
+\subsection{Simulations for memory-one strategies}\label{section:memory_one}
+
+So far we have assumed that individuals adopt reactive strategies.
+To demonstrate that our results hold for more complex strategy sets, 
+in the following we present results when players can choose among all memory-one strategies~\citep{sigmund2010calculus}. 
+Players with these strategies consider the full outcome of the
+previous round to decide on their next action. 
+There are four possible outcome in each
+round; \((C, C), (C, D), (D, C), (D, D)\). 
+A memory-one strategy \(\strategy\) can thus be
+written as a five-dimensional vector \(\strategy=(y, p_1, p_2, p_3, p_4)\). 
+As with reactive strategies, the entry \(y\) is the probability that the strategy opens with a cooperation.
+The other entries \(p_1\), \(p_2\), \(p_3\), \(p_4\) are the probabilities that the strategy
+cooperates in all subsequent rounds, depending on the outcome of the previous round.
+Every reactive strategy $(y,p,q)$ can be written as a memory-one strategy, by using the memory-one representation $(y,p,q,p,q)$.
+Conversely, there are memory-one strategies that cannot be represented by a reactive strategy, such as the strategy \texttt{WSLS}$=(1,1,0,0,1)$. 
+Therefore, the set of memory-1 strategies is a strict superset of the reactive strategies.
+
+To explore the dynamics among memory-1 strategies, we perform four separate simulations, for both perfect and limited memory, and for two different benefit values. 
+Figure~\ref{fig:memory_one_low_benefit} shows the results for a comparably low benefit of cooperation; 
+Figure~\ref{fig:memory_one_high_benefit} shows the corresponding results for a large benefit. 
+These simulations suggest that even when individuals
+are allowed to use memory-one strategies, there is more cooperation when players update their strategies based on perfect payoff memory.
+
+\begin{figure}[!htbp]
+  \centering 
+  \includegraphics[width=\textwidth]{static/memory_one_results_low_benefit.pdf}
+  \caption{\textbf{Evolution of memory-one strategies when the benefit of cooperation is comparably small.}
+  We perform two independent simulations. In one simulation individuals use
+  expected payoffs. 
+  In the other they update their strategy based on the payoff they received in the last round of their last interaction. 
+  We run each simulation for \(T = 10^8\) time steps.
+  For each time step, we have recorded the current resident population, who is
+  now of the form  \((y, p_1, p_2, p_3, p_4)\) In the left panel we report the
+  cooperation rates for each simulation. 
+  Also for memory-one strategies, the use of expected payoffs results in more cooperation. 
+  The right panel reports the most abundant strategy of each
+  simulation. Abundance is the number of mutants a strategy can repel before
+  being invaded. The most abundant strategies have some similarities,
+  namely, \(p_1 \approx 1\), \(p_3 \approx 0\) and \(p_4 > \frac{1}{2}\). 
+  Parameters: \(N =100, c=1, b=3, \beta=1\).}\label{fig:memory_one_low_benefit}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \centering 
+  \includegraphics[width=\textwidth]{static/memory_one_results_high_benefit.pdf}
+  \caption{\textbf{Evolution of memory-one strategies when the benefit is high.}
+  We consider a similar setup as in the previous figure, but now using a larger benefit value $b\!=\!10$.}\label{fig:memory_one_high_benefit}
 \end{figure}

### 2023-10-27 paper/main.tex
fix bibiography error
remove references to strategy inference

@@
 \usepackage{amsmath,amsthm,amssymb,setspace,enumitem,epsfig,titlesec,verbatim,color,array,eurosym,multirow}
-\usepackage[sort&compress,comma,round]{natbib}
+\usepackage[sort&compress,comma,round,numbers]{natbib}
 \usepackage[footnotesize,bf]{caption}
@@
 However, at the outset, it is not clear how easy it is to {\it learn} reciprocal strategies by social imitation. 
-As one obstacle, individuals usually only observe other population members' actions (i.e., whether they cooperate or defect). 
-In general, these observations do not suffice to infer the underlying strategies (i.e., the contingent rules that determine whether to cooperate or defect in any given situation). 
-This problem is equivalent to the difficulties that arise when researchers need to estimate the participants' strategies in behavioral experiments~\citep{Rossetti:ETH:2023}. 
-As another obstacle, even if others' strategies are perfectly observable, individuals might find it difficult to identify which ones have the highest payoff. 
+As one obstacle, even if others' strategies are perfectly observable, individuals might find it difficult to identify which ones have the highest payoff. 
 After all, the payoff of a strategy of direct reciprocity is not determined by the outcome of any single round.
-Rather it is determined by how well this strategy fares over an entire sequence of rounds, against many different population members. 
+Rather, it is determined by how well this strategy fares over an entire sequence of rounds, against many different population members. 
 In practice, such information might be both difficult to obtain and to process.
@@
 Thus, the players' payoffs in a single round are given by the matrix
-
-\begin{equation}
-    \begin{blockarray}{ccc}
-        & \text{cooperate} & \text{defect} \\
-        \begin{block}{c(cc)}
-            \text{cooperate} & b - c & -c \\
-            \text{defect} & b & 0 \\
-        \end{block}
-    \end{blockarray}.
-\end{equation}
+\begin{align}
+\bordermatrix{%
+	& C & D \cr
+C &\ b-c &\ -c\  \cr
+D &\ b &\ 0\ \cr
+} .
+\end{align}
 In particular, payoffs take the form of a prisoner's dilemma:
@@
 }
+}
 

### 2023-10-29 paper/main.tex
add text and references to discussion

@@
 \author{\parbox[c]{16cm}{\centering \onehalfspacing 
-Nikoleta E. Glynatsi$^1$,  Alex McAvoy$^{2,3}$, Christian Hilbe$^1$\\
+Nikoleta E. Glynatsi$^1$,  Alex McAvoy$^{2,3}$, Christian Hilbe$^1$\\ \quad \\
 $^{\rm 1}$Max Planck Research Group on the Dynamics of Social Behavior,\\ Max Planck Institute for Evolutionary Biology, Pl\"{o}n, Germany \\
@@
 For example, when strategies are updated through social learning, it is commonly assumed that individuals compare their respective average payoffs.  
-This would require them to compute (or remember) their payoffs against all other population members.
-Instead, herein, we introduce a theoretical framework to study the evolution of reciprocity when individuals learn based on their most recent experiences.
+This would require them to compute (or remember) their payoffs against everyone else in the population.
+To understand how more realistic constraints influence direct reciprocity, we consider the evolution of conditional behaviors when individuals learn based on more recent experiences.
 Even in the most extreme case that they only take into account their very last interaction, we find that cooperation can still evolve. 
 However, such individuals adopt less generous strategies, and they tend to cooperate less often than in the classical setup with expected payoffs. 
-Interestingly, once individuals remember the payoffs of two or three recent interactions, evolving cooperation rates quickly approach the classical limit. 
+Interestingly, once individuals remember the payoffs of two or three recent interactions, cooperation rates quickly approach the classical limit. 
 These findings contribute to a literature that explores which kind of cognitive capabilities are required for reciprocal cooperation. 
-While our results suggest that some rudimentary form of payoff memory is necessary, it already suffices to remember a few interactions.\end{abstract}
+While our results suggest that some rudimentary form of payoff memory is necessary, it already suffices to remember a few interactions.
+\end{abstract}
 
@@
 \begin{equation} \label{Eq:rho}
-    \rho(\pi_{L}, \pi_{R}) = \frac{1}{1\!+\! e^{- \!\beta (\pi_{R}- \pi_{L})}}.
+    \rho\left(\pi_{L}, \pi_{R}\right) = \frac{1}{1\!+\! e^{- \!\beta\left(\pi_{R}- \pi_{L}\right)}}.
 \end{equation}
@@
 In that model, the players' payoffs $\pi_L$ and $\pi_R$ are taken to be the payoffs that each player received in their very last round prior to making social comparisons. 
-That is, players only consider the very last repeated game they participated in, and there they only take into account the outcome of the very last round ( \FigBaseResults, lower left panel). 
+That is, players only consider the very last repeated game they participated in, and there they only take into account the outcome of the very last round (\FigBaseResults, lower left panel). 
 This assumption could reflect, for example, a strong recency bias in how individuals evaluate payoffs.  
@@
 \section{Discussion}\label{section:discussion}
+%
+In economics, if payoff is measured in terms of money and a decision is to be made at time $t$ in the future, then currency accumulated early on is weighted more in that decision because it has more time to accumulate interest.
+Such a model discounts the future relative to the past.
+It also does not necessarily require ``memory'' because the rewards are accumulated into a factor used in decision-making; the specific time stamps of rewards do not themselves provide better information beyond their effects on total payoff.
+In a similar fashion, the probabilistic interpretation of discounting as a continuation probability \citep{axelrod1981evolution} also effectively discounts the future relative to the past.
+A foraging animal deciding between two behaviors might tend to choose the one that yields a moderate reward sooner relative to a larger reward later, since earlier rewards (e.g., food) contribute to immediate survival, and there is no guarantee that later rewards will happen at all \citep{stephens1986foraging}.
 
-Direct reciprocity is one of the key mechanisms to explain why unrelated individuals might cooperate~\citep{nowak:Science:2006}. 
+Within the context of a single repeated game, the model we consider here is, in some ways, dual to the classical model of temporally-discounted rewards in repeated games.
+Instead of making decisions based on expected rewards in the future, we consider agents who make decisions based on actual rewards in the past.
+Thus, the ability to estimate the future payoff of a strategy is replaced by the memory of how this strategy fared against others in the past.
+This involves two time scales: interaction partners and rounds within those interactions.
+As a result, we are dealing with a model that discounts the past rather than the future.
+Intriguingly, treating payoffs in this manner is reminiscent of the reward-smoothing technique of ``eligibility traces'' in reinforcement learning \citep{sutton:MIT:2018}, which uses past rewards (discounted appropriately) to shape present perceived payoff.
+There is a sound basis for this method in neuroscience, where rewards and (temporal-difference) learning are associated to dopaminergic neurons \citep{schultz:JN:1998} and spike-timing-dependent plasticity \citep{dan:Neuron:2004}.
+This suggests a more biologically-encoded interpretation of memory, which is equally applicable to models of direct reciprocity where rewards have a neurological basis.
+
+Of course, the precise nature of ``memory'' also depends on what payoffs in a game represent, which should be taken into account when applying game-theoretic models.
+For example, a payoff stream of monetary currency might truly accumulate and not require memory on the parts of agents.
+Even in the context of money, however, not all of what was obtained in the past is necessarily available at the time a decision is made, which brings memory into play.
+The serial position effect in human psychology shows that in an ordered list of items (e.g., words), humans tend to have difficulty remembering the entirety of sequences, demonstrating moderate recall for those items coming earlier (primacy effect), substantial recall for those coming later (recency effect), and lower recall for those in between \citep{murdock:JEP:1962}.
+It is therefore reasonable that when presented with a stream of payoffs, whether on the timescale of pairing for repeated interactions or in a stream of one-shot games, players might be able to effectively incorporate only the most recent payoffs.
+
+In fact, even beyond specific psychological considerations, a curious interpretation of payoffs arises from the formula commonly used for expected payoffs in repeated games.
+If $\delta\in\left[0,1\right)$ is the probability of continuing to another round in the game, then the expected payoff to an agent is $\left(1-\delta\right)\sum_{t=0}^{\infty}\delta^{t}\pi_{t}$, where $\pi_{t}$ is the payoff the agent receives in the stage game at time $t$.
+Here, additional stochasticity arises due to uncertainty in the game length, and an agent might not be able to compute his or her expected payoff for use in decision-making.
+The probability that the game terminates after the interaction at time $T$ is $\left(1-\delta\right)\delta^{T}$, in which case $\left(1-\delta\right)\sum_{t=0}^{\infty}\delta^{t}\pi_{t}$ is exactly the expected payoff the agent receives at time $T$, i.e.
+\emph{in the last round}.
+As an unbiased estimator of this expectation, the agent might thus use $\pi_{T}$ as a proxy for ``success'' when evaluating his or her behavior.
+This gives a purely model-driven justification for why considering payoff in the last round of the game can result in more realistic extensions of traditional models.
+
+We note that the expected \emph{total} payoff in the game, $\pi_{0}+\pi_{1}+\cdots +\pi_{T}$, is given by $\sum_{t=0}^{\infty}\delta^{t}\pi_{t}$.
+This version of ``expected payoff'' appears less common in the literature on direct reciprocity than its normalization, $\left(1-\delta\right)\sum_{t=0}^{\infty}\delta^{t}\pi_{t}$, likely owing to the fact that payoffs can grow arbitrarily large with sufficiently long time horizons ($\delta\rightarrow 1^{-}$).
+Non-normalized payoffs interfere with selection intensity ($\beta$) in models of social imitation, which is (presumably) why they appear less frequently in the literature.
+On that point, we note that many of the differences between realized and expected payoffs disappear in the limit of weak selection \citep{mcavoy:NHB:2020}.
+Non-weak selection can introduce substantial differences between models with realized and expected payoffs \citep{mcavoy:PLOSCB:2021}, which is especially important to understand in models of social systems with cultural transmission \citep{cavalli1981cultural}.
+
+Our main contribution is an application of these ideas to direct reciprocity, which is one of the key mechanisms to explain why unrelated individuals might cooperate~\citep{nowak:Science:2006}. 
 According to this mechanism, cooperation pays if it makes the interaction partner more cooperative in future. 
@@
 
-Herein, we explore to which extent perfect payoff memory is in fact required for the evolution of reciprocity. 
+It is natural to ask, then, to what extent perfect payoff memory is in fact required for the evolution of reciprocity. 
 To this end, we consider a model in which individuals only remember the payoff of their very last interaction, or the payoffs of the last few interactions. 
@@
 Our findings also contribute to a wider literature that explores which kinds of cognitive capacities are required for reciprocal altruism to be feasible~\citep[e.g.,][]{Stevens:fip:2011,Volstorf:PlosOne:2011}.
-While more payoff memory is always favorable, reciprocal cooperation can already be sustained if individuals have a record of two or three past outcomes.\\
+While more payoff memory is always favorable, reciprocal cooperation can already be sustained if individuals have a record of two or three past outcomes.
+We believe that this kind of result, derived entirely within a theoretical model, is crucial for making model-informed deductions about reciprocity in natural systems.
 

### 2023-10-29 paper/si.tex
christian should be senior author!

--- 2023-10-19 paper/si.tex
@@
 
-\def\alld{\texttt{ALLD}}
-\def\tft{\texttt{TFT}}
-\def\gtft{\texttt{GTFT}}
-\def\esm{electronic supplementary material}
-
 \definecolor{expectedcolor}{rgb}{0.1791464821222607, 0.49287197231833907, 0.7354248366013072}
@@
 
-\title{{ \sffamily \LARGE Electronic supplementary material}\\{\bf \sffamily \LARGE Evolution of reciprocity with limited payoff memory}}
+\title{\bf  \sffamily \LARGE Supplementary Information: Evolution of cooperation among individuals with
+limited updating payoff memory\\}
 \date{}
@@
 
-%% BRIEF INTRO %%
-
-\noindent
-This document provides further details on our model and our analytical methods, and it contains additional simulation results. 
-Section~\ref{section:model} gives a brief summary of our model. 
-In particular, we provide further details on our implementation of the evolutionary dynamics, and our use of the rare-mutations limit. 
-In Section~\ref{section:analyticalresults} we derive analytical results for the various settings we consider.
-These settings differ in what kind of payoff information individuals take into account when updating their strategies. 
-In the perfect memory setting, individuals take into account all their interactions against all co-players.
-In the limited memory setting, they only consider the very last round of their very last interaction.  
-In addition, we describe several model extensions in which the amount of information taken into account is in between these two extremes. 
-Finally, Section~\ref{section:furthersimulations} presents further simulation results. 
-In particular, we confirm that our main results continue to hold when ({\it i}) players use memory-1 strategies instead of reactive strategies, and  ({\it ii}) mutations are no longer rare.
-
-%%%%%%%%%
-%% MODEL  %%
-%%%%%%%%%
-
-\section{Description of the model}\label{section:model}
-
-%% MODEL: Summary of the game dynamics %%
-
-\noindent
-{\bf Summary of the model.} As described in the main text, we study cooperative behavior in a population of size $N$.
-The dynamics unfolds on two time-scales. 
-The short time scale describes the game dynamics. 
-Here individuals are randomly matched to interact in a repeated prisoner's dilemma.  
-Each round individuals can choose whether to cooperate ($C$) or defect ($D$). 
-In general, the resulting one-shot payoffs can be summarized by the payoff matrix 
-\begin{equation}\label{eq:donation}
-    \begin{blockarray}{ccc}
-        & \text{cooperate} & \text{defect} \\
-        \begin{block}{c(cc)}
-            \text{cooperate} & R & S \\
-            \text{defect} & T & P \\
-        \end{block}
-    \end{blockarray}.
-\end{equation}
-Here, $R$ is interpreted as the reward for mutual cooperation, $S$ is the sucker's payoff, $T$ is the temptation, and $P$ is the punishment payoff~\citep{axelrod1981evolution}. 
-Throughout this work, we parametrize these payoffs as $R\!=\!b\!-\!c$, $S\!=\!-c$, $T\!=\!b$, and $P\!=\!0$, where $b$ and $c$ are the benefit and cost of cooperation, respectively, with $b\!>\!c\!>0$.
-After each round, players learn their co-player's previous action, and the game continues for another round with probability~$\delta$.
-Players make their decisions whether to cooperate in any given round based on their reactive strategies $\mathbf{s}\!=\!(y,p,q)$. 
-The entry $y$ determines a player's first-round cooperation probability. 
-The other entries $p$ and $q$ determine the player's cooperation probability in all subsequent rounds. 
-The probability is $p$ if the co-player cooperated in the previous round, and it is $q$ if the co-player defected. 
-On the short time scale, the players' strategies are fixed.
-
-%% MODEL: Summary of the evolutionary dynamics %%
-
-The long time scale describes the evolutionary dynamics. 
-Here, players are allowed to update their strategies based on the payoffs they yield. 
-We model these strategy updates with a pairwise comparison process~\citep{traulsen2007pairwise}.
-This process assumes that at regular time intervals, one player is randomly selected from the population.
-We refer to this player as the `learner' ($L$). 
-The learner is then given an opportunity to update its strategy. 
-There are two possibilities for how this update may occur. 
-With probability $\mu$, the player's strategy mutates randomly. 
-In that case, the player's new strategy is drawn uniformly from the space of all reactive strategies~$[0,1]^3$.  
-With probability $1\!-\!\mu$, the player engages in a pairwise comparison. 
-In that case, the player randomly picks another individual from the population (referred to as the `role model', $R$). 
-The learner adopts the role model's strategy with a probability \(\rho\) given by
+Section~\ref{section:pairwise_comparison} gives a brief overview of the pairwise
+comparison process. The pairwise comparison process consists of three phases;
+(1) the mutation phase (2) the game phase and (3) the update phase. In the
+update phase an individual adopts the strategy of another individual based on
+their ``updating payoffs''. Here we study different methods for calculating
+updating payoffs and we compare their results using as an example the prisoner's
+dilemma. In Section~\ref{section:perfect_memory} we describe the conventional
+method used in the literature and in Section~\ref{section:limited_memory} we
+present a new method which we refer to as 
+limited payoff memory. In the limited payoff memory method we assume that
+individuals update based on their last round payoff against one other member of
+the population. The framework can easily be extended to consider more rounds,
+more interactions or both. In
+Sections~\ref{section:m_one_n_two}-\ref{section:m_two_n_two} we discuss each of
+these extensions and for each extension we present a special case.
+In Section~\ref{section:simulation_results} we present numerical results of the
+pairwise comparison process using the different updating payoff methods.
+In Sections~\ref{section:memory_one} and~\ref{section:mutation} we
+verify the main result of this work when we no longer assume (i) that
+individuals use reactive strategies but instead memory one strategies (ii) low
+mutation.
+
+\section{Pairwise comparison process}\label{section:pairwise_comparison}
+
+Pairwise comparison process is a stochastic process for modelling the evolution
+of a finite population. The process starts with assigning all individuals of the
+population the same strategy. A strategy is a set of rules of how an individual
+should behave in an interaction with another individual. Each elementary time
+step of the process consists of three phases; (1) the \textbf{mutation phase}
+(2) the \textbf{game phase} and (3) the \textbf{update phase}. These are summarised
+in Figure~\ref{fig:pairwise_phases}.
+
+In the \textbf{mutation phase} one individual is
+chosen to switch to a new mutant strategy with a probability \(\mu\).
+In the \textbf{game phase} individuals are randomly matched with other
+individuals in the population and they engage in a repeated game where each
+subsequent turn occurs with a fixed probability \(\delta\). At each turn the
+individuals decide on an action based on their strategies.
+In repeated games there are infinitely many strategies, however, it is commonly
+assumed that individuals can only choose strategies from a restricted set. One
+such set is that of reactive strategies. A reactive
+strategy considers only the previous action of the other player, and thus, a
+reactive strategy \(s\) can be written as a three-dimensional vector \(s=(y, p,
+q)\). The parameter \(y\) is the probability that the strategy opens with a
+cooperation and \(p\), \(q\) are the probabilities that the strategy cooperates
+given that the opponent cooperated and defected equivalently.
+
+In the \textbf{update stage} two individuals are randomly selected. From
+the two individuals, one  serves as the `learner' and the other as the `role
+model'. The learner adopts the role model's strategy with a probability \(\rho\)
+given by,
+
 \begin{equation} \label{Eq:rho}
-    \rho(\pi_{L}, \pi_{R}) = \frac{1}{1\!+\! e^{\!-\!\beta (\pi_{R} - \pi_{L})}}.
-\end{equation}
-The parameter \(\beta\) is the selection strength
-It determines how important payoff differences are for the learner's decision to imitate the role model. 
-The parameters $\pi_R$ and $\pi_L$ refer to the relevant payoffs of the role model and the learner, respectively. 
-The exact value of these payoffs depend on the players' memory. 
-We say players have {\it perfect memory} when $\pi_R$ and $\pi_L$ are given by the players' expected payoffs (across all rounds and across all possible co-players). 
-We say players have {\it limited memory} when $\pi_R$ and $\pi_L$ are given by the players' realized payoff in the very last round of their very last interaction partner. 
-In addition, we consider several model extensions in which individuals have memory capacities in between these two extreme cases. 
-We provide a detailed description of these different cases and the resulting payoffs in Section~\ref{section:analyticalresults}.\\
-
-%% MODEL: Intro to the rare-mutations limit %%
-
-\noindent
-{\bf Evolutionary simulations for the rare-mutations limit.} 
-To simulate the evolutionary dynamics of the pairwise comparison process described above, it is sometimes useful to assume that mutations are rare, $\mu\!\rightarrow\!0$. 
-In that case, whenever a mutant strategy appears, this mutant either fixes in the population or goes extinct before the next mutant appears. 
-As a result, at any given time there are at most two different strategies present in the population~\citep{fudenberg:JET:2006,wu:JMB:2012,mcavoy:jet:2015}. 
-This assumptions makes computations more efficient, and it makes some of the results easier to interpret.
-In the following, we describe our implementation of the process in the rare-mutations limit in more detail.
-
-%% MODEL: Implementation of the rare-mutations limit %%
-
-Initially, the process starts with a population where all members use the same strategy (called `resident strategy'). 
-Then one individual adopts a mutant strategy selected uniformly at random from the set of feasible strategies.
-The fixation probability \(\phi_{M}\) of the mutant strategy can be calculated explicitly~\citep{nowak:Nature:2004},
+    \rho(\pi_{L}, \pi_{RM}) = \frac{1}{1\!+\! e^{\!-\!\beta (\pi_{RM}\!-\! \pi_{L})}}.
+\end{equation}
+
+\(\pi_{L}\) and \(\pi_{RM}\) are the updating payoffs of the learner and the
+role model respectively. The updating payoffs are a measure of how successful
+individuals are in the current standing of the population. The parameter
+\(\beta\) is known as the selection strength, namely, it shows how important the
+payoff difference is when the learner is considering adopting the strategy of
+the role model.
+
+\begin{figure}[!htbp]
+    \centering
+    \includestandalone[width=\textwidth]{static/pairwise_comparison_process_diagram}
+    \caption{\textbf{Pairwise comparison process phases.} \textbf{A) Initialisation.} The process begins
+    with a finite population where each member is assigned a given strategy.
+    Each color represents a different strategy, and the members are labelled by
+    letters. \textbf{B) Mutation phase.} An individual is selected (in the
+    example individual A) and with a given probability \(\mu\) that individual
+    adopts a new strategy. \textbf{C) Game phase.} Individuals are
+    selected to interact in a repeated social dilemma with other individuals. We
+    demonstrate the case where individuals B and F have been selected to
+    interact. They use the reactive strategies \(s_{B} = (y_B, p_B, q_B)\) and
+    \(s_{F} = (y_F, p_F, q_F)\) respectively. The opening moves depend on their
+    \(y_i\) probability. In turn 1, individual B cooperated,
+    thus, F cooperates with a probability \(p_{F}\) in turn 2. On the opposite,
+    individual F defected in turn 1, and so B cooperates in the next turn with
+    a probability \(q_{B}\). At each turn there is a probability \(\delta\) that a
+    subsequent turn will occur, and with a probability (\(1- \delta\)) the interaction ends.
+    \textbf{D) Update phase.} At the updating phase two individuals are chosen;
+    one serves as the role of the learner and the other one as the 
+    role model. In our example C adopts D's strategy with a probability
+    \(\rho(\pi_{C}, \pi_{D})\) where \(\pi_{C}, \pi_{D}\) denote the updating
+    payoffs of the individuals.}\label{fig:pairwise_phases}
+\end{figure}
+
+This elementary step of the process (mutation, game and update phases) is
+repeated for a large number of time steps, and at each time step we record the
+state of the population.
+
+\subsection{Low mutation \(\mu \rightarrow 0\)}
+
+In the case of low mutation (\(\mu \rightarrow 0\)) we assume that mutations are
+rare. In fact, so rare that only two different strategies can be present in the
+population at any given time. The case of low mutation is vastly adopted because
+it allows us to explicitly calculate the fixation probability of a newly
+introduced mutant.
+
+More specifically, the process again starts with a population where all members are of the same
+strategy. At each step one individual adopts a mutant strategy randomly selected
+from the set of feasible strategies. The fixation probability \(\phi_{M}\) of the
+mutant strategy can be calculated explicitly,
+
 \begin{equation}\label{eq:appendix_fixation_probability}
-    \varphi_{M} = \frac{1}{1+\sum\limits_{i=1}^{N-1}\prod\limits_k^i \frac{\lambda^-_k}{\lambda^+_k}}.
-\end{equation}
-Here, the index $k$ corresponds to the current number of players with the mutant strategy (we refer to these players as `mutants'). 
-The variables \(\lambda^-_k, \lambda^+_k\) are the probabilities that the number of mutants decreases or increases within a single updating step. 
-These probabilities depend on the probability that a mutant and a resident are chosen as the learner and the resident, respectively; 
-in addition, they depend on the respective switching probability~$\rho$, as described by Eq.~\eqref{Eq:rho}. 
-We specify the exact values of  \(\lambda^-_k, \lambda^+_k\) for each memory-setting in the following section.
-
-Depending on the fixation probability \(\varphi_{M}\), the mutant strategy either fixes (becomes the new resident) or goes extinct. 
-Afterwards, another random mutant strategy is introduced into the population. 
-We iterate this elementary population updating process for a large number of mutant strategies. 
-At each step, we record the current resident strategy, and the resulting average cooperation rate. 
-We consider this limit of rare-mutations throughout the main text. 
-The respective process is summarised by Algorithm~\ref{algorithm:pairwise_comparison}.
-However, in Section~\ref{section:furthersimulations} we present additional simulation results to show that our main results continue to hold when the mutation rate is strictly bounded away from zero.
-
-%% MODEL: Algorithm
-
-\begin{algorithm}[t]
+    \varphi_{M} = \frac{1}{1+\sum\limits_{i=1}^{N-1}\prod\limits_k^i \frac{\lambda^-_k}{\lambda^+_k}},
+\end{equation}
+
+where \(\lambda^-_k, \lambda^+_k\) are the probabilities that the number of
+mutants decreases and increases respectively, \(N\) is the size of the
+population, and \(k\) is the number of mutants. The probabilities \(\lambda^-_k
+\text{ and } \lambda^+_k\) depend on the updating payoffs of the mutant and the
+resident strategies. Depending on the fixation probability \(\phi_{M}\) the
+mutant either fixes (becomes the new resident) or goes extinct. Regardless, in
+the elementary time step another mutant strategy is introduced to the
+population. We iterate this elementary population updating process for a large
+number of mutant strategies and we record the resident strategies at each time
+step. The process is summarised by Algorithm~\ref{algorithm:pairwise_comparison}.
+
+\begin{algorithm}[!htbp]
   \SetAlgoLined
   $N \leftarrow$ population size\;
-  %$k \leftarrow 1$\; 
-  resident $\leftarrow$ starting resident\;
+  $k \leftarrow 1$\; resident
+   $\leftarrow$ starting resident\;
    \While{ $t <$ maximum number of steps}
@@
    {resident $\leftarrow$ mutant;}}
-   \caption{Evolutionary process in the limit of rare mutations}\label{algorithm:pairwise_comparison}
+   \caption{Evolutionary process}\label{algorithm:pairwise_comparison}
 \end{algorithm}
 
-\newpage
-
-%%%%%%%%%%%%%%%%%
-%% ANALYTICAL RESULTS %%
-%%%%%%%%%%%%%%%%%
-
-\section{Analytical results}
-\label{section:analyticalresults}
-
-\subsection{Updating Payoffs based on Expected Payoffs (perfect payoff memory)}\label{section:perfect_memory}
-
-Let \(\mathbf{u} = (b-c, -c, b, 0)\)
+Most of the results we present in this work consider the case of low mutation,
+however, we have also verified that the main result holds in the case of
+high mutation rates (Section~\ref{section:mutation}).
+
+\subsection{Updating Payoffs}\label{section:updating_payoffs}
+
+The updating payoffs depend on the interactions of the individuals at the game
+phase. In this work we assume that in the game phase individuals are matched in
+pairs and that they participate in a repeated 2 person donation game. In the
+donation game there are two actions: cooperation (\(C\)) and defection (\(D\)).
+By cooperating a player provides a benefit \(b\) to the other player at their
+cost \(c\), with \(0 < c < b\). Thus the payoffs for a player in each turn are,
+
+\begin{equation}\label{eq:donation}
+    \begin{blockarray}{ccc}
+        & \text{cooperate} & \text{defect} \\
+        \begin{block}{c(cc)}
+            \text{cooperate} & b - c & -c \\
+            \text{defect} & b & 0 \\
+        \end{block}
+    \end{blockarray}.
+\end{equation}
+
+Let \(\mathbf{u} = (b-c, -c, b, 0)\) be Eq.~\ref{eq:donation} in a vector format, and let \(\mathcal{U}
+= \{r, s, t, p\}\) denote the set of feasible payoffs, where \(r\) denotes the
+payoff of mutual cooperation, \(s\) the sucker's payoff, \(t\) the temptation to
+defect payoff, and \(p\) the punishment payoff.
+
+In the following sections we present several methods for calculating the
+updating payoffs. Initially, we discuss the conventional way of the
+expected payoffs and afterwards we present our newly introduced methods.
+
+\section{Updating Payoffs based on Expected Payoffs (perfect payoff memory)}\label{section:perfect_memory}
 
@@
 
-~\\
-{\bf Invasion Analysis}
+\subsubsection*{Invasion Analysis}
 
@@
 
-\subsection{Updating Payoffs based on the Last Round Payoff of One Interaction (limited payoff memory)}\label{section:limited_memory}
+\section{Updating Payoffs based on the Last Round Payoff of One Interaction (limited payoff memory)}\label{section:limited_memory}
 
@@
 
-\subsection{Updating Payoffs based on the last round payoff of $n$ interactions}\label{section:m_one_n_two}
+\section{Updating Payoffs based on the last round payoff of $n$ interactions}\label{section:m_one_n_two}
 
@@
 
-\subsubsection{Last Round Payoff of Two Interactions}
+\subsection{Last Round Payoff of Two Interactions}
 
@@
 
-\subsection{Updating Payoffs based on the Last $m$ Rounds Payoffs of One Interaction}\label{section:m_two_n_one}
+\section{Updating Payoffs based on the Last $m$ Rounds Payoffs of One Interaction}\label{section:m_two_n_one}
 
@@
 
-\subsection{Updating Payoffs based on the last two rounds payoff of two interactions (\(m=2\) and \(n=2\))}\label{section:m_two_n_two}
+\section{Updating Payoffs based on the last two rounds payoff of two interactions (\(m=2\) and \(n=2\))}\label{section:m_two_n_two}
 
@@
 
-%%%%%%%%%%%%%%%%%%
-%% FURTHER SIMULATIONS %% 
-%%%%%%%%%%%%%%%%%%
-
-\section{Further simulation results}~\label{section:furthersimulations}
-
-\subsection{Simulation Results on the Pairwise Comparison Process}\label{section:simulation_results}
+\section{Simulation Results on the Pairwise Comparison Process}\label{section:simulation_results}
 
@@
 
-\subsection{Expected and Last Round Updating Payoffs for Memory One Strategies}\label{section:memory_one}
+\section{Expected and Last Round Updating Payoffs for Memory One Strategies}\label{section:memory_one}
 
@@
 
-\subsection{Expected and Last Round Updating Payoffs for High Mutation ($\mu \neq 0$)}\label{section:mutation}
+\section{Expected and Last Round Updating Payoffs for High Mutation ($\mu \neq 0$)}\label{section:mutation}
 
@@
 independent runs of the pairwise process described in
-Section~\ref{section:model}, and at each time step we record the
+Section~\ref{section:pairwise_comparison}, and at each time step we record the
 average player \(\bar{s}=(\bar{y}, \bar{p}, \bar{q})\). The average cooperation
@@
 
-%%%%%%%%%%%%
-%% REFERENCES %%
-%%%%%%%%%%%%
-
-{
-{\setlength{\bibsep}{0\baselineskip}
-\bibliographystyle{naturemag}
-\bibliography{bibliography.bib}
-}
+\bibliographystyle{unsrt}
+\bibliography{bibliography}
 

### 2023-10-29 paper/si.tex
add symbols for C and D in payoff matrix, remove duplicate bibliography

--- 2023-10-26 paper/si.tex
@@
 In the most general setting, the resulting one-shot payoffs can be summarized by the payoff matrix 
-\begin{equation}\label{eq:prisoner}
-    \begin{blockarray}{ccc}
-        & \text{cooperate} & \text{defect} \\
-        \begin{block}{c(cc)}
-            \text{cooperate} & R & S \\
-            \text{defect} & T & P \\
-        \end{block}
-    \end{blockarray}.
-\end{equation}
+\begin{align}
+	\bordermatrix{%
+		& C & D \cr
+		C &\ R &\ S\  \cr
+		D &\ T &\ P\ \cr
+	} . \label{eq:prisoner}
+\end{align}
 Here, $R$ is interpreted as the reward for mutual cooperation, $S$ is the sucker's payoff, $T$ is the temptation, and $P$ is the punishment payoff~\citep{axelrod1981evolution}. 
@@
 Only as the mutation rate approaches one, and the processes become fully random, the cooperation rate approaches $1/2$ in both simulations.
+>>>>>>> origin/dev
 
@@
 \bibliographystyle{naturemag}
-\bibliography{bibliography.bib}
+\bibliography{bibliography}
 }

### 2023-10-29 paper/main.tex
minor changes after merge
fix reward notation for consistency with si

@@
 In fact, even beyond specific psychological considerations, a curious interpretation of payoffs arises from the formula commonly used for expected payoffs in repeated games.
-If $\delta\in\left[0,1\right)$ is the probability of continuing to another round in the game, then the expected payoff to an agent is $\left(1-\delta\right)\sum_{t=0}^{\infty}\delta^{t}\pi_{t}$, where $\pi_{t}$ is the payoff the agent receives in the stage game at time $t$.
+If $\delta\in\left[0,1\right)$ is the probability of continuing to another round in the game, then the expected payoff to an agent is $\left(1-\delta\right)\sum_{t=0}^{\infty}\delta^{t}r_{t}$, where $r_{t}$ is the reward the agent receives in the stage game at time $t$.
 Here, additional stochasticity arises due to uncertainty in the game length, and an agent might not be able to compute his or her expected payoff for use in decision-making.
-The probability that the game terminates after the interaction at time $T$ is $\left(1-\delta\right)\delta^{T}$, in which case $\left(1-\delta\right)\sum_{t=0}^{\infty}\delta^{t}\pi_{t}$ is exactly the expected payoff the agent receives at time $T$, i.e.
+The probability that the game terminates after the interaction at time $T$ is $\left(1-\delta\right)\delta^{T}$, in which case $\left(1-\delta\right)\sum_{t=0}^{\infty}\delta^{t}r_{t}$ is exactly the expected payoff the agent receives at time $T$, i.e.
 \emph{in the last round}.
-As an unbiased estimator of this expectation, the agent might thus use $\pi_{T}$ as a proxy for ``success'' when evaluating his or her behavior.
+As an unbiased estimator of this expectation, the agent might thus use $r_{T}$ as a proxy for ``success'' when evaluating his or her behavior.
 This gives a purely model-driven justification for why considering payoff in the last round of the game can result in more realistic extensions of traditional models.
 
-We note that the expected \emph{total} payoff in the game, $\pi_{0}+\pi_{1}+\cdots +\pi_{T}$, is given by $\sum_{t=0}^{\infty}\delta^{t}\pi_{t}$.
-This version of ``expected payoff'' appears less common in the literature on direct reciprocity than its normalization, $\left(1-\delta\right)\sum_{t=0}^{\infty}\delta^{t}\pi_{t}$, likely owing to the fact that payoffs can grow arbitrarily large with sufficiently long time horizons ($\delta\rightarrow 1^{-}$).
+We note that the expected \emph{total} payoff in the game, $r_{0}+r_{1}+\cdots +r_{T}$, is given by $\sum_{t=0}^{\infty}\delta^{t}r_{t}$.
+This version of ``expected payoff'' appears less common in the literature on direct reciprocity than its normalization, $\left(1-\delta\right)\sum_{t=0}^{\infty}\delta^{t}r_{t}$, likely owing to the fact that payoffs can grow arbitrarily large with sufficiently long time horizons ($\delta\rightarrow 1^{-}$).
 Non-normalized payoffs interfere with selection intensity ($\beta$) in models of social imitation, which is (presumably) why they appear less frequently in the literature.
@@
 \bibliographystyle{naturemag}
-\bibliography{bibliography.bib}
-}
+\bibliography{bibliography}
 }

### 2023-10-29 paper/main.tex
make "dynamics" plural

@@
 
-In each case, the evolutionary dynamics follows a similar cyclic pattern~\citep[as previously described in Refs.][]{imhof2010stochastic, Nowak1992tit}:
+In each case, the evolutionary dynamics follow a similar cyclic pattern~\citep[as previously described in Refs.][]{imhof2010stochastic, Nowak1992tit}:
 Resident populations of defectors are most likely invaded by strategies close to \tft. 
@@
 According to this mechanism, cooperation pays if it makes the interaction partner more cooperative in future. 
-To describe which strategies of direct reciprocity are most effective, the previous theoretical literature assumes that the evolutionary dynamics is determined by the players' expected payoffs~\citep{brauchli:JTB:1999,brandt:JTB:2006,ohtsuki:JTB:2007b,szolnoki:pre:2009b,imhof2010stochastic,van-segbroeck:prl:2012,grujic:jtb:2012,Martinez2012,stewart:pnas:2013,pinheiro:PLoSCB:2014,stewart:games:2015,Baek2016,McAvoy:ProcA:2019,glynatsi:SCR:2020,Schmid:PlosCB:2022,Murase:SciRep:2022}.
+To describe which strategies of direct reciprocity are most effective, the previous theoretical literature assumes that the evolutionary dynamics are determined by the players' expected payoffs~\citep{brauchli:JTB:1999,brandt:JTB:2006,ohtsuki:JTB:2007b,szolnoki:pre:2009b,imhof2010stochastic,van-segbroeck:prl:2012,grujic:jtb:2012,Martinez2012,stewart:pnas:2013,pinheiro:PLoSCB:2014,stewart:games:2015,Baek2016,McAvoy:ProcA:2019,glynatsi:SCR:2020,Schmid:PlosCB:2022,Murase:SciRep:2022}.
 To the extent that strategies are learned (not inherited), this assumption seems to impose rather stringent requirements on the individuals' cognitive abilities. 

### 2023-10-29 paper/main.tex
reorder probabilities in product

@@
 In fact, even beyond specific psychological considerations, a curious interpretation of payoffs arises from the formula commonly used for expected payoffs in repeated games.
-If $\delta\in\left[0,1\right)$ is the probability of continuing to another round in the game, then the expected payoff to an agent is $\left(1-\delta\right)\sum_{t=0}^{\infty}\delta^{t}r_{t}$, where $r_{t}$ is the reward the agent receives in the stage game at time $t$.
+If $\delta\in\left[0,1\right)$ is the probability of continuing to another round in the game, then the expected payoff to an agent is $\left(1-\delta\right)\sum_{t=0}^{\infty}\delta^{t}u_{t}$, where $u_{t}$ is the reward the agent receives in the stage game at time $t$.
 Here, additional stochasticity arises due to uncertainty in the game length, and an agent might not be able to compute his or her expected payoff for use in decision-making.
-The probability that the game terminates after the interaction at time $T$ is $\left(1-\delta\right)\delta^{T}$, in which case $\left(1-\delta\right)\sum_{t=0}^{\infty}\delta^{t}r_{t}$ is exactly the expected payoff the agent receives at time $T$, i.e.
+The probability that the game terminates after the interaction at time $T$ is $\delta^{T}\left(1-\delta\right)$, in which case $\left(1-\delta\right)\sum_{t=0}^{\infty}\delta^{t}u_{t}$ is exactly the expected payoff the agent receives at time $T$, i.e.
 \emph{in the last round}.
-As an unbiased estimator of this expectation, the agent might thus use $r_{T}$ as a proxy for ``success'' when evaluating his or her behavior.
+As an unbiased estimator of this expectation, the agent might thus use $u_{T}$ as a proxy for ``success'' when evaluating his or her behavior.
 This gives a purely model-driven justification for why considering payoff in the last round of the game can result in more realistic extensions of traditional models.
 
-We note that the expected \emph{total} payoff in the game, $r_{0}+r_{1}+\cdots +r_{T}$, is given by $\sum_{t=0}^{\infty}\delta^{t}r_{t}$.
-This version of ``expected payoff'' appears less common in the literature on direct reciprocity than its normalization, $\left(1-\delta\right)\sum_{t=0}^{\infty}\delta^{t}r_{t}$, likely owing to the fact that payoffs can grow arbitrarily large with sufficiently long time horizons ($\delta\rightarrow 1^{-}$).
+We note that the expected \emph{total} payoff in the game, $u_{0}+u_{1}+\cdots +u_{T}$, is given by $\sum_{t=0}^{\infty}\delta^{t}u_{t}$.
+This version of ``expected payoff'' appears less common in the literature on direct reciprocity than its normalization, $\left(1-\delta\right)\sum_{t=0}^{\infty}\delta^{t}u_{t}$, likely owing to the fact that payoffs can grow arbitrarily large with sufficiently long time horizons ($\delta\rightarrow 1^{-}$).
 Non-normalized payoffs interfere with selection intensity ($\beta$) in models of social imitation, which is (presumably) why they appear less frequently in the literature.

### 2023-10-29 paper/si.tex
remove proposition 1 and tildes

@@
 Section~\ref{section:model} summarizes the model. 
-In particular, we provide further details on our implementation of the evolutionary dynamics, and our use of the rare-mutations limit. 
-In Section~\ref{section:analyticalresults} we derive analytical results for the various settings we consider.
+In particular, we provide further details on our implementation of the evolutionary dynamics, and our use of the rare-mutation limit. 
+In Section~\ref{section:analyticalresults}, we derive analytical results for the various settings we consider.
 These settings differ in what kind of payoff information individuals take into account when updating their strategies. 
-In the perfect memory setting, individuals take into account all their interactions against all co-players.
-In the limited memory setting, they only consider the very last round of their very last interaction.  
+In the perfect-memory setting, individuals take into account all their interactions against all co-players.
+In the limited-memory setting, they only consider the very last round of their very last interaction.  
 In addition, we describe several model extensions in which the amount of information taken into account is in between these two extremes. 
 Finally, Section~\ref{section:furthersimulations} presents further simulation results. 
-In particular, we confirm that our main results continue to hold ({\it i}) when mutations are no longer rare, and  ({\it ii}) when players use memory-1 strategies instead of reactive strategies.
+In particular, we confirm that our main results continue to hold ({\it i}) when mutations are no longer rare, and  ({\it ii}) when players use memory-one strategies instead of reactive strategies.
 
@@
 {\bf Summary of the model.} As described in the main text, we study cooperative behavior in a population of size $N$, with $N$ being even.
-The dynamics unfolds on two time-scales. 
+The dynamics unfold on two time scales. 
 The short time scale describes the game dynamics. 
@@
 In that case, the player's new strategy is drawn uniformly from the space of all reactive strategies~$[0,1]^3$.  
-With probability $1\!-\!\mu$, the player engages in a pairwise comparison. 
+With probability $1\!-\!\mu$, the player compares itself with a neighbor. 
 In that case, the player randomly picks another individual from the population (referred to as the `role model', \rolemodel). 
-The learner adopts the role model's strategy with a probability \(\rho\) given by
+The learner adopts the role model's strategy with a probability \(\rho\), given by
 \begin{equation} \label{Eq:rho}
-    \rho(\pi_{L}, \pi_{R}) = \frac{1}{1\!+\! e^{\!-\!\beta (\pi_\rolemodel - \pi_\learner)}}.
+    \rho(\pi_{\learner}, \pi_{\rolemodel}) = \frac{1}{1\!+\! e^{\!-\!\beta (\pi_\rolemodel - \pi_\learner)}}.
 \end{equation}
@@
 
-%% MODEL: Intro to the rare-mutations limit %%
-
-\noindent
-{\bf Evolutionary simulations for the rare-mutations limit.} 
+%% MODEL: Intro to the rare-mutation limit %%
+
+\noindent
+{\bf Evolutionary simulations for the rare-mutation limit.} 
 To simulate the evolutionary dynamics of the pairwise comparison process, it is sometimes useful to assume that mutations are rare, $\mu\!\rightarrow\!0$. 
@@
 This assumptions makes computations more efficient, and it makes some of the results easier to interpret.
-In the following, we describe our implementation of the process in the rare-mutations limit in more detail.
-
-%% MODEL: Implementation of the rare-mutations limit %%
-
-Initially, the process starts with a population where all members use the same strategy (referred to as the resident strategy~\resident). 
+In the following, we describe our implementation of the process in the rare-mutation limit in more detail.
+
+%% MODEL: Implementation of the rare-mutation limit %%
+
+Initially, the process starts with a population where all members use the same strategy (referred to as the resident strategy, \resident). 
 Then one individual adopts a mutant strategy selected uniformly at random from the set of feasible strategies.
@@
 \begin{equation}\label{eq:fixation_probability}
-    \varphi_{\mutant} = \frac{1}{1+\sum\limits_{i=1}^{N-1}\prod\limits_k^i \frac{\lambda^-_k}{\lambda^+_k}}.
+    \varphi_{\mutant} = \frac{1}{1+\sum\limits_{i=1}^{N-1}\prod\limits_{k=1}^i \frac{\lambda^-_k}{\lambda^+_k}}.
 \end{equation}
@@
 We iterate this elementary population updating process for a large number of mutant strategies. 
-At each step, we record the current resident strategy, and the resulting average cooperation rate.
-
-We consider this limit of rare-mutations throughout the main text. 
+At each step, we record the current resident strategy and the resulting average cooperation rate.
+
+We consider this limit of rare-mutation throughout the main text. 
 The respective process is summarised by Algorithm~\ref{algorithm:pairwise_comparison}.
-In Section~\ref{section:furthersimulations} we present additional simulation results to show that our qualitative results continue to hold when the mutation rate is strictly bounded away from zero.
+In Section~\ref{section:furthersimulations}, we present additional simulation results to show that our qualitative results continue to hold when the mutation rate is strictly bounded away from zero.
 
@@
 Based on this general formula, the four entries of $\mathbf{v}=(v_R, v_S, v_T, v_P)$ can be computed explicitly. 
-Using the shortcut notation $r_i:=p_i\!-\!q_i$, we obtain
+Using the auxiliary notation $r_i:=p_i\!-\!q_i$, we obtain
     \begin{equation} \label{Eq:vAverage}
@@
 {\bf Stochastic stability of cooperation.}
-As an application of this formalism, we can compute when cooperation is stochastically stable in the perfect information setting. 
+As an application of this formalism, we can compute when cooperation is stochastically stable in the perfect-information setting. 
 To this end, suppose there is only a single mutant, $k\!=\!1$.
@@
 To treat the case of limited memory analytically, let $\mathcal{U} \!=\! \{R,S,T,P\}$ be the set of possible one-shot payoffs. 
-We first  compute the probability that a player receives a payoff of $u\!\in\!\mathcal{U}$ in the final round of the repeated game when the strategies of the two players are given.
-
-%% RESULTS: Limited memory: probability distribution for final payoff for two fixed strategies%%
-
-\begin{Prop}\label{proposition:last_round} 
-Consider a repeated game with continuation probability $\delta$, between two players with reactive strategies $\strategy_1\!=\!(y_1, p_1, q_1$)  and $\strategy_2\!=\!(y_2,p_2,q_2)$. 
-Let $\tilde{v}_u\!=\!\tilde{v}_u(\strategy_1,\strategy_2)$ denote the probability that the first player receives the payoff $u\!\in\!\mathcal{U}$ in the final round of the game. 
-Then this probability  $\tilde{v}_u$ is identical to the player's average probability $v_u$ to receive payoff $u$ across all rounds of the game, as specified by Eq.~\eqref{Eq:vAverage}.
-\end{Prop}
-
-\begin{proof}
-Let $\mathbf{\tilde{v}}:=(\tilde{v}_R, \tilde{v}_S, \tilde{v}_T, \tilde{v}_P)$. 
-To compute $\mathbf{\tilde{v}}$, we first compute the probability to receive each outcome $u\!\in\!\mathcal{U}$ given that the last round happens to be round $\tau\!\in\!\{0,1,\ldots\}$. 
-In that case, it follows from Eqs.~\eqref{Eq:v0} and~\eqref{Eq:vt} that the respective probability is
-\begin{equation} \label{Eq:vTildeSimple}
-\mathbf{\tilde v}(\tau) = \mathbf{v_0}\cdot M^\tau.
-\end{equation}
-Round $\tau$ is the last round with probability $\delta^\tau(1\!-\!\delta)$. 
-By the law of total probability, we can obtain $\mathbf{\tilde{v}}$ by summing up over all possible last rounds,
-\begin{equation}
-\tilde{v} = \sum_{\tau=0}^\infty \delta^\tau(1\!-\!\delta)  \cdot \mathbf{v_0}M^\tau 
-=  (1\!-\!\delta)\mathbf{v_0} \sum_{t=0}^\infty \delta^t  M^t. 
-\end{equation}
-This expression for $\mathbf{\tilde{v}}$ coincides with the expression for $\mathbf{v}$ in Eq.~\eqref{Eq:vDefinition}.
-\end{proof}
-
-\noindent
-Note that while we focus on the prisoner's dilemma in this work, the Proposition applies to any repeated matrix game.\\
+Since the game ends after round $T$ with probability $\delta^{T}\left(1-\delta\right)$, if two players use reactive strategies $\strategy_1\!=\!(y_1, p_1, q_1$)  and $\strategy_2\!=\!(y_2,p_2,q_2)$, then, by definition, the probability that the first player receives a payoff of $u\!\in\!\mathcal{U}$ in the final round of the game is given by Eq.~\eqref{Eq:vAverage}.
+To make the strategies explicit, we denote this probability by $v_{u}\!\left(\strategy_{1},\strategy_{2}\right)$.
 
@@
 
+~\\
 \noindent
@@
 In that case, there are four possible cases for their joint final payoffs, $(u_\resident,u_\mutant)\!\in\!\mathcal{U}_F\!:=\!\big\{(R,R),\,(S,T),\,(T,S)\,(P,P)\big\}$.
-These outcomes are distributed according to the distribution $\mathbf{\tilde{v}}(\strategy_\resident,\strategy_\mutant)$, as described in Proposition~\ref{proposition:last_round}.
+These outcomes are distributed according to the distribution $\mathbf{v}(\strategy_\resident,\strategy_\mutant)$ (Eq.~\eqref{Eq:vAverage}).
 \item The learner's last interaction was not with the role model, with probability $(N\!-\!2)/(N\!-\!1)$.
 In this case, there are four different subcases, depending on whether the resident's last interaction partner was a mutant or a resident, and depending on whether the mutant's last interaction partner was a mutant or a resident.
-As a result, the resident's last one-shot payoff is distributed according to $\mathbf{\tilde{v}}(\strategy_\resident,\strategy_\mutant)$ or $\mathbf{\tilde{v}}(\strategy_\resident,\strategy_\resident)$; the mutant's last payoff is distributed according to $\mathbf{\tilde{v}}(\strategy_\mutant,\strategy_\mutant)$ or $\mathbf{\tilde{v}}(\strategy_\mutant,\strategy_\resident)$, respectively.
+As a result, the resident's last one-shot payoff is distributed according to $\mathbf{v}(\strategy_\resident,\strategy_\mutant)$ or $\mathbf{v}(\strategy_\resident,\strategy_\resident)$; the mutant's last payoff is distributed according to $\mathbf{v}(\strategy_\mutant,\strategy_\mutant)$ or $\mathbf{v}(\strategy_\mutant,\strategy_\resident)$, respectively.
 \end{enumerate}
@@
 x(u_\resident,u_\mutant)	 &=
-&\frac{1}{N\!-\!1}\cdot  &\tilde v_{u_\resident}\!(\strategy_\resident,\strategy_\mutant)\cdot 1_{(u_\resident,u_\mutant)\in \mathcal{U}_F}\\[0.5cm]
+&\frac{1}{N\!-\!1}\cdot  & v_{u_\resident}\!(\strategy_\resident,\strategy_\mutant)\cdot 1_{(u_\resident,u_\mutant)\in \mathcal{U}_F}\\[0.5cm]
 &+	
 &\frac{N\!-\!2}{N\!-\!1} \cdot 
-&\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3}\cdot \tilde v_{u_\resident}\!(\strategy_\resident,\strategy_\mutant) \,\tilde v_{u_\mutant}\!(\strategy_\mutant,\strategy_\mutant) + 
- \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3}\cdot \tilde v_{u_\resident}\!(\strategy_\resident,\strategy_\mutant) \,\tilde v_{u_\mutant}\!(\strategy_\mutant,\strategy_\resident)\right.\\[0.5cm]
-&&&\left.
-+\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} \cdot \tilde v_{u_\resident}\!(\strategy_\resident,\strategy_\resident) \, \tilde v_{u_\mutant}\!(\strategy_\mutant,\strategy_\mutant) +
- \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} \cdot \tilde v_{u_\resident}\!(\strategy_\resident,\strategy_\resident) \, \tilde v_{u_\mutant}\!(\strategy_\mutant,\strategy_\resident)\right].
+&\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3}\cdot  v_{u_\resident}\!(\strategy_\resident,\strategy_\mutant) \, v_{u_\mutant}\!(\strategy_\mutant,\strategy_\mutant) + 
+ \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3}\cdot  v_{u_\resident}\!(\strategy_\resident,\strategy_\mutant) \, v_{u_\mutant}\!(\strategy_\mutant,\strategy_\resident)\right.\\[0.5cm]
+&&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} \cdot v_{u_\resident}\!(\strategy_\resident,\strategy_\resident) \,  v_{u_\mutant}\!(\strategy_\mutant,\strategy_\mutant) + 
+ \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} \cdot v_{u_\resident}\!(\strategy_\resident,\strategy_\resident) \, v_{u_\mutant}\!(\strategy_\mutant,\strategy_\resident)\right].
 \end{array}
@@
 There is $k\!=\!1$ mutant with strategy \alld{}. The remaining players use the resident strategy \gtft. 
-When two residents interact, it follows by Proposition~\ref{proposition:last_round} and Eq.~\eqref{Eq:vAverage} that the outcome of the final round is distributed according to the distribution
+When two residents interact, it follows by Eq.~\eqref{Eq:vAverage} that the outcome of the final round is distributed according to the distribution
 \begin{align*}
-\Big(\, \tilde v_R(\gtft,\gtft),~\tilde v_S(\gtft,\gtft),~ \tilde v_T(\gtft,\gtft),~ \tilde v_P(\gtft,\gtft)\, \Big) = (1,0,0,0). 
+\Big(\, v_R(\gtft,\gtft),~ v_S(\gtft,\gtft),~ v_T(\gtft,\gtft),~ v_P(\gtft,\gtft)\, \Big) = (1,0,0,0). 
 \end{align*}
@@
 \begin{align*}
-\Big(\, \tilde v_R(\alld,\gtft),~\tilde v_S(\alld,\gtft),~ \tilde v_T(\alld,\gtft),~ \tilde v_P(\alld,\gtft)\, \Big) = \big(0,~0,~1\!-\!\delta\!+\!\delta q,~\delta(1\! -\! q)\big). 
+\Big(\, v_R(\alld,\gtft),~ v_S(\alld,\gtft),~ v_T(\alld,\gtft),~ v_P(\alld,\gtft)\, \Big) = \big(0,~0,~1\!-\!\delta\!+\!\delta q,~\delta(1\! -\! q)\big). 
 \end{align*}
@@
 {\bf Motivation.} 
-In the limited memory setting, individuals take into account the outcome of one round, and of one interaction. 
+In the limited-memory setting, individuals take into account the outcome of one round, and of one interaction. 
 This setting can be generalized such that an individual
@@
   \begin{array}{lll}
-  \multicolumn{3}{l}{x(u_{\resident1}u_{\resident 2}, u_{\mutant 1}u_{\mutant 2})= \displaystyle \frac{1}{N\!-\!1}\cdot \tilde{v}_{u_{\resident1}}(\strategy_\resident,\strategy_\mutant)\cdot 1_{(u_{\resident1},u_{\mutant 1})\in \mathcal{U}_F} \cdot A }\\[0.4cm]
+  \multicolumn{3}{l}{x(u_{\resident1}u_{\resident 2}, u_{\mutant 1}u_{\mutant 2})= \displaystyle \frac{1}{N\!-\!1}\cdot v_{u_{\resident1}}(\strategy_\resident,\strategy_\mutant)\cdot 1_{(u_{\resident1},u_{\mutant 1})\in \mathcal{U}_F} \cdot A }\\[0.4cm]
 
@@
   \frac{(k - 1)(k - 2)}{(N\!-\!2)(N\!-\!3)}\,
-  \tilde{v}_{u_{\resident1}}\!(\strategy_\resident,\strategy_\mutant) \, \tilde{v}_{u_{\mutant 1}}\!(\strategy_\mutant,\strategy_\mutant) 
-  \left( \frac{\tilde{v}_{u_{\resident 2}}\!(\strategy_\resident,\strategy_\mutant)}{N-2} \, 1_{(u_{\resident 2},u_{\mutant 2})\in \mathcal{U}_F} +\frac{N-3}{N-2} [B_1 \!+\!
+  v_{u_{\resident1}}\!(\strategy_\resident,\strategy_\mutant) \, v_{u_{\mutant 1}}\!(\strategy_\mutant,\strategy_\mutant) 
+  \left( \frac{v_{u_{\resident 2}}\!(\strategy_\resident,\strategy_\mutant)}{N-2} \, 1_{(u_{\resident 2},u_{\mutant 2})\in \mathcal{U}_F} +\frac{N-3}{N-2} [B_1 \!+\!
 B_2 \!+\!
@@
   &&+ \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N\!-\!2)(N\!-\!3)}\,
-  \tilde{v}_{u_{\resident1}}\!(\strategy_\resident,\strategy_\mutant) \, \tilde{v}_{u_{\mutant 1}}\!(\strategy_\mutant,\strategy_\resident)  
-  \left(\frac{\tilde{v}_{u_{\resident 2}}\!(\strategy_\resident,\strategy_\mutant)}{N-2} \, 1_{(u_{\resident 2},u_{\mutant 2})\in \mathcal{U}_F} + \frac{N-3}{N-2} [C_1 \!+\!
+  v_{u_{\resident1}}\!(\strategy_\resident,\strategy_\mutant) \, v_{u_{\mutant 1}}\!(\strategy_\mutant,\strategy_\resident)  
+  \left(\frac{v_{u_{\resident 2}}\!(\strategy_\resident,\strategy_\mutant)}{N-2} \, 1_{(u_{\resident 2},u_{\mutant 2})\in \mathcal{U}_F} + \frac{N-3}{N-2} [C_1 \!+\!
 C_2 \!+\!
@@
   &&+ \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N\!-\!2)(N\!-\!3)} \,
-  \tilde{v}_{u_{\resident1}}\!(\strategy_\resident,\strategy_\resident) \, \tilde{v}_{u_{\mutant 1}}\!(\strategy_\mutant,\strategy_\mutant)  
-  \left(\frac{\tilde{v}_{u_{\resident 2}}\!(\strategy_\resident,\strategy_\mutant)}{N-2} \, 1_{(u_{\resident 2},u_{\mutant 2})\in \mathcal{U}_F} + \frac{N-3}{N-2} [D_1 \!+\!
+  v_{u_{\resident1}}\!(\strategy_\resident,\strategy_\resident) \, v_{u_{\mutant 1}}\!(\strategy_\mutant,\strategy_\mutant)  
+  \left(\frac{v_{u_{\resident 2}}\!(\strategy_\resident,\strategy_\mutant)}{N-2} \, 1_{(u_{\resident 2},u_{\mutant 2})\in \mathcal{U}_F} + \frac{N-3}{N-2} [D_1 \!+\!
 D_2 \!+\!
@@
   &&+ \frac{\left(N - k - 2\right) \left(N - k - 1\right)}{(N\!-\!2)(N\!-\!3)} 
-  \tilde{v}_{u_{\resident1}}\!(\strategy_\resident,\strategy_\resident) \, \tilde{v}_{u_{\mutant 1}}\!(\strategy_\mutant,\strategy_\resident)  
-  \left(\frac{ \tilde{v}_{u_{\resident 2}}\!(\strategy_\resident,\strategy_\mutant)}{N-2} \,1_{(u_{\resident 2},u_{\mutant 2})\in \mathcal{U}_F} + \frac{N-3}{N - 2} [E_1 \!+\!
+  v_{u_{\resident1}}\!(\strategy_\resident,\strategy_\resident) \, v_{u_{\mutant 1}}\!(\strategy_\mutant,\strategy_\resident)  
+  \left(\frac{ v_{u_{\resident 2}}\!(\strategy_\resident,\strategy_\mutant)}{N-2} \,1_{(u_{\resident 2},u_{\mutant 2})\in \mathcal{U}_F} + \frac{N-3}{N - 2} [E_1 \!+\!
 E_2 \!+\!
@@
   \begin{array}{ll}
-  A  = \frac{N-2}{N-1} & \left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} \, \tilde{v}_{u_{\mutant 1}}(\strategy_\resident,\strategy_\mutant)\, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) +
-   \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} \, \tilde{v}_{u_{\mutant 1}}(\strategy_\resident,\strategy_\mutant) \, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident)  \right.
-\\[0.5cm]
-   &+ \left.
-\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} \, \tilde{v}_{u_{\mutant 1}}(\strategy_\resident,\strategy_\resident) \, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) +
-   \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} \, \tilde{v}_{u_{\mutant 1}}(\strategy_\resident,\strategy_\resident) \, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident)\right]
+  A  = \frac{N-2}{N-1} & \left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} \, v_{u_{\mutant 1}}(\strategy_\resident,\strategy_\mutant)\, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) + 
+   \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} \, v_{u_{\mutant 1}}(\strategy_\resident,\strategy_\mutant) \, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident)  \right. \\[0.5cm]
+   &+ \left. \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} \, v_{u_{\mutant 1}}(\strategy_\resident,\strategy_\resident) \, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) + 
+   \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} \, v_{u_{\mutant 1}}(\strategy_\resident,\strategy_\resident) \, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident)\right]
    \end{array} 
@@
   \begin{array}{ll}
-   B_1  = \frac{(k - 2) (k - 3)}{(N - 3) (N - 4)}\, \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant) \,\tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant),
-   &B_2 = \frac{\left(k - 2\right) \left(N - k - 1\right)}{(N - 3) (N - 4)}\, \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident) \,\tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant), \\[0.5cm]
-   B_3  = \frac{\left(k - 2\right) \left(N - k - 1\right)}{(N - 3) (N - 4)}\, \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant) \,\tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident),
-   &B_4 = \frac{\left(N - k - 2\right) \left(N - k - 1\right)}{(N - 3) (N - 4)}\, \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident) \, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident).
+   B_1  = \frac{(k - 2) (k - 3)}{(N - 3) (N - 4)}\, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant) \,v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant),
+   &B_2 = \frac{\left(k - 2\right) \left(N - k - 1\right)}{(N - 3) (N - 4)}\, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident) \,v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant), \\[0.5cm]
+   B_3  = \frac{\left(k - 2\right) \left(N - k - 1\right)}{(N - 3) (N - 4)}\, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant) \,v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident),
+   &B_4 = \frac{\left(N - k - 2\right) \left(N - k - 1\right)}{(N - 3) (N - 4)}\, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident) \, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident).
    \end{array}
@@
   \begin{array}{llrl}
-   C_1  = \frac{\left(k - 3\right) \left(k - 1\right)}{(N - 3) (N - 4)} \, \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant) \, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) \quad
-   &C_2 = \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} \, \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident) \, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) \\[0.5cm]
-   C_3  = \frac{\left(k - 2\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} \, \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant) \, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident) \quad
-   &C_4 = \frac{\left(N - k - 2\right)^{2}}{(N - 3) (N - 4)} \, \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident) \, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident)
+   C_1  = \frac{\left(k - 3\right) \left(k - 1\right)}{(N - 3) (N - 4)} \, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant) \, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) \quad
+   &C_2 = \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} \, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident) \, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) \\[0.5cm]
+   C_3  = \frac{\left(k - 2\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} \, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant) \, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident) \quad
+   &C_4 = \frac{\left(N - k - 2\right)^{2}}{(N - 3) (N - 4)} \, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident) \, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident) 
    \end{array}
@@
   \begin{array}{llrl}
-   D_1  = \frac{\left(k - 2\right)^{2}}{(N - 3) (N - 4)} \, \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant)\, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) \quad
-   &D_2 = \frac{\left(k - 2\right) \left(N - k - 2\right)}{(N - 3) (N - 4)}\,  \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident)\, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) \\[0.5cm]
-   D_3  = \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} \, \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant) \,\tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident) \quad
-   &D_4 = \frac{\left(N - k - 3\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} \, \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident) \,\tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident)
+   D_1  = \frac{\left(k - 2\right)^{2}}{(N - 3) (N - 4)} \, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant)\, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) \quad
+   &D_2 = \frac{\left(k - 2\right) \left(N - k - 2\right)}{(N - 3) (N - 4)}\,  v_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident)\, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) \\[0.5cm]
+   D_3  = \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} \, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant) \,v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident) \quad
+   &D_4 = \frac{\left(N - k - 3\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} \, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident) \,v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident)
       \end{array}
@@
     \begin{array}{llrl}
-   E_1  = \frac{(k - 2)(k - 1)}{(N-3)(N-4)}\, \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant)\, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) \quad
-   &E_2 = \frac{\left(k - 1\right) \left(N - k - 2\right)}{(N - 3) (N - 4)}\, \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident)\, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) \\[0.5cm]
-   E_3  = \frac{\left(k - 1\right) \left(N - k - 2\right)}{(N - 3) (N - 4)}\, \tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant)\, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident) \quad
-   &E_4 = \frac{\left(N - k - 3\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} \,\tilde{v}_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident)\, \tilde{v}_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident)
+   E_1  = \frac{(k - 2)(k - 1)}{(N-3)(N-4)}\, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant)\, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) \quad
+   &E_2 = \frac{\left(k - 1\right) \left(N - k - 2\right)}{(N - 3) (N - 4)}\, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident)\, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) \\[0.5cm]
+   E_3  = \frac{\left(k - 1\right) \left(N - k - 2\right)}{(N - 3) (N - 4)}\, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant)\, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident) \quad
+   &E_4 = \frac{\left(N - k - 3\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} \,v_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident)\, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident)
   \end{array}
@@
 \end{equation}
-Only when \(b/2\!=\!c\) we obtain the same expression as in the limited memory case,~Eq.~\eqref{Eq:TransitionRatioSimple}. 
+Only when \(b/2\!=\!c\) we obtain the same expression as in the limited-memory case,~Eq.~\eqref{Eq:TransitionRatioSimple}. 
 The requirement $q\!\ge\!0$ implies that the minimum continuation probability is $\delta\!>\!1\!-\!\sqrt{2}/2\approx 0.293$ in the first case, $\delta\!>\!\frac{1}{2}$ in the second, and $\delta>\sqrt{2}/2\approx 0.707$ in the third case.
@@
 Instead, we interpret it as the (conditional) probability that the number of mutants increases, given that the learner and the role model happened to interact at least for two rounds during their respective last interaction.
-This new interpretation of transition probabilities affects the time-scale of fixation. 
+This new interpretation of transition probabilities affects the time scale of fixation. 
 After all, we ignore all time steps in which no update occurred because a player's last game only lasted for a single round.
@@
 Let $x(u_\resident u'_\resident, u_\mutant u'_\mutant)$ be the probability that the two players received payoffs $u$ and $u'$ in the last two rounds of their last repeated game.
-We can compute this probability based on the same logic as in the limited memory setting (Section~\ref{section:limited_memory}). 
+We can compute this probability based on the same logic as in the limited-memory setting (Section~\ref{section:limited_memory}). 
 This yields
@@
 The results of these simulations are shown in Supplementary Figure~\ref{fig:expected_payoffs_results}. 
-Qualitatively, all settings lead to a similar dynamics as in the case of perfect and limited memory described in the main text. 
+Qualitatively, all settings lead to similar dynamics as in the case of perfect and limited memory described in the main text. 
 In particular, populations are clustered around \alld, and around a strip of conditionally cooperative strategies, \gtft$=\!(1,1,q)$. 
@@
 Conversely, there are memory-one strategies that cannot be represented by a reactive strategy, such as the strategy \texttt{WSLS}$=(1,1,0,0,1)$. 
-Therefore, the set of memory-1 strategies is a strict superset of the reactive strategies.
-
-To explore the dynamics among memory-1 strategies, we perform four separate simulations, for both perfect and limited memory, and for two different benefit values. 
+Therefore, the set of memory-one strategies is a strict superset of the reactive strategies.
+
+To explore the dynamics among memory-one strategies, we perform four separate simulations, for both perfect and limited memory, and for two different benefit values. 
 Figure~\ref{fig:memory_one_low_benefit} shows the results for a comparably low benefit of cooperation; 

### 2023-10-31 paper/si.tex

@@
     \end{equation}
-Using this distribution $\mathbf{v}$, we compute player 1's expected payoff as the weighted average
+Using this distribution $\mathbf{v}$, we compute the first player's expected payoff as the weighted average
 \begin{equation}
@@
 This happens with probability $1/(N\!-\!1)$.
-In that case, there are four possible cases for their joint final payoffs, $(u_\resident,u_\mutant)\!\in\!\mathcal{U}_F\!:=\!\big\{(R,R),\,(S,T),\,(T,S)\,(P,P)\big\}$.
-These outcomes are distributed according to the distribution $\mathbf{v}(\strategy_\resident,\strategy_\mutant)$ (Eq.~\eqref{Eq:vAverage}).
+In that case, there are four possible cases for their joint final payoffs, $(u_\resident,u_\mutant)\!\in\!\mathcal{U}_F\!:=\!\big\{(R,R),\,(S,T),\,(T,S),\,(P,P)\big\}$.
+These outcomes follow the distribution $\mathbf{v}(\strategy_\resident,\strategy_\mutant)$ (Eq.~\eqref{Eq:vAverage}).
 \item The learner's last interaction was not with the role model, with probability $(N\!-\!2)/(N\!-\!1)$.
@@
 +  \frac{1}{N -1} \left(\frac{ 1 - \delta + \delta q}{1 + e^{-  \beta (b + c)}}
-+  \frac{ \delta  \left(1 - q \right)}{2} \right)}
++  \frac{ \delta  \left(1 - q \right)}{2} \right)} .
 \end{equation*}
@@
 This ratio exceeds one if  \(q <1 \!-\! 1/(2 \delta)\). 
-For such a strategy to be feasible we require $q\!>\!0$, which in turn implies $\delta\!>\!1/2$. 
+For such a strategy to be feasible, we require $q\!>\!0$, which in turn implies $\delta\!>\!1/2$. 
 Moreover, in the special case that games are infinitely repeated, \(\delta\! \rightarrow\! 1\), we conclude that cooperation is stochastically stable if  \(q \!<\! 1/2\). 
@@
 probability that the number of mutants increase by one equals the probability
-that the mutant goes extinct).
+that the mutant goes extinct.)
 
@@
 This setting can be generalized such that an individual
-considers \(m\) rounds and of \(n\) interactions. Here we discuss the case that
+considers \(m\) rounds and of \(n\) interactions. Here, we discuss the case that
 the update depends on the last round of \(n\) interactions.
@@
 We assume that each pair is unique, such that the learner and the role model
-can be matched at most once. 
+can be matched at most once, which is a reasonable assumption in large populations.
 The case of \(n\!=\!1\) corresponds to the previous setting of limited memory. 
-There we have seen that there are five possible combinations to consider. 
+There, we have seen that there are five possible combinations to consider. 
 As we increase $n$, the number of possible combinations increases non-linearly; 
-for a graphical illustration, see Figure~\ref{fig:matching_tree}. 
+for a graphical illustration, see Fig.~\ref{fig:matching_tree}. 
 In the following, we study the case of \(n=2\), such that the learner takes into account the players' final payoffs of two interactions.\\
@@
 These are the same as in the previous setting. 
-Either the learner interacts directly with the role model; otherwise we need to take into account whether the resident interacts with a resident or a mutant, and whether the mutant interacts with a resident or a mutant.
+The learner either interacts directly with the role model or we need to take into account whether the resident interacts with a resident or a mutant, and whether the mutant interacts with a resident or a mutant.
 In case the learner and the role model did not directly interact with each other during the first stage, there are again five possible combinations in the second stage. 
@@
 E_3 \!+\!
-E_4]\right) \biggl].
+E_4]\right) \biggl] .
   \end{array}
@@
    &+ \left. \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} \, v_{u_{\mutant 1}}(\strategy_\resident,\strategy_\resident) \, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) + 
-   \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} \, v_{u_{\mutant 1}}(\strategy_\resident,\strategy_\resident) \, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident)\right]
+   \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} \, v_{u_{\mutant 1}}(\strategy_\resident,\strategy_\resident) \, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident)\right] .
    \end{array} 
@@
  The second line of Eq.~\eqref{Eq:xTwoInteractions} corresponds to the case that during the first stage, both the resident and the mutant interact with a mutant. 
- In the seconds stage, they may interact with each other. 
+ In the second stage, they may interact with each other. 
  If they do not, then again either both of them interact with a mutant, one of them interacts with a resident whereas the other interacts with a mutant, or both interact with a resident.
@@
   \begin{array}{llrl}
-   C_1  = \frac{\left(k - 3\right) \left(k - 1\right)}{(N - 3) (N - 4)} \, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant) \, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) \quad
-   &C_2 = \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} \, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident) \, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) \\[0.5cm]
-   C_3  = \frac{\left(k - 2\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} \, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant) \, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident) \quad
-   &C_4 = \frac{\left(N - k - 2\right)^{2}}{(N - 3) (N - 4)} \, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident) \, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident) 
+   C_1  = \frac{\left(k - 3\right) \left(k - 1\right)}{(N - 3) (N - 4)} \, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant) \, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) , \quad
+   &C_2 = \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} \, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident) \, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) , \\[0.5cm]
+   C_3  = \frac{\left(k - 2\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} \, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant) \, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident) , \quad
+   &C_4 = \frac{\left(N - k - 2\right)^{2}}{(N - 3) (N - 4)} \, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident) \, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident) .
    \end{array}
@@
   \begin{array}{llrl}
-   D_1  = \frac{\left(k - 2\right)^{2}}{(N - 3) (N - 4)} \, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant)\, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) \quad
-   &D_2 = \frac{\left(k - 2\right) \left(N - k - 2\right)}{(N - 3) (N - 4)}\,  v_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident)\, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) \\[0.5cm]
-   D_3  = \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} \, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant) \,v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident) \quad
-   &D_4 = \frac{\left(N - k - 3\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} \, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident) \,v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident)
+   D_1  = \frac{\left(k - 2\right)^{2}}{(N - 3) (N - 4)} \, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant)\, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) , \quad
+   &D_2 = \frac{\left(k - 2\right) \left(N - k - 2\right)}{(N - 3) (N - 4)}\,  v_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident)\, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) , \\[0.5cm]
+   D_3  = \frac{\left(k - 1\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} \, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant) \,v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident) , \quad
+   &D_4 = \frac{\left(N - k - 3\right) \left(N - k - 1\right)}{(N - 3) (N - 4)} \, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident) \,v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident) .
       \end{array}
@@
     \begin{array}{llrl}
-   E_1  = \frac{(k - 2)(k - 1)}{(N-3)(N-4)}\, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant)\, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) \quad
-   &E_2 = \frac{\left(k - 1\right) \left(N - k - 2\right)}{(N - 3) (N - 4)}\, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident)\, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) \\[0.5cm]
-   E_3  = \frac{\left(k - 1\right) \left(N - k - 2\right)}{(N - 3) (N - 4)}\, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant)\, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident) \quad
-   &E_4 = \frac{\left(N - k - 3\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} \,v_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident)\, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident)
+   E_1  = \frac{(k - 2)(k - 1)}{(N-3)(N-4)}\, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant)\, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) , \quad
+   &E_2 = \frac{\left(k - 1\right) \left(N - k - 2\right)}{(N - 3) (N - 4)}\, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident)\, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\mutant) , \\[0.5cm]
+   E_3  = \frac{\left(k - 1\right) \left(N - k - 2\right)}{(N - 3) (N - 4)}\, v_{u_{\resident 2}}(\strategy_\resident,\strategy_\mutant)\, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident) , \quad
+   &E_4 = \frac{\left(N - k - 3\right) \left(N - k - 2\right)}{(N - 3) (N - 4)} \,v_{u_{\resident 2}}(\strategy_\resident,\strategy_\resident)\, v_{u_{\mutant 2}}(\strategy_\mutant,\strategy_\resident) .
   \end{array}
@@
 \begin{array}{ll}
-  x(RR, TT) =  \frac{N - 3}{N - 1}\left(\delta q - \delta + 1\right)^{2} & 
-  x(RR, TP) =  \frac{N - 3}{N - 1} \delta \left(1-q \right) \left(\delta q - \delta + 1\right)\\[0.2cm]
-  x(RR, PT) =  \frac{N - 3}{N - 1} \delta \left(1-q\right) \left(\delta q - \delta + 1\right) &
-  x(RR, PP) =  \frac{N - 3}{N - 1} \delta^{2}  \left(1 - q\right)^{2}\\[0.2cm]
-  x(RS, TT) =  \frac{1}{N-1} \left(\delta q - \delta + 1\right)^{2} &
-  x(RS, PT) =  \frac{1}{N-1} \delta \left(1-q\right) \left(\delta q - \delta + 1\right) \\[0.2cm]
-  x(RP, TP) =  \frac{1}{N-1} \delta \left(1-q\right) \left(\delta q - \delta + 1\right) &
-  x(RP, PP) =  \frac{1}{N-1} \delta^{2} \left(1 - q\right)^{2} \\[0.2cm]
-  x(SR, TT) =  \frac{1}{N-1} \left(\delta q - \delta + 1\right)^{2} &
-  x(SR, TP) =  \frac{1}{N-1} \delta \left(1-q\right) \left(\delta q - \delta + 1\right)\\[0.2cm]
-  x(PR, PT) =  \frac{1}{N-1} \delta \left(1-q\right) \left(\delta q - \delta + 1\right) &
-  x(PR, PP) =  \frac{1}{N-1} \delta^{2} \left(1-q\right)^{2}
+  x(RR, TT) =  \frac{N - 3}{N - 1}\left(\delta q - \delta + 1\right)^{2}, & 
+  x(RR, TP) =  \frac{N - 3}{N - 1} \delta \left(1-q \right) \left(\delta q - \delta + 1\right) ,\\[0.2cm]
+  x(RR, PT) =  \frac{N - 3}{N - 1} \delta \left(1-q\right) \left(\delta q - \delta + 1\right) , &
+  x(RR, PP) =  \frac{N - 3}{N - 1} \delta^{2}  \left(1 - q\right)^{2} , \\[0.2cm]
+  x(RS, TT) =  \frac{1}{N-1} \left(\delta q - \delta + 1\right)^{2} , &
+  x(RS, PT) =  \frac{1}{N-1} \delta \left(1-q\right) \left(\delta q - \delta + 1\right) , \\[0.2cm]
+  x(RP, TP) =  \frac{1}{N-1} \delta \left(1-q\right) \left(\delta q - \delta + 1\right) , &
+  x(RP, PP) =  \frac{1}{N-1} \delta^{2} \left(1 - q\right)^{2} , \\[0.2cm]
+  x(SR, TT) =  \frac{1}{N-1} \left(\delta q - \delta + 1\right)^{2} , &
+  x(SR, TP) =  \frac{1}{N-1} \delta \left(1-q\right) \left(\delta q - \delta + 1\right) , \\[0.2cm]
+  x(PR, PT) =  \frac{1}{N-1} \delta \left(1-q\right) \left(\delta q - \delta + 1\right) ,  &
+  x(PR, PP) =  \frac{1}{N-1} \delta^{2} \left(1-q\right)^{2} .
   \end{array}
@@
 + \frac{2}{N-1} \left(\frac{\delta^{2} \left(1 - q\right)^{2}}{1 + e^{ \beta \left(b-c\right)/2}} + \frac{\delta \left(1 - q\right) \left(\delta q - \delta + 1\right)}{1 + e^{- \beta c/2}} + \frac{\delta \left(1 - q\right) \left(\delta q - \delta + 1\right)}{1 + e^{- \beta c}} + \frac{\left(\delta q - \delta + 1\right)^{2}}{1 + e^{- \beta \left(b/2 + c\right)}}\right)}
-$}
+$} .
 \end{equation}
@@
 \begin{cases}
- \frac{1-(1-\delta(1-q))^2}{(1-\delta(1-q))^2} &\text{if}~ \frac{b}{2} > c \\[0.2cm]
-  \frac{\delta(1-q)}{1-\delta(1-q)}  &\text{if}~ \frac{b}{2} = c \\[0.2cm]
-  \frac{\delta^2(1-q)^2}{1-\delta^2(1-q)^2} &\text{if}~\frac{b}{2} < c
+ \frac{1-(1-\delta(1-q))^2}{(1-\delta(1-q))^2} &\text{if}~ \frac{b}{2} > c , \\[0.2cm]
+  \frac{\delta(1-q)}{1-\delta(1-q)}  &\text{if}~ \frac{b}{2} = c , \\[0.2cm]
+  \frac{\delta^2(1-q)^2}{1-\delta^2(1-q)^2} &\text{if}~\frac{b}{2} < c .
 \end{cases}
@@
 \begin{cases}
-  q <\frac{\delta - 1 + \sqrt{2}/2}{\delta}	&\text{if}~ \frac{b}{2} > c \\[0.15cm]
-  q<1-\frac{1}{2\delta}	&\text{if}~\frac{b}{2}=c\\[0.15cm]
-  q <\frac{\delta - \sqrt{2}/2}{\delta}\ &\text{if}~\frac{b}{2} < c
+  q <\frac{\delta - 1 + \sqrt{2}/2}{\delta}	&\text{if}~ \frac{b}{2} > c , \\[0.15cm]
+  q<1-\frac{1}{2\delta}	&\text{if}~\frac{b}{2}=c , \\[0.15cm]
+  q <\frac{\delta - \sqrt{2}/2}{\delta}\ &\text{if}~\frac{b}{2} < c .
 \end{cases}
@@
 To compute the ratio of transition probabilities, again we need to compute how likely it is that the resident and the mutant received any possible combination of two payoffs during the last two rounds.
-We start by computing the players' payoff distribution in the penultimate round.
+The probability that the game lasts for at least two rounds is $\delta$, since that is the probability of continuing from round zero (the initial round) to round one.
+The distribution of actions in in the penultimate round, conditioned on there being at least two rounds in the game, is then
+\begin{align}
+	\frac{1}{\delta} \sum_{t=1}^{\infty} \delta^{t}\left(1-\delta\right) \mathbf{v_{0}} M^{t-1} &= \left(1-\delta\right)\mathbf{v_{0}}\sum_{t=0}^{\infty}\delta^{t}M^{t} ,
+\end{align}
+which is identical to the expression for $\mathbf{v}$ in Eq.~\eqref{Eq:vDefinition}.
+Thus, the probability that the first player receives the payoff $u\!\in\!\mathcal{U}$ in the penultimate round of the game, conditional on the event that the game lasts at least two rounds, is again identical to the player's average probability $v_u$ to receive payoff $u$ across all rounds of the game, as specified by Eq.~\eqref{Eq:vAverage}.
 
@@
 
-\begin{Prop}\label{proposition:penultimate_round} 
-Consider a repeated game with continuation probability $\delta$, between two players with reactive strategies $\strategy_1\!=\!(y_1, p_1, q_1$)  and $\strategy_2\!=\!(y_2,p_2,q_2)$. 
-Let $\hat{v}_u\!=\!\hat{v}_u(\strategy_1,\strategy_2)$ denote the probability that the first player receives the payoff $u\!\in\!\mathcal{U}$ in the penultimate round of the game, conditional on the event that the game lasts at least two rounds.
-Then this probability  $\hat{v}_u$ is again identical to the player's average probability $v_u$ to receive payoff $u$ across all rounds of the game, as specified by Eq.~\eqref{Eq:vAverage}.
-\end{Prop}
-
-\begin{proof}
-The proof is analogous to the proof of Proposition~\ref{proposition:last_round}. 
-Let $\mathbf{\hat{v}}:=(\hat{v}_R, \hat{v}_S, \hat{v}_T, \hat{v}_P)$, and let $\mathbf{\hat{v}}(\tau)$ denote the conditional distribution given that  the last round of the game happens to be round $\tau\!\in\!\{1,2,\ldots\}$ (note that we refer to the initial round as round zero, so the game is assumed to take at least two rounds).
-Similar to Eq.~\eqref{Eq:vTildeSimple}, we obtain
-\begin{equation}
-\mathbf{\hat v}(\tau) = \mathbf{v_0}\cdot M^{\tau-1}.
-\end{equation}
-The probability that $\tau\!\ge\!1$ is the last round, conditional on the event that the game takes at least two rounds, is 
-$$ \mathbb{P}(\tau ~\text{is final round}~|~\text{there are at least two rounds}\,) = \frac{\delta^\tau(1\!-\!\delta)}{\delta} =  \delta^{\tau-1}(1\!-\!\delta).$$
-By the law of total probability, we again obtain $\mathbf{\hat{v}}$ by summing up over all possible last rounds,
-\begin{equation}
-\hat{v} = \sum_{\tau=1}^\infty \delta^{\tau-1}(1\!-\!\delta)  \cdot \mathbf{v_0}M^{\tau-1} 
-=  (1\!-\!\delta)\mathbf{v_0} \sum_{t=0}^\infty \delta^t  M^t. 
-\end{equation}
-This expression for $\mathbf{\hat{v}}$ coincides with the expression for $\mathbf{v}$ in Eq.~\eqref{Eq:vDefinition}.
-\end{proof}
-
-\noindent
-Based on Proposition~\ref{proposition:penultimate_round}, we can derive a probability distribution for player 1's payoff in the last two rounds of a repeated game. 
+~\\
+\noindent
+We can now derive a probability distribution for the first player's payoff in the last two rounds of a repeated game. 
 To this end, consider the $4\times4$ transition matrix $M(\strategy_1,\strategy_2)=(m_{u,u'})$ according to~Eq.~\eqref{eq:transition_matrix}. 
-Instead of the usual indexing of the four rows by numbers, $i\!\in\!\{1,2,3,4\}$, here we label the four rows of this matrix by player 1's payoffs in the previous round, $u\!\in\!\{R,S,T,P\}$. 
+Instead of the usual indexing of the four rows by numbers, $i\!\in\!\{1,2,3,4\}$, here we label the four rows of this matrix by the first player's payoffs in the previous round, $u\!\in\!\{R,S,T,P\}$.
 Similarly, we label the four columns of this matrix by the resulting payoffs to player~1 in the next round. 
@@
 So by Eq.~\eqref{eq:transition_matrix}, $m_{ST} = \bar{q}_1 p_2=(1\!-\!q_1)p_2$.
-Based on this notation, we can describe the probability $w_{uu'}(\strategy_1,\strategy_2)$ that player~1 obtains a payoff of $u$ in the penultimate round and a payoff of $u'$ in the last round. 
-Based on the vector $\mathbf{\hat{v}}(\strategy_1,\strategy_2):=(\hat{v}_R, \hat{v}_S, \hat{v}_T, \hat{v}_P)$ that describes the outcome of the penultimate round, we obtain
-\begin{equation}
-w_{uu'}(\strategy_1,\strategy_2) = \hat{v}_u \cdot m_{u,u'}.
-\end{equation}
+Based on this notation, we can describe the probability $w_{uu'}(\strategy_1,\strategy_2)$ that player~1 obtains a payoff of $u$ in the penultimate round (probability $v_{u}$) and a payoff of $u'$ in the last round (probability $m_{u,u'}$), giving $w_{uu'}(\strategy_1,\strategy_2) = v_u \cdot m_{u,u'}$.
+
  ~\\
-
  \noindent
@@
  {\bf Stochastic stability of cooperation.}
-We once again calculate how easily a single \alld{} mutant can invade into a
+We once again calculate how easily a single \alld{} mutant can invade a
 resident population of \gtft{} players. When two residents interact,
@@
 As a consequence, we obtain the following probabilities  $x(u_\resident u'_\resident, u_\mutant u'_\mutant)$ that the
-payoffs of a randomly chosen \gtft resident are $u_\resident$ and $u_\resident'$, whereas the payoffs of the mutant are $u_\mutant$ and $u_\mutant'$. 
+payoffs of a randomly chosen \gtft{} resident are $u_\resident$ and $u_\resident'$, whereas the payoffs of the mutant are $u_\mutant$ and $u_\mutant'$:
 \begin{equation*}
@@
 \begin{array}{ll}
-  x(RR, TT) =  \frac{N-2}{N - 1}  q \left(\delta q - \delta + 1\right) 
-  &x(RR, TP) =  \frac{N-2}{N - 1} \left(1-q\right) \left(\delta q - \delta + 1\right)\\[0.2cm]
-  x(RR, PT) =  \frac{N-2}{N - 1} \delta q  \left(1-q\right) 
-  &x(RR, PP) =  \frac{N-2}{N - 1} \delta  \left(1 - q\right)^{2}\\[0.2cm]
-  x(SS, TT) =  \frac{1}{N -1} q \left(\delta q - \delta + 1\right) 
-  &x(SP, TP) =  \frac{1}{N -1} \left(1-q\right) \left(\delta q - \delta + 1\right) \\[0.2cm]
-  x(PS,PT) =   \frac{1}{N -1}\delta q \left(1-q\right) 
-  &x(PP, PP) =  \frac{1}{N -1}\delta \left(1-q\right)^{2}\\[0.2cm]
+  x(RR, TT) =  \frac{N-2}{N - 1}  q \left(\delta q - \delta + 1\right) ,
+  &x(RR, TP) =  \frac{N-2}{N - 1} \left(1-q\right) \left(\delta q - \delta + 1\right) ,\\[0.2cm]
+  x(RR, PT) =  \frac{N-2}{N - 1} \delta q  \left(1-q\right) ,
+  &x(RR, PP) =  \frac{N-2}{N - 1} \delta  \left(1 - q\right)^{2} ,\\[0.2cm]
+  x(SS, TT) =  \frac{1}{N -1} q \left(\delta q - \delta + 1\right) ,
+  &x(SP, TP) =  \frac{1}{N -1} \left(1-q\right) \left(\delta q - \delta + 1\right) , \\[0.2cm]
+  x(PS,PT) =   \frac{1}{N -1}\delta q \left(1-q\right) ,
+  &x(PP, PP) =  \frac{1}{N -1}\delta \left(1-q\right)^{2} ,\\[0.2cm]
   \multicolumn{2}{c}{x(u_\resident u'_\resident, u_\mutant u'_\mutant) =  0~~ \text{ for all other payoff combinations.}}
@@
 \begin{cases}
-  \frac{(1-q) (1+\delta q)}{q \left(\delta q - \delta + 1\right)}  &\text{if}~~\frac{b}{2} > c \\[0.2cm]
-  \frac{\left(1+\delta \right) \left(1-q\right)} {\delta q - \delta + q + 1} &\text{if}~~ \frac{b}{2} = c \\[0.2cm]
-  \frac{\delta (1-q)^{2}}{1-\delta(1-q)^2 } &\text{if}~~ \frac{b}{2} < c
+  \frac{(1-q) (1+\delta q)}{q \left(\delta q - \delta + 1\right)}  &\text{if}~~\frac{b}{2} > c , \\[0.2cm]
+  \frac{\left(1+\delta \right) \left(1-q\right)} {\delta q - \delta + q + 1} &\text{if}~~ \frac{b}{2} = c , \\[0.2cm]
+  \frac{\delta (1-q)^{2}}{1-\delta(1-q)^2 } &\text{if}~~ \frac{b}{2} < c .
 \end{cases}
@@
 \begin{cases}
-  q<\frac{1}{1-\delta+\sqrt{1+\delta^2}}  &\text{if}~~\frac{b}{2} > c \\[0.1cm]
-  q<\frac{\delta} {1+\delta} &\text{if}~~ \frac{b}{2} = c \\[0.1cm]
-  q<1-\frac{1}{\sqrt{2\delta}} &\text{if}~~ \frac{b}{2} < c
+  q<\frac{1}{1-\delta+\sqrt{1+\delta^2}}  &\text{if}~~\frac{b}{2} > c , \\[0.1cm]
+  q<\frac{\delta} {1+\delta} &\text{if}~~ \frac{b}{2} = c , \\[0.1cm]
+  q<1-\frac{1}{\sqrt{2\delta}} &\text{if}~~ \frac{b}{2} < c .
 \end{cases}
@@
 
-As our final extension, we consider a combination of the previous two settings, discussed in Sections~\ref{section:m_one_n_two} and Section~\ref{section:m_two_n_one}. 
-Now individuals take into account two interactions with different co-players, and from each interaction, they take into account the last two rounds. 
+As our final extension, we consider a combination of the previous two settings, discussed in Sections~\ref{section:m_one_n_two} and \ref{section:m_two_n_one}. 
+Now, individuals take into account two interactions with different co-players, and from each interaction, they take into account the last two rounds. 
 The probability that the number of mutants increases or decreases by one is now given by
@@
 \end{equation}
- Though we do not carry any further analytical exploration of this case, in the
+ Though we do not pursue any further analytical exploration of this case, in the
 next section we present simulation results for this setting, as well as for the other settings.\\
@@
     \caption{\textbf{Evolutionary dynamics of our five different settings.}
-    From left to right, we present result on the following cases: individuals update their strategies based on
+    From left to right, we present results on the following cases: individuals update their strategies based on
     ({\it i})~the expected payoffs (perfect memory); ({\it ii}) the last-round payoff
@@
     interactions; ({\it iv})~the payoffs of the last two rounds of one interaction; 
-    and ({\it v} the payoff of the last two rounds of two interactions. We run each simulation for \(T
-    = 10^7\) time steps. For each time step we recorded the current resident
+    and ({\it v}) the payoff of the last two rounds of two interactions. We run each simulation for \(T
+    = 10^7\) time steps. For each time step we record the current resident
     population \((y, p, q)\). The graphs show how often the
@@
 {\bf Strategies favored by selection.} 
-To gain some insights into the evolving strategies in each of the settings described in the previous section, we simulated the evolutionary process according to Algorithm~\ref{algorithm:pairwise_comparison}.
-Each time step, we recorded the current resident population \((y, p, q)\). 
-The results of these simulations are shown in Supplementary Figure~\ref{fig:expected_payoffs_results}. 
+To gain some insights into the strategies evolving in each of the settings described in the previous section, we simulate the evolutionary process according to Algorithm~\ref{algorithm:pairwise_comparison}.
+Each time step, we record the current resident population \((y, p, q)\). 
+The results of these simulations are shown in Supplementary Fig.~\ref{fig:expected_payoffs_results}. 
 Qualitatively, all settings lead to similar dynamics as in the case of perfect and limited memory described in the main text. 
@@
     \caption{\textbf{Invasion analysis for perfect and limited payoff memory.} We consider two residents, \alld{}$=(0,0,0)$ and \gtft$=(1,1,1/3)$ for both perfect and limited payoff memory.
-In each case we simulate the pairwise comparison process 1,000 times and recorded the first mutant strategy that reached fixation (colored dots).
+In each case, we simulate the pairwise comparison process $10^{3}$ times and record the first mutant strategy that reaches fixation (colored dots).
 Black arrows indicate the arithmetic mean of the successful mutant strategies.
-In addition, we recorded how many mutant strategies it takes on average to successfully invade a given resident.
+In addition, we record how many mutant strategies it takes on average to successfully invade a given resident.
 We observe that under perfect memory, defectors are more prone to being invaded, whereas conditional cooperators are more resilient to invasion.
@@
 \noindent
-{\bf Invasion dynamics.} To gain some further into the dynamics over time, and into the differences between perfect and limited payoff memory, we have performed an invasion analysis, see Supplementary Figure~\ref{fig:invasion_analysis}.
+{\bf Invasion dynamics.} To gain some further into the dynamics over time, and into the differences between perfect and limited payoff memory, we perform an invasion analysis, see Supplementary Fig.~\ref{fig:invasion_analysis}.
 Here, we consider two possible resident strategies. 
-Either populations use \alld$=(0,0,0)$, top row, or they start out with \gtft$=(1,1,1/3)$, bottom row. 
-In each case, we simulated  the process until a mutant reached fixation, for both perfect memory (left) and limited memory (right). 
-We iterated this simulation 1,000 times. 
+Either populations use \alld$=(0,0,0)$ (top row) or they start out with \gtft$=(1,1,1/3)$ (bottom row). 
+In each case, we simulate the process until a mutant reaches fixation, for both perfect memory (left) and limited memory (right). 
+We iterate this simulation $10^{3}$ times. 
 Colored dots indicate the position of successful mutants, and arrows indicate the average trajectory of evolution. 
-Moreover, we also recorded how long it takes on average until the first mutant strategy takes over. 
+Moreover, we also record how long it takes on average until the first mutant strategy takes over. 
 This plot shows that \alld{} populations are typically invaded by mutants similar to Tit-for-Tat, with $p\!>\!0$ and $q\!\approx\!0$. 
-In contrast, \gtft{} populations are typically invaded by other \gtft{} mutants, but with higher generosity. 
+In contrast, \gtft{} populations are typically invaded by other \gtft{} mutants, but with higher levels of generosity. 
 A further analysis of the fixation time can shed some light on why perfect memory leads to higher average cooperation rates: 
@@
 %% SIMULATIONS: Impact of mutations %%
+
+\noindent
+{\bf Impact of mutations.} 
+All of our previous simulations are run under the assumption that mutations are rare,~$\mu\!\rightarrow\!0$. 
+This allows us to use the formula~\eqref{eq:fixation_probability} for a mutant's fixation probability. 
+While this limit allows for more efficient computation procedures, it rules out the possibility that two or three strategies might stably coexist in a population~\cite{tkadlec:pnas:2023}.
+To explore the effect of positive mutation rates, we perform five independent runs of the pairwise process described in
+Section~\ref{section:model} for different values of $\mu$.
+At each time step, we record the average cooperation rate of the resident population. 
+The results for different mutation rates are shown in Fig.~\ref{fig:mutation}. 
+The cooperation rate is higher when players have perfect memory (compared to limited memory), regardless of the
+mutation rate. 
+Only as the mutation rate approaches one, and the processes become fully random, does the cooperation rate approach $1/2$ in both simulations.
 
@@
 		We perform five independent simulations. Simulations are run
-  for $T\!=\!4\times 10^7$ time steps for each parameter. In each time step
+		for $T\!=\!4\times 10^7$ time steps for each parameter. In each time step,
 		we introduce a new mutant with a probability \(\mu\), and we then select
@@
 		adopts the strategy of the role model with a probability \(\rho(\pi_\learner, \pi_\rolemodel)\) where the
- payoffs depend on the considered setting. We plot the average cooperation rate
+		payoffs depend on the relevant setting. We plot the average cooperation rate
 		within the resident population for each value of \(\mu\). 
@@
 
-\noindent
-{\bf Impact of mutations.} 
-All of our previous simulations are run under the assumption that mutations are rare,~$\mu\!\rightarrow\!0$. 
-This allows us to use the formula~\eqref{eq:fixation_probability} for a mutant's fixation probability. 
-While this limit allows for more efficient computation procedures, it rules out the possibility that two or three strategies might stably coexist in a population~\cite{tkadlec:pnas:2023}.
-To explore the effect of positive mutation rates, we performed five independent runs of the pairwise process described in
-Section~\ref{section:model} for different values of $\mu$.
-At each time step we record the average cooperation of the resident population. 
-The results for different mutation rates are shown in Figure~\ref{fig:mutation}. 
-The cooperation rate is higher when players have perfect memory (compared to limited memory), regardless of the
-mutation rate. 
-Only as the mutation rate approaches one, and the processes become fully random, the cooperation rate approaches $1/2$ in both simulations.
->>>>>>> origin/dev
-
 \subsection{Simulations for memory-one strategies}\label{section:memory_one}
 
-So far we have assumed that individuals adopt reactive strategies.
+So far, we have assumed that individuals adopt reactive strategies.
 To demonstrate that our results hold for more complex strategy sets, 
@@
 To explore the dynamics among memory-one strategies, we perform four separate simulations, for both perfect and limited memory, and for two different benefit values. 
-Figure~\ref{fig:memory_one_low_benefit} shows the results for a comparably low benefit of cooperation; 
-Figure~\ref{fig:memory_one_high_benefit} shows the corresponding results for a large benefit. 
+Fig.~\ref{fig:memory_one_low_benefit} shows the results for a comparably low benefit of cooperation; 
+Fig.~\ref{fig:memory_one_high_benefit} shows the corresponding results for a large benefit. 
 These simulations suggest that even when individuals
@@
   \caption{\textbf{Evolution of memory-one strategies when the benefit of cooperation is comparably small.}
-  We perform two independent simulations. In one simulation individuals use
+  We perform two independent simulations. In one simulation, individuals use
   expected payoffs. 
-  In the other they update their strategy based on the payoff they received in the last round of their last interaction. 
+  In the other, they update their strategy based on the payoff they received in the last round of their last interaction. 
   We run each simulation for \(T = 10^8\) time steps.
-  For each time step, we have recorded the current resident population, who is
-  now of the form  \((y, p_1, p_2, p_3, p_4)\) In the left panel we report the
+  For each time step, we record the current resident population, which is
+  now of the form  \((y, p_1, p_2, p_3, p_4)\). In the left panel, we report the
   cooperation rates for each simulation. 
-  Also for memory-one strategies, the use of expected payoffs results in more cooperation. 
+  For memory-one strategies, like for reactive strategies, the use of expected payoffs results in more cooperation. 
   The right panel reports the most abundant strategy of each
@@
   \caption{\textbf{Evolution of memory-one strategies when the benefit is high.}
-  We consider a similar setup as in the previous figure, but now using a larger benefit value $b\!=\!10$.}\label{fig:memory_one_high_benefit}
+  We consider a similar setup as in Fig.~\ref{fig:memory_one_low_benefit}, but now using a larger benefit value $b\!=\!10$.}\label{fig:memory_one_high_benefit}
 \end{figure}

### 2023-11-01 paper/si.tex
wip on si

--- 2023-10-29 paper/si.tex
@@
 In the most general setting, the resulting one-shot payoffs can be summarized by the payoff matrix 
-\begin{align}
-	\bordermatrix{%
-		& C & D \cr
-		C &\ R &\ S\  \cr
-		D &\ T &\ P\ \cr
-	} . \label{eq:prisoner}
-\end{align}
+\begin{equation}\label{eq:prisoner}
+    \begin{blockarray}{ccc}
+        & \text{cooperate} & \text{defect} \\
+        \begin{block}{c(cc)}
+            \text{cooperate} & R & S \\
+            \text{defect} & T & P \\
+        \end{block}
+    \end{blockarray}.
+\end{equation}
 Here, $R$ is interpreted as the reward for mutual cooperation, $S$ is the sucker's payoff, $T$ is the temptation, and $P$ is the punishment payoff~\citep{axelrod1981evolution}. 
@@
 Using this distribution $\mathbf{v}$, we compute player 1's expected payoff as the weighted average
-\begin{equation}
+\begin{equation}\label{eq:expected_payoff}
 \pi(\strategy_1,\strategy_2) = R \, v_R + S\,v_S + T\,v_T + P\,v_P. 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% RESULTS: Remembering the last two rounds of one interaction %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Updates based on the realized payoff one repeated game}\label{section:one_interaction_full_memory}
+
+%% RESULTS: Two interactions, last round -- General motivation %%
+
+{\bf Motivation.}
+In the case of limited memory, we have thus far considered that individuals
+update their strategies based solely on the number of interactions and by
+considering only the payoff they received in those interactions in the final
+turns. \\
+
+\noindent
+A player's payoff in such an interaction is calculated based on the
+probabilities. In comparison, in the perfect memory case, individuals consider
+their expected payoff from the interaction. A natural question to ask is what
+happens if individuals update based on the expected payoffs of the interaction.
+
+\noindent
+{\bf Computing the ratio $\lambda^-_k/\lambda^+_k$.}
+We consider the population setup once more, consisting of \(k\) mutants with the
+strategy \(\strategy_\mutant=(y_\mutant,p_\mutant,q_\mutant)\), and \(N-k\)
+residents with strategies \(\strategy_\resident =
+(y_\resident,p_\resident,q_\mutant)\). In each step of the evolutionary process,
+we select a learner and a role model. The learner or the role model can be
+either a resident, and the other player is a mutant. As mentioned previously,
+two major cases must be considered:
+
+\begin{enumerate}
+  \item The learner and the role model have had their most recent interaction
+  with each other.
+  \item The learner's last interaction was not with the role model. In the
+  latter case, there are four different subcases based on whether the resident's
+  last interaction partner was a mutant or a resident, and whether the mutant's
+  last interaction partner was a mutant or a resident.
+\end{enumerate}
+
+\noindent
+In each pairing, we assume that the players interact for a specific number of
+rounds, and the payoff from their interaction is the average of the payoffs
+described in Eq.~\eqref{eq:expected_payoff}. As a result, there are four
+possible payoffs in our population setup: the payoffs an individual with a
+resident strategy can receive when interacting with an individual with the same
+strategy or with a mutant strategy, and similarly, the payoffs for the mutant
+strategy. \\
+
+\noindent
+After these preparations, we can define the probability that the number of
+mutants either increases or decreases by one.
+
+\begin{equation*}
+  \setlength{\arraycolsep}{1pt} 
+  \begin{array}{llll}
+    \lambda^+	 &=
+  &\frac{1}{N\!-\!1}\cdot  & \rho\left(\pi (\strategy_\resident, \strategy_\mutant), \pi (\strategy_\mutant, \strategy_\resident)\right)\\[0.5cm]
+  &+
+  &\frac{N\!-\!2}{N\!-\!1} \cdot 
+  &\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3}\cdot \rho(\pi (\strategy_\resident, \strategy_\mutant), \pi (\strategy_\mutant, \strategy_\mutant)) + 
+   \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3}\cdot \rho(\pi (\strategy_\resident, \strategy_\mutant), \pi (\strategy_\mutant, \strategy_\resident)) \right.\\[0.5cm]
+  &&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} \cdot \rho(\pi (\strategy_\resident, \strategy_\resident), \pi (\strategy_\mutant, \strategy_\mutant)) + 
+   \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} \cdot \rho(\pi (\strategy_\resident, \strategy_\resident), \pi (\strategy_\mutant, \strategy_\resident))\right],
+  \end{array}
+\end{equation*}
+
+\begin{equation*}
+  \setlength{\arraycolsep}{1pt} 
+  \begin{array}{llll}
+    \lambda^+	 &=
+  &\frac{1}{N\!-\!1}\cdot  & \rho\left(\pi (\strategy_\mutant, \strategy_\resident), \pi(\strategy_\resident, \strategy_\mutant)\right)\\[0.5cm]
+  &+
+  &\frac{N\!-\!2}{N\!-\!1} \cdot 
+  &\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3}\cdot \rho(\pi(\strategy_\mutant, \strategy_\mutant), \pi(\strategy_\resident, \strategy_\mutant)) + 
+   \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3}\cdot \rho(\pi(\strategy_\mutant, \strategy_\resident), \pi(\strategy_\resident, \strategy_\mutant)) \right.\\[0.5cm]
+  &&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} \cdot \rho(\pi(\strategy_\mutant, \strategy_\mutant), \pi(\strategy_\resident, \strategy_\resident)) + 
+   \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} \cdot \rho(\pi(\strategy_\mutant, \strategy_\resident), \pi(\strategy_\resident, \strategy_\resident),)\right].
+  \end{array}
+\end{equation*}
+
+\noindent
+{\bf Stochastic stability of cooperation.}
+We once again calculate how easily a single \alld{} mutant can invade into a
+resident population of \gtft{} players. When resident interacts with another
+resident or with a mutant,
+the expected payoffs based on Eq.~\eqref{eq:expected_payoff} of the interactions are,
+\begin{equation*}
+\begin{array}{ll}
+  \pi (\strategy_\resident, \strategy_\resident) =  b - c,
+  &\pi (\strategy_\resident, \strategy_\mutant) =  - c \left(\delta \left(q - 1\right) + 1\right).
+\end{array}
+\end{equation*}
+The expected payoff of the mutant, are
+\begin{align*}
+\begin{array}{ll}
+  \pi (\strategy_\mutant, \strategy_\resident) =  b \left(\delta \left(q - 1\right) + 1\right),
+  &\pi (\strategy_\mutant, \strategy_\mutant) =  0.
+\end{array}
+\end{align*}
+Based on these expected payoffs, we compute the ratio
+\begin{equation}
+  \frac{\lambda^{-}_1}{\lambda^{+}_1} =
+  \frac{\frac{1}{N-1} + \frac{N-2}{N-1}}{\frac{1}{N-1} + \frac{N-2}{N-1}}
+\end{equation}
+
+\begin{equation}
+  \frac{\lambda^{-}_1}{\lambda^{+}_1} =
+  \frac{1 + e^{- \beta \left(b \left(\delta q - \delta\right) + c\right)}}{1 + e^{\beta \left(b \left(\delta q - \delta\right) + c\right)}}.
+\end{equation}
+For stochastic stability of cooperation we require this ratio to be larger than one, which is equivalent to
+\begin{equation} \label{Eq:Condition_OneInteractionExpectedPayoff}
+  q<1 - \frac{c}{\delta b}
+\end{equation}
+Interestingly, this conditions is the same as the condition for stochastic
+stability in the case of perfect memory, see
+Eq.~\eqref{Eq:Condition_PerfectMemory}.
+
 \clearpage
@@
 Only as the mutation rate approaches one, and the processes become fully random, the cooperation rate approaches $1/2$ in both simulations.
->>>>>>> origin/dev
 
@@
 
+\begin{figure}[!htbp]
+  \centering 
+  \includegraphics[width=\textwidth]{static/cooperation_rate_over_b_and_beta_memory_one.pdf}
+  \caption{{\bf Evolution of direct reciprocity for different parameter values.}}
+\end{figure}
+
 %%%%%%%%%%%%
@@
 \bibliographystyle{naturemag}
-\bibliography{bibliography}
+\bibliography{bibliography.bib}
 }

### 2023-11-01 paper/si.tex

--- 2023-10-31 paper/si.tex
@@
 Using this distribution $\mathbf{v}$, we compute the first player's expected payoff as the weighted average
-\begin{equation}
+\begin{equation}\label{eq:expected_payoff}
 \pi(\strategy_1,\strategy_2) = R \, v_R + S\,v_S + T\,v_T + P\,v_P. 

### 2023-11-01 paper/si.tex
include expected payoff of one interaction in si

@@
 In the following, we discuss our different memory settings in more detail. 
-We discuss five cases explicitly. In these cases, updating occurs
+We discuss six cases explicitly. In these cases, updating occurs
 ({\it i})~based on average payoffs based on all interactions (perfect memory), 
@@
 ({\it iii})~based on the last round of two interactions,
-({\it iv})~based on the last two rounds of one interaction, and 
-({\it v})~based on the last two rounds of two interactions. 
+({\it iv})~based on the last two rounds of one interaction,
+({\it v})~based on the last two rounds of two interactions, and
+({\it v})~based on the expected payoff of one interaction.
 In each case, we consider the case that there are only two strategies present in the population (a resident and a mutant strategy). 
@@
 By definition, cooperation is stochastically stable if this ratio exceeds one, which is equivalent to
-\begin{equation*}
+\begin{equation}\label{Eq:Condition_PerfectMemory}
 q<1-\frac{1}{\delta}\cdot \frac{b+(N\!-\!1)c}{(N\!-\!1)b+c}.
-\end{equation*}
+\end{equation}
 For such a strategy to be feasible we require $q\!>\!0$, which implies $\delta \!>\! \big(b+(N\!-\!1)c\big)/\big((N\!-\!1)b\!+\!c\big)$.
@@
 
-\subsection{Updates based on the realized payoff one repeated game}\label{section:one_interaction_full_memory}
+\subsection{Updates based on expected payoff of one repeated game}\label{section:one_interaction_full_memory}
 
@@
 {\bf Motivation.}
-In the case of limited memory, we have thus far considered that individuals
-update their strategies based solely on the number of interactions and by
-considering only the payoff they received in those interactions in the final
-turns. \\
-
-\noindent
-A player's payoff in such an interaction is calculated based on the
-probabilities. In comparison, in the perfect memory case, individuals consider
-their expected payoff from the interaction. A natural question to ask is what
-happens if individuals update based on the expected payoffs of the interaction.
 
@@
 {\bf Computing the ratio $\lambda^-_k/\lambda^+_k$.}
-We consider the population setup once more, consisting of \(k\) mutants with the
-strategy \(\strategy_\mutant=(y_\mutant,p_\mutant,q_\mutant)\), and \(N-k\)
-residents with strategies \(\strategy_\resident =
-(y_\resident,p_\resident,q_\mutant)\). In each step of the evolutionary process,
-we select a learner and a role model. The learner or the role model can be
-either a resident, and the other player is a mutant. As mentioned previously,
+We revisit the population setup, consisting of \(k\) mutants with the strategy
+\(\strategy_\mutant = (y_\mutant, p_\mutant, q_\mutant)\) and \(N-k\) residents
+with strategies \(\strategy_\resident = (y_\resident, p_\resident,
+q_\resident)\). In each step of the evolutionary process, we select a learner and
+a role model. The learner or the role model can be either a resident, and the
+other player is a mutant. We consider that the learner and the
+model update based on their most recent interaction. As mentioned previously,
 two major cases must be considered:
@@
 \begin{enumerate}
-  \item The learner and the role model have had their most recent interaction
-  with each other.
-  \item The learner's last interaction was not with the role model. In the
-  latter case, there are four different subcases based on whether the resident's
-  last interaction partner was a mutant or a resident, and whether the mutant's
-  last interaction partner was a mutant or a resident.
+  \item The learner and the role model have had their most recent interaction with each other.
+  \item The learner's last interaction was not with the role model.
 \end{enumerate}
 
-\noindent
-In each pairing, we assume that the players interact for a specific number of
-rounds, and the payoff from their interaction is the average of the payoffs
-described in Eq.~\eqref{eq:expected_payoff}. As a result, there are four
-possible payoffs in our population setup: the payoffs an individual with a
-resident strategy can receive when interacting with an individual with the same
-strategy or with a mutant strategy, and similarly, the payoffs for the mutant
-strategy. \\
+In the latter case, there are four different subcases based on whether the
+resident's last interaction partner was a mutant or a resident and whether the
+mutant's last interaction partner was a mutant or a resident.
+
+We assume that the players' payoff received from their last interaction is their
+expected payoff given by Eq.~\eqref{eq:expected_payoff}. Thus, a resident
+interacting with a resident receives the payoff $\pi(s_\resident, s_\resident)$,
+and $\pi(s_\resident, s_\mutant)$ when interacting with a mutant. Similarly, a
+mutant interacting with a resident receives the payoff $\pi(s_\mutant,
+s_\resident)$ and $\pi(s_\mutant, s_\mutant)$ when interacting with a mutant.\\
 
@@
   \begin{array}{llll}
-    \lambda^+	 &=
+    \lambda^-	 &=
   &\frac{1}{N\!-\!1}\cdot  & \rho\left(\pi (\strategy_\mutant, \strategy_\resident), \pi(\strategy_\resident, \strategy_\mutant)\right)\\[0.5cm]
@@
   \frac{\lambda^{-}_1}{\lambda^{+}_1} =
-  \frac{\frac{1}{N-1} + \frac{N-2}{N-1}}{\frac{1}{N-1} + \frac{N-2}{N-1}}
-\end{equation}
+  \frac{\frac{1}{N-1}\frac{1}{\left(1 + e^{\beta \left(b \left(\delta q - \delta\right) + c\right)}\right)}
+  + \frac{N - 2}{N-1}\frac{1}{\left(1 + e^{- \beta \left(- b \left(\delta q - \delta + 1\right) + b - c\right)}\right)}}
+  {\frac{1}{N-1}\frac{1}{\left(1 + e^{- \beta \left(b \left(\delta q - \delta\right) + c\right)}\right)}
+  + \frac{N - 2}{N-1}\frac{1}{\left(1 + e^{- \beta \left(b \left(\delta q - \delta + 1\right) - b + c\right)}\right)}}.
+\end{equation}
+
+\noindent
+In the limit of large populations \(N \rightarrow \infty \), this ratio simplifies to
 
@@
     \includegraphics[width=\textwidth]{static/more_memory_heatmaps_donation_game_with_illustrations.pdf}
-    \caption{\textbf{Evolutionary dynamics of our five different settings.}
+    \caption{\textbf{Evolutionary dynamics of our six different settings.}
     From left to right, we present results on the following cases: individuals update their strategies based on
@@
     interactions; ({\it iv})~the payoffs of the last two rounds of one interaction; 
-    and ({\it v}) the payoff of the last two rounds of two interactions. We run each simulation for \(T
-    = 10^7\) time steps. For each time step we record the current resident
+    ({\it v}) the payoff of the last two rounds of two interactions and
+    ({\it vi}) the expected payoff of one interaction.
+    Simulations are run either for a comparably high benefit of cooperation~($b/c\!=\!10$), or for a low benefit~($b/c\!=\!3$).
+    We run each simulation for \(T= 10^7\) time steps. For each time step we record the current resident
     population \((y, p, q)\). The graphs show how often the
@@
 In all settings, we observe that along this strip, strategies are most abundant when they satisfy the condition for stochastic stability. 
-For perfect memory, this means the resident population predominantly adopts a strategy with \(q\! \leq\! 1 \!-\!
-c/b\!=\!0.9\) (for the parameters used in the simulation). 
-For limited memory, the generosity of a conditional cooperator satisfies  \(q\!<\!\frac{1}{2}\). 
-In the next two cases, when players either remember the last round of two interactions, or the last two rounds of one interaction, stochastic stability requires $q\!<\!\sqrt{2}{2}\!\approx\!0.701$. 
+For perfect memory, this means the resident population predominantly adopts a
+strategy with \(q \leq 1 - \frac{c}{b} = 0.9\) for high benefit and \(q \leq 1 -
+\frac{c}{b} = 0.67\) for low benefit.
+For limited memory, the generosity of a conditional cooperator satisfies \(q <
+\frac{1}{2}\) for both values of benefit.
+In the next two cases, when players either remember the last round of two
+interactions or the last two rounds of one interaction, stochastic stability
+requires \(q < \frac{\sqrt{2}}{2} \approx 0.701\).
+In the final case, we observe that the generosity of a conditional cooperator is
+in line with the condition for stochastic stability in the perfect memory case:
+\(q < 1 - \frac{c}{b} = 0.9\) for high benefit and \(q < 1 - \frac{c}{b} =
+0.67\) for low benefit.
 In all cases, the simulation results are well aligned with these theoretical predictions.\\
@@
 		While limited memory allows the evolution of moderate cooperation levels, perfect memory is always more conducive to the evolution of cooperation. 
-		Parameters are the same as before, \(N =100, c=1, b=10, \beta=1\).}\label{fig:mutation}
+		Parameters are the same as before, \(N =100, c=1, b=10, \beta=1, \delta=0.999\).}\label{fig:mutation}
 \end{figure}
@@
 are allowed to use memory-one strategies, there is more cooperation when players update their strategies based on perfect payoff memory.
+
+To further investigate the influence of different parameters, we have
+systematically varied the benefit~$b$ and the selection strength~$\beta$, as
+shown in Fig.~\ref{fig:cooperation_rate_over_benefit_and_beta_memory_one}.
+When compared to the case of reactive strategies, what we observe
+here is that in the case of limited payoff memory, cooperation rates increase
+with an increasing benefit~$b$ and do not remain stable
+(Fig.~\ref{fig:cooperation_rate_over_benefit_and_beta_memory_one}(a)). However,
+even in this scenario, we still observe that perfect memory consistently results
+in a higher cooperation rate.
+Regarding the effect of different selection strengths, as depicted in
+Fig.~\ref{fig:cooperation_rate_over_benefit_and_beta_memory_one}(b), the results
+once again suggest that both perfect and limited payoff memory yield similar
+cooperation rates for weak selection (\(\beta \!<\! 1\)). Beyond weak selection, increasing
+selection has a positive effect under perfect payoff memory but a negative
+effect under limited payoff memory.
 
@@
   namely, \(p_1 \approx 1\), \(p_3 \approx 0\) and \(p_4 > \frac{1}{2}\). 
-  Parameters: \(N =100, c=1, b=3, \beta=1\).}\label{fig:memory_one_low_benefit}
+  Parameters: \(N =100, c=1, b=3, \beta=1, \delta=0.999\).}\label{fig:memory_one_low_benefit}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{static/cooperation_rate_over_b_and_beta_memory_one.pdf}
-  \caption{{\bf Evolution of direct reciprocity for different parameter values.}}
+  \caption{{\bf Evolution of direct reciprocity for different parameter values for memory-one strategies.}
+  To further explore the robustness of our results in the case of memory-one
+  strategies, we run simulations for different benefit values (left panels, A)
+  and for different selection strengths (right panels, B). These simulations
+  suggest that perfect payoff memory consistently leads to more cooperation and
+  more generosity. Unless explicitly varied, the parameters of the simulation
+  are $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.999$.
+  Simulations are run for $T\!=\!5\times 10^7$ time steps for each parameter
+  combination.}
+  \label{fig:cooperation_rate_over_benefit_and_beta_memory_one}
 \end{figure}

### 2023-11-01 paper/si.tex
swaps the notation for the imitation probability and fixation probability

@@
 In that case, the player randomly picks another individual from the population (referred to as the `role model', \rolemodel). 
-The learner adopts the role model's strategy with a probability \(\rho\), given by
+The learner adopts the role model's strategy with a probability \(\varphi\), given by
 \begin{equation} \label{Eq:rho}
-    \rho(\pi_{\learner}, \pi_{\rolemodel}) = \frac{1}{1\!+\! e^{\!-\!\beta (\pi_\rolemodel - \pi_\learner)}}.
+    \varphi(\pi_{\learner}, \pi_{\rolemodel}) = \frac{1}{1\!+\! e^{\!-\!\beta (\pi_\rolemodel - \pi_\learner)}}.
 \end{equation}
@@
 Then one individual adopts a mutant strategy selected uniformly at random from the set of feasible strategies.
-The fixation probability \(\varphi_{\mutant}\) of the mutant strategy can be calculated explicitly~\citep{nowak:Nature:2004},
+The fixation probability \(\rho_{\mutant}\) of the mutant strategy can be calculated explicitly~\citep{nowak:Nature:2004},
 \begin{equation}\label{eq:fixation_probability}
-    \varphi_{\mutant} = \frac{1}{1+\sum\limits_{i=1}^{N-1}\prod\limits_{k=1}^i \frac{\lambda^-_k}{\lambda^+_k}}.
+    \rho_{\mutant} = \frac{1}{1+\sum\limits_{i=1}^{N-1}\prod\limits_{k=1}^i \frac{\lambda^-_k}{\lambda^+_k}}.
 \end{equation}
@@
 These probabilities depend on the probability that a mutant and a resident are chosen as the learner and the resident, respectively. 
-In addition, they depend on the respective switching probability~$\rho$, as described by Eq.~\eqref{Eq:rho}. 
+In addition, they depend on the respective switching probability~$\varphi$, as described by Eq.~\eqref{Eq:rho}. 
 We specify the exact values of  \(\lambda^-_k, \lambda^+_k\) for each memory-setting in the next section.
 
-Depending on the fixation probability \(\varphi_{\mutant}\), the mutant strategy either fixes (becomes the new resident) or goes extinct. 
+Depending on the fixation probability \(\rho_{\mutant}\), the mutant strategy either fixes (becomes the new resident) or goes extinct. 
 Afterwards, another random mutant strategy is introduced into the population. 
@@
    {mutant $\leftarrow$ random strategy\;
-   fixation probability $\leftarrow \varphi_\mutant $\;
-   \If{$\varphi_{\mutant} >$ random: $i \rightarrow [0,1]$}
+   fixation probability $\leftarrow \rho_\mutant $\;
+   \If{$\rho_{\mutant} >$ random: $i \rightarrow [0,1]$}
    {resident $\leftarrow$ mutant;}}
@@
   \lambda^-_k \!=\!
-\frac{N-k}{N}\frac{k}{N-1}\rho\big(\pi_\mutant(k),\,\pi_\resident(k)\big) \quad \text{ and } \quad \lambda^+_k \!=\!\frac{k}{N}\frac{N-k}{N-1}\rho\big(\pi_\resident(k),\,\pi_\mutant(k)\big).
+\frac{N-k}{N}\frac{k}{N-1}\varphi\big(\pi_\mutant(k),\,\pi_\resident(k)\big) \quad \text{ and } \quad \lambda^+_k \!=\!\frac{k}{N}\frac{N-k}{N-1}\varphi\big(\pi_\resident(k),\,\pi_\mutant(k)\big).
 \end{align*}
-For $\rho$ as defined by Eq.~\eqref{Eq:rho}, the ratio of these two transition probabilities simplifies to 
+For $\varphi$ as defined by Eq.~\eqref{Eq:rho}, the ratio of these two transition probabilities simplifies to 
 \begin{equation} \label{eq:LambdaRatio}
- \frac{\lambda^-_k}{\lambda^+_k} \!=\!  \frac{\rho\big(\,\pi_\mutant(k),~\pi_\resident(k)\,\big)}{\rho\big(\,\pi_\resident(k),~\pi_\mutant(k)\,\big)} 
+ \frac{\lambda^-_k}{\lambda^+_k} \!=\!  \frac{\varphi\big(\,\pi_\mutant(k),~\pi_\resident(k)\,\big)}{\varphi\big(\,\pi_\resident(k),~\pi_\mutant(k)\,\big)} 
  =e^{-\beta\big(\pi_\mutant(k)-\pi_\resident(k)\big)}.
@@
 \begin{array}{l}
-\displaystyle \lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_\resident,u_\mutant\,\in\,\mathcal{U}} x(u_\resident,u_\mutant)\cdot \rho(u_\resident,u_\mutant), \\[0.5cm]
-\displaystyle \lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_\resident,u_\mutant \,\in\, \mathcal{U}} x(u_\resident,u_\mutant)\cdot \rho(u_\mutant,u_\resident).
+\displaystyle \lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_\resident,u_\mutant\,\in\,\mathcal{U}} x(u_\resident,u_\mutant)\cdot \varphi(u_\resident,u_\mutant), \\[0.5cm]
+\displaystyle \lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_\resident,u_\mutant \,\in\, \mathcal{U}} x(u_\resident,u_\mutant)\cdot \varphi(u_\mutant,u_\resident).
 \end{array}
@@
 \begin{array}{l}
-\displaystyle \lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} \sum_{u_{\resident 1},u_{\resident 2},u_{\mutant 1},u_{\mutant 2}} \hspace{-0.1cm} x(u_{\resident 1}u_{\resident 2}, u_{\mutant 1}u_{\mutant 2})\cdot \rho\left(\frac{u_{\resident 1} \!+\!
+\displaystyle \lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} \sum_{u_{\resident 1},u_{\resident 2},u_{\mutant 1},u_{\mutant 2}} \hspace{-0.1cm} x(u_{\resident 1}u_{\resident 2}, u_{\mutant 1}u_{\mutant 2})\cdot \varphi\left(\frac{u_{\resident 1} \!+\!
 u_{\resident 2}}{2}, \frac{u_{\mutant 1} \!+\!
 u_{\mutant 2}}{2}\right), \\[0.5cm]
-\displaystyle \lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} \sum_{u_{\resident 1},u_{\resident 2},u_{\mutant 1},u_{\mutant 2}} \hspace{-0.1cm} x(u_{\resident 1}u_{\resident 2},u_{\mutant 1}u_{\mutant 2})\cdot \rho\left(\frac{u_{\mutant 1} \!+\!
+\displaystyle \lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} \sum_{u_{\resident 1},u_{\resident 2},u_{\mutant 1},u_{\mutant 2}} \hspace{-0.1cm} x(u_{\resident 1}u_{\resident 2},u_{\mutant 1}u_{\mutant 2})\cdot \varphi\left(\frac{u_{\mutant 1} \!+\!
 u_{\mutant 2}}{2}, \frac{u_{\resident 1} \!+\!
@@
 \begin{array}{l}
-\displaystyle \lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} \sum_{u_{\resident},u_\resident',u_{\mutant},u_\mutant'} \hspace{-0.1cm} x(u_\resident u_\resident', u_\mutant u_\mutant')\cdot \rho\left(\frac{u_\resident \!+\!
+\displaystyle \lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} \sum_{u_{\resident},u_\resident',u_{\mutant},u_\mutant'} \hspace{-0.1cm} x(u_\resident u_\resident', u_\mutant u_\mutant')\cdot \varphi\left(\frac{u_\resident \!+\!
 u_\resident'}{2}, \frac{u_\mutant \!+\!
 u_\mutant '}{2}\right), \\[0.5cm]
-\displaystyle \lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} \sum_{u_{\resident},u_\resident',u_{\mutant},u_\mutant'} \hspace{-0.1cm} x(u_\resident u_\resident', u_\mutant u_\mutant')\cdot \rho\left( \frac{u_\mutant \!+\!
+\displaystyle \lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N} \sum_{u_{\resident},u_\resident',u_{\mutant},u_\mutant'} \hspace{-0.1cm} x(u_\resident u_\resident', u_\mutant u_\mutant')\cdot \varphi\left( \frac{u_\mutant \!+\!
 u_\mutant '}{2}, \frac{u_\resident \!+\!
@@
 \hspace{-1cm} x(u_{\resident 1} u_{\resident 1}'u_{\resident 2} u_{\resident 2}' ,u_{\mutant 1}u_{\mutant 1}'u_{\mutant 2}u_{\mutant 2}')\cdot 
-\rho\!\left(\frac{u_{\resident 1}\!+\! u_{\resident 1}'\!+\!u_{\resident 2}\!+\!u_{\resident 2}' }{4},
+\varphi\!\left(\frac{u_{\resident 1}\!+\! u_{\resident 1}'\!+\!u_{\resident 2}\!+\!u_{\resident 2}' }{4},
 \frac{u_{\mutant 1}\!+\!u_{\mutant 1}'\!+\!u_{\mutant 2}\!+\!u_{\mutant 2}'}{4}\!\right), \\[1.4cm]
@@
 \hspace{-1cm} x(u_{\resident 1} u_{\resident 1}'u_{\resident 2} u_{\resident 2}' ,u_{\mutant 1}u_{\mutant 1}'u_{\mutant 2}u_{\mutant 2}')\cdot 
-\rho\!\left(\frac{u_{\mutant 1}\!+\!u_{\mutant 1}'\!+\!u_{\mutant 2}\!+\!u_{\mutant 2}'}{4},
+\varphi\!\left(\frac{u_{\mutant 1}\!+\!u_{\mutant 1}'\!+\!u_{\mutant 2}\!+\!u_{\mutant 2}'}{4},
 \frac{u_{\resident 1}\!+\! u_{\resident 1}'\!+\!u_{\resident 2}\!+\!u_{\resident 2}' }{4}\!\right).
@@
     \lambda^+	 &=
-  &\frac{1}{N\!-\!1}\cdot  & \rho\left(\pi (\strategy_\resident, \strategy_\mutant), \pi (\strategy_\mutant, \strategy_\resident)\right)\\[0.5cm]
+  &\frac{1}{N\!-\!1}\cdot  & \varphi\left(\pi (\strategy_\resident, \strategy_\mutant), \pi (\strategy_\mutant, \strategy_\resident)\right)\\[0.5cm]
   &+
   &\frac{N\!-\!2}{N\!-\!1} \cdot 
-  &\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3}\cdot \rho(\pi (\strategy_\resident, \strategy_\mutant), \pi (\strategy_\mutant, \strategy_\mutant)) + 
-   \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3}\cdot \rho(\pi (\strategy_\resident, \strategy_\mutant), \pi (\strategy_\mutant, \strategy_\resident)) \right.\\[0.5cm]
-  &&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} \cdot \rho(\pi (\strategy_\resident, \strategy_\resident), \pi (\strategy_\mutant, \strategy_\mutant)) + 
-   \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} \cdot \rho(\pi (\strategy_\resident, \strategy_\resident), \pi (\strategy_\mutant, \strategy_\resident))\right],
+  &\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3}\cdot \varphi(\pi (\strategy_\resident, \strategy_\mutant), \pi (\strategy_\mutant, \strategy_\mutant)) + 
+   \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3}\cdot \varphi(\pi (\strategy_\resident, \strategy_\mutant), \pi (\strategy_\mutant, \strategy_\resident)) \right.\\[0.5cm]
+  &&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} \cdot \varphi(\pi (\strategy_\resident, \strategy_\resident), \pi (\strategy_\mutant, \strategy_\mutant)) + 
+   \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} \cdot \varphi(\pi (\strategy_\resident, \strategy_\resident), \pi (\strategy_\mutant, \strategy_\resident))\right],
   \end{array}
@@
     \lambda^-	 &=
-  &\frac{1}{N\!-\!1}\cdot  & \rho\left(\pi (\strategy_\mutant, \strategy_\resident), \pi(\strategy_\resident, \strategy_\mutant)\right)\\[0.5cm]
+  &\frac{1}{N\!-\!1}\cdot  & \varphi\left(\pi (\strategy_\mutant, \strategy_\resident), \pi(\strategy_\resident, \strategy_\mutant)\right)\\[0.5cm]
   &+
   &\frac{N\!-\!2}{N\!-\!1} \cdot 
-  &\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3}\cdot \rho(\pi(\strategy_\mutant, \strategy_\mutant), \pi(\strategy_\resident, \strategy_\mutant)) + 
-   \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3}\cdot \rho(\pi(\strategy_\mutant, \strategy_\resident), \pi(\strategy_\resident, \strategy_\mutant)) \right.\\[0.5cm]
-  &&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} \cdot \rho(\pi(\strategy_\mutant, \strategy_\mutant), \pi(\strategy_\resident, \strategy_\resident)) + 
-   \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} \cdot \rho(\pi(\strategy_\mutant, \strategy_\resident), \pi(\strategy_\resident, \strategy_\resident),)\right].
+  &\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3}\cdot \varphi(\pi(\strategy_\mutant, \strategy_\mutant), \pi(\strategy_\resident, \strategy_\mutant)) + 
+   \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3}\cdot \varphi(\pi(\strategy_\mutant, \strategy_\resident), \pi(\strategy_\resident, \strategy_\mutant)) \right.\\[0.5cm]
+  &&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} \cdot \varphi(\pi(\strategy_\mutant, \strategy_\mutant), \pi(\strategy_\resident, \strategy_\resident)) + 
+   \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} \cdot \varphi(\pi(\strategy_\mutant, \strategy_\resident), \pi(\strategy_\resident, \strategy_\resident),)\right].
   \end{array}
@@
 		two random players to serve as the role model and the learner. The learner
-		adopts the strategy of the role model with a probability \(\rho(\pi_\learner, \pi_\rolemodel)\) where the
+		adopts the strategy of the role model with a probability \(\varphi(\pi_\learner, \pi_\rolemodel)\) where the
 		payoffs depend on the relevant setting. We plot the average cooperation rate

### 2023-11-01 paper/main.tex
swaps the notation for the imitation probability and fixation probability
first round of corrections - main paper
Tweak symbols to align with SI, introduce the sixth case and
change referencing to figures

@@
 \usepackage{hyperref}
+\usepackage[all]{hypcap}
 \usepackage{tabularx}
@@
 \def\gtft{\texttt{GTFT}}
+\def\rolemodel{\text{RM}}
+\def\learner{\text{L}}
+\def\resident{\texttt{R}}
+\def\mutant{\texttt{M}}
+\def\strategy{s}
 \def\esm{electronic supplementary material}
@@
 %% Adding shortcut commands to refer to our figures %%
-\newcommand{\FigBaseResults}{{\bf Fig.~1}} 
-\newcommand{\FigDependenceParameters}{{\bf Fig.~2}} 
-\newcommand{\FigHigherMemory}{{\bf Fig.~3}}
+\newcommand{\FigBaseResults}{figure~\ref{fig:expected_and_stochastic_for_donation}}
+\newcommand{\FigDependenceParameters}{figure~\ref{fig:cooperation_rate_over_benefit_and_beta}} 
+\newcommand{\FigHigherMemory}{figure~\ref{fig:cooperation_rate_all_updating_payoffs}}
 
@@
 With the converse probability $1\!-\!\mu$, the learner randomly picks a `role model' from the population. 
-The learner then compares its own payoff $\pi_L$ from the repeated game to the role model's payoff $\pi_R$. 
-The learner adopts the role model's strategy with a probability \(\rho\) described by a Fermi function~\citep{blume:GEB:1995,szabo:PRE:1998}, 
+The learner then compares its own payoff $\pi_\learner$ from the repeated game to the role model's payoff $\pi_\rolemodel$. 
+The learner adopts the role model's strategy with a probability \(\varphi\) described by a Fermi function~\citep{blume:GEB:1995,szabo:PRE:1998}, 
 \begin{equation} \label{Eq:rho}
-    \rho\left(\pi_{L}, \pi_{R}\right) = \frac{1}{1\!+\! e^{- \!\beta\left(\pi_{R}- \pi_{L}\right)}}.
+    \varphi\left(\pi_\learner, \pi_\rolemodel\right) = \frac{1}{1\!+\! e^{- \!\beta\left(\pi_\rolemodel- \pi_\learner\right)}}.
 \end{equation}
@@
 
-We deviate from previous models in how we interpret the payoffs $\pi_L$ and $\pi_R$, which form the basis of the pairwise comparisons in Eq.~\eqref{Eq:rho}. 
+We deviate from previous models in how we interpret the payoffs $\pi_\learner$ and $\pi_\rolemodel$, which form the basis of the pairwise comparisons in Eq.~\eqref{Eq:rho}. 
 In previous work, these payoffs are taken to be the respective players' expected payoffs. 
 We interpret that setup as a model with perfect payoff memory. 
-There, the payoffs  $\pi_L$ and $\pi_R$ represent an average over all possible repeated games the two individuals have played with all population members (\FigBaseResults, upper left panel). 
+There, the payoffs  $\pi_\learner$ and $\pi_\rolemodel$ represent an average over all possible repeated games the two individuals have played with all population members (\FigBaseResults, upper left panel).
 The use of expected payoffs is mathematically convenient, because explicit formulas for these payoffs are available~\citep{hofbauer1998evolutionary}.
 Herein, we compare this model of perfect payoff memory to a model with limited payoff memory. 
-In that model, the players' payoffs $\pi_L$ and $\pi_R$ are taken to be the payoffs that each player received in their very last round prior to making social comparisons. 
+In that model, the players' payoffs $\pi_\learner$ and $\pi_\rolemodel$ are taken to be the payoffs that each player received in their very last round prior to making social comparisons. 
 That is, players only consider the very last repeated game they participated in, and there they only take into account the outcome of the very last round (\FigBaseResults, lower left panel). 
@@
 To further investigate the influence of different parameters, we have systematically varied the benefit~$b$ and the selection strength~$\beta$ in Figure~\FigDependenceParameters.
-According to \FigDependenceParameters \textbf{A}, perfect memory consistently results in a higher cooperation rate, and this relative advantage further increases with an increasing benefit~$b$. 
+According to \FigDependenceParameters{\it a}, perfect memory consistently results in a higher cooperation rate, and this relative advantage further increases with an increasing benefit~$b$. 
 Interestingly, for limited payoff memory, the cooperation rate remains stable at approximately 50\% once \(b \!\ge\! 5\).
 This again reflects our earlier observation that the feasibility of cooperation in this scenario is independent of the exact values of $b$ and $c$, as described by Eq.~\eqref{Eq:LimitedMemory}. 
-With respect to the effect of different selection strengths, \FigDependenceParameters\textbf{B} suggests that both perfect and limited payoff memory yields similar cooperation rates for weak selection \(\beta \!<\!
+With respect to the effect of different selection strengths, \FigDependenceParameters{\it b} suggests that both perfect and limited payoff memory yields similar cooperation rates for weak selection \(\beta \!<\!
 1\).
@@
     \caption{{\bf Evolution of direct reciprocity for different parameter values.} 
-    To explore the robustness of our results, we have run simulations for different benefit values (left panels, A), and for different selection strengths (right panels, B). 
+    To explore the robustness of our results, we have run simulations for different benefit values (left panels, {\it a}), and for different selection strengths (right panels, {\it b}). 
     In each case, we record the resulting average cooperation rate over the entire simulation (upper panels). 
@@
   ({\it iii}) based on the last round of two interactions,
-  ({\it iv}) based on the last two rounds of one interaction, and 
-  ({\it v}) based on the last two rounds of two interactions. 
+  ({\it iv}) based on the last two rounds of one interaction,
+  ({\it v}) based on the last two rounds of two interactions, and
+  ({\it vi}) based on the expected payoff of one interaction.
   Again, simulations are run either for a comparably high benefit of cooperation~($b/c\!=\!10$), or for a low benefit~($b/c\!=\!3$). 
   We observe that perfect memory always yields the highest cooperation rate. 
-  However, when individuals take into account at least two past interactions -- cases ({\it iii}) to ({\it v}) -- evolving cooperation rates are close to this optimum. 
+  However, when individuals take into account at least two past interactions -- cases ({\it iii}) to ({\it vi}) -- evolving cooperation rates are close to this optimum. 
   Baseline parameters are the same as in \FigDependenceParameters.}
@@
 \begin{equation}\label{eq:appendix_fixation_probability}
-    \varphi_{M} = \frac{1}{1+\sum\limits_{i=1}^{N-1}\prod\limits_k^i \frac{\lambda^-_k}{\lambda^+_k}},
+    \rho_{M} = \frac{1}{1+\sum\limits_{i=1}^{N-1}\prod\limits_k^i \frac{\lambda^-_k}{\lambda^+_k}},
 \end{equation}
@@
 Let \(s_M =(y_M, p_M, q_M)\) and \(s_R = (y_R, p_R, q_R)\) denote the
-strategies of a mutant and a resident, the expected payoffs \(\pi_R\) and
+strategies of a mutant and a resident, the expected payoffs \(\pi_\rolemodel\) and
 \(\pi_M\) are given by,
@@
   \begin{array}{lcrcr}
-  \displaystyle \pi_R & = &\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(s_R,s_R),\mathbf{u}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(s_R,s_M),\mathbf{u}\rangle,\\[0.5cm]
+  \displaystyle \pi_\rolemodel & = &\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(s_R,s_R),\mathbf{u}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(s_R,s_M),\mathbf{u}\rangle,\\[0.5cm]
   \displaystyle \pi_M & = &\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(s_M,s_R),\mathbf{u}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(s_M,s_M),\mathbf{u}\rangle.\\
@@
 \begin{align}\label{eq:perfect_memory_lambdas}
-  \lambda^-_k \!=\!\rho(\pi_M, \pi_R) \quad \text{ and } \quad \lambda^+_k \!=\!\rho(\pi_R, \pi_M).
+  \lambda^-_k \!=\!\varphi(\pi_M, \pi_\rolemodel) \quad \text{ and } \quad \lambda^+_k \!=\!\varphi(\pi_\rolemodel, \pi_M).
 \end{align}
@@
 \begin{align}\label{eq:limited_memory_lambdas}
-\lambda^+_k=\frac{N\!-\!k}{N} \cdot \frac{k}{N} \cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M}) \rho(u_{R},u_{M}) \quad \text{and} \quad
-\lambda^-_k=\frac{N\!-\!k}{N} \cdot \frac{k}{N} \cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M}) \rho(u_{M},u_{R}).
+\lambda^+_k=\frac{N\!-\!k}{N} \cdot \frac{k}{N} \cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M}) \varphi(u_{R},u_{M}) \quad \text{and} \quad
+\lambda^-_k=\frac{N\!-\!k}{N} \cdot \frac{k}{N} \cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M}) \varphi(u_{M},u_{R}).
 \end{align}

### 2023-11-01 paper/si.tex
use \strategy in SI

@@
 expected payoff given by Eq.~\eqref{eq:expected_payoff}. Thus, a resident
-interacting with a resident receives the payoff $\pi(s_\resident, s_\resident)$,
-and $\pi(s_\resident, s_\mutant)$ when interacting with a mutant. Similarly, a
-mutant interacting with a resident receives the payoff $\pi(s_\mutant,
-s_\resident)$ and $\pi(s_\mutant, s_\mutant)$ when interacting with a mutant.\\
+interacting with a resident receives the payoff $\pi(\strategy_\resident, \strategy_\resident)$,
+and $\pi(\strategy_\resident, \strategy_\mutant)$ when interacting with a mutant. Similarly, a
+mutant interacting with a resident receives the payoff $\pi(\strategy_\mutant,
+\strategy_\resident)$ and $\pi(\strategy_\mutant, \strategy_\mutant)$ when interacting with a mutant.\\
 

### 2023-11-01 paper/main.tex

@@
 Even in the most extreme case that they only take into account their very last interaction, we find that cooperation can still evolve. 
-However, such individuals adopt less generous strategies, and they tend to cooperate less often than in the classical setup with expected payoffs. 
+However, such individuals adopt less generous strategies, and they tend to cooperate less often than in the classical setup with average payoffs. 
 Interestingly, once individuals remember the payoffs of two or three recent interactions, cooperation rates quickly approach the classical limit. 
@@
 A reactive strategy only depends on the other player's action in the previous round. 
-Such strategies can be written as a three-dimensional tuple \(\mathbf{s}=(y, p,q)\). 
+Such strategies can be written as a three-dimensional tuple \(\strategy=(y, p,q)\).
 The first entry \(y\) is the probability that the player opens with cooperation in the first round. 
@@
 This assumption could reflect, for example, a strong recency bias in how individuals evaluate payoffs.  
-In addition to this extreme case of limited payoff memory, later on we also explore cases in which players take into account the outcome of two, three, or four recent rounds.
+In addition to this extreme case of limited payoff memory, later on we also explore cases in which players take into account the outcome of two, three, four, or more recent rounds.
 
@@
 This assumption is fairly common in evolutionary game theory, because it makes some computations more efficient~\citep{fudenberg:JET:2006,wu:JMB:2012,mcavoy:jet:2015}, and because the results can be interpreted more easily.
-However, in Section~9 of the \esm{} we show that our main results continue to hold for strictly positive mutation rates.
+However, in Section~3 of the \esm{} we show that our main results continue to hold for strictly positive mutation rates.
 
@@
 In particular, individuals may recall a larger number of rounds, they may recall their interactions with several co-players, or both. 
-To gain further insights on the impact of payoff memory, we explore three additional scenarios. 
+To gain further insights on the impact of payoff memory, we explore four additional scenarios. 
 In the first scenario, players recall the payoffs they obtained in the last two rounds against a single co-player. 
 In the second scenario, players recall their last-round payoffs against two co-players. 
-Finally, in the last scenario, they recall the two last rounds against two co-players (further extensions are straightforward, but we do not explore them here).
+In the third scenario, they recall the two last rounds against two co-players. 
+Finally, in the last scenario, players update based on the average payoff they
+receive over all rounds with a single co-player (further extensions are straightforward, but we do not explore them here).
 
@@
 
-For the first two scenarios, we can derive an analytical condition for when cooperation is stochastically stable. 
+For most scenarios, we derive an analytical condition for when cooperation is stochastically stable. 
 To this end, we again assume populations are large and that selection is strong. 
@@
 The details of this analysis can be found in the \esm. 
-Interestingly, we find in both scenarios that for $b\!>\!2c$, cooperation is stochastically stable when $q\!<\! \frac{\sqrt{2}}{2}\approx 0.707$. 
+In the first two scenarios, we interestingly find that for $b\!>\!2c$, cooperation is stochastically stable when $q\!<\! \frac{\sqrt{2}}{2}\approx 0.707$. 
 Comparing this condition with the more stringent condition in Eq.~\eqref{Eq:LimitedMemory} suggests that there are now more conditionally cooperative strategies that can sustain cooperation. 
 Hence, cooperation should evolve more easily.
+In the last scenario, we find that cooperation is stochastically stable when \(q
+< 1 - \frac{c}{b}\), which is the same condition as in
+Eq.~\eqref{Eq:PerfectMemory}, even when only a single co-player is considered
+instead of the whole population. In this case, the evolved cooperation
+approaches the perfect payoff memory case the most.
 
@@
 
-We explore this prediction with additional simulations, again for a low and a high benefit $b$, 
+We explore these prediction with additional simulations, again for a low and a high benefit $b$, 
 see \FigHigherMemory.
 We observe that a minimal increase in the players' payoff memory (compared to the baseline case with a single round recalled) can promote cooperation considerably. 
-Specifically, in all three scenarios with extended memory, we now see similar cooperation rates, and they approach the rates observed under perfect memory. 
+Specifically, in all four scenarios with extended memory, we now see similar cooperation rates, and they approach the rates observed under perfect memory. 
 These results suggest that while it takes {\it some} payoff memory to sustain substantial cooperation rates, the requirements on memory seem to be rather modest. 
@@
   ({\it v}) based on the last two rounds of two interactions, and
-  ({\it vi}) based on the expected payoff of one interaction.
+  ({\it vi}) based on the average payoff of one interaction.
   Again, simulations are run either for a comparably high benefit of cooperation~($b/c\!=\!10$), or for a low benefit~($b/c\!=\!3$). 

### 2023-11-01 paper/si.tex

@@
 \newcommand{\alex}[1]{\textcolor{red}{\textbf{AL}: #1}}
+\newcommand{\nikoleta}[1]{\textcolor{orange}{\textbf{NG}: #1}}
 
@@
 ({\it v})~based on the last two rounds of two interactions, and
-({\it v})~based on the expected payoff of one interaction.
+({\it v})~based on the average payoff of one interaction.
 In each case, we consider the case that there are only two strategies present in the population (a resident and a mutant strategy). 
@@
 
-\subsection{Updates based on expected payoff of one repeated game}\label{section:one_interaction_full_memory}
+\subsection{Updates based on average payoff of one repeated game}\label{section:one_interaction_full_memory}
 
@@
 {\bf Motivation.}
+
+\nikoleta{CH to write a few senteces}
 
@@
 other player is a mutant. We consider that the learner and the
-model update based on their most recent interaction. As mentioned previously,
-two major cases must be considered:
+model update based on a single repeated game with a member of the population.
+As mentioned previously, two major cases must be considered:
 
 \begin{enumerate}
-  \item The learner and the role model have had their most recent interaction with each other.
-  \item The learner's last interaction was not with the role model.
+  \item The learner and the role model play a repeated game with each other.
+  \item The learner's last repeated game was not with the role model.
 \end{enumerate}
@@
 
-We assume that the players' payoff received from their last interaction is their
-expected payoff given by Eq.~\eqref{eq:expected_payoff}. Thus, a resident
+We assume that the players' calculations are based on the average payoffs (as given by
+Eq.~\eqref{eq:expected_payoff})
+received from interacting with their last interaction partners. Thus, a resident
 interacting with a resident receives the payoff $\pi(\strategy_\resident, \strategy_\resident)$,
@@
 We once again calculate how easily a single \alld{} mutant can invade into a
-resident population of \gtft{} players. When resident interacts with another
+resident population of \gtft{} players. When a resident interacts with another
 resident or with a mutant,
-the expected payoffs based on Eq.~\eqref{eq:expected_payoff} of the interactions are,
+the average payoffs based on Eq.~\eqref{eq:expected_payoff} of the interactions are,
 \begin{equation*}
@@
 \end{equation*}
-The expected payoff of the mutant, are
+The average payoffs of the mutant, are
 \begin{align*}
@@
 \end{align*}
-Based on these expected payoffs, we compute the ratio
+Based on these average payoffs, we compute the ratio
 \begin{equation}
@@
 
-\christian{Adjust symbols in main text -- for role model, strategy, etc}
 \end{document}

### 2023-11-02 paper/main.tex

@@
 %% Adding shortcut commands to refer to our figures %%
-\newcommand{\FigBaseResults}{figure~\ref{fig:expected_and_stochastic_for_donation}}
-\newcommand{\FigDependenceParameters}{figure~\ref{fig:cooperation_rate_over_benefit_and_beta}} 
-\newcommand{\FigHigherMemory}{figure~\ref{fig:cooperation_rate_all_updating_payoffs}}
+\newcommand{\FigBaseResults}{Figure~\ref{fig:expected_and_stochastic_for_donation}}
+\newcommand{\FigDependenceParameters}{Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}} 
+\newcommand{\FigHigherMemory}{Figure~\ref{fig:cooperation_rate_all_updating_payoffs}}
+
+\newcommand{\FigInvasionAnalysis}{{Supplementary Figure~3}} 
+\newcommand{\FigMemoryOneParameters}{Supplementary Figure~7}
 
@@
 \noindent
-Direct reciprocity can explain the evolution of cooperative behaviors in repeated interactions. 
-According to this literature, individuals should naturally learn to adopt conditionally cooperative strategies if they have multiple encounters with their interaction partner. 
+Direct reciprocity is a mechanism for the evolution of cooperation in repeated social interactions. 
+According to this literature, individuals naturally learn to adopt conditionally cooperative strategies if they have multiple encounters with their partner. 
 Corresponding models have greatly facilitated our understanding of cooperation, yet they often make strong assumptions on how individuals remember and process payoff information. 
-For example, when strategies are updated through social learning, it is commonly assumed that individuals compare their respective average payoffs.  
+For example, when strategies are updated through social learning, it is commonly assumed that individuals compare their average payoffs.  
 This would require them to compute (or remember) their payoffs against everyone else in the population.
@@
 They may use strategies such as Tit-for-Tat~\citep{rapoport:book:1965,axelrod1981evolution} or Generous Tit-for-Tat~\citep{molander:jcr:1985,Nowak1992tit} to preferentially cooperate with other cooperators.
-Such conditional strategies approximate human behavior fairly well~\citep{fischbacher:EconL:2001,Rand:TCS:2013,DalBo:AER:2019,Rossetti:ETH:2023} and they have also been documented in several other species~\citep{Carter:PRSB:2013,Schweinfurth:AnBehav:2019,Voelkl:PNAS:2015} -- although direct reciprocity is generally more difficult to demonstrate in animals~\citep{CluttonBrock:Nature:2009,Silk:CurrentBiology:2013,taborsky:CurrentBiology:2013}.
+Such conditional strategies approximate human behavior fairly well~\citep{fischbacher:EconL:2001,Rand:TCS:2013,DalBo:AER:2019,Montero-Porras:SciRep:2022,Rossetti:ETH:2023} and they have also been documented in several other species~\citep{Carter:PRSB:2013,Schweinfurth:AnBehav:2019,Voelkl:PNAS:2015} -- although direct reciprocity is generally more difficult to demonstrate in animals~\citep{CluttonBrock:Nature:2009,Silk:CurrentBiology:2013,taborsky:CurrentBiology:2013}.
 However, at the outset, it is not clear how easy it is to {\it learn} reciprocal strategies by social imitation. 
@@
 
-Most models of direct reciprocity abstract from these difficulties~\citep{brauchli:JTB:1999,brandt:JTB:2006,ohtsuki:JTB:2007b,szolnoki:pre:2009b,imhof2010stochastic,van-segbroeck:prl:2012,grujic:jtb:2012,Martinez2012,stewart:pnas:2013,pinheiro:PLoSCB:2014,stewart:games:2015,Baek2016,McAvoy:ProcA:2019,glynatsi:SCR:2020,Schmid:PlosCB:2022,Murase:SciRep:2022}.
+Most models of direct reciprocity abstract from these difficulties~\citep{brauchli:JTB:1999,brandt:JTB:2006,ohtsuki:JTB:2007b,szolnoki:pre:2009b,imhof2010stochastic,van-segbroeck:prl:2012,grujic:jtb:2012,Martinez2012,stewart:pnas:2013,pinheiro:PLoSCB:2014,stewart:games:2015,Baek2016,McAvoy:ProcA:2019,glynatsi:SCR:2020,Schmid:PlosCB:2022,Murase:SciRep:2022,Chen:PNASNexus:2023}.
 They just assume individuals can easily copy the strategies of others. 
@@
 
-To explore the impact of limited payoff memory, we adapt existing models of the evolution of direct reciprocity through social learning. 
+To explore the impact of limited payoff memory, we adapt existing models of the evolution of direct reciprocity.
 These models involve two different time scales. 
@@
 {\bf Description of the game dynamics.} We consider a well-mixed population consisting of $N$~players.
-Players are randomly matched in pairs to participate in a repeated donation game~\citep{sigmund2010calculus}.
-Each round, players can either cooperate (\(C\)) or defect (\(D\)). 
-By cooperating, a player provides a benefit \(b\) to the other player at their own cost \(c\), with \(0 \!<\! c \!<\! b\). 
+Players are randomly matched in pairs to participate in a repeated donation game~\citep{sigmund2010calculus} with their respective co-player.
+Each round, they can either cooperate (\(C\)) or defect (\(D\)). 
+A cooperating player provides a benefit~\(b\) to the other player at their own cost~\(c\), with \(0 \!<\! c \!<\! b\). 
+A defecting player provides no benefit and pays no cost. 
 Thus, the players' payoffs in a single round are given by the matrix
@@
 In the repeated game, strategies become arbitrarily complex. 
-Here, strategies are contingent rules, telling players what to do depending on the outcome of all their previous interactions. 
+Here, strategies are contingent rules, telling players what to do depending on the outcome of all previous rounds. 
 For simplicity, in the following we assume individuals use {\it reactive strategies}~\citep{Nowak1992tit}. 
-A reactive strategy only depends on the other player's action in the previous round. 
+A reactive strategy only depends on the other player's action in the last round. 
 Such strategies can be written as a three-dimensional tuple \(\strategy=(y, p,q)\).
@@
 In the short run, the players' strategies are taken to be fixed.
-Players use their strategies to decide whether to cooperate with any given interaction partner. 
+Players use their strategies to decide whether to cooperate in a series of repeated games against all other population members. 
 In the long run, however, the players' strategies may change depending on the payoffs they yield, as we describe in the following.\\
@@
 As the selection strength~$\beta$ increases, players are increasingly biased to imitate the role model only if it has the higher payoff.
-In either case, we assume in the following that the role model's strategy can be identified.
-That is, if the learner decides to imitate the role model's strategy, the learner copies it exactly, as in previous works on direct reciprocity.
+%In either case, we assume in the following that the role model's strategy can be identified.
+%That is, if the learner decides to imitate the role model's strategy, the learner copies it exactly, as in previous works on direct reciprocity.
 
@@
 Because the population mostly consists of residents, and because residents mutually cooperate with each other, their expected payoff is $\pi_\gtft = b\!-\!c$. 
-The mutant only interacts with residents, and hence receives a benefit in the first round, and in every subsequent round with probability $q$. 
+On the other hand, the defecting mutant only interacts with residents. 
+Given the residents' strategy, the mutant receives a benefit in the first round, and in every subsequent round with probability $q$. 
 As a result, the mutant's expected payoff is $\pi_\alld \!=\! (1\!-\!\delta\!+\!\delta q)b$. 
@@
 For each simulation, we record which strategies ($y,p,q$) the players adopt over time.
-\FigBaseResults{} depicts the conditional cooperation probabilities $p$ and $q$ (we omit the opening move \(y\); because we use a discount factor~\(\delta\) close to one, first-round behavior is largely irrelevant).
+\FigBaseResults{} depicts the conditional cooperation probabilities $p$ and $q$ (we omit the opening move \(y\) because we use a discount factor~\(\delta\) close to one, such that first-round behavior is largely irrelevant).
 In all simulations, we find that the players' strategies cluster in two regions of the strategy space. 
 The first region corresponds to a neighborhood of \alld{} with $(p,q)\!\approx\!(0,0)$.
-The second region corresponds to a thin strip of cooperative strategies with $(p,q)\!=\!(1,q)$. 
+The second region corresponds to a thin strip of cooperative strategies with $(p,q)\!\approx\!(1,q)$. 
 Within this strip, we observe that most strategies satisfy the constraints on $q$ imposed by the inequalities~\eqref{Eq:PerfectMemory} and~\eqref{Eq:LimitedMemory}. 
@@
 
-In each case, the evolutionary dynamics follow a similar cyclic pattern~\citep[as previously described in Refs.][]{imhof2010stochastic, Nowak1992tit}:
+In each case, the evolutionary dynamics follow a similar cyclic pattern~\citep[as described in Refs.][]{imhof2010stochastic, Nowak1992tit}:
 Resident populations of defectors are most likely invaded by strategies close to \tft. 
-Once the population adopts conditionally cooperative strategies  $(1,1,q)$, neutral drift may introduce larger values of generosity $q$. 
+Once the population adopts conditionally cooperative strategies  $(1,1,q)$, neutral drift may introduce larger values of generosity~$q$. 
 If the resident's generosity~$q$ violates the conditions~\eqref{Eq:PerfectMemory} and~\eqref{Eq:LimitedMemory}, defectors can re-invade, and the cycle starts again. 
-The overall cooperation rate, averaged over many time steps, depends on how much time the process spends near \alld{} on the one hand, and near the strip of conditionally cooperative strategies on the other.
-For perfect payoff memory, the overall cooperation rate is substantial. 
-It is 52\% for low benefits, and 98\% for high benefits. 
-For limited payoff memory, the evolving cooperation rates are smaller but strictly positive, with 37\% cooperation for low benefits and 51\% cooperation for high benefits (\FigBaseResults).
+The relative time spent near \alld{} and near the strip of conditionally cooperative strategies depends on the considered memory setting (\FigInvasionAnalysis, depicting the case of high benefits). 
+For perfect memory, we find that \alld{} is replaced relatively quickly by more cooperative strategies. 
+Here, it takes on average 159 invading mutants until \alld{} is successfully replaced. 
+In contrast, for limited memory, \alld{} is more robust, resisting on average 798 mutant strategies. 
+This picture reverses when we consider an initial population that adopts \gtft. 
+Such populations are much more robust under perfect memory than they are under limited memory. 
+Overall, we find that the impact of memory on the population's average cooperation rate is substantial. 
+For perfect memory, this rate is 52\% for low benefits, and 98\% for high benefits. 
+For limited payoff memory, the evolving cooperation rates are smaller but still strictly positive, with 37\% cooperation for low benefits and 51\% cooperation for high benefits (\FigBaseResults).
 
@@
 
-To further investigate the influence of different parameters, we have systematically varied the benefit~$b$ and the selection strength~$\beta$ in Figure~\FigDependenceParameters.
+To further investigate the influence of different parameters, we have systematically varied the benefit~$b$ and the selection strength~$\beta$ in~\FigDependenceParameters.
 According to \FigDependenceParameters{\it a}, perfect memory consistently results in a higher cooperation rate, and this relative advantage further increases with an increasing benefit~$b$. 
 Interestingly, for limited payoff memory, the cooperation rate remains stable at approximately 50\% once \(b \!\ge\! 5\).
-This again reflects our earlier observation that the feasibility of cooperation in this scenario is independent of the exact values of $b$ and $c$, as described by Eq.~\eqref{Eq:LimitedMemory}. 
+This again reflects our earlier observation that the feasibility of cooperation in this scenario is largely independent of the exact values of $b$ and $c$, as described by Eq.~\eqref{Eq:LimitedMemory}.
 With respect to the effect of different selection strengths, \FigDependenceParameters{\it b} suggests that both perfect and limited payoff memory yields similar cooperation rates for weak selection \(\beta \!<\!
@@
 Beyond weak selection, increasing selection has a positive effect under perfect payoff memory, but a negative effect under limited payoff memory.\\
+
+%% RESULTS: Memory-1 strategies %%
+
+\noindent
+{\bf Beyond reactive strategies.} While the results presented in the main text focus on reactive strategies, the patterns we observe do not seem to depend on the considered strategy space. 
+To illustrate this point in more detail, in the \esm{} we consider the dynamics among memory-1 strategies. 
+Here, players take into account both their co-player's and their own last move, see Refs.~\citep{nowak:Nature:1993,imhof:JTB:2007}. 
+Also in that case, we observe that perfect memory leads to systematically higher cooperation rates (Supplementary Figures~5,6).
+Again, this advantage of perfect memory is particularly pronounced for strong selection, or when there is a high benefit of cooperation (\FigMemoryOneParameters).\\
 
@@
 Finally, in the last scenario, players update based on the average payoff they
-receive over all rounds with a single co-player (further extensions are straightforward, but we do not explore them here).
+receive over all rounds with a single co-player (further extensions are possible, but we do not explore them here).
 
@@
 
-For most scenarios, we derive an analytical condition for when cooperation is stochastically stable. 
-To this end, we again assume populations are large and that selection is strong. 
+For most scenarios, we can again derive an analytical condition for when cooperation is stochastically stable. 
+As before, we assume populations are large and that selection is strong. 
 For simplicity, we also assume that the game continues almost certainly after each round (i.e., $\delta$ approaches one). 
@@
 < 1 - \frac{c}{b}\), which is the same condition as in
-Eq.~\eqref{Eq:PerfectMemory}, even when only a single co-player is considered
-instead of the whole population. In this case, the evolved cooperation
-approaches the perfect payoff memory case the most.
+Eq.~\eqref{Eq:PerfectMemory}, even though only a single co-player is considered
+instead of the whole population.
 
@@
 
-We explore these prediction with additional simulations, again for a low and a high benefit $b$, 
-see \FigHigherMemory.
+We complement these analytical results with additional simulations, see \FigHigherMemory.
 We observe that a minimal increase in the players' payoff memory (compared to the baseline case with a single round recalled) can promote cooperation considerably. 
-Specifically, in all four scenarios with extended memory, we now see similar cooperation rates, and they approach the rates observed under perfect memory. 
+Specifically, in all four scenarios with extended memory, we see similar cooperation rates, and they approach the rates observed under perfect memory. 
 These results suggest that while it takes {\it some} payoff memory to sustain substantial cooperation rates, the requirements on memory seem to be rather modest. 
@@
 Such a model discounts the future relative to the past.
-It also does not necessarily require ``memory'' because the rewards are accumulated into a factor used in decision-making; the specific time stamps of rewards do not themselves provide better information beyond their effects on total payoff.
+It also does not necessarily require ``memory'' because rewards are accumulated into a factor used in decision-making; the specific time stamps of rewards do not themselves provide better information beyond their effects on total payoff.
 In a similar fashion, the probabilistic interpretation of discounting as a continuation probability \citep{axelrod1981evolution} also effectively discounts the future relative to the past.
@@
 Within the context of a single repeated game, the model we consider here is, in some ways, dual to the classical model of temporally-discounted rewards in repeated games.
-Instead of making decisions based on expected rewards in the future, we consider agents who make decisions based on actual rewards in the past.
-Thus, the ability to estimate the future payoff of a strategy is replaced by the memory of how this strategy fared against others in the past.
+Instead of making decisions based on expected rewards in the future, we consider individuals who make decisions based on actual rewards in the past.
+Thus, the ability to estimate the future payoff of a strategy is replaced by the memory of how this strategy previously fared against others.
 This involves two time scales: interaction partners and rounds within those interactions.
@@
 According to this mechanism, cooperation pays if it makes the interaction partner more cooperative in future. 
-To describe which strategies of direct reciprocity are most effective, the previous theoretical literature assumes that the evolutionary dynamics are determined by the players' expected payoffs~\citep{brauchli:JTB:1999,brandt:JTB:2006,ohtsuki:JTB:2007b,szolnoki:pre:2009b,imhof2010stochastic,van-segbroeck:prl:2012,grujic:jtb:2012,Martinez2012,stewart:pnas:2013,pinheiro:PLoSCB:2014,stewart:games:2015,Baek2016,McAvoy:ProcA:2019,glynatsi:SCR:2020,Schmid:PlosCB:2022,Murase:SciRep:2022}.
+To describe which strategies are most effective, the previous theoretical literature assumes that the evolutionary dynamics are driven by the players' expected payoffs~\citep{brauchli:JTB:1999,brandt:JTB:2006,ohtsuki:JTB:2007b,szolnoki:pre:2009b,imhof2010stochastic,van-segbroeck:prl:2012,grujic:jtb:2012,Martinez2012,stewart:pnas:2013,pinheiro:PLoSCB:2014,stewart:games:2015,Baek2016,McAvoy:ProcA:2019,glynatsi:SCR:2020,Schmid:PlosCB:2022,Murase:SciRep:2022}.
 To the extent that strategies are learned (not inherited), this assumption seems to impose rather stringent requirements on the individuals' cognitive abilities. 
@@
 While more payoff memory is always favorable, reciprocal cooperation can already be sustained if individuals have a record of two or three past outcomes.
-We believe that this kind of result, derived entirely within a theoretical model, is crucial for making model-informed deductions about reciprocity in natural systems.
+We believe that this kind of result, derived entirely within a theoretical model, is crucial for making model-informed deductions about reciprocity in natural systems.\\[0.5cm]
 
@@
 
-\begin{comment}
 \noindent
@@
 {\bf Data accessibility.}
-All data and code used in this manuscript are openly available at https://github.com/Nikoleta-v3/pd-with-stochastic-payoffs \todo{(Correct?)}.\\
+All data and code used in this manuscript are openly available at\newline
+\texttt{https://zenodo.org/records/10066227}.\\
 
@@
 E-DIRECT, and by the Max Planck Society.\\
-\end{comment}
 
@@
   ({\it vi}) based on the average payoff of one interaction.
-  Again, simulations are run either for a comparably high benefit of cooperation~($b/c\!=\!10$), or for a low benefit~($b/c\!=\!3$). 
+  Again, simulations are run either for a comparably low benefit of cooperation~($b/c\!=\!3$), or for a high benefit~($b/c\!=\!10$). 
   We observe that perfect memory always yields the highest cooperation rate. 
@@
 \end{document}
-
-%% SOME ADDITIONAL TEXT BITS %%
-
-% repeated game markov chain
-given that the opponent cooperated and defected equivalently. The play between a
-pair of reactive strategies $s_1 = (y_1, p_1, q_1)$ and $s_2 = (y_2, p_2, q_2)$
-can be model as a Markov process with the transition matrix \(M\)~\cite{nowak:APC:1989},
-
-\begin{equation}\label{eq:transition_matrix}
-  \input{static/matrix.tex}
-\end{equation}
-
-and the stationary vector \(\mathbf{v}(s_1,s_2)\)
-which is the solution to \(\mathbf{v}(s_1,s_2) \times M 
-= \mathbf{v}(s_1,s_2)\).
-
-% pairwise comparison fixation probability
-
-In fact, so rare that only two different strategies can be
-present in the population at any given time. However, in the Supplementary Information
-Section 9 we show that the main result holds for \(\mu \neq 0\).
-The case of low mutation is vastly adopted because it allows us to explicitly
-calculate the fixation probability of a newly introduced mutant. More specifically,
-at each step
-one individual adopts a mutant strategy randomly selected from the set of
-feasible strategies. The fixation probability \(\phi_{M}\) of the mutant
-strategy can be calculated explicitly as,
-
-\begin{equation}\label{eq:appendix_fixation_probability}
-    \rho_{M} = \frac{1}{1+\sum\limits_{i=1}^{N-1}\prod\limits_k^i \frac{\lambda^-_k}{\lambda^+_k}},
-\end{equation}
-
-where \(k\) is the number of
-mutants and \(\lambda^-_k, \lambda^+_k\) are the probabilities that the number of
-mutants decreases and increases respectively.
-Depending on the fixation probability \(\phi_{M}\) the mutant either
-fixes (becomes the new resident) or goes extinct. Regardless, in the next elementary
-time step another mutant strategy is introduced to the population. We iterate
-this elementary population updating process for a large number of mutant
-strategies and we record the resident strategies at each time step. The
-probabilities \(\lambda^-_k \text{ and } \lambda^+_k\) depend on the fitness
-of the mutant and the resident strategies. In the next section we
-present how fitness is calculated in the cases of perfect and limited payoff memory.
-
-% Explicit formulas for the transition probabilities for perfect and limited memory
-
-\subsection{Fitness based on Perfect and Limited Payoff Memory}
-
-In the perfect payoff memory case an individual updates based on the average payoff
-against each other member of the population, otherwise known as expected payoffs. 
-The payoff of a reactive strategy \(s_1\) against the reactive strategy \(s_2\)
-in an infinitely repeated game ($\delta \rightarrow 1$) is explicitly
-calculated as,
-
-\[\langle\mathbf{v}(s_1,s_2),\mathbf{u}\rangle.\]
-
-In a population of size \(N\) there are \(k\) mutants and \(N - k\) residents.
-Let \(s_M =(y_M, p_M, q_M)\) and \(s_R = (y_R, p_R, q_R)\) denote the
-strategies of a mutant and a resident, the expected payoffs \(\pi_\rolemodel\) and
-\(\pi_M\) are given by,
-
-\begin{equation} \label{Eq:ExpPay}
-  \begin{array}{lcrcr}
-  \displaystyle \pi_\rolemodel & = &\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(s_R,s_R),\mathbf{u}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(s_R,s_M),\mathbf{u}\rangle,\\[0.5cm]
-  \displaystyle \pi_M & = &\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(s_M,s_R),\mathbf{u}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(s_M,s_M),\mathbf{u}\rangle.\\
-  \end{array}
-\end{equation}
-
-The probabilities that the number of mutants decreases and increases,
-\(\lambda^-_k\) and \(\lambda^+_k\), in the perfect payoff memory case are
-defined as,
-
-\begin{align}\label{eq:perfect_memory_lambdas}
-  \lambda^-_k \!=\!\varphi(\pi_M, \pi_\rolemodel) \quad \text{ and } \quad \lambda^+_k \!=\!\varphi(\pi_\rolemodel, \pi_M).
-\end{align}
-
-In the case of limited payoff memory we initially define the probability that a
-reactive strategy receives the payoff $u\!\in\! \mathcal{U}$ in the very last
-round of the game. This is given by Proposition~\ref{proposition:last_round}
-(see Supplementary Information Section 2.2.1 for proof).
-
-\begin{Prop}\label{proposition:last_round} Consider a repeated game, with
-continuation probability $\delta$, between players with reactive strategies
-$s_1\!=\!(y_1, p_1, q_1$)  and $s_2\!=\!(y_2,p_2,q_2)$ respectively. Then
-the probability that the $s_1$ player receives the payoff $u\!\in\!
-\mathcal{U}$ in the very last round of the game is given by
-$v_{u}(s_1,s_2)$, as given by Equation~(\ref{Eq:LastRound}).
-
-\begin{equation} \label{Eq:LastRound}
-  \setlength{\arraycolsep}{1pt}
-  \begin{array}{rcl}
-
-  v_{r}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(q_1+l_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+l_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
-  {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)},\\[1cm]
-
-  v_{s}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(q_1+l_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
-  {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)},\\[1cm]
-
-  v_{t}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+l_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
-  {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)},\\[1cm]
-
-  v_{p}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 l_1 l_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
-  {\displaystyle(1\!-\!\delta l_1l_2)(1\!-\!\delta^2 l_1 l_2)}.
-  \end{array}
-\end{equation}
-
-In these expressions, we have used the notation $l_i:=p_i\!-\!q_i$,
-$\bar{y}_i\!=\!1\!-\!y_i$, $\bar{q}_i:=1\!-\!q_i$, and
-$\bar{l}_i:=\bar{p}_i\!-\!\bar{q}_i=-l_i$ for $i\!\in\!\{1,2\}$.
-\end{Prop}
-
-In the case of limited payoffs memory both the role model and the learner
-estimate their fitness after interacting with a single member of the population.
-At each time step there are five possible pairings. They interact with each
-other with a probability \(\frac{1}{N - 1}\), and they do not interact with
-other with a probability \(1 - \frac{1}{N - 1}\). In the latter case, each of
-them can interact with either a mutant or a resident. Both of them interact with
-a mutant with a probability $\frac{(k-1)(k-2)}{(N-2)(N-3)}$ and both interact
-with a resident with a probability $\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$. The last
-two possible pairings are that either of them interacts with a resident whilst
-the other interacts with a mutant, and this happens with a probability
-$\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$. We define the probability that the randomly
-chosen resident obtained a payoff of $u_R$ in the last round of his respective
-game, and that the mutant obtained a payoff of $u_M$ as $x(u_R,u_M)$.
-
-\begin{equation}\label{eq:Chi}
-\setlength{\arraycolsep}{1pt}
-\begin{array}{llrl}
-x(u_R,u_M)	 &=&\displaystyle \frac{1}{N\!-\!1}\cdot  &v_{u_R}(s_R,s_M)\cdot 1_{(u_R,u_M)\in \mathcal{U}^2_F}\\[0.5cm]
-&+	
-&\displaystyle \left(1\!-\!\frac{1}{N\!-\!1}\right)  
-&\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} v_{u_R}(s_R,s_M) v_{u_M}(s_M,s_M) + 
- \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} v_{u_R}(s_R,s_M) v_{u_M}(s_M,s_R)\right.\\[0.5cm]
-&&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} v_{u_R}(s_R,s_R) v_{u_M}(s_M,s_M) + 
- \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} v_{u_R}(s_R,s_R) v_{u_M}(s_M,s_R)\right].
-\end{array}
-\end{equation}
-
-The probability that the number of mutants increases and decreases by one in the
-case of limited payoff memory are now given by,
-
-\begin{align}\label{eq:limited_memory_lambdas}
-\lambda^+_k=\frac{N\!-\!k}{N} \cdot \frac{k}{N} \cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M}) \varphi(u_{R},u_{M}) \quad \text{and} \quad
-\lambda^-_k=\frac{N\!-\!k}{N} \cdot \frac{k}{N} \cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M}) \varphi(u_{M},u_{R}).
-\end{align}
-
-In this expression, $\frac{(N\!-\!k)}{N}$ is the probability that the randomly
-chosen learner is a resident, and $\frac{k}{N}$ is the probability that the role
-model is a mutant. The sum corresponds to the total probability that the learner
-adopts the role model's strategy over all possible payoffs $u_R$ and $u_M$ that
-the two player may have received in their respective last rounds.

### 2023-11-02 paper/static/person.tex

\documentclass[tikz]{standalone}

\definecolor{solarizedBlue}{HTML}{268BD2}
\definecolor{bostonuniversityred}{rgb}{0.8, 0.0, 0.0}
\definecolor{ao}{rgb}{0.0, 0.5, 0.0}

\usetikzlibrary{positioning, fit, calc,matrix, arrows,scopes}

\begin{document}
\tikzset{man/.pic={
    \node[circle,fill,minimum size=10mm] (head) {};
    \node[rounded corners=4pt,minimum height=1.1cm,minimum width=0.8cm,fill,below = 2pt of head] (body) {};
    \draw[line width=2mm,round cap-round cap] ([shift={(4pt,-2pt)}]body.north east) --++(-90:6mm);
    \draw[line width=2mm,round cap-round cap] ([shift={(-4pt,-2pt)}]body.north west)--++(-90:6mm);
    }}
\begin{tikzpicture}

\pic[ao] at (0, 0) (myman) {man};

\end{tikzpicture}
\end{document}

### 2023-11-02 paper/main.tex
CH: Adding a few references

@@
 In models of biological evolution, individuals with a high fitness produce more offspring; in models of cultural evolution, such individuals are imitated more often. 
-Although biological and cultural evolution are sometimes treated as equivalent, there can be important differences~\citep{Wu2015,Smolla:PTRS:2021}. 
+Although biological and cultural evolution are sometimes treated as equivalent, there can be important differences~\citep{Wu2015,Smolla:PTRS:2021,Denton:TPB:2022}. 
 For example, models of biological evolution do not require individuals to have any particular cognitive abilities.
@@
 
-Most models of direct reciprocity abstract from these difficulties~\citep{brauchli:JTB:1999,brandt:JTB:2006,ohtsuki:JTB:2007b,szolnoki:pre:2009b,imhof2010stochastic,van-segbroeck:prl:2012,grujic:jtb:2012,Martinez2012,stewart:pnas:2013,pinheiro:PLoSCB:2014,stewart:games:2015,Baek2016,McAvoy:ProcA:2019,glynatsi:SCR:2020,Schmid:PlosCB:2022,Murase:SciRep:2022,Chen:PNASNexus:2023}.
+Most models of direct reciprocity abstract from these difficulties~\citep{brauchli:JTB:1999,brandt:JTB:2006,ohtsuki:JTB:2007b,szolnoki:pre:2009b,imhof2010stochastic,van-segbroeck:prl:2012,grujic:jtb:2012,Martinez2012,stewart:pnas:2013,pinheiro:PLoSCB:2014,stewart:games:2015,Baek2016,McAvoy:ProcA:2019,glynatsi:SCR:2020,Schmid:PlosCB:2022,Murase:SciRep:2022,Cooney:BMB:2022,Chen:PNASNexus:2023}.
 They just assume individuals can easily copy the strategies of others. 

### 2023-11-05 paper/main.tex
daggers and removes spaces from abstract

@@
 \usepackage[ruled,vlined]{algorithm2e}
+\usepackage{lineno}
+\linenumbers
 \smallskip % Erlaubt kleine Abstaende zwischen Paragraphen, falls es dem Seitenlayout hilft
@@
 \author{\parbox[c]{16cm}{\centering \onehalfspacing 
-Nikoleta E. Glynatsi$^1$,  Alex McAvoy$^{2,3}$, Christian Hilbe$^1$\\ \quad \\
+Nikoleta E. Glynatsi$^1$,  Alex McAvoy$^{2,3, \dagger}$, Christian Hilbe$^{1, \dagger}$\\ \quad \\
 $^{\rm 1}$Max Planck Research Group on the Dynamics of Social Behavior,\\ Max Planck Institute for Evolutionary Biology, Pl\"{o}n, Germany \\
 $^{\rm 2}$School of Data Science and Society, University of North Carolina at Chapel Hill,\\ Chapel Hill, NC 27599 \\
-$^{\rm 3}$Department of Mathematics, University of North Carolina at Chapel Hill,\\ Chapel Hill, NC 27599}}
+$^{\rm 3}$Department of Mathematics, University of North Carolina at Chapel Hill,\\ Chapel Hill, NC 27599 \\
+$^{\rm \dagger}$ A.M. and C.H. contributed equally to this work.}}
 
@@
 
+\clearpage
+\newpage
+
 \begin{figure}[t]

### 2024-03-11 paper/main.tex
wip NG working on journal revision

@@
 \usepackage{lineno}
-\linenumbers
+% \linenumbers
 \smallskip % Erlaubt kleine Abstaende zwischen Paragraphen, falls es dem Seitenlayout hilft
@@
 
-Most models of direct reciprocity abstract from these difficulties~\citep{brauchli:JTB:1999,brandt:JTB:2006,ohtsuki:JTB:2007b,szolnoki:pre:2009b,imhof2010stochastic,van-segbroeck:prl:2012,grujic:jtb:2012,Martinez2012,stewart:pnas:2013,pinheiro:PLoSCB:2014,stewart:games:2015,Baek2016,McAvoy:ProcA:2019,glynatsi:SCR:2020,Schmid:PlosCB:2022,Murase:SciRep:2022,Cooney:BMB:2022,Chen:PNASNexus:2023}.
+Most models of direct reciprocity abstract from these difficulties~\citep{brauchli:JTB:1999,brandt:JTB:2006,ohtsuki:JTB:2007b,szolnoki:pre:2009b,imhof2010stochastic,van-segbroeck:prl:2012,grujic:jtb:2012,Martinez2012,stewart:pnas:2013,pinheiro:PLoSCB:2014,stewart:games:2015,Baek2016,hilbe:PNAS:2017,McAvoy:ProcA:2019,glynatsi:SCR:2020,Schmid:PlosCB:2022,Murase:SciRep:2022,Cooney:BMB:2022,Chen:PNASNexus:2023}.
 They just assume individuals can easily copy the strategies of others. 
@@
 
-Overall, these findings suggest that while memory is important, already minimal payoff information may suffice for the evolution of direct reciprocity based on social learning. 
-They also suggest that the classical model of reciprocity (based on expected payoffs) can often be interpreted as a useful approximation to more realistic models that include cognitive constraints.
+% Overall, these findings suggest that while memory is important, already minimal payoff information may suffice for the evolution of direct reciprocity based on social learning. 
+% They also suggest that the classical model of reciprocity (based on expected payoffs) can often be interpreted as a useful approximation to more realistic models that include cognitive constraints.
+
+\todo{To address the reviewer's comment, we can either delete the last paragraph or rephrase the second-to-last paragraph as follows:}
+
+To explore this question, we theoretically study imitation dynamics in the
+repeated prisoner's dilemma, employing two extreme scenarios. In the first
+scenario, which follows the conventional modeling approach, individuals update
+their strategies based on their expected payoffs. We contrast this model with an
+alternative scenario where individuals base their strategy updates solely on the
+very last (one-shot) payoff they obtained. We observe that individuals with
+limited payoff memory tend to adopt less generous strategies, yet moderate
+levels of cooperation can still emerge. This suggests that while memory plays a
+significant role, minimal payoff information may suffice for the evolution of
+direct reciprocity through social learning. Moreover, increasing individuals'
+payoff memory to include the last two or three one-shot payoffs leads to
+cooperation rates quickly approaching those observed in the classical baseline
+case. This implies that the classical reciprocity model, based on expected
+payoffs, often serves as a useful approximation to more complex models
+incorporating cognitive constraints.
 

### 2024-03-11 paper/si.tex
wip NG working on journal revision

@@
 %% Adding shortcut commands to refer to our figures %%
-\newcommand{\FigEvoProc}{{\bf Fig.~1}} \newcommand{\FigInvAnalysis}{{\bfFig.~2}} \newcommand{\FigResultsOverPara}{{\bf Fig.~3}}
+\newcommand{\FigEvoProc}{{\bf Fig.~1}} 
+\newcommand{\FigInvAnalysis}{{\bfFig.~2}} 
+\newcommand{\FigResultsOverPara}{{\bf Fig.~3}}
 
@@
 Then the game continues for another round with probability~$\delta$.
-Players make their decisions whether to cooperate in any given round based on their reactive strategies $\mathbf{s}\!=\!(y,p,q)$. 
+Players make their decisions whether to cooperate in any given round based on their reactive strategies $\strategy\!=\!(y,p,q)$. 
 The entry $y$ determines a player's first-round cooperation probability. 
@@
 The probability is $p$ if the co-player cooperated in the previous round, and it is $q$ if the co-player defected. 
-On this short time scale, the players' strategies are fixed, and players are consecutively matched to play many repeated games with randomly changing interaction partners.
+On this short time scale, the players' strategies are fixed, and players are consecutively matched to play repeated games with all other population members.
 
@@
 We refer to this player as the `learner' (\learner). 
-The learner is then given an opportunity to update its strategy. 
+The learner is given an opportunity to update its strategy. 
 There are two possibilities for how this update may occur. 
@@
 In that case, the player's new strategy is drawn uniformly from the space of all reactive strategies~$[0,1]^3$.  
-With probability $1\!-\!\mu$, the player compares itself with a neighbor. 
-In that case, the player randomly picks another individual from the population (referred to as the `role model', \rolemodel). 
+With probability $1\!-\!\mu$, the player compares itself to another population member. 
+To this end, the player randomly picks another individual from the population (referred to as the `role model', \rolemodel). 
 The learner adopts the role model's strategy with a probability \(\varphi\), given by
 \begin{equation} \label{Eq:rho}
-    \varphi(\pi_{\learner}, \pi_{\rolemodel}) = \frac{1}{1\!+\! e^{\!-\!\beta (\pi_\rolemodel - \pi_\learner)}}.
-\end{equation}
-The parameter \(\beta\) is the selection strength.
+    \varphi(\pi_{\learner}, \pi_{\rolemodel}) = \frac{1}{1\!+\! e^{-\!\beta (\pi_\rolemodel - \pi_\learner)}}.
+\end{equation}
+The parameter \(\beta\!\ge\!0\) is the selection strength.
 It determines how important payoff differences are for the learner's decision to imitate the role model. 
-The variables $\pi_\rolemodel$ and $\pi_\learner$ refer to the relevant payoffs of the role model and the learner, respectively. 
+The variables $\pi_\rolemodel$ and $\pi_\learner$ refer to the payoffs of the role model and the learner, respectively. 
 The exact value of these payoffs depend on the players' memory. 
@@
 ({\it v})~based on the average payoff of one interaction.
-In each case, we consider the case that there are only two strategies present in the population (a resident and a mutant strategy). 
+In each case, we assumet there are only two strategies present in the population (a resident and a mutant strategy). 
 We first derive how likely it is that a learner (of any type) assigns a given payoff $\pi_\learner$ to itself, and a payoff of $\pi_\rolemodel$ to the role model. 
@@
 
-{\bf Computing the ratio $\lambda^-_k/\lambda^+_k$.}
+{\bf Computing the players' expected payoffs.}
 The case of perfect payoff memory corresponds to the classical case considered in the previous literature. 
-Here, individuals update their strategies based on the expected payoffs, based on all rounds and all possible interaction partners. 
+Here, individuals update their strategies based on their expected payoffs, taking into account all rounds and all possible interaction partners. 
 When players use reactive strategies (or more generally, strategies with finite memory), these expected payoffs can be computed explicitly, based on a Markov chain approach~\cite{sigmund2010calculus}.
@@
 \end{array}
-\right]
+\right].
 \end{equation}
@@
 \end{equation}
+The second player's payoff can be computed analogously.
 
@@
 \noindent
+{\bf Computing the ratio $\lambda^-_k/\lambda^+_k$.}
 After these preparations, consider now a population with \(k\) mutants and \(N - k\) residents, whose strategies we denote by \(\strategy_\mutant=(y_\mutant,p_\mutant, q_\mutant)\) and \(\strategy_\resident = (y_\resident, p_\resident, q_\resident)\), respectively.
@@
 As an application of this formalism, we can compute when cooperation is stochastically stable in the perfect-information setting. 
-To this end, suppose there is only a single mutant, $k\!=\!1$.
+To this end, suppose there is a single mutant, $k\!=\!1$.
 The residents adopt Generous Tit-for-Tat, $\gtft = (1, 1, q)$ and the mutant adopts $\alld\!
@@
 To treat the case of limited memory analytically, let $\mathcal{U} \!=\! \{R,S,T,P\}$ be the set of possible one-shot payoffs. 
-Since the game ends after round $T$ with probability $\delta^{T}\left(1-\delta\right)$, if two players use reactive strategies $\strategy_1\!=\!(y_1, p_1, q_1$)  and $\strategy_2\!=\!(y_2,p_2,q_2)$, then, by definition, the probability that the first player receives a payoff of $u\!\in\!\mathcal{U}$ in the final round of the game is given by Eq.~\eqref{Eq:vAverage}.
+Since the game ends after round $\tau$ with probability $\delta^{\tau}\left(1-\delta\right)$, if two players use reactive strategies $\strategy_1\!=\!(y_1, p_1, q_1$)  and $\strategy_2\!=\!(y_2,p_2,q_2)$, then, by definition, the probability that the first player receives a payoff of $u\!\in\!\mathcal{U}$ in the final round of the game is given by Eq.~\eqref{Eq:vAverage}.
 To make the strategies explicit, we denote this probability by $v_{u}\!\left(\strategy_{1},\strategy_{2}\right)$.
@@
 {\bf Computing the ratio $\lambda^-_k/\lambda^+_k$.}
-After these preparations, let us again consider the corresponding population setup, with $N\!-\!k$ residents with strategy  \(\strategy_\resident = (y_\resident, p_\resident, q_\resident)\) and $k$ mutants with strategy  \(\strategy_\mutant = (y_\mutant, p_\resident, q_\mutant)\).
+After these preparations, let us again consider the corresponding population setup, with $N\!-\!k$ residents with strategy  \(\strategy_\resident = (y_\resident, p_\resident, q_\resident)\) and $k$ mutants with strategy  \(\strategy_\mutant = (y_\mutant, p_\mutant, q_\mutant)\).
 At each step of the evolutionary process we choose a learner and a role model. 
 The learner compares the performance of its strategy by comparing its own last one-shot payoff with the last one-shot payoff of the role model. 
-In the following, we assume either the learner or the role model is a resident and that the other player is a mutant (otherwise it is certain that the number of mutants does not change). 
+In the following, we assume either the learner or the role model is a resident, and that the other player is a mutant (otherwise the number of mutants does not change). 
 There are two major cases to consider.
@@
 This happens with probability $1/(N\!-\!1)$.
-In that case, there are four possible cases for their joint final payoffs, $(u_\resident,u_\mutant)\!\in\!\mathcal{U}_F\!:=\!\big\{(R,R),\,(S,T),\,(T,S),\,(P,P)\big\}$.
-These outcomes follow the distribution $\mathbf{v}(\strategy_\resident,\strategy_\mutant)$ (Eq.~\eqref{Eq:vAverage}).
+In that case, there are four possible cases for their joint final payoffs, $(u_\resident,u_\mutant)\in\mathcal{U}_F\!:=\!\big\{(R,R),\,(S,T),\,(T,S),\,(P,P)\big\}$.
+By Eq.~\eqref{Eq:vAverage}, these four outcomes follow the distribution~$\mathbf{v}(\strategy_\resident,\strategy_\mutant)$.
 \item The learner's last interaction was not with the role model, with probability $(N\!-\!2)/(N\!-\!1)$.
@@
 \noindent
-The first term on the right side corresponds to the case that the learner and
+The first line on the right side corresponds to the case that the learner and
 the role model happened to be matched directly for their last interaction. 
@@
 That is, it needs to be the case that $(u_\resident,u_\mutant)\in \mathcal{U}_F$, as represented by the respective indicator function. 
-The probability that the number of mutants increase or decreases by one is now given by
-\begin{equation}
+The second and third line of \eqref{eq:Chi} summarize the four possible subcases that can occur when the learner and the role model were not directly matched for their last interaction.
+
+Overall, we now obtain the following expressions for the probability that the number of mutants increases or decreases by one, \begin{equation}
 \begin{array}{l}
@@
 Moreover, in the special case that games are infinitely repeated, \(\delta\! \rightarrow\! 1\), we conclude that cooperation is stochastically stable if  \(q \!<\! 1/2\). 
-(For \(q\! =\! 1/2\), the payoff of the ALLD player is
+(For \(q\! =\! 1/2\), the payoff of the \alld{} player is
 \(T \!>\! R\) for half of the time, and it is \(P\! < \! R\) for the other half. The
 probability that the number of mutants increase by one equals the probability
-that the mutant goes extinct.)
+that the mutant goes extinct).
 
@@
 This setting can be generalized such that an individual
-considers \(m\) rounds and of \(n\) interactions. Here, we discuss the case that
+considers \(m\) rounds and \(n\) interactions. Here, we discuss the case that
 the update depends on the last round of \(n\) interactions.
@@
 As we increase $n$, the number of possible combinations increases non-linearly; 
-for a graphical illustration, see Fig.~\ref{fig:matching_tree}. 
+for a graphical illustration, see Supplementary~Fig.~\ref{fig:matching_tree}. 
 In the following, we study the case of \(n=2\), such that the learner takes into account the players' final payoffs of two interactions.\\
@@
 
-We assume that player the resident receives the payoff \(u_{\resident1}\) with their first interaction and \(u_{\resident 2}\) with their second.
+We assume the resident receives the payoff \(u_{\resident1}\) with their first interaction and \(u_{\resident 2}\) with their second.
 Similarly, the mutant receives payoffs \(u_{\mutant 1}\) and \(u_{\mutant 2}\). 
@@
 As before, we can use this formalism to compute when cooperation is stochastically stable.
-For a single \alld{} mutant in a \gtft{} resident population, we obtain the following non-zero probabilities \(x(u_{\resident1} u_{\resident 2}, u_{\mutant 1},u_{\mutant 2})\)
+For a single \alld{} mutant in a \gtft{} resident population, we obtain the following non-zero probabilities \(x(u_{\resident1} u_{\resident 2}, u_{\mutant 1},u_{\mutant 2})\),
 \begin{equation*}
@@
 \end{equation}
-We note that unlike the case of limited payoff memory, here the exact magnitude of the benefit and the cost of cooperation do have some effect on the transition probabilities. 
+Unlike in the case of limited payoff memory, here the exact magnitude of the benefit and the cost of cooperation do have some effect on the transition probabilities. 
 By requiring that the above ratio exceeds one, we obtain the following conditions for stochastic stability of cooperation, 
@@
 
-{\bf General setup.} As our next setting, we consider a setup in which players update their strategies based on the last two rounds of their last game. 
+{\bf General setup.} As our next setting, we assume players update their strategies based on the last two rounds of their last game. 
 Here, additional complications may arise because some games may only last for one round if $\delta\!<\!1$.
@@
 We can now derive a probability distribution for the first player's payoff in the last two rounds of a repeated game. 
-To this end, consider the $4\times4$ transition matrix $M(\strategy_1,\strategy_2)=(m_{u,u'})$ according to~Eq.~\eqref{eq:transition_matrix}. 
+To this end, consider the $4\times4$ transition matrix $M(\strategy_1,\strategy_2)\!=\!(m_{u,u'})$ according to~Eq.~\eqref{eq:transition_matrix}. 
 Instead of the usual indexing of the four rows by numbers, $i\!\in\!\{1,2,3,4\}$, here we label the four rows of this matrix by the first player's payoffs in the previous round, $u\!\in\!\{R,S,T,P\}$.
@@
 For example, $m_{ST}$ corresponds to the second row and third column of matrix $M$. 
-So by Eq.~\eqref{eq:transition_matrix}, $m_{ST} = \bar{q}_1 p_2=(1\!-\!q_1)p_2$.
-Based on this notation, we can describe the probability $w_{uu'}(\strategy_1,\strategy_2)$ that player~1 obtains a payoff of $u$ in the penultimate round (probability $v_{u}$) and a payoff of $u'$ in the last round (probability $m_{u,u'}$), giving $w_{uu'}(\strategy_1,\strategy_2) = v_u \cdot m_{u,u'}$.
+So by Eq.~\eqref{eq:transition_matrix}, $m_{ST} \!=\! \bar{q}_1 p_2 \!=\!(1\!-\!q_1)p_2$.
+Using this notation, we can describe the probability $w_{uu'}(\strategy_1,\strategy_2)$ that the first player receives the payoffs $u$ and $u'$ in the last two rounds. 
+By combining the probabilities that player~1 obtains a payoff of $u$ in the penultimate round (probability~$v_{u}$) and then a payoff of $u'$ in the last round (conditional probability $m_{u,u'}$), we obtain $w_{uu'}(\strategy_1,\strategy_2) = v_u \cdot m_{u,u'}$.
 
@@
 \end{equation*}
-Based on these payoff probabilities, we compute the ratio
+Based on these probabilities, we compute the ratio
 \begin{equation}
@@
 
-As our final extension, we consider a combination of the previous two settings, discussed in Sections~\ref{section:m_one_n_two} and \ref{section:m_two_n_one}. 
+For our next extension, we consider a combination of the previous two settings, discussed in Sections~\ref{section:m_one_n_two} and \ref{section:m_two_n_one}. 
 Now, individuals take into account two interactions with different co-players, and from each interaction, they take into account the last two rounds. 
@@
 \end{equation}
- Though we do not pursue any further analytical exploration of this case, in the
-next section we present simulation results for this setting, as well as for the other settings.\\
+Due to the large number of cases to consider in this setting, we do not derive conditions for the stochastic stability of cooperation.
 
@@
 
-\subsection{Updates based on average payoff of one repeated game}\label{section:one_interaction_full_memory}
+\subsection{Updates based on the average payoff of one repeated game}\label{section:one_interaction_full_memory}
 
@@
 
-{\bf Motivation.}
-
-\nikoleta{CH to write a few senteces}
+{\bf Motivation.} As our last setting, we consider the case that players update their strategies based on an (entire) repeated game against a single co-player.
+In terms of cognitive complexity, this setting is in between the case of perfect memory and the case considered in Section~\ref{section:m_two_n_one}, where players only take into account two rounds of one repeated game.\\
 
@@
 {\bf Computing the ratio $\lambda^-_k/\lambda^+_k$.}
-We revisit the population setup, consisting of \(k\) mutants with the strategy
-\(\strategy_\mutant = (y_\mutant, p_\mutant, q_\mutant)\) and \(N-k\) residents
+As in the previous sections, we consider \(k\) mutants with strategy
+\(\strategy_\mutant \!=\! (y_\mutant, p_\mutant, q_\mutant)\) and \(N\!-\!k\)~residents
 with strategies \(\strategy_\resident = (y_\resident, p_\resident,
-q_\resident)\). In each step of the evolutionary process, we select a learner and
-a role model. The learner or the role model can be either a resident, and the
-other player is a mutant. We consider that the learner and the
-model update based on a single repeated game with a member of the population.
-As mentioned previously, two major cases must be considered:
-
-\begin{enumerate}
-  \item The learner and the role model play a repeated game with each other.
-  \item The learner's last repeated game was not with the role model.
-\end{enumerate}
-
-In the latter case, there are four different subcases based on whether the
-resident's last interaction partner was a mutant or a resident and whether the
-mutant's last interaction partner was a mutant or a resident.
-
-We assume that the players' calculations are based on the average payoffs (as given by
-Eq.~\eqref{eq:expected_payoff})
-received from interacting with their last interaction partners. Thus, a resident
-interacting with a resident receives the payoff $\pi(\strategy_\resident, \strategy_\resident)$,
-and $\pi(\strategy_\resident, \strategy_\mutant)$ when interacting with a mutant. Similarly, a
-mutant interacting with a resident receives the payoff $\pi(\strategy_\mutant,
-\strategy_\resident)$ and $\pi(\strategy_\mutant, \strategy_\mutant)$ when interacting with a mutant.\\
-
-\noindent
-After these preparations, we can define the probability that the number of
-mutants either increases or decreases by one.
-
+q_\resident)\). Moreover, without loss of generality, we only consider 
+those cases in which either the learner or the role model is a resident, and the other player is a mutant. 
+As in previous cases, two major cases must be considered, ({\it i}) the learner and the role model interact directly, and ({\it ii}) the learner's last repeated game was not with the role model.
+In each case, the learner's decision is now based on the players' expected payoffs, as given by Eq.~\eqref{eq:expected_payoff}, with their last interaction partner. 
+Based on these observations, we can derive the probability that the number of
+mutants either increases or decreases by one,
 \begin{equation*}
@@
     \lambda^+	 &=
-  &\frac{1}{N\!-\!1}\cdot  & \varphi\left(\pi (\strategy_\resident, \strategy_\mutant), \pi (\strategy_\mutant, \strategy_\resident)\right)\\[0.5cm]
+  &\frac{1}{N\!-\!1}\cdot  & \varphi\left(\pi (\strategy_\resident, \strategy_\mutant), \pi (\strategy_\mutant, \strategy_\resident)\right)\\[0.3cm]
   &+
@@
   &\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3}\cdot \varphi(\pi (\strategy_\resident, \strategy_\mutant), \pi (\strategy_\mutant, \strategy_\mutant)) + 
-   \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3}\cdot \varphi(\pi (\strategy_\resident, \strategy_\mutant), \pi (\strategy_\mutant, \strategy_\resident)) \right.\\[0.5cm]
+   \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3}\cdot \varphi(\pi (\strategy_\resident, \strategy_\mutant), \pi (\strategy_\mutant, \strategy_\resident)) \right.\\[0.3cm]
   &&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} \cdot \varphi(\pi (\strategy_\resident, \strategy_\resident), \pi (\strategy_\mutant, \strategy_\mutant)) + 
@@
 \end{equation*}
-
 \begin{equation*}
@@
     \lambda^-	 &=
-  &\frac{1}{N\!-\!1}\cdot  & \varphi\left(\pi (\strategy_\mutant, \strategy_\resident), \pi(\strategy_\resident, \strategy_\mutant)\right)\\[0.5cm]
+  &\frac{1}{N\!-\!1}\cdot  & \varphi\left(\pi (\strategy_\mutant, \strategy_\resident), \pi(\strategy_\resident, \strategy_\mutant)\right)\\[0.3cm]
   &+
@@
   &\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3}\cdot \varphi(\pi(\strategy_\mutant, \strategy_\mutant), \pi(\strategy_\resident, \strategy_\mutant)) + 
-   \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3}\cdot \varphi(\pi(\strategy_\mutant, \strategy_\resident), \pi(\strategy_\resident, \strategy_\mutant)) \right.\\[0.5cm]
+   \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3}\cdot \varphi(\pi(\strategy_\mutant, \strategy_\resident), \pi(\strategy_\resident, \strategy_\mutant)) \right.\\[0.3cm]
   &&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} \cdot \varphi(\pi(\strategy_\mutant, \strategy_\mutant), \pi(\strategy_\resident, \strategy_\resident)) + 
@@
 We once again calculate how easily a single \alld{} mutant can invade into a
-resident population of \gtft{} players. When a resident interacts with another
-resident or with a mutant,
-the average payoffs based on Eq.~\eqref{eq:expected_payoff} of the interactions are,
+resident population of \gtft{} players. 
+Depending on whether a resident interacts with another resident or with a mutant,
+the resident's average payoff based on Eq.~\eqref{eq:expected_payoff} is
 \begin{equation*}
-\begin{array}{ll}
-  \pi (\strategy_\resident, \strategy_\resident) =  b - c,
-  &\pi (\strategy_\resident, \strategy_\mutant) =  - c \left(\delta \left(q - 1\right) + 1\right).
-\end{array}
-\end{equation*}
-The average payoffs of the mutant, are
+  \pi (\strategy_\resident, \strategy_\resident) =  b - c \quad \quad \text{or} \quad \quad \pi (\strategy_\resident, \strategy_\mutant) =  - c \big(1-\delta(1 \!-\! q)\big).
+\end{equation*}
+The mutant can only interact with a resident, yielding the payoff
 \begin{align*}
-\begin{array}{ll}
-  \pi (\strategy_\mutant, \strategy_\resident) =  b \left(\delta \left(q - 1\right) + 1\right),
-  &\pi (\strategy_\mutant, \strategy_\mutant) =  0.
-\end{array}
+  \pi (\strategy_\mutant, \strategy_\resident) =  b \big(1+\delta \left(1\!-\!q\right) \big).
 \end{align*}
@@
 \end{equation}
-Interestingly, this conditions is the same as the condition for stochastic
-stability in the case of perfect memory, see
-Eq.~\eqref{Eq:Condition_PerfectMemory}.
+Interestingly, this condition is the same as condition~\eqref{Eq:Condition_PerfectMemory} for the case of perfect memory when $N\! \rightarrow\! \infty$.
 
@@
 To gain some insights into the strategies evolving in each of the settings described in the previous section, we simulate the evolutionary process according to Algorithm~\ref{algorithm:pairwise_comparison}.
-Each time step, we record the current resident population \((y, p, q)\). 
 The results of these simulations are shown in Supplementary Fig.~\ref{fig:expected_payoffs_results}. 
-Qualitatively, all settings lead to similar dynamics as in the case of perfect and limited memory described in the main text. 
+Qualitatively, all settings lead to similar dynamics as in the case of perfect and limited memory. 
 In particular, populations are clustered around \alld, and around a strip of conditionally cooperative strategies, \gtft$=\!(1,1,q)$. 
@@
 For limited memory, the generosity of a conditional cooperator satisfies \(q <
-\frac{1}{2}\) for both values of benefit.
+\frac{1}{2}\) for both benefit values.
 In the next two cases, when players either remember the last round of two
@@
 requires \(q < \frac{\sqrt{2}}{2} \approx 0.701\).
-In the final case, we observe that the generosity of a conditional cooperator is
-in line with the condition for stochastic stability in the perfect memory case:
+In the last case, we get the same condition for the generosity of a conditional cooperator as in the perfect memory case:
 \(q < 1 - \frac{c}{b} = 0.9\) for high benefit and \(q < 1 - \frac{c}{b} =
@@
 \noindent
-{\bf Impact of mutations.} 
+{\bf Impact of the mutation rate.} 
 All of our previous simulations are run under the assumption that mutations are rare,~$\mu\!\rightarrow\!0$. 
@@
 At each time step, we record the average cooperation rate of the resident population. 
-The results for different mutation rates are shown in Fig.~\ref{fig:mutation}. 
-The cooperation rate is higher when players have perfect memory (compared to limited memory), regardless of the
-mutation rate. 
+The results for different mutation rates are shown in Supplementary~Fig.~\ref{fig:mutation}. 
+For almost all considered mutation rates, the cooperation rate is higher when players have perfect memory (compared to limited memory).
 Only as the mutation rate approaches one, and the processes become fully random, does the cooperation rate approach $1/2$ in both simulations.
 
-\begin{figure}[!htbp]
+\begin{figure}[t!]
 	\centering 
@@
 previous round to decide on their next action. 
-There are four possible outcome in each
-round; \((C, C), (C, D), (D, C), (D, D)\). 
+There are four possible outcomes in each
+round,  \((C, C), (C, D), (D, C), (D, D)\). 
 A memory-one strategy \(\strategy\) can thus be
@@
 To explore the dynamics among memory-one strategies, we perform four separate simulations, for both perfect and limited memory, and for two different benefit values. 
-Fig.~\ref{fig:memory_one_low_benefit} shows the results for a comparably low benefit of cooperation; 
-Fig.~\ref{fig:memory_one_high_benefit} shows the corresponding results for a large benefit. 
+Supplementary~Fig.~\ref{fig:memory_one_low_benefit} shows the results for a comparably low benefit of cooperation; 
+Supplementary~Fig.~\ref{fig:memory_one_high_benefit} shows the corresponding results for a large benefit. 
 These simulations suggest that even when individuals
@@
 systematically varied the benefit~$b$ and the selection strength~$\beta$, as
-shown in Fig.~\ref{fig:cooperation_rate_over_benefit_and_beta_memory_one}.
-When compared to the case of reactive strategies, what we observe
-here is that in the case of limited payoff memory, cooperation rates increase
-with an increasing benefit~$b$ and do not remain stable
-(Fig.~\ref{fig:cooperation_rate_over_benefit_and_beta_memory_one}(a)). However,
-even in this scenario, we still observe that perfect memory consistently results
-in a higher cooperation rate.
+shown in Supplementary~Fig.~\ref{fig:cooperation_rate_over_benefit_and_beta_memory_one}.
+Similar to the case of reactive strategies, we observe that cooperation rates in the setting with limited memory remain stable after $b\!\ge\!4$, see Supplementary~Fig.~\ref{fig:cooperation_rate_over_benefit_and_beta_memory_one}(a).
+Moreover,
+also in this scenario, we observe that perfect memory consistently results
+in more cooperation.
 Regarding the effect of different selection strengths, as depicted in
-Fig.~\ref{fig:cooperation_rate_over_benefit_and_beta_memory_one}(b), the results
+Supplementary~Fig.~\ref{fig:cooperation_rate_over_benefit_and_beta_memory_one}(b), the results
 once again suggest that both perfect and limited payoff memory yield similar
@@
 
-\begin{figure}[!htbp]
+\begin{figure}[t!]
   \centering 
-  \includegraphics[width=\textwidth]{static/memory_one_results_low_benefit.pdf}
+  \includegraphics[width=0.8\textwidth]{static/memory_one_results_low_benefit.pdf}
   \caption{\textbf{Evolution of memory-one strategies when the benefit of cooperation is comparably small.}
@@
 
-\begin{figure}[!htbp]
+\begin{figure}[t!]
   \centering 
-  \includegraphics[width=\textwidth]{static/memory_one_results_high_benefit.pdf}
+  \includegraphics[width=0.8\textwidth]{static/memory_one_results_high_benefit.pdf}
   \caption{\textbf{Evolution of memory-one strategies when the benefit is high.}
-  We consider a similar setup as in Fig.~\ref{fig:memory_one_low_benefit}, but now using a larger benefit value $b\!=\!10$.}\label{fig:memory_one_high_benefit}
+  We consider a similar setup as in Supplementary~Fig.~\ref{fig:memory_one_low_benefit}, but now using a larger benefit value $b\!=\!10$.}\label{fig:memory_one_high_benefit}
 \end{figure}
 
-\begin{figure}[!htbp]
+\begin{figure}[t!]
   \centering 
-  \includegraphics[width=\textwidth]{static/cooperation_rate_over_b_and_beta_memory_one.pdf}
+  \includegraphics[width=0.8\textwidth]{static/cooperation_rate_over_b_and_beta_memory_one.pdf}
   \caption{{\bf Evolution of direct reciprocity for different parameter values for memory-one strategies.}
@@
 
+\clearpage
+
+%%%%%%%%%%%%%%%%%%
+%% FURTHER REPEATED GAMES %% 
+%%%%%%%%%%%%%%%%%%
+
+\section{Further repeated games; Snowdrift game}
+\label{section:furtherrepeatedgames}
+
+So far, we have discussed the prisoner's dilemma. However, this is only one
+example of social dilemmas. Further examples of social dilemmas include the
+snowdrift game~\cite{doebeli:Ecology:2005}, the stag hunt
+game~\cite{pacheco:ProcB:2009}, and the public goods
+game~\cite{pinheiro:PLoSCompBio:2014}. In the following, we discuss how our
+methodology can be applied to other games. More specifically, here we focus on
+the snowdrift game.
+
+Similarly to the prisoner's dilemma, the snowdrift game is a two-player game. In
+each round, individuals can choose whether to cooperate ($C$) or defect ($D$).
+Here, cooperation yields a benefit \(b\) to the cooperator as well as to the
+opposing player and incurs a cost \(c\) if the opponent defects, but only a cost
+\(c/2\) if the opponent cooperates. This results in the following payoffs:
+
+\begin{align}
+	\bordermatrix{%
+		& C & D \cr
+		C &\ b -c/2 &\ b-c \  \cr
+		D &\ b &\ 0\ \cr
+	} . 
+\end{align}\label{eq:snowdrift_payoff}
+If \(b>c>0\), as before, then cooperation is a better strategy than defection if the opponent
+plays defection. On the other hand, if the opponent plays C, then D is still the best
+response. Thus, both strategies can invade when rare, resulting in a mixed
+evolutionarily stable state at which the proportion of cooperators is
+\(1-c/(2b-c)\). It is important to note that in this state, the population
+payoff is smaller than it would be if everybody played C; hence, the SD still
+represents a social dilemma.
+
+In the following, we discuss our methodology for the snowdrift game. We start by
+considering the case of perfect memory, and then we extend our analysis to the
+case of limited memory. More specifically, we look again at which conditions
+cooperation is stochastically stable for the two extreme cases of perfect and
+limited memory. We then present simulation results for all the six cases
+explicitly considered in this work."
+
+\subsection{Stochastic stability of cooperation.}\label{section:stochastic_stability_snowdrift}
+
+{\bf Perfect payoff memory.} We have already discussed the general setup for the
+case of perfect memory in Section~\ref{section:perfect_memory}. Here, we derive
+the conditions under which cooperation is stochastically stable for the
+snowdrift game. As a reminder, we suppose that there is a single mutant,
+\(k\!=\!1\) that adopts the strategy $\alld$, while the residents adopt the
+strategy \(\gtft\). The resulting average distributions between the residents
+and the mutants remain the same as in Section\ref{section:perfect_memory}.
+However, the expected payoffs of the strategies now differ, as we are now
+considering the Snowdrift payoff matrix~\eqref{eq:snowdrift_payoff}.
+Specifically, the expected payoffs are computed as,
+
+\begin{align*}
+  \pi_\gtft =  \frac{N\!-\!2}{N-1} (b - c/2)  \!+\! \frac{1}{N-1}\big(1\!-\!\delta\!+\!\delta q\big)(b - c)
+  \quad \text{and} \quad
+  \pi_\alld  = \big(1\!-\!\delta\!+\!\delta q\big)b.
+\end{align*}
+
+As a consequence, the corresponding ratio of transition probabilities according
+to Eq.~\eqref{eq:LambdaRatio} is now given by,
+
+\begin{align*}
+\frac{\lambda_1^{-}}{\lambda_1^{+}} = e^{-\beta \left( (1-\delta+\delta q)(b - \frac{b-c}{N-1})- \frac{N-2}{N-1}(b-c/2) \right)}
+\end{align*}
+By definition, cooperation is stochastically stable if this ratio exceeds one, which is equivalent to
+\begin{equation}\label{Eq:Condition_PerfectMemory_Snowdrift}
+q<1-\frac{1}{\delta}\cdot \frac{N(c/2)}{(N\!-\!2)b+c}.
+\end{equation}
+For such a strategy to be feasible we require $q\!>\!0$, which implies $\delta
+\!>\! \big((N\!-\!2)b+c\big)/\big(N(c/2)\big)$. In particular, in the limit of
+large populations \(N \!\rightarrow\! \infty\), we obtain that cooperation is
+stochastically stable if  \(q \!<\! 1\! -\! c/(2\delta b)\).
+
+\noindent
+{\bf Limited payoff memory.} Once again, we examine
+the scenario where a single \alld{} mutant can potentially invade a resident
+population of \gtft{} players. However, this time, we consider that individuals
+only take into account the last round of their last interaction.
+Based on the probabilities \(x(u_\resident,u_\mutant)\), where \(u_\resident\)
+represents the payoff of a randomly chosen \gtft{} player and \(u_\mutant\)
+represents the payoff of the \alld{} player, with both \(u_\resident\) and
+\(u_\mutant\) being elements of \(\mathcal{U}\) (see
+Section~\ref{section:limited_memory}), we calculate the ratio of transition
+probabilities as:
+
+\begin{equation*}
+\frac{\lambda^{-}_1}{\lambda^{+}_1} = \frac{ \frac{N - 2}{N - 1}  
+\left(\frac{ 1 - \delta + \delta q}{1+e^{ \beta (c/2)} } 
+  + \frac{ \delta  \left(1 - q \right)}{1+ e^{-  \beta  \left(b - c/2 \right)}} \right)  
+  +  \frac{1}{N-1}  \left(\frac{1 - \delta + \delta q}{1 + e^{\beta c}}
+  + \frac{ \delta  \left(1 - q \right)}{2} \right)}
+  { \frac{N - 2}{N - 1}  
+\left( \frac{ 1 - \delta + \delta q}{1 + e^{-\beta (c/2)}}
++  \frac{ \delta  (1 - q)}{1 +e^{  \beta (b - c/2)}} \right) 
++  \frac{1}{N -1} \left(\frac{ 1 - \delta + \delta q}{1 + e^{-  \beta c}}
++  \frac{ \delta  \left(1 - q \right)}{2} \right)} .
+\end{equation*}
+Cooperation is stochastically stable if $\lambda^-_1/\lambda^+_1\!>\!1$. 
+While one can solve this inequality for $q$, the resulting condition is somewhat lengthy.  
+To obtain a more interpretable condition, we consider the limit of strong selection \(\beta\! \rightarrow\! \infty\) and
+large populations \(N \!\rightarrow \!\infty \). In that case, because $b\!>\!c\!>\!0$, the above ratio simplifies to
+\begin{equation}
+    \frac{\lambda^{-}_1}{\lambda^{+}_1} = \frac{\delta(1 - q)}{1 - \delta + \delta q}.
+\end{equation}
+
+This condition is identical to condition~\eqref{Eq:TransitionRatioSimple} for
+the prisoner's dilemma.
+
+\subsection{Simulation results for reactive strategies}\label{section:simulation_results_snowdrift}
+
 %%%%%%%%%%%%

### 2024-03-13 paper/main.tex
first revision of the manuscript following the reviewers' comments

@@
 
+%% RESULTS: Snowdrift Game
+{\bf Beyond the donation game.}
+So far, we have discussed the donation game, which is the most common
+example of social dilemmas. There are several other examples, which include the
+snowdrift game~\cite{doebeli:Ecology:2005}, the stag hunt
+game~\cite{pacheco:ProcB:2009}, and the public goods
+game~\cite{pinheiro:PLoSCompBio:2014}. However, the pattern we observe is more
+general and still holds for other social dilemmas. To demonstrate this, we
+consider the snowdrift game. In the snowdrift game, individuals can again
+cooperate or defect. However, now if both cooperate, they have the benefit \(b\)
+of getting home while sharing the labor \(c\). Thus for mutual cooperation, they
+receive \(b - \frac{c}{2}\). However, if one defects, they get \(b\) while
+the cooperator gets \(b - c\). By contrast to the donation game, now
+the best action depends on the co-player: to defect if the other cooperates, but
+to cooperate if the other defects.
+
+For most scenarios, we can again derive an analytical condition for when
+cooperation is stochastically stable. Here we focus only on the case of perfect
+and limited memory. The details can be found in the electronic supplementary
+material. In the perfect memory case, we find that cooperation is stochastically
+stable when \(q < 1 - \frac{c}{2 \delta b}\), and interestingly for the limited
+memory case \(q < 1 - \frac{1}{2 \delta}\), which is the same as in
+Eq.~\eqref{Eq:LimitedMemory}. This is because in this analysis, the ordering of
+the payoffs does not matter too much; what matters most is the following.
+
+Apart from the analytical analysis, we evaluate our results with simulations,
+and we do this for all the updating payoff cases considered in this manuscript.
+What we find is that the qualitative results still hold for the snowdrift game.
+Also, in that case, we observe that perfect memory leads to systematically
+higher cooperation rates (Supplementary Figures~8, 9). However, now less memory
+is needed to approximate the classical limit.
+
 %%%%%%%%%%%%
@@
 
-Our main contribution is an application of these ideas to direct reciprocity, which is one of the key mechanisms to explain why unrelated individuals might cooperate~\citep{nowak:Science:2006}. 
-According to this mechanism, cooperation pays if it makes the interaction partner more cooperative in future. 
+Our main contribution in this work is the application of these ideas to direct reciprocity, which is one of the key mechanisms to explain why unrelated individuals might cooperate~\citep{nowak:Science:2006}.
+We discuss the main contributions of our work in the next section.
+
+%%%%%%%%%%%%
+%%  Conclusion %%
+%%%%%%%%%%%%
+
+\section{Conclusion}\label{section:Conclusion}
+
+According to the mechanism of direct reciprocity, cooperation pays if it makes the interaction partner more cooperative in future. 
 To describe which strategies are most effective, the previous theoretical literature assumes that the evolutionary dynamics are driven by the players' expected payoffs~\citep{brauchli:JTB:1999,brandt:JTB:2006,ohtsuki:JTB:2007b,szolnoki:pre:2009b,imhof2010stochastic,van-segbroeck:prl:2012,grujic:jtb:2012,Martinez2012,stewart:pnas:2013,pinheiro:PLoSCB:2014,stewart:games:2015,Baek2016,McAvoy:ProcA:2019,glynatsi:SCR:2020,Schmid:PlosCB:2022,Murase:SciRep:2022}.
@@
     The resulting average cooperation rate depends on which of these two neighborhoods is visited more often.
+    Given the stochastic nature of the evolutionary process, we run long-time simulations to thoroughly explore its behavior and capture significant trends.
     Simulations are run for $T\!=\!10^7$ time steps, using a cost $c\!=\!1$, a continuation probability of $\delta\!=\!0.999$ and a selection strength of $\beta\!=\!1$, in a population of size $N\!=\!100$.}
@@
     Unless explicitly varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$.
-    Simulations are run for $T\!=\!5\times 10^7$ time steps for each parameter combination.}
+    Each point represents a single simulation run. Simulations are run for $T\!=\!5\times 10^7$ time steps for each parameter combination. Each point is a sinlge simualtion run.}
     \label{fig:cooperation_rate_over_benefit_and_beta}
@@
   However, when individuals take into account at least two past interactions -- cases ({\it iii}) to ({\it vi}) -- evolving cooperation rates are close to this optimum. 
-  Baseline parameters are the same as in \FigDependenceParameters.}
+  Simulations are run for $T\!=\!10^7$ time steps, using a cost $c\!=\!1$, a continuation probability of $\delta\!=\!0.999$ and a selection strength of $\beta\!=\!1$, in a population of size $N\!=\!100$.}
 \label{fig:cooperation_rate_all_updating_payoffs}

### 2024-03-14 paper/main.tex
cleaning main paper

@@
 \usepackage{lineno}
-% \linenumbers
+\linenumbers
 \smallskip % Erlaubt kleine Abstaende zwischen Paragraphen, falls es dem Seitenlayout hilft
@@
 %% INTRO: Basic description of our model and our findings 
-
+Suggestion one:
 To explore that question, we theoretically study imitation dynamics in the repeated prisoner's dilemma, using two extreme scenarios. 
@@
 
-%% INTRO: Summary and conclusions
-
-% Overall, these findings suggest that while memory is important, already minimal payoff information may suffice for the evolution of direct reciprocity based on social learning. 
-% They also suggest that the classical model of reciprocity (based on expected payoffs) can often be interpreted as a useful approximation to more realistic models that include cognitive constraints.
-
-\todo{To address the reviewer's comment, we can either delete the last paragraph or rephrase the second-to-last paragraph as follows:}
-
+Suggestion two:
 To explore this question, we theoretically study imitation dynamics in the

### 2024-03-15 paper/si.tex

@@
 \def\alld{\texttt{ALLD}}
+\def\allc{\texttt{ALLC}}
 \def\tft{\texttt{TFT}}
@@
 
-So far, we have discussed the prisoner's dilemma. However, this is only one
-example of social dilemmas. Further examples of social dilemmas include the
-snowdrift game~\cite{doebeli:Ecology:2005}, the stag hunt
-game~\cite{pacheco:ProcB:2009}, and the public goods
-game~\cite{pinheiro:PLoSCompBio:2014}. In the following, we discuss how our
-methodology can be applied to other games. More specifically, here we focus on
-the snowdrift game.
-
-Similarly to the prisoner's dilemma, the snowdrift game is a two-player game. In
-each round, individuals can choose whether to cooperate ($C$) or defect ($D$).
-Here, cooperation yields a benefit \(b\) to the cooperator as well as to the
-opposing player and incurs a cost \(c\) if the opponent defects, but only a cost
-\(c/2\) if the opponent cooperates. This results in the following payoffs:
+So far, we have discussed the repeated donation game, which is only one example
+of social dilemmas. In the following section, we explore how our methodology can
+be applied to other games, with a focus on the snowdrift game.
+
+Similar to the donation game, the snowdrift game is a two-player game where
+individuals choose to cooperate (\(C\)) or defect (\(D\)) each round.
+Cooperation yields a benefit \(b\) to both players, incurring a cost \(c\) if
+the opponent defects, but only \(c/2\) if the opponent cooperates. This results
+in the following payoff matrix
 
@@
 		& C & D \cr
-		C &\ b -c/2 &\ b-c \  \cr
-		D &\ b &\ 0\ \cr
+		C & b - \frac{c}{2} & b - c  \cr
+		D & b & 0 \cr
 	} . 
 \end{align}\label{eq:snowdrift_payoff}
-If \(b>c>0\), as before, then cooperation is a better strategy than defection if the opponent
-plays defection. On the other hand, if the opponent plays C, then D is still the best
-response. Thus, both strategies can invade when rare, resulting in a mixed
-evolutionarily stable state at which the proportion of cooperators is
-\(1-c/(2b-c)\). It is important to note that in this state, the population
-payoff is smaller than it would be if everybody played C; hence, the SD still
-represents a social dilemma.
-
-In the following, we discuss our methodology for the snowdrift game. We start by
-considering the case of perfect memory, and then we extend our analysis to the
-case of limited memory. More specifically, we look again at which conditions
-cooperation is stochastically stable for the two extreme cases of perfect and
-limited memory. We then present simulation results for all the six cases
-explicitly considered in this work."
+
+In the following subsections, we discuss how our results extend to the snowdrift
+game. In Section~\ref{section:stochastic_stability_snowdrift}, we investigate
+under which conditions cooperation is stochastically stable for both perfect and
+limited memory scenarios. In Section~\ref{section:simulation_results_snowdrift},
+we present simulation results for all six cases considered in this study.
 
@@
 However, the expected payoffs of the strategies now differ, as we are now
-considering the Snowdrift payoff matrix~\eqref{eq:snowdrift_payoff}.
+considering the snowdrift payoff matrix~\eqref{eq:snowdrift_payoff}. \\
+
+\noindent
 Specifically, the expected payoffs are computed as,
@@
 large populations \(N \!\rightarrow\! \infty\), we obtain that cooperation is
-stochastically stable if  \(q \!<\! 1\! -\! c/(2\delta b)\).
+stochastically stable if  \(q \!<\! 1\! -\! c/(2\delta b)\). \\
 
@@
 Cooperation is stochastically stable if $\lambda^-_1/\lambda^+_1\!>\!1$. 
-While one can solve this inequality for $q$, the resulting condition is somewhat lengthy.  
-To obtain a more interpretable condition, we consider the limit of strong selection \(\beta\! \rightarrow\! \infty\) and
+We consider the limit of strong selection \(\beta\! \rightarrow\! \infty\) and
 large populations \(N \!\rightarrow \!\infty \). In that case, because $b\!>\!c\!>\!0$, the above ratio simplifies to
-\begin{equation}
+\begin{equation}\label{Eq:Condition_LimitedMemory_Snowdrift}
     \frac{\lambda^{-}_1}{\lambda^{+}_1} = \frac{\delta(1 - q)}{1 - \delta + \delta q}.
@@
 
-This condition is identical to condition~\eqref{Eq:TransitionRatioSimple} for
-the prisoner's dilemma.
+Interestingly, this condition is identical to
+condition~\eqref{Eq:TransitionRatioSimple} for the donation game. To understand
+this result, we first need to discuss the difference between the two games. The
+difference between the two games lies in the relationship between the possible
+payoffs. More specifically, in the donation game, the payoffs satisfy the
+relationship \(T > R > P > S\), whereas in the snowdrift game, it's \(T > R > S
+> P\).
+Now, note that for the above condition, the relationship between all the
+possible payoffs is not important; what is important is the relationship between
+all the feasible pairs of payoffs. In a population of \gtft{} strategies with a
+single mutant of \alld{}, the feasible pairs of payoffs are \((R, T)\), \((R,
+P)\), \((S, T)\), and \((P, P)\). In both games, the payoff \(T\) is higher than
+\(S\) and \(R\), and \(R\) is higher than the punishment payoff, thus the
+condition remains unchanged.
 
 \subsection{Simulation results for reactive strategies}\label{section:simulation_results_snowdrift}
+
+To investigate whether the previous static observations accurately depict the
+dynamics of evolving populations, we once again rely on simulations. Firstly, we
+conducted separate simulations for perfect and limited payoff memory. In each
+scenario, we examined both low and high benefits of cooperation (\(b/c=2\) and
+\(b/c=10\), respectively). Supplementary Fig.
+\ref{fig:expected_and_stochastic_for_snowdrift} illustrates the conditional
+cooperation probabilities \(p\) and \(q\) of the evolving populations. Overall,
+we find that the evolving populations are more cooperative compared to those of
+the donation game. There is no cluster around the neighborhood of \alld{} as
+they are easily outperformed by more cooperative strategies. Moreover, the
+evolving cooperation rates are high for all the cases. Even in the limited
+payoff memory for low benefit, the evolving cooperation is at 71\%. Thus, the
+in the snowdrift game, the memory requirement for cooperation is
+even more modest compared to the donation game.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=.75\textwidth]{static/snowdrift_game_results.pdf}
+  \caption{{\bf Evolutionary dynamics under perfect and limited payoff memory.}
+The four panels depict the outcome of evolutionary simulations for the repeated
+snowdrift game with either a low or high benefit of cooperation. In the top
+panels, individuals have perfect payoff memory, meaning they consider all their
+past interactions against all population members, taking every turn of each
+repeated game into account. In the bottom panels, individuals have limited
+payoff memory, considering only their very last interaction against one specific
+population member, and taking into account only one round of the repeated game.
+Colors represent how often the respective region of the strategy space is
+visited over time.
+In all cases, there is one dominant cluster, and that is the cluster of
+conditionally cooperative strategies with \(p \approx 1\), where \(q\) satisfies
+the constraints \eqref{Eq:Condition_PerfectMemory_Snowdrift} and
+\eqref{Eq:Condition_LimitedMemory_Snowdrift}, respectively.
+Simulations are run for \(T = 10^7\) time steps, using a cost \(c = 1\), a
+continuation probability of \(\delta = 0.999\), and a selection strength of
+\(\beta = 1\), in a population of size \(N = 100\).}
+\label{fig:expected_and_stochastic_for_snowdrift}
+\end{figure}
+
+Furthermore, to understand the impact of payoff memory on the snowdrift game
+even more, we performed simulations for the four additional scenarios considered
+in this work. As a reminder, in the first scenario, players recall the payoffs
+they obtained in the last two rounds against a single co-player. In the second
+scenario, players recall their last-round payoffs against two co-players. In the
+third scenario, they recall the last two rounds against two co-players. Finally,
+in the last scenario, players update based on the average payoff they receive
+over all rounds with a single co-player (further extensions are possible, but we
+do not explore them here). In Supplementary
+Fig.~\ref{fig:cooperation_rate_all_updating_payoffs_snowdrift}, we show the
+evolving cooperation rates for all these scenarios, as well as for the limited
+and perfect memory cases.
+
+We observe that main pattern of our results remains unchanged: the perfect
+memory case always overestimates the cooperation rate. Compared to the baseline
+case with a single round recalled, even a minimal increase in the players'
+payoff memory further increases the cooperation rate. In all four scenarios with
+extended memory, we observe similar cooperation rates, and they quickly approach
+the rates observed under perfect memory. This happens at a faster rate compared
+to the donation game.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\textwidth]{static/snowdrift_game_more_memory.pdf}
+  \caption{{\bf Average cooperation rates in the snowdrift game for different payoff memories.}
+  We vary how much information individuals take into account when updating their strategies.
+  From left to right, we consider the following cases:
+  ({\it i}) Updating occurs based on expected payoffs (perfect memory).
+  ({\it ii}) It occurs based on the last round of one interaction (limited memory).
+  ({\it iii}) Based on the last round of two interactions.
+  ({\it iv}) Based on the last two rounds of one interaction.
+  ({\it v}) Based on the last two rounds of two interactions.
+  ({\it vi}) Based on the average payoff of one interaction.
+  Simulations are run either for a comparably low benefit of cooperation~($b/c\!=\!2$), or for a high benefit~($b/c\!=\!10$). 
+  We observe that perfect memory always yields the highest cooperation rate, and limited memory yields the lowest.
+  However, in the snowdrift game, the difference between the two extreme scenarios is not large, as we have already discussed.
+  When individuals take into account a bit more information—cases (iii) to (vi)—evolving
+  cooperation rates quickly approach the optimum of perfect memory. This increase happens faster compared to the donation game.
+  Simulations are run for $T\!=\!10^7$ time steps, using a cost $c\!=\!1$, a continuation probability of $\delta\!=\!0.999$ and a selection strength of $\beta\!=\!1$, in a population of size $N\!=\!100$.}
+\label{fig:cooperation_rate_all_updating_payoffs_snowdrift}
+\end{figure}
 
@@
 \bibliography{bibliography}
-}
+}}
 

### 2024-03-18 paper/main.tex
CH: Revising the main text

--- 2024-03-11 paper/main.tex
@@
 Even in the most extreme case that they only take into account their very last interaction, we find that cooperation can still evolve.
-However, such individuals adopt less generous strategies, and they tend to cooperate less often than in the classical setup with average payoffs.
+However, such individuals adopt less generous strategies, and they cooperate less often than in the classical setup with average payoffs.
 Interestingly, once individuals remember the payoffs of two or three recent interactions, cooperation rates quickly approach the classical limit.
 These findings contribute to a literature that explores which kind of cognitive capabilities are required for reciprocal cooperation.
-While our results suggest that some rudimentary form of payoff memory is necessary, it already suffices to remember a few interactions.
+While our results suggest that some rudimentary form of payoff memory is necessary, it suffices to remember a few interactions.
 \end{abstract}
@@
 ~\\
-{\it Keywords:} Evolution of cooperation; direct reciprocity; repeated prisoner's dilemma; social learning; evolutionary dynamics
+{\it Keywords:} Evolution of cooperation; evolutionary dynamics; direct reciprocity; repeated prisoner's dilemma; social learning
 
@@
 In contrast, in models of cultural evolution, individuals need to be aware of the different strategies present in the population, and they need to identify those strategies with a higher payoff. 
-As a consequence, evolutionary outcomes may depend on how easily different behaviors can be learned~\citep{Chatterjee:JTB:2012}, and on how easy payoff comparisons are.
+As a consequence, evolutionary outcomes may depend on how easily different behaviors can be learned~\citep{Chatterjee:JTB:2012}, as well as on how readily payoffs can be compared.
 
@@
 They may use strategies such as Tit-for-Tat~\citep{rapoport:book:1965,axelrod1981evolution} or Generous Tit-for-Tat~\citep{molander:jcr:1985,Nowak1992tit} to preferentially cooperate with other cooperators.
-Such conditional strategies approximate human behavior fairly well~\citep{fischbacher:EconL:2001,Rand:TCS:2013,DalBo:AER:2019,Montero-Porras:SciRep:2022,Rossetti:ETH:2023} and they have also been documented in several other species~\citep{Carter:PRSB:2013,Schweinfurth:AnBehav:2019,Voelkl:PNAS:2015} -- although direct reciprocity is generally more difficult to demonstrate in animals~\citep{CluttonBrock:Nature:2009,Silk:CurrentBiology:2013,taborsky:CurrentBiology:2013}.
+Such conditional strategies approximate human behavior fairly well~\citep{fischbacher:EconL:2001,Rand:TCS:2013,DalBo:AER:2019,Montero-Porras:SciRep:2022,Rossetti:ETH:2023} and they have also been documented in several other species~\citep{Carter:PRSB:2013,Schweinfurth:AnBehav:2019,Voelkl:PNAS:2015}--although direct reciprocity is generally more difficult to demonstrate in animals~\citep{CluttonBrock:Nature:2009,Silk:CurrentBiology:2013,taborsky:CurrentBiology:2013}.
 However, at the outset, it is not clear how easy it is to {\it learn} reciprocal strategies by social imitation. 
@@
 Rather, it is determined by how well this strategy fares over an entire sequence of rounds, against many different population members. 
-In practice, such information might be both difficult to obtain and to process.
+In practice, such information might be difficult both to obtain and to process.
 
@@
 
-Most models of direct reciprocity abstract from these difficulties~\citep{brauchli:JTB:1999,brandt:JTB:2006,ohtsuki:JTB:2007b,szolnoki:pre:2009b,imhof2010stochastic,van-segbroeck:prl:2012,grujic:jtb:2012,Martinez2012,stewart:pnas:2013,pinheiro:PLoSCB:2014,stewart:games:2015,Baek2016,hilbe:PNAS:2017,McAvoy:ProcA:2019,glynatsi:SCR:2020,Schmid:PlosCB:2022,Murase:SciRep:2022,Cooney:BMB:2022,Chen:PNASNexus:2023}.
-They just assume individuals can easily copy the strategies of others. 
-Similarly, they just assume that updating decisions are based on the strategies' average (or expected) payoffs, which are based on all rounds and all interactions. 
+Most models of direct reciprocity do not address these difficulties~\citep{brauchli:JTB:1999,brandt:JTB:2006,ohtsuki:JTB:2007b,szolnoki:pre:2009b,imhof2010stochastic,van-segbroeck:prl:2012,grujic:jtb:2012,Martinez2012,stewart:pnas:2013,pinheiro:PLoSCB:2014,stewart:games:2015,Baek2016,McAvoy:ProcA:2019,glynatsi:SCR:2020,Schmid:PlosCB:2022,Murase:SciRep:2022,Cooney:BMB:2022,Chen:PNASNexus:2023} and, instead, assume individuals can easily copy the strategies of others.
+Similarly, they assume that updating decisions are based on the strategies' average (or expected) payoffs, which are based on all rounds and all interactions. 
 These assumptions create a curious inconsistency in how models represent an individual's cognitive abilities. 
 On the one hand, when playing the game, individuals are often assumed to have restricted memory. 
-Respective studies typically assume that individuals make their decisions each round based on the outcome of the last round only~\citep[with only a few exceptions, see Refs.][]{Hauert1997,van-veelen:PNAS:2012,Stewart2016,Li:NatCS:2022,Murase:PLoSCompBio:2023a}.
+Respective studies typically assume that individuals make their decisions each round based on the outcome of the last round only~\citep[with only a few exceptions, see Refs.][]{Hauert1997,van-veelen:PNAS:2012,Stewart2016,hilbe:PNAS:2017,Li:NatCS:2022,Murase:PLoSCompBio:2023a}.
 Yet when learning new strategies, individuals are assumed to remember (or compute) each others' precise average payoff across many rounds and many interaction partners. 
@@
 %% INTRO: Basic description of our model and our findings 
-
 To explore that question, we theoretically study imitation dynamics in the repeated prisoner's dilemma, using two extreme scenarios. 
-The first scenario is the usual modeling approach. 
+The first scenario is the classical modeling approach. 
 Here, individuals update their strategies based on their expected payoffs. 
@@
 Yet moderate levels of cooperation can still evolve. 
-Moreover, as we increase the individuals' payoff memory to include the last two or three one-shot payoffs, cooperation rates quickly approach the rates observed in the classical baseline case.
-
-%% INTRO: Summary and conclusions
-
-% Overall, these findings suggest that while memory is important, already minimal payoff information may suffice for the evolution of direct reciprocity based on social learning. 
-% They also suggest that the classical model of reciprocity (based on expected payoffs) can often be interpreted as a useful approximation to more realistic models that include cognitive constraints.
-
-\todo{To address the reviewer's comment, we can either delete the last paragraph or rephrase the second-to-last paragraph as follows:}
-
-To explore this question, we theoretically study imitation dynamics in the
-repeated prisoner's dilemma, employing two extreme scenarios. In the first
-scenario, which follows the conventional modeling approach, individuals update
-their strategies based on their expected payoffs. We contrast this model with an
-alternative scenario where individuals base their strategy updates solely on the
-very last (one-shot) payoff they obtained. We observe that individuals with
-limited payoff memory tend to adopt less generous strategies, yet moderate
-levels of cooperation can still emerge. This suggests that while memory plays a
-significant role, minimal payoff information may suffice for the evolution of
-direct reciprocity through social learning. Moreover, increasing individuals'
-payoff memory to include the last two or three one-shot payoffs leads to
-cooperation rates quickly approaching those observed in the classical baseline
-case. This implies that the classical reciprocity model, based on expected
-payoffs, often serves as a useful approximation to more complex models
-incorporating cognitive constraints.
+Moreover, as we increase the individuals' payoff memory to include the last two or three one-shot payoffs, cooperation rates quickly approach the rates observed in the classical baseline.
 
@@
 Thus, the players' payoffs in a single round are given by the matrix
-\begin{align}
+\begin{equation}
 \bordermatrix{%
@@
 } .
-\end{align}
+\end{equation}
 In particular, payoffs take the form of a prisoner's dilemma:
-Mutual cooperation yields a better payoff than mutual defection ($b\!-\!c\!>\!0$), but each player individually prefers to defect independent of the co-player's action ($b\!>\!b\!-\!c$ and $0\!>\!-c$).
-To incorporate that individuals interact repeatedly, we assume that after each round, there is a constant continuation probability $\delta$ of interacting for another round. 
+Mutual cooperation yields a better payoff than mutual defection ($b\!-\!c\!>\!0$), but each player individually prefers to defect, independent of the co-player's action ($b\!>\!b\!-\!c$ and $0\!>\!-c$).
+To incorporate repetition, we assume that after each round, there is a constant continuation probability $\delta$ of interacting for another round. 
 For $\delta\!=\!0$, we recover the case of a conventional (one-shot) prisoner's dilemma. 
@@
 This process assumes that at regular time intervals, one population member is randomly selected, and given the chance to revise its strategy.
-We refer to this player as the `learner'. 
+We refer to this player as the ``learner.'' 
 With probability $\mu$ (reflecting a mutation rate), the learner simply adopts a random strategy (all reactive strategies have the same probability to be chosen). 
-With the converse probability $1\!-\!\mu$, the learner randomly picks a `role model' from the population. 
+With the converse probability $1\!-\!\mu$, the learner randomly picks a ``role model'' from the population. 
 The learner then compares its own payoff $\pi_\learner$ from the repeated game to the role model's payoff $\pi_\rolemodel$. 
@@
 The selection strength parameter $\beta\!\ge\!0$ indicates how sensitive players are to payoff differences. 
-For $\beta\!=\!0$, payoff differences are irrelevant, and the learner simply adopts the role model's strategy with probability one half.
+For $\beta\!=\!0$, payoff differences are irrelevant, and the learner simply adopts the role model's strategy with probability $1/2$.
 As the selection strength~$\beta$ increases, players are increasingly biased to imitate the role model only if it has the higher payoff.
@@
 Herein, we compare this model of perfect payoff memory to a model with limited payoff memory. 
-In that model, the players' payoffs $\pi_\learner$ and $\pi_\rolemodel$ are taken to be the payoffs that each player received in their very last round prior to making social comparisons. 
+In the latter model, the players' payoffs $\pi_\learner$ and $\pi_\rolemodel$ are taken to be the payoffs that each player received in their very last round prior to making social comparisons. 
 That is, players only consider the very last repeated game they participated in, and there they only take into account the outcome of the very last round (\FigBaseResults, lower left panel). 
@@
 
-Both in the case of perfect and limited memory, we iterate the elementary strategy updating step described above for many time steps. 
+Both in the case of perfect and limited memory, we iterate the elementary strategy update step described above many times. 
 This gives rise to a stochastic process that describes which strategies players adopt over time. 
-We explore the dynamics of this process mathematically and with computer simulations.
+Provided the selection strength is finite, this process satisfies the mathematical property of being ergodic. 
+This implies that in the long run, the time average of the players' cooperation rates converges, and that this limit is independent of the initial population. 
+We explore the dynamics of this process mathematically and with numerical simulations.
+These simulations are run sufficiently long for convergence to occur (we numerically checked that independent runs are within a 1\% error tolerance). 
 For the results presented in the following, we assume that mutations are rare (\(\mu\!\rightarrow\! 0\)). 
-This assumption is fairly common in evolutionary game theory, because it makes some computations more efficient~\citep{fudenberg:JET:2006,wu:JMB:2012,mcavoy:jet:2015}, and because the results can be interpreted more easily.
+This assumption is fairly common in evolutionary game theory, both because it makes some computations more efficient~\citep{fudenberg:JET:2006,wu:JMB:2012,mcavoy:jet:2015} and because the results can be interpreted more easily.
 However, in Section~3 of the \esm{} we show that our main results continue to hold for strictly positive mutation rates.
@@
 {\bf Stability of cooperative populations.}
-To get some intuition for the differences between perfect and limited payoff memory, we first analyze when cooperation is stable in either scenario.
-To this end, we consider a resident population in which all players but one adopt a strategy of Generous Tit-for-Tat, \gtft{} $=\!(1,1,q)$. 
-The remaining mutant player adopts \alld. 
-We say {\it cooperation is stochastically stable} if the single mutant is more likely to imitate the residents than vice versa. 
+To get some intuition for the differences between perfect and limited payoff memory, we first analyze when full cooperation is stable in either scenario.
+For a population to be fully cooperative, players need to cooperate in the first round (that is, $y\!=\!1$) and after any round with mutual cooperation (that is, $p\!=\!1$). 
+We refer to such a strategy as Generous Tit-for-Tat, \gtft{} $=\!(1,1,q)$. 
+To explore whether such a population consisting of \gtft{} players is stable, we introduce a single mutant who adopts \alld. 
+We say {\it (full) cooperation is stochastically stable} if the single mutant is more likely to imitate the residents than vice versa. 
 For simplicity, we consider a large population~($N\!\rightarrow\!\infty$) and strong selection~($\beta\!\rightarrow\!\infty$).
@@
 On the other hand, the defecting mutant only interacts with residents. 
-Given the residents' strategy, the mutant receives a benefit in the first round, and in every subsequent round with probability $q$. 
+Given the residents' strategy, the mutant receives a benefit~$b$ in the first round, and in every subsequent round with probability $q$. 
 As a result, the mutant's expected payoff is $\pi_\alld \!=\! (1\!-\!\delta\!+\!\delta q)b$. 
-For perfect payoff memory, the requirement for cooperation to be stochastically stable reduces to the condition $\pi_\gtft > \pi_\alld$. 
-This yields
+For perfect payoff memory, the requirement for cooperation to be stochastically stable reduces to the condition $\pi_\gtft > \pi_\alld$, which yields
 \begin{equation} \label{Eq:PerfectMemory}
@@
 In particular, we recover the previous observation that $q\!=\!1-c/(\delta b)$ is the maximum generosity that cooperators should have~\citep{molander:jcr:1985,Nowak1992tit,Schmid:NHB:2021}. 
-Because $q\!\ge\!0$, we also conclude that cooperation can only be stable if $\delta \!>\! c/b$.
-Again, this condition for the feasibility of direct reciprocity is the condition found in the literature~\citep{nowak:Science:2006}.
+Moreover, because $q$ needs to be non-negative, we also conclude that cooperation can only be stable if $\delta \!>\! c/b$.
+Again, this condition for the feasibility of direct reciprocity already exists in the literature~\citep{nowak:Science:2006}.
 
@@
 For the defecting mutant, there are two possibilities.
-({\it i})~ If the mutant's co-player happens to cooperate in the last round, the mutant receives $\pi_\alld\!=\!b$.
+If the mutant's co-player happens to cooperate in the last round, the mutant receives $\pi_\alld\!=\!b$.
 This case occurs with probability $1\!-\!\delta\!+\!\delta q$.
-({\it ii})~If the co-player defects in the last round, the mutant receives $\pi_\alld=\!0$. 
+On the other hand, if the co-player defects in the last round, the mutant receives $\pi_\alld=\!0$.
 This occurs with the converse probability $\delta(1\!-\!q)$.
@@
 Because $b\!-\!c\!>\!0$, mutants tend to imitate the resident in the second case. 
-Cooperation is stochastically stable if the first case is less likely than the second. 
-This yields the condition
+Cooperation is stochastically stable if the first case is less likely than the second, which yields the condition
 \begin{equation} \label{Eq:LimitedMemory}
@@
 \end{equation}
-Interestingly, this condition no longer depends on the exact payoff values $c$ and $b$. 
+Interestingly, this condition no longer depends on the exact payoff values, $b$ and $c$. 
 This independence arises because of our assumption of strong selection, in which case only the payoff ordering $b\!>\!c\!>\!0$ matters.
-Because $q$ is non-negative, condition~\eqref{Eq:LimitedMemory} can only be satisfied if $\delta\!>\!1/2$.
-That is, players need to interact in more than two rounds in expectation.
+Because $q$ is non-negative, condition~\eqref{Eq:LimitedMemory} can be satisfied only if $\delta\!>\!1/2$.
+That is, players need to interact in more than two rounds, on average.
 
@@
 By comparing the two cases, we find that payoff memory affects whether a conditionally cooperative strategy ($1,1,q$) is viable. 
-With perfect memory, the maximum generosity $q$ needs to satisfy Eq.~\eqref{Eq:PerfectMemory}.
+With perfect memory, the maximum generosity $q$ must satisfy Eq.~\eqref{Eq:PerfectMemory}.
 In particular, this generosity can become arbitrarily large, provided the game's benefit-to-cost ratio~$b/c$ and the continuation probability~$\delta$ are sufficiently large. 
@@
 {\bf Evolutionary dynamics of reciprocity.}
-To explore whether the previous static observations describe the dynamics of evolving populations, we turn to simulations.
+The previous static observations describe whether full cooperation, once established, can be sustained. 
+In a next step, we explore in which cases cooperation actually evolves. 
+To this end, we turn to simulations.
 We have run separate simulations for perfect and limited payoff memory. 
@@
 The second region corresponds to a thin strip of cooperative strategies with $(p,q)\!\approx\!(1,q)$. 
-Within this strip, we observe that most strategies satisfy the constraints on $q$ imposed by the inequalities~\eqref{Eq:PerfectMemory} and~\eqref{Eq:LimitedMemory}. 
+Within this strip, we observe that most strategies satisfy the constraints on $q$ suggested by the inequalities~\eqref{Eq:PerfectMemory} and~\eqref{Eq:LimitedMemory}. 
 That is, with perfect memory, most evolving strategies have $q\!<\!1\!-\!c/b$, whereas with limited payoff memory, most strategies have $q\!<\!1/2$. 
@@
 If the resident's generosity~$q$ violates the conditions~\eqref{Eq:PerfectMemory} and~\eqref{Eq:LimitedMemory}, defectors can re-invade, and the cycle starts again. 
-The relative time spent near \alld{} and near the strip of conditionally cooperative strategies depends on the considered memory setting (\FigInvasionAnalysis, depicting the case of high benefits). 
+The relative time spent near \alld{} and near the strip of conditionally cooperative strategies depends on the memory (\FigInvasionAnalysis, depicting the case of high benefits). 
 For perfect memory, we find that \alld{} is replaced relatively quickly by more cooperative strategies. 
@@
 In contrast, for limited memory, \alld{} is more robust, resisting on average 798 mutant strategies. 
-This picture reverses when we consider an initial population that adopts \gtft. 
+This picture reverses for an initial population that adopts \gtft. 
 Such populations are much more robust under perfect memory than they are under limited memory. 
@@
 Interestingly, for limited payoff memory, the cooperation rate remains stable at approximately 50\% once \(b \!\ge\! 5\).
-This again reflects our earlier observation that the feasibility of cooperation in this scenario is largely independent of the exact values of $b$ and $c$, as described by Eq.~\eqref{Eq:LimitedMemory}.
-With respect to the effect of different selection strengths, \FigDependenceParameters{\it b} suggests that both perfect and limited payoff memory yields similar cooperation rates for weak selection \(\beta \!<\!
+This finding is reminiscent of our earlier static results, which also suggested that changes in the game's benefits may have a negligible role on the stochastic stability of full cooperation.  
+With respect to the effect of different selection strengths, \FigDependenceParameters{\it b} suggests that both perfect and limited payoff memory yield similar cooperation rates for weak selection \(\beta \!<\!
 1\).
 Beyond weak selection, increasing selection has a positive effect under perfect payoff memory, but a negative effect under limited payoff memory.\\
-
-%% RESULTS: Memory-1 strategies %%
-
-\noindent
-{\bf Beyond reactive strategies.} While the results presented in the main text focus on reactive strategies, the patterns we observe do not seem to depend on the considered strategy space. 
-To illustrate this point in more detail, in the \esm{} we consider the dynamics among memory-1 strategies. 
-Here, players take into account both their co-player's and their own last move, see Refs.~\citep{nowak:Nature:1993,imhof:JTB:2007}. 
-Also in that case, we observe that perfect memory leads to systematically higher cooperation rates (Supplementary Figures~5,6).
-Again, this advantage of perfect memory is particularly pronounced for strong selection, or when there is a high benefit of cooperation (\FigMemoryOneParameters).\\
 
@@
 So far, we have taken a rather extreme interpretation of limited payoff memory. 
-In the respective scenario, we assumed that individuals update their strategies based on their experience in a single round of the prisoner's dilemma, against a single co-player. 
+We assumed that individuals update their strategies based on their experience in a single round of the prisoner's dilemma, against a single co-player. 
 The limited payoff memory framework can be expanded in various ways. 
@@
 Finally, in the last scenario, players update based on the average payoff they
-receive over all rounds with a single co-player (further extensions are possible, but we do not explore them here).
+receive over all rounds with a single co-player. (Further extensions are possible, but we do not explore them here.)
 
@@
 
-We complement these analytical results with additional simulations, see \FigHigherMemory.
+We complement these analytical results with additional simulations; see \FigHigherMemory.
 We observe that a minimal increase in the players' payoff memory (compared to the baseline case with a single round recalled) can promote cooperation considerably. 
-Specifically, in all four scenarios with extended memory, we see similar cooperation rates, and they approach the rates observed under perfect memory. 
-These results suggest that while it takes {\it some} payoff memory to sustain substantial cooperation rates, the requirements on memory seem to be rather modest. 
-Already remembering a few interactions, either with the same co-player or across different co-players, may provide players with enough information to adopt reciprocal strategies.
+Specifically, in all four scenarios with extended memory, we see similar cooperation rates that approach the rates observed under perfect memory. 
+These results suggest that while it takes {\it some} payoff memory to sustain substantial cooperation rates, the memory requirements are rather modest. 
+Already remembering a few interactions, either with the same co-player or across different co-players, may provide players with enough information to adopt reciprocal strategies.\\
+
+%% RESULTS: Memory-1 strategies %%
+
+\noindent
+{\bf Beyond reactive strategies and the donation game.} 
+While the results presented in the main text focus on reactive strategies and the donation game, the observed patterns hold more generally.
+To illustrate this point in more detail, in the \esm{} we first consider the dynamics of the donation game among memory-1 strategies. 
+Here, players take into account both their co-player's and their own last move, see Refs.~\citep{nowak:Nature:1993,imhof:JTB:2007}. 
+Also in that case, we also observe that perfect memory leads to systematically higher cooperation rates (Supplementary Figures~5,6).
+Again, this advantage of perfect memory is particularly pronounced for strong selection, or when there is a high benefit of cooperation (\FigMemoryOneParameters).
+
+%% RESULTS: Snowdrift Game
+In a second step, we also analyze the evolutionary dynamics of a different social dilemma, the snowdrift game~\cite{doebeli:Ecology:2005}. 
+In this game, individuals can again cooperate or defect. 
+However, now they get the benefit $b$ if at least one of them cooperates. 
+Cooperators in turn pay a cost of $c$ or $c/2$, depending on whether they are the only one to cooperate or not. 
+Compared to the donation game, the snowdrift game represents a weaker form of a social dilemma, because defection is no longer the dominant action in any single round~\citep{nowak:JTB:2012}.
+Again, we derive analytical results in the case of large populations and strong selection. 
+For perfect memory, cooperation is stochastically stable when \(q < 1 \!-\! c/(2 \delta b)\). 
+Interestingly, for limited payoff memory, we obtain the same condition as for the donation game, \(q < 1 \!-\! 1/(2 \delta)\). 
+We explain this invariance in detail in the \esm{}: compared to the donation game, the ordering of one-shot payoffs in the snowdrift game only differs in one instance (the {\it sucker's payoff} now exceeds the {\it punishment payoff}); however, this instance is irrelevant for the stochastic stability of cooperation.
+Further evolutionary simulations suggest that again, perfect memory leads to higher cooperation rates than limited memory; but now even limited memory can result in substantial cooperation (Supplementary Figures~8, 9).
 
@@
 Non-normalized payoffs interfere with selection intensity ($\beta$) in models of social imitation, which is (presumably) why they appear less frequently in the literature.
-On that point, we note that many of the differences between realized and expected payoffs disappear in the limit of weak selection \citep{mcavoy:NHB:2020}.
+On that point, we note that differences between realized and expected payoffs disappear in the limit of weak selection, which is known in a general setting \citep{mcavoy:NHB:2020}.
 Non-weak selection can introduce substantial differences between models with realized and expected payoffs \citep{mcavoy:PLOSCB:2021}, which is especially important to understand in models of social systems with cultural transmission \citep{cavalli1981cultural}.
@@
 To the extent that strategies are learned (not inherited), this assumption seems to impose rather stringent requirements on the individuals' cognitive abilities. 
-In the most extreme case, this would require individuals to remember (or compute) their payoffs against all population members, for all possible ways in which their repeated games may unfold. 
+In the most extreme case, individuals would have to remember (or compute) their payoffs against all population members, for all possible ways in which their repeated games may unfold. 
 This assumption introduces a curious inconsistency in how these models represent an individual's cognitive abilities. 
@@
 
-To address these questions, we combine analytical methods and computer simulations. 
+To explore the effect of payoff memory, we only considered the simplest model of reciprocity. 
+This leaves several possible directions for future work. 
+For example, for both our analytical analysis and our simulations, we assumed that populations are {\it well-mixed}. 
+As a result, any population member is equally likely to be a given player's last co-player. 
+A similar analysis for structured populations would be desirable, as population structure can often promote cooperation without an explicit need for reciprocity~\citep{szabo:PR:2007,van-veelen:PNAS:2012}.
+Similarly, we assumed that after identifying a suitable role model, players can easily imitate that role model's strategy. 
+In the context of our study, this might be a particular simplification, because players with limited payoff memory may also have less accurate information to infer a co-player's strategy. 
+ Finally, we have taken a player's payoff memory to be a given parameter of the model. 
+ Instead, future models could explore how that payoff memory might co-evolve along with the players' strategies \citep[similar to previous work on the evolution of the players' feasible strategy sets, e.g.][]{Stewart2016,Schmid:PlosCB:2022,Murase:PLoSCompBio:2023a}.
+ Here, individuals might experience an interesting trade-off. 
+ Better payoff-memory would help them to make better decisions, yet it might also come with higher cognitive costs. 
+ For such a setup, our results suggest that in many instances, a modest amount of memory may be sufficient to achieve good outcomes.
+
+%%%%%%%%%%%%%
+%%  CONCLUSION  %%
+%%%%%%%%%%%%%
+
+\section{Conclusion}\label{section:discussion}
+
+Herein, we combine analytical methods and computer simulations to explore the impact of payoff memory on the evolution of reciprocal altruism. 
 In the most extreme case, we consider individuals who update their strategies based on only one piece of information:  the last round of a single repeated game. 
-For that case, we find that individuals are less generous, and they tend to be less cooperative overall~(\FigBaseResults). 
+Compared to exisiting models where decisions are made based on expected payoffs (perfect payoff memory), we find that individuals are less generous, and they tend to be less cooperative overall~(\FigBaseResults).
 However, once individuals update their strategies based on two or more recent experiences, overall cooperation rates quickly approach the levels observed under perfect payoff memory~(\FigHigherMemory).
@@
     Unless explicitly varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$.
-    Simulations are run for $T\!=\!5\times 10^7$ time steps for each parameter combination.}
+    Simulations are run for $T\!=\!5\times 10^7$ time steps, and each point represents a single simulation run.}
     \label{fig:cooperation_rate_over_benefit_and_beta}
@@
   However, when individuals take into account at least two past interactions -- cases ({\it iii}) to ({\it vi}) -- evolving cooperation rates are close to this optimum. 
-  Baseline parameters are the same as in \FigDependenceParameters.}
+  Simulations are run for $T\!=\!10^7$ time steps, based on the parameters $N\!=\!100$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.999$.}
 \label{fig:cooperation_rate_all_updating_payoffs}

### 2024-03-18 paper/main.tex
add weak selection comments

--- 2024-03-14 paper/main.tex
@@
 This again reflects our earlier observation that the feasibility of cooperation in this scenario is largely independent of the exact values of $b$ and $c$, as described by Eq.~\eqref{Eq:LimitedMemory}.
-With respect to the effect of different selection strengths, \FigDependenceParameters{\it b} suggests that both perfect and limited payoff memory yields similar cooperation rates for weak selection \(\beta \!<\!
-1\).
+With respect to the effect of different selection strengths, \FigDependenceParameters{\it b} suggests that both perfect and limited payoff memory yield similar cooperation rates for weak selection (\(\beta \!\ll\!
+1\)).
+This is not a coincidence, since it is known that, under weak selection, stochastic payoffs can be replaced by their (deterministic) expectations without altering the evolutionary dynamics \citep{mcavoy:NHB:2020}--and perfect payoff memory corresponds to the expected value of the payoffs in the limited payoff memory model, due to the law of large numbers.
 Beyond weak selection, increasing selection has a positive effect under perfect payoff memory, but a negative effect under limited payoff memory.\\

### 2024-03-18 paper/si.tex
memory-one -> memory-1 (for consistency with main text) + fix typos

@@
 Finally, Section~\ref{section:furthersimulations} presents further simulation results. 
-In particular, we confirm that our main results continue to hold ({\it i}) when mutations are no longer rare, and  ({\it ii}) when players use memory-one strategies instead of reactive strategies.
+In particular, we confirm that our main results continue to hold ({\it i}) when mutations are no longer rare, and  ({\it ii}) when players use memory-1 strategies instead of reactive strategies.
 
@@
 We consider this limit of rare-mutation throughout the main text. 
-The respective process is summarised by Algorithm~\ref{algorithm:pairwise_comparison}.
+The respective process is summarized by Algorithm~\ref{algorithm:pairwise_comparison}.
 In Section~\ref{section:furthersimulations}, we present additional simulation results to show that our qualitative results continue to hold when the mutation rate is strictly bounded away from zero.
@@
 ({\it v})~based on the average payoff of one interaction.
-In each case, we assumet there are only two strategies present in the population (a resident and a mutant strategy). 
+In each case, we assume there are only two strategies present in the population (a resident and a mutant strategy). 
 We first derive how likely it is that a learner (of any type) assigns a given payoff $\pi_\learner$ to itself, and a payoff of $\pi_\rolemodel$ to the role model. 
@@
 
-\subsection{Simulations for memory-one strategies}\label{section:memory_one}
+\subsection{Simulations for memory-1 strategies}\label{section:memory_one}
 
@@
 To demonstrate that our results hold for more complex strategy sets, 
-in the following we present results when players can choose among all memory-one strategies~\citep{sigmund2010calculus}. 
+in the following we present results when players can choose among all memory-1 strategies~\citep{sigmund2010calculus}. 
 Players with these strategies consider the full outcome of the
@@
 round,  \((C, C), (C, D), (D, C), (D, D)\). 
-A memory-one strategy \(\strategy\) can thus be
+A memory-1 strategy \(\strategy\) can thus be
 written as a five-dimensional vector \(\strategy=(y, p_1, p_2, p_3, p_4)\). 
@@
 cooperates in all subsequent rounds, depending on the outcome of the previous round.
-Every reactive strategy $(y,p,q)$ can be written as a memory-one strategy, by using the memory-one representation $(y,p,q,p,q)$.
-Conversely, there are memory-one strategies that cannot be represented by a reactive strategy, such as the strategy \texttt{WSLS}$=(1,1,0,0,1)$. 
-Therefore, the set of memory-one strategies is a strict superset of the reactive strategies.
-
-To explore the dynamics among memory-one strategies, we perform four separate simulations, for both perfect and limited memory, and for two different benefit values. 
+Every reactive strategy $(y,p,q)$ can be written as a memory-1 strategy, by using the memory-1 representation $(y,p,q,p,q)$.
+Conversely, there are memory-1 strategies that cannot be represented by a reactive strategy, such as the strategy \texttt{WSLS}$=(1,1,0,0,1)$. 
+Therefore, the set of memory-1 strategies is a strict superset of the reactive strategies.
+
+To explore the dynamics among memory-1 strategies, we perform four separate simulations, for both perfect and limited memory, and for two different benefit values. 
 Supplementary~Fig.~\ref{fig:memory_one_low_benefit} shows the results for a comparably low benefit of cooperation; 
@@
 These simulations suggest that even when individuals
-are allowed to use memory-one strategies, there is more cooperation when players update their strategies based on perfect payoff memory.
+are allowed to use memory-1 strategies, there is more cooperation when players update their strategies based on perfect payoff memory.
 
@@
   \includegraphics[width=0.8\textwidth]{static/memory_one_results_low_benefit.pdf}
-  \caption{\textbf{Evolution of memory-one strategies when the benefit of cooperation is comparably small.}
+  \caption{\textbf{Evolution of memory-1 strategies when the benefit of cooperation is comparably small.}
   We perform two independent simulations. In one simulation, individuals use
@@
   cooperation rates for each simulation. 
-  For memory-one strategies, like for reactive strategies, the use of expected payoffs results in more cooperation. 
+  For memory-1 strategies, like for reactive strategies, the use of expected payoffs results in more cooperation. 
   The right panel reports the most abundant strategy of each
@@
   \includegraphics[width=0.8\textwidth]{static/memory_one_results_high_benefit.pdf}
-  \caption{\textbf{Evolution of memory-one strategies when the benefit is high.}
+  \caption{\textbf{Evolution of memory-1 strategies when the benefit is high.}
   We consider a similar setup as in Supplementary~Fig.~\ref{fig:memory_one_low_benefit}, but now using a larger benefit value $b\!=\!10$.}\label{fig:memory_one_high_benefit}
@@
   \includegraphics[width=0.8\textwidth]{static/cooperation_rate_over_b_and_beta_memory_one.pdf}
-  \caption{{\bf Evolution of direct reciprocity for different parameter values for memory-one strategies.}
-  To further explore the robustness of our results in the case of memory-one
+  \caption{{\bf Evolution of direct reciprocity for different parameter values for memory-1 strategies.}
+  To further explore the robustness of our results in the case of memory-1
   strategies, we run simulations for different benefit values (left panels, A)

### 2024-03-18 paper/si.tex
CH: Revising the SI

@@
 In addition, we describe several model extensions in which the amount of information taken into account is in between these two extremes. 
-Finally, Section~\ref{section:furthersimulations} presents further simulation results. 
-In particular, we confirm that our main results continue to hold ({\it i}) when mutations are no longer rare, and  ({\it ii}) when players use memory-1 strategies instead of reactive strategies.
+Finally, Section~\ref{section:furthersimulations} presents further simulation results and model extensions. 
+In particular, we confirm that our main results continue to hold ({\it i}) when mutations are no longer rare, ({\it ii}) when players use memory-1 strategies instead of reactive strategies, and ({\it iii}) when players interact in a snowdrift game.
 
@@
 Here, $R$ is interpreted as the reward for mutual cooperation, $S$ is the sucker's payoff, $T$ is the temptation, and $P$ is the punishment payoff~\citep{axelrod1981evolution}. 
-Throughout this work, we parametrize these payoffs as $R\!=\!b\!-\!c$, $S\!=\!-c$, $T\!=\!b$, and $P\!=\!0$, where $b$ and $c$ are the benefit and cost of cooperation, respectively, with $b\!>\!c\!>0$.
+For most of this work, we parametrize these payoffs as $R\!=\!b\!-\!c$, $S\!=\!-c$, $T\!=\!b$, and $P\!=\!0$, where $b$ and $c$ are the benefit and cost of cooperation, respectively, with $b\!>\!c\!>0$.
 After each round, players learn their co-player's previous action. 
@@
 As a result, at any given time there are at most two different strategies present in the population~\citep{fudenberg:JET:2006,wu:JMB:2012,mcavoy:jet:2015}. 
-This assumptions makes computations more efficient, and it makes some of the results easier to interpret.
+This assumption makes computations more efficient, and it makes some of the results easier to interpret.
 In the following, we describe our implementation of the process in the rare-mutation limit in more detail.
@@
 ({\it v})~based on the last two rounds of two interactions, and
-({\it v})~based on the average payoff of one interaction.
+({\it vi})~based on the average payoff of one interaction.
 In each case, we assume there are only two strategies present in the population (a resident and a mutant strategy). 
@@
 \end{equation}
-Interestingly, this condition is the same as condition~\eqref{Eq:Condition_PerfectMemory} for the case of perfect memory when $N\! \rightarrow\! \infty$.
-
+This condition is the same as condition~\eqref{Eq:Condition_PerfectMemory} for the case of perfect memory when $N\!
+\rightarrow\!
+\infty$.
+This result is intuitive: in infinitely large populations, we can ignore the possibility that the learner and the role model interacted directly immediately prior to the updating stage.
+Given that, their expected payoff in a single repeated game with their last interaction partner is identical to their expected payoff across all repeated games, with all possible interaction partners.
 \clearpage
@@
 
-\section{Further simulation results}
+\section{Further simulation results and model extensions}
 \label{section:furthersimulations}
@@
 
-\section{Further repeated games; Snowdrift game}
+\subsection{An analysis of the snowdrift game}
 \label{section:furtherrepeatedgames}
@@
 So far, we have discussed the repeated donation game, which is only one example
-of social dilemmas. In the following section, we explore how our methodology can
-be applied to other games, with a focus on the snowdrift game.
-
-Similar to the donation game, the snowdrift game is a two-player game where
-individuals choose to cooperate (\(C\)) or defect (\(D\)) each round.
-Cooperation yields a benefit \(b\) to both players, incurring a cost \(c\) if
-the opponent defects, but only \(c/2\) if the opponent cooperates. This results
+of a social dilemma. In the following section, we apply our methodology to a different dilemma, the snowdrift game.
+Similar to the donation game, the snowdrift game is a two-player game where each round, 
+individuals can choose to cooperate~(\(C\)) or to defect~(\(D\)).
+If at least one player cooperates, both players get a benefit \(b\). 
+A cooperator pays a cost of \(c\) if the opponent defects, and of \(c/2\) if the opponent cooperates. This results
 in the following payoff matrix
-
-\begin{align}
+\begin{equation} \label{eq:snowdrift_payoff}
 	\bordermatrix{%
@@
 	} . 
-\end{align}\label{eq:snowdrift_payoff}
-
+\end{equation}
 In the following subsections, we discuss how our results extend to the snowdrift
@@
 under which conditions cooperation is stochastically stable for both perfect and
-limited memory scenarios. In Section~\ref{section:simulation_results_snowdrift},
-we present simulation results for all six cases considered in this study.
-
-\subsection{Stochastic stability of cooperation.}\label{section:stochastic_stability_snowdrift}
-
-{\bf Perfect payoff memory.} We have already discussed the general setup for the
-case of perfect memory in Section~\ref{section:perfect_memory}. Here, we derive
-the conditions under which cooperation is stochastically stable for the
-snowdrift game. As a reminder, we suppose that there is a single mutant,
+limited memory. In Section~\ref{section:simulation_results_snowdrift},
+we present simulation results.
+
+\subsubsection{Stochastic stability of cooperation.}\label{section:stochastic_stability_snowdrift}
+
+{\bf Perfect payoff memory.} As before, we suppose that there is a single mutant,
 \(k\!=\!1\) that adopts the strategy $\alld$, while the residents adopt the
 strategy \(\gtft\). The resulting average distributions between the residents
-and the mutants remain the same as in Section\ref{section:perfect_memory}.
+and the mutants remain the same as in Section~\ref{section:perfect_memory}.
 However, the expected payoffs of the strategies now differ, as we are now
@@
 Specifically, the expected payoffs are computed as,
-
-\begin{align*}
+\begin{equation*}
   \pi_\gtft =  \frac{N\!-\!2}{N-1} (b - c/2)  \!+\! \frac{1}{N-1}\big(1\!-\!\delta\!+\!\delta q\big)(b - c)
@@
   \pi_\alld  = \big(1\!-\!\delta\!+\!\delta q\big)b.
-\end{align*}
-
+\end{equation*}
 As a consequence, the corresponding ratio of transition probabilities according
-to Eq.~\eqref{eq:LambdaRatio} is now given by,
-
+to Eq.~\eqref{eq:LambdaRatio} is now given by
 \begin{align*}
@@
 \end{equation}
-For such a strategy to be feasible we require $q\!>\!0$, which implies $\delta
-\!>\! \big((N\!-\!2)b+c\big)/\big(N(c/2)\big)$. In particular, in the limit of
+For such a strategy to be feasible we require $q\!\ge\!0$, which implies $\delta
+\!\ge\! \big((N\!-\!2)b+c\big)/\big(N(c/2)\big)$. In particular, in the limit of
 large populations \(N \!\rightarrow\! \infty\), we obtain that cooperation is
@@
 \noindent
-{\bf Limited payoff memory.} Once again, we examine
-the scenario where a single \alld{} mutant can potentially invade a resident
-population of \gtft{} players. However, this time, we consider that individuals
+{\bf Limited payoff memory.} Now we assume that individuals
 only take into account the last round of their last interaction.
@@
 \end{equation}
-
 Interestingly, this condition is identical to
 condition~\eqref{Eq:TransitionRatioSimple} for the donation game. To understand
-this result, we first need to discuss the difference between the two games. The
-difference between the two games lies in the relationship between the possible
-payoffs. More specifically, in the donation game, the payoffs satisfy the
-relationship \(T > R > P > S\), whereas in the snowdrift game, it's \(T > R > S
-> P\).
-Now, note that for the above condition, the relationship between all the
+this result, we first need to discuss the difference between the two games. 
+In the donation game, the payoffs satisfy the
+relationship \(T \!>\! R \!>\! P \!>\! S\), whereas in the snowdrift game, the respective ordering is \(T \!>\! R \!>\! S
+\!>\! P\).
+Note that for the stochastic stability of \gtft, the relationship between all the
 possible payoffs is not important; what is important is the relationship between
-all the feasible pairs of payoffs. In a population of \gtft{} strategies with a
-single mutant of \alld{}, the feasible pairs of payoffs are \((R, T)\), \((R,
-P)\), \((S, T)\), and \((P, P)\). In both games, the payoff \(T\) is higher than
-\(S\) and \(R\), and \(R\) is higher than the punishment payoff, thus the
-condition remains unchanged.
-
-\subsection{Simulation results for reactive strategies}\label{section:simulation_results_snowdrift}
-
-To investigate whether the previous static observations accurately depict the
-dynamics of evolving populations, we once again rely on simulations. Firstly, we
-conducted separate simulations for perfect and limited payoff memory. In each
+all the feasible pairs of payoffs. In a \gtft{} population  with a
+single \alld{} mutant, the feasible payoff pairs are \((R, T)\), \((R,
+P)\), \((S, T)\), \((P, P)\). In all these four cases, the relative ranking of the two payoffs remains unchanged.
+
+\subsubsection{Simulation results for reactive strategies}\label{section:simulation_results_snowdrift}
+
+In addition to these static observations, we again explore the evolutionary dynamics with simulations. 
+Firstly, we conducted separate simulations for perfect and limited payoff memory. In each
 scenario, we examined both low and high benefits of cooperation (\(b/c=2\) and
@@
 they are easily outperformed by more cooperative strategies. Moreover, the
-evolving cooperation rates are high for all the cases. Even in the limited
-payoff memory for low benefit, the evolving cooperation is at 71\%. Thus, the
+evolving cooperation rates are high for all the cases. Even for limited
+payoff memory and a low benefit, the evolving cooperation is at 71\%. Thus, 
 in the snowdrift game, the memory requirement for cooperation is
@@
   \includegraphics[width=.75\textwidth]{static/snowdrift_game_results.pdf}
-  \caption{{\bf Evolutionary dynamics under perfect and limited payoff memory.}
+  \caption{{\bf Evolutionary dynamics for the snowdrift game under perfect and limited payoff memory.}
 The four panels depict the outcome of evolutionary simulations for the repeated
-snowdrift game with either a low or high benefit of cooperation. In the top
-panels, individuals have perfect payoff memory, meaning they consider all their
-past interactions against all population members, taking every turn of each
-repeated game into account. In the bottom panels, individuals have limited
-payoff memory, considering only their very last interaction against one specific
-population member, and taking into account only one round of the repeated game.
+snowdrift game with either a low or high benefit of cooperation. 
 Colors represent how often the respective region of the strategy space is
@@
 
-Furthermore, to understand the impact of payoff memory on the snowdrift game
-even more, we performed simulations for the four additional scenarios considered
-in this work. As a reminder, in the first scenario, players recall the payoffs
-they obtained in the last two rounds against a single co-player. In the second
-scenario, players recall their last-round payoffs against two co-players. In the
-third scenario, they recall the last two rounds against two co-players. Finally,
-in the last scenario, players update based on the average payoff they receive
-over all rounds with a single co-player (further extensions are possible, but we
-do not explore them here). In Supplementary
-Fig.~\ref{fig:cooperation_rate_all_updating_payoffs_snowdrift}, we show the
-evolving cooperation rates for all these scenarios, as well as for the limited
-and perfect memory cases.
-
-We observe that main pattern of our results remains unchanged: the perfect
-memory case always overestimates the cooperation rate. Compared to the baseline
+Again, we performed similar simulations when the players' payoff memory is 
+in between the two extremes of perfect and limited memory (Supplementary
+Fig.~\ref{fig:cooperation_rate_all_updating_payoffs_snowdrift}).
+The main patterns of our previous results remain unchanged: ({\it i})~Perfect
+memory always yields the highest cooperation rate; ({\it ii})~Compared to the 
 case with a single round recalled, even a minimal increase in the players'
-payoff memory further increases the cooperation rate. In all four scenarios with
-extended memory, we observe similar cooperation rates, and they quickly approach
-the rates observed under perfect memory. This happens at a faster rate compared
-to the donation game.
+memory further increases cooperation; ({\it iii})~In all scenarios with
+intermediate memory, we observe similar cooperation rates, and they quickly approach
+the rates observed under perfect memory.
 
@@
   \caption{{\bf Average cooperation rates in the snowdrift game for different payoff memories.}
-  We vary how much information individuals take into account when updating their strategies.
-  From left to right, we consider the following cases:
-  ({\it i}) Updating occurs based on expected payoffs (perfect memory).
-  ({\it ii}) It occurs based on the last round of one interaction (limited memory).
-  ({\it iii}) Based on the last round of two interactions.
-  ({\it iv}) Based on the last two rounds of one interaction.
-  ({\it v}) Based on the last two rounds of two interactions.
-  ({\it vi}) Based on the average payoff of one interaction.
-  Simulations are run either for a comparably low benefit of cooperation~($b/c\!=\!2$), or for a high benefit~($b/c\!=\!10$). 
+  Again, we vary how much information individuals take into account when updating their strategies.
+  Simulations are run either for a low benefit of cooperation~($b/c\!=\!2$), or for a high benefit~($b/c\!=\!10$). 
   We observe that perfect memory always yields the highest cooperation rate, and limited memory yields the lowest.
-  However, in the snowdrift game, the difference between the two extreme scenarios is not large, as we have already discussed.
-  When individuals take into account a bit more information—cases (iii) to (vi)—evolving
-  cooperation rates quickly approach the optimum of perfect memory. This increase happens faster compared to the donation game.
+  However, in the snowdrift game, the difference between the two extreme scenarios is smaller than in the donation game. 
+  When individuals take into account a bit more information, evolving cooperation rates quickly approach the optimum of perfect memory. 
   Simulations are run for $T\!=\!10^7$ time steps, using a cost $c\!=\!1$, a continuation probability of $\delta\!=\!0.999$ and a selection strength of $\beta\!=\!1$, in a population of size $N\!=\!100$.}

### 2024-03-18 paper/main.tex
alphabetize keywords

@@
 
-Most models of direct reciprocity abstract from these difficulties~\citep{brauchli:JTB:1999,brandt:JTB:2006,ohtsuki:JTB:2007b,szolnoki:pre:2009b,imhof2010stochastic,van-segbroeck:prl:2012,grujic:jtb:2012,Martinez2012,stewart:pnas:2013,pinheiro:PLoSCB:2014,stewart:games:2015,Baek2016,hilbe:PNAS:2017,McAvoy:ProcA:2019,glynatsi:SCR:2020,Schmid:PlosCB:2022,Murase:SciRep:2022,Cooney:BMB:2022,Chen:PNASNexus:2023}.
-They just assume individuals can easily copy the strategies of others. 
-Similarly, they just assume that updating decisions are based on the strategies' average (or expected) payoffs, which are based on all rounds and all interactions. 
+Most models of direct reciprocity do not address these difficulties~\citep{brauchli:JTB:1999,brandt:JTB:2006,ohtsuki:JTB:2007b,szolnoki:pre:2009b,imhof2010stochastic,van-segbroeck:prl:2012,grujic:jtb:2012,Martinez2012,stewart:pnas:2013,pinheiro:PLoSCB:2014,stewart:games:2015,Baek2016,hilbe:PNAS:2017,McAvoy:ProcA:2019,glynatsi:SCR:2020,Schmid:PlosCB:2022,Murase:SciRep:2022,Cooney:BMB:2022,Chen:PNASNexus:2023} and, instead, assume individuals can easily copy the strategies of others.
+Similarly, they assume that updating decisions are based on the strategies' average (or expected) payoffs, which are based on all rounds and all interactions. 
 These assumptions create a curious inconsistency in how models represent an individual's cognitive abilities. 
@@
 In particular, payoffs take the form of a prisoner's dilemma:
-Mutual cooperation yields a better payoff than mutual defection ($b\!-\!c\!>\!0$), but each player individually prefers to defect independent of the co-player's action ($b\!>\!b\!-\!c$ and $0\!>\!-c$).
-To incorporate that individuals interact repeatedly, we assume that after each round, there is a constant continuation probability $\delta$ of interacting for another round. 
+Mutual cooperation yields a better payoff than mutual defection ($b\!-\!c\!>\!0$), but each player individually prefers to defect, independent of the co-player's action ($b\!>\!b\!-\!c$ and $0\!>\!-c$).
+To incorporate repetition, we assume that after each round, there is a constant continuation probability $\delta$ of interacting for another round. 
 For $\delta\!=\!0$, we recover the case of a conventional (one-shot) prisoner's dilemma. 
@@
 
-Both in the case of perfect and limited memory, we iterate the elementary strategy updating step described above for many time steps. 
+Both in the case of perfect and limited memory, we iterate the elementary strategy update step described above many times. 
 This gives rise to a stochastic process that describes which strategies players adopt over time. 
-We explore the dynamics of this process mathematically and with computer simulations.
+We explore the dynamics of this process mathematically and with numerical simulations.
 For the results presented in the following, we assume that mutations are rare (\(\mu\!\rightarrow\! 0\)). 
-This assumption is fairly common in evolutionary game theory, because it makes some computations more efficient~\citep{fudenberg:JET:2006,wu:JMB:2012,mcavoy:jet:2015}, and because the results can be interpreted more easily.
+This assumption is fairly common in evolutionary game theory, both because it makes some computations more efficient~\citep{fudenberg:JET:2006,wu:JMB:2012,mcavoy:jet:2015} and because the results can be interpreted more easily.
 However, in Section~3 of the \esm{} we show that our main results continue to hold for strictly positive mutation rates.
@@
 As a result, the mutant's expected payoff is $\pi_\alld \!=\! (1\!-\!\delta\!+\!\delta q)b$. 
-For perfect payoff memory, the requirement for cooperation to be stochastically stable reduces to the condition $\pi_\gtft > \pi_\alld$. 
-This yields
+For perfect payoff memory, the requirement for cooperation to be stochastically stable reduces to the condition $\pi_\gtft > \pi_\alld$, which yields
 \begin{equation} \label{Eq:PerfectMemory}
@@
 In particular, we recover the previous observation that $q\!=\!1-c/(\delta b)$ is the maximum generosity that cooperators should have~\citep{molander:jcr:1985,Nowak1992tit,Schmid:NHB:2021}. 
-Because $q\!\ge\!0$, we also conclude that cooperation can only be stable if $\delta \!>\! c/b$.
-Again, this condition for the feasibility of direct reciprocity is the condition found in the literature~\citep{nowak:Science:2006}.
+Because $q\!\ge\!0$, we also conclude that cooperation can be stable only if $\delta \!>\! c/b$.
+Again, this condition for the feasibility of direct reciprocity already exists in the literature~\citep{nowak:Science:2006}.
 
@@
 {\bf Evolutionary dynamics of reciprocity.}
-To explore whether the previous static observations describe the dynamics of evolving populations, we turn to simulations.
+To explore whether the previous observations describe the dynamics of evolving populations, we turn to simulations.
 We have run separate simulations for perfect and limited payoff memory. 
@@
 1\)).
-This is not a coincidence, since it is known that, under weak selection, stochastic payoffs can be replaced by their (deterministic) expectations without altering the evolutionary dynamics \citep{mcavoy:NHB:2020}--and perfect payoff memory corresponds to the expected value of the payoffs in the limited payoff memory model, due to the law of large numbers.
+This is not a coincidence: it is known that, under weak selection, stochastic payoffs can be replaced by their (deterministic) expectations without altering the evolutionary dynamics \citep{mcavoy:NHB:2020}--and perfect payoff memory corresponds to the expected value of the payoffs in the limited payoff memory model, due to the law of large numbers.
 Beyond weak selection, increasing selection has a positive effect under perfect payoff memory, but a negative effect under limited payoff memory.\\
@@
 \noindent
-{\bf Beyond reactive strategies.} While the results presented in the main text focus on reactive strategies, the patterns we observe do not seem to depend on the considered strategy space. 
+{\bf Beyond reactive strategies.} While the results presented in the main text focus on reactive strategies, the patterns we observe do not appear to be sensitive to the strategy space. 
 To illustrate this point in more detail, in the \esm{} we consider the dynamics among memory-1 strategies. 
 Here, players take into account both their co-player's and their own last move, see Refs.~\citep{nowak:Nature:1993,imhof:JTB:2007}. 
-Also in that case, we observe that perfect memory leads to systematically higher cooperation rates (Supplementary Figures~5,6).
+In that case, we also observe that perfect memory leads to systematically higher cooperation rates (Supplementary Figures~5,6).
 Again, this advantage of perfect memory is particularly pronounced for strong selection, or when there is a high benefit of cooperation (\FigMemoryOneParameters).\\
@@
 
-We complement these analytical results with additional simulations, see \FigHigherMemory.
+We complement these analytical results with additional simulations; see \FigHigherMemory.
 We observe that a minimal increase in the players' payoff memory (compared to the baseline case with a single round recalled) can promote cooperation considerably. 
@@
 So far, we have discussed the donation game, which is the most common
-example of social dilemmas. There are several other examples, which include the
+example of a social dilemma. There are several other examples, which include the
 snowdrift game~\cite{doebeli:Ecology:2005}, the stag hunt
 game~\cite{pacheco:ProcB:2009}, and the public goods
-game~\cite{pinheiro:PLoSCompBio:2014}. However, the pattern we observe is more
+game~\cite{pinheiro:PLoSCompBio:2014}. The pattern we observe is more
 general and still holds for other social dilemmas. To demonstrate this, we
@@
 Apart from the analytical analysis, we evaluate our results with simulations,
-and we do this for all the updating payoff cases considered in this manuscript.
+which we do for all cases considered in this manuscript.
 What we find is that the qualitative results still hold for the snowdrift game.
@@
     Unless explicitly varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$.
-    Each point represents a single simulation run. Simulations are run for $T\!=\!5\times 10^7$ time steps for each parameter combination. Each point is a sinlge simualtion run.}
+    Each point represents a single simulation run. Simulations are run for $T\!=\!5\times 10^7$ time steps for each parameter combination. Each point is a single simulation run.}
     \label{fig:cooperation_rate_over_benefit_and_beta}

### 2024-03-18 paper/First submission/si_old.tex

--- 2024-03-15 paper/si.tex
@@
 \def\alld{\texttt{ALLD}}
-\def\allc{\texttt{ALLC}}
 \def\tft{\texttt{TFT}}
@@
 
-\clearpage
-
-%%%%%%%%%%%%%%%%%%
-%% FURTHER REPEATED GAMES %% 
-%%%%%%%%%%%%%%%%%%
-
-\section{Further repeated games; Snowdrift game}
-\label{section:furtherrepeatedgames}
-
-So far, we have discussed the repeated donation game, which is only one example
-of social dilemmas. In the following section, we explore how our methodology can
-be applied to other games, with a focus on the snowdrift game.
-
-Similar to the donation game, the snowdrift game is a two-player game where
-individuals choose to cooperate (\(C\)) or defect (\(D\)) each round.
-Cooperation yields a benefit \(b\) to both players, incurring a cost \(c\) if
-the opponent defects, but only \(c/2\) if the opponent cooperates. This results
-in the following payoff matrix
-
-\begin{align}
-	\bordermatrix{%
-		& C & D \cr
-		C & b - \frac{c}{2} & b - c  \cr
-		D & b & 0 \cr
-	} . 
-\end{align}\label{eq:snowdrift_payoff}
-
-In the following subsections, we discuss how our results extend to the snowdrift
-game. In Section~\ref{section:stochastic_stability_snowdrift}, we investigate
-under which conditions cooperation is stochastically stable for both perfect and
-limited memory scenarios. In Section~\ref{section:simulation_results_snowdrift},
-we present simulation results for all six cases considered in this study.
-
-\subsection{Stochastic stability of cooperation.}\label{section:stochastic_stability_snowdrift}
-
-{\bf Perfect payoff memory.} We have already discussed the general setup for the
-case of perfect memory in Section~\ref{section:perfect_memory}. Here, we derive
-the conditions under which cooperation is stochastically stable for the
-snowdrift game. As a reminder, we suppose that there is a single mutant,
-\(k\!=\!1\) that adopts the strategy $\alld$, while the residents adopt the
-strategy \(\gtft\). The resulting average distributions between the residents
-and the mutants remain the same as in Section\ref{section:perfect_memory}.
-However, the expected payoffs of the strategies now differ, as we are now
-considering the snowdrift payoff matrix~\eqref{eq:snowdrift_payoff}. \\
-
-\noindent
-Specifically, the expected payoffs are computed as,
-
-\begin{align*}
-  \pi_\gtft =  \frac{N\!-\!2}{N-1} (b - c/2)  \!+\! \frac{1}{N-1}\big(1\!-\!\delta\!+\!\delta q\big)(b - c)
-  \quad \text{and} \quad
-  \pi_\alld  = \big(1\!-\!\delta\!+\!\delta q\big)b.
-\end{align*}
-
-As a consequence, the corresponding ratio of transition probabilities according
-to Eq.~\eqref{eq:LambdaRatio} is now given by,
-
-\begin{align*}
-\frac{\lambda_1^{-}}{\lambda_1^{+}} = e^{-\beta \left( (1-\delta+\delta q)(b - \frac{b-c}{N-1})- \frac{N-2}{N-1}(b-c/2) \right)}
-\end{align*}
-By definition, cooperation is stochastically stable if this ratio exceeds one, which is equivalent to
-\begin{equation}\label{Eq:Condition_PerfectMemory_Snowdrift}
-q<1-\frac{1}{\delta}\cdot \frac{N(c/2)}{(N\!-\!2)b+c}.
-\end{equation}
-For such a strategy to be feasible we require $q\!>\!0$, which implies $\delta
-\!>\! \big((N\!-\!2)b+c\big)/\big(N(c/2)\big)$. In particular, in the limit of
-large populations \(N \!\rightarrow\! \infty\), we obtain that cooperation is
-stochastically stable if  \(q \!<\! 1\! -\! c/(2\delta b)\). \\
-
-\noindent
-{\bf Limited payoff memory.} Once again, we examine
-the scenario where a single \alld{} mutant can potentially invade a resident
-population of \gtft{} players. However, this time, we consider that individuals
-only take into account the last round of their last interaction.
-Based on the probabilities \(x(u_\resident,u_\mutant)\), where \(u_\resident\)
-represents the payoff of a randomly chosen \gtft{} player and \(u_\mutant\)
-represents the payoff of the \alld{} player, with both \(u_\resident\) and
-\(u_\mutant\) being elements of \(\mathcal{U}\) (see
-Section~\ref{section:limited_memory}), we calculate the ratio of transition
-probabilities as:
-
-\begin{equation*}
-\frac{\lambda^{-}_1}{\lambda^{+}_1} = \frac{ \frac{N - 2}{N - 1}  
-\left(\frac{ 1 - \delta + \delta q}{1+e^{ \beta (c/2)} } 
-  + \frac{ \delta  \left(1 - q \right)}{1+ e^{-  \beta  \left(b - c/2 \right)}} \right)  
-  +  \frac{1}{N-1}  \left(\frac{1 - \delta + \delta q}{1 + e^{\beta c}}
-  + \frac{ \delta  \left(1 - q \right)}{2} \right)}
-  { \frac{N - 2}{N - 1}  
-\left( \frac{ 1 - \delta + \delta q}{1 + e^{-\beta (c/2)}}
-+  \frac{ \delta  (1 - q)}{1 +e^{  \beta (b - c/2)}} \right) 
-+  \frac{1}{N -1} \left(\frac{ 1 - \delta + \delta q}{1 + e^{-  \beta c}}
-+  \frac{ \delta  \left(1 - q \right)}{2} \right)} .
-\end{equation*}
-Cooperation is stochastically stable if $\lambda^-_1/\lambda^+_1\!>\!1$. 
-We consider the limit of strong selection \(\beta\! \rightarrow\! \infty\) and
-large populations \(N \!\rightarrow \!\infty \). In that case, because $b\!>\!c\!>\!0$, the above ratio simplifies to
-\begin{equation}\label{Eq:Condition_LimitedMemory_Snowdrift}
-    \frac{\lambda^{-}_1}{\lambda^{+}_1} = \frac{\delta(1 - q)}{1 - \delta + \delta q}.
-\end{equation}
-
-Interestingly, this condition is identical to
-condition~\eqref{Eq:TransitionRatioSimple} for the donation game. To understand
-this result, we first need to discuss the difference between the two games. The
-difference between the two games lies in the relationship between the possible
-payoffs. More specifically, in the donation game, the payoffs satisfy the
-relationship \(T > R > P > S\), whereas in the snowdrift game, it's \(T > R > S
-> P\).
-Now, note that for the above condition, the relationship between all the
-possible payoffs is not important; what is important is the relationship between
-all the feasible pairs of payoffs. In a population of \gtft{} strategies with a
-single mutant of \alld{}, the feasible pairs of payoffs are \((R, T)\), \((R,
-P)\), \((S, T)\), and \((P, P)\). In both games, the payoff \(T\) is higher than
-\(S\) and \(R\), and \(R\) is higher than the punishment payoff, thus the
-condition remains unchanged.
-
-\subsection{Simulation results for reactive strategies}\label{section:simulation_results_snowdrift}
-
-To investigate whether the previous static observations accurately depict the
-dynamics of evolving populations, we once again rely on simulations. Firstly, we
-conducted separate simulations for perfect and limited payoff memory. In each
-scenario, we examined both low and high benefits of cooperation (\(b/c=2\) and
-\(b/c=10\), respectively). Supplementary Fig.
-\ref{fig:expected_and_stochastic_for_snowdrift} illustrates the conditional
-cooperation probabilities \(p\) and \(q\) of the evolving populations. Overall,
-we find that the evolving populations are more cooperative compared to those of
-the donation game. There is no cluster around the neighborhood of \alld{} as
-they are easily outperformed by more cooperative strategies. Moreover, the
-evolving cooperation rates are high for all the cases. Even in the limited
-payoff memory for low benefit, the evolving cooperation is at 71\%. Thus, the
-in the snowdrift game, the memory requirement for cooperation is
-even more modest compared to the donation game.
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=.75\textwidth]{static/snowdrift_game_results.pdf}
-  \caption{{\bf Evolutionary dynamics under perfect and limited payoff memory.}
-The four panels depict the outcome of evolutionary simulations for the repeated
-snowdrift game with either a low or high benefit of cooperation. In the top
-panels, individuals have perfect payoff memory, meaning they consider all their
-past interactions against all population members, taking every turn of each
-repeated game into account. In the bottom panels, individuals have limited
-payoff memory, considering only their very last interaction against one specific
-population member, and taking into account only one round of the repeated game.
-Colors represent how often the respective region of the strategy space is
-visited over time.
-In all cases, there is one dominant cluster, and that is the cluster of
-conditionally cooperative strategies with \(p \approx 1\), where \(q\) satisfies
-the constraints \eqref{Eq:Condition_PerfectMemory_Snowdrift} and
-\eqref{Eq:Condition_LimitedMemory_Snowdrift}, respectively.
-Simulations are run for \(T = 10^7\) time steps, using a cost \(c = 1\), a
-continuation probability of \(\delta = 0.999\), and a selection strength of
-\(\beta = 1\), in a population of size \(N = 100\).}
-\label{fig:expected_and_stochastic_for_snowdrift}
-\end{figure}
-
-Furthermore, to understand the impact of payoff memory on the snowdrift game
-even more, we performed simulations for the four additional scenarios considered
-in this work. As a reminder, in the first scenario, players recall the payoffs
-they obtained in the last two rounds against a single co-player. In the second
-scenario, players recall their last-round payoffs against two co-players. In the
-third scenario, they recall the last two rounds against two co-players. Finally,
-in the last scenario, players update based on the average payoff they receive
-over all rounds with a single co-player (further extensions are possible, but we
-do not explore them here). In Supplementary
-Fig.~\ref{fig:cooperation_rate_all_updating_payoffs_snowdrift}, we show the
-evolving cooperation rates for all these scenarios, as well as for the limited
-and perfect memory cases.
-
-We observe that main pattern of our results remains unchanged: the perfect
-memory case always overestimates the cooperation rate. Compared to the baseline
-case with a single round recalled, even a minimal increase in the players'
-payoff memory further increases the cooperation rate. In all four scenarios with
-extended memory, we observe similar cooperation rates, and they quickly approach
-the rates observed under perfect memory. This happens at a faster rate compared
-to the donation game.
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=\textwidth]{static/snowdrift_game_more_memory.pdf}
-  \caption{{\bf Average cooperation rates in the snowdrift game for different payoff memories.}
-  We vary how much information individuals take into account when updating their strategies.
-  From left to right, we consider the following cases:
-  ({\it i}) Updating occurs based on expected payoffs (perfect memory).
-  ({\it ii}) It occurs based on the last round of one interaction (limited memory).
-  ({\it iii}) Based on the last round of two interactions.
-  ({\it iv}) Based on the last two rounds of one interaction.
-  ({\it v}) Based on the last two rounds of two interactions.
-  ({\it vi}) Based on the average payoff of one interaction.
-  Simulations are run either for a comparably low benefit of cooperation~($b/c\!=\!2$), or for a high benefit~($b/c\!=\!10$). 
-  We observe that perfect memory always yields the highest cooperation rate, and limited memory yields the lowest.
-  However, in the snowdrift game, the difference between the two extreme scenarios is not large, as we have already discussed.
-  When individuals take into account a bit more information—cases (iii) to (vi)—evolving
-  cooperation rates quickly approach the optimum of perfect memory. This increase happens faster compared to the donation game.
-  Simulations are run for $T\!=\!10^7$ time steps, using a cost $c\!=\!1$, a continuation probability of $\delta\!=\!0.999$ and a selection strength of $\beta\!=\!1$, in a population of size $N\!=\!100$.}
-\label{fig:cooperation_rate_all_updating_payoffs_snowdrift}
-\end{figure}
-
 %%%%%%%%%%%%
@@
 \bibliography{bibliography}
-}}
+}
 

### 2024-03-18 paper/main.tex
adding missing section

--- 2024-03-18 paper/main.tex
@@
 This process assumes that at regular time intervals, one population member is randomly selected, and given the chance to revise its strategy.
-We refer to this player as the ``learner.'' 
+We refer to this player as the ``learner''.
 With probability $\mu$ (reflecting a mutation rate), the learner simply adopts a random strategy (all reactive strategies have the same probability to be chosen). 
@@
 This finding is reminiscent of our earlier static results, which also suggested that changes in the game's benefits may have a negligible role on the stochastic stability of full cooperation.  
-With respect to the effect of different selection strengths, \FigDependenceParameters{\it b} suggests that both perfect and limited payoff memory yield similar cooperation rates for weak selection \(\beta \!<\!
-1\).
+With respect to the effect of different selection strengths, \FigDependenceParameters{\it b} suggests that both perfect and limited payoff memory yield similar cooperation rates for weak selection (\(\beta \!\ll\!
+1\)).
+This is not a coincidence: it is known that, under weak selection, stochastic payoffs can be replaced by their (deterministic) expectations without altering the evolutionary dynamics \citep{mcavoy:NHB:2020}--and perfect payoff memory corresponds to the expected value of the payoffs in the limited payoff memory model, due to the law of large numbers.
 Beyond weak selection, increasing selection has a positive effect under perfect payoff memory, but a negative effect under limited payoff memory.\\

### 2024-03-20 matlab/evolSimulation.m
cleaning and documentating the MATLAB code

@@ -1,5 +1,5 @@
-function [xDat, AvCoop, AvPay,payoff_type, Data]=evolSimulation(starting_resident, u, N, delta, beta, numberIterations, payoff_type, filename);
-rng('default')
+function [xDat, payoff_type, Data]=evolSimulation(starting_resident, u, N, delta, beta, numberIterations, payoff_type, filename, seed);
+rng(seed)
 %% Preparations for the output
 Data=['R=',num2str(u(1)),'; S=',num2str(u(2)),'; T=',num2str(u(3)), '; P=',num2str(u(4)),'; N=',num2str(N),'; beta=',num2str(beta), '; nIt=',num2str(numberIterations), '; payoff=',payoff_type];
 AvCoop=0; AvPay=0; Res=starting_resident;
@@ -8,56 +8,59 @@ AvCoop=0; AvPay=0; Res=starting_resident;
 sdim=3;
 xDat=zeros(numberIterations/100,6);
 xDat(1,:)=[Res, 0, u(4), 0];
-Rho=calcRho(u, beta);
+Phi=calcPhi(u, beta);
 
 %% Running the evolutionary process
 j = 2;
 for t = progress(1:numberIterations)
     Mut=rand(1, sdim);
-    [phi, coopM, piM]=calcPhi(Mut, Res, Rho, N, u, delta, beta, payoff_type);
-    if rand(1)<phi
+    [rho, coopM, piM]=calcRho(Mut, Res, Phi, N, u, delta, beta, payoff_type);
+    if rand(1)<rho
         Res=Mut; xDat(j,:)=[Res, coopM, piM, t]; j=j+1;
     end
 end
 
 dlmwrite(filename + ".csv", xDat, 'precision', 9);
 writematrix(Data, filename + ".txt");
-
-AvCoop = mean(xDat(:,end-2));
-AvPay = mean(xDat(:,end-1));
 end
 
-function [Rho]=calcRho(u, beta);
+function [Phi]=calcPhi(u, beta);
 %% Calculates all possible pairwise imitation probabilities based on one payoff
-    Rho=zeros(4,4);
+    Phi=zeros(4,4);
     for i1=1:4
         for i2=1:4
-            Rho(i1,i2) = 1 / (1 + exp(-beta * (u(i2) - u(i1))));
[…]

### 2024-03-20 matlab/rhoOneOpponent.m

@@ -1,0 +1,23 @@
+function [rho]=rhoOneOpponent(N, piMM, piMR, piRR, piRM, beta);
+%% Calculates the fixation probability based on the last round payoffs
+
+rhos = [1 / (1 + exp(-beta * (piRM - piMR)));
+        1 / (1 + exp(-beta * (piRM - piMM))); 
+        1 / (1 + exp(-beta * (piRR - piMM))); 
+        1 / (1 + exp(-beta * (piRR - piMR)));
+        1 / (1 + exp(-beta * (piMR - piRM)));
+        1 / (1 + exp(-beta * (piMM - piRM)));
+        1 / (1 + exp(-beta * (piMM - piRR)));
+        1 / (1 + exp(-beta * (piMR - piRR)));];
+
+laplus=zeros(1, N-1); laminus=laplus;
+for k=1:N-1
+    laminus(k) = 1/(N-1)*rhos(1)...
+                + (1-1/(N-1))/(N-2)/(N-3) * ((k-1)*(k-2)*rhos(2) + (k-1)*(N-k-1)*rhos(1) + (N-k-1)*(k-1)*rhos(3) + (N-k-1)*(N-k-2)*rhos(4));
+   
+    laplus(k) = 1/(N-1)*rhos(5)...
+                + (1-1/(N-1))/(N-2)/(N-3) * ((k-1)*(k-2)*rhos(6) + (k-1)*(N-k-1)*rhos(5) + (N-k-1)*(k-1)*rhos(7) + (N-k-1)*(N-k-2)*rhos(8));
+end
+
+rho = 1 / (1 + sum(cumprod(laminus./laplus)));
+end

### 2024-03-20 matlab/tools/calcPhiSixteen.m

@@ -1,4 +1,4 @@
-function [Rho]=calcRhoSixteen(u, beta);
+function [Phi]=calcPhiSixteen(u, beta);
 %% Calculates all possible pairwise imitation probabilities based on the last two payoffs
     Us = zeros(16, 16);
 
@@ -9,11 +9,11 @@ function [Rho]=calcRhoSixteen(u, beta);
     end
 
     Us = Us / 2;
-    Rho = zeros(16, 16);
+    Phi = zeros(16, 16);
 
     for i=1:16
         for j=1:16
-            Rho(i, j) = 1 / (1 + exp(-beta * Us(i, j)));
+            Phi(i, j) = 1 / (1 + exp(-beta * Us(i, j)));
         end
     end
 end

### 2024-03-20 scripts/clean_data.py
cleaning notebooks

@@ -8,7 +8,7 @@ import pandas as pd
 
 if __name__ == "__main__":  # pragma: no cover
     folder = sys.argv[1]
-    starting_characters = "test_low_benefit_limited"  # sys.argv[2]
+    starting_characters = sys.argv[2]
 
     in_columns = [
         "y",

### 2024-03-20 matlab/evolRun.m
Code base (#17)
* cleaning README and documenting script folder

* create tools folder

* cleaning and documentating the MATLAB code

* cleaning notebooks

@@ -1,16 +1,66 @@
-function evolRun();
+function evolRun(payoff_type);
+% `payoff_type` is an input to the evolutionary simulation function. The payoff
+% type can take six different values:
+%
+% 1. "expected" - Updating occurs based on expected payoffs (perfect memory).
+% 2. "last_round" - Updating occurs based on the last round of one interaction
+%    (limited memory).
+% 3. "one_opponent" - Updating occurs based on the average payoff of one
+%    interaction.
+% 4. "two_rounds" - Updating occurs based on the last two rounds of one
+%    interaction.
+% 5. "two_opponents" - Updating occurs based on the last round of two
+%    interactions.
+% 6. "two_rounds_opponents" - Updating occurs based on the last two rounds of two
+%    interactions.
 
 starting_resident = [0, 0, 0];
 N = 100;
 delta = 0.999;
 beta = 1;
-numberIterations = 10 ^ 7;
-payoff_type = "expected";
-payoffs = readmatrix("two_by_two_games_values.csv");
+numberIterations = 10^7;
 
-parfor (i = 1:11)
-    u = payoffs(i, :);
-    filename = "../data/two_by_two_games_values_S" + u(2) + "_T_" + u(2);
+% `b` is the benefit of cooperation. For the results, we consider different values of `b`.
+b = 10;
+
+% `u` captures the payoff matrix of the game.
+% We can change this to capture the snowdrift game.
+u = [b - 1, -1, b, 0];
+
+if payoff_type == "expected"
+    filename = 'expected_payoff_benefit_' + string(b);
     evolSimulation(starting_resident, u, N, delta, beta, numberIterations, payoff_type, filename);
+
+elseif payoff_type == "last_round"
+    filename = 'limited_payoff_benefit_' + string(b);
+    evolSimulation(starting_resident, u, N, delta, beta, numberIterations, payoff_type, filename);
+
+elseif payoff_type == "one_opponent"
+    filename = 'one_opponent_payoff_benefit_' + string(b);
+    evolSimulation(starting_resident, u, N, delta, beta, numberIterations, payoff_type, filename);
+
+elseif payoff_type == "two_rounds"
+    filename = 'two_rounds_payoff_benefit_' + string(b);
+    evolSimulationUpdatingMemoryTwo(starting_resident, u, N, delta, beta, numberIterations, payoff_type, filename);
+
+elseif payoff_type == "two_opponents"
[…]

### 2024-03-20 matlab/evolRunRoundsTwo.m

@@ -1,15 +1,0 @@
-function evolRunRoundsTwo();
-
-starting_resident = [0, 0, 0];
-N = 100;
-delta = 0.999;
-beta = 1;
-numberIterations = 10 ^ 7;
-payoff_type = "two_rounds";
-u = [2, -1, 3, 0];
-
-filename = "../data/two_rounds_low_iterations";
-
-evolSimulationTwoRounds(starting_resident, u, N, delta, beta, numberIterations, payoff_type, filename);
-
-end

### 2024-03-20 matlab/evolSimulationUpdatingMemoryFour.m

@@ -1,4 +1,4 @@
-function [xDat, AvCoop, AvPay,payoff_type, Data]=evolSimulationTwoRoundsOpponents(starting_resident, u, N, delta, beta, numberIterations, payoff_type, filename);
+function [xDat, payoff_type, Data]=evolSimulationUpdatingMemoryFour(starting_resident, u, N, delta, beta, numberIterations, payoff_type, filename);
 %% Preparations for the output
 Data=['R=',num2str(u(1)),'; S=',num2str(u(2)),'; T=',num2str(u(3)), '; P=',num2str(u(4)),'; N=',num2str(N),'; beta=',num2str(beta), '; nIt=',num2str(numberIterations), '; payoff=',payoff_type];
 AvCoop=0; AvPay=0; Res=starting_resident;
@@ -7,7 +7,7 @@ AvCoop=0; AvPay=0; Res=starting_resident;
 sdim=3;
 xDat=zeros(round(numberIterations/(100 * 0.6)), 5);
 xDat(1,:)=[Res, 0, 0];
-Rho=calcRhoTwoFiftySix(u, beta);
+Phi=calcPhiTwoFiftySix(u, beta);
 
 ps = zeros(256, 2);
 for i1=1:256
@@ -34,34 +34,31 @@ end
 j = 2;
 for t = progress(1:numberIterations)
     Mut=rand(1, sdim);
-    [phi, coopM]=calcPhi(Mut, Res, Rho, N, delta,  payoff_type, condition_round_one, condition_round_two, ps);
-    if rand(1)<phi
+    [rho, coopM]=calcPhi(Mut, Res, Phi, N, delta,  payoff_type, condition_round_one, condition_round_two, ps);
+    if rand(1)<rho
         Res=Mut; xDat(j,:)=[Res, coopM, t]; j=j+1;
     end
 end
 
 csvwrite(filename + ".csv", xDat);
 writematrix(Data, filename + ".txt");
-
-AvCoop = mean(xDat(:,end-2));
-AvPay = mean(xDat(:,end-1));
 end
 
-function [Rho]=calcRhoTwoFiftySix(u, beta);
+function [Phi]=calcPhiTwoFiftySix(u, beta);
     %% Calculates all possible pairwise imitation probabilities based on the last two payoffs
         new_u = log(kron(exp(u),exp(u)));
-        Rho = zeros(256, 256);
+        Phi = zeros(256, 256);
     
         for i=1:256
             for j=1:256
[…]

### 2024-03-20 matlab/rhoExpected.m

@@ -1,4 +1,4 @@
-function [phi]=phiExpected(N, piMM, piMR, piRR, piRM, beta);
+function [rho]=rhoExpected(N, piMM, piMR, piRR, piRM, beta);
 %% Calculates the fixation probability based on the expected payoffs
 
 laplus=zeros(1, N-1); laminus=laplus;
@@ -10,5 +10,5 @@ for k=1:N-1
     laminus(k) = 1 / (1 + exp(-beta * (piR - piM)));
 end
 
-phi = 1 / (1 + sum(cumprod(laminus./laplus)));
+rho = 1 / (1 + sum(cumprod(laminus./laplus)));
 end

### 2024-03-20 matlab/rhoLastRound.m

@@ -1,4 +1,4 @@
-function [phi]=phiLastRound(N, vRM, vMM, vMR, vRR, Rho);
+function [rho]=rhoLastRound(N, vRM, vMM, vMR, vRR, Rho);
 %% Calculates the fixation probability based on the last round payoffs
 
 laplus=zeros(1, N-1); laminus=laplus;
@@ -14,5 +14,5 @@ for k=1:N-1
     laminus(k)=sum(sum(x.*Rho'));
 end
 
-phi = 1 / (1 + sum(cumprod(laminus./laplus)));
+rho = 1 / (1 + sum(cumprod(laminus./laplus)));
 end

### 2024-03-20 matlab/rhoTwoRounds.m

@@ -1,4 +1,4 @@
-function [laplus, laminus]=phiTwoRounds(N, vRM, vMM, vMR, vRR, Rho);
+function [rho]=rhoTwoRounds(N, vRM, vMM, vMR, vRR, Rho);
 %% Calculates the fixation probability based on the last two rounds payoff
 
 laplus = zeros(1, 1); laminus=laplus;
@@ -16,4 +16,6 @@ x(i1,i2) = 1 / (N-1) * vRM(i1) * ((i1==1 & i2==1)  |(i1==2 & i2==3)  |(i1==3 &i2
 end
 laplus(k) = sum(sum(x.*Rho));
 laminus(k) = sum(sum(x.*Rho'));
+
+rho = 1 / (1 + sum(cumprod(laminus./laplus)));
 end

### 2024-03-20 matlab/tools/payoffsExpected.m

@@ -1,6 +1,5 @@
 function [payoffs]=payoffsExpected(player_indices, population, u, N, delta);
-
-
+% The expected payoff for each member of the population given the population
 payoffs = zeros(N - 1, 2);
 
 for i=1:N 

### 2024-03-20 matlab/tools/payoffsLastRound.m

@@ -1,5 +1,5 @@
 function [payoffs]=payoffsLastRound(players_indices, population, u, N, delta);
-
+% The realised payoff 
 payoffs = zeros(1, 2);
 exclude = zeros(1, 3);
 

### 2024-03-20 matlab/tools/stationaryMemoryOne.m

@@ -1,7 +1,6 @@
 function v=stationaryMemoryOne(p1,p2,delta);
-% Calculates the invariant distribution v
-% for the game between two memory-1 players with strategies p1, p2.
-% p1=(pCC, pCD, pDC, pDD, p0);
+% Calculates the probabilities of being at CC, CD, DC, DD at the last round
+% given that player 1 and 2 use memory-1 strategies
 
 M = [p1(1)*p2(1), p1(1)*(1-p2(1)), (1-p1(1))*p2(1), (1-p1(1))*(1-p2(1));
      p1(2)*p2(3), p1(2)*(1-p2(3)), (1-p1(2))*p2(3), (1-p1(2))*(1-p2(3));

### 2024-03-20 matlab/tools/stationaryRoundTwo.m

@@ -1,4 +1,6 @@
 function v=stationaryRoundTwo(s1,s2,delta);
+% Calculates the probabilities of being at CC|CC, CC|CD, CC|DC, .. etc at the last
+% two rounds.
 
 M=[s1(2) * s2(2), s1(2) * (1-s2(2)), (1-s1(2)) * s2(2), (1-s1(2)) * (1-s2(2));
     s1(3) * s2(2), s1(3) * (1-s2(2)), (1-s1(3)) * s2(2), (1-s1(3)) * (1-s2(2));

### 2024-03-20 scripts/clean_data.py

@@ -8,7 +8,7 @@ import pandas as pd
 
 if __name__ == "__main__":  # pragma: no cover
     folder = sys.argv[1]
-    starting_characters = "beta_0.01_stochastic_1"  # sys.argv[2]
+    starting_characters = sys.argv[2]
 
     in_columns = [
         "y",

### 2024-03-20 paper/main.tex
fix keyword order
submit revisions

@@
 ~\\
-{\it Keywords:} Evolution of cooperation; evolutionary dynamics; direct reciprocity; repeated prisoner's dilemma; social learning
+{\it Keywords:} Direct reciprocity; evolution of cooperation; evolutionary dynamics; repeated prisoner's dilemma; social learning
 
@@
 
-Evolutionary game theory describes the dynamics of populations when an individual's fitness depends on the traits or strategies of other population members~\cite{hofbauer1998evolutionary, nowak:Nature:2004, hauert2005game,Traulsen:PhilTrans:2022}.
+Evolutionary game theory describes the dynamics of populations when an individual's fitness depends on the traits or strategies of other population members~\citep{hofbauer1998evolutionary, nowak:Nature:2004, hauert2005game,Traulsen:PhilTrans:2022}.
 This theory can be used to describe the dynamics of animal conflict~\citep{maynard-smith:Nature:1973}, cancer cells~\citep{Stein:PTRS:2023}, and of cooperation~\citep{nowak:Science:2006}. 
-Respective models translate strategic interactions into games~\cite{smith1982evolution}. 
+Respective models translate strategic interactions into games~\citep{smith1982evolution}. 
 These games specify how individuals (players) interact, which strategies individuals can choose, and what fitness consequences (or payoffs) the different strategies have. 
@@
 
-Most models of direct reciprocity do not address these difficulties~\citep{brauchli:JTB:1999,brandt:JTB:2006,ohtsuki:JTB:2007b,szolnoki:pre:2009b,imhof2010stochastic,van-segbroeck:prl:2012,grujic:jtb:2012,Martinez2012,stewart:pnas:2013,pinheiro:PLoSCB:2014,stewart:games:2015,Baek2016,McAvoy:ProcA:2019,glynatsi:SCR:2020,Schmid:PlosCB:2022,Murase:SciRep:2022,Cooney:BMB:2022,Chen:PNASNexus:2023} and, instead, assume individuals can easily copy the strategies of others.
+Most models of direct reciprocity do not address these difficulties~\citep{brauchli:JTB:1999,brandt:JTB:2006,ohtsuki:JTB:2007b,szolnoki:pre:2009b,imhof2010stochastic,vansegbroeck:prl:2012,grujic:jtb:2012,Martinez2012,stewart:pnas:2013,pinheiro:PLoSCB:2014,stewart:games:2015,Baek2016,McAvoy:ProcA:2019,glynatsi:SCR:2020,Schmid:PlosCB:2022,Murase:SciRep:2022,Cooney:BMB:2022,Chen:PNASNexus:2023} and, instead, assume individuals can easily copy the strategies of others.
 Similarly, they assume that updating decisions are based on the strategies' average (or expected) payoffs, which are based on all rounds and all interactions. 
@@
 On the one hand, when playing the game, individuals are often assumed to have restricted memory. 
-Respective studies typically assume that individuals make their decisions each round based on the outcome of the last round only~\citep[with only a few exceptions, see Refs.][]{Hauert1997,van-veelen:PNAS:2012,Stewart2016,hilbe:PNAS:2017,Li:NatCS:2022,Murase:PLoSCompBio:2023a}.
+Respective studies typically assume that individuals make their decisions each round based on the outcome of the last round only~\citep[with only a few exceptions, see Refs.]{Hauert1997,vanveelen:PNAS:2012,Stewart2016,hilbe:PNAS:2017,Li:NatCS:2022,Murase:PLoSCompBio:2023a}.
 Yet when learning new strategies, individuals are assumed to remember (or compute) each others' precise average payoff across many rounds and many interaction partners. 
@@
 %% RESULTS: Snowdrift Game
-In a second step, we also analyze the evolutionary dynamics of a different social dilemma, the snowdrift game~\cite{doebeli:Ecology:2005}. 
+In a second step, we also analyze the evolutionary dynamics of a different social dilemma, the snowdrift game~\citep{doebeli:Ecology:2005}. 
 In this game, individuals can again cooperate or defect. 
@@
 According to this mechanism, cooperation pays if it makes the interaction partner more cooperative in future. 
-To describe which strategies are most effective, the previous theoretical literature assumes that the evolutionary dynamics are driven by the players' expected payoffs~\citep{brauchli:JTB:1999,brandt:JTB:2006,ohtsuki:JTB:2007b,szolnoki:pre:2009b,imhof2010stochastic,van-segbroeck:prl:2012,grujic:jtb:2012,Martinez2012,stewart:pnas:2013,pinheiro:PLoSCB:2014,stewart:games:2015,Baek2016,McAvoy:ProcA:2019,glynatsi:SCR:2020,Schmid:PlosCB:2022,Murase:SciRep:2022}.
+To describe which strategies are most effective, the previous theoretical literature assumes that the evolutionary dynamics are driven by the players' expected payoffs~\citep{brauchli:JTB:1999,brandt:JTB:2006,ohtsuki:JTB:2007b,szolnoki:pre:2009b,imhof2010stochastic,vansegbroeck:prl:2012,grujic:jtb:2012,Martinez2012,stewart:pnas:2013,pinheiro:PLoSCB:2014,stewart:games:2015,Baek2016,McAvoy:ProcA:2019,glynatsi:SCR:2020,Schmid:PlosCB:2022,Murase:SciRep:2022}.
 To the extent that strategies are learned (not inherited), this assumption seems to impose rather stringent requirements on the individuals' cognitive abilities. 
@@
 As a result, any population member is equally likely to be a given player's last co-player. 
-A similar analysis for structured populations would be desirable, as population structure can often promote cooperation without an explicit need for reciprocity~\citep{szabo:PR:2007,van-veelen:PNAS:2012}.
+A similar analysis for structured populations would be desirable, as population structure can often promote cooperation without an explicit need for reciprocity~\citep{szabo:PR:2007,vanveelen:PNAS:2012}.
 Similarly, we assumed that after identifying a suitable role model, players can easily imitate that role model's strategy. 
@@
 {\bf Data accessibility.}
-All data and code used in this manuscript are openly available at\newline
-\texttt{https://zenodo.org/records/10066227}.\\
+All data generated for this manuscript are openly available at:
+\href{https://zenodo.org/records/10844104}{https://zenodo.org/records/10844104}.
+The code for producing the results and generating the figures is available
+on the online repository
+\href{https://github.com/Nikoleta-v3/evolution-of-cooperation-with-limited-updating-payoff-memory}{evolution-of-cooperation-with-limited-updating-payoff-memory}.\\
 
@@
     \centering
-    \includegraphics[width=\textwidth]{static/donation_expected_last_round_summary_results.pdf}
+    \includegraphics[width=\textwidth]{donation_expected_last_round_summary_results.pdf}
     \caption{{\bf Evolutionary dynamics under perfect and limited payoff memory.}
@@
     \centering
-    \includegraphics[width=.75\textwidth]{static/cooperation_rate_over_b_and_beta.pdf}
+    \includegraphics[width=.75\textwidth]{cooperation_rate_over_b_and_beta.pdf}
     \caption{{\bf Evolution of direct reciprocity for different parameter values.} 
@@
   \centering
-  \includegraphics[width=\textwidth]{static/more_memory_summary_results.pdf}
+  \includegraphics[width=\textwidth]{more_memory_summary_results.pdf}
   \caption{{\bf Average cooperation rates for different payoff memories.}

### 2024-05-04 paper/main.tex
final fixes

@@
 \usepackage{lineno}
-\linenumbers
 \smallskip % Erlaubt kleine Abstaende zwischen Paragraphen, falls es dem Seitenlayout hilft
@@
 A similar analysis for structured populations would be desirable, as population structure can often promote cooperation without an explicit need for reciprocity~\citep{szabo:PR:2007,vanveelen:PNAS:2012}.
-Similarly, we assumed that after identifying a suitable role model, players can easily imitate that role model's strategy. 
-In the context of our study, this might be a particular simplification, because players with limited payoff memory may also have less accurate information to infer a co-player's strategy. 
+Similarly, we assumed that after identifying a suitable role model, players can
+easily imitate that role model's strategy. Especially in repeated games, this
+assumption is less innocuous than it may appear. After all, a player can
+typically only observe the co-players' actions, but not their underlying
+strategies. This player may only attempt to infer the co-players' strategies
+from their past reactions. Once players only have a restricted recollection of
+the past, this task seems particularly difficult. In the context of our study,
+this might be a particular simplification, because players with limited payoff
+memory may also have less accurate information to infer a co-player's strategy.
 Finally, we have taken a player's payoff memory to be a given parameter of the model. 
@@
 The code for producing the results and generating the figures is available
-on the online repository
+on the online GitHub repository
 \href{https://github.com/Nikoleta-v3/evolution-of-cooperation-with-limited-updating-payoff-memory}{evolution-of-cooperation-with-limited-updating-payoff-memory}.\\
@@
     \centering
-    \includegraphics[width=\textwidth]{donation_expected_last_round_summary_results.pdf}
+    \includegraphics[width=\textwidth]{static/donation_expected_last_round_summary_results.pdf}
     \caption{{\bf Evolutionary dynamics under perfect and limited payoff memory.}
@@
     \centering
-    \includegraphics[width=.75\textwidth]{cooperation_rate_over_b_and_beta.pdf}
+    \includegraphics[width=.75\textwidth]{static/cooperation_rate_over_b_and_beta.pdf}
     \caption{{\bf Evolution of direct reciprocity for different parameter values.} 
@@
   \centering
-  \includegraphics[width=\textwidth]{more_memory_summary_results.pdf}
+  \includegraphics[width=\textwidth]{static/more_memory_summary_results.pdf}
   \caption{{\bf Average cooperation rates for different payoff memories.}

### 2024-05-04 paper/si.tex
final fixes

@@
 \date{}
-\author{Nikoleta E. Glynatsi, Alex McAvoy, Christian Hilbe}
+\author{\parbox[c]{16cm}{\centering \onehalfspacing 
+Nikoleta E. Glynatsi$^1$,  Alex McAvoy$^{2,3, \dagger}$, Christian Hilbe$^{1, \dagger}$\\ \quad \\
+$^{\rm 1}$Max Planck Research Group on the Dynamics of Social Behavior,\\ Max Planck Institute for Evolutionary Biology, Pl\"{o}n, Germany \\
+$^{\rm 2}$School of Data Science and Society, University of North Carolina at Chapel Hill,\\ Chapel Hill, NC 27599 \\
+$^{\rm 3}$Department of Mathematics, University of North Carolina at Chapel Hill,\\ Chapel Hill, NC 27599 \\
+$^{\rm \dagger}$ A.M. and C.H. contributed equally to this work.}}
 
